### 2019-05-25 main_new.tex
Initial Overleaf Import

% Template for the submission to:
%   The Annals of Probability           [aop]
%   The Annals of Applied Probability   [aap]
%   The Annals of Statistics            [aos] 
%   The Annals of Applied Statistics    [aoas]
%   Stochastic Systems                  [ssy]
%
%Author: In this template, the places where you need to add information
%        (or delete line) are indicated by {???}.  Mostly the information
%        required is obvious, but some explanations are given in lines starting
%Author:
%All other lines should be ignored.  After editing, there should be
%no instances of ??? after this line.

% use option [preprint] to remove info line at bottom
% journal options: aop,aap,aos,aoas,ssy
% natbib option: authoryear
\documentclass[aoas]{imsart}

\input{commands.tex}
\usepackage{amsmath}
\usepackage{pstricks,pst-grad}
\usepackage{graphicx}
\usepackage{floatrow}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\floatsetup[table]{capposition=top}
\usepackage{subfig}
\usepackage[utf8]{inputenc}

\usepackage[tableposition = top]{caption} % captions
\usepackage{booktabs}                     % horizontal lines in tables

% == Enable text degree
\usepackage{textcomp}

\usepackage{amsthm,amsmath,natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

% provide arXiv number if available:
%\arxiv{arXiv:0000.0000}

% == Trygve Test
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

% put your definitions there:
\startlocaldefs
\endlocaldefs

\begin{document}

\begin{frontmatter}

% "Title of the paper"
\title{Informative Sampling of Oceanographic Features using Excursion Probabilities for Multivariate Random Fields}
\runtitle{Excursion Probabilities for Informative Sampling}

% indicate corresponding author with \corref{}
% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
% \thankstext{t1}{Thanks to somebody} 
% \address{line 1\\ line 2\\ printead{e1}}
% \affiliation{Some University}

\author{\fnms{Trygve Olav} \snm{Fossum}\ead[label=e1]{trygve.o.fossum@ntnu.no}, \fnms{Jo} \snm{Eidsvik}\ead[label=e2]{jo.eidsvik@math.ntnu.no}, \fnms{David} \snm{Ginsbourger} \ead[label=e3]{david.ginsbourger@idiap.ch}  \and \fnms{Kanna} \snm{Rajan}\ead[label=e4]{kanna.rajan@ntnu.no}}

\address{Department of Marine Technology\\ Trondheim, Norway\printead{e1}}
\affiliation{Department of Marine Technology, The Norwegian University of Science and Technology.}

%\address{Department of Mathematical Sciences \printead{e2}}
%\affiliation{Department of Mathematical Sciences, Norwegian University of Science and Technology (NTNU), Trondheim, Norway.}

%\address{Department of Engineering Cybernetics,\printead{e4}}
%\affiliation{Department of Engineering Cybernetics, Norwegian University of Science and Technology (NTNU), Trondheim, Norway.}

\runauthor{TO. Fossum et al.}

\begin{abstract}
We examine designs of experiments for informative sampling in the context of characterizing oceanographic phenomena. Methodologies are developed for efficient spatial sampling using autonomous robotic platforms, where the goal is to map the regions of cold and fresh water near a river mouth. Specifically, we define a design criterion based on reducing uncertainty of excursion sets in the context of multivariate random fields. We develop analytical expressions for the expected variance reduction in excursion sets, and show how these can be applicable for static designs and sequential designs building on myopic and look-ahead planning. We use simulations to study the properties of the suggested methods. Results from full-scale deployment with an autonomous underwater vehicle are also presented, for the case of river plume characterization.

%Motivated by the challenges related to efficient allocation of sampling resources in environmental sensing, the combination of Excursion Probabilities and Gaussian process modeling is explored for autonomous robotic sampling of ocean features; enabling information driven measures  sampling efforts to high-interest regions. These regions are usually characterized by gradients of measurable environmental variables, e.g., temperature or salinity gradients, on which EPs subsequently can be used...Correlation among samples and multivariate requirements are typical in environmental studies.
\end{abstract}

%\begin{keyword}[class=MSC]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\begin{keyword}
\kwd{Ocean Sampling}
\kwd{Excursion Sets}
\kwd{Gaussian Processes}
\kwd{Design of experiments}
\kwd{Adaptive information gathering}
\end{keyword}

\end{frontmatter}

\section{Introduction}

The pressure on marine resources is growing and increased accuracy, resolution, and persistent monitoring of the oceans is crucial for long-term sustainable management. A central problem for this type of environmental sensing is the large spatial extent and spatio-temporal variation inherent to ocean processes, prompting the need for effective allocation of sampling resources. 
%\begin{em}
Sampling can, in this context, not simply be distributed uniformly ---along simple transect, grid, or ``lawn-mover" patterns--- but must instead be prioritized to relevant regions to save resources and avoid redundant data.
%\end{em}

Traditional data collection at sea is to a large degree based on buoy or ship sampling, with both economic and logistic limitations that directly affect sensor coverage. Satellite data provide large-scale coverage, but have limited resolution, are limited to the surface and not available on cloudy days. A sustained focus on prioritized and efficient data collection strategies have therefore started to emerge. The advent of marine robotic platforms, especially autonomous underwater vehicles (AUVs), have provided means to execute this prioritization through the capacity of autonomy and data-driven sampling, where data collection in principle can be optimized. These capabilities have made robotics a crucial part of the emerging sensing practice for ocean science, allowing scientists to increase the observational efficiency and resolution beyond what was previously possible. But how should a robotic platform, such as an AUV, effectively prioritize and identify important regions for sampling? The answer to this question relates to statistical methodologies for design and sequential data assimilation, and the application domain is clearly an arena where statisticians can contribute.

There has recently been some statistical attention in oceanography:
\cite{wikle2013modern} present an overview of approaches for modeling uncertainties in the ocean sciences with focus on hierarchical statistical models. 
%\cite{hopkins2010new} use statistical modeling to extract features from satellite data. 
\cite{sahu2008space} study spatio-temporal models for sea surface temperature and salinity data, which will also are our key variables of interest in the current paper.
\cite{mellucci2018oceanic} looks at the statistical prediction of features using an autonomous underwater glider. 
%From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest. Each phenomenon can in turn be characterized by a set of process specific conditions expressed through different measures of key environmental variables, such as temperature or salinity. One such measure is the gradient, that can be associated with a number of important processes, such as the vertical location of the thermocline and pycnocline, location of upwelling systems, vertical mixing, eddies, fronts, and currents \cite{sverdrup2006}, as well as distribution, growth, and accumulation of biological activity \cite{SatOceanSoci00, Ryan2014}. These gradients create boundaries separating the ocean into process specific regions which are of profound interest to both identify and map effectively. Quantification of gradient features is therefore a much needed competence in robotic sampling.
The combination of statistical tools and robotic platforms is a natural one, enabling information-based robotic sensing systems. Central to this is the ability to model spatially correlated variables and provide formal measures of uncertainty. The by far most common model for practical use builds on Gaussian processes (GPs), which will also be used in our formulation, because this approach enables evaluation and prioritization of sampling efforts to be formalized into decision making systems on-board the robotic platform. 
%The imperfect understanding of causality coupled with the underlying fundamental uncertainty of the ocean implies that we need to obtain more measurements and use statistical models and methods to determine what and how phenomena form, to increase the quality of predictive statements.

Toward this end, we study excursion sets (ES) and excursion probabilities (EP) for spatial GPs, and the goal is to design sampling strategies for reducing the variance in the ES. The main contributions involve: i) extending ES-based uncertainty criteria to the situation with multivariate random fields, allowing the evaluation to include multiple environmental parameters such as temperature and salinity, ii) develop methods for informative environmental sensing and autonomous data collection, and iii) demonstrate the approach in a real case scenario, sampling a river plume using an AUV.

The remainder of the paper is organized as follows: Section \ref{sec:bg} introduces background content relating to ocean sampling, and basic notation for ESs and EPs. Section \ref{sec:GP_EP} sets up the required assumptions of GPs, including bi-variate random fields for temperature and salinity. Section \ref{sec:sur} formulates the sampling design criteria for variance reduction in ESs, with analytical derivation for multivariate spatial processes. Section \ref{sec:heuristics} builds on these sampling criteria to present heuristics for adaptive sampling strategies. Section \ref{sec:simulations} shows results and discusses properties of the methods on simulated examples. Section \ref{sec:case_study} demonstrates a full-scale implementation using an AUV to characterize a river plume. Section \ref{sec:concl_disc} contains a summary and discussion with links to future work.

\section{Background and Motivation}\label{sec:bg}

\subsection{Observing the Upper Ocean Column}

The processes occurring in the upper water column and the photic zone (upper sunlit zone) are of vital importance for long-term sustainable management of the greater marine ecosystem, climate, and human health. With the onset of the warming oceans and increased anthropogenic pressure, it is critical to improve our understanding of the associated ecosystem response. Central to this is the understanding of biophysical interaction (see Fig. \ref{fig:envir2}) driven by an agglomeration of physical forcings (wind, topography, bathymetry, tidal influences and local geography such as fresh water inflows) mixed with incipient micro-biology driven by planktonic and coastal anthropogenic input, such as pollution and agricultural runoff, which results in a range of ecosystem phenomena (e.g. blooms, anoxic zones, plumes) ---with significant impact to society. Renowned oceanographer Walter Munk called the 20\textsuperscript{th} century, the \emph{century of undersampling} \citep{munk2002}, pointing out that lack of sufficient sampling resolution in time and space, owing to current observation practices. The lack of resolution, arising from sparsely distributed sensing methods, is an issue for characterizing, understanding, and predicting ocean phenomena. In later decades, this view has been driving ocean sensing practices towards more networked and distributed use of robotic assets in order to fill the gaps in resolution (See Fig. \ref{fig:envir}). 
\begin{figure}[!h]
\centering
\subfloat[Modern ocean sensing network, where robotic assets augment ship based measurements to increase coverage of the phenomena.]{\includegraphics[width = 0.48\textwidth]{Figures/envir.pdf}\label{fig:envir1}}
\hfill
\subfloat[Different important coastal processes directly or indirectly associated with gradients and boundaries.]{\includegraphics[width = 0.48\textwidth]{Figures/envir_ocean.pdf}\label{fig:envir2}}
\caption{Ocean observation is moving away from single-ship sampling towards more collaborative networked operations in order to resolve the numerous processes and their interaction.}
\label{fig:envir}
\end{figure}

Our focus in the current paper is limited to spatial characterization of a river plume system. 
Among other plumes and front systems, Fig. \ref{fig:river_fronts} shows the survey area (Nidelva outlet, Norway) taken as cold freshwater enters from the river. The image shows the separation of water masses (different haphazard patterns and surface slicks) and strong gradients for both temperature and salinity, where cold fresh water from the river meets saline ocean water. Because of the local topography and the Coreolis force \citep{coriolis1835memoire} the cold fresh water tends to stay near land to the east. Depending on the river discharge, tidal effects, and temperature differences, this boundary in water masses will get distorted. Prior knowledge about the location and evolution of these features is therefore highly uncertain, making deterministic planning sub-optimal, and one would rather demonstrate the ability of AUVs to map the plume area efficiently. 
\begin{figure}[!h]
\centering
 \subfloat[River front - Columbia River, Astoria, Oregon, US.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/a.png}\label{fig:river1}}
	\hfill
	\subfloat[Tidal front - Korsfjorden, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/d.png}\label{fig:river2}}
 \hfill
 \subfloat[River plume - Rio de la Plata, Buenos Aires, Argentina.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/b.png}\label{fig:river3}}
 \hfill
 \subfloat[Frontal patterns - Nidelva, Trondheim, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/c.png}\label{fig:river4}}
	\caption{Different types of frontal features associated with river-ocean interaction. (\ref{fig:river1}) The river plume front of the Columbia River meeting the Pacific Ocean. (\ref{fig:river2}) Superimposed images (10 minutes) of the surface tidal front in Korsfjorden, Norway. (\ref{fig:river3}) The river plume of the Rio de la Plata, taken by European Space Agency (ESA)'s Envisat platform and MEdium Resolution Imaging Spectrometer (MERIS) sensor. (\ref{fig:river4}) Aerial photo of the Nidelva outlet in Trondheim, where fresh river water meets the salty Fjord water and creates frontal patterns. Image courtesy: (\ref{fig:river1}) \cite{zamon2014marine}, (\ref{fig:river3}) ESA, CC BY-SA 3.0 IGO.}
 \label{fig:river_fronts}
\end{figure}

%The use of \emph{robotic sampling} is motivated by the fundamental challenges of ocean observation, where a limited set of resources, a highly dynamic ocean, and the associated

%This challenge can be addressed by employing more elaborate and adaptive sampling strategies that can capitalize on both prior and current observations in order to locate and map these areas.  
River plumes belong to a class of ocean processes that are local and usually act --compared to large ocean scale processes-- on smaller scales from 10 m -- 5 km (\emph{sub-mesoscale}), where spatial lateral variability tends to dominate. Remote sensing and synthetic ocean models cannot provide detailed inputs at this scale \citep{Lermusiaux:2006} and the only way to resolve fine resolution is by direct observation. At large scale, usually referred to as above \emph{mesoscale} ($>50$ km), both three-dimensional space and time dynamics are important and can shift substantially, but in the case of river plumes one can often limit scope to cover only lateral spatial elements ([north, east] domain). For this case, time-effects can be regarded as static when acquiring AUV data over a short time window and limited region. However, the methodological framework can be extended to three-dimensional spatial variability and temporal elements in a similar manner.

\subsection{Autonomous Vehicles}
AUV is a term that describes autonomous sensor platforms that operate underwater used to gather spatio-temporal data from a targeted domain, carrying payload to survey the water column or the seafloor. They run at $\sim 1-3$ m/s, can reach depths from 100--6000m, and have battery capacity $\sim$ 8--42 hrs, depending on survey speed, payload sensors, and manufacture. By deploying AUVs one can in principle fill the gaps and augment data obtained from either ocean models, earth observing satellite data, or fixed-location buoys due to their survey speed and mobility. Traditional operations and surveys with AUVs are usually limited to planning sampling along fixed transects or survey lines set by human operators, programmed as a sequential plan consisting of different waypoints and depths. But by using on-board computers to continuously evaluate, update, and refine future sampling locations, making information gathering adaptive \citep{das11b}. For instance, an AUV could reconstruct or modify a survey line based on what temperatures it is measuring by using embedded methods in decision-theoretic planning \citep{ryan10}.

%To improve the state of sampling, modern tools and methods, including the use of autonomous platforms, oceanographic models and satellite remote sensing, needs to be combined with traditional data acquired by surface vessels or buoys (See Fig. \ref{fig:envir}). However, without adequate understanding of the theoretical underpinnings of how, when, and where to gather data, these tools and methods are insufficient in the vast and harsh oceans.

% retaining an advantageous strategy for information recovery, online during execution.
%The pressure on marine resources is growing and increased accuracy, resolution, and persistent monitoring of the oceans is crucial for long-term sustainable management. The impact of this research provides cost effective tools, techniques, and processes for doing ocean based measurements using robotic platforms. 
%adaptive design of experiments using robotic assets identifying and prioritizing relevant sampling locations on an information-theoretic basis. Spatial statistics naturally enters here through the ability to both model spatially correlated parameters and provide formal measures of uncertainty, on which this basis can be formed. 
%for oceanic sensing applications, where sensors are sparsely distributed and capitalizing on all available information is

%This imperfect understanding of causality coupled with uncertainty implies that we need to obtain more measurements and use statistical models and methods to determine what and how phenomena form, to increase the quality of predictive statements. To improve the state of sampling, modern tools and methods, including the use of autonomous platforms, oceanographic models and satellite remote sensing, augment the more traditional data acquired by surface vessels or buoys (See Fig. \ref{fig:envir}). However, without adequate understanding of the theoretical underpinnings of how, when, and where to gather data, these tools and methods are insufficient in the vast and harsh oceans.

%There are several oceanographic data sources that can be used together with statistical tools to improve data collection in ocean science. In the following section we briefly discuss buoy data, satellite data, ocean models data, and AUV data.

%Buoy data provide very accurate information of oceanographic variables, but in most situations they are local, giving information only at one (north, east) coordinate, possibly with opportunities for conducting measurements at different depths. Gliders and other surface vessels are a kind of floating buoys that drift and can measure variables at many locations, but still with limited spatial coverage.

%Satellite data are important for the mapping of oceanographic variables. They can also be indicative of variables such as temperature and salinity which we look at here, especially if the data are calibrated to for instance buoy data from the same spatial domain. However, the resolution of satellite data is relatively large, say $1 \times 1$ km $^2$ grid cells, and it only provides accurate information at the sea surface. Moreover, one cannot get useful satellite information on a cloudy day.

%What is often done in practice is to run ocean models based on the complex differential equations governing oceanographic phenomena, where also satellite data can be assimilated and used as input, and possibly playing with different forcing mechanisms to capture some of the uncertainty of the models. Initial conditions, and can be used t the goal is to guide the sampling towards regions these regions, where we can reduce the uncertainty in the ES. These models provide insight that can be understood from practitioners, but they tend to be biased, and must also be calibrated to buoy data in one way or the other.

%to provide an effective specification of regions of phenomenological interest resolving the boundaries and the position of phenomena, on which evaluation of future sampling can be constructed.

%; some of these phenomena are illustrated in Fig. \ref{fig:envir2}. Figure \ref{fig:river_fronts} illustrates the %phenomena of interest in this paper, which is river plumes and their interaction with the ocean.

%Feature driven sampling
%processes acting across a wide range of spatio-temporal scales makes prioritizing sampling efforts necessary. %Autonomous robotic platforms, can address these issues by providing the ability to focus sampling efforts to %high-interest regions (``hotspots").

Determining paths for mobile robotic sensors in order to maximize the information gained about an environment is formally known as informative path planning. This, and related problems such as the orienteering problem \citep{Golden87}, has been studied in the context of graphs, where the potential measurement are assigned to nodes on which evaluation of different routes can be conducted. 
%In this context, a fundamental result from \cite{nemhauser1978analysis} proves that a simple greedy algorithm (iteratively selecting the location which most increases the utility) can achieve a near-optimal solution if the problem can be shown to be \emph{submodular}\footnote{an intuitive diminishing returns property, where the informative value of adding sensors decrease with the number of sensors added.}. Building on this result, \cite{chekuri2005recursive} explored this in the setting of a graph using a recursive greedy algorithm, providing near-optimal solutions depending on the planning horizon and graph resolution. 
Adaptive sampling of an evolving frontal feature has been explored in \cite{Zhang2012} and \cite{Pinto2018} using state-based autonomous agents. Both approaches use a reactive-adaptive scheme, meaning that the selection does not use a statistical model of the environment in its strategic sampling effort, but rather adapts based on detection and state switching.
Myopic sampling, i.e. step-wise selection of the path (on the graph) which reduces the information criterion the most, has been used for adaptive surveying \citep{singh2009efficient,Binney2013}. These exploration strategies focus largely on reducing predictive variance or mutual information (entropy). Because the variance and entropy reduction are independent of the actual data realizations under the assumption of GP models, the use of data-driven (adaptive) criteria was introduced to include more targeted sampling of interesting regions (hotspot sampling) by \cite{Low2009} and \cite{fossuminformation}.

%\begin{figure}[h]
%\centering
%\includegraphics[width=0.4\textwidth]{Data-driven.png}
%\label{fig:auv_alone}
%\caption{The concept of data-driven sampling.}
%\end{figure}

%The goal is effectively map the extension of the river plume given this

%, we can formulate this question as follows: the ocean region  defined by the boundary between water bodies, separated by a characteristic temperature gradient that is smaller than $T=5$\textdegree C and a salinity concentration lower than $S=30$ [g/kg]. A goal is then to improve our predictive capabilities to answer this question better.
%For example,  (examples are shown in Fig. \ref{fig:river_fronts}). However,

% In doing so, the aim is to increase the knowledge about the uncertain ocean environment through adaptive sampling

\subsection{Notation and excursion sets}

In our setting we study two variables; temperature and salinity, which together characterize the phenomenon taking place in the river plume. At location $\bx \in \mathcal{M} \subset \mathbf{R^2}$, we define the two process variables by $\bxi(\bx)=(\xi_a(\bx),\xi_b(\bx))$. %In practice there is of course time variability, but we assume this to be negligible during the AUV operation. Depth variability could also be important in some cases, but in the current application the with depths only a few meters this is considered irrelevant for mapping the phenomenon.

In the context illustrated by Figure \ref{fig:river4}, interest is on spatial separation of cold freshwater. 
Because a main goal is classifying the water masses, ES variability is a useful sampling criterion for guiding the observation effort of the AUV. 
A bivariate ES is here defined via joint variables being below thresholds, at individual locations, for the process over the spatial domain $\mathcal{M}$.
Denoting the threshold for temperature by $T_a$ and the threshold for salinity by $T_b$,
we use the following definition;
 \begin{equation}\label{ES}
     \mbox{ES} = \{\bx \in \mathcal{M} | \xi_a(\bx) < T_a,\xi_b(\bx) < T_b\}.
 \end{equation}
The excursions could also be defined as larger than threshold, or within a boundary of limits for the two variables. Calculations and interpretations will be similar in these different cases.

The ES is defined by random indicator variables, and the associated EPs become 
\begin{equation}\label{eq:prob}
 p(\bx) = P(\xi_a(\bx) < T_a, \xi_b(\bx) < T_b), \hspace{3mm} \bx \in \mathcal{M}.
\end{equation}
The actual calculation of the EP would require a model specification. As we demonstrate below, Gaussian modeling assumptions facilitate the computation. 
If the EPs in Eq. (\ref{eq:prob}) are close to $1$ or $0$, it is easy to classify water masses. It is more difficult if the probabilities are close to $0.5$. In this case it will likely be worthwhile gathering data to improve the classification skill.

There has lately been much statistical literature on ESs and EPs connected to spatial statistics \citep{french2013spatio,bolin2015excursion,french2016credible}. 
Most of the work done is based on GPs for the spatial phenomenon, as we use here.
Our focus in the current paper is on ES as defined in Eq. (\ref{ES}) and location-wise EPs and the uncertainty reduction that can be achieved by informative sampling. 
In this sense our work is similar to \cite{chevalier2014fast} and \cite{azzimonti2016quantifying} who described analytical results for uncertainty reduction in ESs for univariate processes.
One contribution of the current paper is to derive closed-form results for expectations of ESs in the situation with multivariate processes.

\section{Gaussian processes and excursion probabilities}\label{sec:GP_EP}

\subsection{Bivariate Gaussian processes}

We represent the bivariate random field of ocean temperature and salinity by a Gaussian model.
At location $\bx \in \mathcal{M}$, the bivariate distribution is 
\begin{equation}\label{gp_bar}
\begin{bmatrix}\xi_a(\bx) \\
\xi_b(\bx) 
\end{bmatrix}
 \sim N \left( 
\begin{bmatrix} \mu_{a}(\bx)\\
\mu_{b}(\bx)
\end{bmatrix},\begin{bmatrix}
\sigma_{a}^2(\bx) & \sigma_{a}(\bx) \sigma_b(\bx) \gamma  \\
\sigma_{a}(\bx) \sigma_b(\bx) \gamma  & \sigma_{b}^2(\bx) 
\end{bmatrix}
\right).
\end{equation}
Here, the means and standard deviations are allowed to vary with location, which can be important for capturing the trends and variabilities near a river mouth. The correlation in salinity and temperature is assumed to be the same for all locations.

There exist several multivariate spatial covariance functions \citep{gneiting2010matern,genton2015cross}. But in practice it has shown difficult to estimate all parameters from data. Like is often done, we assume the same correlation decay for salinity and temperature, which is not unrealistic for a river plume. We use a separable covariance function such that
\begin{eqnarray}\label{gp_corr}
\mbox{Cov}(\xi_a(\bx),\xi_b(\bx')) &=& \rho(\bx,\bx') \gamma \sigma_a(\bx) \sigma_b(\bx'), \\
\mbox{Cov}(\xi_a(\bx),\xi_a(\bx')) &=& \rho(\bx,\bx') \sigma_a(\bx) \sigma_a(\bx'), \nonumber \\
\mbox{Cov}(\xi_b(\bx),\xi_b(\bx')) &=& \rho(\bx,\bx') \sigma_b(\bx) \sigma_b(\bx'). \nonumber
\end{eqnarray}
Moreover, we assume a stationary isotropic correlation function $\rho(\bx,\bx')=\rho(h)$, for distance $h=\sqrt{(\bx-\bx')^t (\bx-\bx')}$.

For implementation on the computer, we discretize the spatial domain $\mathcal{M}$ on a set of $n$ grid locations; $\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$.
We let $\Delta$ be the area of each cell in the regular grid.
Using vector notation, we collect the temperature and salinity variables on the grid as
\begin{equation}
    \bxi=(\xi_a(\bx_1),\xi_b(\bx_1),\ldots,\xi_a(\bx_n),\xi_b(\bx_n))^t \sim \mbox{GP} (\bmu,\bSigma), \nonumber
\end{equation}
where the entries in length $2 n$ vector $\bmu$ hold the expected values of temperature and salinity at the grid locations, and the $2n \times 2n$ matrix $\bSigma$ holds all the covariances between temperature and salinity variables at all grid locations. 
 The joint probability density function is then
\begin{equation}\label{prior}
\pi(\bxi) = N(\bxi;\bmu,\bSigma) = \frac{1}{(2\pi)^{\frac{n}{2}}|\bSigma|^{\frac{1}{2}}}e^{-\frac{1}{2}(\bxi -\bmu)^T\bSigma^{-1}(\bxi - \bmu)}.
\end{equation}

\subsection{Updating Gaussian processes}

The GP mean and covariance in Eq. (\ref{prior}) are updated when more information gets available. 
In our setting, the autonomous vehicle will collect a batch of temperature and salinity measurements along a survey path. After each batch, the data are assimilated to get an updated model.
We here describe one round of updating. In Section \ref{sec:heuristics} we discuss sequential data gathering and model updating.

Denote the AUV data at $m$ locations by $\by=(y_{a,1},y_{b,1},\ldots,y_{a,m},y_{b,m})^t$, representing the temperature and salinity measurements gathered along a survey line.
Restricting attention to the grid representation, the likelihood model is described by 
\begin{equation}\label{likelihood}
\by | \bxi \sim GP(\by; \bA \bxi,\bR),
\end{equation}
where $\bA$ is an $2m \times 2n$ matrix holding $1$ entries at the measurement indexes on the grid, and otherwise $0$ entries. The size $2m \times 2m$ covariance matrix $\bR$ is diagonal with entries $r_a$ and $r_b$ which are the measurement error variances of temperature and salinity observations.

The updated distribution for temperature and salinity variables on the spatial grid, given survey data $\by$, is Gaussian with mean and covariance
\begin{eqnarray}\label{gp_upd}
\bm &=& \bmu+\bSigma \bA^t (\bA \Sigma \bA^t+\bR)^{-1}(\by-\bA \bmu),  \\
\bS &=& \bSigma - \bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.\nonumber
\end{eqnarray}
The updated bivariate distribution at a grid location $\bx$ is then 
\begin{equation}\label{gp_hat}
\begin{bmatrix}
\xi_a(\bx) \\
\xi_b(\bx)
\end{bmatrix}
 |\by
 \sim N \left( 
\begin{bmatrix} m_{a}(\bx)\\
m_{b}(\bx)
\end{bmatrix},\begin{bmatrix}
s_{a}^2(\bx) & s_{a,b}(\bx)  \\
s_{a,b}(\bx)  & s_{b}^2(\bx)  
\end{bmatrix}
\right),
\end{equation}
with entries extracted from the conditional mean and covariance expressions in Eq. (\ref{gp_upd}).

The conditional mean $\bm$ will be important in the following derivation in Section \ref{sec:sur}. This mean is a linear function of the data $\by$, and since the marginal distribution of the data is $\by \sim N( \bA \bmu,\bA \bSigma \bA^t+\bR)$, we get
\begin{equation}\label{distxi}
\bm \sim N(\bmu , \bPsi), \hspace{3mm} \bPsi=\bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.  
\end{equation}
When considering only grid location $\bx$, we can extract elements from the mean vector and covariance matrix to get the bivariate distribution as follows 
\begin{equation}\label{dist_mxi}
\bm(\bx) \sim N 
\left( 
\begin{bmatrix}
\mu_{a}(\bx) \\
\mu_{b}(\bx)
\end{bmatrix},
\begin{bmatrix}
\psi^2_{a}(\bx) & \psi_{a,b}(\bx)\\
\psi_{a,b}(\bx) & \psi^2_{b}(\bx)
\end{bmatrix}
\right).
\end{equation}

\subsection{Excursion probabilities for Gaussian processes}

Based on the Gaussian modeling assumptions, the EPs in Eq. (\ref{eq:prob}) can be computed at any location $\bx$. Without additional data, we have
\begin{eqnarray}\label{eq:prob_mv0}
 p(\bx) &=& P(\xi_a(\bx) < T_a, \xi_b(\bx) < T_b) \nonumber \\
 &=& \Phi_2 \begin{pmatrix} 
\begin{bmatrix} T_a\\
T_b
\end{bmatrix};
\begin{bmatrix} \mu_{a}(\bx)\\
\mu_{b}(\bx)
\end{bmatrix},\begin{bmatrix}
\sigma_{a}^2(\bx) & \sigma_{a}(\bx)\sigma_{b}(\bx) \gamma  \\
\sigma_{a}(\bx)\sigma_{b}(\bx) \gamma  & \sigma_{b}^2(\bx)  
\end{bmatrix}\end{pmatrix},
\end{eqnarray}
where $\Phi_2$ is the bivariate Gaussian cumulative distribution function, and the mean and covariance entries are defined in Eq. (\ref{gp_bar}).

When data $\by$ is available, we similarly get
\begin{eqnarray}\label{eq:prob_mv}
 p(\bx;\by) &=& P(\xi_a(\bx) < T_a, \xi_b(\bx) < T_b |\by) 
 \nonumber \\
 &=& \Phi_2 \begin{pmatrix} 
\begin{bmatrix} T_a\\
T_b
\end{bmatrix};
\begin{bmatrix} m_{a}(\bx)\\
m_{b}(\bx)
\end{bmatrix},\begin{bmatrix}
s_{a}^2(\bx) & s_{a,b}(\bx)  \\
s_{a,b}(\bx)  & s_{b}^2(\bx)  
\end{bmatrix}\end{pmatrix},
\end{eqnarray}
with parameters defined in Eq. (\ref{gp_upd}) and (\ref{gp_hat}).

For one particular dataset, the probability $p(\bx;\by)$ can be closer to $0.5$ than the initial probability $p(\bx)$. This means that the variance in the ES is increased. On average, taking expectations over the data $\by$, the ES variance will be reduced by gathering data. We next explain how AUVs can be used to efficiently gather data and reduce the overall marginal spatial uncertainties in the ES.

\section{Uncertainty reduction on excursion sets}\label{sec:sur}

\subsection{Variance in excursion set}

Regions with temperatures or salinity that surpasses the thresholds with a large margin (e.g. very high/low temperatures) are not interesting, unless the uncertainty at these locations is very large. The objective will therefore be to explore regions where the process variables are close to the thresholds. With this mind, it is valuable to explore locations $\bx$ where $p(\bx) \approx 0.5$. This is what gives the maximum variance $p(\bx) (1-p(\bx))$ in the indicator for the ES.

Consider then a measure as follows:
\begin{equation}\label{measVAR}
    V = \int_{\bx \in \mathcal{M}} p(\bx) \left(1-p(\bx)\right) d\bx \approx \sum_{x \in \mathcal{M}_g} p(\bx) \left(1-p(\bx)\right) \Delta.
\end{equation}
The measure is dominated by location having probabilities close to $0.5$.

Considering one location only, we study the effect of different correlation $\gamma$ and prior variance. 
Figure \ref{illus_bivarDens} shows contour plots of three different densities. Here, the temperature mean is $T_a=5^o$ and salinity mean is $T_b=30$ mg/l. The thresholds are identical to these means. The variables have unit variance and different correlation (low, medium and high).
\begin{figure}[h!]
\centering
\includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
\caption{Density contour plots having different correlations for temperature and salinity.}\label{illus_bivarDens}
\end{figure}
Table \ref{tab:sim_rhoab} shows the EPs and the associated variances in the ES (top rows). The results are computed for different correlations $\gamma$ indicated in Figure \ref{illus_bivarDens}, and for standard deviations $1$ (left) and $2$ (right). The variance in the ES is largest for high correlation, as there is more tendency of joint small or high occurrences for temperature and salinity in this case. EPs are the same for standard deviation $1$ or $2$. 
\begin{table}[!h]
\centering
\caption{EP and variance of ES for different correlations (top rows), and for data of both variables and only for temperature (bottom rows).}
\begin{tabular}{c|ccc|ccc}
 &\multicolumn{3}{c}{$\sigma_b=\sigma_a=1$} & \multicolumn{3}{c}{$\sigma_b=\sigma_a=2$} \\
\hline
Correlation $\gamma$ & 0.2 & 0.6 & 0.8 & 0.2 & 0.6 & 0.8 \\
\hline
$p$ & 0.28 & 0.35 & 0.40 & 0.28 & 0.35 & 0.40 \\ 
$p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
$E_{(y_a,y_b)}(p (1-p))$ & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
$E_{y_a}(p (1-p))$ & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
\hline
\end{tabular}
\label{tab:sim_rhoab}
\end{table}

\subsection{Expected variance criteria}

When designing AUV sampling paths, we aim to reduce the variances in the ES and in this way efficiently characterize the ES throughout the spatial domain. We saw that the measure defined by Eq. (\ref{measVAR}) does not discriminate between high or low variability in the temperature and salinity (two top rows of Table \ref{tab:sim_rhoab}).
Also, focusing only on locations with EPs near $0.5$ would not account for the spatial correlation in the model.  
The same has been seen for related measures learning graphical models \citep{lilleborge2016information}.

To evaluate AUV data sampling paths, it is hence natural to evaluate the expected variance reduction in the ES  \citep{chevalier2014fast}, as provided by the survey data. In this way, we take the expectation with respect to the planned measurements $\by$:
\begin{eqnarray}\label{sur}
    V_{\mbox{upd}} &=& \int_{\bx \in \mathcal{M}} E_{\by} \left\{ p(\bx;\by)\left( 1-p(\bx;\by)\right) \right\} d\bx \\
    & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by} \left\{ p(\bx;\by)\left( 1-p(\bx;\by)\right) \right\} \Delta, \nonumber \\
    p(\bx;\by) &=& P(\xi_a(\bx) < T_a, \xi_b(\bx) < T_b |\by), \nonumber
\end{eqnarray}
where the conditional probability is defined in Eq. (\ref{eq:prob_mv}).

We will next derive closed form solutions to the expectation in (\ref{sur}). A critical
element in the derivation is the GP model, and the linear dependence on data $\by$, which means that we have a closed form Gaussian distribution for the conditional mean in Eq. (\ref{dist_mxi}). This means that the high-dimensional inner integral in Eq. (\ref{sur}) reduces to a univariate integral \citep{bhattacharjya2013value, chevalier2014fast}.

Table \ref{tab:sim_rhoab} (bottom two rows) shows results of doing such calculations, for different correlation $\gamma$ and standard deviations (left and right). Here, the expected variances are computed for both data types $(y_a,y_b)$, and for temperature measurements $y_a$ alone. With both data; $(y_a,y_b)^t=(\xi_a,\xi_b)^t+N(0,0.5^2\bI)$, while $y_a=\xi_a+N(0,0.5^2)$ when only temperature is measured.
The results in Table \ref{tab:sim_rhoab} shows that 
the expected variance depends on the (prior) uncertainty in temperature and salinity. The expected variance in the ES is lower with larger standard deviations $\sigma_a$ and $\sigma_b$ (right columns). Moreover, the relative reduction is largest for the case with high correlation $\gamma$. There is also large uncertainty reduction when only temperature is measured (bottom row), especially when the temperature and salinity are very dependent. When correlation is low ($\gamma=0.2$) there is hardly any information about salinity in the temperature data. In the application with fresh cold water from the river, the temperature and salinity variables will not only be interdependent, but they will be dependent in the spatial dimension, so this kind of location-wise dependence will also impact design criteria when we evaluate the information measure by integrating over several locations $\bx$.

\subsection{Analytical results for bivariate ES}

In the rest of this section, the location index $\bx$ is suppressed to avoid overly complex notation. We present new results for computing the expectation $E_{\by}\left\{ \cdot \right\}$ which is the integrand of Eq. (\ref{sur}).

The first key result is that the expectation only depends on the data through the linear combination $\bm=(m_{a},m_{b})^t$. 
To see this we standardize the two variables i.e.
$Z_a=\frac{\xi_a-m_{a}}{s_{a}}$,
$Z_b=\frac{\xi_b-m_{b}}{s_{b}}$; 
then
\begin{eqnarray}
   P(\xi_a < T_a, \xi_b < T_b|\by) &=& P \left( Z_a <\frac{T_a-m_{a}}{s_{a}}, Z_b <\frac{T_b-m_{b}}{s_{b}}|\by \right) \nonumber \\
   &=& \Phi_2 \begin{pmatrix} 
\begin{bmatrix} \frac{T_a-m_{a}}{s_{a}}\\
\frac{T_b-m_{b}}{s_{b}}
\end{bmatrix};
 \begin{bmatrix} 0\\
0
\end{bmatrix},\begin{bmatrix}
1 & \eta_{s,t}  \\
\eta_{s,t}   & 1  
\end{bmatrix}\end{pmatrix} \nonumber \\
&=& P(\xi_a < T_a, \xi_b < T_b|\bm),
\end{eqnarray}
where $\eta_{a,b} =s_{a,b}/(s_{a} s_{b})$ is defined in Eq. (\ref{eq:prob_mv}).
Since the expression only depends on $\by$ via $\bm$, the expectation in Eq. (\ref{sur}) is reduced to a bivariate integral over $\bm$ (which will vary for each integrand term, because the locations differ).

The expected variance can now be written as
\begin{eqnarray}\label{eq:var2}
E_{\by}(p[1-p]) &=& E_{\bm}(p[1-p]), \hspace{3mm} p=P(\xi_a < T_a, \xi_b < T_b|\bm) \nonumber \\
E_{\bm}(p[1-p]) & = & \int p[1-p] p(\bm) d\bm, \nonumber \\
 &=& \int P(\xi_a < T_a, \xi_b < T_b|\bm)  p(\bm) d\bm \nonumber  \\
&-& \int P(\xi_a < T_a, \xi_b < T_b|\bm) P(\xi_a < T_a, \xi_b < T_b|\bm) p(\bm) d\bm, 
\end{eqnarray}
where the integral is split in two components.
We will next extend results of \cite{chevalier2014fast} to solve for these two parts.

Considering part I of expression (\ref{eq:var2}), we have
\begin{equation}\label{part1:phi2}
 \int P(\xi_a < T_a, \xi_b < T_b|\bm) p(\bm) d\bm= 
P \left( Z_{a} < \frac{T_a-m_{a}}{s_{a}}, 
Z_{b} < \frac{T_b-m_{b}}{s_{b}} \right), \nonumber
\end{equation}
where $\bZ=(Z_{a},Z_{b})$ is a bivariate zero-mean and unit-variance Gaussian vector, with $\mbox{Corr}(Z_{a},Z_{b})=\eta_{a,b}$. Moreover, $\bZ$ is chosen to be independent of $\bm$. Using Eq. (\ref{dist_mxi}) we have
\begin{equation}
    E(s_{a} Z_{a}+m_{a}-T_a) = \mu_{a}-T_a, \hspace{3mm}
    \mbox{Var}(s_{a} Z_{a}+m_{a}-T_a) = s^2_a+ \psi^2_{a}, 
\end{equation}
and the same holds for the salinity part with subscript $b$. From this calculation, we get
\begin{eqnarray}\label{two_parts0}
& P & \left( Z_{a} < \frac{T_a-m_{a}}{s_{a}}, 
Z_{b} < \frac{T_b-m_{b}}{s_{b}} \right) \\
&=& \Phi_2 \begin{pmatrix} 
\begin{bmatrix} 0\\
0
\end{bmatrix};
\begin{bmatrix} \mu_{a}-T_a\\
\mu_{b}-T_b
\end{bmatrix},\begin{bmatrix}
s^2_a+ \psi^2_{a} & s_{a,b}+\psi_{a,b}  \\
s_{a,b}+\psi_{a,b}   & s^2_a+ \psi^2_{a} 
\end{bmatrix}\end{pmatrix} \nonumber.
\end{eqnarray}

For part II of expression (\ref{eq:var2}), standardization means that
\begin{eqnarray}\label{part1:phi4}
&& \int P(\xi_a < T_a, \xi_b < T_b|\bm) P(\xi_a < T_a, \xi_b < T_b|\bm) p(\bm) d\bm_{\xi}= \nonumber \\
&& P\left( Z_{1,a} < \frac{T_a-m_{a}}{s_{a}}, 
Z_{1,b} < \frac{T_b-m_{b}}{s_{b}},Z_{2,b} < \frac{T_a-m_{a}}{s_{a}}, 
Z_{2,b} < \frac{T_b-m_{b}}{s_{b}} \right). \nonumber
\end{eqnarray}
Here,
$\bZ_1=(Z_{1,a},Z_{1,b})$, $\bZ_2=(Z_{2,a},Z_{2,b})$ are independent bivariate zero-mean and unit-variance Gaussian vectors, both with element-wise correlation
$\eta_{a,b}$. Moreover, $\bZ_1$ and $\bZ_2$ are chosen to be independent of $\bm$. Hence,
\begin{equation}
    \left(
    \begin{array}{cccccc}
         Z_{1,a} \\
         Z_{1,b} \\
         Z_{2,a} \\
         Z_{2,b} \\
         m_{a} \\
         m_{b}
    \end{array}
    \right)
\sim N \left(
\left(
    \begin{array}{cccccc}
    0 \\
    0 \\
    0 \\
    0 \\
          \mu_{a} \\
         \mu_{b} 
    \end{array}
    \right),
   \left[ 
      \begin{array}{cccccc}
        1 & \eta_{a,b} & 0 & 0 & 0 & 0  \\
        \eta_{a,b} & 1 & 0 & 0 & 0 & 0 \\
        0 & 0 & 1 & \eta_{a,b} & 0 & 0 \\
        0 & 0 & \eta_{a,b} & 1 & 0 & 0 \\
        0 & 0 & 0 & 0 & \psi^2_{a} & \psi_{a,b} \\
        0 & 0 & 0 & 0 & \psi_{a,b} & \psi^2_{b} 
    \end{array}
\right]
\right).
\vspace{0.2cm}
\end{equation}

The solution to part II is then to evaluate a four-variable cumulative distribution function, accounting for the mean, variance and covariances of $\bZ_1$, $\bZ_2$ and $\bm$ in the appropriate linear combinations. Using vector-matrix formulations, we define linear combinations
\begin{equation}\label{Vcomb}
    \bV=\left[
    \begin{array}{cccccc}
        s_a & 0 & 0 & 0 & 1 & 0 \\
         0 & s_b & 0 & 0 & 0 & 1 \\
         0 & 0 & s_a & 0 & 1 & 0\\
         0 & 0 & 0 & s_b & 0 & 1\\
    \end{array}
    \right] 
    \left(
    \begin{array}{cccccc}
         Z_{1,a} \\
         Z_{1,b} \\
         Z_{2,a} \\
         Z_{2,b} \\
         m_{a} \\
         m_{b}
    \end{array}
    \right),
\end{equation}
and the part II calculation becomes
\begin{eqnarray}\label{part1:phi4}
&& P \left( Z_{1,a} < \frac{T_a-m_{a}}{s_{a}}, 
Z_{1,b} < \frac{T_b-m_{b}}{s_{b}},Z_{2,b} < \frac{T_a-m_{a}}{s_{a}}, 
Z_{2,b} < \frac{T_b-m_{b}}{s_{b}} \right)  \\
&=&  \Phi_4 
\begin{pmatrix} 
\begin{bmatrix} T_a\\
T_b \\
T_a \\
T_b
\end{bmatrix};
E( \bV),Var(\bV)
\end{pmatrix} \nonumber.
\end{eqnarray}

The multivariate cumulative probabilities in two and four dimensions are fast to compute using methods of \cite{genz2009computation}.
In the end, the part I solution in Eq. (\ref{two_parts0}) and part II solution in Eq. (\ref{part1:phi4}) are computed for each location on the grid, and summing over the spatial domain $x \in \mathcal{M}_g$, gives the design evaluation criterion in Eq. (\ref{sur}).

\subsection{Generalization to the multivariate case}

In our application with temperature and salinity, the solutions to the complex integral equations for the expected variance are reduced to calculating bi-variate and four-variate cumulative distribution functions for Gaussian vectors. For the more general situation one would be interested in $K$ random fields. Say, in the ocean sciences the variables could include chlorophyll or plankton variables, in addition to temperature and salinity.

The equations derived in the previous section can be generalized to such $K$-variable cases. We then define $\bxi(\bx)=(\xi_1(\bx),\ldots,\xi_K(\bx))$, and the expected variance in the ES is
\begin{equation}\label{two_partsK}
E_y(p[1-p]) = \int p[1-p] p(\by) d\by, \hspace{3mm} P(\xi_k < T_k ; k=1,\ldots,K|\by). 
\end{equation}
By the first key result of linear combination of Gaussian variables, we reduce the integral to a $K$ dimensional integral of the relevant $\bm=E(\bxi(\bx)|\by)$. Next, reducing the integral to two parts, we standard the variables in each and compute cumulative distributions. 
In summary, we get 
\begin{equation}\label{two_parts}
E_y(p[1-p]) =  \Phi_K 
\begin{pmatrix}
\begin{bmatrix} T_1\\
\vdots \\
T_K 
\end{bmatrix};
\bmu,\bS+\bPsi 
\end{pmatrix}
- \Phi_{2K} 
\begin{pmatrix}
\begin{bmatrix} T_1\\
\vdots \\
T_K \\
T_1\\
\vdots \\
T_K 
\end{bmatrix};
E(\bV),Var(\bV) 
\end{pmatrix},\nonumber
\end{equation}
where $\bS$ and $\bPsi$ are direct $K$-variate extensions of the matrices entries in Eq. (\ref{two_parts0}), and the matrix $\bV$ finds the relevant linear combination of two independent vectors $\bZ_l=(Z_{l,1},\ldots,Z_{l,K})$, $l=1,2$ and $\bm$, as an extension of Eq. (\ref{Vcomb}).

%\subsection{Expected classification criteria}

%{\bf{I have not done anything here - skip this, I think}}.

%Rather than minimizing expected variance one can aim at minimizing the classification probability of the excursion set: 
%$\min [p,(1-p) ]$, see e.g. \cite{lilleborge2016information}.

%Again, since the conditional mean is a linear in the data, we only have to look at the %relevant linear combination via $\bm_{\xi}=E(\bxi)$.
%See also \cite{bhattacharjya2013value}.

%\begin{equation}
%E(\min \{ p,(1-p)\})=\int \min P(\xi_a < t, \xi_b < s),[1-P(\xi_a < t, \xi_b < s)] p(E(\bxi)=) dE(\bxi)=,
%\end{equation}
%for the complementary probabilities, this is again evaluated by the corner regions for the block, and $\Phi_4()$ evaluations are required.

%In the end, this result is integrated over the spatial domain $x \in X$.

\section{Sequential updating and heuristic path planning}\label{sec:heuristics}

\subsection{Optimal sequential design}
\label{myopic}

The derived closed form solutions for the expected variance (in the multivariate ES) can now be used to evaluate and select among different survey designs for sampling applications with robotic platforms. 
Considering a set $\mathcal{J}$ of possible designs for a survey. A criterion for the AUV path selection is
\begin{equation}\label{crit}
    j^* = \mbox{argmin}_{j \in \mathcal{J}} V_{\mbox{upd}}(j),
\end{equation}
where the criterion $V_{\mbox{upd}}(j)$ in Eq. (\ref{sur}) is computed for each of the possible designs.
In practice, a survey is split in many stages, giving opportunities for sequential adaption of the AUV path. This means that the optimal survey design is found by considering not only the best current AUV line, but also what data gathered along this line would lead to for future sampling at the next stages.

Let $d^{j,n}$ denote design number $j$ at stage $s$ of sequential surveying. If this design is selected, data $\by^{j,s}$ will be gathered. The optimal path selection situation is then depicted in Figure \ref{fig:PathSelOpt}, 
\begin{figure}[h!]
\centering
\includegraphics[width=0.85\textwidth]{Figures/sequent_select.pdf}
\caption{Optimal sequential path selection.}\label{fig:PathSelOpt}
\end{figure}
where design choices are indicated by squares while data realizations are indicated by circles. 
In mathematical terms this optimal design is defined as
\begin{equation}\label{opt_crit}
    \bd^* = \mbox{argmin}_{d^{j,1}} \left\{ \int_{\by^{j,1}} \mbox{argmin}_{d^{j,2}} \left\{ \int_{\by^{j,2}} \ldots p(\by^{j,2}|\by^{j,1}) d\by^{j,2} \right\} p(\by^{j,1}) d \by^{j,1} \right\},
\end{equation}
where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages.

In practice the optimal solution is not tractable because of the enormous growth in maximization and integral terms over stages, see e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}. Instead, we next outline two heuristic strategies. In terms of notation, let $\mathcal{Y}^{s-1} =\{d^{j,r}, \by^{j,r} ; r=1,\ldots,s-1 \}$ denote the data gathered until stage $s-1$, using a design indexed $j$ (which is suppressed in the data set $\mathcal{Y}^{s-1}$.

\subsection{Naive path planning}
\label{naive}

The simplest heuristic for adaptive sampling is to choose the next sampling location based on current EPs. The variance is largest for EP equal to $0.5$, so one selects the graph location where $\bx$ with average probabilities $p(\bx)$ closest to $0.5$.

At stage $s$, based on the currently available data $\mathcal{Y}^{s-1}$, we fit an updated Gaussian model from Eq. (\ref{gp_upd}), with mean $\bm_{s-1}$ and covariance matrix $\bS_{s-1}$, represented on the grid locations covering the spatial domain. 
The next batch of measurements $\by^{j,s}$, can be  gathered with design $j=1,\ldots,J$, indicating all possible AUV lines at the current stage. The selected design is then
\begin{equation}\label{critNaive}
    d^{*,n} = \mbox{argmin}_{j \in \mathcal{J}} \int_{\bx \in \mathcal{J}} |p(\bx;\mathcal{Y}^{s-1})-0.5| d\bx. 
\end{equation}
This strategy does not account for the uncertainty in the ES, if one design has EPs closer to $0.5$, this is chosen, even though the variance along that line is already very small compared with other locations. Hence, this strategy lacks memory of where it has been, where the uncertainty has been reduced. For this reason it can easily get stuck in local regions.

\subsection{Myopic path planning}
\label{myopic}

The myopic (greedy) strategy which we present here is optimal if we imagine taking only one more survey line of measurements. In this selection strategy there is no anticipation of what the subsequent survey lines might offer.

Based on the currently available data $\mathcal{Y}^{s-1}$, we again fit an updated Gaussian model, represented on the grid locations covering the spatial domain. 
The next batch of measurements $\by^{j,s}$, can be  gathered with design $j=1,\ldots,J$, indicating all possible AUV lines at the current stage. The selected design is then
\begin{eqnarray}\label{critSEQ}
    d^{*,s} &=& \mbox{argmin}_{j \in \mathcal{J}} \left\{ V_{m,\mbox{upd,j}} \right\},  \\
V_{m,\mbox{upd}} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j,s}|\mathcal{Y}^{s-1}} \left\{ p(\bx;\mathcal{Y}^{j,s})\left( 1-p(\bx;\mathcal{Y}^{j,s})\right) \right\} \Delta \nonumber \\
    p(\bx;\mathcal{Y}^{j,s}) &=& P(\xi_a(\bx) < T_a, \xi_b(\bx) < T_b |\by^{j,s},\mathcal{Y}^{s-1}). \nonumber
\end{eqnarray}

Note that this strategy gives a sequential conditional version of the formula in Eq. (\ref{sur}). Now $\mathcal{Y}^{s-1}$ is available, and the expectation is with respect to the conditional distribution $p(\by^{j,s}|\mathcal{Y}^{s-1})$. The same closed form calculation is hence applicable in Eq. (\ref{critSEQ}), using the updated GP model from step $s-1$. 
Once the data are collected along this best line, the GP model is updated again. The mean $\bm_{s}$ and covariance matrix $\bS_{s}$ are used to compute the next survey line at stage $s+1$, using the same strategy, and so on.

This myopic strategy is non-anticipative and hence sub-optimal, but it still gives a reasonable approach for creating designs in many applications. Moreover, it is easily implemented on-board, requiring only an efficient approach for data updating of the GP model and the calculation of the closed form variance expressions for each next survey line.

\subsection{Look-ahead path planning}
\label{LA}

We now extend the myopic strategy to a look-ahead strategy which is optimal when one can gather only two more survey lines of measurements. In its selection this strategy for the current survey line, it anticipates what the next survey line might offer, but not anything beyond this line.

In addition to the next batch of measurements $\by^{j,s}$, one now considers the subsequent design $j_2$ with data $\by^{j_2,s+1}$ when choosing the next survey line.

The selected design is then
\begin{eqnarray}\label{critLA}
    d^{*,n} &=& \mbox{argmin}_{j \in \mathcal{J}} \left\{ U_{la,\mbox{upd,j}} \right\},  \\
    U_{la,\mbox{upd},j} & = &  E_{\by^{j,s}|\mathcal{Y}^{s-1}} \left\{ \mbox{argmin}_{j_2 \in \mathcal{J}_2} \left[ V_{la,\mbox{upd},j_2} \right] \right\} \nonumber \\
V_{la,\mbox{upd},j_2} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j_2,s+1}|\mathcal{Y}^{j,s}} \left\{ p(\bx;\mathcal{Y}^{j,s+1})\left( 1-p(\bx;\mathcal{Y}^{j,s+1})\right) \right\} \Delta \nonumber \\
    p(\bx;\mathcal{Y}^{j,s+1}) &=& P(\xi_a(\bx) < T_a, \xi_b(\bx) < T_b |\by^{j_2,s+1},\mathcal{Y}^{j,s}). \nonumber
\end{eqnarray}
Here, $\mathcal{Y}^{j,s}=\{\by^{j,s},\mathcal{Y}^{s-1}\}$ and $\mathcal{Y}^{j,s+1}=\{\by^{j_2,s+1},\mathcal{Y}^{j,s}\}$ represent the set of data variables, and the expectations are with respect to the conditional distributions $p(\by^{j,s}|\mathcal{Y}^{s-1})$ and $p(\by^{j_2,s+1}|\by^{j,s},\mathcal{Y}^{s-1})$.

Because of the intermixed optimization and expectation, there is no longer a closed form for the required integrals. Instead, we solve the first expectation by Monte Carlo sampling of data $\by^{j,s}$ from its conditional distribution. For each data sample, the second expectation is solved using the closed form expressions from Section \ref{sec:sur}.

Even though the strategy looks at two stages, it is only used to find the current best survey line. When data are collected along this best line, the GP model is updated, and the mean $\bm_{s}$ and covariance matrix $\bS_{s}$ are used to compute the next survey line at stage $s+1$, now anticipating what stage $s+2$ could give after this survey, and so on.

This look-ahead approach is much more computer-demanding than the myopic strategy, and for practical implementation we prune paths in the evaluation of Eq. (\ref{critLA}). This means that we do not compute all possible branches of the first two stages, as they are indicated in Figure \ref{fig:PathSelOpt}. Instead, we use the myopic strategy to rank the three best paths on the first stage alone, and for each of these we go through with the look-ahead calculation.

\section{Simulation study}\label{sec:simulations}
We here show results from tests performed to study the properties of different static and sequential survey designs in a realistic simulated case, mapping a freshwater plume defined by a temperature and salinity field.

\subsection{Modeling}

We use a bivariate GP model for temperature and salinity, with prior mean specified by
\begin{equation}\label{m}
    \bmu(\bx)=E 
    \begin{bmatrix}
    \xi_a(\bx) \\
    \xi_b(\bx) 
    \end{bmatrix}=\begin{bmatrix} \mu_{a}(\bx)\\
\mu_{b}(\bx)
\end{bmatrix} 
= \begin{bmatrix} \beta_{a,0} + \beta_{a,1} x_{\mbox{East}} \\
\beta_{a,1} + \beta_{b,1} x_{\mbox{East}}
\end{bmatrix}.
\end{equation}
We expect the eastern part of the domain to be cold freshwater. The situation mimics that of a river mouth entering from the south, and the water masses are pulled to the east. This is imposed by setting slopes to $\beta_{a,1}=0.065$ and $\beta_{b,1}=0.1$. Further, the intercepts of the temperature and salinity variables are set to $\beta_{a,0}=5.8$ and $\beta_{b,1}=29.0$.

%The prior assumption used by the agent are $\tilde{\beta_{a,1}}=0.085$ and $\tilde{\beta_{b,1}}=0.138$, which assumes that the process is centered.

The covariance is assumed to be stationary and separable for the two variables. 
The $2 \times 2$ pointwise covariance matrix is set to
\begin{equation}\label{v0}
\bSigma(\bx)=\mbox{Var} 
\begin{bmatrix}
    \xi_a(\bx) \\
    \xi_b(\bx) 
    \end{bmatrix}=
\begin{bmatrix}
0.25^2 & 0.6 \cdot 0.25^2 \\
0.6 \cdot 0.25^2 & 0.25^2
\end{bmatrix}.
\end{equation}
The spatial correlation is of a Matern type giving
\begin{equation}\label{v}
\mbox{Corr} 
\left(
\begin{bmatrix}
    \xi_a(\bx) \\
    \xi_b(\bx) 
    \end{bmatrix},
    \begin{bmatrix}
    \xi_a(\bx') \\
    \xi_b(\bx') 
    \end{bmatrix}
    \right)
    = \begin{bmatrix}
1 & 0.6  \\
0.6  & 1
\end{bmatrix}(1+\phi |\bx-\bx'|)\exp (-\phi |\bx-\bx'|),
\end{equation}
and $\phi=0.3$ indicates an effective correlation range of about $1200$ m. 
%Figure \ref{fig:stat_design} shows the contour lines for the EP for the reference. We notice the trend of increasing salinity and temperature to the west.  
We study sensitivity to the parameter specification in the results below.
One realization of the random fields representing salt and temperature is shown in Figure \ref{fig:true_temp} and \ref{fig:true_sal}.
\begin{figure}[!h]
\centering
\subfloat[Simulated temperature.]{\includegraphics[height = 0.41\textwidth]{Figures/sim/true_temp.pdf}\label{fig:true_temp}}
\hfill
\subfloat[Simulated salinity.]{\includegraphics[height = 0.41\textwidth]{Figures/sim/true_sal.pdf}\label{fig:true_sal}}
\hfill
\subfloat[Excursion set given the realization of temperature and salinity.]{\includegraphics[height = 0.41\textwidth]{Figures/sim/es_ts_true.pdf}\label{fig:ESet}}
\hfill
\subfloat[Excursion probabilities given collected data along the strategy ``static\_north".]{\includegraphics[height = 0.41\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
\caption{(\ref{fig:true_temp}) and (\ref{fig:true_sal}) shows realisations of the simulated temperature and salinity fields, as well as the associated excursion set (\ref{fig:ESet}). (\ref{fig:Eprob}) show the estimated excursion probabilities after performing data collection along the N-S survey line.}
\label{fig:realisations}
\end{figure}
The true ES from this prior realization is plotted in Figure \ref{fig:ESet}.
%while the EP ({\bf{what information is this based on? Skip EP here? only ES true? }} )

\subsection{Static and sequential sampling designs}

%Three different designs are considered as indicated in Figure \ref{fig:stat_design}. 
%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.65\textwidth]{Figures/Des3.eps}
%\caption{Three different static survey designs plotted on the initial EP.}\label{fig:stat_design}
%\end{figure}
%In this display the designs are plotted along with the prior probability contours of the ES for the reference parameter inputs.

We compare three different static designs denoted \texttt{static\_north}, \texttt{static\_east and} \texttt{static\_zigzag} with the three sequential approaches \texttt{naive}, \texttt{myopic} and \texttt{look-ahead}. The static sampling paths are pre-determined, and cannot adapt to data.

For a fixed length, the expected variance after static surveying is analytically available, but here we focus on the properties of the sequential designs which are evaluated using Monte Carlo sampling over several replicates of realizations from the prior, and conducting simulated sequential surveying for each one. For comparison, the same setup is used for the static designs. Figure \ref{fig:Eprob} shows the conditional EP, given data gathered along the north-south survey line for the realization indicated in these displays. In the Monte Carlo replicates, such results are repeatedly computed to approximate the expected variance reduction in the ES. Along with this criterion we also compare predictive performance measured by root mean square error (RMSE) in the temperature and salinity fields and variance reduction in these two field variables. Note that objective function used by the agents is focused on reducing the expected variance in the ES, but we nevertheless hope to achieve good predictive performance for the other criteria as well. Another non-statistical criteria that is important for practical purposes is the computing time of the strategy.

The agent of the sequential designs starts at the center East-West coordinate at the southern end of the domain (node 53), see Fig. \ref{fig:wp_graph}.

\begin{figure}[h]
\centering
\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
\caption{The equilateral waypoint graph used to discretize the route choices over the $31\times31$ grid.}
\label{fig:wp_graph}
\end{figure}

At each step, the AUV will choose to move along an edge in this waypoint graph. While doing so, it gathers three measurements. These are assimilated in the model, and then a new choice is done at the end of the edge. The paths of will differ for the various sequential designs. The designs will also differ between the runs, because the measurements vary among replicates.

Results of the replicate runs are shown in Figure \ref{fig:sim_results}, where the different criteria are plotted as a function of surveying distance. Here, expected ES variance in Figure \ref{fig:avg_ev} is reduced most when using the \texttt{myopic} and \texttt{look-ahead} strategies. There is not much difference between these two.

\begin{figure}[!h]
\centering
\subfloat[Excursion set variance $E_y(p[1-p])$ .]{\includegraphics[height = 0.49\textwidth]{Figures/sim/avg_EV.pdf}\label{fig:avg_ev}}
\hfill
\subfloat[RMSE between estimated field and truth.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/avg_RMSE.pdf}\label{fig:avg_rmse}}
\hfill
\subfloat[Explained variance $\bR^{2}$.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/avg_R2.pdf}\label{fig:avg_r2}}
\hfill
\subfloat[Time used to do inference.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/avg_Time.pdf}\label{fig:avg_time}}
\caption{Simulation results from 100 replicate simulations for 10 sampling choices/steps on the grid. (Updated:01.05.2019)}
\label{fig:sim_results}
\end{figure}

The north-south design is also doing well, because it moves near the boundary of the water masses. 
In Figure \ref{fig:avg_rmse} and \ref{fig:avg_r2} we show results of RMSE and explained variance. Both \texttt{myopic} and \texttt{look-ahead} still perform well, but some of the static east and zigzag are also achieving good results because they are set to always cover large parts of the domain, and in this way visit locations that are not considered interesting by the sequential strategies targeting ES variance. Regarding computing time in Fig. \ref{fig:avg_time}, the \texttt{naive} strategy is on par with the static designs, while \texttt{myopic} is slower and \texttt{look-ahead} much slower. The \texttt{look-ahead} strategy is reaching levels that are near impractical for on-board execution. Of course, with harder pruning of branches in the graph, it would be faster to run. Then again, this pruning is likely the reason that this strategy does not outperforms the myopic approach.

In Figure \ref{fig:route_choices} we plot the realized sampling paths of the sequential schemes and static designs. 
\begin{figure}[!h]
\centering
\subfloat[Look-ahead strategy.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_look-ahead.pdf}\label{fig:avg_look-ahead}}
\hfill
\subfloat[Myopic strategy.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_myopic.pdf}\label{fig:avg_myopic}}
\hfill
\subfloat[Naive strategy.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_naive.pdf}\label{fig:route_naive}}
\hfill
\subfloat[Static strategies.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_static.pdf}\label{fig:route_static}}
\hfill
\caption{The overlaid route choices, superimposed for the 100 replicate surveys with 10 sampling choices/steps.}
\label{fig:route_choices}
\end{figure}
The naive strategy appears to get stuck in the southern part of the domain in some cases, because it is too focused on the probabilities near $0.5$. The myopic strategy seems to be covering a wider domain than the look-ahead strategy, possibly because the look-ahead strategy sees less value of going far west because it account for having to return east again.

We further go on to study sensitivity of the result to different prior correlation values between temperature and salinity, prior standard deviations and spatial correlations, as well as those of measuring only temperature on-board the AUV (as indicated in the rows). We compare results at the end of the surveying distance. Expected ES variances are shown in Table \ref{{tab:sim_res_ev}}. Results of RMSE are in Table \ref{tab:sim_res_rmse} and explained variance in Table \ref{tab:sim_res_r2}. 
In all runs, the myopic and look-ahead strategies perform the best in terms of expected variance in the ES. Surprisingly, the myopic is on-par and even slightly better than the look-ahead strategy. Part of the reason is pruning of paths for the look-ahead strategy, but even with less pruning, the look-ahead does not outperform the myopic approach. Of course, we are able to make up cases where look-ahead beats myopic, but within our ranges of realistic input parameter values it was not easy.
The static north design is clearly better than the other two static designs. The others do not focus on the areas where the ES is really uncertain. 
By measuring temperature alone the expected variances in the ES is only slightly higher. This occurs because the ES is almost the same and because of the correlation of $0.6$ which means that temperature observations give information about salinity as well.

For the other predictive measures, myopic and look-ahead are still doing well, clearly beating the naive approach. The static zigzag design is usually the best among the static designs for these predictive performance measures because its pattern gives more effective coverage of the domain when there is significant spatial correlation as in this case.
\begin{table}[!h]
    \centering
    \scalebox{0.87}{
    \begin{tabular}{lrrrrrr}
    \toprule
    Parameter: $E_y(p[1-p])$ &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
    \midrule
                    \rowcolor{Gray}
ts. cor. low: 0.2  &  29.65 &   27.08 &       \textbf{26.44} &         32.33 &        40.31 &          36.23 \\
ts. cor. high: 0.8 &  36.18 &   \textbf{28.42} &       30.30 &         32.99 &        37.18 &          36.04 \\
                    \rowcolor{Gray}
std. low: 0.1      &  18.26 &   15.79 &       \textbf{15.35} &         18.89 &        26.65 &          26.10 \\
std. high: 0.5     &  47.01 &   \textbf{41.31} &       43.15 &         47.92 &        52.63 &          49.71 \\
                    \rowcolor{Gray}
cor. low: 0.8      &  44.42 &   \textbf{42.02} &       42.47 &         43.84 &        46.05 &          47.82 \\
cor. high: 0.2     &  29.73 &   21.09 &       \textbf{20.43} &         27.91 &        37.60 &          34.35 \\
                    \rowcolor{Gray}
temp. only         &  38.25 &   29.16 &       \textbf{28.69} &         34.32 &        39.81 &          36.17 \\
basecase           &  35.21 &   \textbf{28.56} &       28.94 &         33.50 &        39.84 &          39.26 \\
    \bottomrule
    \end{tabular}}
    \caption{Simulation results for the final mean excursion set variance (Eq. \eqref{two_partsK}) for 20 replicates, 8 configurations, and 6 strategies.}
    \label{tab:sim_res_ev}
\end{table}

\begin{table}[!h]
    \centering
    \scalebox{0.92}{
        \begin{tabular}{lrrrrrr}
        \toprule
        Parameter: RMSE &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
        \midrule
                \rowcolor{Gray}
ts. cor. low: 0.2  &   0.63 &    \textbf{0.57} &        0.59 &          0.62 &         0.57 &           0.59 \\
ts. cor. high: 0.8 &   0.63 &    \textbf{0.59} &        0.61 &          0.64 &         0.60 &           0.60 \\
                \rowcolor{Gray}
std. low: 0.1      &   0.39 &    0.37 &        0.38 &          0.38 &         0.40 &           \textbf{0.35} \\
std. high: 0.5     &   0.85 &    0.82 &        0.84 &          0.82 &         0.83 &           \textbf{0.81} \\  
                \rowcolor{Gray}
cor. low: 0.8      &   0.68 &    0.66 &        0.67 &          0.68 &         0.66 &           \textbf{0.65} \\
cor. high: 0.2     &   0.58 &    0.54 &        0.55 &          0.58 &         0.54 &           \textbf{0.49} \\
                \rowcolor{Gray}
temp. only         &   0.65 &    0.63 &        0.62 &          0.65 &         \textbf{0.61} &           0.62 \\
basecase           &  0.61 &   \textbf{0.56} &       \textbf{0.56} &         0.60 &        0.57 &          0.57 \\
        \bottomrule
        \end{tabular}}
    \caption{Simulation results for the final mean RMSE ($\frac{1}{B} \sum_{b=1}^{B} \sqrt{(\bmu-\tilde{\bmu})^2}$) for 20 replicates, 8 configurations, and 6 strategies.}
    \label{tab:sim_res_rmse}
\end{table}

\begin{table}[!h]
    \centering
    \scalebox{0.95}{
        \begin{tabular}{lrrrrrr}
        \toprule
        Parameter: $\bR^{2}$ &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
        \midrule
                        \rowcolor{Gray}
ts. cor. low: 0.2  &  27.00 &   33.29 &       32.24 &         26.71 &        30.56 &          \textbf{33.50} \\
ts. cor. high: 0.8 &  23.68 &   \textbf{33.87} &       32.42 &         26.72 &        30.73 &          33.61 \\
                        \rowcolor{Gray}
std. low: 0.1      &  25.38 &   30.80 &       30.29 &         26.09 &        28.75 &          \textbf{31.66} \\
std. high: 0.5     &  26.11 &   34.22 &       32.48 &         26.95 &        31.77 &          \textbf{34.48} \\
                        \rowcolor{Gray}
cor. low: 0.8      &   8.80 &   11.50 &       11.23 &          9.56 &        12.31 &          \textbf{12.58} \\
cor. high: 0.2     &  37.57 &   \textbf{48.68} &       48.11 &         39.61 &        41.98 &          48.04 \\
                        \rowcolor{Gray}
temp. only         &  11.67 &   \textbf{22.82} &       21.98 &         18.16 &        20.77 &          22.78 \\
basecase           &  24.71 &   33.45 &       32.25 &         26.71 &        30.62 &          \textbf{33.54}\\
        \bottomrule
        \end{tabular}}
    \caption{Simulation results for the final mean explained variance ($\bR^{2}=100*(1-(\bSigma_{posterior}/\bSigma_{initial}))$) for 20 replicates, 8 configurations, and 6 strategies.}
    \label{tab:sim_res_r2}
\end{table}

\section{Case study}\label{sec:case_study}

\begin{itemize}
    \item Tune prior mean, standard deviations and correlation parameters from ocean model simulations and initial runs. 
    \item Run tests 16 May or 21 May.
\end{itemize}

\newpage

\section{Closing remarks}\label{sec:concl_disc}

Additional flexibility can be gained by having two AUVs exploring together. It might then be useful to sample different parts of the space, and maybe move in parallel to best capture the excursion set.

\section*{Acknowledgements}
Thanks to SINTEF Ocean, AMOS, etc..

%\begin{supplement}
%\sname{Supplement A}\label{suppA}
%\stitle{Title of the Supplement A}
%\slink[url]{http://www.e-publications.org/ims/support/dowload/imsart-ims.zip}
%\sdescription{Dum esset rex in
%accubitu suo, nardus mea dedit odorem suavitatis. Quoniam confortavit
%seras portarum tuarum, benedixit filiis tuis in te. Qui posuit fines tuos}
%\end{supplement}

% == Adding references
\footnotesize
\bibliographystyle{apalike}
\bibliography{ref}

% AOS,AOAS: If there are supplements please fill:
%\begin{supplement}[id=suppA]
%  \sname{Supplement A}
%  \stitle{Title}
%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
%  \sdatatype{.pdf}" 
%  \sdescription{Some text}
%\end{supplement}

\end{document}

%\begin{enumerate}
%\item  This is the first item of an enumerated list.  Each item
%      in the list is marked with a ``tick.''  The document
%      style determines what kind of tick mark is used.
%\item  This is the second item of the list.  It contains another
%      list nested inside it.  The three inner lists are an {\em enumerated}
%      list.
%    \begin{enumerate}
%       \item This is the first item of an enumerated list that
%            is nested within the enumerated list.
%          \item This is the second item of the inner list.  \LaTeX\
%            allows you to nest lists deeper than you really should.
%      \end{enumerate}
%      This is the rest of the second item of the outer list.  It
%      is no more interesting than any other part of the item.
%   \item  This is the third item of the list.
%\end{enumerate}

### 2019-05-26 main_new.tex
updates

@@
 \floatsetup[table]{capposition=top}
-\usepackage{subfig}
+% \usepackage{subfig}
+\usepackage{subfigure}
 \usepackage[utf8]{inputenc}
@@
 % "Title of the paper"
-\title{Informative Sampling of Oceanographic Features using Excursion Probabilities for Multivariate Random Fields}
+\title{Informative Oceanographic Sampling using Excursion Probabilities for Multivariate Random Fields}
 \runtitle{Excursion Probabilities for Informative Sampling}
@@
 \address{Department of Marine Technology\\ Trondheim, Norway\printead{e1}}
-\affiliation{Department of Marine Technology, The Norwegian University of Science and Technology.}
-
-%\address{Department of Mathematical Sciences \printead{e2}}
-%\affiliation{Department of Mathematical Sciences, Norwegian University of Science and Technology (NTNU), Trondheim, Norway.}
-
-%\address{Department of Engineering Cybernetics,\printead{e4}}
-%\affiliation{Department of Engineering Cybernetics, Norwegian University of Science and Technology (NTNU), Trondheim, Norway.}
+\affiliation{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway}
+
+\address{Department of Mathematical Sciences \printead{e2}}
+\affiliation{Department of Mathematical Sciences, NTNU}
+
+\address{Idiap \printead{e3}}
+\affiliation{Uncertainty Quantification and Optimal Design group , Idiap Research Institute, Martigny, Switzerland}
+
+\address{Department of Engineering Cybernetics,\printead{e4}}
+\affiliation{Department of Engineering Cybernetics, NTNU}
 
@@
 \begin{abstract}
-We examine designs of experiments for informative sampling in the context of characterizing oceanographic phenomena.
-Methodologies are developed for efficient spatial sampling using autonomous robotic platforms, where the goal is to map the regions of cold and fresh water near a river mouth.
-Specifically, we define a design criterion based on reducing uncertainty of excursion sets in the context of multivariate random fields.
-We develop analytical expressions for the expected variance reduction in excursion sets, and show how these can be applicable for static designs and sequential designs building on myopic and look-ahead planning.
-We use simulations to study the properties of the suggested methods.
-Results from full-scale deployment with an autonomous underwater vehicle are also presented, for the case of river plume characterization.
+  We examine experiment design for informative sampling for
+  characterizing oceanographic phenomena. Methods are developed
+  for efficient spatial sampling using autonomous robotic platforms,
+  where the goal is to map the regions of cold and fresh water near a
+  river mouth. Specifically, we define a design criterion based on
+  reducing uncertainty of excursion sets in the context of multivariate
+  random fields. We develop analytical expressions for the expected
+  variance reduction in excursion sets, and show how these can be
+  applicable for static designs and sequential designs building on
+  myopic and look-ahead planning. We use simulations to study the
+  properties of the suggested methods. Results from field
+  deployments with an autonomous underwater vehicle are also presented.
 
@@
 \kwd{Gaussian Processes}
-\kwd{Design of experiments}
+\kwd{Experimental design}
 \kwd{Adaptive information gathering}
@@
 
-The pressure on marine resources is growing and increased accuracy, resolution, and persistent monitoring of the oceans is crucial for long-term sustainable management.
-A central problem for this type of environmental sensing is the large spatial extent and spatio-temporal variation inherent to ocean processes, prompting the need for effective allocation of sampling resources.
+Monitoring the worlds oceans has gained increased importance in the
+light of the changing climate. While the focus has often been on
+maga-fauna and the anthropogenic impact from micro-plastics to
+pollution, biological oceanographers have focused intently on studying
+micro-organisms at the base of the human food web. These organisms are
+critically impacted by the changing dynamics in the upper water-column,
+especially in coastal zones which are complex and often hard to observe
+in space and time. By studying the bio-geochemical processes in the
+upper water-column scientists can measure the impact of change, natural
+or anthropomorphic, and provide an informed opinion to policy makers to
+effect changes in preserving the environment. However, the challenge of 
+% The pressure on marine resources is growing and increased accuracy,
+% resolution, and persistent monitoring of the oceans is crucial for
+% long-term sustainable management. A central problem for this type of
+% environmental 
+sensing in the small% in the large spatial extent
+while studying the large spatio-temporal variation inherent to ocean
+processes, prompts the need for effective means of sampling.
+% allocation of sampling resources.
 %\begin{em}
-Sampling can, in this context, not simply be distributed uniformly ---along simple transect, grid, or ``lawn-mover" patterns--- but must instead be prioritized to relevant regions to save resources and avoid redundant data.
+Sampling can, in this context, not simply be distributed uniformly
+---along simple transects, grids, or ``lawn-mover" patterns--- but must
+instead be prioritized to relevant regions to ensure it is
+cost-effective while providing adequate coverage of the area of
+scientific interest.
+% save resources and avoid
+% redundant data.
 %\end{em}
 
-Traditional data collection at sea is to a large degree based on buoy or ship sampling, with both economic and logistic limitations that directly affect sensor coverage.
-Satellite data provide large-scale coverage, but have limited resolution, are limited to the surface and not available on cloudy days.
-A sustained focus on prioritized and efficient data collection strategies have therefore started to emerge.
-The advent of marine robotic platforms, especially autonomous underwater vehicles (AUVs), have provided means to execute this prioritization through the capacity of autonomy and data-driven sampling, where data collection in principle can be optimized.
-These capabilities have made robotics a crucial part of the emerging sensing practice for ocean science, allowing scientists to increase the observational efficiency and resolution beyond what was previously possible.
-But how should a robotic platform, such as an AUV, effectively prioritize and identify important regions for sampling?
-The answer to this question relates to statistical methodologies for design and sequential data assimilation, and the application domain is clearly an arena where statisticians can contribute.
-
-There has recently been some statistical attention in oceanography:
-\cite{wikle2013modern} present an overview of approaches for modeling uncertainties in the ocean sciences with focus on hierarchical statistical models. 
-%\cite{hopkins2010new} use statistical modeling to extract features from satellite data. 
-\cite{sahu2008space} study spatio-temporal models for sea surface temperature and salinity data, which will also are our key variables of interest in the current paper.
-\cite{mellucci2018oceanic} looks at the statistical prediction of features using an autonomous underwater glider. 
+Traditional data collection at sea has typically been based on static
+buoys, lagrangian floats or ship-centric methods, with significant
+logistical limitations that directly impact coverage and sampling
+resolution. Modern methods using satellite remote-sensing provide
+large-scale coverage, but have limited resolution, are limited to the
+sensing the surface of the ocean and impacted by cloud cover. With the
+advent of robust mobile robotic platforms \cite{Bellingham07}, more
+effective and higher resolution methods are making significant
+contributions to environmental monitoring and sampling. 
+% A
+% sustained focus on prioritized and efficient data collection strategies
+% have therefore started to emerge. The advent of marine robotic
+% platforms, especially
+In particular Autonomous underwater vehicles (AUVs), have made
+significant strides in advancing the state of sampling
+% provided means to execute this prioritization through the capacity of
+% autonomy and data-driven sampling, where data collection in principle
+% can be optimized. These capabilities have made
+and consequently have made robotics an integral part of ocean
+observation in large part driven by using 
+% of the emerging sensing practice for ocean science, allowing scientists
+% to increase the observational efficiency and resolution beyond what was
+% previously possible. But how should a robotic platform, such as an AUV,
+% effectively prioritize and identify important regions for sampling? The
+% answer to this question relates to 
+statistical methods in design and data assimilation
+\cite{das11b,Graham2013,Das2015,das15,fossuminformation,fossum18b}.% , and
+% the application domain is clearly an arena where statisticians can
+% contribute.
+% There has recently been some statistical attention in oceanography:
+Others include \cite{wikle2013modern} which % present an overview of approaches for modeling
+% uncertainties in the ocean sciences with
+focuses on hierarchical statistical models; \cite{sahu2008space} studies
+spatio-temporal models for sea surface temperature and salinity data,
+while
+% which will also are our key variables of interest in the current paper.
+\cite{mellucci2018oceanic} looks at the statistical prediction of
+features using an underwater glider.
 %From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest.
@@
 Quantification of gradient features is therefore a much needed competence in robotic sampling.
-The combination of statistical tools and robotic platforms is a natural one, enabling information-based robotic sensing systems.
-Central to this is the ability to model spatially correlated variables and provide formal measures of uncertainty.
-The by far most common model for practical use builds on Gaussian processes (GPs), which will also be used in our formulation, because this approach enables evaluation and prioritization of sampling efforts to be formalized into decision making systems on-board the robotic platform.
+The combination of statistical tools and robotic platforms is a natural
+symbiosis which enables information-based sensing. Central to this is
+the ability to model spatially correlated variables and provide formal
+measures of uncertainty. By far the most practical models use Gaussian
+processes (GPs), which will also be used in our formulation, in large
+part because this approach enables evaluation and prioritization of
+sampling efforts to be formalized into decision making systems on-board
+a robotic platform.
 %The imperfect understanding of causality coupled with the underlying fundamental uncertainty of the ocean implies that we need to obtain more measurements and use statistical models and methods to determine what and how phenomena form, to increase the quality of predictive statements.
 
-Toward this end, we study excursion sets (ES) and excursion probabilities (EP) for spatial GPs, and the goal is to design sampling strategies for reducing the variance in the ES.
-The main contributions involve: i) extending ES-based uncertainty criteria to the situation with multivariate random fields, allowing the evaluation to include multiple environmental parameters such as temperature and salinity, ii) develop methods for informative environmental sensing and autonomous data collection, and iii) demonstrate the approach in a real case scenario, sampling a river plume using an AUV.
-
-The remainder of the paper is organized as follows: Section \ref{sec:bg} introduces background content relating to ocean sampling, and basic notation for ESs and EPs.
-Section \ref{sec:GP_EP} sets up the required assumptions of GPs, including bi-variate random fields for temperature and salinity.
-Section \ref{sec:sur} formulates the sampling design criteria for variance reduction in ESs, with analytical derivation for multivariate spatial processes.
-Section \ref{sec:heuristics} builds on these sampling criteria to present heuristics for adaptive sampling strategies.
-Section \ref{sec:simulations} shows results and discusses properties of the methods on simulated examples.
-Section \ref{sec:case_study} demonstrates a full-scale implementation using an AUV to characterize a river plume.
-Section \ref{sec:concl_disc} contains a summary and discussion with links to future work.
+In this work, we study excursion sets (ES) and excursion probabilities
+(EP) for spatial GPs, and the goal is to design sampling strategies for
+reducing the variance in the ES. The main contributions involve: i)
+extending ES-based uncertainty criteria to multivariate random fields,
+allowing the evaluation to include multiple environmental parameters
+such as temperature and salinity, ii) develop methods for informative
+environmental sensing and autonomous data collection, and iii)
+demonstrating the approach in a real world scenario of sampling a river
+plume using an AUV.
+
+The remainder of the paper is organized as follows: Section \ref{sec:bg}
+provides some background on ocean sampling, as also some notation for
+ESs and EPs. Section \ref{sec:GP_EP} sets up the required assumptions of
+GPs, including bi-variate random fields for temperature and salinity.
+Section \ref{sec:sur} formulates the sampling design criteria for
+variance reduction in ESs, with analytical derivation for multivariate
+spatial processes. Section \ref{sec:heuristics} builds on these sampling
+criteria to present heuristics for adaptive sampling. Section
+\ref{sec:simulations} shows results and discusses properties of the
+methods in simulation while Section \ref{sec:case_study} instantiates
+the work in the form of characterizing a river plume. We conclude with
+Section \ref{sec:concl_disc} with a summary and a discussion of future
+work.
 
@@
 
-The processes occurring in the upper water column and the photic zone (upper sunlit zone) are of vital importance for long-term sustainable management of the greater marine ecosystem, climate, and human health.
-With the onset of the warming oceans and increased anthropogenic pressure, it is critical to improve our understanding of the associated ecosystem response.
-Central to this is the understanding of biophysical interaction (see Fig.
-\ref{fig:envir2}) driven by an agglomeration of physical forcings (wind, topography, bathymetry, tidal influences and local geography such as fresh water inflows) mixed with incipient micro-biology driven by planktonic and coastal anthropogenic input, such as pollution and agricultural runoff, which results in a range of ecosystem phenomena (e.g.
-blooms, anoxic zones, plumes) ---with significant impact to society.
-Renowned oceanographer Walter Munk called the 20\textsuperscript{th} century, the \emph{century of undersampling} \citep{munk2002}, pointing out that lack of sufficient sampling resolution in time and space, owing to current observation practices.
-The lack of resolution, arising from sparsely distributed sensing methods, is an issue for characterizing, understanding, and predicting ocean phenomena.
-In later decades, this view has been driving ocean sensing practices towards more networked and distributed use of robotic assets in order to fill the gaps in resolution (See Fig.
-\ref{fig:envir}).
+% The processes occurring in the upper water-column and the photic zone
+% (upper sunlit zone) are of vital importance for long-term sustainable
+% management of the greater marine ecosystem, climate, and human health.
+% With the onset of the warming oceans and increased anthropogenic
+% pressure, it is critical to improve our understanding of the associated
+% ecosystem response.
+Central to understanding the changes taking place in the upper
+water-column is the understanding of biophysical interaction (see Fig.
+\ref{fig:envir2}) driven by an agglomeration of physical forcing (wind,
+topography, bathymetry, tidal influences and local geography such as
+fresh water inflows) mixed with incipient micro-biology driven by
+planktonic and coastal anthropogenic input, such as pollution and
+agricultural runoff, which results in a range of ecosystem phenomena
+(e.g. blooms, anoxic zones, plumes) ---with significant impact to
+society. Renowned oceanographer Walter Munk called the
+20\textsuperscript{th} century, the \emph{century of undersampling}
+\citep{munk2002}, pointing out that lack of sufficient sampling
+resolution in time and space, % owing to current observation practices.
+% The lack of resolution, 
+% arising from sparsely distributed sensing methods, is
+as a critical issue for characterizing ocean phenomena. This view has
+been driving ocean sensing practices towards more networked and
+distributed use of robotic assets in order to fill the gaps in
+our understanding (See Fig. \ref{fig:envir}).
+
 \begin{figure}[!h] 
   \centering 
-\subfloat[Modern ocean sensing network, where robotic assets augment ship based measurements to increase coverage of the phenomena.]{\includegraphics[width = 0.48\textwidth]{Figures/envir.pdf}\label{fig:envir1}}
+  \subfigure[Modern ocean sensing network, where robotic assets augment
+  ship based measurements to increase coverage of the phenomena.]{\includegraphics[width =
+    0.48\textwidth]{Figures/envir.pdf}\label{fig:envir1}}
   \hfill
-\subfloat[Different important coastal processes directly or indirectly associated with gradients and boundaries.]{\includegraphics[width = 0.48\textwidth]{Figures/envir_ocean.pdf}\label{fig:envir2}}
-\caption{Ocean observation is moving away from single-ship sampling towards more collaborative networked operations in order to resolve the numerous processes and their interaction.}
-\label{fig:envir}
+  \subfigure[Different important coastal processes directly or indirectly
+  associated with gradients and boundaries.]{\includegraphics[width =
+    0.48\textwidth]{Figures/envir_ocean.pdf}\label{fig:envir2}}
+  \caption{Ocean observation is moving away from single-ship sampling
+    towards more collaborative networked operations in order to resolve
+    the numerous processes and their interaction.} \label{fig:envir}
 \end{figure}
 
-Our focus in the current paper is limited to spatial characterization of a river plume system. 
-Among other plumes and front systems, Fig.
-\ref{fig:river_fronts} shows the survey area (Nidelva outlet, Norway) taken as cold freshwater enters from the river.
-The image shows the separation of water masses (different haphazard patterns and surface slicks) and strong gradients for both temperature and salinity, where cold fresh water from the river meets saline ocean water.
-Because of the local topography and the Coreolis force \citep{coriolis1835memoire} the cold fresh water tends to stay near land to the east.
-Depending on the river discharge, tidal effects, and temperature differences, this boundary in water masses will get distorted.
-Prior knowledge about the location and evolution of these features is therefore highly uncertain, making deterministic planning sub-optimal, and one would rather demonstrate the ability of AUVs to map the plume area efficiently.
+Our focus in this work is limited to spatial characterization of a river
+plume system. Fig. \ref{fig:river_fronts} shows the survey area (the
+outlet of the Nidelva river, Trondheim, Norway) taken as cold freshwater
+enters from the river showing the separation of water masses and strong
+gradients for both temperature and salinity. Because of the local
+topography and the Coreolis force \citep{coriolis1835memoire} the cold
+fresh water tends to stay near land to the east. Depending on the river
+discharge, tidal effects, and temperature differences, this boundary in
+water masses will get distorted. Prior knowledge about the location and
+evolution of these features is therefore highly uncertain, making
+deterministic planning sub-optimal.% , and one would rather demonstrate the
+% ability of AUVs to map the plume area efficiently.
+
 \begin{figure}[!h]
 \centering
- \subfloat[River front - Columbia River, Astoria, Oregon, US.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/a.png}\label{fig:river1}}
-	\hfill
-	\subfloat[Tidal front - Korsfjorden, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/d.png}\label{fig:river2}}
- \hfill
- \subfloat[River plume - Rio de la Plata, Buenos Aires, Argentina.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/b.png}\label{fig:river3}}
- \hfill
- \subfloat[Frontal patterns - Nidelva, Trondheim, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/c.png}\label{fig:river4}}
-	\caption{Different types of frontal features associated with river-ocean interaction.
-(\ref{fig:river1}) The river plume front of the Columbia River meeting the Pacific Ocean.
-(\ref{fig:river2}) Superimposed images (10 minutes) of the surface tidal front in Korsfjorden, Norway.
-(\ref{fig:river3}) The river plume of the Rio de la Plata, taken by European Space Agency (ESA)'s Envisat platform and MEdium Resolution Imaging Spectrometer (MERIS) sensor.
-(\ref{fig:river4}) Aerial photo of the Nidelva outlet in Trondheim, where fresh river water meets the salty Fjord water and creates frontal patterns.
-Image courtesy: (\ref{fig:river1}) \cite{zamon2014marine}, (\ref{fig:river3}) ESA, CC BY-SA 3.0 IGO.}
+ % \subfigure[River front - Columbia River, Astoria, Oregon, US.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/a.png}\label{fig:river1}}
+ % \hfill
+ \subfigure[Tidal front - Korsfjorden, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/d.png}\label{fig:river2}}
+ % \hfill
+ % \subfigure[River plume - Rio de la Plata, Buenos Aires, Argentina.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/b.png}\label{fig:river3}}
+ % \hfill
+ \subfigure[Frontal patterns - Nidelva, Trondheim, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/c.png}\label{fig:river4}}
+ \caption{Two examples of frontal features associated with
+   river-ocean interaction. % (\ref{fig:river1}) The river plume front of
+   % the Columbia River meeting the Pacific Ocean.
+   (\ref{fig:river2})  Superimposed images (over 10 minutes) of a surface tidal front in
+   Korsfjorden, Norway. % (\ref{fig:river3}) The river plume of the Rio de
+   % la Plata, taken by European Space Agency (ESA)'s Envisat platform and
+   % MEdium Resolution Imaging Spectrometer (MERIS) sensor.
+   (\ref{fig:river4}) Aerial image of the mouth of the Nidelva river in
+   Trondheim, Norway where fresh water meets the salty fjord and creates
+   frontal patterns.} % Image courtesy: (\ref{fig:river1})
+   % \cite{zamon2014marine}, (\ref{fig:river3}) ESA, CC BY-SA 3.0 IGO.}
  \label{fig:river_fronts}
@@
 %This challenge can be addressed by employing more elaborate and adaptive sampling strategies that can capitalize on both prior and current observations in order to locate and map these areas.
-River plumes belong to a class of ocean processes that are local and usually act --compared to large ocean scale processes-- on smaller scales from 10 m -- 5 km (\emph{sub-mesoscale}), where spatial lateral variability tends to dominate.
-Remote sensing and synthetic ocean models cannot provide detailed inputs at this scale \citep{Lermusiaux:2006} and the only way to resolve fine resolution is by direct observation.
-At large scale, usually referred to as above \emph{mesoscale} ($>50$ km), both three-dimensional space and time dynamics are important and can shift substantially, but in the case of river plumes one can often limit scope to cover only lateral spatial elements ([north, east] domain).
-For this case, time-effects can be regarded as static when acquiring AUV data over a short time window and limited region.
-However, the methodological framework can be extended to three-dimensional spatial variability and temporal elements in a similar manner.
+
+River plumes belong to a class of ocean processes that are local and
+usually act on smaller \emph{sub-mesoscale}, from 10 m -- 5 km where
+spatial lateral variability tends to dominate. Remote sensing and
+synthetic ocean models find it challenging to provide detailed at this
+scale \citep{Lermusiaux:2006} and the principal way to resolve for fine
+resolution is by direct observation. At larger \emph{mesoscale} ($>50$
+km), both three-dimensional space and time dynamics are important and
+can shift substantially, while in the case of river plumes one can often
+limit scope to cover only lateral spatial elements ([north, east]
+domain). For this case, time-effects can be regarded as static when
+acquiring AUV data over a short time window and limited region. However,
+the methodological framework can be extended to three-dimensional
+spatial variability and temporal elements in a similar manner.
 
 \subsection{Autonomous Vehicles}
-AUV is a term that describes autonomous sensor platforms that operate underwater used to gather spatio-temporal data from a targeted domain, carrying payload to survey the water column or the seafloor.
-They run at $\sim 1-3$ m/s, can reach depths from 100--6000m, and have battery capacity $\sim$ 8--42 hrs, depending on survey speed, payload sensors, and manufacture.
-By deploying AUVs one can in principle fill the gaps and augment data obtained from either ocean models, earth observing satellite data, or fixed-location buoys due to their survey speed and mobility.
-Traditional operations and surveys with AUVs are usually limited to planning sampling along fixed transects or survey lines set by human operators, programmed as a sequential plan consisting of different waypoints and depths.
-But by using on-board computers to continuously evaluate, update, and refine future sampling locations, making information gathering adaptive \citep{das11b}.
-For instance, an AUV could reconstruct or modify a survey line based on what temperatures it is measuring by using embedded methods in decision-theoretic planning \citep{ryan10}.
+
+AUVs operate underwater used to gather spatio-temporal data from a
+targeted domain, carrying a scientific payload to survey the water
+column or the seafloor. They run at $\sim 1-3$ m/s, can reach depths
+from 100--6000m. In water staying capacity depends on survey speed,
+payload sensors and mission design. By deploying AUVs one can in
+principle fill gaps and augment data obtained from either ocean models,
+earth observing satellite data, or fixed-location sensors on buoys due
+to their speed and mobility. Traditional operations and surveys with
+AUVs are usually limited to planning samples along fixed transects or
+survey lines set by human operators, programmed as a sequential plan
+consisting of different waypoints and depths. But by using on-board
+algorithms to continuously evaluate, update, and refine future sampling
+locations, making information gathering adaptive \citep{das11b} is
+feasible. For instance, an AUV could reconstruct or modify a survey line
+based on what temperatures it is measuring by using embedded methods in
+decision-theoretic planning \citep{ryan10}.
 
@@
 
-Determining paths for mobile robotic sensors in order to maximize the information gained about an environment is formally known as informative path planning.
-This, and related problems such as the orienteering problem \citep{Golden87}, has been studied in the context of graphs, where the potential measurement are assigned to nodes on which evaluation of different routes can be conducted.
+Determining paths for mobile robotic sensors in order to maximize the
+information gained about an environment is formally known as informative
+path planning. This, and related problems such as the orienteering
+problem \citep{Golden87}, has been studied in the context of graphs,
+where the potential measurement are assigned to nodes on which
+evaluation of different routes can be conducted.
 %In this context, a fundamental result from \cite{nemhauser1978analysis} proves that a simple greedy algorithm (iteratively selecting the location which most increases the utility) can achieve a near-optimal solution if the problem can be shown to be \emph{submodular}\footnote{an intuitive diminishing returns property, where the informative value of adding sensors decrease with the number of sensors added.}.
@@
 One realization of the random fields representing salt and temperature is shown in Figure \ref{fig:true_temp} and \ref{fig:true_sal}.
+
 \begin{figure}[!h]
   \centering
-\subfloat[Simulated temperature.]{\includegraphics[height = 0.41\textwidth]{Figures/sim/true_temp.pdf}\label{fig:true_temp}}
+  \subfigure[Simulated temperature.]{\includegraphics[height = 0.41\textwidth]{Figures/sim/true_temp.pdf}\label{fig:true_temp}}
   \hfill
-\subfloat[Simulated salinity.]{\includegraphics[height = 0.41\textwidth]{Figures/sim/true_sal.pdf}\label{fig:true_sal}}
+  \subfigure[Simulated salinity.]{\includegraphics[height = 0.41\textwidth]{Figures/sim/true_sal.pdf}\label{fig:true_sal}}
   \hfill
-\subfloat[Excursion set given the realization of temperature and salinity.]{\includegraphics[height = 0.41\textwidth]{Figures/sim/es_ts_true.pdf}\label{fig:ESet}}
+  \subfigure[Excursion set given the realization of temperature and salinity.]{\includegraphics[height = 0.41\textwidth]{Figures/sim/es_ts_true.pdf}\label{fig:ESet}}
   \hfill
-\subfloat[Excursion probabilities given collected data along the strategy ``static\_north".]{\includegraphics[height = 0.41\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
-\caption{(\ref{fig:true_temp}) and (\ref{fig:true_sal}) shows realisations of the simulated temperature and salinity fields, as well as the associated excursion set (\ref{fig:ESet}).
-(\ref{fig:Eprob}) show the estimated excursion probabilities after performing data collection along the N-S survey line.}
+  \subfigure[Excursion probabilities given collected data along the strategy ``static\_north".]{\includegraphics[height = 0.41\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
+  \caption{(\ref{fig:true_temp}) and (\ref{fig:true_sal}) shows
+    realisations of the simulated temperature and salinity fields, as
+    well as the associated excursion set (\ref{fig:ESet}).
+    (\ref{fig:Eprob}) show the estimated excursion probabilities after
+    performing data collection along the N-S survey line.} 
 \label{fig:realisations}
 \end{figure}
+
 The true ES from this prior realization is plotted in Figure \ref{fig:ESet}.
@@
 
-At each step, the AUV will choose to move along an edge in this waypoint graph.
-While doing so, it gathers three measurements.
-These are assimilated in the model, and then a new choice is done at the end of the edge.
-The paths of will differ for the various sequential designs.
-The designs will also differ between the runs, because the measurements vary among replicates.
-
-Results of the replicate runs are shown in Figure \ref{fig:sim_results}, where the different criteria are plotted as a function of surveying distance.
-Here, expected ES variance in Figure \ref{fig:avg_ev} is reduced most when using the \texttt{myopic} and \texttt{look-ahead} strategies.
-There is not much difference between these two.
+At each step, the AUV will choose to move along an edge in this waypoint
+graph. While doing so, it gathers three measurements. These are
+assimilated in the model, and then a new choice is done at the end of
+the edge. The paths of will differ for the various sequential designs.
+The designs will also differ between the runs, because the measurements
+vary among replicates.
+
+Results of the replicate runs are shown in Figure \ref{fig:sim_results},
+where the different criteria are plotted as a function of surveying
+distance. Here, expected ES variance in Figure \ref{fig:avg_ev} is
+reduced most when using the \texttt{myopic} and \texttt{look-ahead}
+strategies. There is not much difference between these two.
+
+\begin{figure}[!th]
+  \centering
+  % \subfigure[Excursion set variance $E_y(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
+  \subfigure[Excursion set $E_y$(p\lbrack 1-p\rbrack) variance.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
+  \hfill
+  \subfigure[RMSE between estimated field and truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
+  \hfill 
+  \subfigure[Explained variance $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
+  \hfill 
+  \subfigure[Time used to do inference.]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}}
+\caption{Simulation results from 100 replicate simulations for 10
+  sampling choices/steps on the grid. (Updated:01.05.2019)} 
+\label{fig:sim_results}
+\end{figure}
+
+The north-south design is also doing well, because it moves near the
+boundary of the water masses. In Figure \ref{fig:avg_rmse} and
+\ref{fig:avg_r2} we show results of RMSE and explained variance. Both
+\texttt{myopic} and \texttt{look-ahead} still perform well, but some of
+the static east and zigzag are also achieving good results because they
+are set to always cover large parts of the domain, and in this way visit
+locations that are not considered interesting by the sequential
+strategies targeting ES variance. Regarding computing time in Fig.
+\ref{fig:avg_time}, the \texttt{naive} strategy is on par with the
+static designs, while \texttt{myopic} is slower and \texttt{look-ahead}
+much slower. The \texttt{look-ahead} strategy is reaching levels that
+are near impractical for on-board execution. Of course, with harder
+pruning of branches in the graph, it would be faster to run. Then again,
+this pruning is likely the reason that this strategy does not
+outperforms the myopic approach.
+
+In Figure \ref{fig:route_choices} we plot the realized sampling paths of
+the sequential schemes and static designs.
 
@@
   \centering
-\subfloat[Excursion set variance $E_y(p[1-p])$ .]{\includegraphics[height = 0.49\textwidth]{Figures/sim/avg_EV.pdf}\label{fig:avg_ev}}
+  \subfigure[Look-ahead strategy.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_look-ahead.pdf}\label{fig:avg_look-ahead}}
   \hfill
-\subfloat[RMSE between estimated field and truth.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/avg_RMSE.pdf}\label{fig:avg_rmse}}
+  \subfigure[Myopic strategy.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_myopic.pdf}\label{fig:avg_myopic}}
   \hfill
-\subfloat[Explained variance $\bR^{2}$.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/avg_R2.pdf}\label{fig:avg_r2}}
+  \subfigure[Naive strategy.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_naive.pdf}\label{fig:route_naive}}
   \hfill
-\subfloat[Time used to do inference.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/avg_Time.pdf}\label{fig:avg_time}}
-\caption{Simulation results from 100 replicate simulations for 10 sampling choices/steps on the grid. (Updated:01.05.2019)}
-\label{fig:sim_results}
-\end{figure}
-
-The north-south design is also doing well, because it moves near the boundary of the water masses. 
-In Figure \ref{fig:avg_rmse} and \ref{fig:avg_r2} we show results of RMSE and explained variance.
-Both \texttt{myopic} and \texttt{look-ahead} still perform well, but some of the static east and zigzag are also achieving good results because they are set to always cover large parts of the domain, and in this way visit locations that are not considered interesting by the sequential strategies targeting ES variance.
-Regarding computing time in Fig.
-\ref{fig:avg_time}, the \texttt{naive} strategy is on par with the static designs, while \texttt{myopic} is slower and \texttt{look-ahead} much slower.
-The \texttt{look-ahead} strategy is reaching levels that are near impractical for on-board execution.
-Of course, with harder pruning of branches in the graph, it would be faster to run.
-Then again, this pruning is likely the reason that this strategy does not outperforms the myopic approach.
-
-In Figure \ref{fig:route_choices} we plot the realized sampling paths of the sequential schemes and static designs. 
-\begin{figure}[!h]
-\centering
-\subfloat[Look-ahead strategy.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_look-ahead.pdf}\label{fig:avg_look-ahead}}
+  \subfigure[Static strategies.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_static.pdf}\label{fig:route_static}}
   \hfill
-\subfloat[Myopic strategy.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_myopic.pdf}\label{fig:avg_myopic}}
-\hfill
-\subfloat[Naive strategy.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_naive.pdf}\label{fig:route_naive}}
-\hfill
-\subfloat[Static strategies.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_static.pdf}\label{fig:route_static}}
-\hfill
-\caption{The overlaid route choices, superimposed for the 100 replicate surveys with 10 sampling choices/steps.}
+\caption{The overlaid route choices, superimposed for the 100 replicate
+  surveys with 10 sampling choices/steps.} 
 \label{fig:route_choices}
 \end{figure}
-The naive strategy appears to get stuck in the southern part of the domain in some cases, because it is too focused on the probabilities near $0.5$.
-The myopic strategy seems to be covering a wider domain than the look-ahead strategy, possibly because the look-ahead strategy sees less value of going far west because it account for having to return east again.
+
+The naive strategy appears to get stuck in the southern part of the
+domain in some cases, because it is too focused on the probabilities
+near $0.5$. The myopic strategy seems to be covering a wider domain than
+the look-ahead strategy, possibly because the look-ahead strategy sees
+less value of going far west because it account for having to return
+east again.
 

### 2019-05-27 main_new.tex
updates

@@
   \hfill
-  \subfigure[Different important coastal processes directly or indirectly
+  \subfigure[Different coastal processes directly or indirectly
   associated with gradients and boundaries.]{\includegraphics[width =

### 2019-05-27 main_new.tex
updates

@@
 
-\runauthor{TO. Fossum et al.}
+\runauthor{Fossum et al.}
 
@@
 column or the seafloor. They run at $\sim 1-3$ m/s, can reach depths
-from 100--6000m. In water staying capacity depends on survey speed,
-payload sensors and mission design. By deploying AUVs one can in
+from 100--6000m with in-water staying capacity depending on survey
+speed, payload sensors and mission design. By deploying AUVs one can in
 principle fill gaps and augment data obtained from either ocean models,
@@
 based on what temperatures it is measuring by using embedded methods in
-decision-theoretic planning \citep{ryan10}.
+decision-theoretic planning \citep{Rajan12b}.
 
@@
 
-Determining paths for mobile robotic sensors in order to maximize the
-information gained about an environment is formally known as informative
-path planning. This, and related problems such as the orienteering
-problem \citep{Golden87}, has been studied in the context of graphs,
-where the potential measurement are assigned to nodes on which
-evaluation of different routes can be conducted.
+% Determining paths for mobile robotic sensors in order to maximize the
+% information gained about an environment is formally known as informative
+% path planning. This, and related problems such as the orienteering
+% problem \citep{Golden87}, has been studied in the context of graphs,
+% where the potential measurement are assigned to nodes on which
+% evaluation of different routes can be conducted.
 %In this context, a fundamental result from \cite{nemhauser1978analysis} proves that a simple greedy algorithm (iteratively selecting the location which most increases the utility) can achieve a near-optimal solution if the problem can be shown to be \emph{submodular}\footnote{an intuitive diminishing returns property, where the informative value of adding sensors decrease with the number of sensors added.}.
 Building on this result, \cite{chekuri2005recursive} explored this in the setting of a graph using a recursive greedy algorithm, providing near-optimal solutions depending on the planning horizon and graph resolution.
-Adaptive sampling of an evolving frontal feature has been explored in \cite{Zhang2012} and \cite{Pinto2018} using state-based autonomous agents.
-Both approaches use a reactive-adaptive scheme, meaning that the selection does not use a statistical model of the environment in its strategic sampling effort, but rather adapts based on detection and state switching.
-Myopic sampling, i.e.
-step-wise selection of the path (on the graph) which reduces the information criterion the most, has been used for adaptive surveying \citep{singh2009efficient,Binney2013}.
-These exploration strategies focus largely on reducing predictive variance or mutual information (entropy).
-Because the variance and entropy reduction are independent of the actual data realizations under the assumption of GP models, the use of data-driven (adaptive) criteria was introduced to include more targeted sampling of interesting regions (hotspot sampling) by \cite{Low2009} and \cite{fossuminformation}.
+Adaptive sampling of an evolving frontal feature has been explored in
+\cite{fronts11,Zhang2012,Pinto2018,costa19} using state-based autonomous
+agents. These approaches typically use a reactive-adaptive scheme,
+whereby exploration does not rely on a statistical model of the
+environment, but rather adapts based on closing the sensing and
+actuation loop. Myopic sampling, i.e. step-wise selection of the path
+(on the graph) which reduces the information criterion the most, has
+been used for adaptive surveys \citep{singh2009efficient,Binney2013} which
+focus largely on reducing predictive variance or mutual information
+(entropy). Because the variance and entropy reduction are independent of
+the actual data realizations under the assumption of GP models, the use
+of data-driven (adaptive) criteria was introduced to include more
+targeted sampling of interesting regions (hotspot sampling) by
+\cite{Low2009} and \cite{fossuminformation}.
 
@@
 
-In our setting we study two variables; temperature and salinity, which together characterize the phenomenon taking place in the river plume.
-At location $\bx \in \mathcal{M} \subset \mathbf{R^2}$, we define the two process variables by $\bxi(\bx)=(\xi_a(\bx),\xi_b(\bx))$.
+In our setting we study two variables; temperature and salinity, which
+together characterize the phenomenon taking place in the river plume. At
+location $\bx \in \mathcal{M} \subset \mathbf{R^2}$, we define the two
+process variables by
+$\bxi(\bx)=(\xi_a(\bx),\xi_b(\bx))$.
 %In practice there is of course time variability, but we assume this to be negligible during the AUV operation.

### 2019-05-27 main_new.tex
updates

@@
 (on the graph) which reduces the information criterion the most, has
-been used for adaptive surveys \citep{singh2009efficient,Binney2013} which
-focus largely on reducing predictive variance or mutual information
-(entropy). Because the variance and entropy reduction are independent of
-the actual data realizations under the assumption of GP models, the use
-of data-driven (adaptive) criteria was introduced to include more
-targeted sampling of interesting regions (hotspot sampling) by
+been used for adaptive surveys \citep{singh2009efficient,Binney2013}
+which focus largely on reducing predictive variance or mutual
+information (entropy). Because the variance and entropy reduction are
+independent of the actual data realizations under the assumption of GP
+models, the use of data-driven (adaptive) criteria was introduced to
+include more targeted sampling of regions of scientific interest by
 \cite{Low2009} and \cite{fossuminformation}.

### 2019-05-27 main_new.tex
updates

@@
 \definecolor{Gray}{gray}{0.9}
+\newcommand{\kcomment}[1]{{\color{blue}{\{KR: #1\}}}}
 
@@
 In our setting we study two variables; temperature and salinity, which
-together characterize the phenomenon taking place in the river plume. At
-location $\bx \in \mathcal{M} \subset \mathbf{R^2}$, we define the two
-process variables by
+together characterize the phenomenon taking place in the river plume as
+they define the gradients which define the frontal zone we wish to
+observe. At location $\bx \in \mathcal{M} \subset \mathbf{R^2}$, we
+define the two process variables by
 $\bxi(\bx)=(\xi_a(\bx),\xi_b(\bx))$.
@@
 
-In the context illustrated by Figure \ref{fig:river4}, interest is on spatial separation of cold freshwater. 
-Because a main goal is classifying the water masses, ES variability is a useful sampling criterion for guiding the observation effort of the AUV. 
-A bivariate ES is here defined via joint variables being below thresholds, at individual locations, for the process over the spatial domain $\mathcal{M}$.
-Denoting the threshold for temperature by $T_a$ and the threshold for salinity by $T_b$,
-we use the following definition;
+% In the context illustrated by Figure \ref{fig:river4},
+Our focus is on spatial separation of cold freshwater from the saline
+waters of the fjord and to be able to classify these water masses. In
+this context, ES variability is a useful sampling criterion for guiding
+the observation effort of the AUV. A bivariate ES is defined via joint
+variables being below thresholds, at individual locations, for the
+process over the spatial domain $\mathcal{M}$. Denoting the threshold
+for temperature by $T_a$ and the threshold for salinity by $T_b$, we use
+the following definition;
  \begin{equation}\label{ES}

### 2019-05-28 main_new.tex
updates

@@
 together characterize the phenomenon taking place in the river plume as
-they define the gradients which define the frontal zone we wish to
-observe. At location $\bx \in \mathcal{M} \subset \mathbf{R^2}$, we
-define the two process variables by
+they define the gradients defining the frontal zone. At location
+$\bx \in \mathcal{M} \subset \mathbf{R^2}$, we define the two process
+variables by
 $\bxi(\bx)=(\xi_a(\bx),\xi_b(\bx))$.
@@
 this context, ES variability is a useful sampling criterion for guiding
-the observation effort of the AUV. A bivariate ES is defined via joint
-variables being below thresholds, at individual locations, for the
-process over the spatial domain $\mathcal{M}$. Denoting the threshold
-for temperature by $T_a$ and the threshold for salinity by $T_b$, we use
-the following definition;
+the observational effort of the AUV. A bivariate ES is defined via joint
+variables \kcomment{'being below thresholds' -- is this correct?}, at
+individual locations, for the process over $\mathcal{M}$. If $T_a$ is
+the threshold for temperature and $T_b$ the threshold for salinity,
+we use the following definition:
+
  \begin{equation}\label{ES}
@@
  \end{equation}
+
 The excursions could also be defined as larger than threshold, or within a boundary of limits for the two variables. Calculations and interpretations will be similar in these different cases.

### 2019-05-29 main_new.tex
Updates from Overleaf

--- 2019-05-27 main_new.tex
@@
   river mouth. Specifically, we define a design criterion based on
-  reducing uncertainty of excursion sets in the context of multivariate
+  reducing uncertainty in excursion sets of multivariate
   random fields. We develop analytical expressions for the expected
-  variance reduction in excursion sets, and show how these can be
+  variance reduction in excursion set indicators, and show how these can be
   applicable for static designs and sequential designs building on
@@
 %\begin{em}
-Sampling can, in this context, not simply be distributed uniformly
+Sampling can, in this context, not simply be distributed evenly
 ---along simple transects, grids, or ``lawn-mover" patterns--- but must
@@
 % There has recently been some statistical attention in oceanography:
-Others include \cite{wikle2013modern} which % present an overview of approaches for modeling
+Others include \cite{wikle2013modern} who % present an overview of approaches for modeling
 % uncertainties in the ocean sciences with
-focuses on hierarchical statistical models; \cite{sahu2008space} studies
+focus on hierarchical statistical models; \cite{sahu2008space} study
 spatio-temporal models for sea surface temperature and salinity data,
@@
 % which will also are our key variables of interest in the current paper.
-\cite{mellucci2018oceanic} looks at the statistical prediction of
+\cite{mellucci2018oceanic} look at the statistical prediction of
 features using an underwater glider.
@@
 processes (GPs), which will also be used in our formulation, in large
-part because this approach enables evaluation and prioritization of
+part because this approach allows efficient implementation and evaluation, enabling prioritization of
 sampling efforts to be formalized into decision making systems on-board
@@
 (EP) for spatial GPs, and the goal is to design sampling strategies for
-reducing the variance in the ES. The main contributions involve: i)
+reducing the variance in the ES indicators. The main contributions involve: i)
 extending ES-based uncertainty criteria to multivariate random fields,
@@
 GPs, including bi-variate random fields for temperature and salinity.
-Section \ref{sec:sur} formulates the sampling design criteria for
-variance reduction in ESs, with analytical derivation for multivariate
+Section \ref{sec:sur} formulates the choice of sampling design criteria for
+variance reduction in ES indicators, with analytical derivation for multivariate
 spatial processes. Section \ref{sec:heuristics} builds on these sampling
@@
 limit scope to cover only lateral spatial elements ([north, east]
-domain). For this case, time-effects can be regarded as static when
+domain). Morover, for this case, time-effects can be regarded as static when
 acquiring AUV data over a short time window and limited region. However,
-the methodological framework can be extended to three-dimensional
-spatial variability and temporal elements in a similar manner.
+the methodological framework can be extended to higher dimensional situations in a similar manner.
 
@@
  \begin{equation}\label{ES}
-     \mbox{ES} = \{\bx \in \mathcal{M} | \xi_a(\bx) < T_a,\xi_b(\bx) < T_b\}.
+     \mbox{ES} = \{\bx \in \mathcal{M} | \xi_a(\bx) \leq T_a,\xi_b(\bx) \leq T_b\}.
  \end{equation}
@@
 \begin{equation}\label{eq:prob}
- p(\bx) = P(\xi_a(\bx) < T_a, \xi_b(\bx) < T_b), \hspace{3mm} \bx \in \mathcal{M}.
+ p(\bx) = P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b), \hspace{3mm} \bx \in \mathcal{M}.
 \end{equation}
@@
 Using vector notation, we collect the temperature and salinity variables on the grid as
-\begin{equation}
-    \bxi=(\xi_a(\bx_1),\xi_b(\bx_1),\ldots,\xi_a(\bx_n),\xi_b(\bx_n))^t \sim \mbox{GP} (\bmu,\bSigma), \nonumber
-\end{equation}
-where the entries in length $2 n$ vector $\bmu$ hold the expected values of temperature and salinity at the grid locations, and the $2n \times 2n$ matrix $\bSigma$ holds all the covariances between temperature and salinity variables at all grid locations.
- The joint probability density function is then
-\begin{equation}\label{prior}
-\pi(\bxi) = N(\bxi;\bmu,\bSigma) = \frac{1}{(2\pi)^{\frac{n}{2}}|\bSigma|^{\frac{1}{2}}}e^{-\frac{1}{2}(\bxi -\bmu)^T\bSigma^{-1}(\bxi - \bmu)}.
-\end{equation}
+\begin{eqnarray}
+    \bxi &=& (\xi_a(\bx_1),\xi_b(\bx_1),\ldots,\xi_a(\bx_n),\xi_b(\bx_n))^t, \nonumber \\
+    \pi(\bxi) &=& N(\bmu,\bSigma) = \frac{1}{(2\pi)^{n}|\bSigma|^{\frac{1}{2}}}e^{-\frac{1}{2}(\bxi -\bmu)^T\bSigma^{-1}(\bxi - \bmu)},
+\end{eqnarray}
+where $\pi(\bxi)$ is the joint density function of $\bxi$, entries in length $2 n$ vector $\bmu$ hold the expected values of temperature and salinity at the grid locations, and the $2n \times 2n$ matrix $\bSigma$ holds all the variance-covariances of temperature and salinity variables at all grid locations.
 
@@
 \begin{equation}\label{likelihood}
-\by | \bxi \sim GP(\by; \bA \bxi,\bR),
+\pi(\by | \bxi)=N(\bA \bxi,\bR)= \frac{1}{(2\pi)^{m}|\bR|^{\frac{1}{2}}}e^{-\frac{1}{2}(\by -\bA\bxi)^T\bR^{-1}(\by-\bA\bxi)},
 \end{equation}
@@
 \begin{equation}\label{gp_hat}
+\pi 
+\left(
 \begin{bmatrix}
@@
  |\by
- \sim N \left( 
+ \right)
+ = N \left( 
 \begin{bmatrix} m_{a}(\bx)\\
@@
 The conditional mean $\bm$ will be important in the following derivation in Section \ref{sec:sur}.
-This mean is a linear function of the data $\by$, and since the marginal distribution of the data is $\by \sim N( \bA \bmu,\bA \bSigma \bA^t+\bR)$, we get
+This mean is a linear function of the data $\by$, and since the marginal distribution of the data is $\pi (\by)= N( \bA \bmu,\bA \bSigma \bA^t+\bR)$, we get
 \begin{equation}\label{distxi}
-\bm \sim N(\bmu , \bPsi), \hspace{3mm} \bPsi=\bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.  
+\pi(\bm)= N(\bmu , \bPsi), \hspace{3mm} \bPsi=\bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.  
 \end{equation}
@@
 \begin{equation}\label{dist_mxi}
-\bm(\bx) \sim N 
+\pi \left( \bm(\bx) \right) 
+= N 
 \left( 
@@
 \begin{eqnarray}\label{eq:prob_mv0}
- p(\bx) &=& P(\xi_a(\bx) < T_a, \xi_b(\bx) < T_b) \nonumber \\
+ p(\bx) &=& P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b) \nonumber \\
  &=& \Phi_2 \begin{pmatrix} 
@@
 \begin{eqnarray}\label{eq:prob_mv}
- p(\bx;\by) &=& P(\xi_a(\bx) < T_a, \xi_b(\bx) < T_b |\by) 
+ p(\bx;\by) &=& P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b |\by) 
  \nonumber \\
@@
 
-\subsection{Variance in excursion set}
+\subsection{Variance in excursion set indicators}
 
@@
 
-Consider then a measure as follows:
+Consider then an integrated measure as follows:
 \begin{equation}\label{measVAR}
@@
 The measure is dominated by location having probabilities close to $0.5$. 
+(Note that \cite{Bect2019} call this the integrated Bernoulli variance.)
 
@@
     & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by} \left\{ p(\bx;\by)\left( 1-p(\bx;\by)\right) \right\} \Delta, \nonumber \\
-    p(\bx;\by) &=& P(\xi_a(\bx) < T_a, \xi_b(\bx) < T_b |\by), \nonumber
+    p(\bx;\by) &=& P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b |\by), \nonumber
 \end{eqnarray}
@@
 This means that the high-dimensional inner integral in Eq.
-(\ref{sur}) reduces to a univariate integral \citep{bhattacharjya2013value, chevalier2014fast}.
+(\ref{sur}) reduces to a bivariate integral \citep{bhattacharjya2013value, chevalier2014fast}.
+The end result then relies on calls to the multivariate Gaussian cumulative distribution function for calculation of orthant probabilities.
 
@@
 \begin{eqnarray}
-   P(\xi_a < T_a, \xi_b < T_b|\by) &=& P \left( Z_a <\frac{T_a-m_{a}}{s_{a}}, Z_b <\frac{T_b-m_{b}}{s_{b}}|\by \right) \nonumber \\
+   P(\xi_a \leq T_a, \xi_b \leq T_b|\by) &=& P \left( Z_a \leq\frac{T_a-m_{a}}{s_{a}}, Z_b \leq\frac{T_b-m_{b}}{s_{b}}|\by \right) \nonumber \\
    &=& \Phi_2 \begin{pmatrix} 
@@
 \end{bmatrix}\end{pmatrix} \nonumber \\
-&=& P(\xi_a < T_a, \xi_b < T_b|\bm),
+&=& P(\xi_a \leq T_a, \xi_b \leq T_b|\bm),
 \end{eqnarray}
@@
 \begin{eqnarray}\label{eq:var2}
-E_{\by}(p[1-p]) &=& E_{\bm}(p[1-p]), \hspace{3mm} p=P(\xi_a < T_a, \xi_b < T_b|\bm) \nonumber \\
+E_{\by}(p[1-p]) &=& E_{\bm}(p[1-p]), \hspace{3mm} p=P(\xi_a \leq T_a, \xi_b \leq T_b|\bm) \nonumber \\
 E_{\bm}(p[1-p]) & = & \int p[1-p] p(\bm) d\bm, \nonumber \\
- &=& \int P(\xi_a < T_a, \xi_b < T_b|\bm)  p(\bm) d\bm \nonumber  \\
-&-& \int P(\xi_a < T_a, \xi_b < T_b|\bm) P(\xi_a < T_a, \xi_b < T_b|\bm) p(\bm) d\bm, 
+ &=& \int P(\xi_a \leq T_a, \xi_b \leq T_b|\bm)  p(\bm) d\bm \nonumber  \\
+&-& \int P(\xi_a \leq T_a, \xi_b \leq T_b|\bm) P(\xi_a \leq T_a, \xi_b \leq T_b|\bm) p(\bm) d\bm, 
 \end{eqnarray}
 where the integral is split in two components.
-We will next extend results of \cite{chevalier2014fast} to solve for these two parts.
+We will next extend results of \cite{chevalier2014fast} to solve for these two parts (see also \cite{stroh} who mentioned the possibilities of such extensions).
 
@@
 \begin{equation}\label{part1:phi2}
- \int P(\xi_a < T_a, \xi_b < T_b|\bm) p(\bm) d\bm= 
-P \left( Z_{a} < \frac{T_a-m_{a}}{s_{a}}, 
-Z_{b} < \frac{T_b-m_{b}}{s_{b}} \right), \nonumber
+ \int P(\xi_a \leq T_a, \xi_b \leq T_b|\bm) p(\bm) d\bm= 
+P \left( Z_{a} \leq \frac{T_a-m_{a}}{s_{a}}, 
+Z_{b} \leq \frac{T_b-m_{b}}{s_{b}} \right), \nonumber
 \end{equation}
@@
 \begin{eqnarray}\label{two_parts0}
-& P & \left( Z_{a} < \frac{T_a-m_{a}}{s_{a}}, 
-Z_{b} < \frac{T_b-m_{b}}{s_{b}} \right) \\
+& P & \left( Z_{a} \leq \frac{T_a-m_{a}}{s_{a}}, 
+Z_{b} \leq \frac{T_b-m_{b}}{s_{b}} \right) \\
 &=& \Phi_2 \begin{pmatrix} 
@@
 \begin{eqnarray}\label{part1:phi4}
-&& \int P(\xi_a < T_a, \xi_b < T_b|\bm) P(\xi_a < T_a, \xi_b < T_b|\bm) p(\bm) d\bm_{\xi}= \nonumber \\
-&& P\left( Z_{1,a} < \frac{T_a-m_{a}}{s_{a}}, 
-Z_{1,b} < \frac{T_b-m_{b}}{s_{b}},Z_{2,b} < \frac{T_a-m_{a}}{s_{a}}, 
-Z_{2,b} < \frac{T_b-m_{b}}{s_{b}} \right). \nonumber
+&& \int P(\xi_a \leq T_a, \xi_b \leq T_b|\bm) P(\xi_a \leq T_a, \xi_b \leq T_b|\bm) p(\bm) d\bm_{\xi}= \nonumber \\
+&& P\left( Z_{1,a} \leq \frac{T_a-m_{a}}{s_{a}}, 
+Z_{1,b} \leq \frac{T_b-m_{b}}{s_{b}},Z_{2,b} \leq \frac{T_a-m_{a}}{s_{a}}, 
+Z_{2,b} \leq \frac{T_b-m_{b}}{s_{b}} \right). \nonumber
 \end{eqnarray}
@@
 \begin{eqnarray}\label{part1:phi4}
-&& P \left( Z_{1,a} < \frac{T_a-m_{a}}{s_{a}}, 
-Z_{1,b} < \frac{T_b-m_{b}}{s_{b}},Z_{2,b} < \frac{T_a-m_{a}}{s_{a}}, 
-Z_{2,b} < \frac{T_b-m_{b}}{s_{b}} \right)  \\
+&& P \left( Z_{1,a} \leq \frac{T_a-m_{a}}{s_{a}}, 
+Z_{1,b} \leq \frac{T_b-m_{b}}{s_{b}},Z_{2,b} \leq \frac{T_a-m_{a}}{s_{a}}, 
+Z_{2,b} \leq \frac{T_b-m_{b}}{s_{b}} \right)  \\
 &=&  \Phi_4 
@@
 \begin{equation}\label{two_partsK}
-E_y(p[1-p]) = \int p[1-p] p(\by) d\by, \hspace{3mm} P(\xi_k < T_k ; k=1,\ldots,K|\by). 
+E_y(p[1-p]) = \int p[1-p] p(\by) d\by, \hspace{3mm} P(\xi_k \leq T_k ; k=1,\ldots,K|\by). 
 \end{equation}
@@
 %\begin{equation}
-%E(\min \{ p,(1-p)\})=\int \min P(\xi_a < t, \xi_b < s),[1-P(\xi_a < t, \xi_b < s)] p(E(\bxi)=) dE(\bxi)=,
+%E(\min \{ p,(1-p)\})=\int \min P(\xi_a \leq t, \xi_b \leq s),[1-P(\xi_a \leq t, \xi_b \leq s)] p(E(\bxi)=) dE(\bxi)=,
 %\end{equation}
@@
 V_{m,\mbox{upd}} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j,s}|\mathcal{Y}^{s-1}} \left\{ p(\bx;\mathcal{Y}^{j,s})\left( 1-p(\bx;\mathcal{Y}^{j,s})\right) \right\} \Delta \nonumber \\
-    p(\bx;\mathcal{Y}^{j,s}) &=& P(\xi_a(\bx) < T_a, \xi_b(\bx) < T_b |\by^{j,s},\mathcal{Y}^{s-1}). \nonumber
+    p(\bx;\mathcal{Y}^{j,s}) &=& P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b |\by^{j,s},\mathcal{Y}^{s-1}). \nonumber
 \end{eqnarray}
@@
 V_{la,\mbox{upd},j_2} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j_2,s+1}|\mathcal{Y}^{j,s}} \left\{ p(\bx;\mathcal{Y}^{j,s+1})\left( 1-p(\bx;\mathcal{Y}^{j,s+1})\right) \right\} \Delta \nonumber \\
-    p(\bx;\mathcal{Y}^{j,s+1}) &=& P(\xi_a(\bx) < T_a, \xi_b(\bx) < T_b |\by^{j_2,s+1},\mathcal{Y}^{j,s}). \nonumber
+    p(\bx;\mathcal{Y}^{j,s+1}) &=& P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b |\by^{j_2,s+1},\mathcal{Y}^{j,s}). \nonumber
 \end{eqnarray}
@@
 
-We use a bivariate GP model for temperature and salinity, with prior mean specified by
+We use a bivariate GP model for temperature and salinity, where we specify the prior mean 
 \begin{equation}\label{m}
@@
 \end{equation}
-We expect the eastern part of the domain to be cold freshwater.
-The situation mimics that of a river mouth entering from the south, and the water masses are pulled to the east.
-This is imposed by setting slopes to $\beta_{a,1}=0.065$ and $\beta_{b,1}=0.1$.
+In this simulation setup, which is motivated by the real case, we expect the eastern part of the domain to be cold freshwater.
+This situation mimics that of a river mouth entering from the south, and the water masses are pulled to the east.
+The actual trends are in practiece set from preliminary information or runs of oceanographic models solving the differential equations for fluid flow.
+In this example the mean is specified by $\beta_{a,1}=0.065$ and $\beta_{b,1}=0.1$.
 Further, the intercepts of the temperature and salinity variables are set to $\beta_{a,0}=5.8$ and $\beta_{b,1}=29.0$.

### 2019-05-29 main_new.tex
updates

@@
 
-The excursions could also be defined as larger than threshold, or within a boundary of limits for the two variables. Calculations and interpretations will be similar in these different cases.
+ The excursions could also be defined as larger than a threshold, or
+ within a boundary of limits for the two variables. Calculations and
+ interpretations will be similar in these different cases.
 
@@
 \end{equation}
-The actual calculation of the EP would require a model specification. As we demonstrate below, Gaussian modeling assumptions facilitate the computation. 
-If the EPs in Eq.
-(\ref{eq:prob}) are close to $1$ or $0$, it is easy to classify water masses.
-It is more difficult if the probabilities are close to $0.5$.
-In this case it will likely be worthwhile gathering data to improve the classification skill.
-
-There has lately been much statistical literature on ESs and EPs connected to spatial statistics \citep{french2013spatio,bolin2015excursion,french2016credible}. 
-Most of the work done is based on GPs for the spatial phenomenon, as we use here.
-Our focus in the current paper is on ES as defined in Eq. (\ref{ES}) and location-wise EPs and the uncertainty reduction that can be achieved by informative sampling. 
-In this sense our work is similar to \cite{chevalier2014fast} and \cite{azzimonti2016quantifying} who described analytical results for uncertainty reduction in ESs for univariate processes.
-One contribution of the current paper is to derive closed-form results for expectations of ESs in the situation with multivariate processes.
-
-\section{Gaussian processes and excursion probabilities}\label{sec:GP_EP}
+
+The actual calculation of the EP would require a model
+specification. As we demonstrate below, Gaussian modeling assumptions
+facilitate the computation.  If the EPs in Eq. (\ref{eq:prob}) are
+close to $1$ or $0$, it is easy to classify water masses; it is
+challenging if the probabilities are close to $0.5$. In such a case it
+would be worthwhile to gather data to improve classification skill.
+
+Recent work in statistical literature on ESs and EPs connected to
+spatial statistics includes
+\citep{french2013spatio,bolin2015excursion,french2016credible}.  Most
+are based on GPs for spatial phenomena, as in our work.  Our own focus
+is on ES as defined in Eq. (\ref{ES}) and location-wise EPs and the
+uncertainty reduction achieved by informative sampling.  In this sense
+our work is similar to \cite{chevalier2014fast} and
+\cite{azzimonti2016quantifying} who describe analytical results for
+uncertainty reduction in ESs for univariate processes.  A major
+contribution of our work is to derive closed-form results for
+expectations of ESs in the situation with multivariate processes.
+
+\section{Gaussian processes and excursion probabilities}
+\label{sec:GP_EP}
 
@@
 
-We represent the bivariate random field of ocean temperature and salinity by a Gaussian model.
-At location $\bx \in \mathcal{M}$, the bivariate distribution is 
+We represent the bivariate random field of ocean temperature and
+salinity by a Gaussian model.  At location $\bx \in \mathcal{M}$, the
+bivariate distribution is
+
 \begin{equation}\label{gp_bar}

### 2019-05-30 main_new.tex
updates

@@
 \end{equation}
-Here, the means and standard deviations are allowed to vary with location, which can be important for capturing the trends and variabilities near a river mouth.
-The correlation in salinity and temperature is assumed to be the same for all locations.
-
-There exist several multivariate spatial covariance functions \citep{gneiting2010matern,genton2015cross}.
-But in practice it has shown difficult to estimate all parameters from data.
-Like is often done, we assume the same correlation decay for salinity and temperature, which is not unrealistic for a river plume.
-We use a separable covariance function such that
+
+\kcomment{Explain notion, even if this is a for a Statistics journal}
+Here, the mean and standard deviation are allowed to vary with
+location, which is important for capturing the variability such as
+near a river mouth even as the correlation between salinity and
+temperature is assumed to be uniform across all locations. While there
+exist several multivariate spatial covariance functions
+\citep{gneiting2010matern,genton2015cross}, in practice it has been
+shown to be difficult to estimate all parameters from data. Typically
+we assume the same correlation decay for salinity and temperature,
+which is not unrealistic for a river plume. As a result, We use a
+separable covariance function such that:
+
 \begin{eqnarray}\label{gp_corr}
@@
 \end{eqnarray}
-Moreover, we assume a stationary isotropic correlation function $\rho(\bx,\bx')=\rho(h)$, for distance $h=\sqrt{(\bx-\bx')^t (\bx-\bx')}$.
-
-For implementation on the computer, we discretize the spatial domain $\mathcal{M}$ on a set of $n$ grid locations; $\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$.
-We let $\Delta$ be the area of each cell in the regular grid.
-Using vector notation, we collect the temperature and salinity variables on the grid as
+
+\kcomment{Ditto above; presumably x' is location separate from where an
+  earlier measurement was taken?}
+
+Moreover, we assume a stationary isotropic correlation function:
+
+$\rho(\bx,\bx')=\rho(h)$ where h is the distance between $x$ and $x'$ 
+($h=\sqrt{(\bx-\bx')^t (\bx-\bx')}$).
+
+For implementation we discretize the spatial domain $\mathcal{M}$ on a
+set of $n$ grid locations; $\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$.
+We let $\Delta$ be the area of each cell in the regular grid.  Using
+vector notation, we collect the temperature and salinity variables on
+the grid as:
+
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $\pi(\bxi)$ is the joint density function of $\bxi$, entries in length $2 n$ vector $\bmu$ hold the expected values of temperature and salinity at the grid locations, and the $2n \times 2n$ matrix $\bSigma$ holds all the variance-covariances of temperature and salinity variables at all grid locations.
+
+where $\pi(\bxi)$ \kcomment{suggest you try a different (e.g. upper
+  case) notation for $\pi$ given it is being used in different ways
+  above} is the joint density function of $\bxi$, entries in length
+$2 n$ vector $\bmu$ hold the expected values of temperature and
+salinity at the grid locations, and the $2n \times 2n$ matrix
+$\bSigma$ holds all the variance-covariances of temperature and
+salinity variables at all grid locations.
 

### 2019-05-30 main_new.tex
updates

@@
 
-The GP mean and covariance in Eq. (\ref{prior}) are updated when more information gets available. 
-In our setting, the autonomous vehicle will collect a batch of temperature and salinity measurements along a survey path. After each batch, the data are assimilated to get an updated model.
-We here describe one round of updating. In Section \ref{sec:heuristics} we discuss sequential data gathering and model updating.
-
-Denote the AUV data at $m$ locations by $\by=(y_{a,1},y_{b,1},\ldots,y_{a,m},y_{b,m})^t$, representing the temperature and salinity measurements gathered along a survey line.
-Restricting attention to the grid representation, the likelihood model is described by 
+The GP mean and covariance in Eq. (\ref{prior} \kcomment{citation
+  incorrect}) are updated when more information is available. In our
+setting, the autonomous vehicle will collect a batch \kcomment{what is
+  a 'batch' be explicit or use another term; 'iteration'?} of
+temperature and salinity measurements along a survey path; after each
+batch, the data is assimilated to get an updated model. We describe
+one round of updating \kcomment{where? In this section?}. In Section
+\ref{sec:heuristics} we discuss sequential data gathering and the
+model update process.
+
+We denote the AUV data at $m$ locations by
+$\by=(y_{a,1},y_{b,1},\ldots,y_{a,m},y_{b,m})^t$, representing the
+temperature and salinity measurements gathered along a survey
+transect. In the grid representation, the likelihood model is
+described by:
+
 \begin{equation}\label{likelihood}
@@
 \end{equation}
-where $\bA$ is an $2m \times 2n$ matrix holding $1$ entries at the measurement indexes on the grid, and otherwise $0$ entries.
-The size $2m \times 2m$ covariance matrix $\bR$ is diagonal with entries $r_a$ and $r_b$ which are the measurement error variances of temperature and salinity observations.
-
-The updated distribution for temperature and salinity variables on the spatial grid, given survey data $\by$, is Gaussian with mean and covariance
+
+where $\bA$ is an $2m \times 2n$ matrix holding '$1$'s on the
+measurement indexes on the grid, otherwise '$0$'. The size
+$2m \times 2m$ covariance matrix $\bR$ has diagonal entries $r_a$ and
+$r_b$ which are measurement error variances of temperature and
+salinity observations.
+
+The updated distribution for temperature and salinity variables on the
+spatial grid, given survey data $\by$, is Gaussian with mean and
+covariance
+
 \begin{eqnarray}\label{gp_upd}
@@
 \end{eqnarray}
-The updated bivariate distribution at a grid location $\bx$ is then 
+
+The updated bivariate distribution at grid location $\bx$ is then
+
 \begin{equation}\label{gp_hat}
@@
 \end{equation}
-with entries extracted from the conditional mean and covariance expressions in Eq. (\ref{gp_upd}).
-
-The conditional mean $\bm$ will be important in the following derivation in Section \ref{sec:sur}.
-This mean is a linear function of the data $\by$, and since the marginal distribution of the data is $\pi (\by)= N( \bA \bmu,\bA \bSigma \bA^t+\bR)$, we get
-\begin{equation}\label{distxi}
-\pi(\bm)= N(\bmu , \bPsi), \hspace{3mm} \bPsi=\bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.  
-\end{equation}
-When considering only grid location $\bx$, we can extract elements from the mean vector and covariance matrix to get the bivariate distribution as follows 
-\begin{equation}\label{dist_mxi}
-\pi \left( \bm(\bx) \right) 
-= N 
-\left( 
-\begin{bmatrix}
+
+with entries extracted from the conditional mean and covariance
+expressions in Eq. (\ref{gp_upd}).
+
+% The conditional mean $\bm$ will be important in the following
+% derivation in Section \ref{sec:sur}.
+This conditional mean is a linear function of the data $\by$, and
+since the marginal distribution of the data is
+$\pi (\by)= N( \bA \bmu,\bA \bSigma \bA^t+\bR)$, we get
+\begin{equation}\label{distxi} \pi(\bm)= N(\bmu , \bPsi), \hspace{3mm}
+  \bPsi=\bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.
+\end{equation} When considering only grid location $\bx$, we can
+extract elements from the mean vector and covariance matrix to get the
+bivariate distribution as follows \begin{equation}\label{dist_mxi} \pi
+  \left( \bm(\bx) \right) = N \left( \begin{bmatrix}
       \mu_{a}(\bx) \\
-\mu_{b}(\bx)
-\end{bmatrix},
-\begin{bmatrix}
+      \mu_{b}(\bx) \end{bmatrix}, \begin{bmatrix}
       \psi^2_{a}(\bx) & \psi_{a,b}(\bx)\\
-\psi_{a,b}(\bx) & \psi^2_{b}(\bx)
-\end{bmatrix}
-\right).
+      \psi_{a,b}(\bx) & \psi^2_{b}(\bx) \end{bmatrix} \right).
 \end{equation}

### 2019-05-30 main_new.tex
updates

@@
 
-Based on the Gaussian modeling assumptions, the EPs in Eq. (\ref{eq:prob}) can be computed at any location $\bx$. Without additional data, we have
+Based on the Gaussian modeling assumptions, the EPs in Eq.
+(\ref{eq:prob}) can be computed at any location $\bx$. Without
+additional data, we have:
+
 \begin{eqnarray}\label{eq:prob_mv0}
@@
 \end{eqnarray}
-where $\Phi_2$ is the bivariate Gaussian cumulative distribution function, and the mean and covariance entries are defined in Eq. (\ref{gp_bar}).
-
-When data $\by$ is available, we similarly get
+
+where $\Phi_2$ is the bivariate Gaussian cumulative distribution
+function, and the mean and covariance entries are as defined in Eq.
+(\ref{gp_bar}).
+
+When $\by$ is available, we similarly get:
+
 \begin{eqnarray}\label{eq:prob_mv}
@@
 \end{eqnarray}
-with parameters defined in Eq. (\ref{gp_upd}) and (\ref{gp_hat}).
-
-For one particular dataset, the probability $p(\bx;\by)$ can be closer to $0.5$ than the initial probability $p(\bx)$.
-This means that the variance in the ES is increased.
-On average, taking expectations over the data $\by$, the ES variance will be reduced by gathering data.
-We next explain how AUVs can be used to efficiently gather data and reduce the overall marginal spatial uncertainties in the ES.
+
+with parameters as defined in Eq. (\ref{gp_upd}) and (\ref{gp_hat}).
+
+For one particular dataset \kcomment{This is unclear, rephrase}, the
+probability $p(\bx;\by)$ can be closer to $0.5$ than the initial
+probability $p(\bx)$. This means that the variance in the ES is
+increased. On average, taking expectations over the data $\by$, the ES
+variance will be reduced by gathering data. We next explain how AUVs
+can be used to efficiently gather data and reduce the overall marginal
+spatial uncertainties in the ES.
 

### 2019-06-01 main_new.tex
updates

@@
 
-\section{Uncertainty reduction on excursion sets}\label{sec:sur}
+\section{Uncertainty reduction on excursion sets}
+\label{sec:sur}
 
@@
 
-Regions with temperatures or salinity that surpasses the thresholds with a large margin (e.g.
-very high/low temperatures) are not interesting, unless the uncertainty at these locations is very large.
-The objective will therefore be to explore regions where the process variables are close to the thresholds.
-With this mind, it is valuable to explore locations $\bx$ where $p(\bx) \approx 0.5$.
-This is what gives the maximum variance $p(\bx) (1-p(\bx))$ in the indicator for the ES.
+Regions with high variability in temperature or salinity % that surpasses the thresholds
+% with a large margin (e.g. very high/low temperatures)
+are not interesting, unless the uncertainty at these locations is very
+large \kcomment{Huh? Not sure I understand. Are you saying the
+  \emph{value} of a parameter if very large/small it is of less
+  interest? Why?}. The objective is to explore regions where the
+process variables are close to the thresholds \kcomment{You want to
+  justify this some more
+  -- why is it so? You can't make a categorical statement and leave it
+  at that; the reviewers will want to know.}. With this mind, it is
+valuable to explore locations $\bx$ where $p(\bx) \approx 0.5$. This
+is what gives the maximum variance $p(\bx) (1-p(\bx))$ in the
+indicator for the ES. \kcomment{This para is poorly structured. You
+  should make a positive case for what you want to measure, justify
+  and leave it as it is. Starting with what you \textbf{don't} want
+  and then stating the positive is less convincing.}
 
@@
 \end{equation}
-The measure is dominated by location having probabilities close to $0.5$. 
-(Note that \cite{Bect2019} call this the integrated Bernoulli variance.)
-
-Considering one location only, we study the effect of different correlation $\gamma$ and prior variance. 
-Figure \ref{illus_bivarDens} shows contour plots of three different densities.
-Here, the temperature mean is $T_a=5^o$ and salinity mean is $T_b=30$ mg/l.
-The thresholds are identical to these means.
-The variables have unit variance and different correlation (low, medium and high).
-\begin{figure}[h!]
-\centering
+
+The measure is dominated by locations with probabilities \kcomment{Of
+  what?} close to $0.5$, or the integrated Bernoulli variance
+\cite{Bect2019}.
+
+Considering a single location, we explore the effect of different
+correlation ($\gamma$) and prior variance. Fig. \ref{illus_bivarDens}
+shows contour plots of three different densities where, the mean
+temperature $T_a=5^o$ and mean salinity $T_b=30$ mg/l. The thresholds
+are identical to these means. The variables have unit variance and
+different correlation (low, medium and high). \kcomment{I don't
+  understand these figures. Where did you get the three different
+  correlations from? Add a note about how they were constructed
+  and what is on the X/Y axis. Are the contours based on data from the
+  table below?}
+
+\begin{figure}[h!] \centering
   \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
-\caption{Density contour plots having different correlations for temperature and salinity.}\label{illus_bivarDens}
+  \caption{Density contour plots having different correlations for
+    temperature and salinity.}\label{illus_bivarDens} 
 \end{figure}
-Table \ref{tab:sim_rhoab} shows the EPs and the associated variances in the ES (top rows).
-The results are computed for different correlations $\gamma$ indicated in Figure \ref{illus_bivarDens}, and for standard deviations $1$ (left) and $2$ (right).
-The variance in the ES is largest for high correlation, as there is more tendency of joint small or high occurrences for temperature and salinity in this case.
-EPs are the same for standard deviation $1$ or $2$.
-\begin{table}[!h]
-\centering
-\caption{EP and variance of ES for different correlations (top rows), and for data of both variables and only for temperature (bottom rows).}
+
+Table \ref{tab:sim_rhoab} shows the EPs and the associated variances
+in the ES (top rows). The results are computed for different
+correlations $\gamma$, indicated in Fig. \ref{illus_bivarDens}, and
+for standard deviations $1$ (left) and $2$ (right). The variance in
+the ES is largest for high correlation, as there is more tendency of
+joint small or high occurrences \kcomment{What is ``joint small''?}for
+temperature and salinity in this case. EPs are the same for standard
+deviation $1$ or $2$. \kcomment{Reformulate; this entire para is
+  unclear and its connection to Fig. 3 above, is tenuous and vague. If
+  there is a connection, then make it explicit.}
+
+\begin{table}[!h] \centering \caption{EP and variance of ES for
+    different correlations (top rows), and for data of both variables
+    and only for temperature (bottom rows).}
   \begin{tabular}{c|ccc|ccc}

### 2019-08-25 main_new.tex
Updates - synced Kanna and overleaf
Updates from Overleaf

--- 2019-05-30 main_new.tex
@@
 \newcommand{\kcomment}[1]{{\color{blue}{\{KR: #1\}}}}
+\newcommand{\tofcomment}[1]{{\color{green}{\{TF: #1\}}}}
 
@@
 \begin{abstract}
-  We examine experiment design for informative sampling for
-  characterizing oceanographic phenomena. Methods are developed
-  for efficient spatial sampling using autonomous robotic platforms,
-  where the goal is to map the regions of cold and fresh water near a
-  river mouth. Specifically, we define a design criterion based on
-  reducing uncertainty in excursion sets of multivariate
-  random fields. We develop analytical expressions for the expected
-  variance reduction in excursion set indicators, and show how these can be
-  applicable for static designs and sequential designs building on
-  myopic and look-ahead planning. We use simulations to study the
-  properties of the suggested methods. Results from field
-  deployments with an autonomous underwater vehicle are also presented.
+Improving and optimizing oceanographic sampling is a crucial task for modern marine management and science faced with limited resources and highly dynamic processes.
+The combination of statistics and autonomous robotics provides opportunities for studying aspects related to experimental designs for effective characterization of oceanographic phenomena.
+In this paper methods are developed for efficient spatial sampling applied to the mapping of dynamic coastal processes by providing informative descriptions of spatial characteristics of ocean phenomena.
+Specifically, we define a design criterion based on reducing uncertainty in the excursion set of multivariate random fields, and derive analytical expressions for the expected Bernoulli variance reduction in the excursion set of spatial temperature and salinity variables.
+We show how these can be applicable in static designs and sequential designs, building on myopic and look-ahead sampling strategies.
+We use simulations to study the properties and differences between state-of-the-art approaches.
+Results from field deployments with an autonomous underwater vehicle are also presented, for a case study mapping the boundary of a river plume.
+The results demonstrate the ability of the method to effectively inform and execute data-driven environmental sampling.
 
@@
 
-Monitoring the worlds oceans has gained increased importance in the
-light of the changing climate. While the focus has often been on
-maga-fauna and the anthropogenic impact from micro-plastics to
-pollution, biological oceanographers have focused intently on studying
-micro-organisms at the base of the human food web. These organisms are
-critically impacted by the changing dynamics in the upper water-column,
-especially in coastal zones which are complex and often hard to observe
-in space and time. By studying the bio-geochemical processes in the
-upper water-column scientists can measure the impact of change, natural
-or anthropomorphic, and provide an informed opinion to policy makers to
-effect changes in preserving the environment. However, the challenge of 
+Monitoring the worlds oceans has gained increased importance in the light of the changing climate and increasing anthropogenic impact.
+A central problem for understanding these factors is the lack of representative data with sufficient resolution.
+Most of this \emph{undersampling} can be contributed to the large spatio-temporal variation on which ocean processes transpire, prompting the need for effective means of sampling.
+By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use of enabling autonomous platforms can be combined with statistical methods to pursuit measurements with high scientific relevance.
+
+Traditional data collection at sea has typically been based on static buoys, lagrangian floats or ship-centric methods, with significant logistical limitations that directly impact coverage and sampling resolution.
+Modern methods using satellite remote-sensing provide large-scale coverage, but have limited resolution, are limited to the sensing the surface of the ocean, and is impacted by cloud cover.
+However, the advent of robust mobile robotic platforms \cite{Bellingham07} are making significant contributions to environmental monitoring and sampling.
+In particular autonomous underwater vehicles (AUVs), have made significant strides in advancing the state of sampling and consequently have made robotics an integral part of ocean observation in large part driven by using statistical methods in design and data assimilation \citep{das11b,Graham2013,Das2015,das15,fossuminformation,fossum18b}.
+Others include \cite{wikle2013modern} who focus on hierarchical statistical models; \cite{sahu2008space} study spatio-temporal models for sea surface temperature and salinity data, while \cite{mellucci2018oceanic} look at the statistical prediction of features using an underwater glider.
+Common to these approaches is the use of information-based methods to infer environmental sampling and a few important characteristics should be mentioned, namely:
+
+\begin{itemize}
+    \item Decisions on where to sample happen sequentially, meaning recent information can be assimilated and used to improve statistical models and reduce uncertainty associated with the  choices. 
+    \item As a priori information about the state of the ocean is very uncertain, there lies a substantial efficiency gain in reacting to information obtained from measurements taken \emph{in-situ}, this is usually referred to as the \emph{adaptivity gap}.
+    \item For this type of sampling problem, the number of choices (i.e.
+locations, paths, and candidate designs) makes the problem combinatorially large, creating a trade-off between optimization (finding the optimal design) and computability (arriving at a solution in reasonable time), which has to be considered in practice.
+\end{itemize}{}
+
+The combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based sensing.
+Central to this is the ability to model spatially correlated variables and provide formal measures of uncertainty.
+By far the most practical models use Gaussian processes (GPs), which will also be used in our formulation, in large  part because this approach allows efficient implementation and evaluation, enabling prioritization of sampling efforts to be formalized into decision making systems on-board a robotic platform.
+
+%Sampling can, in this context, not simply be distributed evenly ---along simple transects or ``lawn-mover" patterns--- but must instead be prioritized to relevant regions to ensure it is cost-effective while providing adequate coverage and resolution of the area of scientific interest.
+
+%While the focus has often been on
+%biological and anthropogenic impact from micro-plastics to
+%pollution, biological oceanographers have focused intently on studying
+%micro-organisms at the base of the human food web. These organisms are
+%critically impacted by the changing dynamics in the upper water-column,
+%especially in coastal zones which are complex and often hard to observe
+%in space and time. By studying the bio-geochemical processes in the
+%upper water-column scientists can measure the impact of change, natural
+%or anthropomorphic, and provide an informed opinion to policy makers to
+%effect changes in preserving the environment. However, the challenge of 
 % The pressure on marine resources is growing and increased accuracy,
 % resolution, and persistent monitoring of the oceans is crucial for
-% long-term sustainable management. A central problem for this type of
-% environmental 
-sensing in the small% in the large spatial extent
-while studying the large spatio-temporal variation inherent to ocean
-processes, prompts the need for effective means of sampling.
-% allocation of sampling resources.
-%\begin{em}
-Sampling can, in this context, not simply be distributed evenly
----along simple transects, grids, or ``lawn-mover" patterns--- but must
-instead be prioritized to relevant regions to ensure it is
-cost-effective while providing adequate coverage of the area of
-scientific interest.
-% save resources and avoid
-% redundant data.
-%\end{em}
-
-Traditional data collection at sea has typically been based on static
-buoys, lagrangian floats or ship-centric methods, with significant
-logistical limitations that directly impact coverage and sampling
-resolution. Modern methods using satellite remote-sensing provide
-large-scale coverage, but have limited resolution, are limited to the
-sensing the surface of the ocean and impacted by cloud cover. With the
-advent of robust mobile robotic platforms \cite{Bellingham07}, more
-effective and higher resolution methods are making significant
-contributions to environmental monitoring and sampling. 
+% long-term sustainable management.
+
 % A
@@
 % platforms, especially
-In particular Autonomous underwater vehicles (AUVs), have made
-significant strides in advancing the state of sampling
+
 % provided means to execute this prioritization through the capacity of
@@
 % can be optimized. These capabilities have made
-and consequently have made robotics an integral part of ocean
-observation in large part driven by using 
+
 % of the emerging sensing practice for ocean science, allowing scientists
@@
 % answer to this question relates to
-statistical methods in design and data assimilation
-\cite{das11b,Graham2013,Das2015,das15,fossuminformation,fossum18b}.% , and
+
+% , and
 % the application domain is clearly an arena where statisticians can
@@
 % There has recently been some statistical attention in oceanography:
-Others include \cite{wikle2013modern} who % present an overview of approaches for modeling
-% uncertainties in the ocean sciences with
-focus on hierarchical statistical models; \cite{sahu2008space} study
-spatio-temporal models for sea surface temperature and salinity data,
-while
-% which will also are our key variables of interest in the current paper.
-\cite{mellucci2018oceanic} look at the statistical prediction of
-features using an underwater glider.
+
 %From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest.
@@
 Quantification of gradient features is therefore a much needed competence in robotic sampling.
-The combination of statistical tools and robotic platforms is a natural
-symbiosis which enables information-based sensing. Central to this is
-the ability to model spatially correlated variables and provide formal
-measures of uncertainty. By far the most practical models use Gaussian
-processes (GPs), which will also be used in our formulation, in large
-part because this approach allows efficient implementation and evaluation, enabling prioritization of
-sampling efforts to be formalized into decision making systems on-board
-a robotic platform.
+
 %The imperfect understanding of causality coupled with the underlying fundamental uncertainty of the ocean implies that we need to obtain more measurements and use statistical models and methods to determine what and how phenomena form, to increase the quality of predictive statements.
 
-In this work, we study excursion sets (ES) and excursion probabilities
-(EP) for spatial GPs, and the goal is to design sampling strategies for
-reducing the variance in the ES indicators. The main contributions involve: i)
-extending ES-based uncertainty criteria to multivariate random fields,
-allowing the evaluation to include multiple environmental parameters
-such as temperature and salinity, ii) develop methods for informative
-environmental sensing and autonomous data collection, and iii)
-demonstrating the approach in a real world scenario of sampling a river
-plume using an AUV.
-
-The remainder of the paper is organized as follows: Section \ref{sec:bg}
-provides some background on ocean sampling, as also some notation for
-ESs and EPs. Section \ref{sec:GP_EP} sets up the required assumptions of
-GPs, including bi-variate random fields for temperature and salinity.
-Section \ref{sec:sur} formulates the choice of sampling design criteria for
-variance reduction in ES indicators, with analytical derivation for multivariate
-spatial processes. Section \ref{sec:heuristics} builds on these sampling
-criteria to present heuristics for adaptive sampling. Section
-\ref{sec:simulations} shows results and discusses properties of the
-methods in simulation while Section \ref{sec:case_study} instantiates
-the work in the form of characterizing a river plume. We conclude with
-Section \ref{sec:concl_disc} with a summary and a discussion of future
-work.
+In this work, we build on excursion sets (ES) and excursion probabilities (EP) for spatial design of experiments.
+The goal is to design sampling strategies for reducing the spatially integrated Bernoulli variance (IBV) in the ES indicators.
+The main contributions involve: i) extending ES-based uncertainty criteria to multivariate GPs, allowing the evaluation to include multiple environmental parameters such as temperature and salinity, ii) developing methods for informative environmental sensing and autonomous data collection, and iii) demonstrating the approach in a real world scenario of sampling a river plume using an AUV.
+
+The remainder of the paper is organized as follows: Section \ref{sec:bg} provides some background on ocean sampling.
+Section \ref{sec:ESEP} defines ESs and EPs, and the design criteria of IBV.
+Section \ref{sec:GP_EP} sets up the required statistical modeling assumptions of the bi-variate GPs for temperature and salinity.
+Section \ref{sec:sur} builds on these when deriving the analytical derivation of the design criteria.
+Section \ref{sec:heuristics} extends results to present heuristics for adaptive sampling.
+Section \ref{sec:simulations} shows results and discusses properties of the methods in simulation studies.
+Section \ref{sec:case_study} demonstrates the methodology in field work characterizing a river plume.
+Section \ref{sec:concl_disc} contains summary and a discussion of future work.
 
@@
 
-\subsection{Observing the Upper Ocean Column}
+\subsection{Observing Processes in the Upper Ocean}
 
@@
 % ecosystem response.
-Central to understanding the changes taking place in the upper
-water-column is the understanding of biophysical interaction (see Fig.
-\ref{fig:envir2}) driven by an agglomeration of physical forcing (wind,
-topography, bathymetry, tidal influences and local geography such as
-fresh water inflows) mixed with incipient micro-biology driven by
-planktonic and coastal anthropogenic input, such as pollution and
-agricultural runoff, which results in a range of ecosystem phenomena
-(e.g. blooms, anoxic zones, plumes) ---with significant impact to
-society. Renowned oceanographer Walter Munk called the
-20\textsuperscript{th} century, the \emph{century of undersampling}
-\citep{munk2002}, pointing out that lack of sufficient sampling
-resolution in time and space, % owing to current observation practices.
-% The lack of resolution, 
-% arising from sparsely distributed sensing methods, is
-as a critical issue for characterizing ocean phenomena. This view has
-been driving ocean sensing practices towards more networked and
-distributed use of robotic assets in order to fill the gaps in
-our understanding (See Fig. \ref{fig:envir}).
+
+% owing to current observation practices.
+
+Central to understanding the changes taking place in the upper water-column is the understanding of biophysical interaction  driven by an agglomeration of physical forcing (wind, topography, bathymetry, tidal influences and local geography such as fresh water inflows) mixed with incipient micro-biology driven by planktonic and coastal anthropogenic input, such as pollution and agricultural runoff, which results in a range of ecosystem phenomena (e.g.
+blooms, anoxic zones, plumes) ---all influences with significant impact both directly and indirectly on society.
+Reverberating the requirement of more effective ocean sensing practices, renowned oceanographer Walter Munk referred to the 20\textsuperscript{th} century as the \emph{century of undersampling} \citep{munk2002}, pointing out that lack of sufficient sampling resolution in time and space, as a critical issue for characterizing ocean phenomena.
+Still today, decision makers struggle with limited observational efforts sometimes leading to severe environmental problems.
+The recent (May 2018) outbreak of harmful algal blooms in Northern Norway\footnote{Report from the Institute of Marine Research:https://bit.ly/2KZGJfI}, with losses exceeding 12 million tons of salmon valued at \$ 80 million (Norwegian Directorate of Fisheries), is a case in point in how adequate sampling efforts could have provided a more accurate warning assessment.
+These challenges of undersampling have been driving ocean sensing practices towards more networked and distributed use of robotic assets in order to fill the gaps in our understanding (See Fig.
+\ref{fig:envir}), where the surveying is guided by a model and the measurements made onboard the robot are used to update the model.
 
@@
   \centering 
-  \subfigure[Modern ocean sensing network, where robotic assets augment
-  ship based measurements to increase coverage of the phenomena.]{\includegraphics[width =
+  \subfigure[Modern ocean sensing network, where robotic assets augment ship based measurements to increase coverage of the phenomena.]{\includegraphics[width =
     0.48\textwidth]{Figures/envir.pdf}\label{fig:envir1}}
   \hfill
-  \subfigure[Different coastal processes directly or indirectly
-  associated with gradients and boundaries.]{\includegraphics[width =
-    0.48\textwidth]{Figures/envir_ocean.pdf}\label{fig:envir2}}
-  \caption{Ocean observation is moving away from single-ship sampling
-    towards more collaborative networked operations in order to resolve
-    the numerous processes and their interaction.} \label{fig:envir}
+  \subfigure[Data-driven sampling under the sense-plan-act paradigm.]{\includegraphics[width =
+    0.48\textwidth]{Figures/adaptive_cycle.pdf}\label{fig:sense-plan-act}}
+  \caption{Ocean observation is moving away from single-ship sampling towards more autonomous and collaborative networked operations in order to resolve the numerous processes and their interaction.} \label{fig:envir}
 \end{figure}
 
-Our focus in this work is limited to spatial characterization of a river
-plume system. Fig. \ref{fig:river_fronts} shows the survey area (the
-outlet of the Nidelva river, Trondheim, Norway) taken as cold freshwater
-enters from the river showing the separation of water masses and strong
-gradients for both temperature and salinity. Because of the local
-topography and the Coreolis force \citep{coriolis1835memoire} the cold
-fresh water tends to stay near land to the east. Depending on the river
-discharge, tidal effects, and temperature differences, this boundary in
-water masses will get distorted. Prior knowledge about the location and
-evolution of these features is therefore highly uncertain, making
-deterministic planning sub-optimal.% , and one would rather demonstrate the
-% ability of AUVs to map the plume area efficiently.
+Focusing on coupling statistics and robotic sampling, our attention is in this work towards spatial characterization of a frontal systems, and more specifically fronts created by river plume systems.
+Fig.
+\ref{fig:nidelven} shows the survey area (the outlet of the Nidelva river, Trondheim, Norway) taken as cold freshwater enters from the river creating a strong separation gradient in both temperature and salinity.
+Because of the local topography and the Coreolis force \citep{coriolis1835memoire} the cold fresh water tends to flow near land to the east.
+Depending on the river discharge, tidal effects, wind and temperature differences, this boundary in water masses will get distorted.
+Prior knowledge about the location and evolution of these features is therefore highly uncertain, making deterministic planning sub-optimal.
 
@@
 \centering 
+\includegraphics[width=0.85\textwidth]{Figures/pictures/c-updated.pdf}
+\caption{Frontal patterns - Nidelven, Trondheim, Norway.}
+\label{fig:nidelven}
+\end{figure}
+
+%\begin{figure}[!h]
+%\centering
  % \subfigure[River front - Columbia River, Astoria, Oregon, US.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/a.png}\label{fig:river1}}
  % \hfill
- \subfigure[Tidal front - Korsfjorden, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/d.png}\label{fig:river2}}
+ %\subfigure[Tidal front - Korsfjorden, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/d.png}\label{fig:river2}}
  % \hfill
@@
  % \hfill
- \subfigure[Frontal patterns - Nidelva, Trondheim, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/c.png}\label{fig:river4}}
- \caption{Two examples of frontal features associated with
-   river-ocean interaction. % (\ref{fig:river1}) The river plume front of
+ %\subfigure[Frontal patterns - Nidelven, Trondheim, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/c.png}\label{fig:river4}}
+ %\caption{Two examples of frontal features associated with
+%   river-ocean interaction. % (\ref{fig:river1}) The river plume front of
    % the Columbia River meeting the Pacific Ocean.
-   (\ref{fig:river2})  Superimposed images (over 10 minutes) of a surface tidal front in
-   Korsfjorden, Norway. % (\ref{fig:river3}) The river plume of the Rio de
+%   (\ref{fig:river2})  Superimposed images (over 10 minutes) of a surface tidal front in
+%   Korsfjorden, Norway. % (\ref{fig:river3}) The river plume of the Rio de
    % la Plata, taken by European Space Agency (ESA)'s Envisat platform and
    % MEdium Resolution Imaging Spectrometer (MERIS) sensor.
-   (\ref{fig:river4}) Aerial image of the mouth of the Nidelva river in
-   Trondheim, Norway where fresh water meets the salty fjord and creates
-   frontal patterns.} % Image courtesy: (\ref{fig:river1})
+%   (\ref{fig:river4}) Aerial image of the mouth of the Nidelva river in
+%   Trondheim, Norway where fresh water meets the salty fjord and creates
+%   frontal patterns.} % Image courtesy: (\ref{fig:river1})
    % \cite{zamon2014marine}, (\ref{fig:river3}) ESA, CC BY-SA 3.0 IGO.}
- \label{fig:river_fronts}
-\end{figure}
+% \label{fig:river_fronts}
+%\end{figure}
 
@@
 
-River plumes belong to a class of ocean processes that are local and
-usually act on smaller \emph{sub-mesoscale}, from 10 m -- 5 km where
-spatial lateral variability tends to dominate. Remote sensing and
-synthetic ocean models find it challenging to provide detailed at this
-scale \citep{Lermusiaux:2006} and the principal way to resolve for fine
-resolution is by direct observation. At larger \emph{mesoscale} ($>50$
-km), both three-dimensional space and time dynamics are important and
-can shift substantially, while in the case of river plumes one can often
-limit scope to cover only lateral spatial elements ([north, east]
-domain). Morover, for this case, time-effects can be regarded as static when
-acquiring AUV data over a short time window and limited region. However,
-the methodological framework can be extended to higher dimensional situations in a similar manner.
+River plumes belong to a class of ocean processes that are local and usually act on smaller \emph{sub-mesoscale}, from 10 m -- 5 km where spatial lateral variability tends to dominate.
+Remote sensing and synthetic ocean models find it challenging to provide detailed at this scale \citep{Lermusiaux:2006} and the principal way to resolve for fine resolution is by direct observation.
+At larger \emph{mesoscale} ($>50$ km), both three-dimensional space and time dynamics are important and can shift substantially, while in the case of river plumes one can often limit scope to cover only lateral spatial elements ([north, east] domain).
+Moreover, for this case, time-effects can be regarded as static when acquiring AUV data over a short time window and limited region.
+However, the methodological framework can be extended to higher dimensional situations in a similar manner.
 
@@
 to their speed and mobility. Traditional operations and surveys with
-AUVs are usually limited to planning samples along fixed transects or
-survey lines set by human operators, programmed as a sequential plan
-consisting of different waypoints and depths. But by using on-board
+AUVs are usually limited to planning samples along fixed transects or survey lines on a spatial grid, pre-programmed by the human operator to 
+consist of different waypoints on the grid. But by using on-board
 algorithms to continuously evaluate, update, and refine future sampling
-locations, making information gathering adaptive \citep{das11b} is
-feasible. For instance, an AUV could reconstruct or modify a survey line
+locations, one makes information gathering adaptive \citep{das11b}.
+The space of sampling opportunities is still limited to the spatial grid, but now there is much more flexibility in automatic adaption.
+For instance, an AUV could reconstruct or modify a survey line
 based on what temperatures it is measuring by using embedded methods in
@@
 environment, but rather adapts based on closing the sensing and
-actuation loop. Myopic sampling, i.e. step-wise selection of the path
+actuation loop (Fig. \ref{fig:sense-plan-act}). Myopic sampling, i.e. step-wise selection of the path
 (on the graph) which reduces the information criterion the most, has
@@
 which focus largely on reducing predictive variance or mutual
-information (entropy). Because the variance and entropy reduction are
-independent of the actual data realizations under the assumption of GP
-models, the use of data-driven (adaptive) criteria was introduced to
-include more targeted sampling of regions of scientific interest by
+information (entropy). Variance and entropy reduction are
+independent of the actual data realizations under the assumption of GP models, and the use of data-driven (adaptive) criteria was introduced to include more targeted sampling of regions of scientific interest by
 \cite{Low2009} and \cite{fossuminformation}.
+In this paper, with the focus on mapping the river plume, designs that improve the classification of water masses should be rewarded.
 
@@
 
-\subsection{Notation and excursion sets}
-
-In our setting we study two variables; temperature and salinity, which
-together characterize the phenomenon taking place in the river plume as
-they define the gradients defining the frontal zone. At location
-$\bx \in \mathcal{M} \subset \mathbf{R^2}$, we define the two process
-variables by
-$\bxi(\bx)=(\xi_a(\bx),\xi_b(\bx))$.
-%In practice there is of course time variability, but we assume this to be negligible during the AUV operation.
-Depth variability could also be important in some cases, but in the current application the with depths only a few meters this is considered irrelevant for mapping the phenomenon.
-
-% In the context illustrated by Figure \ref{fig:river4},
-Our focus is on spatial separation of cold freshwater from the saline
-waters of the fjord and to be able to classify these water masses. In
-this context, ES variability is a useful sampling criterion for guiding
-the observational effort of the AUV. A bivariate ES is defined via joint
-variables \kcomment{'being below thresholds' -- is this correct?}, at
-individual locations, for the process over $\mathcal{M}$. If $T_a$ is
-the threshold for temperature and $T_b$ the threshold for salinity,
-we use the following definition:
-
+\section{Excursion sets and Integrated Bernoulli variance}
+\label{sec:ESEP}
+
+The objective is to characterize the river plume, focusing on spatial separation of cold freshwater from the warmer saline waters of the fjord.
+ESs and EPs are useful starting points to measure the ability to classify water masses, and the goal of experimental sampling is to improve this characterization, as quantified via changes in the EPs.
+
+We consider a lateral domain, near sea surface, with locations $\bx \in \mathcal{M} \subset \mathbf{R^2}$.
+The two variables of interest, temperature and salinity, are defined as random processes; $\xi_a(\bx)$ denotes the temperature (in $^o C$) and $\xi_b(\bx)$ denotes the salinity (in $mg/l$) at location $\bx$.
+We let $t_a$ be
+a threshold for temperature and $t_b$ a threshold for salinity, the bivariate ES is then defined by
  \begin{equation}\label{ES}
-     \mbox{ES} = \{\bx \in \mathcal{M} | \xi_a(\bx) \leq T_a,\xi_b(\bx) \leq T_b\}.
- \end{equation}
-
+     \mbox{ES} = \{\bx \in \mathcal{M} | \xi_a(\bx) \leq t_a,\xi_b(\bx) \leq t_b\}.
+ \end{equation}
  The excursions could also be defined as larger than a threshold, or
@@
 
-The ES is defined by random indicator variables, and the associated EPs become 
+The ES holds random indicator variables, while the associated EPs are 
 \begin{equation}\label{eq:prob}
- p(\bx) = P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b), \hspace{3mm} \bx \in \mathcal{M}.
-\end{equation}
-
-The actual calculation of the EP would require a model
-specification. As we demonstrate below, Gaussian modeling assumptions
-facilitate the computation.  If the EPs in Eq. (\ref{eq:prob}) are
+ p(\bx) = P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b), \hspace{3mm} \bx \in \mathcal{M}.
+\end{equation}
+If the EPs in Eq. (\ref{eq:prob}) are
 close to $1$ or $0$, it is easy to classify water masses; it is
-challenging if the probabilities are close to $0.5$. In such a case it
-would be worthwhile to gather data to improve classification skill.
-
-Recent work in statistical literature on ESs and EPs connected to
-spatial statistics includes
-\citep{french2013spatio,bolin2015excursion,french2016credible}.  Most
-are based on GPs for spatial phenomena, as in our work.  Our own focus
-is on ES as defined in Eq. (\ref{ES}) and location-wise EPs and the
-uncertainty reduction achieved by informative sampling.  In this sense
-our work is similar to \cite{chevalier2014fast} and
+challenging if the probabilities are close to $0.5$. 
+The integrated Bernoulli variance (IBV), see \cite{bect2019}, is here defined as 
+\begin{equation}\label{measVAR}
+    V = \int_{\bx \in \mathcal{M}} p(\bx) \left(1-p(\bx)\right) d\bx.
+\end{equation}
+The measure is dominated by locations with EPs close to $0.5$.
+
+If prior knowledge strongly indicates cold freshwater in a part of the domain, it is easy to classify, and there is likely little value in exploring that part much.
+Similarly with strong prior knowledge about warm saltwater.
+Data can be gathered at selected locations of the domain $\mathcal{M}$, to improve classification skill.
+We denote temperature measurements by $y_{a}(\bx_i)$ and salinity measurements by $y_{b}(\bx_i)$, at a sampling location $\bx_i \in \mathcal{M}$.
+Given a set of measurements, denoted $\by$, the conditional EPs are
+\begin{equation}\label{eq:post_ep}
+ p(\bx;\by) = P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b |\by). 
+\end{equation}
+
+Our goal is to construct designs for the data gathering, that are expected to improve the mapping of the water masses. 
+When designing AUV sampling designs, it is hence natural to evaluate the expected reduction in the IBV, as provided by the survey data.
+In this way, we take the expectation with respect to the planned measurements $\by$;
+\begin{equation}\label{sur}
+    V_{\mbox{upd}} = \int_{\bx \in \mathcal{M}} E_{\by} \left\{ p(\bx;\by)\left( 1-p(\bx;\by)\right) \right\} d\bx, 
+\end{equation}
+where the probability is defined in Eq. (\ref{eq:post_ep}).
+Considering a set $\mathcal{J}$ of possible designs for an AUV survey. A criterion for the selection is
+\begin{equation}\label{crit}
+    j^* = \mbox{argmin}_{j \in \mathcal{J}} V_{\mbox{upd}}(j),
+\end{equation}
+where the criterion $V_{\mbox{upd}}(j)$ in Eq. (\ref{sur}) is computed for each of the possible designs.
+
+Recent works in statistical literature on ESs and EPs connected to
+spatial statistics include
+\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}. Our focus
+is on ES as defined in Eq. (\ref{ES}), and the
+uncertainty reduction achieved by sampling as in Eq. (\ref{sur}).  In this sense
+our work is similar to \cite{bect2012}, \cite{chevalier2014fast} and
 \cite{azzimonti2016quantifying} who describe analytical results for
-uncertainty reduction in ESs for univariate processes.  A major
-contribution of our work is to derive closed-form results for
-expectations of ESs in the situation with multivariate processes.
+uncertainty reduction in ESs for univariate processes.  
+A major contribution of our work is to derive closed-form results for this design criteria in the situation with multivariate GPs.
 
@@
 
+The actual calculations of the EPs and the IBV require a model
+specification. Here, Gaussian modeling assumptions are used, as described next. They allow closed form evaluation of the IBVs, as derived in Section \ref{sec:sur}.
+
 \subsection{Bivariate Gaussian processes}
 
-We represent the bivariate random field of ocean temperature and
-salinity by a Gaussian model.  At location $\bx \in \mathcal{M}$, the
-bivariate distribution is
-
+We represent the bivariate random fields of sea temperature and
+salinity by a Gaussian model.  At location $\bx \in \mathcal{M}$, the bivariate distribution is
 \begin{equation}\label{gp_bar}
@@
 \end{bmatrix},\begin{bmatrix}
-\sigma_{a}^2(\bx) & \sigma_{a}(\bx) \sigma_b(\bx) \gamma  \\
-\sigma_{a}(\bx) \sigma_b(\bx) \gamma  & \sigma_{b}^2(\bx) 
+\sigma_{a}^2(\bx) & \sigma_{a}(\bx) \sigma_b(\bx) \gamma(\bx)  \\
+\sigma_{a}(\bx) \sigma_b(\bx) \gamma(\bx)  & \sigma_{b}^2(\bx) 
 \end{bmatrix}
-\right).
-\end{equation}
-
-\kcomment{Explain notion, even if this is a for a Statistics journal}
-Here, the mean and standard deviation are allowed to vary with
-location, which is important for capturing the variability such as
-near a river mouth even as the correlation between salinity and
-temperature is assumed to be uniform across all locations. While there
-exist several multivariate spatial covariance functions
-\citep{gneiting2010matern,genton2015cross}, in practice it has been
-shown to be difficult to estimate all parameters from data. Typically
-we assume the same correlation decay for salinity and temperature,
-which is not unrealistic for a river plume. As a result, We use a
-separable covariance function such that:
-
-\begin{eqnarray}\label{gp_corr}
-\mbox{Cov}(\xi_a(\bx),\xi_b(\bx')) &=& \rho(\bx,\bx') \gamma \sigma_a(\bx) \sigma_b(\bx'), \\
-\mbox{Cov}(\xi_a(\bx),\xi_a(\bx')) &=& \rho(\bx,\bx') \sigma_a(\bx) \sigma_a(\bx'), \nonumber \\
-\mbox{Cov}(\xi_b(\bx),\xi_b(\bx')) &=& \rho(\bx,\bx') \sigma_b(\bx) \sigma_b(\bx'). \nonumber
-\end{eqnarray}
-
-\kcomment{Ditto above; presumably x' is location separate from where an
-  earlier measurement was taken?}
-
-Moreover, we assume a stationary isotropic correlation function:
-
-$\rho(\bx,\bx')=\rho(h)$ where h is the distance between $x$ and $x'$ 
-($h=\sqrt{(\bx-\bx')^t (\bx-\bx')}$).
-
-For implementation we discretize the spatial domain $\mathcal{M}$ on a
-set of $n$ grid locations; $\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$.
-We let $\Delta$ be the area of each cell in the regular grid.  Using
-vector notation, we collect the temperature and salinity variables on
-the grid as:
-
-\begin{eqnarray}
-    \bxi &=& (\xi_a(\bx_1),\xi_b(\bx_1),\ldots,\xi_a(\bx_n),\xi_b(\bx_n))^t, \nonumber \\
-    \pi(\bxi) &=& N(\bmu,\bSigma) = \frac{1}{(2\pi)^{n}|\bSigma|^{\frac{1}{2}}}e^{-\frac{1}{2}(\bxi -\bmu)^T\bSigma^{-1}(\bxi - \bmu)},
-\end{eqnarray}
-
-where $\pi(\bxi)$ \kcomment{suggest you try a different (e.g. upper
-  case) notation for $\pi$ given it is being used in different ways
-  above} is the joint density function of $\bxi$, entries in length
-$2 n$ vector $\bmu$ hold the expected values of temperature and
-salinity at the grid locations, and the $2n \times 2n$ matrix
-$\bSigma$ holds all the variance-covariances of temperature and
-salinity variables at all grid locations.
-
-\subsection{Updating Gaussian processes}
-
-The GP mean and covariance in Eq. (\ref{prior} \kcomment{citation
-  incorrect}) are updated when more information is available. In our
-setting, the autonomous vehicle will collect a batch \kcomment{what is
-  a 'batch' be explicit or use another term; 'iteration'?} of
-temperature and salinity measurements along a survey path; after each
-batch, the data is assimilated to get an updated model. We describe
-one round of updating \kcomment{where? In this section?}. In Section
-\ref{sec:heuristics} we discuss sequential data gathering and the
-model update process.
-
-We denote the AUV data at $m$ locations by
-$\by=(y_{a,1},y_{b,1},\ldots,y_{a,m},y_{b,m})^t$, representing the
-temperature and salinity measurements gathered along a survey
-transect. In the grid representation, the likelihood model is
-described by:
-
+\right),
+\end{equation}
+where the notation means Gaussian (Normal) distributed variables specified by the mean vector and covariance matrix.
+In the application below, the model parameters are specified from preliminary data: The mean varies with location, capturing the variability near the river mouth, while the variance parameters and correlation between salinity and
+temperature are assumed to be constant for all locations.
+
+For the spatial dependence among variables, 
+we assume a separable correlation function with the same decay for salinity and temperature,
+which is not unrealistic for a river plume. We have
+\begin{equation}\label{gp_corr}
+\mbox{Cov}(\xi_i(\bx),\xi_j(\bx')) = \rho(\bx,\bx') \sigma_i \sigma_j (\gamma +(1-\gamma )\delta_{ij}), \\
+\end{equation}
+for $i,j \in {a,b}$ and $\delta_{ij}=1$ if $i=j$ and $\delta_{ij}=0$ otherwise.
+Moreover, we assume a stationary isotropic where the correlation function
+$\rho(\bx,\bx')$ solely depends on $\bx$ and $\bx'$ through their Euclidean distance $h=\sqrt{|\bx-\bx'|^2}$.
+With lots of data and prior knowledge one could possibly fit and estimate parameters of non-stationary or non-isotropic covariance function and more complex multivariate spatial covariance functions
+\citep{gneiting2010matern,genton2015cross}, but we do not consider that here.
+
+For implementation on the computer we discretize the spatial domain $\mathcal{M}$ to a
+set of $n$ grid locations $\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell has area $\Delta$.
+The integral expression in Eq.
+(\ref{measVAR}) and (\ref{sur}) are then approximated by sums over all grid cells.
+This grid is identical to the possible set of design locations, as mentioned previously, defined as a waypoint graph.
+Using vector notation, we denote the Gaussian temperature and salinity variables on the grid and its density function by
+\begin{equation}\label{prior}
+    \bxi = (\xi_a(\bx_1),\xi_b(\bx_1),\ldots,\xi_a(\bx_n),\xi_b(\bx_n))^t, \hspace{3mm}
+    \bxi  \sim  N(\bmu, \bSigma) %\pi (\bxi) & =& \frac{1}{(2\pi)^{n}|\bSigma|^{\frac{1}{2}}}e^{-\frac{1}{2}(\bxi -\bmu)^T\bSigma^{-1}(\bxi-\bmu)}.
+\end{equation}
+%Here, $\pi(\bxi)$ is the joint density function of length $2 n$ vector $\bxi$, and
+Here, the length
+$2 n$ vector $\bmu$ holds the expected values of temperature and
+salinity at the grid locations, while the $2n \times 2n$ positive definite  matrix $\bSigma$ holds the variance-covariances of temperature and salinity variables at all grid locations.
+
+\subsection{Data model and updating of Gaussian processes}
+
+Data are in our setting gathered by an autonomous vehicle that can measure
+temperature and salinity at a subset of locations on the defined grid.
+This can be done in a static manner, where the survey design is pre-planned, or in a sequential way with several rounds of data gathering.
+For the latter, a round of data is gathered first, and then assimilated to get an updated model.
+Then the updated model is used to design and select the next round of measurements, and so on.
+We here describe the model for one round of data and the Gaussian formula for one round of updating. In Section
+\ref{sec:heuristics} we discuss sequential data gathering in the context of adaptive design of AUV measurements.
+
+Denote the data at $n_y$ locations by
+$\by=(y_{a,1},y_{b,1},\ldots,y_{a,n_y},y_{b,n_y})^t$, representing the
+temperature and salinity measurements gathered at one round grid nodes. The conditional model for the data, given the true temperature and salinity, is described by
 \begin{equation}\label{likelihood}
-\pi(\by | \bxi)=N(\bA \bxi,\bR)= \frac{1}{(2\pi)^{m}|\bR|^{\frac{1}{2}}}e^{-\frac{1}{2}(\by -\bA\bxi)^T\bR^{-1}(\by-\bA\bxi)},
-\end{equation}
-
-where $\bA$ is an $2m \times 2n$ matrix holding '$1$'s on the
-measurement indexes on the grid, otherwise '$0$'. The size
-$2m \times 2m$ covariance matrix $\bR$ has diagonal entries $r_a$ and
-$r_b$ which are measurement error variances of temperature and
+\by | \bxi \sim N( \bA \bxi, \bR), %\pi(\by | \bxi)= \frac{1}{(2\pi)^{m}|\bR|^{\frac{1}{2}}}e^{-\frac{1}{2}(\by -\bA\bxi)^T\bR^{-1}(\by-\bA\bxi)},
+\end{equation}
+where $\bA$ is an $2n_y \times 2n$ matrix holding '$1$'s on the
+measurement indices on the grid, otherwise '$0$'. The size
+$2n_y \times 2n_y$ covariance matrix $\bR$ contains diagonal entries $r^2_a$ and
+$r^2_b$ which are measurement error variances of temperature and
 salinity observations.
 
+The GP mean and covariance in Eq. (\ref{prior}) are moderated when more information gets available. 
 The updated distribution for temperature and salinity variables on the
@@
 covariance 
-
 \begin{eqnarray}\label{gp_upd}
-\bm &=& \bmu+\bSigma \bA^t (\bA \Sigma \bA^t+\bR)^{-1}(\by-\bA \bmu),  \\
+\bm &=& \bm(\by) = \bmu+\bSigma \bA^t (\bA \Sigma \bA^t+\bR)^{-1}(\by-\bA \bmu),  \\
 \bS &=& \bSigma - \bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.\nonumber
@@
 
-The updated bivariate distribution at grid location $\bx$ is then
-
+The updated bivariate distribution at a grid location $\bx \in \mathcal{M}_g$ is then 
 \begin{equation}\label{gp_hat}
-\pi 
-\left(
 \begin{bmatrix}
@@
  |\by
- \right)
- = N \left( 
+ \sim N \left( 
 \begin{bmatrix} m_{a}(\bx)\\
@@
 \end{equation}
-
 with entries extracted from the conditional mean and covariance
@@
 % derivation in Section \ref{sec:sur}.
-This conditional mean is a linear function of the data $\by$, and
-since the marginal distribution of the data is
-$\pi (\by)= N( \bA \bmu,\bA \bSigma \bA^t+\bR)$, we get
-\begin{equation}\label{distxi} \pi(\bm)= N(\bmu , \bPsi), \hspace{3mm}
+The calculation of the design criterion in Eq. (\ref{sur}) requires the marginal distribution of the data which is defined via marginalization over $\bxi$ to be
+$\by \sim N( \bA \bmu,\bA \bSigma \bA^t+\bR)$. 
+However, it turns out that a simplified form of the calculation is possible because the 
+conditional mean in Eq (\ref{gp_hat}) is a linear (affine) function of the data $\by$ and the covariance is not a function of the data.
+Before knowing the data, the distribution of the conditional mean is
+\begin{equation}\label{distxi} \bm \sim N(\bmu , \bPsi), \hspace{3mm}
   \bPsi=\bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.
-\end{equation} When considering only grid location $\bx$, we can
-extract elements from the mean vector and covariance matrix to get the
-bivariate distribution as follows \begin{equation}\label{dist_mxi} \pi
-  \left( \bm(\bx) \right) = N \left( \begin{bmatrix}
+\end{equation} 
+When considering only grid location $\bx$, we can
+extract elements from its mean vector and covariance matrix to get the
+bivariate distribution \begin{equation}\label{dist_mxi} 
+  \bm(\bx) \sim N \left( \begin{bmatrix}
       \mu_{a}(\bx) \\
@@
       \psi^2_{a}(\bx) & \psi_{a,b}(\bx)\\
-      \psi_{a,b}(\bx) & \psi^2_{b}(\bx) \end{bmatrix} \right).
-\end{equation}
+      \psi_{a,b}(\bx) & \psi^2_{b}(\bx) \end{bmatrix} \right),
+\end{equation}
+which is important for the derivation of the closed form solution of the design criteria.
+
+\section{Uncertainty reduction on excursion sets}
+\label{sec:sur}
 
@@
 
-Based on the Gaussian modeling assumptions, the EPs in Eq. (\ref{eq:prob}) can be computed at any location $\bx$. Without additional data, we have
+Based on the Gaussian modeling assumptions, the EPs can be computed at any location $\bx$. Without
+additional data, Eq.
+(\ref{eq:prob}), we have
 \begin{eqnarray}\label{eq:prob_mv0}
- p(\bx) &=& P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b) \nonumber \\
+ p(\bx) &=& P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b) \nonumber \\
  &=& \Phi_2 \begin{pmatrix} 
-\begin{bmatrix} T_a\\
-T_b
+\begin{bmatrix} t_a\\
+t_b
 \end{bmatrix};
@@
 \end{bmatrix},\begin{bmatrix}
-\sigma_{a}^2(\bx) & \sigma_{a}(\bx)\sigma_{b}(\bx) \gamma  \\
-\sigma_{a}(\bx)\sigma_{b}(\bx) \gamma  & \sigma_{b}^2(\bx)  
+\sigma_{a}^2(\bx) & \sigma_{a}(\bx)\sigma_{b}(\bx) \gamma(\bx)  \\
+\sigma_{a}(\bx)\sigma_{b}(\bx) \gamma(\bx)  & \sigma_{b}^2(\bx)  
 \end{bmatrix}\end{pmatrix},
 \end{eqnarray}
-where $\Phi_2$ is the bivariate Gaussian cumulative distribution function, and the mean and covariance entries are defined in Eq. (\ref{gp_bar}).
-
-When data $\by$ is available, we similarly get
+where $\Phi_2$ is the bivariate Gaussian cumulative distribution
+function, and the mean and covariance entries are as defined in Eq. (\ref{gp_bar}).
+
+When data $\by$ are available, Eq. (\ref{eq:post_ep}), we similarly get
 \begin{eqnarray}\label{eq:prob_mv}
- p(\bx;\by) &=& P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b |\by) 
+ p(\bx;\by) &=& P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b |\by)
  \nonumber \\
  &=& \Phi_2 \begin{pmatrix} 
-\begin{bmatrix} T_a\\
-T_b
+\begin{bmatrix} t_a\\
+t_b
 \end{bmatrix};
@@
 \end{eqnarray}
-with parameters defined in Eq. (\ref{gp_upd}) and (\ref{gp_hat}).
-
-For one particular dataset, the probability $p(\bx;\by)$ can be closer to $0.5$ than the initial probability $p(\bx)$.
-This means that the variance in the ES is increased.
-On average, taking expectations over the data $\by$, the ES variance will be reduced by gathering data.
-We next explain how AUVs can be used to efficiently gather data and reduce the overall marginal spatial uncertainties in the ES.
-
-\section{Uncertainty reduction on excursion sets}\label{sec:sur}
-
-\subsection{Variance in excursion set indicators}
-
-Regions with temperatures or salinity that surpasses the thresholds with a large margin (e.g.
-very high/low temperatures) are not interesting, unless the uncertainty at these locations is very large.
-The objective will therefore be to explore regions where the process variables are close to the thresholds.
-With this mind, it is valuable to explore locations $\bx$ where $p(\bx) \approx 0.5$.
-This is what gives the maximum variance $p(\bx) (1-p(\bx))$ in the indicator for the ES.
-
-Consider then an integrated measure as follows:
-\begin{equation}\label{measVAR}
-    V = \int_{\bx \in \mathcal{M}} p(\bx) \left(1-p(\bx)\right) d\bx \approx \sum_{x \in \mathcal{M}_g} p(\bx) \left(1-p(\bx)\right) \Delta.
-\end{equation}
-The measure is dominated by location having probabilities close to $0.5$. 
-(Note that \cite{Bect2019} call this the integrated Bernoulli variance.)
-
-Considering one location only, we study the effect of different correlation $\gamma$ and prior variance. 
-Figure \ref{illus_bivarDens} shows contour plots of three different densities.
-Here, the temperature mean is $T_a=5^o$ and salinity mean is $T_b=30$ mg/l.
-The thresholds are identical to these means.
-The variables have unit variance and different correlation (low, medium and high).
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
-\caption{Density contour plots having different correlations for temperature and salinity.}\label{illus_bivarDens}
-\end{figure}
-Table \ref{tab:sim_rhoab} shows the EPs and the associated variances in the ES (top rows).
-The results are computed for different correlations $\gamma$ indicated in Figure \ref{illus_bivarDens}, and for standard deviations $1$ (left) and $2$ (right).
-The variance in the ES is largest for high correlation, as there is more tendency of joint small or high occurrences for temperature and salinity in this case.
-EPs are the same for standard deviation $1$ or $2$.
-\begin{table}[!h]
-\centering
-\caption{EP and variance of ES for different correlations (top rows), and for data of both variables and only for temperature (bottom rows).}
-\begin{tabular}{c|ccc|ccc}
- &\multicolumn{3}{c}{$\sigma_b=\sigma_a=1$} & \multicolumn{3}{c}{$\sigma_b=\sigma_a=2$} \\
-\hline
-Correlation $\gamma$ & 0.2 & 0.6 & 0.8 & 0.2 & 0.6 & 0.8 \\
-\hline
-$p$ & 0.28 & 0.35 & 0.40 & 0.28 & 0.35 & 0.40 \\ 
-$p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
-$E_{(y_a,y_b)}(p (1-p))$ & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
-$E_{y_a}(p (1-p))$ & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
-\hline
-\end{tabular}
-\label{tab:sim_rhoab}
-\end{table}
-
-\subsection{Expected variance criteria}
-
-When designing AUV sampling paths, we aim to reduce the variances in the ES and in this way efficiently characterize the ES throughout the spatial domain.
-We saw that the measure defined by Eq.
-(\ref{measVAR}) does not discriminate between high or low variability in the temperature and salinity (two top rows of Table \ref{tab:sim_rhoab}).
-Also, focusing only on locations with EPs near $0.5$ would not account for the spatial correlation in the model.  
-The same has been seen for related measures learning graphical models \citep{lilleborge2016information}.
-
-To evaluate AUV data sampling paths, it is hence natural to evaluate the expected variance reduction in the ES  \citep{chevalier2014fast}, as provided by the survey data.
-In this way, we take the expectation with respect to the planned measurements $\by$:
-\begin{eqnarray}\label{sur}
-    V_{\mbox{upd}} &=& \int_{\bx \in \mathcal{M}} E_{\by} \left\{ p(\bx;\by)\left( 1-p(\bx;\by)\right) \right\} d\bx \\
-    & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by} \left\{ p(\bx;\by)\left( 1-p(\bx;\by)\right) \right\} \Delta, \nonumber \\
-    p(\bx;\by) &=& P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b |\by), \nonumber
-\end{eqnarray}
-where the conditional probability is defined in Eq. (\ref{eq:prob_mv}).
-
-We will next derive closed form solutions to the expectation in (\ref{sur}). A critical
-element in the derivation is the GP model, and the linear dependence on data $\by$, which means that we have a closed form Gaussian distribution for the conditional mean in Eq.
-(\ref{dist_mxi}).
-This means that the high-dimensional inner integral in Eq.
-(\ref{sur}) reduces to a bivariate integral \citep{bhattacharjya2013value, chevalier2014fast}.
-The end result then relies on calls to the multivariate Gaussian cumulative distribution function for calculation of orthant probabilities.
-
-Table \ref{tab:sim_rhoab} (bottom two rows) shows results of doing such calculations, for different correlation $\gamma$ and standard deviations (left and right).
-Here, the expected variances are computed for both data types $(y_a,y_b)$, and for temperature measurements $y_a$ alone.
-With both data; $(y_a,y_b)^t=(\xi_a,\xi_b)^t+N(0,0.5^2\bI)$, while $y_a=\xi_a+N(0,0.5^2)$ when only temperature is measured.
-The results in Table \ref{tab:sim_rhoab} shows that 
-the expected variance depends on the (prior) uncertainty in temperature and salinity.
-The expected variance in the ES is lower with larger standard deviations $\sigma_a$ and $\sigma_b$ (right columns).
-Moreover, the relative reduction is largest for the case with high correlation $\gamma$.
-There is also large uncertainty reduction when only temperature is measured (bottom row), especially when the temperature and salinity are very dependent.
-When correlation is low ($\gamma=0.2$) there is hardly any information about salinity in the temperature data.
-In the application with fresh cold water from the river, the temperature and salinity variables will not only be interdependent, but they will be dependent in the spatial dimension, so this kind of location-wise dependence will also impact design criteria when we evaluate the information measure by integrating over several locations $\bx$.
-
-\subsection{Analytical results for bivariate ES}
-
-In the rest of this section, the location index $\bx$ is suppressed to avoid overly complex notation.
+with parameters as defined in Eq. (\ref{gp_upd}) and (\ref{gp_hat}).
+
+\subsection{Expected Integrated Bernoulli variance for bivariate models}
+
+We will next derive closed form solutions to the expectation in (\ref{sur}).
+The location index $\bx$ is suppressed to avoid overly complex notation.
 We present new results for computing the expectation $E_{\by}\left\{ \cdot \right\}$ which is the integrand of Eq.
@@
 
-The first key result is that the expectation only depends on the data through the linear combination $\bm=(m_{a},m_{b})^t$. 
-To see this we standardize the two variables i.e.
+A critical first
+element in the derivation is the GP model, and the linear dependence on data $\by$ in the conditional mean in Eq.
+(\ref{gp_upd}), which means that we have a closed form Gaussian distribution for the conditional mean in Eq.
+(\ref{dist_mxi}).
+The high-dimensional inner integral in Eq.
+(\ref{sur}) then reduces to a bivariate integral \citep{bhattacharjya2013value, chevalier2014fast}, with expectation over $\bm=(m_{a},m_{b})^t$.
+We standardize the two variables in Eq. (\ref{sur}), i.e.
 $Z_a=\frac{\xi_a-m_{a}}{s_{a}}$,
-$Z_b=\frac{\xi_b-m_{b}}{s_{b}}$; 
-then
-\begin{eqnarray}
-   P(\xi_a \leq T_a, \xi_b \leq T_b|\by) &=& P \left( Z_a \leq\frac{T_a-m_{a}}{s_{a}}, Z_b \leq\frac{T_b-m_{b}}{s_{b}}|\by \right) \nonumber \\
+$Z_b=\frac{\xi_b-m_{b}}{s_{b}}$, and then
+\begin{eqnarray}\label{ytom}
+   P(\xi_a \leq t_a, \xi_b \leq t_b|\by) &=& P \left( Z_a \leq\frac{t_a-m_{a}}{s_{a}}, Z_b \leq\frac{t_b-m_{b}}{s_{b}}|\by \right) \nonumber \\
    &=& \Phi_2 \begin{pmatrix} 
-\begin{bmatrix} \frac{T_a-m_{a}}{s_{a}}\\
-\frac{T_b-m_{b}}{s_{b}}
+\begin{bmatrix} \frac{t_a-m_{a}}{s_{a}}\\
+\frac{t_b-m_{b}}{s_{b}}
 \end{bmatrix};
@@
 \end{bmatrix}\end{pmatrix} \nonumber \\
-&=& P(\xi_a \leq T_a, \xi_b \leq T_b|\bm),
+&=& P(\xi_a \leq t_a, \xi_b \leq t_b|\bm),
 \end{eqnarray}
-where $\eta_{a,b} =s_{a,b}/(s_{a} s_{b})$ is defined in Eq. (\ref{eq:prob_mv}).
-Since the expression only depends on $\by$ via $\bm$, the expectation in Eq.
+where $\eta_{a,b} =s_{a,b}/(s_{a} s_{b})$.
+%Since the expression only depends on $\by$ via $\bm$, the expectation in Eq.
 (\ref{sur}) is reduced to a bivariate integral over $\bm$ (which will vary for each integrand term, because the locations differ).
 
-The expected variance can now be written as
+Using Eq. (\ref{ytom}) and
+introducing $\pi(\bm)$ for the density function of $\bm$, the expected Bernoulli variance becomes
 \begin{eqnarray}\label{eq:var2}
-E_{\by}(p[1-p]) &=& E_{\bm}(p[1-p]), \hspace{3mm} p=P(\xi_a \leq T_a, \xi_b \leq T_b|\bm) \nonumber \\
-E_{\bm}(p[1-p]) & = & \int p[1-p] p(\bm) d\bm, \nonumber \\
- &=& \int P(\xi_a \leq T_a, \xi_b \leq T_b|\bm)  p(\bm) d\bm \nonumber  \\
-&-& \int P(\xi_a \leq T_a, \xi_b \leq T_b|\bm) P(\xi_a \leq T_a, \xi_b \leq T_b|\bm) p(\bm) d\bm, 
+E_{\by}(p(1-p)) &=& E_{\bm}(p(1-p)), \hspace{3mm} p=P(\xi_a \leq t_a, \xi_b \leq t_b|\bm) \nonumber \\
+E_{\bm}(p(1-p)) & = & \int p[1-p] \pi(\bm) d\bm, \nonumber \\
+ &=& \int P(\xi_a \leq t_a, \xi_b \leq t_b|\bm)  \pi(\bm) d\bm \nonumber  \\
+&-& \int P(\xi_a \leq t_a, \xi_b \leq t_b|\bm) P(\xi_a \leq t_a, \xi_b \leq t_b|\bm) \pi(\bm) d\bm. 
 \end{eqnarray}
-where the integral is split in two components.
 We will next extend results of \cite{chevalier2014fast} to solve for these two parts (see also \cite{stroh} who mentioned the possibilities of such extensions).
 
-Considering part I of expression (\ref{eq:var2}), we have
+Considering part I of the, we have
 \begin{equation}\label{part1:phi2}
- \int P(\xi_a \leq T_a, \xi_b \leq T_b|\bm) p(\bm) d\bm= 
-P \left( Z_{a} \leq \frac{T_a-m_{a}}{s_{a}}, 
-Z_{b} \leq \frac{T_b-m_{b}}{s_{b}} \right), \nonumber
+ \int P(\xi_a \leq t_a, \xi_b \leq t_b|\bm) \pi(\bm) d\bm= 
+P \left( Z_{a} \leq \frac{t_a-m_{a}}{s_{a}}, 
+Z_{b} \leq \frac{t_b-m_{b}}{s_{b}} \right), \nonumber
 \end{equation}
@@
 \begin{equation}
-    E(s_{a} Z_{a}+m_{a}-T_a) = \mu_{a}-T_a, \hspace{3mm}
-    \mbox{Var}(s_{a} Z_{a}+m_{a}-T_a) = s^2_a+ \psi^2_{a}, 
+    E(s_{a} Z_{a}+m_{a}-t_a) = \mu_{a}-t_a, \hspace{3mm}
+    \mbox{Var}(s_{a} Z_{a}+m_{a}-t_a) = s^2_a+ \psi^2_{a}, 
 \end{equation}
@@
 \begin{eqnarray}\label{two_parts0}
-& P & \left( Z_{a} \leq \frac{T_a-m_{a}}{s_{a}}, 
-Z_{b} \leq \frac{T_b-m_{b}}{s_{b}} \right) \\
+& P & \left( Z_{a} \leq \frac{t_a-m_{a}}{s_{a}}, 
+Z_{b} \leq \frac{t_b-m_{b}}{s_{b}} \right) \\
 &=& \Phi_2 \begin{pmatrix} 
@@
 \end{bmatrix};
-\begin{bmatrix} \mu_{a}-T_a\\
-\mu_{b}-T_b
+\begin{bmatrix} \mu_{a}-t_a\\
+\mu_{b}-t_b
 \end{bmatrix},\begin{bmatrix}
@@
 \begin{eqnarray}\label{part1:phi4}
-&& \int P(\xi_a \leq T_a, \xi_b \leq T_b|\bm) P(\xi_a \leq T_a, \xi_b \leq T_b|\bm) p(\bm) d\bm_{\xi}= \nonumber \\
-&& P\left( Z_{1,a} \leq \frac{T_a-m_{a}}{s_{a}}, 
-Z_{1,b} \leq \frac{T_b-m_{b}}{s_{b}},Z_{2,b} \leq \frac{T_a-m_{a}}{s_{a}}, 
-Z_{2,b} \leq \frac{T_b-m_{b}}{s_{b}} \right). \nonumber
+&& \int P(\xi_a \leq t_a, \xi_b \leq t_b|\bm) P(\xi_a \leq t_a, \xi_b \leq t_b|\bm) p(\bm) d\bm =  \\
+&& P\left( Z_{1,a} \leq \frac{t_a-m_{a}}{s_{a}}, 
+Z_{1,b} \leq \frac{t_b-m_{b}}{s_{b}},Z_{2,b} \leq \frac{t_a-m_{a}}{s_{a}}, 
+Z_{2,b} \leq \frac{t_b-m_{b}}{s_{b}} \right). \nonumber
 \end{eqnarray}
@@
 
-The solution to part II is then to evaluate a four-variable cumulative distribution function, accounting for the mean, variance and covariances of $\bZ_1$, $\bZ_2$ and $\bm$ in the appropriate linear combinations.
+The solution to part II is then to evaluate a four-variable cumulative distribution function associated with Eqn.
+(\ref{part1:phi4}), accounting for the mean, variance and covariances of $\bZ_1$, $\bZ_2$ and $\bm$ in the appropriate linear combinations.
 Using vector-matrix formulations, we define linear combinations
@@
 \begin{eqnarray}\label{part1:phi4}
-&& P \left( Z_{1,a} \leq \frac{T_a-m_{a}}{s_{a}}, 
-Z_{1,b} \leq \frac{T_b-m_{b}}{s_{b}},Z_{2,b} \leq \frac{T_a-m_{a}}{s_{a}}, 
-Z_{2,b} \leq \frac{T_b-m_{b}}{s_{b}} \right)  \\
+&& P \left( Z_{1,a} \leq \frac{t_a-m_{a}}{s_{a}}, 
+Z_{1,b} \leq \frac{t_b-m_{b}}{s_{b}},Z_{2,b} \leq \frac{t_a-m_{a}}{s_{a}}, 
+Z_{2,b} \leq \frac{t_b-m_{b}}{s_{b}} \right)  \\
 &=&  \Phi_4 
 \begin{pmatrix} 
-\begin{bmatrix} T_a\\
-T_b \\
-T_a \\
-T_b
+\begin{bmatrix} t_a\\
+t_b \\
+t_a \\
+t_b
 \end{bmatrix};
@@
 (\ref{two_parts0}) and part II solution in Eq.
-(\ref{part1:phi4}) are computed for each location on the grid, and summing over the spatial domain $x \in \mathcal{M}_g$, gives the design evaluation criterion in Eq.
+(\ref{part1:phi4}) are computed for each location on the grid, and summing over the $x \in \mathcal{M}_g$, gives the design evaluation criterion in Eq.
 (\ref{sur}).
@@
 
-In our application with temperature and salinity, the solutions to the complex integral equations for the expected variance are reduced to calculating bi-variate and four-variate cumulative distribution functions for Gaussian vectors.
+In our application with temperature and salinity, the solutions to the complex integral equations for the expected IBV are reduced to calculating bi-variate and four-variate cumulative distribution functions for Gaussian vectors.
 For the more general situation one would be interested in $K$ random fields.
-Say, in the ocean sciences the variables could include chlorophyll or plankton variables, in addition to temperature and salinity.
-
-The equations derived in the previous section can be generalized to such $K$-variable cases. We then define $\bxi(\bx)=(\xi_1(\bx),\ldots,\xi_K(\bx))$, and the expected variance in the ES is
+Say, in the ocean sciences the variables could include chlorophyll or other bio-chemical quantities, in addition to temperature and salinity.
+
+The equations derived in the previous section can then be generalized. Let $\bxi(\bx)=(\xi_1(\bx),\ldots,\xi_K(\bx))$, and the expected Bernoulli variance is
 \begin{equation}\label{two_partsK}
-E_y(p[1-p]) = \int p[1-p] p(\by) d\by, \hspace{3mm} P(\xi_k \leq T_k ; k=1,\ldots,K|\by). 
+E_{\by}(p(\bx;\by)(1-p(\bx;\by))), \hspace{3mm} p(\bx;\by)=P(\xi_k(\bx) \leq t_k ; k=1,\ldots,K|\by). 
 \end{equation}
 By the first key result of linear combination of Gaussian variables, we reduce the integral to a $K$ dimensional integral of the relevant $\bm=E(\bxi(\bx)|\by)$.
-Next, reducing the integral to two parts, we standard the variables in each and compute cumulative distributions.
+Next, reducing the integral to two parts, we standardize the variables and compute cumulative distributions.
 In summary, we get 
 \begin{equation}\label{two_parts}
-E_y(p[1-p]) =  \Phi_K 
+E_{\by}(p(\bx)(1-p(\bx))) =  \Phi_K 
 \begin{pmatrix}
-\begin{bmatrix} T_1\\
+\begin{bmatrix} t_1\\
 \vdots \\
-T_K 
+t_K 
 \end{bmatrix};
@@
 \begin{pmatrix}
-\begin{bmatrix} T_1\\
+\begin{bmatrix} t_1\\
 \vdots \\
-T_K \\
-T_1\\
+t_K \\
+t_1\\
 \vdots \\
-T_K 
+t_K 
 \end{bmatrix};
@@
 \end{equation}
-where $\bS$ and $\bPsi$ are direct $K$-variate extensions of the matrices entries in Eq.
+where $\bS$ and $\bPsi$ are direct $K$-variate extensions of the matrices in Eq.
 (\ref{two_parts0}), and the matrix $\bV$ finds the relevant linear combination of two independent vectors $\bZ_l=(Z_{l,1},\ldots,Z_{l,K})$, $l=1,2$ and $\bm$, as an extension of Eq.
 (\ref{Vcomb}).
+
+\subsection{Illustration}
+
+We calculate EPs and the Bernoulli variance for a few bivariate Gaussian distributions to illustrate concepts. 
+Fig. \ref{illus_bivarDens}
+shows contour plots of three different densities with increasing correlation $\gamma$ between temperature and salinity. 
+\begin{figure}[h!] \centering
+  \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
+  \caption{Density contour plots having different correlations between
+    temperature and salinity. The densities have unit variance and the thresholds are identical to the mean values $5^o C$ and $30 mg/l$.}\label{illus_bivarDens} 
+\end{figure}
+Here, the thresholds are set equal to the mean, $\mu_a=t_a=5^o C$ and salinity $\mu_b=t_b=30$ mg/l.
+The displayed densities have unit standard deviations for both temperature and salinity, but we also study the effect of doubling the standard deviations.
+
+Table \ref{tab:sim_rhoab} shows the EPs and the associated Bernoulli variance (second row) for the examples indicated in Fig.
+\ref{illus_bivarDens}.
+The EPs increase with the correlation as there is a stronger tendency of having jointly small temperature and salinity.
+The Bernoulli variance is similarly largest for high correlation.
+EPs and Bernoulli variances are the same for standard deviation $1$ or $2$, and this means that high variability in the temperature and salinity variables is not captured in the $p(1-p)$ expression.
+
+\begin{table}[!h] \centering \caption{EP and Bernoulli variance for different correlations and variances (top rows), and expected Bernoulli variances for both data 
+    and only for temperature (bottom rows).}
+  \begin{tabular}{c|ccc|ccc}
+ &\multicolumn{3}{c}{$\sigma_b=\sigma_a=1$} & \multicolumn{3}{c}{$\sigma_b=\sigma_a=2$} \\
+\hline
+Correlation $\gamma$ & 0.2 & 0.6 & 0.8 & 0.2 & 0.6 & 0.8 \\
+\hline
+$p$ & 0.28 & 0.35 & 0.40 & 0.28 & 0.35 & 0.40 \\ 
+$p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
+$E_{(y_a,y_b)}(p (1-p))$ & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
+$E_{y_a}(p (1-p))$ & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
+\hline
+\end{tabular}
+\label{tab:sim_rhoab}
+\end{table}
+
+Table \ref{tab:sim_rhoab} (bottom two rows) shows results of expected Bernoulli variance calculations.
+This is presented for a design gathering both data types $(y_a,y_b)$, and for a design with temperature measurements $y_a$ alone.
+Having both data; $(y_a,y_b)^t=(\xi_a,\xi_b)^t+N(0,0.5^2\bI)$, while $y_a=\xi_a+N(0,0.5^2)$ when only temperature is measured.
+The results in Table \ref{tab:sim_rhoab} shows that 
+the expected Bernoulli variance gets lower with larger standard deviations $\sigma_a$ and $\sigma_b$ (right columns).
+The reduction of Bernoulli variance is largest for the cases with high correlation $\gamma$.
+Albeit smaller, there is also uncertainty reduction when only temperature is measured (bottom row), especially when the temperature and salinity are very dependent.
+When correlation is low ($\gamma=0.2$) there is hardly any information about salinity in the temperature data, and hence less uncertainty reduction.
+In the application with fresh cold water from the river, the temperature and salinity variables will not only be interdependent, but they will be dependent in the spatial dimension.
+This dependence will also impact design criteria when we evaluate the information measure by integrating over several locations $\bx$.
 
@@
 
+Recall that the AUV can adapt its survey plan based on what it is measuring, in accordance with the sense, plan, act loop (Fig.
+\ref{fig:sense-plan-act}).
+Here we present adaptive strategies for sampling.
+
 \subsection{Optimal sequential design}
@@
 
-The derived closed form solutions for the expected variance (in the multivariate ES) can now be used to evaluate and select among different survey designs for sampling applications with robotic platforms.
-Considering a set $\mathcal{J}$ of possible designs for a survey. A criterion for the AUV path selection is
-\begin{equation}\label{crit}
-    j^* = \mbox{argmin}_{j \in \mathcal{J}} V_{\mbox{upd}}(j),
-\end{equation}
-where the criterion $V_{\mbox{upd}}(j)$ in Eq. (\ref{sur}) is computed for each of the possible designs. 
-In practice, a survey is split in many stages, giving opportunities for sequential adaption of the AUV path.
-This means that the optimal survey design is found by considering not only the best current AUV line, but also what data gathered along this line would lead to for future sampling at the next stages.
-
-Let $d^{j,n}$ denote design number $j$ at stage $s$ of sequential surveying.
+An adaptive AUV survey is split in many rounds (stages), and at each round, the survey design is found by considering the largest reduction in uncertainty of the ES.
+In our setting the selection is made on the graph defined by the nearest grid nodes in the domain $\mathcal{M}_g$.
+
+The optimal sequential design does not only consider the best current AUV grid node, but also what data gathered at this node would lead to for future sampling at the next stages. 
+Let $d^{j,s}$ denote design number $j$ at stage $s$ of sequential surveying.
 If this design is selected, data $\by^{j,s}$ will be gathered.
@@
 where design choices are indicated by squares while data realizations are indicated by circles. 
-In mathematical terms this optimal design is defined as
-\begin{equation}\label{opt_crit}
-    \bd^* = \mbox{argmin}_{d^{j,1}} \left\{ \int_{\by^{j,1}} \mbox{argmin}_{d^{j,2}} \left\{ \int_{\by^{j,2}} \ldots p(\by^{j,2}|\by^{j,1}) d\by^{j,2} \right\} p(\by^{j,1}) d \by^{j,1} \right\},
-\end{equation}
-where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages.
-
-In practice the optimal solution is not tractable because of the enormous growth in maximization and integral terms over stages, see e.g.
-\cite{sucar2015probabilistic} and \cite{powell2016perspectives}.
-Instead, we next outline two heuristic strategies.
-In terms of notation, let $\mathcal{Y}^{s-1} =\{d^{j,r}, \by^{j,r} ; r=1,\ldots,s-1 \}$ denote the data gathered until stage $s-1$, using a design indexed $j$ (which is suppressed in the data set $\mathcal{Y}^{s-1}$.
+%terms this optimal design is defined as
+%\begin{equation}\label{opt_crit}
+%    \bd^* = \mbox{argmin}_{d^{j,1}} \left\{ \int_{\by^{j,1}} \mbox{argmin}_{d^{j,2}} \left\{ \int_{\by^{j,2}} \ldots \pi(\by^{j,2}|\by^{j,1}) d\by^{j,2} \right\} \pi(\by^{j,1}) d \by^{j,1} \right\},
+%\end{equation}
+%where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages. 
+The mathematical expression for the optimal design then involves a series of intermixed maximizations over designs and integrals over data.
+In practice the optimal solution is not tractable because of the enormous growth over stages, see e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}.
+
+Instead, we next outline heuristic strategies.
+In terms of notation, let $\mathcal{Y}^{s-1} =\{d^{j,r}, \by^{j,r} ; r=1,\ldots,s-1 \}$ denote the data gathered until stage $s-1$, using a selected design $d^{j,r}$, $r=1,\ldots,s-1$.
+
+The derived closed form solutions for the expected IBV are still important building blocks when constructing the sampling designs as they will be used to score the different adaptive design.
+Efficient calculation is important for using adaptive survey designs with robotic platforms.
 
@@
 The simplest heuristic for adaptive sampling is to choose the next sampling location based on current EPs.
-The variance is largest for EP equal to $0.5$, so one selects the graph location where $\bx$ with average probabilities $p(\bx)$ closest to $0.5$.
+The variance is largest for EPs equal to $0.5$, so one selects the design location with EP closest to $0.5$.
 
 At stage $s$, based on the currently available data $\mathcal{Y}^{s-1}$, we fit an updated Gaussian model from Eq.
-(\ref{gp_upd}), with mean $\bm_{s-1}$ and covariance matrix $\bS_{s-1}$, represented on the grid locations covering the spatial domain.
-The next batch of measurements $\by^{j,s}$, can be  gathered with design $j=1,\ldots,J$, indicating all possible AUV lines at the current stage. The selected design is then
-\begin{equation}\label{critNaive}
-    d^{*,n} = \mbox{argmin}_{j \in \mathcal{J}} \int_{\bx \in \mathcal{J}} |p(\bx;\mathcal{Y}^{s-1})-0.5| d\bx. 
-\end{equation}
-This strategy does not account for the uncertainty in the ES, if one design has EPs closer to $0.5$, this is chosen, even though the variance along that line is already very small compared with other locations.
-Hence, this strategy lacks memory of where it has been, where the uncertainty has been reduced.
+(\ref{gp_upd}), with mean $\bm_{s-1}=E(\bxi|\mathcal{Y}^{s-1})$ and covariance matrix $\bS_{s-1}=\mbox{Var}(\bxi|\mathcal{Y}^{s-1})$.
+The next round of measurements $\by^{j,s}$, can be gathered with design $d^{j,s}$, $j=1,\ldots,J$, where $J$ indicates all possible AUV directions or paths at the current stage.
+The {\it{naive}} strategy selects the design according to
+\begin{eqnarray}\label{critNaive}
+    d^{*,s} &=& \mbox{argmin}_{j \in \{1,\ldots,J\}} |p(\bx_{d^{j,s}};\mathcal{Y}^{s-1})-0.5|, \\
+    p(\bx;\mathcal{Y}^{s-1}) &=& P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b | \mathcal{Y}^{s-1}). \nonumber
+\end{eqnarray}
+This strategy does not account for the uncertainty in the temperature or salinity variables, only if one design (node) has EPs closer to $0.5$ (see Table \ref{tab:sim_rhoab}, line two).
+Neither does it account for spatial correlation.
+This strategy lacks memory of where it has been and where the uncertainty has been reduced.
 For this reason it can easily get stuck in local regions.
@@
 
-The myopic (greedy) strategy which we present here is optimal if we imagine taking only one more survey line of measurements.
-In this selection strategy there is no anticipation of what the subsequent survey lines might offer.
+The myopic (greedy) strategy which we present here is optimal if we imagine taking only one more round of measurements.
+In this selection strategy there is no anticipation of what the subsequent designs might offer, beyond the first round.
 
 Based on the currently available data $\mathcal{Y}^{s-1}$, we again fit an updated Gaussian model, represented on the grid locations covering the spatial domain. 
-The next batch of measurements $\by^{j,s}$, can be  gathered with design $j=1,\ldots,J$, indicating all possible AUV lines at the current stage. The selected design is then
+The next round of measurements $\by^{j,s}$, can be gathered with design $d^{j,s}$, $j=1,\ldots,J$, indicating all possible AUV directions / paths at the current stage. The selected design is then
 \begin{eqnarray}\label{critSEQ}
-    d^{*,s} &=& \mbox{argmin}_{j \in \mathcal{J}} \left\{ V_{m,\mbox{upd,j}} \right\},  \\
-V_{m,\mbox{upd}} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j,s}|\mathcal{Y}^{s-1}} \left\{ p(\bx;\mathcal{Y}^{j,s})\left( 1-p(\bx;\mathcal{Y}^{j,s})\right) \right\} \Delta \nonumber \\
-    p(\bx;\mathcal{Y}^{j,s}) &=& P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b |\by^{j,s},\mathcal{Y}^{s-1}). \nonumber
+    d^{*,s} &=& \mbox{argmin}_{j \in \{1,\ldots,J\}} \left\{ V_{m,\mbox{upd,j}} \right\},  \\
+V_{m,\mbox{upd}} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j,s}|\mathcal{Y}^{s-1}} \left\{ p(\bx;\mathcal{Y}^{j,s})\left( 1-p(\bx;\mathcal{Y}^{j,s})\right) \right\} \Delta, \nonumber \\
+    p(\bx;\mathcal{Y}^{j,s}) &=& P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b |\by^{j,s},\mathcal{Y}^{s-1}). \nonumber
 \end{eqnarray}
@@
 (\ref{sur}).
-Now $\mathcal{Y}^{s-1}$ is available, and the expectation is with respect to the conditional distribution $p(\by^{j,s}|\mathcal{Y}^{s-1})$.
-The same closed form calculation is hence applicable in Eq.
+Now $\mathcal{Y}^{s-1}$ is available, and the expectation is with respect to the conditional density $\pi(\by^{j,s}|\mathcal{Y}^{s-1})$.
+A similar closed form calculation for expected IBV is hence applicable in Eq.
 (\ref{critSEQ}), using the updated GP model from step $s-1$.
-Once the data are collected along this best line, the GP model is updated again.
-The mean $\bm_{s}$ and covariance matrix $\bS_{s}$ are used to compute the next survey line at stage $s+1$, using the same strategy, and so on.
-
-This myopic strategy is non-anticipative and hence sub-optimal, but it still gives a reasonable approach for creating designs in many applications.
-Moreover, it is easily implemented on-board, requiring only an efficient approach for data updating of the GP model and the calculation of the closed form variance expressions for each next survey line.
+Once the data are collected for the best design, the GP model is updated again. The mean $\bm_{s}$ and covariance matrix $\bS_{s}$ are used to compute the next design at stage $s+1$, and so on.
+
+Even though this myopic strategy is non-anticipative, it still gives a reasonable approach for creating designs in many applications.
+Moreover, it is easily implemented on-board an AUV, using the efficient approach for data updating of the GP model and the calculation of the closed form expected IBV expressions for each next survey line.
 
@@
 
-We now extend the myopic strategy to a look-ahead strategy which is optimal when one can gather only two more survey lines of measurements.
-In its selection this strategy for the current survey line, it anticipates what the next survey line might offer, but not anything beyond this line.
-
-In addition to the next batch of measurements $\by^{j,s}$, one now considers the subsequent design $j_2$ with data $\by^{j_2,s+1}$ when choosing the next survey line.
-
-The selected design is then
+We now extend the myopic strategy to a look-ahead strategy which is optimal when one can gather only two more rounds of measurements.
+In addition to the next round of measurements $\by^{j,s}$, this look-ahead strategy hence anticipates the subsequent design $j_2$ with data $\by^{j_2,s+1}$, when choosing the current design $d^{j,s}$.
+The selected design is
 \begin{eqnarray}\label{critLA}
-    d^{*,n} &=& \mbox{argmin}_{j \in \mathcal{J}} \left\{ U_{la,\mbox{upd,j}} \right\},  \\
-    U_{la,\mbox{upd},j} & = &  E_{\by^{j,s}|\mathcal{Y}^{s-1}} \left\{ \mbox{argmin}_{j_2 \in \mathcal{J}_2} \left[ V_{la,\mbox{upd},j_2} \right] \right\} \nonumber \\
-V_{la,\mbox{upd},j_2} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j_2,s+1}|\mathcal{Y}^{j,s}} \left\{ p(\bx;\mathcal{Y}^{j,s+1})\left( 1-p(\bx;\mathcal{Y}^{j,s+1})\right) \right\} \Delta \nonumber \\
-    p(\bx;\mathcal{Y}^{j,s+1}) &=& P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b |\by^{j_2,s+1},\mathcal{Y}^{j,s}). \nonumber
+    d^{*,s} &=& \mbox{argmin}_{j \in \{1,\ldots,J\}} \left\{ U_{la,\mbox{upd,j}} \right\},  \\
+    U_{la,\mbox{upd},j} & = &  E_{\by^{j,s}|\mathcal{Y}^{s-1}} \left\{ \mbox{argmin}_{j_2 \in \{1,\ldots,{J}_2\}} \left[ V_{la,\mbox{upd},j_2} \right] \right\}, \nonumber \\
+V_{la,\mbox{upd},j_2} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j_2,s+1}|\mathcal{Y}^{j,s}} \left\{ p(\bx;\mathcal{Y}^{j,s+1})\left( 1-p(\bx;\mathcal{Y}^{j,s+1})\right) \right\} \Delta, \nonumber \\
+    p(\bx;\mathcal{Y}^{j,s+1}) &=& P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b |\by^{j_2,s+1},\mathcal{Y}^{j,s}). \nonumber
 \end{eqnarray}
-Here, $\mathcal{Y}^{j,s}=\{\by^{j,s},\mathcal{Y}^{s-1}\}$ and $\mathcal{Y}^{j,s+1}=\{\by^{j_2,s+1},\mathcal{Y}^{j,s}\}$ represent the set of data variables, and the expectations are with respect to the conditional distributions $p(\by^{j,s}|\mathcal{Y}^{s-1})$ and $p(\by^{j_2,s+1}|\by^{j,s},\mathcal{Y}^{s-1})$.
+Here, $\mathcal{Y}^{j,s}=\{\by^{j,s},\mathcal{Y}^{s-1}\}$ and $\mathcal{Y}^{j,s+1}=\{\by^{j_2,s+1},\mathcal{Y}^{j,s}\}$ represent the sets of data variables, and the expectations are with respect to the conditional densities $\pi(\by^{j,s}|\mathcal{Y}^{s-1})$ and $\pi(\by^{j_2,s+1}|\by^{j,s},\mathcal{Y}^{s-1})$.
 
@@
 Instead, we solve the first expectation by Monte Carlo sampling of data $\by^{j,s}$ from its conditional distribution.
-For each data sample, the second expectation is solved using the closed form expressions from Section \ref{sec:sur}.
-
-Even though the strategy looks at two stages, it is only used to find the current best survey line.
-When data are collected along this best line, the GP model is updated, and the mean $\bm_{s}$ and covariance matrix $\bS_{s}$ are used to compute the next survey line at stage $s+1$, now anticipating what stage $s+2$ could give after this survey, and so on.
+For each data sample, the second expectation is solved using the closed form expressions for expected IBV from Section \ref{sec:sur}.
+
+Even though the strategy looks at two stages, it is only used to find the current best design.
+When data are collected, the GP model is updated, and the mean $\bm_{s}$ and covariance matrix $\bS_{s}$ are used to compute the next design at stage $s+1$, now anticipating what stage $s+2$ could offer, and so on.
 
@@
 This means that we do not compute all possible branches of the first two stages, as they are indicated in Figure \ref{fig:PathSelOpt}.
-Instead, we use the myopic strategy to rank the three best paths on the first stage alone, and for each of these we go through with the look-ahead calculation.
+Instead, we use the myopic strategy to rank the three best designs on the first stage alone, and for each of these we go through with the look-ahead calculation.
 
 \section{Simulation study}\label{sec:simulations}
-We here show results from tests performed to study the properties of different static and sequential survey designs in a realistic simulated case, mapping a freshwater plume defined by a temperature and salinity field.
+
+Results from tests performed to study the properties of different static and sequential survey designs, in a realistic simulated case, is shown in this section.
+The context is mapping a freshwater plume defined by a temperature and salinity field using an AUV.
 
@@
 One realization of the random fields representing salt and temperature is shown in Figure \ref{fig:true_temp} and \ref{fig:true_sal}.
+The true ES from this prior realization is plotted in Figure \ref{fig:ESet}.
 
@@
   \centering
-  \subfigure[Simulated temperature.]{\includegraphics[height = 0.41\textwidth]{Figures/sim/true_temp.pdf}\label{fig:true_temp}}
-  \hfill
-  \subfigure[Simulated salinity.]{\includegraphics[height = 0.41\textwidth]{Figures/sim/true_sal.pdf}\label{fig:true_sal}}
-  \hfill
-  \subfigure[Excursion set given the realization of temperature and salinity.]{\includegraphics[height = 0.41\textwidth]{Figures/sim/es_ts_true.pdf}\label{fig:ESet}}
-  \hfill
-  \subfigure[Excursion probabilities given collected data along the strategy ``static\_north".]{\includegraphics[height = 0.41\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
+  \subfigure[Simulated temperature.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/true_temp.pdf}\label{fig:true_temp}}
+  \hfill
+  \subfigure[Simulated salinity.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/true_sal.pdf}\label{fig:true_sal}}
+  \hfill
+  \subfigure[Excursion set given the realization of temperature and salinity.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_true.pdf}\label{fig:ESet}}
+  \hfill
+  \subfigure[Excursion probabilities given collected data along the strategy ``static\_north".]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
   \caption{(\ref{fig:true_temp}) and (\ref{fig:true_sal}) shows
@@
 
-The true ES from this prior realization is plotted in Figure \ref{fig:ESet}.
-%while the EP ({\bf{what information is this based on? Skip EP here? only ES true? }} )
-
-\subsection{Static and sequential sampling designs}
+\subsection{Static and sequential sampling designs}\label{sec:sampling_designs}
 
@@
   \centering
-  % \subfigure[Excursion set variance $E_y(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
-  \subfigure[Excursion set $E_y$(p\lbrack 1-p\rbrack) variance.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
+  % \subfigure[Excursion set variance $E_{\by}(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
+  \subfigure[Excursion set $E_{\by}$(p\lbrack 1-p\rbrack) variance.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
   \hfill
@@
 this pruning is likely the reason that this strategy does not
-outperforms the myopic approach.
-
-In Figure \ref{fig:route_choices} we plot the realized sampling paths of
-the sequential schemes and static designs.
+outperforms the myopic approach. In Figure \ref{fig:route_choices} we plot the realized sampling paths of the sequential schemes and static designs.
 
@@
   \centering
-  \subfigure[Look-ahead strategy.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_look-ahead.pdf}\label{fig:avg_look-ahead}}
-  \hfill
-  \subfigure[Myopic strategy.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_myopic.pdf}\label{fig:avg_myopic}}
-  \hfill
-  \subfigure[Naive strategy.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_naive.pdf}\label{fig:route_naive}}
-  \hfill
-  \subfigure[Static strategies.]{\includegraphics[height = 0.49\textwidth]{Figures/sim/route_static.pdf}\label{fig:route_static}}
+  \subfigure[Look-ahead strategy.]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_look-ahead.pdf}\label{fig:avg_look-ahead}}
+  \hfill
+  \subfigure[Myopic strategy.]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_myopic.pdf}\label{fig:avg_myopic}}
+  \hfill
+  \subfigure[Naive strategy.]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_naive.pdf}\label{fig:route_naive}}
+  \hfill
+  \subfigure[Static strategies.]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_static.pdf}\label{fig:route_static}}
   \hfill
@@
     \toprule
-    Parameter: $E_y(p[1-p])$ &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
+    Parameter: $E_{\by}(p[1-p])$ &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
     \midrule
@@
 
-\section{Case study}\label{sec:case_study}
-
-\begin{itemize}
-    \item Tune prior mean, standard deviations and correlation parameters from ocean model simulations and initial runs. 
-    \item Run tests 16 May or 21 May.
-\end{itemize}
-
 \newpage
 
+\section{Case study - Mapping a river plume}\label{sec:case_study}
+
+To demonstrate the applicability of multivariate EP to informative oceanographic sampling, a case study mapping river plumes with an AUV is presented.
+This test is done in the near Trondheim, Norway, where the Nidelva river (see Figure \ref{fig:river1}).
+The test focuses along the eastern shoreline due to the aforementioned Coriolis force.%, see Fig.
+\ref{fig:map}.
+The frontal zone therefore runs parallel to the eastern shore as indicated in pink.
+enters the Trondheimsfjord.
+The tests were done in the late Spring of 2019, and because of snow melting in the mountains, the fresh river water is then substantially colder than the fjord water.
+
+\subsection{Statistical model specification}
+
+Statistical model parameters were specified based on preliminary runs.
+A short preliminary survey was therefore conducted upon arriving at the site, where the AUV did an initial transect to determine the environmental conditions and correlation structure.
+
+In these runs the trend parameters were estimated by linear regression, where both temperature and salinity are assumed to increase linearly, going west from the river mouth.
+Next, the residuals from the regression analysis are analyzed to study the fit of the Gaussian model and to specify the covariance parameter.
+
+Figure \ref{fig:parest} shows the diagnostic plots of this residual analysis.
+The left display show a cross-plot of temperature and salinity residuals, indicating larger variability in the salinity than in temperature, and a positive correlation ($0.5$) between the two variables.
+Based on the fitted bivariate Gaussian model we can compute the quadratic form of the residuals.
+If the Gaussian model is adequate the residuals should be approximately $\chi^2_2$ distributed.
+Figure \ref{fig:parest} (middle) shows the empirical cdf of these quadratic forms (solid) together with the theoretical cdf of the $\chi^2_2$ distribution.
+Even though there appears to be some clustering in both left and middle displays, the fit fit appears quite reasonable, and justifies the use of a Gaussian bivariate model in this case.
+Figure \ref{fig:parest} (right) shows the empirical variogram of the scaled residuals for temperature and salinity.
+The decay is similar, and seems to be negligible after about $150$ m.
+The resulting parameters are given in Table \ref{tab:experiment_param} below.
+
+%Mapping the spatial extent of a frontal zones is an important problem for studying many bio-physical interactions in the ocean.
+The frontal zone is determined by the boundary where plumes of sediments, nutrients, and possibly pollutants spreading from the river outlet meet and interact with adjacent coastal water.
+Due to the lower density the plumes spread on the surface, creating a front with an sharp gradient in both temperature and salinity.
+
+\begin{figure}[!h] 
+ \centering 
+\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.eps}
+\caption{Data analysis from a preliminary trial experiment using the AUV.
+Left) Residual plot of temperature and salinity along with Gaussian contours.
+Middle) Empirical cdf of the quadratic form of the residuals along with the theoretical cdf of the Kji-square distribution with two degrees of freedom.
+Right) Empirical variogram of the salinity and temperature data.} \label{fig:parest}
+\end{figure}
+
+\begin{table}[!h]
+\centering
+\begin{tabular}{lrr}
+\toprule
+Parameter & Value & Source\\
+\midrule
+\rowcolor{Gray}
+Cross correlation temp. and sal. & 0.5 & AUV observations\\
+Temp. variance &  0.20 & AUV observations (variogram)\\
+\rowcolor{Gray}
+Sal. variance &  5.76 & AUV observations (variogram)\\
+Corr. range  & 0.15 km & AUV observations (variogram)\\
+\rowcolor{Gray}
+River temp. $T_{river}$ & $10.0\,^{\circ}\mathrm{C}$ & AUV observations\\
+Ocean temp. $T_{ocean}$ & $11.0\,^{\circ}\mathrm{C}$ & AUV observations\\
+\rowcolor{Gray}
+River sal. $S_{river}$ & $14.0$g/kg & AUV observations\\
+Ocean sal. $S_{ocean}$ & $22.0$g/kg & AUV observations\\
+\rowcolor{Gray}
+Threshold temp. & $10.5\,^{\circ}\mathrm{C}$ & $(T_{ocean}-T_{river})/2+T_{river}$\\
+Threshold sal. & $22.0$g/kg & $(S_{ocean}-S_{river})/2+S_{river}$\\
+\rowcolor{Gray}
+\bottomrule
+\end{tabular}
+\caption{Model and threshold parameters found from initial AUV survey. Observations were taken across the front crossing from fresh and cold river water to saline and warmer ocean water.}
+\label{tab:experiment_param}
+\end{table}
+
+These mean and covariance parameters values were then used in a field trial, where we explored the algorithm's ability to explore the front between the river and fjord water masses.
+
+\subsection{Experiment setup}
+
+Two full-scale deployments were conducted with adaptive AUV sampling.
+%on the 2$^{nd}$ of July 2019 spatially mapping the location of the plume from the Nidelven river (Trondheim, Norway) spreading into the Trondheimsfjord 
+The AUV used the onboard autonomous agent T-REX \citep{Mcgann}, running an instance of the myopic strategy from Section \ref{sec:sampling_designs}, to control the AUV and decide between sampling locations.
+
+The sampling locations are distributed over an equilateral grid, as shown in Fig.
+\ref{fig:map} in a grey colored lattice.
+The robotic platform consisted of a Light AUV \citep{sousa2012lauv} equipped with a 16 Hz Seabird Fastcat-49 CTD (conductivity, temperature, and depth) sensor providing temperature and salinity measurements to the bi-variate GP model.
+The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
+
+%\begin{figure}[!h] 
+%\centering 
+%\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
+%\caption{The light AUV platform (Harald) for upper water-
+%column exploration used in the experiment.}
+%\label{fig:harald_auv}
+%\end{figure}
+
+The sampling strategy is designed around sequentially visiting waypoints, where arrival at a desired waypoint subsequently triggers the myopic strategy to evaluate and find the next waypoint based on the ES variance formulated in Eq.
+\eqref{critSEQ}.
+The waypoint --and associated sampling-- that results in lowering the ES variance the most is selected.
+For the field trials, the AUV started in the lower middle of the graph with a GP prior estimate of the environment given as a linear gradient from west to east, between the initially measured temperature and salinity.
+Each AUV survey/mission was set to take approx.
+40 min, visiting 15 waypoints on the grid, with the AUV running at the surface to better capture the plume.
+The GP model assimilated temperature and salinity data continuously with an update frequency of 30 sec.
+
+\subsection{Results}
+
+Two missions, Survey 1 and 2, was run successively from 11:00 AM to 01:00 PM, with a short break in between.
+The resulting path between the selected waypoints are shown in spatial context in Fig.
+\ref{fig:map}, both within the expected frontal region.
+The recorded temperatures are shown as colored trails in Fig.
+\ref{fig:res_both}, clearly showing the temperature difference between ocean and river waters.
+The temperatures are then shown separately overlaid with the estimated ES for each survey in Fig.
+\ref{fig:res1} and \ref{fig:res2}.
+
+Both Survey 1 and 2, successfully estimates and navigates the separation zone crossing the front multiple times.
+As conditions change between the two deployments the resulting path (after waypoint 5) deviate.
+Survey 1 continue northwards tracking the north-eastern part of the front, while Survey 2 turns west mapping the south-western part.
+The final prediction of the front location given Fig.
+\ref{fig:res3} and \ref{fig:res4}, correspond with one another yielding a picture of the front gradually bending off towards north east.
+The amount of exploration done by Survey 1 is greater than Survey 2, coming close to the survey area boarders in the south-western corner.
+A longer planning horizon would identify and discourage such choices.
+
+\begin{figure}[!h]
+\centering
+\subfigure[Survey map.]{\includegraphics[height = 0.31\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
+\hfill
+\subfigure[Temperature tracks.]{\includegraphics[height = 0.31\textwidth]{Figures/field-trials/auv.pdf}\label{fig:res_both}}
+\hfill
+\subfigure[Survey 1]{\includegraphics[height = 0.31\textwidth]{Figures/field-trials/auv1_es.pdf}\label{fig:res1}}
+\hfill
+\subfigure[Survey 2]{\includegraphics[height = 0.31\textwidth]{Figures/field-trials/auv4_es.pdf}\label{fig:res2}}
+\hfill
+\subfigure[ES for Survey 1.]{\includegraphics[height = 0.31\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}
+\hfill
+\subfigure[ES for Survey 2.]{\includegraphics[height = 0.31\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
+\caption{Results from mapping the Nidelven river} 
+\label{fig:results}
+\end{figure}
+
+%\begin{figure}[!h] 
+% \centering 
+%\includegraphics[width=0.48\textwidth]{Figures/envir_ocean.pdf}
+%\caption{Ocean observation is moving away from %single-ship sampling
+%towards more collaborative networked operations in %order to resolve
+%the numerous processes and their interaction.} %\label{fig:envir}
+%\end{figure}
+
 \section{Closing remarks}\label{sec:concl_disc}
 
-Additional flexibility can be gained by having two AUVs exploring together.
+The methods presented in this paper are shown to be useful in characterizing the frontal water masses where cold river water is separated from the warmer saltwater.
+The study did not consider any temporal effects, which would surely be relevant on a longer time scale.
+We consider the extension to spatio-temporal modeling future work, and we envision that advection-diffusion equations could be useful in this kind of modeling \cite{sigrist2015stochastic}.
+For more complicated oceanographic phenomena, we the methods must maybe be extended to non-Gaussian phenomena and with more complex dynamical models.
+
+Regarding the autonomous sampling, more effort can be spent on approximating the look-ahead steps, but it is perhaps more interesting to explore the additional flexibility that can be gained by having two or more AUVs exploring a spatial or spatio-temporal domain together.
 It might then be useful to sample different parts of the space, and maybe move in parallel to best capture the excursion set.

### 2019-08-26 main_new.tex
updates

@@
 \definecolor{Gray}{gray}{0.9}
-\newcommand{\kcomment}[1]{{\color{blue}{\{KR: #1\}}}}
+\newcommand{\kcomment}[1]{{\color{red}{\{KR: #1\}}}}
 \newcommand{\tofcomment}[1]{{\color{green}{\{TF: #1\}}}}
-
+\newcommand{\kc}{\kcomment}
 % put your definitions there:
@@
 \address{Idiap \printead{e3}}
-\affiliation{Uncertainty Quantification and Optimal Design group , Idiap Research Institute, Martigny, Switzerland}
+\affiliation{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland}
 
 \address{Department of Engineering Cybernetics,\printead{e4}}
-\affiliation{Department of Engineering Cybernetics, NTNU}
+\affiliation{Department of Engineering Cybernetics, NTNU \&\\ Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal}
 
@@
 \begin{abstract}
-Improving and optimizing oceanographic sampling is a crucial task for modern marine management and science faced with limited resources and highly dynamic processes.
-The combination of statistics and autonomous robotics provides opportunities for studying aspects related to experimental designs for effective characterization of oceanographic phenomena.
-In this paper methods are developed for efficient spatial sampling applied to the mapping of dynamic coastal processes by providing informative descriptions of spatial characteristics of ocean phenomena.
-Specifically, we define a design criterion based on reducing uncertainty in the excursion set of multivariate random fields, and derive analytical expressions for the expected Bernoulli variance reduction in the excursion set of spatial temperature and salinity variables.
-We show how these can be applicable in static designs and sequential designs, building on myopic and look-ahead sampling strategies.
-We use simulations to study the properties and differences between state-of-the-art approaches.
-Results from field deployments with an autonomous underwater vehicle are also presented, for a case study mapping the boundary of a river plume.
-The results demonstrate the ability of the method to effectively inform and execute data-driven environmental sampling.
+  Improving and optimizing oceanographic sampling is a crucial task for
+  modern marine management and science faced with limited resources to
+  understand highly dynamic processes in the water-column. The
+  combination of statistics and autonomous
+  robotics provides opportunities for studying aspects related to
+  experimental designs for effective characterization of oceanographic
+  phenomena. In this work we develop methods for efficient spatial
+  sampling applied to the mapping of dynamic coastal processes by
+  providing informative descriptions of spatial characteristics of ocean
+  phenomena. Specifically, we define a design criterion based on
+  reducing uncertainty in the excursion set of multivariate random
+  fields, and derive analytical expressions for the expected Bernoulli
+  variance reduction in the excursion set of spatial temperature and
+  salinity variables. We show how these can be applicable in static
+  designs and sequential designs, building on myopic and look-ahead
+  sampling strategies. We use simulations to study the properties and
+  differences between state-of-the-art approaches. Results from field
+  deployments with an autonomous underwater vehicle are also presented,
+  for a case study mapping the boundary of a river plume. The results
+  demonstrate the ability of the method to effectively inform and
+  execute data-driven environmental sampling.
 
@@
 
-Monitoring the worlds oceans has gained increased importance in the light of the changing climate and increasing anthropogenic impact.
-A central problem for understanding these factors is the lack of representative data with sufficient resolution.
-Most of this \emph{undersampling} can be contributed to the large spatio-temporal variation on which ocean processes transpire, prompting the need for effective means of sampling.
-By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use of enabling autonomous platforms can be combined with statistical methods to pursuit measurements with high scientific relevance.
-
-Traditional data collection at sea has typically been based on static buoys, lagrangian floats or ship-centric methods, with significant logistical limitations that directly impact coverage and sampling resolution.
-Modern methods using satellite remote-sensing provide large-scale coverage, but have limited resolution, are limited to the sensing the surface of the ocean, and is impacted by cloud cover.
-However, the advent of robust mobile robotic platforms \cite{Bellingham07} are making significant contributions to environmental monitoring and sampling.
-In particular autonomous underwater vehicles (AUVs), have made significant strides in advancing the state of sampling and consequently have made robotics an integral part of ocean observation in large part driven by using statistical methods in design and data assimilation \citep{das11b,Graham2013,Das2015,das15,fossuminformation,fossum18b}.
-Others include \cite{wikle2013modern} who focus on hierarchical statistical models; \cite{sahu2008space} study spatio-temporal models for sea surface temperature and salinity data, while \cite{mellucci2018oceanic} look at the statistical prediction of features using an underwater glider.
-Common to these approaches is the use of information-based methods to infer environmental sampling and a few important characteristics should be mentioned, namely:
+Monitoring the worlds oceans has gained increased importance in the
+light of the changing climate and increasing anthropogenic impact. A
+central problem for understanding these factors is the lack of
+representative data with sufficient resolution. Most of this
+\emph{undersampling} can be attributed to the large spatio-temporal
+variations on which ocean processes transpire, prompting the need for
+effective means of sampling. By \emph{sampling}, we refer to the design
+of observational strategies in the spatial domain, where the use of
+autonomous platforms can be combined with statistical methods to pursue
+measurements with high scientific relevance.
+
+Traditional data collection at sea has typically been based on static
+buoys, lagrangian floats or ship-based methods, with significant
+logistical limitations that directly impact coverage and sampling
+resolution. Modern methods using satellite remote-sensing provide
+large-scale coverage, but have limited resolution, are limited to
+sensing the surface of the ocean, and is impacted by cloud cover. The
+advent of robust mobile robotic platforms \cite{Bellingham07} has
+resulted in significant contributions to environmental monitoring and
+sampling. In particular autonomous underwater vehicles (AUVs), have made
+strides in advancing the state of sampling and consequently have made
+robotics an integral part of ocean observation in large part by our own
+work % driven by
+% using statistical methods in design and data assimilation
+\citep{das11b,Graham2013,Das2015,das15,fossuminformation,fossum18b}.
+Others include \cite{wikle2013modern} focusing on hierarchical
+statistical models, \cite{sahu2008space} which studies spatio-temporal
+models for sea surface temperature and salinity data, while
+\cite{mellucci2018oceanic} looks at the statistical prediction of
+features using an underwater glider. Common to these approaches is the
+use of information-based methods to infer environmental sampling. A few
+important characteristics should be mentioned, namely:
 
 \begin{itemize}
-    \item Decisions on where to sample happen sequentially, meaning recent information can be assimilated and used to improve statistical models and reduce uncertainty associated with the  choices. 
-    \item As a priori information about the state of the ocean is very uncertain, there lies a substantial efficiency gain in reacting to information obtained from measurements taken \emph{in-situ}, this is usually referred to as the \emph{adaptivity gap}.
-    \item For this type of sampling problem, the number of choices (i.e.
-locations, paths, and candidate designs) makes the problem combinatorially large, creating a trade-off between optimization (finding the optimal design) and computability (arriving at a solution in reasonable time), which has to be considered in practice.
+\item Decisions on where to sample happen sequentially, i.e.  how
+  recently acquired information can be assimilated and used to improve
+  statistical models and reduce uncertainty associated with where and
+  when to make measurements.
+\item As a priori information about the state of the ocean is very
+  uncertain, there lies a substantial efficiency gain in reacting to
+  information obtained from measurements taken \emph{in-situ}, this is
+  usually referred to as the \emph{adaptivity gap} \kc{citation
+    needed}.
+\item For such types of sampling problems, the number of choices (i.e.
+  locations, trajectories, and candidate designs) makes the problem
+  combinatorially large, creating a trade-off between optimization
+  (finding the optimal design) and computability (arriving at a
+  solution in reasonable time), which has to be considered in
+  practice.
 \end{itemize}{}
 
-The combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based sensing.
-Central to this is the ability to model spatially correlated variables and provide formal measures of uncertainty.
-By far the most practical models use Gaussian processes (GPs), which will also be used in our formulation, in large  part because this approach allows efficient implementation and evaluation, enabling prioritization of sampling efforts to be formalized into decision making systems on-board a robotic platform.
+The combination of statistical tools and robotic platforms is a
+natural symbiosis which enables information-based sensing. Central to
+this is the ability to model spatially correlated variables and
+provide formal measures of uncertainty. By far the most practical
+models use Gaussian processes (GPs), which will also be used in our
+formulation, in large part because it allows efficient implementation
+and evaluation, in turn enabling prioritization of sampling efforts to
+be formalized into decision making systems on-board a robotic
+platform.
 
@@
 
-In this work, we build on excursion sets (ES) and excursion probabilities (EP) for spatial design of experiments.
-The goal is to design sampling strategies for reducing the spatially integrated Bernoulli variance (IBV) in the ES indicators.
-The main contributions involve: i) extending ES-based uncertainty criteria to multivariate GPs, allowing the evaluation to include multiple environmental parameters such as temperature and salinity, ii) developing methods for informative environmental sensing and autonomous data collection, and iii) demonstrating the approach in a real world scenario of sampling a river plume using an AUV.
-
-The remainder of the paper is organized as follows: Section \ref{sec:bg} provides some background on ocean sampling.
-Section \ref{sec:ESEP} defines ESs and EPs, and the design criteria of IBV.
-Section \ref{sec:GP_EP} sets up the required statistical modeling assumptions of the bi-variate GPs for temperature and salinity.
-Section \ref{sec:sur} builds on these when deriving the analytical derivation of the design criteria.
-Section \ref{sec:heuristics} extends results to present heuristics for adaptive sampling.
-Section \ref{sec:simulations} shows results and discusses properties of the methods in simulation studies.
-Section \ref{sec:case_study} demonstrates the methodology in field work characterizing a river plume.
-Section \ref{sec:concl_disc} contains summary and a discussion of future work.
+In this work, we build on excursion sets (ES) and excursion
+probabilities (EP) for spatial design of experiments. The goal is to
+design sampling strategies for reducing the spatially integrated
+Bernoulli variance (IBV) in the ES indicators.  The main contributions
+involve: i) extending ES-based uncertainty criteria to multivariate
+GPs, allowing the evaluation to include multiple environmental
+parameters such as temperature and salinity, ii) developing methods
+for informative environmental sensing and autonomous data collection,
+and iii) demonstrating the approach in a real world scenario of
+sampling a river plume using an AUV.
+
+The remainder of the paper is organized as follows: Section
+\ref{sec:bg} provides some background on ocean sampling. Section
+\ref{sec:ESEP} defines ESs and EPs, and the design criteria of
+IBV. Section \ref{sec:GP_EP} sets up the required statistical modeling
+assumptions of the bi-variate GPs for temperature and
+salinity. Section \ref{sec:sur} builds on these when deriving the
+analytical derivation of the design criteria. Section
+\ref{sec:heuristics} extends results to present heuristics for
+adaptive sampling. Section \ref{sec:simulations} shows results and
+discusses properties of the methods in simulation studies. Section
+\ref{sec:case_study} demonstrates the methodology in field work
+characterizing a river plume.  Section \ref{sec:concl_disc} contains
+summary and a discussion of future work.
 
@@
 
-Central to understanding the changes taking place in the upper water-column is the understanding of biophysical interaction  driven by an agglomeration of physical forcing (wind, topography, bathymetry, tidal influences and local geography such as fresh water inflows) mixed with incipient micro-biology driven by planktonic and coastal anthropogenic input, such as pollution and agricultural runoff, which results in a range of ecosystem phenomena (e.g.
-blooms, anoxic zones, plumes) ---all influences with significant impact both directly and indirectly on society.
-Reverberating the requirement of more effective ocean sensing practices, renowned oceanographer Walter Munk referred to the 20\textsuperscript{th} century as the \emph{century of undersampling} \citep{munk2002}, pointing out that lack of sufficient sampling resolution in time and space, as a critical issue for characterizing ocean phenomena.
-Still today, decision makers struggle with limited observational efforts sometimes leading to severe environmental problems.
-The recent (May 2018) outbreak of harmful algal blooms in Northern Norway\footnote{Report from the Institute of Marine Research:https://bit.ly/2KZGJfI}, with losses exceeding 12 million tons of salmon valued at \$ 80 million (Norwegian Directorate of Fisheries), is a case in point in how adequate sampling efforts could have provided a more accurate warning assessment.
-These challenges of undersampling have been driving ocean sensing practices towards more networked and distributed use of robotic assets in order to fill the gaps in our understanding (See Fig.
-\ref{fig:envir}), where the surveying is guided by a model and the measurements made onboard the robot are used to update the model.
+Central to understanding the changes taking place in the upper
+water-column is the understanding of biophysical interaction driven by
+an agglomeration of physical forcing (wind, topography, bathymetry,
+tidal influences and local geography such as fresh water inflows)
+mixed with incipient micro-biology driven by planktonic and coastal
+anthropogenic input, such as pollution and agricultural runoff. These
+often result in a range of ecosystem phenomena (e.g. blooms, anoxic
+zones, plumes) all with significant impact both directly and
+indirectly on society. Reverberating the requirement of more effective
+ocean sensing practices, Walter Munk referred to the
+20\textsuperscript{th} century as the \emph{century of undersampling}
+\citep{munk2002}, pointing out that lack of sufficient sampling
+resolution in time and space, as a critical issue for characterizing
+ocean phenomena. Even today policy makers struggle with limited
+observational efforts sometimes leading to severe environmental
+problems such as the recent (May 2018) outbreak of harmful algal
+blooms in Northern Norway\footnote{Report from the Institute of Marine
+  Research:https://bit.ly/2KZGJfI}, with losses exceeding 12 million
+tons of salmon valued at \$ 80 million (Norwegian Directorate of
+Fisheries). This is a case in point in how adequate sampling efforts
+could have provided a more accurate assessment and prior
+warning. These challenges related to undersampling have been driving
+ocean sensing practices towards more autonomous, networked and
+distributed use of robotic assets in order to fill the gaps in our
+understanding (See Fig. \ref{fig:envir}), where the surveying is
+guided by a dynamical ocean model and the measurements made onboard
+the robot are often used to update the model.
 
@@
   \centering 
-  \subfigure[Modern ocean sensing network, where robotic assets augment ship based measurements to increase coverage of the phenomena.]{\includegraphics[width =
+  \subfigure[Illustration of an ocean sensing network, where robotic
+  assets augment ship based measurements to increase coverage and
+  resolution of phenomena of scientific interest.]{\includegraphics[width =
     0.48\textwidth]{Figures/envir.pdf}\label{fig:envir1}}
@@
     0.48\textwidth]{Figures/adaptive_cycle.pdf}\label{fig:sense-plan-act}}
-  \caption{Ocean observation is moving away from single-ship sampling towards more autonomous and collaborative networked operations in order to resolve the numerous processes and their interaction.} \label{fig:envir}
+  \caption{Ocean observation is moving away from single-ship sampling
+    towards more autonomous and collaborative networked operations in
+    order to resolve the numerous bio-geochemical processes and their
+    interactions.}
+\label{fig:envir}
 \end{figure}
 
-Focusing on coupling statistics and robotic sampling, our attention is in this work towards spatial characterization of a frontal systems, and more specifically fronts created by river plume systems.
-Fig.
-\ref{fig:nidelven} shows the survey area (the outlet of the Nidelva river, Trondheim, Norway) taken as cold freshwater enters from the river creating a strong separation gradient in both temperature and salinity.
-Because of the local topography and the Coreolis force \citep{coriolis1835memoire} the cold fresh water tends to flow near land to the east.
-Depending on the river discharge, tidal effects, wind and temperature differences, this boundary in water masses will get distorted.
-Prior knowledge about the location and evolution of these features is therefore highly uncertain, making deterministic planning sub-optimal.
+Our attention in this work is towards spatial characterization of a
+frontal system, and more specifically fronts generated by river
+plumes. Fig. \ref{fig:nidelven} shows the survey area (the outlet of
+the Nidelva river, Trondheim, Norway) taken as cold freshwater enters
+from the river creating a strong gradient in both temperature and
+salinity. Because of the local topography and the Coreolis force
+\citep{coriolis1835memoire} the cold fresh water tends to flow near
+land to the east. Depending on the river discharge, tidal effects,
+wind and temperature differences, this boundary in water masses often
+gets distorted. Prior knowledge about the location and evolution of
+these features is therefore highly uncertain, making deterministic
+planning challenging.
 
@@
 \includegraphics[width=0.85\textwidth]{Figures/pictures/c-updated.pdf}
-\caption{Frontal patterns - Nidelven, Trondheim, Norway.}
+\caption{Frontal patterns off of the Nidelva river, Trondheim, Norway.}
 \label{fig:nidelven}
@@
 
-River plumes belong to a class of ocean processes that are local and usually act on smaller \emph{sub-mesoscale}, from 10 m -- 5 km where spatial lateral variability tends to dominate.
-Remote sensing and synthetic ocean models find it challenging to provide detailed at this scale \citep{Lermusiaux:2006} and the principal way to resolve for fine resolution is by direct observation.
-At larger \emph{mesoscale} ($>50$ km), both three-dimensional space and time dynamics are important and can shift substantially, while in the case of river plumes one can often limit scope to cover only lateral spatial elements ([north, east] domain).
-Moreover, for this case, time-effects can be regarded as static when acquiring AUV data over a short time window and limited region.
-However, the methodological framework can be extended to higher dimensional situations in a similar manner.
+River plumes belong to a class of ocean processes that are local and
+at smaller \emph{sub-mesoscale}, from 10 m -- 5 km where spatial
+lateral variability tends to dominate. Remote sensing and synthetic
+ocean models find it challenging to provide detail at this scale
+\citep{Lermusiaux:2006} and the principal way to resolve for fine
+resolution is by direct observation. At larger \emph{mesoscale}
+($>50 km^2$), both three-dimensional space and time dynamics are
+important and can shift substantially, while in the case of river
+plumes one can often limit scope to cover only lateral spatial
+elements ([north, east] domain). Moreover, for this case, time-effects
+can be regarded as static when acquiring AUV data over a short time
+window and limited region. However, the methodological framework can
+be extended to higher dimensional situations in a similar manner.
 
@@
 from 100--6000m with in-water staying capacity depending on survey
-speed, payload sensors and mission design. By deploying AUVs one can in
-principle fill gaps and augment data obtained from either ocean models,
-earth observing satellite data, or fixed-location sensors on buoys due
-to their speed and mobility. Traditional operations and surveys with
-AUVs are usually limited to planning samples along fixed transects or survey lines on a spatial grid, pre-programmed by the human operator to 
-consist of different waypoints on the grid. But by using on-board
-algorithms to continuously evaluate, update, and refine future sampling
-locations, one makes information gathering adaptive \citep{das11b}.
-The space of sampling opportunities is still limited to the spatial grid, but now there is much more flexibility in automatic adaption.
-For instance, an AUV could reconstruct or modify a survey line
-based on what temperatures it is measuring by using embedded methods in
-decision-theoretic planning \citep{Rajan12b}.
+speed, payload sensors and mission design. By deploying AUVs one can
+in principle fill gaps and augment data obtained from either ocean
+models, earth observing satellite data, or fixed-location sensors on
+buoys due to their speed and mobility. Traditional operations and
+surveys with AUVs are usually limited to planning samples along fixed
+transects or survey lines on a spatial grid, pre-programmed by the
+human operator to consist of different waypoints on the grid. But by
+using on-board algorithms to continuously evaluate, update, and refine
+future sampling locations, one makes information gathering adaptive
+\citep{das11b}.  The space of sampling opportunities is still limited
+to the spatial grid, but now there is much more flexibility in
+automatic adaption.  For instance, an AUV could reconstruct or modify
+a survey line based on what temperatures it is measuring by using
+embedded methods in decision-theoretic planning \citep{Rajan12b}.
 

### 2019-08-26 main_new.tex
updates

@@
 River plumes belong to a class of ocean processes that are local and
-at smaller \emph{sub-mesoscale}, from 10 m -- 5 km where spatial
-lateral variability tends to dominate. Remote sensing and synthetic
-ocean models find it challenging to provide detail at this scale
-\citep{Lermusiaux:2006} and the principal way to resolve for fine
-resolution is by direct observation. At larger \emph{mesoscale}
+at smaller \emph{sub-mesoscale} (from $10 m^2$ -- $5 km^2$ where
+spatial lateral variability tends to dominate. Remote sensing and
+synthetic ocean models find it challenging to provide detail at this
+scale \citep{Lermusiaux:2006} and the principal way to resolve for
+fine resolution is by direct observation. At larger \emph{mesoscale}
 ($>50 km^2$), both three-dimensional space and time dynamics are

### 2019-09-05 main_new.tex
updates

@@
 anthropogenic input, such as pollution and agricultural runoff. These
-often result in a range of ecosystem phenomena (e.g. blooms, anoxic
-zones, plumes) all with significant impact both directly and
-indirectly on society. Reverberating the requirement of more effective
-ocean sensing practices, Walter Munk referred to the
+often result in a range of ecosystem related phenomena such as blooms,
+anoxic zones and plumes, all with significant impact both directly and
+indirectly on society. Walter Munk referred to the
 20\textsuperscript{th} century as the \emph{century of undersampling}
@@
 ocean phenomena. Even today policy makers struggle with limited
-observational efforts sometimes leading to severe environmental
-problems such as the recent (May 2018) outbreak of harmful algal
-blooms in Northern Norway\footnote{Report from the Institute of Marine
+observability of severe environmental problems such as the recent (May
+2018) outbreak of harmful algal blooms in Northern
+Norway\footnote{Report from the Institute of Marine
   Research:https://bit.ly/2KZGJfI}, with losses exceeding 12 million
@@
 
-Our attention in this work is towards spatial characterization of a
+Our focus in this work is towards spatial characterization of a
 frontal system, and more specifically fronts generated by river
 plumes. Fig. \ref{fig:nidelven} shows the survey area (the outlet of
-the Nidelva river, Trondheim, Norway) taken as cold freshwater enters
+the Nidelva river, Trondheim, Norway) where cold freshwater enters
 from the river creating a strong gradient in both temperature and
@@
 land to the east. Depending on the river discharge, tidal effects,
-wind and temperature differences, this boundary in water masses often
-gets distorted. Prior knowledge about the location and evolution of
-these features is therefore highly uncertain, making deterministic
-planning challenging.
+wind and temperature differences, this boundary often gets
+distorted. Prior knowledge about the location and evolution of these
+features is therefore highly uncertain, making deterministic planning
+challenging.
 
@@
 River plumes belong to a class of ocean processes that are local and
-at smaller \emph{sub-mesoscale} (from $10 m^2$ -- $5 km^2$ where
+at smaller \emph{sub-mesoscale} (from $10 m^2$ -- $5 km^2$) where
 spatial lateral variability tends to dominate. Remote sensing and
@@
 plumes one can often limit scope to cover only lateral spatial
-elements ([north, east] domain). Moreover, for this case, time-effects
-can be regarded as static when acquiring AUV data over a short time
-window and limited region. However, the methodological framework can
-be extended to higher dimensional situations in a similar manner.
+elements. Moreover, for this case, time-effects can be regarded as
+static when acquiring AUV data over a short time window and limited
+region. However, the methodological framework can be extended to
+higher dimensional situations in a similar manner.
 
@@
 
-AUVs operate underwater used to gather spatio-temporal data from a
-targeted domain, carrying a scientific payload to survey the water
-column or the seafloor. They run at $\sim 1-3$ m/s, can reach depths
-from 100--6000m with in-water staying capacity depending on survey
-speed, payload sensors and mission design. By deploying AUVs one can
-in principle fill gaps and augment data obtained from either ocean
-models, earth observing satellite data, or fixed-location sensors on
-buoys due to their speed and mobility. Traditional operations and
-surveys with AUVs are usually limited to planning samples along fixed
-transects or survey lines on a spatial grid, pre-programmed by the
-human operator to consist of different waypoints on the grid. But by
-using on-board algorithms to continuously evaluate, update, and refine
-future sampling locations, one makes information gathering adaptive
-\citep{das11b}.  The space of sampling opportunities is still limited
-to the spatial grid, but now there is much more flexibility in
-automatic adaption.  For instance, an AUV could reconstruct or modify
-a survey line based on what temperatures it is measuring by using
-embedded methods in decision-theoretic planning \citep{Rajan12b}.
+AUVs are used to gather in-situ spatio-temporal data from a targeted
+domain while carrying a range of scientific payloads to survey the
+water column.  Typically, they operate at $\sim 1-3$ m/s, can reach
+depths from 100--6000m with an in-water staying capacity depending on
+survey speed, payload sensors and mission design. By deploying AUVs
+one can fill gaps in data gathering and augment information obtained
+from either ocean models, remote-sensing data, or fixed-location
+sensors on buoys. Traditional operations and surveys with AUVs are
+usually limited to observations along fixed transects on a spatial
+grid, pre-programmed by the human operator.  By using on-board
+algorithms to continuously evaluate, update, and refine future
+sampling locations, one can make information gathering adaptive and
+contextual to the survey requirements
+\citep{das11b,fossuminformation,fossum18b}.  The space of sampling
+opportunities is still limited to the spatial grid, but now there is
+much more flexibility because of adaptation.  For instance, an AUV
+could reconstruct or modify a survey line based on what temperatures
+it is measuring by using embedded methods in decision-theoretic
+planning \citep{py10,Rajan12,Rajan12b}.
 
@@
 Adaptive sampling of an evolving frontal feature has been explored in
-\cite{fronts11,Zhang2012,Pinto2018,costa19} using state-based autonomous
-agents. These approaches typically use a reactive-adaptive scheme,
-whereby exploration does not rely on a statistical model of the
-environment, but rather adapts based on closing the sensing and
-actuation loop (Fig. \ref{fig:sense-plan-act}). Myopic sampling, i.e. step-wise selection of the path
-(on the graph) which reduces the information criterion the most, has
-been used for adaptive surveys \citep{singh2009efficient,Binney2013}
-which focus largely on reducing predictive variance or mutual
-information (entropy). Variance and entropy reduction are
-independent of the actual data realizations under the assumption of GP models, and the use of data-driven (adaptive) criteria was introduced to include more targeted sampling of regions of scientific interest by
-\cite{Low2009} and \cite{fossuminformation}.
-In this paper, with the focus on mapping the river plume, designs that improve the classification of water masses should be rewarded.
+\cite{fronts11,Zhang2012,Pinto2018,costa19}. These approaches
+typically use a reactive-adaptive scheme, whereby exploration does not
+rely on a statistical model of the environment, but rather adapts
+based on closing the sensing and actuation loop
+(Fig. \ref{fig:sense-plan-act}). Myopic sampling, i.e. step-wise
+selection of the path (on the graph) which reduces the information
+criterion the most, has been used for adaptive surveys
+\citep{singh2009efficient,Binney2013} which focus largely on reducing
+predictive variance or mutual information (entropy). Variance and
+entropy reduction are independent of the actual data realizations
+under the assumption of GP models, and the use of data-driven adaptive
+criteria was introduced to include more targeted sampling of regions
+of scientific interest by \cite{Low2009} and \cite{fossuminformation}.
+In this paper, with the focus on mapping the river plume, we reward
+the designs that improve the classification of water masses, as a
+means to characterize the frontal zone.
 
@@
 
-The objective is to characterize the river plume, focusing on spatial separation of cold freshwater from the warmer saline waters of the fjord.
-ESs and EPs are useful starting points to measure the ability to classify water masses, and the goal of experimental sampling is to improve this characterization, as quantified via changes in the EPs.
-
-We consider a lateral domain, near sea surface, with locations $\bx \in \mathcal{M} \subset \mathbf{R^2}$.
-The two variables of interest, temperature and salinity, are defined as random processes; $\xi_a(\bx)$ denotes the temperature (in $^o C$) and $\xi_b(\bx)$ denotes the salinity (in $mg/l$) at location $\bx$.
-We let $t_a$ be
-a threshold for temperature and $t_b$ a threshold for salinity, the bivariate ES is then defined by
+Our objective is to characterize a river plume, focusing on spatial
+separation of cold freshwater from the warmer saline waters of the
+fjord. ESs and EPs are useful starting points to measure the ability
+to classify water masses, and the goal of experimental sampling is to
+improve this characterization, as quantified via changes in the EPs.
+
+We consider a lateral domain, near sea surface, with locations
+$\bx \in \mathcal{M} \subset \mathbf{R^2}$. The two variables of
+interest, temperature and salinity, are defined as random processes;
+$\xi_a(\bx)$ denotes the temperature (in $^o C$) and $\xi_b(\bx)$
+denotes the salinity (in $mg/l$) at location $\bx$.  We let $t_a$ be a
+threshold for temperature and $t_b$ a threshold for salinity, the
+bivariate ES is then defined by
  \begin{equation}\label{ES}

### 2019-09-06 main_new.tex
updates

@@
 separation of cold freshwater from the warmer saline waters of the
-fjord. ESs and EPs are useful starting points to measure the ability
-to classify water masses, and the goal of experimental sampling is to
-improve this characterization, as quantified via changes in the EPs.
-
-We consider a lateral domain, near sea surface, with locations
+fjord. Excursion sets (ES) and excursion probabilities (EP) are useful
+starting points to measure the ability to classify water masses. The
+goal of experimental sampling is to improve this characterization, as
+quantified via changes in the EPs.
+
+We consider a lateral domain, near the sea surface, with locations
 $\bx \in \mathcal{M} \subset \mathbf{R^2}$. The two variables of
@@
 bivariate ES is then defined by
+
 \begin{equation}\label{ES}
@@
 The excursions could also be defined as larger than a threshold, or
- within a boundary of limits for the two variables. Calculations and
- interpretations will be similar in these different cases.
+within a boundary of limits for the two variables; calculations and
+interpretations will be similar in these cases.
 
@@
 challenging if the probabilities are close to $0.5$. 
-The integrated Bernoulli variance (IBV), see \cite{bect2019}, is here defined as 
+The integrated Bernoulli variance (IBV) \cite{bect2019}, is defined as 
 \begin{equation}\label{measVAR}
@@
 The measure is dominated by locations with EPs close to $0.5$.
-
-If prior knowledge strongly indicates cold freshwater in a part of the domain, it is easy to classify, and there is likely little value in exploring that part much.
-Similarly with strong prior knowledge about warm saltwater.
-Data can be gathered at selected locations of the domain $\mathcal{M}$, to improve classification skill.
-We denote temperature measurements by $y_{a}(\bx_i)$ and salinity measurements by $y_{b}(\bx_i)$, at a sampling location $\bx_i \in \mathcal{M}$.
-Given a set of measurements, denoted $\by$, the conditional EPs are
+\kc{This sentence is a bit strange; just about, you state that EPs
+  close to 0.5 can be problematic since they do not categorize the
+  water masses well. And here you state that there is a strong
+  dominance of the average. So now you should say how the biasing can
+  change to make ``V'' more effective. Otherwise this sentence is
+  dangling.}
+
+If prior knowledge provides a strong indication of cold freshwater
+(for instance) in a part of the domain, it is easy to classify, and
+there is likely little value in exploring those locations. Data can be
+gathered at selected locations of the domain $\mathcal{M}$, to improve
+classification skill. We denote temperature measurements by
+$y_{a}(\bx_i)$ and salinity measurements by $y_{b}(\bx_i)$, at a
+sampling location $\bx_i \in \mathcal{M}$. Given a set of
+measurements, denoted $\by$, the conditional EPs are
 \begin{equation}\label{eq:post_ep}
@@
 
-Our goal is to construct designs for the data gathering, that are expected to improve the mapping of the water masses. 
-When designing AUV sampling designs, it is hence natural to evaluate the expected reduction in the IBV, as provided by the survey data.
-In this way, we take the expectation with respect to the planned measurements $\by$;
+Our goal is to construct designs for data gathering, that are expected
+to improve the mapping and separation of water masses. When designing
+AUV sampling designs, it is therefore natural to evaluate the expected
+reduction in the IBV, as provided by the survey data. In this way, we
+take the expectation with respect to the planned measurements $\by$;
 \begin{equation}\label{sur}
@@
 
-Recent works in statistical literature on ESs and EPs connected to
-spatial statistics include
-\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}. Our focus
-is on ES as defined in Eq. (\ref{ES}), and the
-uncertainty reduction achieved by sampling as in Eq. (\ref{sur}).  In this sense
+Recent work in literature on ESs and EPs connected to spatial
+statistics include
+\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
+Our focus is on ES as defined in Eq. (\ref{ES}), and the uncertainty
+reduction achieved by sampling as in Eq. (\ref{sur}). In this sense
 our work is similar to \cite{bect2012}, \cite{chevalier2014fast} and
 \cite{azzimonti2016quantifying} who describe analytical results for
-uncertainty reduction in ESs for univariate processes.  
-A major contribution of our work is to derive closed-form results for this design criteria in the situation with multivariate GPs.
+uncertainty reduction in ESs for univariate processes. A major
+contribution of our work is to derive closed-form results for this
+design criteria in the situation with multivariate GPs.
 

### 2019-09-06 main_new.tex
updates

@@
 $\xi_a(\bx)$ denotes the temperature (in $^o C$) and $\xi_b(\bx)$
-denotes the salinity (in $mg/l$) at location $\bx$.  We let $t_a$ be a
+denotes the salinity (in $mg/l$) at location $\bx$. \kc{might be
+  clearer if you used 't' for temp and 's' for salinity. 'a' and 'b'
+  are somewhat opaque especially further in the text}.  We let $t_a$ be a
 threshold for temperature and $t_b$ a threshold for salinity, the
-bivariate ES is then defined by
+bivariate ES is then defined by:
 
@@
 
-The ES holds random indicator variables, while the associated EPs are 
+The ES holds random indicator variables, while the associated EPs are: 
 \begin{equation}\label{eq:prob}
@@
 challenging if the probabilities are close to $0.5$. 
-The integrated Bernoulli variance (IBV) \cite{bect2019}, is defined as 
+The integrated Bernoulli variance (IBV) \cite{bect2019}, is defined as: 
 \begin{equation}\label{measVAR}
@@
 sampling location $\bx_i \in \mathcal{M}$. Given a set of
-measurements, denoted $\by$, the conditional EPs are
+measurements, denoted $\by$, the conditional EPs are:
 \begin{equation}\label{eq:post_ep}
@@
 where the probability is defined in Eq. (\ref{eq:post_ep}).
-Considering a set $\mathcal{J}$ of possible designs for an AUV survey. A criterion for the selection is
+Considering a set $\mathcal{J}$ of possible designs for an AUV survey. A criterion for the selection is:
 \begin{equation}\label{crit}
@@
 \end{equation}
-where the criterion $V_{\mbox{upd}}(j)$ in Eq. (\ref{sur}) is computed for each of the possible designs.
-
-Recent work in literature on ESs and EPs connected to spatial
-statistics include
+where the criterion $V_{\mbox{upd}}(j)$ in Eq. (\ref{sur}) is computed
+for each of the possible designs.
+
+Recent work on ESs and EPs connected to spatial statistics include
 \cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
@@
 
-The actual calculations of the EPs and the IBV require a model
-specification. Here, Gaussian modeling assumptions are used, as described next. They allow closed form evaluation of the IBVs, as derived in Section \ref{sec:sur}.
+Calculations of the EPs and the IBV require a model specification.
+Here, we use Gaussian modeling assumptions which allow closed form
+evaluation of the IBVs, as derived in Section \ref{sec:sur}.
 
@@
 
-We represent the bivariate random fields of sea temperature and
-salinity by a Gaussian model.  At location $\bx \in \mathcal{M}$, the bivariate distribution is
+We represent the bivariate random fields of temperature and salinity
+with a Gaussian model. At location $\bx \in \mathcal{M}$, the
+bivariate distribution is:
 \begin{equation}\label{gp_bar}
   \begin{bmatrix}\xi_a(\bx) \\
-\xi_b(\bx) 
-\end{bmatrix}
+    \xi_b(\bx) \end{bmatrix}
  \sim N \left( 
@@
 \end{equation}
-where the notation means Gaussian (Normal) distributed variables specified by the mean vector and covariance matrix.
-In the application below, the model parameters are specified from preliminary data: The mean varies with location, capturing the variability near the river mouth, while the variance parameters and correlation between salinity and
-temperature are assumed to be constant for all locations.
+where the notation refers to Gaussian (Normal) distributed variables
+specified by the mean vector and covariance matrix. In the application
+below, the model parameters are specified from preliminary data: The
+mean varies with location, capturing the variability near the river
+mouth, while the variance parameters and correlation between salinity
+and temperature are assumed to be constant for all locations. \kc{?
+  Isn't that too strong an assumption? T/S tend to be correlated,
+  however for uniform bodies of water; here we're dealing with coastal
+  ecosystems where the two are not necessarily so.}
 
@@
 we assume a separable correlation function with the same decay for salinity and temperature,
-which is not unrealistic for a river plume. We have
+which is not unrealistic for a river plume \kc{why so?}. We have:
 \begin{equation}\label{gp_corr}
@@
 \end{equation}
-for $i,j \in {a,b}$ and $\delta_{ij}=1$ if $i=j$ and $\delta_{ij}=0$ otherwise.
-Moreover, we assume a stationary isotropic where the correlation function
-$\rho(\bx,\bx')$ solely depends on $\bx$ and $\bx'$ through their Euclidean distance $h=\sqrt{|\bx-\bx'|^2}$.
-With lots of data and prior knowledge one could possibly fit and estimate parameters of non-stationary or non-isotropic covariance function and more complex multivariate spatial covariance functions
-\citep{gneiting2010matern,genton2015cross}, but we do not consider that here.
-
-For implementation on the computer we discretize the spatial domain $\mathcal{M}$ to a
-set of $n$ grid locations $\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell has area $\Delta$.
-The integral expression in Eq.
-(\ref{measVAR}) and (\ref{sur}) are then approximated by sums over all grid cells.
-This grid is identical to the possible set of design locations, as mentioned previously, defined as a waypoint graph.
-Using vector notation, we denote the Gaussian temperature and salinity variables on the grid and its density function by
+for $i,j \in {a,b}$ and $\delta_{ij}=1$ if $i=j$ and $\delta_{ij}=0$
+otherwise. Moreover, we assume a stationary isotropic property
+\kc{This is probably related to the assumption as above where T/S are
+  assumed to be constant in all locations? But why?} where the
+correlation function $\rho(\bx,\bx')$ solely depends on $\bx$ and
+$\bx'$ via a Euclidean distance measure $h=\sqrt{|\bx-\bx'|^2}$. With
+data and prior knowledge one could possibly fit and estimate
+parameters of non-stationary or non-isotropic covariance functions and
+more complex multivariate spatial covariance functions
+\citep{gneiting2010matern,genton2015cross}; however that is left for
+future work.
+
+For implementation purposes, we discretize the spatial domain
+$\mathcal{M}$ to a set of $n$ grid locations
+$\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell has area
+$\Delta$. The integral expression in Eq. (\ref{measVAR}) and
+(\ref{sur}) are then approximated by sums over all grid cells. This
+grid is identical to the possible set of design locations and defined
+as a waypoint graph \kc{You mean the
+  centroids of the cells are part of the graph?}. Using vector
+notation, we denote the Gaussian temperature and salinity variables on
+the grid and its density function by: 
 \begin{equation}\label{prior}
@@
 %Here, $\pi(\bxi)$ is the joint density function of length $2 n$ vector $\bxi$, and
-Here, the length
-$2 n$ vector $\bmu$ holds the expected values of temperature and
-salinity at the grid locations, while the $2n \times 2n$ positive definite  matrix $\bSigma$ holds the variance-covariances of temperature and salinity variables at all grid locations.
+The length $2 n$ vector $\bmu$ holds the expected values of
+temperature and salinity at the grid locations, while the
+$2n \times 2n$ positive definite matrix $\bSigma$ holds the
+variance-covariances of temperature and salinity variables at all grid
+locations.
 

### 2019-09-11 main_new.tex
updates

@@
 
-Data are in our setting gathered by an autonomous vehicle that can measure
-temperature and salinity at a subset of locations on the defined grid.
-This can be done in a static manner, where the survey design is pre-planned, or in a sequential way with several rounds of data gathering.
-For the latter, a round of data is gathered first, and then assimilated to get an updated model.
-Then the updated model is used to design and select the next round of measurements, and so on.
-We here describe the model for one round of data and the Gaussian formula for one round of updating. In Section
-\ref{sec:heuristics} we discuss sequential data gathering in the context of adaptive design of AUV measurements.
-
-Denote the data at $n_y$ locations by
+Data is gathered by an AUV that can measure temperature and salinity
+at a subset of locations on the defined grid. This can be done in a
+static manner, where the survey design is pre-planned, or in a
+sequential way with several rounds of data gathering. For the latter,
+a round of data is gathered first, and then assimilated to get an
+updated model. Then the updated model is used to design and select the
+next round of measurements, and so on. We describe the model for one
+round of data and the Gaussian formula for a single update. In Section
+\ref{sec:heuristics} we discuss sequential data gathering in the
+context of adaptive design of AUV measurements.
+
+We denote the data at $n_y$ locations by
 $\by=(y_{a,1},y_{b,1},\ldots,y_{a,n_y},y_{b,n_y})^t$, representing the
-temperature and salinity measurements gathered at one round grid nodes. The conditional model for the data, given the true temperature and salinity, is described by
+temperature and salinity measurements gathered at the first round of
+nodes in the grid. The conditional model for the data, given the true temperature
+and salinity, is described by: 
 \begin{equation}\label{likelihood}
@@
 where $\bA$ is an $2n_y \times 2n$ matrix holding '$1$'s on the
-measurement indices on the grid, otherwise '$0$'. The size
-$2n_y \times 2n_y$ covariance matrix $\bR$ contains diagonal entries $r^2_a$ and
-$r^2_b$ which are measurement error variances of temperature and
-salinity observations.
-
-The GP mean and covariance in Eq. (\ref{prior}) are moderated when more information gets available. 
-The updated distribution for temperature and salinity variables on the
+measurement indices on the grid and '$0$' otherwise. The size
+$2n_y \times 2n_y$ covariance matrix $\bR$ contains diagonal entries
+$r^2_a$ and $r^2_b$ which are measurement error variances of
+temperature and salinity observations.
+
+The GP mean and covariance in Eq. (\ref{prior}) are moderated \kc{What
+  does 'moderated' mean?} when more information is available. The
+updated distribution for temperature and salinity variables on the
 spatial grid, given survey data $\by$, is Gaussian with mean and
-covariance 
+covariance: 
 \begin{eqnarray}\label{gp_upd}
   \bm &=& \bm(\by) = \bmu+\bSigma \bA^t (\bA \Sigma \bA^t+\bR)^{-1}(\by-\bA \bmu),  \\
-\bS &=& \bSigma - \bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.\nonumber
+  \bS &=& \bSigma - \bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA
+          \bSigma.\nonumber
 \end{eqnarray}
 
-The updated bivariate distribution at a grid location $\bx \in \mathcal{M}_g$ is then 
+The updated bivariate distribution at a grid location $\bx \in
+\mathcal{M}_g$ is then: 
 \begin{equation}\label{gp_hat}

### 2019-09-12 main_new.tex
Small changes Simulation study

@@
   information obtained from measurements taken \emph{in-situ}, this is
-  usually referred to as the \emph{adaptivity gap} \kc{citation
-    needed}.
+  usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.
 \item For such types of sampling problems, the number of choices (i.e.
@@
 Building on this result, \cite{chekuri2005recursive} explored this in the setting of a graph using a recursive greedy algorithm, providing near-optimal solutions depending on the planning horizon and graph resolution.
+
 Adaptive sampling of an evolving frontal feature has been explored in
@@
 This situation mimics that of a river mouth entering from the south, and the water masses are pulled to the east.
-The actual trends are in practiece set from preliminary information or runs of oceanographic models solving the differential equations for fluid flow.
+The actual trends are in practice set from preliminary information or runs of oceanographic models solving the differential equations for fluid flow.
 In this example the mean is specified by $\beta_{a,1}=0.065$ and $\beta_{b,1}=0.1$.
@@
   \subfigure[Excursion probabilities given collected data along the strategy ``static\_north".]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
-  \caption{(\ref{fig:true_temp}) and (\ref{fig:true_sal}) shows
+  \caption{\ref{fig:true_temp} and \ref{fig:true_sal} shows
     realisations of the simulated temperature and salinity fields, as
-    well as the associated excursion set (\ref{fig:ESet}).
-    (\ref{fig:Eprob}) show the estimated excursion probabilities after
+    well as the associated excursion set \ref{fig:ESet}.
+    \ref{fig:Eprob} show the estimated excursion probabilities after
     performing data collection along the N-S survey line.} 
@@
 
+\newpage
 \subsection{Static and sequential sampling designs}\label{sec:sampling_designs}
@@
 
-We compare three different static designs denoted \texttt{static\_north}, \texttt{static\_east and} \texttt{static\_zigzag} with the three sequential approaches \texttt{naive}, \texttt{myopic} and \texttt{look-ahead}.
-The static sampling paths are pre-determined, and cannot adapt to data.
-
-For a fixed length, the expected variance after static surveying is analytically available, but here we focus on the properties of the sequential designs which are evaluated using Monte Carlo sampling over several replicates of realizations from the prior, and conducting simulated sequential surveying for each one.
+We compare three different static designs denoted \texttt{static\_north}, \texttt{static\_east and} \texttt{static\_zigzag} with the three sequential approaches \texttt{naive}, \texttt{myopic}, and \texttt{look-ahead}.
+The static sampling paths are pre-determined and cannot be changed on the basis of data, representing the archetypical pre-planned observation strategies used in current observational practices.
+For a fixed survey length, the expected variance after static surveying is analytically available (equal coverage each time), but here we focus on the properties from the sequential designs which are evaluated using Monte Carlo sampling over several replicates of realizations from the prior, and conducting simulated sequential surveying for each one; making the result dependent on the observed data.
 For comparison, the same setup is used for the static designs.
@@
 Along with this criterion we also compare predictive performance measured by root mean square error (RMSE) in the temperature and salinity fields and variance reduction in these two field variables.
-Note that objective function used by the agents is focused on reducing the expected variance in the ES, but we nevertheless hope to achieve good predictive performance for the other criteria as well.
-Another non-statistical criteria that is important for practical purposes is the computing time of the strategy.
-
-The agent of the sequential designs starts at the center East-West coordinate at the southern end of the domain (node 53), see Fig. \ref{fig:wp_graph}.
+It is important to note that the objective function used by the agents is focused on reducing the expected variance in the ES, but we nevertheless hope to achieve good predictive performance for criteria such as RMSE as well.
+Another non-statistical criteria that is important for practical purposes is the computing time of the strategy, as this will impact the performance of the approach in the field for an embedded system.
+
+The sequential sampling agents start at the center East-West coordinate at the southern end of the domain (node 53), see Fig. \ref{fig:wp_graph}.
 
@@
 
-At each step, the AUV will choose to move along an edge in this waypoint
-graph. While doing so, it gathers three measurements. These are
-assimilated in the model, and then a new choice is done at the end of
-the edge. The paths of will differ for the various sequential designs.
-The designs will also differ between the runs, because the measurements
-vary among replicates.
+Maintaining the AUV example, the AUV will move along edges in the Fig.
+\ref{fig:wp_graph} waypoint graph.
+While doing so, the AUV will gather measurements.
+These are assimilated in the GP model, before an evaluation of the next node to sample will be conducted at the end of the edge.
+The paths of will differ for the various sequential designs due to the individual design criteria and the simulated variability in the environment among the replicates.
 
@@
 distance. Here, expected ES variance in Figure \ref{fig:avg_ev} is
-reduced most when using the \texttt{myopic} and \texttt{look-ahead}
-strategies. There is not much difference between these two.
+reduced most under the \texttt{myopic} and \texttt{look-ahead}
+strategies, performing almost equally.
 

### 2019-09-13 main_new.tex
updates

@@
   clearer if you used 't' for temp and 's' for salinity. 'a' and 'b'
-  are somewhat opaque especially further in the text}.  We let $t_a$ be a
+  are somewhat opaque especially further in the text. Also there
+  should be a statement about why we're considering only 2 variables
+  and the impact to considering (algebraically) more in the following
+  formulations. This will surely be asked by any reviewer.}.  We let $t_a$ be a
 threshold for temperature and $t_b$ a threshold for salinity, the
@@
 % derivation in Section \ref{sec:sur}.
-The calculation of the design criterion in Eq. (\ref{sur}) requires the marginal distribution of the data which is defined via marginalization over $\bxi$ to be
-$\by \sim N( \bA \bmu,\bA \bSigma \bA^t+\bR)$. 
-However, it turns out that a simplified form of the calculation is possible because the 
-conditional mean in Eq (\ref{gp_hat}) is a linear (affine) function of the data $\by$ and the covariance is not a function of the data.
+The calculation of the design criterion in Eq. (\ref{sur}) requires
+the marginal distribution of the data which is defined over $\bxi$ to
+be $\by \sim N( \bA \bmu,\bA \bSigma \bA^t+\bR)$.  However, it turns
+out that a simplified form of the calculation is possible because the
+conditional mean in Eq (\ref{gp_hat}) is a linear (affine) function of
+the data $\by$ and the covariance is not a function of the data.
 Before knowing the data, the distribution of the conditional mean is
@@
 
-\section{Uncertainty reduction on excursion sets}
+\section{Uncertainty reduction in excursion sets}
 \label{sec:sur}
@@
 
-Based on the Gaussian modeling assumptions, the EPs can be computed at any location $\bx$. Without
-additional data, Eq.
-(\ref{eq:prob}), we have
+Based on Gaussian modeling assumptions, the EPs can be computed at any
+location $\bx$. Without additional data, Eq.  (\ref{eq:prob}) gives us:
+
 \begin{eqnarray}\label{eq:prob_mv0}
@@
 
-When data $\by$ are available, Eq. (\ref{eq:post_ep}), we similarly get
+When data $\by$ are available, Eq. (\ref{eq:post_ep}) similarly gives
+us:
 \begin{eqnarray}\label{eq:prob_mv}
@@
 
-We will next derive closed form solutions to the expectation in (\ref{sur}).
-The location index $\bx$ is suppressed to avoid overly complex notation.
-We present new results for computing the expectation $E_{\by}\left\{ \cdot \right\}$ which is the integrand of Eq.
-(\ref{sur}).
-
-A critical first
-element in the derivation is the GP model, and the linear dependence on data $\by$ in the conditional mean in Eq.
-(\ref{gp_upd}), which means that we have a closed form Gaussian distribution for the conditional mean in Eq.
-(\ref{dist_mxi}).
-The high-dimensional inner integral in Eq.
-(\ref{sur}) then reduces to a bivariate integral \citep{bhattacharjya2013value, chevalier2014fast}, with expectation over $\bm=(m_{a},m_{b})^t$.
-We standardize the two variables in Eq. (\ref{sur}), i.e.
-$Z_a=\frac{\xi_a-m_{a}}{s_{a}}$,
+We will next derive closed form solutions to the expectation in
+(\ref{sur}). The location index $\bx$ is suppressed to avoid overly
+complex notation. We present new results for computing the expectation
+$E_{\by}\left\{ \cdot \right\}$ which is the integrand of
+Eq. (\ref{sur}).
+
+A critical first element in the derivation is the GP model, and the
+linear dependence on $\by$ in the conditional mean in
+Eq. (\ref{gp_upd}), which results in a closed form Gaussian
+distribution for the conditional mean in Eq. (\ref{dist_mxi}). The
+high-dimensional inner integral in Eq. (\ref{sur}) then reduces to a
+bivariate integral \citep{bhattacharjya2013value, chevalier2014fast},
+with expectation over $\bm=(m_{a},m_{b})^t$.  We standardize the two
+variables in Eq. (\ref{sur}), i.e.  $Z_a=\frac{\xi_a-m_{a}}{s_{a}}$,
 $Z_b=\frac{\xi_b-m_{b}}{s_{b}}$, and then
@@
 
-Using Eq. (\ref{ytom}) and
-introducing $\pi(\bm)$ for the density function of $\bm$, the expected Bernoulli variance becomes
+Using Eq. (\ref{ytom}) and introducing $\pi(\bm)$ for the density
+function of $\bm$, the expected Bernoulli variance becomes:
 \begin{eqnarray}\label{eq:var2}
@@
 \end{eqnarray}
-We will next extend results of \cite{chevalier2014fast} to solve for these two parts (see also \cite{stroh} who mentioned the possibilities of such extensions).
-
-Considering part I of the, we have
+We will next extend results of \cite{chevalier2014fast} to solve for
+these two parts (see also \cite{stroh} for such an extensions).
+
+For the first part of the above, we have:
 \begin{equation}\label{part1:phi2}
@@
 \end{equation}
-where $\bZ=(Z_{a},Z_{b})$ is a bivariate zero-mean and unit-variance Gaussian vector, with $\mbox{Corr}(Z_{a},Z_{b})=\eta_{a,b}$.
-Moreover, $\bZ$ is chosen to be independent of $\bm$.
-Using Eq.
-(\ref{dist_mxi}) we have
+where $\bZ=(Z_{a},Z_{b})$ is a bivariate zero-mean and unit-variance
+Gaussian vector, with $\mbox{Corr}(Z_{a},Z_{b})=\eta_{a,b}$. Moreover,
+$\bZ$ is chosen to be independent of $\bm$. Using Eq. (\ref{dist_mxi})
+we have
 \begin{equation}
@@
 \end{equation}
-and the same holds for the salinity part with subscript $b$. From this calculation, we get
+and the same holds for the salinity part with subscript $b$ \kc{I
+  really think this 'a' 'b' business should go and be more descriptive
+  or intuitive.}. From this we get:
 \begin{eqnarray}\label{two_parts0}
@@
 
-For part II of expression (\ref{eq:var2}), standardization means that
+For the second part of the expression (\ref{eq:var2}), standardization
+implies that: 
 \begin{eqnarray}\label{part1:phi4}
@@
 \end{eqnarray}
-Here,
-$\bZ_1=(Z_{1,a},Z_{1,b})$, $\bZ_2=(Z_{2,a},Z_{2,b})$ are independent bivariate zero-mean and unit-variance Gaussian vectors, both with element-wise correlation
-$\eta_{a,b}$. Moreover, $\bZ_1$ and $\bZ_2$ are chosen to be independent of $\bm$. Hence,
+Here, $\bZ_1=(Z_{1,a},Z_{1,b})$, $\bZ_2=(Z_{2,a},Z_{2,b})$ are
+independent bivariate zero-mean and unit-variance Gaussian vectors,
+both with element-wise correlation $\eta_{a,b}$. Moreover, $\bZ_1$ and
+$\bZ_2$ are chosen to be independent of $\bm$. Hence:
 \begin{equation}
@@
 
-The solution to part II is then to evaluate a four-variable cumulative distribution function associated with Eqn.
-(\ref{part1:phi4}), accounting for the mean, variance and covariances of $\bZ_1$, $\bZ_2$ and $\bm$ in the appropriate linear combinations.
-Using vector-matrix formulations, we define linear combinations
+The solution to the second part of the equation above, is then to
+evaluate a four-variable cumulative distribution function associated
+with Eqn. (\ref{part1:phi4}), accounting for the mean, variance and
+covariances of $\bZ_1$, $\bZ_2$ and $\bm$ in the appropriate linear
+combinations. Using vector-matrix formulations, we define linear
+combinations:
 \begin{equation}\label{Vcomb}
@@
 \end{equation}
-and the part II calculation becomes
+and this part of the equation then becomes:
 \begin{eqnarray}\label{part1:phi4}
@@
 
-The multivariate cumulative probabilities in two and four dimensions are fast to compute using methods of \cite{genz2009computation}.
-In the end, the part I solution in Eq.
-(\ref{two_parts0}) and part II solution in Eq.
-(\ref{part1:phi4}) are computed for each location on the grid, and summing over the $x \in \mathcal{M}_g$, gives the design evaluation criterion in Eq.
-(\ref{sur}).
+The multivariate cumulative probabilities in two and four dimensions
+are fast to compute using methods of \cite{genz2009computation}.  In
+the end, the first part in Eq. (\ref{two_parts0}) and the second in
+Eq. (\ref{part1:phi4}) are computed for each location on the grid, and
+summing over the $x \in \mathcal{M}_g$, gives the design evaluation
+criterion in Eq. (\ref{sur}). \kc{This might be a prevalent style in
+  Math/Statistics, but this entire formulation is way too dense. I
+  think it might be a good idea to have some basic intuition worked
+  out early before the start of Sec 3}
 

### 2019-09-13 main_new.tex
updates

@@
 \end{equation}
-and this part of the equation then becomes:
+and this in turn then becomes:
 \begin{eqnarray}\label{part1:phi4}
@@
 Eq. (\ref{part1:phi4}) are computed for each location on the grid, and
-summing over the $x \in \mathcal{M}_g$, gives the design evaluation
+summed over $x \in \mathcal{M}_g$, gives the design evaluation
 criterion in Eq. (\ref{sur}). \kc{This might be a prevalent style in
@@
   think it might be a good idea to have some basic intuition worked
-  out early before the start of Sec 3}
+  out early before the start of Sec 3.}
 
@@
 
-In our application with temperature and salinity, the solutions to the complex integral equations for the expected IBV are reduced to calculating bi-variate and four-variate cumulative distribution functions for Gaussian vectors.
-For the more general situation one would be interested in $K$ random fields.
-Say, in the ocean sciences the variables could include chlorophyll or other bio-chemical quantities, in addition to temperature and salinity.
-
-The equations derived in the previous section can then be generalized. Let $\bxi(\bx)=(\xi_1(\bx),\ldots,\xi_K(\bx))$, and the expected Bernoulli variance is
+In our application domain with the variables temperature and salinity,
+the solutions to the complex integral equations for the expected IBV
+are reduced to calculating bi-variate and four-variate cumulative
+distribution functions for Gaussian vectors. For the more general
+situation one would be interested in $K$ random fields. Say, in the
+ocean sciences the variables could include chlorophyll or other
+bio-geochemical quantities, in addition to temperature and salinity.
+
+The equations derived in the previous section can then be generalized
+\kc{Aha! Looks like this is where the generalization across multiple
+  variables can be brought out. If so, state this up front that we're
+  starting off with 2 and then generalizing to more variables.}. Let
+$\bxi(\bx)=(\xi_1(\bx),\ldots,\xi_K(\bx))$, and the expected Bernoulli
+variance is:
 \begin{equation}\label{two_partsK}
@@
 \end{equation}
-By the first key result of linear combination of Gaussian variables, we reduce the integral to a $K$ dimensional integral of the relevant $\bm=E(\bxi(\bx)|\by)$.
-Next, reducing the integral to two parts, we standardize the variables and compute cumulative distributions.
-In summary, we get 
+By the first key result of linear combination of Gaussian variables,
+we reduce the integral to a $K$ dimensional integral of the relevant
+$\bm=E(\bxi(\bx)|\by)$. Next, reducing the integral to two parts, we
+standardize the variables and compute cumulative distributions.  In
+summary, we get:
 \begin{equation}\label{two_parts}

### 2019-09-14 main_new.tex
updates

@@
 
-We calculate EPs and the Bernoulli variance for a few bivariate Gaussian distributions to illustrate concepts. 
-Fig. \ref{illus_bivarDens}
-shows contour plots of three different densities with increasing correlation $\gamma$ between temperature and salinity. 
+We calculate EPs and the Bernoulli variance for a few bivariate
+Gaussian distributions to illustrate the concepts articulated above.
+Fig. \ref{illus_bivarDens} shows contour plots of three different
+densities with increasing correlation $\gamma$ between temperature and
+salinity.
 \begin{figure}[h!] \centering
@@
   \caption{Density contour plots having different correlations between
-    temperature and salinity. The densities have unit variance and the thresholds are identical to the mean values $5^o C$ and $30 mg/l$.}\label{illus_bivarDens} 
+    temperature and salinity. The densities have unit variance and the
+    thresholds are identical to the mean values $5^o C$ and
+    $30 mg/l$. x-axis is temperature and y-axis is salinity.}
+\label{illus_bivarDens}
 \end{figure}
-Here, the thresholds are set equal to the mean, $\mu_a=t_a=5^o C$ and salinity $\mu_b=t_b=30$ mg/l.
-The displayed densities have unit standard deviations for both temperature and salinity, but we also study the effect of doubling the standard deviations.
-
-Table \ref{tab:sim_rhoab} shows the EPs and the associated Bernoulli variance (second row) for the examples indicated in Fig.
-\ref{illus_bivarDens}.
-The EPs increase with the correlation as there is a stronger tendency of having jointly small temperature and salinity.
-The Bernoulli variance is similarly largest for high correlation.
-EPs and Bernoulli variances are the same for standard deviation $1$ or $2$, and this means that high variability in the temperature and salinity variables is not captured in the $p(1-p)$ expression.
-
-\begin{table}[!h] \centering \caption{EP and Bernoulli variance for different correlations and variances (top rows), and expected Bernoulli variances for both data 
-    and only for temperature (bottom rows).}
+Here, the thresholds are set equal to the mean, $\mu_a=t_a=5^o C$ and
+salinity $\mu_b=t_b=30$ mg/l. The displayed densities have unit
+standard deviations for both temperature and salinity, but we also
+study the effect of doubling the standard deviations.
+
+Table \ref{tab:sim_rhoab} shows the EPs and the associated Bernoulli
+variance (second row) for the examples indicated in
+Fig. \ref{illus_bivarDens}. The EPs increase with the correlation as
+there is a stronger tendency of having jointly small temperature and
+salinity. The Bernoulli variance is similarly largest for high
+correlation. EPs and Bernoulli variances are the same for standard
+deviation $1$ or $2$, and this means that high variability in the
+temperature and salinity variables is not captured in the $p(1-p)$
+expression.
+
+\begin{table}[!h] \centering \caption{EP and Bernoulli variance for
+    different correlations and variances (top rows), and expected
+    Bernoulli variances for both data \kc{Do you mean an update of the
+      data?} and only for temperature (bottom
+    rows).}
   \begin{tabular}{c|ccc|ccc}

### 2019-09-15 main_new.tex
updates

@@
 \end{equation}
-where $\bS$ and $\bPsi$ are direct $K$-variate extensions of the matrices in Eq.
-(\ref{two_parts0}), and the matrix $\bV$ finds the relevant linear combination of two independent vectors $\bZ_l=(Z_{l,1},\ldots,Z_{l,K})$, $l=1,2$ and $\bm$, as an extension of Eq.
-(\ref{Vcomb}).
-
-\subsection{Illustration}
+where $\bS$ and $\bPsi$ are direct $K$-variate extensions of the
+matrices in Eq. (\ref{two_parts0}), and the matrix $\bV$ finds the
+relevant linear combination of two independent vectors
+$\bZ_l=(Z_{l,1},\ldots,Z_{l,K})$, $l=1,2$ and $\bm$, as an extension
+of Eq. (\ref{Vcomb}).
+
+\subsection{An Example}
 
@@
 correlation. EPs and Bernoulli variances are the same for standard
-deviation $1$ or $2$, and this means that high variability in the
-temperature and salinity variables is not captured in the $p(1-p)$
-expression.
+deviation $1$ or $2$, and this implies that high variability in
+temperature and salinity is not captured in the $p(1-p)$ expression.
 
@@
     Bernoulli variances for both data \kc{Do you mean an update of the
-      data?} and only for temperature (bottom
-    rows).}
+      data? What is 'both' in this context? T and S? I believe its the
+    latter, and so should be explicit.} and only for
+    temperature (bottom rows).}
   \begin{tabular}{c|ccc|ccc}
@@
 
-Table \ref{tab:sim_rhoab} (bottom two rows) shows results of expected Bernoulli variance calculations.
-This is presented for a design gathering both data types $(y_a,y_b)$, and for a design with temperature measurements $y_a$ alone.
-Having both data; $(y_a,y_b)^t=(\xi_a,\xi_b)^t+N(0,0.5^2\bI)$, while $y_a=\xi_a+N(0,0.5^2)$ when only temperature is measured.
-The results in Table \ref{tab:sim_rhoab} shows that 
-the expected Bernoulli variance gets lower with larger standard deviations $\sigma_a$ and $\sigma_b$ (right columns).
-The reduction of Bernoulli variance is largest for the cases with high correlation $\gamma$.
-Albeit smaller, there is also uncertainty reduction when only temperature is measured (bottom row), especially when the temperature and salinity are very dependent.
-When correlation is low ($\gamma=0.2$) there is hardly any information about salinity in the temperature data, and hence less uncertainty reduction.
-In the application with fresh cold water from the river, the temperature and salinity variables will not only be interdependent, but they will be dependent in the spatial dimension.
-This dependence will also impact design criteria when we evaluate the information measure by integrating over several locations $\bx$.
+Table \ref{tab:sim_rhoab} (bottom two rows) shows results of expected
+Bernoulli variance calculations. This is presented for a design
+gathering both data types $(y_a,y_b)$, and for a design with
+temperature measurements $y_a$ alone. Having both data;
+$(y_a,y_b)^t=(\xi_a,\xi_b)^t+N(0,0.5^2\bI)$, while
+$y_a=\xi_a+N(0,0.5^2)$ when only temperature is measured.  The results
+in Table \ref{tab:sim_rhoab} show that the expected Bernoulli variance
+gets lower with larger standard deviations $\sigma_a$ and $\sigma_b$
+(right columns) \kc{With just 2 data points isn't this a too strong
+  statment to make?}. The reduction of Bernoulli variance is largest
+for the cases with high correlation $\gamma$. Albeit smaller, there is
+also uncertainty reduction when only temperature is measured (bottom
+row), especially when temperature and salinity are very
+dependent. When correlation is low ($\gamma=0.2$) there is little
+information about salinity in the temperature data, and hence less
+uncertainty reduction. In an application with fresh cold water from
+river source, the temperature and salinity variables will not only be
+interdependent, but will show dependence in the spatial
+dimension. This in turn will impact the design criteria when we
+evaluate the information measure by integrating over several locations
+$\bx$.
 
@@
 
-Recall that the AUV can adapt its survey plan based on what it is measuring, in accordance with the sense, plan, act loop (Fig.
-\ref{fig:sense-plan-act}).
-Here we present adaptive strategies for sampling.
+Recall AUV's can adapt survey plans based on what is being measured,
+in accordance with the sense, plan, act loop
+(Fig. \ref{fig:sense-plan-act}). Here we present adaptive strategies
+for sampling.
 
@@
 
-An adaptive AUV survey is split in many rounds (stages), and at each round, the survey design is found by considering the largest reduction in uncertainty of the ES.
-In our setting the selection is made on the graph defined by the nearest grid nodes in the domain $\mathcal{M}_g$.
-
-The optimal sequential design does not only consider the best current AUV grid node, but also what data gathered at this node would lead to for future sampling at the next stages. 
-Let $d^{j,s}$ denote design number $j$ at stage $s$ of sequential surveying.
-If this design is selected, data $\by^{j,s}$ will be gathered.
-The optimal path selection situation is then depicted in Figure \ref{fig:PathSelOpt},
+An adaptive AUV survey is split in many rounds (stages), and at each
+round, the survey design is found by using a myopic approach of
+considering the largest reduction in uncertainty of the ES. In our
+setting the selection is made on the graph defined by the nearest grid
+nodes in the domain $\mathcal{M}_g$.
+
+The optimal sequential design not only considers the best current AUV
+grid node, but also what data gathered at this node would lead to for
+future sampling at successive stages.  Let $d^{j,s}$ denote design
+number $j$ at stage $s$ of a sequential survey. If this design is
+selected, data $\by^{j,s}$ will be gathered. The optimal path
+selection situation is then depicted in Figure \ref{fig:PathSelOpt},
 \begin{figure}[h!]
@@
 %where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages. 
-The mathematical expression for the optimal design then involves a series of intermixed maximizations over designs and integrals over data.
-In practice the optimal solution is not tractable because of the enormous growth over stages, see e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}.
-
-Instead, we next outline heuristic strategies.
-In terms of notation, let $\mathcal{Y}^{s-1} =\{d^{j,r}, \by^{j,r} ; r=1,\ldots,s-1 \}$ denote the data gathered until stage $s-1$, using a selected design $d^{j,r}$, $r=1,\ldots,s-1$.
-
-The derived closed form solutions for the expected IBV are still important building blocks when constructing the sampling designs as they will be used to score the different adaptive design.
-Efficient calculation is important for using adaptive survey designs with robotic platforms.
-
-\subsection{Naive path planning}
+The mathematical expression for the optimal design then involves a
+series of intermixed maximizations over designs and integrals over
+data.  In practice the optimal solution is not tractable because of
+the enormous growth over stages, see
+e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}.
+Instead, we outline heuristic strategies. In terms of notation, let
+$\mathcal{Y}^{s-1} =\{d^{j,r}, \by^{j,r} ; r=1,\ldots,s-1 \}$ denote
+the data gathered until stage $s-1$, using a selected design
+$d^{j,r}$, $r=1,\ldots,s-1$.
+
+The derived closed form solutions for the expected IBV are still
+important building blocks when constructing the sampling designs as
+they will be used to score the different adaptive designs. % Efficient
+% calculation is important for using adaptive survey designs with
+% robotic platforms.
+
+\subsection{A Naive Sampling Strategy}
 \label{naive}
 
-The simplest heuristic for adaptive sampling is to choose the next sampling location based on current EPs.
-The variance is largest for EPs equal to $0.5$, so one selects the design location with EP closest to $0.5$.
-
-At stage $s$, based on the currently available data $\mathcal{Y}^{s-1}$, we fit an updated Gaussian model from Eq.
-(\ref{gp_upd}), with mean $\bm_{s-1}=E(\bxi|\mathcal{Y}^{s-1})$ and covariance matrix $\bS_{s-1}=\mbox{Var}(\bxi|\mathcal{Y}^{s-1})$.
-The next round of measurements $\by^{j,s}$, can be gathered with design $d^{j,s}$, $j=1,\ldots,J$, where $J$ indicates all possible AUV directions or paths at the current stage.
-The {\it{naive}} strategy selects the design according to
+The simplest heuristic for adaptive sampling is to choose the next
+sampling location based on current EPs. The largest variance in
+sampling locations is and indication of higher uncertainty; therefore
+one selects the design location with an EP closest to $0.5$.
+
+At stage $s$, based on the currently available data
+$\mathcal{Y}^{s-1}$, we fit an updated Gaussian model from
+Eq. (\ref{gp_upd}), with mean $\bm_{s-1}=E(\bxi|\mathcal{Y}^{s-1})$
+and covariance matrix $\bS_{s-1}=\mbox{Var}(\bxi|\mathcal{Y}^{s-1})$.
+The next round of measurements $\by^{j,s}$, can be gathered with
+design $d^{j,s}$, $j=1,\ldots,J$, where $J$ indicates all possible AUV
+trajectories from the current stage. The {\it{naive}} strategy then,
+selects the design according to:
 \begin{eqnarray}\label{critNaive}
@@
 \end{eqnarray}
-This strategy does not account for the uncertainty in the temperature or salinity variables, only if one design (node) has EPs closer to $0.5$ (see Table \ref{tab:sim_rhoab}, line two).
-Neither does it account for spatial correlation.
-This strategy lacks memory of where it has been and where the uncertainty has been reduced.
-For this reason it can easily get stuck in local regions.
+This strategy does not account for the uncertainty in the temperature
+or salinity variables, only if one design (node) has EPs closer to
+$0.5$ (see Table \ref{tab:sim_rhoab}, line two). Neither does it
+account for spatial correlation. This strategy lacks memory of where
+it has been and where the uncertainty has been reduced and therefore
+susceptible to local minimas.
 
@@
 
-The myopic (greedy) strategy which we present here is optimal if we imagine taking only one more round of measurements.
-In this selection strategy there is no anticipation of what the subsequent designs might offer, beyond the first round.
-
-Based on the currently available data $\mathcal{Y}^{s-1}$, we again fit an updated Gaussian model, represented on the grid locations covering the spatial domain. 
-The next round of measurements $\by^{j,s}$, can be gathered with design $d^{j,s}$, $j=1,\ldots,J$, indicating all possible AUV directions / paths at the current stage. The selected design is then
+The myopic (greedy) strategy which we present here is optimal if we
+imagine taking only one more round of measurements. In this selection
+strategy there is no anticipation of what the subsequent designs might
+offer, beyond the first round.
+
+Based on the currently available data $\mathcal{Y}^{s-1}$, we again
+fit an updated Gaussian model, represented on the grid locations
+covering the spatial domain.  The next round of measurements
+$\by^{j,s}$, can be gathered with design $d^{j,s}$, $j=1,\ldots,J$,
+indicating all possible AUV trajectories at the current stage. The
+selected design is then:
 \begin{eqnarray}\label{critSEQ}
@@
 
-Note that this strategy gives a sequential conditional version of the formula in Eq.
-(\ref{sur}).
-Now $\mathcal{Y}^{s-1}$ is available, and the expectation is with respect to the conditional density $\pi(\by^{j,s}|\mathcal{Y}^{s-1})$.
-A similar closed form calculation for expected IBV is hence applicable in Eq.
-(\ref{critSEQ}), using the updated GP model from step $s-1$.
-Once the data are collected for the best design, the GP model is updated again. The mean $\bm_{s}$ and covariance matrix $\bS_{s}$ are used to compute the next design at stage $s+1$, and so on.
-
-Even though this myopic strategy is non-anticipative, it still gives a reasonable approach for creating designs in many applications.
-Moreover, it is easily implemented on-board an AUV, using the efficient approach for data updating of the GP model and the calculation of the closed form expected IBV expressions for each next survey line.
-
-\subsection{Look-ahead path planning}
+Note that this strategy gives a sequential conditional version of the
+formula in Eq. (\ref{sur}). Now $\mathcal{Y}^{s-1}$ is available, and
+the expectation is with respect to the conditional density
+$\pi(\by^{j,s}|\mathcal{Y}^{s-1})$. A similar closed form calculation
+for expected IBV is hence applicable in Eq. (\ref{critSEQ}), using the
+updated GP model from step $s-1$.  Once the data is collected for the
+best design, the GP model is updated again. The mean $\bm_{s}$ and
+covariance matrix $\bS_{s}$ are used to compute the next design at
+stage $s+1$, and so on.
+
+Even though this myopic strategy is non-anticipative, it still gives a
+reasonable approach for creating designs in many
+applications. Moreover, it is easily implemented on-board an AUV,
+using the efficient approach for data updating of the GP model and the
+calculation of the closed form expected IBV expressions for each
+subsequent trajectory.
+
+\subsection{Look-ahead trajectory planning}
 \label{LA}
 
-We now extend the myopic strategy to a look-ahead strategy which is optimal when one can gather only two more rounds of measurements.
-In addition to the next round of measurements $\by^{j,s}$, this look-ahead strategy hence anticipates the subsequent design $j_2$ with data $\by^{j_2,s+1}$, when choosing the current design $d^{j,s}$.
-The selected design is
+We now extend the myopic strategy to a look-ahead strategy which is
+optimal when one can gather only two more rounds of measurements. In
+addition to the next round of measurements $\by^{j,s}$, this
+look-ahead strategy, anticipates the subsequent design $j_2$ with
+data $\by^{j_2,s+1}$, when choosing the current design $d^{j,s}$.  The
+selected design is:
 \begin{eqnarray}\label{critLA}
@@
 \end{eqnarray}
-Here, $\mathcal{Y}^{j,s}=\{\by^{j,s},\mathcal{Y}^{s-1}\}$ and $\mathcal{Y}^{j,s+1}=\{\by^{j_2,s+1},\mathcal{Y}^{j,s}\}$ represent the sets of data variables, and the expectations are with respect to the conditional densities $\pi(\by^{j,s}|\mathcal{Y}^{s-1})$ and $\pi(\by^{j_2,s+1}|\by^{j,s},\mathcal{Y}^{s-1})$.
-
-Because of the intermixed optimization and expectation, there is no longer a closed form for the required integrals.
-Instead, we solve the first expectation by Monte Carlo sampling of data $\by^{j,s}$ from its conditional distribution.
-For each data sample, the second expectation is solved using the closed form expressions for expected IBV from Section \ref{sec:sur}.
-
-Even though the strategy looks at two stages, it is only used to find the current best design.
-When data are collected, the GP model is updated, and the mean $\bm_{s}$ and covariance matrix $\bS_{s}$ are used to compute the next design at stage $s+1$, now anticipating what stage $s+2$ could offer, and so on.
-
-This look-ahead approach is much more computer-demanding than the myopic strategy, and for practical implementation we prune paths in the evaluation of Eq.
-(\ref{critLA}).
-This means that we do not compute all possible branches of the first two stages, as they are indicated in Figure \ref{fig:PathSelOpt}.
-Instead, we use the myopic strategy to rank the three best designs on the first stage alone, and for each of these we go through with the look-ahead calculation.
-
-\section{Simulation study}\label{sec:simulations}
-
-Results from tests performed to study the properties of different static and sequential survey designs, in a realistic simulated case, is shown in this section.
-The context is mapping a freshwater plume defined by a temperature and salinity field using an AUV.
+Here, $\mathcal{Y}^{j,s}=\{\by^{j,s},\mathcal{Y}^{s-1}\}$ and
+$\mathcal{Y}^{j,s+1}=\{\by^{j_2,s+1},\mathcal{Y}^{j,s}\}$ represent
+the sets of data variables, and the expectations are with respect to
+the conditional densities $\pi(\by^{j,s}|\mathcal{Y}^{s-1})$ and
+$\pi(\by^{j_2,s+1}|\by^{j,s},\mathcal{Y}^{s-1})$.
+
+Because of the intermixed optimization and expectation, there is no
+longer a closed form for the required integrals. Instead, we solve the
+first expectation by Monte Carlo sampling of data $\by^{j,s}$ from its
+conditional distribution. For each data sample, the second expectation
+is solved using the closed form expressions for expected IBV from
+Section \ref{sec:sur}.
+
+Even though the strategy looks at two stages, it is only used to find
+the current best design. When data is collected, the GP model is
+updated, and the mean $\bm_{s}$ and covariance matrix $\bS_{s}$ are
+used to compute the next design at stage $s+1$, now anticipating what
+stage $s+2$ could offer, and so on.
+
+This look-ahead approach is much more computationally demanding than
+the myopic strategy, and for practical implementation we prune paths
+in the evaluation of Eq. (\ref{critLA}). This means that we do not
+compute all possible branches of the first two stages, as they are
+indicated in Figure \ref{fig:PathSelOpt}. Instead, we use the myopic
+strategy to rank the three best designs on the first stage alone, and
+for each of these we undertake the look-ahead calculations.
+
+\section{Simulation study}
+\label{sec:simulations}
+
+We now show results from tests performed to study the properties of
+different static and sequential survey designs, in a realistic
+simulated case. The context is mapping a freshwater plume defined by a
+temperature and salinity gradient using an AUV.
 
@@
 
-We use a bivariate GP model for temperature and salinity, where we specify the prior mean 
+We use a bivariate GP model for temperature and salinity, where we
+specify the prior mean:
 \begin{equation}\label{m}
@@
 \end{equation}
-In this simulation setup, which is motivated by the real case, we expect the eastern part of the domain to be cold freshwater.
-This situation mimics that of a river mouth entering from the south, and the water masses are pulled to the east.
-The actual trends are in practice set from preliminary information or runs of oceanographic models solving the differential equations for fluid flow.
-In this example the mean is specified by $\beta_{a,1}=0.065$ and $\beta_{b,1}=0.1$.
-Further, the intercepts of the temperature and salinity variables are set to $\beta_{a,0}=5.8$ and $\beta_{b,1}=29.0$.
+In this simulation setup, which is motivated by a real case, we expect
+the eastern part of the domain to be cold freshwater. This situation
+mimics that of a river mouth entering from the south, and the water
+masses are pulled to the east. The actual trends are, in practice, set
+from preliminary information or runs of oceanographic models \kc{a
+  citation of such model use, e.g. the JFR paper might be in order
+  here} solving differential equations for hydrodynamic flow. In this
+example the mean is specified by $\beta_{a,1}=0.065$ and
+$\beta_{b,1}=0.1$. Further, the intercepts \kc{Whats an intercept?} of
+the temperature and salinity variables are set to $\beta_{a,0}=5.8$
+and $\beta_{b,1}=29.0$.
 
@@
 The covariance is assumed to be stationary and separable for the two variables. 
-The $2 \times 2$ pointwise covariance matrix is set to
+The $2 \times 2$ pointwise covariance matrix is set to:
 \begin{equation}\label{v0}
@@
 \end{equation}
-The spatial correlation is of a Matern type giving
+The spatial correlation is of a Matern type \kc{? Define or cite} giving:
 \begin{equation}\label{v}
@@
 %Figure \ref{fig:stat_design} shows the contour lines for the EP for the reference. We notice the trend of increasing salinity and temperature to the west.  
-We study sensitivity to the parameter specification in the results below.
-One realization of the random fields representing salt and temperature is shown in Figure \ref{fig:true_temp} and \ref{fig:true_sal}.
-The true ES from this prior realization is plotted in Figure \ref{fig:ESet}.
-
-\begin{figure}[!h]
+We study sensitivity to the parameter specification in the results
+below.  One realization of the random fields representing salinity and
+temperature is shown in Figure \ref{fig:true_temp} and
+\ref{fig:true_sal}. The true ES from this prior realization is shown
+in Figure \ref{fig:ESet}.
+
+\begin{figure}[!ht]
   \centering
@@
   \hfill
-  \subfigure[Excursion probabilities given collected data along the strategy ``static\_north".]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
-  \caption{\ref{fig:true_temp} and \ref{fig:true_sal} shows
+  \subfigure[Excursion probabilities given collected data using the
+  `static\_north" strategy.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
+  \caption{\ref{fig:true_temp} and \ref{fig:true_sal} show
     realisations of the simulated temperature and salinity fields, as
     well as the associated excursion set \ref{fig:ESet}.
-    \ref{fig:Eprob} show the estimated excursion probabilities after
+    \ref{fig:Eprob} shows the estimated excursion probabilities after
     performing data collection along the N-S survey line.} 
@@
 
-We compare three different static designs denoted \texttt{static\_north}, \texttt{static\_east and} \texttt{static\_zigzag} with the three sequential approaches \texttt{naive}, \texttt{myopic}, and \texttt{look-ahead}.
-The static sampling paths are pre-determined and cannot be changed on the basis of data, representing the archetypical pre-planned observation strategies used in current observational practices.
-For a fixed survey length, the expected variance after static surveying is analytically available (equal coverage each time), but here we focus on the properties from the sequential designs which are evaluated using Monte Carlo sampling over several replicates of realizations from the prior, and conducting simulated sequential surveying for each one; making the result dependent on the observed data.
-For comparison, the same setup is used for the static designs.
-Figure \ref{fig:Eprob} shows the conditional EP, given data gathered along the north-south survey line for the realization indicated in these displays.
-In the Monte Carlo replicates, such results are repeatedly computed to approximate the expected variance reduction in the ES.
-Along with this criterion we also compare predictive performance measured by root mean square error (RMSE) in the temperature and salinity fields and variance reduction in these two field variables.
-It is important to note that the objective function used by the agents is focused on reducing the expected variance in the ES, but we nevertheless hope to achieve good predictive performance for criteria such as RMSE as well.
-Another non-statistical criteria that is important for practical purposes is the computing time of the strategy, as this will impact the performance of the approach in the field for an embedded system.
-
-The sequential sampling agents start at the center East-West coordinate at the southern end of the domain (node 53), see Fig. \ref{fig:wp_graph}.
-
-\begin{figure}[h]
+We compare three different static designs denoting
+\texttt{static\_north}, \texttt{static\_east} and
+\texttt{static\_zigzag} with the three sequential approaches
+\texttt{naive}, \texttt{myopic}, and \texttt{look-ahead}. The static
+sampling paths are pre-determined and cannot be changed on the basis
+of data, representing the archetypical pre-planned observation
+strategies used in typical AUV operational survey designs.
+
+For a fixed survey length, the expected variance after static
+surveying is analytically available (equal coverage each time), but
+here we focus on the properties from the sequential designs which are
+evaluated using Monte Carlo sampling over several replicates of
+realizations from the prior, and conducting simulated sequential
+surveying for each one, making the result dependent on observed
+data. For comparison, the same setup is used for static
+designs. Figure \ref{fig:Eprob} shows the conditional EP, given data
+gathered along the north-south survey line for the realization
+indicated in these displays. In the Monte Carlo replicates, such
+results are repeatedly computed to approximate the expected variance
+reduction in the ES. Along with this criterion we also compare
+predictive performance measured by root mean square error (RMSE) in
+the temperature and salinity fields and variance reduction in these
+two variables. It is important to note that the objective function
+used by the agents is focused on reducing the expected variance in the
+ES, but we nevertheless hope to achieve good predictive performance
+for criteria such as RMSE as well. Another non-statistical criteria
+that is important for practical purposes is the computational time of
+the strategy, as this will impact the performance of the approach in
+the field for an embedded system. In Fig. \ref{fig:wp_graph}, the
+sequential sampling agents start at the center East-West coordinate at
+the southern end of the domain (node 53).
+
+\begin{figure}[t]
 \centering
 \includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
-\caption{The equilateral waypoint graph used to discretize the route choices over the $31\times31$ grid.}
+\caption{The equilateral waypoint graph used to discretize the
+  trajectory choices over the $31\times31$ grid.}
 \label{fig:wp_graph}
@@
 
-Maintaining the AUV example, the AUV will move along edges in the Fig.
-\ref{fig:wp_graph} waypoint graph.
-While doing so, the AUV will gather measurements.
-These are assimilated in the GP model, before an evaluation of the next node to sample will be conducted at the end of the edge.
-The paths of will differ for the various sequential designs due to the individual design criteria and the simulated variability in the environment among the replicates.
+Maintaining the example above, the AUV will move along edges in the
+Fig. \ref{fig:wp_graph} waypoint graph while the vehicle gathers
+measurements. These are assimilated in the GP model, before an
+evaluation of the next node to sample is conducted at the end of the
+edge. The paths will differ for the various sequential designs due to
+the individual design criteria and the simulated variability in the
+environment among the replicates.
 
@@
 
-The north-south design is also doing well, because it moves near the
-boundary of the water masses. In Figure \ref{fig:avg_rmse} and
-\ref{fig:avg_r2} we show results of RMSE and explained variance. Both
-\texttt{myopic} and \texttt{look-ahead} still perform well, but some of
-the static east and zigzag are also achieving good results because they
-are set to always cover large parts of the domain, and in this way visit
-locations that are not considered interesting by the sequential
-strategies targeting ES variance. Regarding computing time in Fig.
-\ref{fig:avg_time}, the \texttt{naive} strategy is on par with the
-static designs, while \texttt{myopic} is slower and \texttt{look-ahead}
-much slower. The \texttt{look-ahead} strategy is reaching levels that
-are near impractical for on-board execution. Of course, with harder
-pruning of branches in the graph, it would be faster to run. Then again,
-this pruning is likely the reason that this strategy does not
-outperforms the myopic approach. In Figure \ref{fig:route_choices} we plot the realized sampling paths of the sequential schemes and static designs.
-
-\begin{figure}[!h]
+The north-south design also does well since it moves near the boundary
+of the water masses. In Figure \ref{fig:avg_rmse} and \ref{fig:avg_r2}
+we show results of RMSE and explained variance. Both \texttt{myopic}
+and \texttt{look-ahead} perform well, but some of the
+\texttt{static\_east} and \texttt{static\_zigzag} also achieve good
+results because they are set to cover large parts of the domain, and
+in doing so visit locations that are not considered interesting by the
+sequential strategies targeting ES variance.
+
+\begin{figure}[!b]
   \centering
-  \subfigure[Look-ahead strategy.]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_look-ahead.pdf}\label{fig:avg_look-ahead}}
+  \subfigure[Look-ahead]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_look-ahead.pdf}\label{fig:avg_look-ahead}}
   \hfill
-  \subfigure[Myopic strategy.]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_myopic.pdf}\label{fig:avg_myopic}}
+  \subfigure[Myopic]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_myopic.pdf}\label{fig:avg_myopic}}
   \hfill
-  \subfigure[Naive strategy.]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_naive.pdf}\label{fig:route_naive}}
+  \subfigure[Naive]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_naive.pdf}\label{fig:route_naive}}
   \hfill
-  \subfigure[Static strategies.]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_static.pdf}\label{fig:route_static}}
+  \subfigure[Static]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_static.pdf}\label{fig:route_static}}
   \hfill
-\caption{The overlaid route choices, superimposed for the 100 replicate
-  surveys with 10 sampling choices/steps.} 
+  \caption{The overlaid route choices (strategies), superimposed for
+    the 100 replicate surveys with 10 sampling choices/steps.}
 \label{fig:route_choices}
@@
 
-The naive strategy appears to get stuck in the southern part of the
-domain in some cases, because it is too focused on the probabilities
-near $0.5$. The myopic strategy seems to be covering a wider domain than
-the look-ahead strategy, possibly because the look-ahead strategy sees
-less value of going far west because it account for having to return
-east again.
-
-We further go on to study sensitivity of the result to different prior correlation values between temperature and salinity, prior standard deviations and spatial correlations, as well as those of measuring only temperature on-board the AUV (as indicated in the rows).
-We compare results at the end of the surveying distance.
-Expected ES variances are shown in Table \ref{{tab:sim_res_ev}}.
-Results of RMSE are in Table \ref{tab:sim_res_rmse} and explained variance in Table \ref{tab:sim_res_r2}.
-In all runs, the myopic and look-ahead strategies perform the best in terms of expected variance in the ES.
-Surprisingly, the myopic is on-par and even slightly better than the look-ahead strategy.
-Part of the reason is pruning of paths for the look-ahead strategy, but even with less pruning, the look-ahead does not outperform the myopic approach.
-Of course, we are able to make up cases where look-ahead beats myopic, but within our ranges of realistic input parameter values it was not easy.
-The static north design is clearly better than the other two static designs. The others do not focus on the areas where the ES is really uncertain. 
-By measuring temperature alone the expected variances in the ES is only slightly higher.
-This occurs because the ES is almost the same and because of the correlation of $0.6$ which means that temperature observations give information about salinity as well.
-
-For the other predictive measures, myopic and look-ahead are still doing well, clearly beating the naive approach.
-The static zigzag design is usually the best among the static designs for these predictive performance measures because its pattern gives more effective coverage of the domain when there is significant spatial correlation as in this case.
+Fig.  \ref{fig:avg_time} shows the computational effort; the
+\texttt{naive} strategy is on par with the static designs, while the
+\texttt{myopic} strategy is slower and \texttt{look-ahead} more so,
+with the latter reaching levels that are nearly impractical for
+execution on a vehicle -- with more pruning of branches in the graph,
+it would likely be faster. Then again, this pruning is likely the
+reason that this strategy does not outperform the myopic approach. In
+Figure \ref{fig:route_choices} we plot the realized sampling paths of
+the sequential schemes and static designs.
+
+The naive strategy gets stuck in the southern part of the domain in
+some cases, because it is too focused on the probabilities near
+$0.5$. The myopic strategy meanwhile, covers a wider domain than the
+look-ahead, possibly because the look-ahead strategy sees less value
+of going far west because it accounts for having to return east again.
+
+We further go on to study sensitivity of the result to different prior
+correlation values between temperature and salinity, prior standard
+deviations and spatial correlations, as well as those of measuring
+only temperature on-board the AUV (as indicated in the rows). We
+compare results at the end of the surveying distance. Expected ES
+variances are shown in Table \ref{tab:sim_res_ev}, results of RMSE are
+in Table \ref{tab:sim_res_rmse} and explained variance in Table
+\ref{tab:sim_res_r2}.  In all runs, the myopic and look-ahead
+strategies perform the best in terms of expected variance in the
+ES. Surprisingly, the myopic is on-par and even slightly better than
+the look-ahead strategy. Part of the reason is pruning of paths for
+the look-ahead strategy, but even with less pruning, the look-ahead
+does not outperform the myopic approach. % Of course, we are able to
+% make up cases where look-ahead beats myopic, but within our ranges of
+% realistic input parameter values it was not easy.
+The \texttt{static\_north} design is clearly better than the other two
+static designs, the reason being that the others do not focus on the
+areas where the ES is really uncertain.  By measuring temperature
+alone the expected variances in the ES is only slightly higher. This
+occurs because the ES is almost the same and because of the
+correlation of $0.6$ which means that temperature observations give
+information about salinity as well.
+
+For the other predictive measures, myopic and look-ahead do well,
+clearly beating the naive approach. The \texttt{static\_zigzag} design
+is usually the best among the static designs for these predictive
+performance measures because its pattern gives more effective coverage
+of the domain when there is significant spatial correlation as in this
+case.
 \begin{table}[!h]
@@
     \end{tabular}}
-    \caption{Simulation results for the final mean excursion set variance (Eq. \eqref{two_partsK}) for 20 replicates, 8 configurations, and 6 strategies.}
+  \caption{Simulation results for the final mean excursion set
+    variance (Eq. \eqref{two_partsK}) for 20 replicates, 8
+    configurations, and 6 strategies.}
     \label{tab:sim_res_ev}
@@
         \end{tabular}}
-    \caption{Simulation results for the final mean RMSE ($\frac{1}{B} \sum_{b=1}^{B} \sqrt{(\bmu-\tilde{\bmu})^2}$) for 20 replicates, 8 configurations, and 6 strategies.}
+      \caption{Simulation results for the final mean RMSE
+        ($\frac{1}{B} \sum_{b=1}^{B} \sqrt{(\bmu-\tilde{\bmu})^2}$)
+        for 20 replicates, 8 configurations, and 6 strategies.}
     \label{tab:sim_res_rmse}
@@
         \end{tabular}}
-    \caption{Simulation results for the final mean explained variance ($\bR^{2}=100*(1-(\bSigma_{posterior}/\bSigma_{initial}))$) for 20 replicates, 8 configurations, and 6 strategies.}
+      \caption{Simulation results for the final mean explained
+        variance
+        ($\bR^{2}=100*(1-(\bSigma_{posterior}/\bSigma_{initial}))$)
+        for 20 replicates, 8 configurations, and 6 strategies.}
     \label{tab:sim_res_r2}
@@
 
-\section{Case study - Mapping a river plume}\label{sec:case_study}
-
-To demonstrate the applicability of multivariate EP to informative oceanographic sampling, a case study mapping river plumes with an AUV is presented.
-This test is done in the near Trondheim, Norway, where the Nidelva river (see Figure \ref{fig:river1}).
-The test focuses along the eastern shoreline due to the aforementioned Coriolis force.%, see Fig.
-\ref{fig:map}.
-The frontal zone therefore runs parallel to the eastern shore as indicated in pink.
-enters the Trondheimsfjord.
-The tests were done in the late Spring of 2019, and because of snow melting in the mountains, the fresh river water is then substantially colder than the fjord water.
-
-\subsection{Statistical model specification}
-
-Statistical model parameters were specified based on preliminary runs.
-A short preliminary survey was therefore conducted upon arriving at the site, where the AUV did an initial transect to determine the environmental conditions and correlation structure.
-
-In these runs the trend parameters were estimated by linear regression, where both temperature and salinity are assumed to increase linearly, going west from the river mouth.
-Next, the residuals from the regression analysis are analyzed to study the fit of the Gaussian model and to specify the covariance parameter.
-
-Figure \ref{fig:parest} shows the diagnostic plots of this residual analysis.
-The left display show a cross-plot of temperature and salinity residuals, indicating larger variability in the salinity than in temperature, and a positive correlation ($0.5$) between the two variables.
-Based on the fitted bivariate Gaussian model we can compute the quadratic form of the residuals.
-If the Gaussian model is adequate the residuals should be approximately $\chi^2_2$ distributed.
-Figure \ref{fig:parest} (middle) shows the empirical cdf of these quadratic forms (solid) together with the theoretical cdf of the $\chi^2_2$ distribution.
-Even though there appears to be some clustering in both left and middle displays, the fit fit appears quite reasonable, and justifies the use of a Gaussian bivariate model in this case.
-Figure \ref{fig:parest} (right) shows the empirical variogram of the scaled residuals for temperature and salinity.
-The decay is similar, and seems to be negligible after about $150$ m.
-The resulting parameters are given in Table \ref{tab:experiment_param} below.
+\section{Case study - Mapping a river plume}
+\label{sec:case_study}
+
+To demonstrate the applicability of multivariate EP to informative
+oceanographic sampling, we present a case study mapping river plumes
+with an AUV.  This experiment was performed in Trondheim, Norway,
+where the Nidelva river (Figure \ref{fig:nidelven}) meets water from
+the Trondheimsfjord. The experiment is focused along the eastern
+shoreline due to the aforementioned Coriolis
+force.%, see Fig. \ref{fig:map}.
+The frontal zone therefore runs parallel to the eastern shore as
+indicated in purple as it enters the fjord. The experiments were
+conducted in late Spring 2019, and because of snow melt with
+consequent fresh river water substantially colder than fjord waters.
+
+\subsection{Model specification}
+
+The statistical model parameters were specified based on preliminary
+run where the AUV did an initial transect to determine the
+environmental conditions and correlation structure. In these runs the
+trend parameters were estimated by linear regression, where both
+temperature and salinity are assumed to increase linearly, going west
+from the river mouth. Next, the residuals from the regression analysis
+are analyzed to study the fit of the Gaussian model and to specify the
+covariance parameter.
+
+Figure \ref{fig:parest} shows the diagnostic plots of this residual
+analysis. The figure on the left shows a cross-plot of temperature and
+salinity residuals, indicating larger variability in salinity than in
+temperature, and a positive correlation ($0.5$) between the two
+variables. Based on the fitted bivariate Gaussian model we can then
+compute the quadratic form of the residuals. If the Gaussian model is
+adequate the residuals should be approximately $\chi^2_2$
+distributed. Figure \ref{fig:parest} (middle) shows the empirical
+(cumulative distribution function) CDF of these quadratic forms
+(solid) together with the theoretical CDF of the $\chi^2_2$
+distribution. Even though there appears to be some clustering in both
+left and middle displays, the fit appears reasonable, and justifies
+the use of a Gaussian bivariate model. Figure \ref{fig:parest} (right)
+shows the empirical variogram of the scaled residuals for temperature
+and salinity. The decay is similar, and seems to be negligible after
+about $150$ m. The resulting parameters are given in Table
+\ref{tab:experiment_param} below.
 
@@
 \includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.eps}
-\caption{Data analysis from a preliminary trial experiment using the AUV.
-Left) Residual plot of temperature and salinity along with Gaussian contours.
-Middle) Empirical cdf of the quadratic form of the residuals along with the theoretical cdf of the Kji-square distribution with two degrees of freedom.
-Right) Empirical variogram of the salinity and temperature data.} \label{fig:parest}
+\caption{Data analysis from a preliminary trial experiment using the
+  AUV. Left: Residual plot of temperature and salinity along with
+  Gaussian contours. Middle: Empirical CDF of the quadratic form of
+  the residuals along with the theoretical CDF of the Kji-square
+  distribution with two degrees of freedom. Right: Empirical variogram
+  of the salinity and temperature data.} \label{fig:parest}
 \end{figure}
@@
 \end{tabular}
-\caption{Model and threshold parameters found from initial AUV survey. Observations were taken across the front crossing from fresh and cold river water to saline and warmer ocean water.}
+\caption{Model and threshold parameters found from an initial AUV
+  survey. Observations were taken across the front while crossing from
+  fresh cold river water to saline and warmer ocean waters.}
 \label{tab:experiment_param}
@@
 
-These mean and covariance parameters values were then used in a field trial, where we explored the algorithm's ability to explore the front between the river and fjord water masses.
-
-\subsection{Experiment setup}
+These mean and covariance parameters values were then used in a field
+trial, where we explored the algorithm's ability to characterize the
+front between the river and fjord water masses.
+
+\subsection{Experimental setup}
 
@@
 %on the 2$^{nd}$ of July 2019 spatially mapping the location of the plume from the Nidelven river (Trondheim, Norway) spreading into the Trondheimsfjord 
-The AUV used the onboard autonomous agent T-REX \citep{Mcgann}, running an instance of the myopic strategy from Section \ref{sec:sampling_designs}, to control the AUV and decide between sampling locations.
-
-The sampling locations are distributed over an equilateral grid, as shown in Fig.
-\ref{fig:map} in a grey colored lattice.
-The robotic platform consisted of a Light AUV \citep{sousa2012lauv} equipped with a 16 Hz Seabird Fastcat-49 CTD (conductivity, temperature, and depth) sensor providing temperature and salinity measurements to the bi-variate GP model.
-The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
+The AUV used the onboard autonomous agent \texttt{T-REX}
+\citep{py10,Rajan12,Rajan12b}, running an instance of the myopic
+strategy from Section \ref{sec:sampling_designs}, to control the AUV
+and decide between sampling locations.
+
+\begin{figure}[!b] 
+ \centering 
+\includegraphics[width=0.98\textwidth]{Figures/LAUV.jpg}
+\caption{The commercially available Light Autonomous Underwater
+  Vehicle (LAUV) used in the experiments in Trondheim.}
+\label{fig:lauv}
+\end{figure}
+
+The sampling locations are distributed over an equilateral grid, as
+shown in Fig. \ref{fig:map} in a grey colored lattice. The robotic
+platform consisted of a Light AUV \citep{sousa2012lauv} (Figure
+\ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49 CTD
+(conductivity, temperature, and depth) sensor providing temperature
+and salinity measurements to the bi-variate GP model. The accuracy of
+the CTD instrument is $\pm 0.0003$ S/m (conductivity) and
+$\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
 
@@
 
-The sampling strategy is designed around sequentially visiting waypoints, where arrival at a desired waypoint subsequently triggers the myopic strategy to evaluate and find the next waypoint based on the ES variance formulated in Eq.
-\eqref{critSEQ}.
-The waypoint --and associated sampling-- that results in lowering the ES variance the most is selected.
-For the field trials, the AUV started in the lower middle of the graph with a GP prior estimate of the environment given as a linear gradient from west to east, between the initially measured temperature and salinity.
-Each AUV survey/mission was set to take approx.
-40 min, visiting 15 waypoints on the grid, with the AUV running at the surface to better capture the plume.
-The GP model assimilated temperature and salinity data continuously with an update frequency of 30 sec.
+The sampling strategy was designed around sequentially visiting
+waypoints, where arrival at a desired waypoint subsequently triggers
+the myopic strategy in order to evaluate and find the next waypoint
+based on the ES variance formulated in Eq. \eqref{critSEQ}. The
+waypoint --and associated sampling-- that results in lowering the ES
+variance the most then, is selected. For the field trials, the AUV
+started in the lower middle of the graph with a GP prior estimate of
+the environment given as a linear gradient from west to east, between
+the initially measured temperature and salinity. Each AUV
+survey/mission was set to take approx. 40 min, visiting 15 waypoints
+on the grid, with the AUV running at the surface to better capture the
+plume. The GP model assimilated temperature and salinity data
+continuously with an update frequency of 30 sec.
 
@@
 
-Two missions, Survey 1 and 2, was run successively from 11:00 AM to 01:00 PM, with a short break in between.
-The resulting path between the selected waypoints are shown in spatial context in Fig.
-\ref{fig:map}, both within the expected frontal region.
-The recorded temperatures are shown as colored trails in Fig.
-\ref{fig:res_both}, clearly showing the temperature difference between ocean and river waters.
-The temperatures are then shown separately overlaid with the estimated ES for each survey in Fig.
-\ref{fig:res1} and \ref{fig:res2}.
-
-Both Survey 1 and 2, successfully estimates and navigates the separation zone crossing the front multiple times.
-As conditions change between the two deployments the resulting path (after waypoint 5) deviate.
-Survey 1 continue northwards tracking the north-eastern part of the front, while Survey 2 turns west mapping the south-western part.
-The final prediction of the front location given Fig.
-\ref{fig:res3} and \ref{fig:res4}, correspond with one another yielding a picture of the front gradually bending off towards north east.
-The amount of exploration done by Survey 1 is greater than Survey 2, coming close to the survey area boarders in the south-western corner.
-A longer planning horizon would identify and discourage such choices.
+Two missions, Survey 1 and 2, was run successively from 11:00 AM to
+01:00 PM, with a short break in between. The resulting path between
+the selected waypoints are shown in spatial context in
+Fig. \ref{fig:map}, both within the expected frontal region. The
+recorded temperatures are shown as colored trails in
+Fig. \ref{fig:res_both}, clearly showing the temperature difference
+between fjord and river waters. The temperatures are then shown
+separately overlaid with the estimated ES for each survey in
+Fig. \ref{fig:res1} and \ref{fig:res2}.
+
+Both Survey 1 and 2, successfully estimate and navigate the separation
+zone while crossing the front multiple times. As conditions changed
+between the two deployments the resulting path (after waypoint 5
+\kc{This should be marked in the figure. And all figures need to be
+  zoomed up, since they're too small}) deviates. Survey 1 continues
+northwards tracking the north-eastern part of the front, while Survey
+2 turns west mapping the south-western part. The final prediction of
+the front location given Fig. \ref{fig:res3} and \ref{fig:res4},
+correspond with one another yielding a picture of the front gradually
+bending off towards north east. The amount of exploration done by
+Survey 1 is greater than Survey 2, coming close to the survey area
+borders in the south-western corner. A longer planning horizon would
+identify and discourage such choices.
 
@@
 
-The methods presented in this paper are shown to be useful in characterizing the frontal water masses where cold river water is separated from the warmer saltwater.
-The study did not consider any temporal effects, which would surely be relevant on a longer time scale.
-We consider the extension to spatio-temporal modeling future work, and we envision that advection-diffusion equations could be useful in this kind of modeling \cite{sigrist2015stochastic}.
-For more complicated oceanographic phenomena, we the methods must maybe be extended to non-Gaussian phenomena and with more complex dynamical models.
-
-Regarding the autonomous sampling, more effort can be spent on approximating the look-ahead steps, but it is perhaps more interesting to explore the additional flexibility that can be gained by having two or more AUVs exploring a spatial or spatio-temporal domain together.
-It might then be useful to sample different parts of the space, and maybe move in parallel to best capture the excursion set.
+The methods presented in this paper are shown to be useful in
+characterizing the frontal water masses where cold river water is
+separated from warmer saline conditions. The study did not consider
+any temporal effects, which would be relevant on a larger time
+scale. We consider the extension to spatio-temporal modeling as future
+work, and we envision that advection-diffusion equations could be
+useful in this kind of modeling \cite{sigrist2015stochastic}.  For
+more complex oceanographic phenomena, the methods will need to be
+extended to non-Gaussian phenomena and augmented by dynamical models.
+
+While more effort could be spent on approximating the look-ahead
+steps, it is perhaps more interesting to explore the additional
+flexibility that can be gained by having two or more AUVs exploring a
+spatial or spatio-temporal domain together. It might then be useful to
+sample different parts of the space, and maybe move in parallel to
+best capture the excursion set.
 
 \section*{Acknowledgements}
-Thanks to SINTEF Ocean, AMOS, etc..
+Thanks to SINTEF Ocean, AMOS, etc..\kc{Complete}
 

### 2019-09-23 main_new.tex
Updates from Jo Eidsvik

@@
   robotics provides opportunities for studying aspects related to
-  experimental designs for effective characterization of oceanographic
+  experimental design for effective characterization of oceanographic
   phenomena. In this work we develop methods for efficient spatial
-  sampling applied to the mapping of dynamic coastal processes by
+  sampling applied to the mapping of coastal processes by
   providing informative descriptions of spatial characteristics of ocean
   phenomena. Specifically, we define a design criterion based on
-  reducing uncertainty in the excursion set of multivariate random
+  reducing uncertainty in the joint excursions of random
   fields, and derive analytical expressions for the expected Bernoulli
@@
   recently acquired information can be assimilated and used to improve
-  statistical models and reduce uncertainty associated with where and
-  when to make measurements.
+  statistical models and reduce uncertainty associated with where to make measurements.
 \item As a priori information about the state of the ocean is very
@@
 
-In this work, we build on excursion sets (ES) and excursion
-probabilities (EP) for spatial design of experiments. The goal is to
+In this work, we build on joint excursion sets (ES) and excursion probabilities (EP) for spatial design of experiments. The goal is to
 design sampling strategies for reducing the spatially integrated
@@
 
-AUVs are used to gather in-situ spatio-temporal data from a targeted
-domain while carrying a range of scientific payloads to survey the
+AUVs are used to gather in-situ spatio-temporal data while carrying a range of scientific payloads to survey the
 water column. Typically, they operate at $\sim 1-3$ m/s, can reach
-depths from 100--6000m with an in-water staying capacity depending on
+depths from 100--6000 m with an in-water operation time capacity depending on
 survey speed, payload sensors and mission design. By deploying AUVs
-one can fill gaps in data gathering and augment information obtained
+together with other resources one can augment information obtained
 from either ocean models, remote-sensing data, or fixed-location
@@
 usually limited to observations along fixed transects on a spatial
-grid, pre-programmed by the human operator.  By using on-board
+grid, pre-programmed by the human operator. By instead using on-board
 algorithms to continuously evaluate, update, and refine future
@@
 
-Our objective is to characterize a river plume, focusing on spatial
-separation of cold freshwater from the warmer saline waters of the
-fjord. Excursion sets (ES) and excursion probabilities (EP) are useful
-starting points to measure the ability to classify water masses. The
-goal of experimental sampling is to improve this characterization, as
-quantified via changes in the EPs.
+Our objective is to leverage statistical tools to characterize a river plume, focusing on spatial separation of cold freshwater (from the river) and warmer saline waters of the fjord.
+Excursion sets (ES) and excursion probabilities (EP) are useful starting points to measure the ability to classify water masses.
+The goal of experimental sampling is to improve this characterization, as quantified via changes in the EPs.
 
@@
 interest, temperature and salinity, are defined as random processes;
-$\xi_a(\bx)$ denotes the temperature (in $^o C$) and $\xi_b(\bx)$
-denotes the salinity (in $mg/l$) at location $\bx$. \kc{might be
-  clearer if you used 't' for temp and 's' for salinity. 'a' and 'b'
-  are somewhat opaque especially further in the text. Also there
-  should be a statement about why we're considering only 2 variables
-  and the impact to considering (algebraically) more in the following
-  formulations. This will surely be asked by any reviewer.}.  We let $t_a$ be a
-threshold for temperature and $t_b$ a threshold for salinity, the
-bivariate ES is then defined by:
-
+$\xi_T(\bx)$ denotes the temperature (in $^o C$) and $\xi_S(\bx)$
+denotes the salinity (in $mg/l$) at location $\bx$. The generalization to more than two variables is given in Section \ref{gen_k}.
+
+We let $t_T$ be a threshold for temperature and $t_S$ a threshold for salinity, the joint ES is then defined by:
 \begin{equation}\label{ES}
-     \mbox{ES} = \{\bx \in \mathcal{M} | \xi_a(\bx) \leq t_a,\xi_b(\bx) \leq t_b\}.
+     \mbox{ES} = \{\bx \in \mathcal{M} | \xi_T(\bx) \leq t_T,\xi_S(\bx) \leq t_S\}.
 \end{equation}
@@
 \begin{equation}\label{eq:prob}
-  p(\bx) = P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b), \hspace{3mm} \bx \in \mathcal{M}.
+  p(\bx) = P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S), \hspace{3mm} \bx \in \mathcal{M}.
 \end{equation}
 If the EPs in Eq. (\ref{eq:prob}) are
-close to $1$ or $0$, it is easy to classify water masses; it is
-challenging if the probabilities are close to $0.5$. 
+close to $1$ or $0$ at a given location, it is easy to classify water mass at this point as either river or ocean; it is however
+challenging if the probability is close to $0.5$. 
 The integrated Bernoulli variance (IBV) \cite{bect2019}, is defined as: 
@@
 \end{equation}
-The measure is dominated by locations with EPs close to $0.5$.
-\kc{This sentence is a bit strange; just about, you state that EPs
-  close to 0.5 can be problematic since they do not categorize the
-  water masses well. And here you state that there is a strong
-  dominance of the average. So now you should say how the biasing can
-  change to make ``V'' more effective. Otherwise this sentence is
-  dangling.}
 
@@
 classification skill. We denote temperature measurements by
-$y_{a}(\bx_i)$ and salinity measurements by $y_{b}(\bx_i)$, at a
+$y_{T}(\bx_i)$ and salinity measurements by $y_{S}(\bx_i)$, at a
 sampling location $\bx_i \in \mathcal{M}$. Given a set of
@@
 \begin{equation}\label{eq:post_ep}
- p(\bx;\by) = P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b |\by). 
-\end{equation}
-
-Our goal is to construct designs for data gathering, that are expected
-to improve the mapping and separation of water masses. When designing
-AUV sampling designs, it is therefore natural to evaluate the expected
-reduction in the IBV, as provided by the survey data. In this way, we
+ p(\bx;\by) = P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\by). 
+\end{equation}
+
+Our goal is to construct adaptive sampling strategies based on continuous evaluation of the EPs.
+As the EPs provide a description of the boundary uncertainty, the strategies can effectively use this description to prioritize sampling efforts to locations that are ambiguous; making exploration more effective.
+As the AUV gathers data throughout its mission, it is natural to evaluate the expected reduction in the IBV, conditioned on the survey data.
+In this way, we can
 take the expectation with respect to the planned measurements $\by$;
@@
 \end{equation}
-where the probability is defined in Eq. (\ref{eq:post_ep}).
+where the expectation is over the planned data, and the probability is defined in Eq. (\ref{eq:post_ep}).
 Considering a set $\mathcal{J}$ of possible designs for an AUV survey. A criterion for the selection is:
@@
 contribution of our work is to derive closed-form results for this
-design criteria in the situation with multivariate GPs.
+design criteria for situations where the underlying model is based on multivariate GPs.
 
@@
 Calculations of the EPs and the IBV require a model specification.
-Here, we use Gaussian modeling assumptions which allow closed form
+We use Gaussian modeling, which provides a practical and popular probabilistic approach to modeling environmental data, allowing a closed form
 evaluation of the IBVs, as derived in Section \ref{sec:sur}.
@@
 
-We represent the bivariate random fields of temperature and salinity
-with a Gaussian model. At location $\bx \in \mathcal{M}$, the
-bivariate distribution is:
+The Gaussian model is used to represent the bivariate random fields of temperature and salinity. At location $\bx \in \mathcal{M}$, the prior bivariate distribution is:
 \begin{equation}\label{gp_bar}
-  \begin{bmatrix}\xi_a(\bx) \\
-    \xi_b(\bx) \end{bmatrix}
+  \begin{bmatrix}\xi_T(\bx) \\
+    \xi_S(\bx) \end{bmatrix}
  \sim N \left( 
-\begin{bmatrix} \mu_{a}(\bx)\\
-\mu_{b}(\bx)
+\begin{bmatrix} \mu_{T}(\bx)\\
+\mu_{S}(\bx)
 \end{bmatrix},\begin{bmatrix}
-\sigma_{a}^2(\bx) & \sigma_{a}(\bx) \sigma_b(\bx) \gamma(\bx)  \\
-\sigma_{a}(\bx) \sigma_b(\bx) \gamma(\bx)  & \sigma_{b}^2(\bx) 
+\sigma_{T}^2(\bx) & \sigma_{T}(\bx) \sigma_S(\bx) \gamma(\bx)  \\
+\sigma_{T}(\bx) \sigma_S(\bx) \gamma(\bx)  & \sigma_{S}^2(\bx) 
 \end{bmatrix}
@@
 where the notation refers to Gaussian (Normal) distributed variables
-specified by the mean vector and covariance matrix. In the application
-below, the model parameters are specified from preliminary data: The
-mean varies with location, capturing the variability near the river
-mouth, while the variance parameters and correlation between salinity
-and temperature are assumed to be constant for all locations. \kc{?
-  Isn't that too strong an assumption? T/S tend to be correlated,
-  however for uniform bodies of water; here we're dealing with coastal
-  ecosystems where the two are not necessarily so.}
-
-For the spatial dependence among variables, 
-we assume a separable correlation function with the same decay for salinity and temperature,
-which is not unrealistic for a river plume \kc{why so?}. We have:
+specified by the mean vector and covariance matrix. For the application
+of tracking river plumes, the model parameters are specified from a short preliminary survey.
+The mean varies with location, capturing the variability near the river mouth, while the variance parameters and correlation between salinity and temperature are assumed to be constant for all locations and within each water mass.
+
+For the spatial dependence among variables, we assume a separable correlation function with the same decay for salinity and temperature,
+which is not unrealistic for our case with a river plume as the decay is related to the common water masses (Sect \ref{sec:case_study}). We have:
 \begin{equation}\label{gp_corr}
@@
 \end{equation}
-for $i,j \in {a,b}$ and $\delta_{ij}=1$ if $i=j$ and $\delta_{ij}=0$
-otherwise. Moreover, we assume a stationary isotropic property
-\kc{This is probably related to the assumption as above where T/S are
-  assumed to be constant in all locations? But why?} where the
+for $i,j \in {T,S}$ and $\delta_{ij}=1$ if $i=j$ and $\delta_{ij}=0$
+otherwise. Moreover, we assume stationary isotropic random fields a priori where the
 correlation function $\rho(\bx,\bx')$ solely depends on $\bx$ and
-$\bx'$ via a Euclidean distance measure $h=\sqrt{|\bx-\bx'|^2}$. With
+$\bx'$ via a Euclidean distance measure $h=\sqrt{||\bx-\bx'||^2}$. With
 data and prior knowledge one could possibly fit and estimate
@@
 
-For implementation purposes, we discretize the spatial domain
+We discretize the spatial domain
 $\mathcal{M}$ to a set of $n$ grid locations
-$\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell has area
-$\Delta$. The integral expression in Eq. (\ref{measVAR}) and
-(\ref{sur}) are then approximated by sums over all grid cells. This
-grid is identical to the possible set of design locations and defined
-as a waypoint graph \kc{You mean the
-  centroids of the cells are part of the graph?}. Using vector
+$\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell has area $\Delta$. The
+grid is used for the waypoint graph setting the possible design locations. It is also used when solving integrals expressions in Eq. (\ref{measVAR}) and
+(\ref{sur}) which are then approximated by sums over all grid cells.  Using vector
 notation, we denote the Gaussian temperature and salinity variables on
@@
 \begin{equation}\label{prior}
-    \bxi = (\xi_a(\bx_1),\xi_b(\bx_1),\ldots,\xi_a(\bx_n),\xi_b(\bx_n))^t, \hspace{3mm}
+    \bxi = (\xi_T(\bx_1),\xi_S(\bx_1),\ldots,\xi_T(\bx_n),\xi_S(\bx_n))^t, \hspace{3mm}
     \bxi  \sim  N(\bmu, \bSigma) %\pi (\bxi) & =& \frac{1}{(2\pi)^{n}|\bSigma|^{\frac{1}{2}}}e^{-\frac{1}{2}(\bxi -\bmu)^T\bSigma^{-1}(\bxi-\bmu)}.
@@
 temperature and salinity at the grid locations, while the
-$2n \times 2n$ positive definite matrix $\bSigma$ holds the
+$2n \times 2n$ matrix $\bSigma$ holds the
 variance-covariances of temperature and salinity variables at all grid
@@
 We denote the data at $n_y$ locations by
-$\by=(y_{a,1},y_{b,1},\ldots,y_{a,n_y},y_{b,n_y})^t$, representing the
+$\by=(y_{T,1},y_{S,1},\ldots,y_{T,n_y},y_{S,n_y})^t$, representing the
 temperature and salinity measurements gathered at the first round of
@@
 \begin{equation}\label{likelihood}
-\by | \bxi \sim N( \bA \bxi, \bR), %\pi(\by | \bxi)= \frac{1}{(2\pi)^{m}|\bR|^{\frac{1}{2}}}e^{-\frac{1}{2}(\by -\bA\bxi)^T\bR^{-1}(\by-\bA\bxi)},
-\end{equation}
-where $\bA$ is an $2n_y \times 2n$ matrix holding '$1$'s on the
+\by | \bxi \sim N( \bG \bxi, \bR), %\pi(\by | \bxi)= \frac{1}{(2\pi)^{m}|\bR|^{\frac{1}{2}}}e^{-\frac{1}{2}(\by -\bG\bxi)^T\bR^{-1}(\by-\bG\bxi)},
+\end{equation}
+where $\bG$ is an $2n_y \times 2n$ matrix holding '$1$'s on the
 measurement indices on the grid and '$0$' otherwise. The size
 $2n_y \times 2n_y$ covariance matrix $\bR$ contains diagonal entries
-$r^2_a$ and $r^2_b$ which are measurement error variances of
+$r^2_T$ and $r^2_S$ which are measurement error variances of
 temperature and salinity observations.
 
-The GP mean and covariance in Eq. (\ref{prior}) are moderated \kc{What
-  does 'moderated' mean?} when more information is available. The
+The GP mean and covariance in Eq. (\ref{prior}) are updated when more information is available. The
 updated distribution for temperature and salinity variables on the
@@
 \begin{eqnarray}\label{gp_upd}
-  \bm &=& \bm(\by) = \bmu+\bSigma \bA^t (\bA \Sigma \bA^t+\bR)^{-1}(\by-\bA \bmu),  \\
-  \bS &=& \bSigma - \bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA
+  \bm &=& \bm(\by) = \bmu+\bSigma \bG^t (\bG \Sigma \bG^t+\bR)^{-1}(\by-\bG \bmu),  \\
+  \bA &=& \bSigma - \bSigma \bG^t (\bG \bSigma \bG^t+\bR)^{-1} \bG
           \bSigma.\nonumber
@@
 \begin{bmatrix}
-\xi_a(\bx) \\
-\xi_b(\bx)
+\xi_T(\bx) \\
+\xi_S(\bx)
 \end{bmatrix}
@@
  \sim N \left( 
-\begin{bmatrix} m_{a}(\bx)\\
-m_{b}(\bx)
+\begin{bmatrix} m_{T}(\bx)\\
+m_{S}(\bx)
 \end{bmatrix},\begin{bmatrix}
-s_{a}^2(\bx) & s_{a,b}(\bx)  \\
-s_{a,b}(\bx)  & s_{b}^2(\bx)  
+a_{T}^2(\bx) & a_{T,S}(\bx)  \\
+a_{T,S}(\bx)  & a_{S}^2(\bx)  
 \end{bmatrix}
@@
 \end{equation}
-with entries extracted from the conditional mean and covariance
+with entries $m_{\cdot}$ and $a_{\cdot}$ extracted from the conditional mean and covariance
 expressions in Eq. (\ref{gp_upd}).
@@
 the marginal distribution of the data which is defined over $\bxi$ to
-be $\by \sim N( \bA \bmu,\bA \bSigma \bA^t+\bR)$.  However, it turns
+be $\by \sim N( \bG \bmu,\bG \bSigma \bG^t+\bR)$.  However, it turns
 out that a simplified form of the calculation is possible because the
@@
 \begin{equation}\label{distxi} \bm \sim N(\bmu , \bPsi), \hspace{3mm}
-  \bPsi=\bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.
+  \bPsi=\bSigma \bG^t (\bG \bSigma \bG^t+\bR)^{-1} \bG \bSigma.
 \end{equation} 
@@
   \bm(\bx) \sim N \left( \begin{bmatrix}
-      \mu_{a}(\bx) \\
-      \mu_{b}(\bx) \end{bmatrix}, \begin{bmatrix}
-      \psi^2_{a}(\bx) & \psi_{a,b}(\bx)\\
-      \psi_{a,b}(\bx) & \psi^2_{b}(\bx) \end{bmatrix} \right),
+      \mu_{T}(\bx) \\
+      \mu_{S}(\bx) \end{bmatrix}, \begin{bmatrix}
+      \psi^2_{T}(\bx) & \psi_{T,S}(\bx)\\
+      \psi_{T,S}(\bx) & \psi^2_{S}(\bx) \end{bmatrix} \right),
 \end{equation}
@@
 \begin{eqnarray}\label{eq:prob_mv0}
- p(\bx) &=& P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b) \nonumber \\
+ p(\bx) &=& P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S) \nonumber \\
  &=& \Phi_2 \begin{pmatrix} 
-\begin{bmatrix} t_a\\
-t_b
+\begin{bmatrix} t_T\\
+t_S
 \end{bmatrix};
-\begin{bmatrix} \mu_{a}(\bx)\\
-\mu_{b}(\bx)
+\begin{bmatrix} \mu_{T}(\bx)\\
+\mu_{S}(\bx)
 \end{bmatrix},\begin{bmatrix}
-\sigma_{a}^2(\bx) & \sigma_{a}(\bx)\sigma_{b}(\bx) \gamma(\bx)  \\
-\sigma_{a}(\bx)\sigma_{b}(\bx) \gamma(\bx)  & \sigma_{b}^2(\bx)  
+\sigma_{T}^2(\bx) & \sigma_{T}(\bx)\sigma_{S}(\bx) \gamma(\bx)  \\
+\sigma_{T}(\bx)\sigma_{S}(\bx) \gamma(\bx)  & \sigma_{S}^2(\bx)  
 \end{bmatrix}\end{pmatrix},
@@
 \begin{eqnarray}\label{eq:prob_mv}
- p(\bx;\by) &=& P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b |\by)
+ p(\bx;\by) &=& P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\by)
  \nonumber \\
  &=& \Phi_2 \begin{pmatrix} 
-\begin{bmatrix} t_a\\
-t_b
+\begin{bmatrix} t_T\\
+t_S
 \end{bmatrix};
-\begin{bmatrix} m_{a}(\bx)\\
-m_{b}(\bx)
+\begin{bmatrix} m_{T}(\bx)\\
+m_{S}(\bx)
 \end{bmatrix},\begin{bmatrix}
-s_{a}^2(\bx) & s_{a,b}(\bx)  \\
-s_{a,b}(\bx)  & s_{b}^2(\bx)  
+a_{T}^2(\bx) & a_{T,S}(\bx)  \\
+a_{T,S}(\bx)  & a_{S}^2(\bx)  
 \end{bmatrix}\end{pmatrix},
 \end{eqnarray}
-with parameters as defined in Eq. (\ref{gp_upd}) and (\ref{gp_hat}).
+with parameters as defined in Eq.
+(\ref{gp_upd}) and (\ref{gp_hat}).
+Here, the probabilities would not only depend on the data but also on the data design defined via matrix $\bG$ in Eq.
+(\ref{likelihood}).
+We next derive results that ease comparison of several designs.
 
@@
 
-A critical first element in the derivation is the GP model, and the
+A critical first element in the derivation is using the GP model and the
 linear dependence on $\by$ in the conditional mean in
-Eq. (\ref{gp_upd}), which results in a closed form Gaussian
+Eq. (\ref{gp_upd}). This results in a closed form Gaussian
 distribution for the conditional mean in Eq. (\ref{dist_mxi}). The
@@
 bivariate integral \citep{bhattacharjya2013value, chevalier2014fast},
-with expectation over $\bm=(m_{a},m_{b})^t$.  We standardize the two
-variables in Eq. (\ref{sur}), i.e.  $Z_a=\frac{\xi_a-m_{a}}{s_{a}}$,
-$Z_b=\frac{\xi_b-m_{b}}{s_{b}}$, and then
+with expectation over $\bm=(m_{T},m_{S})^t$.  We standardize the two
+variables in Eq. (\ref{sur}), i.e.  $Z_T=\frac{\xi_T-m_{T}}{a_{T}}$,
+$Z_S=\frac{\xi_S-m_{S}}{a_{S}}$, and then
 \begin{eqnarray}\label{ytom}
-   P(\xi_a \leq t_a, \xi_b \leq t_b|\by) &=& P \left( Z_a \leq\frac{t_a-m_{a}}{s_{a}}, Z_b \leq\frac{t_b-m_{b}}{s_{b}}|\by \right) \nonumber \\
+   P(\xi_T \leq t_T, \xi_S \leq t_S|\by) &=& P \left( Z_T \leq\frac{t_T-m_{T}}{a_{T}}, Z_S \leq\frac{t_S-m_{S}}{a_{S}}|\by \right) \nonumber \\
    &=& \Phi_2 \begin{pmatrix} 
-\begin{bmatrix} \frac{t_a-m_{a}}{s_{a}}\\
-\frac{t_b-m_{b}}{s_{b}}
+\begin{bmatrix} \frac{t_T-m_{T}}{a_{T}}\\
+\frac{t_S-m_{S}}{a_{S}}
 \end{bmatrix};
@@
 \end{bmatrix}\end{pmatrix} \nonumber \\
-&=& P(\xi_a \leq t_a, \xi_b \leq t_b|\bm),
+&=& P(\xi_T \leq t_T, \xi_S \leq t_S|\bm),
 \end{eqnarray}
-where $\eta_{a,b} =s_{a,b}/(s_{a} s_{b})$.
-%Since the expression only depends on $\by$ via $\bm$, the expectation in Eq.
-(\ref{sur}) is reduced to a bivariate integral over $\bm$ (which will vary for each integrand term, because the locations differ).
-
-Using Eq. (\ref{ytom}) and introducing $\pi(\bm)$ for the density
+where $\eta_{T,S} =a_{T,S}/(a_{T} a_{S})$.
+Since the expression only depends on $\by$ via its affine function $\bm$, the expectation in Eq. (\ref{sur}) is reduced to a bivariate integral over $\bm$. 
+Introducing $\pi(\bm)$ for the density
 function of $\bm$, the expected Bernoulli variance becomes:
 \begin{eqnarray}\label{eq:var2}
-E_{\by}(p(1-p)) &=& E_{\bm}(p(1-p)), \hspace{3mm} p=P(\xi_a \leq t_a, \xi_b \leq t_b|\bm) \nonumber \\
+E_{\by}(p(1-p)) &=& E_{\bm}(p(1-p)), \hspace{3mm} p=P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) \nonumber \\
 E_{\bm}(p(1-p)) & = & \int p[1-p] \pi(\bm) d\bm, \nonumber \\
- &=& \int P(\xi_a \leq t_a, \xi_b \leq t_b|\bm)  \pi(\bm) d\bm \nonumber  \\
-&-& \int P(\xi_a \leq t_a, \xi_b \leq t_b|\bm) P(\xi_a \leq t_a, \xi_b \leq t_b|\bm) \pi(\bm) d\bm. 
+ &=& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm)  \pi(\bm) d\bm \nonumber  \\
+&-& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) \pi(\bm) d\bm. 
 \end{eqnarray}
@@
 \begin{equation}\label{part1:phi2}
- \int P(\xi_a \leq t_a, \xi_b \leq t_b|\bm) \pi(\bm) d\bm= 
-P \left( Z_{a} \leq \frac{t_a-m_{a}}{s_{a}}, 
-Z_{b} \leq \frac{t_b-m_{b}}{s_{b}} \right), \nonumber
-\end{equation}
-where $\bZ=(Z_{a},Z_{b})$ is a bivariate zero-mean and unit-variance
-Gaussian vector, with $\mbox{Corr}(Z_{a},Z_{b})=\eta_{a,b}$. Moreover,
-$\bZ$ is chosen to be independent of $\bm$. Using Eq. (\ref{dist_mxi})
-we have
+ \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) \pi(\bm) d\bm= 
+P \left( Z_{T} \leq \frac{t_T-m_{T}}{a_{T}}, 
+Z_{S} \leq \frac{t_S-m_{S}}{a_{S}} \right), \nonumber
+\end{equation}
+where $\bZ=(Z_{T},Z_{S})$ is a bivariate zero-mean and unit-variance
+Gaussian vector, with $\mbox{Corr}(Z_{T},Z_{S})=\eta_{T,S}$. Moreover,
+$\bZ$ is chosen to be independent of $\bm$. Using Eq. (\ref{dist_mxi}) we have
 \begin{equation}
-    E(s_{a} Z_{a}+m_{a}-t_a) = \mu_{a}-t_a, \hspace{3mm}
-    \mbox{Var}(s_{a} Z_{a}+m_{a}-t_a) = s^2_a+ \psi^2_{a}, 
-\end{equation}
-and the same holds for the salinity part with subscript $b$ \kc{I
-  really think this 'a' 'b' business should go and be more descriptive
-  or intuitive.}. From this we get:
+    E(a_{T} Z_{T}+m_{T}-t_T) = \mu_{T}-t_T, \hspace{3mm}
+    \mbox{Var}(a_{T} Z_{T}+m_{T}-t_T) = a^2_T+ \psi^2_{T}, 
+\end{equation}
+for the temperature part and the same holds for the salinity part. From this we get:
 \begin{eqnarray}\label{two_parts0}
-& P & \left( Z_{a} \leq \frac{t_a-m_{a}}{s_{a}}, 
-Z_{b} \leq \frac{t_b-m_{b}}{s_{b}} \right) \\
+& P & \left( Z_{T} \leq \frac{t_T-m_{T}}{a_{T}}, 
+Z_{S} \leq \frac{t_S-m_{S}}{a_{S}} \right) \\
 &=& \Phi_2 \begin{pmatrix} 
@@
 \end{bmatrix};
-\begin{bmatrix} \mu_{a}-t_a\\
-\mu_{b}-t_b
+\begin{bmatrix} \mu_{T}-t_T\\
+\mu_{S}-t_S
 \end{bmatrix},\begin{bmatrix}
-s^2_a+ \psi^2_{a} & s_{a,b}+\psi_{a,b}  \\
-s_{a,b}+\psi_{a,b}   & s^2_a+ \psi^2_{a} 
+a^2_T+ \psi^2_{T} & a_{T,S}+\psi_{T,S}  \\
+a_{T,S}+\psi_{T,S}   & a^2_T+ \psi^2_{T} 
 \end{bmatrix}\end{pmatrix} \nonumber.
@@
 \begin{eqnarray}\label{part1:phi4}
-&& \int P(\xi_a \leq t_a, \xi_b \leq t_b|\bm) P(\xi_a \leq t_a, \xi_b \leq t_b|\bm) p(\bm) d\bm =  \\
-&& P\left( Z_{1,a} \leq \frac{t_a-m_{a}}{s_{a}}, 
-Z_{1,b} \leq \frac{t_b-m_{b}}{s_{b}},Z_{2,b} \leq \frac{t_a-m_{a}}{s_{a}}, 
-Z_{2,b} \leq \frac{t_b-m_{b}}{s_{b}} \right). \nonumber
+&& \hspace{6mm} \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) p(\bm) d\bm =  \\
+&P& \hspace{-3mm} \left( Z_{1,T} \leq \frac{t_T-m_{T}}{a_{T}}, 
+Z_{1,S} \leq \frac{t_S-m_{S}}{a_{S}},Z_{2,S} \leq \frac{t_T-m_{T}}{a_{T}}, 
+Z_{2,S} \leq \frac{t_S-m_{S}}{a_{S}} \right). \nonumber
 \end{eqnarray}
-Here, $\bZ_1=(Z_{1,a},Z_{1,b})$, $\bZ_2=(Z_{2,a},Z_{2,b})$ are
+Here, $\bZ_1=(Z_{1,T},Z_{1,S})$, $\bZ_2=(Z_{2,T},Z_{2,S})$ are
 independent bivariate zero-mean and unit-variance Gaussian vectors,
-both with element-wise correlation $\eta_{a,b}$. Moreover, $\bZ_1$ and
+both with element-wise correlation $\eta_{T,S}$. Moreover, $\bZ_1$ and
 $\bZ_2$ are chosen to be independent of $\bm$. Hence:
@@
     \begin{array}{cccccc}
-         Z_{1,a} \\
-         Z_{1,b} \\
-         Z_{2,a} \\
-         Z_{2,b} \\
-         m_{a} \\
-         m_{b}
+         Z_{1,T} \\
+         Z_{1,S} \\
+         Z_{2,T} \\
+         Z_{2,S} \\
+         m_{T} \\
+         m_{S}
     \end{array}
@@
     0 \\
-          \mu_{a} \\
-         \mu_{b} 
+          \mu_{T} \\
+         \mu_{S} 
     \end{array}
@@
       \begin{array}{cccccc}
-        1 & \eta_{a,b} & 0 & 0 & 0 & 0  \\
-        \eta_{a,b} & 1 & 0 & 0 & 0 & 0 \\
-        0 & 0 & 1 & \eta_{a,b} & 0 & 0 \\
-        0 & 0 & \eta_{a,b} & 1 & 0 & 0 \\
-        0 & 0 & 0 & 0 & \psi^2_{a} & \psi_{a,b} \\
-        0 & 0 & 0 & 0 & \psi_{a,b} & \psi^2_{b} 
+        1 & \eta_{T,S} & 0 & 0 & 0 & 0  \\
+        \eta_{T,S} & 1 & 0 & 0 & 0 & 0 \\
+        0 & 0 & 1 & \eta_{T,S} & 0 & 0 \\
+        0 & 0 & \eta_{T,S} & 1 & 0 & 0 \\
+        0 & 0 & 0 & 0 & \psi^2_{T} & \psi_{T,S} \\
+        0 & 0 & 0 & 0 & \psi_{T,S} & \psi^2_{S} 
     \end{array}
@@
     \begin{array}{cccccc}
-        s_a & 0 & 0 & 0 & 1 & 0 \\
-         0 & s_b & 0 & 0 & 0 & 1 \\
-         0 & 0 & s_a & 0 & 1 & 0\\
-         0 & 0 & 0 & s_b & 0 & 1\\
+        a_T & 0 & 0 & 0 & 1 & 0 \\
+         0 & a_S & 0 & 0 & 0 & 1 \\
+         0 & 0 & a_T & 0 & 1 & 0\\
+         0 & 0 & 0 & a_S & 0 & 1\\
     \end{array}
@@
 and this in turn then becomes:
-\begin{eqnarray}\label{part1:phi4}
-&& P \left( Z_{1,a} \leq \frac{t_a-m_{a}}{s_{a}}, 
-Z_{1,b} \leq \frac{t_b-m_{b}}{s_{b}},Z_{2,b} \leq \frac{t_a-m_{a}}{s_{a}}, 
-Z_{2,b} \leq \frac{t_b-m_{b}}{s_{b}} \right)  \\
+\begin{eqnarray}\label{part1:phi4_2}
+&& P \left( Z_{1,T} \leq \frac{t_T-m_{T}}{a_{T}}, 
+Z_{1,S} \leq \frac{t_S-m_{S}}{a_{S}},Z_{2,S} \leq \frac{t_T-m_{T}}{a_{T}}, 
+Z_{2,S} \leq \frac{t_S-m_{S}}{a_{S}} \right)  \\
 &=&  \Phi_4 
 \begin{pmatrix} 
-\begin{bmatrix} t_a\\
-t_b \\
-t_a \\
-t_b
+\begin{bmatrix} t_T\\
+t_S \\
+t_T \\
+t_S
 \end{bmatrix};
@@
 Eq. (\ref{part1:phi4}) are computed for each location on the grid, and
-summed over $x \in \mathcal{M}_g$, gives the design evaluation
-criterion in Eq. (\ref{sur}). \kc{This might be a prevalent style in
-  Math/Statistics, but this entire formulation is way too dense. I
-  think it might be a good idea to have some basic intuition worked
-  out early before the start of Sec 3.}
+summed over $\bx \in \mathcal{M}_g$, gives the design evaluation
+criterion in Eq. (\ref{sur}).
 
 \subsection{Generalization to the multivariate case}
+\label{gen_k}
 
@@
 
-The equations derived in the previous section can then be generalized
-\kc{Aha! Looks like this is where the generalization across multiple
-  variables can be brought out. If so, state this up front that we're
-  starting off with 2 and then generalizing to more variables.}. Let
+The equations derived in the previous section can then be generalized. Let
 $\bxi(\bx)=(\xi_1(\bx),\ldots,\xi_K(\bx))$, and the expected Bernoulli
@@
 \end{bmatrix};
-\bmu,\bS+\bPsi 
+\bmu,\bA+\bPsi 
 \end{pmatrix}
@@
 \end{equation}
-where $\bS$ and $\bPsi$ are direct $K$-variate extensions of the
+where $\bA$ and $\bPsi$ are direct $K$-variate extensions of the
 matrices in Eq. (\ref{two_parts0}), and the matrix $\bV$ finds the
@@
 \end{figure}
-Here, the thresholds are set equal to the mean, $\mu_a=t_a=5^o C$ and
-salinity $\mu_b=t_b=30$ mg/l. The displayed densities have unit
+Here, the thresholds are set equal to the mean, $\mu_T=t_T=5^o C$ and
+salinity $\mu_S=t_S=30$ mg/l. The displayed densities have unit
 standard deviations for both temperature and salinity, but we also
@@
     different correlations and variances (top rows), and expected
-    Bernoulli variances for both data \kc{Do you mean an update of the
-      data? What is 'both' in this context? T and S? I believe its the
-    latter, and so should be explicit.} and only for
-    temperature (bottom rows).}
+    Bernoulli variances for both temperature and salinity data $\by$ and only for
+    temperature $y_T$ (bottom rows).}
   \begin{tabular}{c|ccc|ccc}
- &\multicolumn{3}{c}{$\sigma_b=\sigma_a=1$} & \multicolumn{3}{c}{$\sigma_b=\sigma_a=2$} \\
+ &\multicolumn{3}{c}{$\sigma_S=\sigma_T=1$} & \multicolumn{3}{c}{$\sigma_S=\sigma_T=2$} \\
 \hline
@@
 $p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
-$E_{(y_a,y_b)}(p (1-p))$ & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
-$E_{y_a}(p (1-p))$ & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
+$E_{\by}(p(\by) (1-p(\by)))$ & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
+$E_{y_T}(p(y_T) (1-p(y_T)))$ & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
 \hline
@@
 Bernoulli variance calculations. This is presented for a design
-gathering both data types $(y_a,y_b)$, and for a design with
-temperature measurements $y_a$ alone. Having both data;
-$(y_a,y_b)^t=(\xi_a,\xi_b)^t+N(0,0.5^2\bI)$, while
-$y_a=\xi_a+N(0,0.5^2)$ when only temperature is measured.  The results
-in Table \ref{tab:sim_rhoab} show that the expected Bernoulli variance
-gets lower with larger standard deviations $\sigma_a$ and $\sigma_b$
-(right columns) \kc{With just 2 data points isn't this a too strong
-  statment to make?}. The reduction of Bernoulli variance is largest
+gathering both data types $(y_T,y_S)$, and for a design with
+temperature measurements $y_T$ alone. Having both data;
+$(y_T,y_S)^t=(\xi_T,\xi_S)^t+N(0,0.5^2\bI)$, while
+$y_T=\xi_T+N(0,0.5^2)$ when only temperature is measured.  For this illustration, Table \ref{tab:sim_rhoab} shows that the expected Bernoulli variance
+gets lower with larger standard deviations $\sigma_T$ and $\sigma_S$
+(right columns). The reduction of Bernoulli variance is largest
 for the cases with high correlation $\gamma$. Albeit smaller, there is
@@
 %\begin{equation}
-%E(\min \{ p,(1-p)\})=\int \min P(\xi_a \leq t, \xi_b \leq s),[1-P(\xi_a \leq t, \xi_b \leq s)] p(E(\bxi)=) dE(\bxi)=,
+%E(\min \{ p,(1-p)\})=\int \min P(\xi_T \leq t, \xi_S \leq s),[1-P(\xi_T \leq t, \xi_S \leq s)] p(E(\bxi)=) dE(\bxi)=,
 %\end{equation}
@@
 
-Recall AUV's can adapt survey plans based on what is being measured,
-in accordance with the sense, plan, act loop
-(Fig. \ref{fig:sense-plan-act}). Here we present adaptive strategies
-for sampling.
+An important synergy underlining this work is the potential of robotic sampling to adapt survey plans based on data and statistical inference,
+in accordance with the sense-plan-act loop
+(Fig. \ref{fig:sense-plan-act}). We will therefore present a few adaptive strategies
+for robotic sampling that builds on the notion of multivariate ES. Simulations, analysis, and  results from full-scale trials are presented.
 
 \subsection{Optimal sequential design}
-\label{myopic}
+\label{Optdes}
 
@@
 Eq. (\ref{gp_upd}), with mean $\bm_{s-1}=E(\bxi|\mathcal{Y}^{s-1})$
-and covariance matrix $\bS_{s-1}=\mbox{Var}(\bxi|\mathcal{Y}^{s-1})$.
+and covariance matrix $\bA_{s-1}=\mbox{Var}(\bxi|\mathcal{Y}^{s-1})$.
 The next round of measurements $\by^{j,s}$, can be gathered with
@@
     d^{*,s} &=& \mbox{argmin}_{j \in \{1,\ldots,J\}} |p(\bx_{d^{j,s}};\mathcal{Y}^{s-1})-0.5|, \\
-    p(\bx;\mathcal{Y}^{s-1}) &=& P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b | \mathcal{Y}^{s-1}). \nonumber
+    p(\bx;\mathcal{Y}^{s-1}) &=& P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S | \mathcal{Y}^{s-1}). \nonumber
 \end{eqnarray}
@@
 V_{m,\mbox{upd}} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j,s}|\mathcal{Y}^{s-1}} \left\{ p(\bx;\mathcal{Y}^{j,s})\left( 1-p(\bx;\mathcal{Y}^{j,s})\right) \right\} \Delta, \nonumber \\
-    p(\bx;\mathcal{Y}^{j,s}) &=& P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b |\by^{j,s},\mathcal{Y}^{s-1}). \nonumber
+    p(\bx;\mathcal{Y}^{j,s}) &=& P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\by^{j,s},\mathcal{Y}^{s-1}). \nonumber
 \end{eqnarray}
@@
 best design, the GP model is updated again. The mean $\bm_{s}$ and
-covariance matrix $\bS_{s}$ are used to compute the next design at
+covariance matrix $\bA_{s}$ are used to compute the next design at
 stage $s+1$, and so on.
@@
 V_{la,\mbox{upd},j_2} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j_2,s+1}|\mathcal{Y}^{j,s}} \left\{ p(\bx;\mathcal{Y}^{j,s+1})\left( 1-p(\bx;\mathcal{Y}^{j,s+1})\right) \right\} \Delta, \nonumber \\
-    p(\bx;\mathcal{Y}^{j,s+1}) &=& P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b |\by^{j_2,s+1},\mathcal{Y}^{j,s}). \nonumber
+    p(\bx;\mathcal{Y}^{j,s+1}) &=& P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\by^{j_2,s+1},\mathcal{Y}^{j,s}). \nonumber
 \end{eqnarray}
@@
 the current best design. When data is collected, the GP model is
-updated, and the mean $\bm_{s}$ and covariance matrix $\bS_{s}$ are
+updated, and the mean $\bm_{s}$ and covariance matrix $\bA_{s}$ are
 used to compute the next design at stage $s+1$, now anticipating what
@@
     \begin{bmatrix}
-    \xi_a(\bx) \\
-    \xi_b(\bx) 
-    \end{bmatrix}=\begin{bmatrix} \mu_{a}(\bx)\\
-\mu_{b}(\bx)
+    \xi_T(\bx) \\
+    \xi_S(\bx) 
+    \end{bmatrix}=\begin{bmatrix} \mu_{T}(\bx)\\
+\mu_{S}(\bx)
 \end{bmatrix} 
-= \begin{bmatrix} \beta_{a,0} + \beta_{a,1} x_{\mbox{East}} \\
-\beta_{a,1} + \beta_{b,1} x_{\mbox{East}}
+= \begin{bmatrix} \beta_{T,0} + \beta_{T,1} x_{\mbox{East}} \\
+\beta_{T,1} + \beta_{S,1} x_{\mbox{East}}
 \end{bmatrix}.
@@
 masses are pulled to the east. The actual trends are, in practice, set
-from preliminary information or runs of oceanographic models \kc{a
-  citation of such model use, e.g. the JFR paper might be in order
-  here} solving differential equations for hydrodynamic flow. In this
-example the mean is specified by $\beta_{a,1}=0.065$ and
-$\beta_{b,1}=0.1$. Further, the intercepts \kc{Whats an intercept?} of
-the temperature and salinity variables are set to $\beta_{a,0}=5.8$
-and $\beta_{b,1}=29.0$.
-
-%The prior assumption used by the agent are $\tilde{\beta_{a,1}}=0.085$ and $\tilde{\beta_{b,1}}=0.138$, which assumes that the process is centered.
+from preliminary information or runs of oceanographic models \cite{fossuminformation} solving differential equations for hydrodynamic flow. In this
+example the mean is specified by regression parameters $\beta_{T,1}=0.065$ and
+$\beta_{S,1}=0.1$ for the slope with west coordinate and $\beta_{T,0}=5.8$
+and $\beta_{S,1}=29.0$ for the intercept at a reference start map coordinate.
+
+%The prior assumption used by the agent are $\tilde{\beta_{T,1}}=0.085$ and $\tilde{\beta_{S,1}}=0.138$, which assumes that the process is centered.
 
@@
 \begin{bmatrix}
-    \xi_a(\bx) \\
-    \xi_b(\bx) 
+    \xi_T(\bx) \\
+    \xi_S(\bx) 
     \end{bmatrix}=
@@
 \end{equation}
-The spatial correlation is of a Matern type \kc{? Define or cite} giving:
+The spatial correlation is of a Matern type:
 \begin{equation}\label{v}
@@
 \begin{bmatrix}
-    \xi_a(\bx) \\
-    \xi_b(\bx) 
+    \xi_T(\bx) \\
+    \xi_S(\bx) 
     \end{bmatrix},
     \begin{bmatrix}
-    \xi_a(\bx') \\
-    \xi_b(\bx') 
+    \xi_T(\bx') \\
+    \xi_S(\bx') 
     \end{bmatrix}
@@
     \end{tabular}}
-  \caption{Simulation results for the final mean excursion set
-    variance (Eq. \eqref{two_partsK}) for 20 replicates, 8
+  \caption{Simulation results for the final mean IBV (Eq. \eqref{two_partsK}) for 20 replicates, 8
     configurations, and 6 strategies.}
@@
 
-To demonstrate the applicability of multivariate EP to informative
-oceanographic sampling, we present a case study mapping river plumes
+To demonstrate the applicability of multivariate EPs and the IBV to informative
+oceanographic sampling, we present a case study mapping a river plume
 with an AUV.  This experiment was performed in Trondheim, Norway,
 where the Nidelva river (Figure \ref{fig:nidelven}) meets water from
-the Trondheimsfjord. The experiment is focused along the eastern
+the Trondheim fjord. The experiment is focused along the eastern
 shoreline due to the aforementioned Coriolis
-force.%, see Fig. \ref{fig:map}.
-The frontal zone therefore runs parallel to the eastern shore as
-indicated in purple as it enters the fjord. The experiments were
-conducted in late Spring 2019, and because of snow melt with
-consequent fresh river water substantially colder than fjord waters.
+force, and the frontal zone should therefore run more or less parallel to the eastern shore as
+indicated in purple in Figure \ref{fig:nidelven}. The experiments were
+conducted in late Spring 2019, when there is still snow melting in the mountains so that the fresh river water is substantially colder than the water in the fjord.
 
@@
 
-The statistical model parameters were specified based on preliminary
-run where the AUV did an initial transect to determine the
+The statistical model parameters were specified based on a preliminary run where the AUV did an initial transect to determine the
 environmental conditions and correlation structure. In these runs the
@@
 
-Figure \ref{fig:parest} shows the diagnostic plots of this residual
-analysis. The figure on the left shows a cross-plot of temperature and
-salinity residuals, indicating larger variability in salinity than in
+Figure \ref{fig:parest} shows diagnostic plots of this
+analysis. The display on the left shows a cross-plot of temperature and
+salinity residuals, after the westernly trends in both salinity and temperature are subtracted from the data. This scatterplot of joint residuals indicates larger variability in salinity than in
 temperature, and a positive correlation ($0.5$) between the two
-variables. Based on the fitted bivariate Gaussian model we can then
-compute the quadratic form of the residuals. If the Gaussian model is
-adequate the residuals should be approximately $\chi^2_2$
+variables. Based on the fitted bivariate Gaussian model (ellipses in left display), we can compute the modeled quadratic form of the residuals, and if the model is
+adequate they should be approximately $\chi^2_2$
 distributed. Figure \ref{fig:parest} (middle) shows the empirical
@@
 (solid) together with the theoretical CDF of the $\chi^2_2$
-distribution. Even though there appears to be some clustering in both
-left and middle displays, the fit appears reasonable, and justifies
-the use of a Gaussian bivariate model. Figure \ref{fig:parest} (right)
-shows the empirical variogram of the scaled residuals for temperature
-and salinity. The decay is similar, and seems to be negligible after
-about $150$ m. The resulting parameters are given in Table
-\ref{tab:experiment_param} below.
+distribution.
+The modeled and theoretical curves are very similar and this indicates that the Gaussian model fits reasonable well.
+Even though there appears to be some clustering in both left and middle displays of Figure \ref{fig:parest}, the bivariate diagnostic plots look reasonable and justify a Gaussian model.
+Figure \ref{fig:parest} (right)
+shows the empirical variogram of the scaled residuals for temperature and salinity. The decay is similar for the two, and seems to be negligible after
+about $150$ m.
+
+The resulting parameters are given in Table
+\ref{tab:experiment_param}.
+The regression parameters shown here are scaled to represent the east and west boundaries of the domain as seen in the preliminary transect data, and the thresholds are intermediate values.
+These parameter values were then used in field trials, where we explored the algorithm's ability to characterize the river plume front separating the river and fjord water masses.
 
@@
   Gaussian contours. Middle: Empirical CDF of the quadratic form of
-  the residuals along with the theoretical CDF of the Kji-square
+  the residuals along with the theoretical CDF of the $\chi^2$
   distribution with two degrees of freedom. Right: Empirical variogram
@@
 
-These mean and covariance parameters values were then used in a field
-trial, where we explored the algorithm's ability to characterize the
-front between the river and fjord water masses.
-
 \subsection{Experimental setup}
@@
 \citep{py10,Rajan12,Rajan12b}, running an instance of the myopic
-strategy from Section \ref{sec:sampling_designs}, to control the AUV
-and decide between sampling locations.
+strategy from Section \ref{sec:sampling_designs}, to control the AUV and decide between sampling locations.
 
@@
 (conductivity, temperature, and depth) sensor providing temperature
-and salinity measurements to the bi-variate GP model. The accuracy of
+and salinity measurements. The accuracy of
 the CTD instrument is $\pm 0.0003$ S/m (conductivity) and
@@
 
-The sampling strategy was designed around sequentially visiting
-waypoints, where arrival at a desired waypoint subsequently triggers
-the myopic strategy in order to evaluate and find the next waypoint
-based on the ES variance formulated in Eq. \eqref{critSEQ}. The
-waypoint --and associated sampling-- that results in lowering the ES
-variance the most then, is selected. For the field trials, the AUV
-started in the lower middle of the graph with a GP prior estimate of
-the environment given as a linear gradient from west to east, between
-the initially measured temperature and salinity. Each AUV
-survey/mission was set to take approx. 40 min, visiting 15 waypoints
-on the grid, with the AUV running at the surface to better capture the
-plume. The GP model assimilated temperature and salinity data
-continuously with an update frequency of 30 sec.
+The sampling strategy was designed around sequential visiting of waypoints defined by the grid.
+Arriving at a desired waypoint, with new measurements and an updated model, the AUV triggers the myopic strategy to evaluate all possible design alternatives as in Eq.
+\eqref{critSEQ}.
+The design which is expected to reduce the IBV the most is selected and defines the next waypoint.
+This procedure is repeated for several stages in both surveys.
+At each stage, it takes the AUV about 30 sec to evaluate the expected IBV all possible design strategies.
+
+For the field trials, the AUV
+started in the south center part of the waypoint graph with the outlined GP prior model of
+the environment. Each AUV
+survey was set to take approximately 40 min, visiting 15 waypoints on the grid, with the AUV running near the surface to better capture the
+plume.
+On its path from one waypoint to the next, the AUV gathered data regularly, and the GP model assimilated these temperature and salinity data with an update frequency of 30 sec.
+This means about 3 measurements of temperature and salinity per stage.
 
@@
 
-Two missions, Survey 1 and 2, was run successively from 11:00 AM to
-01:00 PM, with a short break in between. The resulting path between
-the selected waypoints are shown in spatial context in
+Two missions, Survey 1 and 2, were run successively from 11:00 AM to
+01:00 PM, with a short break in between. The resulting path of
+the selected waypoints are shown in the map in
 Fig. \ref{fig:map}, both within the expected frontal region. The
@@
 between fjord and river waters. The temperatures are then shown
-separately overlaid with the estimated ES for each survey in
+separately overlaid with the estimated EP for each survey in
 Fig. \ref{fig:res1} and \ref{fig:res2}.
 
-Both Survey 1 and 2, successfully estimate and navigate the separation
-zone while crossing the front multiple times. As conditions changed
+Both Survey 1 and 2 successfully estimate and navigate the separation
+zone while crossing the front multiple times. As conditions changed slightly 
 between the two deployments the resulting path (after waypoint 5
@@
 northwards tracking the north-eastern part of the front, while Survey
-2 turns west mapping the south-western part. The final prediction of
-the front location given Fig. \ref{fig:res3} and \ref{fig:res4},
+2 turns west mapping the south-western part.
+
+The final prediction of
+the front location given in Fig. \ref{fig:res3} and \ref{fig:res4},
 correspond with one another yielding a picture of the front gradually
 bending off towards north east. The amount of exploration done by
-Survey 1 is greater than Survey 2, coming close to the survey area
-borders in the south-western corner. A longer planning horizon would
+Survey 1 is greater than Survey 2. In Survey 1 the AUV gets overall more detail by going north, while Survey 2 coming close to the survey area
+borders in the south-western corner get poorer understanding of the northern parts. A longer planning horizon would likely 
 identify and discourage such choices.
@@
 
-The methods presented in this paper are shown to be useful in
-characterizing the frontal water masses where cold river water is
-separated from warmer saline conditions. The study did not consider
+This paper builds on multidisciplinary work combining statistical modeling and methods with new robotic surveying technologies for an oceanographic application.
+We show how the application can gain from new statistical approaches to achieve more effective spatial monitoring.
+At the same time, we show that novel autonomous sensing techniques give opportunities for multivariable spatial data gathering which statisticians can learn from to develop new general-purpose toolkits.
+
+In particular, we derive and show results for
+characterizing river front water masses with cold river water separating from warmer saline fjord conditions.
+The characterization is based on joint excursion sets and we derive new results for the expected integrated Bernoulli variance reduction achieved by the spatial sampling design.
+This result is first derived for the situation with a static design, and then extended to the adaptive situation which provides new insight in our application with the autonomous vehicle.
+
+The study did not consider
 any temporal effects, which would be relevant on a larger time
-scale. We consider the extension to spatio-temporal modeling as future
-work, and we envision that advection-diffusion equations could be
-useful in this kind of modeling \cite{sigrist2015stochastic}.  For
-more complex oceanographic phenomena, the methods will need to be
-extended to non-Gaussian phenomena and augmented by dynamical models.
-
-While more effort could be spent on approximating the look-ahead
-steps, it is perhaps more interesting to explore the additional
-flexibility that can be gained by having two or more AUVs exploring a
+scale.
+We consider the extension to spatio-temporal modeling as future work, and we envision that advection-diffusion equations could be useful in this kind of modeling \citep{sigrist2015stochastic}.
+For more complex oceanographic phenomena, the methods will need to be
+extended to non-Gaussian phenomena, possibly feature-based mixtures of GPs which could still be run onboard, and augmented by dynamical models.
+
+The spatial statistical design criterion building on excursion sets is relevant in our setting with the different water masses and could also be useful in other oceanographic settings such as algea blooms or frontal systems.
+Other criteria more directly related to decision making would also be relevant, especially with much spatio-temporal effects.
+For instance, managers must make difficult decisions related to fish farming or other marine resources operations, or environmental projects.
+Value of information analysis \citep{Eidsvik:15} would then be used in a similar vein as the integrated Bernoulli variance in the current paper, now studying which, when and where information is likely to materialize in improved decisions.
+
+While more effort could be spent on approximating the look-ahead steps, it is perhaps more interesting to explore the additional
+flexibility that can be gained by having two or more AUVs \citep{ferreira2019advancing} exploring a
 spatial or spatio-temporal domain together. It might then be useful to
@@
 \section*{Acknowledgements}
-Thanks to SINTEF Ocean, AMOS, etc..\kc{Complete}
+Thanks to SINTEF Ocean, AMOS, etc.JE acknowledge support from Norwegian research council project 294404.
+DG acknowledges support from the Swiss National Science Foundation project number 178858.\kc{Complete}
 

### 2019-09-23 main_new.tex
Updates Trygve

@@
 % natbib option: authoryear
+
 \documentclass[aoas]{imsart}
@@
 
-\usepackage[tableposition = top]{caption} % captions
+%\usepackage[tableposition = top]{caption} % captions
 \usepackage{booktabs}                     % horizontal lines in tables
@@
 % \affiliation{Some University}
-
-\author{\fnms{Trygve Olav} \snm{Fossum}\ead[label=e1]{trygve.o.fossum@ntnu.no}, \fnms{Jo} \snm{Eidsvik}\ead[label=e2]{jo.eidsvik@math.ntnu.no}, \fnms{David} \snm{Ginsbourger} \ead[label=e3]{david.ginsbourger@idiap.ch}  \and \fnms{Kanna} \snm{Rajan}\ead[label=e4]{kanna.rajan@ntnu.no}}
+\begin{aug}
+\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{a}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
+\author{\fnms{Jo} \snm{Eidsvik}\thanksref{b}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
+\author{\fnms{David} \snm{Ginsbourger}\thanksref{c}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
+\and
+\author{\fnms{Kanna} \snm{Rajan}\thanksref{d}. \ead[label=e4]{kanna.rajan@ntnu.no}}
 
 \address{Department of Marine Technology\\ Trondheim, Norway\printead{e1}}
-\affiliation{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway}
+\affiliation[a]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU) \&\\ Centre for Autonomous Marine Operations and Systems (AMOS), Trondheim, Norway.}
 
 \address{Department of Mathematical Sciences \printead{e2}}
-\affiliation{Department of Mathematical Sciences, NTNU}
+\affiliation[b]{Department of Mathematical Sciences, NTNU.}
 
 \address{Idiap \printead{e3}}
-\affiliation{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland}
+\affiliation[c]{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland.}
 
 \address{Department of Engineering Cybernetics,\printead{e4}}
-\affiliation{Department of Engineering Cybernetics, NTNU \&\\ Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal}
-
-\runauthor{Fossum et al.}
+\affiliation[d]{Department of Engineering Cybernetics, NTNU \&\\ Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
+
+\runauthor{TO. Fossum et al.}
+\end{aug}
 
@@
 Norway\footnote{Report from the Institute of Marine
-  Research:https://bit.ly/2KZGJfI}, with losses exceeding 12 million
+  Research: https://bit.ly/2KZGJfI}, with losses exceeding 12 million
 tons of salmon valued at \$ 80 million (Norwegian Directorate of
@@
 
-\newpage
 \subsection{Static and sequential sampling designs}\label{sec:sampling_designs}
@@
 We compare three different static designs denoting
-\texttt{static\_north}, \texttt{static\_east} and
+\texttt{static\_north}, \texttt{static\_east}, and
 \texttt{static\_zigzag} with the three sequential approaches
@@
 sampling paths are pre-determined and cannot be changed on the basis
-of data, representing the archetypical pre-planned observation
+of data, representing the typical pre-planned observation
 strategies used in typical AUV operational survey designs.
 
-For a fixed survey length, the expected variance after static
-surveying is analytically available (equal coverage each time), but
-here we focus on the properties from the sequential designs which are
-evaluated using Monte Carlo sampling over several replicates of
-realizations from the prior, and conducting simulated sequential
-surveying for each one, making the result dependent on observed
-data. For comparison, the same setup is used for static
-designs. Figure \ref{fig:Eprob} shows the conditional EP, given data
-gathered along the north-south survey line for the realization
-indicated in these displays. In the Monte Carlo replicates, such
-results are repeatedly computed to approximate the expected variance
-reduction in the ES. Along with this criterion we also compare
-predictive performance measured by root mean square error (RMSE) in
-the temperature and salinity fields and variance reduction in these
-two variables. It is important to note that the objective function
-used by the agents is focused on reducing the expected variance in the
-ES, but we nevertheless hope to achieve good predictive performance
-for criteria such as RMSE as well. Another non-statistical criteria
-that is important for practical purposes is the computational time of
-the strategy, as this will impact the performance of the approach in
-the field for an embedded system. In Fig. \ref{fig:wp_graph}, the
-sequential sampling agents start at the center East-West coordinate at
-the southern end of the domain (node 53).
-
-\begin{figure}[t]
+For a fixed survey length, the expected variance is analytically available (equal coverage and reduction in variance each time).
+However, for the sequential approaches this is not the case, as the result and path are dependent on the observed data and the expected reduction in IBV.
+The properties is therefore evaluated using Monte Carlo sampling over several replicates of realizations from the prior, conducting simulated sequential surveying for each one.
+Figure \ref{fig:Eprob} shows the conditional EP, given data gathered along the north-south survey line for the realization indicated in these displays.
+In the Monte Carlo replicates, such results are repeatedly computed to approximate the expected variance reduction in the ES.
+We also compare predictive performance measured by root mean square error (RMSE) for the temperature and salinity estimates and the variance reduction in these two variables.
+It is important to note that the objective function used by the agents is focused on reducing the expected variance in the ES, but we nevertheless hope to achieve good predictive performance for criteria such as RMSE as well.
+Another non-statistical criteria that is important for practical purposes is the computational time of the strategy, as this will impact the performance of the approach in the field for an embedded system.
+
+\begin{figure}[h!]
 \centering
@@
 \caption{The equilateral waypoint graph used to discretize the
-  trajectory choices over the $31\times31$ grid.}
+  trajectory choices over the $31\times31$ grid used to discretize the GP.}
 \label{fig:wp_graph}
@@
 
-Maintaining the example above, the AUV will move along edges in the
-Fig. \ref{fig:wp_graph} waypoint graph while the vehicle gathers
-measurements. These are assimilated in the GP model, before an
-evaluation of the next node to sample is conducted at the end of the
-edge. The paths will differ for the various sequential designs due to
-the individual design criteria and the simulated variability in the
-environment among the replicates.
-
-Results of the replicate runs are shown in Figure \ref{fig:sim_results},
-where the different criteria are plotted as a function of surveying
-distance. Here, expected ES variance in Figure \ref{fig:avg_ev} is
-reduced most under the \texttt{myopic} and \texttt{look-ahead}
-strategies, performing almost equally.
-
-\begin{figure}[!th]
+Each strategy is conducted on an equilateral grid shown in Fig.
+\ref{fig:wp_graph}.
+The sequential sampling agents start at the center East-West coordinate at the southern end of the domain (node 53).
+The AUV will move along edges in the waypoint graph while the vehicle gathers measurements.
+The data is assimilated into the GP model, before an evaluation of the next node to sample is conducted at the end of the edge.
+The paths will differ for the various sequential designs due to the individual design criteria and the simulated variability in the environment among the replicates.
+
+A total of 100 replicate simulation was conducted and the results are shown in Figure \ref{fig:sim_results},
+where the different criteria are plotted as a function of survey distance.
+Figure \ref{fig:avg_ev} shows the resulting drop in ES variance for each of the six strategies.
+The ES variance is reduced most under the \texttt{myopic} and \texttt{look-ahead} strategies, performing almost equally; this is expected as the two criteria (Eq.
+\eqref{critSEQ} and \eqref{critLA}) are much more sensitive to differences in IBV.
+The north-south design also does well in terms of ES variance reduction since the path is parallel to the  boundary between the water masses.
+
+\begin{figure}[h!]
   \centering
@@
 
-The north-south design also does well since it moves near the boundary
-of the water masses. In Figure \ref{fig:avg_rmse} and \ref{fig:avg_r2}
-we show results of RMSE and explained variance. Both \texttt{myopic}
-and \texttt{look-ahead} perform well, but some of the
-\texttt{static\_east} and \texttt{static\_zigzag} also achieve good
-results because they are set to cover large parts of the domain, and
-in doing so visit locations that are not considered interesting by the
-sequential strategies targeting ES variance.
+Figure \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop in RMSE and increase in explained variance.
+Both \texttt{myopic} and \texttt{look-ahead} perform well here, but some of the \texttt{static\_east} and \texttt{static\_zigzag} also achieve good results because they are pre-determined to cover large parts of the domain without re-visitation.
+Sequential strategies targeting ES variance will sometimes not reach similar coverage as interesting data may draw the path into turns and twists that may result in re-visitation.
 
@@
 
-Fig.  \ref{fig:avg_time} shows the computational effort; the
-\texttt{naive} strategy is on par with the static designs, while the
-\texttt{myopic} strategy is slower and \texttt{look-ahead} more so,
-with the latter reaching levels that are nearly impractical for
-execution on a vehicle -- with more pruning of branches in the graph,
-it would likely be faster. Then again, this pruning is likely the
-reason that this strategy does not outperform the myopic approach. In
-Figure \ref{fig:route_choices} we plot the realized sampling paths of
-the sequential schemes and static designs.
-
-The naive strategy gets stuck in the southern part of the domain in
-some cases, because it is too focused on the probabilities near
-$0.5$. The myopic strategy meanwhile, covers a wider domain than the
-look-ahead, possibly because the look-ahead strategy sees less value
-of going far west because it accounts for having to return east again.
-
-We further go on to study sensitivity of the result to different prior
-correlation values between temperature and salinity, prior standard
-deviations and spatial correlations, as well as those of measuring
-only temperature on-board the AUV (as indicated in the rows). We
-compare results at the end of the surveying distance. Expected ES
-variances are shown in Table \ref{tab:sim_res_ev}, results of RMSE are
-in Table \ref{tab:sim_res_rmse} and explained variance in Table
-\ref{tab:sim_res_r2}.  In all runs, the myopic and look-ahead
-strategies perform the best in terms of expected variance in the
-ES. Surprisingly, the myopic is on-par and even slightly better than
-the look-ahead strategy. Part of the reason is pruning of paths for
-the look-ahead strategy, but even with less pruning, the look-ahead
-does not outperform the myopic approach. % Of course, we are able to
-% make up cases where look-ahead beats myopic, but within our ranges of
-% realistic input parameter values it was not easy.
-The \texttt{static\_north} design is clearly better than the other two
-static designs, the reason being that the others do not focus on the
-areas where the ES is really uncertain.  By measuring temperature
-alone the expected variances in the ES is only slightly higher. This
-occurs because the ES is almost the same and because of the
-correlation of $0.6$ which means that temperature observations give
-information about salinity as well.
-
-For the other predictive measures, myopic and look-ahead do well,
-clearly beating the naive approach. The \texttt{static\_zigzag} design
-is usually the best among the static designs for these predictive
-performance measures because its pattern gives more effective coverage
-of the domain when there is significant spatial correlation as in this
-case.
+Fig.
+\ref{fig:avg_time} shows the computational effort; the \texttt{naive} strategy is on par with the static designs, while the \texttt{myopic} strategy is slower and \texttt{look-ahead} more so, with the latter reaching levels that are nearly impractical for execution on a vehicle.
+Some pruning of the graph is performed to improve the computational speed, such as ruling out repeated visitations and back-and-forth routes.
+Some of the intermediate results are also stored for longer planning horizons.
+Still, more pruning of branches or inclusion of other heuristics could be included to make the strategy faster.
+Then again, the inclusion of heuristics is likely a contributing factor that the \texttt{look-ahead} strategy is not outperforming the \texttt{myopic} strategy.
+In Figure \ref{fig:route_choices} the realized sampling paths for each sequential schemes and static designs is shown.
+
+The \texttt{naive} strategy gets stuck in the southern part of the domain in some cases, because it is too focused on the probabilities near $0.5$.
+The \texttt{myopic} strategy meanwhile, covers a wider domain than the look-ahead, there are several reasons for this.
+First, a greedy approach will tend to do more put more emphasis on locations close to the agent, which may lead away from the centre.
+Secondly, as the agent evaluates the impact of locations further away (look-ahead) where assimilated data has less effect, the prior (which is centered here) will act to restrict paths deviating from the central zone.
+
+We further go on to study sensitivity of the result to different prior correlation values between temperature and salinity, prior standard deviations, spatial correlations, as well as measuring only temperature on-board the AUV (as indicated in the rows).
+We only compare results at the end of each survey.
+The expected ES variances are shown in Table \ref{tab:sim_res_ev}, results of RMSE are in Table \ref{tab:sim_res_rmse}, and explained variance in Table \ref{tab:sim_res_r2}.
+In all runs, both the \texttt{myopic} and \texttt{look-ahead} strategies perform the best in terms of expected variance in the ES.
+Surprisingly, the \texttt{myopic} is on-par and even slightly better than the \texttt{look-ahead} strategy.
+Part of the reason can be traced to the act of pruning some of the paths for the \texttt{look-ahead} strategy, but even with less pruning, the \texttt{look-ahead} does not outperform the \texttt{myopic} approach.
+The \texttt{static\_north} design is clearly better than the other two static designs, the reason being that the others do not focus on the areas where the ES is really uncertain (centre locations are the most uncertain).
+By measuring temperature alone the expected variances in the ES is only slightly higher.
+This occurs because the ES is almost the same and because of the correlation of $0.6$, which means that temperature observations provide information about salinity as well.
+For the other predictive measures, \texttt{myopic} and \texttt{look-ahead} do well, clearly beating the \texttt{naive} strategy.
+The \texttt{static\_zigzag} design is usually the best among the static designs for these predictive performance measures because its pattern gives more effective coverage of the domain when there is significant spatial correlation as in this case.
+
 \begin{table}[!h]
@@
     \end{tabular}}
-  \caption{Simulation results for the final mean IBV (Eq. \eqref{two_partsK}) for 20 replicates, 8
-    configurations, and 6 strategies.}
+  \caption{Simulation results for the final mean IBV (Eq. \eqref{two_partsK}).}
     \label{tab:sim_res_ev}
 \end{table}
-
+\vspace{-0.4cm}
 \begin{table}[!h]
@@
         \end{tabular}}
-      \caption{Simulation results for the final mean RMSE
-        ($\frac{1}{B} \sum_{b=1}^{B} \sqrt{(\bmu-\tilde{\bmu})^2}$)
-        for 20 replicates, 8 configurations, and 6 strategies.}
+      \caption{Simulation results for the final mean RMSE ($\frac{1}{B} \sum_{b=1}^{B} \sqrt{(\bmu-\tilde{\bmu})^2}$).}
     \label{tab:sim_res_rmse}
 \end{table}
-
+\vspace{-0.4cm}
 \begin{table}[!h]
@@
         \end{tabular}}
-      \caption{Simulation results for the final mean explained
-        variance
-        ($\bR^{2}=100*(1-(\bSigma_{posterior}/\bSigma_{initial}))$)
-        for 20 replicates, 8 configurations, and 6 strategies.}
+      \caption{Simulation results for the final mean explained variance $\bR^{2}=100*(1-(\bSigma_{posterior}/\bSigma_{initial}))$).}
     \label{tab:sim_res_r2}
@@
 
-To demonstrate the applicability of multivariate EPs and the IBV to informative
-oceanographic sampling, we present a case study mapping a river plume
-with an AUV.  This experiment was performed in Trondheim, Norway,
-where the Nidelva river (Figure \ref{fig:nidelven}) meets water from
-the Trondheim fjord. The experiment is focused along the eastern
-shoreline due to the aforementioned Coriolis
-force, and the frontal zone should therefore run more or less parallel to the eastern shore as
-indicated in purple in Figure \ref{fig:nidelven}. The experiments were
-conducted in late Spring 2019, when there is still snow melting in the mountains so that the fresh river water is substantially colder than the water in the fjord.
+To demonstrate the applicability of using multivariate EPs and the IBV to inform oceanographic sampling, we present a case study mapping a river plume with an AUV.
+The experiment was performed in Trondheim, Norway,
+surveying the Nidelva river (Figure \ref{fig:nidelven}).
+Here cold and fresh water from the river meet the warm and saline ocean water in the Trondheim fjord.
+The experiment is focused along the eastern shoreline due to the aforementioned Coriolis force.
+The frontal zone therefore runs more or less parallel to the eastern shore as
+indicated in purple in Figure \ref{fig:nidelven}.
+The experiments were conducted in late Spring 2019, when there is still snow melting in the surrounding mountains so that the river water is substantially colder than the water in the fjord.
 
@@
 
-The statistical model parameters were specified based on a preliminary run where the AUV did an initial transect to determine the
-environmental conditions and correlation structure. In these runs the
-trend parameters were estimated by linear regression, where both
-temperature and salinity are assumed to increase linearly, going west
-from the river mouth. Next, the residuals from the regression analysis
-are analyzed to study the fit of the Gaussian model and to specify the
-covariance parameter.
-
-Figure \ref{fig:parest} shows diagnostic plots of this
-analysis. The display on the left shows a cross-plot of temperature and
-salinity residuals, after the westernly trends in both salinity and temperature are subtracted from the data. This scatterplot of joint residuals indicates larger variability in salinity than in
-temperature, and a positive correlation ($0.5$) between the two
-variables. Based on the fitted bivariate Gaussian model (ellipses in left display), we can compute the modeled quadratic form of the residuals, and if the model is
-adequate they should be approximately $\chi^2_2$
-distributed. Figure \ref{fig:parest} (middle) shows the empirical
-(cumulative distribution function) CDF of these quadratic forms
-(solid) together with the theoretical CDF of the $\chi^2_2$
-distribution.
-The modeled and theoretical curves are very similar and this indicates that the Gaussian model fits reasonable well.
-Even though there appears to be some clustering in both left and middle displays of Figure \ref{fig:parest}, the bivariate diagnostic plots look reasonable and justify a Gaussian model.
-Figure \ref{fig:parest} (right)
-shows the empirical variogram of the scaled residuals for temperature and salinity. The decay is similar for the two, and seems to be negligible after
-about $150$ m.
-
-The resulting parameters are given in Table
-\ref{tab:experiment_param}.
-The regression parameters shown here are scaled to represent the east and west boundaries of the domain as seen in the preliminary transect data, and the thresholds are intermediate values.
-These parameter values were then used in field trials, where we explored the algorithm's ability to characterize the river plume front separating the river and fjord water masses.
-
-%Mapping the spatial extent of a frontal zones is an important problem for studying many bio-physical interactions in the ocean.
-The frontal zone is determined by the boundary where plumes of sediments, nutrients, and possibly pollutants spreading from the river outlet meet and interact with adjacent coastal water.
-Due to the lower density the plumes spread on the surface, creating a front with an sharp gradient in both temperature and salinity.
+The statistical model parameters were specified based on a short preliminary survey where the AUV did an initial transect to determine the environmental conditions and correlation structure.
+Based on the initial data the trend parameters were estimated by linear regression, where both temperature and salinity are assumed to increase linearly, going west from the river mouth.
+Next, the residuals from the regression analysis are analyzed to study the fit of the Gaussian model and to specify the covariance parameters.
 
@@
  \centering 
-\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.eps}
+\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
 \caption{Data analysis from a preliminary trial experiment using the
@@
 
+Figure \ref{fig:parest} shows diagnostic plots of this analysis.
+The display on the left shows a cross-plot of temperature and salinity residuals, after the westernly trends in both salinity and temperature are subtracted from the data.
+This scatterplot of joint residuals indicates larger variability in salinity than in temperature, and a positive correlation ($0.5$) between the two variables.
+Based on the fitted bivariate Gaussian model (ellipses in left display), we can compute the modeled quadratic form of the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed.
+Figure \ref{fig:parest} (middle) shows the empirical (cumulative distribution function) CDF of these quadratic forms (solid) together with the theoretical CDF of the $\chi^2_2$ distribution.
+The modeled and theoretical curves are very similar and this indicates that the Gaussian model fits reasonable well.
+Even though there appears to be some clustering in both left and middle displays of Figure \ref{fig:parest}, the bivariate diagnostic plots look reasonable and justify a Gaussian model.
+Figure \ref{fig:parest} (right) shows the empirical variogram of the scaled residuals for temperature and salinity.
+The decay is similar for the two, and seems to be negligible after about $150$ m.
+
+Based on the analysis in Fig.
+\ref{fig:parest} the resulting parameters are given in Table \ref{tab:experiment_param}.
+The regression parameters shown here are scaled to represent the east and west boundaries of the domain as seen in the preliminary transect data, and the thresholds are intermediate values.
+These parameter values were then used in field trials, where we explored the algorithm's ability to characterize the river plume front separating the river and fjord water masses, providing a spatial map of this boundary.
+
+%Mapping the spatial extent of a frontal zones is an important problem for studying many bio-physical interactions in the ocean.
+The frontal zone is determined by the boundary where plumes of sediments, nutrients, and possibly pollutants spreading from the river outlet meet and interact with adjacent coastal water.
+Due to the lower density the plumes spread on the surface, creating a front with an sharp gradient in both temperature and salinity.
+
 \begin{table}[!h]
@@
 \rowcolor{Gray}
-River sal. $S_{river}$ & $14.0$g/kg & AUV observations\\
-Ocean sal. $S_{ocean}$ & $22.0$g/kg & AUV observations\\
+River sal. $S_{river}$ & $14.0$ g/kg & AUV observations\\
+Ocean sal. $S_{ocean}$ & $22.0$ g/kg & AUV observations\\
 \rowcolor{Gray}
 Threshold temp. & $10.5\,^{\circ}\mathrm{C}$ & $(T_{ocean}-T_{river})/2+T_{river}$\\
-Threshold sal. & $22.0$g/kg & $(S_{ocean}-S_{river})/2+S_{river}$\\
+Threshold sal. & $18.0$ g/kg & $(S_{ocean}-S_{river})/2+S_{river}$\\
 \rowcolor{Gray}
@@
 
-Two full-scale deployments were conducted with adaptive AUV sampling.
-%on the 2$^{nd}$ of July 2019 spatially mapping the location of the plume from the Nidelven river (Trondheim, Norway) spreading into the Trondheimsfjord 
-The AUV used the onboard autonomous agent \texttt{T-REX}
-\citep{py10,Rajan12,Rajan12b}, running an instance of the myopic
-strategy from Section \ref{sec:sampling_designs}, to control the AUV and decide between sampling locations.
-
-\begin{figure}[!b] 
+The sampling locations was distributed over an equilateral grid, as shown in Fig.
+\ref{fig:map} in a grey colored lattice.
+The robotic platform consisted of a Light AUV \citep{sousa2012lauv} (Figure \ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49 CTD (conductivity, temperature, and depth) sensor providing temperature and salinity measurements.
+The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
+The sampling agent was built on top of the autonomous agent framework \texttt{T-REX} \citep{py10,Rajan12,Rajan12b}, running an instance of the \texttt{myopic} strategy from Section \ref{sec:sampling_designs}, to control the AUV and decide between sampling locations.
+
+\begin{figure}[!h] 
 \centering 
-\includegraphics[width=0.98\textwidth]{Figures/LAUV.jpg}
-\caption{The commercially available Light Autonomous Underwater
-  Vehicle (LAUV) used in the experiments in Trondheim.}
+\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
+\caption{The commercially available Light Autonomous Underwater Vehicle (LAUV) platform (Harald) for upper water-
+column exploration used in the experiment.}
 \label{fig:lauv}
@@
 
-The sampling locations are distributed over an equilateral grid, as
-shown in Fig. \ref{fig:map} in a grey colored lattice. The robotic
-platform consisted of a Light AUV \citep{sousa2012lauv} (Figure
-\ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49 CTD
-(conductivity, temperature, and depth) sensor providing temperature
-and salinity measurements. The accuracy of
-the CTD instrument is $\pm 0.0003$ S/m (conductivity) and
-$\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
-
-%\begin{figure}[!h] 
-%\centering 
-%\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
-%\caption{The light AUV platform (Harald) for upper water-
-%column exploration used in the experiment.}
-%\label{fig:harald_auv}
-%\end{figure}
-
-The sampling strategy was designed around sequential visiting of waypoints defined by the grid.
-Arriving at a desired waypoint, with new measurements and an updated model, the AUV triggers the myopic strategy to evaluate all possible design alternatives as in Eq.
+The sampling strategy was designed around the concept of visiting waypoints sequentially.
+Arriving at a desired waypoint, with new measurements and an updated model, the AUV triggers the myopic strategy to evaluate all possible design criteria, e.g Eq.
 \eqref{critSEQ}.
-The design which is expected to reduce the IBV the most is selected and defines the next waypoint.
-This procedure is repeated for several stages in both surveys.
-At each stage, it takes the AUV about 30 sec to evaluate the expected IBV all possible design strategies.
-
-For the field trials, the AUV
-started in the south center part of the waypoint graph with the outlined GP prior model of
-the environment. Each AUV
-survey was set to take approximately 40 min, visiting 15 waypoints on the grid, with the AUV running near the surface to better capture the
-plume.
-On its path from one waypoint to the next, the AUV gathered data regularly, and the GP model assimilated these temperature and salinity data with an update frequency of 30 sec.
-This means about 3 measurements of temperature and salinity per stage.
-
-\subsection{Results}
-
-Two missions, Survey 1 and 2, were run successively from 11:00 AM to
-01:00 PM, with a short break in between. The resulting path of
-the selected waypoints are shown in the map in
-Fig. \ref{fig:map}, both within the expected frontal region. The
-recorded temperatures are shown as colored trails in
-Fig. \ref{fig:res_both}, clearly showing the temperature difference
-between fjord and river waters. The temperatures are then shown
-separately overlaid with the estimated EP for each survey in
-Fig. \ref{fig:res1} and \ref{fig:res2}.
-
-Both Survey 1 and 2 successfully estimate and navigate the separation
-zone while crossing the front multiple times. As conditions changed slightly 
-between the two deployments the resulting path (after waypoint 5
-\kc{This should be marked in the figure. And all figures need to be
-  zoomed up, since they're too small}) deviates. Survey 1 continues
-northwards tracking the north-eastern part of the front, while Survey
-2 turns west mapping the south-western part.
-
-The final prediction of
-the front location given in Fig. \ref{fig:res3} and \ref{fig:res4},
-correspond with one another yielding a picture of the front gradually
-bending off towards north east. The amount of exploration done by
-Survey 1 is greater than Survey 2. In Survey 1 the AUV gets overall more detail by going north, while Survey 2 coming close to the survey area
-borders in the south-western corner get poorer understanding of the northern parts. A longer planning horizon would likely 
-identify and discourage such choices.
-
-\begin{figure}[!h]
+The waypoint-and-path combination which is expected to reduce the IBV the most is selected; upon arrival this procedure is then repeated.
+At each stage, it takes the AUV about 30 sec to evaluate the expected IBV for all the possible waypoint-and-path alternatives.
+
+The AUV was set to start in the south center part of the waypoint graph, with the previously outlined GP prior model of the environment.
+Each AUV survey was set to take approximately 40 min, visiting 15 waypoints on the grid, with the AUV running near the surface to better capture the plume.
+On its path from one waypoint to the next, the AUV gathered data regularly, and the GP model assimilated temperature and salinity data with an update frequency of 30 sec, about three updates per stage.
+
+\begin{figure*}[!h]
 \centering
-\subfigure[Survey map.]{\includegraphics[height = 0.31\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
-\hfill
-\subfigure[Temperature tracks.]{\includegraphics[height = 0.31\textwidth]{Figures/field-trials/auv.pdf}\label{fig:res_both}}
-\hfill
-\subfigure[Survey 1]{\includegraphics[height = 0.31\textwidth]{Figures/field-trials/auv1_es.pdf}\label{fig:res1}}
-\hfill
-\subfigure[Survey 2]{\includegraphics[height = 0.31\textwidth]{Figures/field-trials/auv4_es.pdf}\label{fig:res2}}
-\hfill
-\subfigure[ES for Survey 1.]{\includegraphics[height = 0.31\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}
-\hfill
-\subfigure[ES for Survey 2.]{\includegraphics[height = 0.31\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
+\subfigure[Survey map.]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
+\hspace{0.3cm}
+\subfigure[Temperature tracks.]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/auv.pdf}\label{fig:res_both}}
+
+\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv1_es_sal.pdf}\label{fig:res1}}
+\hspace{0.2cm}
+\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv4_es_sal.pdf}\label{fig:res2}}
+
+\subfigure[ES for Survey 1.]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
+\subfigure[ES for Survey 2.]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
 \caption{Results from mapping the Nidelven river} 
 \label{fig:results}
-\end{figure}
+\end{figure*}
+
+\subsection{Results}
+
+Two missions, Survey 1 and 2, were run successively from 11:00 AM to 01:00 PM, with a short break in between.
+The resulting path of the selected waypoints are shown in the map in Fig.
+\ref{fig:map}, both within the expected frontal region (shaded pink region).
+The recorded temperatures are shown as colored trails in Fig.
+\ref{fig:res_both}, clearly showing the temperature difference between fjord and river waters.
+The salinity data are then shown separately overlaid with the estimated EP for each survey in Fig.
+\ref{fig:res1} and \ref{fig:res2}.
+
+Both Survey 1 and 2 successfully estimate and navigate the separation zone crossing the frontal boundary multiple times.
+As conditions changed slightly between the two deployments the resulting path (after waypoint 5) deviates.
+Survey 1 continues northwards tracking the north-eastern part of the front, while Survey 2 turns west mapping the south-western part.
+
+The final prediction of the front location given in Fig.
+\ref{fig:res3} and \ref{fig:res4}, correspond with one another yielding a picture of the front gradually bending off towards north east.
+The amount of exploration done by Survey 1 is greater than Survey 2.
+In Survey 1 the AUV gets overall more detail by going north, while Survey 2 coming close to the survey area borders in the south-western corner get poorer understanding of the northern parts.
+A longer planning horizon would likely identify and discourage such choices.
 
@@
 
-This paper builds on multidisciplinary work combining statistical modeling and methods with new robotic surveying technologies for an oceanographic application.
-We show how the application can gain from new statistical approaches to achieve more effective spatial monitoring.
-At the same time, we show that novel autonomous sensing techniques give opportunities for multivariable spatial data gathering which statisticians can learn from to develop new general-purpose toolkits.
-
-In particular, we derive and show results for
-characterizing river front water masses with cold river water separating from warmer saline fjord conditions.
+This paper builds on multidisciplinary work combining statistical modeling and methods with robotic surveying technologies for oceanographic sensing applications.
+We show how the observation practices can gain from development of statistical techniques to achieve more effective spatial monitoring.
+We further demonstrate the opportunities available for real-time multivariable spatial data gathering and analysis onboard autonomous sensing platforms, upon which statisticians can exploit towards creating general-purpose toolkits for similar applications (enabeling theory).
+
+In particular, we derive and show results for characterizing phenomena connected to water masses properties.
 The characterization is based on joint excursion sets and we derive new results for the expected integrated Bernoulli variance reduction achieved by the spatial sampling design.
-This result is first derived for the situation with a static design, and then extended to the adaptive situation which provides new insight in our application with the autonomous vehicle.
-
-The study did not consider
-any temporal effects, which would be relevant on a larger time
-scale.
+This result is first derived for the situation with a static design, and then extended to the adaptive situation, which provides new insight in our application with the autonomous vehicle.
+
+The study did not consider any temporal effects, which would be relevant on a larger time scale.
 We consider the extension to spatio-temporal modeling as future work, and we envision that advection-diffusion equations could be useful in this kind of modeling \citep{sigrist2015stochastic}.
-For more complex oceanographic phenomena, the methods will need to be
-extended to non-Gaussian phenomena, possibly feature-based mixtures of GPs which could still be run onboard, and augmented by dynamical models.
-
-The spatial statistical design criterion building on excursion sets is relevant in our setting with the different water masses and could also be useful in other oceanographic settings such as algea blooms or frontal systems.
+For more complex oceanographic phenomena, the methods will need to be extended to non-Gaussian phenomena, possibly feature-based mixtures of GPs which could still be run onboard, and augmented by dynamical models.
+
+The spatial statistical design criterion building on excursion sets is relevant in our setting with the different water mass properties and could also be useful in other oceanographic settings such as algae-blooms or frontal systems.
 Other criteria more directly related to decision making would also be relevant, especially with much spatio-temporal effects.
@@
 
-While more effort could be spent on approximating the look-ahead steps, it is perhaps more interesting to explore the additional
-flexibility that can be gained by having two or more AUVs \citep{ferreira2019advancing} exploring a
-spatial or spatio-temporal domain together. It might then be useful to
-sample different parts of the space, and maybe move in parallel to
-best capture the excursion set.
+While more effort could be spent on approximating the look-ahead steps, it is perhaps more interesting to explore the additional flexibility that can be gained by having two or more AUVs \citep{ferreira2019advancing} exploring a spatial or spatio-temporal domain together.
+It might then be useful to sample different parts of the space, and maybe move in parallel to best capture the excursion set.
 
 \section*{Acknowledgements}
-Thanks to SINTEF Ocean, AMOS, etc.JE acknowledge support from Norwegian research council project 294404.
-DG acknowledges support from the Swiss National Science Foundation project number 178858.\kc{Complete}
+TOF and KR acknowledge support from the Centre for Autonomous Marine Operations and Systems (AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of Excellence, project number 223254, and the Applied Underwater Robotics Labortatory (AURLab).
+JE acknowledge support from Norwegian research council project 294404.
+DG acknowledges support from the Swiss National Science Foundation project number 178858.
 
@@
 \end{document}
-
-%\begin{enumerate}
-%\item  This is the first item of an enumerated list.  Each item
-%      in the list is marked with a ``tick.''  The document
-%      style determines what kind of tick mark is used.
-%\item  This is the second item of the list.  It contains another
-%      list nested inside it.  The three inner lists are an {\em enumerated}
-%      list.
-%    \begin{enumerate}
-%       \item This is the first item of an enumerated list that
-%            is nested within the enumerated list.
-%          \item This is the second item of the inner list.  \LaTeX\
-%            allows you to nest lists deeper than you really should.
-%      \end{enumerate}
-%      This is the rest of the second item of the outer list.  It
-%      is no more interesting than any other part of the item.
-%   \item  This is the third item of the list.
-%\end{enumerate}

### 2019-09-23 main_new.tex
updates

@@
 \begin{abstract}
-  Improving and optimizing oceanographic sampling is a crucial task for
-  modern marine management and science faced with limited resources to
-  understand highly dynamic processes in the water-column. The
-  combination of statistics and autonomous
-  robotics provides opportunities for studying aspects related to
-  experimental design for effective characterization of oceanographic
-  phenomena. In this work we develop methods for efficient spatial
-  sampling applied to the mapping of coastal processes by
-  providing informative descriptions of spatial characteristics of ocean
-  phenomena. Specifically, we define a design criterion based on
-  reducing uncertainty in the joint excursions of random
-  fields, and derive analytical expressions for the expected Bernoulli
-  variance reduction in the excursion set of spatial temperature and
-  salinity variables. We show how these can be applicable in static
-  designs and sequential designs, building on myopic and look-ahead
-  sampling strategies. We use simulations to study the properties and
-  differences between state-of-the-art approaches. Results from field
-  deployments with an autonomous underwater vehicle are also presented,
-  for a case study mapping the boundary of a river plume. The results
-  demonstrate the ability of the method to effectively inform and
-  execute data-driven environmental sampling.
+  Improving and optimizing oceanographic sampling is a crucial task
+  for modern marine management and science faced with limited
+  resources to understand highly dynamic processes in the
+  water-column. The combination of statistics and autonomous robotics
+  provides opportunities for studying aspects related to experimental
+  design for effective characterization of oceanographic phenomena. We
+  develop methods for efficient spatial sampling applied to the
+  mapping of coastal processes by providing informative descriptions
+  of spatial characteristics of ocean phenomena. Specifically, we
+  define a design criterion based on reducing uncertainty in the joint
+  excursions of random fields, and derive analytical expressions for
+  the expected Bernoulli variance reduction in the excursion set of
+  spatial temperature and salinity variables. We show how these can be
+  applicable in static and sequential designs, building on myopic and
+  look-ahead sampling strategies. We use simulations to study the
+  properties and differences between state-of-the-art
+  approaches. Results from field deployments with an autonomous
+  underwater vehicle are also presented, for a case study mapping the
+  boundary of a river plume. The results demonstrate the ability of
+  the method to effectively inform and execute data-driven
+  environmental sampling.
 
@@
 large-scale coverage, but have limited resolution, are limited to
-sensing the surface of the ocean, and is impacted by cloud cover. The
+sensing the surface of the ocean, and are impacted by cloud cover. The
 advent of robust mobile robotic platforms \cite{Bellingham07} has
 resulted in significant contributions to environmental monitoring and
-sampling. In particular autonomous underwater vehicles (AUVs), have made
-strides in advancing the state of sampling and consequently have made
-robotics an integral part of ocean observation in large part by our own
-work % driven by
+sampling. In particular autonomous underwater vehicles (AUVs), have
+made strides in advancing the state of sampling and consequently have
+made robotics an integral part of ocean observation; our own work
+sheds has contributed to this effort % driven by
 % using statistical methods in design and data assimilation
@@
   recently acquired information can be assimilated and used to improve
-  statistical models and reduce uncertainty associated with where to make measurements.
+  statistical models and reduce uncertainty associated with where to
+  make measurements.
 \item As a priori information about the state of the ocean is very
@@
 
-In this work, we build on joint excursion sets (ES) and excursion probabilities (EP) for spatial design of experiments. The goal is to
+In this work, we build on joint excursion sets (ES) and excursion
+probabilities (EP) for spatial design of experiments. The goal is to
 design sampling strategies for reducing the spatially integrated
@@
 
-\begin{figure}[!h] 
+\begin{figure}[!t] 
   \centering 
@@
 
-\begin{figure}[!h] 
+\begin{figure}[!t] 
 \centering 
@@
 
-AUVs are used to gather in-situ spatio-temporal data while carrying a range of scientific payloads to survey the
-water column. Typically, they operate at $\sim 1-3$ m/s, can reach
-depths from 100--6000 m with an in-water operation time capacity depending on
-survey speed, payload sensors and mission design. By deploying AUVs
-together with other resources one can augment information obtained
-from either ocean models, remote-sensing data, or fixed-location
-sensors on buoys. Traditional operations and surveys with AUVs are
-usually limited to observations along fixed transects on a spatial
-grid, pre-programmed by the human operator. By instead using on-board
-algorithms to continuously evaluate, update, and refine future
-sampling locations, one can make information gathering adaptive and
-contextual to the survey requirements
-\citep{das11b,fossuminformation,fossum18b}. The space of sampling
-opportunities is still limited to the spatial grid, but now there is
-much more flexibility because of adaptation. For instance, an AUV
-could reconstruct or modify a survey line based on what temperatures
-it is measuring by using embedded methods in decision-theoretic
-planning \citep{py10,Rajan12,Rajan12b}.
+AUVs are used to gather in-situ spatio-temporal data while carrying a
+range of scientific payloads to survey the water column. Typically,
+they operate at $\sim 1-3$ m/s, can reach depths from 100--6000 m with
+an in-water operation time capacity depending on survey speed, payload
+sensors and mission design. By deploying AUVs together with other
+resources one can augment information obtained from either ocean
+models, remote-sensing data, or fixed-location sensors on
+buoys. Traditional operations and surveys with AUVs are usually
+limited to observations along fixed transects on a spatial grid,
+pre-programmed by the human operator. By using on-board algorithms to
+continuously evaluate, update, and refine future sampling locations,
+one can make information gathering adaptive and contextual to the
+survey requirements \citep{das11b,fossuminformation,fossum18b}. The
+space of sampling opportunities while still limited to the spatial
+grid, allows for adaptation. For instance, an AUV could reconstruct or
+modify a survey line based on what temperatures it is measuring by
+using embedded methods in decision-theoretic planning
+\citep{py10,Rajan12,Rajan12b}.
 
@@
 
-Our objective is to leverage statistical tools to characterize a river plume, focusing on spatial separation of cold freshwater (from the river) and warmer saline waters of the fjord.
-Excursion sets (ES) and excursion probabilities (EP) are useful starting points to measure the ability to classify water masses.
-The goal of experimental sampling is to improve this characterization, as quantified via changes in the EPs.
+Our objective is to leverage statistical tools to characterize a river
+plume, focusing on spatial separation of cold freshwater (from the
+river) and warmer saline waters of the fjord. Excursion sets (ES) and
+excursion probabilities (EP) are useful starting points to measure the
+ability to classify water masses. The goal of experimental sampling is
+to improve this characterization, as quantified via changes in the
+EPs.
 
@@
 $\xi_T(\bx)$ denotes the temperature (in $^o C$) and $\xi_S(\bx)$
-denotes the salinity (in $mg/l$) at location $\bx$. The generalization to more than two variables is given in Section \ref{gen_k}.
-
-We let $t_T$ be a threshold for temperature and $t_S$ a threshold for salinity, the joint ES is then defined by:
+denotes the salinity (in $mg/l$) at location $\bx$. The generalization
+to more than two variables is given in Section \ref{gen_k}.
+
+We let $t_T$ be a threshold for temperature and $t_S$ a threshold for
+salinity, the joint ES is then defined by:
 \begin{equation}\label{ES}
@@
 \end{equation}
-If the EPs in Eq. (\ref{eq:prob}) are
-close to $1$ or $0$ at a given location, it is easy to classify water mass at this point as either river or ocean; it is however
-challenging if the probability is close to $0.5$. 
-The integrated Bernoulli variance (IBV) \cite{bect2019}, is defined as: 
+If the EPs in Eq. (\ref{eq:prob}) are close to $1$ or $0$ at a given
+location, it is easy to classify water mass at this point as either
+river or ocean; it is however challenging if the probability is close
+to $0.5$.  The integrated Bernoulli variance (IBV) \cite{bect2019}, is
+defined as:
 \begin{equation}\label{measVAR}
@@
 
-Calculations of the EPs and the IBV require a model specification.
-We use Gaussian modeling, which provides a practical and popular probabilistic approach to modeling environmental data, allowing a closed form
-evaluation of the IBVs, as derived in Section \ref{sec:sur}.
+Calculations of the EPs and the IBV require a model specification. We
+use Gaussian modeling, which provides a practical and popular
+probabilistic approach to modeling environmental data, allowing a
+closed form evaluation of the IBVs, as derived in Section
+\ref{sec:sur}.
 
@@
 
-The Gaussian model is used to represent the bivariate random fields of temperature and salinity. At location $\bx \in \mathcal{M}$, the prior bivariate distribution is:
+The Gaussian model is used to represent the bivariate random fields of
+temperature and salinity. At location $\bx \in \mathcal{M}$, the prior
+bivariate distribution is:
 \begin{equation}\label{gp_bar}
@@
 where the notation refers to Gaussian (Normal) distributed variables
-specified by the mean vector and covariance matrix. For the application
-of tracking river plumes, the model parameters are specified from a short preliminary survey.
-The mean varies with location, capturing the variability near the river mouth, while the variance parameters and correlation between salinity and temperature are assumed to be constant for all locations and within each water mass.
-
-For the spatial dependence among variables, we assume a separable correlation function with the same decay for salinity and temperature,
-which is not unrealistic for our case with a river plume as the decay is related to the common water masses (Sect \ref{sec:case_study}). We have:
+specified by the mean vector and covariance matrix. For the
+application of tracking river plumes, the model parameters are
+specified from a short preliminary survey. The mean varies with
+location, capturing the variability near the river mouth, while the
+variance parameters and correlation between salinity and temperature
+are assumed to be constant for all locations and within each water
+mass.
+
+For the spatial dependence among variables, we assume a separable
+correlation function with the same decay for salinity and temperature,
+which is not unrealistic for our case with a river plume as the decay
+is related to the common water masses (Sect \ref{sec:case_study}). We
+have:
 \begin{equation}\label{gp_corr}
@@
 for $i,j \in {T,S}$ and $\delta_{ij}=1$ if $i=j$ and $\delta_{ij}=0$
-otherwise. Moreover, we assume stationary isotropic random fields a priori where the
-correlation function $\rho(\bx,\bx')$ solely depends on $\bx$ and
-$\bx'$ via a Euclidean distance measure $h=\sqrt{||\bx-\bx'||^2}$. With
-data and prior knowledge one could possibly fit and estimate
-parameters of non-stationary or non-isotropic covariance functions and
-more complex multivariate spatial covariance functions
+otherwise. Moreover, we assume stationary isotropic random fields a
+priori where the correlation function $\rho(\bx,\bx')$ solely depends
+on $\bx$ and $\bx'$ via a Euclidean distance measure
+$h=\sqrt{||\bx-\bx'||^2}$. With data and prior knowledge one could
+possibly fit and estimate parameters of non-stationary or
+non-isotropic covariance functions and more complex multivariate
+spatial covariance functions
 \citep{gneiting2010matern,genton2015cross}; however that is left for
@@
 
-We discretize the spatial domain
-$\mathcal{M}$ to a set of $n$ grid locations
-$\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell has area $\Delta$. The
-grid is used for the waypoint graph setting the possible design locations. It is also used when solving integrals expressions in Eq. (\ref{measVAR}) and
-(\ref{sur}) which are then approximated by sums over all grid cells.  Using vector
-notation, we denote the Gaussian temperature and salinity variables on
-the grid and its density function by: 
+We discretize the spatial domain $\mathcal{M}$ to a set of $n$ grid
+locations $\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell
+has area $\Delta$. The grid is used for the waypoint graph setting the
+possible design locations. It is also used when solving integrals
+expressions in Eq. (\ref{measVAR}) and (\ref{sur}) which are then
+approximated by sums over all grid cells.  Using vector notation, we
+denote the Gaussian temperature and salinity variables on the grid and
+its density function by:
 \begin{equation}\label{prior}
@@
 temperature and salinity measurements gathered at the first round of
-nodes in the grid. The conditional model for the data, given the true temperature
-and salinity, is described by: 
+nodes in the grid. The conditional model for the data, given the true
+temperature and salinity, is described by:
 \begin{equation}\label{likelihood}
@@
 \end{equation}
-with entries $m_{\cdot}$ and $a_{\cdot}$ extracted from the conditional mean and covariance
-expressions in Eq. (\ref{gp_upd}).
+with entries $m_{\cdot}$ and $a_{\cdot}$ extracted from the
+conditional mean and covariance expressions in Eq. (\ref{gp_upd}).
 
@@
 \end{eqnarray}
-with parameters as defined in Eq.
-(\ref{gp_upd}) and (\ref{gp_hat}).
-Here, the probabilities would not only depend on the data but also on the data design defined via matrix $\bG$ in Eq.
-(\ref{likelihood}).
-We next derive results that ease comparison of several designs.
+with parameters as defined in Eq. (\ref{gp_upd}) and
+(\ref{gp_hat}). Here, the probabilities would not only depend on the
+data but also on the data design defined via matrix $\bG$ in
+Eq. (\ref{likelihood}). We next derive results that ease comparison of
+several designs.
 
@@
 salinity.
-\begin{figure}[h!] \centering
+
+\begin{figure}[!b] \centering
   \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
@@
 selection situation is then depicted in Figure \ref{fig:PathSelOpt},
-\begin{figure}[h!]
+\begin{figure}[!t]
 \centering
@@
 \includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
-\caption{The commercially available Light Autonomous Underwater Vehicle (LAUV) platform (Harald) for upper water-
-column exploration used in the experiment.}
+\caption{The commercially available Light Autonomous Underwater
+  Vehicle (LAUV) platform for upper water- column exploration used in
+  the experiment. Image Courtesy LSTS, Univ. of Porto, Portugal.}
 \label{fig:lauv}

### 2019-09-23 main_new.tex
Updates from Overleaf

--- 2019-09-23 main_new.tex
@@
   Improving and optimizing oceanographic sampling is a crucial task for
-  modern marine management and science faced with limited resources to
-  understand highly dynamic processes in the water-column. The
+  modern marine management and science. Faced with limited resources to
+  understand highly dynamic processes in the water-column, the
   combination of statistics and autonomous
@@
 Traditional data collection at sea has typically been based on static
-buoys, lagrangian floats or ship-based methods, with significant
+buoys, lagrangian float,s or ship-based methods, with significant
 logistical limitations that directly impact coverage and sampling
@@
 large-scale coverage, but have limited resolution, are limited to
-sensing the surface of the ocean, and is impacted by cloud cover. The
+sensing the surface of the ocean, and are impacted by cloud cover. The
 advent of robust mobile robotic platforms \cite{Bellingham07} has
 resulted in significant contributions to environmental monitoring and
-sampling. In particular autonomous underwater vehicles (AUVs), have made
+sampling. In particular, autonomous underwater vehicles (AUVs), have made
 strides in advancing the state of sampling and consequently have made
@@
 features using an underwater glider. Common to these approaches is the
-use of information-based methods to infer environmental sampling. A few
-important characteristics should be mentioned, namely:
+use of information-based methods to infer environmental sampling and overcome common challenges, namely:
 
 \begin{itemize}
-\item Decisions on where to sample happen sequentially, i.e.  how
+\item Static sampling paths fail to capture necessary features in dynamic ocean conditions. Incontrast, these approaches make decisions about where to sample  sequentially, i.e.  
   recently acquired information can be assimilated and used to improve
@@
 \item As a priori information about the state of the ocean is very
-  uncertain, there lies a substantial efficiency gain in reacting to
-  information obtained from measurements taken \emph{in-situ}, this is
+  uncertain, there is a substantial efficiency gain in reacting to
+  information obtained from measurements taken \emph{in-situ}; this is
   usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.

### 2019-09-24 main_new.tex
Updates from Overleaf

@@
 Traditional data collection at sea has typically been based on static
-buoys, lagrangian float,s or ship-based methods, with significant
+buoys, lagrangian floats or ship-based methods, with significant
 logistical limitations that directly impact coverage and sampling
@@
 strides in advancing the state of sampling and consequently have made
-robotics an integral part of ocean observation in large part by our own
-work % driven by
+robotics an integral part of ocean observation; our own work
+has contributed to this effort.
+
 % using statistical methods in design and data assimilation
@@
 \begin{itemize}
-\item Static sampling paths fail to capture necessary features in dynamic ocean conditions. Incontrast, these approaches make decisions about where to sample  sequentially, i.e.  
+\item Decisions on where to sample happen sequentially, i.e.  how
   recently acquired information can be assimilated and used to improve
@@
   locations, trajectories, and candidate designs) makes the problem
-  combinatorially large, creating a trade-off between optimization
-  (finding the optimal design) and computability (arriving at a
+  combinatorially large, creating a trade-off between optimization (finding the most resource-efficient design to collect necessary data), and computability (arriving at a
   solution in reasonable time), which has to be considered in
@@
 natural symbiosis which enables information-based sensing. Central to
-this is the ability to model spatially correlated variables and
+this is the ability to model spatially-correlated variables and
 provide formal measures of uncertainty. By far the most practical
 models use Gaussian processes (GPs), which will also be used in our
-formulation, in large part because it allows efficient implementation
+formulation, in large part because they allow efficient implementation
 and evaluation, in turn enabling prioritization of sampling efforts to

### 2019-09-24 main_new.tex
Front end updates

@@
-% Template for the submission to:
-%   The Annals of Probability           [aop]
-%   The Annals of Applied Probability   [aap]
-%   The Annals of Statistics            [aos] 
-%   The Annals of Applied Statistics    [aoas]
-%   Stochastic Systems                  [ssy]
-%
-%Author: In this template, the places where you need to add information
-%        (or delete line) are indicated by {???}.  Mostly the information
-%        required is obvious, but some explanations are given in lines starting
-%Author:
-%All other lines should be ignored.  After editing, there should be
-%no instances of ??? after this line.
-
-% use option [preprint] to remove info line at bottom
-% journal options: aop,aap,aos,aoas,ssy
-% natbib option: authoryear
-
 \documentclass[aoas]{imsart}
@@
 \floatsetup[table]{capposition=top}
-% \usepackage{subfig}
 \usepackage{subfigure}
 \usepackage[utf8]{inputenc}
-
-%\usepackage[tableposition = top]{caption} % captions
 \usepackage{booktabs}                     % horizontal lines in tables
@@
 \RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
-
-% provide arXiv number if available:
-%\arxiv{arXiv:0000.0000}
 
@@
 
-% indicate corresponding author with \corref{}
-% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
-% \thankstext{t1}{Thanks to somebody} 
-% \address{line 1\\ line 2\\ printead{e1}}
-% \affiliation{Some University}
 \begin{aug}
-\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{a}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
-\author{\fnms{Jo} \snm{Eidsvik}\thanksref{b}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
-\author{\fnms{David} \snm{Ginsbourger}\thanksref{c}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
+\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{a,b}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
+\author{\fnms{Jo} \snm{Eidsvik}\thanksref{c}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
+\author{\fnms{David} \snm{Ginsbourger}\thanksref{d}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
 \and
-\author{\fnms{Kanna} \snm{Rajan}\thanksref{d}. \ead[label=e4]{kanna.rajan@ntnu.no}}
+\author{\fnms{Kanna} \snm{Rajan}\thanksref{e,f}. \ead[label=e4]{kanna.rajan@ntnu.no}}
 
 \address{Department of Marine Technology\\ Trondheim, Norway\printead{e1}}
-\affiliation[a]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU) \&\\ Centre for Autonomous Marine Operations and Systems (AMOS), Trondheim, Norway.}
+\affiliation[a]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
+\affiliation[b]{Centre for Autonomous Marine Operations and Systems (AMOS).}
 
 \address{Department of Mathematical Sciences \printead{e2}}
-\affiliation[b]{Department of Mathematical Sciences, NTNU.}
+\affiliation[c]{Department of Mathematical Sciences, NTNU.}
 
 \address{Idiap \printead{e3}}
-\affiliation[c]{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland.}
+\affiliation[d]{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland.}
 
 \address{Department of Engineering Cybernetics,\printead{e4}}
-\affiliation[d]{Department of Engineering Cybernetics, NTNU \&\\ Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
+\affiliation[e]{Department of Engineering Cybernetics, NTNU}
+\affiliation[f]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
 

### 2019-09-24 main_new.tex
updates

@@
 \and
-\author{\fnms{Kanna} \snm{Rajan}\thanksref{e,f}. \ead[label=e4]{kanna.rajan@ntnu.no}}
+\author{\fnms{Kanna} \snm{Rajan}\thanksref{b,e,f}. \ead[label=e4]{kanna.rajan@ntnu.no}}
 
@@
 \affiliation[e]{Department of Engineering Cybernetics, NTNU}
-\affiliation[f]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
+\affiliation[f]{Underwater Systems and Technology Laboratory (LSTS),
+  Faculty of Engineering, University of Porto, Portugal.}
 
@@
 \begin{abstract}
-  Improving and optimizing oceanographic sampling is a crucial task for
-  modern marine management and science. Faced with limited resources to
-  understand highly dynamic processes in the water-column, the
-  combination of statistics and autonomous
-  robotics provides opportunities for studying aspects related to
-  experimental design for effective characterization of oceanographic
-  phenomena. In this work we develop methods for efficient spatial
-  sampling applied to the mapping of coastal processes by
-  providing informative descriptions of spatial characteristics of ocean
+  Improving and optimizing oceanographic sampling is a crucial task
+  for modern marine management and science. Faced with limited
+  resources to understand highly dynamic processes in the
+  water-column, the combination of statistics and autonomous robotics
+  provides opportunities for studying aspects related to experimental
+  design for effective characterization of oceanographic phenomena. In
+  this work, we develop methods for efficient spatial sampling applied
+  to the mapping of coastal processes by providing informative
+  descriptions of spatial characteristics of ocean
   phenomena. Specifically, we define a design criterion based on
-  reducing uncertainty in the joint excursions of random
-  fields, and derive analytical expressions for the expected Bernoulli
-  variance reduction in the excursion set of spatial temperature and
-  salinity variables. We show how these can be applicable in static
-  designs and sequential designs, building on myopic and look-ahead
-  sampling strategies. We use simulations to study the properties and
+  reducing uncertainty in the joint excursions of random fields, and
+  derive analytical expressions for the expected Bernoulli variance
+  reduction in the excursion set of spatial temperature and salinity
+  variables. We show how these can be applicable in static designs and
+  sequential designs, building on myopic and look-ahead sampling
+  strategies. We use simulations to study the properties and
   differences between state-of-the-art approaches. Results from field
-  deployments with an autonomous underwater vehicle are also presented,
-  for a case study mapping the boundary of a river plume. The results
-  demonstrate the ability of the method to effectively inform and
-  execute data-driven environmental sampling.
+  deployments with an autonomous underwater vehicle are also
+  presented, for a case study mapping the boundary of a river
+  plume. The results demonstrate the ability of the method to
+  effectively inform and execute data-driven environmental sampling.
 
@@
 
-An important synergy underlining this work is the potential of robotic sampling to adapt survey plans based on data and statistical inference,
-in accordance with the sense-plan-act loop
-(Fig. \ref{fig:sense-plan-act}). We will therefore present a few adaptive strategies
-for robotic sampling that builds on the notion of multivariate ES. Simulations, analysis, and  results from full-scale trials are presented.
+An important synergy underlining this work is the connection to
+robotic sampling to adapt survey plans based on data and statistical
+inference, in accordance with the sense-plan-act loop
+(Fig. \ref{fig:sense-plan-act}). To do so we build on the notion of
+multivariate ES and simulations, analysis, and results from full-scale
+robotic trials are presented.
 
@@
 selection situation is then depicted in Figure \ref{fig:PathSelOpt},
-\begin{figure}[h!]
+\begin{figure}[!b]
 \centering
@@
 \caption{The equilateral waypoint graph used to discretize the
-  trajectory choices over the $31\times31$ grid used to discretize the GP.}
+  trajectory choices over the $31\times31$ grid used to discretize the
+  Gaussian Process.}
 \label{fig:wp_graph}
@@
 
-\begin{figure}[!h] 
+\begin{figure}[!b] 
  \centering 
@@
 
-The sampling locations was distributed over an equilateral grid, as shown in Fig.
-\ref{fig:map} in a grey colored lattice.
-The robotic platform consisted of a Light AUV \citep{sousa2012lauv} (Figure \ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49 CTD (conductivity, temperature, and depth) sensor providing temperature and salinity measurements.
-The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
-The sampling agent was built on top of the autonomous agent framework \texttt{T-REX} \citep{py10,Rajan12,Rajan12b}, running an instance of the \texttt{myopic} strategy from Section \ref{sec:sampling_designs}, to control the AUV and decide between sampling locations.
+The sampling locations was distributed over an equilateral grid, as
+shown in Fig. \ref{fig:map} in a grey colored lattice. The robotic
+platform consisted of a Light AUV \citep{sousa2012lauv} (Figure
+\ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49 CTD
+(conductivity, temperature, and depth) sensor providing temperature
+and salinity measurements. The accuracy of the CTD instrument is
+$\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$
+(temperature). The sampling agent was built on top of the autonomous
+agent framework \texttt{T-REX} \citep{py10,Rajan12,Rajan12b}, running
+an instance of the \texttt{myopic} strategy from Section
+\ref{sec:sampling_designs}, to control the AUV and decide between
+sampling locations.
 
@@
 \includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
-\caption{The commercially available Light Autonomous Underwater Vehicle (LAUV) platform (Harald) for upper water-
-column exploration used in the experiment.}
+\caption{The commercially available Light Autonomous Underwater
+  Vehicle (LAUV) platform for upper water- column exploration used in
+  the experiment.}
 \label{fig:lauv}
@@
 \section*{Acknowledgements}
-TOF and KR acknowledge support from the Centre for Autonomous Marine Operations and Systems (AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of Excellence, project number 223254, and the Applied Underwater Robotics Labortatory (AURLab).
-JE acknowledge support from Norwegian research council project 294404.
-DG acknowledges support from the Swiss National Science Foundation project number 178858.
+TOF and KR acknowledge support from the Centre for Autonomous Marine
+Operations and Systems
+(AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of
+Excellence, project number 223254, and the Applied Underwater Robotics
+Labortatory (AURLab). JE acknowledge support from Norwegian research
+council project 294404. DG acknowledges support from the Swiss
+National Science Foundation project number 178858.
 

### 2019-09-24 main_new.tex
updates

@@
 
-The sampling strategy was designed around the concept of visiting waypoints sequentially.
-Arriving at a desired waypoint, with new measurements and an updated model, the AUV triggers the myopic strategy to evaluate all possible design criteria, e.g Eq.
-\eqref{critSEQ}.
-The waypoint-and-path combination which is expected to reduce the IBV the most is selected; upon arrival this procedure is then repeated.
-At each stage, it takes the AUV about 30 sec to evaluate the expected IBV for all the possible waypoint-and-path alternatives.
-
-The AUV was set to start in the south center part of the waypoint graph, with the previously outlined GP prior model of the environment.
-Each AUV survey was set to take approximately 40 min, visiting 15 waypoints on the grid, with the AUV running near the surface to better capture the plume.
-On its path from one waypoint to the next, the AUV gathered data regularly, and the GP model assimilated temperature and salinity data with an update frequency of 30 sec, about three updates per stage.
+The sampling strategy was designed around the concept of visiting
+waypoints sequentially. Arriving at a desired waypoint, with new
+measurements and an updated model, the AUV triggers the myopic
+strategy to evaluate all possible design criteria, e.g
+Eq. \eqref{critSEQ}. The waypoint-and-path combination which is
+expected to reduce the IBV the most is selected; upon arrival this
+procedure is then repeated. At each stage, it takes the AUV about 30
+sec to evaluate the expected IBV for all the possible
+waypoint-and-path alternatives.
+
+The AUV was set to start in the south center part of the waypoint
+graph, with the previously outlined GP prior model of the
+environment. Each AUV survey was set to take approximately 40 min,
+visiting 15 waypoints on the grid, with the AUV running near the
+surface to better capture the plume. On its path from one waypoint to
+the next, the AUV gathered data regularly, and the GP model
+assimilated temperature and salinity data with an update frequency of
+30 sec, about three updates per stage.
 
@@
 
-Two missions, Survey 1 and 2, were run successively from 11:00 AM to 01:00 PM, with a short break in between.
-The resulting path of the selected waypoints are shown in the map in Fig.
-\ref{fig:map}, both within the expected frontal region (shaded pink region).
-The recorded temperatures are shown as colored trails in Fig.
-\ref{fig:res_both}, clearly showing the temperature difference between fjord and river waters.
-The salinity data are then shown separately overlaid with the estimated EP for each survey in Fig.
-\ref{fig:res1} and \ref{fig:res2}.
-
-Both Survey 1 and 2 successfully estimate and navigate the separation zone crossing the frontal boundary multiple times.
-As conditions changed slightly between the two deployments the resulting path (after waypoint 5) deviates.
-Survey 1 continues northwards tracking the north-eastern part of the front, while Survey 2 turns west mapping the south-western part.
-
-The final prediction of the front location given in Fig.
-\ref{fig:res3} and \ref{fig:res4}, correspond with one another yielding a picture of the front gradually bending off towards north east.
-The amount of exploration done by Survey 1 is greater than Survey 2.
-In Survey 1 the AUV gets overall more detail by going north, while Survey 2 coming close to the survey area borders in the south-western corner get poorer understanding of the northern parts.
-A longer planning horizon would likely identify and discourage such choices.
+Two missions, Survey 1 and 2, were run successively from 11:00 AM to
+01:00 PM \kc{date? month/year}, with a short break in between. The
+resulting path of the selected waypoints are shown in the map in
+Fig. \ref{fig:map}, both within the expected frontal region (shaded
+pink region). The recorded temperatures are shown as colored trails in
+Fig. \ref{fig:res_both}, clearly showing the temperature difference
+between fjord and river waters. The salinity data are then shown
+separately overlaid with the estimated EP for each survey in
+Fig. \ref{fig:res1} and \ref{fig:res2}.
+
+Both Survey 1 and 2 successfully estimate and navigate the separation
+zone crossing the frontal boundary multiple times. As conditions
+changed slightly between the two deployments the resulting path (after
+waypoint 5) deviates. Survey 1 continues northwards tracking the
+north-eastern part of the front, while Survey 2 turns west mapping the
+south-western part.
+
+The final prediction of the front location given in
+Fig. \ref{fig:res3} and \ref{fig:res4}, correspond with one another
+yielding a picture of the front gradually bending off towards north
+east. The amount of exploration done by Survey 1 is greater than
+Survey 2. In Survey 1 the AUV gets overall more detail by going north,
+while Survey 2 coming close to the survey area borders in the
+south-western corner get poorer understanding of the northern parts. A
+longer planning horizon would likely identify and discourage such
+choices.
 
@@
 
-This paper builds on multidisciplinary work combining statistical modeling and methods with robotic surveying technologies for oceanographic sensing applications.
-We show how the observation practices can gain from development of statistical techniques to achieve more effective spatial monitoring.
-We further demonstrate the opportunities available for real-time multivariable spatial data gathering and analysis onboard autonomous sensing platforms, upon which statisticians can exploit towards creating general-purpose toolkits for similar applications (enabeling theory).
-
-In particular, we derive and show results for characterizing phenomena connected to water masses properties.
-The characterization is based on joint excursion sets and we derive new results for the expected integrated Bernoulli variance reduction achieved by the spatial sampling design.
-This result is first derived for the situation with a static design, and then extended to the adaptive situation, which provides new insight in our application with the autonomous vehicle.
-
-The study did not consider any temporal effects, which would be relevant on a larger time scale.
-We consider the extension to spatio-temporal modeling as future work, and we envision that advection-diffusion equations could be useful in this kind of modeling \citep{sigrist2015stochastic}.
-For more complex oceanographic phenomena, the methods will need to be extended to non-Gaussian phenomena, possibly feature-based mixtures of GPs which could still be run onboard, and augmented by dynamical models.
-
-The spatial statistical design criterion building on excursion sets is relevant in our setting with the different water mass properties and could also be useful in other oceanographic settings such as algae-blooms or frontal systems.
-Other criteria more directly related to decision making would also be relevant, especially with much spatio-temporal effects.
-For instance, managers must make difficult decisions related to fish farming or other marine resources operations, or environmental projects.
-Value of information analysis \citep{Eidsvik:15} would then be used in a similar vein as the integrated Bernoulli variance in the current paper, now studying which, when and where information is likely to materialize in improved decisions.
-
-While more effort could be spent on approximating the look-ahead steps, it is perhaps more interesting to explore the additional flexibility that can be gained by having two or more AUVs \citep{ferreira2019advancing} exploring a spatial or spatio-temporal domain together.
-It might then be useful to sample different parts of the space, and maybe move in parallel to best capture the excursion set.
+This work builds on multidisciplinary work combining statistical
+modeling and methods with robotic surveying technologies for
+oceanographic sensing applications. We show how the observation
+practices can gain from development of statistical techniques to
+achieve more effective spatial monitoring. We further demonstrate the
+opportunities available for real-time multivariable spatial data
+gathering and analysis onboard autonomous sensing platforms, upon
+which statisticians can exploit towards creating general-purpose
+toolkits for similar applications (enabeling theory).
+
+In particular, we derive and show results for characterizing phenomena
+connected to water masses properties. The characterization is based on
+joint excursion sets and we derive new results for the expected
+integrated Bernoulli variance reduction achieved by the spatial
+sampling design. This result is first derived for the situation with a
+static design, and then extended to the adaptive situation, which
+provides new insight in our application with the autonomous vehicle.
+
+The study did not consider any temporal effects, which would be
+relevant on a larger time scale. We consider the extension to
+spatio-temporal modeling as future work, and we envision that
+advection-diffusion equations could be useful in this kind of modeling
+\citep{sigrist2015stochastic}. For more complex oceanographic
+phenomena, the methods will need to be extended to non-Gaussian
+phenomena, possibly feature-based mixtures of GPs which could still be
+run onboard, and augmented by dynamical models.
+
+The spatial statistical design criterion building on excursion sets is
+relevant in our setting with the different water mass properties and
+could also be useful in other oceanographic settings such as
+algae-blooms or frontal systems. Other criteria more directly related
+to decision making would also be relevant, especially with much
+spatio-temporal effects. For instance, managers must make difficult
+decisions related to fish farming or other marine resources
+operations, or environmental projects. Value of information analysis
+\citep{Eidsvik:15} would then be used in a similar vein as the
+integrated Bernoulli variance in the current paper, now studying
+which, when and where information is likely to materialize in improved
+decisions.
+
+While more effort could be spent on approximating the look-ahead
+steps, it is perhaps more interesting to explore the additional
+flexibility that can be gained by having two or more AUVs
+\citep{ferreira2019advancing} exploring a spatial or spatio-temporal
+domain together. It might then be useful to sample different parts of
+the space, and maybe move in parallel to best capture the excursion
+set.
 

### 2019-09-25 main_new.tex
Updates from Overleaf

--- 2019-09-24 main_new.tex
@@
   differences between state-of-the-art approaches. Results from field
-  deployments with an autonomous underwater vehicle are also presented,
-  for a case study mapping the boundary of a river plume. The results
+  deployments with an autonomous underwater vehicle are also presented as part of a case study mapping the boundary of a river plume. The results
   demonstrate the ability of the method to effectively inform and
@@
 
-Monitoring the worlds oceans has gained increased importance in the
+Monitoring the world's oceans has gained increased importance in 
 light of the changing climate and increasing anthropogenic impact. A
@@
 Traditional data collection at sea has typically been based on static
-buoys, lagrangian floats or ship-based methods, with significant
+buoys, lagrangian floats, or ship-based methods, with significant
 logistical limitations that directly impact coverage and sampling
 resolution. Modern methods using satellite remote-sensing provide
-large-scale coverage, but have limited resolution, are limited to
+large-scale coverage but have limited resolution, are limited to
 sensing the surface of the ocean, and are impacted by cloud cover. The
@@
 strides in advancing the state of sampling and consequently have made
-robotics an integral part of ocean observation; our own work
-has contributed to this effort.
-
+robotics an integral part of ocean observation; our previous work
+has contributed to this effort
 % using statistical methods in design and data assimilation
 \citep{das11b,Graham2013,Das2015,das15,fossuminformation,fossum18b}.
-Others include \cite{wikle2013modern} focusing on hierarchical
-statistical models, \cite{sahu2008space} which studies spatio-temporal
-models for sea surface temperature and salinity data, while
-\cite{mellucci2018oceanic} looks at the statistical prediction of
+Others working in this space include \cite{wikle2013modern}, focusing on hierarchical
+statistical models; \cite{sahu2008space}, studying spatio-temporal
+models for sea surface temperature and salinity data; and
+\cite{mellucci2018oceanic} looking at the statistical prediction of
 features using an underwater glider. Common to these approaches is the
@@
 \begin{itemize}
-\item Decisions on where to sample happen sequentially, i.e.  how
-  recently acquired information can be assimilated and used to improve
-  statistical models and reduce uncertainty associated with where to make measurements.
+\item Fully fledged numerical ocean models cannot be run online on the sensing platforms due to the computational load they require to provide accurate results, hence statistical proxy models of the environment must be used.
+\item Decisions on where to sample happen sequentially, i.e.  
+  recently acquired information can be assimilated into and used to improve statistical models and reduce uncertainty associated with where to make measurements.
 \item As a priori information about the state of the ocean is very
@@
   combinatorially large, creating a trade-off between optimization (finding the most resource-efficient design to collect necessary data), and computability (arriving at a
-  solution in reasonable time), which has to be considered in
-  practice.
+  solution in reasonable time); to engage in successful sampling, this trade-off has to be considered in sample design and practice.
 \end{itemize}{}
 
-The combination of statistical tools and robotic platforms is a
-natural symbiosis which enables information-based sensing. Central to
-this is the ability to model spatially-correlated variables and
-provide formal measures of uncertainty. By far the most practical
-models use Gaussian processes (GPs), which will also be used in our
-formulation, in large part because they allow efficient implementation
-and evaluation, in turn enabling prioritization of sampling efforts to
-be formalized into decision making systems on-board a robotic
-platform.
+The combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based sensing.
+Central to this is the ability to model spatially-correlated variables and provide formal measures of uncertainty.
+In our formulation we used Gaussian Processes (GPs), in large part because they allow for efficient implementation and evaluation, in turn enabling prioritization of sampling efforts which can be be formalized into decision-making systems on board a robotic platform.
 
@@
 
-In this work, we build on joint excursion sets (ES) and excursion probabilities (EP) for spatial design of experiments. The goal is to
-design sampling strategies for reducing the spatially integrated
-Bernoulli variance (IBV) in the ES indicators.  The main contributions
-involve: i) extending ES-based uncertainty criteria to multivariate
-GPs, allowing the evaluation to include multiple environmental
-parameters such as temperature and salinity, ii) developing methods
-for informative environmental sensing and autonomous data collection,
-and iii) demonstrating the approach in a real world scenario of
+In this work, we build on joint excursion sets (ES) and excursion probabilities (EP) for the spatial design of experiments. The goal is to
+design sampling strategies for reducing the spatially-integrated
+Bernoulli variance (IBV) in the ES indicators. Our main contributions
+include: i) extending ES-based uncertainty criteria to multivariate
+GPs, thereby allowing the evaluation to include multiple environmental
+parameters such as temperature and salinity; ii) developing methods
+for information-based environmental sensing and autonomous data collection;
+and iii) demonstrating the approach in the real-world scenario of
 sampling a river plume using an AUV.
 
-The remainder of the paper is organized as follows: Section
-\ref{sec:bg} provides some background on ocean sampling. Section
-\ref{sec:ESEP} defines ESs and EPs, and the design criteria of
-IBV. Section \ref{sec:GP_EP} sets up the required statistical modeling
+The remainder of this paper is organized as follows: Section
+\ref{sec:bg} provides select background on ocean sampling; Section
+\ref{sec:ESEP} defines ESs, EPs, and the design criteria of
+IBV; Section \ref{sec:GP_EP} sets up the required statistical modeling
 assumptions of the bi-variate GPs for temperature and
-salinity. Section \ref{sec:sur} builds on these when deriving the
-analytical derivation of the design criteria. Section
-\ref{sec:heuristics} extends results to present heuristics for
-adaptive sampling. Section \ref{sec:simulations} shows results and
+salinity; Section \ref{sec:sur} builds on these assumptions when deriving the analytical derivation of the design criteria; Section
+\ref{sec:heuristics} extends these results and present heuristics for
+adaptive sampling; Section \ref{sec:simulations} shows our results and
 discusses properties of the methods in simulation studies. Section
-\ref{sec:case_study} demonstrates the methodology in field work
-characterizing a river plume.  Section \ref{sec:concl_disc} contains
-summary and a discussion of future work.
+\ref{sec:case_study} demonstrates the methodology used in field work
+characterizing a river plume; and Section \ref{sec:concl_disc} contains a summary and a discussion of future work.
 
@@
 Central to understanding the changes taking place in the upper
-water-column is the understanding of biophysical interaction driven by
+water-column is an understanding of biophysical interaction driven by
 an agglomeration of physical forcing (wind, topography, bathymetry,
-tidal influences and local geography such as fresh water inflows)
-mixed with incipient micro-biology driven by planktonic and coastal
+tidal influences and local geography, such as fresh water inflows)
+and incipient micro-biology driven by planktonic and coastal
 anthropogenic input, such as pollution and agricultural runoff. These
-often result in a range of ecosystem related phenomena such as blooms,
-anoxic zones and plumes, all with significant impact both directly and
-indirectly on society. Walter Munk referred to the
+often result in a range of ecosystem-related phenomena such as blooms,
+anoxic zones and plumes, all of which have significant direct and indirect effects  on society. Walter Munk referred to the
 20\textsuperscript{th} century as the \emph{century of undersampling}
-\citep{munk2002}, pointing out that lack of sufficient sampling
-resolution in time and space, as a critical issue for characterizing
-ocean phenomena. Even today policy makers struggle with limited
-observability of severe environmental problems such as the recent (May
-2018) outbreak of harmful algal blooms in Northern
+\citep{munk2002}, pointing out a lack of sufficient sampling
+resolution in time and space as a critical issue for characterizing
+ocean phenomena. Even today, policymakers struggle with limited
+observability of severe environmental problems such as the recent outbreak of harmful algal blooms in Northern
 Norway\footnote{Report from the Institute of Marine
-  Research: https://bit.ly/2KZGJfI}, with losses exceeding 12 million
-tons of salmon valued at \$ 80 million (Norwegian Directorate of
-Fisheries). This is a case in point in how adequate sampling efforts
+  Research: https://bit.ly/2KZGJfI} (May 2018), with losses exceeding 12 million
+tons of salmon valued at \$80 million (Norwegian Directorate of
+Fisheries). This is a case in point of how adequate sampling efforts
 could have provided a more accurate assessment and prior
-warning. These challenges related to undersampling have been driving
-ocean sensing practices towards more autonomous, networked and
+warning of changing environmental conditions. These challenges related to undersampling have been driving
+ocean sensing practices towards more autonomous, networked, and
 distributed use of robotic assets in order to fill the gaps in our
-understanding (See Fig. \ref{fig:envir}), where the surveying is
-guided by a dynamical ocean model and the measurements made onboard
-the robot are often used to update the model.
+understanding (See Fig. \ref{fig:envir}), where observation is
+guided by assimilation of measurements and statistical models.
 
@@
   \subfigure[Illustration of an ocean sensing network, where robotic
-  assets augment ship based measurements to increase coverage and
+  assets augment ship-based measurements to increase coverage and
   resolution of phenomena of scientific interest.]{\includegraphics[width =
@@
 Our focus in this work is towards spatial characterization of a
-frontal system, and more specifically fronts generated by river
+frontal system, more specifically, fronts generated by river
 plumes. Fig. \ref{fig:nidelven} shows the survey area (the outlet of
@@
 land to the east. Depending on the river discharge, tidal effects,
-wind and temperature differences, this boundary often gets
+wind, and temperature differences, this boundary often gets
 distorted. Prior knowledge about the location and evolution of these
@@
 River plumes belong to a class of ocean processes that are local and
-at smaller \emph{sub-mesoscale} (from $10 m^2$ -- $5 km^2$) where
+at smaller \emph{sub-mesoscale} (from $5 m^2$ -- $10 km^2$) where
 spatial lateral variability tends to dominate. Remote sensing and
@@
 important and can shift substantially, while in the case of river
-plumes one can often limit scope to cover only lateral spatial
-elements. Moreover, for this case, time-effects can be regarded as
+plumes one can often limit the scope to cover only lateral spatial
+elements. Moreover, in this case, time-effects can be regarded as
 static when acquiring AUV data over a short time window and limited
@@
 water column. Typically, they operate at $\sim 1-3$ m/s, can reach
-depths from 100--6000 m with an in-water operation time capacity depending on
+depths from 100--6000 m with an in-water operation time capacity dependent on
 survey speed, payload sensors and mission design. By deploying AUVs
-together with other resources one can augment information obtained
-from either ocean models, remote-sensing data, or fixed-location
+together with other resources, one can augment information obtained
+from ocean models, remote-sensing data, or fixed-location
 sensors on buoys. Traditional operations and surveys with AUVs are
@@
 based on closing the sensing and actuation loop
-(Fig. \ref{fig:sense-plan-act}). Myopic sampling, i.e. step-wise
+(Fig. \ref{fig:sense-plan-act}). Myopic sampling—i.e. step-wise
 selection of the path (on the graph) which reduces the information
-criterion the most, has been used for adaptive surveys
-\citep{singh2009efficient,Binney2013} which focus largely on reducing
-predictive variance or mutual information (entropy). Variance and
+criterion the most—has been used for adaptive surveys
+\citep{singh2009efficient,Binney2013} that focus largely on reducing
+predictive variance or mutual information (e.g. entropy). Variance and
 entropy reduction are independent of the actual data realizations
-under the assumption of GP models, and the use of data-driven adaptive
+under the assumptions of GP models; the use of data-driven adaptive
 criteria was introduced to include more targeted sampling of regions
-of scientific interest by \cite{Low2009} and \cite{fossuminformation}.
-In this paper, with the focus on mapping the river plume, we reward
-the designs that improve the classification of water masses, as a
+of scientific interest in \cite{Low2009} and \cite{fossuminformation}.
+In this paper, with a focus on mapping the river plume, we reward
+the designs that improve the classification of water masses as a
 means to characterize the frontal zone.
@@
 
-\section{Excursion sets and Integrated Bernoulli variance}
+\section{Excursion Sets and Integrated Bernoulli Variance}
 \label{sec:ESEP}
 
-Our objective is to leverage statistical tools to characterize a river plume, focusing on spatial separation of cold freshwater (from the river) and warmer saline waters of the fjord.
+Our objective is to leverage statistical tools to characterize a river plume, focusing on spatial separation of cold freshwater from the river and warmer saline waters of the fjord.
 Excursion sets (ES) and excursion probabilities (EP) are useful starting points to measure the ability to classify water masses.
@@
 
-We consider a lateral domain, near the sea surface, with locations
+We consider a lateral domain near the sea surface, with locations
 $\bx \in \mathcal{M} \subset \mathbf{R^2}$. The two variables of
-interest, temperature and salinity, are defined as random processes;
+interest, temperature and salinity, are defined as random processes:
 $\xi_T(\bx)$ denotes the temperature (in $^o C$) and $\xi_S(\bx)$
@@
 
-We let $t_T$ be a threshold for temperature and $t_S$ a threshold for salinity, the joint ES is then defined by:
+We let $t_T$ be a threshold for temperature and $t_S$ a threshold for salinity. The joint ES is then defined by:
 \begin{equation}\label{ES}
@@
 
-The ES holds random indicator variables, while the associated EPs are: 
+The ES holds random indicator variables while the associated EPs are: 
 \begin{equation}\label{eq:prob}
@@
 If the EPs in Eq. (\ref{eq:prob}) are
-close to $1$ or $0$ at a given location, it is easy to classify water mass at this point as either river or ocean; it is however
+close to $1$ or $0$ at a given location, it is easy to classify water mass at this point as either river or ocean; it is, however,
 challenging if the probability is close to $0.5$. 
@@
 If prior knowledge provides a strong indication of cold freshwater
-(for instance) in a part of the domain, it is easy to classify, and
+ in a part of the domain, it is easy to classify and
 there is likely little value in exploring those locations. Data can be
 gathered at selected locations of the domain $\mathcal{M}$, to improve
-classification skill. We denote temperature measurements by
-$y_{T}(\bx_i)$ and salinity measurements by $y_{S}(\bx_i)$, at a
+classification performance. We denote temperature measurements by
+$y_{T}(\bx_i)$ and salinity measurements by $y_{S}(\bx_i)$ at a
 sampling location $\bx_i \in \mathcal{M}$. Given a set of
-measurements, denoted $\by$, the conditional EPs are:
+measurements denoted $\by$, the conditional EPs are:
 \begin{equation}\label{eq:post_ep}
@@
 Our goal is to construct adaptive sampling strategies based on continuous evaluation of the EPs.
-As the EPs provide a description of the boundary uncertainty, the strategies can effectively use this description to prioritize sampling efforts to locations that are ambiguous; making exploration more effective.
+As the EPs provide a description of the boundary uncertainty, the strategies can effectively use this description to prioritize sampling efforts at locations that are ambiguous, making exploration more effective.
 As the AUV gathers data throughout its mission, it is natural to evaluate the expected reduction in the IBV, conditioned on the survey data.
 In this way, we can
-take the expectation with respect to the planned measurements $\by$;
+take the expectation with respect to the planned measurements $\by$
 \begin{equation}\label{sur}
@@
 where the expectation is over the planned data, and the probability is defined in Eq. (\ref{eq:post_ep}).
-Considering a set $\mathcal{J}$ of possible designs for an AUV survey. A criterion for the selection is:
+Considering a set $\mathcal{J}$ of possible designs for an AUV survey, a criterion for the selection is
 \begin{equation}\label{crit}
@@
 \cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
-Our focus is on ES as defined in Eq. (\ref{ES}), and the uncertainty
-reduction achieved by sampling as in Eq. (\ref{sur}). In this sense
+Our focus is on ES as defined in Eq. (\ref{ES}) and the uncertainty
+reduction achieved by sampling as in Eq. (\ref{sur}). In this sense,
 our work is similar to \cite{bect2012}, \cite{chevalier2014fast} and
@@
 
-\section{Gaussian processes and excursion probabilities}
+\section{Gaussian Processes and Excursion Probabilities}
 \label{sec:GP_EP}
@@
 
-\subsection{Bivariate Gaussian processes}
+\subsection{Bivariate Gaussian Processes}
 

### 2019-09-25 main_new.tex
Manual merge after unsuccsessfull auto-merge
updates

@@
 \begin{abstract}
-  Improving and optimizing oceanographic sampling is a crucial task for
-  modern marine management and science. Faced with limited resources to
-  understand highly dynamic processes in the water-column, the
-  combination of statistics and autonomous
-  robotics provides opportunities for studying aspects related to
-  experimental design for effective characterization of oceanographic
-  phenomena. In this work we develop methods for efficient spatial
-  sampling applied to the mapping of coastal processes by
-  providing informative descriptions of spatial characteristics of ocean
+  Improving and optimizing oceanographic sampling is a crucial task
+  for modern marine management and science. Faced with limited
+  resources to understand highly dynamic processes in the
+  water-column, the combination of statistics and autonomous robotics
+  provides opportunities for studying aspects related to experimental
+  design for effective characterization of oceanographic phenomena. In
+  this work we develop methods for efficient spatial sampling applied
+  to the mapping of coastal processes by providing informative
+  descriptions of spatial characteristics of ocean
   phenomena. Specifically, we define a design criterion based on
-  reducing uncertainty in the joint excursions of random
-  fields, and derive analytical expressions for the expected Bernoulli
-  variance reduction in the excursion set of spatial temperature and
-  salinity variables. We show how these can be applicable in static
-  designs and sequential designs, building on myopic and look-ahead
-  sampling strategies. We use simulations to study the properties and
-  differences between state-of-the-art approaches. Results from field
-  deployments with an autonomous underwater vehicle are also presented as part of a case study mapping the boundary of a river plume. The results
+  reducing uncertainty in the joint excursions of random fields, and
+  derive analytical expressions for the expected Bernoulli variance
+  reduction in the excursion set of spatial temperature and salinity
+  variables. We show how these can be applicable in static designs and
+  sequential designs, building on myopic and look-ahead sampling
+  strategies. We use simulations to study the properties and
+  differences between state-of-the-art approaches followed by results
+  from field deployments with an autonomous underwater vehicle as part
+  of a case study mapping the boundary of a river plume. The results
   demonstrate the ability of the method to effectively inform and
@@
 % using statistical methods in design and data assimilation
-\citep{das11b,Graham2013,Das2015,das15,fossuminformation,fossum18b}.
-Others working in this space include \cite{wikle2013modern}, focusing on hierarchical
-statistical models; \cite{sahu2008space}, studying spatio-temporal
-models for sea surface temperature and salinity data; and
-\cite{mellucci2018oceanic} looking at the statistical prediction of
-features using an underwater glider. Common to these approaches is the
-use of information-based methods to infer environmental sampling and overcome common challenges, namely:
+\citep{das11b,Graham2013,das15,Das2015,fossum18b,fossuminformation}.
+Others working in this space include \cite{wikle2013modern}, focusing
+on hierarchical statistical models; \cite{sahu2008space}, studying
+spatio-temporal models for sea surface temperature and salinity data;
+and \cite{mellucci2018oceanic} looking at the statistical prediction
+of features using an underwater glider. Common to these approaches is
+the use of information-based methods to infer environmental sampling
+and overcome common challenges, namely:
 
 \begin{itemize}
-\item Fully fledged numerical ocean models cannot be run online on the sensing platforms due to the computational load they require to provide accurate results, hence statistical proxy models of the environment must be used.
-\item Decisions on where to sample happen sequentially, i.e.  
-  recently acquired information can be assimilated into and used to improve statistical models and reduce uncertainty associated with where to make measurements.
-\item As a priori information about the state of the ocean is very
+\item Full numerical ocean models cannot be run online on the sensing
+  platforms due to their computational load for accurate results,
+  hence statistical proxy models of the environment must be used.
+\item Decisions on where to sample happen sequentially, i.e.  recently
+  acquired information can be assimilated into and used to improve
+  statistical models and reduce uncertainty associated with where to
+  make measurements.
+\item As a-priori information about the state of the ocean is very
   uncertain, there is a substantial efficiency gain in reacting to
   information obtained from measurements taken \emph{in-situ}; this is
-  usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.
+  usually referred to as the \emph{adaptivity gap}
+  \citep{Krause2008phd}.
 \item For such types of sampling problems, the number of choices (i.e.
   locations, trajectories, and candidate designs) makes the problem
-  combinatorially large, creating a trade-off between optimization (finding the most resource-efficient design to collect necessary data), and computability (arriving at a
-  solution in reasonable time); to engage in successful sampling, this trade-off has to be considered in sample design and practice.
-\end{itemize}{}
-
-The combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based sensing.
-Central to this is the ability to model spatially-correlated variables and provide formal measures of uncertainty.
-In our formulation we used Gaussian Processes (GPs), in large part because they allow for efficient implementation and evaluation, in turn enabling prioritization of sampling efforts which can be be formalized into decision-making systems on board a robotic platform.
+  combinatorially large, creating a trade-off between optimization
+  (finding the most resource-efficient design to collect necessary
+  data), and computability (arriving at a solution in reasonable
+  time). To successfully resolve features, this trade-off has to be
+  considered in sample design and practice.
+\end{itemize}
+
+The combination of statistical tools and robotic platforms is a
+natural symbiosis which enables information-based sensing. Central to
+this is the ability to model spatially-correlated variables and
+provide formal measures of uncertainty. In our formulation we used
+Gaussian Processes (GPs) in large part because they allow for
+efficient implementation and evaluation, in turn enabling
+prioritization of sampling efforts which can be be formalized into
+decision-making systems on board a robotic platform.
 
@@
 
-In this work, we build on joint excursion sets (ES) and excursion probabilities (EP) for the spatial design of experiments. The goal is to
-design sampling strategies for reducing the spatially-integrated
+In this work, we build on joint excursion sets (ES) and excursion
+probabilities (EP) for the spatial design of experiments. The goal is
+to design sampling strategies for reducing the spatially-integrated
 Bernoulli variance (IBV) in the ES indicators. Our main contributions
@@
 parameters such as temperature and salinity; ii) developing methods
-for information-based environmental sensing and autonomous data collection;
-and iii) demonstrating the approach in the real-world scenario of
-sampling a river plume using an AUV.
+for information-based environmental sensing and autonomous data
+collection; and iii) demonstrating the approach in the real-world
+scenario of sampling a river plume using an AUV.
 
@@
 \ref{sec:bg} provides select background on ocean sampling; Section
-\ref{sec:ESEP} defines ESs, EPs, and the design criteria of
-IBV; Section \ref{sec:GP_EP} sets up the required statistical modeling
-assumptions of the bi-variate GPs for temperature and
-salinity; Section \ref{sec:sur} builds on these assumptions when deriving the analytical derivation of the design criteria; Section
+\ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV;
+Section \ref{sec:GP_EP} sets up the required statistical modeling
+assumptions of the bi-variate GPs for temperature and salinity;
+Section \ref{sec:sur} builds on these assumptions when deriving the
+analytical derivation of the design criteria; Section
 \ref{sec:heuristics} extends these results and present heuristics for
@@
 \ref{sec:case_study} demonstrates the methodology used in field work
-characterizing a river plume; and Section \ref{sec:concl_disc} contains a summary and a discussion of future work.
+characterizing a river plume; and Section \ref{sec:concl_disc}
+contains a summary and a discussion of future work.
 
@@
 
-AUVs are used to gather in-situ spatio-temporal data while carrying a range of scientific payloads to survey the
-water column. Typically, they operate at $\sim 1-3$ m/s, can reach
-depths from 100--6000 m with an in-water operation time capacity dependent on
-survey speed, payload sensors and mission design. By deploying AUVs
-together with other resources, one can augment information obtained
-from ocean models, remote-sensing data, or fixed-location
-sensors on buoys. Traditional operations and surveys with AUVs are
-usually limited to observations along fixed transects on a spatial
-grid, pre-programmed by the human operator. By instead using on-board
-algorithms to continuously evaluate, update, and refine future
-sampling locations, one can make information gathering adaptive and
-contextual to the survey requirements
-\citep{das11b,fossuminformation,fossum18b}. The space of sampling
-opportunities is still limited to the spatial grid, but now there is
-much more flexibility because of adaptation. For instance, an AUV
-could reconstruct or modify a survey line based on what temperatures
-it is measuring by using embedded methods in decision-theoretic
-planning \citep{py10,Rajan12,Rajan12b}.
+AUVs are used to gather in-situ spatio-temporal data while carrying a
+range of scientific payloads to survey the water column. Typically,
+they operate at $\sim 1-3$ m/s, can reach depths from 100--6000 m with
+an in-water operation time capacity dependent on survey speed, payload
+sensors and mission design. By deploying AUVs together with other
+resources, one can augment information obtained from ocean models,
+remote-sensing data, or fixed-location sensors on buoys. Traditional
+operations and surveys with AUVs are usually limited to observations
+along fixed transects on a spatial grid, pre-programmed by the human
+operator. By instead using on-board algorithms to continuously
+evaluate, update, and refine future sampling locations, one can make
+information gathering adaptive and contextual to the survey
+requirements \citep{das11b,fossum18b,fossuminformation}. The space of
+sampling opportunities is still limited to the spatial grid, but now
+there is much more flexibility because of adaptation. For instance, an
+AUV could reconstruct or modify a survey line based on what
+temperatures it is measuring by using embedded methods in
+decision-theoretic planning \citep{py10,Rajan12,Rajan12b}.
 
@@
 \includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
-\caption{The commercially available Light Autonomous Underwater Vehicle (LAUV) platform (Harald) for upper water-
-column exploration used in the experiment.}
+\caption{The commercially available Light Autonomous Underwater
+  Vehicle (LAUV) platform for upper water-column exploration used in
+  the experiment.}
 \label{fig:lauv}

### 2019-09-26 main_new.tex

@@
 Our objective is to leverage statistical tools to characterize a river plume, focusing on spatial separation of cold freshwater from the river and warmer saline waters of the fjord.
-Excursion sets (ES) and excursion probabilities (EP) are useful starting points to measure the ability to classify water masses.
+ES and EPs are useful starting points to measure the ability to classify water masses.
 The goal of experimental sampling is to improve this characterization, as quantified via changes in the EPs.
@@
 Calculations of the EPs and the IBV require a model specification.
-We use Gaussian modeling, which provides a practical and popular probabilistic approach to modeling environmental data, allowing a closed form
-evaluation of the IBVs, as derived in Section \ref{sec:sur}.
+We use Gaussian modeling, which provides a practical and popular probabilistic approach to modeling environmental data, and allows a closed form
+evaluation of the IBVs as derived in Section \ref{sec:sur}.
 
@@
 
-The Gaussian model is used to represent the bivariate random fields of temperature and salinity. At location $\bx \in \mathcal{M}$, the prior bivariate distribution is:
+The Gaussian model is used to represent the bivariate random fields of temperature and salinity. At location $\bx \in \mathcal{M}$, the prior bivariate distribution is
 \begin{equation}\label{gp_bar}
@@
 \end{equation}
-where the notation refers to Gaussian (Normal) distributed variables
+where the notation refers to Gaussian (i.e. normal) distributed variables
 specified by the mean vector and covariance matrix. For the application
@@
 For the spatial dependence among variables, we assume a separable correlation function with the same decay for salinity and temperature,
-which is not unrealistic for our case with a river plume as the decay is related to the common water masses (Sect \ref{sec:case_study}). We have:
+which is consistent with expectations for  a river plume where the decay is related to the common water masses (Sect \ref{sec:case_study}). In the resulting function, we have
 \begin{equation}\label{gp_corr}
@@
 \end{equation}
-for $i,j \in {T,S}$ and $\delta_{ij}=1$ if $i=j$ and $\delta_{ij}=0$
+where for $i,j \in {T,S}$ and $\delta_{ij}=1$ if $i=j$, and $\delta_{ij}=0$
 otherwise. Moreover, we assume stationary isotropic random fields a priori where the
@@
 $\bx'$ via a Euclidean distance measure $h=\sqrt{||\bx-\bx'||^2}$. With
-data and prior knowledge one could possibly fit and estimate
+data and prior knowledge, one could possibly fit and estimate
 parameters of non-stationary or non-isotropic covariance functions and
 more complex multivariate spatial covariance functions
-\citep{gneiting2010matern,genton2015cross}; however that is left for
+\citep{gneiting2010matern,genton2015cross}; however, that is left for
 future work.
@@
 grid is used for the waypoint graph setting the possible design locations. It is also used when solving integrals expressions in Eq. (\ref{measVAR}) and
-(\ref{sur}) which are then approximated by sums over all grid cells.  Using vector
+(\ref{sur}), which are then approximated by sums over all grid cells.  Using vector
 notation, we denote the Gaussian temperature and salinity variables on
-the grid and its density function by: 
+the grid and its density function by the formula: 
 \begin{equation}\label{prior}
@@
 %Here, $\pi(\bxi)$ is the joint density function of length $2 n$ vector $\bxi$, and
-The length $2 n$ vector $\bmu$ holds the expected values of
+The length $2 n$ vector $\bmu$ contains the expected values of
 temperature and salinity at the grid locations, while the
-$2n \times 2n$ matrix $\bSigma$ holds the
+$2n \times 2n$ matrix $\bSigma$ contains the
 variance-covariances of temperature and salinity variables at all grid
@@
 at a subset of locations on the defined grid. This can be done in a
-static manner, where the survey design is pre-planned, or in a
+static manner (where the survey design is pre-planned), or in a
 sequential way with several rounds of data gathering. For the latter,
@@
 nodes in the grid. The conditional model for the data, given the true temperature
-and salinity, is described by: 
+and salinity, is described by 
 \begin{equation}\label{likelihood}
@@
 \end{equation}
-where $\bG$ is an $2n_y \times 2n$ matrix holding '$1$'s on the
-measurement indices on the grid and '$0$' otherwise. The size
+where $\bG$ is a $2n_y \times 2n$ matrix containing '$1$'s on the
+measurement indices of the grid and '$0$' otherwise. The size
 $2n_y \times 2n_y$ covariance matrix $\bR$ contains diagonal entries
@@
 
-The GP mean and covariance in Eq. (\ref{prior}) are updated when more information is available. The
+The GP mean and covariance in Eq. (\ref{prior}) are updated in the model when more information is available. The
 updated distribution for temperature and salinity variables on the
@@
 the data $\by$ and the covariance is not a function of the data.
-Before knowing the data, the distribution of the conditional mean is
+Before inputting the data, the distribution of the conditional mean is
 \begin{equation}\label{distxi} \bm \sim N(\bmu , \bPsi), \hspace{3mm}
@@
 \end{equation}
-which is important for the derivation of the closed form solution of the design criteria.
+which is important for the derivation of the closed-form solution of the design criteria.
 

### 2019-09-28 main_new.tex
Language input - Ginny

@@
 updated model. Then the updated model is used to design and select the
-next round of measurements, and so on. We describe the model for one
-round of data and the Gaussian formula for a single update. In Section
-\ref{sec:heuristics} we discuss sequential data gathering in the
-context of adaptive design of AUV measurements.
+next round of measurements, and so on.
+Below, we describe the model for one round of data and the Gaussian formula for a single update; in Section \ref{sec:heuristics}, we discuss sequential data gathering in the context of adaptive design of AUV measurements.
 
@@
 
-\subsection{Excursion probabilities for Gaussian processes}
+\subsection{Excursion Probabilities for Gaussian Processes}
 
 Based on Gaussian modeling assumptions, the EPs can be computed at any
-location $\bx$. Without additional data, Eq.  (\ref{eq:prob}) gives us:
+location $\bx$. Without additional data, Eq.  (\ref{eq:prob}) gives us
 
@@
 When data $\by$ are available, Eq. (\ref{eq:post_ep}) similarly gives
-us:
+us
 \begin{eqnarray}\label{eq:prob_mv}
@@
 (\ref{likelihood}).
-We next derive results that ease comparison of several designs.
+We next derive results that simplify the comparison of several designs.
 
@@
 
-We will next derive closed form solutions to the expectation in
+We will next derive closed-form solutions for the expectation in
 (\ref{sur}). The location index $\bx$ is suppressed to avoid overly
@@
 bivariate integral \citep{bhattacharjya2013value, chevalier2014fast},
-with expectation over $\bm=(m_{T},m_{S})^t$.  We standardize the two
-variables in Eq. (\ref{sur}), i.e.  $Z_T=\frac{\xi_T-m_{T}}{a_{T}}$,
-$Z_S=\frac{\xi_S-m_{S}}{a_{S}}$, and then
+with an expectation over $\bm=(m_{T},m_{S})^t$.  We standardize the two
+variables in Eq. (\ref{sur}) as   $Z_T=\frac{\xi_T-m_{T}}{a_{T}}$,
+$Z_S=\frac{\xi_S-m_{S}}{a_{S}}$; we can then rephrase the probability as
 \begin{eqnarray}\label{ytom}
@@
 
-For the first part of the above, we have:
+For the first part of the integral above, we have
 \begin{equation}\label{part1:phi2}
@@
 Gaussian vector, with $\mbox{Corr}(Z_{T},Z_{S})=\eta_{T,S}$. Moreover,
-$\bZ$ is chosen to be independent of $\bm$. Using Eq. (\ref{dist_mxi}) we have
+$\bZ$ is chosen to be independent of $\bm$. Using Eq. (\ref{dist_mxi}), we have
 \begin{equation}
     E(a_{T} Z_{T}+m_{T}-t_T) = \mu_{T}-t_T, \hspace{3mm}
-    \mbox{Var}(a_{T} Z_{T}+m_{T}-t_T) = a^2_T+ \psi^2_{T}, 
-\end{equation}
-for the temperature part and the same holds for the salinity part. From this we get:
+    \mbox{Var}(a_{T} Z_{T}+m_{T}-t_T) = a^2_T+ \psi^2_{T} 
+\end{equation}
+for the temperature part; the same holds for the salinity part. From this we get:
 \begin{eqnarray}\label{two_parts0}
@@
 both with element-wise correlation $\eta_{T,S}$. Moreover, $\bZ_1$ and
-$\bZ_2$ are chosen to be independent of $\bm$. Hence:
+$\bZ_2$ are chosen to be independent of $\bm$, hence:
+
 \begin{equation}
+\small
+\label{eq:22}
     \left(
@@
 
-The solution to the second part of the equation above, is then to
-evaluate a four-variable cumulative distribution function associated
-with Eqn. (\ref{part1:phi4}), accounting for the mean, variance and
+The solution to the second part of Eq. \eqref{eq:22} can be found by evaluating a four-variable cumulative distribution function associated
+with Eq. (\ref{part1:phi4}), accounting for the mean, variance and
 covariances of $\bZ_1$, $\bZ_2$ and $\bm$ in the appropriate linear
-combinations. Using vector-matrix formulations, we define linear
-combinations:
-\begin{equation}\label{Vcomb}
+combinations. Using vector-matrix formulations, we define the linear
+combinations
+\begin{equation}
+\label{Vcomb}
     \bV=\left[
@@
 \end{equation}
-and this in turn then becomes:
-\begin{eqnarray}\label{part1:phi4_2}
-&& P \left( Z_{1,T} \leq \frac{t_T-m_{T}}{a_{T}}, 
-Z_{1,S} \leq \frac{t_S-m_{S}}{a_{S}},Z_{2,S} \leq \frac{t_T-m_{T}}{a_{T}}, 
-Z_{2,S} \leq \frac{t_S-m_{S}}{a_{S}} \right)  \\
+which in turn then becomes:
+{\small
+\begin{eqnarray}
+\label{part1:phi4_2}
+&&P \left( Z_{1,T} \leq \frac{t_T-m_{T}}{a_{T}}, 
+Z_{1,S} \leq \frac{t_S-m_{S}}{a_{S}}, Z_{2,S} \leq \frac{t_T-m_{T}}{a_{T}}, Z_{2,S} \leq \frac{t_S-m_{S}}{a_{S}} \right)\\
 &=&  \Phi_4 
@@
 \end{eqnarray}
-
+}
 The multivariate cumulative probabilities in two and four dimensions
-are fast to compute using methods of \cite{genz2009computation}.  In
+are fast to compute using methods developed by \cite{genz2009computation}.  In
 the end, the first part in Eq. (\ref{two_parts0}) and the second in
-Eq. (\ref{part1:phi4}) are computed for each location on the grid, and
-summed over $\bx \in \mathcal{M}_g$, gives the design evaluation
+Eq. (\ref{part1:phi4}) are computed for each location on the grid and
+summed over $\bx \in \mathcal{M}_g$, which gives the design evaluation
 criterion in Eq. (\ref{sur}).
@@
 distribution functions for Gaussian vectors. For the more general
-situation one would be interested in $K$ random fields. Say, in the
-ocean sciences the variables could include chlorophyll or other
-bio-geochemical quantities, in addition to temperature and salinity.
+situation, one would be interested in $K$ random fields. In the
+ocean sciences, the variables could include chlorophyll or other
+bio-geochemical quantities in addition to temperature and salinity.
 
@@
 \end{equation}
-By the first key result of linear combination of Gaussian variables,
+From the first key result of linear combination of Gaussian variables,
 we reduce the integral to a $K$ dimensional integral of the relevant
@@
 standardize the variables and compute cumulative distributions.  In
-summary, we get:
+summary, we get
 \begin{equation}\label{two_parts}
@@
 
-\subsection{An Example}
-
-We calculate EPs and the Bernoulli variance for a few bivariate
+\subsection{An Example of IBV using Temperature and Salinity}
+
+In this section, we calculate EPs and the Bernoulli variance for a few bivariate
 Gaussian distributions to illustrate the concepts articulated above.
@@
   \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
-  \caption{Density contour plots having different correlations between
+  \caption{Density contour plots with different correlations between
     temperature and salinity. The densities have unit variance and the
     thresholds are identical to the mean values $5^o C$ and
-    $30 mg/l$. x-axis is temperature and y-axis is salinity.}
+    $30 mg/l$. X-axis is temperature and y-axis is salinity.}
 \label{illus_bivarDens}
 \end{figure}
-Here, the thresholds are set equal to the mean, $\mu_T=t_T=5^o C$ and
+Here, the thresholds are set equal to the mean $\mu_T=t_T=5^o C$, and
 salinity $\mu_S=t_S=30$ mg/l. The displayed densities have unit
@@
 Fig. \ref{illus_bivarDens}. The EPs increase with the correlation as
-there is a stronger tendency of having jointly small temperature and
-salinity. The Bernoulli variance is similarly largest for high
-correlation. EPs and Bernoulli variances are the same for standard
-deviation $1$ or $2$, and this implies that high variability in
+there is a strong tendency to have concurrently low temperature and
+salinity. The Bernoulli variance is similarly large for high
+correlations. EPs and Bernoulli variances are the same for standard
+deviation $1$ or $2$, which implies that high variability in
 temperature and salinity is not captured in the $p(1-p)$ expression.
@@
     different correlations and variances (top rows), and expected
-    Bernoulli variances for both temperature and salinity data $\by$ and only for
+    Bernoulli variances for both temperature and salinity data $\by$ and 
     temperature $y_T$ (bottom rows).}
@@
 gathering both data types $(y_T,y_S)$, and for a design with
-temperature measurements $y_T$ alone. Having both data;
+temperature measurements $y_T$ alone. With both data,
 $(y_T,y_S)^t=(\xi_T,\xi_S)^t+N(0,0.5^2\bI)$, while
@@
 also uncertainty reduction when only temperature is measured (bottom
-row), especially when temperature and salinity are very
-dependent. When correlation is low ($\gamma=0.2$) there is little
-information about salinity in the temperature data, and hence less
+row), especially when temperature and salinity are highly correlated. When correlation is low ($\gamma=0.2$), there is little
+information about salinity in the temperature data, and therefore less
 uncertainty reduction. In an application with fresh cold water from
@@
 
-An important synergy underlining this work is the potential of robotic sampling to adapt survey plans based on data and statistical inference,
+An important concept underlying this work is the potential of robotic sampling to adapt survey plans based on data and statistical inference
 in accordance with the sense-plan-act loop
 (Fig. \ref{fig:sense-plan-act}). We will therefore present a few adaptive strategies
-for robotic sampling that builds on the notion of multivariate ES. Simulations, analysis, and  results from full-scale trials are presented.
+for robotic sampling that build on the notion of multivariate ES. Simulations, analysis, and  results from full-scale trials are presented.
 
@@
 
-An adaptive AUV survey is split in many rounds (stages), and at each
-round, the survey design is found by using a myopic approach of
-considering the largest reduction in uncertainty of the ES. In our
-setting the selection is made on the graph defined by the nearest grid
+An adaptive AUV survey is split into many rounds (stages). In each
+round, the survey design is created using a myopic approach that considers the largest reduction in uncertainty of the ES. In our
+setting, the selection is made on the graph and defined by the nearest grid
 nodes in the domain $\mathcal{M}_g$.
@@
 The optimal sequential design not only considers the best current AUV
-grid node, but also what data gathered at this node would lead to for
-future sampling at successive stages.  Let $d^{j,s}$ denote design
+grid node, but also how the data gathered at this node could inform future sampling at successive stages. Let $d^{j,s}$ denote design
 number $j$ at stage $s$ of a sequential survey. If this design is
 selected, data $\by^{j,s}$ will be gathered. The optimal path
-selection situation is then depicted in Figure \ref{fig:PathSelOpt},
+selection situation is  depicted in Figure \ref{fig:PathSelOpt},
 \begin{figure}[h!]
@@
 series of intermixed maximizations over designs and integrals over
-data.  In practice the optimal solution is not tractable because of
-the enormous growth over stages, see
-e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}.
+data.  In practice, the optimal solution is intractable because of
+the enormous growth over stages (see
+e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}).
 Instead, we outline heuristic strategies. In terms of notation, let
@@
 
-The derived closed form solutions for the expected IBV are still
-important building blocks when constructing the sampling designs as
+The derived closed-form solutions for the expected IBV are still
+important building blocks when constructing the sampling designs, as
 they will be used to score the different adaptive designs. % Efficient
@@
 sampling location based on current EPs. The largest variance in
-sampling locations is and indication of higher uncertainty; therefore
+sampling locations is an indication of higher uncertainty; therefore
 one selects the design location with an EP closest to $0.5$.

### 2019-09-28 main_new.tex
Trygve updates

@@
 \begin{itemize}
-\item Full numerical ocean models cannot be run online on the sensing
-  platforms due to their computational load for accurate results,
-  hence statistical proxy models of the environment must be used.
+\item Full numerical ocean models cannot provide accurate results online if run on robotic sensing
+  platforms, as on board computers cannot deliver the computational power required,
+  hence statistical proxy models of the environment must be used instead.
 \item Decisions on where to sample happen sequentially, i.e.  recently
@@
   resolution of phenomena of scientific interest.]{\includegraphics[width =
-    0.48\textwidth]{Figures/envir.pdf}\label{fig:envir1}}
+    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
   \hfill
-  \subfigure[Data-driven sampling under the sense-plan-act paradigm.]{\includegraphics[width =
-    0.48\textwidth]{Figures/adaptive_cycle.pdf}\label{fig:sense-plan-act}}
+  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
+    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
   \caption{Ocean observation is moving away from single-ship sampling
@@
 
-\begin{figure}[!h] 
-\centering 
-\includegraphics[width=0.85\textwidth]{Figures/pictures/c-updated.pdf}
-\caption{Frontal patterns off of the Nidelva river, Trondheim, Norway.}
-\label{fig:nidelven}
-\end{figure}
+%\begin{figure}[!h] 
+%\centering 
+%\includegraphics[width=0.85\textwidth]{Figures/pictures/c-updated.pdf}
+%\caption{Frontal patterns off of the Nidelva river, Trondheim, Norway.}
+%\label{fig:nidelven}
+%\end{figure}
 
@@
 sampling location based on current EPs. The largest variance in
-sampling locations is an indication of higher uncertainty; therefore
+sampling locations indicates higher levels of uncertainty; therefore
 one selects the design location with an EP closest to $0.5$.
@@
 $\mathcal{Y}^{s-1}$, we fit an updated Gaussian model from
-Eq. (\ref{gp_upd}), with mean $\bm_{s-1}=E(\bxi|\mathcal{Y}^{s-1})$
+Eq. (\ref{gp_upd}) with mean $\bm_{s-1}=E(\bxi|\mathcal{Y}^{s-1})$
 and covariance matrix $\bA_{s-1}=\mbox{Var}(\bxi|\mathcal{Y}^{s-1})$.
-The next round of measurements $\by^{j,s}$, can be gathered with
+The next round of measurements $\by^{j,s}$ can be gathered with
 design $d^{j,s}$, $j=1,\ldots,J$, where $J$ indicates all possible AUV
-trajectories from the current stage. The {\it{naive}} strategy then,
+trajectories from the current stage. The {\it{naive}} strategy, then,
 selects the design according to:
@@
 This strategy does not account for the uncertainty in the temperature
-or salinity variables, only if one design (node) has EPs closer to
-$0.5$ (see Table \ref{tab:sim_rhoab}, line two). Neither does it
+or salinity variables, but is instead applicable only if one design (node) has EPs closer to
+$0.5$ (see Table \ref{tab:sim_rhoab}, line two). Additionally, this strategy does not
 account for spatial correlation. This strategy lacks memory of where
-it has been and where the uncertainty has been reduced and therefore
+it has been and where the uncertainty has been reduced, and is therefore
 susceptible to local minimas.
@@
 imagine taking only one more round of measurements. In this selection
-strategy there is no anticipation of what the subsequent designs might
-offer, beyond the first round.
+strategy, there is no anticipation of what the subsequent designs might
+offer beyond the first round.
 
@@
 the expectation is with respect to the conditional density
-$\pi(\by^{j,s}|\mathcal{Y}^{s-1})$. A similar closed form calculation
-for expected IBV is hence applicable in Eq. (\ref{critSEQ}), using the
+$\pi(\by^{j,s}|\mathcal{Y}^{s-1})$. A similar closed-form calculation
+for expected IBV is hence applicable in Eq. (\ref{critSEQ}) using the
 updated GP model from step $s-1$.  Once the data is collected for the
@@
 applications. Moreover, it is easily implemented on-board an AUV,
-using the efficient approach for data updating of the GP model and the
-calculation of the closed form expected IBV expressions for each
+using an efficient approach for data updating the GP model and the
+calculation of the closed-form expected IBV expressions for each
 subsequent trajectory.
@@
 
-We now extend the myopic strategy to a look-ahead strategy which is
+We now extend the myopic strategy to a look-ahead strategy, which is
 optimal when one can gather only two more rounds of measurements. In
@@
 look-ahead strategy, anticipates the subsequent design $j_2$ with
-data $\by^{j_2,s+1}$, when choosing the current design $d^{j,s}$.  The
+data $\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$.  The
 selected design is:
@@
 $\mathcal{Y}^{j,s+1}=\{\by^{j_2,s+1},\mathcal{Y}^{j,s}\}$ represent
-the sets of data variables, and the expectations are with respect to
+the sets of data variables, and the expectations are calculated with respect to
 the conditional densities $\pi(\by^{j,s}|\mathcal{Y}^{s-1})$ and
@@
 conditional distribution. For each data sample, the second expectation
-is solved using the closed form expressions for expected IBV from
+is solved using the closed-form expressions for expected IBV from
 Section \ref{sec:sur}.
 
-Even though the strategy looks at two stages, it is only used to find
-the current best design. When data is collected, the GP model is
+Even though the strategy looks at two sampling stages, it is only used to find
+the current best design. When data are collected, the GP model is
 updated, and the mean $\bm_{s}$ and covariance matrix $\bA_{s}$ are
@@
 This look-ahead approach is much more computationally demanding than
-the myopic strategy, and for practical implementation we prune paths
+the myopic strategy, so for practical implementation, we prune paths
 in the evaluation of Eq. (\ref{critLA}). This means that we do not

### 2019-09-29 main_new.tex
Language updates Ginny

@@
 based on closing the sensing and actuation loop
-(Fig. \ref{fig:sense-plan-act}). Myopic sampling—i.e. step-wise
+(Fig. \ref{fig:envir1}). Myopic sampling—i.e. step-wise
 selection of the path (on the graph) which reduces the information
@@
 challenging if the probability is close to $0.5$. 
-The integrated Bernoulli variance (IBV) \cite{bect2019}, is defined as: 
+The IBV \cite{bect2019}, is defined as: 
 \begin{equation}\label{measVAR}
@@
 
-\subsection{Data model and updating of Gaussian processes}
+\subsection{Data Model and Updating of Gaussian Processes}
 
@@
 
-\subsection{Expected Integrated Bernoulli variance for bivariate models}
+\subsection{Expected Integrated Bernoulli Variance for Bivariate Models}
 
@@
 We will next extend results of \cite{chevalier2014fast} to solve for
-these two parts (see also \cite{stroh} for such an extensions).
+these two parts (see also \cite{stroh}).
 
@@
 
-For the second part of the expression (\ref{eq:var2}), standardization
+For the second part of the expression Eq.(\ref{eq:var2}), standardization
 implies that: 
@@
 
-\subsection{Generalization to the multivariate case}
+\subsection{Generalization to the Multivariate Case}
 \label{gen_k}
@@
 in accordance with the sense-plan-act loop
-(Fig. \ref{fig:sense-plan-act}). We will therefore present a few adaptive strategies
+(Fig. \ref{fig:envir1}). We will therefore present a few adaptive strategies
 for robotic sampling that build on the notion of multivariate ES. Simulations, analysis, and  results from full-scale trials are presented.
 
-\subsection{Optimal sequential design}
+\subsection{Optimal Sequential Design}
 \label{Optdes}
@@
 it has been and where the uncertainty has been reduced, and is therefore
-susceptible to local minimas.
-
-\subsection{Myopic path planning}
+susceptible to local minima.
+
+\subsection{Myopic Path Planning}
 \label{myopic}
@@
 
-\subsection{Look-ahead trajectory planning}
+\subsection{Look-ahead Trajectory Planning}
 \label{LA}
@@
 indicated in Figure \ref{fig:PathSelOpt}. Instead, we use the myopic
-strategy to rank the three best designs on the first stage alone, and
-for each of these we undertake the look-ahead calculations.
+strategy to rank the three best designs following the first stage alone, and
+for each of these preferred designs, we undertake the look-ahead calculations.
 
@@
 different static and sequential survey designs, in a realistic
-simulated case. The context is mapping a freshwater plume defined by a
+simulated case: mapping a freshwater plume defined by a
 temperature and salinity gradient using an AUV.
@@
 the eastern part of the domain to be cold freshwater. This situation
-mimics that of a river mouth entering from the south, and the water
-masses are pulled to the east. The actual trends are, in practice, set
+mimics that of a river mouth opening from the south, with the water
+masses pulled to the east. The actual trends are, in practice, set
 from preliminary information or runs of oceanographic models \cite{fossuminformation} solving differential equations for hydrodynamic flow. In this
-example the mean is specified by regression parameters $\beta_{T,1}=0.065$ and
-$\beta_{S,1}=0.1$ for the slope with west coordinate and $\beta_{T,0}=5.8$
+example, the mean is specified by regression parameters $\beta_{T,1}=0.065$ and
+$\beta_{S,1}=0.1$ for the slope, with west coordinate and $\beta_{T,0}=5.8$
 and $\beta_{S,1}=29.0$ for the intercept at a reference start map coordinate.
@@
 \begin{equation}\label{v}
+\small
 \mbox{Corr} 
@@
 below.  One realization of the random fields representing salinity and
-temperature is shown in Figure \ref{fig:true_temp} and
+temperature is shown in Figures \ref{fig:true_temp} and
 \ref{fig:true_sal}. The true ES from this prior realization is shown
@@
 
-\subsection{Static and sequential sampling designs}\label{sec:sampling_designs}
+\subsection{Static and Sequential Sampling Designs}\label{sec:sampling_designs}
 
@@
 We compare three different static designs denoting
-\texttt{static\_north}, \texttt{static\_east}, and
-\texttt{static\_zigzag} with the three sequential approaches
-\texttt{naive}, \texttt{myopic}, and \texttt{look-ahead}. The static
+\textit{static\_north}, \textit{static\_east}, and
+\textit{static\_zigzag} with the three sequential approaches
+\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
 sampling paths are pre-determined and cannot be changed on the basis
-of data, representing the typical pre-planned observation
+of data, representing the pre-planned observation
 strategies used in typical AUV operational survey designs.
@@
 However, for the sequential approaches this is not the case, as the result and path are dependent on the observed data and the expected reduction in IBV.
-The properties is therefore evaluated using Monte Carlo sampling over several replicates of realizations from the prior, conducting simulated sequential surveying for each one.
+The properties are therefore evaluated using Monte Carlo sampling over several replicates of realizations from the prior, conducting simulated sequential surveying for each one.
 Figure \ref{fig:Eprob} shows the conditional EP, given data gathered along the north-south survey line for the realization indicated in these displays.
@@
 We also compare predictive performance measured by root mean square error (RMSE) for the temperature and salinity estimates and the variance reduction in these two variables.
-It is important to note that the objective function used by the agents is focused on reducing the expected variance in the ES, but we nevertheless hope to achieve good predictive performance for criteria such as RMSE as well.
+It is important to note that the objective function used by the agents is focused on reducing the expected variance in the ES, but we nevertheless expect that we will achieve good predictive performance for criteria such as RMSE as well.
 Another non-statistical criteria that is important for practical purposes is the computational time of the strategy, as this will impact the performance of the approach in the field for an embedded system.
@@
 
-Each strategy is conducted on an equilateral grid shown in Fig.
+Each strategy is conducted on an equilateral grid as shown in Fig.
 \ref{fig:wp_graph}.
@@
 The AUV will move along edges in the waypoint graph while the vehicle gathers measurements.
-The data is assimilated into the GP model, before an evaluation of the next node to sample is conducted at the end of the edge.
+The data is assimilated into the GP model before an evaluation of the next node to sample is conducted at the end of the edge.
 The paths will differ for the various sequential designs due to the individual design criteria and the simulated variability in the environment among the replicates.
 
-A total of 100 replicate simulation was conducted and the results are shown in Figure \ref{fig:sim_results},
+\subsection{Simulation Results}
+A total of 100 replicate simulations were conducted; the results are shown in Figure \ref{fig:sim_results},
 where the different criteria are plotted as a function of survey distance.
 Figure \ref{fig:avg_ev} shows the resulting drop in ES variance for each of the six strategies.
-The ES variance is reduced most under the \texttt{myopic} and \texttt{look-ahead} strategies, performing almost equally; this is expected as the two criteria (Eq.
+The ES variance is reduced most under the \textit{myopic} and \textit{look-ahead} strategies, each performing almost equally; this is expected as the two criteria (Eq.
 \eqref{critSEQ} and \eqref{critLA}) are much more sensitive to differences in IBV.
@@
 
-Figure \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop in RMSE and increase in explained variance.
-Both \texttt{myopic} and \texttt{look-ahead} perform well here, but some of the \texttt{static\_east} and \texttt{static\_zigzag} also achieve good results because they are pre-determined to cover large parts of the domain without re-visitation.
-Sequential strategies targeting ES variance will sometimes not reach similar coverage as interesting data may draw the path into turns and twists that may result in re-visitation.
+Figure \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop in RMSE and increase in explained variance, respectively.
+Both \textit{myopic} and \textit{look-ahead} strategies perform well here, but some of the \textit{static\_east} and \textit{static\_zigzag} also achieve good results because they are pre-determined to cover large parts of the domain without re-visitation.
+Sequential strategies targeting ES variance will sometimes not reach similar coverage, as interesting data may draw the path into twists and turns that may result in re-visitation.
 
@@
   \hfill
-  \caption{The overlaid route choices (strategies), superimposed for
+  \caption{The overlaid route choices (strategies), superimposed on
     the 100 replicate surveys with 10 sampling choices/steps.}
@@
 Fig.
-\ref{fig:avg_time} shows the computational effort; the \texttt{naive} strategy is on par with the static designs, while the \texttt{myopic} strategy is slower and \texttt{look-ahead} more so, with the latter reaching levels that are nearly impractical for execution on a vehicle.
+\ref{fig:avg_time} shows the computational effort: the \textit{naive} strategy is on par with the static designs, while the \textit{myopic} strategy is slower.
+The \textit{look-ahead} is even slower, reaching levels that are nearly impractical for execution on a vehicle.
 Some pruning of the graph is performed to improve the computational speed, such as ruling out repeated visitations and back-and-forth routes.
@@
 Still, more pruning of branches or inclusion of other heuristics could be included to make the strategy faster.
-Then again, the inclusion of heuristics is likely a contributing factor that the \texttt{look-ahead} strategy is not outperforming the \texttt{myopic} strategy.
-In Figure \ref{fig:route_choices} the realized sampling paths for each sequential schemes and static designs is shown.
-
-The \texttt{naive} strategy gets stuck in the southern part of the domain in some cases, because it is too focused on the probabilities near $0.5$.
-The \texttt{myopic} strategy meanwhile, covers a wider domain than the look-ahead, there are several reasons for this.
-First, a greedy approach will tend to do more put more emphasis on locations close to the agent, which may lead away from the centre.
-Secondly, as the agent evaluates the impact of locations further away (look-ahead) where assimilated data has less effect, the prior (which is centered here) will act to restrict paths deviating from the central zone.
-
-We further go on to study sensitivity of the result to different prior correlation values between temperature and salinity, prior standard deviations, spatial correlations, as well as measuring only temperature on-board the AUV (as indicated in the rows).
+Then again, the inclusion of heuristics is likely a contributing factor to the \textit{look-ahead} strategy failing to outperform the \textit{myopic} strategy.
+In Figure \ref{fig:route_choices}, the realized sampling paths for each of the sequential schemes and static designs is shown.
+
+In this model, the \textit{naive} strategy often gets stuck in the southern part of the domain because it is too focused on the probabilities near $0.5$.
+The \textit{myopic} strategy, meanwhile, covers a wider domain than the naive or look-ahead.
+There are several reasons for this; first, a greedy approach will tend to do more put more emphasis on locations close to the agent, which may lead away from the centre.
+Second, as the agent evaluates the impact of locations further away (look-ahead) where assimilated data has less predictive power, the GP prior (which is centered here) will act to restrict paths deviating from the central zone.
+
+\subsection{Performance Metrics}
+The tables below explore the sensitivity of the result to different prior correlation values between temperature measured on-board the AUV, temperature and salinity, prior standard deviations, and spatial correlations (see table rows).
 We only compare results at the end of each survey.
-The expected ES variances are shown in Table \ref{tab:sim_res_ev}, results of RMSE are in Table \ref{tab:sim_res_rmse}, and explained variance in Table \ref{tab:sim_res_r2}.
-In all runs, both the \texttt{myopic} and \texttt{look-ahead} strategies perform the best in terms of expected variance in the ES.
-Surprisingly, the \texttt{myopic} is on-par and even slightly better than the \texttt{look-ahead} strategy.
-Part of the reason can be traced to the act of pruning some of the paths for the \texttt{look-ahead} strategy, but even with less pruning, the \texttt{look-ahead} does not outperform the \texttt{myopic} approach.
-The \texttt{static\_north} design is clearly better than the other two static designs, the reason being that the others do not focus on the areas where the ES is really uncertain (centre locations are the most uncertain).
-By measuring temperature alone the expected variances in the ES is only slightly higher.
-This occurs because the ES is almost the same and because of the correlation of $0.6$, which means that temperature observations provide information about salinity as well.
-For the other predictive measures, \texttt{myopic} and \texttt{look-ahead} do well, clearly beating the \texttt{naive} strategy.
-The \texttt{static\_zigzag} design is usually the best among the static designs for these predictive performance measures because its pattern gives more effective coverage of the domain when there is significant spatial correlation as in this case.
+The expected ES variances are shown in Table \ref{tab:sim_res_ev}, results of RMSE are in Table \ref{tab:sim_res_rmse}, and explained variance is in Table \ref{tab:sim_res_r2}.
+In all runs, both the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of expected variance in the ES.
+Surprisingly, the \textit{myopic} performs slightly better than the \textit{look-ahead} strategy.
+Part of the reason for this performance discrepancy can be traced to the act of pruning some of the paths for the \textit{look-ahead} strategy; however, even with less pruning, the \textit{look-ahead} does not outperform the \textit{myopic} approach.
+The \textit{static\_north} design is clearly better than the other two static designs, the reason being that the others do not focus on the areas where the ES is really uncertain (centre locations are the most uncertain).
+By measuring temperature alone, the expected variances in the ES are only slightly higher than the combined temperature-salinity estimate.
+This occurs because the correlation of $0.6$ means that temperature observations provide information about salinity as well.
+For the other predictive measures, \textit{myopic} and \textit{look-ahead} strategies do well, clearly beating the \textit{naive} strategy.
+The \textit{static\_zigzag} design is usually the best among the static designs for these predictive performance measures because its pattern gives more effective coverage of the domain when there is significant spatial correlation, as indicated in this case.
 
@@
 
-\section{Case study - Mapping a river plume}
+\section{Case Study - Mapping a River Plume}
 \label{sec:case_study}
@@
 
-\subsection{Model specification}
+\subsection{Model Specification}
 
 The statistical model parameters were specified based on a short preliminary survey where the AUV did an initial transect to determine the environmental conditions and correlation structure.
-Based on the initial data the trend parameters were estimated by linear regression, where both temperature and salinity are assumed to increase linearly, going west from the river mouth.
-Next, the residuals from the regression analysis are analyzed to study the fit of the Gaussian model and to specify the covariance parameters.
+Based on the initial data, the trend parameters were estimated by linear regression, where both temperature and salinity are assumed to increase linearly, going west from the river mouth.
+Next, the residuals from the regression analysis were analyzed to study the fit of the Gaussian model and to specify the covariance parameters.
 
@@
 Figure \ref{fig:parest} shows diagnostic plots of this analysis.
-The display on the left shows a cross-plot of temperature and salinity residuals, after the westernly trends in both salinity and temperature are subtracted from the data.
+The display on the left shows a cross-plot of temperature and salinity residuals after the westernly trends in both salinity and temperature are subtracted from the data.
 This scatterplot of joint residuals indicates larger variability in salinity than in temperature, and a positive correlation ($0.5$) between the two variables.
 Based on the fitted bivariate Gaussian model (ellipses in left display), we can compute the modeled quadratic form of the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed.
-Figure \ref{fig:parest} (middle) shows the empirical (cumulative distribution function) CDF of these quadratic forms (solid) together with the theoretical CDF of the $\chi^2_2$ distribution.
-The modeled and theoretical curves are very similar and this indicates that the Gaussian model fits reasonable well.
+Figure \ref{fig:parest} (middle) shows the empirical cumulative distribution function (CDF) of these quadratic forms (solid) together with the theoretical CDF of the $\chi^2_2$ distribution.
+The modeled and theoretical curves are very similar, which indicates that the Gaussian model fits reasonably well.
 Even though there appears to be some clustering in both left and middle displays of Figure \ref{fig:parest}, the bivariate diagnostic plots look reasonable and justify a Gaussian model.
@@
 Based on the analysis in Fig.
-\ref{fig:parest} the resulting parameters are given in Table \ref{tab:experiment_param}.
+\ref{fig:parest}, the resulting parameters are given in Table \ref{tab:experiment_param}.
 The regression parameters shown here are scaled to represent the east and west boundaries of the domain as seen in the preliminary transect data, and the thresholds are intermediate values.
-These parameter values were then used in field trials, where we explored the algorithm's ability to characterize the river plume front separating the river and fjord water masses, providing a spatial map of this boundary.
+These parameter values were then used in field trials where we explored the algorithm's ability to characterize the river plume front separating the river and fjord water masses, providing a spatial map of this boundary.
 
@@
 \end{tabular}
-\caption{Model and threshold parameters found from an initial AUV
+\caption{Model and threshold parameters from an initial AUV
   survey. Observations were taken across the front while crossing from
-  fresh cold river water to saline and warmer ocean waters.}
+  fresh, cold river water to saline and warmer ocean waters.}
 \label{tab:experiment_param}
@@
 
-\subsection{Experimental setup}
-
-The sampling locations was distributed over an equilateral grid, as shown in Fig.
-\ref{fig:map} in a grey colored lattice.
-The robotic platform consisted of a Light AUV \citep{sousa2012lauv} (Figure \ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49 CTD (conductivity, temperature, and depth) sensor providing temperature and salinity measurements.
+\subsection{Experimental Setup}
+
+The sampling locations were distributed over an equilateral grid, as shown in the grey-colored lattice in Fig.
+\ref{fig:map}.
+The robotic platform consisted of a Light AUV \citep{sousa2012lauv} (Figure \ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49 conductivity, temperature, and depth (CTD) sensor providing temperature and salinity measurements.
 The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
-The sampling agent was built on top of the autonomous agent framework \texttt{T-REX} \citep{py10,Rajan12,Rajan12b}, running an instance of the \texttt{myopic} strategy from Section \ref{sec:sampling_designs}, to control the AUV and decide between sampling locations.
+The sampling agent was built on top of the autonomous agent framework Teleo-Reactive EXecutive (\textit{T-REX}) \citep{py10,Rajan12,Rajan12b}, running an instance of the \textit{myopic} strategy from Section \ref{sec:sampling_designs} to control the AUV and decide between sampling locations.
 
@@
 The sampling strategy was designed around the concept of visiting waypoints sequentially.
-Arriving at a desired waypoint, with new measurements and an updated model, the AUV triggers the myopic strategy to evaluate all possible design criteria, e.g Eq.
-\eqref{critSEQ}.
-The waypoint-and-path combination which is expected to reduce the IBV the most is selected; upon arrival this procedure is then repeated.
-At each stage, it takes the AUV about 30 sec to evaluate the expected IBV for all the possible waypoint-and-path alternatives.
-
-The AUV was set to start in the south center part of the waypoint graph, with the previously outlined GP prior model of the environment.
-Each AUV survey was set to take approximately 40 min, visiting 15 waypoints on the grid, with the AUV running near the surface to better capture the plume.
-On its path from one waypoint to the next, the AUV gathered data regularly, and the GP model assimilated temperature and salinity data with an update frequency of 30 sec, about three updates per stage.
+Arriving at a desired waypoint with new measurements and an updated model, the AUV triggers the myopic strategy to evaluate all possible design criteria (see Eq.
+\eqref{critSEQ}).
+The waypoint-and-path combination that is expected to reduce the IBV the most is selected; upon arrival this procedure is then repeated.
+At each stage, it takes the AUV about 30 seconds to evaluate the expected IBV for all the possible waypoint-and-path alternatives.
+
+The AUV was set to start in the south-center part of the waypoint graph, with the previously outlined GP model of the environment.
+Each AUV survey was set to take approximately 40 minutes to visit 15 waypoints on the grid, with the AUV running near the surface to better capture the plume.
+On its path from one waypoint to the next, the AUV gathered data regularly, and the GP model assimilated temperature and salinity data with an update frequency of 30 seconds, with about three updates per stage.
 
@@
 \centering
-\subfigure[Survey map.]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
+\subfigure[Survey map]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
 \hspace{0.3cm}
-\subfigure[Temperature tracks.]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/auv.pdf}\label{fig:res_both}}
+\subfigure[Temperature tracks]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/auv.pdf}\label{fig:res_both}}
 
@@
 
-\subfigure[ES for Survey 1.]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
-\subfigure[ES for Survey 2.]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
-\caption{Results from mapping the Nidelven river} 
+\subfigure[ES for Survey 1]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
+\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
+\caption{Results from mapping the Nidelva river} 
 \label{fig:results}
@@
 
-Two missions, Survey 1 and 2, were run successively from 11:00 AM to 01:00 PM, with a short break in between.
+Two missions, Surveys 1 and 2, were run successively from 11:00 AM to 01:00 PM, with a short break in between.
 The resulting path of the selected waypoints are shown in the map in Fig.
-\ref{fig:map}, both within the expected frontal region (shaded pink region).
+\ref{fig:map}, both within the expected frontal region (shaded pink).
 The recorded temperatures are shown as colored trails in Fig.
-\ref{fig:res_both}, clearly showing the temperature difference between fjord and river waters.
-The salinity data are then shown separately overlaid with the estimated EP for each survey in Fig.
+\ref{fig:res_both}, clearly indicating the temperature difference between fjord and river waters.
+The salinity data are then shown separately, overlaid with the estimated EP for each survey in Fig.
 \ref{fig:res1} and \ref{fig:res2}.
 
-Both Survey 1 and 2 successfully estimate and navigate the separation zone crossing the frontal boundary multiple times.
-As conditions changed slightly between the two deployments the resulting path (after waypoint 5) deviates.
-Survey 1 continues northwards tracking the north-eastern part of the front, while Survey 2 turns west mapping the south-western part.
+Both Surveys 1 and 2 successfully estimated and navigated the separation zone, crossing the frontal boundary multiple times.
+As conditions changed slightly between the two deployments, the resulting path (after waypoint 5) deviates.
+Survey 1 continued northwards, tracking the north-eastern part of the front, while Survey 2 turned west, mapping the south-western part.
 
 The final prediction of the front location given in Fig.
-\ref{fig:res3} and \ref{fig:res4}, correspond with one another yielding a picture of the front gradually bending off towards north east.
+\ref{fig:res3} and \ref{fig:res4}, correspond with one another, yielding a picture of the front gradually bending off toward the north east.
 The amount of exploration done by Survey 1 is greater than Survey 2.
-In Survey 1 the AUV gets overall more detail by going north, while Survey 2 coming close to the survey area borders in the south-western corner get poorer understanding of the northern parts.
+In Survey 1, the AUV obtained more detail by going north, while Survey 2, coming close to the survey area borders in the south-western corner, obtained a poorer understanding of the northern parts.
 A longer planning horizon would likely identify and discourage such choices.
@@
 This paper builds on multidisciplinary work combining statistical modeling and methods with robotic surveying technologies for oceanographic sensing applications.
-We show how the observation practices can gain from development of statistical techniques to achieve more effective spatial monitoring.
-We further demonstrate the opportunities available for real-time multivariable spatial data gathering and analysis onboard autonomous sensing platforms, upon which statisticians can exploit towards creating general-purpose toolkits for similar applications (enabeling theory).
-
-In particular, we derive and show results for characterizing phenomena connected to water masses properties.
-The characterization is based on joint excursion sets and we derive new results for the expected integrated Bernoulli variance reduction achieved by the spatial sampling design.
-This result is first derived for the situation with a static design, and then extended to the adaptive situation, which provides new insight in our application with the autonomous vehicle.
+We show how  observation practices can gain efficiency and accuracy from the development of statistical techniques to achieve more effective spatial monitoring.
+We further demonstrate the opportunities available for real-time multivariable spatial data-gathering and analysis onboard autonomous sensing platforms, which statisticians can exploit to create general-purpose toolkits for similar applications (This, in turn, could enable the further development of theory and provide a framework for future experiments.)
+
+In particular, we derive and show results for characterizing phenomena connected to the properties of water masses.
+The characterization is based on joint excursion sets, whereby we derive new results for the expected integrated Bernoulli variance reduction achieved by the spatial sampling design.
+This result is first derived for the situation with a static design, and then extended to the adaptive situation; the sequential derivations  provide new insights into efficient applications of adaptive sampling, as demonstrated in our application of these models with an autonomous vehicle in Trondheim fjord.
 
 The study did not consider any temporal effects, which would be relevant on a larger time scale.
-We consider the extension to spatio-temporal modeling as future work, and we envision that advection-diffusion equations could be useful in this kind of modeling \citep{sigrist2015stochastic}.
-For more complex oceanographic phenomena, the methods will need to be extended to non-Gaussian phenomena, possibly feature-based mixtures of GPs which could still be run onboard, and augmented by dynamical models.
-
-The spatial statistical design criterion building on excursion sets is relevant in our setting with the different water mass properties and could also be useful in other oceanographic settings such as algae-blooms or frontal systems.
-Other criteria more directly related to decision making would also be relevant, especially with much spatio-temporal effects.
-For instance, managers must make difficult decisions related to fish farming or other marine resources operations, or environmental projects.
-Value of information analysis \citep{Eidsvik:15} would then be used in a similar vein as the integrated Bernoulli variance in the current paper, now studying which, when and where information is likely to materialize in improved decisions.
-
-While more effort could be spent on approximating the look-ahead steps, it is perhaps more interesting to explore the additional flexibility that can be gained by having two or more AUVs \citep{ferreira2019advancing} exploring a spatial or spatio-temporal domain together.
-It might then be useful to sample different parts of the space, and maybe move in parallel to best capture the excursion set.
+We consider the extension to spatio-temporal modeling as future work, and envision that advection-diffusion equations could be useful in this kind of modeling \citep{sigrist2015stochastic}.
+For more complex oceanographic phenomena, the methods will need to be extended to non-Gaussian phenomena, possibly feature-based mixtures of GPs which could still be run onboard and augmented by dynamical models.
+
+The spatial-statistical design criterion, building on excursion sets, is relevant in our setting with different water mass properties; this could also be useful in other oceanographic settings, such as algae-blooms or frontal systems.
+Of course, other criteria that incorporate spatio-temporal effects could be included in this approach, making it especially relevant for more long-term decision-making.
+For instance, managers and regulators must make difficult decisions related to fish farming, other marine resources operations, or environmental projects.
+Value of information analysis (VOI) \citep{Eidsvik:15} could be used in a similar vein as the integrated Bernoulli variance in the current paper, as both analytic methods explore which, when, and where information is likely to result in improved decision-making.
+
+While more effort could be spent on approximating the look-ahead steps, it is perhaps more interesting to explore the additional flexibility that can be gained by having two or more AUVs  exploring a spatial or spatio-temporal domain together \citep{ferreira2019advancing}.
+Such an approach would enable concurrent sampling in  different parts of the space, or opportunities to move in parallel to best capture the excursion set.
 
@@
 TOF and KR acknowledge support from the Centre for Autonomous Marine Operations and Systems (AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of Excellence, project number 223254, and the Applied Underwater Robotics Labortatory (AURLab).
-JE acknowledge support from Norwegian research council project 294404.
-DG acknowledges support from the Swiss National Science Foundation project number 178858.
+JE acknowledges support from Norwegian research council project 294404.
+DG acknowledges support from the Swiss National Science Foundation, project number 178858.
 

### 2019-09-30 main_new.tex
Updates Jo

@@
   for modern marine management and science. Faced with limited
-  resources to understand highly dynamic processes in the
+  resources to understand highly processes in the
   water-column, the combination of statistics and autonomous robotics
@@
 \kwd{Gaussian Processes}
-\kwd{Experimental design}
-\kwd{Adaptive information gathering}
+\kwd{Experimental Design}
+\kwd{Adaptive Information Gathering}
 \end{keyword}
@@
 Traditional data collection at sea has typically been based on static
-buoys, lagrangian floats, or ship-based methods, with significant
+buoys, Lagrangian floats, or ship-based methods, with significant
 logistical limitations that directly impact coverage and sampling
@@
 sensing the surface of the ocean, and are impacted by cloud cover. The
-advent of robust mobile robotic platforms \cite{Bellingham07} has
+advent of robust mobile robotic platforms \citep{Bellingham07} has
 resulted in significant contributions to environmental monitoring and
-sampling. In particular, autonomous underwater vehicles (AUVs), have made
-strides in advancing the state of sampling and consequently have made
+sampling. In particular, autonomous underwater vehicles (AUVs), have advanced the state of sampling and consequently have made
 robotics an integral part of ocean observation; our previous work
@@
 % using statistical methods in design and data assimilation
-\citep{das11b,Graham2013,das15,Das2015,fossum18b,fossuminformation}.
-Others working in this space include \cite{wikle2013modern}, focusing
+\citep{das11b,Das2015,fossum18b,fossuminformation}.
+Other statistical work in the oceanographic domain include \cite{wikle2013modern} focusing
 on hierarchical statistical models; \cite{sahu2008space}, studying
@@
 and \cite{mellucci2018oceanic} looking at the statistical prediction
-of features using an underwater glider. Common to these approaches is
-the use of information-based methods to infer environmental sampling
-and overcome common challenges, namely:
+of features using an underwater glider.
+
+The goal of the information-based methods that we highlight in the current paper, is to do efficient environmental sampling and overcome common challenges, namely:
 
@@
 \item Full numerical ocean models cannot provide accurate results online if run on robotic sensing
-  platforms, as on board computers cannot deliver the computational power required,
+  platforms, as onboard computers cannot deliver the computational power required,
   hence statistical proxy models of the environment must be used instead.
@@
   make measurements.
-\item As a-priori information about the state of the ocean is very
+\item As a priori information about the state of the ocean is rather
   uncertain, there is a substantial efficiency gain in reacting to
-  information obtained from measurements taken \emph{in-situ}; this is
-  usually referred to as the \emph{adaptivity gap}
-  \citep{Krause2008phd}.
+  information obtained from measurements taken in-situ. %\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.
 \item For such types of sampling problems, the number of choices (i.e.
@@
 this is the ability to model spatially-correlated variables and
-provide formal measures of uncertainty. In our formulation we used
+provide formal measures of uncertainty. In our formulation we use
 Gaussian Processes (GPs) in large part because they allow for
 efficient implementation and evaluation, in turn enabling
-prioritization of sampling efforts which can be be formalized into
-decision-making systems on board a robotic platform.
+prioritization of sampling efforts which can be formalized into decision-making systems onboard a robotic platform.
 
@@
 Bernoulli variance (IBV) in the ES indicators. Our main contributions
-include: i) extending ES-based uncertainty criteria to multivariate
+include: i) extending the IBV criteria to multivariate
 GPs, thereby allowing the evaluation to include multiple environmental
@@
 \ref{sec:heuristics} extends these results and present heuristics for
-adaptive sampling; Section \ref{sec:simulations} shows our results and
+adaptive sampling; Section \ref{sec:simulations}
 discusses properties of the methods in simulation studies. Section
@@
 Fisheries). This is a case in point of how adequate sampling efforts
-could have provided a more accurate assessment and prior
+could have provided a more accurate assessment and early
 warning of changing environmental conditions. These challenges related to undersampling have been driving
@@
 plumes one can often limit the scope to cover only lateral spatial
-elements. Moreover, in this case, time-effects can be regarded as
-static when acquiring AUV data over a short time window and limited
-region. However, the methodological framework can be extended to
+elements. However, the methodological framework can be extended to
 higher dimensional situations in a similar manner.
@@
 sensors and mission design. By deploying AUVs together with other
-resources, one can augment information obtained from ocean models,
+resources, one can augment information obtained from output from synthetic ocean models,
 remote-sensing data, or fixed-location sensors on buoys. Traditional
@@
 along fixed transects on a spatial grid, pre-programmed by the human
-operator. By instead using on-board algorithms to continuously
+operator. By instead using onboard algorithms to continuously
 evaluate, update, and refine future sampling locations, one can make
@@
 ES and EPs are useful starting points to measure the ability to classify water masses.
-The goal of experimental sampling is to improve this characterization, as quantified via changes in the EPs.
+The goal of experimental sampling is to improve this characterization.
+%, as quantified via changes in the EPs.
 
@@
 $\xi_T(\bx)$ denotes the temperature (in $^o C$) and $\xi_S(\bx)$
-denotes the salinity (in $mg/l$) at location $\bx$. The generalization to more than two variables is given in Section \ref{gen_k}.
+denotes the salinity (in $mg/l$) at location $\bx$.
+The generalization to more than two variables is given in Section \ref{gen_k}.
+The adaptive sampling strategies are described in Section \ref{sec:heuristics}.
 
@@
 challenging if the probability is close to $0.5$. 
-The IBV \cite{bect2019}, is defined as: 
+The IBV, see \cite{bect2019}, is defined as: 
 \begin{equation}\label{measVAR}
@@
 
-If prior knowledge provides a strong indication of cold freshwater
- in a part of the domain, it is easy to classify and
-there is likely little value in exploring those locations. Data can be
+Data can be
 gathered at selected locations of the domain $\mathcal{M}$, to improve
@@
 As the EPs provide a description of the boundary uncertainty, the strategies can effectively use this description to prioritize sampling efforts at locations that are ambiguous, making exploration more effective.
-As the AUV gathers data throughout its mission, it is natural to evaluate the expected reduction in the IBV, conditioned on the survey data.
-In this way, we can
-take the expectation with respect to the planned measurements $\by$
+It is natural to evaluate the expected reduction in the IBV, conditioned on the survey data.
+The expected updated IBV is
 \begin{equation}\label{sur}
-    V_{\mbox{upd}} = \int_{\bx \in \mathcal{M}} E_{\by} \left\{ p(\bx;\by)\left( 1-p(\bx;\by)\right) \right\} d\bx, 
+    V_{\mbox{upd}} = \int_{\bx \in \mathcal{M}} E_{\by} \left( p(\bx;\by)\left( 1-p(\bx;\by)\right) \right) d\bx, 
 \end{equation}
@@
 our work is similar to \cite{bect2012}, \cite{chevalier2014fast} and
-\cite{azzimonti2016quantifying} who describe analytical results for
-uncertainty reduction in ESs for univariate processes. A major
+\cite{azzimonti2016quantifying} who describe analytical results for the IBV for univariate processes. A major
 contribution of our work is to derive closed-form results for this
@@
 Calculations of the EPs and the IBV require a model specification.
-We use Gaussian modeling, which provides a practical and popular probabilistic approach to modeling environmental data, and allows a closed form
-evaluation of the IBVs as derived in Section \ref{sec:sur}.
-
-\subsection{Bivariate Gaussian Processes}
+We use GP models, which provides a practical and popular probabilistic approach to modeling environmental data, and allows a closed form
+evaluation of the IBVs as derived in Section \ref{sec:sur}. GPs can also be fitted to run onboard robotic vehicles with limited computational assets, such as AUVs.
+
+\subsection{Bivariate GPs}
 
@@
 specified by the mean vector and covariance matrix. For the application
-of tracking river plumes, the model parameters are specified from a short preliminary survey.
-The mean varies with location, capturing the variability near the river mouth, while the variance parameters and correlation between salinity and temperature are assumed to be constant for all locations and within each water mass.
+of a river plume, the model parameters are specified from a short preliminary survey.
+The means vary with location, capturing the variability near the river mouth, while the variance parameters and correlation between salinity and temperature are assumed to be constant for all locations.
+%and within each water mass.
 
 For the spatial dependence among variables, we assume a separable correlation function with the same decay for salinity and temperature,
-which is consistent with expectations for  a river plume where the decay is related to the common water masses (Sect \ref{sec:case_study}). In the resulting function, we have
+which is consistent with expectations for  a river plume where the decay is related to the common water masses (Section \ref{sec:case_study}). In the resulting function, we have
 \begin{equation}\label{gp_corr}
@@
 correlation function $\rho(\bx,\bx')$ solely depends on $\bx$ and
-$\bx'$ via a Euclidean distance measure $h=\sqrt{||\bx-\bx'||^2}$. With
+$\bx'$ via Euclidean distance $h=\sqrt{||\bx-\bx'||^2}$. With
 data and prior knowledge, one could possibly fit and estimate
@@
 $\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell has area $\Delta$. The
-grid is used for the waypoint graph setting the possible design locations. It is also used when solving integrals expressions in Eq. (\ref{measVAR}) and
+grid is used for the waypoint graph setting the possible design locations. It is also used when solving integral expressions in Eq. (\ref{measVAR}) and
 (\ref{sur}), which are then approximated by sums over all grid cells.  Using vector
@@
     \bxi = (\xi_T(\bx_1),\xi_S(\bx_1),\ldots,\xi_T(\bx_n),\xi_S(\bx_n))^t, \hspace{3mm}
-    \bxi  \sim  N(\bmu, \bSigma) %\pi (\bxi) & =& \frac{1}{(2\pi)^{n}|\bSigma|^{\frac{1}{2}}}e^{-\frac{1}{2}(\bxi -\bmu)^T\bSigma^{-1}(\bxi-\bmu)}.
+    \bxi  \sim  N(\bmu, \bSigma). %\pi (\bxi) & =& \frac{1}{(2\pi)^{n}|\bSigma|^{\frac{1}{2}}}e^{-\frac{1}{2}(\bxi -\bmu)^T\bSigma^{-1}(\bxi-\bmu)}.
 \end{equation}
@@
 
-\subsection{Data Model and Updating of Gaussian Processes}
-
-Data is gathered by an AUV that can measure temperature and salinity
+\subsection{Data Model and Updating of GPs}
+
+Data are gathered by an AUV that can measure temperature and salinity
 at a subset of locations on the defined grid. This can be done in a
@@
 updated model. Then the updated model is used to design and select the
-next round of measurements, and so on.
-Below, we describe the model for one round of data and the Gaussian formula for a single update; in Section \ref{sec:heuristics}, we discuss sequential data gathering in the context of adaptive design of AUV measurements.
+next round of measurements, and so on (Section \ref{sec:heuristics}). Below, we describe the model for one round of data and the Gaussian formula for a single update.
 
@@
 \end{equation}
-with entries $m_{\cdot}$ and $a_{\cdot}$ extracted from the conditional mean and covariance
+with entries extracted from the conditional mean and covariance
 expressions in Eq. (\ref{gp_upd}).
@@
 out that a simplified form of the calculation is possible because the
-conditional mean in Eq (\ref{gp_hat}) is a linear (affine) function of
+conditional mean in Eq. (\ref{gp_hat}) is a linear (affine) function of
 the data $\by$ and the covariance is not a function of the data.
@@
 
-\subsection{Excursion Probabilities for Gaussian Processes}
+\subsection{EPs for GPs}
 
 Based on Gaussian modeling assumptions, the EPs can be computed at any
-location $\bx$. Without additional data, Eq.  (\ref{eq:prob}) gives us
-
+location $\bx$. Without additional data, Eq.  (\ref{eq:prob}) gives
 \begin{eqnarray}\label{eq:prob_mv0}
@@
 When data $\by$ are available, Eq. (\ref{eq:post_ep}) similarly gives
-us
 \begin{eqnarray}\label{eq:prob_mv}
@@
 
-\subsection{Expected Integrated Bernoulli Variance for Bivariate Models}
-
-We will next derive closed-form solutions for the expectation in
-(\ref{sur}). The location index $\bx$ is suppressed to avoid overly
-complex notation. We present new results for computing the expectation
-$E_{\by}\left\{ \cdot \right\}$ which is the integrand of
+\subsection{Expected IBV for Bivariate Models}
+
+We will next derive closed-form solutions for the expectation in Eq. (\ref{sur}). The location index $\bx$ is suppressed to avoid overly
+complex notation. We present new results for computing the expectation in the integral of
 Eq. (\ref{sur}).
@@
 variables in Eq. (\ref{sur}) as   $Z_T=\frac{\xi_T-m_{T}}{a_{T}}$,
-$Z_S=\frac{\xi_S-m_{S}}{a_{S}}$; we can then rephrase the probability as
+$Z_S=\frac{\xi_S-m_{S}}{a_{S}}$, with $\mbox{Corr}(Z_T,Z_S)=a_{T,S}/(a_{T} a_{S})=\eta_{T,S}$, and we can then rephrase the probability as
 \begin{eqnarray}\label{ytom}
@@
 \end{bmatrix}\end{pmatrix} \nonumber \\
-&=& P(\xi_T \leq t_T, \xi_S \leq t_S|\bm),
+&=& P(\xi_T \leq t_T, \xi_S \leq t_S|\bm).
 \end{eqnarray}
-where $\eta_{T,S} =a_{T,S}/(a_{T} a_{S})$.
 Since the expression only depends on $\by$ via its affine function $\bm$, the expectation in Eq. (\ref{sur}) is reduced to a bivariate integral over $\bm$. 
@@
 E_{\by}(p(1-p)) &=& E_{\bm}(p(1-p)), \hspace{3mm} p=P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) \nonumber \\
-E_{\bm}(p(1-p)) & = & \int p[1-p] \pi(\bm) d\bm, \nonumber \\
+E_{\bm}(p(1-p)) & = & \int p(1-p) \pi(\bm) d\bm, \nonumber \\
  &=& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm)  \pi(\bm) d\bm \nonumber  \\
@@
 P \left( Z_{T} \leq \frac{t_T-m_{T}}{a_{T}}, 
-Z_{S} \leq \frac{t_S-m_{S}}{a_{S}} \right), \nonumber
-\end{equation}
-where $\bZ=(Z_{T},Z_{S})$ is a bivariate zero-mean and unit-variance
-Gaussian vector, with $\mbox{Corr}(Z_{T},Z_{S})=\eta_{T,S}$. Moreover,
-$\bZ$ is chosen to be independent of $\bm$. Using Eq. (\ref{dist_mxi}), we have
+Z_{S} \leq \frac{t_S-m_{S}}{a_{S}} \right). \nonumber
+\end{equation}
+The standardized variable $\bZ=(Z_{T},Z_{S})$ is chosen to be independent of $\bm$. Using Eq. (\ref{dist_mxi}), we have
 \begin{equation}
@@
 Eq. (\ref{part1:phi4}) are computed for each location on the grid and
-summed over $\bx \in \mathcal{M}_g$, which gives the design evaluation
+summed over $\bx \in \mathcal{M}_g$, to get the design evaluation
 criterion in Eq. (\ref{sur}).
@@
 
-Table \ref{tab:sim_rhoab} shows the EPs and the associated Bernoulli
+Table \ref{tab:sim_rhoab} shows the EPs and the associated prior Bernoulli
 variance (second row) for the examples indicated in
@@
 dimension. This in turn will impact the design criteria when we
-evaluate the information measure by integrating over several locations
-$\bx$.
+evaluate the information measure by integrating over several locations (Section \ref{sec:case_study}).
 
@@
 
-An important concept underlying this work is the potential of robotic sampling to adapt survey plans based on data and statistical inference
+An important concept underlying this work is the potential of robotic sampling to adapt survey plans based on data and statistical methods
 in accordance with the sense-plan-act loop
 (Fig. \ref{fig:envir1}). We will therefore present a few adaptive strategies
-for robotic sampling that build on the notion of multivariate ES. Simulations, analysis, and  results from full-scale trials are presented.
+for robotic sampling that build on the notion of multivariate ES.
 
@@
 
-An adaptive AUV survey is split into many rounds (stages). In each
-round, the survey design is created using a myopic approach that considers the largest reduction in uncertainty of the ES. In our
-setting, the selection is made on the graph and defined by the nearest grid
+An adaptive AUV survey is split into many rounds (stages). The selection is made on the graph and defined by the nearest grid
 nodes in the domain $\mathcal{M}_g$.
@@
 selected, data $\by^{j,s}$ will be gathered. The optimal path
-selection situation is  depicted in Figure \ref{fig:PathSelOpt},
+selection situation is  depicted in Fig. \ref{fig:PathSelOpt},
 \begin{figure}[h!]
@@
 $\mathcal{Y}^{s-1}$, we fit an updated Gaussian model from
-Eq. (\ref{gp_upd}) with mean $\bm_{s-1}=E(\bxi|\mathcal{Y}^{s-1})$
-and covariance matrix $\bA_{s-1}=\mbox{Var}(\bxi|\mathcal{Y}^{s-1})$.
+Eq. (\ref{gp_upd}) with mean $\bm^{s-1}=E(\bxi|\mathcal{Y}^{s-1})$
+and covariance matrix $\bA^{s-1}=\mbox{Var}(\bxi|\mathcal{Y}^{s-1})$.
 The next round of measurements $\by^{j,s}$ can be gathered with
-design $d^{j,s}$, $j=1,\ldots,J$, where $J$ indicates all possible AUV
-trajectories from the current stage. The {\it{naive}} strategy, then,
+design $d^{j,s}$, $j=1,\ldots,J$. The {\it{naive}} strategy then
 selects the design according to:
@@
 
-Based on the currently available data $\mathcal{Y}^{s-1}$, we again
-fit an updated Gaussian model, represented on the grid locations
-covering the spatial domain.  The next round of measurements
-$\by^{j,s}$, can be gathered with design $d^{j,s}$, $j=1,\ldots,J$,
-indicating all possible AUV trajectories at the current stage. The
+Based on the currently available data $\mathcal{Y}^{s-1}$, we
+fit an updated GP model.  The next round of measurements
+$\by^{j,s}$, can be gathered with design $d^{j,s}$, $j=1,\ldots,J$. The
 selected design is then:
@@
     d^{*,s} &=& \mbox{argmin}_{j \in \{1,\ldots,J\}} \left\{ V_{m,\mbox{upd,j}} \right\},  \\
-V_{m,\mbox{upd}} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j,s}|\mathcal{Y}^{s-1}} \left\{ p(\bx;\mathcal{Y}^{j,s})\left( 1-p(\bx;\mathcal{Y}^{j,s})\right) \right\} \Delta, \nonumber \\
+V_{m,\mbox{upd}} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j,s}|\mathcal{Y}^{s-1}} \left( p(\bx;\mathcal{Y}^{j,s})\left( 1-p(\bx;\mathcal{Y}^{j,s})\right) \right) \Delta, \nonumber \\
     p(\bx;\mathcal{Y}^{j,s}) &=& P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\by^{j,s},\mathcal{Y}^{s-1}). \nonumber
@@
 updated GP model from step $s-1$.  Once the data is collected for the
-best design, the GP model is updated again. The mean $\bm_{s}$ and
-covariance matrix $\bA_{s}$ are used to compute the next design at
+best design, the GP model is updated again. The mean $\bm^{s}$ and
+covariance matrix $\bA^{s}$ are used to compute the next design at
 stage $s+1$, and so on.
@@
 reasonable approach for creating designs in many
-applications. Moreover, it is easily implemented on-board an AUV,
+applications. Moreover, it is easily implemented onboard an AUV,
 using an efficient approach for data updating the GP model and the
@@
 
-Because of the intermixed optimization and expectation, there is no
-longer a closed form for the required integrals. Instead, we solve the
+We solve the
 first expectation by Monte Carlo sampling of data $\by^{j,s}$ from its
-conditional distribution. For each data sample, the second expectation
+conditional distribution. For each of these data samples, the second expectation
 is solved using the closed-form expressions for expected IBV from
@@
 the current best design. When data are collected, the GP model is
-updated, and the mean $\bm_{s}$ and covariance matrix $\bA_{s}$ are
+updated, and the mean $\bm^{s}$ and covariance matrix $\bA^{s}$ are
 used to compute the next design at stage $s+1$, now anticipating what
@@
 compute all possible branches of the first two stages, as they are
-indicated in Figure \ref{fig:PathSelOpt}. Instead, we use the myopic
+indicated in Fig. \ref{fig:PathSelOpt}. Instead, we use the myopic
 strategy to rank the three best designs following the first stage alone, and
@@
 
-We now show results from tests performed to study the properties of
-different static and sequential survey designs, in a realistic
-simulated case: mapping a freshwater plume defined by a
-temperature and salinity gradient using an AUV.
+We now study the properties of
+the different static and sequential survey designs in a realistic
+simulated case that emulates the spatial properties of a river plume interacting with coastal water.
 
@@
 \end{bmatrix} 
-= \begin{bmatrix} \beta_{T,0} + \beta_{T,1} x_{\mbox{East}} \\
-\beta_{T,1} + \beta_{S,1} x_{\mbox{East}}
+= \begin{bmatrix} \beta_{T,0} + \beta_{T,1} x_{\mbox{West}} \\
+\beta_{T,1} + \beta_{S,1} x_{\mbox{West}}
 \end{bmatrix}.
 \end{equation}
-In this simulation setup, which is motivated by a real case, we expect
-the eastern part of the domain to be cold freshwater. This situation
+This situation
 mimics that of a river mouth opening from the south, with the water
-masses pulled to the east. The actual trends are, in practice, set
-from preliminary information or runs of oceanographic models \cite{fossuminformation} solving differential equations for hydrodynamic flow. In this
-example, the mean is specified by regression parameters $\beta_{T,1}=0.065$ and
+masses pulled to the east. The regression parameters are  $\beta_{T,1}=0.065$ and
 $\beta_{S,1}=0.1$ for the slope, with west coordinate and $\beta_{T,0}=5.8$
@@
 0.6  & 1
-\end{bmatrix}(1+\phi |\bx-\bx'|)\exp (-\phi |\bx-\bx'|),
-\end{equation}
-and $\phi=0.3$ indicates an effective correlation range of about $1200$ m. 
+\end{bmatrix}(1+\phi h)\exp (-\phi h),
+\end{equation}
+where $h=\sqrt{||\bx-\bx'||^2}$ and $\phi=0.3$ indicates an effective correlation range of about $1200$ m. 
 %Figure \ref{fig:stat_design} shows the contour lines for the EP for the reference. We notice the trend of increasing salinity and temperature to the west.  
-We study sensitivity to the parameter specification in the results
-below.  One realization of the random fields representing salinity and
-temperature is shown in Figures \ref{fig:true_temp} and
+One realization of the random fields representing salinity and
+temperature is shown in Fig. \ref{fig:true_temp} and
 \ref{fig:true_sal}. The true ES from this prior realization is shown
-in Figure \ref{fig:ESet}.
+in Fig. \ref{fig:ESet}.
 
@@
   \caption{\ref{fig:true_temp} and \ref{fig:true_sal} show
-    realisations of the simulated temperature and salinity fields, as
-    well as the associated excursion set \ref{fig:ESet}.
-    \ref{fig:Eprob} shows the estimated excursion probabilities after
-    performing data collection along the N-S survey line.} 
+    a realisation of the simulated temperature and salinity fields. The associated ES is in \ref{fig:ESet}, while 
+    \ref{fig:Eprob} shows the estimated EPs after
+    performing data collection along the static N-S survey line.} 
 \label{fig:realisations}
@@
 
-We compare three different static designs denoting
+We compare three different static designs denoted
 \textit{static\_north}, \textit{static\_east}, and
 \textit{static\_zigzag} with the three sequential approaches
-\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
+\textit{naive}, \textit{myopic}, and \textit{look-ahead}. Recall that the static
 sampling paths are pre-determined and cannot be changed on the basis
-of data, representing the pre-planned observation
-strategies used in typical AUV operational survey designs.
-
-For a fixed survey length, the expected variance is analytically available (equal coverage and reduction in variance each time).
-However, for the sequential approaches this is not the case, as the result and path are dependent on the observed data and the expected reduction in IBV.
-The properties are therefore evaluated using Monte Carlo sampling over several replicates of realizations from the prior, conducting simulated sequential surveying for each one.
-Figure \ref{fig:Eprob} shows the conditional EP, given data gathered along the north-south survey line for the realization indicated in these displays.
-In the Monte Carlo replicates, such results are repeatedly computed to approximate the expected variance reduction in the ES.
+of data. They are representing the pre-planned observation
+strategies used in current-day AUV operational survey designs.
+
+For a fixed survey length, the expected IBV is analytically available.
+However, for the sequential approaches this is not the case.
+For comparison the properties are therefore evaluated using Monte Carlo sampling over several replicates of realizations from the prior, conducting simulated sequential surveying for each one.
+Fig.
+\ref{fig:Eprob} shows the conditional EP, given data gathered along the north-south survey line for the realization indicated in these displays.
+In the Monte Carlo replicates, such results are repeatedly computed to approximate the expected IBV.
 We also compare predictive performance measured by root mean square error (RMSE) for the temperature and salinity estimates and the variance reduction in these two variables.
-It is important to note that the objective function used by the agents is focused on reducing the expected variance in the ES, but we nevertheless expect that we will achieve good predictive performance for criteria such as RMSE as well.
+It is important to note that the objective function used by the agents is focused on reducing the expected IBV, but we nevertheless expect that we will achieve good predictive performance for criteria such as RMSE as well.
 Another non-statistical criteria that is important for practical purposes is the computational time of the strategy, as this will impact the performance of the approach in the field for an embedded system.
@@
 \ref{fig:wp_graph}.
-The sequential sampling agents start at the center East-West coordinate at the southern end of the domain (node 53).
+The sequential sampling agent starts at the center East-West coordinate at the southern end of the domain (node 53).
 The AUV will move along edges in the waypoint graph while the vehicle gathers measurements.
-The data is assimilated into the GP model before an evaluation of the next node to sample is conducted at the end of the edge.
-The paths will differ for the various sequential designs due to the individual design criteria and the simulated variability in the environment among the replicates.
+The data are assimilated into the GP model before an evaluation of the next node to sample is conducted at the end of the edge.
 
 \subsection{Simulation Results}
-A total of 100 replicate simulations were conducted; the results are shown in Figure \ref{fig:sim_results},
+A total of 100 replicate simulations were conducted. For each simulation all strategies are run. The results are shown in Fig. \ref{fig:sim_results},
 where the different criteria are plotted as a function of survey distance.
-Figure \ref{fig:avg_ev} shows the resulting drop in ES variance for each of the six strategies.
-The ES variance is reduced most under the \textit{myopic} and \textit{look-ahead} strategies, each performing almost equally; this is expected as the two criteria (Eq.
+Fig.
+\ref{fig:avg_ev} shows the resulting drop in IBV for each of the six strategies.
+The IBV is reduced most under the \textit{myopic} and \textit{look-ahead} strategies, each performing almost equally; this is expected as the two criteria (Eq.
 \eqref{critSEQ} and \eqref{critLA}) are much more sensitive to differences in IBV.
-The north-south design also does well in terms of ES variance reduction since the path is parallel to the  boundary between the water masses.
+The \textit{static\_north} design also does well here because the path is parallel to the  boundary between the water masses.
 
@@
   % \subfigure[Excursion set variance $E_{\by}(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
-  \subfigure[Excursion set $E_{\by}$(p\lbrack 1-p\rbrack) variance.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
+  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
   \hfill
@@
 \caption{Simulation results from 100 replicate simulations for 10
-  sampling choices/steps on the grid. (Updated:01.05.2019)} 
+  sampling choices/steps on the grid.} 
 \label{fig:sim_results}
@@
 
-Figure \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop in RMSE and increase in explained variance, respectively.
+Fig.
+\ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop in RMSE and increase in explained variance, respectively.
 Both \textit{myopic} and \textit{look-ahead} strategies perform well here, but some of the \textit{static\_east} and \textit{static\_zigzag} also achieve good results because they are pre-determined to cover large parts of the domain without re-visitation.
-Sequential strategies targeting ES variance will sometimes not reach similar coverage, as interesting data may draw the path into twists and turns that may result in re-visitation.
+Sequential strategies targeting IBV will sometimes not reach similar coverage, as interesting data may draw the path into twists and turns.
+There is relatively large variety in the replicate results as indicated by the vertical lines.
+Still, the ordering of strategies is similar.
 
@@
 Then again, the inclusion of heuristics is likely a contributing factor to the \textit{look-ahead} strategy failing to outperform the \textit{myopic} strategy.
-In Figure \ref{fig:route_choices}, the realized sampling paths for each of the sequential schemes and static designs is shown.
-
-In this model, the \textit{naive} strategy often gets stuck in the southern part of the domain because it is too focused on the probabilities near $0.5$.
-The \textit{myopic} strategy, meanwhile, covers a wider domain than the naive or look-ahead.
-There are several reasons for this; first, a greedy approach will tend to do more put more emphasis on locations close to the agent, which may lead away from the centre.
-Second, as the agent evaluates the impact of locations further away (look-ahead) where assimilated data has less predictive power, the GP prior (which is centered here) will act to restrict paths deviating from the central zone.
+
+In Fig.
+\ref{fig:route_choices}, the realized sampling paths for each of the sequential schemes and static designs are shown.
+The \textit{naive} strategy often gets stuck in the southern part of the domain because it is too focused on the probabilities near $0.5$.
+The \textit{myopic} strategy covers a wider domain than the naive or look-ahead.
+There are several reasons for this; first, a greedy approach will tend to put more emphasis on locations close to the agent, which may lead away from the centre.
+Second, as the agent evaluates the impact of locations further away (look-ahead) where assimilated data has less predictive power, the GP prior model (which is centered here) will act to restrict paths deviating from the central zone.
+
+%We studied the sensitivity of the results by modifying the input parameters to different prior correlations between temperature and salinity, standard deviations, and spatial correlation range.
+In all runs,  the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of IBV, and much better than \textit{naive}.
+The \textit{look-ahead} strategy seems to be substantially better than the \textit{myopic} only for very small prior standard deviation or very large spatial correlation range.
+The \textit{static\_north} design continues to be the best static design for IBV, while \textit{static\_zigzag} is the best design for the other predictive performance measures, especially so with large spatial correlation range.
+We also ran simulation studies with only temperature data, and for realistic correlation levels between temperature and salinity the effects are not much worse in terms of IBV when only temperature data are available.
+In addition to the comparison made in Table \ref{tab:sim_rhoab}, the current setting includes spatial correlation and this likely reduces the additional influence of having bivariate data.
+However, it seems that having temperature data alone does a substantially worse job in terms of explained variance
 
 \subsection{Performance Metrics}
-The tables below explore the sensitivity of the result to different prior correlation values between temperature measured on-board the AUV, temperature and salinity, prior standard deviations, and spatial correlations (see table rows).
+The tables below explore the sensitivity of the result to different prior correlation values between temperature measured onboard the AUV, temperature and salinity, prior standard deviations, and spatial correlations (see table rows).
 We only compare results at the end of each survey.
@@
 The experiment was performed in Trondheim, Norway,
-surveying the Nidelva river (Figure \ref{fig:nidelven}).
-Here cold and fresh water from the river meet the warm and saline ocean water in the Trondheim fjord.
-The experiment is focused along the eastern shoreline due to the aforementioned Coriolis force.
-The frontal zone therefore runs more or less parallel to the eastern shore as
-indicated in purple in Figure \ref{fig:nidelven}.
+surveying the Nidelva river (Fig.
+\ref{fig:nidelven}).
 The experiments were conducted in late Spring 2019, when there is still snow melting in the surrounding mountains so that the river water is substantially colder than the water in the fjord.
+The experiment is focused along the frontal zone that runs more or less parallel to the eastern shore as
+indicated in purple in Fig. \ref{fig:nidelven}.
 
@@
 Based on the initial data, the trend parameters were estimated by linear regression, where both temperature and salinity are assumed to increase linearly, going west from the river mouth.
-Next, the residuals from the regression analysis were analyzed to study the fit of the Gaussian model and to specify the covariance parameters.
+Next, the residuals from the regression analysis were analyzed to study the fit of the GP model and to specify the covariance parameters.
 
@@
   AUV. Left: Residual plot of temperature and salinity along with
-  Gaussian contours. Middle: Empirical CDF of the quadratic form of
-  the residuals along with the theoretical CDF of the $\chi^2$
+  Gaussian contours. Middle: Empirical CDF (solid) of the quadratic form of
+  the residuals along with the theoretical CDF (dashed) of the $\chi^2$
   distribution with two degrees of freedom. Right: Empirical variogram
@@
 
-Figure \ref{fig:parest} shows diagnostic plots of this analysis.
+Fig.
+\ref{fig:parest} shows diagnostic plots of this analysis.
 The display on the left shows a cross-plot of temperature and salinity residuals after the westernly trends in both salinity and temperature are subtracted from the data.
@@
 Based on the fitted bivariate Gaussian model (ellipses in left display), we can compute the modeled quadratic form of the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed.
-Figure \ref{fig:parest} (middle) shows the empirical cumulative distribution function (CDF) of these quadratic forms (solid) together with the theoretical CDF of the $\chi^2_2$ distribution.
+Fig.
+\ref{fig:parest} (middle) shows the empirical cumulative distribution function (CDF) of these quadratic forms (solid) together with the theoretical CDF of the $\chi^2_2$ distribution.
 The modeled and theoretical curves are very similar, which indicates that the Gaussian model fits reasonably well.
-Even though there appears to be some clustering in both left and middle displays of Figure \ref{fig:parest}, the bivariate diagnostic plots look reasonable and justify a Gaussian model.
-Figure \ref{fig:parest} (right) shows the empirical variogram of the scaled residuals for temperature and salinity.
+Even though there appears to be some clustering in both left and middle displays of Fig.
+\ref{fig:parest}, the bivariate diagnostic plots look reasonable and justify a Gaussian model.
+Fig.
+\ref{fig:parest} (right) shows the empirical variogram of the scaled residuals for temperature and salinity.
 The decay is similar for the two, and seems to be negligible after about $150$ m.
@@
 \ref{fig:map}.
-The robotic platform consisted of a Light AUV \citep{sousa2012lauv} (Figure \ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49 conductivity, temperature, and depth (CTD) sensor providing temperature and salinity measurements.
-The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
+The robotic platform consisted of a Light AUV \citep{sousa2012lauv} (Fig.
+\ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49 conductivity, temperature, and depth (CTD) sensor providing temperature and salinity measurements.
+%The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
 The sampling agent was built on top of the autonomous agent framework Teleo-Reactive EXecutive (\textit{T-REX}) \citep{py10,Rajan12,Rajan12b}, running an instance of the \textit{myopic} strategy from Section \ref{sec:sampling_designs} to control the AUV and decide between sampling locations.
@@
 The AUV was set to start in the south-center part of the waypoint graph, with the previously outlined GP model of the environment.
-Each AUV survey was set to take approximately 40 minutes to visit 15 waypoints on the grid, with the AUV running near the surface to better capture the plume.
+A survey was set to take approximately 40 minutes to visit 15 waypoints on the grid, with the AUV running near the surface to better capture the plume.
 On its path from one waypoint to the next, the AUV gathered data regularly, and the GP model assimilated temperature and salinity data with an update frequency of 30 seconds, with about three updates per stage.
@@
 We show how  observation practices can gain efficiency and accuracy from the development of statistical techniques to achieve more effective spatial monitoring.
-We further demonstrate the opportunities available for real-time multivariable spatial data-gathering and analysis onboard autonomous sensing platforms, which statisticians can exploit to create general-purpose toolkits for similar applications (This, in turn, could enable the further development of theory and provide a framework for future experiments.)
+We further demonstrate the opportunities available for real-time multivariable spatial data-gathering and analysis onboard autonomous sensing platforms, which statisticians can exploit to create general-purpose toolkits for similar applications.
 
 In particular, we derive and show results for characterizing phenomena connected to the properties of water masses.
-The characterization is based on joint excursion sets, whereby we derive new results for the expected integrated Bernoulli variance reduction achieved by the spatial sampling design.
-This result is first derived for the situation with a static design, and then extended to the adaptive situation; the sequential derivations  provide new insights into efficient applications of adaptive sampling, as demonstrated in our application of these models with an autonomous vehicle in Trondheim fjord.
+The characterization is based on joint ESs, whereby we derive new results for the expected IBV reduction achieved by the spatial sampling design.
+This result is first calculated in closed form for the situation with a static design, and then extended to the adaptive situation; the sequential derivations  provide new insights into efficient applications of adaptive sampling, as demonstrated in our application of these models with an autonomous vehicle in Trondheim fjord.
 
@@
 
-The spatial-statistical design criterion, building on excursion sets, is relevant in our setting with different water mass properties; this could also be useful in other oceanographic settings, such as algae-blooms or frontal systems.
-Of course, other criteria that incorporate spatio-temporal effects could be included in this approach, making it especially relevant for more long-term decision-making.
+The spatial-statistical design criterion, building on ESs, is relevant in our setting with different water mass properties; this could also be useful in other oceanographic settings, such as algae-blooms or frontal systems.
+Of course, other criteria are also relevant for decision-making.
 For instance, managers and regulators must make difficult decisions related to fish farming, other marine resources operations, or environmental projects.
-Value of information analysis (VOI) \citep{Eidsvik:15} could be used in a similar vein as the integrated Bernoulli variance in the current paper, as both analytic methods explore which, when, and where information is likely to result in improved decision-making.
-
-While more effort could be spent on approximating the look-ahead steps, it is perhaps more interesting to explore the additional flexibility that can be gained by having two or more AUVs  exploring a spatial or spatio-temporal domain together \citep{ferreira2019advancing}.
+Value of information analysis (VOI) \citep{Eidsvik:15} could be used in a similar vein as the IBV in the current paper, as both analytic methods explore when, where and which information is likely to result in improved decision-making.
+
+While more effort could be spent on approximating the look-ahead steps in the adaptive designs, it is perhaps more interesting to explore the additional flexibility that can be gained by having two or more AUVs  exploring a spatial or spatio-temporal domain together \citep{ferreira2019advancing}.
 Such an approach would enable concurrent sampling in  different parts of the space, or opportunities to move in parallel to best capture the excursion set.

### 2019-09-30 main_new.tex
Trygve updates

@@
 \definecolor{Gray}{gray}{0.9}
-\newcommand{\kcomment}[1]{{\color{red}{\{KR: #1\}}}}
-\newcommand{\tofcomment}[1]{{\color{green}{\{TF: #1\}}}}
-\newcommand{\kc}{\kcomment}
+
 % put your definitions there:
@@
 \begin{aug}
-\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{a,b}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
-\author{\fnms{Jo} \snm{Eidsvik}\thanksref{c}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
-\author{\fnms{David} \snm{Ginsbourger}\thanksref{d}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
+\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
+\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t3}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
+\author{\fnms{David} \snm{Ginsbourger}\thanksref{t4}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
 \and
-\author{\fnms{Kanna} \snm{Rajan}\thanksref{b,e,f}. \ead[label=e4]{kanna.rajan@ntnu.no}}
-
-\address{Department of Marine Technology\\ Trondheim, Norway\printead{e1}}
-\affiliation[a]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
-\affiliation[b]{Centre for Autonomous Marine Operations and Systems (AMOS).}
-
-\address{Department of Mathematical Sciences \printead{e2}}
-\affiliation[c]{Department of Mathematical Sciences, NTNU.}
-
-\address{Idiap \printead{e3}}
-\affiliation[d]{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland.}
-
-\address{Department of Engineering Cybernetics,\printead{e4}}
-\affiliation[e]{Department of Engineering Cybernetics, NTNU}
-\affiliation[f]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
+\author{\fnms{Kanna} \snm{Rajan}\thanksref{t2,t5,t6}. \ead[label=e4]{kanna.rajan@ntnu.no}}
+
+\affiliation[t1]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
+\affiliation[t2]{Centre for Autonomous Marine Operations and Systems (AMOS).}
+\affiliation[t3]{Department of Mathematical Sciences, NTNU.}
+\affiliation[t4]{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland.}
+\affiliation[t5]{Department of Engineering Cybernetics, NTNU}
+\affiliation[t6]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
+
+\address{ \\Department of Marine Technology,\\ Trondheim, Norway \\
+\printead{e1}}
+\address{Department of Mathematical Sciences,\\ Trondheim, Norway\\ \printead{e2}}
+\address{Idiap, \\ Martigny, Switzerland, \\
+\printead{e3}}
+\address{Department of Engineering Cybernetics,\\ Trondheim, Norway \\
+\printead{e4}}
 
@@
 
-The goal of the information-based methods that we highlight in the current paper, is to do efficient environmental sampling and overcome common challenges, namely:
+The potential that lies in using information-based methods, which we highlight in the current paper, is to do efficient environmental sampling and overcome three common challenges, namely:
 
 \begin{itemize}
-\item Full numerical ocean models cannot provide accurate results online if run on robotic sensing
-  platforms, as onboard computers cannot deliver the computational power required,
-  hence statistical proxy models of the environment must be used instead.
-\item Decisions on where to sample happen sequentially, i.e. recently
-  acquired information can be assimilated into and used to improve
-  statistical models and reduce uncertainty associated with where to
-  make measurements.
-\item As a priori information about the state of the ocean is rather
-  uncertain, there is a substantial efficiency gain in reacting to
-  information obtained from measurements taken in-situ. %\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.
-\item For such types of sampling problems, the number of choices (i.e.
-  locations, trajectories, and candidate designs) makes the problem
-  combinatorially large, creating a trade-off between optimization
-  (finding the most resource-efficient design to collect necessary
-  data), and computability (arriving at a solution in reasonable
-  time). To successfully resolve features, this trade-off has to be
-  considered in sample design and practice.
+\item Full numerical ocean models cannot provide accurate results online if run on robotic sensing platforms, as onboard computers cannot deliver the computational power required, hence statistical proxy models of the environment must be used for doing inference about where to sample.
+\item As a priori information about the state of the ocean is uncertain, there is a substantial efficiency gain in reacting to information obtained from measurements taken in-situ.
+Recently acquired information can be assimilated into statistical models that is used to inform decisions on where to sample sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.
+\item For sampling problems related to environmental sensing, the number of choices (i.e.
+locations, trajectories, and candidate designs) makes the problem combinatorially large, creating a trade-off between optimization (finding the most resource-efficient design to collect necessary data), and computability (arriving at a solution in reasonable time).
+To successfully resolve features, this trade-off has to be considered in sample design and practice.
 \end{itemize}
@@
 distributed use of robotic assets in order to fill the gaps in our
-understanding (See Fig. \ref{fig:envir}), where observation is
+understanding (See Fig. \ref{fig:envir1}), where observation is
 guided by assimilation of measurements and statistical models.
@@
   \centering 
-  \subfigure[Illustration of an ocean sensing network, where robotic
-  assets augment ship-based measurements to increase coverage and
-  resolution of phenomena of scientific interest.]{\includegraphics[width =
+  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
     0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
@@
     0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
-  \caption{Ocean observation is moving away from single-ship sampling
-    towards more autonomous and collaborative networked operations in
-    order to resolve the numerous bio-geochemical processes and their
-    interactions.}
+  \caption{\ref{fig:envir1} Ocean observation is moving away from single-ship sampling towards more autonomous and collaborative networked operations in order to resolve the numerous bio-geochemical processes and their interactions; this requires autonomy that can reason and make decisions on their own following the sense-plan-act control methodology.
+\ref{fig:nidelven} The interaction of river and ocean creates processes that are challenging to map, where the combination of robotics and statistics can play a vital role in enabling more effective observation practices.}
 \label{fig:envir}
@@
 \begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
+  \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.pdf}
   \caption{Density contour plots with different correlations between
@@
 
+%\begin{figure}[!h] 
+% \centering 
+%\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
+%\caption{Data analysis from a preliminary trial experiment using the
+%  AUV. Left: Residual plot of temperature and salinity along with
+%  Gaussian contours. Middle: Empirical CDF (solid) of the quadratic form of
+%  the residuals along with the theoretical CDF (dashed) of the $\chi^2$
+%  distribution with two degrees of freedom. Right: Empirical variogram
+%  of the salinity and temperature data.} \label{fig:parest}
+%\end{figure}
+
 \begin{figure}[!h]
   \centering
-\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
-\caption{Data analysis from a preliminary trial experiment using the
-  AUV. Left: Residual plot of temperature and salinity along with
-  Gaussian contours. Middle: Empirical CDF (solid) of the quadratic form of
-  the residuals along with the theoretical CDF (dashed) of the $\chi^2$
-  distribution with two degrees of freedom. Right: Empirical variogram
-  of the salinity and temperature data.} \label{fig:parest}
+  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_a.pdf}\label{fig:parest_a}}
+  \hfill
+  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_b.pdf}\label{fig:parest_b}}
+  \hfill
+  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_c.pdf}\label{fig:parest_c}}
+  \caption{Data analysis from a preliminary trial experiment using the AUV.
+\ref{fig:parest_a} Residual plot of temperature and salinity along with Gaussian contours.
+\ref{fig:parest_b} Empirical CDF (solid) of the quadratic form of the residuals along with the theoretical CDF (dashed) of the $\chi^2$  distribution with two degrees of freedom.
+\ref{fig:parest_c} Empirical variogram of the salinity and temperature data.}
+\label{fig:parest}
 \end{figure}
@@
 \ref{fig:parest} shows diagnostic plots of this analysis.
-The display on the left shows a cross-plot of temperature and salinity residuals after the westernly trends in both salinity and temperature are subtracted from the data.
+Fig.
+\ref{fig:parest_a} shows a cross-plot of temperature and salinity residuals after the westernly trends in both salinity and temperature are subtracted from the data.
 This scatterplot of joint residuals indicates larger variability in salinity than in temperature, and a positive correlation ($0.5$) between the two variables.
-Based on the fitted bivariate Gaussian model (ellipses in left display), we can compute the modeled quadratic form of the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed.
+Based on the fitted bivariate Gaussian model (ellipses in Fig.
+\ref{fig:parest_a}), we can compute the modeled quadratic form of the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed.
 Fig.
-\ref{fig:parest} (middle) shows the empirical cumulative distribution function (CDF) of these quadratic forms (solid) together with the theoretical CDF of the $\chi^2_2$ distribution.
+\ref{fig:parest_b} shows the empirical cumulative distribution function (CDF) of these quadratic forms (solid) together with the theoretical CDF of the $\chi^2_2$ distribution.
 The modeled and theoretical curves are very similar, which indicates that the Gaussian model fits reasonably well.
-Even though there appears to be some clustering in both left and middle displays of Fig.
-\ref{fig:parest}, the bivariate diagnostic plots look reasonable and justify a Gaussian model.
+Even though there appears to be some clustering in both Fig.
+\ref{fig:parest_a} and \ref{fig:parest_b}, the bivariate diagnostic plots look reasonable and justify a Gaussian model.
 Fig.
-\ref{fig:parest} (right) shows the empirical variogram of the scaled residuals for temperature and salinity.
+\ref{fig:parest_c} shows the empirical variogram of the scaled residuals for temperature and salinity.
 The decay is similar for the two, and seems to be negligible after about $150$ m.
@@
 
-\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv1_es_sal.pdf}\label{fig:res1}}
+\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv1_es_sal_ep.pdf}\label{fig:res1}}
 \hspace{0.2cm}
-\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv4_es_sal.pdf}\label{fig:res2}}
-
-\subfigure[ES for Survey 1]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
-\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
-\caption{Results from mapping the Nidelva river} 
+\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv4_es_sal_ep.pdf}\label{fig:res2}}
+
+%\subfigure[ES for Survey 1]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
+%\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
+\caption{Results from mapping the Nidelva river.
+\ref{fig:map} Show an overview of the survey area overlaid with the AUV path in black and dashed line.
+Note the shaded region indicating the usual frontal region.
+\ref{fig:res_both} Presents the collected temperature data as colored trails.
+Note waypoint 5 (WP5) which indicate where the two surveys deviate.
+\ref{fig:res1} and \ref{fig:res2} Shows the collected salinity data (as a colored trail) overlaid on the final EP, which indicate the AUVs statistical impression of the front.
+For both missions the temperature and salinity data correspond with the EP front indication.}
 \label{fig:results}
@@
 The final prediction of the front location given in Fig.
-\ref{fig:res3} and \ref{fig:res4}, correspond with one another, yielding a picture of the front gradually bending off toward the north east.
+\ref{fig:res1} and \ref{fig:res2} as a dashed line, correspond with one another, yielding a picture of the front gradually bending off toward the north east.
 The amount of exploration done by Survey 1 is greater than Survey 2.
@@
 \footnotesize
-\bibliographystyle{apalike}
+\bibliographystyle{imsart-nameyear}
 \bibliography{ref}

### 2019-10-01 main_new.tex
updates

@@
   Improving and optimizing oceanographic sampling is a crucial task
-  for modern marine management and science. Faced with limited
-  resources to understand highly processes in the
-  water-column, the combination of statistics and autonomous robotics
-  provides opportunities for studying aspects related to experimental
-  design for effective characterization of oceanographic phenomena. In
-  this work we develop methods for efficient spatial sampling applied
-  to the mapping of coastal processes by providing informative
-  descriptions of spatial characteristics of ocean
-  phenomena. Specifically, we define a design criterion based on
-  reducing uncertainty in the joint excursions of random fields, and
-  derive analytical expressions for the expected Bernoulli variance
-  reduction in the excursion set of spatial temperature and salinity
-  variables. We show how these can be applicable in static designs and
-  sequential designs, building on myopic and look-ahead sampling
-  strategies. We use simulations to study the properties and
-  differences between state-of-the-art approaches followed by results
-  from field deployments with an autonomous underwater vehicle as part
-  of a case study mapping the boundary of a river plume. The results
-  demonstrate the ability of the method to effectively inform and
-  execute data-driven environmental sampling.
+  for marine science and maritime management. Faced with limited
+  resources to understand dynamic processes in the water-column, the
+  combination of statistics and autonomous robotics can provide
+  opportunities for studying experimental design for effective
+  characterization of oceanographic phenomena. In this work we develop
+  methods for efficient spatial sampling applied to the mapping of
+  coastal processes by providing informative descriptions of spatial
+  characteristics of ocean phenomena. Specifically, we define a design
+  criterion based on reducing uncertainty in the joint excursions of
+  random fields, and derive analytical expressions for the expected
+  Bernoulli variance reduction in the excursion set of spatial
+  temperature and salinity variables. We show how these can be
+  applicable in static and sequential designs, building on myopic and
+  look-ahead sampling strategies. We use simulations to study the
+  properties and differences between state-of-the-art approaches
+  followed by results from field deployments with an autonomous
+  underwater vehicle as part of a case study mapping the boundary of a
+  river plume. The results demonstrate the ability of the methods to
+  effectively inform and execute data-driven environmental sampling.
 
@@
 large-scale coverage but have limited resolution, are limited to
-sensing the surface of the ocean, and are impacted by cloud cover. The
+sensing the surface of the ocean and are impacted by cloud cover. The
 advent of robust mobile robotic platforms \citep{Bellingham07} has
 resulted in significant contributions to environmental monitoring and
-sampling. In particular, autonomous underwater vehicles (AUVs), have advanced the state of sampling and consequently have made
-robotics an integral part of ocean observation; our previous work
-has contributed to this effort
+sampling. In particular, autonomous underwater vehicles (AUVs), have
+advanced the state of sampling and consequently have made robotics an
+integral part of ocean observation; our previous work has contributed
+to this effort
 % using statistical methods in design and data assimilation
-\citep{das11b,Das2015,fossum18b,fossuminformation}.
-Other statistical work in the oceanographic domain include \cite{wikle2013modern} focusing
-on hierarchical statistical models; \cite{sahu2008space}, studying
-spatio-temporal models for sea surface temperature and salinity data;
-and \cite{mellucci2018oceanic} looking at the statistical prediction
-of features using an underwater glider.
-
-The potential that lies in using information-based methods, which we highlight in the current paper, is to do efficient environmental sampling and overcome three common challenges, namely:
+\citep{das11b,Das2015,fossum18b,fossuminformation}.  Other statistical
+work in the oceanographic domain include \cite{wikle2013modern}
+focusing on hierarchical statistical models; \cite{sahu2008space},
+studying spatio-temporal models for sea surface temperature and
+salinity data; and \cite{mellucci2018oceanic} looking at the
+statistical prediction of features using an underwater glider.
+
+The potential that lies in using information-based methods, which we
+highlight in the paper, is to do efficient environmental sampling and
+overcome three common challenges, namely:
 
 \begin{itemize}
-\item Full numerical ocean models cannot provide accurate results online if run on robotic sensing platforms, as onboard computers cannot deliver the computational power required, hence statistical proxy models of the environment must be used for doing inference about where to sample.
-\item As a priori information about the state of the ocean is uncertain, there is a substantial efficiency gain in reacting to information obtained from measurements taken in-situ.
-Recently acquired information can be assimilated into statistical models that is used to inform decisions on where to sample sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.
-\item For sampling problems related to environmental sensing, the number of choices (i.e.
-locations, trajectories, and candidate designs) makes the problem combinatorially large, creating a trade-off between optimization (finding the most resource-efficient design to collect necessary data), and computability (arriving at a solution in reasonable time).
-To successfully resolve features, this trade-off has to be considered in sample design and practice.
+
+\item Full numerical ocean models cannot provide accurate results
+  online if run on robotic sensing platforms, as onboard computers
+  cannot deliver the computational power required, hence statistical
+  proxy models of the environment must be used for doing inference
+  about where to sample.
+
+\item As a priori information about the state of the ocean is
+  uncertain, there is a substantial efficiency gain in reacting to
+  information obtained from measurements taken in-situ. Recently
+  acquired information can be assimilated into statistical models that
+  can be used to inform decisions on where to sample
+  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.
+
+\item For sampling problems related to environmental sensing, the
+  number of choices (i.e. locations, trajectories, and candidate
+  designs) makes the problem combinatorially large, creating a
+  trade-off between optimization (finding the most resource-efficient
+  design to collect necessary data), and computability (arriving at a
+  solution in reasonable time). To successfully resolve features, this
+  trade-off has to be considered in sample design and the practice.
+
 \end{itemize}
@@
 Gaussian Processes (GPs) in large part because they allow for
-efficient implementation and evaluation, in turn enabling
-prioritization of sampling efforts which can be formalized into decision-making systems onboard a robotic platform.
+efficient implementation and evaluation onboard a robotic platform.
 
@@
 an agglomeration of physical forcing (wind, topography, bathymetry,
-tidal influences and local geography, such as fresh water inflows)
-and incipient micro-biology driven by planktonic and coastal
-anthropogenic input, such as pollution and agricultural runoff. These
-often result in a range of ecosystem-related phenomena such as blooms,
-anoxic zones and plumes, all of which have significant direct and indirect effects  on society. Walter Munk referred to the
-20\textsuperscript{th} century as the \emph{century of undersampling}
-\citep{munk2002}, pointing out a lack of sufficient sampling
-resolution in time and space as a critical issue for characterizing
-ocean phenomena. Even today, policymakers struggle with limited
-observability of severe environmental problems such as the recent outbreak of harmful algal blooms in Northern
-Norway\footnote{Report from the Institute of Marine
-  Research: https://bit.ly/2KZGJfI} (May 2018), with losses exceeding 12 million
-tons of salmon valued at \$80 million (Norwegian Directorate of
-Fisheries). This is a case in point of how adequate sampling efforts
-could have provided a more accurate assessment and early
-warning of changing environmental conditions. These challenges related to undersampling have been driving
-ocean sensing practices towards more autonomous, networked, and
-distributed use of robotic assets in order to fill the gaps in our
-understanding (See Fig. \ref{fig:envir1}), where observation is
-guided by assimilation of measurements and statistical models.
-
-\begin{figure}[!h] 
+tidal influences and local geography, such as fresh water inflows) and
+incipient micro-biology driven by planktonic and coastal anthropogenic
+input, such as pollution and agricultural runoff. These often result
+in a range of ecosystem-related phenomena such as blooms, anoxic zones
+and plumes, all of which have significant direct and indirect effects
+on society. Walter Munk referred to the 20\textsuperscript{th} century
+as the \emph{century of undersampling} \citep{munk2002}, pointing out
+a lack of sufficient sampling resolution in time and space as a
+critical issue for characterizing ocean phenomena. Even today,
+policymakers struggle with limited observability of severe
+environmental problems; a prominent example is the recent (May 2018)
+outbreak of harmful algal blooms in Northern Norway\footnote{Report
+  from the Institute of Marine Research: https://bit.ly/2KZGJfI}, with
+losses exceeding 12 million tons of salmon valued at \$80 million
+(Norwegian Directorate of Fisheries). This is a case in point of how
+adequate sampling efforts could have provided a more accurate
+assessment and early warning of changing environmental
+conditions. These challenges related to undersampling have been
+driving ocean sensing practices towards more autonomous, networked,
+and distributed use of robotic assets in order to fill the gaps in our
+understanding (See Fig. \ref{fig:envir1}), where observation is guided
+by assimilation of measurements and statistical models.
+
+\begin{figure}[!b] 
   \centering 
@@
     0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
-  \caption{\ref{fig:envir1} Ocean observation is moving away from single-ship sampling towards more autonomous and collaborative networked operations in order to resolve the numerous bio-geochemical processes and their interactions; this requires autonomy that can reason and make decisions on their own following the sense-plan-act control methodology.
-\ref{fig:nidelven} The interaction of river and ocean creates processes that are challenging to map, where the combination of robotics and statistics can play a vital role in enabling more effective observation practices.}
+  \caption{\ref{fig:envir1} Traditional ocean observation based on
+    ship-based sampling has recently been augmented by autonomous and
+    collaborative networked operations of robotic vehicles to resolve
+    numerous bio-geochemical processes and their interactions. The
+    sense-plan-act control approach to autonomous control can reason
+    and make decisions for efficient onboard adaptive sampling.
+    \ref{fig:nidelven} The interaction of river and ocean creates
+    processes that are challenging to map, where the combination of
+    statistics and robotics can play a vital role in enabling more
+    effective observation practices.}
 \label{fig:envir}
@@
 Our focus in this work is towards spatial characterization of a
-frontal system, more specifically, fronts generated by river
+frontal system specifically, fronts generated by river
 plumes. Fig. \ref{fig:nidelven} shows the survey area (the outlet of
 the Nidelva river, Trondheim, Norway) where cold freshwater enters
-from the river creating a strong gradient in both temperature and
+from the river, creating a strong gradient in both temperature and
 salinity. Because of the local topography and the Coreolis force
@@
 at smaller \emph{sub-mesoscale} (from $5 m^2$ -- $10 km^2$) where
-spatial lateral variability tends to dominate. Remote sensing and
-synthetic ocean models find it challenging to provide detail at this
-scale \citep{Lermusiaux:2006} and the principal way to resolve for
-fine resolution is by direct observation. At larger \emph{mesoscale}
-($>50 km^2$), both three-dimensional space and time dynamics are
-important and can shift substantially, while in the case of river
-plumes one can often limit the scope to cover only lateral spatial
-elements. However, the methodological framework can be extended to
-higher dimensional situations in a similar manner.
+lateral spatial variability tends to dominate. Remote sensing and
+synthetic ocean models often find it challenging to provide detail at
+this scale \citep{Lermusiaux:2006} and the principal way to resolve
+for fine resolution is by direct in-situ observation. At larger
+\emph{mesoscale} ($>50 km^2$), both three-dimensional space and time
+dynamics are important and can shift substantially, while in the case
+of river plumes one can often limit the scope to cover only lateral
+spatial elements. However, the methodological framework can be
+extended to higher dimensional situations in a similar manner.
 
@@
 sensors and mission design. By deploying AUVs together with other
-resources, one can augment information obtained from output from synthetic ocean models,
+resources, one can augment information from synthetic ocean models,
 remote-sensing data, or fixed-location sensors on buoys. Traditional
@@
 along fixed transects on a spatial grid, pre-programmed by the human
-operator. By instead using onboard algorithms to continuously
-evaluate, update, and refine future sampling locations, one can make
-information gathering adaptive and contextual to the survey
+operator. Instead using onboard algorithms to continuously evaluate,
+update, and refine future sampling locations, one can make information
+gathering \emph{adaptive} and \emph{contextual} to the survey
 requirements \citep{das11b,fossum18b,fossuminformation}. The space of
-sampling opportunities is still limited to the spatial grid, but now
-there is much more flexibility because of adaptation. For instance, an
-AUV could reconstruct or modify a survey line based on what
-temperatures it is measuring by using embedded methods in
-decision-theoretic planning \citep{py10,Rajan12,Rajan12b}.
+sampling opportunities is still limited to a grid, but adaptation
+allows more flexibility. For instance, an AUV could reconstruct or
+modify a survey line based on what temperatures it is measuring by
+using embedded methods in decision-theoretic planning
+\citep{py10,Rajan12,Rajan12b}.
 
@@
 based on closing the sensing and actuation loop
-(Fig. \ref{fig:envir1}). Myopic sampling—i.e. step-wise
-selection of the path (on the graph) which reduces the information
-criterion the most—has been used for adaptive surveys
+(Fig. \ref{fig:envir1}). Myopic sampling, i.e. step-wise selection of
+the path (on the graph) which reduces the information criterion the
+most, has been used for adaptive surveys
 \citep{singh2009efficient,Binney2013} that focus largely on reducing
@@
 of scientific interest in \cite{Low2009} and \cite{fossuminformation}.
-In this paper, with a focus on mapping the river plume, we reward
-the designs that improve the classification of water masses as a
-means to characterize the frontal zone.
+In this paper, with a focus on mapping the river plume, we reward the
+designs that improve the classification of water masses as a means to
+characterize the frontal zone.
 
@@
 
-Our objective is to leverage statistical tools to characterize a river plume, focusing on spatial separation of cold freshwater from the river and warmer saline waters of the fjord.
-ES and EPs are useful starting points to measure the ability to classify water masses.
-The goal of experimental sampling is to improve this characterization.
-%, as quantified via changes in the EPs.
+Our objective is to leverage statistical tools to characterize a river
+plume, focusing on spatial separation of cold freshwater from the
+river and warmer saline waters of the fjord. ES and EPs are useful
+starting points to measure the ability to classify water masses. The
+goal of experimental sampling is to improve this
+characterization. %, as quantified via changes in the EPs.
 
@@
 $\xi_T(\bx)$ denotes the temperature (in $^o C$) and $\xi_S(\bx)$
-denotes the salinity (in $mg/l$) at location $\bx$.
-The generalization to more than two variables is given in Section \ref{gen_k}.
-The adaptive sampling strategies are described in Section \ref{sec:heuristics}.
-
-We let $t_T$ be a threshold for temperature and $t_S$ a threshold for salinity. The joint ES is then defined by:
+denotes the salinity (in $mg/l$) at location $\bx$. The generalization
+to more than two variables is given in Section \ref{gen_k}. The
+adaptive sampling strategies are described in Section
+\ref{sec:heuristics}.
+
+We let $t_T$ be a threshold for temperature and $t_S$ a threshold for
+salinity. The joint ES is then defined by:
 \begin{equation}\label{ES}
@@
 
-Our goal is to construct adaptive sampling strategies based on continuous evaluation of the EPs.
-As the EPs provide a description of the boundary uncertainty, the strategies can effectively use this description to prioritize sampling efforts at locations that are ambiguous, making exploration more effective.
-It is natural to evaluate the expected reduction in the IBV, conditioned on the survey data.
-The expected updated IBV is
+Our goal is to construct adaptive sampling strategies based on
+continuous evaluation of the EPs. As the EPs provide a description of
+the boundary uncertainty, the strategies can effectively use this
+description to prioritize sampling efforts at locations that are
+ambiguous, making exploration more effective. It is natural to
+evaluate the expected reduction in the IBV, conditioned on the survey
+data. The expected updated IBV is:
+
 \begin{equation}\label{sur}
@@
 \end{equation}
-where the expectation is over the planned data, and the probability is defined in Eq. (\ref{eq:post_ep}).
-Considering a set $\mathcal{J}$ of possible designs for an AUV survey, a criterion for the selection is
+where the expectation is over the planned data, and the probability is
+defined in Eq. (\ref{eq:post_ep}).  Considering a set $\mathcal{J}$ of
+possible designs for an AUV survey, a criterion for the selection is:
 \begin{equation}\label{crit}
@@
 our work is similar to \cite{bect2012}, \cite{chevalier2014fast} and
-\cite{azzimonti2016quantifying} who describe analytical results for the IBV for univariate processes. A major
-contribution of our work is to derive closed-form results for this
-design criteria for situations where the underlying model is based on multivariate GPs.
+\cite{azzimonti2016quantifying} who describe analytical results for
+the IBV for univariate processes. A major contribution of our work is
+to derive closed-form results for this design criteria for situations
+where the underlying model is based on multivariate GPs.
 
@@
 
-Calculations of the EPs and the IBV require a model specification.
-We use GP models, which provides a practical and popular probabilistic approach to modeling environmental data, and allows a closed form
-evaluation of the IBVs as derived in Section \ref{sec:sur}. GPs can also be fitted to run onboard robotic vehicles with limited computational assets, such as AUVs.
+Calculations of the EPs and the IBV require a model specification. We
+use GP models, which provide a practical and popular probabilistic
+approach to modeling environmental data, and allows a closed form
+evaluation of the IBVs as derived in Section \ref{sec:sur}. GPs can
+also be fitted to run onboard robotic vehicles with limited
+computational assets, such as AUVs.
 
@@
 
-The Gaussian model is used to represent the bivariate random fields of temperature and salinity. At location $\bx \in \mathcal{M}$, the prior bivariate distribution is
+The Gaussian model is used to represent the bivariate random fields of
+temperature and salinity. At location $\bx \in \mathcal{M}$, the prior
+bivariate distribution is:
 \begin{equation}\label{gp_bar}
@@
 \end{equation}
-where the notation refers to Gaussian (i.e. normal) distributed variables
-specified by the mean vector and covariance matrix. For the application
-of a river plume, the model parameters are specified from a short preliminary survey.
-The means vary with location, capturing the variability near the river mouth, while the variance parameters and correlation between salinity and temperature are assumed to be constant for all locations.
-%and within each water mass.
-
-For the spatial dependence among variables, we assume a separable correlation function with the same decay for salinity and temperature,
-which is consistent with expectations for  a river plume where the decay is related to the common water masses (Section \ref{sec:case_study}). In the resulting function, we have
+where the notation refers to Gaussian (i.e. normal) distributed
+variables specified by the mean vector and covariance matrix. For the
+application of a river plume, the model parameters are specified from
+a short preliminary survey. The means vary with location, capturing
+the variability near the river mouth, while the variance parameters
+and correlation between salinity and temperature are assumed to be
+constant for all locations. %and within each water mass.
+
+For the spatial dependence among variables, we assume a separable
+correlation function with the same decay for salinity and temperature,
+which is consistent with expectations for a river plume where the
+decay is related to common water masses (Section
+\ref{sec:case_study}). In the resulting function, we have:
 \begin{equation}\label{gp_corr}
@@
 \end{equation}
-where for $i,j \in {T,S}$ and $\delta_{ij}=1$ if $i=j$, and $\delta_{ij}=0$
-otherwise. Moreover, we assume stationary isotropic random fields a priori where the
-correlation function $\rho(\bx,\bx')$ solely depends on $\bx$ and
-$\bx'$ via Euclidean distance $h=\sqrt{||\bx-\bx'||^2}$. With
-data and prior knowledge, one could possibly fit and estimate
-parameters of non-stationary or non-isotropic covariance functions and
-more complex multivariate spatial covariance functions
+where for $i,j \in {T,S}$ and $\delta_{ij}=1$ if $i=j$, and
+$\delta_{ij}=0$ otherwise. Moreover, we assume stationary isotropic
+random fields a priori where the correlation function $\rho(\bx,\bx')$
+solely depends on $\bx$ and $\bx'$ via Euclidean distance
+$h=\sqrt{||\bx-\bx'||^2}$. With data and prior knowledge, one could
+possibly fit and estimate parameters of non-stationary or
+non-isotropic covariance functions and more complex multivariate
+spatial covariance functions
 \citep{gneiting2010matern,genton2015cross}; however, that is left for
@@
 
-We discretize the spatial domain
-$\mathcal{M}$ to a set of $n$ grid locations
-$\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell has area $\Delta$. The
-grid is used for the waypoint graph setting the possible design locations. It is also used when solving integral expressions in Eq. (\ref{measVAR}) and
-(\ref{sur}), which are then approximated by sums over all grid cells.  Using vector
-notation, we denote the Gaussian temperature and salinity variables on
-the grid and its density function by the formula: 
+We discretize the spatial domain $\mathcal{M}$ to a set of $n$ grid
+locations $\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell
+has area $\Delta$; the grid is used for the waypoint graph setting the
+possible design locations. It is also used when solving integral
+expressions in Eq. (\ref{measVAR}) and (\ref{sur}), which are then
+approximated by sums over all grid cells.  Using vector notation, we
+denote the Gaussian temperature and salinity variables on the grid and
+its density function by the formula:
 \begin{equation}\label{prior}
@@
 updated model. Then the updated model is used to design and select the
-next round of measurements, and so on (Section \ref{sec:heuristics}). Below, we describe the model for one round of data and the Gaussian formula for a single update.
+next round of measurements, and so on (Section
+\ref{sec:heuristics}). Below, we describe the model for one round of
+data and the Gaussian formula for a single update.
 
@@
 nodes in the grid. The conditional model for the data, given the true temperature
-and salinity, is described by 
+and salinity, is described by: 
 \begin{equation}\label{likelihood}
@@
 
-The GP mean and covariance in Eq. (\ref{prior}) are updated in the model when more information is available. The
-updated distribution for temperature and salinity variables on the
-spatial grid, given survey data $\by$, is Gaussian with mean and
-covariance: 
+The GP mean and covariance in Eq. (\ref{prior}) are updated in the
+model when more information is available. The updated distribution for
+temperature and salinity variables on the spatial grid, given survey
+data $\by$, is Gaussian with mean and covariance:
 \begin{eqnarray}\label{gp_upd}
@@
 the data $\by$ and the covariance is not a function of the data.
-Before inputting the data, the distribution of the conditional mean is
+Before inputting the data, the distribution of the conditional mean is:
 \begin{equation}\label{distxi} \bm \sim N(\bmu , \bPsi), \hspace{3mm}
@@
 extract elements from its mean vector and covariance matrix to get the
-bivariate distribution \begin{equation}\label{dist_mxi} 
+bivariate distribution:
+\begin{equation}\label{dist_mxi}
   \bm(\bx) \sim N \left( \begin{bmatrix}
@@
 Based on Gaussian modeling assumptions, the EPs can be computed at any
-location $\bx$. Without additional data, Eq.  (\ref{eq:prob}) gives
+location $\bx$. Without additional data, Eq.  (\ref{eq:prob}) gives:
 \begin{eqnarray}\label{eq:prob_mv0}
@@
 where $\Phi_2$ is the bivariate Gaussian cumulative distribution
-function, and the mean and covariance entries are as defined in Eq. (\ref{gp_bar}).
-
-When data $\by$ are available, Eq. (\ref{eq:post_ep}) similarly gives
+function, and the mean and covariance entries are as defined in
+Eq. (\ref{gp_bar}).
+
+When data $\by$ are available, Eq. (\ref{eq:post_ep}) similarly gives:
 \begin{eqnarray}\label{eq:prob_mv}
@@
 \end{eqnarray}
-with parameters as defined in Eq.
-(\ref{gp_upd}) and (\ref{gp_hat}).
-Here, the probabilities would not only depend on the data but also on the data design defined via matrix $\bG$ in Eq.
-(\ref{likelihood}).
-We next derive results that simplify the comparison of several designs.
+with parameters as defined in Eq. (\ref{gp_upd}) and
+(\ref{gp_hat}). Here, the probabilities would not only depend on the
+data but also on the data design defined via matrix $\bG$ in
+Eq. (\ref{likelihood}). We next derive results that simplify the
+comparison of several designs.
 
@@
 
-We will next derive closed-form solutions for the expectation in Eq. (\ref{sur}). The location index $\bx$ is suppressed to avoid overly
-complex notation. We present new results for computing the expectation in the integral of
-Eq. (\ref{sur}).
+We will next derive closed-form solutions for the expectation in
+Eq. (\ref{sur}). The location index $\bx$ is suppressed to avoid
+overly complex notation. We present new results for computing the
+expectation in the integral of Eq. (\ref{sur}).
 
@@
 variables in Eq. (\ref{sur}) as   $Z_T=\frac{\xi_T-m_{T}}{a_{T}}$,
-$Z_S=\frac{\xi_S-m_{S}}{a_{S}}$, with $\mbox{Corr}(Z_T,Z_S)=a_{T,S}/(a_{T} a_{S})=\eta_{T,S}$, and we can then rephrase the probability as
+$Z_S=\frac{\xi_S-m_{S}}{a_{S}}$, with
+$\mbox{Corr}(Z_T,Z_S)=a_{T,S}/(a_{T} a_{S})=\eta_{T,S}$, and we can
+then rephrase the probability as: 
 \begin{eqnarray}\label{ytom}
@@
 \end{eqnarray}
-Since the expression only depends on $\by$ via its affine function $\bm$, the expectation in Eq. (\ref{sur}) is reduced to a bivariate integral over $\bm$. 
-Introducing $\pi(\bm)$ for the density
-function of $\bm$, the expected Bernoulli variance becomes:
-\begin{eqnarray}\label{eq:var2}
+Since the expression only depends on $\by$ via its affine function
+$\bm$, the expectation in Eq. (\ref{sur}) is reduced to a bivariate
+integral over $\bm$.  Introducing $\pi(\bm)$ for the density function
+of $\bm$, the expected Bernoulli variance becomes:
+\begin{eqnarray}
+\label{eq:var2}
 E_{\by}(p(1-p)) &=& E_{\bm}(p(1-p)), \hspace{3mm} p=P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) \nonumber \\
@@
 
-For the first part of the integral above, we have
-\begin{equation}\label{part1:phi2}
+For the first part of the integral above, we have:
+\begin{equation}
+\label{part1:phi2}
  \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) \pi(\bm) d\bm= 
@@
 \end{equation}
-The standardized variable $\bZ=(Z_{T},Z_{S})$ is chosen to be independent of $\bm$. Using Eq. (\ref{dist_mxi}), we have
+The standardized variable $\bZ=(Z_{T},Z_{S})$ is chosen to be independent of $\bm$. Using Eq. (\ref{dist_mxi}), we have:
 \begin{equation}
@@
 \end{equation}
-for the temperature part; the same holds for the salinity part. From this we get:
+for the temperature part; the same holds for salinity. From this we
+get:
 \begin{eqnarray}\label{two_parts0}
@@
 
-The solution to the second part of Eq. \eqref{eq:22} can be found by evaluating a four-variable cumulative distribution function associated
+The solution to the second part of Eq. \eqref{eq:22} can be found by
+evaluating a four-variable cumulative distribution function associated
 with Eq. (\ref{part1:phi4}), accounting for the mean, variance and
@@
 combinations. Using vector-matrix formulations, we define the linear
-combinations
+combinations:
 \begin{equation}
@@
 The multivariate cumulative probabilities in two and four dimensions
-are fast to compute using methods developed by \cite{genz2009computation}.  In
-the end, the first part in Eq. (\ref{two_parts0}) and the second in
-Eq. (\ref{part1:phi4}) are computed for each location on the grid and
-summed over $\bx \in \mathcal{M}_g$, to get the design evaluation
-criterion in Eq. (\ref{sur}).
+are computationally efficient using methods developed by
+\cite{genz2009computation}.  In the end, the first part in
+Eq. (\ref{two_parts0}) and the second in Eq. (\ref{part1:phi4}) are
+computed for each location on the grid and summed over
+$\bx \in \mathcal{M}_g$, to get the design evaluation criterion in
+Eq. (\ref{sur}).
 
@@
 distribution functions for Gaussian vectors. For the more general
-situation, one would be interested in $K$ random fields. In the
-ocean sciences, the variables could include chlorophyll or other
+case, one would be interested in $K$ random fields. In the ocean
+sciences, the variables could include chlorophyll or other
 bio-geochemical quantities in addition to temperature and salinity.
@@
 standardize the variables and compute cumulative distributions.  In
-summary, we get
+summary, we get:
 \begin{equation}\label{two_parts}
@@
 $(y_T,y_S)^t=(\xi_T,\xi_S)^t+N(0,0.5^2\bI)$, while
-$y_T=\xi_T+N(0,0.5^2)$ when only temperature is measured.  For this illustration, Table \ref{tab:sim_rhoab} shows that the expected Bernoulli variance
-gets lower with larger standard deviations $\sigma_T$ and $\sigma_S$
-(right columns). The reduction of Bernoulli variance is largest
-for the cases with high correlation $\gamma$. Albeit smaller, there is
-also uncertainty reduction when only temperature is measured (bottom
-row), especially when temperature and salinity are highly correlated. When correlation is low ($\gamma=0.2$), there is little
-information about salinity in the temperature data, and therefore less
-uncertainty reduction. In an application with fresh cold water from
-river source, the temperature and salinity variables will not only be
-interdependent, but will show dependence in the spatial
-dimension. This in turn will impact the design criteria when we
-evaluate the information measure by integrating over several locations (Section \ref{sec:case_study}).
+$y_T=\xi_T+N(0,0.5^2)$ when only temperature is measured.  For this
+illustration, Table \ref{tab:sim_rhoab} shows that the expected
+Bernoulli variance gets lower with larger standard deviations
+$\sigma_T$ and $\sigma_S$ (right columns). The reduction of Bernoulli
+variance is largest for the cases with high correlation
+$\gamma$. Albeit smaller, there is also uncertainty reduction when
+only temperature is measured (bottom row), especially when temperature
+and salinity are highly correlated. When correlation is low
+($\gamma=0.2$), there is little information about salinity in the
+temperature data, and therefore less uncertainty reduction. In an
+application with fresh cold water from a river source, the temperature
+and salinity variables will not only be interdependent, but will also
+likely show dependence in the spatial dimension. This in turn will
+impact the design criteria when we evaluate the information measure by
+integrating over several locations (Section \ref{sec:case_study}).
 
@@
 
-An important concept underlying this work is the potential of robotic sampling to adapt survey plans based on data and statistical methods
-in accordance with the sense-plan-act loop
-(Fig. \ref{fig:envir1}). We will therefore present a few adaptive strategies
-for robotic sampling that build on the notion of multivariate ES.
+An important concept underlying this work is the potential of robotic
+sampling to adapt survey plans based on data and statistical methods
+in accordance with the sense-plan-act loop (Fig. \ref{fig:envir1}). We
+will therefore present a few adaptive strategies for robotic sampling
+that build on the notion of multivariate ES.
 
@@
 The optimal sequential design not only considers the best current AUV
-grid node, but also how the data gathered at this node could inform future sampling at successive stages. Let $d^{j,s}$ denote design
+grid node, but also how the data gathered at this node could inform
+future sampling at successive stages. Let $d^{j,s}$ denote design
 number $j$ at stage $s$ of a sequential survey. If this design is
@@
 selection situation is depicted in Fig. \ref{fig:PathSelOpt},
-\begin{figure}[h!]
+\begin{figure}[b!]
 \centering

### 2019-10-02 main_new.tex
Updates from Overleaf

--- 2019-09-30 main_new.tex
@@
 
-\address{ \\Department of Marine Technology,\\ Trondheim, Norway \\
+\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
 \printead{e1}}
-\address{Department of Mathematical Sciences,\\ Trondheim, Norway\\ \printead{e2}}
-\address{Idiap, \\ Martigny, Switzerland, \\
+\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e2}}
+\address{David Ginsbourger\\ Idiap Research Institute\\ Rue Marconi 19, 1920 Martigny\\ Switzerland\\
 \printead{e3}}
-\address{Department of Engineering Cybernetics,\\ Trondheim, Norway \\
+\address{Kanna Rajan\\Department of Engineering Cybernetics\\ O. S. Bragstads Plass 2D, 7034 Trondheim\\ Norway\\
 \printead{e4}}

### 2019-10-02 cover_letter/cover_letter.tex
added cover letter

\documentclass[10pt,stdletter]{newlfm}
\usepackage{charter}
\usepackage{times}

\widowpenalty=1000
\clubpenalty=1000

\newsavebox{\Luiuc}
\sbox{\Luiuc}{%
	\parbox[b]{1.75in}{%
		\vspace{0.5in}%
		\includegraphics[scale=0.8]
		{hovedlogo_eng.pdf}%
	}%
}%
\makeletterhead{Uiuc}{\Lheader{\usebox{\Luiuc}}}

\newlfmP{headermarginskip=10pt}
\newlfmP{sigsize=10pt}
\newlfmP{dateskipafter=30pt}
\newlfmP{addrfromphone}
\newlfmP{addrfromemail}
\PhrPhone{Phone}
\PhrEmail{Email}

\lthUiuc

\namefrom{\vspace{-1.2cm}Trygve Olav Fossum}
\addrfrom{Department of Marine Technology\\
	Otto Nielsens Veg 10\\
	Trondheim, Norway
}
\phonefrom{(+47) 98 678 309}
\emailfrom{\emph{trygve.o.fossum@ntnu.no}}

\addrto{%
Editor-in-Chief\\
Annals of Applied Statistics\\
}

\greetto{Dear Editor,}
\closeline{Yours faithfully,}

\Lfooter{\vspace{-2cm}Norwegian University of Science and Technology\\
Department of Marine Technology, Applied Underwater Robotics Laboratory\\
NO-7491, Trondheim, Norway\\ 
}

\Rheader{1 of 1}

\begin{document}
\begin{newlfm}

Please see our manuscript \textbf{Informative Oceanographic Sampling using Excursion Probabilities for Multivariate Random Fields} submitted for possible publication in the \emph{Annals of Applied Statistics} (AOAS).\\*

This work focuses on tackling the challenges related to oceanographic sampling, where observations are expensive, few, and far apart. A key question in this regard is

\begin{quote}
\textit{where and when should we measure in order to effectively measure a phenomenon or process taking place in the ocean?}      
\end{quote}{}

The research presented aims to address this central question by combining statistical models and methodology with autonomous robotic sensing platforms. A central effort is the introduction of informative descriptions of spatial characteristics of ocean phenomena that greatly improve how sequential data collection can be conducted. Specifically, we present a design criterion based on reducing uncertainty in the joint excursions of random fields, and derive analytical expressions for the expected Bernoulli variance reduction in the excursion set of spatial temperature and salinity variables. The proposed methodology is deployed and tested in a full-scale deployment where an autonomous underwater vehicle is tasked with mapping a river plume.

The combined study demonstrate the potential of statistics applied to support autonomous robotic decision making, representing methods and approaches with a substantial potential for further integration, which are necessary and critical to understand our changing oceans. 
\vspace{1em}

% combining new methods and tools to achieve increased sampling efficiency.
%Towards this end, the work introduces informative descriptions of spatial characteristics of ocean phenomena by define a design criterion based on reducing uncertainty in the joint excursions of random fields, and derive analytical expressions for the expected Bernoulli variance reduction in the excursion set of spatial temperature and salinity variables. 
%Resolving this type of sensing problem require theory and assets that can capitalize on the 
%Combining oceanographic sampling, combining compact models and autonomous robotic platforms for upper water-column observation. The work is novel in multiple ways. First, it uses a combination of unsupervised machine learning and hierarchical clustering methods to distill a compact ocean model based on remote sensing images. The method is capable of characterizing important dominant sea conditions using these images by grouping them into distinct classes. The classification is then verified using preceding wind data to ensure that the model is physically correct in oceanographical sense. Second, the work is supplemented with theory for doing forecast and predictions, as well as evaluation of model error. The combined study demonstrates a novel sampling perspective, combining new methods and tools to achieve increased sampling efficiency.\par \vspace{1em}
%We believe such a work to be representative of new methods and approaches which are necessary and critical to understand our changing oceans.

\end{newlfm}
\end{document}

### 2019-10-03 cover_letter/cover_letter.tex

@@
 
-Please see our manuscript \textbf{Informative Oceanographic Sampling using Excursion Probabilities for Multivariate Random Fields} submitted for possible publication in the \emph{Annals of Applied Statistics} (AOAS).\\*
+Please see our manuscript \textbf{Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets} submitted for possible publication in the \emph{Annals of Applied Statistics} (AOAS).\\*
 
@@
 \begin{quote}
-\textit{where and when should we measure in order to effectively measure a phenomenon or process taking place in the ocean?}      
+\textit{where, when, and how should we gather information in order to effectively characterize a phenomenon in the ocean?}  
 \end{quote}{}
@@
 Specifically, we present a design criterion based on reducing uncertainty in the joint excursions of random fields, and derive analytical expressions for the expected Bernoulli variance reduction in the excursion set of spatial temperature and salinity variables.
-The proposed methodology is deployed and tested in a full-scale deployment where an autonomous underwater vehicle is tasked with mapping a river plume.
+The proposed methodology is deployed and tested in a full-scale deployment where an autonomous underwater vehicle is used to map a river plume.
 
-The combined study demonstrate the potential of statistics applied to support autonomous robotic decision making, representing methods and approaches with a substantial potential for further integration, which are necessary and critical to understand our changing oceans.
+This interdisciplinary study demonstrates the potential of statistics applied to support autonomous robotic decision making, representing methods and approaches with a substantial potential for further integration, which are necessary and critical to understand our changing oceans.
 \vspace{1em}

### 2019-10-07 main_new.tex
Updates from Overleaf
updated abstract

@@
 % "Title of the paper"
-\title{Informative Oceanographic Sampling using Excursion Probabilities for Multivariate Random Fields}
+\title{Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets}
 \runtitle{Excursion Probabilities for Informative Sampling}
@@
   resources to understand dynamic processes in the water-column, the
-  combination of statistics and autonomous robotics can provide
-  opportunities for studying experimental design for effective
-  characterization of oceanographic phenomena. In this work we develop
+  combination of statistics and autonomous robotics provide
+  opportunities for studying experimental design. In this work we develop
   methods for efficient spatial sampling applied to the mapping of
@@
   applicable in static and sequential designs, building on myopic and
-  look-ahead sampling strategies. We use simulations to study the
+  non-myopic sampling strategies. We use simulations to study the
   properties and differences between state-of-the-art approaches
@@
   underwater vehicle as part of a case study mapping the boundary of a
-  river plume. The results demonstrate the ability of the methods to
-  effectively inform and execute data-driven environmental sampling.
+  river plume.
+The results demonstrate the potential of combining statistical methods and robotic platforms, and the abilities of the derived methods to effectively inform and execute data-driven environmental sampling.
 

### 2019-10-08 main_new.tex
updates

@@
 \end{equation}
-The spatial correlation is of a Matern type:
+The spatial correlation is of a Matern type \kc{cite or define}:
 \begin{equation}\label{v}
@@
 \textit{static\_zigzag} with the three sequential approaches
-\textit{naive}, \textit{myopic}, and \textit{look-ahead}. Recall that the static
-sampling paths are pre-determined and cannot be changed on the basis
-of data. They are representing the pre-planned observation
-strategies used in current-day AUV operational survey designs.
-
-For a fixed survey length, the expected IBV is analytically available.
-However, for the sequential approaches this is not the case.
-For comparison the properties are therefore evaluated using Monte Carlo sampling over several replicates of realizations from the prior, conducting simulated sequential surveying for each one.
-Fig.
-\ref{fig:Eprob} shows the conditional EP, given data gathered along the north-south survey line for the realization indicated in these displays.
-In the Monte Carlo replicates, such results are repeatedly computed to approximate the expected IBV.
-We also compare predictive performance measured by root mean square error (RMSE) for the temperature and salinity estimates and the variance reduction in these two variables.
-It is important to note that the objective function used by the agents is focused on reducing the expected IBV, but we nevertheless expect that we will achieve good predictive performance for criteria such as RMSE as well.
-Another non-statistical criteria that is important for practical purposes is the computational time of the strategy, as this will impact the performance of the approach in the field for an embedded system.
+\textit{naive}, \textit{myopic}, and \textit{look-ahead}. Recall that
+the static sampling paths are pre-determined and cannot be altered.
+These represent the pre-planned strategies used in most current AUV
+operational survey designs.
+
+For a fixed survey length, the expected IBV is analytically available
+\kc{I don't understand what ``analytically available''
+  means}. However, for the sequential approaches this is not the
+case. For comparison the properties are therefore evaluated using
+Monte Carlo sampling over several replicates of realizations from the
+prior while conducting simulated sequential surveys for each
+one. Fig. \ref{fig:Eprob} shows the conditional EP, given data
+gathered along the north-south survey lines in these figures.  In the
+Monte Carlo replicates, such results are repeatedly computed to
+approximate the expected IBV. We also compare predictive performance
+measured by root mean square error (RMSE) for temperature and salinity
+estimates as also the variance reduction in these two variables. It is
+important to note that the objective function used by the agent is
+focused on reducing the expected IBV, but we nevertheless expect that
+we will achieve good predictive performance for criteria such as RMSE
+as well. Another non-statistical criteria that is important for
+practical purposes is the computational time needed for the strategy,
+as this will impact the performance for an embedded system.
 
@@
 
-Each strategy is conducted on an equilateral grid as shown in Fig.
-\ref{fig:wp_graph}.
-The sequential sampling agent starts at the center East-West coordinate at the southern end of the domain (node 53).
-The AUV will move along edges in the waypoint graph while the vehicle gathers measurements.
-The data are assimilated into the GP model before an evaluation of the next node to sample is conducted at the end of the edge.
+Each strategy is conducted on an equilateral grid as shown in
+Fig. \ref{fig:wp_graph}. The sequential sampling agent starts at the
+center East-West coordinate at the southern end of the domain (node
+53). The AUV moves along edges in the waypoint graph while the vehicle
+gathers measurements. The data is assimilated into the GP model before
+an evaluation of the next node to sample is conducted at the end of
+the edge. \kc{Note there are plenty of language errors related to
+  tense. Pl. do not mess with any more changes from here onwards.}
 
 \subsection{Simulation Results}
-A total of 100 replicate simulations were conducted. For each simulation all strategies are run. The results are shown in Fig. \ref{fig:sim_results},
-where the different criteria are plotted as a function of survey distance.
-Fig.
-\ref{fig:avg_ev} shows the resulting drop in IBV for each of the six strategies.
-The IBV is reduced most under the \textit{myopic} and \textit{look-ahead} strategies, each performing almost equally; this is expected as the two criteria (Eq.
-\eqref{critSEQ} and \eqref{critLA}) are much more sensitive to differences in IBV.
-The \textit{static\_north} design also does well here because the path is parallel to the  boundary between the water masses.
+
+A total of 100 replicate simulations were conducted with all
+strategies. The results are shown in Fig. \ref{fig:sim_results}, where
+the different criteria are plotted as a function of survey
+distance. Fig. \ref{fig:avg_ev} shows the resulting drop in IBV for
+each of the six strategies. IBV reduction occurs most under the
+\textit{myopic} and \textit{look-ahead} strategies, each performing
+almost equally; this is expected as the two criteria
+(Eq. \eqref{critSEQ} and \eqref{critLA}) are sensitive to differences
+in IBV. The \textit{static\_north} design also does well here because
+the path is parallel to the boundary between the water masses.
 
@@
 
-Fig.
-\ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop in RMSE and increase in explained variance, respectively.
-Both \textit{myopic} and \textit{look-ahead} strategies perform well here, but some of the \textit{static\_east} and \textit{static\_zigzag} also achieve good results because they are pre-determined to cover large parts of the domain without re-visitation.
-Sequential strategies targeting IBV will sometimes not reach similar coverage, as interesting data may draw the path into twists and turns.
-There is relatively large variety in the replicate results as indicated by the vertical lines.
-Still, the ordering of strategies is similar.
+Fig. \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop
+in RMSE and increase in explained variance, respectively. Both
+\textit{myopic} and \textit{look-ahead} strategies perform well here,
+but some of the \textit{static\_east} and \textit{static\_zigzag} also
+achieve good results because they are pre-determined to cover large
+parts of the domain without re-visitation. Sequential strategies
+targeting IBV will sometimes not reach similar coverage, as
+interesting data may draw the path into twists and turns. There is
+relatively large variety in the replicate results as indicated by the
+vertical lines. Nevertheless, the ordering of strategies is similar.
 
@@
 
-Fig.
-\ref{fig:avg_time} shows the computational effort: the \textit{naive} strategy is on par with the static designs, while the \textit{myopic} strategy is slower.
-The \textit{look-ahead} is even slower, reaching levels that are nearly impractical for execution on a vehicle.
-Some pruning of the graph is performed to improve the computational speed, such as ruling out repeated visitations and back-and-forth routes.
-Some of the intermediate results are also stored for longer planning horizons.
-Still, more pruning of branches or inclusion of other heuristics could be included to make the strategy faster.
-Then again, the inclusion of heuristics is likely a contributing factor to the \textit{look-ahead} strategy failing to outperform the \textit{myopic} strategy.
-
-In Fig.
-\ref{fig:route_choices}, the realized sampling paths for each of the sequential schemes and static designs are shown.
-The \textit{naive} strategy often gets stuck in the southern part of the domain because it is too focused on the probabilities near $0.5$.
-The \textit{myopic} strategy covers a wider domain than the naive or look-ahead.
-There are several reasons for this; first, a greedy approach will tend to put more emphasis on locations close to the agent, which may lead away from the centre.
-Second, as the agent evaluates the impact of locations further away (look-ahead) where assimilated data has less predictive power, the GP prior model (which is centered here) will act to restrict paths deviating from the central zone.
+Fig.  \ref{fig:avg_time} shows the computational effort: the
+\textit{naive} strategy is on par with the static designs, while the
+\textit{myopic} strategy is slower. The \textit{look-ahead} is even
+slower, reaching levels that are nearly impractical for execution on a
+vehicle. Some pruning of the graph is performed to improve the
+computational speed, such as ruling out repeated visitations and
+back-and-forth routes. Some of the intermediate results are also
+stored for longer planning horizons. Further pruning of branches or
+inclusion of other heuristics could be included to make the strategy
+faster. Then again, the inclusion of heuristics is likely a
+contributing factor for the \textit{look-ahead} strategy failing to
+outperform the \textit{myopic} strategy.
+
+In Fig. \ref{fig:route_choices}, the realized sampling paths for each
+of the sequential schemes and static designs are shown. The
+\textit{naive} strategy often gets stuck in the southern part of the
+domain because it is too focused on the probabilities near $0.5$. The
+\textit{myopic} strategy covers a wider domain than the naive or
+look-ahead. There are several reasons for this. First, a greedy
+approach will tend to put more emphasis on locations close to the
+agent, which may lead away from the centre. Second, as the agent
+evaluates the impact of locations further away (look-ahead) where
+assimilated data has less predictive power, the GP prior model (which
+is centered here) will act to restrict paths deviating from the
+central zone.
 
@@
 \subsection{Performance Metrics}
-The tables below explore the sensitivity of the result to different prior correlation values between temperature measured onboard the AUV, temperature and salinity, prior standard deviations, and spatial correlations (see table rows).
-We only compare results at the end of each survey.
-The expected ES variances are shown in Table \ref{tab:sim_res_ev}, results of RMSE are in Table \ref{tab:sim_res_rmse}, and explained variance is in Table \ref{tab:sim_res_r2}.
-In all runs, both the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of expected variance in the ES.
-Surprisingly, the \textit{myopic} performs slightly better than the \textit{look-ahead} strategy.
-Part of the reason for this performance discrepancy can be traced to the act of pruning some of the paths for the \textit{look-ahead} strategy; however, even with less pruning, the \textit{look-ahead} does not outperform the \textit{myopic} approach.
-The \textit{static\_north} design is clearly better than the other two static designs, the reason being that the others do not focus on the areas where the ES is really uncertain (centre locations are the most uncertain).
-By measuring temperature alone, the expected variances in the ES are only slightly higher than the combined temperature-salinity estimate.
-This occurs because the correlation of $0.6$ means that temperature observations provide information about salinity as well.
-For the other predictive measures, \textit{myopic} and \textit{look-ahead} strategies do well, clearly beating the \textit{naive} strategy.
-The \textit{static\_zigzag} design is usually the best among the static designs for these predictive performance measures because its pattern gives more effective coverage of the domain when there is significant spatial correlation, as indicated in this case.
+
+The tables below explore the sensitivity of the result to different
+prior correlation values between temperature measured onboard the AUV,
+temperature and salinity, prior standard deviations, and spatial
+correlations (see table rows). We only compare results at the end of
+each survey. The expected ES variances are shown in Table
+\ref{tab:sim_res_ev}, results of RMSE are in Table
+\ref{tab:sim_res_rmse}, and explained variance is in Table
+\ref{tab:sim_res_r2}. In all runs, both the \textit{myopic} and
+\textit{look-ahead} strategies perform the best in terms of expected
+variance in the ES. Surprisingly, the \textit{myopic} strategy
+performs slightly better than the \textit{look-ahead}. Part of the
+reason for this discrepancy can be traced to the act of pruning some
+of the paths for the \textit{look-ahead} strategy; however, even with
+less pruning, the \textit{look-ahead} does not outperform the
+\textit{myopic} approach. The \textit{static\_north} design is clearly
+better than the other two static designs, the reason being that the
+others do not focus on the areas where the ES is really uncertain
+(centre locations are the most uncertain). By measuring temperature
+alone, the expected variances in the ES are only slightly higher than
+the combined temperature-salinity estimate. This occurs because the
+correlation of $0.6$ means that temperature observations provide
+information about salinity as well. For the other predictive measures,
+\textit{myopic} and \textit{look-ahead} strategies do well, clearly
+beating the \textit{naive} strategy. The \textit{static\_zigzag}
+design is usually the best among the static designs for these
+predictive performance measures because its pattern gives more
+effective coverage of the domain when there is significant spatial
+correlation, as indicated in this case.
 
@@
 
-To demonstrate the applicability of using multivariate EPs and the IBV to inform oceanographic sampling, we present a case study mapping a river plume with an AUV.
-The experiment was performed in Trondheim, Norway,
-surveying the Nidelva river (Fig.
-\ref{fig:nidelven}).
-The experiments were conducted in late Spring 2019, when there is still snow melting in the surrounding mountains so that the river water is substantially colder than the water in the fjord.
-The experiment is focused along the frontal zone that runs more or less parallel to the eastern shore as
-indicated in purple in Fig. \ref{fig:nidelven}.
+To demonstrate the applicability of using multivariate EPs and the IBV
+to inform oceanographic sampling, we present a case study mapping a
+river plume with an AUV. The experiment was performed in Trondheim,
+Norway, surveying the Nidelva river (Fig. \ref{fig:nidelven}). The
+experiments were conducted in late Spring 2019, when there is still
+snow melting in the surrounding mountains so that the river water is
+substantially colder than the water in the fjord. The experiment was
+focused along the frontal zone that runs more or less parallel to the
+eastern shore as noted in Fig. \ref{fig:nidelven}.
 
@@
 
-The statistical model parameters were specified based on a short preliminary survey where the AUV did an initial transect to determine the environmental conditions and correlation structure.
-Based on the initial data, the trend parameters were estimated by linear regression, where both temperature and salinity are assumed to increase linearly, going west from the river mouth.
-Next, the residuals from the regression analysis were analyzed to study the fit of the GP model and to specify the covariance parameters.
+The statistical model parameters were specified based on a short
+preliminary survey where the AUV made an initial transect to determine
+environmental conditions and correlation structure. Based on the
+initial data, the trend parameters were estimated by linear
+regression, where both temperature and salinity are assumed to
+increase linearly, going west from the river mouth. Next, the
+residuals from the regression analysis were analyzed to study the fit
+of the GP model and to specify the covariance parameters.
 
@@
   \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_c.pdf}\label{fig:parest_c}}
-  \caption{Data analysis from a preliminary trial experiment using the AUV.
-\ref{fig:parest_a} Residual plot of temperature and salinity along with Gaussian contours.
-\ref{fig:parest_b} Empirical CDF (solid) of the quadratic form of the residuals along with the theoretical CDF (dashed) of the $\chi^2$  distribution with two degrees of freedom.
-\ref{fig:parest_c} Empirical variogram of the salinity and temperature data.}
+  \caption{Data analysis from a preliminary trial experiment using the
+    AUV. \ref{fig:parest_a} Residual plot of temperature and salinity
+    along with Gaussian contours. \ref{fig:parest_b} Empirical CDF
+    (solid) of the quadratic form of the residuals along with the
+    theoretical CDF (dashed) of the $\chi^2$ distribution with two
+    degrees of freedom. \ref{fig:parest_c} Empirical variogram of the
+    salinity and temperature data.}
 \label{fig:parest}
@@
 
-Fig.
-\ref{fig:parest} shows diagnostic plots of this analysis.
-Fig.
-\ref{fig:parest_a} shows a cross-plot of temperature and salinity residuals after the westernly trends in both salinity and temperature are subtracted from the data.
-This scatterplot of joint residuals indicates larger variability in salinity than in temperature, and a positive correlation ($0.5$) between the two variables.
-Based on the fitted bivariate Gaussian model (ellipses in Fig.
-\ref{fig:parest_a}), we can compute the modeled quadratic form of the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed.
-Fig.
-\ref{fig:parest_b} shows the empirical cumulative distribution function (CDF) of these quadratic forms (solid) together with the theoretical CDF of the $\chi^2_2$ distribution.
-The modeled and theoretical curves are very similar, which indicates that the Gaussian model fits reasonably well.
-Even though there appears to be some clustering in both Fig.
-\ref{fig:parest_a} and \ref{fig:parest_b}, the bivariate diagnostic plots look reasonable and justify a Gaussian model.
-Fig.
-\ref{fig:parest_c} shows the empirical variogram of the scaled residuals for temperature and salinity.
-The decay is similar for the two, and seems to be negligible after about $150$ m.
-
-Based on the analysis in Fig.
-\ref{fig:parest}, the resulting parameters are given in Table \ref{tab:experiment_param}.
-The regression parameters shown here are scaled to represent the east and west boundaries of the domain as seen in the preliminary transect data, and the thresholds are intermediate values.
-These parameter values were then used in field trials where we explored the algorithm's ability to characterize the river plume front separating the river and fjord water masses, providing a spatial map of this boundary.
+Fig. \ref{fig:parest} shows diagnostic plots of this
+analysis. Fig. \ref{fig:parest_a} shows a cross-plot of temperature
+and salinity residuals after the westerly trends in both salinity and
+temperature are subtracted from the data. This scatterplot of joint
+residuals indicates larger variability in salinity than in
+temperature, and a positive correlation ($0.5$) between the two
+variables. Based on the fitted bivariate Gaussian model (ellipses in
+Fig. \ref{fig:parest_a}), we can compute the modeled quadratic form of
+the residuals, and if the model is adequate they should be
+approximately $\chi^2_2$ distributed. Fig. \ref{fig:parest_b} shows
+the empirical cumulative distribution function (CDF) of these
+quadratic forms (solid) together with the theoretical CDF of the
+$\chi^2_2$ distribution. The modeled and theoretical curves are very
+similar, which indicates that the Gaussian model fits reasonably
+well. Even though there appears to be some clustering in both
+Fig. \ref{fig:parest_a} and \ref{fig:parest_b}, the bivariate
+diagnostic plots look reasonable and justify a Gaussian
+model. Fig. \ref{fig:parest_c} shows the empirical variogram of the
+scaled residuals for temperature and salinity. The decay is similar
+for the two, and seems to be negligible after about $150$ m.
+
+Based on the analysis in Fig. \ref{fig:parest}, the resulting
+parameters are given in Table \ref{tab:experiment_param}. The
+regression parameters shown here are scaled to represent the east and
+west boundaries of the domain as seen in the preliminary transect
+data, and the thresholds are intermediate values. These parameter
+values were then used in field trials where we explored the
+algorithm's ability to characterize the river plume front separating
+the river and fjord water masses, providing a spatial map of this
+boundary.
 
@@
 
-The sampling locations were distributed over an equilateral grid, as shown in the grey-colored lattice in Fig.
-\ref{fig:map}.
-The robotic platform consisted of a Light AUV \citep{sousa2012lauv} (Fig.
-\ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49 conductivity, temperature, and depth (CTD) sensor providing temperature and salinity measurements.
-%The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
-The sampling agent was built on top of the autonomous agent framework Teleo-Reactive EXecutive (\textit{T-REX}) \citep{py10,Rajan12,Rajan12b}, running an instance of the \textit{myopic} strategy from Section \ref{sec:sampling_designs} to control the AUV and decide between sampling locations.
+The sampling locations were distributed over an equilateral grid, as
+shown in the grey-colored lattice in Fig. \ref{fig:map}. The robotic
+platform consisted of a Light AUV \citep{sousa2012lauv}
+(Fig. \ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49
+conductivity, temperature, and depth (CTD) sensor providing
+temperature and salinity
+measurements. %The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
+The sampling agent was built on top of the autonomous agent framework
+Teleo-Reactive EXecutive (\textit{T-REX})
+\citep{py10,Rajan12,Rajan12b}, running an instance of the
+\textit{myopic} strategy from Section \ref{sec:sampling_designs} to
+control the AUV and decide between sampling locations.
 
@@
   Vehicle (LAUV) platform for upper water-column exploration used in
-  the experiment.}
+  our experiments.}
 \label{fig:lauv}
@@
 
-The sampling strategy was designed around the concept of visiting waypoints sequentially.
-Arriving at a desired waypoint with new measurements and an updated model, the AUV triggers the myopic strategy to evaluate all possible design criteria (see Eq.
-\eqref{critSEQ}).
-The waypoint-and-path combination that is expected to reduce the IBV the most is selected; upon arrival this procedure is then repeated.
-At each stage, it takes the AUV about 30 seconds to evaluate the expected IBV for all the possible waypoint-and-path alternatives.
-
-The AUV was set to start in the south-center part of the waypoint graph, with the previously outlined GP model of the environment.
-A survey was set to take approximately 40 minutes to visit 15 waypoints on the grid, with the AUV running near the surface to better capture the plume.
-On its path from one waypoint to the next, the AUV gathered data regularly, and the GP model assimilated temperature and salinity data with an update frequency of 30 seconds, with about three updates per stage.
+The sampling strategy was designed around the concept of visiting
+waypoints sequentially. Arriving at a desired waypoint with new
+measurements and an updated model, the AUV triggers the myopic
+strategy to evaluate all possible design criteria (see
+Eq. \eqref{critSEQ}). The waypoint-and-path combination that is
+expected to reduce the IBV the most is selected; upon arrival this
+procedure is then repeated. At each stage, it takes the AUV about 30
+seconds to evaluate the expected IBV for all the possible
+waypoint-and-path alternatives.
+
+The AUV was set to start in the south-center part of the waypoint
+graph, with the previously outlined GP model of the environment. A
+survey was set to take approximately 40 minutes to visit 15 waypoints
+on the grid, with the AUV running near the surface to capture the
+plume. On its path from one waypoint to the next, the AUV gathered
+data regularly, and the GP model assimilated temperature and salinity
+data with an update frequency of 30 seconds, with about three updates
+per stage.
 
@@
 %\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
-\caption{Results from mapping the Nidelva river.
-\ref{fig:map} Show an overview of the survey area overlaid with the AUV path in black and dashed line.
-Note the shaded region indicating the usual frontal region.
-\ref{fig:res_both} Presents the collected temperature data as colored trails.
-Note waypoint 5 (WP5) which indicate where the two surveys deviate.
-\ref{fig:res1} and \ref{fig:res2} Shows the collected salinity data (as a colored trail) overlaid on the final EP, which indicate the AUVs statistical impression of the front.
-For both missions the temperature and salinity data correspond with the EP front indication.}
+\caption{Results from mapping the Nidelva river. \ref{fig:map} Show an
+  overview of the survey area overlaid with the AUV path in black and
+  dashed line. Note the shaded region indicating the usual frontal
+  region. \ref{fig:res_both} Presents the collected temperature data
+  as colored trails. Note waypoint 5 (WP5) which indicate where the
+  two surveys deviate. \ref{fig:res1} and \ref{fig:res2} Shows the
+  collected salinity data (as a colored trail) overlaid on the final
+  EP, which indicate the AUVs statistical impression of the front. For
+  both missions the temperature and salinity data correspond with the
+  EP front indication.}
 \label{fig:results}
@@
 
-Two missions, Surveys 1 and 2, were run successively from 11:00 AM to 01:00 PM, with a short break in between.
-The resulting path of the selected waypoints are shown in the map in Fig.
-\ref{fig:map}, both within the expected frontal region (shaded pink).
-The recorded temperatures are shown as colored trails in Fig.
-\ref{fig:res_both}, clearly indicating the temperature difference between fjord and river waters.
-The salinity data are then shown separately, overlaid with the estimated EP for each survey in Fig.
-\ref{fig:res1} and \ref{fig:res2}.
-
-Both Surveys 1 and 2 successfully estimated and navigated the separation zone, crossing the frontal boundary multiple times.
-As conditions changed slightly between the two deployments, the resulting path (after waypoint 5) deviates.
-Survey 1 continued northwards, tracking the north-eastern part of the front, while Survey 2 turned west, mapping the south-western part.
-
-The final prediction of the front location given in Fig.
-\ref{fig:res1} and \ref{fig:res2} as a dashed line, correspond with one another, yielding a picture of the front gradually bending off toward the north east.
-The amount of exploration done by Survey 1 is greater than Survey 2.
-In Survey 1, the AUV obtained more detail by going north, while Survey 2, coming close to the survey area borders in the south-western corner, obtained a poorer understanding of the northern parts.
-A longer planning horizon would likely identify and discourage such choices.
+Two missions, Surveys 1 and 2, were run successively from 11:00 AM to
+01:00 PM, with a short break in between. The resulting path of the
+selected waypoints are shown in the map in Fig. \ref{fig:map}, both
+within the expected frontal region (shaded pink). The recorded
+temperatures are shown as colored trails in Fig. \ref{fig:res_both},
+clearly indicating the temperature difference between fjord and river
+waters. The salinity data are then shown separately, overlaid with the
+estimated EP for each survey in Fig. \ref{fig:res1} and
+\ref{fig:res2}.
+
+Both surveys successfully estimated and navigated the separation zone,
+crossing the frontal boundary multiple times. As conditions changed
+slightly between the two deployments, the resulting path (after
+waypoint 5) is shown to deviate. Survey 1 continued northwards,
+tracking the north-eastern part of the front, while Survey 2 turned
+west, mapping the south-western part.
+
+The final predictions of the front location given in
+Fig. \ref{fig:res1} and \ref{fig:res2} as a dashed line, correspond
+with one another, yielding a picture of the front gradually bending
+off toward the north east. The amount of exploration done by Survey 1
+is greater than Survey 2. In Survey 1, the AUV obtained more detail by
+going north, while Survey 2, coming close to the survey area borders
+in the south-western corner, obtained a poorer understanding of the
+northern parts. A longer planning horizon would likely identify and
+discourage such choices.
 
@@
 
-This paper builds on multidisciplinary work combining statistical modeling and methods with robotic surveying technologies for oceanographic sensing applications.
-We show how  observation practices can gain efficiency and accuracy from the development of statistical techniques to achieve more effective spatial monitoring.
-We further demonstrate the opportunities available for real-time multivariable spatial data-gathering and analysis onboard autonomous sensing platforms, which statisticians can exploit to create general-purpose toolkits for similar applications.
-
-In particular, we derive and show results for characterizing phenomena connected to the properties of water masses.
-The characterization is based on joint ESs, whereby we derive new results for the expected IBV reduction achieved by the spatial sampling design.
-This result is first calculated in closed form for the situation with a static design, and then extended to the adaptive situation; the sequential derivations  provide new insights into efficient applications of adaptive sampling, as demonstrated in our application of these models with an autonomous vehicle in Trondheim fjord.
-
-The study did not consider any temporal effects, which would be relevant on a larger time scale.
-We consider the extension to spatio-temporal modeling as future work, and envision that advection-diffusion equations could be useful in this kind of modeling \citep{sigrist2015stochastic}.
-For more complex oceanographic phenomena, the methods will need to be extended to non-Gaussian phenomena, possibly feature-based mixtures of GPs which could still be run onboard and augmented by dynamical models.
-
-The spatial-statistical design criterion, building on ESs, is relevant in our setting with different water mass properties; this could also be useful in other oceanographic settings, such as algae-blooms or frontal systems.
-Of course, other criteria are also relevant for decision-making.
-For instance, managers and regulators must make difficult decisions related to fish farming, other marine resources operations, or environmental projects.
-Value of information analysis (VOI) \citep{Eidsvik:15} could be used in a similar vein as the IBV in the current paper, as both analytic methods explore when, where and which information is likely to result in improved decision-making.
-
-While more effort could be spent on approximating the look-ahead steps in the adaptive designs, it is perhaps more interesting to explore the additional flexibility that can be gained by having two or more AUVs  exploring a spatial or spatio-temporal domain together \citep{ferreira2019advancing}.
-Such an approach would enable concurrent sampling in  different parts of the space, or opportunities to move in parallel to best capture the excursion set.
+This work builds on a multidisciplinary effort combining statistical
+modeling with robotic surveying techniques for oceanographic
+applications. We show how observation practices can gain efficiency
+and accuracy from the development of statistical techniques to achieve
+more effective spatial monitoring. We further demonstrate the
+opportunities available for real-time multivariable spatial
+data-gathering and analysis onboard autonomous sensing platforms,
+which statisticians can exploit to create general-purpose toolkits for
+similar applications.
+
+In particular, we derive and show results for characterizing phenomena
+connected to the properties of water masses. The characterization is
+based on joint ESs, whereby we derive new results for the expected IBV
+reduction achieved by the spatial sampling design. This result is
+first calculated in closed form for the situation with a static
+design, and then extended to the adaptive situation; the sequential
+derivations provide new insights into efficient applications of
+adaptive sampling, as demonstrated in our application of these models
+with an autonomous vehicle in Trondheim fjord.
+
+The study did not consider any temporal effects, which would be
+relevant on a larger time scale. We consider the extension to
+spatio-temporal modeling as future work, and envision that
+advection-diffusion equations could be useful in this kind of modeling
+\citep{sigrist2015stochastic}. For more complex oceanographic
+phenomena, the methods will need to be extended to non-Gaussian
+phenomena, possibly feature-based mixtures of GPs which could still be
+run onboard and augmented by dynamical models.
+
+The spatial-statistical design criterion, building on ESs, is relevant
+in our setting with different water mass properties; this could also
+be useful in other oceanographic settings, such as algal-blooms or
+open water fronts. Of course, other criteria are also relevant for
+decision-making.  For instance, managers and regulators must make
+difficult decisions related to fish farming, other marine resources
+operations, or environmental projects. Value of information analysis
+(VOI) \citep{Eidsvik:15} could be used in a similar vein as the IBV in
+the current paper, as both analytic methods explore when, where and
+which information is likely to result in improved decision-making.
+
+While more effort could be spent on approximating the look-ahead steps
+in the adaptive designs, it is perhaps more interesting to explore the
+additional flexibility that can be gained by having two or more AUVs
+co-temporally exploring a spatial or spatio-temporal domain together
+\citep{ferreira2019advancing}. Such an approach would enable
+concurrent sampling in different parts of the space, or opportunities
+to move in parallel to best capture the excursion set.
 
 \section*{Acknowledgements}
-TOF and KR acknowledge support from the Centre for Autonomous Marine Operations and Systems (AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of Excellence, project number 223254, and the Applied Underwater Robotics Labortatory (AURLab).
-JE acknowledges support from Norwegian research council project 294404.
-DG acknowledges support from the Swiss National Science Foundation, project number 178858.
+
+TOF and KR acknowledge support from the Centre for Autonomous Marine
+Operations and Systems
+(AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of
+Excellence, project number 223254, and the Applied Underwater Robotics
+Labortatory (AURLab). JE acknowledges support from Norwegian research
+council project 294404. DG acknowledges support from the Swiss
+National Science Foundation, project number 178858.
 

### 2019-10-09 main_new.tex
Small updates

@@
   resources to understand dynamic processes in the water-column, the
-  combination of statistics and autonomous robotics provide
+  combination of statistics and autonomous robotics provide new
   opportunities for studying experimental design. In this work we develop
@@
   Bernoulli variance reduction in the excursion set of spatial
-  temperature and salinity variables. We show how these can be
+  temperature and salinity variables, providing a description of
+the boundary uncertainty that can be used prioritize sampling efforts at locations that are
+ambiguous, making exploration more effective. We show how these can be
   applicable in static and sequential designs, building on myopic and
   non-myopic sampling strategies. We use simulations to study the
-  properties and differences between state-of-the-art approaches
+  properties and differences between state-of-the-art approaches,
   followed by results from field deployments with an autonomous
@@
 large-scale coverage but have limited resolution, are limited to
-sensing the surface of the ocean and are impacted by cloud cover. The
+sensing the surface of the ocean, and are impacted by cloud cover. The
 advent of robust mobile robotic platforms \citep{Bellingham07} has
@@
 integral part of ocean observation; our previous work has contributed
-to this effort
-% using statistical methods in design and data assimilation
-\citep{das11b,Das2015,fossum18b,fossuminformation}.  Other statistical
+to this effort \citep{das11b,Das2015,fossum18b,fossuminformation}. Other statistical
 work in the oceanographic domain include \cite{wikle2013modern}
@@
 The potential that lies in using information-based methods, which we
-highlight in the paper, is to do efficient environmental sampling and
+highlight in this paper, is to do efficient environmental sampling and
 overcome three common challenges, namely:
@@
   information obtained from measurements taken in-situ. Recently
-  acquired information can be assimilated into statistical models that
+  acquired information can be used to resolve this by assimilating up-to-date data into statistical models that
   can be used to inform decisions on where to sample
@@
   solution in reasonable time). To successfully resolve features, this
-  trade-off has to be considered in sample design and the practice.
+  trade-off has to be considered in development and practice.
 
@@
 
-The combination of statistical tools and robotic platforms is a
+Addressing this, the combination of statistical tools and robotic platforms is a
 natural symbiosis which enables information-based sensing. Central to
@@
 Gaussian Processes (GPs) in large part because they allow for
-efficient implementation and evaluation onboard a robotic platform.
+efficient implementation and evaluation in real time onboard a robotic platform.
 
@@
 an agglomeration of physical forcing (wind, topography, bathymetry,
-tidal influences and local geography, such as fresh water inflows) and
+tidal influences, and local geography, such as fresh water inflows) and
 incipient micro-biology driven by planktonic and coastal anthropogenic
-input, such as pollution and agricultural runoff. These often result
+input, such as pollution and agricultural runoff transported into the ocean by rivers and streams. These often result
 in a range of ecosystem-related phenomena such as blooms, anoxic zones
@@
 understanding (See Fig. \ref{fig:envir1}), where observation is guided
-by assimilation of measurements and statistical models.
+by assimilation of recent measurements and statistical models.
 
@@
 River plumes belong to a class of ocean processes that are local and
-at smaller \emph{sub-mesoscale} (from $5 m^2$ -- $10 km^2$) where
+act on smaller \emph{sub-mesoscale} extent (from $5 m^2$ -- $10 km^2$) where
 lateral spatial variability tends to dominate. Remote sensing and
@@
 along fixed transects on a spatial grid, pre-programmed by the human
-operator. Instead using onboard algorithms to continuously evaluate,
-update, and refine future sampling locations, one can make information
+operator. A more effective approach is to instead use onboard algorithms to continuously evaluate,
+update, and refine future sampling locations (sense-plan-act cycle in Fig. \ref{fig:envir1}), making the information
 gathering \emph{adaptive} and \emph{contextual} to the survey
 requirements \citep{das11b,fossum18b,fossuminformation}. The space of
-sampling opportunities is still limited to a grid, but adaptation
+sampling opportunities can still be limited to a grid, but adaptation
 allows more flexibility. For instance, an AUV could reconstruct or
@@
 rely on a statistical model of the environment, but rather adapts
-based on closing the sensing and actuation loop
-(Fig. \ref{fig:envir1}). Myopic sampling, i.e. step-wise selection of
+based on closing the sensing and actuation loop. Myopic sampling, i.e. step-wise selection of
 the path (on the graph) which reduces the information criterion the
@@
 designs that improve the classification of water masses as a means to
-characterize the frontal zone.
+characterize the frontal zone and thereby the extent of the river plume.
 
@@
 $\xi_T(\bx)$ denotes the temperature (in $^o C$) and $\xi_S(\bx)$
-denotes the salinity (in $mg/l$) at location $\bx$. The generalization
-to more than two variables is given in Section \ref{gen_k}. The
+denotes the salinity (in $g/kg$) at location $\bx$. The generalization
+to more than two variables is given in Section \ref{gen_k}, and the
 adaptive sampling strategies are described in Section
@@
 reduction achieved by sampling as in Eq. (\ref{sur}). In this sense,
-our work is similar to \cite{bect2012}, \cite{chevalier2014fast} and
+our work is similar to \cite{bect2012}, \cite{chevalier2014fast}, and
 \cite{azzimonti2016quantifying} who describe analytical results for
@@
 %Here, $\pi(\bxi)$ is the joint density function of length $2 n$ vector $\bxi$, and
-The length $2 n$ vector $\bmu$ contains the expected values of
+The length $2n$ vector $\bmu$ contains the expected values of
 temperature and salinity at the grid locations, while the
@@
 sampling to adapt survey plans based on data and statistical methods
-in accordance with the sense-plan-act loop (Fig. \ref{fig:envir1}). We
+in accordance with the sense-plan-act control loop (Fig. \ref{fig:envir1}). We
 will therefore present a few adaptive strategies for robotic sampling
@@
 
-An adaptive AUV survey is split into many rounds (stages). The selection is made on the graph and defined by the nearest grid
-nodes in the domain $\mathcal{M}_g$.
+An adaptive AUV survey is split into many rounds (stages). The selection --on where to sample-- is made on the graph and defined by the nearest grid nodes in the domain $\mathcal{M}_g$.
 
@@
 The mathematical expression for the optimal design then involves a
-series of intermixed maximizations over designs and integrals over
+series of intermixed maximisations over designs and integrals over
 data. In practice, the optimal solution is intractable because of
@@
 \subsection{Myopic Path Planning}
-\label{myopic}
+\label{sec:myopic}
 
@@
 \subsection{Look-ahead Trajectory Planning}
-\label{LA}
+\label{sec:LA}
 
@@
 $\beta_{S,1}=0.1$ for the slope, with west coordinate and $\beta_{T,0}=5.8$
-and $\beta_{S,1}=29.0$ for the intercept at a reference start map coordinate.
+and $\beta_{S,1}=29.0$ for the intercept at a reference map coordinate.
 
@@
 \textit{static\_north}, \textit{static\_east}, and
-\textit{static\_zigzag} with the three sequential approaches
-\textit{naive}, \textit{myopic}, and \textit{look-ahead}. Recall that
-the static sampling paths are pre-determined and cannot be altered.
-These represent the pre-planned strategies used in most current AUV
+\textit{static\_zigzag} with the three described sequential approaches
+\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static sampling paths are pre-determined and cannot be altered and represent the pre-planned strategies used in most current AUV
 operational survey designs.
@@
 \textit{myopic} and \textit{look-ahead} strategies do well, clearly
-beating the \textit{naive} strategy. The \textit{static\_zigzag}
-design is usually the best among the static designs for these
-predictive performance measures because its pattern gives more
-effective coverage of the domain when there is significant spatial
-correlation, as indicated in this case.
+beating the \textit{naive} strategy.
 
@@
 \subsection{Model Specification}
-
+\label{sec:exp_modeling}
 The statistical model parameters were specified based on a short
@@
 and salinity residuals after the westerly trends in both salinity and
-temperature are subtracted from the data. This scatterplot of joint
+temperature are subtracted from the data. This scatter-plot of joint
 residuals indicates larger variability in salinity than in
@@
 \citep{py10,Rajan12,Rajan12b}, running an instance of the
-\textit{myopic} strategy from Section \ref{sec:sampling_designs} to
+\textit{myopic} strategy from Section \ref{sec:myopic} to
 control the AUV and decide between sampling locations.
@@
 measurements and an updated model, the AUV triggers the myopic
-strategy to evaluate all possible design criteria (see
+strategy to evaluate the different design criteria (see
 Eq. \eqref{critSEQ}). The waypoint-and-path combination that is
-expected to reduce the IBV the most is selected; upon arrival this
+expected to reduce the IBV the most is selected, and upon arrival this
 procedure is then repeated. At each stage, it takes the AUV about 30
@@
 The AUV was set to start in the south-center part of the waypoint
-graph, with the previously outlined GP model of the environment. A
-survey was set to take approximately 40 minutes to visit 15 waypoints
+graph, with the previously outlined GP model of the environment (Section \ref{sec:exp_modeling}). A
+survey was set to take approximately 40 minutes, visiting 15 waypoints
 on the grid, with the AUV running near the surface to capture the
@@
 data regularly, and the GP model assimilated temperature and salinity
-data with an update frequency of 30 seconds, with about three updates
+data with an update frequency of 30 seconds, giving about three updates
 per stage.
@@
 first calculated in closed form for the situation with a static
-design, and then extended to the adaptive situation; the sequential
+design, and then extended to the adaptive situation. The sequential
 derivations provide new insights into efficient applications of

### 2019-10-09 main_new.tex
Minor changes & rephrased parts around sensitivity on simulation results.

@@
 \usepackage{booktabs}                     % horizontal lines in tables
+\usepackage{comment}
 
@@
   temperature and salinity variables, providing a description of
-the boundary uncertainty that can be used prioritize sampling efforts at locations that are
+the boundary uncertainty that can be used to prioritize sampling efforts at locations that are
 ambiguous, making exploration more effective. We show how these can be
@@
 goal of experimental sampling is to improve this
-characterization. %, as quantified via changes in the EPs.
+characterization, as quantified via changes in the EPs.
 
@@
 
-The derived closed-form solutions for the expected IBV are still
-important building blocks when constructing the sampling designs, as
-they will be used to score the different adaptive designs. % Efficient
+%The derived closed-form solutions for the expected IBV are still important building blocks when constructing the sampling designs, as they will be used to score the different adaptive designs.
+% Efficient
 % calculation is important for using adaptive survey designs with
@@
 = \begin{bmatrix} \beta_{T,0} + \beta_{T,1} x_{\mbox{West}} \\
-\beta_{T,1} + \beta_{S,1} x_{\mbox{West}}
+\beta_{S,0} + \beta_{S,1} x_{\mbox{West}}
 \end{bmatrix}.
@@
 masses pulled to the east. The regression parameters are  $\beta_{T,1}=0.065$ and
-$\beta_{S,1}=0.1$ for the slope, with west coordinate and $\beta_{T,0}=5.8$
-and $\beta_{S,1}=29.0$ for the intercept at a reference map coordinate.
+$\beta_{S,1}=0.1$ for the slopes with west coordinate, and $\beta_{T,0}=5.8$
+and $\beta_{S,0}=29.0$ for the intercept at a chosen reference map coordinate.
 
@@
 
-For a fixed survey length, the expected IBV is analytically available
-\kc{I don't understand what ``analytically available''
-  means}. However, for the sequential approaches this is not the
+For a fixed survey length, a closed form expression for the expected IBV is available as in Eq. \eqref{sur}. However, for the sequential approaches this is not the
 case. For comparison the properties are therefore evaluated using
@@
 an evaluation of the next node to sample is conducted at the end of
-the edge. \kc{Note there are plenty of language errors related to
-  tense. Pl. do not mess with any more changes from here onwards.}
+the edge.
 
@@
 look-ahead. There are several reasons for this. First, a greedy
-approach will tend to put more emphasis on locations close to the
+approach will tend to put more emphasis on promising locations close to the
 agent, which may lead away from the centre. Second, as the agent
@@
 
-%We studied the sensitivity of the results by modifying the input parameters to different prior correlations between temperature and salinity, standard deviations, and spatial correlation range.
+We studied the sensitivity of the results by modifying the input parameters to different prior correlations between temperature and salinity, standard deviations, and spatial correlation range.
 In all runs,  the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of IBV, and much better than \textit{naive}.
@@
 
+\begin{comment}
 \subsection{Performance Metrics}
@@
 
-\newpage
+\end{comment}
 

### 2019-10-09 main_new.tex
Updated small fixes on wordings

@@
 rely on a statistical model of the environment, but rather adapts
-based on closing the sensing and actuation loop. Myopic sampling, i.e. step-wise selection of
+based on closing the sensing and actuation loop. Myopic sampling, i.e. stage-wise selection of
 the path (on the graph) which reduces the information criterion the
@@
 static manner (where the survey design is pre-planned), or in a
-sequential way with several rounds of data gathering. For the latter,
-a round of data is gathered first, and then assimilated to get an
+sequential way with several stages of data gathering. For the latter,
+a stage of data is gathered first, and then assimilated to get an
 updated model. Then the updated model is used to design and select the
-next round of measurements, and so on (Section
-\ref{sec:heuristics}). Below, we describe the model for one round of
+next stage of measurements, and so on (Section
+\ref{sec:heuristics}). Below, we describe the model for one stage of
 data and the Gaussian formula for a single update.
@@
 $\by=(y_{T,1},y_{S,1},\ldots,y_{T,n_y},y_{S,n_y})^t$, representing the
-temperature and salinity measurements gathered at the first round of
+temperature and salinity measurements gathered at the first stage of
 nodes in the grid. The conditional model for the data, given the true temperature
@@
 
-An adaptive AUV survey is split into many rounds (stages). The selection --on where to sample-- is made on the graph and defined by the nearest grid nodes in the domain $\mathcal{M}_g$.
+An adaptive AUV survey is split into many stages. The selection --on where to sample-- is made on the graph and defined by the nearest grid nodes in the domain $\mathcal{M}_g$.
 
@@
 and covariance matrix $\bA^{s-1}=\mbox{Var}(\bxi|\mathcal{Y}^{s-1})$.
-The next round of measurements $\by^{j,s}$ can be gathered with
+The next stage of measurements $\by^{j,s}$ can be gathered with
 design $d^{j,s}$, $j=1,\ldots,J$. The {\it{naive}} strategy then
@@
 The myopic (greedy) strategy which we present here is optimal if we
-imagine taking only one more round of measurements. In this selection
+imagine taking only one more stage of measurements. In this selection
 strategy, there is no anticipation of what the subsequent designs might
-offer beyond the first round.
+offer beyond the first stage.
 
 Based on the currently available data $\mathcal{Y}^{s-1}$, we
-fit an updated GP model.  The next round of measurements
+fit an updated GP model.  The next stage of measurements
 $\by^{j,s}$, can be gathered with design $d^{j,s}$, $j=1,\ldots,J$. The
@@
 for expected IBV is hence applicable in Eq. (\ref{critSEQ}) using the
-updated GP model from step $s-1$. Once the data is collected for the
+updated GP model from stage $s-1$. Once the data is collected for the
 best design, the GP model is updated again. The mean $\bm^{s}$ and
@@
 We now extend the myopic strategy to a look-ahead strategy, which is
-optimal when one can gather only two more rounds of measurements. In
-addition to the next round of measurements $\by^{j,s}$, this
+optimal when one can gather only two more stages of measurements. In
+addition to the next stage of measurements $\by^{j,s}$, this
 look-ahead strategy, anticipates the subsequent design $j_2$ with
@@
 \caption{Simulation results from 100 replicate simulations for 10
-  sampling choices/steps on the grid.} 
+  sampling choices/stages on the grid.} 
 \label{fig:sim_results}
@@
   \caption{The overlaid route choices (strategies), superimposed on
-    the 100 replicate surveys with 10 sampling choices/steps.}
+    the 100 replicate surveys with 10 sampling choices/stages.}
 \label{fig:route_choices}
@@
 \textit{look-ahead} strategies perform the best in terms of expected
-variance in the ES. Surprisingly, the \textit{myopic} strategy
+variance in the ES. The \textit{myopic} strategy
 performs slightly better than the \textit{look-ahead}. Part of the
 reason for this discrepancy can be traced to the act of pruning some
-of the paths for the \textit{look-ahead} strategy; however, even with
-less pruning, the \textit{look-ahead} does not outperform the
-\textit{myopic} approach. The \textit{static\_north} design is clearly
+of the paths for the \textit{look-ahead} strategy. The \textit{static\_north} design is clearly
 better than the other two static designs, the reason being that the
@@
 
-While more effort could be spent on approximating the look-ahead steps
+While more effort could be spent on approximating the look-ahead stages
 in the adaptive designs, it is perhaps more interesting to explore the

### 2019-10-10 main_new.tex
updates

@@
   combination of statistics and autonomous robotics provide new
-  opportunities for studying experimental design. In this work we develop
-  methods for efficient spatial sampling applied to the mapping of
-  coastal processes by providing informative descriptions of spatial
-  characteristics of ocean phenomena. Specifically, we define a design
-  criterion based on reducing uncertainty in the joint excursions of
-  random fields, and derive analytical expressions for the expected
-  Bernoulli variance reduction in the excursion set of spatial
-  temperature and salinity variables, providing a description of
-the boundary uncertainty that can be used to prioritize sampling efforts at locations that are
-ambiguous, making exploration more effective. We show how these can be
-  applicable in static and sequential designs, building on myopic and
-  non-myopic sampling strategies. We use simulations to study the
-  properties and differences between state-of-the-art approaches,
-  followed by results from field deployments with an autonomous
-  underwater vehicle as part of a case study mapping the boundary of a
-  river plume.
-The results demonstrate the potential of combining statistical methods and robotic platforms, and the abilities of the derived methods to effectively inform and execute data-driven environmental sampling.
+  opportunities for studying experimental design. In this work we
+  develop methods for efficient spatial sampling applied to the
+  mapping of coastal processes by providing informative descriptions
+  of spatial characteristics of ocean phenomena. \kc{Specifically, we
+    define a design criterion based on reducing uncertainty in the
+    joint excursions of random fields, and derive analytical
+    expressions for the expected Bernoulli variance reduction in the
+    excursion set of spatial temperature and salinity variables,
+    providing a description of the boundary uncertainty that can be
+    used to prioritize sampling efforts at locations that are
+    ambiguous, making exploration more effective ---- TOO LONG. Break
+    it up}. We show how these can be applicable in static and
+  sequential designs, building on myopic and non-myopic sampling
+  strategies. We use simulations to study the properties and
+  differences between state-of-the-art approaches, followed by results
+  from field deployments with an autonomous underwater vehicle as part
+  of a case study mapping the boundary of a river plume. The results
+  demonstrate the potential of combining statistical methods and
+  robotic platforms, and the abilities of the derived methods to
+  effectively inform and execute data-driven environmental sampling.
 

### 2019-10-14 main_new.tex
Minor edits in Intro+near Fig 10. Trygve; fix p(x;y) in legends of Fig 10c,d

--- 2019-10-09 main_new.tex
@@
 \author{\fnms{Jo} \snm{Eidsvik}\thanksref{t3}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
-\author{\fnms{David} \snm{Ginsbourger}\thanksref{t4}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
+\author{\fnms{David} \snm{Ginsbourger}\thanksref{t4,t5}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
 \and
-\author{\fnms{Kanna} \snm{Rajan}\thanksref{t2,t5,t6}. \ead[label=e4]{kanna.rajan@ntnu.no}}
+\author{\fnms{Kanna} \snm{Rajan}\thanksref{t2,t6,t7}. \ead[label=e4]{kanna.rajan@ntnu.no}}
 
@@
 \affiliation[t4]{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland.}
-\affiliation[t5]{Department of Engineering Cybernetics, NTNU}
-\affiliation[t6]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
+\affiliation[t5]{Department of Mathematics and Statistics, Uncertainty of Bern, Switzerland.}
+\affiliation[t6]{Department of Engineering Cybernetics, NTNU}
+\affiliation[t7]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
 
@@
   for marine science and maritime management. Faced with limited
-  resources to understand dynamic processes in the water-column, the
+  resources to understand processes in the water-column, the
   combination of statistics and autonomous robotics provide new
-  opportunities for studying experimental design. In this work we develop
-  methods for efficient spatial sampling applied to the mapping of
-  coastal processes by providing informative descriptions of spatial
+  opportunities for experimental designs. In this work we develop
+  methods for efficient spatial sampling applied to the mapping of coastal processes by providing informative descriptions of spatial
   characteristics of ocean phenomena. Specifically, we define a design
-  criterion based on reducing uncertainty in the joint excursions of
-  random fields, and derive analytical expressions for the expected
-  Bernoulli variance reduction in the excursion set of spatial
-  temperature and salinity variables, providing a description of
-the boundary uncertainty that can be used to prioritize sampling efforts at locations that are
-ambiguous, making exploration more effective. We show how these can be
-  applicable in static and sequential designs, building on myopic and
-  non-myopic sampling strategies. We use simulations to study the
+  criterion based on reducing uncertainty in the joint excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected
+  Bernoulli variance reduction in such a framework. We demonstrate how these provide a description of
+the boundary uncertainty in temperature and salinity variables, which is used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective by building on myopic and
+  non-myopic strategies. We use simulations to study the
   properties and differences between state-of-the-art approaches,
@@
   underwater vehicle as part of a case study mapping the boundary of a
-  river plume.
-The results demonstrate the potential of combining statistical methods and robotic platforms, and the abilities of the derived methods to effectively inform and execute data-driven environmental sampling.
+  river plume. The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.
 
@@
   cannot deliver the computational power required, hence statistical
-  proxy models of the environment must be used for doing inference
-  about where to sample.
-
-\item As a priori information about the state of the ocean is
-  uncertain, there is a substantial efficiency gain in reacting to
-  information obtained from measurements taken in-situ. Recently
-  acquired information can be used to resolve this by assimilating up-to-date data into statistical models that
+  proxy models of the environment must be used for learning where to sample.
+
+\item With limited available information about the state of the ocean, there is substantial value in reacting to
+  information obtained from measurements taken in-situ. This
+  acquired information must be assimilated into statistical models that
   can be used to inform decisions on where to sample
@@
   number of choices (i.e. locations, trajectories, and candidate
-  designs) makes the problem combinatorially large, creating a
+  designs) is enormous, creating a
   trade-off between optimization (finding the most resource-efficient
@@
 probabilities (EP) for the spatial design of experiments. The goal is
-to design sampling strategies for reducing the spatially-integrated
+to design sampling strategies for for improving the estimation of joint ES with respect to temperature and salinity by reducing the spatially integrated
 Bernoulli variance (IBV) in the ES indicators. Our main contributions
@@
 
-Our objective is to leverage statistical tools to characterize a river
+Our objective is to leverage statistical approaches to characterize a river
 plume, focusing on spatial separation of cold freshwater from the
@@
 $\delta_{ij}=0$ otherwise. Moreover, we assume stationary isotropic
-random fields a priori where the correlation function $\rho(\bx,\bx')$
+random fields where the correlation function $\rho(\bx,\bx')$
 solely depends on $\bx$ and $\bx'$ via Euclidean distance
@@
 The multivariate cumulative probabilities in two and four dimensions 
-are computationally efficient using methods developed by
+are available using methods developed by
 \cite{genz2009computation}.  In the end, the first part in
@@
 %\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
-\caption{Results from mapping the Nidelva river. \ref{fig:map} Show an
-  overview of the survey area overlaid with the AUV path in black and
-  dashed line. Note the shaded region indicating the usual frontal
-  region. \ref{fig:res_both} Presents the collected temperature data
-  as colored trails. Note waypoint 5 (WP5) which indicate where the
-  two surveys deviate. \ref{fig:res1} and \ref{fig:res2} Shows the
-  collected salinity data (as a colored trail) overlaid on the final
-  EP, which indicate the AUVs statistical impression of the front. For
+\caption{Results from mapping the Nidelva river. \ref{fig:map} shows an
+  overview of the survey area overlaid with the AUV paths in black and
+  dashed line. The shaded region indicates the usual frontal
+  region. \ref{fig:res_both} presents the collected temperature data
+  as colored trails. Waypoint 5 (WP5) is where the
+  two surveys deviate. \ref{fig:res1} and \ref{fig:res2} show the
+  collected salinity data in the two surveys (as colored trails) overlaid on the final
+  EPs, which indicate the AUVs statistical impression of the front in the surveys. For
   both missions the temperature and salinity data correspond with the
@@
 Two missions, Surveys 1 and 2, were run successively from 11:00 AM to
-01:00 PM, with a short break in between. The resulting path of the
+01:00 PM, with a break in between. The resulting path of the
 selected waypoints are shown in the map in Fig. \ref{fig:map}, both
 within the expected frontal region (shaded pink). The recorded
-temperatures are shown as colored trails in Fig. \ref{fig:res_both},
+temperatures in the two surveys are shown as colored trails in Fig. \ref{fig:res_both},
 clearly indicating the temperature difference between fjord and river
 waters. The salinity data are then shown separately, overlaid with the
-estimated EP for each survey in Fig. \ref{fig:res1} and
+estimated EPs for each survey in Fig. \ref{fig:res1} and
 \ref{fig:res2}.
@@
 crossing the frontal boundary multiple times. As conditions changed
-slightly between the two deployments, the resulting path (after
-waypoint 5) is shown to deviate. Survey 1 continued northwards,
+slightly between the two deployments, the resulting paths (after
+waypoint 5) are shown to deviate. Survey 1 continued northwards,
 tracking the north-eastern part of the front, while Survey 2 turned
@@
 
-The final predictions of the front location given in
-Fig. \ref{fig:res1} and \ref{fig:res2} as a dashed line, correspond
-with one another, yielding a picture of the front gradually bending
-off toward the north east. The amount of exploration done by Survey 1
+The final predictions of the front location, represented by conditional EPs in
+Fig. \ref{fig:res1} and \ref{fig:res2} as dashed lines, correspond
+with one another.
+In both surveys they yield a picture of the front being a bit to the west in the southern parts of the domain and gradually bending off toward the north east.
+The amount of exploration done by Survey 1
 is greater than Survey 2. In Survey 1, the AUV obtained more detail by
-going north, while Survey 2, coming close to the survey area borders
+going north from waypoint 5, while Survey 2, coming close to the survey area borders
 in the south-western corner, obtained a poorer understanding of the
-northern parts. A longer planning horizon would likely identify and
+northern parts. A look-ahead strategy might identify and
 discourage such choices.

### 2019-10-14 main_new.tex
Updated abstract

--- 2019-10-09 main_new.tex
@@
 \begin{abstract}
-  Improving and optimizing oceanographic sampling is a crucial task
-  for marine science and maritime management. Faced with limited
-  resources to understand dynamic processes in the water-column, the
-  combination of statistics and autonomous robotics provide new
-  opportunities for studying experimental design. In this work we develop
-  methods for efficient spatial sampling applied to the mapping of
-  coastal processes by providing informative descriptions of spatial
-  characteristics of ocean phenomena. Specifically, we define a design
-  criterion based on reducing uncertainty in the joint excursions of
-  random fields, and derive analytical expressions for the expected
-  Bernoulli variance reduction in the excursion set of spatial
-  temperature and salinity variables, providing a description of
-the boundary uncertainty that can be used to prioritize sampling efforts at locations that are
-ambiguous, making exploration more effective. We show how these can be
-  applicable in static and sequential designs, building on myopic and
-  non-myopic sampling strategies. We use simulations to study the
-  properties and differences between state-of-the-art approaches,
-  followed by results from field deployments with an autonomous
-  underwater vehicle as part of a case study mapping the boundary of a
-  river plume.
-The results demonstrate the potential of combining statistical methods and robotic platforms, and the abilities of the derived methods to effectively inform and execute data-driven environmental sampling.
+ \begin{abstract}
+  Improving and optimizing oceanographic sampling is a crucial task for marine science and maritime management.
+Faced with limited resources to understand processes in the water-column, the combination of statistics and autonomous robotics provide new opportunities for experimental designs.
+In this work we develop methods for efficient spatial sampling applied to the mapping of coastal processes by providing informative descriptions of spatial characteristics of ocean phenomena.
+Specifically, we define a design criterion based on reducing uncertainty in the joint excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected Bernoulli variance reduction in such a framework.
+We demonstrate how these provide a description of the boundary uncertainty in temperature and salinity variables, which is used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective by building on myopic and non-myopic strategies.
+We use simulations to study the properties and differences between state-of-the-art approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a case study mapping the boundary of a river plume.
+The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.
 

### 2019-10-15 main_new.tex
Updated figure 10
Updates Jo

@@
 \begin{abstract}
- \begin{abstract}
+
   Improving and optimizing oceanographic sampling is a crucial task for marine science and maritime management.
-Faced with limited resources to understand processes in the water-column, the combination of statistics and autonomous robotics provide new opportunities for experimental designs.
+Faced with limited resources to understand processes in the water-column, the combination of statistics and autonomous robotics provides new opportunities for experimental designs.
 In this work we develop methods for efficient spatial sampling applied to the mapping of coastal processes by providing informative descriptions of spatial characteristics of ocean phenomena.
 Specifically, we define a design criterion based on reducing uncertainty in the joint excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected Bernoulli variance reduction in such a framework.
-We demonstrate how these provide a description of the boundary uncertainty in temperature and salinity variables, which is used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective by building on myopic and non-myopic strategies.
-We use simulations to study the properties and differences between state-of-the-art approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a case study mapping the boundary of a river plume.
+We demonstrate how this criterion gives a description of the boundary uncertainty in temperature and salinity variables, which is used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective by building on myopic and non-myopic strategies.
+We use simulations to study the properties of methods and to compare them with state-of-the-art approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a case study mapping the boundary of a river plume.
 The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.
@@
   online if run on robotic sensing platforms, as onboard computers
-  cannot deliver the computational power required, hence statistical
-  proxy models of the environment must be used for doing inference
-  about where to sample.
-
-\item As a priori information about the state of the ocean is
-  uncertain, there is a substantial efficiency gain in reacting to
-  information obtained from measurements taken in-situ. Recently
-  acquired information can be used to resolve this by assimilating up-to-date data into statistical models that
+  cannot deliver the computational power required. Hence statistical
+  proxy models of the environment must be used for learning where to sample.
+
+\item With limited available information about the state of the ocean, there is substantial value in reacting to
+  information obtained from measurements taken in-situ. This
+  acquired information must be assimilated into statistical models that
   can be used to inform decisions on where to sample
@@
   number of choices (i.e. locations, trajectories, and candidate
-  designs) makes the problem combinatorially large, creating a
+  designs) is enormous, creating a
   trade-off between optimization (finding the most resource-efficient
-  design to collect necessary data), and computability (arriving at a
+  design to collect necessary data) and computability (arriving at a
   solution in reasonable time). To successfully resolve features, this
@@
 
-\begin{figure}[!b] 
+\begin{figure}[!h] 
   \centering 
@@
 wind, and temperature differences, this boundary often gets
-distorted. Prior knowledge about the location and evolution of these
+distorted. Initial knowledge about the location and evolution of these
 features is therefore highly uncertain, making deterministic planning
@@
 The Gaussian model is used to represent the bivariate random fields of
-temperature and salinity. At location $\bx \in \mathcal{M}$, the prior
+temperature and salinity. At location $\bx \in \mathcal{M}$, the
 bivariate distribution is:
@@
 
-Table \ref{tab:sim_rhoab} shows the EPs and the associated prior Bernoulli
+Table \ref{tab:sim_rhoab} shows the initial EPs and the associated Bernoulli
 variance (second row) for the examples indicated in
@@
 We use a bivariate GP model for temperature and salinity, where we
-specify the prior mean:
+specify the mean:
 \begin{equation}\label{m}
@@
 temperature is shown in Fig. \ref{fig:true_temp} and
-\ref{fig:true_sal}. The true ES from this prior realization is shown
+\ref{fig:true_sal}. The true ES is a realization from this model as shown
 in Fig. \ref{fig:ESet}.
@@
 case. For comparison the properties are therefore evaluated using
-Monte Carlo sampling over several replicates of realizations from the
-prior while conducting simulated sequential surveys for each
+Monte Carlo sampling over several replicates of realizations from the model while conducting simulated sequential surveys for each
 one. Fig. \ref{fig:Eprob} shows the conditional EP, given data
@@
 evaluates the impact of locations further away (look-ahead) where
-assimilated data has less predictive power, the GP prior model (which
+assimilated data has less predictive power, the GP model (which
 is centered here) will act to restrict paths deviating from the
@@
 
-We studied the sensitivity of the results by modifying the input parameters to different prior correlations between temperature and salinity, standard deviations, and spatial correlation range.
+We studied the sensitivity of the results by modifying the input parameters to have different correlations between temperature and salinity, standard deviations, and spatial correlation range.
 In all runs,  the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of IBV, and much better than \textit{naive}.
-The \textit{look-ahead} strategy seems to be substantially better than the \textit{myopic} only for very small prior standard deviation or very large spatial correlation range.
+The \textit{look-ahead} strategy seems to be substantially better than the \textit{myopic} only for very small initial standard deviation or very large spatial correlation range.
 The \textit{static\_north} design continues to be the best static design for IBV, while \textit{static\_zigzag} is the best design for the other predictive performance measures, especially so with large spatial correlation range.
-We also ran simulation studies with only temperature data, and for realistic correlation levels between temperature and salinity the effects are not much worse in terms of IBV when only temperature data are available.
+We also ran simulation studies with only temperature data, and for realistic correlation levels between temperature and salinity the IBV results are not much worse when only temperature data are available.
 In addition to the comparison made in Table \ref{tab:sim_rhoab}, the current setting includes spatial correlation and this likely reduces the additional influence of having bivariate data.
@@
 
-The tables below explore the sensitivity of the result to different
-prior correlation values between temperature measured onboard the AUV,
-temperature and salinity, prior standard deviations, and spatial
+The tables below explore the sensitivity of the result to different input correlation values between temperature and salinity measured onboard the AUV, standard deviations, and spatial
 correlations (see table rows). We only compare results at the end of
@@
 The statistical model parameters were specified based on a short
-preliminary survey where the AUV made an initial transect to determine
-environmental conditions and correlation structure. Based on the
+preliminary survey where the AUV made an initial transect to determine the trends in environmental conditions and correlation structures. Based on the
 initial data, the trend parameters were estimated by linear
@@
 
-Fig. \ref{fig:parest} shows diagnostic plots of this
+Fig. \ref{fig:parest} summarizes diagnostic plots of this
 analysis. Fig. \ref{fig:parest_a} shows a cross-plot of temperature
@@
 %\end{figure}
-
+\newpage
 \section{Closing remarks}\label{sec:concl_disc}

### 2019-10-17 main_new.tex
Update

@@
 Specifically, we define a design criterion based on reducing uncertainty in the joint excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected Bernoulli variance reduction in such a framework.
-We demonstrate how this criterion gives a description of the boundary uncertainty in temperature and salinity variables, which is used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective by building on myopic and non-myopic strategies.
+We demonstrate how this criterion gives a description of the boundary uncertainty, that can be used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective.
 We use simulations to study the properties of methods and to compare them with state-of-the-art approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a case study mapping the boundary of a river plume.
@@
   can be used to inform decisions on where to sample
-  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.
+  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.
 
@@
 \end{equation}
-The spatial correlation is of a Matern type \kc{cite or define}:
+The spatial correlation is of a Matern type \citep{Matern2013}:
 \begin{equation}\label{v}

### 2020-04-23 ans_to_reviewers/answers-to-reviewers.tex
Updated with Ans-to-Rev

\documentclass[a4paper]{article}
\topmargin	-10mm
\oddsidemargin	0mm
\textwidth	150mm
\textheight	230mm

\newcommand{\ba}{ {\boldsymbol a} }
\newcommand{\bA}{ {\boldsymbol A} }
\newcommand{\bb}{ {\boldsymbol b} }
\newcommand{\bB}{ {\boldsymbol B} }
\newcommand{\bc}{ {\boldsymbol c} }
\newcommand{\bC}{ {\boldsymbol C} }
\newcommand{\bd}{ {\boldsymbol d} }
\newcommand{\bD}{ {\boldsymbol D} }
\newcommand{\be}{ {\boldsymbol e} }
\newcommand{\bE}{ {\boldsymbol E} }
\newcommand{\boldf}{ {\boldsymbol f} }
\newcommand{\bF}{ {\boldsymbol F} }
\newcommand{\bg}{ {\boldsymbol g} }
\newcommand{\bG}{ {\boldsymbol G} }
\newcommand{\bh}{ {\boldsymbol h} }
\newcommand{\bH}{ {\boldsymbol H} }
\newcommand{\bi}{ {\boldsymbol i} }
\newcommand{\bI}{ {\boldsymbol I} }
\newcommand{\bj}{ {\boldsymbol j} }
\newcommand{\bJ}{ {\boldsymbol J} }
\newcommand{\bk}{ {\boldsymbol k} }
\newcommand{\bK}{ {\boldsymbol K} }
\newcommand{\bl}{ {\boldsymbol l} }
\newcommand{\bL}{ {\boldsymbol L} }
\newcommand{\bm}{ {\boldsymbol m} }
\newcommand{\bM}{ {\boldsymbol M} }
\newcommand{\bn}{ {\boldsymbol n} }
\newcommand{\bN}{ {\boldsymbol N} }
\newcommand{\bo}{ {\boldsymbol o} }
\newcommand{\bO}{ {\boldsymbol O} }
\newcommand{\bp}{ {\boldsymbol p} }
\newcommand{\bP}{ {\boldsymbol P} }
\newcommand{\bq}{ {\boldsymbol q} }
\newcommand{\bQ}{ {\boldsymbol Q} }
\newcommand{\br}{ {\boldsymbol r} }
\newcommand{\bR}{ {\boldsymbol R} }
\newcommand{\bs}{ {\boldsymbol s} }
\newcommand{\bS}{ {\boldsymbol S} }
\newcommand{\bt}{ {\boldsymbol t} }
\newcommand{\bT}{ {\boldsymbol T} }
\newcommand{\bu}{ {\boldsymbol u} }
\newcommand{\bU}{ {\boldsymbol U} }
\newcommand{\bv}{ {\boldsymbol v} }
\newcommand{\bV}{ {\boldsymbol V} }
\newcommand{\bw}{ {\boldsymbol w} }
\newcommand{\bW}{ {\boldsymbol W} }
\newcommand{\bx}{ {\boldsymbol x} }
\newcommand{\bX}{ {\boldsymbol X} }
\newcommand{\by}{ {\boldsymbol y} }
\newcommand{\bY}{ {\boldsymbol Y} }
\newcommand{\bz}{ {\boldsymbol z} }
\newcommand{\bZ}{ {\boldsymbol Z} }
\newcommand{\vc}[1]{\mbox{\boldmath $#1$}}
\newcommand{\balph}{ {\boldsymbol \alpha} }
\newcommand{\balpha}{ {\boldsymbol \alpha} }
\newcommand{\bbet}{ {\boldsymbol \beta} }
\newcommand{\bbeta}{ {\boldsymbol \beta} }
\newcommand{\bgam}{ {\boldsymbol \gamma} }
\newcommand{\bgamma}{ {\boldsymbol \gamma} }
\newcommand{\bGamma}{ {\boldsymbol \Gamma} }
\newcommand{\bdelta}{ {\boldsymbol \delta} }
\newcommand{\bDelta}{ {\boldsymbol \Delta} }
\newcommand{\beps}{ {\boldsymbol \epsilon} }
\newcommand{\bepsilon}{ {\boldsymbol \epsilon} }
\newcommand{\bphi}{ {\boldsymbol \phi} }
\newcommand{\bPhi}{ {\boldsymbol \Phi} }
\newcommand{\bpi}{ {\boldsymbol \pi} }
\newcommand{\bpsi}{ {\boldsymbol \psi} }
\newcommand{\bkap}{ {\boldsymbol \kappa} }
\newcommand{\bkappa}{ {\boldsymbol \kappa} }
\newcommand{\bKappa}{ {\boldsymbol \Kappa} }
\newcommand{\blam}{ {\boldsymbol \lambda} }
\newcommand{\blambda}{ {\boldsymbol \lambda} }
\newcommand{\bLambda}{ {\boldsymbol \Lambda} }
\newcommand{\bmu}{ {\boldsymbol \mu} }
\newcommand{\bMu}{ {\boldsymbol \Mu} }
\newcommand{\bet}{ {\boldsymbol \eta} }
\newcommand{\bome}{ {\boldsymbol \omega} }
\newcommand{\bomega}{ {\boldsymbol \omega} }
\newcommand{\bOmega}{ {\boldsymbol \Omega} }
\newcommand{\bnabla}{ {\boldsymbol \nabla} }
\newcommand{\brho}{ {\boldsymbol \rho} }
\newcommand{\bsigma}{ {\boldsymbol \sigma} }
\newcommand{\bSig}{ {\boldsymbol \Sigma} }
\newcommand{\bSigma}{ {\boldsymbol \Sigma} }
\newcommand{\btheta}{ {\boldsymbol \theta} }
\newcommand{\bTheta}{ {\boldsymbol \Theta} }
\newcommand{\bzeta}{ {\boldsymbol \zeta} }
\newcommand{\bPsi}{ {\boldsymbol \Psi} }
\newcommand{\btau}{ {\boldsymbol \tau} }
\newcommand{\bzero}{ {\boldsymbol 0} }
\newcommand{\bones}{ {\boldsymbol 1} }
\newcommand{\given}{\,|\,}
\newcommand{\sS}{{\cal S}}
\newcommand{\Ss}{{\cal S}}
\newcommand{\Field}{{\cal F}}
\newcommand{\colsp}{{\cal C}}
\newcommand{\nullsp}{{\cal N}}
\newcommand{\rowsp}{{\cal R}}
\newcommand{\tildeC}{\tilde{C}}
\newcommand{\tildeK}{\tilde{K}}
\newcommand{\tildew}{\tilde{w}}
\newcommand{\tildebw}{\tilde{\bw}}
\newcommand{\tildebW}{\tilde{\bW}}
\newcommand{\calC}{{\cal C}}
\newcommand{\calcbC}{{\bf {\cal C}}}

\def\rx{{\texttt{T-REX\ }}}
\def\rxe{{\texttt{T-REX}}}
\def\du{{\texttt{DUNE}\ }}
\def\due{{\texttt{DUNE}}}

% == Maths
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage[normalem]{ulem} %to strike the words

\newcounter{reviewer}
\def\reply{\textbf{Reviewer query}}
\def\action{\textbf{Action taken}}

\newlist{answers}{enumerate}{10}
\setlist[answers]{leftmargin=*, label=\arabic{reviewer}.\arabic*}

\newcommand{\edcomment}[1]{{\color{green}{\{Editor: #1\}}}}
\newcommand{\frevcomment}[1]{{\color{blue}{\{Rev 1: #1\}}}}
\newcommand{\srevcomment}[1]{{\color{red}{\{Rev 2: #1\}}}}
\newcommand{\trevcomment}[1]{{\color{violet}{\{Rev 3: #1\}}}}

\begin{document}

%----------------------------------------------------------------------
\title{Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets
  \\\vspace{5mm}
 Authors response
  \\\vspace{5mm}
\small{Manuscript: AOAS1910-027}}
\author{ }

\date{\today}

\maketitle

The authors are grateful to the reviewers for their comments and
suggestions. They have all been carefully considered in this revised
version of this manuscript, which we believe has significantly
improved as a result. Detailed responses to the feedback, from each reviewer and the editorial comments follow.\\

The principal revisions made to the manuscript are:
\vspace{0.5cm}

\begin{itemize}[noitemsep, topsep=0pt, parsep=0pt, partopsep=0pt]
\item \textbf{item1}: bla bla.
\item \textbf{item2}: bla bla.
\item \textbf{item3}: bla bla.
\end{itemize}
\par \vspace{1em}
Note that the reference to page $X$ is made relative to
the original submitted manuscript given to the reviewers.

%======================================================================
\section*{Answers to referee 1}
%======================================================================
All changes requested from Reviewer 1 are marked as \frevcomment{blue text} in the updated manuscript. \setcounter{reviewer}{1}
\begin{answers}

%----------------------------------------------------------------------
\item{Major comment 1}\label{q1}

\reply: Reviewer comment 4: Adjustments and improvements towards the intended audience?

\action: Our action

\begin{quote}
\frevcomment{a change in the text}
\end{quote}

\begin{quote}
  \frevcomment{another change in the text} \end{quote}

\item{Major comment 2}\label{q2}

\reply: Reviewer comment 4: On including relevant code.

\action: Our actions

\item{Major comment 3}\label{q3}

\reply: Reviewer comment 5: On shortening the paper and moving important sentences earlier.

\action: Our actions

\item{Major comment 4}\label{q4}

\reply: Reviewer comment 6: It is not clear why the binary classification and excursion sets are being used. Is there really a sharp boundary at the edge of a plume? Why not mapping the salinity and temperature fields instead? After all, on page 19, “achieving good predictive performance” is mentioned as a desirable feature of a good design. A lot of work has been done on that topic. And what about indicator kriging (along with logit regression) that generalizes indicator kriging that has been used to identify plumes? Modelling the p(x) is more like prohibit analysis-what is the advantage.

\action: Our actions

\item{Major comment 5}\label{q5}

\reply: Reviewer comment 7: The sequential sampling approaches are not novel. they are mentioned as examples in the cited paper by Bect et al. What is novel is the study of how all these methods work first in simulation studies as well as well as in the case study. That work will be a valuable contribution.

\action: Our actions

\item{Major comment 6}\label{q6}

\reply: Reviewer comment 8: While I can see some advantages of simplicity in the IBV of Bect et al for binary data, one could have used entropy as it has information-theoretical credentials. Furthermore it could cover design both for spatial mapping as well as indicator mapping. Finally model uncertainty could be included in the overall assessment of uncertainty. The more general issue here is what objective function should be optimized in the design? Are we in a multi-attribute criterion situation?

\action: Our actions

\item{Major comment 7}\label{q7}

\reply: Reviewer comment 9: On the issue of covariance separability, objections have always been raised when I have used it as its unrealistic. My rejoinder has always worked, that of citing the “all models are wrong”, George Box’s famous phrase. But what I was able to show in each case is a cross-validatory assessment that showed the model was “useful”. I wonder if such a model assessment could be made here to support the assumption of separability and other things as well. By the way, in spatial mapping at least, point predictors are pretty robust against mis-specification of the covariance. Where things go sideways is in the confidence bands or ellipsoids they produce. What are the error bands when modelling binary responses? Surely the predicted plumes have fuzzy boundaries.

\action: Our actions

\item{Major comment 8}\label{q8}

\reply: Reviewer comment 10: The strategy at the bottom of page 17 seems ad hoc. Could you not use the data collected with the two ahead design to get better estimates of parameters and then with these parameters and thereafter a better myopic design?

\action: Our actions

\item{Major comment 9}\label{q9}

\reply: Reviewer comment 11: Are these results generalizable? In other words, how much plume-to-plume variability is there? One of the problems with modelling using a lot of parameters, is there instability over replicate experiments. To put it another way which model parameters represent plume-population parameters and which represent random plume effects?

\action: Our actions

\item{Major comment 10}\label{q10}

\reply: Reviewer comment 12: The separability assumption is to be relaxed in future work. Should publication of this case study await the proposed improvement? It would make for a much better paper.

\action: Our actions

\item{Specific comments}\label{q11}

\reply: See the numbered list below:
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

\item[1] “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.
\action: Our actions
\item[2] Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?
\action: Our actions
\item[3] I don’t think Figure 3 conveys too much additional benefit and eliminating it for space might be considered.
\action: Our actions
\item[4] Why are the ES’s called ”excursion sets? Why excursion probabilities?
\action: Our actions
\item[5] I did not spot a definition for the “waypoint graph” a phrase used several times in the paper. Again points to the need to identify the audience. I doubt that most statisticians would know that phrase. 
\action: Our actions
\item[6] In statistics journals, random variables are commonly labelled by Roman characters e.g. Z, parameters in Greek e.g. $\xi$. I realize that in other scientific domains that distinction is not always made but I do recommend the change if the paper is to appear in a stats journal. 
\action: Our actions
\item[7] There is not much of novelty in the Gaussian Process theory used in the paper- its pretty much standard. I suggest that the derivations stated in detail be summarized by just the results. Statisticians will know these things and users will not care.
\action: Our actions

\end{itemize}

\end{answers}

% \newpage
%======================================================================
\section*{Answers to referee 2}
%======================================================================
All changes requested from Reviewer 2 are marked as \srevcomment{red text}
in the updated manuscript.

\setcounter{reviewer}{2}

\begin{answers}
\item{Major comment 1 - Details on AUVs}\label{q12}

\reply: I know very little about the AUVs you discuss, and I suspect many of your readers will be in a similar position. Additional contextual information would be helpful

\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

\item[1] What sort of onboard computing do they have? It may not be something like an Intel 2.1 GHz, 4-thread processor, but a more informative description of something most readers relate to would be helpful. Similar information related to RAM and hard drive space would be helpful. What are the computational limits of these machines?

\item[2] How long do the AUVs typically explore in one go? Is there a minimum/maximum time?

\item[3] What kind of movements can the AUVs make? Chess-like moves (horizontal, vertical, diagonal)?

\item[4] Is there a minimum/maximum number of points you want to sample?

\item[5] What is the size of the grid that is typically considered? i.e., what is a typical value of $n$ in Eq.(9)?

\item[6] Does the AUV use a selection algorithm from the very beginning or is there a minimum number of initial measurements that are made before the optimization algorithm begins?

\end{itemize}

\action: Our action

\begin{quote}
\srevcomment{a change in the text}
\end{quote}

\begin{quote}
  \srevcomment{another change in the text} \end{quote}

\item{Major comment 2}\label{q13}

\reply: p. 9, Eq. (8): the dependency on $\bx$ and $\bx'$ has been omitted for some of the parameters without mention. Or are the parameters constant across locations?

\action: Our actions

\item{Major comment 3}\label{q14}

\reply: p. 9, How reasonable is the assumption of stationary and isotropic random fields? You may be limited by the computational abilities of the AUVs so that more complex models are computationally infeasible. That is a valid reason to choose a simpler, less accurate model. I’m just wondering if that is the reason here.

\action: Our actions

\item{Major comment 4}\label{q15}

\reply: p. 9, Eq. (10). Are the temperature and salinity processes at a specific location effectively independent of one another? (I truly don’t know). That is what is implied by Eq. (10).

\action: Our actions

\item{Major comment 5}\label{q16}

\reply: p. 9, Eq. (11). The form of the equation is reminiscent of the Sherman-Morrison-Woodbury solution for fixed-rank kriging. $\bSigma$ may not be small here, but I wonder if there may be a computational advantage to considering an alternative computational solution.  See e.g., Cressie, N. and Johannesson, G. (2008), Fixed rank kriging for very large spatial data sets. Journal of the Royal Statistical Society: Series B (Statistical Methodology), 70: 209-226. doi:10.1111/j.1467-9868.2007.00633.x.

\action: Our actions

\item{Major comment 6}\label{q17}

\reply: p. 9, Eq. (13). Why does $\bm$ have a distribution before seeing the data? I can see how $\bm(\by)$ would have a distribution since it depends on $\by$. It almost feels Bayesian, but nothing indicates that this analysis is being pursued from a Bayesian perspective. Why is it valid for $\bm$ to have a distribution prior to seeing data?

\action: Our actions

\item{Major comment 7}\label{q18}

\reply: p. 11, Eq. (17). I’m not understanding the advantage to conditioning on $\bm$ instead of $\by$? $\bm$ is a function of $\by$. The probability is the same two-dimensional integral either way. It seems to be related to the standardization that results in $\bZ$? Please clarify.

\action: Our actions

\item{Major comment 8}\label{q19}

\reply: p. 11, Is $\bpi(\bm)$ the density associated with the distribution in Eq. (14)?

\action: Our actions

\item{Major comment 9}\label{q20}

\reply: p. 11, Is it purely for stylistic reasons that $P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\bm)P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\bm)$ isn’t written as $P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\bm)^2$? It seems to be a more natural way of expanding Eq. (18).

\action: Our actions

\item{Major comment 10}\label{q21}

\reply: p. 11, It is stated that $\bZ$ is chosen to be independent of $\bm$, so presumably the joint CDF $F_{z,m}=F_{Z}F_m$ which would certainly be convenient, but is not obvious to me why this would be true. 
Instead, it appears that $\bZ$ is chosen to be a pivotal quantity to get the simple CDF in Eq. (20). Please clarify.

\action: Our actions

\item{Major comment 11}\label{q22}

\reply: p. 11, Eq. (21). The trailing $p(\bm)$ should be $\bpi(\bm)$. Should the right side of the equality be squared?

\action: Our actions

\item{Major comment 12}\label{q23}

\reply: This method relies on computing the marginal probabilities in each grid cell. This is computationally convenient and is likely required because of the limited computing power of the AUVs. However, the optimization should technically be performed with respect to the joint distribution across all grid cells. While the context may not allow this in practice, it would be helpful to get an idea of how much error is introduced by this approximation. I recommend creating a small but realistic-as-feasible simulation study comparing the accuracy of the probabilities, a comparison of the computational expense, and examining if/how navigation decisions change based on whether one uses the marginal/joint computations.

\action: Our actions

\item{Major comment 13}\label{q24}

\reply: Fig. 7 is difficult to understand. The goal is to show the various paths taken by the AUV in the
simulation example. But it is very difficult to determine which paths are very common and which
are uncommon. Some ideas that may make this graphic more useful:

\begin{quote}
-Don’t worry about the paths but focus on the nodes visited. Represent each node using a circle or hexagon.\par
-Make the node shading darker depending on whether that node is visited in a
simulation. Nodes with no visits would be white. Nodes with 100 visits would be black.\par
-The true excursion set is a random variable, but it would be informative to overlay the
average position of the excursion set over each of the panels so the reader can see how
the simulated AUV moves back and forth across the excursion set.
\end{quote}

\action: Our actions

\item{Major comment 14}\label{q25}

\reply: While the context doesn’t allow full reproducibility of the paper’s results, no implementation details are provided. What software was used to perform the bulk of analysis? Are there scripts available that can be adapted by others for similar purposes? Please provide resources for the practical implementation of the methodology discussed in the paper.

\action: Our actions

\item{Minor comments}\label{q26}

\reply: \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

\item[-] “closed-form” is sometimes written as “closed form” (e.g., p. 7, p. 18)

\item[-] p. 9, `1s` should be ‘1s’. Similar for 0s.

\item[-] p. 15, my manuscript has -on where to sample-, which is an unusual use of dashes.

\end{itemize}

\action: Our actions

\end{answers}
%======================================================================
\section*{Answers to referee 3}
%======================================================================
All changes requested from Reviewer 3 are marked as \trevcomment{violet text}
in the updated manuscript.

\setcounter{reviewer}{3}

\begin{answers}

\item{Minor comment 1}\label{q27}

\reply: The authors indicate several paths forward for work in this area, which I think pose interesting questions. They may want to elaborate
a bit more on those.

\action: Our action

\item{Minor comment 2}\label{q28}

\reply: I think the equation $V_{m,upd}$ needs a j as a part of that subscript.

\action: Our action

\item{Major comment 3}\label{q29}

\reply: Figure 4: Why not show excursion probabilities for, say, myopic selection as well?

\action: Our actions

\end{answers}

%======================================================================
 \section*{Associate Editor}
%======================================================================
All changes requested from the editor are marked as \edcomment{green text}
in the updated manuscript.

\begin{answers}

\item{Comments to the Author}\label{q30}

\reply: Specific comments:

\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

\item[1.] General: The inferential problem considered is the paper is that of estimating the “boundary”
of a plume. However, the authors never define exactly what a “boundary” is. Such a definition
should be provided, as well as some discussion of why determining the location of a boundary
is so important. Why is it not sufficient to create a map of predictions (with an accompanying
map of prediction error variances) for the variable(s) under study throughout the region of
interest (a problem for which there is a huge literature on sampling design)?

\item[2.] Page 8, second paragraph of Section 4.1: The assumptions of separability and isotropy, and
even stationarity, are not really necessary for the methodological development of Sections 4–6.
I suggest that you wait to make these assumptions until you get to the simulation studies and
example.

\item[3.] Page 9, second paragraph of Section 4.2: It has been assumed here that the measurement
errors for temperature and salinity at any given location are uncorrelated. Is this a reasonable
assumption? This could easily be checked by taking replicate measurements at each site.

\item[4.] General: Related to the previous point, I do not understand why sampling locations are being
restricted to a regular grid. In fact, it seems unwise because it is well known (see, for example,
Chapters 9 and 22 of Handbook of Environmental and Ecological Statistics, 2019, CRC Press)
that the quality of inferences for spatial fields can be improved substantially by taking some
observations very close to others.

\item[5.] Page 13, last line: Specifically, what was the design? We aren’t even told what the sample size
($n_y$) was used in these calculations.

\item[6.] Page 18, first paragraph of Section 7.1: The numerical values of the regression parameters
given here are not meaningful without knowing the units of “West.”

\item[7.] Page 24, first paragraph of Fig. 8: Please define “the modeled quadratic form of the residuals”
more precisely. Perhaps a mathematical expression is needed for clarity.

\end{itemize}

\action: Changes made.

\item{Minor comments to the Author}\label{q31}

\reply: 
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

\item[-] Page 3, line 19: Change “bi-variate” to “bivariate”. The same change is needed on the next-
to-last line of Page 12.

\item[-] Page 9, three lines after (10): I presume that all off-diagonal elements of R are zeros, but this
should be stated explicitly.

\item[-] Page 19, fourth line from bottom: Change “as also” to “as well as”

\item[-] Page 30, second and third references listed: “et al.” is listed after the authors’ names for each
of these two references, but actually there are no additional authors beyond the two listed.

\end{itemize}

\action: Changes made.

\end{answers}

% == Adding references
\footnotesize
\bibliographystyle{apalike}
\bibliography{ref}

%======================================================================
\end{document}
%======================================================================

### 2020-04-23 cover_letter/cover_letter.tex
Small updates
Minor changes & rephrased parts around sensitivity on simulation results.
Updated with Ans-to-Rev

@@
 
-Please see our manuscript \textbf{Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets} submitted for possible publication in the \emph{Annals of Applied Statistics} (AOAS).\\*
+Please see our manuscript \textbf{Sequential uncertainty reduction on joint excursion sets of oceanographic variables with an autonomous underwater vehicle} submitted for possible publication in the \emph{Annals of Applied Statistics} (AOAS).\\*
 
@@
 The proposed methodology is deployed and tested in a full-scale deployment where an autonomous underwater vehicle is used to map a river plume.
+This particular application is relevant as rivers are central to understanding the increasing anthropogenic impact that is affecting the health of our oceans.
 
 This interdisciplinary study demonstrates the potential of statistics applied to support autonomous robotic decision making, representing methods and approaches with a substantial potential for further integration, which are necessary and critical to understand our changing oceans.
+
+%Note that additional simulation results relating to Section 7 is available together with python code for all algorithms.
+%This material can be added in the supplementary section of the paper if requested.
+
 \vspace{1em}

### 2020-04-23 ans_to_reviewers/answers-to-reviewers.tex
Started answering and reformatted the answers into bullets

@@
 
+\usepackage{amsthm,amsmath,natbib}
+\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+
 \begin{document}
@@
 
-\item[1] “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.
+\item[1.1.1] “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.\par
 \action: Our actions
-\item[2] Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?
+\vspace{1em}
+
+\item[1.1.2] Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?\par
 \action: Our actions
-\item[3] I don’t think Figure 3 conveys too much additional benefit and eliminating it for space might be considered.
+\vspace{1em}
+
+\item[1.1.3] I don’t think Figure 3 conveys too much additional benefit and eliminating it for space might be considered.\par
 \action: Our actions
-\item[4] Why are the ES’s called ”excursion sets? Why excursion probabilities?
+\vspace{1em}
+
+\item[1.1.4] Why are the ES’s called ”excursion sets? Why excursion probabilities?\par
 \action: Our actions
-\item[5] I did not spot a definition for the “waypoint graph” a phrase used several times in the paper.
+\vspace{1em}
+
+\item[1.1.5] I did not spot a definition for the “waypoint graph” a phrase used several times in the paper.
 Again points to the need to identify the audience.
-I doubt that most statisticians would know that phrase.
+I doubt that most statisticians would know that phrase.\par
 \action: Our actions
-\item[6] In statistics journals, random variables are commonly labelled by Roman characters e.g.
+\vspace{1em}
+
+\item[1.1.6] In statistics journals, random variables are commonly labelled by Roman characters e.g.
 Z, parameters in Greek e.g.
 $\xi$.
-I realize that in other scientific domains that distinction is not always made but I do recommend the change if the paper is to appear in a stats journal.
+I realize that in other scientific domains that distinction is not always made but I do recommend the change if the paper is to appear in a stats journal.\par
 \action: Our actions
-\item[7] There is not much of novelty in the Gaussian Process theory used in the paper- its pretty much standard.
+\vspace{1em}
+
+\item[1.1.7] There is not much of novelty in the Gaussian Process theory used in the paper- its pretty much standard.
 I suggest that the derivations stated in detail be summarized by just the results.
-Statisticians will know these things and users will not care.
+Statisticians will know these things and users will not care.\par
 \action: Our actions
+\vspace{1em}
 
@@
 
-\item[1] What sort of onboard computing do they have?
+\item[2.1.1] What sort of onboard computing do they have?
 It may not be something like an Intel 2.1 GHz, 4-thread processor, but a more informative description of something most readers relate to would be helpful.
 Similar information related to RAM and hard drive space would be helpful.
-What are the computational limits of these machines?
+What are the computational limits of these machines?\par
+\action: Our actions
+\vspace{1em}
 
-\item[2] How long do the AUVs typically explore in one go? Is there a minimum/maximum time?
+\item[2.1.2] How long do the AUVs typically explore in one go? Is there a minimum/maximum time?\par 
+\action: Our actions
+\vspace{1em}
 
-\item[3] What kind of movements can the AUVs make? Chess-like moves (horizontal, vertical, diagonal)?
+\item[2.1.3] What kind of movements can the AUVs make? Chess-like moves (horizontal, vertical, diagonal)?\par  
+\action: Our actions
+\vspace{1em}
 
-\item[4] Is there a minimum/maximum number of points you want to sample?
+\item[2.1.4] Is there a minimum/maximum number of points you want to sample?\par
+\action: Our actions
+\vspace{1em}
 
-\item[5] What is the size of the grid that is typically considered? i.e., what is a typical value of $n$ in Eq.(9)?
+\item[2.1.5] What is the size of the grid that is typically considered? i.e., what is a typical value of $n$ in Eq.(9)?\par 
+\action: Our actions
+\vspace{1em}
 
-\item[6] Does the AUV use a selection algorithm from the very beginning or is there a minimum number of initial measurements that are made before the optimization algorithm begins?
+\item[2.1.6] Does the AUV use a selection algorithm from the very beginning or is there a minimum number of initial measurements that are made before the optimization algorithm begins?\par
+\action: Our actions
+\vspace{1em}
 
 \end{itemize}
-
-\action: Our action
 
@@
 
-\item[-] “closed-form” is sometimes written as “closed form” (e.g., p. 7, p. 18)
+\item[2.15.1] “closed-form” is sometimes written as “closed form” (e.g., p. 7, p. 18)
 
-\item[-] p. 9, `1s` should be ‘1s’. Similar for 0s.
+\item[2.15.2] p. 9, `1s` should be ‘1s’. Similar for 0s.
 
-\item[-] p. 15, my manuscript has -on where to sample-, which is an unusual use of dashes.
+\item[2.15.3] p. 15, my manuscript has -on where to sample-, which is an unusual use of dashes.
 
@@
 
-\action: Our actions
+\action:
+\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
+
+\item[2.15.1] 
+\item[2.15.2] 
+\item[2.15.3]
+
+\end{itemize}
 
@@
 
-\item[1.] General: The inferential problem considered is the paper is that of estimating the “boundary”
+\item[3.1.1] General: The inferential problem considered is the paper is that of estimating the “boundary”
 of a plume. However, the authors never define exactly what a “boundary” is. Such a definition
@@
 map of prediction error variances) for the variable(s) under study throughout the region of
-interest (a problem for which there is a huge literature on sampling design)?
+interest (a problem for which there is a huge literature on sampling design)?\par
+\action: Our actions
+\vspace{1em}
 
-\item[2.] Page 8, second paragraph of Section 4.1: The assumptions of separability and isotropy, and
-even stationarity, are not really necessary for the methodological development of Sections 4–6.
+\item[3.1.2] Page 8, second paragraph of Section 4.1: The assumptions of separability and isotropy, and even stationarity, are not really necessary for the methodological development of Sections 4–6.
 I suggest that you wait to make these assumptions until you get to the simulation studies and
-example.
+example.\par
+\action: Our actions
+\vspace{1em}
 
-\item[3.] Page 9, second paragraph of Section 4.2: It has been assumed here that the measurement
+\item[3.1.3] Page 9, second paragraph of Section 4.2: It has been assumed here that the measurement
 errors for temperature and salinity at any given location are uncorrelated. Is this a reasonable
-assumption? This could easily be checked by taking replicate measurements at each site.
+assumption? This could easily be checked by taking replicate measurements at each site.\par
+\action: Our actions
+\vspace{1em}
 
-\item[4.] General: Related to the previous point, I do not understand why sampling locations are being
-restricted to a regular grid. In fact, it seems unwise because it is well known (see, for example,
-Chapters 9 and 22 of Handbook of Environmental and Ecological Statistics, 2019, CRC Press)
+\item[3.1.4] General: Related to the previous point, I do not understand why sampling locations are being restricted to a regular grid.
+In fact, it seems unwise because it is well known (see, for example, Chapters 9 and 22 of Handbook of Environmental and Ecological Statistics, 2019, CRC Press)
 that the quality of inferences for spatial fields can be improved substantially by taking some
-observations very close to others.
+observations very close to others.\par
+\action: Our actions
+\vspace{1em}
 
-\item[5.] Page 13, last line: Specifically, what was the design? We aren’t even told what the sample size
-($n_y$) was used in these calculations.
+\item[3.1.5] Page 13, last line: Specifically, what was the design? We aren’t even told what the sample size ($n_y$) was used in these calculations.\par
+\action: Our actions
+\vspace{1em}
 
-\item[6.] Page 18, first paragraph of Section 7.1: The numerical values of the regression parameters
-given here are not meaningful without knowing the units of “West.”
+\item[3.1.6] Page 18, first paragraph of Section 7.1: The numerical values of the regression parameters given here are not meaningful without knowing the units of “West.”\par
+\action: Our actions
+\vspace{1em}
 
-\item[7.] Page 24, first paragraph of Fig. 8: Please define “the modeled quadratic form of the residuals”
-more precisely. Perhaps a mathematical expression is needed for clarity.
+\item[3.1.7] Page 24, first paragraph of Fig. 8: Please define “the modeled quadratic form of the residuals” more precisely. Perhaps a mathematical expression is needed for clarity.\par
+\action: Our actions
+\vspace{1em}
 
 \end{itemize}
-
-\action: Changes made.
 
@@
 
-\item[-] Page 3, line 19: Change “bi-variate” to “bivariate”. The same change is needed on the next-
+\item[3.2.1] Page 3, line 19: Change “bi-variate” to “bivariate”. The same change is needed on the next-
 to-last line of Page 12.
 
-\item[-] Page 9, three lines after (10): I presume that all off-diagonal elements of R are zeros, but this
+\item[3.2.2] Page 9, three lines after (10): I presume that all off-diagonal elements of R are zeros, but this
 should be stated explicitly.
 
-\item[-] Page 19, fourth line from bottom: Change “as also” to “as well as”
+\item[3.2.3] Page 19, fourth line from bottom: Change ``as also" to ``as well as"
 
-\item[-] Page 30, second and third references listed: “et al.” is listed after the authors’ names for each
+\item[3.2.4] Page 30, second and third references listed: ``et al." is listed after the authors’ names for each
 of these two references, but actually there are no additional authors beyond the two listed.
@@
 
-\action: Changes made.
+\action: 
+\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
+
+\item[3.2.1] we have changed this in the text - page 3 and 12.
+
+\item[3.2.2] we have added a statement on page 9: $\bR$ contains \edcomment{zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations}.
+
+\item[3.2.3] we have changed this in the text - page 19.
+
+\item[3.2.4] we have made this change to the citations \cite{genton2015cross} and \cite{french2016credible} - page 30.
+
+\end{itemize}
 

### 2020-04-23 main_new.tex
Updated with Ans-to-Rev
Started answering and reformatted the answers into bullets

@@
 Section \ref{sec:GP_EP} sets up the required statistical modeling
-assumptions of the bi-variate GPs for temperature and salinity;
+assumptions of the bivariate GPs for temperature and salinity;
 Section \ref{sec:sur} builds on these assumptions when deriving the
@@
 measurement indices of the grid and '$0$' otherwise. The size
-$2n_y \times 2n_y$ covariance matrix $\bR$ contains diagonal entries
-$r^2_T$ and $r^2_S$ which are measurement error variances of
-temperature and salinity observations.
+$2n_y \times 2n_y$ covariance matrix $\bR$ contains \edcomment{zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations}.
 
@@
 the solutions to the complex integral equations for the expected IBV
-are reduced to calculating bi-variate and four-variate cumulative
+are reduced to calculating bivariate and four-variate cumulative
 distribution functions for Gaussian vectors. For the more general
@@
 \end{equation}
-The spatial correlation is of a Matern type \citep{Matern2013}:
+The spatial correlation is of a Matern type \kc{cite or define}:
 \begin{equation}\label{v}
@@
 measured by root mean square error (RMSE) for temperature and salinity
-estimates as also the variance reduction in these two variables. It is
+estimates \edcomment{as well as} also the variance reduction in these two variables. It is
 important to note that the objective function used by the agent is

### 2020-04-23 ans_to_reviewers/answers-to-reviewers.tex
Added ans to 3.1.4

@@
 observations very close to others.\par
-\action: Our actions
+\action: The AUV and the sampling happens regularly at 1 Hz, the collected measurements are then assimilated to this regular grid.
+The sampling density is therefore higher than the grid resolution.
+Still, there is a limitation on restricting the sampling locations to finite number of fixed locations.
+However, in order to handle the computational load and resources available on the AUV this simplification is necessary.
+% Comment Trygve: Should we add some more here?
 \vspace{1em}

### 2020-04-29 main_new.tex
updates
Updated missing commands

@@
 \startlocaldefs
+\newcommand{\edcomment}[1]{{\color{green}{\{Editor: #1\}}}}
+\newcommand{\frevcomment}[1]{{\color{blue}{\{Rev 1: #1\}}}}
+\newcommand{\srevcomment}[1]{{\color{red}{\{Rev 2: #1\}}}}
+\newcommand{\trevcomment}[1]{{\color{violet}{\{Rev 3: #1\}}}}
 \endlocaldefs
@@
 \and
-\author{\fnms{Kanna} \snm{Rajan}\thanksref{t2,t5,t6}. \ead[label=e4]{kanna.rajan@ntnu.no}}
+\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e4]{kanna.rajan@fe.up.pt}}
 
@@
 \affiliation[t4]{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland.}
-\affiliation[t5]{Department of Engineering Cybernetics, NTNU}
-\affiliation[t6]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
+% \affiliation[t5]{Department of Engineering Cybernetics, NTNU}
+\affiliation[t5]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
 
@@
 \printead{e3}}
-\address{Kanna Rajan\\Department of Engineering Cybernetics\\ O. S. Bragstads Plass 2D, 7034 Trondheim\\ Norway\\
+\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
+  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
+% Department of Engineering Cybernetics\\ O. S. Bragstads Plass 2D, 7034 Trondheim\\ Norway\\
 \printead{e4}}

### 2020-05-08 ans_to_reviewers/answers-to-reviewers.tex
Added "answer" instead of "action taken"
Updates after meeting with Jo, David, Trygve, and Cedric

@@
 \def\reply{\textbf{Reviewer query}}
-\def\action{\textbf{Action taken}}
+\def\action{\textbf{Answer}}
 
@@
 \begin{itemize}[noitemsep, topsep=0pt, parsep=0pt, partopsep=0pt]
-\item \textbf{item1}: bla bla.
-\item \textbf{item2}: bla bla.
-\item \textbf{item3}: bla bla.
+\item \textbf{item1}: Generalized
+\item \textbf{item2}: Simulation code
+\item \textbf{item3}: Introduction more succinct 
 \end{itemize}
 \par \vspace{1em}
+
 Note that the reference to page $X$ is made relative to
@@
 %======================================================================
+ \section*{Associate Editor}
+%======================================================================
+All changes requested from the editor are are listed below and answered on a point-by-point basis.
+
+\setcounter{reviewer}{1}
+
+\begin{answers}
+
+\item{Comments to the Author}\label{ec}
+
+\reply: Specific comments:
+
+\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
+
+\item[1.1.1]\label{ec1} General: The inferential problem considered is the paper is that of estimating the “boundary” of a plume.
+However, the authors never define exactly what a “boundary” is.
+Such a definition should be provided, as well as some discussion of why determining the location of a boundary is so important.
+Why is it not sufficient to create a map of predictions (with an accompanying map of prediction error variances) for the variable(s) under study throughout the region of interest (a problem for which there is a huge literature on sampling design)?\par
+\action: Added references \cite{adler2000excursion,adler2009random}.
+
+\vspace{1em}
+
+\item[1.1.2]\label{ec2} Page 8, second paragraph of Section 4.1: The assumptions of separability and isotropy, and even stationarity, are not really necessary for the methodological development of Sections 4–6.
+I suggest that you wait to make these assumptions until you get to the simulation studies and example.\par
+\action: Do this! See also earlier Rev comments. Tone down this.
+\vspace{1em}
+
+\item[1.1.3]\label{ec3} Page 9, second paragraph of Section 4.2: It has been assumed here that the measurement errors for temperature and salinity at any given location are uncorrelated.
+Is this a reasonable assumption?
+This could easily be checked by taking replicate measurements at each site.\par
+\action: Conditional independent sensors. 
+\vspace{1em}
+
+\item[1.1.4]\label{ec4} General: Related to the previous point, I do not understand why sampling locations are being restricted to a regular grid.
+In fact, it seems unwise because it is well known (see, for example, Chapters 9 and 22 of Handbook of Environmental and Ecological Statistics, 2019, CRC Press)
+that the quality of inferences for spatial fields can be improved substantially by taking some
+observations very close to others.\par
+\action: The AUV and the sampling happens regularly at 1 Hz, the collected measurements are then assimilated to this regular grid.
+The sampling density is therefore higher than the grid resolution.
+Still, there is a limitation on restricting the sampling locations to finite number of fixed locations.
+However, in order to handle the computational load and resources available on the AUV this simplification is necessary.
+% Comment Trygve: Should we add some more here?
+\vspace{1em}
+
+\item[1.1.5]\label{ec5} Page 13, last line: Specifically, what was the design? We aren’t even told what the sample size ($n_y$) was used in these calculations.\par
+\action: Our actions
+\vspace{1em}
+
+\item[1.1.6]\label{ec1} Page 18, first paragraph of Section 7.1: The numerical values of the regression parameters given here are not meaningful without knowing the units of “West.”\par
+\action: Our actions. Clarify. In a larger scale this could be improved by having auxiliary data for instance from ocean models or satellite data.
+
+\vspace{1em}
+
+\item[1.1.7]\label{ec7} Page 24, first paragraph of Fig. 8: Please define “the modeled quadratic form of the residuals” more precisely. Perhaps a mathematical expression is needed for clarity.\par
+\action: Our actions
+\vspace{1em}
+
+\end{itemize}
+
+\item{Minor comments to the Author}\label{emc}
+
+\reply: 
+\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
+
+\item[1.2.1]\label{emc1} Page 3, line 19: Change “bi-variate” to “bivariate”. The same change is needed on the next-
+to-last line of Page 12.
+
+\item[1.2.2]\label{emc2} Page 9, three lines after (10): I presume that all off-diagonal elements of R are zeros, but this
+should be stated explicitly.
+
+\item[1.2.3]\label{emc3} Page 19, fourth line from bottom: Change ``as also" to ``as well as"
+
+\item[1.2.4]\label{emc4} Page 30, second and third references listed: ``et al." is listed after the authors’ names for each
+of these two references, but actually there are no additional authors beyond the two listed.
+
+\end{itemize}
+
+\action: 
+\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
+
+\item[1.2.1] we have changed this in the text - page 3 and 12.
+
+\item[1.2.2] we have changed the statement on page 9 to: ``$\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations."
+
+\item[1.2.3] we have changed this in the text - page 19.
+
+\item[1.2.4] we have made this change to the citations \cite{genton2015cross} and \cite{french2016credible} - page 30.
+
+\end{itemize}
+
+\end{answers}
+
+%======================================================================
 \section*{Answers to referee 1}
 %======================================================================
-All changes requested from Reviewer 1 are marked as \frevcomment{blue text} in the updated manuscript. \setcounter{reviewer}{1}
+All changes requested from Reviewer 1 are listed below and answered on a point-by-point basis.
+
+\setcounter{reviewer}{2}
+
 \begin{answers}
@@
 %----------------------------------------------------------------------
-\item{Major comment 1}\label{q1}
+\item{Reviewer comment 4}\label{r1c4}
 
-\reply: Reviewer comment 4: Adjustments and improvements towards the intended audience?
+\reply: Adjustments and improvements towards the intended audience?
 
@@
 
-\item{Major comment 2}\label{q2}
+\item{Reviewer comment 4 cont.}\label{r1c4}
 
-\reply: Reviewer comment 4: On including relevant code.
+\reply: On including relevant code.
+
+\action: We have made an open-source repository available for the public.
+This includes commented code examples for exploration and analysis of the material presented in the paper.
+The code is optimized for speed to allow fast computation using Pytorch (citation).
+The code is available here: (link).
+
+\item{Reviewer comment 5}\label{r1c5}
+
+\reply: On shortening the paper and moving important sentences earlier.
+Reduced oceanographic background, while highlighting elements that reviewers did not understand about the AUV planning, waypoint graph, etc.
 
@@
 
-\item{Major comment 3}\label{q3}
+\item{Reviewer comment 6}\label{r1c6}
 
-\reply: Reviewer comment 5: On shortening the paper and moving important sentences earlier.
-
-\action: Our actions
-
-\item{Major comment 4}\label{q4}
-
-\reply: Reviewer comment 6: It is not clear why the binary classification and excursion sets are being used.
+\reply: It is not clear why the binary classification and excursion sets are being used.
 Is there really a sharp boundary at the edge of a plume?
@@
 
-\action: Our actions
+\action: Add reference to Adler.
+Motivate environmental sensing.
+Plants need BOTH oxygen and sun.
+Medical doctors diagnose with MORE THAN ONE symptom, not enough to have a fever.
+Plankton and sea life (relate to lice in salmon industry) need high temperatures and high salinity.
 
-\item{Major comment 5}\label{q5}
+\item{Reviewer comment 7}\label{r1c7}
 
-\reply: Reviewer comment 7: The sequential sampling approaches are not novel.
+\reply: The sequential sampling approaches are not novel.
 they are mentioned as examples in the cited paper by Bect et al.
@@
 
-\item{Major comment 6}\label{q6}
+\item{Reviewer comment 8}\label{r1c8}
 
-\reply: Reviewer comment 8: While I can see some advantages of simplicity in the IBV of Bect et al for binary data, one could have used entropy as it has information-theoretical credentials.
+\reply: While I can see some advantages of simplicity in the IBV of Bect et al for binary data, one could have used entropy as it has information-theoretical credentials.
 Furthermore it could cover design both for spatial mapping as well as indicator mapping.
@@
 
-\action: Our actions
+\action: Our actions.
+The traditional entropy calculations for this GP aim to reduce the log determinant of the updated covariance matrix at the grid locations.
+It will not vary with the data realizations, so in this setting this information criteria will always give the same path.
+The IBV criterion is aligned with our ES motivation for this study.
+When focusing on the expected IBV, the paths will vary with the data as we see in our simulation study and in the real-data example.
+We see this as a clear benefit of the approach compared with entropy-based calculations.
+Still, sampling approaches using hybrid criteria or multi-attribute criteria (Multi-objective optimization and Pareto fronts) might balance goals in such situations, but we leave that for future research (closing remarks).
 
-\item{Major comment 7}\label{q7}
+\item{Reviewer comment 9}\label{r1c9}
 
-\reply: Reviewer comment 9: On the issue of covariance separability, objections have always been raised when I have used it as its unrealistic.
+\reply: On the issue of covariance separability, objections have always been raised when I have used it as its unrealistic.
 My rejoinder has always worked, that of citing the “all models are wrong”, George Box’s famous phrase.
@@
 
-\action: Our actions
+\action: Our actions.
+We focus on point-wise predictions in the criterion.
+Our trends are simple, so another covariance might be overfitting too much to data.
+We could try to fit different ranges to temperature and salinity, while maintaining separability.
+Could also refer to non-separable models, but pulls focus away from the current paper's focus.
+There are even tests for separability, but not worth so much, what do you do if you reject?
+Spatio-temporal variations are more interesting here, e.g.
+\cite{sigrist2015stochastic,richardson2017sparsity}.
+Ocean models, though rarely available on the scale of river plumes, might be highly useful for training - more complex trends and covariance model could then maybe be fit from the data.
 
-\item{Major comment 8}\label{q8}
+\item{Reviewer comment 10}\label{r1c10}
 
-\reply: Reviewer comment 10: The strategy at the bottom of page 17 seems ad hoc.
+\reply: The strategy at the bottom of page 17 seems ad hoc.
 Could you not use the data collected with the two ahead design to get better estimates of parameters and then with these parameters and thereafter a better myopic design?
 
-\action: Our actions
+\action: LA and LA2 improved in revision! + code. Also studied benefits of look-ahead strategies in sequential sampling in \cite{martinelli2013dynamic}.
 
-\item{Major comment 9}\label{q9}
+\item{Reviewer comment 11}\label{r1c11}
 
-\reply: Reviewer comment 11: Are these results generalizable?
+\reply: Are these results generalizable?
 In other words, how much plume-to-plume variability is there?
@@
 
-\action: Our actions
+\action: This is a simple model. Ocean models for training?
 
-\item{Major comment 10}\label{q10}
+\item{Reviewer comment 12}\label{r1c12}
 
-\reply: Reviewer comment 12: The separability assumption is to be relaxed in future work. Should publication of this case study await the proposed improvement? It would make for a much better paper.
+\reply: The separability assumption is to be relaxed in future work. Should publication of this case study await the proposed improvement? It would make for a much better paper.
 
-\action: Our actions
+\action: As pointed out above, we believe that spatio-temporal variations would be a more realistic extension, especially if repeated satellite data or ocean modeling is available.
+But this is likely relevant at a larger scale, such as that of fronts in a fjord or ocean interface.
+The methodological approach for sampling designs would however still be the same and this is the main contribution of the current paper.
 
@@
 
-\item[1.1.1] “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.\par
+\item[2.1.1] “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.\par
 \action: Our actions
@@
 
-\item[1.1.2] Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?\par
+\item[2.1.2] Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?\par
+\action: Preliminary runs set these trends. In a larger scale, one could here use ocean model inputs or satellite data.
+
+\vspace{1em}
+
+\item[2.1.3] I don’t think Figure 3 conveys too much additional benefit and eliminating it for space might be considered.\par
 \action: Our actions
@@
 
-\item[1.1.3] I don’t think Figure 3 conveys too much additional benefit and eliminating it for space might be considered.\par
+\item[2.1.4] Why are the ES’s called ”excursion sets? Why excursion probabilities?\par
 \action: Our actions
@@
 
-\item[1.1.4] Why are the ES’s called ”excursion sets? Why excursion probabilities?\par
-\action: Our actions
+\item[2.1.5] I did not spot a definition for the “waypoint graph” a phrase used several times in the paper.
+Again points to the need to identify the audience.
+I doubt that most statisticians would know that phrase.\par
+\action: Clarified. 
 \vspace{1em}
 
-\item[1.1.5] I did not spot a definition for the “waypoint graph” a phrase used several times in the paper.
-Again points to the need to identify the audience.
-I doubt that most statisticians would know that phrase.\par
-\action: Our actions
-\vspace{1em}
-
-\item[1.1.6] In statistics journals, random variables are commonly labelled by Roman characters e.g.
+\item[2.1.6] In statistics journals, random variables are commonly labelled by Roman characters e.g.
 Z, parameters in Greek e.g.
@@
 
-\item[1.1.7] There is not much of novelty in the Gaussian Process theory used in the paper- its pretty much standard.
+\item[2.1.7] There is not much of novelty in the Gaussian Process theory used in the paper- its pretty much standard.
 I suggest that the derivations stated in detail be summarized by just the results.
@@
 % \newpage
+
 %======================================================================
@@
 %======================================================================
-All changes requested from Reviewer 2 are marked as \srevcomment{red text}
-in the updated manuscript.
+All changes requested from Reviewer 2 are listed below and answered on a point-by-point basis.
 
-\setcounter{reviewer}{2}
+\setcounter{reviewer}{3}
 
 \begin{answers}
-\item{Major comment 1 - Details on AUVs}\label{q12}
+\item{Major comment 1 - Details on AUVs}\label{r2c1}
 
@@
 
-\item[2.1.1] What sort of onboard computing do they have?
+\item[3.1.1] What sort of onboard computing do they have?
 It may not be something like an Intel 2.1 GHz, 4-thread processor, but a more informative description of something most readers relate to would be helpful.
@@
 
-\item[2.1.2] How long do the AUVs typically explore in one go? Is there a minimum/maximum time?\par 
-\action: Our actions
+\item[3.1.2] How long do the AUVs typically explore in one go? Is there a minimum/maximum time?\par 
+\action: Max speed is limit for coverage. Time constraints for operations related to costs of man-hours in operation. Battery capacity is not a challenge for this AUV. 
 \vspace{1em}
 
-\item[2.1.3] What kind of movements can the AUVs make? Chess-like moves (horizontal, vertical, diagonal)?\par  
-\action: Our actions
+\item[3.1.3] What kind of movements can the AUVs make? Chess-like moves (horizontal, vertical, diagonal)?\par  
+\action: We clarified the descriptions of a way-point graph.
 \vspace{1em}
 
-\item[2.1.4] Is there a minimum/maximum number of points you want to sample?\par
-\action: Our actions
+\item[3.1.4] Is there a minimum/maximum number of points you want to sample?\par
+\action: Assuming a fixed time of sampling.
 \vspace{1em}
 
-\item[2.1.5] What is the size of the grid that is typically considered? i.e., what is a typical value of $n$ in Eq.(9)?\par 
-\action: Our actions
+\item[3.1.5] What is the size of the grid that is typically considered? i.e., what is a typical value of $n$ in Eq.(9)?\par 
+\action: Only surface considered here. No 3D with depth. 
 \vspace{1em}
 
-\item[2.1.6] Does the AUV use a selection algorithm from the very beginning or is there a minimum number of initial measurements that are made before the optimization algorithm begins?\par
-\action: Our actions
+\item[3.1.6] Does the AUV use a selection algorithm from the very beginning or is there a minimum number of initial measurements that are made before the optimization algorithm begins?\par
+\action: Batch initialization. Then adaptive sampling.
 \vspace{1em}
@@
 
-\item{Major comment 2}\label{q13}
+\item{Major comment 2}\label{r2c2}
 
@@
 
-\item{Major comment 3}\label{q14}
+\item{Major comment 3}\label{r2c3}
 
@@
 
-\action: Our actions
+\action: No data to fit a complex model. See Ocean model comments to Rev 1.
 
-\item{Major comment 4}\label{q15}
+\item{Major comment 4}\label{r2c4}
 
@@
 
-\action: Our actions
+\action: Conditionally independent measurements.
 
-\item{Major comment 5}\label{q16}
+\item{Major comment 5}\label{r2c5}
 
@@
 
-\action: Our actions
+\action: Only the surface is considered here, so computational requirements are not mainly driven by the size of the domain, but rather by the number of combinations for sampling designs (computational complexity).
+No 3D with depth, which would require simplified spatial models like fixed-rank-kriging or similar approaches.
 
-\item{Major comment 6}\label{q17}
+\item{Major comment 6}\label{r2c6}
 
@@
 
-\action: Our actions
+\action: Appendix?
 
-\item{Major comment 7}\label{q18}
+\item{Major comment 7}\label{r2c7}
 
@@
 
-\action: Our actions
+\action: Appendix?
 
-\item{Major comment 8}\label{q19}
+\item{Major comment 8}\label{r2c8}
 
@@
 
-\action: Our actions
+\action: Appendix?
 
-\item{Major comment 9}\label{q20}
+\item{Major comment 9}\label{r2c9}
 
@@
 
-\action: Our actions
+\action: Appendix?
 
-\item{Major comment 10}\label{q21}
+\item{Major comment 10}\label{r2c10}
 
@@
 
-\action: Our actions
+\action: Appendix?
 
-\item{Major comment 11}\label{q22}
+\item{Major comment 11}\label{r2c11}
 
@@
 
-\item{Major comment 12}\label{q23}
+\item{Major comment 12}\label{r2c12}
 
@@
 
-\action: Our actions
+\action: Joints would be too demanding here. There is more of a question of the longer planning horizon for the AUV to improve the sampling design, so the domain can be more efficiently explored.
 
-\item{Major comment 13}\label{q24}
+\item{Major comment 13}\label{r2c13}
 
@@
 
-\item{Major comment 14}\label{q25}
+\item{Major comment 14}\label{r2c14}
 
@@
 
-\action: Our actions
+\action: Code is improved in the revision and provided for open-source usage.
 
-\item{Minor comments}\label{q26}
+\item{Minor comments}\label{r2cminor}
 
@@
 
-\item[2.15.1] “closed-form” is sometimes written as “closed form” (e.g., p. 7, p. 18)
+\item[3.15.1] “closed-form” is sometimes written as “closed form” (e.g., p. 7, p. 18)
 
-\item[2.15.2] p. 9, `1s` should be ‘1s’. Similar for 0s.
+\item[3.15.2] p. 9, `1s` should be ‘1s’. Similar for 0s.
 
-\item[2.15.3] p. 15, my manuscript has -on where to sample-, which is an unusual use of dashes.
+\item[3.15.3] p. 15, my manuscript has -on where to sample-, which is an unusual use of dashes.
 
@@
 
-\item[2.15.1] 
-\item[2.15.2] 
-\item[2.15.3]
+\item[3.15.1] 
+\item[3.15.2] 
+\item[3.15.3]
 
@@
 \end{answers}
+
 %======================================================================
@@
 %======================================================================
-All changes requested from Reviewer 3 are marked as \trevcomment{violet text}
-in the updated manuscript.
+All changes requested from Reviewer 3 are listed below and answered on a point-by-point basis.
 
-\setcounter{reviewer}{3}
+\setcounter{reviewer}{4}
 
@@
 
-\item{Minor comment 1}\label{q27}
+\item{Minor comment 1}\label{r3c1}
 
-\reply: The authors indicate several paths forward for work in this area, which I think pose interesting questions. They may want to elaborate
-a bit more on those.
+\reply: The authors indicate several paths forward for work in this area, which I think pose interesting questions. They may want to elaborate a bit more on those.
 
-\action: Our action
+\action: Space-time \cite{sigrist2015stochastic,richardson2017sparsity}.
+More complex sampling Look-ahead, going further.
+3D sampling, requiring more computationally efficient spatial statistical models.
 
-\item{Minor comment 2}\label{q28}
+\item{Minor comment 2}\label{r3c2}
 
@@
 
-\item{Major comment 3}\label{q29}
+\item{Major comment 3}\label{r3c3}
 
@@
 \action: Our actions
-
-\end{answers}
-
-%======================================================================
- \section*{Associate Editor}
-%======================================================================
-All changes requested from the editor are marked as \edcomment{green text}
-in the updated manuscript.
-
-\begin{answers}
-
-\item{Comments to the Author}\label{q30}
-
-\reply: Specific comments:
-
-\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
-
-\item[3.1.1] General: The inferential problem considered is the paper is that of estimating the “boundary”
-of a plume. However, the authors never define exactly what a “boundary” is. Such a definition
-should be provided, as well as some discussion of why determining the location of a boundary
-is so important. Why is it not sufficient to create a map of predictions (with an accompanying
-map of prediction error variances) for the variable(s) under study throughout the region of
-interest (a problem for which there is a huge literature on sampling design)?\par
-\action: Our actions
-\vspace{1em}
-
-\item[3.1.2] Page 8, second paragraph of Section 4.1: The assumptions of separability and isotropy, and even stationarity, are not really necessary for the methodological development of Sections 4–6.
-I suggest that you wait to make these assumptions until you get to the simulation studies and
-example.\par
-\action: Our actions
-\vspace{1em}
-
-\item[3.1.3] Page 9, second paragraph of Section 4.2: It has been assumed here that the measurement
-errors for temperature and salinity at any given location are uncorrelated. Is this a reasonable
-assumption? This could easily be checked by taking replicate measurements at each site.\par
-\action: Our actions
-\vspace{1em}
-
-\item[3.1.4] General: Related to the previous point, I do not understand why sampling locations are being restricted to a regular grid.
-In fact, it seems unwise because it is well known (see, for example, Chapters 9 and 22 of Handbook of Environmental and Ecological Statistics, 2019, CRC Press)
-that the quality of inferences for spatial fields can be improved substantially by taking some
-observations very close to others.\par
-\action: The AUV and the sampling happens regularly at 1 Hz, the collected measurements are then assimilated to this regular grid.
-The sampling density is therefore higher than the grid resolution.
-Still, there is a limitation on restricting the sampling locations to finite number of fixed locations.
-However, in order to handle the computational load and resources available on the AUV this simplification is necessary.
-% Comment Trygve: Should we add some more here?
-\vspace{1em}
-
-\item[3.1.5] Page 13, last line: Specifically, what was the design? We aren’t even told what the sample size ($n_y$) was used in these calculations.\par
-\action: Our actions
-\vspace{1em}
-
-\item[3.1.6] Page 18, first paragraph of Section 7.1: The numerical values of the regression parameters given here are not meaningful without knowing the units of “West.”\par
-\action: Our actions
-\vspace{1em}
-
-\item[3.1.7] Page 24, first paragraph of Fig. 8: Please define “the modeled quadratic form of the residuals” more precisely. Perhaps a mathematical expression is needed for clarity.\par
-\action: Our actions
-\vspace{1em}
-
-\end{itemize}
-
-\item{Minor comments to the Author}\label{q31}
-
-\reply: 
-\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
-
-\item[3.2.1] Page 3, line 19: Change “bi-variate” to “bivariate”. The same change is needed on the next-
-to-last line of Page 12.
-
-\item[3.2.2] Page 9, three lines after (10): I presume that all off-diagonal elements of R are zeros, but this
-should be stated explicitly.
-
-\item[3.2.3] Page 19, fourth line from bottom: Change ``as also" to ``as well as"
-
-\item[3.2.4] Page 30, second and third references listed: ``et al." is listed after the authors’ names for each
-of these two references, but actually there are no additional authors beyond the two listed.
-
-\end{itemize}
-
-\action: 
-\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
-
-\item[3.2.1] we have changed this in the text - page 3 and 12.
-
-\item[3.2.2] we have added a statement on page 9: $\bR$ contains \edcomment{zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations}.
-
-\item[3.2.3] we have changed this in the text - page 19.
-
-\item[3.2.4] we have made this change to the citations \cite{genton2015cross} and \cite{french2016credible} - page 30.
-
-\end{itemize}
 
@@
 
-%======================================================================
+%=====================================================================
 \end{document}
-%======================================================================
+%=====================================================================

### 2020-05-08 main_new.tex
Update to remove color-coding. Adding colors will be too complicated for this paper, so we will not use coloring to indicate changes.

@@
 measurement indices of the grid and '$0$' otherwise. The size
-$2n_y \times 2n_y$ covariance matrix $\bR$ contains \edcomment{zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations}.
+$2n_y \times 2n_y$ covariance matrix $\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations.
 
@@
 measured by root mean square error (RMSE) for temperature and salinity
-estimates \edcomment{as well as} also the variance reduction in these two variables. It is
+estimates as well as also the variance reduction in these two variables. It is
 important to note that the objective function used by the agent is

### 2020-05-11 ans_to_reviewers/answers-to-reviewers.tex
Update to remove color-coding. Adding colors will be too complicated for this paper, so we will not use coloring to indicate changes.
First draft of answer-letter. to be continued.

@@
 
-The authors are grateful to the reviewers for their comments and
-suggestions. They have all been carefully considered in this revised
-version of this manuscript, which we believe has significantly
-improved as a result. Detailed responses to the feedback, from each reviewer and the editorial comments follow.\\
+We are grateful to the reviewers for their comments and
+suggestions. They have all been considered in this revised
+version of this manuscript. The principal revisions made to the manuscript are as follows and are done to address the main comments by Editors and Reviewers:
 
-The principal revisions made to the manuscript are:
-\vspace{0.5cm}
+\begin{enumerate}
+\item The motivation for excursion sets (ESs) and in particular that of {\bf{multivariate ESs}} has been strengthened.
+Motivating examples are abundant in medicine where doctors do not solely rely on a single symptom but must see several combined effects before making a diagnosis.
+Another example is that of photosynthesis where plants and algae produce oxygen only when given enough of every component; light, carbon dioxide and water.
+Our motivation in this paper is that temperature and salinity, combined, drive several processes in the oceans.
+For instance, salmon lice larvae are known to depend on relatively high temperature and high salinities to survive.
+Salmon lice are parasites to salmon, and the bivariate ES for temperature and salinity is hence important to understand for sustainable activities in salmon fisheries.
 
-\begin{itemize}[noitemsep, topsep=0pt, parsep=0pt, partopsep=0pt]
-\item \textbf{item1}: Generalized
-\item \textbf{item2}: Simulation code
-\item \textbf{item3}: Introduction more succinct 
-\end{itemize}
+{\it{Action:}} In the revised manuscript these aspects are largely rewritten in the Introduction, Background, ES definitions and Case Study. This change addresses the comments of AE1, R1 (3,) MORE.
+
+\item The {\bf{statistical methodological contribution}} have been clarified and improved, including definitions of ESs and having more general results related to the integrated Bernoulli variance (IBV) for multivariate processes, using marginal (one-location) entries and for volume calculations requiring multiple locations.
+The revised version highlights these novel statistical results as a main contribution of the paper.
+
+{\it{Action:}} In the revised manuscript these aspects are largely rewritten in the main body containing novel results for IBV for sums of marginal variances as well as for volumes, while the new Appendix derives the closed form results.
+This change addresses the comments of E2, AE1, R2 (7), R3 (1), MORE.
+
+\item We realized that a contribution of the current work is available code for statisticians who could try this themselves or even extend it for their own statistical cases.
+The computer code could of course also be useful to practitioners having related sampling design problems.
+The {\bf{simulation results and code}} are thoroughly re-worked.
+Open-source python code is now provided for all the examples used in the paper and also for simulation studies using more nuances statistical models (non-stationary, non-isotropic and non-separable).
+
+{\it{Action:}} The simulation study section is largely modified and it contains pointers to the associated open-source code. This change addresses the comments of R1 (2,3),  MORE.
+
+\item In the initial submission we tried to motivate these kinds of applications carefully, but based on the comments, we see that the {\bf{introduction and background}} were too long.
+We were not selling the paper enough to professional statisticians, and maybe too much to other scientists who are interested in particular areas of these statistical application.
+In the revised version we motivate more for statisticians, in particular we emphasize the statistical design problem which is sought solved by AUV with its constraints on navigation and computing resources.
+
+{\it{Action:}} The Introduction and Background sections have been largely rewritten.
+We now get quicker to the point and engage readers in its both applied statistical focus and the methodological contributions.
+This change addresses the comments of E1, E4, AE1, R2 (7), R3 (1), MORE.
+
+\item We have further included Cedrid Traveletti as co-author. He has contributed on deriving statistical properties of the methods and on improving the code for open-source quality.
+
+\end{enumerate}
 \par \vspace{1em}
@@
 %======================================================================
+\section*{Editor}
+%======================================================================
+All changes requested from the editor are listed below and answered on a point-by-point basis.
+
+\begin{answers}
+
+	\item{Comment:}\label{ec1}
+
+Thank you for submitting your manuscript to The Annals of 
+Applied Statistics.  Three referees, an Associate Editor (AE),
+and I have read it carefully; reports from all four reviewers,
+including particularly constructive reports from Referee 1
+(R1R1.pdf; marked manuscript in R1R2.pdf) and the AE (R1AE1.pdf), 
+can be downloaded via the links below. We all agree that your 
+manuscript offers a potentially useful approach to estimating 
+plume boundaries and a sequential design for the data collection
+by autonomous underwater vehicles (AUVs). Citing from Referee 3
+(R1E1.txt), "the authors have motivated the application well."
+In this sense, the paper is consistent with the mission of AOAS.
+
+However, the reports also identify serious shortcomings that 
+prevent me from accepting your manuscript for publication at
+this time. The reports are rather consistent in the nature of 
+the problems that would need to be addressed for an AOAS publication.  
+Briefly, these issues include:
+
+\item{1. Intended audience:}\label{r3c1}
+
+As you know from reading the Manuscript Submission page of our
+web site (from where you submitted your manuscript), "AOAS is 
+intended to be read by all professional statisticians, as well 
+as by other scientists who are interested in particular areas 
+of statistical application. It is important for the abstract
+and introduction to say: Why the topic is of interest, and What 
+is to be shown, in as non-technical a manner as possible."
+Remarks from the AE (point 1), Referee 1 (point 4), and Referee 2 
+(R1R3.pdf, 1) refer to this question: "Who is the intended
+audience?"  Quoting from Referee 1, "I am not clear about the 
+intended audience for this paper.  Or to put it another way, 
+for whom is this paper important?" The reports have some 
+suggestions that may guide you in making the necessary critical 
+revisions along these lines.
+
+\reply: The comment by you, AE and reviewers made us realize that the initial submission had a motivation which was too applied.
+We were not selling the paper contributions in early enough, and it was not targeted sufficiently to professional statisticians.
+In the revised version we motivate more for statisticians.
+(See also comment 4 in the list of main changes above.)
+
+\action: Our actions are to i) rewrite Introduction and Background sections to catch our intended audience and highlight our contributions, ii) restructure the presentation to highlight both statistical contributions as well as the impact such technologies and applied questions could provide for statisticians.
+
+\item{2. Novelty:}\label{r3c2}
+
+As noted in the AOAS mission statement, papers are expected
+to include some degree of "methodological innovation."
+In separate comments to the Editors, one reviewer noted,
+"Methodological novelty is an issue with this paper.  Much of 
+it is either standard Gaussian process calculations/formulae 
+or sequential designs that were, in all but one case, previously 
+proposed. Two things that could add a lot would be good code 
+that others could run for the simulation studies like the authors' 
+(if someone else were contemplating a similar experiment), and 
+good code for finding the optimum design. That could lead to a 
+fairly substantial impact of their work."
+
+\reply: We realize that we did not stress the statistical novelty sufficiently in the submitted version.
+The main statistical novelty is in the closed form calculation for the design criteria of the integrated Bernoulli variance (IBV) associated with excursion sets (ESs) for multivariate processes.
+In the revised version this is extended to volumes (going beyond single location).
+We believe that part of the problem with the initial submission was the organization of these new results in the structure of the paper.
+This is hence re-structured in the revision.
+
+\action: Our actions are to i) extend the closed form calculations to a criterion working on ES volume calculations (going beyond single location), ii) restructure the presentation to work directly on $p$-variate random fields and hence generalize the presentation, iii) restructure the closed form results to be presented in the main body, but with proofs in the new Appendix.
+
+\item{3. Assumptions }\label{r3c3}
+
+Referees 1 and 2, as well as the AE, call attention to some of
+the assumptions (e.g., stationary and isotropic random fields),
+and ask you to address how realistic they are.
+
+\reply: At this spatial scale there is often not so much data to infer a detailed statistical model, and this is the reason for our parsimonious construction for the real-data case.
+In other related situations, for instance that of larger fjord, ocean fronts, which are at a larger scale, there would usually be more data available for instance in the form of satellite data or from ensembles of computer model simulations.
+For such cases, we agree that it would make sense to embed more nuanced models, for instance that of larger variances at certain locations and stronger correlations in certain directions, but for our field trial here we attempted to have as few parameters as possible and had no reason to assume more complicated forms.
+
+That said, in the updated simulation study, which we have generalized and made open-source there are opportunities and examples of more complex covariance models (non-stationary and non-separable and non-isotropic), but we do not conduct any inference for parameters in these models from ocean model data.
+
+\action: Our actions include i) a better description of our parameter choices for the synthetic case and the specification for the real-data case, and ii) brief notes of the extended opportunities available for the associated open-access code, iii) references to related papers (including one of our own) that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here.
+
+\item{Major comment 4. Exposition}\label{r3c4}
+
+Both the AE and Referee 1 strongly advise shortening the first
+five pages - or, at least, getting to the point (currently on p6)
+much sooner.  Quoting from  Referee 1, 5: "I slogged through the 
+first six [pages] before I discovered what the paper was to be 
+about, on page 6."  You do not want to lose your readers, so I
+advise you to engage them early on.  All four reviewers note
+potential technical or typographical errors, so please proofread
+carefully.
+
+\reply: Based on the comments by yourself, AE and referees, we realize that the motivation part of the paper was too long.
+
+\action: Our actions include a much shorter Section 2 with background. We now aim to engage readers earlier on with the applied statistical focus and methodological contribution of our work.
+
+\end{answers}
+
+%======================================================================
  \section*{Associate Editor}
 %======================================================================
-All changes requested from the editor are are listed below and answered on a point-by-point basis.
+All changes requested from the editor are listed below and answered on a point-by-point basis.
 
@@
 \item[1.1.4]\label{ec4} General: Related to the previous point, I do not understand why sampling locations are being restricted to a regular grid.
-In fact, it seems unwise because it is well known (see, for example, Chapters 9 and 22 of Handbook of Environmental and Ecological Statistics, 2019, CRC Press)
-that the quality of inferences for spatial fields can be improved substantially by taking some
+In fact, it seems unwise because it is well known (see, for example, Chapters 9 and 22 of Handbook of Environmental and Ecological Statistics, 2019, CRC Press) that the quality of inferences for spatial fields can be improved substantially by taking some
 observations very close to others.\par
@@
 
-%======================================================================
+%=====================================================================
 \section*{Answers to referee 1}
-%======================================================================
+%=====================================================================
 All changes requested from Reviewer 1 are listed below and answered on a point-by-point basis.
@@
 
-\begin{quote}
-\srevcomment{a change in the text}
-\end{quote}
-
-\begin{quote}
-  \srevcomment{another change in the text} \end{quote}
-
 \item{Major comment 2}\label{r2c2}
@@
 
-\item{Minor comment 1}\label{r3c1}
+\item{Comment:}\label{r3c1}
 
-\reply: The authors indicate several paths forward for work in this area, which I think pose interesting questions. They may want to elaborate a bit more on those.
+\reply: I think this is an interesting paper. The paper is well written, 
+and the authors have motivated the application well. 
+The authors indicate several paths forward for work in this area, which I think pose interesting questions. They may want to elaborate a bit more on those.
 
@@
 
-\item{Minor comment 2}\label{r3c2}
+\item{Minor comment 1}\label{r3c2}
 
@@
 
-\item{Major comment 3}\label{r3c3}
+\item{Major comment 2}\label{r3c3}
 

### 2020-05-11 main_new.tex
First draft of answer-letter. to be continued.

@@
 variations on which ocean processes transpire, prompting the need for
-effective means of sampling. By \emph{sampling}, we refer to the design
-of observational strategies in the spatial domain, where the use of
-autonomous platforms can be combined with statistical methods to pursue
-measurements with high scientific relevance.
-
-Traditional data collection at sea has typically been based on static
-buoys, Lagrangian floats, or ship-based methods, with significant
-logistical limitations that directly impact coverage and sampling
-resolution. Modern methods using satellite remote-sensing provide
-large-scale coverage but have limited resolution, are limited to
-sensing the surface of the ocean, and are impacted by cloud cover. The
-advent of robust mobile robotic platforms \citep{Bellingham07} has
-resulted in significant contributions to environmental monitoring and
-sampling. In particular, autonomous underwater vehicles (AUVs), have
-advanced the state of sampling and consequently have made robotics an
-integral part of ocean observation; our previous work has contributed
-to this effort \citep{das11b,Das2015,fossum18b,fossuminformation}. Other statistical
-work in the oceanographic domain include \cite{wikle2013modern}
-focusing on hierarchical statistical models; \cite{sahu2008space},
-studying spatio-temporal models for sea surface temperature and
-salinity data; and \cite{mellucci2018oceanic} looking at the
-statistical prediction of features using an underwater glider.
-
-The potential that lies in using information-based methods, which we
-highlight in this paper, is to do efficient environmental sampling and
-overcome three common challenges, namely:
+effective means of sampling.
+By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use of autonomous platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
+The potential that lies in using information-based methods \citep{das11b,Das2015,fossum18b,fossuminformation}, which we highlight in this paper, is to do efficient environmental sampling and overcome three common challenges, namely:
+
+% After reviews: - This is moved to Section 2.1.
+
+%Traditional data collection at sea has typically been based on static
+%buoys, Lagrangian floats, or ship-based methods, with significant
+%logistical limitations that directly impact coverage and sampling
+%resolution. Modern methods using satellite remote-sensing provide
+%large-scale coverage but have limited resolution, are limited to
+%sensing the surface of the ocean, and are impacted by cloud cover. The
+%advent of robust mobile robotic platforms \citep{Bellingham07} has
+%resulted in significant contributions to environmental monitoring and
+%sampling. In particular, autonomous underwater vehicles (AUVs), have
+%advanced the state of sampling and consequently have made robotics an
+%integral part of ocean observation; our previous work has contributed
+%to this effort \citep{das11b,Das2015,fossum18b,fossuminformation}. Other ¤ %statistical work in the oceanographic domain include \cite{wikle2013modern}
+%focusing on hierarchical statistical models; \cite{sahu2008space},
+%studying spatio-temporal models for sea surface temperature and
+%salinity data; and \cite{mellucci2018oceanic} looking at the
+%statistical prediction of features using an underwater glider.
 
@@
 this is the ability to model spatially-correlated variables and
-provide formal measures of uncertainty. In our formulation we use
-Gaussian Processes (GPs) in large part because they allow for
-efficient implementation and evaluation in real time onboard a robotic platform.
+provide formal measures of uncertainty. Our formulation is based on
+Gaussian Processes (GPs) as they allow efficient implementation 
+and evaluation in real time onboard a robotic platform.
 
@@
 % ecosystem response.
-
 % owing to current observation practices.
@@
 incipient micro-biology driven by planktonic and coastal anthropogenic
-input, such as pollution and agricultural runoff transported into the ocean by rivers and streams. These often result
-in a range of ecosystem-related phenomena such as blooms, anoxic zones
-and plumes, all of which have significant direct and indirect effects
-on society. Walter Munk referred to the 20\textsuperscript{th} century
-as the \emph{century of undersampling} \citep{munk2002}, pointing out
-a lack of sufficient sampling resolution in time and space as a
-critical issue for characterizing ocean phenomena. Even today,
-policymakers struggle with limited observability of severe
-environmental problems; a prominent example is the recent (May 2018)
-outbreak of harmful algal blooms in Northern Norway\footnote{Report
-  from the Institute of Marine Research: https://bit.ly/2KZGJfI}, with
-losses exceeding 12 million tons of salmon valued at \$80 million
-(Norwegian Directorate of Fisheries). This is a case in point of how
-adequate sampling efforts could have provided a more accurate
-assessment and early warning of changing environmental
-conditions. These challenges related to undersampling have been
-driving ocean sensing practices towards more autonomous, networked,
-and distributed use of robotic assets in order to fill the gaps in our
-understanding (See Fig. \ref{fig:envir1}), where observation is guided
-by assimilation of recent measurements and statistical models.
+input, such as pollution and agricultural runoff transported into the ocean by rivers and streams.
+These often result in a range of ecosystem-related phenomena such as blooms, anoxic zones and plumes, all of which have significant direct and indirect effects on society.
+
+% On observation and why robotics
+Traditional data collection at sea has typically been based on static
+buoys, Lagrangian floats, or ship-based methods, with significant
+logistical limitations that directly impact coverage and sampling
+resolution. Modern methods using satellite remote-sensing provide
+large-scale coverage but have limited resolution, are limited to
+sensing the surface of the ocean, and are impacted by cloud cover. The
+advent of mobile robotic platforms \citep{Bellingham07} has
+resulted in significant contributions to environmental monitoring and
+sampling. In particular, autonomous underwater vehicles (AUVs), have
+advanced the state of sampling and consequently have made robotics an
+integral part of ocean observation.
+The previously listed challenges related to undersampling have also been driving ocean sensing practices towards more autonomous, networked, and distributed use of robotic assets (See Fig.
+\ref{fig:envir1}), where observation is guided by assimilation of recent measurements and statistical models.
+
+%Walter Munk referred to the 20\textsuperscript{th} century
+%as the \emph{century of undersampling} \citep{munk2002}, pointing out
+%a lack of sufficient sampling resolution in time and space as a
+%critical issue for characterizing ocean phenomena. Even today,
+%policymakers struggle with limited observability of severe
+%environmental problems; a prominent example is the recent (May 2018)
+%outbreak of harmful algal blooms in Northern Norway\footnote{Report
+%  from the Institute of Marine Research: https://bit.ly/2KZGJfI}, with
+%losses exceeding 12 million tons of salmon valued at \$80 million
+%(Norwegian Directorate of Fisheries). This is a case in point of how
+%adequate sampling efforts could have provided a more accurate
+%assessment and early warning of changing environmental conditions.
 
@@
     collaborative networked operations of robotic vehicles to resolve
-    numerous bio-geochemical processes and their interactions. The
-    sense-plan-act control approach to autonomous control can reason
-    and make decisions for efficient onboard adaptive sampling.
+    numerous bio-geochemical processes and their interactions.
+AUV platforms are an integral part of this network being able to reason and make decisions for efficient onboard adaptive sampling using the sense-plan-act control approach to autonomous control.
     \ref{fig:nidelven} The interaction of river and ocean creates
@@
 
-Our focus in this work is towards spatial characterization of a
-frontal system specifically, fronts generated by river
+The application that is of focus in this work is towards spatial characterization of a frontal system specifically, fronts generated by river
 plumes. Fig. \ref{fig:nidelven} shows the survey area (the outlet of
@@
 features is therefore highly uncertain, making deterministic planning
-challenging.
+challenging. River plumes belong to a class of ocean processes that are local and
+act on smaller \emph{sub-mesoscale} extent (from $5 m^2$ -- $10 km^2$) where
+lateral spatial variability tends to dominate. Remote sensing and
+synthetic ocean models often find it challenging to provide detail at
+this scale \citep{Lermusiaux:2006} and the principal way to resolve
+for fine resolution is by direct in-situ observation. At larger
+\emph{mesoscale} ($>50 km^2$), both three-dimensional space and time
+dynamics are important and can shift substantially, while in the case
+of river plumes one can often limit the scope to cover only lateral
+spatial elements. However, the methodological framework can be
+extended to higher dimensional situations in a similar manner.
 
@@
 
-River plumes belong to a class of ocean processes that are local and
-act on smaller \emph{sub-mesoscale} extent (from $5 m^2$ -- $10 km^2$) where
-lateral spatial variability tends to dominate. Remote sensing and
-synthetic ocean models often find it challenging to provide detail at
-this scale \citep{Lermusiaux:2006} and the principal way to resolve
-for fine resolution is by direct in-situ observation. At larger
-\emph{mesoscale} ($>50 km^2$), both three-dimensional space and time
-dynamics are important and can shift substantially, while in the case
-of river plumes one can often limit the scope to cover only lateral
-spatial elements. However, the methodological framework can be
-extended to higher dimensional situations in a similar manner.
-
 \subsection{Autonomous Vehicles}
@@
 an in-water operation time capacity dependent on survey speed, payload
-sensors and mission design. By deploying AUVs together with other
+sensors and mission design.
+AUVs have limited computational capacity due to their limited energy and payload space.
+The computations is usually done on a single-board computer (SBC), like the Raspberry Pi or an multicore GPU NVIDIA Jetson TX1.
+
+By deploying AUVs together with other
 resources, one can augment information from synthetic ocean models,
@@
 
-\begin{figure}[h!]
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
+%\caption{The equilateral waypoint graph used to discretize the
+%  trajectory choices over the $31\times31$ grid used to discretize the GP.}
+%\label{fig:wp_graph}
+%\end{figure}
+
+\begin{figure}[!h] 
 \centering 
-\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
-\caption{The equilateral waypoint graph used to discretize the
-  trajectory choices over the $31\times31$ grid used to discretize the GP.}
+\subfigure[The waypoint graph.]{\includegraphics[width =
+0.49\textwidth]{Figures/sim/wp_graph_paper.pdf}\label{fig:wp_graph_a}}
+\hfill
+\subfigure[The waypoint graph in 3D.]{\includegraphics[width =
+0.49\textwidth]{Figures/sim/wp_graph_3d.pdf}\label{fig:wp_graph_b}}
+\caption{\ref{fig:wp_graph_a} The equilateral waypoint graph used to discretize the
+trajectory choices over the $31\times31$ grid used to discretize the GP.
+\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment.}
 \label{fig:wp_graph}

### 2020-05-12 ans_to_reviewers/answers-to-reviewers.tex
answers modified

@@
 \newcounter{reviewer}
-\def\reply{\textbf{Reviewer query}}
-\def\action{\textbf{Answer}}
+\def\revcom{\textbf{Reviewer comment}}
+\def\aecom{\textbf{Associate Editor comment}}
+\def\reply{\textbf{Reply}}
+\def\action{\textbf{Action}}
 
@@
 
-\begin{answers}
+%\begin{answers}
 
-	\item{Comment:}\label{ec1}
+%	\item{Comment:}\label{ec1}
+\vspace{5mm}
+\noindent {\bf{Comment:}}
 
@@
 
-\item{1. Intended audience:}\label{r3c1}
+%\item{1. Intended audience:}\label{r3c1}
+\vspace{5mm}
+\noindent {\bf{Main comment 1. Intended audience:}}
 
@@
 
-\item{2. Novelty:}\label{r3c2}
+%\item{2. Novelty:}\label{r3c2}
+\vspace{5mm}
+\noindent {\bf{Main comment 2. Novelty:}}
 
@@
 
-\item{3. Assumptions }\label{r3c3}
+%\item{3. Assumptions }\label{r3c3}
+\vspace{5mm}
+\noindent {\bf{Main comment 3. Assumptions }}
 
@@
 
-\item{Major comment 4. Exposition}\label{r3c4}
+%\item{4. Exposition}\label{r3c4}
+
+\vspace{5mm}
+\noindent {\bf{Main comment 4. Exposition}}
 
@@
 
-\end{answers}
+%\end{answers}
 
@@
 
-\begin{answers}
+%\begin{answers}
 
-\item{Comments to the Author}\label{ec}
+%\item{Comments to the Author}\label{ec}
 
-\reply: Specific comments:
+%\aecom: Specific comments:
 
-\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
+%\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
 
-\item[1.1.1]\label{ec1} General: The inferential problem considered is the paper is that of estimating the “boundary” of a plume.
+%\item[1.1.1]\label{ec1} 
+\vspace{5mm}
+\noindent \aecom
+
+General: The inferential problem considered is the paper is that of estimating the “boundary” of a plume.
 However, the authors never define exactly what a “boundary” is.
@@
 Why is it not sufficient to create a map of predictions (with an accompanying map of prediction error variances) for the variable(s) under study throughout the region of interest (a problem for which there is a huge literature on sampling design)?\par
+
+\reply
+
 \action: Added references \cite{adler2000excursion,adler2009random}.
@@
 
-\item[1.1.2]\label{ec2} Page 8, second paragraph of Section 4.1: The assumptions of separability and isotropy, and even stationarity, are not really necessary for the methodological development of Sections 4–6.
+%\item[1.1.2]\label{ec2} 
+\vspace{5mm}
+\noindent \aecom
+
+Page 8, second paragraph of Section 4.1: The assumptions of separability and isotropy, and even stationarity, are not really necessary for the methodological development of Sections 4–6.
 I suggest that you wait to make these assumptions until you get to the simulation studies and example.\par
-\action: Do this! See also earlier Rev comments. Tone down this.
+
+\reply We agree that they are not necessary.
+In the open-source that is now available, we implement non-stationary, non-separable and non-isotropic covariance models.
+But we do not conduct any inference for parameters.
+This might be possible for abundant data, but not in our small-scale resolution example.
+
+\action: No description of modeling assumptions in the methodological sections.
+A better description of our parameter choices for the synthetic case and the specification for the real-data case.
+Brief notes of the extended opportunities available for the associated open-access code.
+References to related papers (including one of our own) that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here.
+
 \vspace{1em}
 
-\item[1.1.3]\label{ec3} Page 9, second paragraph of Section 4.2: It has been assumed here that the measurement errors for temperature and salinity at any given location are uncorrelated.
+%\item[1.1.3]\label{ec3} 
+\vspace{5mm}
+\noindent \aecom
+
+Page 9, second paragraph of Section 4.2: It has been assumed here that the measurement errors for temperature and salinity at any given location are uncorrelated.
 Is this a reasonable assumption?
 This could easily be checked by taking replicate measurements at each site.\par
-\action: Conditional independent sensors. 
+
+\reply
+
+\action: The assumptions are conditionally independent measurements of salinity and temperature.
+We believe this is realistic as they are gathered from different sensors.
+In the latent Gaussian random field, the variables are dependent.
 \vspace{1em}
 
-\item[1.1.4]\label{ec4} General: Related to the previous point, I do not understand why sampling locations are being restricted to a regular grid.
+%\item[1.1.4]\label{ec4} 
+\vspace{5mm}
+\noindent \aecom
+
+General: Related to the previous point, I do not understand why sampling locations are being restricted to a regular grid.
 In fact, it seems unwise because it is well known (see, for example, Chapters 9 and 22 of Handbook of Environmental and Ecological Statistics, 2019, CRC Press) that the quality of inferences for spatial fields can be improved substantially by taking some
 observations very close to others.\par
+
+\reply
+
 \action: The AUV and the sampling happens regularly at 1 Hz, the collected measurements are then assimilated to this regular grid.
@@
 
-\item[1.1.5]\label{ec5} Page 13, last line: Specifically, what was the design? We aren’t even told what the sample size ($n_y$) was used in these calculations.\par
+%\item[1.1.5]\label{ec5} 
+\vspace{5mm}
+\noindent \aecom
+
+Page 13, last line: Specifically, what was the design? We aren’t even told what the sample size ($n_y$) was used in these calculations.\par
+
+\reply
+
 \action: Our actions
@@
 
-\item[1.1.6]\label{ec1} Page 18, first paragraph of Section 7.1: The numerical values of the regression parameters given here are not meaningful without knowing the units of “West.”\par
+%\item[1.1.6]\label{ec1} 
+\vspace{5mm}
+\noindent \aecom
+
+Page 18, first paragraph of Section 7.1: The numerical values of the regression parameters given here are not meaningful without knowing the units of “West.”\par
+
+\reply
+
 \action: Our actions. Clarify. In a larger scale this could be improved by having auxiliary data for instance from ocean models or satellite data.
@@
 
-\item[1.1.7]\label{ec7} Page 24, first paragraph of Fig. 8: Please define “the modeled quadratic form of the residuals” more precisely. Perhaps a mathematical expression is needed for clarity.\par
+%\item[1.1.7]\label{ec7} 
+\vspace{5mm}
+\noindent \aecom
+
+Page 24, first paragraph of Fig. 8: Please define “the modeled quadratic form of the residuals” more precisely. Perhaps a mathematical expression is needed for clarity.\par
+
+\reply
+
 \action: Our actions
@@
 
-\end{itemize}
+%\end{itemize}
 
-\item{Minor comments to the Author}\label{emc}
+%\item{Minor comments to the Author}\label{emc}
 
-\reply: 
+\vspace{5mm}
+\noindent {\bf{Minor AE comments to the Authors}}
+%\reply: 
 \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
 
-\item[1.2.1]\label{emc1} Page 3, line 19: Change “bi-variate” to “bivariate”. The same change is needed on the next-
+%\item[1.2.1]\label{emc1} 
+\item Page 3, line 19: Change “bi-variate” to “bivariate”. The same change is needed on the next-
 to-last line of Page 12.
 
-\item[1.2.2]\label{emc2} Page 9, three lines after (10): I presume that all off-diagonal elements of R are zeros, but this
+\action: We have changed this in the text - page 3 and 12.
+
+%\item[1.2.2]\label{emc2} 
+\item Page 9, three lines after (10): I presume that all off-diagonal elements of R are zeros, but this
 should be stated explicitly.
 
-\item[1.2.3]\label{emc3} Page 19, fourth line from bottom: Change ``as also" to ``as well as"
+\action: We have changed the statement on page 9 to: ``$\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations."
 
-\item[1.2.4]\label{emc4} Page 30, second and third references listed: ``et al." is listed after the authors’ names for each
+%\item[1.2.3]\label{emc3} 
+\item Page 19, fourth line from bottom: Change ``as also" to ``as well as"
+
+\action: We have changed this in the text - page 19.
+
+%\item[1.2.4]\label{emc4} 
+\item Page 30, second and third references listed: ``et al." is listed after the authors’ names for each
 of these two references, but actually there are no additional authors beyond the two listed.
 
+\action: we have made this change to the citations \cite{genton2015cross} and \cite{french2016credible} - page 30.
 \end{itemize}
-
-\action: 
-\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
-
-\item[1.2.1] we have changed this in the text - page 3 and 12.
-
-\item[1.2.2] we have changed the statement on page 9 to: ``$\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations."
-
-\item[1.2.3] we have changed this in the text - page 19.
-
-\item[1.2.4] we have made this change to the citations \cite{genton2015cross} and \cite{french2016credible} - page 30.
-
-\end{itemize}
-
-\end{answers}
 
@@
 
-\setcounter{reviewer}{2}
+\setcounter{reviewer}{1}
 
@@
 %----------------------------------------------------------------------
-\item{Reviewer comment 4}\label{r1c4}
+\item{\revcom:  4}\label{r1c4}
 
-\reply: Adjustments and improvements towards the intended audience?
+Adjustments and improvements towards the intended audience?
 
@@
 
-\begin{quote}
-\frevcomment{a change in the text}
-\end{quote}
+%\begin{quote}
+%\frevcomment{a change in the text}
+%\end{quote}
 
-\begin{quote}
-  \frevcomment{another change in the text} \end{quote}
+%\begin{quote}
+%  \frevcomment{another change in the text} 
+%\end{quote}
 
-\item{Reviewer comment 4 cont.}\label{r1c4}
+\item{\revcom: 4 cont.}\label{r1c4}
 
-\reply: On including relevant code.
+On including relevant code.
 
@@
 
-\item{Reviewer comment 5}\label{r1c5}
+\item{\revcom: 5}\label{r1c5}
 
-\reply: On shortening the paper and moving important sentences earlier.
+On shortening the paper and moving important sentences earlier.
 Reduced oceanographic background, while highlighting elements that reviewers did not understand about the AUV planning, waypoint graph, etc.
@@
 
-\item{Reviewer comment 6}\label{r1c6}
+\item{\revcom: 6}\label{r1c6}
 
-\reply: It is not clear why the binary classification and excursion sets are being used.
+It is not clear why the binary classification and excursion sets are being used.
 Is there really a sharp boundary at the edge of a plume?
@@
 
-\item{Reviewer comment 7}\label{r1c7}
+\item{\revcom: 7}\label{r1c7}
 
-\reply: The sequential sampling approaches are not novel.
+ The sequential sampling approaches are not novel.
 they are mentioned as examples in the cited paper by Bect et al.
@@
 
-\item{Reviewer comment 8}\label{r1c8}
+\item{\revcom:  8}\label{r1c8}
 
-\reply: While I can see some advantages of simplicity in the IBV of Bect et al for binary data, one could have used entropy as it has information-theoretical credentials.
+While I can see some advantages of simplicity in the IBV of Bect et al for binary data, one could have used entropy as it has information-theoretical credentials.
 Furthermore it could cover design both for spatial mapping as well as indicator mapping.
@@
 
-\item{Reviewer comment 9}\label{r1c9}
+\item{\revcom:  9}\label{r1c9}
 
-\reply: On the issue of covariance separability, objections have always been raised when I have used it as its unrealistic.
+On the issue of covariance separability, objections have always been raised when I have used it as its unrealistic.
 My rejoinder has always worked, that of citing the “all models are wrong”, George Box’s famous phrase.
@@
 
-\item{Reviewer comment 10}\label{r1c10}
+\item{\revcom:  10}\label{r1c10}
 
-\reply: The strategy at the bottom of page 17 seems ad hoc.
+The strategy at the bottom of page 17 seems ad hoc.
 Could you not use the data collected with the two ahead design to get better estimates of parameters and then with these parameters and thereafter a better myopic design?
 
-\action: LA and LA2 improved in revision! + code. Also studied benefits of look-ahead strategies in sequential sampling in \cite{martinelli2013dynamic}.
+\action: LA improved in revision! + code. %Also studied benefits of look-ahead strategies in sequential sampling in \cite{martinelli2013dynamic}.
 
-\item{Reviewer comment 11}\label{r1c11}
+\item{\revcom:  11}\label{r1c11}
 
-\reply: Are these results generalizable?
+Are these results generalizable?
 In other words, how much plume-to-plume variability is there?
@@
 
-\item{Reviewer comment 12}\label{r1c12}
+\item{\revcom:  12}\label{r1c12}
 
-\reply: The separability assumption is to be relaxed in future work. Should publication of this case study await the proposed improvement? It would make for a much better paper.
+The separability assumption is to be relaxed in future work. Should publication of this case study await the proposed improvement? It would make for a much better paper.
 
@@
 
-\item{Specific comments}\label{q11}
+\item{Reviewer specific comments}\label{q11}
 
@@
 
-\item[2.1.1] “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.\par
+\item[1.1.1] “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.\par
 \action: Our actions
@@
 
-\item[2.1.2] Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?\par
+\item[1.1.2] Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?\par
 \action: Preliminary runs set these trends. In a larger scale, one could here use ocean model inputs or satellite data.
@@
 
-\item[2.1.3] I don’t think Figure 3 conveys too much additional benefit and eliminating it for space might be considered.\par
+\item[1.1.3] I don’t think Figure 3 conveys too much additional benefit and eliminating it for space might be considered.\par
 \action: Our actions
@@
 
-\item[2.1.4] Why are the ES’s called ”excursion sets? Why excursion probabilities?\par
+\item[1.1.4] Why are the ES’s called ”excursion sets? Why excursion probabilities?\par
 \action: Our actions
@@
 
-\item[2.1.5] I did not spot a definition for the “waypoint graph” a phrase used several times in the paper.
+\item[1.1.5] I did not spot a definition for the “waypoint graph” a phrase used several times in the paper.
 Again points to the need to identify the audience.
@@
 
-\item[2.1.6] In statistics journals, random variables are commonly labelled by Roman characters e.g.
+\item[1.1.6] In statistics journals, random variables are commonly labelled by Roman characters e.g.
 Z, parameters in Greek e.g.
@@
 
-\item[2.1.7] There is not much of novelty in the Gaussian Process theory used in the paper- its pretty much standard.
+\item[1.1.7] There is not much of novelty in the Gaussian Process theory used in the paper- its pretty much standard.
 I suggest that the derivations stated in detail be summarized by just the results.
@@
 
-\setcounter{reviewer}{3}
+\setcounter{reviewer}{2}
 
 \begin{answers}
-\item{Major comment 1 - Details on AUVs}\label{r2c1}
+\item{\revcom:  1 - Details on AUVs}\label{r2c1}
 
-\reply: I know very little about the AUVs you discuss, and I suspect many of your readers will be in a similar position. Additional contextual information would be helpful
+I know very little about the AUVs you discuss, and I suspect many of your readers will be in a similar position. Additional contextual information would be helpful
 
@@
 
-\item[3.1.1] What sort of onboard computing do they have?
+\item[2.1.1] What sort of onboard computing do they have?
 It may not be something like an Intel 2.1 GHz, 4-thread processor, but a more informative description of something most readers relate to would be helpful.
@@
 
-\item[3.1.2] How long do the AUVs typically explore in one go? Is there a minimum/maximum time?\par 
+\item[2.1.2] How long do the AUVs typically explore in one go? Is there a minimum/maximum time?\par 
 \action: Max speed is limit for coverage. Time constraints for operations related to costs of man-hours in operation. Battery capacity is not a challenge for this AUV. 
@@
 
-\item[3.1.3] What kind of movements can the AUVs make? Chess-like moves (horizontal, vertical, diagonal)?\par  
+\item[2.1.3] What kind of movements can the AUVs make? Chess-like moves (horizontal, vertical, diagonal)?\par  
 \action: We clarified the descriptions of a way-point graph.
@@
 
-\item[3.1.4] Is there a minimum/maximum number of points you want to sample?\par
+\item[2.1.4] Is there a minimum/maximum number of points you want to sample?\par
 \action: Assuming a fixed time of sampling.
@@
 
-\item[3.1.5] What is the size of the grid that is typically considered? i.e., what is a typical value of $n$ in Eq.(9)?\par 
+\item[2.1.5] What is the size of the grid that is typically considered? i.e., what is a typical value of $n$ in Eq.(9)?\par 
 \action: Only surface considered here. No 3D with depth. 
@@
 
-\item[3.1.6] Does the AUV use a selection algorithm from the very beginning or is there a minimum number of initial measurements that are made before the optimization algorithm begins?\par
+\item[2.1.6] Does the AUV use a selection algorithm from the very beginning or is there a minimum number of initial measurements that are made before the optimization algorithm begins?\par
 \action: Batch initialization. Then adaptive sampling.
@@
 
-\item{Major comment 2}\label{r2c2}
+\item{\revcom:  2}\label{r2c2}
 
-\reply: p. 9, Eq. (8): the dependency on $\bx$ and $\bx'$ has been omitted for some of the parameters without mention. Or are the parameters constant across locations?
+p. 9, Eq. (8): the dependency on $\bx$ and $\bx'$ has been omitted for some of the parameters without mention. Or are the parameters constant across locations?
 
@@
 
-\item{Major comment 3}\label{r2c3}
+\item{\revcom:  3}\label{r2c3}
 
-\reply: p.
+p.
 9, How reasonable is the assumption of stationary and isotropic random fields?
@@
 
-\item{Major comment 4}\label{r2c4}
+\item{\revcom:  4}\label{r2c4}
 
-\reply: p. 9, Eq. (10). Are the temperature and salinity processes at a specific location effectively independent of one another? (I truly don’t know). That is what is implied by Eq. (10).
+p. 9, Eq. (10). Are the temperature and salinity processes at a specific location effectively independent of one another? (I truly don’t know). That is what is implied by Eq. (10).
 
@@
 
-\item{Major comment 5}\label{r2c5}
+\item{\revcom:  5}\label{r2c5}
 
-\reply: p.
+p.
 9, Eq.
@@
 
-\item{Major comment 6}\label{r2c6}
+\item{\revcom:  6}\label{r2c6}
 
-\reply: p.
+p.
 9, Eq.
@@
 
-\item{Major comment 7}\label{r2c7}
+\item{\revcom:  7}\label{r2c7}
 
-\reply: p.
+p.
 11, Eq.
@@
 
-\item{Major comment 8}\label{r2c8}
+\item{\revcom:  8}\label{r2c8}
 
-\reply: p. 11, Is $\bpi(\bm)$ the density associated with the distribution in Eq. (14)?
+p. 11, Is $\bpi(\bm)$ the density associated with the distribution in Eq. (14)?
 
@@
 
-\item{Major comment 9}\label{r2c9}
+\item{\revcom:  9}\label{r2c9}
 
-\reply: p.
+p.
 11, Is it purely for stylistic reasons that $P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\bm)P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\bm)$ isn’t written as $P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\bm)^2$?
@@
 
-\item{Major comment 10}\label{r2c10}
+\item{\revcom:  10}\label{r2c10}
 
-\reply: p.
+p.
 11, It is stated that $\bZ$ is chosen to be independent of $\bm$, so presumably the joint CDF $F_{z,m}=F_{Z}F_m$ which would certainly be convenient, but is not obvious to me why this would be true.
@@
 
-\item{Major comment 11}\label{r2c11}
+\item{\revcom:  11}\label{r2c11}
 
-\reply: p. 11, Eq. (21). The trailing $p(\bm)$ should be $\bpi(\bm)$. Should the right side of the equality be squared?
+p. 11, Eq. (21). The trailing $p(\bm)$ should be $\bpi(\bm)$. Should the right side of the equality be squared?
 
@@
 
-\item{Major comment 12}\label{r2c12}
+\item{\revcom:  12}\label{r2c12}
 
-\reply: This method relies on computing the marginal probabilities in each grid cell.
+This method relies on computing the marginal probabilities in each grid cell.
 This is computationally convenient and is likely required because of the limited computing power of the AUVs.
@@
 
-\item{Major comment 13}\label{r2c13}
+\item{\revcom:  13}\label{r2c13}
 
-\reply: Fig. 7 is difficult to understand. The goal is to show the various paths taken by the AUV in the
+Fig. 7 is difficult to understand. The goal is to show the various paths taken by the AUV in the
 simulation example. But it is very difficult to determine which paths are very common and which
@@
 
-\item{Major comment 14}\label{r2c14}
+\item{\revcom:  14}\label{r2c14}
 
-\reply: While the context doesn’t allow full reproducibility of the paper’s results, no implementation details are provided.
+While the context doesn’t allow full reproducibility of the paper’s results, no implementation details are provided.
 What software was used to perform the bulk of analysis?
@@
 
-\item{Minor comments}\label{r2cminor}
+\item{\revcom: Minor comments}\label{r2cminor}
 
-\reply: \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
-
-\item[3.15.1] “closed-form” is sometimes written as “closed form” (e.g., p. 7, p. 18)
-
-\item[3.15.2] p. 9, `1s` should be ‘1s’. Similar for 0s.
-
-\item[3.15.3] p. 15, my manuscript has -on where to sample-, which is an unusual use of dashes.
-
-\end{itemize}
-
-\action:
 \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
 
-\item[3.15.1] 
-\item[3.15.2] 
-\item[3.15.3]
+\item[2.15.1] “closed-form” is sometimes written as “closed form” (e.g., p. 7, p. 18)
+\action:
 
+\item[2.15.2] p. 9, `1s` should be ‘1s’. Similar for 0s.
+\action:
+
+\item[2.15.3] p. 15, my manuscript has -on where to sample-, which is an unusual use of dashes.
+\action:
 \end{itemize}
@@
 
-\setcounter{reviewer}{4}
+\setcounter{reviewer}{3}
 
@@
 
-\item{Comment:}\label{r3c1}
+\item{\revcom : Major 1:}\label{r3c1}
 
-\reply: I think this is an interesting paper. The paper is well written, 
+I think this is an interesting paper. The paper is well written, 
 and the authors have motivated the application well. 
@@
 
-\item{Minor comment 1}\label{r3c2}
+\item{\revcom : Minor 1}\label{r3c2}
 
@@
 
-\item{Major comment 2}\label{r3c3}
+\item{\revcom : Major 2}\label{r3c3}
 

### 2020-05-13 main_new.tex
Jo: revised sect 1 and 2 (ongoing)

@@
 \author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
-\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t3}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
-\author{\fnms{David} \snm{Ginsbourger}\thanksref{t4}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
+\author{\fnms{Cédric} \snm{Travelletti}\thanksref{t3}, \corref{} \ead[label=e1]{cedric.travelletti@stat.unibe.ch}}
+\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
+\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
 \and
@@
 \affiliation[t2]{Centre for Autonomous Marine Operations and Systems (AMOS).}
-\affiliation[t3]{Department of Mathematical Sciences, NTNU.}
-\affiliation[t4]{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland.}
-% \affiliation[t5]{Department of Engineering Cybernetics, NTNU}
+\affiliation[t3]{Institute of Mathematical Statistics and Actuarial Science, University of Bern, Switzerland.}
+\affiliation[t4]{Department of Mathematical Sciences, NTNU.}
 \affiliation[t5]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
@@
 effective means of sampling.
-By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use of autonomous platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
-The potential that lies in using information-based methods \citep{das11b,Das2015,fossum18b,fossuminformation}, which we highlight in this paper, is to do efficient environmental sampling and overcome three common challenges, namely:
+By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use of new technologies including autonomous platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
+In this paper we build upon the potential that lies in using autonomous underwater vehicles (AUVs) to conduct efficient sampling in the ocean \citep{das11b,Das2015,fossuminformation,fossum18b}.
+Addressing this, the combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based environmental sensing.
+
+There are several obstacles when doing sampling with an AUV in the ocean, and we discuss some of these to motivate our setting, and also the modeling and methodological choices that will be made.
+First, full numerical ocean models based on complex physical processes cannot provide accurate results for online operations because the onboard computers cannot deliver the computational power required.
+Hence statistical proxy models of the environment must be used on the AUV.
+For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest.
+Second, with limited available information about the state of the ocean, one must usually plan for a targeted operation where a relevant criterion is used to extract the most valuable designs, preferable in such a way that there is substantial value in reacting to information obtained from measurements taken in-situ.
+The acquired information must then be assimilated into the statistical model onboard the AUV, so that it can be used to re-compute the sampling criterion and in this way actively inform decisions on where to sample sequentially.
+We use a criterion focusing on the prediction of excursion sets (ESs) for multivariate processes of interest.
+Third, the number of sampling choices (i.e.
+locations, trajectories, and candidate
+designs) is enormous, creating a trade-off between optimization (finding the most resource-efficient design to collect necessary data) and computability (arriving at a
+solution in reasonable time).
+One must often compromise in both development and practical operations, and we will rely on short-horizon optimization routines among sequential designs that encapsulate nearest neighbors sampling in a waypoint graph created for the AUV operation.
 
@@
 
-\begin{itemize}
-
-\item Full numerical ocean models cannot provide accurate results
-  online if run on robotic sensing platforms, as onboard computers
-  cannot deliver the computational power required. Hence statistical
-  proxy models of the environment must be used for learning where to sample.
-
-\item With limited available information about the state of the ocean, there is substantial value in reacting to
-  information obtained from measurements taken in-situ. This
-  acquired information must be assimilated into statistical models that
-  can be used to inform decisions on where to sample
-  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.
-
-\item For sampling problems related to environmental sensing, the
-  number of choices (i.e. locations, trajectories, and candidate
-  designs) is enormous, creating a
-  trade-off between optimization (finding the most resource-efficient
-  design to collect necessary data) and computability (arriving at a
-  solution in reasonable time). To successfully resolve features, this
-  trade-off has to be considered in development and practice.
-
-\end{itemize}
-
-Addressing this, the combination of statistical tools and robotic platforms is a
-natural symbiosis which enables information-based sensing. Central to
-this is the ability to model spatially-correlated variables and
-provide formal measures of uncertainty. Our formulation is based on
-Gaussian Processes (GPs) as they allow efficient implementation 
-and evaluation in real time onboard a robotic platform.
+%\begin{itemize}
+%
+%\item Full numerical ocean models cannot provide accurate results
+%  online if run on robotic sensing platforms, as onboard computers
+%  cannot deliver the computational power required. Hence statistical
+%  proxy models of the environment must be used for learning where to sample.
+%
+%\item With limited available information about the state of the ocean, there is substantial value in reacting to
+%  information obtained from measurements taken in-situ. This
+%  acquired information must be assimilated into statistical models that
+%  can be used to inform decisions on where to sample
+%  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.
+%
+%\item For sampling problems related to environmental sensing, the
+%  number of choices (i.e. locations, trajectories, and candidate
+%  designs) is enormous, creating a
+%  trade-off between optimization (finding the most resource-efficient
+%  design to collect necessary data) and computability (arriving at a
+%  solution in reasonable time). To successfully resolve features, this
+%  trade-off has to be considered in development and practice.
+%
+%\end{itemize}
+
+%Addressing this, the combination of statistical tools and robotic platforms is a
+%natural symbiosis which enables information-based sensing. Central to
+%this is the ability to model spatially-correlated variables and
+%provide formal measures of uncertainty. Our formulation is based on
+%Gaussian Processes (GPs) as they allow efficient implementation 
+%and evaluation in real time onboard a robotic platform.
 
@@
 
-In this work, we build on joint excursion sets (ES) and excursion
-probabilities (EP) for the spatial design of experiments. The goal is
+We build on ESs and excursion
+probabilities (EPs) for the spatial design of experiments. The goal is
 to design sampling strategies for reducing the spatially-integrated
-Bernoulli variance (IBV) in the ES indicators. Our main contributions
-include: i) extending the IBV criteria to multivariate
-GPs, thereby allowing the evaluation to include multiple environmental
-parameters such as temperature and salinity; ii) developing methods
-for information-based environmental sensing and autonomous data
-collection; and iii) demonstrating the approach in the real-world
-scenario of sampling a river plume using an AUV.
+Bernoulli variance (IBV) in the ES indicator variables. 
+Recent work on ESs and EPs connected to spatial statistics include
+\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
+Our focus is on uncertainty
+characterization and how it is influenced by sampling. In this sense,
+our work is similar to \cite{bect2012}, \cite{chevalier2014fast}, and
+\cite{azzimonti2016quantifying} who describe analytical results for
+the IBV for univariate processes. A major contribution of our work is
+to derive closed-form results for this design criteria for situations
+where the underlying model is based on multivariate GPs. Another is that we embed this approach in a sequential strategy that runs onboard an AUV, and in applying this in the real-world
+scenario of autonomously sampling temperature and salinity variables in a river plume.
+
+Motivating examples relevant for ESs of multivariate processes are abundant.
+In medicine, doctors do not solely rely on a single symptom but must see several combined effects before making a diagnosis.
+In our context of spatial and environmental applications, a famous example is that of photosynthesis where plants and algae produce oxygen only when given enough of every component; light, carbon dioxide and water.
+Our key motivation in this paper is that temperature and salinity, combined, drive several processes in the oceans.
+For instance, salmon lice larvae are known to depend on relatively high temperature and high salinities to survive.
+Salmon lice are parasites to salmon, and the bivariate ES for temperature and salinity is hence important to understand for sustainable activities in salmon fisheries.
+The salinity and temperature excursions of a river plume similarly characterize its oceanographic and bio-geochemical processes \cite{hopkins2013detection,Pinto2018}, and sampling should be aimed at mapping relevant parts to reduce uncertainty in the temperature and salinity ES.
 
@@
 water-column is an understanding of biophysical interaction driven by
-an agglomeration of physical forcing (wind, topography, bathymetry,
-tidal influences, and local geography, such as fresh water inflows) and
-incipient micro-biology driven by planktonic and coastal anthropogenic
-input, such as pollution and agricultural runoff transported into the ocean by rivers and streams.
-These often result in a range of ecosystem-related phenomena such as blooms, anoxic zones and plumes, all of which have significant direct and indirect effects on society.
+an agglomeration of physical forcing %(wind, topography, bathymetry,
+%tidal influences, and local geography, such as fresh water inflows) 
+and incipient micro-biology driven by planktonic and coastal anthropogenic
+input, such as pollution and agricultural runoff transported into the ocean by rivers and streams. These often result in a range of ecosystem-related phenomena such as blooms%, anoxic zones
+and plumes. %, all of which have significant direct and indirect effects on society.
 
@@
 Traditional data collection at sea has typically been based on static
-buoys, Lagrangian floats, or ship-based methods, with significant
+buoys, floats, or ship-based methods, with significant
 logistical limitations that directly impact coverage and sampling
@@
 resulted in significant contributions to environmental monitoring and
-sampling. In particular, autonomous underwater vehicles (AUVs), have
+sampling. In particular, AUVs have
 advanced the state of sampling and consequently have made robotics an
@@
     collaborative networked operations of robotic vehicles to resolve
-    numerous bio-geochemical processes and their interactions.
-AUV platforms are an integral part of this network being able to reason and make decisions for efficient onboard adaptive sampling using the sense-plan-act control approach to autonomous control.
+    numerous bio-geochemical processes. % and their interactions. 
+    AUV platforms are an integral part of this network being able to reason and make decisions for efficient onboard adaptive sampling.
+% using the sense-plan-act control approach to autonomous control.
     \ref{fig:nidelven} The interaction of river and ocean creates
@@
 salinity. Because of the local topography and the Coreolis force
-\citep{coriolis1835memoire} the cold fresh water tends to flow near
-land to the east. Depending on the river discharge, tidal effects,
-wind, and temperature differences, this boundary often gets
-distorted. Initial knowledge about the location and evolution of these
-features is therefore highly uncertain, making deterministic planning
-challenging. River plumes belong to a class of ocean processes that are local and
-act on smaller \emph{sub-mesoscale} extent (from $5 m^2$ -- $10 km^2$) where
-lateral spatial variability tends to dominate. Remote sensing and
+%\citep{coriolis1835memoire} 
+the cold fresh water tends to flow near
+land to the east, %. Depending on the river discharge, tidal effects,
+%wind, and temperature differences, this boundary often gets
+%distorted. Initial knowledge about the location and evolution of these
+but the locations of these features are highly uncertain, making deterministic planning
+challenging. 
+%River plumes belong to a class of ocean processes that are local and
+%act on smaller \emph{sub-mesoscale} extent (from $5 m^2$ -- $10 km^2$) where
+%lateral spatial variability tends to dominate. 
+Remote sensing and
 synthetic ocean models often find it challenging to provide detail at
 this scale \citep{Lermusiaux:2006} and the principal way to resolve
-for fine resolution is by direct in-situ observation. At larger
-\emph{mesoscale} ($>50 km^2$), both three-dimensional space and time
-dynamics are important and can shift substantially, while in the case
-of river plumes one can often limit the scope to cover only lateral
-spatial elements. However, the methodological framework can be
-extended to higher dimensional situations in a similar manner.
+for fine resolution is by direct in-situ observation. 
+%At larger \emph{mesoscale} ($>50 km^2$), both three-dimensional space and time
+%dynamics are important and can shift substantially, while in the case
+At this scale of river plumes one can further often limit scope to covering only lateral near surface
+spatial elements, and the case appears very amenable for AUVs usage. 
+%However, the methodological framework can be extended to higher dimensional situations in a similar manner.
 
@@
 AUVs are used to gather in-situ spatio-temporal data while carrying a
-range of scientific payloads to survey the water column. Typically,
-they operate at $\sim 1-3$ m/s, can reach depths from 100--6000 m with
-an in-water operation time capacity dependent on survey speed, payload
-sensors and mission design.
+range of scientific payloads to survey the water column. The ones we considere here
+operate at $\sim 1-3$ m/s, covering the upper water column, with an in-water operation time capacity dependent on survey speed, payload sensors and mission design.
 AUVs have limited computational capacity due to their limited energy and payload space.
@@
 
-By deploying AUVs together with other
-resources, one can augment information from synthetic ocean models,
-remote-sensing data, or fixed-location sensors on buoys. Traditional
-operations and surveys with AUVs are usually limited to observations
-along fixed transects on a spatial grid, pre-programmed by the human
-operator. A more effective approach is to instead use onboard algorithms to continuously evaluate,
-update, and refine future sampling locations (sense-plan-act cycle in Fig. \ref{fig:envir1}), making the information
+%By deploying AUVs together with other resources, one can augment information from synthetic ocean models, remote-sensing data, or fixed-location sensors on buoys. 
+Surveys with AUVs are usually limited to observations
+along fixed transects on a spatial grid, called a waypoint graph, which is pre-programmed by the human operator.
+A more effective approach is to instead use onboard algorithms to continuously evaluate, update, and refine future sampling locations (sense-plan-act cycle in Fig.
+\ref{fig:envir1}), making the information
 gathering \emph{adaptive} and \emph{contextual} to the survey
 requirements \citep{das11b,fossum18b,fossuminformation}. The space of
-sampling opportunities can still be limited to a grid, but adaptation
+sampling opportunities is still limited by the waypoint graph, but adaptation
 allows more flexibility. For instance, an AUV could reconstruct or
@@
 
+\cite{adler2000excursion,adler2009random}.
+
 Our objective is to leverage statistical tools to characterize a river
@@
 
-Recent work on ESs and EPs connected to spatial statistics include
+%Recent work on ESs and EPs connected to spatial statistics include
 \cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
-Our focus is on ES as defined in Eq. (\ref{ES}) and the uncertainty
-reduction achieved by sampling as in Eq. (\ref{sur}). In this sense,
-our work is similar to \cite{bect2012}, \cite{chevalier2014fast}, and
-\cite{azzimonti2016quantifying} who describe analytical results for
-the IBV for univariate processes. A major contribution of our work is
-to derive closed-form results for this design criteria for situations
-where the underlying model is based on multivariate GPs.
+%Our focus is on ES as defined in Eq. (\ref{ES}) and the uncertainty
+%reduction achieved by sampling as in Eq. (\ref{sur}). In this sense,
+%our work is similar to \cite{bect2012}, \cite{chevalier2014fast}, and
+%\cite{azzimonti2016quantifying} who describe analytical results for
+%the IBV for univariate processes. A major contribution of our work is
+%to derive closed-form results for this design criteria for situations
+%where the underlying model is based on multivariate GPs.
 

### 2020-05-13 main_new.tex
Sect 1-2 again. (Jo)

@@
 In this paper we build upon the potential that lies in using autonomous underwater vehicles (AUVs) to conduct efficient sampling in the ocean \citep{das11b,Das2015,fossuminformation,fossum18b}.
-Addressing this, the combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based environmental sensing.
+%Addressing this, the combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based environmental sensing.
 
 There are several obstacles when doing sampling with an AUV in the ocean, and we discuss some of these to motivate our setting, and also the modeling and methodological choices that will be made.
-First, full numerical ocean models based on complex physical processes cannot provide accurate results for online operations because the onboard computers cannot deliver the computational power required.
+First, full numerical ocean models based on complex differential equations cannot provide accurate results for online operations because the onboard computers lack the computational power required.
 Hence statistical proxy models of the environment must be used on the AUV.
 For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest.
-Second, with limited available information about the state of the ocean, one must usually plan for a targeted operation where a relevant criterion is used to extract the most valuable designs, preferable in such a way that there is substantial value in reacting to information obtained from measurements taken in-situ.
-The acquired information must then be assimilated into the statistical model onboard the AUV, so that it can be used to re-compute the sampling criterion and in this way actively inform decisions on where to sample sequentially.
-We use a criterion focusing on the prediction of excursion sets (ESs) for multivariate processes of interest.
-Third, the number of sampling choices (i.e.
-locations, trajectories, and candidate
-designs) is enormous, creating a trade-off between optimization (finding the most resource-efficient design to collect necessary data) and computability (arriving at a
-solution in reasonable time).
-One must often compromise in both development and practical operations, and we will rely on short-horizon optimization routines among sequential designs that encapsulate nearest neighbors sampling in a waypoint graph created for the AUV operation.
+Second, %with limited available information about the state of the ocean,
+with undersampling issues one must usually plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs.
+%, preferable in such a way that there is substantial value in reacting to information obtained from measurements taken in-situ.
+The acquired in-situ information must then be assimilated into the statistical model onboard the AUV, so that it can be used to re-compute the sampling criterion and in this way inform decisions on where to sample next.
+We use a criterion focusing on the prediction of excursion sets (ESs) for the multivariate spatial processes of interest.
+Third, the number of sampling designs is enormous, creating a trade-off between optimization and computability.
+One must often compromise in both development and practical operations, and we will rely on short-horizon optimization routines for sequential designs among nearest neighbors points in a spatial grid created for the AUV operation.
 
@@
 
-We build on ESs and excursion
-probabilities (EPs) for the spatial design of experiments. The goal is
-to design sampling strategies for reducing the spatially-integrated
+We design AUV sampling strategies for reducing the spatially-integrated
 Bernoulli variance (IBV) in the ES indicator variables. 
-Recent work on ESs and EPs connected to spatial statistics include
-\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
+Recent work on ESs and associated excursion probabilities (EPs) connected to spatial statistics include
+\cite{french2013spatio,bolin2015excursion,french2016credible}.
 Our focus is on uncertainty
 characterization and how it is influenced by sampling. In this sense,
-our work is similar to \cite{bect2012}, \cite{chevalier2014fast}, and
-\cite{azzimonti2016quantifying} who describe analytical results for
+our work is similar to \cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who describe analytical results for
 the IBV for univariate processes. A major contribution of our work is
 to derive closed-form results for this design criteria for situations
-where the underlying model is based on multivariate GPs. Another is that we embed this approach in a sequential strategy that runs onboard an AUV, and in applying this in the real-world
+where the underlying model is based on multivariate GPs.
+Another contribution is that we embed this approach in a sequential strategy to produce algorithms, and in applying this on an AUV in the real-world
 scenario of autonomously sampling temperature and salinity variables in a river plume.
@@
 Motivating examples relevant for ESs of multivariate processes are abundant.
-In medicine, doctors do not solely rely on a single symptom but must see several combined effects before making a diagnosis.
-In our context of spatial and environmental applications, a famous example is that of photosynthesis where plants and algae produce oxygen only when given enough of every component; light, carbon dioxide and water.
+In medicine, doctors do not rely solely on a single symptom but must see several combined effects before making a diagnosis.
+In our context of environmental applications, an example is that of photosynthesis where plants and algae produce oxygen only when exposed to enough light, carbon dioxide and water.
 Our key motivation in this paper is that temperature and salinity, combined, drive several processes in the oceans.
 For instance, salmon lice larvae are known to depend on relatively high temperature and high salinities to survive.
-Salmon lice are parasites to salmon, and the bivariate ES for temperature and salinity is hence important to understand for sustainable activities in salmon fisheries.
-The salinity and temperature excursions of a river plume similarly characterize its oceanographic and bio-geochemical processes \cite{hopkins2013detection,Pinto2018}, and sampling should be aimed at mapping relevant parts to reduce uncertainty in the temperature and salinity ES.
+Salmon lice are parasites to salmon, and the bivariate ES for temperature and salinity is hence important to understand for sustainable salmon fisheries management.
+The salinity and temperature excursions of a river plume can similarly help characterize its oceanographic and bio-geochemical processes \cite{hopkins2013detection,Pinto2018}, and sampling should be aimed at mapping relevant parts to reduce uncertainty in the temperature and salinity ES.
+Unlike what has been done in previous plume exploration studies, we emphasize the links to spatial statistical modeling and targeted multivaraite sampling criteria for sequential sampling.
+Addressing this, the combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based environmental sensing.
 
@@
 
-\section{Background and Motivation}\label{sec:bg}
-
-\subsection{Observing Processes in the Upper Ocean}
+\section{Background on Ocean Processes and Observations}\label{sec:bg}
+
+%\subsection{Observing Processes in the Upper Ocean}
 
@@
 Central to understanding the changes taking place in the upper
-water-column is an understanding of biophysical interaction driven by
-an agglomeration of physical forcing %(wind, topography, bathymetry,
+water-column is knowledge of the biophysical interaction driven by
+an agglomeration of physical forcings %(wind, topography, bathymetry,
 %tidal influences, and local geography, such as fresh water inflows) 
@@
 % On observation and why robotics
-Traditional data collection at sea has typically been based on static
+Traditional data collection to understand such phenomena at sea has typically been based on static
 buoys, floats, or ship-based methods, with significant
@@
 large-scale coverage but have limited resolution, are limited to
-sensing the surface of the ocean, and are impacted by cloud cover. The
-advent of mobile robotic platforms \citep{Bellingham07} has
+sensing the surface, and are impacted by cloud cover. Numerical ocean models similarly find it challenging to provide detail at
+smaller scale \citep{Lermusiaux:2006}. 
+The advent of mobile robotic platforms \citep{Bellingham07} has
 resulted in significant contributions to environmental monitoring and
 sampling. In particular, AUVs have
-advanced the state of sampling and consequently have made robotics an
-integral part of ocean observation.
-The previously listed challenges related to undersampling have also been driving ocean sensing practices towards more autonomous, networked, and distributed use of robotic assets (See Fig.
+advanced the state of in-situ sampling, and made robotics an
+integral part of ocean observation, filling parts of the undersampling gap (See Fig.
+\ref{fig:envir1}).
+%The previously listed challenges related to undersampling have also been driving ocean sensing practices towards more autonomous, networked, and distributed use of robotic assets (See Fig.
 \ref{fig:envir1}), where observation is guided by assimilation of recent measurements and statistical models.
@@
   \caption{\ref{fig:envir1} Traditional ocean observation based on
-    ship-based sampling has recently been augmented by autonomous and
-    collaborative networked operations of robotic vehicles to resolve
-    numerous bio-geochemical processes. % and their interactions. 
+    ship-based sampling has recently been augmented by autonomous robotic vehicles. % and their interactions. 
     AUV platforms are an integral part of this network being able to reason and make decisions for efficient onboard adaptive sampling.
@@
 
-The application that is of focus in this work is towards spatial characterization of a frontal system specifically, fronts generated by river
-plumes. Fig. \ref{fig:nidelven} shows the survey area (the outlet of
-the Nidelva river, Trondheim, Norway) where cold freshwater enters
+The AUVs considered here
+operate at $1$-$3$ m/s in the upper water column.
+The in-water operation time capacity depends on survey speed, payload sensors and navigation.
+AUVs have limited computational capacity, and they are usually done on a single-board computer (SBC), like the Raspberry Pi or an multicore GPU NVIDIA Jetson TX1.
+
+%By deploying AUVs together with other resources, one can augment information from synthetic ocean models, remote-sensing data, or fixed-location sensors on buoys. 
+Surveys with AUVs are usually limited to observations
+along fixed transects that are pre-programmed by the human operator.
+This occurs on a spatial grid, called a waypoint graph.
+A more effective approach is to instead use onboard algorithms to continuously evaluate, update, and refine future sampling locations (sense-plan-act cycle in Fig.
+\ref{fig:envir1}), making the information
+gathering \emph{adaptive} \citep{das11b,fossum18b,fossuminformation}. In doing so, the space of sampling opportunities is still limited by the waypoint graph, but the AUV has flexibility to 
+modify its path at each waypoint based on what has been measured onboard
+\citep{py10,Rajan12,Rajan12b}.
+
+The application that is of focus in this work is towards spatial characterization of a frontal system generated by river
+plumes. Fig. \ref{fig:nidelven} shows the survey area (Nidelva, Norway) where cold freshwater enters
 from the river, creating a strong gradient in both temperature and
@@
 %distorted. Initial knowledge about the location and evolution of these
-but the locations of these features are highly uncertain, making deterministic planning
-challenging. 
+but the locations of these features are highly uncertain, making deterministic planning challenging.
+
 %River plumes belong to a class of ocean processes that are local and
@@
 %lateral spatial variability tends to dominate. 
-Remote sensing and
-synthetic ocean models often find it challenging to provide detail at
-this scale \citep{Lermusiaux:2006} and the principal way to resolve
-for fine resolution is by direct in-situ observation. 
 %At larger \emph{mesoscale} ($>50 km^2$), both three-dimensional space and time
 %dynamics are important and can shift substantially, while in the case
-At this scale of river plumes one can further often limit scope to covering only lateral near surface
-spatial elements, and the case appears very amenable for AUVs usage. 
+%At this scale of river plumes one can further often limit scope to covering only lateral near surface spatial elements, and the case appears very amenable for AUVs usage. 
 %However, the methodological framework can be extended to higher dimensional situations in a similar manner.
@@
 
-\subsection{Autonomous Vehicles}
-
-AUVs are used to gather in-situ spatio-temporal data while carrying a
-range of scientific payloads to survey the water column. The ones we considere here
-operate at $\sim 1-3$ m/s, covering the upper water column, with an in-water operation time capacity dependent on survey speed, payload sensors and mission design.
-AUVs have limited computational capacity due to their limited energy and payload space.
-The computations is usually done on a single-board computer (SBC), like the Raspberry Pi or an multicore GPU NVIDIA Jetson TX1.
-
-%By deploying AUVs together with other resources, one can augment information from synthetic ocean models, remote-sensing data, or fixed-location sensors on buoys. 
-Surveys with AUVs are usually limited to observations
-along fixed transects on a spatial grid, called a waypoint graph, which is pre-programmed by the human operator.
-A more effective approach is to instead use onboard algorithms to continuously evaluate, update, and refine future sampling locations (sense-plan-act cycle in Fig.
-\ref{fig:envir1}), making the information
-gathering \emph{adaptive} and \emph{contextual} to the survey
-requirements \citep{das11b,fossum18b,fossuminformation}. The space of
-sampling opportunities is still limited by the waypoint graph, but adaptation
-allows more flexibility. For instance, an AUV could reconstruct or
-modify a survey line based on what temperatures it is measuring by
-using embedded methods in decision-theoretic planning
-\citep{py10,Rajan12,Rajan12b}.
+%\subsection{Autonomous Vehicles}
 
@@
 
-Adaptive sampling of an evolving frontal feature has been explored in
+Adaptive in-situ AUV sampling of an evolving frontal feature has been explored in
 \cite{fronts11,Zhang2012,Pinto2018,costa19}. These approaches
@@
 based on closing the sensing and actuation loop. Myopic sampling, i.e. stage-wise selection of
-the path (on the graph) which reduces the information criterion the
-most, has been used for adaptive surveys
+the path (on the waypoint graph), has been used for surveys
 \citep{singh2009efficient,Binney2013} that focus largely on reducing
-predictive variance or mutual information (e.g. entropy). Variance and
+predictive variance or entropy. But variance and
 entropy reduction are independent of the actual data realizations
-under the assumptions of GP models; the use of data-driven adaptive
+under the assumptions of GP models. The use of data-driven adaptive
 criteria was introduced to include more targeted sampling of regions
 of scientific interest in \cite{Low2009} and \cite{fossuminformation}.
-In this paper, with a focus on mapping the river plume, we reward the
-designs that improve the classification of water masses as a means to
-characterize the frontal zone and thereby the extent of the river plume.
+In this paper, with a focus on mapping the river plume by rewarding the
+designs that improve the classification of ES in temperature and salinity.
 

### 2020-05-15 ans_to_reviewers/answers-to-reviewers.tex
Reply letter structure improved.

@@
 
-\item We have further included Cedrid Traveletti as co-author. He has contributed on deriving statistical properties of the methods and on improving the code for open-source quality.
+\item We have further included Cedric Traveletti as co-author. He has contributed on deriving the more general statistical properties of the methods and on improving the code for open-source quality.
 
@@
 \vspace{5mm}
-\noindent \aecom
+\noindent \aecom 1:
 
@@
 
-\reply
+\reply:
 
@@
 \vspace{5mm}
-\noindent \aecom
+\noindent \aecom 2:
 
@@
 
-\reply We agree that they are not necessary.
+\reply: We agree that they are not necessary for the .
 In the open-source that is now available, we implement non-stationary, non-separable and non-isotropic covariance models.
@@
 \vspace{5mm}
-\noindent \aecom
+\noindent \aecom 3:
 
@@
 \vspace{5mm}
-\noindent \aecom
+\noindent \aecom 4:
 
@@
 \vspace{5mm}
-\noindent \aecom
+\noindent \aecom 5:
 
@@
 \vspace{5mm}
-\noindent \aecom
+\noindent \aecom 6:
 
@@
 \vspace{5mm}
-\noindent \aecom
+\noindent \aecom 7:
 
@@
 %----------------------------------------------------------------------
-\item{\revcom:  4}\label{r1c4}
 
-Adjustments and improvements towards the intended audience?
+\item{\revcom  :}\label{r1c3}
+
+Overall I found this paper interesting.
+Moreover it addresses the topic of spatial design a topic that does not receive the attention it deserves.
+It is driven by the practical need to direct autonomous underwater vehicles where to take samples.
+The sequential sampling strategy may be a valuable addition to the designer’s toolkit.
+
+\action: Thank you. We have improved and documented the code which is now open-source available.
+
+\item{\revcom  :}\label{r1c3}
+
+The things that stand out as novel in this paper are: the need for a se- quential design for driving an autonomous under water vehicle; a way of characterizing water plumes that are washed down from the land into the sea.
+Papers co-authored by one of the authors (Ginsbourger) of this paper inevitably means some overlap of ideas with the other papers, although the two step ahead greedy design algorithm seems novel.
+I seem to recall that Jon Lee published that algorithm albeit for determining the maximum entropy optimum by finding optimal points sequentially over space not time as in this paper.
+The literature on spatial design is huge and often involves spatial sequential designs.
+The handbook chapters by Zidek and Zimmerman (or is the other around) might be worth citing for the benefit of readers inter- ested in some of the background.
+
+\action: We have improved the augmented the discussion on other design criteria and why the ES related IBV is used in our case.
+In this part, we cite the handbook reference on design \cite{zidek2019monitoring} and the paper we believe you refer to with a comment from Jon Lee \cite{bueso1998state}.
+
+\item{\revcom  :}\label{r1c3}
+
+Overall the paper is remarkably free of a need for editorial improve- ments. I will attach a marked up copy indicating a few errors as well as suggestions for minor changes.
+
+\action: We incorporated the suggested minor changes.
+
+\item{\revcom  :}\label{r1c4}
+
+I am not clear about the intended audience for this paper. Or to put it another way, for whom is this paper important? The case study or the
+design methodology?
+Case studies offer concern specific events deemed of importance e.g.
+the much–studied O-ring failure long ago on a US rocket launch.
+Such a case study may then drive the innovation of statistical methods for an analysis of data that were collected because of that case’s importance.
+More likely the case study was intended as a tutorial on the use of a novel design methodology.
+In fact the Closing Remarks section points “the opportunities available real-time multivariable spatial data-gathering” that “statistician can exploit”.
 
@@
 
-%\begin{quote}
-%\frevcomment{a change in the text}
-%\end{quote}
-
-%\begin{quote}
-%  \frevcomment{another change in the text} 
-%\end{quote}
-
-\item{\revcom: 4 cont.}\label{r1c4}
-
-On including relevant code.
+However, the user community is pretty tech-savvy nowadays, and that community might well be the primary beneficiary and more motivated than the statisticians to use the tools for sequentially sampling rivers for mapping plumes.
+That suggests including the relevant code for running the analyses, especially the simulation studies.
+An alterna- tive journal to a statistical one might be more appropriate from that perspective.
 
@@
 
-\item{\revcom: 5}\label{r1c5}
+\item{\revcom :}\label{r1c5}
 
@@
 
-\item{\revcom: 6}\label{r1c6}
+\item{\revcom :}\label{r1c6}
 
@@
 
-\item{\revcom: 7}\label{r1c7}
+\item{\revcom :}\label{r1c7}
 
@@
 
-\item{\revcom:  8}\label{r1c8}
+\item{\revcom  :}\label{r1c8}
 
@@
 
-\item{\revcom:  9}\label{r1c9}
+\item{\revcom  :}\label{r1c9}
 
@@
 
-\item{\revcom:  10}\label{r1c10}
+\item{\revcom  :}\label{r1c10}
 
@@
 
-\item{\revcom:  11}\label{r1c11}
+\item{\revcom :}\label{r1c11}
 
@@
 
-\item{\revcom:  12}\label{r1c12}
+\item{\revcom :}\label{r1c12}
 
@@
 
-\item{Reviewer specific comments}\label{q11}
+\item{\revcom : (specific comments)}\label{q11}
 
-\reply: See the numbered list below:
 \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
 
-\item[1.1.1] “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.\par
+\item “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.\par
 \action: Our actions
@@
 
-\item[1.1.2] Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?\par
+\item Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?\par
 \action: Preliminary runs set these trends. In a larger scale, one could here use ocean model inputs or satellite data.
@@
 
-\item[1.1.3] I don’t think Figure 3 conveys too much additional benefit and eliminating it for space might be considered.\par
+\item I don’t think Figure 3 conveys too much additional benefit and eliminating it for space might be considered.\par
 \action: Our actions
@@
 
-\item[1.1.4] Why are the ES’s called ”excursion sets? Why excursion probabilities?\par
+\item Why are the ES’s called ”excursion sets? Why excursion probabilities?\par
 \action: Our actions
@@
 
-\item[1.1.5] I did not spot a definition for the “waypoint graph” a phrase used several times in the paper.
+\item I did not spot a definition for the “waypoint graph” a phrase used several times in the paper.
 Again points to the need to identify the audience.
@@
 
-\item[1.1.6] In statistics journals, random variables are commonly labelled by Roman characters e.g.
+\item In statistics journals, random variables are commonly labelled by Roman characters e.g.
 Z, parameters in Greek e.g.
@@
 
-\item[1.1.7] There is not much of novelty in the Gaussian Process theory used in the paper- its pretty much standard.
+\item There is not much of novelty in the Gaussian Process theory used in the paper- its pretty much standard.
 I suggest that the derivations stated in detail be summarized by just the results.
@@
 \begin{answers}
-\item{\revcom:  1 - Details on AUVs}\label{r2c1}
+\item{\revcom 1: - Details on AUVs}\label{r2c1}
 
@@
 
-\item{\revcom:  2}\label{r2c2}
+\item{\revcom 2:}\label{r2c2}
 
@@
 
-\item{\revcom:  3}\label{r2c3}
+\item{\revcom 3:}\label{r2c3}
 
@@
 
-\item{\revcom:  4}\label{r2c4}
+\item{\revcom  4:}\label{r2c4}
 
@@
 
-\item{\revcom:  5}\label{r2c5}
+\item{\revcom 5:}\label{r2c5}
 
@@
 
-\item{\revcom:  6}\label{r2c6}
+\item{\revcom  6:}\label{r2c6}
 
@@
 
-\item{\revcom:  7}\label{r2c7}
+\item{\revcom 7:}\label{r2c7}
 
@@
 
-\item{\revcom:  8}\label{r2c8}
+\item{\revcom  8:}\label{r2c8}
 
@@
 
-\item{\revcom:  9}\label{r2c9}
+\item{\revcom 9:}\label{r2c9}
 
@@
 
-\item{\revcom:  10}\label{r2c10}
+\item{\revcom 10:}\label{r2c10}
 
@@
 
-\item{\revcom:  11}\label{r2c11}
+\item{\revcom 11:}\label{r2c11}
 
@@
 
-\item{\revcom:  12}\label{r2c12}
+\item{\revcom 12:}\label{r2c12}
 
@@
 
-\item{\revcom:  13}\label{r2c13}
+\item{\revcom 13:}\label{r2c13}
 
@@
 
-\item{\revcom:  14}\label{r2c14}
+\item{\revcom  14:}\label{r2c14}
 
@@
 
-\item{\revcom : Major 1:}\label{r3c1}
+\item{\revcom - Major 1:}\label{r3c1}
 
@@
 
-\item{\revcom : Minor 1}\label{r3c2}
+\item{\revcom - Minor 1:}\label{r3c2}
 
@@
 
-\item{\revcom : Major 2}\label{r3c3}
+\item{\revcom - Major 2:}\label{r3c3}
 

### 2020-05-15 main_new.tex
Reply letter structure improved.

@@
 \citep{singh2009efficient,Binney2013} that focus largely on reducing
-predictive variance or entropy. But variance and
+predictive variance or entropy. These criteria are widely adopted in the statistics literature on spatio-temporal design as well \cite{bueso1998state,zidek2019monitoring}, but variance and
 entropy reduction are independent of the actual data realizations
-under the assumptions of GP models. The use of data-driven adaptive
+under the assumptions of GP models, so limited decision flexibility is created for sequential designs. The use of data-driven adaptive
 criteria was introduced to include more targeted sampling of regions

### 2020-05-15 ans_to_reviewers/answers-to-reviewers.tex
Reply-letter and 1-2

@@
 \begin{answers}
-\item{\revcom 1: - Details on AUVs}\label{r2c1}
+\item{\revcom : }\label{r2c1}
 
@@
 
-\item{\revcom 2:}\label{r2c2}
+\item{\revcom :}\label{r2c2}
 
@@
 
-\item{\revcom 3:}\label{r2c3}
+\item{\revcom :}\label{r2c3}
 
@@
 
-\item{\revcom  4:}\label{r2c4}
+\item{\revcom  :}\label{r2c4}
 
@@
 
-\item{\revcom 5:}\label{r2c5}
+\item{\revcom :}\label{r2c5}
 
@@
 
-\item{\revcom  6:}\label{r2c6}
+\item{\revcom  :}\label{r2c6}
 
@@
 
-\item{\revcom 7:}\label{r2c7}
+\item{\revcom :}\label{r2c7}
 
@@
 
-\item{\revcom  8:}\label{r2c8}
+\item{\revcom  :}\label{r2c8}
 
@@
 
-\item{\revcom 9:}\label{r2c9}
+\item{\revcom :}\label{r2c9}
 
@@
 
-\item{\revcom 10:}\label{r2c10}
+\item{\revcom :}\label{r2c10}
 
@@
 
-\item{\revcom 11:}\label{r2c11}
+\item{\revcom :}\label{r2c11}
 
@@
 
-\item{\revcom 12:}\label{r2c12}
+\item{\revcom :}\label{r2c12}
 
@@
 
-\item{\revcom 13:}\label{r2c13}
+\item{\revcom :}\label{r2c13}
 
@@
 
-\item{\revcom  14:}\label{r2c14}
+\item{\revcom :}\label{r2c14}
 
@@
 \item[2.15.1] “closed-form” is sometimes written as “closed form” (e.g., p. 7, p. 18)
+
 \action:
@@
 \item[2.15.2] p. 9, `1s` should be ‘1s’. Similar for 0s.
+
 \action:
@@
 \item[2.15.3] p. 15, my manuscript has -on where to sample-, which is an unusual use of dashes.
+
 \action:
@@
 
-\reply: I think the equation $V_{m,upd}$ needs a j as a part of that subscript.
+I think the equation $V_{m,upd}$ needs a j as a part of that subscript.
 
@@
 
-\reply: Figure 4: Why not show excursion probabilities for, say, myopic selection as well?
+Figure 4: Why not show excursion probabilities for, say, myopic selection as well?
 

### 2020-05-23 main_new.tex
Updates from Jo
Updated introduction

@@
   Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
-% Department of Engineering Cybernetics\\ O. S. Bragstads Plass 2D, 7034 Trondheim\\ Norway\\
+
 \printead{e4}}
@@
 These regions are usually characterized by gradients of measurable environmental variables, e.g., temperature or salinity gradients, on which EPs subsequently can be used...Correlation among samples and multivariate requirements are typical in environmental studies.
+
 \end{abstract}
-
-%\begin{keyword}[class=MSC]
-%\kwd[Primary ]{}
-%\kwd{}
-%\kwd[; secondary ]{}
-%\end{keyword}
 
@@
 effective means of sampling.
-By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use of new technologies including autonomous platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
-In this paper we build upon the potential that lies in using autonomous underwater vehicles (AUVs) to conduct efficient sampling in the ocean \citep{das11b,Das2015,fossuminformation,fossum18b}.
-%Addressing this, the combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based environmental sensing.
-
-There are several obstacles when doing sampling with an AUV in the ocean, and we discuss some of these to motivate our setting, and also the modeling and methodological choices that will be made.
+By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
+This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this the paper.
+
+There are several obstacles when doing sampling with autonomous platforms in the ocean.
 First, full numerical ocean models based on complex differential equations cannot provide accurate results for online operations because the onboard computers lack the computational power required.
-Hence statistical proxy models of the environment must be used on the AUV.
+Hence statistical proxy models of the environment must be used.
 For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest.
-Second, %with limited available information about the state of the ocean,
-with undersampling issues one must usually plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs.
-%, preferable in such a way that there is substantial value in reacting to information obtained from measurements taken in-situ.
-The acquired in-situ information must then be assimilated into the statistical model onboard the AUV, so that it can be used to re-compute the sampling criterion and in this way inform decisions on where to sample next.
+Second, with undersampling issues one must usually plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs.
+The acquired in-situ information must then be assimilated into the statistical model onboard the platform, so that it can be used to re-compute the sampling criterion and in this way inform decisions on where to sample next.
 We use a criterion focusing on the prediction of excursion sets (ESs) for the multivariate spatial processes of interest.
 Third, the number of sampling designs is enormous, creating a trade-off between optimization and computability.
-One must often compromise in both development and practical operations, and we will rely on short-horizon optimization routines for sequential designs among nearest neighbors points in a spatial grid created for the AUV operation.
-
-% After reviews: - This is moved to Section 2.1.
-
-%Traditional data collection at sea has typically been based on static
-%buoys, Lagrangian floats, or ship-based methods, with significant
-%logistical limitations that directly impact coverage and sampling
-%resolution. Modern methods using satellite remote-sensing provide
-%large-scale coverage but have limited resolution, are limited to
-%sensing the surface of the ocean, and are impacted by cloud cover. The
-%advent of robust mobile robotic platforms \citep{Bellingham07} has
-%resulted in significant contributions to environmental monitoring and
-%sampling. In particular, autonomous underwater vehicles (AUVs), have
-%advanced the state of sampling and consequently have made robotics an
-%integral part of ocean observation; our previous work has contributed
-%to this effort \citep{das11b,Das2015,fossum18b,fossuminformation}. Other ¤ %statistical work in the oceanographic domain include \cite{wikle2013modern}
-%focusing on hierarchical statistical models; \cite{sahu2008space},
-%studying spatio-temporal models for sea surface temperature and
-%salinity data; and \cite{mellucci2018oceanic} looking at the
-%statistical prediction of features using an underwater glider.
-
-%\begin{itemize}
-%
-%\item Full numerical ocean models cannot provide accurate results
-%  online if run on robotic sensing platforms, as onboard computers
-%  cannot deliver the computational power required. Hence statistical
-%  proxy models of the environment must be used for learning where to sample.
-%
-%\item With limited available information about the state of the ocean, there is substantial value in reacting to
-%  information obtained from measurements taken in-situ. This
-%  acquired information must be assimilated into statistical models that
-%  can be used to inform decisions on where to sample
-%  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.
-%
-%\item For sampling problems related to environmental sensing, the
-%  number of choices (i.e. locations, trajectories, and candidate
-%  designs) is enormous, creating a
-%  trade-off between optimization (finding the most resource-efficient
-%  design to collect necessary data) and computability (arriving at a
-%  solution in reasonable time). To successfully resolve features, this
-%  trade-off has to be considered in development and practice.
-%
-%\end{itemize}
-
-%Addressing this, the combination of statistical tools and robotic platforms is a
-%natural symbiosis which enables information-based sensing. Central to
-%this is the ability to model spatially-correlated variables and
-%provide formal measures of uncertainty. Our formulation is based on
-%Gaussian Processes (GPs) as they allow efficient implementation 
-%and evaluation in real time onboard a robotic platform.
-
-%Sampling can, in this context, not simply be distributed evenly ---along simple transects or ``lawn-mover" patterns--- but must instead be prioritized to relevant regions to ensure it is cost-effective while providing adequate coverage and resolution of the area of scientific interest.
-
-%While the focus has often been on
-%biological and anthropogenic impact from micro-plastics to
-%pollution, biological oceanographers have focused intently on studying
-%micro-organisms at the base of the human food web. These organisms are
-%critically impacted by the changing dynamics in the upper water-column,
-%especially in coastal zones which are complex and often hard to observe
-%in space and time. By studying the bio-geochemical processes in the
-%upper water-column scientists can measure the impact of change, natural
-%or anthropomorphic, and provide an informed opinion to policy makers to
-%effect changes in preserving the environment. However, the challenge of 
-% The pressure on marine resources is growing and increased accuracy,
-% resolution, and persistent monitoring of the oceans is crucial for
-% long-term sustainable management.
-
-% A
-% sustained focus on prioritized and efficient data collection strategies
-% have therefore started to emerge. The advent of marine robotic
-% platforms, especially
-
-% provided means to execute this prioritization through the capacity of
-% autonomy and data-driven sampling, where data collection in principle
-% can be optimized. These capabilities have made
-
-% of the emerging sensing practice for ocean science, allowing scientists
-% to increase the observational efficiency and resolution beyond what was
-% previously possible. But how should a robotic platform, such as an AUV,
-% effectively prioritize and identify important regions for sampling? The
-% answer to this question relates to
-
-% , and
-% the application domain is clearly an arena where statisticians can
-% contribute.
-% There has recently been some statistical attention in oceanography:
-
-%From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest.
-Each phenomenon can in turn be characterized by a set of process specific conditions expressed through different measures of key environmental variables, such as temperature or salinity.
-One such measure is the gradient, that can be associated with a number of important processes, such as the vertical location of the thermocline and pycnocline, location of upwelling systems, vertical mixing, eddies, fronts, and currents \cite{sverdrup2006}, as well as distribution, growth, and accumulation of biological activity \cite{SatOceanSoci00, Ryan2014}.
-These gradients create boundaries separating the ocean into process specific regions which are of profound interest to both identify and map effectively.
-Quantification of gradient features is therefore a much needed competence in robotic sampling.
-
-%The imperfect understanding of causality coupled with the underlying fundamental uncertainty of the ocean implies that we need to obtain more measurements and use statistical models and methods to determine what and how phenomena form, to increase the quality of predictive statements.
-
-We design AUV sampling strategies for reducing the spatially-integrated
-Bernoulli variance (IBV) in the ES indicator variables. 
-Recent work on ESs and associated excursion probabilities (EPs) connected to spatial statistics include
-\cite{french2013spatio,bolin2015excursion,french2016credible}.
-Our focus is on uncertainty
-characterization and how it is influenced by sampling. In this sense,
-our work is similar to \cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who describe analytical results for
-the IBV for univariate processes. A major contribution of our work is
-to derive closed-form results for this design criteria for situations
-where the underlying model is based on multivariate GPs.
-Another contribution is that we embed this approach in a sequential strategy to produce algorithms, and in applying this on an AUV in the real-world
-scenario of autonomously sampling temperature and salinity variables in a river plume.
+One must often compromise in both development and practical operations, and we will rely on short-horizon optimization routines for sequential designs among nearest neighbors points in a spatial grid (waypoint graph).
+
+The imperfect understanding of causality coupled with the underlying fundamental uncertainty of the ocean implies that we need to obtain more measurements and use statistical models and methods to determine what and how phenomena form, to increase the quality of predictive statements.
+With this in mind, we design AUV sampling strategies for reducing the spatially-integrated Bernoulli variance (IBV) in the ES indicator variables.
+Recent work on ESs and associated excursion probabilities (EPs) connected to spatial statistics include \cite{french2013spatio,bolin2015excursion,french2016credible}.
+Our focus is on uncertainty characterization and how this is influenced by sampling.
+In this sense, our work is similar to \cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who describe analytical results for the IBV for univariate processes.
+A major contribution of our work is to derive closed-form results for this design criteria for situations where the underlying model is based on multivariate GPs.
+Another contribution is that we embed this approach in a sequential strategy to produce sampling algorithms, and apply this to a real-world scenario of autonomously sampling temperature and salinity variables using an autonomous underwater vehicle (AUV), with the intention of understanding the between the ocean and a river plume.
 
@@
 In medicine, doctors do not rely solely on a single symptom but must see several combined effects before making a diagnosis.
-In our context of environmental applications, an example is that of photosynthesis where plants and algae produce oxygen only when exposed to enough light, carbon dioxide and water.
-Our key motivation in this paper is that temperature and salinity, combined, drive several processes in the oceans.
-For instance, salmon lice larvae are known to depend on relatively high temperature and high salinities to survive.
-Salmon lice are parasites to salmon, and the bivariate ES for temperature and salinity is hence important to understand for sustainable salmon fisheries management.
-The salinity and temperature excursions of a river plume can similarly help characterize its oceanographic and bio-geochemical processes \cite{hopkins2013detection,Pinto2018}, and sampling should be aimed at mapping relevant parts to reduce uncertainty in the temperature and salinity ES.
-Unlike what has been done in previous plume exploration studies, we emphasize the links to spatial statistical modeling and targeted multivaraite sampling criteria for sequential sampling.
-Addressing this, the combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based environmental sensing.
-
-The remainder of this paper is organized as follows: Section
-\ref{sec:bg} provides select background on ocean sampling; Section
-\ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV;
-Section \ref{sec:GP_EP} sets up the required statistical modeling
-assumptions of the bivariate GPs for temperature and salinity;
-Section \ref{sec:sur} builds on these assumptions when deriving the
-analytical derivation of the design criteria; Section
-\ref{sec:heuristics} extends these results and present heuristics for
-adaptive sampling; Section \ref{sec:simulations}
-discusses properties of the methods in simulation studies. Section
-\ref{sec:case_study} demonstrates the methodology used in field work
-characterizing a river plume; and Section \ref{sec:concl_disc}
-contains a summary and a discussion of future work.
+In our context of environmental sampling applications, the salinity and temperature excursions of a river plume can similarly help characterize the underlying oceanographic and bio-geochemical processes \cite{hopkins2013detection,Pinto2018}, where sampling should be aimed at mapping relevant parts of the domain to reduce uncertainty in the temperature and salinity ES.
+Unlike what has been done in previous plume exploration studies, we emphasize the links to spatial statistical modeling and targeted multivariate sampling criteria for sequential sampling.
+
+The remainder of this paper is organized as follows: Section \ref{sec:bg} provides select background on ocean sampling; Section \ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV; Section \ref{sec:GP_EP} sets up the required statistical modeling assumptions of the bivariate GPs for temperature and salinity; Section \ref{sec:sur} builds on these assumptions when deriving the analytical derivation of the design criteria; Section \ref{sec:heuristics} extends these results and present heuristics for adaptive sampling; Section \ref{sec:simulations} discusses properties of the methods in simulation studies.
+Section \ref{sec:case_study} demonstrates the methodology used in field work characterizing a river plume; and Section \ref{sec:concl_disc} contains a summary and a discussion of future work.
 
 \section{Background on Ocean Processes and Observations}\label{sec:bg}
-
-%\subsection{Observing Processes in the Upper Ocean}
-
-% The processes occurring in the upper water-column and the photic zone
-% (upper sunlit zone) are of vital importance for long-term sustainable
-% management of the greater marine ecosystem, climate, and human health.
-% With the onset of the warming oceans and increased anthropogenic
-% pressure, it is critical to improve our understanding of the associated
-% ecosystem response.
-% owing to current observation practices.
 
@@
 water-column is knowledge of the biophysical interaction driven by
-an agglomeration of physical forcings %(wind, topography, bathymetry,
-%tidal influences, and local geography, such as fresh water inflows) 
-and incipient micro-biology driven by planktonic and coastal anthropogenic
-input, such as pollution and agricultural runoff transported into the ocean by rivers and streams. These often result in a range of ecosystem-related phenomena such as blooms%, anoxic zones
-and plumes. %, all of which have significant direct and indirect effects on society.
-
-% On observation and why robotics
-Traditional data collection to understand such phenomena at sea has typically been based on static
-buoys, floats, or ship-based methods, with significant
-logistical limitations that directly impact coverage and sampling
-resolution. Modern methods using satellite remote-sensing provide
-large-scale coverage but have limited resolution, are limited to
-sensing the surface, and are impacted by cloud cover. Numerical ocean models similarly find it challenging to provide detail at
-smaller scale \citep{Lermusiaux:2006}. 
-The advent of mobile robotic platforms \citep{Bellingham07} has
-resulted in significant contributions to environmental monitoring and
-sampling. In particular, AUVs have
-advanced the state of in-situ sampling, and made robotics an
-integral part of ocean observation, filling parts of the undersampling gap (See Fig.
+an agglomeration of physical forcings (e.g.
+wind, topography, bathymetry, tidal influences, etc.) and incipient micro-biology driven by planktonic and coastal anthropogenic input, such as pollution and agricultural runoff transported into the ocean by rivers and streams.
+These often result in a range of ecosystem-related phenomena such as blooms and plumes, with direct and indirect effects on society.
+
+Traditional data collection to understand such phenomena at sea has typically been based on static buoys, floats, or ship-based methods, with significant logistical limitations that directly impact coverage and sampling resolution.
+Modern methods using satellite remote-sensing provide large-scale coverage but have limited resolution, are limited to sensing the surface, and are impacted by cloud cover.
+Numerical ocean models similarly find it challenging to provide detail at smaller scale \citep{Lermusiaux:2006}.
+The advent of mobile robotic platforms \citep{Bellingham07} has resulted in significant contributions to environmental monitoring and sampling.
+In particular, AUVs have advanced the state of in-situ sampling, and made robotics an integral part of ocean observation, filling parts of the undersampling gap (See Fig.
 \ref{fig:envir1}).
-%The previously listed challenges related to undersampling have also been driving ocean sensing practices towards more autonomous, networked, and distributed use of robotic assets (See Fig.
-\ref{fig:envir1}), where observation is guided by assimilation of recent measurements and statistical models.
 
@@
 
-The AUVs considered here
-operate at $1$-$3$ m/s in the upper water column.
-The in-water operation time capacity depends on survey speed, payload sensors and navigation.
+The AUVs considered here operate at $1$-$3$ m/s in the upper water column.
+The in-water operation time capacity depends on survey speed, payload sensors and navigation, see Fig.
+\ref{fig:lauv}.
 AUVs have limited computational capacity, and they are usually done on a single-board computer (SBC), like the Raspberry Pi or an multicore GPU NVIDIA Jetson TX1.
 
-%By deploying AUVs together with other resources, one can augment information from synthetic ocean models, remote-sensing data, or fixed-location sensors on buoys. 
-Surveys with AUVs are usually limited to observations
-along fixed transects that are pre-programmed by the human operator.
+Surveys with AUVs are usually limited to observations along fixed transects that are pre-programmed by the human operator.
 This occurs on a spatial grid, called a waypoint graph.
 A more effective approach is to instead use onboard algorithms to continuously evaluate, update, and refine future sampling locations (sense-plan-act cycle in Fig.
-\ref{fig:envir1}), making the information
-gathering \emph{adaptive} \citep{das11b,fossum18b,fossuminformation}. In doing so, the space of sampling opportunities is still limited by the waypoint graph, but the AUV has flexibility to 
-modify its path at each waypoint based on what has been measured onboard
-\citep{py10,Rajan12,Rajan12b}.
-
-The application that is of focus in this work is towards spatial characterization of a frontal system generated by river
-plumes. Fig. \ref{fig:nidelven} shows the survey area (Nidelva, Norway) where cold freshwater enters
-from the river, creating a strong gradient in both temperature and
-salinity. Because of the local topography and the Coreolis force
-%\citep{coriolis1835memoire} 
-the cold fresh water tends to flow near
-land to the east, %. Depending on the river discharge, tidal effects,
-%wind, and temperature differences, this boundary often gets
-%distorted. Initial knowledge about the location and evolution of these
-but the locations of these features are highly uncertain, making deterministic planning challenging.
-
-%River plumes belong to a class of ocean processes that are local and
-%act on smaller \emph{sub-mesoscale} extent (from $5 m^2$ -- $10 km^2$) where
-%lateral spatial variability tends to dominate. 
-%At larger \emph{mesoscale} ($>50 km^2$), both three-dimensional space and time
-%dynamics are important and can shift substantially, while in the case
-%At this scale of river plumes one can further often limit scope to covering only lateral near surface spatial elements, and the case appears very amenable for AUVs usage. 
-%However, the methodological framework can be extended to higher dimensional situations in a similar manner.
-
-%\begin{figure}[!h] 
-%\centering 
-%\includegraphics[width=0.85\textwidth]{Figures/pictures/c-updated.pdf}
-%\caption{Frontal patterns off of the Nidelva river, Trondheim, Norway.}
-%\label{fig:nidelven}
-%\end{figure}
-
-%\begin{figure}[!h]
-%\centering
- % \subfigure[River front - Columbia River, Astoria, Oregon, US.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/a.png}\label{fig:river1}}
- % \hfill
- %\subfigure[Tidal front - Korsfjorden, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/d.png}\label{fig:river2}}
- % \hfill
- % \subfigure[River plume - Rio de la Plata, Buenos Aires, Argentina.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/b.png}\label{fig:river3}}
- % \hfill
- %\subfigure[Frontal patterns - Nidelven, Trondheim, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/c.png}\label{fig:river4}}
- %\caption{Two examples of frontal features associated with
-%   river-ocean interaction. % (\ref{fig:river1}) The river plume front of
-   % the Columbia River meeting the Pacific Ocean.
-%   (\ref{fig:river2})  Superimposed images (over 10 minutes) of a surface tidal front in
-%   Korsfjorden, Norway. % (\ref{fig:river3}) The river plume of the Rio de
-   % la Plata, taken by European Space Agency (ESA)'s Envisat platform and
-   % MEdium Resolution Imaging Spectrometer (MERIS) sensor.
-%   (\ref{fig:river4}) Aerial image of the mouth of the Nidelva river in
-%   Trondheim, Norway where fresh water meets the salty fjord and creates
-%   frontal patterns.} % Image courtesy: (\ref{fig:river1})
-   % \cite{zamon2014marine}, (\ref{fig:river3}) ESA, CC BY-SA 3.0 IGO.}
-% \label{fig:river_fronts}
-%\end{figure}
-
-%The use of \emph{robotic sampling} is motivated by the fundamental challenges of ocean observation, where a limited set of resources, a highly dynamic ocean, and the associated
-
-%This challenge can be addressed by employing more elaborate and adaptive sampling strategies that can capitalize on both prior and current observations in order to locate and map these areas.
-
-%\subsection{Autonomous Vehicles}
-
-%To improve the state of sampling, modern tools and methods, including the use of autonomous platforms, oceanographic models and satellite remote sensing, needs to be combined with traditional data acquired by surface vessels or buoys (See Fig.
-\ref{fig:envir}).
-However, without adequate understanding of the theoretical underpinnings of how, when, and where to gather data, these tools and methods are insufficient in the vast and harsh oceans.
-
-% retaining an advantageous strategy for information recovery, online during execution.
-%The pressure on marine resources is growing and increased accuracy, resolution, and persistent monitoring of the oceans is crucial for long-term sustainable management.
-The impact of this research provides cost effective tools, techniques, and processes for doing ocean based measurements using robotic platforms.
-%adaptive design of experiments using robotic assets identifying and prioritizing relevant sampling locations on an information-theoretic basis.
-Spatial statistics naturally enters here through the ability to both model spatially correlated parameters and provide formal measures of uncertainty, on which this basis can be formed.
-%for oceanic sensing applications, where sensors are sparsely distributed and capitalizing on all available information is
-
-%This imperfect understanding of causality coupled with uncertainty implies that we need to obtain more measurements and use statistical models and methods to determine what and how phenomena form, to increase the quality of predictive statements.
-To improve the state of sampling, modern tools and methods, including the use of autonomous platforms, oceanographic models and satellite remote sensing, augment the more traditional data acquired by surface vessels or buoys (See Fig.
-\ref{fig:envir}).
-However, without adequate understanding of the theoretical underpinnings of how, when, and where to gather data, these tools and methods are insufficient in the vast and harsh oceans.
-
-%There are several oceanographic data sources that can be used together with statistical tools to improve data collection in ocean science.
-In the following section we briefly discuss buoy data, satellite data, ocean models data, and AUV data.
-
-%Buoy data provide very accurate information of oceanographic variables, but in most situations they are local, giving information only at one (north, east) coordinate, possibly with opportunities for conducting measurements at different depths.
-Gliders and other surface vessels are a kind of floating buoys that drift and can measure variables at many locations, but still with limited spatial coverage.
-
-%Satellite data are important for the mapping of oceanographic variables.
-They can also be indicative of variables such as temperature and salinity which we look at here, especially if the data are calibrated to for instance buoy data from the same spatial domain.
-However, the resolution of satellite data is relatively large, say $1 \times 1$ km $^2$ grid cells, and it only provides accurate information at the sea surface.
-Moreover, one cannot get useful satellite information on a cloudy day.
-
-%What is often done in practice is to run ocean models based on the complex differential equations governing oceanographic phenomena, where also satellite data can be assimilated and used as input, and possibly playing with different forcing mechanisms to capture some of the uncertainty of the models.
-Initial conditions, and can be used t the goal is to guide the sampling towards regions these regions, where we can reduce the uncertainty in the ES.
-These models provide insight that can be understood from practitioners, but they tend to be biased, and must also be calibrated to buoy data in one way or the other.
-
-%to provide an effective specification of regions of phenomenological interest resolving the boundaries and the position of phenomena, on which evaluation of future sampling can be constructed.
-
-%; some of these phenomena are illustrated in Fig.
-\ref{fig:envir2}.
-Figure \ref{fig:river_fronts} illustrates the %phenomena of interest in this paper, which is river plumes and their interaction with the ocean.
-
-%Feature driven sampling
-%processes acting across a wide range of spatio-temporal scales makes prioritizing sampling efforts necessary.
-%Autonomous robotic platforms, can address these issues by providing the ability to focus sampling efforts to %high-interest regions (``hotspots").
-
-% Determining paths for mobile robotic sensors in order to maximize the
-% information gained about an environment is formally known as informative
-% path planning. This, and related problems such as the orienteering
-% problem \citep{Golden87}, has been studied in the context of graphs,
-% where the potential measurement are assigned to nodes on which
-% evaluation of different routes can be conducted.
-%In this context, a fundamental result from \cite{nemhauser1978analysis} proves that a simple greedy algorithm (iteratively selecting the location which most increases the utility) can achieve a near-optimal solution if the problem can be shown to be \emph{submodular}\footnote{an intuitive diminishing returns property, where the informative value of adding sensors decrease with the number of sensors added.}.
-Building on this result, \cite{chekuri2005recursive} explored this in the setting of a graph using a recursive greedy algorithm, providing near-optimal solutions depending on the planning horizon and graph resolution.
-
-Adaptive in-situ AUV sampling of an evolving frontal feature has been explored in
-\cite{fronts11,Zhang2012,Pinto2018,costa19}. These approaches
-typically use a reactive-adaptive scheme, whereby exploration does not
-rely on a statistical model of the environment, but rather adapts
-based on closing the sensing and actuation loop. Myopic sampling, i.e. stage-wise selection of
-the path (on the waypoint graph), has been used for surveys
-\citep{singh2009efficient,Binney2013} that focus largely on reducing
-predictive variance or entropy. These criteria are widely adopted in the statistics literature on spatio-temporal design as well \cite{bueso1998state,zidek2019monitoring}, but variance and
-entropy reduction are independent of the actual data realizations
-under the assumptions of GP models, so limited decision flexibility is created for sequential designs. The use of data-driven adaptive
-criteria was introduced to include more targeted sampling of regions
-of scientific interest in \cite{Low2009} and \cite{fossuminformation}.
-In this paper, with a focus on mapping the river plume by rewarding the
-designs that improve the classification of ES in temperature and salinity.
+\ref{fig:envir1}), making the information gathering \emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
+In doing so, the space of sampling opportunities is still limited by the waypoint graph, but the AUV has flexibility to  modify its path at each waypoint based on what has been measured onboard \citep{py10,Rajan12,Rajan12b}.
+
+The application that is of focus in this work is towards spatial characterization of a frontal system generated by river plumes.
+Fig.
+\ref{fig:nidelven} shows the survey area (Nidelva, Norway) where cold freshwater enters from the river, creating a strong gradient in both temperature and salinity.
+Because of the local topography and the Coreolis force \citep{coriolis1835memoire} the cold fresh water tends to flow near
+land to the east.
+Depending on the river discharge, tidal effects, wind, and temperature differences, this boundary often gets distorted.
+Initial knowledge about the location and evolution of these but the locations of these features are highly uncertain, making deterministic planning challenging.
+
+Adaptive in-situ AUV sampling of an evolving frontal feature has been explored in \cite{fronts11,Zhang2012,Pinto2018,costa19}.
+These approaches typically use a reactive-adaptive scheme, whereby exploration does not rely on a statistical model of the environment, but rather adapts based on closing the sensing and actuation loop.
+Myopic sampling, i.e.
+stage-wise selection of the path (on the waypoint graph), has been used for surveys \citep{singh2009efficient,Binney2013} that focus largely on reducing predictive variance or entropy.
+These criteria are widely adopted in the statistics literature on spatio-temporal design as well \cite{bueso1998state,zidek2019monitoring}, but variance and entropy reduction are independent of the actual data realizations under the assumptions of GP models, so limited decision flexibility is created for sequential designs.
+The use of data-driven adaptive criteria was introduced to include more targeted sampling of regions of scientific interest in \cite{Low2009} and \cite{fossuminformation}.
+In this paper, with a focus on mapping the river plume by rewarding the designs that improve the classification of ES in temperature and salinity.
 
@@
 %Recent work on ESs and EPs connected to spatial statistics include
-\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
+%\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
 %Our focus is on ES as defined in Eq. (\ref{ES}) and the uncertainty
@@
 
+% === Not used
+% After reviews: - This is moved to Section 2.1.
+
+%Traditional data collection at sea has typically been based on static
+%buoys, Lagrangian floats, or ship-based methods, with significant
+%logistical limitations that directly impact coverage and sampling
+%resolution. Modern methods using satellite remote-sensing provide
+%large-scale coverage but have limited resolution, are limited to
+%sensing the surface of the ocean, and are impacted by cloud cover. The
+%advent of robust mobile robotic platforms \citep{Bellingham07} has
+%resulted in significant contributions to environmental monitoring and
+%sampling. In particular, autonomous underwater vehicles (AUVs), have
+%advanced the state of sampling and consequently have made robotics an
+%integral part of ocean observation; our previous work has contributed
+%to this effort \citep{das11b,Das2015,fossum18b,fossuminformation}. Other ¤ %statistical work in the oceanographic domain include \cite{wikle2013modern}
+%focusing on hierarchical statistical models; \cite{sahu2008space},
+%studying spatio-temporal models for sea surface temperature and
+%salinity data; and \cite{mellucci2018oceanic} looking at the
+%statistical prediction of features using an underwater glider.
+
+%\begin{itemize}
+%
+%\item Full numerical ocean models cannot provide accurate results
+%  online if run on robotic sensing platforms, as onboard computers
+%  cannot deliver the computational power required. Hence statistical
+%  proxy models of the environment must be used for learning where to sample.
+%
+%\item With limited available information about the state of the ocean, there is substantial value in reacting to
+%  information obtained from measurements taken in-situ. This
+%  acquired information must be assimilated into statistical models that
+%  can be used to inform decisions on where to sample
+%  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.
+%
+%\item For sampling problems related to environmental sensing, the
+%  number of choices (i.e. locations, trajectories, and candidate
+%  designs) is enormous, creating a
+%  trade-off between optimization (finding the most resource-efficient
+%  design to collect necessary data) and computability (arriving at a
+%  solution in reasonable time). To successfully resolve features, this
+%  trade-off has to be considered in development and practice.
+%
+%\end{itemize}
+
+%Addressing this, the combination of statistical tools and robotic platforms is a
+%natural symbiosis which enables information-based sensing. Central to
+%this is the ability to model spatially-correlated variables and
+%provide formal measures of uncertainty. Our formulation is based on
+%Gaussian Processes (GPs) as they allow efficient implementation 
+%and evaluation in real time onboard a robotic platform.
+
+%Sampling can, in this context, not simply be distributed evenly ---along simple transects or ``lawn-mover" patterns--- but must instead be prioritized to relevant regions to ensure it is cost-effective while providing adequate coverage and resolution of the area of scientific interest.
+
+%While the focus has often been on
+%biological and anthropogenic impact from micro-plastics to
+%pollution, biological oceanographers have focused intently on studying
+%micro-organisms at the base of the human food web. These organisms are
+%critically impacted by the changing dynamics in the upper water-column,
+%especially in coastal zones which are complex and often hard to observe
+%in space and time. By studying the bio-geochemical processes in the
+%upper water-column scientists can measure the impact of change, natural
+%or anthropomorphic, and provide an informed opinion to policy makers to
+%effect changes in preserving the environment. However, the challenge of 
+% The pressure on marine resources is growing and increased accuracy,
+% resolution, and persistent monitoring of the oceans is crucial for
+% long-term sustainable management.
+
+% A
+% sustained focus on prioritized and efficient data collection strategies
+% have therefore started to emerge. The advent of marine robotic
+% platforms, especially
+
+% provided means to execute this prioritization through the capacity of
+% autonomy and data-driven sampling, where data collection in principle
+% can be optimized. These capabilities have made
+
+% of the emerging sensing practice for ocean science, allowing scientists
+% to increase the observational efficiency and resolution beyond what was
+% previously possible. But how should a robotic platform, such as an AUV,
+% effectively prioritize and identify important regions for sampling? The
+% answer to this question relates to
+
+% , and
+% the application domain is clearly an arena where statisticians can
+% contribute.
+% There has recently been some statistical attention in oceanography:
+
+%From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest.
+Each phenomenon can in turn be characterized by a set of process specific conditions expressed through different measures of key environmental variables, such as temperature or salinity.
+One such measure is the gradient, that can be associated with a number of important processes, such as the vertical location of the thermocline and pycnocline, location of upwelling systems, vertical mixing, eddies, fronts, and currents \cite{sverdrup2006}, as well as distribution, growth, and accumulation of biological activity \cite{SatOceanSoci00, Ryan2014}.
+These gradients create boundaries separating the ocean into process specific regions which are of profound interest to both identify and map effectively.
+Quantification of gradient features is therefore a much needed competence in robotic sampling.
 \end{document}

### 2020-05-24 ans_to_reviewers/answers-to-reviewers.tex
Updated introduction
Updated with macros

@@
 \def\revcom{\textbf{Reviewer comment}}
-\def\aecom{\textbf{Associate Editor comment}}
+\def\aecom{\textbf{Associate Editor comment }}
 \def\reply{\textbf{Reply}}
@@
 \vspace{5mm}
-\noindent \aecom 1:
+\noindent \aecom \textbf{1}:
 
@@
 \vspace{5mm}
-\noindent \aecom 2:
+\noindent \aecom \textbf{2}:
 
@@
 \vspace{5mm}
-\noindent \aecom 3:
+\noindent \aecom \textbf{3}:
 
@@
 \vspace{5mm}
-\noindent \aecom 4:
+\noindent \aecom \textbf{4}:
 
@@
 \vspace{5mm}
-\noindent \aecom 5:
+\noindent \aecom \textbf{5}:
 
@@
 \vspace{5mm}
-\noindent \aecom 6:
+\noindent \aecom \textbf{6}:
 
@@
 \vspace{5mm}
-\noindent \aecom 7:
+\noindent \aecom \textbf{7}:
 
@@
 What are the computational limits of these machines?\par
-\action: Our actions
+\action: We have added more information about AUVs and their computational capabilities in Section 2.1 (Background on Ocean Processes and Observations) to address this. 
+\begin{quote}
+The AUVs, see Fig.
+9, considered here operate at $1$-$3$ m/s in the upper water column, where they are free to move in six degrees of freedom (6 DOF).
+The in-water operation time capacity depends on survey speed, payload sensors and navigation.
+Generally this amounts to 8-24 hrs.
+operation for smaller crafts, and 24-48 hrs.
+for larger crafts.
+AUVs have limited computational capacity, as they typically use single-board computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM) to do computation.
+\end{quote}
 \vspace{1em}
@@
 \item[2.1.2] How long do the AUVs typically explore in one go? Is there a minimum/maximum time?\par 
-\action: Max speed is limit for coverage. Time constraints for operations related to costs of man-hours in operation. Battery capacity is not a challenge for this AUV. 
+\action: Maximum speed and payload are limiting for coverage.
+Generally this amounts to 8-24 hrs.
+operation for smaller crafts, and 24-48 hrs.
+for larger crafts.
+Time constraints for operations are typically related to costs of ship time and man-hours during operation.
+Battery capacity is not a challenge for the AUV described in the paper.
 \vspace{1em}

### 2020-05-24 main_new.tex
Updated with macros

@@
 
+% New Marcos.
+\usepackage{amssymb}
+\usepackage{hyperref}
+\usepackage{cleveref}
+\input{macros}
+
 \begin{document}
@@
   Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
-
 \printead{e4}}
@@
 
-The AUVs considered here operate at $1$-$3$ m/s in the upper water column.
-The in-water operation time capacity depends on survey speed, payload sensors and navigation, see Fig.
-\ref{fig:lauv}.
-AUVs have limited computational capacity, and they are usually done on a single-board computer (SBC), like the Raspberry Pi or an multicore GPU NVIDIA Jetson TX1.
-
-Surveys with AUVs are usually limited to observations along fixed transects that are pre-programmed by the human operator.
-This occurs on a spatial grid, called a waypoint graph.
-A more effective approach is to instead use onboard algorithms to continuously evaluate, update, and refine future sampling locations (sense-plan-act cycle in Fig.
+The AUVs, see Fig.
+\ref{fig:lauv}, considered here operate at $1$-$3$ m/s in the upper water column, where they are free to move in six degrees of freedom (6 DOF).
+The in-water operation time capacity depends on survey speed, payload sensors and navigation.
+Generally this amounts to 8-24 hrs.
+operation for smaller crafts, and 24-48 hrs.
+for larger crafts.
+AUVs have limited computational capacity, as they typically use single-board computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM) to do computation.
+
+Surveys with AUVs are usually limited to observations along fixed transects that are pre-scripted in mission plans created manually by a human operator.
+The mission can be specified operating on a scale of hundreds of meters to tens of kilometers depending on the scientific context.
+Faced with limited coverage capacity, a more effective approach is to instead use onboard algorithms to continuously evaluate, update, and refine future sampling locations (sense-plan-act cycle in Fig.
 \ref{fig:envir1}), making the information gathering \emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
+This usually occurs on a spatial grid, called a waypoint graph.
 In doing so, the space of sampling opportunities is still limited by the waypoint graph, but the AUV has flexibility to modify its path at each waypoint based on what has been measured onboard \citep{py10,Rajan12,Rajan12b}.
@@
 
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.4\textwidth]{Data-driven.png}
-%\label{fig:auv_alone}
-%\caption{The concept of data-driven sampling.}
-%\end{figure}
-
-%The goal is effectively map the extension of the river plume given this
-
-%, we can formulate this question as follows: the ocean region  defined by the boundary between water bodies, separated by a characteristic temperature gradient that is smaller than $T=5$\textdegree C and a salinity concentration lower than $S=30$ [g/kg].
-A goal is then to improve our predictive capabilities to answer this question better.
-%For example,  (examples are shown in Fig. \ref{fig:river_fronts}). However,
-
-% In doing so, the aim is to increase the knowledge about the uncertain ocean environment through adaptive sampling
-
-\section{Excursion Sets and Integrated Bernoulli Variance}
+\section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
 \label{sec:ESEP}
 
+\subsection{Excursion sets and Integrated Bernoulli variance}
+\label{ES}
+
+Our objective is to leverage statistical tools onboard autonomous vehicles to characterize a river
+plume, focusing on spatial separation of cold freshwater from the
+river and warmer saline waters of the fjord. The characterization of uncertainties on the ES is a useful
+goal of this experimental sampling.
+
+In order to formalize the problem in mathematical terms,
+let us consider a multivariate (i.e., vector-valued) Gaussian random field $(\gp[\x])_{\x \in D}$, where $D$ is an arbitrary set (typically in geostatistical studies, a subset of interest of the ambient Euclidean space or of a sphere therein) and the dimensionality of the output $\no \geq 1$ is arbitrary.
+Assume furthermore that the set of interest is characterized in terms of a subset of values of the variables, say $\T \subset \R^{\no}$, and denote the corresponding set of index points by
+$$
+\es=\gp^{-1}(\T)=\{\x \in D: \gp[\x] \in \T\}.
+$$
+Our goal here is to develop approaches to quantify and improve the characterization of uncertainties on such sets in the context of multivariate random fields. 
+While some aspects of the developed
+approaches do not call for a specific form of $\T$, we will often stick for
+simplicity to the case of orthants ($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where $t_1,\dots, t_{\no} \in \R$) as this will allow efficiently calculating several key quantities.
+Note that changing some $\leq$ inequalities to $\geq$ ones would lead to immediate adaptations.
+We will keep the abstract notation $\T$ and it shall be clear from the context when specific properties following the orthant assumption come into play.
+Note that assuming that $\gp$ has continuous trajectories (almost surely) and $T$ to be closed, then $\es$ can be cast as a Random Closed Set \cite{Molchanov2005}.
+In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
+
+For our application, we let $t_1$ be a threshold for temperature and $t_2$ a threshold for
+salinity. The ES is then defined by $\mbox{ES} = \{\bx \in \mathcal{M} | \xi_T(\bx) \leq t_T,\xi_S(\bx) \leq t_S\}$.
+This ES holds random indicator variables while the associated EPs are $p(\bx) = P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S)$, $\bx \in \mathcal{M}$.
+If the EPs are
+close to $1$ or $0$ at a given location, it is easy to classify water mass at this point as either river or ocean. It is, however,
+challenging if the probability is close to $0.5$.
+
+Relate to ES and refer to 
 \cite{adler2000excursion,adler2009random}.
 
-Our objective is to leverage statistical tools to characterize a river
-plume, focusing on spatial separation of cold freshwater from the
-river and warmer saline waters of the fjord. ES and EPs are useful
-starting points to measure the ability to classify water masses. The
-goal of experimental sampling is to improve this
-characterization, as quantified via changes in the EPs.
-
-We consider a lateral domain near the sea surface, with locations
-$\bx \in \mathcal{M} \subset \mathbf{R^2}$. The two variables of
-interest, temperature and salinity, are defined as random processes:
-$\xi_T(\bx)$ denotes the temperature (in $^o C$) and $\xi_S(\bx)$
-denotes the salinity (in $g/kg$) at location $\bx$. The generalization
-to more than two variables is given in Section \ref{gen_k}, and the
-adaptive sampling strategies are described in Section
-\ref{sec:heuristics}.
-
-We let $t_T$ be a threshold for temperature and $t_S$ a threshold for
-salinity. The joint ES is then defined by:
-\begin{equation}\label{ES}
-     \mbox{ES} = \{\bx \in \mathcal{M} | \xi_T(\bx) \leq t_T,\xi_S(\bx) \leq t_S\}.
+Let us in particular consider a locally finite (Borel) measure $\mes$ on $D$ and investigate the probability distribution of $\mes(\es)$ through its moments.
+
+\medskip
+
+%That $\mu(\es)$ is a well-defined random variable follows from
+\begin{propo}
+	\label{propo1}
+	%Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
+	For a measurable random field $\gp$ and a locally finite measure $\mes$ on $D$, $\mes(\es)$ is a random variable and for any $r\geq 1$, 
+	\begin{equation*}
+	\begin{split}
+	\mathbb{E}[\mes(\es)^r]
+	&=\int_{D^{r}} \jointExcuProb 
+	\productMeasure
+	,
+	\end{split}
+	\end{equation*}
+
+	where the product measure is denoted as
+	$\nu^{\otimes}:=\bigotimes_{i=1}^r \nu$. 
+	%Note that we will use boldface to denote concatenated variables. % TO RELOCATE
+	Here $\gp$ is defined on $D$, and for
+	$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in D^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
+	\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.
+	\medskip
+
+	In the particular case where $\gp$ is a multivariate Gaussian random field 
+	%and denoting the mean and covariance matrix of $\gp[\bm{u}]$ by $\meanUU\in\mathbb{R}^{p\times r}$ and $\covUU$, respectively, 
+	we have 
+	\begin{align*}
+	\jointExcuProb = \mathcal{N}_{\no r}(T^r; \meanUU, \covUU),
+	\end{align*}
+	where $\mathcal{N}_{\no r}(\cdot ; \meanUU, \covUU)$ is the Gaussian measure on $\mathbb{R}^{\no r}$ with mean $\meanUU$ and covariance matrix $\covUU$, respectively defined blockwise by  
+	%where both are defined blockwise by 
+	\begin{align*}
+	\meanUU&=\begin{pmatrix}\mu(u^{(1)})\\ \vdots\\ \mu(u^{(r)})\end{pmatrix}
+	\in \mathbb{R}^{\no r}, \\
+	\text{and } \covUU &= \begin{pmatrix}
+	\cov(\gp[u^{(1)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(1)}],
+	\gp[u^{(r)}])\\
+	\vdots & & \vdots\\
+	\cov(\gp[u^{(r)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(r)}],
+	\gp[u^{(r)}])\\
+	\end{pmatrix}\in \mathbb{R}^{pr\times pr},
+	\end{align*}
+	each of the $r\times r$ blocks of the latter matrix being itself a (cross-)covariance matrix of dimension $\no \times \no$. 
+	%dimensional Gaussian random vector. 
+	Assuming further that $\covUU$ is non-singular, the probability of interest can be formulated in terms of the $\no r$-dimensional Gaussian probability density function 
+	$\varphi_{\no  r}(\cdot;~\meanUU, \covUU)$ as 
+	%$\varphi_{\no \times r}$ as
+	%$\bm{u}\in D^r$ such that $\gp[\bm{u}]$ be non-degenerate, the integrand can be expressed in terms of the $r \times \no$-dimensional Gaussian distribution, namely 
+	%cumulative distribution function, namely 
+	\begin{equation*}
+	\begin{split}
+	&\jointExcuProb
+	=
+	\int_{T^r} \varphi_{\no  r}\left(\bm{v};
+	~\meanUU, \covUU\right) 
+	\mathrm{d}\bm{v},
+	\end{split}
+	\end{equation*}
+	In the particular orthant case with $\T=(-\infty, t_1] \times \dots \times (-\infty, t_{r}]$,
+	the latter probability directly writes in terms of the multivariate Gaussian
+	cumulative distribution, % with mean $\meanUU$ and covariance matrix $\covUU$,
+	this time by the way without requiring $\covUU$ %the latter 
+	to be non-singular: 
+	\begin{equation*}
+	\begin{split}
+	\jointExcuProb 
+	&=
+	\varPhi_{\no r}\left(\bm{t};~\meanUU, \covUU\right),
+	\end{split}
+	\end{equation*}
+	where we have used the notations 
+	$t=(t_1,\dots,t_{\no}%,..., t_1, ..., t_p
+	)\in\mathbb{R}^{\no}$, $1_{r}=(1,\dots,1)\in \R^{r}$, and 
+	$\bm{t}=1_{r}\otimes \bm{t}=(t_1,\dots,t_{\no},\dots,t_1,\dots,t_{\no}) 
+	\in \R^{\no r}$.
+\end{propo}
+
+The multivariate cumulative probabilities in two and four dimensions
+are computationally efficient using methods developed by
+\cite{genz2009computation}.
+
+Proposition~\ref{propo1} enables in turn calculating centred moments of $\mes(\es)$. 
+The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ is probably the most relevant of them for the considered purposes. We have
+\begin{equation*}
+\begin{split}
+\emv
+&=\int_{D^2} \mathbb{P}\left(
+\gp[u]\in T, \gp[v]\in T \right) 
+d\mes^{\otimes}(u, v)\\
+&-\left( \int_{D} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
+\end{split}
+\end{equation*} 
+which boils down in the excursion/sojourn case where $\T=(-\infty, t_1] \times
+\dots \times (-\infty, t_{\no}]$ to 
+\begin{equation*}
+\begin{split}
+\emv
+%\operatorname{Var}[\mes(\es)]
+&=\int_{D^2} 
+\varPhi_{2\no}
+\left(
+(\bt, \bt); \mu((u,v)), 
+K((u,v),(u,v))
+\right) 
+\
+\mathrm{d}\mes^{\otimes} %\mes 
+%\productMeasure 
+(u,v)\\
+&-\left( \int_{D} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
+\end{split}
+\end{equation*}
+%
+Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $D^2$.
+In contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves solely and integral on $D$ and can be expanded in our present settings as follows
+\begin{equation*}
+\begin{split}
+\operatorname{IBV} %(\es) %;\mes)
+&=\int_{D}
+\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
+d\mes(u) \\
+&=\int_{D}
+\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) 
+-\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
+\mathrm{d}\mes(u). 
+\end{split}
+\end{equation*}
+
+\begin{propo}
+	\label{propo2}
+	Let $p, q, h \geq 1$, $a \in \R^p$, $B \in \R^{p\times q}$,
+	and $\covN, \covV$ be two covariance matrices in
+	$\R^{p\times p}$ and $\R^{q\times q}$, respectively.
+	Then, for $V \sim \mathcal{N}_{q}(0_q, \covV)$,
+	\begin{equation*}
+	\mathbb{E}\left[ \varPhi_{p}\left( a + BV; \covN \right)^h \right]
+	=
+	\varPhi_{ph}
+	\left(
+	\bm{a}
+	;~
+	\bm{\Sigma}
+	\right),
+	\end{equation*}
+	where the vector $\bm{a} \in \R^{p h}$ is defined as 
+	$\bm{a} := 1_h\otimes a = 
+	\left(a, \dots , a
+	\right)'$
+	and the $p h\times p h$ covariance matrix is given by
+	$\bm{\Sigma} := 
+	1_h 1_h'\otimes B\covV B' + I_h\otimes \covN$.
+\end{propo}
+
+	In blockwise representation, $\bm{\Sigma}$ can be expressed as follows: 
+	\begin{align*}
+	% \bm{\Sigma}&=
+	\begin{pmatrix}
+	\covN & &\\
+	& \ddots &\\
+	&   & \covN
+	\end{pmatrix}
+	+
+	\begin{pmatrix}
+	B\covV B' & \dots & B\covV  B'\\
+	\vdots & & \vdots\\
+	B\covV B' & \dots & B\covV B'\\
+	\end{pmatrix}
+	\end{align*}
+
+In the Appendix we also provide a more general derivation which holds for more general criterian than IBV, in partical for volumes of sets, meaning correlations between many locations meaning sums of indicator variables in Prop 1.
+
+\subsection{Sampling Designs and Expected Integrated Bernoulli Variance}
+
+Next we investigate how sampling designs are chosen to minimize the expected Integrated Bernoulli variance (EIBV).
+
+Data are gathered by an AUV that can measure temperature and salinity
+at a subset of locations on the defined grid. This can be done in a
+static manner (where the survey design is pre-planned), or in a
+sequential way with several stages of data gathering. For the latter,
+a stage of data is gathered first, and then assimilated to get an
+updated model. Then the updated model is used to design and select the
+next stage of measurements, and so on (Section
+\ref{sec:heuristics}). Below, we describe the model for one stage of
+data and the Gaussian formula for a single update.
+
+In order to be able to calculate them in semi-analytical forms relying on integrals of multivariate Gaussian cumulative distribution functions, we will appeal to some key intermediate results presented next.
+
+(MODIFY NOTATION HERE - BUT NEED THIS $y$ IN HERE)
+
+We denote the data at $n_y$ locations by
+$\by=(y_{T,1},y_{S,1},\ldots,y_{T,n_y},y_{S,n_y})^t$, representing the
+temperature and salinity measurements gathered at the first stage of
+nodes in the grid. The conditional model for the data, given the true temperature
+and salinity, is described by: 
+\begin{equation}\label{likelihood}
+\by | \bxi \sim N( \bG \bxi, \bR), %\pi(\by | \bxi)= \frac{1}{(2\pi)^{m}|\bR|^{\frac{1}{2}}}e^{-\frac{1}{2}(\by -\bG\bxi)^T\bR^{-1}(\by-\bG\bxi)},
 \end{equation}
-The excursions could also be defined as larger than a threshold, or
-within a boundary of limits for the two variables; calculations and
-interpretations will be similar in these cases.
-
-The ES holds random indicator variables while the associated EPs are: 
-\begin{equation}\label{eq:prob}
-  p(\bx) = P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S), \hspace{3mm} \bx \in \mathcal{M}.
-\end{equation}
-If the EPs in Eq. (\ref{eq:prob}) are
-close to $1$ or $0$ at a given location, it is easy to classify water mass at this point as either river or ocean; it is, however,
-challenging if the probability is close to $0.5$. 
-The IBV, see \cite{bect2019}, is defined as:
-\begin{equation}\label{measVAR}
-    V = \int_{\bx \in \mathcal{M}} p(\bx) \left(1-p(\bx)\right) d\bx.
-\end{equation}
+where $\bG$ is a $2n_y \times 2n$ matrix containing '$1$'s on the
+measurement indices of the grid and '$0$' otherwise. The size
+$2n_y \times 2n_y$ covariance matrix $\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations.
 
@@
 evaluate the expected reduction in the IBV, conditioned on the survey
-data. The expected updated IBV is:
+data. The EIBV is:
+
+(MODIFY NOTATION HERE)
 
@@
 
+REUSE PROPOSITION RESULTS
+
 %Recent work on ESs and EPs connected to spatial statistics include
@@
 
-\section{Gaussian Processes and Excursion Probabilities}
-\label{sec:GP_EP}
-
 Calculations of the EPs and the IBV require a model specification. We
@@
 
-\subsection{Bivariate GPs}
-
-The Gaussian model is used to represent the bivariate random fields of
-temperature and salinity. At location $\bx \in \mathcal{M}$, the
-bivariate distribution is:
-\begin{equation}\label{gp_bar}
-  \begin{bmatrix}\xi_T(\bx) \\
-    \xi_S(\bx) \end{bmatrix}
- \sim N \left( 
-\begin{bmatrix} \mu_{T}(\bx)\\
-\mu_{S}(\bx)
-\end{bmatrix},\begin{bmatrix}
-\sigma_{T}^2(\bx) & \sigma_{T}(\bx) \sigma_S(\bx) \gamma(\bx)  \\
-\sigma_{T}(\bx) \sigma_S(\bx) \gamma(\bx)  & \sigma_{S}^2(\bx) 
-\end{bmatrix}
-\right),
-\end{equation}
-where the notation refers to Gaussian (i.e. normal) distributed
-variables specified by the mean vector and covariance matrix. For the
-application of a river plume, the model parameters are specified from
-a short preliminary survey. The means vary with location, capturing
-the variability near the river mouth, while the variance parameters
-and correlation between salinity and temperature are assumed to be
-constant for all locations. %and within each water mass.
-
-For the spatial dependence among variables, we assume a separable
-correlation function with the same decay for salinity and temperature,
-which is consistent with expectations for a river plume where the
-decay is related to common water masses (Section
-\ref{sec:case_study}). In the resulting function, we have:
-\begin{equation}\label{gp_corr}
-\mbox{Cov}(\xi_i(\bx),\xi_j(\bx')) = \rho(\bx,\bx') \sigma_i \sigma_j (\gamma +(1-\gamma )\delta_{ij}), \\
-\end{equation}
-where for $i,j \in {T,S}$ and $\delta_{ij}=1$ if $i=j$, and
-$\delta_{ij}=0$ otherwise. Moreover, we assume stationary isotropic
-random fields where the correlation function $\rho(\bx,\bx')$
-solely depends on $\bx$ and $\bx'$ via Euclidean distance
-$h=\sqrt{||\bx-\bx'||^2}$. With data and prior knowledge, one could
-possibly fit and estimate parameters of non-stationary or
-non-isotropic covariance functions and more complex multivariate
-spatial covariance functions
-\citep{gneiting2010matern,genton2015cross}; however, that is left for
-future work.
-
-We discretize the spatial domain $\mathcal{M}$ to a set of $n$ grid
-locations $\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell
-has area $\Delta$; the grid is used for the waypoint graph setting the
-possible design locations. It is also used when solving integral
-expressions in Eq. (\ref{measVAR}) and (\ref{sur}), which are then
-approximated by sums over all grid cells. Using vector notation, we
-denote the Gaussian temperature and salinity variables on the grid and
-its density function by the formula:
-\begin{equation}\label{prior}
-    \bxi = (\xi_T(\bx_1),\xi_S(\bx_1),\ldots,\xi_T(\bx_n),\xi_S(\bx_n))^t, \hspace{3mm}
-    \bxi  \sim  N(\bmu, \bSigma). %\pi (\bxi) & =& \frac{1}{(2\pi)^{n}|\bSigma|^{\frac{1}{2}}}e^{-\frac{1}{2}(\bxi -\bmu)^T\bSigma^{-1}(\bxi-\bmu)}.
-\end{equation}
-%Here, $\pi(\bxi)$ is the joint density function of length $2 n$ vector $\bxi$, and
-The length $2n$ vector $\bmu$ contains the expected values of
-temperature and salinity at the grid locations, while the
-$2n \times 2n$ matrix $\bSigma$ contains the
-variance-covariances of temperature and salinity variables at all grid
-locations.
-
-\subsection{Data Model and Updating of GPs}
-
-Data are gathered by an AUV that can measure temperature and salinity
-at a subset of locations on the defined grid. This can be done in a
-static manner (where the survey design is pre-planned), or in a
-sequential way with several stages of data gathering. For the latter,
-a stage of data is gathered first, and then assimilated to get an
-updated model. Then the updated model is used to design and select the
-next stage of measurements, and so on (Section
-\ref{sec:heuristics}). Below, we describe the model for one stage of
-data and the Gaussian formula for a single update.
-
-We denote the data at $n_y$ locations by
-$\by=(y_{T,1},y_{S,1},\ldots,y_{T,n_y},y_{S,n_y})^t$, representing the
-temperature and salinity measurements gathered at the first stage of
-nodes in the grid. The conditional model for the data, given the true temperature
-and salinity, is described by: 
-\begin{equation}\label{likelihood}
-\by | \bxi \sim N( \bG \bxi, \bR), %\pi(\by | \bxi)= \frac{1}{(2\pi)^{m}|\bR|^{\frac{1}{2}}}e^{-\frac{1}{2}(\by -\bG\bxi)^T\bR^{-1}(\by-\bG\bxi)},
-\end{equation}
-where $\bG$ is a $2n_y \times 2n$ matrix containing '$1$'s on the
-measurement indices of the grid and '$0$' otherwise. The size
-$2n_y \times 2n_y$ covariance matrix $\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations.
+(MODIFY NOTATION HERE)
 
@@
 data $\by$, is Gaussian with mean and covariance:
-\begin{eqnarray}\label{gp_upd}
-  \bm &=& \bm(\by) = \bmu+\bSigma \bG^t (\bG \Sigma \bG^t+\bR)^{-1}(\by-\bG \bmu),  \\
-  \bA &=& \bSigma - \bSigma \bG^t (\bG \bSigma \bG^t+\bR)^{-1} \bG
-          \bSigma.\nonumber
-\end{eqnarray}
+%\begin{eqnarray}\label{gp_upd}
+%  \bm &=& \bm(\by) = \bmu+\bSigma \bG^t (\bG \Sigma \bG^t+\bR)^{-1}(\by-\bG \bmu),  %\\
+%  \bA &=& \bSigma - \bSigma \bG^t (\bG \bSigma \bG^t+\bR)^{-1} \bG
+%          \bSigma.\nonumber
+%\end{eqnarray}
 
@@
 
-\section{Uncertainty reduction in excursion sets}
-\label{sec:sur}
-
-\subsection{EPs for GPs}
-
 Based on Gaussian modeling assumptions, the EPs can be computed at any
@@
 
-\subsection{Expected IBV for Bivariate Models}
-
-We will next derive closed-form solutions for the expectation in
-Eq. (\ref{sur}). The location index $\bx$ is suppressed to avoid
-overly complex notation. We present new results for computing the
-expectation in the integral of Eq. (\ref{sur}).
+MODIFY AND REUSE PROPOSITIONS
 
@@
 of $\bm$, the expected Bernoulli variance becomes:
-\begin{eqnarray}
-\label{eq:var2}
-E_{\by}(p(1-p)) &=& E_{\bm}(p(1-p)), \hspace{3mm} p=P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) \nonumber \\
-E_{\bm}(p(1-p)) & = & \int p(1-p) \pi(\bm) d\bm, \nonumber \\
- &=& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm)  \pi(\bm) d\bm \nonumber  \\
-&-& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) \pi(\bm) d\bm. 
-\end{eqnarray}
+%\begin{eqnarray}
+%\label{eq:var2}
+%E_{\by}(p(1-p)) &=& E_{\bm}(p(1-p)), \hspace{3mm} p=P(\xi_T \leq t_T, \xi_S \leq %t_S|\bm) \nonumber \\
+%E_{\bm}(p(1-p)) & = & \int p(1-p) \pi(\bm) d\bm, \nonumber \\
+% &=& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm)  \pi(\bm) d\bm \nonumber  \\
+%&-& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) %\pi(\bm) d\bm. 
+%\end{eqnarray}
 We will next extend results of \cite{chevalier2014fast} to solve for
@@
 \end{eqnarray}
-
-For the second part of the expression Eq.(\ref{eq:var2}), standardization
-implies that: 
-\begin{eqnarray}\label{part1:phi4}
-&& \hspace{6mm} \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) p(\bm) d\bm =  \\
-&P& \hspace{-3mm} \left( Z_{1,T} \leq \frac{t_T-m_{T}}{a_{T}}, 
-Z_{1,S} \leq \frac{t_S-m_{S}}{a_{S}},Z_{2,S} \leq \frac{t_T-m_{T}}{a_{T}}, 
-Z_{2,S} \leq \frac{t_S-m_{S}}{a_{S}} \right). \nonumber
-\end{eqnarray}
-Here, $\bZ_1=(Z_{1,T},Z_{1,S})$, $\bZ_2=(Z_{2,T},Z_{2,S})$ are
-independent bivariate zero-mean and unit-variance Gaussian vectors,
-both with element-wise correlation $\eta_{T,S}$. Moreover, $\bZ_1$ and
-$\bZ_2$ are chosen to be independent of $\bm$, hence:
-
-\begin{equation}
-\small
-\label{eq:22}
-    \left(
-    \begin{array}{cccccc}
-         Z_{1,T} \\
-         Z_{1,S} \\
-         Z_{2,T} \\
-         Z_{2,S} \\
-         m_{T} \\
-         m_{S}
-    \end{array}
-    \right)
-\sim N \left(
-\left(
-    \begin{array}{cccccc}
-    0 \\
-    0 \\
-    0 \\
-    0 \\
-          \mu_{T} \\
-         \mu_{S} 
-    \end{array}
-    \right),
-   \left[ 
-      \begin{array}{cccccc}
-        1 & \eta_{T,S} & 0 & 0 & 0 & 0  \\
-        \eta_{T,S} & 1 & 0 & 0 & 0 & 0 \\
-        0 & 0 & 1 & \eta_{T,S} & 0 & 0 \\
-        0 & 0 & \eta_{T,S} & 1 & 0 & 0 \\
-        0 & 0 & 0 & 0 & \psi^2_{T} & \psi_{T,S} \\
-        0 & 0 & 0 & 0 & \psi_{T,S} & \psi^2_{S} 
-    \end{array}
-\right]
-\right).
-\vspace{0.2cm}
-\end{equation}
-
-The solution to the second part of Eq. \eqref{eq:22} can be found by
-evaluating a four-variable cumulative distribution function associated
-with Eq. (\ref{part1:phi4}), accounting for the mean, variance and
-covariances of $\bZ_1$, $\bZ_2$ and $\bm$ in the appropriate linear
-combinations. Using vector-matrix formulations, we define the linear
-combinations:
-\begin{equation}
-\label{Vcomb}
-    \bV=\left[
-    \begin{array}{cccccc}
-        a_T & 0 & 0 & 0 & 1 & 0 \\
-         0 & a_S & 0 & 0 & 0 & 1 \\
-         0 & 0 & a_T & 0 & 1 & 0\\
-         0 & 0 & 0 & a_S & 0 & 1\\
-    \end{array}
-    \right] 
-    \left(
-    \begin{array}{cccccc}
-         Z_{1,T} \\
-         Z_{1,S} \\
-         Z_{2,T} \\
-         Z_{2,S} \\
-         m_{T} \\
-         m_{S}
-    \end{array}
-    \right),
-\end{equation}
-which in turn then becomes:
-{\small
-\begin{eqnarray}
-\label{part1:phi4_2}
-&&P \left( Z_{1,T} \leq \frac{t_T-m_{T}}{a_{T}}, 
-Z_{1,S} \leq \frac{t_S-m_{S}}{a_{S}}, Z_{2,S} \leq \frac{t_T-m_{T}}{a_{T}}, Z_{2,S} \leq \frac{t_S-m_{S}}{a_{S}} \right)\\
-&=&  \Phi_4 
-\begin{pmatrix} 
-\begin{bmatrix} t_T\\
-t_S \\
-t_T \\
-t_S
-\end{bmatrix};
-E( \bV),Var(\bV)
-\end{pmatrix} \nonumber.
-\end{eqnarray}
-}
-The multivariate cumulative probabilities in two and four dimensions
-are computationally efficient using methods developed by
-\cite{genz2009computation}.  In the end, the first part in
-Eq. (\ref{two_parts0}) and the second in Eq. (\ref{part1:phi4}) are
-computed for each location on the grid and summed over
-$\bx \in \mathcal{M}_g$, to get the design evaluation criterion in
-Eq. (\ref{sur}).
-
-\subsection{Generalization to the Multivariate Case}
-\label{gen_k}
-
-In our application domain with the variables temperature and salinity,
-the solutions to the complex integral equations for the expected IBV
-are reduced to calculating bivariate and four-variate cumulative
-distribution functions for Gaussian vectors. For the more general
-case, one would be interested in $K$ random fields. In the ocean
-sciences, the variables could include chlorophyll or other
-bio-geochemical quantities in addition to temperature and salinity.
 
@@
 
-\subsection{An Example of IBV using Temperature and Salinity}
+\subsection{An Example of EIBV using Temperature and Salinity}
 
@@
 \subsection{Modeling}
+
+The Gaussian model is used to represent the bivariate random fields of
+temperature and salinity. At location $\bx \in \mathcal{M}$, the
+bivariate distribution is:
+\begin{equation}\label{gp_bar}
+\begin{bmatrix}\xi_T(\bx) \\
+\xi_S(\bx) \end{bmatrix}
+\sim N \left( 
+\begin{bmatrix} \mu_{T}(\bx)\\
+\mu_{S}(\bx)
+\end{bmatrix},\begin{bmatrix}
+\sigma_{T}^2(\bx) & \sigma_{T}(\bx) \sigma_S(\bx) \gamma(\bx)  \\
+\sigma_{T}(\bx) \sigma_S(\bx) \gamma(\bx)  & \sigma_{S}^2(\bx) 
+\end{bmatrix}
+\right),
+\end{equation}
+where the notation refers to Gaussian (i.e. normal) distributed
+variables specified by the mean vector and covariance matrix. For the
+application of a river plume, the model parameters are specified from
+a short preliminary survey. The means vary with location, capturing
+the variability near the river mouth, while the variance parameters
+and correlation between salinity and temperature are assumed to be
+constant for all locations. %and within each water mass.
+
+For the spatial dependence among variables, we assume a separable
+correlation function with the same decay for salinity and temperature,
+which is consistent with expectations for a river plume where the
+decay is related to common water masses (Section
+\ref{sec:case_study}). In the resulting function, we have:
+\begin{equation}\label{gp_corr}
+\mbox{Cov}(\xi_i(\bx),\xi_j(\bx')) = \rho(\bx,\bx') \sigma_i \sigma_j (\gamma +(1-\gamma )\delta_{ij}), \\
+\end{equation}
+where for $i,j \in {T,S}$ and $\delta_{ij}=1$ if $i=j$, and
+$\delta_{ij}=0$ otherwise. Moreover, we assume stationary isotropic
+random fields where the correlation function $\rho(\bx,\bx')$
+solely depends on $\bx$ and $\bx'$ via Euclidean distance
+$h=\sqrt{||\bx-\bx'||^2}$. With data and prior knowledge, one could
+possibly fit and estimate parameters of non-stationary or
+non-isotropic covariance functions and more complex multivariate
+spatial covariance functions
+\citep{gneiting2010matern,genton2015cross}; however, that is left for
+future work.
+
+We discretize the spatial domain $\mathcal{M}$ to a set of $n$ grid
+locations $\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell
+has area $\Delta$; the grid is used for the waypoint graph setting the
+possible design locations. It is also used when solving integral
+expressions in Eq. (\ref{measVAR}) and (\ref{sur}), which are then
+approximated by sums over all grid cells. Using vector notation, we
+denote the Gaussian temperature and salinity variables on the grid and
+its density function by the formula:
+\begin{equation}\label{prior}
+\bxi = (\xi_T(\bx_1),\xi_S(\bx_1),\ldots,\xi_T(\bx_n),\xi_S(\bx_n))^t, \hspace{3mm}
+\bxi  \sim  N(\bmu, \bSigma). %\pi (\bxi) & =& \frac{1}{(2\pi)^{n}|\bSigma|^{\frac{1}{2}}}e^{-\frac{1}{2}(\bxi -\bmu)^T\bSigma^{-1}(\bxi-\bmu)}.
+\end{equation}
+%Here, $\pi(\bxi)$ is the joint density function of length $2 n$ vector $\bxi$, and
+The length $2n$ vector $\bmu$ contains the expected values of
+temperature and salinity at the grid locations, while the
+$2n \times 2n$ matrix $\bSigma$ contains the
+variance-covariances of temperature and salinity variables at all grid
+locations.
 
@@
 
+\section*{Appendix}
+
+Proof of Proposition 1:
+
+\begin{proof}
+	That $\mes(\es)$ defines indeed a random variable follows from Fubini's theorem 
+	relying on the joint measurability of 
+	$(\x, \omega) \to \mathbbm{1}_{\es(\omega)}(\x)$, 
+	%= \mathbf{1}_{\xi_{\x}(\omega) \in T}$ and Fubini's theorem.  
+	itself inherited from the assumed measurability for  
+	$(\x, \omega) \to \gp[\x](\omega)$ and $T$, respectively. From there, following the steps of Robbins' theorem \cite{Robins1944}, we find that 
+	\begin{equation*}
+	\begin{split}
+	\mathbb{E}[\mes(\es)^{r}]
+	&=\mathbb{E}\left[\left(\int_{D} \mathbbm{1}_{\gp[u] \in T} ~d\mes(u) \right)^{r} \right]
+	=\mathbb{E}\left[ \prod_{i=1}^{r} \left(
+	\int_{D} \mathbbm{1}_{\gp[\uu^{(i)}] \in T} ~d\mes(\uu^{(i)}) 
+	\right) \right] \\
+	&
+	=
+	%\mathbb{E}\left[
+	%\int_{D^{r}}
+	%\prod_{i=1}^{r}
+	%\mathbf{1}_{\gp[\uu^{(i)}] \in T}
+	%\productMeasure
+	%\right]
+	%=
+	\mathbb{E}\left[
+	\int_{D^{r}}
+	\mathbbm{1}_{\gp[\uu^{(1)}] \in T,\dots, \gp[\uu^{(r)}]  \in T}
+	\productMeasure
+	\right]
+	%\\
+	%&=\int_{D^{r}} \mathbb{E}\left[\mathbf{1}_{\gp[\bm{u}] \in T^r} \right] \productMeasure \\
+	%&
+	=\int_{D^{r}}
+	\jointExcuProb 
+	\productMeasure.
+	\end{split}
+	\end{equation*}   
+	The rest consists in expliciting the probability of $T\times \dots \times T$ under the multivariate Gaussian distribution of  
+	$\left(\gp[u^{(1)}], \dots,  \gp[u^{(r)}] \right)$. 
+\end{proof}
+
+Proof of Proposition 2:
+
+\begin{proof}
+	By definition of $\Phi_{p}$, for $N\sim \mathcal{N}_{p}(0_{p},\covN)$, 
+	$$
+	\mathbb{P}(N\leq a + BV | V)
+	=
+	\varPhi_{p}\left( a + BV; \covN \right).
+	$$
+	Now for $\varPhi_{p}\left( a + BV; \covN \right)^h$, provided that the probability space is sufficiently large to accomodate $h$ independent Gaussian random vectors $N_i\sim \mathcal{N}_{p}(0,\covN)$ (which is silently assumed here), using the former equality delivers
+	$$
+	\varPhi_{p}\left( a + BV; \covN \right)^h
+	=
+	\prod_{i=1}^h \mathbb{P}(N_i\leq a + BV | V).
+	$$
+	Now by independence of the $N_i$'s we obtain the joint conditional probability 
+	$$
+	\prod_{i=1}^h \mathbb{P}(N_i\leq a + BV | V)
+	=
+	\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV| V),
+	$$
+	whereof, by virtue of the law of total expectation, 
+	\begin{equation*}
+	\begin{split}
+	\mathbb{E}\left[ \varPhi_{p}\left( a + BV; \covN \right)^h \right]
+	&=\mathbb{E}\left[\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV| V)\right]\\
+	&=\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV)\\
+	&=\mathbb{P}(W_1 \leq a, \dots, W_h\leq a)\\
+	&=\varPhi_{ph}
+	\left(
+	1_{h} \otimes a
+	;
+	(1_{h}1_{h}')\otimes (B\Sigma_{V} B') + 
+	I_{h}\otimes \covN
+	\right),
+	\end{split}
+	\end{equation*}
+	where $\mathbf{W}=(W_1,\dots,W_h)$ with $W_i=N_i- BV \ (1\leq i \leq h)$
+	and the last line follows $\mathbf{W}$ forming a Gaussian vector (by global independence of the $N_i$'s and $V$) and from the definition of $\varPhi_{p h}$.
+The covariance matrix $\mathbf{\Sigma}$ of $\mathbf{W}$ is obtained by noting that $\operatorname{cov}(W_i,W_j)=B \covV B' + \delta_{ij} \covN \
+	(i,j \in \{1,\dots,h\})$.
+	%the latter concatenated Gaussian vector is obtained by straightfoward calculation. 
+\end{proof}
+
+We the case of multivariate monomials in orthant probabilities with thresholds affine in a common Gaussian vector. 
+%Namely, for monomials of degree $k$, we have.
+
+\begin{propo}%[\textcolor{red}{Notation in progress, to be reviewed/revised}]
+	\label{propo3}
+	Let $g, p, q\geq 1$, $h_{1},\dots, h_{g}\geq 1$ with $H=\sum_{i=1}^g h_i$, $a_{i} \in \R^{p}$, $B_{i}\in \R^{p \times q}$, and covariance matrices $\covN_i \in \R^{p \times p}$ $(1\leq i \leq g)$.
+Then, for any covariance matrix $\covV \in \R^{q\times q}$ and $V\sim\mathcal{N}_{q}(0_q,\covV)$,
+	%	Let $g$ and $k$ be integers, and consider a $g$-terms monomial of degree $k$
+	%	\[
+	%	    \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \Sigma_{i}
+	%	    \right)^{h_i},
+	%    \]
+	%    where $\sum_{i=1}^g h_i = k$, each $a_i$, $B_i$ and $\Sigma_i$ is as in
+	%    \cref{propo2} and $V$ is a $q$-dimensional Gaussian vector distributed as
+	%    $V\sim\mathcal{N}(0,\covV)$.
+	%
+	% Then the expectation of the monomial is given by
+	\begin{equation}
+	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
+	=
+	\varPhi_{p H}
+	\left(
+	\bm{a}
+	;
+	\mathbf{\Sigma}
+	\right),
+	\end{equation}
+	with $\bm{a}=(1_{h_1}\otimes a_1, \dots, 1_{h_g}\otimes a_{g}) \in \R^{p H}$
+	and $\mathbf{\Sigma}\in \R^{p H \times p H}$ is defined blockwise by $(\Sigma_{i,j})_{i,j \in \{1,\dots, g\}}$ where, for any $i,j \in \{1,\dots, g\}$, %the blocks are given by 
+	\begin{equation}
+	\Sigma_{i,j}=
+	(1_{h_{i}}1_{h_{j}}')\otimes (B_{i}\Sigma_{V} B_{j}') + \delta_{i,j}(I_{h_{i}}\otimes \covN_{i}) \in \R^{p h_{i} \times p h_{j}}.
+	\end{equation}
+	%with $\bm{a}=\sum_{j=1}^g \mathbf{1}_{J_{j}} \otimes a_j \in \R^{p \times k}$ 
+	%where $J_{j}=\{\sum_{i=1}^{j-1}h_i + 1, \dots, \sum_{i=1}^{j}h_i\}$,
+	%$\mathbf{1}_{I_{j}}$ denotes the vector of $\R^{k}$ with components equal to $1$ for indices in $I_{j}$ and $0$ else ($1\leq j \leq g$), 
+	%and $\mathbf{\Sigma}\in \R^{p k \times p k}$ is defined blockwise by 
+	%\begin{equation}
+	%\mathbf{\Sigma}_{\widetilde{J}_{j_{1}},\widetilde{J}_{j_{2}}}=
+	%(\mathbf{1}_{h_{j_1}}\mathbf{1}_{h_{j_2}}')\otimes (B_{j_{1}}\Sigma_{V} B_{j_2}') + \delta_{j_{1},j_{2}}(I_{j_{1}}\otimes \Sigma_{j_{1}}),
+	%\end{equation}
+	%where $\widetilde{J}_{j}=\{p\sum_{i=1}^{j-1}h_i + 1, \dots, p\sum_{i=1}^{j}h_i\}$ ($1\leq j \leq g$).
+\end{propo}
+
+\begin{remark}
+	%    Note that the vector $\bm{a}$ is obtained by stacking $h_i$ times each $a_i$
+	%\begin{equation*}
+	%\bm{a} = \begin{pmatrix}
+	%    \bm{a}_1\\ \vdots\\ \bm{a}_g\end{pmatrix},~
+	%\bm{a}_i =
+	%\begin{pmatrix}
+	%    a_i\\ \vdots\\ a_i\end{pmatrix}
+	%\in\mathbb{R}^{l\times h_i}\end{equation*}
+	%and the covariance matrix is obtained by similar stacking
+	Using blockwise representation for the blocks themselves delivers %$\bm{\Sigma}$ can be expressed as follows
+	\begin{equation*}
+	%    \bm{\Sigma} = \begin{pmatrix}
+	%        \Sigma_{11} & & \Sigma_{1g}\\
+	%        & \ddots &\\
+	%        \Sigma_{g1}&   & \Sigma_{gg}
+	%\end{pmatrix},~
+	\Sigma_{ij} =
+	\begin{pmatrix}
+	B_i\Sigma_{V} B_j' & \dots & B_i\Sigma_{V} B_j'\\
+	\vdots & & \vdots\\
+	B_i\Sigma_{V} B_j' & \dots & B_i\Sigma_{V} B_j'\\
+	\end{pmatrix}
+	+
+	\delta_{ij}
+	\begin{pmatrix}
+	\covN_i & &\\
+	& \ddots &\\
+	&   & \covN_i
+	\end{pmatrix}%\in\mathbb{R}^{lh_i\times lh_j}
+	\end{equation*}
+	Here each $\Sigma_{ij}$ is made of $h_i$ times $h_j$
+	(vertically/horizontally) $p \times p$ sub-blocks, hence possesses $ph_i$ lines and $ph_j$ columns.
+\end{remark}
+
+%Proof TODO along with notation polish.
+
+\begin{proof}
+	The proof relies (again) heavily on the fact that, by definition of $\Phi_{p}$, for any covariance matrix $\covN \in \R^{p \times p}$, $a\in \R^p$, $B\in \R^{p \times q}$, and $N\sim \mathcal{N}_{p}(0_{p},\covN)$,
+	$$
+	\mathbb{P}(N\leq a + BV | V)
+	=
+	\varPhi_{p}\left( a + BV; \covN \right).
+	$$
+	In particular, for globally independent $N_{i,j} \sim \mathcal{N}_{p}(0_{p},\covN_i)$ $(1\leq j \leq h_i, 1\leq i \leq g)$, 
+	\begin{equation*}
+	\begin{split}
+	\prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i}
+	&=
+	\prod_{i=1}^{g}
+	\prod_{j=1}^{h_{i}}
+	\mathbb{P}(N_{i,j}\leq a_i + B_i V | V)\\
+	&=\mathbb{P}(N_{1,1}\leq a_1 + B_1 V, \dots, N_{g,h_{g}}\leq a_g + B_g V | V),
+	\end{split} 
+	\end{equation*}
+	so that, by the law of total expectation, 
+	\begin{equation*}
+	\begin{split}
+	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
+	=
+	\mathbb{P}(W_{1} \leq 1_{h_{1}} \otimes a_1, \dots, W_{g} \leq 1_{h_{g}} \otimes a_g)
+	\end{split}
+	\end{equation*}
+	where 
+	$W_{1}=(N_{1,1}- B_1 V, \dots, N_{1,h_{1}}- B_1 V), 
+	W_{2}=(N_{2,1}- B_2 V, \dots, N_{2,h_{2}}- B_2 V), 
+	\dots, W_{g}=(N_{g,1}- B_g V, \dots, N_{g,h_{g}}- B_g V)$. Noting that $\mathbf{W}=(W_1,\dots, W_g)$ is a centred $p H$-dimensional Gaussian random vector, we finally obtain that 
+	\begin{equation*}
+	\begin{split}
+	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
+	=
+	\varPhi_{p H}\left(\bm{a};\mathbf{\Sigma}\right),
+	\end{split}
+	\end{equation*}
+	with $\bm{a}=(1_{h_{1}} \otimes a_1, \dots, 1_{h_{g}} \otimes a_g)$ and $\bm{\Sigma}=(\operatorname{cov}(W_i,W_j))_{i,j \in \{1,\dots, g\}}$.
+\end{proof}
+
 % AOS,AOAS: If there are supplements please fill:

### 2020-05-29 ans_to_reviewers/answers-to-reviewers.tex
Revised paper and reply letter.

@@
 suggestions. They have all been considered in this revised
-version of this manuscript. The principal revisions made to the manuscript are as follows and are done to address the main comments by Editors and Reviewers:
+version of this manuscript.
+We have worked hard on developing both theoretical results, computational resources, and finally on the presentation in the paper.
+To help us on these efforts, we included Cedric Traveletti as co-author because he has contributed significantly on the revised version of the paper.
+
+The principal revisions made to the manuscript are as follows and are done to address the main comments by Editors and Reviewers:
 
@@
 Another example is that of photosynthesis where plants and algae produce oxygen only when given enough of every component; light, carbon dioxide and water.
-Our motivation in this paper is that temperature and salinity, combined, drive several processes in the oceans.
+Our motivation in this paper is that joint temperature and salinity excursions are proxies for several processes in the oceans.
 For instance, salmon lice larvae are known to depend on relatively high temperature and high salinities to survive.
@@
 
-{\it{Action:}} In the revised manuscript these aspects are largely rewritten in the Introduction, Background, ES definitions and Case Study. This change addresses the comments of AE1, R1 (3,) MORE.
+In the revised manuscript these aspects are largely rewritten in the Introduction, Background, expected integrated Bernoulli variance (EIBV) and Case Study.
+This change addresses the comments of AE1, R1 (3,) MORE.
 
-\item The {\bf{statistical methodological contribution}} have been clarified and improved, including definitions of ESs and having more general results related to the integrated Bernoulli variance (IBV) for multivariate processes, using marginal (one-location) entries and for volume calculations requiring multiple locations.
+In the revised version we motivate more for statisticians, in particular we emphasize the statistical design problem which is sought solved by AUV with its constraints on navigation and computing resources.
+
+\item The {\bf{statistical methodological contribution}} have been clarified and improved, including definitions of ESs and having more general results related to the EIBV for vector-valued processes, using marginal (one-location) entries and for volume calculations requiring multiple locations.
 The revised version highlights these novel statistical results as a main contribution of the paper.
 
-{\it{Action:}} In the revised manuscript these aspects are largely rewritten in the main body containing novel results for IBV for sums of marginal variances as well as for volumes, while the new Appendix derives the closed form results.
+In the revised manuscript these aspects are largely rewritten in the main body containing novel results for EIBV for sums of marginal variances as well as for more general results inluding that of volumes of random sets, while the Appendix derives the closed-form results.
 This change addresses the comments of E2, AE1, R2 (7), R3 (1), MORE.
@@
 The {\bf{simulation results and code}} are thoroughly re-worked.
-Open-source python code is now provided for all the examples used in the paper and also for simulation studies using more nuances statistical models (non-stationary, non-isotropic and non-separable).
+Open-source python code is now provided for all the examples used in the paper and also for simulation studies using more nuanced statistical models (non-stationary and non-isotropic).
 
-{\it{Action:}} The simulation study section is largely modified and it contains pointers to the associated open-source code. This change addresses the comments of R1 (2,3),  MORE.
-
-\item In the initial submission we tried to motivate these kinds of applications carefully, but based on the comments, we see that the {\bf{introduction and background}} were too long.
-We were not selling the paper enough to professional statisticians, and maybe too much to other scientists who are interested in particular areas of these statistical application.
-In the revised version we motivate more for statisticians, in particular we emphasize the statistical design problem which is sought solved by AUV with its constraints on navigation and computing resources.
-
-{\it{Action:}} The Introduction and Background sections have been largely rewritten.
-We now get quicker to the point and engage readers in its both applied statistical focus and the methodological contributions.
-This change addresses the comments of E1, E4, AE1, R2 (7), R3 (1), MORE.
-
-\item We have further included Cedric Traveletti as co-author. He has contributed on deriving the more general statistical properties of the methods and on improving the code for open-source quality.
+The simulation study section is largely modified and it contains pointers to the associated open-source code. This change addresses the comments of R1 (2,3),  MORE.
 
@@
 \par \vspace{1em}
-
-Note that the reference to page $X$ is made relative to
-the original submitted manuscript given to the reviewers.
 
@@
 In the revised version we motivate more for statisticians.
-(See also comment 4 in the list of main changes above.)
 
-\action: Our actions are to i) rewrite Introduction and Background sections to catch our intended audience and highlight our contributions, ii) restructure the presentation to highlight both statistical contributions as well as the impact such technologies and applied questions could provide for statisticians.
+Our actions are to i) rewrite Introduction and Background sections to catch our intended audience and highlight our contributions, ii) restructure the presentation to highlight both statistical contributions as well as the impact such technologies and applied questions could provide for statisticians.
+(See also comment 1 in the list of main changes above.)
 
@@
 \reply: We realize that we did not stress the statistical novelty sufficiently in the submitted version.
-The main statistical novelty is in the closed form calculation for the design criteria of the integrated Bernoulli variance (IBV) associated with excursion sets (ESs) for multivariate processes.
+The main statistical novelty is in the closed-form calculation for the design criteria of the expected integrated Bernoulli variance (EIBV) associated with excursion sets (ESs) for multivariate processes.
 In the revised version this is extended to volumes (going beyond single location).
@@
 
-\action: Our actions are to i) extend the closed form calculations to a criterion working on ES volume calculations (going beyond single location), ii) restructure the presentation to work directly on $p$-variate random fields and hence generalize the presentation, iii) restructure the closed form results to be presented in the main body, but with proofs in the new Appendix.
+Our actions are to i) extend the closed-form calculations to a criterion working on ES volume calculations (going beyond single location), ii) restructure the presentation to work directly on $p$-variate random fields and hence generalize the presentation, iii) restructure the closed-form results to be presented in the main body, but with proofs in the new Appendix.
 
@@
 
-That said, in the updated simulation study, which we have generalized and made open-source there are opportunities and examples of more complex covariance models (non-stationary and non-separable and non-isotropic), but we do not conduct any inference for parameters in these models from ocean model data.
+That said, in the updated simulation study, which we have generalized and made open-source there are opportunities and examples of more complex covariance models (non-stationary and non-isotropic), but we do not conduct any inference for parameters in these models because it would require high resolution data from ocean models or satellites.
 
-\action: Our actions include i) a better description of our parameter choices for the synthetic case and the specification for the real-data case, and ii) brief notes of the extended opportunities available for the associated open-access code, iii) references to related papers (including one of our own) that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here.
+Our actions include i) a better description of our parameter choices for the synthetic case and the specification for the real-data case, and ii) brief notes of the extended opportunities available for the associated open-access code, iii) references to related papers (including one of our own) that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here.
+MORE
 
@@
 
-\action: Our actions include a much shorter Section 2 with background. We now aim to engage readers earlier on with the applied statistical focus and methodological contribution of our work.
+Our actions include a much shorter Section 2 with background. We now aim to engage readers earlier on with the applied statistical focus and methodological contribution of our work.
 
@@
 
-\reply:
-
-\action: Added references \cite{adler2000excursion,adler2009random}.
+\reply: The revised version has an entirely new Section 3 which includes general background for random sets and excursion sets.
+We added references, for instance \cite{adler2000excursion,Molchanov2005}.
 
@@
 
-\reply: We agree that they are not necessary for the .
-In the open-source that is now available, we implement non-stationary, non-separable and non-isotropic covariance models.
-But we do not conduct any inference for parameters.
-This might be possible for abundant data, but not in our small-scale resolution example.
-
-\action: No description of modeling assumptions in the methodological sections.
-A better description of our parameter choices for the synthetic case and the specification for the real-data case.
-Brief notes of the extended opportunities available for the associated open-access code.
-References to related papers (including one of our own) that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here.
+\reply: We agree that they are not necessary for the methodological contributions, so moved this specification to the simulation study and example. 
+In the open-source that is now available, we implement non-stationary and non-isotropic covariance models.
+These represent situations that have relevance for our case.
+The code we post along with the paper facilitate testing so users can input more complex models to see how it influences designs.
 
@@
 
-\reply:
+\reply: The assumptions are conditionally independent measurements of salinity and temperature.
+We believe this is realistic as they are gathered from different sensors.
+We added a comment about this not necessarily being the case with other data that are extracted and processed by combining sensors.
+In the latent Gaussian random field, the variables are dependent.
 
-\action: The assumptions are conditionally independent measurements of salinity and temperature.
-We believe this is realistic as they are gathered from different sensors.
-In the latent Gaussian random field, the variables are dependent.
 \vspace{1em}
@@
 
-\reply:
-
-\action: The AUV and the sampling happens regularly at 1 Hz, the collected measurements are then assimilated to this regular grid.
+\reply: The AUV and the sampling happens regularly at 1 Hz, the collected measurements are then assimilated to this regular grid.
 The sampling density is therefore higher than the grid resolution.
 Still, there is a limitation on restricting the sampling locations to finite number of fixed locations.
-However, in order to handle the computational load and resources available on the AUV this simplification is necessary.
+In order to handle the computational load and navigation resources available on the AUV this simplification is necessary.
 % Comment Trygve: Should we add some more here?
@@
 
-\reply:
+\reply: We clarify this. TO DO!
 
-\action: Our actions
 \vspace{1em}
@@
 
-\reply:
+\reply: We clarified the parameter specification. TO CHECK.
 
-\action: Our actions. Clarify. In a larger scale this could be improved by having auxiliary data for instance from ocean models or satellite data.
+In a larger scale this could be improved by having auxiliary data for instance from ocean models or satellite data. (Mentioned under further work in closing remarks.)
 
@@
 
-\reply:
+\reply: This part has been rewritten. TO DO.
 
-\action: Our actions
 \vspace{1em}
@@
 
-\action: We have changed this in the text - page 3 and 12.
+\reply: We have changed this to be consistent in the revised text.
 
@@
 
-\action: We have changed the statement on page 9 to: ``$\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations."
+\reply: We have changed the statement on page 9 to: ``$\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations."
 
@@
 
-\action: We have changed this in the text - page 19.
+\reply: We have changed this in the text - page 19.
 
@@
 
-\action: we have made this change to the citations \cite{genton2015cross} and \cite{french2016credible} - page 30.
+\reply: we have made this change to the citations \cite{genton2015cross} and \cite{french2016credible} - page 30.
 \end{itemize}
@@
 
-\action: Thank you. We have improved and documented the code which is now open-source available.
+\reply: Thank you. We have improved and documented the code which is now open-source available.
 
@@
 
-\action: We have improved the augmented the discussion on other design criteria and why the ES related IBV is used in our case.
-In this part, we cite the handbook reference on design \cite{zidek2019monitoring} and the paper we believe you refer to with a comment from Jon Lee \cite{bueso1998state}.
+\reply: We have improved the augmented the discussion on other design criteria and why the ES related IBV is used in our case.
+In this part, we cite the handbook reference on design \cite{zidek2019monitoring} and the paper we believe that contains a comment from Jon Lee \citep{bueso1998state}.
 
@@
 
-\action: We incorporated the suggested minor changes.
+\reply: We incorporated the suggested minor changes.
 
@@
 
-\action: Our action
+\reply: TO DO
 
@@
 
-\action: We have made an open-source repository available for the public.
+\reply: We have made an open-source repository available for the public.
 This includes commented code examples for exploration and analysis of the material presented in the paper.
 The code is optimized for speed to allow fast computation using Pytorch (citation).
-The code is available here: (link).
+The code is available here: (TO DO link).
 
@@
 
-\action: Our actions
+\reply: TO DO
 
@@
 
-\action: Add reference to Adler.
-Motivate environmental sensing.
-Plants need BOTH oxygen and sun.
-Medical doctors diagnose with MORE THAN ONE symptom, not enough to have a fever.
-Plankton and sea life (relate to lice in salmon industry) need high temperatures and high salinity.
+\reply: We added reference to \cite{adler2009random} on random sets and motivated this topic for environmental sensing. See our general comment 1 on top.
 
@@
 
-\action: Our actions
+\reply: We improved our statistical contribution. See our general comment 2 on top.
 
@@
 
-\action: Our actions.
-The traditional entropy calculations for this GP aim to reduce the log determinant of the updated covariance matrix at the grid locations.
+\reply: The traditional entropy calculations for this GP aim to reduce the log determinant of the updated covariance matrix at the grid locations.
 It will not vary with the data realizations, so in this setting this information criteria will always give the same path.
-The IBV criterion is aligned with our ES motivation for this study.
-When focusing on the expected IBV, the paths will vary with the data as we see in our simulation study and in the real-data example.
+The EIBV criterion is aligned with our ES motivation for this study.
+When focusing on the expected EIBV, the paths will vary with the data as we see in our simulation study and in the real-data example.
 We see this as a clear benefit of the approach compared with entropy-based calculations.
@@
 
-\action: Our actions.
-We focus on point-wise predictions in the criterion.
+\reply: We focus on point-wise predictions in the criterion.
 Our trends are simple, so another covariance might be overfitting too much to data.
@@
 
-\action: LA improved in revision! + code. %Also studied benefits of look-ahead strategies in sequential sampling in \cite{martinelli2013dynamic}.
+\reply: TO DO LA improved in revision! + code. %Also studied benefits of look-ahead strategies in sequential sampling in \cite{martinelli2013dynamic}.
 
@@
 
-\action: This is a simple model. Ocean models for training?
+\reply: This is a simple model for plume temperature and salinity variation.
+However, it is not common to have ocean model data or satellite data at this resolution, which would be required for training.
+
+We have chosen to set few parameters, which also means that it is relatively easy to tune them at a break in the operation, if so is necessary.
 
@@
 
-\action: As pointed out above, we believe that spatio-temporal variations would be a more realistic extension, especially if repeated satellite data or ocean modeling is available.
+\reply: As pointed out above, we believe that spatio-temporal variations would be a more realistic extension, especially if repeated satellite data or ocean modeling is available.
 But this is likely relevant at a larger scale, such as that of fronts in a fjord or ocean interface.
@@
 \item “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.\par
-\action: Our actions
+\reply: We avoid the term' joint excursion sets' in the revised version.
+Instead we have introduced a mo more general random sets notation, but link to the ES for temperature and salinity whenever it is appropriate to exemplify concepts and show results.
+
 \vspace{1em}
@@
 \item Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?\par
-\action: Preliminary runs set these trends. In a larger scale, one could here use ocean model inputs or satellite data.
+
+\reply: Preliminary runs set these trends. In a larger scale, one could here use ocean model inputs or satellite data.
 
@@
 \item I don’t think Figure 3 conveys too much additional benefit and eliminating it for space might be considered.\par
-\action: Our actions
+
+\reply: We removed this Figure. It took much space and it is sufficient to explain the optimal design in text as we did. 
 \vspace{1em}
@@
 \item Why are the ES’s called ”excursion sets? Why excursion probabilities?\par
-\action: Our actions
+
+\reply: The revised Section 3 contains a general formulation for random sets, which is then exemplified by our excursion sets for the multivariate process.
+In the Introduction we motivate the excursion set for multivariate processes for this application, because it is important to find zones with joint high values of many variables.
+In the revised version we believe that the sets are introduced both more in the line with the practical application as well as for statistical novelty.
 \vspace{1em}
@@
 I doubt that most statisticians would know that phrase.\par
-\action: Clarified. 
+\reply: TO DO. Clarify. 
 \vspace{1em}
@@
 I realize that in other scientific domains that distinction is not always made but I do recommend the change if the paper is to appear in a stats journal.\par
-\action: Our actions
+
+\reply: We use $Z$ in the revised version.
 \vspace{1em}
@@
 Statisticians will know these things and users will not care.\par
-\action: Our actions
+\reply: We collapsed the previous sections 3, 4 and 5 to one new section 3 containing the statistical novelty on EIBV closed-form expressions. The proofs are in a new Appendix.
+
 \vspace{1em}
@@
 What are the computational limits of these machines?\par
-\action: We have added more information about AUVs and their computational capabilities in Section 2.1 (Background on Ocean Processes and Observations) to address this. 
+
+\reply: We have added more information about AUVs and their computational capabilities in Section 2.1 (Background on Ocean Processes and Observations) to address this. 
 \begin{quote}
@@
 \item[2.1.2] How long do the AUVs typically explore in one go? Is there a minimum/maximum time?\par
-\action: Maximum speed and payload are limiting for coverage.
+
+\reply: Maximum speed and payload are limiting for coverage.
 Generally this amounts to 8-24 hrs.
@@
 Time constraints for operations are typically related to costs of ship time and man-hours during operation.
-Battery capacity is not a challenge for the AUV described in the paper.
+This information is contained in the changes made in point 2.1.1 above.
 \vspace{1em}
@@
 \item[2.1.3] What kind of movements can the AUVs make? Chess-like moves (horizontal, vertical, diagonal)?\par
-\action: We clarified the descriptions of a way-point graph.
+
+\reply: AUVs are free to move in 6 degrees of freedom on the triangulated graph.
+They are essentially flying under water.
+However, one usually always proceed forward while pitching up and down.
+We have clarified this in the updated section shown in point 2.1.1.
+This is also shown in some of the figures (Fig.1 and Fig.
+4) as well as the waypoint graph.
 \vspace{1em}
@@
 \item[2.1.4] Is there a minimum/maximum number of points you want to sample?\par
-\action: Assuming a fixed time of sampling.
+
+\reply: The coverage available is usually limited on battery capacity, which also determines the maximum number of points to sample.
+The important question here is "How do one prioritize the correct points to visit?", which is central to the development of the IBV and the application shown in the paper.
 \vspace{1em}
@@
 \item[2.1.5] What is the size of the grid that is typically considered? i.e., what is a typical value of $n$ in Eq.(9)?\par
-\action: Only surface considered here. No 3D with depth. 
+
+\reply: The waypoint grid size $n$ should be small enough to resolve the process that is of interest and to avoid aliasing.
+A guiding rule of thumb provided by the the Nyquist theorem is to sample at least twice in time for the shortest significant record period, and twice in space for the shortest significant length to either resolve or eliminate (by filtering) scales of oceanic variability shorter than those being studied.
+In practice, this means mapping at an adequate spatial resolution faster than significant changes – in the phenomena – occur.
 \vspace{1em}
@@
 \item[2.1.6] Does the AUV use a selection algorithm from the very beginning or is there a minimum number of initial measurements that are made before the optimization algorithm begins?\par
-\action: Batch initialization. Then adaptive sampling.
+
+\reply: As is common in such field trials with AUVs, we gathered data in an initial run that also tests the depth and bearings navigation systems.
+From this data, the model parameters were tuned and the AUV conducted adaptive sampling using the optimization strategy.
 \vspace{1em}
@@
 
-\action: Our actions
+\reply: In the revised version the notation has been improved.
 
@@
 
-\action: No data to fit a complex model. See Ocean model comments to Rev 1.
+\reply: We agree that this is a simplified model.
+However, one would need high-resolution ocean model data or satellite information to fit a more complex model.
+But in the accompanying code we run cases with more complex covariances, so the interested user can test if / how it matters for the design.
+We believe spatio-temporal models would be a very interesting next step and present that as further work.
 
@@
 
-\action: Conditionally independent measurements.
+\reply: In our setting the temperature and salinity data are from different sensors and it is naturally to say they are conditionally independent measurements.
+That said, for other measurements, they might be extracted or processed from common sensor readings and would then be conditionally dependent.
+We mentioned this briefly in the data model description.
 
@@
 
-\action: Only the surface is considered here, so computational requirements are not mainly driven by the size of the domain, but rather by the number of combinations for sampling designs (computational complexity).
+\reply: Only the surface is considered here, so computational requirements are not mainly driven by the size of the domain, but rather by the number of combinations for sampling designs (computational complexity).
 No 3D with depth, which would require simplified spatial models like fixed-rank-kriging or similar approaches.
@@
 
-\action: Appendix?
+\reply: We have generalized the closed-form solutions.
+(See point 2 at the start of reply letter.) The main body should contain results of general interest as well as provide insight about the statistical novelty.
+The proofs are now in Appendix with a succinct notation.
 
@@
 
-\action: Appendix?
+\reply: We have generalized the closed-form solutions. (See reply above.)
 
@@
 
-\action: Appendix?
+\reply: We have generalized the closed-form solutions. (See above.)
 
@@
 
-\action: Our actions
+\reply: We have generalized the closed-form solutions. (See reply above.)
 
@@
 
-\action: Joints would be too demanding here. There is more of a question of the longer planning horizon for the AUV to improve the sampling design, so the domain can be more efficiently explored.
+\reply: In the revised version we provide closed-form solutions to expected uncertainty reduction on volumes of random sets.
+However, for real-world application this involves many more terms in the evaluation criteria, so for the application, we still used marginals.
 
@@
 
-While the context doesn’t allow full reproducibility of the paper’s results, no implementation details are provided.
+While the context does not allow full reproducibility of the paper results, no implementation details are provided.
 What software was used to perform the bulk of analysis?
@@
 
-\action: Code is improved in the revision and provided for open-source usage.
+\reply: Code is improved in the revision and provided for open-source usage: LINK
 
@@
 
-\item[2.15.1] “closed-form” is sometimes written as “closed form” (e.g., p. 7, p. 18)
+\item[2.15.1] “closed-form” is sometimes written as “closed-form” (e.g., p. 7, p. 18)
 
-\action:
+\reply: Now consistent with 'closed-form'.
 
@@
 
-\action:
+\reply:
 
@@
 
-\action:
+\reply: Modified in revised version.
+
 \end{itemize}
@@
 
-\action: Space-time \cite{sigrist2015stochastic,richardson2017sparsity}.
-More complex sampling Look-ahead, going further.
-3D sampling, requiring more computationally efficient spatial statistical models.
+\reply: We elaborated on space-time applications \cite{sigrist2015stochastic,richardson2017sparsity}. TO DO: 
+More complex sampling Look-ahead, going further. 3D sampling, requiring more computationally efficient spatial statistical models.
 
@@
 
-\action: Our action
+\reply: This is changed in revised version.
 
@@
 
-\action: Our actions
+\reply: We show EPs for myopic selection. Also, the accompanying code provides more cases.
 

### 2020-05-29 main_new.tex
Revised paper and reply letter.

@@
 % "Title of the paper"
-\title{Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets}
+\title{Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields}
 \runtitle{Excursion Probabilities for Informative Sampling}
@@
 \author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
-\author{\fnms{Cédric} \snm{Travelletti}\thanksref{t3}, \corref{} \ead[label=e1]{cedric.travelletti@stat.unibe.ch}}
-\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
-\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
+\author{\fnms{Cédric} \snm{Travelletti}\thanksref{t3}, \corref{} \ead[label=e2]{cedric.travelletti@stat.unibe.ch}}
+\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e3]{jo.eidsvik@ntnu.no}}
+\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e4]{david.ginsbourger@idiap.ch}}
 \and
-\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e4]{kanna.rajan@fe.up.pt}}
+\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e5]{kanna.rajan@fe.up.pt}}
 
@@
 \printead{e1}}
-\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e2}}
-\address{David Ginsbourger\\ Idiap Research Institute\\ Rue Marconi 19, 1920 Martigny\\ Switzerland\\
-\printead{e3}}
+\address{Cédric Travelletti\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
+Switzerland.
+\printead{e2}}
+\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e3}}
+\address{David Ginsbourger\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
+Switzerland.
+\printead{e4}}
 \address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
   Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
-\printead{e4}}
+\printead{e5}}
 
@@
 In this work we develop methods for efficient spatial sampling applied to the mapping of coastal processes by providing informative descriptions of spatial characteristics of ocean phenomena.
-Specifically, we define a design criterion based on reducing uncertainty in the joint excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected Bernoulli variance reduction in such a framework.
-We demonstrate how this criterion gives a description of the boundary uncertainty, that can be used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective.
+Specifically, we define a design criterion based on improved characterization of the uncertainty in the excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected Bernoulli variance reduction in such a framework.
+We demonstrate how this criterion can be used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective.
 We use simulations to study the properties of methods and to compare them with state-of-the-art approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a case study mapping the boundary of a river plume.
@@
 In medicine, doctors do not rely solely on a single symptom but must see several combined effects before making a diagnosis.
-In our context of environmental sampling applications, the salinity and temperature excursions of a river plume can similarly help characterize the underlying oceanographic and bio-geochemical processes \cite{hopkins2013detection,Pinto2018}, where sampling should be aimed at mapping relevant parts of the domain to reduce uncertainty in the temperature and salinity ES.
+In our context of environmental sampling applications, the salinity and temperature excursions of a river plume can similarly help characterize the underlying oceanographic and bio-geochemical processes \citep{hopkins2013detection,Pinto2018}, where sampling should be aimed at mapping relevant parts of the domain to reduce uncertainty in the temperature and salinity ES.
 Unlike what has been done in previous plume exploration studies, we emphasize the links to spatial statistical modeling and targeted multivariate sampling criteria for sequential sampling.
 
-The remainder of this paper is organized as follows: Section \ref{sec:bg} provides select background on ocean sampling; Section \ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV; Section \ref{sec:GP_EP} sets up the required statistical modeling assumptions of the bivariate GPs for temperature and salinity; Section \ref{sec:sur} builds on these assumptions when deriving the analytical derivation of the design criteria; Section \ref{sec:heuristics} extends these results and present heuristics for adaptive sampling; Section \ref{sec:simulations} discusses properties of the methods in simulation studies.
+The remainder of this paper is organized as follows: Section \ref{sec:bg} provides select background on ocean sampling; Section \ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for vector-valued GPs; Section \ref{sec:heuristics} builds on these assumptions when deriving the sequential design criteria for adaptive sampling; Section \ref{sec:simulations} discusses properties of the methods in simulation studies.
 Section \ref{sec:case_study} demonstrates the methodology used in field work characterizing a river plume; and Section \ref{sec:concl_disc} contains a summary and a discussion of future work.
@@
 
-The AUVs, see Fig.
-\ref{fig:lauv}, considered here operate at $1$-$3$ m/s in the upper water column, where they are free to move in six degrees of freedom (6 DOF).
+The AUVs considered here operate at $1$-$3$ m/s in the upper water column, where they are free to move in six degrees of freedom (6 DOF).
 The in-water operation time capacity depends on survey speed, payload sensors and navigation.
@@
 \ref{fig:nidelven} shows the survey area (Nidelva, Norway) where cold freshwater enters from the river, creating a strong gradient in both temperature and salinity.
-Because of the local topography and the Coreolis force \citep{coriolis1835memoire} the cold fresh water tends to flow near
+Because of the local topography and the Coreolis force %\citep{coriolis1835memoire}
+the cold fresh water tends to flow near
 land to the east.
@@
 In order to formalize the problem in mathematical terms,
-let us consider a multivariate (i.e., vector-valued) Gaussian random field $(\gp[\x])_{\x \in D}$, where $D$ is an arbitrary set (typically in geostatistical studies, a subset of interest of the ambient Euclidean space or of a sphere therein) and the dimensionality of the output $\no \geq 1$ is arbitrary.
+let us consider a multivariate (i.e., vector-valued) Gaussian random field $(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary set (typically in geostatistical studies, a subset of interest of the ambient Euclidean space or of a sphere therein) and the dimensionality of the output $\no \geq 1$ is arbitrary.
 Assume furthermore that the set of interest is characterized in terms of a subset of values of the variables, say $\T \subset \R^{\no}$, and denote the corresponding set of index points by
 $$
-\es=\gp^{-1}(\T)=\{\x \in D: \gp[\x] \in \T\}.
+\es=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
 $$
 Our goal here is to develop approaches to quantify and improve the characterization of uncertainties on such sets in the context of multivariate random fields.
+
 While some aspects of the developed
@@
 Note that changing some $\leq$ inequalities to $\geq$ ones would lead to immediate adaptations.
-We will keep the abstract notation $\T$ and it shall be clear from the context when specific properties following the orthant assumption come into play.
-Note that assuming that $\gp$ has continuous trajectories (almost surely) and $T$ to be closed, then $\es$ can be cast as a Random Closed Set \cite{Molchanov2005}.
-In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
-
-For our application, we let $t_1$ be a threshold for temperature and $t_2$ a threshold for
-salinity. The ES is then defined by $\mbox{ES} = \{\bx \in \mathcal{M} | \xi_T(\bx) \leq t_T,\xi_S(\bx) \leq t_S\}$.
-This ES holds random indicator variables while the associated EPs are $p(\bx) = P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S)$, $\bx \in \mathcal{M}$.
-If the EPs are
-close to $1$ or $0$ at a given location, it is easy to classify water mass at this point as either river or ocean. It is, however,
-challenging if the probability is close to $0.5$.
-
-Relate to ES and refer to 
-\cite{adler2000excursion,adler2009random}.
-
+In the application specific properties following the orthant assumption come into play: We let $t_1$ be a threshold for temperature and $t_2$ a threshold for salinity.
+The set of index point then defines $\mbox{ES} = \{\x \in D : \gp[\x,1] \leq t_1,\gp[\x,2]  \leq t_2\}$.
+The associated EPs are $p(\x) = P(\gp[\x,1] \leq t_1,\gp[\x,2]  \leq t_2)$, $\x \in \mathcal{M}$.
+
+For statistical generality, we will however keep the abstract notation with $\T$ and general $\es$ for most of the derivations.
+Note that assuming that $\gp$ has continuous trajectories (almost surely) and $T$ to be closed, then $\es$ can be cast as a Random Closed Set \citep{Molchanov2005,adler2009random}.
+%In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
 Let us in particular consider a locally finite (Borel) measure $\mes$ on $D$ and investigate the probability distribution of $\mes(\es)$ through its moments.
@@
 	%Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
-	For a measurable random field $\gp$ and a locally finite measure $\mes$ on $D$, $\mes(\es)$ is a random variable and for any $r\geq 1$, 
+	For a measurable random field $\gp$ and a locally finite measure $\mes$ on $\mathcal{M}$, $\mes(\es)$ is a random variable and for any $r\geq 1$, 
 	\begin{equation*}
@@
 	\mathbb{E}[\mes(\es)^r]
-	&=\int_{D^{r}} \jointExcuProb 
+	&=\int_{\mathcal{M}^{r}} \jointExcuProb 
 	\productMeasure
@@
 	%Note that we will use boldface to denote concatenated variables. % TO RELOCATE
-	Here $\gp$ is defined on $D$, and for
-	$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in D^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
+	Here $\gp$ is defined on $\mathcal{M}$, and for
+	$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in \mathcal{M}^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
 	\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.
@@
 
-The multivariate cumulative probabilities in two and four dimensions
-are computationally efficient using methods developed by
-\cite{genz2009computation}.
-
-Proposition~\ref{propo1} enables in turn calculating centred moments of $\mes(\es)$. 
+Proposition~\ref{propo1} enables one to calculate centered moments of $\mes(\es)$. 
 The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ is probably the most relevant of them for the considered purposes. We have
@@
 \left(
-(\bt, \bt); \mu((u,v)), 
-K((u,v),(u,v))
+(\bt, \bt); \bmu((u,v)), 
+\bK((u,v),(u,v))
 \right) 
@@
 (u,v)\\
-&-\left( \int_{D} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
+&-\left( \int_{D} \varPhi_{\no}\left(\bt;\bmu(u), \bK(u)\right) d\mes(u) \right)^2.
 \end{split}
 \end{equation*}
-%
-Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $D^2$.
-In contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves solely and integral on $D$ and can be expanded in our present settings as follows
-\begin{equation*}
+This quantity requires to work out an integral over $D^2$.
+In contrast, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves solely an integral on $\mathcal{M}$ and can be expanded in our present setting as follows
+\begin{equation}\label{IBV}
 \begin{split}
 \operatorname{IBV} %(\es) %;\mes)
-&=\int_{D}
+&=\int_{\mathcal{M}}
 \mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
 d\mes(u) \\
-&=\int_{D}
-\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) 
--\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
+&=\int_{\mathcal{M}}
+\varPhi_{\no}\left(\bt;\bmu(\uu), \bK(\uu)\right) 
+-\left(\varPhi_{\no}\left(\bt;\bmu(\uu), \bK(\uu)\right) \right)^2
 \mathrm{d}\mes(u). 
 \end{split}
-\end{equation*}
+\end{equation}
+
+\subsection{Expected Integrated Bernoulli Variance and Sampling Designs}
+
+With limited resources, the goal is to construct informative sampling designs.
+For our situation, data are gathered by an AUV that measure temperature and salinity at chosen locations on a defined waypoint graph covering the domain of interest.
+This discretized grid foster reliable navigation and mission performance.
+The waypoints further enable a structure to the design problem in the sense that sampling flexibility is limited to operation on the graph.
+This gives lots of sampling opportunities, and it is important to spend the sampling efforts wisely.
+We focus on sampling designs that reduce the expected integrated Bernoulli variance (EIBV).
+Note that this sampling can be done in a static manner where the survey design is pre-planned, or in an adaptive setting.
+For the latter case, data can be processed sequentially, and the updated model is then used to design and select the
+next waypoint graph to visit for this stage of data gathering, which are again used to update the model, and so on (Section
+\ref{sec:heuristics}).
+We next describe the data modeling assumptions for one stage and develop design criteria of semi-analytical forms relying on integrals of multivariate Gaussian cumulative distribution functions, we will appeal to some key intermediate results presented next.
+
+We let $d \in \mathcal{D}$ denote a sampling design, with $q$ sampling locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$. The measurement model is defined by
+\begin{equation}\label{eq:meas}
+    \bY_d=\bG_d \bz(\bu_d) + \epsilon(\bu_d), \hspace{3mm} \epsilon(\bu_d) \sim N(0,\bR_d).
+\end{equation}
+Here, $\bz(\bu_d)$ is the size $pq$ vector of GP variables at the sampling locations, $\bG_d$ is the design matrix which in principle allows for observing subsets of the vector-valued field at some of the $q$ locations (i.e.
+the matrix has fewer rows than columns).
+If the additive error has independent elements, the covariance matrix $\bR_d$ is diagonal.
+In our setting temperature and salinity measurements are extracted from different sensors, so the independence assumption between variables is not unrealistic, but this is not necessarily the case for other kinds of processed measurements on the AUV.
+It is further common in geostatistics to assume spatially independent measurement errors, and we also do so.
+But our methodology for constructing sampling designs does not rely on these independence assumptions.
+
+For a given design $d$ the evaluation is based on the expected information contained in the data to influence the criterion.  
+Considering the EIBV criteria, we have
+\begin{equation}\label{eq:eibv}
+    \mbox{EIBV} (d)= E_{\bY(\bu_d)} \int_{M}
+\mathbb{P}\left(\gp[\uu]\in T|\by(\bu_d) \right)(1-\mathbb{P}\left(\gp[\uu]\in T |\by(\bu_d)\right))
+d\mes(u),
+\end{equation}
+where the expectation is over the planned data.  The optimal design for an AUV survey is then
+\begin{equation}\label{crit}
+    d^* = \mbox{argmin}_{d \in \mathcal{D}} \mbox{EIBV}(d).
+\end{equation}
+For another criterion using random moments of random sets similar expectations would be required, but as we will see, the EIBV has a closed-form expression which is relatively fast to compute and it is hence useful for onboard design selection on AUVs.
+
+Assuming now that $q$ observations are available, one wishes to predict $\gp[\bu]$, the multivariate Kriging equations holds with size $p$ vector of conditional means: 
+\begin{equation}\label{eq:cokrig}
+\bbm(\x)=\bmu(\x)+\blambda(\x,\bu_d)^T \left( \by_d-\bG_d \bmu(\bu_d)\right),
+\end{equation}
+where $\bmu$ is size $p$ unconditional mean function of $Z$, $\by_d$ is the realized vector of observed responses, and $\blambda(\x,\bu_d)=\bK^{-1}_y(\bu_d,\bu_d) \bk(\bu_d, \bu)$.
+Here, the data covariance is $K_y=\mbox{Cov}(\bY_d,\bY_d)=\bG_d\bK(\bu_d,\bu_d)\bG_d^T+\bR_d$ and the cross-covariance is $\mbox{Cov}(\bY_d,Z(\x)) \bk(\bu_d, \x)=$, assuming the measurement error is independent of the GP.
+The Kriging (cross-)covariance function can also be expressed in the same vein via $p \times p$ matrix
+\begin{equation}\label{eq:krig:cov}
+\bSigma(\x,\x')=\bK(\x,\x')-\blambda(\x,\bu_d)^T \bK_y(\bu_d, \bu_d) \lambda(\x,\bu_d).
+\end{equation}
+
+Equations \eqref{eq:cokrig} and \eqref{eq:krig:cov} are used to provide informative, updated statements about sets of interest. 
+For our case with temperature and salinity, the conditional EPs, given measurements $\by_d$, can then be computed from 
+\begin{eqnarray}\label{eq:post_ep}
+ p(\x;\by) &=& P(Z_1(\x) \leq t_1, Z_{2}(\x) \leq t_2 |\by(\bu_d)),  \nonumber \\
+ &=& \Phi_2 ((t_1,t_2)^T-\bmu(\x)-\blambda(\x,\bu_d)^T \left( \by(\bu_d)-\bG_d \bmu(\bu_d)\right);0_2, \bSigma(\bu,\bu)). 
+\end{eqnarray}
+
+A critical element in the derivation of a closed-form for the EIBV is that the
+conditional mean in Eq. \eqref{eq:cokrig} is a linear (affine) function of
+the data $\by_d$ and the covariance is not a function of the data.
+This in turn means that the probabilities in expression \eqref{eq:post_ep} involve inequality statements for linear combinations of Gaussian variables.
+Related closed-form solutions have been noted in similar contexts \citep{bhattacharjya2013value, chevalier2014fast,stroh}, but not generalized to our situation with random sets for vector-valued GPs.
+Notably, the mean in equation \eqref{eq:cokrig}, is in the form of $\ba+\bB \bV$, where $\bV \sim N(0_q, \bK_y (\bu_d,\bu_d))$, $\ba=(t_1,t_2)^t-\bmu(\bu)-\blambda(\bu,\bu_d)^T \bG_d \bmu(\bu_d)$ and $\bB=\blambda(\bu,\bu_d)^T$.
+The following then applies to give a closed-form solution to EIBV calculation in this case.
 
@@
 	\begin{equation*}
-	\mathbb{E}\left[ \varPhi_{p}\left( a + BV; \covN \right)^h \right]
+	\mathbb{E}\left[ \varPhi_{p}\left( a + BV; 0_p,\covN \right)^h \right]
 	=
@@
 	\bm{a}
-	;~
+	;0_{ph},
 	\bm{\Sigma}
@@
 \end{propo}
-
-	In blockwise representation, $\bm{\Sigma}$ can be expressed as follows: 
+The proof of Proposition \ref{propo2} is provided in the Appendix.
+
+In block-wise representation, the following relation holds for the covariance in Proposition \ref{propo2}:
 \begin{align*}
-	% \bm{\Sigma}&=
+	\bm{\Sigma}&=
 	\begin{pmatrix}
@@
 	B\covV B' & \dots & B\covV B'\\
-	\end{pmatrix}
+	\end{pmatrix}.
 \end{align*}
-
-In the Appendix we also provide a more general derivation which holds for more general criterian than IBV, in partical for volumes of sets, meaning correlations between many locations meaning sums of indicator variables in Prop 1.
-
-\subsection{Sampling Designs and Expected Integrated Bernoulli Variance}
-
-Next we investigate how sampling designs are chosen to minimize the expected Integrated Bernoulli variance (EIBV).
-
-Data are gathered by an AUV that can measure temperature and salinity
-at a subset of locations on the defined grid. This can be done in a
-static manner (where the survey design is pre-planned), or in a
-sequential way with several stages of data gathering. For the latter,
-a stage of data is gathered first, and then assimilated to get an
-updated model. Then the updated model is used to design and select the
-next stage of measurements, and so on (Section
-\ref{sec:heuristics}). Below, we describe the model for one stage of
-data and the Gaussian formula for a single update.
-
-In order to be able to calculate them in semi-analytical forms relying on integrals of multivariate Gaussian cumulative distribution functions, we will appeal to some key intermediate results presented next.
-
-(MODIFY NOTATION HERE - BUT NEED THIS $y$ IN HERE)
-
-We denote the data at $n_y$ locations by
-$\by=(y_{T,1},y_{S,1},\ldots,y_{T,n_y},y_{S,n_y})^t$, representing the
-temperature and salinity measurements gathered at the first stage of
-nodes in the grid. The conditional model for the data, given the true temperature
-and salinity, is described by: 
-\begin{equation}\label{likelihood}
-\by | \bxi \sim N( \bG \bxi, \bR), %\pi(\by | \bxi)= \frac{1}{(2\pi)^{m}|\bR|^{\frac{1}{2}}}e^{-\frac{1}{2}(\by -\bG\bxi)^T\bR^{-1}(\by-\bG\bxi)},
-\end{equation}
-where $\bG$ is a $2n_y \times 2n$ matrix containing '$1$'s on the
-measurement indices of the grid and '$0$' otherwise. The size
-$2n_y \times 2n_y$ covariance matrix $\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations.
-
-Data can be
-gathered at selected locations of the domain $\mathcal{M}$, to improve
-classification performance. We denote temperature measurements by
-$y_{T}(\bx_i)$ and salinity measurements by $y_{S}(\bx_i)$ at a
-sampling location $\bx_i \in \mathcal{M}$. Given a set of
-measurements denoted $\by$, the conditional EPs are:
-\begin{equation}\label{eq:post_ep}
- p(\bx;\by) = P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\by). 
-\end{equation}
-
-Our goal is to construct adaptive sampling strategies based on
-continuous evaluation of the EPs. As the EPs provide a description of
-the boundary uncertainty, the strategies can effectively use this
-description to prioritize sampling efforts at locations that are
-ambiguous, making exploration more effective. It is natural to
-evaluate the expected reduction in the IBV, conditioned on the survey
-data. The EIBV is:
-
-(MODIFY NOTATION HERE)
-
-\begin{equation}\label{sur}
-    V_{\mbox{upd}} = \int_{\bx \in \mathcal{M}} E_{\by} \left( p(\bx;\by)\left( 1-p(\bx;\by)\right) \right) d\bx, 
-\end{equation}
-where the expectation is over the planned data, and the probability is
-defined in Eq. (\ref{eq:post_ep}).  Considering a set $\mathcal{J}$ of
-possible designs for an AUV survey, a criterion for the selection is:
-\begin{equation}\label{crit}
-    j^* = \mbox{argmin}_{j \in \mathcal{J}} V_{\mbox{upd}}(j),
-\end{equation}
-where the criterion $V_{\mbox{upd}}(j)$ in Eq. (\ref{sur}) is computed
-for each of the possible designs.
-
-REUSE PROPOSITION RESULTS
-
-%Recent work on ESs and EPs connected to spatial statistics include
-%\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
-%Our focus is on ES as defined in Eq. (\ref{ES}) and the uncertainty
-%reduction achieved by sampling as in Eq. (\ref{sur}). In this sense,
-%our work is similar to \cite{bect2012}, \cite{chevalier2014fast}, and
-%\cite{azzimonti2016quantifying} who describe analytical results for
-%the IBV for univariate processes. A major contribution of our work is
-%to derive closed-form results for this design criteria for situations
-%where the underlying model is based on multivariate GPs.
-
-Calculations of the EPs and the IBV require a model specification. We
-use GP models, which provide a practical and popular probabilistic
-approach to modeling environmental data, and allows a closed form
-evaluation of the IBVs as derived in Section \ref{sec:sur}. GPs can
-also be fitted to run onboard robotic vehicles with limited
-computational assets, such as AUVs.
-
-(MODIFY NOTATION HERE)
-
-The GP mean and covariance in Eq. (\ref{prior}) are updated in the
-model when more information is available. The updated distribution for
-temperature and salinity variables on the spatial grid, given survey
-data $\by$, is Gaussian with mean and covariance:
-%\begin{eqnarray}\label{gp_upd}
-%  \bm &=& \bm(\by) = \bmu+\bSigma \bG^t (\bG \Sigma \bG^t+\bR)^{-1}(\by-\bG \bmu),  %\\
-%  \bA &=& \bSigma - \bSigma \bG^t (\bG \bSigma \bG^t+\bR)^{-1} \bG
-%          \bSigma.\nonumber
-%\end{eqnarray}
-
-The updated bivariate distribution at a grid location $\bx \in
-\mathcal{M}_g$ is then: 
-\begin{equation}\label{gp_hat}
-\begin{bmatrix}
-\xi_T(\bx) \\
-\xi_S(\bx)
-\end{bmatrix}
- |\by
- \sim N \left( 
-\begin{bmatrix} m_{T}(\bx)\\
-m_{S}(\bx)
-\end{bmatrix},\begin{bmatrix}
-a_{T}^2(\bx) & a_{T,S}(\bx)  \\
-a_{T,S}(\bx)  & a_{S}^2(\bx)  
-\end{bmatrix}
+The calculation of the multivariate Gaussian cumulative distribution is done effectively using code such as that of \cite{genz2009computation}.
+
+In the more general case, for instance that of volumes of random sets, the criterion will include couplings  with different locations and the result must be generalized.
+For the case of multivariate monomials in orthant probabilities with thresholds affine in a common Gaussian vector.
+%Namely, for monomials of degree $k$, we have.
+
+\begin{propo}%[\textcolor{red}{Notation in progress, to be reviewed/revised}]
+	\label{propo3}
+	Let $g, p, q\geq 1$, $h_{1},\dots, h_{g}\geq 1$ with $H=\sum_{i=1}^g h_i$, $a_{i} \in \R^{p}$, $B_{i}\in \R^{p \times q}$, and covariance matrices $\covN_i \in \R^{p \times p}$ $(1\leq i \leq g)$.
+Then, for any covariance matrix $\covV \in \R^{q\times q}$ and $V\sim\mathcal{N}_{q}(0_q,\covV)$,
+	%	Let $g$ and $k$ be integers, and consider a $g$-terms monomial of degree $k$
+	%	\[
+	%	    \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \Sigma_{i}
+	%	    \right)^{h_i},
+	%    \]
+	%    where $\sum_{i=1}^g h_i = k$, each $a_i$, $B_i$ and $\Sigma_i$ is as in
+	%    \cref{propo2} and $V$ is a $q$-dimensional Gaussian vector distributed as
+	%    $V\sim\mathcal{N}(0,\covV)$.
+	%
+	% Then the expectation of the monomial is given by
+	\begin{equation}
+	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V;0_p, \covN_{i} \right)^{h_i} \right]
+	=
+	\varPhi_{p H}
+	\left(
+	\bm{a}
+	;0_{pH},
+	\mathbf{\Sigma}
 	\right),
 	\end{equation}
-with entries extracted from the conditional mean and covariance
-expressions in Eq. (\ref{gp_upd}).
-
-% The conditional mean $\bm$ will be important in the following
-% derivation in Section \ref{sec:sur}.
-The calculation of the design criterion in Eq. (\ref{sur}) requires
-the marginal distribution of the data which is defined over $\bxi$ to
-be $\by \sim N( \bG \bmu,\bG \bSigma \bG^t+\bR)$.  However, it turns
-out that a simplified form of the calculation is possible because the
-conditional mean in Eq. (\ref{gp_hat}) is a linear (affine) function of
-the data $\by$ and the covariance is not a function of the data.
-Before inputting the data, the distribution of the conditional mean is:
-\begin{equation}\label{distxi} \bm \sim N(\bmu , \bPsi), \hspace{3mm}
-  \bPsi=\bSigma \bG^t (\bG \bSigma \bG^t+\bR)^{-1} \bG \bSigma.
+	with $\bm{a}=(1_{h_1}\otimes a_1, \dots, 1_{h_g}\otimes a_{g}) \in \R^{p H}$
+	and $\mathbf{\Sigma}\in \R^{p H \times p H}$ is defined blockwise by $(\Sigma_{i,j})_{i,j \in \{1,\dots, g\}}$ where, for any $i,j \in \{1,\dots, g\}$, %the blocks are given by 
+	\begin{equation}
+	\Sigma_{i,j}=
+	(1_{h_{i}}1_{h_{j}}')\otimes (B_{i}\Sigma_{V} B_{j}') + \delta_{i,j}(I_{h_{i}}\otimes \covN_{i}) \in \R^{p h_{i} \times p h_{j}}.
 	\end{equation}
-When considering only grid location $\bx$, we can
-extract elements from its mean vector and covariance matrix to get the
-bivariate distribution:
-\begin{equation}\label{dist_mxi}
-  \bm(\bx) \sim N \left( \begin{bmatrix}
-      \mu_{T}(\bx) \\
-      \mu_{S}(\bx) \end{bmatrix}, \begin{bmatrix}
-      \psi^2_{T}(\bx) & \psi_{T,S}(\bx)\\
-      \psi_{T,S}(\bx) & \psi^2_{S}(\bx) \end{bmatrix} \right),
-\end{equation}
-which is important for the derivation of the closed-form solution of the design criteria.
-
-Based on Gaussian modeling assumptions, the EPs can be computed at any
-location $\bx$. Without additional data, Eq.  (\ref{eq:prob}) gives:
-\begin{eqnarray}\label{eq:prob_mv0}
- p(\bx) &=& P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S) \nonumber \\
- &=& \Phi_2 \begin{pmatrix} 
-\begin{bmatrix} t_T\\
-t_S
-\end{bmatrix};
-\begin{bmatrix} \mu_{T}(\bx)\\
-\mu_{S}(\bx)
-\end{bmatrix},\begin{bmatrix}
-\sigma_{T}^2(\bx) & \sigma_{T}(\bx)\sigma_{S}(\bx) \gamma(\bx)  \\
-\sigma_{T}(\bx)\sigma_{S}(\bx) \gamma(\bx)  & \sigma_{S}^2(\bx)  
-\end{bmatrix}\end{pmatrix},
-\end{eqnarray}
-where $\Phi_2$ is the bivariate Gaussian cumulative distribution
-function, and the mean and covariance entries are as defined in
-Eq. (\ref{gp_bar}).
-
-When data $\by$ are available, Eq. (\ref{eq:post_ep}) similarly gives:
-\begin{eqnarray}\label{eq:prob_mv}
- p(\bx;\by) &=& P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\by)
- \nonumber \\
- &=& \Phi_2 \begin{pmatrix} 
-\begin{bmatrix} t_T\\
-t_S
-\end{bmatrix};
-\begin{bmatrix} m_{T}(\bx)\\
-m_{S}(\bx)
-\end{bmatrix},\begin{bmatrix}
-a_{T}^2(\bx) & a_{T,S}(\bx)  \\
-a_{T,S}(\bx)  & a_{S}^2(\bx)  
-\end{bmatrix}\end{pmatrix},
-\end{eqnarray}
-with parameters as defined in Eq. (\ref{gp_upd}) and
-(\ref{gp_hat}). Here, the probabilities would not only depend on the
-data but also on the data design defined via matrix $\bG$ in
-Eq. (\ref{likelihood}). We next derive results that simplify the
-comparison of several designs.
-
-MODIFY AND REUSE PROPOSITIONS
-
-A critical first element in the derivation is using the GP model and the
-linear dependence on $\by$ in the conditional mean in
-Eq. (\ref{gp_upd}). This results in a closed form Gaussian
-distribution for the conditional mean in Eq. (\ref{dist_mxi}). The
-high-dimensional inner integral in Eq. (\ref{sur}) then reduces to a
-bivariate integral \citep{bhattacharjya2013value, chevalier2014fast},
-with an expectation over $\bm=(m_{T},m_{S})^t$.  We standardize the two
-variables in Eq. (\ref{sur}) as   $Z_T=\frac{\xi_T-m_{T}}{a_{T}}$,
-$Z_S=\frac{\xi_S-m_{S}}{a_{S}}$, with
-$\mbox{Corr}(Z_T,Z_S)=a_{T,S}/(a_{T} a_{S})=\eta_{T,S}$, and we can
-then rephrase the probability as: 
-\begin{eqnarray}\label{ytom}
-   P(\xi_T \leq t_T, \xi_S \leq t_S|\by) &=& P \left( Z_T \leq\frac{t_T-m_{T}}{a_{T}}, Z_S \leq\frac{t_S-m_{S}}{a_{S}}|\by \right) \nonumber \\
-   &=& \Phi_2 \begin{pmatrix} 
-\begin{bmatrix} \frac{t_T-m_{T}}{a_{T}}\\
-\frac{t_S-m_{S}}{a_{S}}
-\end{bmatrix};
- \begin{bmatrix} 0\\
-0
-\end{bmatrix},\begin{bmatrix}
-1 & \eta_{s,t}  \\
-\eta_{s,t}   & 1  
-\end{bmatrix}\end{pmatrix} \nonumber \\
-&=& P(\xi_T \leq t_T, \xi_S \leq t_S|\bm).
-\end{eqnarray}
-Since the expression only depends on $\by$ via its affine function
-$\bm$, the expectation in Eq. (\ref{sur}) is reduced to a bivariate
-integral over $\bm$.  Introducing $\pi(\bm)$ for the density function
-of $\bm$, the expected Bernoulli variance becomes:
-%\begin{eqnarray}
-%\label{eq:var2}
-%E_{\by}(p(1-p)) &=& E_{\bm}(p(1-p)), \hspace{3mm} p=P(\xi_T \leq t_T, \xi_S \leq %t_S|\bm) \nonumber \\
-%E_{\bm}(p(1-p)) & = & \int p(1-p) \pi(\bm) d\bm, \nonumber \\
-% &=& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm)  \pi(\bm) d\bm \nonumber  \\
-%&-& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) %\pi(\bm) d\bm. 
-%\end{eqnarray}
-We will next extend results of \cite{chevalier2014fast} to solve for
-these two parts (see also \cite{stroh}).
-
-For the first part of the integral above, we have:
-\begin{equation}
-\label{part1:phi2}
- \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) \pi(\bm) d\bm= 
-P \left( Z_{T} \leq \frac{t_T-m_{T}}{a_{T}}, 
-Z_{S} \leq \frac{t_S-m_{S}}{a_{S}} \right). \nonumber
-\end{equation}
-The standardized variable $\bZ=(Z_{T},Z_{S})$ is chosen to be independent of $\bm$. Using Eq. (\ref{dist_mxi}), we have:
-\begin{equation}
-    E(a_{T} Z_{T}+m_{T}-t_T) = \mu_{T}-t_T, \hspace{3mm}
-    \mbox{Var}(a_{T} Z_{T}+m_{T}-t_T) = a^2_T+ \psi^2_{T} 
-\end{equation}
-for the temperature part; the same holds for salinity. From this we
-get:
-\begin{eqnarray}\label{two_parts0}
-& P & \left( Z_{T} \leq \frac{t_T-m_{T}}{a_{T}}, 
-Z_{S} \leq \frac{t_S-m_{S}}{a_{S}} \right) \\
-&=& \Phi_2 \begin{pmatrix} 
-\begin{bmatrix} 0\\
-0
-\end{bmatrix};
-\begin{bmatrix} \mu_{T}-t_T\\
-\mu_{S}-t_S
-\end{bmatrix},\begin{bmatrix}
-a^2_T+ \psi^2_{T} & a_{T,S}+\psi_{T,S}  \\
-a_{T,S}+\psi_{T,S}   & a^2_T+ \psi^2_{T} 
-\end{bmatrix}\end{pmatrix} \nonumber.
-\end{eqnarray}
-
-The equations derived in the previous section can then be generalized. Let
-$\bxi(\bx)=(\xi_1(\bx),\ldots,\xi_K(\bx))$, and the expected Bernoulli
-variance is:
-\begin{equation}\label{two_partsK}
-E_{\by}(p(\bx;\by)(1-p(\bx;\by))), \hspace{3mm} p(\bx;\by)=P(\xi_k(\bx) \leq t_k ; k=1,\ldots,K|\by). 
-\end{equation}
-From the first key result of linear combination of Gaussian variables,
-we reduce the integral to a $K$ dimensional integral of the relevant
-$\bm=E(\bxi(\bx)|\by)$. Next, reducing the integral to two parts, we
-standardize the variables and compute cumulative distributions.  In
-summary, we get:
-\begin{equation}\label{two_parts}
-E_{\by}(p(\bx)(1-p(\bx))) =  \Phi_K 
-\begin{pmatrix}
-\begin{bmatrix} t_1\\
-\vdots \\
-t_K 
-\end{bmatrix};
-\bmu,\bA+\bPsi 
-\end{pmatrix}
-- \Phi_{2K} 
-\begin{pmatrix}
-\begin{bmatrix} t_1\\
-\vdots \\
-t_K \\
-t_1\\
-\vdots \\
-t_K 
-\end{bmatrix};
-E(\bV),Var(\bV) 
-\end{pmatrix},\nonumber
-\end{equation}
-where $\bA$ and $\bPsi$ are direct $K$-variate extensions of the
-matrices in Eq. (\ref{two_parts0}), and the matrix $\bV$ finds the
-relevant linear combination of two independent vectors
-$\bZ_l=(Z_{l,1},\ldots,Z_{l,K})$, $l=1,2$ and $\bm$, as an extension
-of Eq. (\ref{Vcomb}).
+	%with $\bm{a}=\sum_{j=1}^g \mathbf{1}_{J_{j}} \otimes a_j \in \R^{p \times k}$ 
+	%where $J_{j}=\{\sum_{i=1}^{j-1}h_i + 1, \dots, \sum_{i=1}^{j}h_i\}$,
+	%$\mathbf{1}_{I_{j}}$ denotes the vector of $\R^{k}$ with components equal to $1$ for indices in $I_{j}$ and $0$ else ($1\leq j \leq g$), 
+	%and $\mathbf{\Sigma}\in \R^{p k \times p k}$ is defined blockwise by 
+	%\begin{equation}
+	%\mathbf{\Sigma}_{\widetilde{J}_{j_{1}},\widetilde{J}_{j_{2}}}=
+	%(\mathbf{1}_{h_{j_1}}\mathbf{1}_{h_{j_2}}')\otimes (B_{j_{1}}\Sigma_{V} B_{j_2}') + \delta_{j_{1},j_{2}}(I_{j_{1}}\otimes \Sigma_{j_{1}}),
+	%\end{equation}
+	%where $\widetilde{J}_{j}=\{p\sum_{i=1}^{j-1}h_i + 1, \dots, p\sum_{i=1}^{j}h_i\}$ ($1\leq j \leq g$).
+\end{propo}
+The proof of Proposition \ref{propo2} is provided in the Appendix.
 
@@
 \end{figure}
-Here, the thresholds are set equal to the mean $\mu_T=t_T=5^o C$, and
-salinity $\mu_S=t_S=30$ mg/l. The displayed densities have unit
+Here, the thresholds are set equal to the mean $\mu_1=t_1=5^o C$, and
+salinity $\mu_2=t_2=30$ mg/l, and the EP is $p=P(Z_1 < 5,Z_2 < 30)$. The displayed densities have unit
 standard deviations for both temperature and salinity, but we also
@@
     Bernoulli variances for both temperature and salinity data $\by$ and 
-    temperature $y_T$ (bottom rows).}
+    temperature $y_2$ (bottom rows).}
   \begin{tabular}{c|ccc|ccc}
- &\multicolumn{3}{c}{$\sigma_S=\sigma_T=1$} & \multicolumn{3}{c}{$\sigma_S=\sigma_T=2$} \\
+ &\multicolumn{3}{c}{$\sigma_1=\sigma_2=1$} & \multicolumn{3}{c}{$\sigma_1=\sigma_2=2$} \\
 \hline
@@
 $p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
-$E_{\by}(p(\by) (1-p(\by)))$ & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
-$E_{y_T}(p(y_T) (1-p(y_T)))$ & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
+EIBV, Temp and Salinity data & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
+EIBV, Temperature data only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
 \hline
@@
 Bernoulli variance calculations. This is presented for a design
-gathering both data types $(y_T,y_S)$, and for a design with
-temperature measurements $y_T$ alone. With both data,
-$(y_T,y_S)^t=(\xi_T,\xi_S)^t+N(0,0.5^2\bI)$, while
-$y_T=\xi_T+N(0,0.5^2)$ when only temperature is measured.  For this
+gathering both data types, and for a design with
+temperature measurements alone. Both data are modeled by $(Y_1,Y_2)^t=(Z_1,Z_2)^t+N(0,0.5^2\bI)$, while
+$Y_2=Z_2+N(0,0.5^2)$ when only temperature is measured.  For this
 illustration, Table \ref{tab:sim_rhoab} shows that the expected
 Bernoulli variance gets lower with larger standard deviations
-$\sigma_T$ and $\sigma_S$ (right columns). The reduction of Bernoulli
+$\sigma_1$ and $\sigma_2$ (right columns). The reduction of Bernoulli
 variance is largest for the cases with high correlation
@@
 impact the design criteria when we evaluate the information measure by
-integrating over several locations (Section \ref{sec:case_study}).
+integrating over several locations (Section \ref{sec:simulations}).
 
@@
 %\begin{equation}
-%E(\min \{ p,(1-p)\})=\int \min P(\xi_T \leq t, \xi_S \leq s),[1-P(\xi_T \leq t, \xi_S \leq s)] p(E(\bxi)=) dE(\bxi)=,
+%E(\min \{ p,(1-p)\})=\int \min P(\xi_1 \leq t, \xi_2 \leq s),[1-P(\xi_1 \leq t, \xi_2 \leq s)] p(E(\bxi)=) dE(\bxi)=,
 %\end{equation}
@@
 
-\section{Sequential updating and heuristic path planning}\label{sec:heuristics}
+\section{Sequential designs and heuristic path planning}\label{sec:heuristics}
 
@@
 sampling to adapt survey plans based on data and statistical methods
-in accordance with the sense-plan-act control loop (Fig. \ref{fig:envir1}). We
-will therefore present a few adaptive strategies for robotic sampling
-that build on the notion of multivariate ES.
+in accordance with the sense-plan-act control loop (Fig. \ref{fig:envir1}). We now present a few adaptive strategies for robotic sampling
+that build on the notion of EIBV.
 
@@
 
-An adaptive AUV survey is split into many stages. The selection --on where to sample-- is made on the graph and defined by the nearest grid nodes in the domain $\mathcal{M}_g$.
-
-The optimal sequential design not only considers the best current AUV
-grid node, but also how the data gathered at this node could inform
-future sampling at successive stages. Let $d^{j,s}$ denote design
+An adaptive AUV survey is split into many stages. The selection on where to sample is made with restriction to the waypoint graph and defined by the nearest grid nodes in the domain $\mathcal{M}_g$.
+
+Let $d^{j,s}$ denote design
 number $j$ at stage $s$ of a sequential survey. If this design is
-selected, data $\by^{j,s}$ will be gathered. The optimal path
-selection situation is depicted in Fig. \ref{fig:PathSelOpt},
-\begin{figure}[b!]
-\centering
-\includegraphics[width=0.85\textwidth]{Figures/sequent_select.pdf}
-\caption{Optimal sequential path selection.}\label{fig:PathSelOpt}
-\end{figure}
-where design choices are indicated by squares while data realizations are indicated by circles. 
+selected, data $\by_{d^{j,s}}$ will be gathered. 
+We further let
+$\mathcal{Y}^{s-1} =\{(d^{j,1}, \by_{d^{j,1}}),\ldots, (d^{j,s}, \by_{d^{j,s}}) \}$ denote
+the set of all data gathered until stage $s-1$ according to selected design $d^{j,1},\ldots,d^{j,s}$. 
+This defines the AUV position at the stage $s$. 
+The data further defines the conditional GP  with mean $\bbm^{s-1}(\x)=E(\gp[u]|\mathcal{Y}^{s-1})$ from
+Eq. (\ref{eq:cokrig})
+and (cross) covariance matrix $\bSigma^{s-1}=\mbox{Var}(\gp[u]|\mathcal{Y}^{s-1})$ from Eq. (\ref{eq:krig:cov}).
+
+At the next stage, we choose where measurements $\by_{d^{j,s}}$ should be gathered with
+design $d^{j,s}$. For the triangularized waypoint graph that we use in our AUV application, there are six possible directions for $j$, and we must evaluate the design criterion for all these. 
+The goal of the sampling design is to minimize the expected uncertainty of random sets, and specifically the EIBV for our application.
+We hence focus on choosing designs for data modeled by equation \eqref{eq:meas}, to conduct efficient sequential uncertainty reduction according to the EIBV criterion.
+Notably, compared with the static design $d$ in equation \eqref{eq:eibv}, the optimal sequential design not only considers the best current AUV
+design grid point, but also how the data gathered at this node could inform future sampling at successive stages.
+
+%The optimal path
+%selection situation is depicted in Fig. %\ref{fig:PathSelOpt},
+%\begin{figure}[b!]
+%\centering
+%\includegraphics[width=0.85\textwidth]{Figures/sequent_select.pdf}
+%\caption{Optimal sequential path selection.}\label{fig:PathSelOpt}
+%\end{figure}
+%where design choices are indicated by squares while %data realizations are indicated by circles. 
 %terms this optimal design is defined as
@@
 %where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages. 
-The mathematical expression for the optimal design then involves a
-series of intermixed maximisations over designs and integrals over
+The mathematical expression for the optimal design involves a
+series of intermixed maximizations over designs and integrals over
 data. In practice, the optimal solution is intractable because of
@@
 e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}).
-Instead, we outline heuristic strategies. In terms of notation, let
-$\mathcal{Y}^{s-1} =\{d^{j,r}, \by^{j,r} ; r=1,\ldots,s-1 \}$ denote
-the data gathered until stage $s-1$, using a selected design
-$d^{j,r}$, $r=1,\ldots,s-1$.
-
-%The derived closed-form solutions for the expected IBV are still important building blocks when constructing the sampling designs, as they will be used to score the different adaptive designs.
-% Efficient
+Instead, we outline heuristic strategies.
+
+%The derived closed-form solutions for the EIBV are still important building blocks when constructing the sampling designs, as they will be used to score the different adaptive designs. % Efficient
 % calculation is important for using adaptive survey designs with
@@
 The simplest heuristic for adaptive sampling is to choose the next
-sampling location based on current EPs. The largest variance in
-sampling locations indicates higher levels of uncertainty; therefore
-one selects the design location with an EP closest to $0.5$.
-
-At stage $s$, based on the currently available data
-$\mathcal{Y}^{s-1}$, we fit an updated Gaussian model from
-Eq. (\ref{gp_upd}) with mean $\bm^{s-1}=E(\bxi|\mathcal{Y}^{s-1})$
-and covariance matrix $\bA^{s-1}=\mbox{Var}(\bxi|\mathcal{Y}^{s-1})$.
-The next stage of measurements $\by^{j,s}$ can be gathered with
-design $d^{j,s}$, $j=1,\ldots,J$. The {\it{naive}} strategy then
+sampling location based on current probabilities. One reduces the largest levels of uncertainty by selecting the design $j$ with $P(Z(\x_j) \in T)$ closest to $0.5$.
+For our situation with temperature and salinity this {\it{naive}} strategy then
 selects the design according to:
 \begin{eqnarray}\label{critNaive}
-    d^{*,s} &=& \mbox{argmin}_{j \in \{1,\ldots,J\}} |p(\bx_{d^{j,s}};\mathcal{Y}^{s-1})-0.5|, \\
-    p(\bx;\mathcal{Y}^{s-1}) &=& P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S | \mathcal{Y}^{s-1}). \nonumber
+    d^{*,s} &=& \mbox{argmin}_{j} |P(Z_1(\x_j) \leq t_1, Z_2(\x_j) \leq t_2 | \mathcal{Y}^{s-1})-0.5|. \nonumber
 \end{eqnarray}
-This strategy does not account for the uncertainty in the temperature
-or salinity variables, but is instead applicable only if one design (node) has EPs closer to
+
+This strategy does not account for the uncertainty in the GP variables, but is instead applicable only if one design (node) has EPs closer to
 $0.5$ (see Table \ref{tab:sim_rhoab}, line two). Additionally, this strategy does not
@@
 The myopic (greedy) strategy which we present here is optimal if we
-imagine taking only one more stage of measurements. In this selection
-strategy, there is no anticipation of what the subsequent designs might
+imagine taking only one more stage of measurements. This selection
+strategy is based on expected uncertainty reduction, but it is heuristic because there is no anticipation of what the subsequent designs might
 offer beyond the first stage.
@@
 Based on the currently available data $\mathcal{Y}^{s-1}$, we
-fit an updated GP model.  The next stage of measurements
-$\by^{j,s}$, can be gathered with design $d^{j,s}$, $j=1,\ldots,J$. The
-selected design is then:
+fit an updated GP model. Focusing on the EIBV criterion, at the next stage $s$, the selected design is:
 \begin{eqnarray}\label{critSEQ}
-    d^{*,s} &=& \mbox{argmin}_{j \in \{1,\ldots,J\}} \left\{ V_{m,\mbox{upd,j}} \right\},  \\
-V_{m,\mbox{upd}} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j,s}|\mathcal{Y}^{s-1}} \left( p(\bx;\mathcal{Y}^{j,s})\left( 1-p(\bx;\mathcal{Y}^{j,s})\right) \right) \Delta, \nonumber \\
-    p(\bx;\mathcal{Y}^{j,s}) &=& P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\by^{j,s},\mathcal{Y}^{s-1}). \nonumber
+    d^{*,s} &=& \mbox{argmin}_{j} \left\{ \sum_{\x \in \mathcal{M}_g} E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( p(\x;\mathcal{Y}^{j,s})\left( 1-p(\x;\mathcal{Y}^{j,s})\right) \right) \Delta, \right\} \\
+    p(\x;\mathcal{Y}^{j,s}) &=& P(Z_u \in T |\by_{d^{j,s}},\mathcal{Y}^{s-1}). \nonumber
 \end{eqnarray}
 
-Note that this strategy gives a sequential conditional version of the
-formula in Eq. (\ref{sur}). Now $\mathcal{Y}^{s-1}$ is available, and
-the expectation is with respect to the conditional density
-$\pi(\by^{j,s}|\mathcal{Y}^{s-1})$. A similar closed-form calculation
-for expected IBV is hence applicable in Eq. (\ref{critSEQ}) using the
+Note that this strategy gives a sequential conditional version of
+expression (\ref{eq:eibv}). Now $\mathcal{Y}^{s-1}$ is available, and
+the expectation is conditional to this data. A similar closed-form calculation
+for EIBV is hence applicable in Eq. (\ref{critSEQ}) using the
 updated GP model from stage $s-1$. Once the data is collected for the
-best design, the GP model is updated again. The mean $\bm^{s}$ and
-covariance matrix $\bA^{s}$ are used to compute the next design at
+best design at stage $s$, the GP model is updated again. The mean $\bbm^{s}$ and
+covariance matrix $\bSigma^{s}$ are used to compute the next design at
 stage $s+1$, and so on.
@@
 using an efficient approach for data updating the GP model and the
-calculation of the closed-form expected IBV expressions for each
+calculation of the closed-form EIBV expressions for each
 subsequent trajectory.
@@
 \begin{eqnarray}\label{critLA}
-    d^{*,s} &=& \mbox{argmin}_{j \in \{1,\ldots,J\}} \left\{ U_{la,\mbox{upd,j}} \right\},  \\
-    U_{la,\mbox{upd},j} & = &  E_{\by^{j,s}|\mathcal{Y}^{s-1}} \left\{ \mbox{argmin}_{j_2 \in \{1,\ldots,{J}_2\}} \left[ V_{la,\mbox{upd},j_2} \right] \right\}, \nonumber \\
-V_{la,\mbox{upd},j_2} & \approx & \sum_{\bx \in \mathcal{M}_g} E_{\by^{j_2,s+1}|\mathcal{Y}^{j,s}} \left\{ p(\bx;\mathcal{Y}^{j,s+1})\left( 1-p(\bx;\mathcal{Y}^{j,s+1})\right) \right\} \Delta, \nonumber \\
-    p(\bx;\mathcal{Y}^{j,s+1}) &=& P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S |\by^{j_2,s+1},\mathcal{Y}^{j,s}). \nonumber
+    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left[ \mbox{argmin}_{j_2}  V_{j_2} \right] \right\}, \nonumber \\
+V_{j_2} & = & \sum_{\x \in \mathcal{M}_g} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1})\left( 1-p(\x;\mathcal{Y}^{j,s+1})\right) \right) \Delta, \nonumber \\
+    p(\x;\mathcal{Y}^{j,s+1}) &=& P(Z(\x) \in T |\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}). \nonumber
 \end{eqnarray}
-Here, $\mathcal{Y}^{j,s}=\{\by^{j,s},\mathcal{Y}^{s-1}\}$ and
-$\mathcal{Y}^{j,s+1}=\{\by^{j_2,s+1},\mathcal{Y}^{j,s}\}$ represent
+Here, $\mathcal{Y}^{j,s}=\{\by_{d^{j,s}},\mathcal{Y}^{s-1}\}$ and
+$\mathcal{Y}^{j,s+1}=\{\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}\}$ represent
 the sets of data variables, and the expectations are calculated with respect to
-the conditional densities $\pi(\by^{j,s}|\mathcal{Y}^{s-1})$ and
-$\pi(\by^{j_2,s+1}|\by^{j,s},\mathcal{Y}^{s-1})$.
-
+the conditional densities given data at that stage, meaning we have two levels of intermixed minimizations and integrals.
 We solve the
-first expectation by Monte Carlo sampling of data $\by^{j,s}$ from its
+first expectation by Monte Carlo sampling of data $\by_{d^{j,s}}$ from its
 conditional distribution. For each of these data samples, the second expectation
-is solved using the closed-form expressions for expected IBV from
-Section \ref{sec:sur}.
+is solved using the closed-form expressions for EIBV from
+Section \ref{sec:ESEP}.
 
@@
 the current best design. When data are collected, the GP model is
-updated, and the mean $\bm^{s}$ and covariance matrix $\bA^{s}$ are
+updated, and the mean $\bbm^{s}$ and covariance matrix $\bSigma^{s}$ are
 used to compute the next design at stage $s+1$, now anticipating what
 stage $s+2$ could offer, and so on.
-
 This look-ahead approach is much more computationally demanding than
 the myopic strategy, so for practical implementation, we prune paths
-in the evaluation of Eq. (\ref{critLA}). This means that we do not
-compute all possible branches of the first two stages, as they are
-indicated in Fig. \ref{fig:PathSelOpt}. Instead, we use the myopic
+in the evaluation of Eq. (\ref{critLA}). This means that we use the myopic
 strategy to rank the three best designs following the first stage alone, and
@@
 
-The Gaussian model is used to represent the bivariate random fields of
-temperature and salinity. At location $\bx \in \mathcal{M}$, the
+A GP model is used to represent the bivariate random fields of
+temperature and salinity. At location $\x \in \mathcal{M}$, the
 bivariate distribution is:
 \begin{equation}\label{gp_bar}
-\begin{bmatrix}\xi_T(\bx) \\
-\xi_S(\bx) \end{bmatrix}
+\begin{bmatrix}Z_1(\x) \\
+Z_2(\x) \end{bmatrix}
 \sim N \left( 
-\begin{bmatrix} \mu_{T}(\bx)\\
-\mu_{S}(\bx)
+\begin{bmatrix} \mu_{1}(\x)\\
+\mu_{2}(\x)
 \end{bmatrix},\begin{bmatrix}
-\sigma_{T}^2(\bx) & \sigma_{T}(\bx) \sigma_S(\bx) \gamma(\bx)  \\
-\sigma_{T}(\bx) \sigma_S(\bx) \gamma(\bx)  & \sigma_{S}^2(\bx) 
+\sigma_{1}^2 & \sigma_{1} \sigma_2 \gamma \\
+\sigma_{1} \sigma_2 \gamma  & \sigma_{2}^2 
 \end{bmatrix}
-\right),
+\right).
 \end{equation}
-where the notation refers to Gaussian (i.e. normal) distributed
-variables specified by the mean vector and covariance matrix. For the
-application of a river plume, the model parameters are specified from
-a short preliminary survey. The means vary with location, capturing
-the variability near the river mouth, while the variance parameters
-and correlation between salinity and temperature are assumed to be
-constant for all locations. %and within each water mass.
-
-For the spatial dependence among variables, we assume a separable
-correlation function with the same decay for salinity and temperature,
-which is consistent with expectations for a river plume where the
-decay is related to common water masses (Section
-\ref{sec:case_study}). In the resulting function, we have:
-\begin{equation}\label{gp_corr}
-\mbox{Cov}(\xi_i(\bx),\xi_j(\bx')) = \rho(\bx,\bx') \sigma_i \sigma_j (\gamma +(1-\gamma )\delta_{ij}), \\
-\end{equation}
-where for $i,j \in {T,S}$ and $\delta_{ij}=1$ if $i=j$, and
-$\delta_{ij}=0$ otherwise. Moreover, we assume stationary isotropic
-random fields where the correlation function $\rho(\bx,\bx')$
-solely depends on $\bx$ and $\bx'$ via Euclidean distance
-$h=\sqrt{||\bx-\bx'||^2}$. With data and prior knowledge, one could
-possibly fit and estimate parameters of non-stationary or
-non-isotropic covariance functions and more complex multivariate
-spatial covariance functions
-\citep{gneiting2010matern,genton2015cross}; however, that is left for
-future work.
-
-We discretize the spatial domain $\mathcal{M}$ to a set of $n$ grid
-locations $\mathcal{M}_g = \{\bx_i, i=1,\ldots,n \}$, where each cell
-has area $\Delta$; the grid is used for the waypoint graph setting the
-possible design locations. It is also used when solving integral
-expressions in Eq. (\ref{measVAR}) and (\ref{sur}), which are then
-approximated by sums over all grid cells. Using vector notation, we
-denote the Gaussian temperature and salinity variables on the grid and
-its density function by the formula:
-\begin{equation}\label{prior}
-\bxi = (\xi_T(\bx_1),\xi_S(\bx_1),\ldots,\xi_T(\bx_n),\xi_S(\bx_n))^t, \hspace{3mm}
-\bxi  \sim  N(\bmu, \bSigma). %\pi (\bxi) & =& \frac{1}{(2\pi)^{n}|\bSigma|^{\frac{1}{2}}}e^{-\frac{1}{2}(\bxi -\bmu)^T\bSigma^{-1}(\bxi-\bmu)}.
-\end{equation}
-%Here, $\pi(\bxi)$ is the joint density function of length $2 n$ vector $\bxi$, and
-The length $2n$ vector $\bmu$ contains the expected values of
-temperature and salinity at the grid locations, while the
-$2n \times 2n$ matrix $\bSigma$ contains the
-variance-covariances of temperature and salinity variables at all grid
-locations.
-
-We use a bivariate GP model for temperature and salinity, where we
-specify the mean:
-\begin{equation}\label{m}
-    \bmu(\bx)=E 
-    \begin{bmatrix}
-    \xi_T(\bx) \\
-    \xi_S(\bx) 
-    \end{bmatrix}=\begin{bmatrix} \mu_{T}(\bx)\\
-\mu_{S}(\bx)
-\end{bmatrix} 
-= \begin{bmatrix} \beta_{T,0} + \beta_{T,1} x_{\mbox{West}} \\
-\beta_{S,0} + \beta_{S,1} x_{\mbox{West}}
-\end{bmatrix}.
-\end{equation}
+We specify the mean by 
+$\mu_{1}(\x)=5.8 + 0.065 u_{\mbox{West}}$ and
+$\mu_{2}(\x)=29 + 0.1 u_{\mbox{West}}$, which means that the expected levels of both temperature and salinity increase linearly in the western direction.
 This situation
-mimics that of a river mouth opening from the south, with the water
-masses pulled to the east. The regression parameters are  $\beta_{T,1}=0.065$ and
-$\beta_{S,1}=0.1$ for the slopes with west coordinate, and $\beta_{T,0}=5.8$
-and $\beta_{S,0}=29.0$ for the intercept at a chosen reference map coordinate.
-
-%The prior assumption used by the agent are $\tilde{\beta_{T,1}}=0.085$ and $\tilde{\beta_{S,1}}=0.138$, which assumes that the process is centered.
-
-The covariance is assumed to be stationary and separable for the two variables. 
-The $2 \times 2$ pointwise covariance matrix is set to:
-\begin{equation}\label{v0}
-\bSigma(\bx)=\mbox{Var} 
-\begin{bmatrix}
-    \xi_T(\bx) \\
-    \xi_S(\bx) 
-    \end{bmatrix}=
-\begin{bmatrix}
-0.25^2 & 0.6 \cdot 0.25^2 \\
-0.6 \cdot 0.25^2 & 0.25^2
-\end{bmatrix}.
-\end{equation}
-The spatial correlation is of a Matern type \kc{cite or define}:
+mimics that of a river mouth opening from the south, with the water masses pulled to the east by the Coriolis force. 
+The variances are set to $\sigma_1=0.25$ and $\sigma_2=0.6$, while the correlation is $0.6$. 
+The spatial correlation is of separable Matern type (see e.g. \cite{Cressie:11}) so that
 \begin{equation}\label{v}
@@
 \begin{bmatrix}
-    \xi_T(\bx) \\
-    \xi_S(\bx) 
+    Z_1(\x) \\
+    Z_2(\x) 
     \end{bmatrix},
     \begin{bmatrix}
-    \xi_T(\bx') \\
-    \xi_S(\bx') 
+    Z_1(\x') \\
+    Z_2(\x') 
     \end{bmatrix}
@@
 \end{equation}
-where $h=\sqrt{||\bx-\bx'||^2}$ and $\phi=0.3$ indicates an effective correlation range of about $1200$ m. 
-%Figure \ref{fig:stat_design} shows the contour lines for the EP for the reference. We notice the trend of increasing salinity and temperature to the west.  
+where $h=\sqrt{||\x-\x'||^2}$ and $\phi=0.3$ indicates an effective correlation range of about $1200$ m.
+
+In the accompanying Python code, these modeling assumptions can be generalized to anisotropic covariance and changing variance levels across the spatial domain.
+One can hence easily run the code to see if different models result in other sampling designs.
+Anisotropy and non-stationary variance are both relevant for the setting with river plumes, but in practice this requires more parameters to be specified.
+With extensive data and prior knowledge, one could also possibly fit and estimate parameters of more complex multivariate spatial covariance functions
+\citep{gneiting2010matern,genton2015cross}, but that is outside the scope of the current paper.
+
+We discretize the spatial domain $\mathcal{M}$ to a set of $n$ grid
+locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,n \}$, where each cell
+has area $\Delta$; the same grid is used for the waypoint graph setting the
+possible design locations. The EIBV is approximated by sums over all grid cells.
+
 One realization of the random fields representing salinity and
@@
 
-For a fixed survey length, a closed form expression for the expected IBV is available as in Eq. \eqref{sur}. However, for the sequential approaches this is not the
+For a fixed survey length, a closed-form expression for the EIBV is available as in Eq. \eqref{eq:eibv}. However, for the sequential approaches this is not the
 case. For comparison the properties are therefore evaluated using
@@
 Monte Carlo replicates, such results are repeatedly computed to
-approximate the expected IBV. We also compare predictive performance
+approximate the EIBV. We also compare predictive performance
 measured by root mean square error (RMSE) for temperature and salinity
@@
 important to note that the objective function used by the agent is
-focused on reducing the expected IBV, but we nevertheless expect that
+focused on reducing the EIBV, but we nevertheless expect that
 we will achieve good predictive performance for criteria such as RMSE
@@
 procedure is then repeated. At each stage, it takes the AUV about 30
-seconds to evaluate the expected IBV for all the possible
+seconds to evaluate the EIBV for all the possible
 waypoint-and-path alternatives.
@@
 %\end{figure}
-\newpage
+%\newpage
 \section{Closing remarks}\label{sec:concl_disc}
@@
 In particular, we derive and show results for characterizing phenomena
-connected to the properties of water masses. The characterization is
-based on joint ESs, whereby we derive new results for the expected IBV
-reduction achieved by the spatial sampling design. This result is
-first calculated in closed form for the situation with a static
+connected to the properties of water masses. The characterization of uncertainties in random sets is extended with new results for the EIBV uncertainty
+reduction achieved by the spatial sampling design. This is provided in closed-form for the situation with a static
 design, and then extended to the adaptive situation. The sequential
@@
 
-The spatial-statistical design criterion, building on ESs, is relevant
-in our setting with different water mass properties; this could also
-be useful in other oceanographic settings, such as algal-blooms or
-open water fronts. Of course, other criteria are also relevant for
-decision-making. For instance, managers and regulators must make
-difficult decisions related to fish farming, other marine resources
+The spatial-statistical design criterion, building on random sets is relevant
+in our setting with different water mass properties, and we show mathematical generality beyond the EIBV which is convenient to implement for onboard planning. Such criteria could also
+be useful in other oceanographic settings as for algal-blooms or
+open water fronts.
+Of course, other criteria are also relevant for, possibly hybrid criteria or multi-attribute criteria that could balance goals of exploration and exploitation in this situation.
+One further has cases where there is a clear downstream
+decision-making, for instance managers and regulators that must make difficult decisions related to fish farming, other marine resources
 operations, or environmental projects. Value of information analysis
@@
 Excellence, project number 223254, and the Applied Underwater Robotics
-Labortatory (AURLab). JE acknowledges support from Norwegian research
-council project 294404. DG acknowledges support from the Swiss
-National Science Foundation, project number 178858.
+Labortatory (AURLab). JE and KR acknowledges support from Norwegian research
+council, project number 305445. CT and DG acknowledges support from the Swiss National Science Foundation, project number 178858.
 
@@
 
-Proof of Proposition 1:
+Proof of Proposition \ref{propo1}:
 
@@
 
-Proof of Proposition 2:
+Proof of Proposition \ref{propo2}:
 
@@
 
-We the case of multivariate monomials in orthant probabilities with thresholds affine in a common Gaussian vector. 
-%Namely, for monomials of degree $k$, we have.
-
-\begin{propo}%[\textcolor{red}{Notation in progress, to be reviewed/revised}]
-	\label{propo3}
-	Let $g, p, q\geq 1$, $h_{1},\dots, h_{g}\geq 1$ with $H=\sum_{i=1}^g h_i$, $a_{i} \in \R^{p}$, $B_{i}\in \R^{p \times q}$, and covariance matrices $\covN_i \in \R^{p \times p}$ $(1\leq i \leq g)$.
-Then, for any covariance matrix $\covV \in \R^{q\times q}$ and $V\sim\mathcal{N}_{q}(0_q,\covV)$,
-	%	Let $g$ and $k$ be integers, and consider a $g$-terms monomial of degree $k$
-	%	\[
-	%	    \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \Sigma_{i}
-	%	    \right)^{h_i},
-	%    \]
-	%    where $\sum_{i=1}^g h_i = k$, each $a_i$, $B_i$ and $\Sigma_i$ is as in
-	%    \cref{propo2} and $V$ is a $q$-dimensional Gaussian vector distributed as
-	%    $V\sim\mathcal{N}(0,\covV)$.
-	%
-	% Then the expectation of the monomial is given by
-	\begin{equation}
-	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
-	=
-	\varPhi_{p H}
-	\left(
-	\bm{a}
-	;
-	\mathbf{\Sigma}
-	\right),
-	\end{equation}
-	with $\bm{a}=(1_{h_1}\otimes a_1, \dots, 1_{h_g}\otimes a_{g}) \in \R^{p H}$
-	and $\mathbf{\Sigma}\in \R^{p H \times p H}$ is defined blockwise by $(\Sigma_{i,j})_{i,j \in \{1,\dots, g\}}$ where, for any $i,j \in \{1,\dots, g\}$, %the blocks are given by 
-	\begin{equation}
-	\Sigma_{i,j}=
-	(1_{h_{i}}1_{h_{j}}')\otimes (B_{i}\Sigma_{V} B_{j}') + \delta_{i,j}(I_{h_{i}}\otimes \covN_{i}) \in \R^{p h_{i} \times p h_{j}}.
-	\end{equation}
-	%with $\bm{a}=\sum_{j=1}^g \mathbf{1}_{J_{j}} \otimes a_j \in \R^{p \times k}$ 
-	%where $J_{j}=\{\sum_{i=1}^{j-1}h_i + 1, \dots, \sum_{i=1}^{j}h_i\}$,
-	%$\mathbf{1}_{I_{j}}$ denotes the vector of $\R^{k}$ with components equal to $1$ for indices in $I_{j}$ and $0$ else ($1\leq j \leq g$), 
-	%and $\mathbf{\Sigma}\in \R^{p k \times p k}$ is defined blockwise by 
-	%\begin{equation}
-	%\mathbf{\Sigma}_{\widetilde{J}_{j_{1}},\widetilde{J}_{j_{2}}}=
-	%(\mathbf{1}_{h_{j_1}}\mathbf{1}_{h_{j_2}}')\otimes (B_{j_{1}}\Sigma_{V} B_{j_2}') + \delta_{j_{1},j_{2}}(I_{j_{1}}\otimes \Sigma_{j_{1}}),
-	%\end{equation}
-	%where $\widetilde{J}_{j}=\{p\sum_{i=1}^{j-1}h_i + 1, \dots, p\sum_{i=1}^{j}h_i\}$ ($1\leq j \leq g$).
-\end{propo}
+Proof of Proposition \ref{propo3}.
 

### 2020-06-04 main_new.tex
updates

@@
 
-  Improving and optimizing oceanographic sampling is a crucial task for marine science and maritime management.
-Faced with limited resources to understand processes in the water-column, the combination of statistics and autonomous robotics provides new opportunities for experimental designs.
-In this work we develop methods for efficient spatial sampling applied to the mapping of coastal processes by providing informative descriptions of spatial characteristics of ocean phenomena.
-Specifically, we define a design criterion based on improved characterization of the uncertainty in the excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected Bernoulli variance reduction in such a framework.
-We demonstrate how this criterion can be used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective.
-We use simulations to study the properties of methods and to compare them with state-of-the-art approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a case study mapping the boundary of a river plume.
-The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.
+  Improving and optimizing oceanographic sampling is a crucial task
+  for marine science and maritime management. Faced with limited
+  resources to understand processes in the water-column, the
+  combination of statistics and autonomous robotics provides new
+  opportunities for experimental design. In this work we develop
+  methods for efficient spatial sampling applied to the mapping of
+  coastal ocean processes by providing informative descriptions of
+  spatial characteristics of ocean phenomena. Specifically, we define
+  a design
+  criterion based on improved characterization of the uncertainty in
+  the excursions of vector-valued Gaussian random fields, and derive
+  tractable expressions for the expected Bernoulli variance reduction
+  in such a framework. We demonstrate how this criterion can be used
+  to prioritize sampling efforts at locations that are ambiguous,
+  making exploration more effective. We use simulations to study the
+  properties of methods and to compare them with state-of-the-art
+  approaches, followed by results from field deployments with an
+  autonomous underwater vehicle (AUV) as part of a study mapping the
+  boundary of a river plume. The results demonstrate the potential of
+  combining statistical methods and robotic platforms to effectively
+  inform and execute data-driven environmental sampling.
 
@@
 \kwd{Experimental Design}
+\kwd{Autonomous robots}
 \kwd{Adaptive Information Gathering}
@@
 
-Monitoring the world's oceans has gained increased importance in 
-light of the changing climate and increasing anthropogenic impact. A
-central problem for understanding these factors is the lack of
-representative data with sufficient resolution. Most of this
-\emph{undersampling} can be attributed to the large spatio-temporal
-variations on which ocean processes transpire, prompting the need for
-effective means of sampling.
-By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
-This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this the paper.
-
-There are several obstacles when doing sampling with autonomous platforms in the ocean.
-First, full numerical ocean models based on complex differential equations cannot provide accurate results for online operations because the onboard computers lack the computational power required.
-Hence statistical proxy models of the environment must be used.
-For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest.
-Second, with undersampling issues one must usually plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs.
-The acquired in-situ information must then be assimilated into the statistical model onboard the platform, so that it can be used to re-compute the sampling criterion and in this way inform decisions on where to sample next.
-We use a criterion focusing on the prediction of excursion sets (ESs) for the multivariate spatial processes of interest.
-Third, the number of sampling designs is enormous, creating a trade-off between optimization and computability.
-One must often compromise in both development and practical operations, and we will rely on short-horizon optimization routines for sequential designs among nearest neighbors points in a spatial grid (waypoint graph).
-
-The imperfect understanding of causality coupled with the underlying fundamental uncertainty of the ocean implies that we need to obtain more measurements and use statistical models and methods to determine what and how phenomena form, to increase the quality of predictive statements.
-With this in mind, we design AUV sampling strategies for reducing the spatially-integrated Bernoulli variance (IBV) in the ES indicator variables.
-Recent work on ESs and associated excursion probabilities (EPs) connected to spatial statistics include \cite{french2013spatio,bolin2015excursion,french2016credible}.
-Our focus is on uncertainty characterization and how this is influenced by sampling.
-In this sense, our work is similar to \cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who describe analytical results for the IBV for univariate processes.
-A major contribution of our work is to derive closed-form results for this design criteria for situations where the underlying model is based on multivariate GPs.
-Another contribution is that we embed this approach in a sequential strategy to produce sampling algorithms, and apply this to a real-world scenario of autonomously sampling temperature and salinity variables using an autonomous underwater vehicle (AUV), with the intention of understanding the between the ocean and a river plume.
-
-Motivating examples relevant for ESs of multivariate processes are abundant.
-In medicine, doctors do not rely solely on a single symptom but must see several combined effects before making a diagnosis.
-In our context of environmental sampling applications, the salinity and temperature excursions of a river plume can similarly help characterize the underlying oceanographic and bio-geochemical processes \citep{hopkins2013detection,Pinto2018}, where sampling should be aimed at mapping relevant parts of the domain to reduce uncertainty in the temperature and salinity ES.
-Unlike what has been done in previous plume exploration studies, we emphasize the links to spatial statistical modeling and targeted multivariate sampling criteria for sequential sampling.
-
-The remainder of this paper is organized as follows: Section \ref{sec:bg} provides select background on ocean sampling; Section \ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for vector-valued GPs; Section \ref{sec:heuristics} builds on these assumptions when deriving the sequential design criteria for adaptive sampling; Section \ref{sec:simulations} discusses properties of the methods in simulation studies.
-Section \ref{sec:case_study} demonstrates the methodology used in field work characterizing a river plume; and Section \ref{sec:concl_disc} contains a summary and a discussion of future work.
-
-\section{Background on Ocean Processes and Observations}\label{sec:bg}
+Monitoring the world's oceans has gained increased importance in light
+of the changing climate and increasing anthropogenic impact. A central
+problem for understanding these factors is the lack of representative
+data with sufficient resolution. Most of this \emph{undersampling} can
+be attributed to the large spatio-temporal variations in which ocean
+processes transpire, prompting the need for effective means of
+sampling. By \emph{sampling}, we refer to the design of observational
+strategies in the spatial domain, where the use autonomous robotic
+platforms can be combined with statistical methods to pursue
+measurements with high scientific relevance. This combination, of
+statistical tools and robotic platforms, constitute a methodological
+basis for addressing the problem of efficient sampling of the ocean,
+which is at the core of this work.
+
+There are several challenges with sampling with autonomous platforms.
+\kc{You seem to jump the gun in stating a challenge without
+  connecting it to the sampling problem. Introduce a sentence as to
+  why models (as a proxy) are important for \emph{understanding} ocean
+  processes} First, full numerical ocean models based on complex
+differential equations cannot be run onboard a marine robot with
+limited computational capacity.
+% provide accurate results for online
+% operations because the onboard computers lack the computational power
+% required.
+Hence statistical proxy models of the environment must be
+used. For our purpose, we rely on Gaussian process (GP)
+representations of the ocean variables of interest.\kc{Sorry....but
+  this makes no sense whatsoever. What is the connection of you using
+  a GP with sampling? You need to make a case first as to the
+  viability of using GP in a marine robot first as a potential viable
+  option, before you state why you want to use a GP}. Second, with
+undersampling issues \kc{what issues? If there are issues, you state
+  them first and then suggest solutions.} one must usually plan for
+active learning during the operation, where a relevant criterion is
+used to extract the most valuable designs. The acquired in-situ
+information must then be assimilated into the statistical model
+onboard the platform, so that it can be used to re-compute the
+sampling criterion and in this way inform decisions on where to sample
+next. We use a criterion focusing on the prediction of excursion sets
+(ESs) for the multivariate spatial processes of interest. Third, the
+number of sampling designs is enormous, creating a trade-off between
+optimization and computability. One must often compromise in both
+development and practical operations, and we will rely on
+short-horizon optimization routines for sequential designs among
+nearest neighbors points in a spatial grid (waypoint graph).
+
+\kc{The para above needs a rewrite to me. It starts off by stating
+  there are challenges. But then jumps to models without stating the
+  connection with sampling, and then jumps into learning without
+  stating why learning is important. Typically the English in formal
+  text starts with a statement which is supported or clarified. In the
+  two cases above, the connections are simply not obvious. If this was
+  there in the first version of the m/s submitted, I'm not sure why
+  the reviewers didn't catch it. This is not just a stylistic issue;
+  it has to do with making sense and bring the reader/reviewer along.}
+
+\kc{Second, I don't see a motivation for GPs here and as I read
+  further into the text. Even for stats folks, are GPs the ``best''
+  ways to go? Are they obviously the tools one would use in
+  environmental sampling? And even if they are, I think we ought to
+  make the case for it strongly and generally, so this paper is
+  actually accessible and understandable to those not in the Spatial
+  Stats field. I would suggest adding a short para on GPs and why
+  they're important -- i.e. motivate the use of GPs early, so we can
+  freely make attributions to them later, as in the next section, for
+  example.}
+
+The imperfect understanding of causality coupled with the underlying
+fundamental uncertainty of the ocean implies that we need to make a
+large number of measurements. In addition, the use statistical models
+and methods to determine what and how phenomena form, is one way to
+increase the quality of predictive power aiding our understanding of
+upper water-column processes.
+
+With this in mind, we have designed AUV sampling strategies for
+reducing the spatially-integrated Bernoulli variance (IBV) in the ES
+indicator variables. Recent work on ESs and associated excursion
+probabilities (EPs) connected to spatial statistics include
+\cite{french2013spatio,bolin2015excursion,french2016credible}. Our
+focus is on uncertainty characterization and how this is influenced by
+sampling. In this sense, our work is similar to
+\cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who
+describe analytical results for the IBV for univariate processes.
+
+A major contribution of our work is to derive closed-form results for
+this design criteria for situations where the underlying model is
+based on multivariate GPs. Another contribution is that we embed this
+approach in a sequential strategy to produce sampling algorithms, and
+apply this to a real-world scenario of autonomously sampling
+temperature and salinity gradients using an AUV, with the intention of
+characterizing the variability of oceanic and riverine waters in
+typical mixed coastal environments.
+
+Motivating examples relevant for ESs of multivariate processes are
+abundant. In medicine, doctors do not rely solely on a single symptom
+but must see several combined effects before making a diagnosis. In
+our context of environmental sampling, the salinity and temperature
+excursions of a river plume can similarly help characterize the
+underlying bio-geochemical processes
+\citep{hopkins2013detection,Pinto2018}, where sampling can be aimed
+towards reducing uncertainty in the temperature and salinity ES.
+Unlike what has been done in previous plume exploration studies, we
+emphasize the links to spatial statistical modeling and targeted
+multivariate sampling criteria for sequential sampling.
+
+The remainder of this paper is organized as follows: Section
+\ref{sec:bg} provides select background on ocean sampling. Section
+\ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for
+vector-valued GPs. Section \ref{sec:heuristics} builds on these
+assumptions when deriving the sequential design criteria for adaptive
+sampling. Section \ref{sec:simulations} discusses properties of the
+methods in simulation studies. Section \ref{sec:case_study}
+demonstrates the methodology used in field work characterizing a river
+plume and finally, Section \ref{sec:concl_disc} contains a summary and
+a discussion of future work.
+
+\section{Ocean Processes and Observation}
+\label{sec:bg}
 
 Central to understanding the changes taking place in the upper
-water-column is knowledge of the biophysical interaction driven by
-an agglomeration of physical forcings (e.g.
-wind, topography, bathymetry, tidal influences, etc.) and incipient micro-biology driven by planktonic and coastal anthropogenic input, such as pollution and agricultural runoff transported into the ocean by rivers and streams.
-These often result in a range of ecosystem-related phenomena such as blooms and plumes, with direct and indirect effects on society.
-
-Traditional data collection to understand such phenomena at sea has typically been based on static buoys, floats, or ship-based methods, with significant logistical limitations that directly impact coverage and sampling resolution.
-Modern methods using satellite remote-sensing provide large-scale coverage but have limited resolution, are limited to sensing the surface, and are impacted by cloud cover.
-Numerical ocean models similarly find it challenging to provide detail at smaller scale \citep{Lermusiaux:2006}.
-The advent of mobile robotic platforms \citep{Bellingham07} has resulted in significant contributions to environmental monitoring and sampling.
-In particular, AUVs have advanced the state of in-situ sampling, and made robotics an integral part of ocean observation, filling parts of the undersampling gap (See Fig.
-\ref{fig:envir1}).
+water-column is knowledge of the bio-geophysical interaction driven by
+an agglomeration of physical forcings (e.g. wind, topography,
+bathymetry, tidal influences, etc.) and incipient micro-biology driven
+by planktonic and coastal anthropogenic input, such as pollution and
+agricultural runoff transported into the ocean by rivers and streams.
+These often result in a range of ecosystem-related phenomena such as
+blooms and plumes, with direct and indirect effects on society. \kc{we
+  should look for a citation here. Check with John Ryan}
+
+Traditional data collection to understand such phenomena at sea has
+typically been based on static buoys, floats, or ship-based methods,
+with significant logistical limitations that directly impact coverage
+and sampling resolution. Modern methods using satellite remote-sensing
+provide large-scale coverage but have limited resolution, are limited
+to sensing the surface, and are impacted by cloud cover. Numerical
+ocean models similarly find it challenging to provide detail at fine
+scale \citep{Lermusiaux:2006}. The advent of mobile robotic platforms
+\citep{Bellingham07} has resulted in significant contributions to
+environmental monitoring and sampling. In particular, AUVs have
+advanced the state of in-situ sampling, and made robotics an integral
+part of ocean observation, filling parts of the undersampling gap
+\cite{rudnick03,rudnick18}(See Fig. \ref{fig:envir1}).
 
@@
   \caption{\ref{fig:envir1} Traditional ocean observation based on 
-    ship-based sampling has recently been augmented by autonomous robotic vehicles. % and their interactions. 
-    AUV platforms are an integral part of this network being able to reason and make decisions for efficient onboard adaptive sampling.
+    ship-based sampling has been augmented by autonomous
+    robotic vehicles. % and their interactions.  
+    AUV platforms are an integral part of this network being able to
+    reason and make decisions for efficient onboard adaptive sampling.
     % using the sense-plan-act control approach to autonomous control. 
@@
     statistics and robotics can play a vital role in enabling more
-    effective observation practices.}
-\label{fig:envir}
-\end{figure}
-
-The AUVs considered here operate at $1$-$3$ m/s in the upper water column, where they are free to move in six degrees of freedom (6 DOF).
-The in-water operation time capacity depends on survey speed, payload sensors and navigation.
-Generally this amounts to 8-24 hrs.
-operation for smaller crafts, and 24-48 hrs.
-for larger crafts.
-AUVs have limited computational capacity, as they typically use single-board computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM) to do computation.
-
-Surveys with AUVs are usually limited to observations along fixed transects that are pre-scripted in mission plans created manually by a human operator.
-The mission can be specified operating on a scale of hundreds of meters to tens of kilometers depending on the scientific context.
-Faced with limited coverage capacity, a more effective approach is to instead use onboard algorithms to continuously evaluate, update, and refine future sampling locations (sense-plan-act cycle in Fig.
-\ref{fig:envir1}), making the information gathering \emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
-This usually occurs on a spatial grid, called a waypoint graph.
-In doing so, the space of sampling opportunities is still limited by the waypoint graph, but the AUV has flexibility to modify its path at each waypoint based on what has been measured onboard \citep{py10,Rajan12,Rajan12b}.
-
-The application that is of focus in this work is towards spatial characterization of a frontal system generated by river plumes.
-Fig.
-\ref{fig:nidelven} shows the survey area (Nidelva, Norway) where cold freshwater enters from the river, creating a strong gradient in both temperature and salinity.
-Because of the local topography and the Coreolis force %\citep{coriolis1835memoire}
-the cold fresh water tends to flow near
-land to the east.
-Depending on the river discharge, tidal effects, wind, and temperature differences, this boundary often gets distorted.
-Initial knowledge about the location and evolution of these but the locations of these features are highly uncertain, making deterministic planning challenging.
-
-Adaptive in-situ AUV sampling of an evolving frontal feature has been explored in \cite{fronts11,Zhang2012,Pinto2018,costa19}.
-These approaches typically use a reactive-adaptive scheme, whereby exploration does not rely on a statistical model of the environment, but rather adapts based on closing the sensing and actuation loop.
-Myopic sampling, i.e.
-stage-wise selection of the path (on the waypoint graph), has been used for surveys \citep{singh2009efficient,Binney2013} that focus largely on reducing predictive variance or entropy.
-These criteria are widely adopted in the statistics literature on spatio-temporal design as well \cite{bueso1998state,zidek2019monitoring}, but variance and entropy reduction are independent of the actual data realizations under the assumptions of GP models, so limited decision flexibility is created for sequential designs.
-The use of data-driven adaptive criteria was introduced to include more targeted sampling of regions of scientific interest in \cite{Low2009} and \cite{fossuminformation}.
-In this paper, with a focus on mapping the river plume by rewarding the designs that improve the classification of ES in temperature and salinity.
+    effective oceanographic observation.\kc{The inset ``AUV robot''
+      looks very tacky and ad-hoc; might be best to remove it. Why not
+      add a subfigure of an AUV instead and not lump it into a figure
+      which is already quite busy? It can't hurt. Change ``ships'' to
+      ``Ship'', remove the aerial drone, since we didn't use it and
+      don't plan to in this work.}}
+  \label{fig:envir} \end{figure}
+
+The AUVs considered here powered untethered platforms that operate at
+$1$-$3$ m/s in the upper water column, where they are free to move in
+six degrees of freedom (6 DOF). The in-water operation time capability
+depends on survey speed, payload sensors and navigation; typically
+this amounts to 8-48 hrs. AUVs typically use single-board computers
+(SBCs), like the Raspberry Pi or a multicore GPU NVIDIA Jetson TX1
+(quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4
+GB of 1600 MHz DRAM) for computation onboard.
+
+Surveys with AUVs are usually limited to observations along fixed
+transects that are pre-scripted in mission plans created manually by a
+human operator. The mission can be specified operating on a scale of
+hundreds of meters to tens of kilometers depending on the scientific
+context. Faced with limited coverage capacity, a more effective
+approach is to instead use onboard algorithms to continuously
+evaluate, update, and refine future sampling locations (sense-plan-act
+cycle in Fig. \ref{fig:envir1}), making the information gathering
+\emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
+% This usually occurs on a spatial grid, called a waypoint graph. 
+In doing so, the space of sampling opportunities is still limited by
+the waypoint graph, but the AUV has flexibility to modify its path at
+each waypoint based on in-situ sensing and measurements onboard
+\citep{py10,Rajan12,Rajan12b}.
+
+We focus on spatial characterization of a frontal system generated by
+river plumes. Fig. \ref{fig:nidelven} shows the survey area in
+Trondheim, Norway, where cold freshwater enters from the river,
+creating a strong gradient in both temperature and salinity. Because
+of the local topography and the Coriolis force the cold fresh water
+tends to flow to the east. Depending on the river discharge, tidal
+effects, wind, and temperature differences, this boundary often gets
+distorted. Initial knowledge about the location and evolution of these
+features are highly uncertain, making deterministic planning
+challenging.
+
+Adaptive in-situ AUV sampling of an evolving frontal feature has been
+explored in \cite{fronts11,Zhang2012,Pinto2018,costa19}. These
+approaches typically use a reactive-adaptive scheme, whereby
+exploration does not rely on a statistical model of the environment,
+but rather adaptation is based on closing the sensing and actuation
+loop. Myopic sampling, i.e. stage-wise selection of the path (on the
+waypoint graph), has been used for surveys
+\citep{singh2009efficient,Binney2013} that focus largely on reducing
+predictive variance or entropy. These criteria are widely adopted in
+the statistics literature on spatio-temporal design as well
+\cite{bueso1998state,zidek2019monitoring}, but variance and entropy
+reduction are independent of the actual data realizations under the
+assumptions of GP models, so limited decisional flexibility is created
+for sequential designs \kc{should this be, ``...so it has limited
+  decisional flexibility''?}. The use of data-driven adaptive criteria
+was introduced to include more targeted sampling of regions of
+scientific interest in \cite{Low2009} and \cite{fossuminformation}. In
+this paper, we focus on mapping the river plume by rewarding designs
+that improve the classification of ES in temperature and salinity.
 
@@
 
+\kc{Typically, I have been told in the past, it is bad form to start a section and
+  then a subsection right away. I sort of see it, so, it might be good to
+  provide a 1-2 sentence introduction here perhaps?}
+
 \subsection{Excursion sets and Integrated Bernoulli variance}
@@
 
-Our objective is to leverage statistical tools onboard autonomous vehicles to characterize a river
-plume, focusing on spatial separation of cold freshwater from the
-river and warmer saline waters of the fjord. The characterization of uncertainties on the ES is a useful
-goal of this experimental sampling.
-
-In order to formalize the problem in mathematical terms,
-let us consider a multivariate (i.e., vector-valued) Gaussian random field $(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary set (typically in geostatistical studies, a subset of interest of the ambient Euclidean space or of a sphere therein) and the dimensionality of the output $\no \geq 1$ is arbitrary.
-Assume furthermore that the set of interest is characterized in terms of a subset of values of the variables, say $\T \subset \R^{\no}$, and denote the corresponding set of index points by
+Our objective is to leverage statistical tools onboard an autonomous
+robotic platform to characterize a river plume, focusing on spatial
+separation of cold freshwater from a river and warmer saline waters
+of a fjord. The characterization of uncertainties on the ES is a
+useful goal of this sampling strategy.
+
+In order to formalize the problem in mathematical terms, let us
+consider a multivariate (i.e., vector-valued) Gaussian random field
+$(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary
+set (typically in geostatistical studies, a subset of interest of the
+ambient Euclidean space or of a sphere therein \kc{is this obvious to
+  the readership why a sphere? It isn't for me certainly.}) and the
+dimensionality of the output $\no \geq 1$ is arbitrary \kc{Is the
+  definition of such a 'p' and its dimensionality, obvious? This to me
+  springs out of nowhere.}. Assume furthermore that the set of interest
+is characterized in terms of a subset of values of the variables, say
+$\T \subset \R^{\no}$, and denote the corresponding set of index
+points by
 $$
@@
 $$
-Our goal here is to develop approaches to quantify and improve the characterization of uncertainties on such sets in the context of multivariate random fields.
-
-While some aspects of the developed
-approaches do not call for a specific form of $\T$, we will often stick for
-simplicity to the case of orthants ($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where $t_1,\dots, t_{\no} \in \R$) as this will allow efficiently calculating several key quantities.
-Note that changing some $\leq$ inequalities to $\geq$ ones would lead to immediate adaptations.
-In the application specific properties following the orthant assumption come into play: We let $t_1$ be a threshold for temperature and $t_2$ a threshold for salinity.
-The set of index point then defines $\mbox{ES} = \{\x \in D : \gp[\x,1] \leq t_1,\gp[\x,2]  \leq t_2\}$.
-The associated EPs are $p(\x) = P(\gp[\x,1] \leq t_1,\gp[\x,2]  \leq t_2)$, $\x \in \mathcal{M}$.
-
-For statistical generality, we will however keep the abstract notation with $\T$ and general $\es$ for most of the derivations.
-Note that assuming that $\gp$ has continuous trajectories (almost surely) and $T$ to be closed, then $\es$ can be cast as a Random Closed Set \citep{Molchanov2005,adler2009random}.
-%In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
-Let us in particular consider a locally finite (Borel) measure $\mes$ on $D$ and investigate the probability distribution of $\mes(\es)$ through its moments.
+Our goal here is to develop approaches to quantify and improve the
+characterization of uncertainties on such sets \kc{$\es$?} in the
+context of multivariate random fields.
+
+While some aspects of the developed approaches do not call for a
+specific form of $\T$, we will often stick for simplicity to the case
+of orthants
+($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
+$t_1,\dots, t_{\no} \in \R$) as this will allow efficient calculation
+of several key quantities. Note that changing some $\leq$ inequalities
+to $\geq$ ones would lead to immediate adaptations. In the application
+specific properties following the orthant assumption come into play
+\kc{This last phrase doesn't parse; is it ``..., the following
+  orthant....''?}: We let $t_1$ be a threshold for temperature and
+$t_2$ a threshold for salinity \kc{Might be easier to remember with
+  symbols 't' and 's' in some
+  form.}. The set of index point then defines
+$\mbox{ES} = \{\x \in D : \gp[\x,1] \leq t_1,\gp[\x,2] \leq t_2\}$.
+The associated EPs are
+$p(\x) = P(\gp[\x,1] \leq t_1,\gp[\x,2] \leq t_2)$,
+$\x \in \mathcal{M}$.
+
+For statistical generality, we will however keep the abstract notation
+with $\T$ and general $\es$ \kc{Again, previous phrase doesn't parse.}
+for most of the derivations. Note that assuming that $\gp$ has
+continuous trajectories (almost surely) and $T$ to be closed, then
+$\es$ can be cast as a Random Closed Set
+\citep{Molchanov2005,adler2009random}. %In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
+Let us in particular consider a locally finite (Borel) measure $\mes$
+on $D$ and investigate the probability distribution of $\mes(\es)$
+through its moments.
 
@@
 	%Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
-	For a measurable random field $\gp$ and a locally finite measure $\mes$ on $\mathcal{M}$, $\mes(\es)$ is a random variable and for any $r\geq 1$, 
-	\begin{equation*}
-	\begin{split}
-	\mathbb{E}[\mes(\es)^r]
-	&=\int_{\mathcal{M}^{r}} \jointExcuProb 
-	\productMeasure
-	,
-	\end{split}
-	\end{equation*}
+	For a measurable random field $\gp$ and a locally finite
+        measure $\mes$ on $\mathcal{M}$, $\mes(\es)$ is a random
+        variable and for any $r\geq 1$, \begin{equation*}
+          \begin{split} \mathbb{E}[\mes(\es)^r]
+            &=\int_{\mathcal{M}^{r}} \jointExcuProb \productMeasure ,
+          \end{split} \end{equation*}
 
@@
 	$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in \mathcal{M}^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
-	\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.
+	\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.\kc{Is $\no r$ a
+        product of the $\no$ from above? Is that obvious?}
 	\medskip
@@
 	cumulative distribution, % with mean $\meanUU$ and covariance matrix $\covUU$,
-	this time by the way without requiring $\covUU$ %the latter 
+	this time without requiring $\covUU$ %the latter 
 	to be non-singular: 
@@
 Proposition~\ref{propo1} enables one to calculate centered moments of $\mes(\es)$. 
-The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ is probably the most relevant of them for the considered purposes. We have
+The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$
+is probably the most relevant of these for our purposes. We
+have 
 \begin{equation*}
@@
 \end{equation*} 
-which boils down in the excursion/sojourn case where $\T=(-\infty, t_1] \times
-\dots \times (-\infty, t_{\no}]$ to 
-\begin{equation*}
-\begin{split}
+which evaluates in the excursion/sojourn case where
+$\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ to
+\begin{equation*} \begin{split}
   \emv
@@
 \end{equation*}
-This quantity requires to work out an integral over $D^2$.
-In contrast, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves solely an integral on $\mathcal{M}$ and can be expanded in our present setting as follows
+This quantity requires  an integral over $D^2$. In
+contrast, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal}
+involves solely an integral on $\mathcal{M}$ and can be expanded in
+our present setting as follows 
 \begin{equation}\label{IBV}
@@
 
-With limited resources, the goal is to construct informative sampling designs.
-For our situation, data are gathered by an AUV that measure temperature and salinity at chosen locations on a defined waypoint graph covering the domain of interest.
-This discretized grid foster reliable navigation and mission performance.
-The waypoints further enable a structure to the design problem in the sense that sampling flexibility is limited to operation on the graph.
-This gives lots of sampling opportunities, and it is important to spend the sampling efforts wisely.
-We focus on sampling designs that reduce the expected integrated Bernoulli variance (EIBV).
-Note that this sampling can be done in a static manner where the survey design is pre-planned, or in an adaptive setting.
-For the latter case, data can be processed sequentially, and the updated model is then used to design and select the
-next waypoint graph to visit for this stage of data gathering, which are again used to update the model, and so on (Section
-\ref{sec:heuristics}).
-We next describe the data modeling assumptions for one stage and develop design criteria of semi-analytical forms relying on integrals of multivariate Gaussian cumulative distribution functions, we will appeal to some key intermediate results presented next.
-
-We let $d \in \mathcal{D}$ denote a sampling design, with $q$ sampling locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$. The measurement model is defined by
+Our goal is to construct informative sampling designs, given limited
+resources \kc{information?}. In our domain, data is gathered by an AUV
+that can measure temperature and salinity at chosen locations on a
+defined waypoint graph. Such a discretized grid fosters reliable
+navigation and mission performance. The waypoints further enable a
+structure to the design problem in that, that sampling flexibility is
+limited to operation on the graph. % This gives lots of sampling
+% opportunities, and it is important to spend the sampling efforts
+% wisely.
+We focus on sampling designs that reduce the expected integrated
+Bernoulli variance (EIBV). Note that this sampling can be done in a
+static manner where the survey design is pre-planned, or in an
+adaptive setting. For the latter case, data can be processed
+sequentially and the updated model is then used to design and select
+the next waypoint graph to visit, which in turn are used again to
+update the model, and so on (Section \ref{sec:heuristics}). We next
+describe the data modeling assumptions for one stage and develop
+design criteria of semi-analytical forms relying on integrals of
+multivariate Gaussian cumulative distribution functions; we will
+appeal to some key intermediate results presented next.
+
+We let $d \in \mathcal{D}$ denote a sampling design, with $q$ sampling 
+locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$. The
+measurement model is defined by 
 \begin{equation}\label{eq:meas}
@@
 \end{equation}
-Here, $\bz(\bu_d)$ is the size $pq$ vector of GP variables at the sampling locations, $\bG_d$ is the design matrix which in principle allows for observing subsets of the vector-valued field at some of the $q$ locations (i.e.
-the matrix has fewer rows than columns).
-If the additive error has independent elements, the covariance matrix $\bR_d$ is diagonal.
-In our setting temperature and salinity measurements are extracted from different sensors, so the independence assumption between variables is not unrealistic, but this is not necessarily the case for other kinds of processed measurements on the AUV.
-It is further common in geostatistics to assume spatially independent measurement errors, and we also do so.
-But our methodology for constructing sampling designs does not rely on these independence assumptions.
-
-For a given design $d$ the evaluation is based on the expected information contained in the data to influence the criterion.  
+Here, $\bz(\bu_d)$ is the size $pq$ vector of GP variables at the
+sampling locations, $\bG_d$ is the design matrix which in principle
+allows for observing subsets of the vector-valued field at some of the
+$q$ locations (i.e. the matrix has fewer rows than columns). If the
+additive error has independent elements, the covariance matrix $\bR_d$
+is diagonal. In our setting temperature and salinity measurements are
+extracted from different sensors, so the independence assumption
+between variables is not unrealistic, but this is not necessarily the
+case for other kinds of measurements on the AUV. It is further common
+in geostatistics to assume spatially independent measurement errors,
+and we also do so. But our methodology for constructing sampling
+designs does not rely on these independence assumptions.
+
+For a given design $d$ the evaluation is based on the expected
+information contained in the data to influence the criterion.   
 Considering the EIBV criteria, we have
@@
 \end{equation}
-where the expectation is over the planned data.  The optimal design for an AUV survey is then
+where the expectation is over the planned data.  The optimal design
+for an AUV survey is then 
 \begin{equation}\label{crit}
@@
 \end{equation}
-For another criterion using random moments of random sets similar expectations would be required, but as we will see, the EIBV has a closed-form expression which is relatively fast to compute and it is hence useful for onboard design selection on AUVs.
-
-Assuming now that $q$ observations are available, one wishes to predict $\gp[\bu]$, the multivariate Kriging equations holds with size $p$ vector of conditional means: 
+For another criterion using random moments of random sets similar
+expectations would be required, but as we will see, the EIBV has a
+closed-form expression which is relatively fast to compute and hence
+useful for onboard design selection.
+
+Assuming now that $q$ observations are available, one wishes to
+predict $\gp[\bu]$, the multivariate Kriging equations holds with size
+$p$ vector of conditional means \kc{This last sentence doesn't parse.
+  Are they two separate statements?}:  
 \begin{equation}\label{eq:cokrig}

### 2020-06-04 main_new.tex
updates

@@
     effective oceanographic observation.\kc{The inset ``AUV robot''
-      looks very tacky and ad-hoc; might be best to remove it. Why not
-      add a subfigure of an AUV instead and not lump it into a figure
-      which is already quite busy? It can't hurt. Change ``ships'' to
+      looks very tacky and ad-hoc; might be best to remove it. Change
+      ``ships'' to
       ``Ship'', remove the aerial drone, since we didn't use it and
@@
 \end{equation}
-where $\bmu$ is size $p$ unconditional mean function of $Z$, $\by_d$ is the realized vector of observed responses, and $\blambda(\x,\bu_d)=\bK^{-1}_y(\bu_d,\bu_d) \bk(\bu_d, \bu)$.
-Here, the data covariance is $K_y=\mbox{Cov}(\bY_d,\bY_d)=\bG_d\bK(\bu_d,\bu_d)\bG_d^T+\bR_d$ and the cross-covariance is $\mbox{Cov}(\bY_d,Z(\x)) \bk(\bu_d, \x)=$, assuming the measurement error is independent of the GP.
-The Kriging (cross-)covariance function can also be expressed in the same vein via $p \times p$ matrix
+where $\bmu$ is size $p$ unconditional mean function of $Z$, $\by_d$
+is the realized vector of observed responses, and
+$\blambda(\x,\bu_d)=\bK^{-1}_y(\bu_d,\bu_d) \bk(\bu_d, \bu)$. Here,
+the data covariance is
+$K_y=\mbox{Cov}(\bY_d,\bY_d)=\bG_d\bK(\bu_d,\bu_d)\bG_d^T+\bR_d$ and
+the cross-covariance is $\mbox{Cov}(\bY_d,Z(\x)) \bk(\bu_d, \x)=$,
+\kc{There is a dangling '='. Is this legit?} assuming the measurement
+error is independent of the GP.  The Kriging 
+(cross-)covariance function can also be expressed in the same vein via
+a $p \times p$ matrix 
 \begin{equation}\label{eq:krig:cov}
@@
 
-Equations \eqref{eq:cokrig} and \eqref{eq:krig:cov} are used to provide informative, updated statements about sets of interest. 
-For our case with temperature and salinity, the conditional EPs, given measurements $\by_d$, can then be computed from 
+Equations \eqref{eq:cokrig} and \eqref{eq:krig:cov} are used to
+provide informative, updated statements about sets of interest. For
+our case with temperature and salinity, the conditional EPs, given
+measurements $\by_d$, can then be computed from
 \begin{eqnarray}\label{eq:post_ep}
@@
 
-A critical element in the derivation of a closed-form for the EIBV is that the
-conditional mean in Eq. \eqref{eq:cokrig} is a linear (affine) function of
-the data $\by_d$ and the covariance is not a function of the data.
-This in turn means that the probabilities in expression \eqref{eq:post_ep} involve inequality statements for linear combinations of Gaussian variables.
-Related closed-form solutions have been noted in similar contexts \citep{bhattacharjya2013value, chevalier2014fast,stroh}, but not generalized to our situation with random sets for vector-valued GPs.
-Notably, the mean in equation \eqref{eq:cokrig}, is in the form of $\ba+\bB \bV$, where $\bV \sim N(0_q, \bK_y (\bu_d,\bu_d))$, $\ba=(t_1,t_2)^t-\bmu(\bu)-\blambda(\bu,\bu_d)^T \bG_d \bmu(\bu_d)$ and $\bB=\blambda(\bu,\bu_d)^T$.
-The following then applies to give a closed-form solution to EIBV calculation in this case.
+A critical element in the derivation of a closed-form for the EIBV is
+that the conditional mean in Eq. \eqref{eq:cokrig} is a linear
+(affine) function of the data $\by_d$ and the covariance is not a
+function of the data. This in turn means that the probabilities in
+\eqref{eq:post_ep} involve inequality statements for linear
+combinations of Gaussian variables. Related closed-form solutions have
+been noted in similar contexts \citep{bhattacharjya2013value,
+  chevalier2014fast,stroh}, but not generalized to our situation with
+random sets for vector-valued GPs. Notably, the mean in equation
+\eqref{eq:cokrig}, is in the form of $\ba+\bB \bV$, where
+$\bV \sim N(0_q, \bK_y (\bu_d,\bu_d))$,
+$\ba=(t_1,t_2)^t-\bmu(\bu)-\blambda(\bu,\bu_d)^T \bG_d \bmu(\bu_d)$
+and $\bB=\blambda(\bu,\bu_d)^T$. The following then applies to give a
+closed-form solution to EIBV calculation in this case.
 
@@
 
-In block-wise representation, the following relation holds for the covariance in Proposition \ref{propo2}:
+In block-wise representation, the following relation holds for the
+covariance in Proposition \ref{propo2}: 
 \begin{align*}
@@
 \end{align*}
-The calculation of the multivariate Gaussian cumulative distribution is done effectively using code such as that of \cite{genz2009computation}.
-
-In the more general case, for instance that of volumes of random sets, the criterion will include couplings  with different locations and the result must be generalized.
-For the case of multivariate monomials in orthant probabilities with thresholds affine in a common Gaussian vector.
+The calculation of the multivariate Gaussian cumulative distribution
+is done effectively using code such as that of
+\cite{genz2009computation}.
+
+In the more general case, for instance that of volumes of random sets,
+the criterion will include couplings with different locations and the
+result must be generalized. For the case of multivariate monomials in
+orthant probabilities with thresholds affine in a common Gaussian
+vector \kc{This sentence is dangling. Doesn't have a conclusion.}.
 %Namely, for monomials of degree $k$, we have.

### 2020-06-04 main_new.tex
updates

@@
 
-The AUVs considered here powered untethered platforms that operate at
-$1$-$3$ m/s in the upper water column, where they are free to move in
-six degrees of freedom (6 DOF). The in-water operation time capability
-depends on survey speed, payload sensors and navigation; typically
-this amounts to 8-48 hrs. AUVs typically use single-board computers
-(SBCs), like the Raspberry Pi or a multicore GPU NVIDIA Jetson TX1
-(quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4
-GB of 1600 MHz DRAM) for computation onboard.
+The AUVs considered here are powered untethered platforms, that
+operate at $1$-$3$ m/s in the upper water column, where they are free
+to move in six degrees of freedom (6 DOF). The in-water operation time
+capability depends on survey speed, payload sensors and navigation;
+typically this amounts to 8-48 hrs. AUVs typically use single-board
+computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA
+Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared
+cache, and 4 GB of 1600 MHz DRAM) for computation onboard.
 

### 2020-06-04 main_new.tex
Updates from Overleaf

@@
 reducing the spatially-integrated Bernoulli variance (IBV) in the ES
-indicator variables. Recent work on ESs and associated excursion
+indicator variables. 
+ES of random fields are thoroughly presented in \citep{adler2009random}. Recent work on ESs and associated excursion
 probabilities (EPs) connected to spatial statistics include
@@
 
-\kc{Typically, I have been told in the past, it is bad form to start a section and
-  then a subsection right away. I sort of see it, so, it might be good to
-  provide a 1-2 sentence introduction here perhaps?}
+The definitions of the sets of interest are introduced in Section \ref{sec:ES}, while the closed-form expressions for the expected uncertainty reduction are derived in Section \ref{sec:EIBV}.
+Section \ref{Sec:UnivarEx} provides an instructive bivariate example relevant for sampling in the temperature and salinity application.
 
 \subsection{Excursion sets and Integrated Bernoulli variance}
-\label{ES}
+\label{sec:ES}
 
@@
 $(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary
-set (typically in geostatistical studies, a subset of interest of the
-ambient Euclidean space or of a sphere therein \kc{is this obvious to
-  the readership why a sphere? It isn't for me certainly.}) and the
+set (typically in geostatistical studies, a geographical subset of interest) and the
 dimensionality of the output $\no \geq 1$ is arbitrary \kc{Is the
@@
 $\es$ can be cast as a Random Closed Set
-\citep{Molchanov2005,adler2009random}. %In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
+\citep{Molchanov2005}. %In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
 Let us in particular consider a locally finite (Borel) measure $\mes$
@@
 \emv
-&=\int_{D^2} \mathbb{P}\left(
+&=\int_{\mathcal{M}^2} \mathbb{P}\left(
 \gp[u]\in T, \gp[v]\in T \right) 
 d\mes^{\otimes}(u, v)\\
-&-\left( \int_{D} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
+&-\left( \int_{\mathcal{M}} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
 \end{split}
@@
 %\operatorname{Var}[\mes(\es)]
-&=\int_{D^2} 
+&=\int_{\mathcal{M}^2} 
 \varPhi_{2\no}
@@
 (\bt, \bt); \bmu((u,v)), 
-\bK((u,v),(u,v))
+K((u,v),(u,v))
 \right) 
@@
 (u,v)\\
-&-\left( \int_{D} \varPhi_{\no}\left(\bt;\bmu(u), \bK(u)\right) d\mes(u) \right)^2.
+&-\left( \int_{\mathcal{M}} \varPhi_{\no}\left(\bt;\bmu(u), K(u)\right) d\mes(u) \right)^2.
 \end{split}
@@
 &=\int_{\mathcal{M}}
-\varPhi_{\no}\left(\bt;\bmu(\uu), \bK(\uu)\right) 
--\left(\varPhi_{\no}\left(\bt;\bmu(\uu), \bK(\uu)\right) \right)^2
+\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) 
+-\left(\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) \right)^2
 \mathrm{d}\mes(u). 
@@
 \subsection{Expected Integrated Bernoulli Variance and Sampling Designs}
+\label{sec:EIBV}
 
@@
 
-We let $d \in \mathcal{D}$ denote a sampling design, with $q$ sampling 
-locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$. The
+We let $d \in \mathcal{D}$ denote a concatenation of sampling designs with $q$ sampling locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$. The
 measurement model is defined by 
 \begin{equation}\label{eq:meas}
-    \bY_d=\bG_d \bz(\bu_d) + \epsilon(\bu_d), \hspace{3mm} \epsilon(\bu_d) \sim N(0,\bR_d).
+    \bY_d=G_d \bz(\bu_d) + \epsilon, \hspace{3mm} \epsilon \sim N(0,R_d).
 \end{equation}
 Here, $\bz(\bu_d)$ is the size $pq$ vector of GP variables at the
-sampling locations, $\bG_d$ is the design matrix which in principle
+sampling locations, $G_d$ is the design matrix which in principle
 allows for observing subsets of the vector-valued field at some of the
 $q$ locations (i.e. the matrix has fewer rows than columns). If the
-additive error has independent elements, the covariance matrix $\bR_d$
+additive error has independent elements, the covariance matrix $R_d$
 is diagonal. In our setting temperature and salinity measurements are
@@
 Considering the EIBV criteria, we have
-\begin{equation}\label{eq:eibv}
-    \mbox{EIBV} (d)= E_{\bY(\bu_d)} \int_{M}
-\mathbb{P}\left(\gp[\uu]\in T|\by(\bu_d) \right)(1-\mathbb{P}\left(\gp[\uu]\in T |\by(\bu_d)\right))
-d\mes(u),
-\end{equation}
+\begin{eqnarray}\label{eq:eibv}
+    \mbox{EIBV} (d) &=& E_{\bY_d} \left( \int_{\mathcal{M}}
+p(\x;\by_d)(1-p(\x;\by_d))
+d\mes(u) \right), \nonumber \\
+p(\x;\by_d) &=& \mathbb{P}\left(\gp[\uu]\in T|\bY_d=\by_d \right)
+\end{eqnarray}
 where the expectation is over the planned data.  The optimal design
@@
 Assuming now that $q$ observations are available, one wishes to
-predict $\gp[\bu]$, the multivariate Kriging equations holds with size
-$p$ vector of conditional means \kc{This last sentence doesn't parse.
-  Are they two separate statements?}:  
+predict $\gp[\x]$. The multivariate Kriging equations holds with size
+$p$ vector of conditional means:  
 \begin{equation}\label{eq:cokrig}
-\bbm(\x)=\bmu(\x)+\blambda(\x,\bu_d)^T \left( \by_d-\bG_d \bmu(\bu_d)\right),
+\bbm(\x)=\bmu(\x)+\lambda(\x,\bu_d)^T \left( \by_d-G_d \bmu(\bu_d)\right),
 \end{equation}
@@
 is the realized vector of observed responses, and
-$\blambda(\x,\bu_d)=\bK^{-1}_y(\bu_d,\bu_d) \bk(\bu_d, \bu)$. Here,
+$\lambda(\x,\bu_d)=K^{-1}_y(\bu_d,\bu_d) k(\bu_d, \bu)$. Here,
 the data covariance is
-$K_y=\mbox{Cov}(\bY_d,\bY_d)=\bG_d\bK(\bu_d,\bu_d)\bG_d^T+\bR_d$ and
-the cross-covariance is $\mbox{Cov}(\bY_d,Z(\x)) \bk(\bu_d, \x)=$,
-\kc{There is a dangling '='. Is this legit?} assuming the measurement
-error is independent of the GP.  The Kriging 
-(cross-)covariance function can also be expressed in the same vein via
+$K_y=\mbox{Cov}(\bY_d,\bY_d)=G_d K(\bu_d,\bu_d)G_d^T+R_d$ while
+the cross-covariance is $\mbox{Cov}(\bY_d,Z(\x))=G_d k(\bu_d, \x)$, assuming the measurement
+error is independent of the random fields.  The Kriging (cross-)covariance function can also be expressed in the same vein via
 a $p \times p$ matrix 
 \begin{equation}\label{eq:krig:cov}
-\bSigma(\x,\x')=\bK(\x,\x')-\blambda(\x,\bu_d)^T \bK_y(\bu_d, \bu_d) \lambda(\x,\bu_d).
+\Sigma(\x,\x')=K(\x,\x')-\lambda(\x,\bu_d)^T K_y(\bu_d, \bu_d) \lambda(\x,\bu_d).
 \end{equation}
 
-Equations \eqref{eq:cokrig} and \eqref{eq:krig:cov} are used to
+Eq. \eqref{eq:cokrig} and \eqref{eq:krig:cov} are used to
 provide informative, updated statements about sets of interest. For
@@
 \begin{eqnarray}\label{eq:post_ep}
- p(\x;\by) &=& P(Z_1(\x) \leq t_1, Z_{2}(\x) \leq t_2 |\by(\bu_d)),  \nonumber \\
- &=& \Phi_2 ((t_1,t_2)^T-\bmu(\x)-\blambda(\x,\bu_d)^T \left( \by(\bu_d)-\bG_d \bmu(\bu_d)\right);0_2, \bSigma(\bu,\bu)). 
+ p(\x;\by_d) &=& P(Z_1(\x) \leq t_1, Z_{2}(\x) \leq t_2 |\bY_d=\by_d),  \nonumber \\
+ &=& \Phi_2 ((t_1,t_2)^T-\bbm(\x);0_2, \Sigma(\x,\x)). 
 \end{eqnarray}
@@
 (affine) function of the data $\by_d$ and the covariance is not a
-function of the data. This in turn means that the probabilities in
+function of the data. This in turn means that the probabilities in Eq.
 \eqref{eq:post_ep} involve inequality statements for linear
@@
 random sets for vector-valued GPs. Notably, the mean in equation
-\eqref{eq:cokrig}, is in the form of $\ba+\bB \bV$, where
-$\bV \sim N(0_q, \bK_y (\bu_d,\bu_d))$,
-$\ba=(t_1,t_2)^t-\bmu(\bu)-\blambda(\bu,\bu_d)^T \bG_d \bmu(\bu_d)$
-and $\bB=\blambda(\bu,\bu_d)^T$. The following then applies to give a
+\eqref{eq:cokrig}, is in the form of $\ba+B V$, where
+$V \sim N(0_q, K_y (\bu_d,\bu_d))$,
+$\ba=(t_1,t_2)^t-\bmu(\x)+\lambda(\x,\bu_d)^T G_d \bmu(\bu_d)$
+and $B=\lambda(\x,\bu_d)^T$. The following then applies to give a
 closed-form solution to EIBV calculation in this case.
@@
 	\begin{equation*}
-	\mathbb{E}\left[ \varPhi_{p}\left( a + BV; 0_p,\covN \right)^h \right]
+	\mathbb{E}\left[ \varPhi_{p}\left( \ba + BV; 0_p,\covN \right)^h \right]
 	=
@@
 	\left(
-	\bm{a}
+	\ba
 	;0_{ph},
-	\bm{\Sigma}
+	\Sigma
 	\right),
 	\end{equation*}
-	where the vector $\bm{a} \in \R^{p h}$ is defined as 
-	$\bm{a} := 1_h\otimes a = 
+	where the vector $\ba \in \R^{p h}$ is defined as 
+	$\ba := 1_h\otimes a = 
 	\left(a, \dots , a
@@
 	and the $p h\times p h$ covariance matrix is given by
-	$\bm{\Sigma} := 
+	$\Sigma := 
 	1_h 1_h'\otimes B\covV B' + I_h\otimes \covN$.
@@
 \begin{align*}
-	\bm{\Sigma}&=
+	\Sigma&=
 	\begin{pmatrix}
@@
 \subsection{An Example of EIBV using Temperature and Salinity}
+\label{Sec:UnivarEx}
 
@@
 gathering both data types, and for a design with
-temperature measurements alone. Both data are modeled by $(Y_1,Y_2)^t=(Z_1,Z_2)^t+N(0,0.5^2\bI)$, while
+temperature measurements alone. Both data are modeled by $(Y_1,Y_2)^t=(Z_1,Z_2)^t+N(0,0.5^2I_2)$, while
 $Y_2=Z_2+N(0,0.5^2)$ when only temperature is measured.  For this
@@
 
-An adaptive AUV survey is split into many stages. The selection on where to sample is made with restriction to the waypoint graph and defined by the nearest grid nodes in the domain $\mathcal{M}_g$.
+An adaptive AUV survey is split into many stages. The selection on where to sample is made with restriction to the triangular waypoint graph and defined by its nearest grid nodes.
 
@@
 This defines the AUV position at the stage $s$. 
-The data further defines the conditional GP  with mean $\bbm^{s-1}(\x)=E(\gp[u]|\mathcal{Y}^{s-1})$ from
+The data further defines the conditional GP  with mean $\bbm^{s-1}(\x)=E(\gp[\x]|\mathcal{Y}^{s-1})$ from
 Eq. (\ref{eq:cokrig})
-and (cross) covariance matrix $\bSigma^{s-1}=\mbox{Var}(\gp[u]|\mathcal{Y}^{s-1})$ from Eq. (\ref{eq:krig:cov}).
+and (cross) covariance matrix $\Sigma^{s-1}=\mbox{Var}(\gp[\x]|\mathcal{Y}^{s-1})$ from Eq. (\ref{eq:krig:cov}).
 
@@
 sampling location based on current probabilities. One reduces the largest levels of uncertainty by selecting the design $j$ with $P(Z(\x_j) \in T)$ closest to $0.5$.
-For our situation with temperature and salinity this {\it{naive}} strategy then
-selects the design according to:
+For our situation with temperature and salinity variables, this {\it{naive}} strategy selects the stage $s$ sampling design according to:
 \begin{eqnarray}\label{critNaive}
-    d^{*,s} &=& \mbox{argmin}_{j} |P(Z_1(\x_j) \leq t_1, Z_2(\x_j) \leq t_2 | \mathcal{Y}^{s-1})-0.5|. \nonumber
+    d^{*,s} &=& \mbox{argmin}_{j} |P(Z_1(\x_j) \leq t_1, Z_2(\x_j) \leq t_2 | \mathcal{Y}^{s-1})-0.5|, \nonumber
 \end{eqnarray}
+where $j$ is indexing all neighboring locations on the waypoint graph.
 
@@
 \begin{eqnarray}\label{critSEQ}
-    d^{*,s} &=& \mbox{argmin}_{j} \left\{ \sum_{\x \in \mathcal{M}_g} E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( p(\x;\mathcal{Y}^{j,s})\left( 1-p(\x;\mathcal{Y}^{j,s})\right) \right) \Delta, \right\} \\
+    d^{*,s} &=& \mbox{argmin}_{j} \left\{ \int_{\mathcal{M}} E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( p(\x;\mathcal{Y}^{j,s})\left( 1-p(\x;\mathcal{Y}^{j,s})\right) \right) d\mes(u), \right\} \\
     p(\x;\mathcal{Y}^{j,s}) &=& P(Z_u \in T |\by_{d^{j,s}},\mathcal{Y}^{s-1}). \nonumber
@@
 best design at stage $s$, the GP model is updated again. The mean $\bbm^{s}$ and
-covariance matrix $\bSigma^{s}$ are used to compute the next design at
+covariance matrix $\Sigma^{s}$ are used to compute the next design at
 stage $s+1$, and so on.
@@
 addition to the next stage of measurements $\by^{j,s}$, this
-look-ahead strategy, anticipates the subsequent design $j_2$ with
+look-ahead strategy, anticipates the subsequent design indexed $j_2$ with
 data $\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$.  The
@@
 \begin{eqnarray}\label{critLA}
-    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left[ \mbox{argmin}_{j_2}  V_{j_2} \right] \right\}, \nonumber \\
-V_{j_2} & = & \sum_{\x \in \mathcal{M}_g} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1})\left( 1-p(\x;\mathcal{Y}^{j,s+1})\right) \right) \Delta, \nonumber \\
+    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left[ \mbox{argmin}_{j_2}  V_{j_2} \right] \right\}, \\
+V_{j_2} & = & \int_{\mathcal{M}} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1})\left( 1-p(\x;\mathcal{Y}^{j,s+1})\right) \right) d\mes(u), \nonumber \\
     p(\x;\mathcal{Y}^{j,s+1}) &=& P(Z(\x) \in T |\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}). \nonumber
@@
 the current best design. When data are collected, the GP model is
-updated, and the mean $\bbm^{s}$ and covariance matrix $\bSigma^{s}$ are
+updated, and the mean $\bbm^{s}$ and covariance matrix $\Sigma^{s}$ are
 used to compute the next design at stage $s+1$, now anticipating what
@@
 the myopic strategy, so for practical implementation, we prune paths
-in the evaluation of Eq. (\ref{critLA}). This means that we use the myopic
+in the evaluation of Eq. \eqref{critLA}. This means that we use the myopic
 strategy to rank the three best designs following the first stage alone, and
@@
         \end{tabular}}
-      \caption{Simulation results for the final mean explained variance $\bR^{2}=100*(1-(\bSigma_{posterior}/\bSigma_{initial}))$).}
+      \caption{Simulation results for the final mean explained variance $\bR^{2}=100*(1-(\Sigma_{posterior}/\Sigma_{initial}))$).}
     \label{tab:sim_res_r2}

### 2020-06-04 main_new.tex
updates

--- 2020-06-04 main_new.tex
@@
 
-In this section, we calculate EPs and the Bernoulli variance for a few bivariate
+% In this section, 
+We calculate EPs and the Bernoulli variance for a few bivariate
 Gaussian distributions to illustrate the concepts articulated above.
@@
 densities with increasing correlation $\gamma$ between temperature and
-salinity.
+salinity. \kc{What is the source of this data?}
 \begin{figure}[h!] \centering
@@
 Here, the thresholds are set equal to the mean $\mu_1=t_1=5^o C$, and
-salinity $\mu_2=t_2=30$ mg/l, and the EP is $p=P(Z_1 < 5,Z_2 < 30)$. The displayed densities have unit
-standard deviations for both temperature and salinity, but we also
-study the effect of doubling the standard deviations.
-
-Table \ref{tab:sim_rhoab} shows the initial EPs and the associated Bernoulli
-variance (second row) for the examples indicated in
-Fig. \ref{illus_bivarDens}. The EPs increase with the correlation as
-there is a strong tendency to have concurrently low temperature and
+salinity $\mu_2=t_2=30$ mg/l, and the EP is $p=P(Z_1 < 5,Z_2 < 30)$.
+The displayed densities have unit standard deviations for both
+temperature and salinity, but we also study the effect of doubling the
+standard deviations.
+
+Table \ref{tab:sim_rhoab} shows the initial EPs and the associated
+Bernoulli variance (second row) for the examples indicated in Fig.
+\ref{illus_bivarDens}. The EPs increase with the correlation as there
+is a strong tendency to have concurrently low temperature and
 salinity. The Bernoulli variance is similarly large for high

### 2020-06-04 main_new.tex
Updates from Overleaf

--- 2020-06-04 main_new.tex
@@
 the expectation is conditional to this data. A similar closed-form calculation
-for EIBV is hence applicable in Eq. (\ref{critSEQ}) using the
+for EIBV is hence applicable in Eq.
+\eqref{critSEQ} using the
 updated GP model from stage $s-1$. Once the data is collected for the
@@
 \begin{eqnarray}\label{critLA}
-    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left[ \mbox{argmin}_{j_2}  V_{j_2} \right] \right\}, \\
-V_{j_2} & = & \int_{\mathcal{M}} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1})\left( 1-p(\x;\mathcal{Y}^{j,s+1})\right) \right) d\mes(u), \nonumber \\
+    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( \mbox{argmin}_{j_2}  V_{j_2} \right) \right\}, \\
+V_{j_2} & = & \int_{\mathcal{M}} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1}) ( 1-p(\x;\mathcal{Y}^{j,s+1}) ) \right) d\mes(u), \nonumber \\
     p(\x;\mathcal{Y}^{j,s+1}) &=& P(Z(\x) \in T |\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}). \nonumber
@@
 one. Fig. \ref{fig:Eprob} shows the conditional EP, given data
-gathered along the north-south survey lines in these figures. In the
+gathered along the north-south survey lines in these displays. In the
 Monte Carlo replicates, such results are repeatedly computed to
@@
 We studied the sensitivity of the results by modifying the input parameters to have different correlations between temperature and salinity, standard deviations, and spatial correlation range.
-In all runs,  the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of IBV, and much better than \textit{naive}.
+In all runs,  the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of realized IBV, and much better than \textit{naive}.
 The \textit{look-ahead} strategy seems to be substantially better than the \textit{myopic} only for very small initial standard deviation or very large spatial correlation range.

### 2020-06-04 main_new.tex
updates

--- 2020-06-04 main_new.tex
@@
 	In the particular orthant case with $\T=(-\infty, t_1] \times \dots \times (-\infty, t_{r}]$,
-	the latter probability directly writes in terms of the multivariate Gaussian
+	the latter probability directly writes \kc{What does
+          ``writes'' mean here? Connotes?} in terms of the multivariate Gaussian
 	cumulative distribution, % with mean $\meanUU$ and covariance matrix $\covUU$,
@@
 densities with increasing correlation $\gamma$ between temperature and
-salinity. \kc{What is the source of this data?}
+salinity. \kc{What is the source of this data and this figure,
+  i.e. how did you derive this plot?}
 \begin{figure}[h!] \centering
@@
 \ref{illus_bivarDens}. The EPs increase with the correlation as there
-is a strong tendency to have concurrently low temperature and
-salinity. The Bernoulli variance is similarly large for high
+is a strong tendency to have concurrently low temperature and salinity
+\kc{Same question as above. What is the basis for this correlation? Is
+  this just an example or derived from actual data in the mixed layer
+  in the front?}. The Bernoulli variance is similarly large for high
 correlations. EPs and Bernoulli variances are the same for standard
@@
 Bernoulli variance calculations. This is presented for a design
-gathering both data types, and for a design with
-temperature measurements alone. Both data are modeled by $(Y_1,Y_2)^t=(Z_1,Z_2)^t+N(0,0.5^2I_2)$, while
-$Y_2=Z_2+N(0,0.5^2)$ when only temperature is measured.  For this
-illustration, Table \ref{tab:sim_rhoab} shows that the expected
-Bernoulli variance gets lower with larger standard deviations
+gathering both data types, and for a design with temperature
+measurements alone. Both data are modeled by
+$(Y_1,Y_2)^t=(Z_1,Z_2)^t+N(0,0.5^2I_2)$, while $Y_2=Z_2+N(0,0.5^2)$
+when only temperature is measured \kc{I find this notation somewhat
+  confusing. Earlier you had set up t1 and t2 to denote temp and
+  salinity respectively. It might help to have \emph{some} notational
+  similarity with that; otherwise it becomes mind bending I believe.}.
+For this illustration, Table \ref{tab:sim_rhoab} shows that the
+expected Bernoulli variance gets lower with larger standard deviations
 $\sigma_1$ and $\sigma_2$ (right columns). The reduction of Bernoulli
@@
 
+\kc{I feel this section above, which was present in the earlier
+  submitted version too, but to which I did not comment, is
+  significant as an example to the reader. But, as noted above, I
+  don't understand the provenance of the data. It isn't made up,
+  right?}
+
 %\subsection{Expected classification criteria}
@@
 sampling to adapt survey plans based on data and statistical methods
-in accordance with the sense-plan-act control loop (Fig. \ref{fig:envir1}). We now present a few adaptive strategies for robotic sampling
-that build on the notion of EIBV.
+in accordance with the sense-plan-act control loop
+(Fig. \ref{fig:envir1}). We now present a few adaptive strategies for
+robotic sampling that build on the notion of EIBV.
 
@@
 
-An adaptive AUV survey is split into many stages. The selection on where to sample is made with restriction to the triangular waypoint graph and defined by its nearest grid nodes.
-
-Let $d^{j,s}$ denote design
-number $j$ at stage $s$ of a sequential survey. If this design is
-selected, data $\by_{d^{j,s}}$ will be gathered. 
-We further let
-$\mathcal{Y}^{s-1} =\{(d^{j,1}, \by_{d^{j,1}}),\ldots, (d^{j,s}, \by_{d^{j,s}}) \}$ denote
-the set of all data gathered until stage $s-1$ according to selected design $d^{j,1},\ldots,d^{j,s}$. 
-This defines the AUV position at the stage $s$. 
-The data further defines the conditional GP  with mean $\bbm^{s-1}(\x)=E(\gp[\x]|\mathcal{Y}^{s-1})$ from
-Eq. (\ref{eq:cokrig})
-and (cross) covariance matrix $\Sigma^{s-1}=\mbox{Var}(\gp[\x]|\mathcal{Y}^{s-1})$ from Eq. (\ref{eq:krig:cov}).
-
-At the next stage, we choose where measurements $\by_{d^{j,s}}$ should be gathered with
-design $d^{j,s}$. For the triangularized waypoint graph that we use in our AUV application, there are six possible directions for $j$, and we must evaluate the design criterion for all these. 
-The goal of the sampling design is to minimize the expected uncertainty of random sets, and specifically the EIBV for our application.
-We hence focus on choosing designs for data modeled by equation \eqref{eq:meas}, to conduct efficient sequential uncertainty reduction according to the EIBV criterion.
-Notably, compared with the static design $d$ in equation \eqref{eq:eibv}, the optimal sequential design not only considers the best current AUV
-design grid point, but also how the data gathered at this node could inform future sampling at successive stages.
+Our adaptive AUV survey is split into many stages. The selection on
+where to sample is constrained by a triangular waypoint graph and
+defined by its nearest grid nodes.
+
+Let $d^{j,s}$ denote design number $j$ at stage $s$ of a sequential
+survey. If this design is selected, data $\by_{d^{j,s}}$ will be
+gathered.  We further let
+$\mathcal{Y}^{s-1} =\{(d^{j,1}, \by_{d^{j,1}}),\ldots, (d^{j,s},
+\by_{d^{j,s}}) \}$ denote the set of all data gathered until stage
+$s-1$ according to selected design $d^{j,1},\ldots,d^{j,s}$.  This
+defines the AUV position at the stage $s$.  The data further defines
+the conditional GP with mean
+$\bbm^{s-1}(\x)=E(\gp[\x]|\mathcal{Y}^{s-1})$ from
+Eq. (\ref{eq:cokrig}) and (cross) covariance matrix
+$\Sigma^{s-1}=\mbox{Var}(\gp[\x]|\mathcal{Y}^{s-1})$ from
+Eq. (\ref{eq:krig:cov}).
+
+At the next stage, we choose where measurements $\by_{d^{j,s}}$ should
+be gathered with design $d^{j,s}$. For the triangularized waypoint
+graph that we use in our application, there are six possible
+directions for $j$, and we must evaluate the design criterion for all
+these.  The goal of the sampling design is to minimize the expected
+uncertainty of random sets, and specifically the EIBV for our
+application. We hence focus on choosing designs for data modeled by
+equation \eqref{eq:meas}, to conduct efficient sequential uncertainty
+reduction according to the EIBV criterion.  Notably, compared with the
+static design $d$ in equation \eqref{eq:eibv}, the optimal sequential
+design not only considers the best current design grid point, but also
+how the data gathered at this node could inform future sampling at
+successive stages.
 
@@
 %where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages. 
-The mathematical expression for the optimal design involves a
-series of intermixed maximizations over designs and integrals over
-data. In practice, the optimal solution is intractable because of
-the enormous growth over stages (see
-e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}).
-Instead, we outline heuristic strategies.
+The mathematical expression for the optimal design involves a series
+of intermixed maximizations over designs and integrals over data. In
+practice, the optimal solution is intractable because of the enormous
+growth over stages (see e.g. \cite{sucar2015probabilistic} and
+\cite{powell2016perspectives}).  Instead, we outline heuristic
+strategies.
 
@@
 The simplest heuristic for adaptive sampling is to choose the next
-sampling location based on current probabilities. One reduces the largest levels of uncertainty by selecting the design $j$ with $P(Z(\x_j) \in T)$ closest to $0.5$.
-For our situation with temperature and salinity variables, this {\it{naive}} strategy selects the stage $s$ sampling design according to:
+sampling location based on current probabilities. One reduces the
+largest levels of uncertainty by selecting the design $j$ with
+$P(Z(\x_j) \in T)$ closest to $0.5$.  With temperature and salinity
+variables, this {\it{naive}} strategy selects the stage $s$ sampling
+design according to:
 \begin{eqnarray}\label{critNaive}
@@
 
-This strategy does not account for the uncertainty in the GP variables, but is instead applicable only if one design (node) has EPs closer to
-$0.5$ (see Table \ref{tab:sim_rhoab}, line two). Additionally, this strategy does not
-account for spatial correlation. This strategy lacks memory of where
-it has been and where the uncertainty has been reduced, and is therefore
-susceptible to local minima.
+This strategy does not account for the uncertainty in the GP
+variables, but is instead applicable only if one design (node) has EPs
+closer to $0.5$ (see Table \ref{tab:sim_rhoab}, line \kc{row?}
+two). Additionally, this strategy does not account for spatial
+correlation. This strategy lacks memory of where it has been and where
+the uncertainty has been reduced, and therefore susceptible to local
+minima.
 
@@
 imagine taking only one more stage of measurements. This selection
-strategy is based on expected uncertainty reduction, but it is heuristic because there is no anticipation of what the subsequent designs might
-offer beyond the first stage.
-
-Based on the currently available data $\mathcal{Y}^{s-1}$, we
-fit an updated GP model. Focusing on the EIBV criterion, at the next stage $s$, the selected design is:
+strategy is based on expected uncertainty reduction, but it is a
+heuristic because there is no anticipation of what the subsequent
+designs might offer beyond the first stage.
+
+Based on the currently available data $\mathcal{Y}^{s-1}$, we fit an
+updated GP model. Focusing on the EIBV criterion, at the next stage
+$s$, the selected design is:
 \begin{eqnarray}\label{critSEQ}

### 2020-06-05 main_new.tex
updates

@@
 expression (\ref{eq:eibv}). Now $\mathcal{Y}^{s-1}$ is available, and
-the expectation is conditional to this data. A similar closed-form calculation
-for EIBV is hence applicable in Eq.
-\eqref{critSEQ} using the
-updated GP model from stage $s-1$. Once the data is collected for the
-best design at stage $s$, the GP model is updated again. The mean $\bbm^{s}$ and
-covariance matrix $\Sigma^{s}$ are used to compute the next design at
-stage $s+1$, and so on.
+the expectation is conditional to this data. A similar closed-form
+calculation for EIBV is hence applicable in Eq.  \eqref{critSEQ}
+\kc{Strangely in my TeX compilation this points to the earlier
+  definition of $d^{*,s}$. Seems like some index is messed up in the
+  sources; all indexes should perhaps be checked?} using the updated
+GP model from stage $s-1$. Once the data is collected for the best
+\kc{??}  design at stage $s$, the GP model is updated again. The mean
+$\bbm^{s}$ and covariance matrix $\Sigma^{s}$ are used to compute the
+next design at stage $s+1$, and so on.
 
@@
 reasonable approach for creating designs in many
-applications. Moreover, it is easily implemented onboard an AUV,
-using an efficient approach for data updating the GP model and the
-calculation of the closed-form EIBV expressions for each
-subsequent trajectory.
+applications. Moreover, it is easily implemented, in our case, onboard
+an AUV, using an efficient data update of the GP model and the
+calculation of the closed-form EIBV expressions for each subsequent
+trajectory.
 
@@
 We now extend the myopic strategy to a look-ahead strategy, which is
-optimal when one can gather only two more stages of measurements. In
-addition to the next stage of measurements $\by^{j,s}$, this
-look-ahead strategy, anticipates the subsequent design indexed $j_2$ with
-data $\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$.  The
+optimal when one can gather only two more stages of
+measurements. \kc{How can you be sure of this?} In addition to the
+next stage of measurements $\by^{j,s}$, this look-ahead strategy,
+anticipates the subsequent design indexed $j_2$ with data
+$\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$.  The
 selected design is:
@@
 Here, $\mathcal{Y}^{j,s}=\{\by_{d^{j,s}},\mathcal{Y}^{s-1}\}$ and
-$\mathcal{Y}^{j,s+1}=\{\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}\}$ represent
-the sets of data variables, and the expectations are calculated with respect to
-the conditional densities given data at that stage, meaning we have two levels of intermixed minimizations and integrals.
-We solve the
-first expectation by Monte Carlo sampling of data $\by_{d^{j,s}}$ from its
-conditional distribution. For each of these data samples, the second expectation
-is solved using the closed-form expressions for EIBV from
-Section \ref{sec:ESEP}.
+$\mathcal{Y}^{j,s+1}=\{\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}\}$
+represent the sets of data variables, and the expectations are
+calculated with respect to the conditional densities given data at
+that stage, implying that we have two levels of intermixed
+minimizations and integrals.  We solve the first expectation by Monte
+Carlo sampling of data $\by_{d^{j,s}}$ from its conditional
+distribution. For each of these data samples, the second expectation
+is solved using the closed-form expressions for EIBV from Section
+\ref{sec:ESEP}.
 
@@
 
-We now study the properties of
-the different static and sequential survey designs in a realistic
-simulated case that emulates the spatial properties of a river plume interacting with coastal water.
+We now study the properties of the different static and sequential
+survey designs in a realistic case that emulates the spatial
+properties of a river plume interacting with coastal water.
 
@@
 \end{equation}
-We specify the mean by 
-$\mu_{1}(\x)=5.8 + 0.065 u_{\mbox{West}}$ and
-$\mu_{2}(\x)=29 + 0.1 u_{\mbox{West}}$, which means that the expected levels of both temperature and salinity increase linearly in the western direction.
-This situation
-mimics that of a river mouth opening from the south, with the water masses pulled to the east by the Coriolis force. 
-The variances are set to $\sigma_1=0.25$ and $\sigma_2=0.6$, while the correlation is $0.6$. 
-The spatial correlation is of separable Matern type (see e.g. \cite{Cressie:11}) so that
+We specify the mean by $\mu_{1}(\x)=5.8 + 0.065 u_{\mbox{West}}$ and
+$\mu_{2}(\x)=29 + 0.1 u_{\mbox{West}}$, which means that the expected
+levels of both temperature and salinity increase linearly in the
+western direction.  This situation mimics that of a river mouth
+opening from the south, with the water masses pulled to the east by
+the Coriolis force \kc{This seems overly specific to the Nidelva use
+  case. Can it be generalized, after all that is the overall intent of
+  this manuscript?}.  The variances are set to $\sigma_1=0.25$ and
+$\sigma_2=0.6$, while the correlation is $0.6$.  The spatial
+correlation is of separable Matern type (see e.g. \cite{Cressie:11})
+so that
 \begin{equation}\label{v}
@@
 \end{equation}
-where $h=\sqrt{||\x-\x'||^2}$ and $\phi=0.3$ indicates an effective correlation range of about $1200$ m.
-
-In the accompanying Python code, these modeling assumptions can be generalized to anisotropic covariance and changing variance levels across the spatial domain.
-One can hence easily run the code to see if different models result in other sampling designs.
-Anisotropy and non-stationary variance are both relevant for the setting with river plumes, but in practice this requires more parameters to be specified.
-With extensive data and prior knowledge, one could also possibly fit and estimate parameters of more complex multivariate spatial covariance functions
-\citep{gneiting2010matern,genton2015cross}, but that is outside the scope of the current paper.
+where $h=\sqrt{||\x-\x'||^2}$ and $\phi=0.3$ indicates an effective
+correlation range of about $1200$ m.
+
+In the accompanying Python code, these modeling assumptions can be
+generalized to anisotropic covariance and changing variance levels
+across the spatial domain. One can hence, easily run the code to see
+if different models result in other sampling designs. Anisotropy and
+non-stationary variance are both relevant for the setting with river
+plumes, but in practice this requires more parameters to be
+specified. With extensive data and prior knowledge, one could also
+possibly fit and estimate parameters of more complex multivariate
+spatial covariance functions
+\citep{gneiting2010matern,genton2015cross}, but that is outside the
+scope of the current paper.
 
 We discretize the spatial domain $\mathcal{M}$ to a set of $n$ grid
-locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,n \}$, where each cell
-has area $\Delta$; the same grid is used for the waypoint graph setting the
-possible design locations. The EIBV is approximated by sums over all grid cells.
+locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,n \}$, \kc{is the index
+  correct? TeX shows it as u or g for $\mathcal{M}$} where each cell
+has area $\Delta$; the same grid is used for the waypoint graph for
+possible design locations. The EIBV is approximated by sums over all
+grid cells.
 
@@
 temperature is shown in Fig. \ref{fig:true_temp} and
-\ref{fig:true_sal}. The true ES is a realization from this model as shown
-in Fig. \ref{fig:ESet}.
-
-\begin{figure}[!ht]
+\ref{fig:true_sal}. The true ES is a realization from this model as
+shown in Fig. \ref{fig:ESet}.
+
+\begin{figure}[!h]
   \centering
@@
   `static\_north" strategy.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
-  \caption{\ref{fig:true_temp} and \ref{fig:true_sal} show
-    a realisation of the simulated temperature and salinity fields. The associated ES is in \ref{fig:ESet}, while 
-    \ref{fig:Eprob} shows the estimated EPs after
-    performing data collection along the static N-S survey line.} 
+  \caption{\ref{fig:true_temp} and \ref{fig:true_sal} show a
+    realization of the simulated temperature and salinity fields. The
+    associated ES is in \ref{fig:ESet}, while \ref{fig:Eprob} shows
+    the estimated EPs after performing data collection along the
+    static N-S survey line.}
 \label{fig:realisations}
@@
 
-\subsection{Static and Sequential Sampling Designs}\label{sec:sampling_designs}
+\subsection{Static and Sequential Sampling Designs}
+\label{sec:sampling_designs}
 
@@
 \textit{static\_zigzag} with the three described sequential approaches
-\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static sampling paths are pre-determined and cannot be altered and represent the pre-planned strategies used in most current AUV
-operational survey designs.
-
-For a fixed survey length, a closed-form expression for the EIBV is available as in Eq. \eqref{eq:eibv}. However, for the sequential approaches this is not the
-case. For comparison the properties are therefore evaluated using
-Monte Carlo sampling over several replicates of realizations from the model while conducting simulated sequential surveys for each
-one. Fig. \ref{fig:Eprob} shows the conditional EP, given data
-gathered along the north-south survey lines in these displays. In the
-Monte Carlo replicates, such results are repeatedly computed to
-approximate the EIBV. We also compare predictive performance
-measured by root mean square error (RMSE) for temperature and salinity
-estimates as well as also the variance reduction in these two variables. It is
-important to note that the objective function used by the agent is
-focused on reducing the EIBV, but we nevertheless expect that
-we will achieve good predictive performance for criteria such as RMSE
-as well. Another non-statistical criteria that is important for
-practical purposes is the computational time needed for the strategy,
-as this will impact the performance for an embedded system.
+\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
+sampling paths are pre-determined and cannot be altered and represent
+the pre-planned strategies used in most current AUV operational survey
+designs.
+
+For a fixed survey length, a closed-form expression for the EIBV is
+available as in Eq. \eqref{eq:eibv}. However, for the sequential
+approaches this is not the case. For comparison the properties are
+therefore evaluated using Monte Carlo sampling over several replicates
+of realizations from the model while conducting simulated sequential
+surveys for each one. Fig. \ref{fig:Eprob} shows the conditional EP,
+given data gathered along the north-south survey lines in these
+displays. In the Monte Carlo replicates, such results are repeatedly
+computed to approximate the EIBV. We also compare predictive
+performance measured by root mean square error (RMSE) for temperature
+and salinity estimates as well as also the variance reduction in these
+two variables. It is important to note that the objective function
+used by the agent \kc{First of 'agent'. Define it or change it
+  to 'robot'? Note further uses below.} is focused on reducing the
+EIBV, but we nevertheless expect that we will achieve good predictive
+performance for criteria such as RMSE as well. Another non-statistical
+criteria that is important for practical purposes is the computational
+time needed for the strategy, as this will impact the performance for
+an embedded system.
 
@@
 Fig. \ref{fig:wp_graph}. The sequential sampling agent starts at the
-center East-West coordinate at the southern end of the domain (node
-53). The AUV moves along edges in the waypoint graph while the vehicle
-gathers measurements. The data is assimilated into the GP model before
-an evaluation of the next node to sample is conducted at the end of
-the edge.
+center East-West coordinate at the southern end of the domain (node 53
+\kc{Maybe color this node differently and modify the caption?}). The
+AUV moves along edges in the waypoint graph while the vehicle makes
+measurements. The data is assimilated into the GP model before an
+evaluation of the next node to sample is conducted at the end of the
+edge.
 
@@
   \hfill 
-  \subfigure[Time used to do inference.]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}}
+  \subfigure[Computational time for inferencing.\kc{I see only 
+    lines associated with 4 variables showing in the graph. Where is
+    static\_north and static\_east?}]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}} 
 \caption{Simulation results from 100 replicate simulations for 10
-  sampling choices/stages on the grid.} 
+  sampling choices/stages on the grid. \kc{Vertical lines show large
+    variation in replicate results.}}  
 \label{fig:sim_results}
@@
   \caption{The overlaid route choices (strategies), superimposed on
-    the 100 replicate surveys with 10 sampling choices/stages.}
+    the 100 replicate surveys with 10 sampling choices/stages. \kc{I
+      would reorder these figurs above based on the order the material
+      was discussed -- Static, naive, myopic and then Look-ahead. Also
+      do you really need a key associated with each subfigure, given
+      the caption? An abridged analysis as in the text, might also be
+      in order here in the caption. Also refer to
+      Fig. \ref{fig:realisations} as the reference, where the
+      temp/salinity fields are shown.}}
 \label{fig:route_choices}
@@
 vehicle. Some pruning of the graph is performed to improve the
-computational speed, such as ruling out repeated visitations and
+performance, such as ruling out repeated visitations and
 back-and-forth routes. Some of the intermediate results are also
 stored for longer planning horizons. Further pruning of branches or
-inclusion of other heuristics could be included to make the strategy
-faster. Then again, the inclusion of heuristics is likely a
+inclusion of other heuristics could be included for better
+performance. Then again, the inclusion of heuristics is likely a
 contributing factor for the \textit{look-ahead} strategy failing to
@@
 look-ahead. There are several reasons for this. First, a greedy
-approach will tend to put more emphasis on promising locations close to the
-agent, which may lead away from the centre. Second, as the agent
-evaluates the impact of locations further away (look-ahead) where
-assimilated data has less predictive power, the GP model (which
+approach will tend to put more emphasis on promising locations close
+to the agent, which may lead away from the centre. Second, as the
+agent evaluates the impact of locations further away (look-ahead)
+where assimilated data has less predictive power, the GP model (which
 is centered here) will act to restrict paths deviating from the
@@
 
-We studied the sensitivity of the results by modifying the input parameters to have different correlations between temperature and salinity, standard deviations, and spatial correlation range.
-In all runs,  the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of realized IBV, and much better than \textit{naive}.
-The \textit{look-ahead} strategy seems to be substantially better than the \textit{myopic} only for very small initial standard deviation or very large spatial correlation range.
-The \textit{static\_north} design continues to be the best static design for IBV, while \textit{static\_zigzag} is the best design for the other predictive performance measures, especially so with large spatial correlation range.
-We also ran simulation studies with only temperature data, and for realistic correlation levels between temperature and salinity the IBV results are not much worse when only temperature data are available.
-In addition to the comparison made in Table \ref{tab:sim_rhoab}, the current setting includes spatial correlation and this likely reduces the additional influence of having bivariate data.
-However, it seems that having temperature data alone does a substantially worse job in terms of explained variance
+We studied the sensitivity of the results by modifying the input
+parameters to have different correlations between temperature and
+salinity, standard deviations, and spatial correlation range.  In all
+runs, the \textit{myopic} and \textit{look-ahead} strategies perform
+the best in terms of realized IBV, and much better than
+\textit{naive}. The \textit{look-ahead} strategy seems to be
+substantially better than the \textit{myopic} design only for very
+small initial standard deviation or very large spatial correlation
+range. \textit{static\_north} continues to be the best static design
+for IBV, while \textit{static\_zigzag} is the best design for the
+other predictive performance measures, especially so with large
+spatial correlation range. We also ran simulation studies with only
+temperature data, and for realistic correlation levels between
+temperature and salinity, the IBV results are not much worse when only
+temperature data are available. In addition to the comparison made in
+Table \ref{tab:sim_rhoab}, the current setting includes spatial
+correlation and this likely reduces the additional influence of having
+bivariate data. However, it seems that having temperature data alone
+does a substantially worse job in terms of explained variance
 
@@
 \label{sec:exp_modeling}
+
 The statistical model parameters were specified based on a short
-preliminary survey where the AUV made an initial transect to determine the trends in environmental conditions and correlation structures. Based on the
-initial data, the trend parameters were estimated by linear
-regression, where both temperature and salinity are assumed to
-increase linearly, going west from the river mouth. Next, the
-residuals from the regression analysis were analyzed to study the fit
-of the GP model and to specify the covariance parameters.
+preliminary survey where the AUV made an initial transect to determine
+the trends in environmental conditions and correlation
+structures. Based on the initial data, the trend parameters were
+estimated by linear regression, where both temperature and salinity
+are assumed to increase linearly, going west from the river
+mouth. Next, the residuals from the regression analysis were analyzed
+to study the fit of the GP model and to specify the covariance
+parameters.
 
@@
   survey. Observations were taken across the front while crossing from
-  fresh, cold river water to saline and warmer ocean waters.}
+  fresh, cold river water to saline and warmer ocean waters. \kc{Might
+  be good to expand the terms and not have abbreviations in the table
+  above; there's plenty of space.}}
 \label{tab:experiment_param}
@@
 \citep{py10,Rajan12,Rajan12b}, running an instance of the
-\textit{myopic} strategy from Section \ref{sec:myopic} to
-control the AUV and decide between sampling locations.
+\textit{myopic} strategy from Section \ref{sec:myopic} to control the
+AUV and decide between sampling locations.
 
@@
 procedure is then repeated. At each stage, it takes the AUV about 30
-seconds to evaluate the EIBV for all the possible
-waypoint-and-path alternatives.
+seconds to evaluate the EIBV for all the possible waypoint-and-path
+alternatives.
 
 The AUV was set to start in the south-center part of the waypoint
-graph, with the previously outlined GP model of the environment (Section \ref{sec:exp_modeling}). A
-survey was set to take approximately 40 minutes, visiting 15 waypoints
-on the grid, with the AUV running near the surface to capture the
-plume. On its path from one waypoint to the next, the AUV gathered
-data regularly, and the GP model assimilated temperature and salinity
-data with an update frequency of 30 seconds, giving about three updates
-per stage.
+graph, with the previously outlined GP model of the environment
+(Section \ref{sec:exp_modeling}). A survey was set to take
+approximately 40 minutes, visiting 15 waypoints on the grid, with the
+AUV running near the surface to capture the plume. On its path from
+one waypoint to the next, the AUV gathered data regularly, and the GP
+model assimilated temperature and salinity data with an update
+frequency of 30 seconds, giving about three updates per stage.
 
@@
 \centering
-\subfigure[Survey map]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
+\subfigure[AUV survey area]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
 \hspace{0.3cm}
@@
 %\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
-\caption{Results from mapping the Nidelva river. \ref{fig:map} Show an
-  overview of the survey area overlaid with the AUV path in black and
-  dashed line. Note the shaded region indicating the usual frontal
-  region. \ref{fig:res_both} Presents the collected temperature data
-  as colored trails. Note waypoint 5 (WP5) which indicate where the
-  two surveys deviate. \ref{fig:res1} and \ref{fig:res2} Shows the
-  collected salinity data (as a colored trail) overlaid on the final
-  EP, which indicate the AUVs statistical impression of the front. For
-  both missions the temperature and salinity data correspond with the
-  EP front indication.}
+\caption{Results from mapping the Nidelva river, Trondheim, Norway
+  over two survey missions. \ref{fig:map} shows an overview of the
+  survey area overlaid with the AUV path in black and dashed
+  line. Note the shaded region indicating a typical frontal
+  region. \ref{fig:res_both} shows the collected temperature data as
+  colored trails. Note waypoint 5 (WP5) which indicates where the two
+  surveys diverge. \ref{fig:res1} and \ref{fig:res2} shows the
+  collected salinity data overlaid on the final EP, which indicate the
+  AUVs statistical impression of the front. For both missions the
+  temperature and salinity data correspond with an indication of the
+  EP front. \kc{Might want to indicate when the surveys (dates) were
+    conducted so people can understand the variability in the EP with
+    respect to time.}}
 \label{fig:results}
@@
 
-Two missions, Surveys 1 and 2, were run successively from 11:00 AM to
+Two surveys missions (1 and 2), were run successively from 11:00 AM to
 01:00 PM, with a short break in between. The resulting path of the
@@
 temperatures are shown as colored trails in Fig. \ref{fig:res_both},
-clearly indicating the temperature difference between fjord and river
-waters. The salinity data are then shown separately, overlaid with the
-estimated EP for each survey in Fig. \ref{fig:res1} and
+clearly indicating the temperature difference between fjord and
+riverine waters. The salinity data are then shown separately, overlaid
+with the estimated EP for each survey in Fig. \ref{fig:res1} and
 \ref{fig:res2}.
@@
 crossing the frontal boundary multiple times. As conditions changed
-slightly between the two deployments, the resulting path (after
-waypoint 5) is shown to deviate. Survey 1 continued northwards,
-tracking the north-eastern part of the front, while Survey 2 turned
-west, mapping the south-western part.
-
-The final predictions of the front location, represented by conditional EPs in
-Fig. \ref{fig:res1} and \ref{fig:res2} as dashed lines, correspond
-with one another.
-In both surveys they yield a picture of the front being a bit to the west in the southern parts of the domain and gradually bending off toward the north east.
-The amount of exploration done by Survey 1
-is greater than Survey 2. In Survey 1, the AUV obtained more detail by
-going north from waypoint 5, while Survey 2, coming close to the survey area borders
-in the south-western corner, obtained a poorer understanding of the
-northern parts. A look-ahead strategy might identify and
-discourage such choices.
+slightly between the two surveys, the resulting path (after waypoint
+5) is shown to deviate. Survey 1 continued northwards, tracking the
+north-eastern portion of the front, while Survey 2 turned west,
+mapping the south-western region.
+
+The final predictions of the front location, represented by
+conditional EPs in Fig. \ref{fig:res1} and \ref{fig:res2} as dashed
+lines, correspond with one another. In both surveys they yield a
+picture of the front being to the west in the southern portions of the
+region and gradually bending off toward the north east. The amount of
+exploration done by Survey 1 is greater than Survey 2. In Survey 1,
+the AUV obtained more detail by going north from waypoint 5, while
+Survey 2, coming close to the survey area borders in the south-western
+corner, obtained a poorer understanding of the northern parts. A
+look-ahead strategy might identify and discourage such choices.
 
@@
 opportunities available for real-time multivariable spatial
-data-gathering and analysis onboard autonomous sensing platforms,
-which statisticians can exploit to create general-purpose toolkits for
+data-gathering and analysis onboard autonomous platforms, which
+statisticians can exploit to create general-purpose toolkits for
 similar applications.
@@
 In particular, we derive and show results for characterizing phenomena
-connected to the properties of water masses. The characterization of uncertainties in random sets is extended with new results for the EIBV uncertainty
-reduction achieved by the spatial sampling design. This is provided in closed-form for the situation with a static
-design, and then extended to the adaptive situation. The sequential
-derivations provide new insights into efficient applications of
-adaptive sampling, as demonstrated in our application of these models
-with an autonomous vehicle in Trondheim fjord.
+connected to the properties of water masses. The characterization of
+uncertainties in random sets is extended with new results for the EIBV
+uncertainty reduction achieved by spatial sampling design. This is
+provided in closed-form for the situation with a static design, and
+then extended to the adaptive situation. The sequential derivations
+provide new insights into efficient applications of adaptive sampling,
+as demonstrated in our application.
 
@@
 phenomena, possibly feature-based mixtures of GPs which could still be
-run onboard and augmented by dynamical models.
-
-The spatial-statistical design criterion, building on random sets is relevant
-in our setting with different water mass properties, and we show mathematical generality beyond the EIBV which is convenient to implement for onboard planning. Such criteria could also
-be useful in other oceanographic settings as for algal-blooms or
-open water fronts.
-Of course, other criteria are also relevant for, possibly hybrid criteria or multi-attribute criteria that could balance goals of exploration and exploitation in this situation.
-One further has cases where there is a clear downstream
-decision-making, for instance managers and regulators that must make difficult decisions related to fish farming, other marine resources
-operations, or environmental projects. Value of information analysis
-(VOI) \citep{Eidsvik:15} could be used in a similar vein as the IBV in
-the current paper, as both analytic methods explore when, where and
-which information is likely to result in improved decision-making.
-
-While more effort could be spent on approximating the look-ahead stages
-in the adaptive designs, it is perhaps more interesting to explore the
-additional flexibility that can be gained by having two or more AUVs
-co-temporally exploring a spatial or spatio-temporal domain together
-\citep{ferreira2019advancing}. Such an approach would enable
+run onboard and augmented by dynamical models. \kc{How about talking
+  about removing any grids or waypoints altogether?}
+
+The spatial-statistical design criterion, building on random sets is
+relevant in our setting with different water properties, and we show
+mathematical generality beyond the EIBV which is convenient to
+implement for onboard robotic planning. Such criteria could also be
+useful in other oceanographic settings as for algal-blooms, anoxic
+zones or open water fronts \cite{costa19}. Of course, other criteria
+are also relevant for instance, hybrid or multi-attribute criteria
+that could balance goals of exploration and exploitation in this
+situation. Equally, such techniques have significant use cases in
+downstream decision-making, with policy makers and regulators who need
+to make difficult decisions related to fish farming, other marine
+resource operations. Value of information analysis (VOI)
+\citep{Eidsvik:15} could be used in a similar vein as the IBV in the
+current work, as both analytic methods explore when, where and which
+information is likely to result in improved decision-making.
+
+While more effort could be spent on approximating the look-ahead
+stages in the adaptive designs, it is perhaps more interesting to
+explore the additional flexibility that can be gained by having
+multiple AUVs co-temporally exploring a spatial or spatio-temporal
+domain \citep{ferreira2019advancing}. Such an approach would enable
 concurrent sampling in different parts of the space, or opportunities
@@
 
-TOF and KR acknowledge support from the Centre for Autonomous Marine
+TOF acknowledges support from the Centre for Autonomous Marine
 Operations and Systems
@@
 Excellence, project number 223254, and the Applied Underwater Robotics
-Labortatory (AURLab). JE and KR acknowledges support from Norwegian research
-council, project number 305445. CT and DG acknowledges support from the Swiss National Science Foundation, project number 178858.
+Labortatory (AURLab). JE and KR acknowledge support from Norwegian
+research council (RCN), project number 305445. CT and DG acknowledge
+support from the Swiss National Science Foundation, project number
+178858.
 

### 2020-06-05 main_new.tex
Jo incorporated some changes in main file.

@@
 
-There are several challenges with sampling with autonomous platforms.
-\kc{You seem to jump the gun in stating a challenge without
-  connecting it to the sampling problem. Introduce a sentence as to
-  why models (as a proxy) are important for \emph{understanding} ocean
-  processes} First, full numerical ocean models based on complex
+The
+advent of robust mobile robotic platforms \citep{Bellingham07} has
+resulted in significant contributions to environmental monitoring and
+sampling in the ocean. In particular, autonomous underwater vehicles (AUVs), have
+advanced the state of sampling and consequently have made robotics an
+integral part of ocean observation; our previous work has contributed
+to this effort \citep{das11b,Das2015,fossuminformation,fossum18b}.
+
+The potential that lies in using information-based sampling methods, which we
+highlight in this paper, is to do efficient environmental sampling and
+overcome common challenges as discussed next. First, full numerical ocean models based on complex
 differential equations cannot be run onboard a marine robot with
@@
 % required.
-Hence statistical proxy models of the environment must be
-used. For our purpose, we rely on Gaussian process (GP)
-representations of the ocean variables of interest.\kc{Sorry....but
-  this makes no sense whatsoever. What is the connection of you using
-  a GP with sampling? You need to make a case first as to the
-  viability of using GP in a marine robot first as a potential viable
-  option, before you state why you want to use a GP}. Second, with
-undersampling issues \kc{what issues? If there are issues, you state
-  them first and then suggest solutions.} one must usually plan for
+Hence statistical proxy models of the environment must be used. For our purpose, we rely on Gaussian process (GP)
+representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
+Second, because of limited resources to sample the large oceanographic domains, one must plan for
 active learning during the operation, where a relevant criterion is
@@
 
-With this in mind, we have designed AUV sampling strategies for
-reducing the spatially-integrated Bernoulli variance (IBV) in the ES
+With this in mind, we design AUV sampling strategies for reducing the spatially-integrated Bernoulli variance (IBV) in the ES
 indicator variables. 
@@
 
+Other statistical
+work in the oceanographic domain include \cite{wikle2013modern}
+focusing on hierarchical statistical models; \cite{sahu2008space},
+studying spatio-temporal models for sea surface temperature and
+salinity data; and \cite{mellucci2018oceanic} looking at the
+statistical prediction of features using an underwater glider.
+In this paper the focus is not on statistical modeling, but rather on statistical results and computations for efficient sampling designs.
+
 The remainder of this paper is organized as follows: Section
@@
 ocean models similarly find it challenging to provide detail at fine
-scale \citep{Lermusiaux:2006}. The advent of mobile robotic platforms
-\citep{Bellingham07} has resulted in significant contributions to
-environmental monitoring and sampling. In particular, AUVs have
+scale \citep{Lermusiaux:2006}. Mobile robotic platforms
+\citep{Bellingham07} have contributed significantly to environmental monitoring and sampling. In particular, AUVs have
 advanced the state of in-situ sampling, and made robotics an integral
@@
 reduction are independent of the actual data realizations under the
-assumptions of GP models, so limited decisional flexibility is created
-for sequential designs \kc{should this be, ``...so it has limited
-  decisional flexibility''?}. The use of data-driven adaptive criteria
+assumptions of GP models, so it has limited
+decisional flexibility. The use of data-driven adaptive criteria
 was introduced to include more targeted sampling of regions of
@@
 
-The definitions of the sets of interest are introduced in Section \ref{sec:ES}, while the closed-form expressions for the expected uncertainty reduction are derived in Section \ref{sec:EIBV}.
+We define the random sets of interest in Section \ref{sec:ES}, while the closed-form expressions for the expected uncertainty reduction are derived in Section \ref{sec:EIBV}.
 Section \ref{Sec:UnivarEx} provides an instructive bivariate example relevant for sampling in the temperature and salinity application.
@@
 set (typically in geostatistical studies, a geographical subset of interest) and the
-dimensionality of the output $\no \geq 1$ is arbitrary \kc{Is the
-  definition of such a 'p' and its dimensionality, obvious? This to me
-  springs out of nowhere.}. Assume furthermore that the set of interest
+dimensionality of the output $\no \geq 1$ is arbitrary. We have $\no = 2$ in the example with temperature and salinity in the river plume.
+
+Assume furthermore that the set of interest
 is characterized in terms of a subset of values of the variables, say
@@
 Our goal here is to develop approaches to quantify and improve the
-characterization of uncertainties on such sets \kc{$\es$?} in the
+characterization of uncertainties on such sets $\es$, in the
 context of multivariate random fields.
@@
 
-We let $d \in \mathcal{D}$ denote a concatenation of sampling designs with $q$ sampling locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$. The
+We denote a design by $d \in \mathcal{D}$, where the concatenation of designs with $q$ sampling locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$ is $\mathcal{D}$. The
 measurement model is defined by 
@@
 Here, $\bz(\bu_d)$ is the size $pq$ vector of GP variables at the
-sampling locations, $G_d$ is the design matrix which in principle
+sampling locations, and $G_d$ is the design matrix which in principle
 allows for observing subsets of the vector-valued field at some of the
@@
 is the realized vector of observed responses, and
-$\lambda(\x,\bu_d)=K^{-1}_y(\bu_d,\bu_d) k(\bu_d, \bu)$. Here,
+$\lambda(\x,\bu_d)=K^{-1}_y k(\bu_d, \bu)$. Here,
 the data covariance is
@@
 We calculate EPs and the Bernoulli variance for a few bivariate
-Gaussian distributions to illustrate the concepts articulated above.
+Gaussian distributions for temperature and salinity to illustrate the concepts articulated above.
+The parameter inputs are made up, but they are in a realistic range as we have in the simulation study and real-world example below.
+The thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
+
 Fig. \ref{illus_bivarDens} shows contour plots of three different
 densities with increasing correlation $\gamma$ between temperature and
-salinity. \kc{What is the source of this data and this figure,
-  i.e. how did you derive this plot?}
+salinity. 
 \begin{figure}[h!] \centering
@@
 \end{figure}
-Here, the thresholds are set equal to the mean $\mu_1=t_1=5^o C$, and
-salinity $\mu_2=t_2=30$ mg/l, and the EP is $p=P(Z_1 < 5,Z_2 < 30)$.
 The displayed densities have unit standard deviations for both
@@
 \ref{illus_bivarDens}. The EPs increase with the correlation as there
-is a strong tendency to have concurrently low temperature and salinity
-\kc{Same question as above. What is the basis for this correlation? Is
-  this just an example or derived from actual data in the mixed layer
-  in the front?}. The Bernoulli variance is similarly large for high
+is a strong tendency to have concurrently low temperature and salinity. The Bernoulli variance is similarly large for high
 correlations. EPs and Bernoulli variances are the same for standard
@@
 gathering both data types, and for a design with temperature
-measurements alone. Both data are modeled by
-$(Y_1,Y_2)^t=(Z_1,Z_2)^t+N(0,0.5^2I_2)$, while $Y_2=Z_2+N(0,0.5^2)$
-when only temperature is measured \kc{I find this notation somewhat
-  confusing. Earlier you had set up t1 and t2 to denote temp and
-  salinity respectively. It might help to have \emph{some} notational
-  similarity with that; otherwise it becomes mind bending I believe.}.
+measurements alone. When both data are gathered, the measurement model is
+$(Y_{d,1},Y_{d,2})^t=(Z_1,Z_2)^t+\bepsilon$, with $\bepsilon \sim N(0,0.5^2I_2)$, while $Y_d=Z_1+\epsilon$, $\epsilon \sim N(0,0.5^2)$ when only temperature is measured.
 For this illustration, Table \ref{tab:sim_rhoab} shows that the
@@
 
-\kc{I feel this section above, which was present in the earlier
-  submitted version too, but to which I did not comment, is
-  significant as an example to the reader. But, as noted above, I
-  don't understand the provenance of the data. It isn't made up,
-  right?}
-
 %\subsection{Expected classification criteria}
@@
 Let $d^{j,s}$ denote design number $j$ at stage $s$ of a sequential
-survey. If this design is selected, data $\by_{d^{j,s}}$ will be
-gathered.  We further let
-$\mathcal{Y}^{s-1} =\{(d^{j,1}, \by_{d^{j,1}}),\ldots, (d^{j,s},
-\by_{d^{j,s}}) \}$ denote the set of all data gathered until stage
-$s-1$ according to selected design $d^{j,1},\ldots,d^{j,s}$.  This
-defines the AUV position at the stage $s$.  The data further defines
+survey.
+For the AUV sampling situation, the choice of design entails a commitment to one of many possible navigation directions on the waypoint graph.
+Once the choice has been made, the AUV steers in the chosen direction and this entails sampling locations $\bu_{d^{j,s}}$ and data $\by_{d^{j,s}}$ is gathered.
+We let
+$\mathcal{Y}^{s-1} =\{(d^{*,1}, \by_{d^{*,1}}),\ldots, (d^{*,s-1},
+\by_{d^{*,s-1}}) \}$ denote the set of all data gathered until stage
+$s-1$ according to selected design $d^{*,1},\ldots,d^{*,s-1}$.
+
+The design and data define
 the conditional GP with mean
-$\bbm^{s-1}(\x)=E(\gp[\x]|\mathcal{Y}^{s-1})$ from
-Eq. (\ref{eq:cokrig}) and (cross) covariance matrix
-$\Sigma^{s-1}=\mbox{Var}(\gp[\x]|\mathcal{Y}^{s-1})$ from
-Eq. (\ref{eq:krig:cov}).
-
-At the next stage, we choose where measurements $\by_{d^{j,s}}$ should
-be gathered with design $d^{j,s}$. For the triangularized waypoint
-graph that we use in our application, there are six possible
-directions for $j$, and we must evaluate the design criterion for all
-these.  The goal of the sampling design is to minimize the expected
-uncertainty of random sets, and specifically the EIBV for our
+$E(\gp[\x]|\mathcal{Y}^{s-1})$ and (cross) covariance matrix
+$\mbox{Var}(\gp[\x]|\mathcal{Y}^{s-1})$.
+These are efficiently updated in a sequential manner keeping track of the mean and covariance kernel of the GP over the stages.
+As extensions of Eq.
+\eqref{eq:cokrig} and \eqref{eq:krig:cov}, we have
+\begin{eqnarray}\label{eq:krigSUR}
+\mu^{s}(\x) &=&\mu^{s-1}(\x)+\lambda^{[s,s]}(\x,\bu_{d^{j,s}})^T (\by_{d^{j,s}}-G_{d^{j,s}} \mu(\x)), \nonumber \\
+k^{s}(\x,\x') &=& k^{s-1}(\x,\x')-\lambda^{[s,s]}(\x,\bu_{d^{j,s}})^T K_{Y_{d^{j,s}}} \lambda^{[s,s]}(\x',\bu_{d^{j,s}}),
+\end{eqnarray}
+where $\lambda^{[s,s]}(\x,\bu_{d^{j,s}})=K^{-1}_{Y_{d^{j,s}}} k(\bu_{d^{j,s}},\x)$ and the data covariance is $K^{-1}_{Y_{d^{j,s}}}=G_{d^{j,s}} K^s(\bu_{d^{j,s}},\bu_{d^{j,s}})G^T_{d^{j,s}}+R_{d^{j,s}}$, with measurement model and design matrix defined as in equation \eqref{eq:meas}
+
+At stage $s$, the robot must then choose design $d^{j,s}$ where measurements $\by_{d^{j,s}}$ should
+be gathered. For the triangularized waypoint
+graph all six possible
+directions must be evaluated to minimize the expected uncertainty of random sets, and in particular the EIBV for our 
 application. We hence focus on choosing designs for data modeled by
@@
 variables, but is instead applicable only if one design (node) has EPs
-closer to $0.5$ (see Table \ref{tab:sim_rhoab}, line \kc{row?}
+closer to $0.5$ (see Table \ref{tab:sim_rhoab}, row
 two). Additionally, this strategy does not account for spatial
@@
 the expectation is conditional to this data. A similar closed-form
-calculation for EIBV is hence applicable in Eq.  \eqref{critSEQ}
-\kc{Strangely in my TeX compilation this points to the earlier
-  definition of $d^{*,s}$. Seems like some index is messed up in the
-  sources; all indexes should perhaps be checked?} using the updated
-GP model from stage $s-1$. Once the data is collected for the best
-\kc{??}  design at stage $s$, the GP model is updated again. The mean
-$\bbm^{s}$ and covariance matrix $\Sigma^{s}$ are used to compute the
-next design at stage $s+1$, and so on.
+calculation for EIBV is hence applicable in Eq.  \eqref{critSEQ} using the updated
+GP model from stage $s-1$ as defined via Eq.
+\eqref{eq:krigSUR} in the results from Proposition \ref{propo2}.
+Once the data are collected for the best design $d^{*s}$ at stage $s$, the GP model is updated again.
+The mean and covariance are used to compute the next design at stage $s+1$, and so on.
 
@@
 
-We now extend the myopic strategy to a look-ahead strategy, which is
-optimal when one can gather only two more stages of
-measurements. \kc{How can you be sure of this?} In addition to the
-next stage of measurements $\by^{j,s}$, this look-ahead strategy,
-anticipates the subsequent design indexed $j_2$ with data
+We now extend the myopic strategy to a look-ahead strategy. We consider two stages of
+measurements, and the strategy is then optimal in the sense that it accounts consistently for the expectations and minimizations in these two stages, but still without including any planning beyond those two steps.
+In addition to the
+next stage of measurements $\by^{j,s}$, this look-ahead strategy anticipates the subsequent design indexed $j_2$ with data
 $\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$. The
@@
 calculated with respect to the conditional densities given data at
-that stage, implying that we have two levels of intermixed
-minimizations and integrals.  We solve the first expectation by Monte
+that stage.  We solve the first expectation by Monte
 Carlo sampling of data $\by_{d^{j,s}}$ from its conditional
 distribution. For each of these data samples, the second expectation
-is solved using the closed-form expressions for EIBV from Section
-\ref{sec:ESEP}.
+is solved using the closed-form expressions for EIBV using the results from Proposition \ref{propo2}, now with conditioning to the first stage data already going into the mean and covariances in Eq.
+\eqref{eq:krigSUR}.
 
 Even though the strategy looks at two sampling stages, it is only used to find
-the current best design. When data are collected, the GP model is
-updated, and the mean $\bbm^{s}$ and covariance matrix $\Sigma^{s}$ are
-used to compute the next design at stage $s+1$, now anticipating what
-stage $s+2$ could offer, and so on.
+the current best design. When data are collected at stage $s$, the GP model is
+updated as in Eqn. \eqref{eq:krigSUR}, and the mean and (cross) covariance are
+used to compute the next design for stage $s+1$, now anticipating what
+stage $s+1$ and $s+2$ could offer, and so on.
 This look-ahead approach is much more computationally demanding than
-the myopic strategy, so for practical implementation, we prune paths
-in the evaluation of Eq. \eqref{critLA}. This means that we use the myopic
+the myopic strategy, so for practical implementation onboard the AUV this is reaching the limits allowed for calculation.
+We conduct some pruning of the least valuable design paths in the evaluation of Eq.
+\eqref{critLA}, to get a satisfactory computation time below one minute per waypoint.
+This means that we use the myopic
 strategy to rank the three best designs following the first stage alone, and

### 2020-06-08 main_new.tex
Updates from Overleaf

@@
 
-% "Title of the paper"
-\title{Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields}
-\runtitle{Excursion Probabilities for Informative Sampling}
+% "New Title of the paper"
+\title{Autonomous Oceanographic Data Collection via Uncertainty Reduction for  Excursion Sets of Vector-valued Random Fields} 
+\runtitle{Autonomous Oceanographic Uncertainty Reduction}
+
+% "Former Title of the paper" (DG 08.06.2020)
+%\title{Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields} \runtitle{Excursion Probabilities for Informative Sampling}
 
@@
 \affiliation[t1]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
-\affiliation[t2]{Centre for Autonomous Marine Operations and Systems (AMOS).}
+\affiliation[t2]{Centre for Autonomous Marine Operations and Systems, NTNU.}
+% (AMOS).} Is the acronym needed now that this affiliation appears once only?
+% Added NTNU, though. 
 \affiliation[t3]{Institute of Mathematical Statistics and Actuarial Science, University of Bern, Switzerland.}
@@
 \end{frontmatter}
-
-\section{Introduction}
+\section{Wannabe new introduction structure}
+
+NB: The subsections are used as a temporary instrument to highlight the proposed structure and should probably be abolished upon convergence to a stable version.
+
+\subsection{Oceanic data collection and spatial design of experiments}
+
+\begin{itemize}
+    \item Oceanic data collection as a timely and crucial challenge 
+    \item Uneasy to measure state variables varying continuously across space
+    \item Some lend themselves well to random field modelling 
+    \item Where to make measurements is a question related to experimental design
+    \item Yet with AUVs measurements ought to be made along a trajectory 
+    \item Furthermore, in several problems such as mapping Fjord versus ocean water, the goal is to reduce uncertainties on regions implicitly defined by several state variables such as temperature and salinity.
+    \item Besides this, AUV are able to revise their trajectory after intermediate measurements, calling for experimental design strategies that are adaptive and ideally also account for multiple steps ahead.
+    \item All this in the framework where several responses are modelled within a vector-valued random field framework and under observation noise calls for sophisticated design criteria and computational strategies for their optimization.
+    \item Here we aim to leverage recent progress in sequential design of computer experiments and adapt them to the specifics of guiding AUV measurements for parcimoniously distinguishing fjord from ocean waters.
+    \item While our problem involves multivariate observational responses to be observed along a continuous trajectory, our approach will take substantial advantage of existing work in the field of excursion set estimation mainly dedicated to computer experiments in the scalar-valued case, and where space exploration can be performed without constraints regarding the distance between successive design points.
+    \item Let us briefly review random field modelling and recent advances in sequential design of experiments before detailing further our proposed approach and outline.  
+\end{itemize}
+
+\subsection{Random field modelling and sequential design of experiments}
+
+\begin{itemize}
+    \item 
+\end{itemize}
+
+\subsection{Proposed approach in a nutshell}
+
+\begin{itemize}
+    \item 
+\end{itemize}
+
+\subsection{Outline of the paper and its appendices / supplemental material}
+
+\begin{itemize}
+    \item 
+\end{itemize}
+
+%\subsection{}
+
+%\begin{itemize}
+%    \item 
+%\end{itemize}
+
+\section{Former Introduction (Before 08.06.2020)}
 

### 2020-06-09 main_new.tex
Updates from Overleaf

--- 2020-06-05 main_new.tex
@@
 
-% "Title of the paper"
-\title{Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields}
-\runtitle{Excursion Probabilities for Informative Sampling}
+% "New Title of the paper"
+\title{Autonomous Oceanographic Data Collection towards Estimating Excursion Sets of Vector-valued Random Fields} 
+\runtitle{Autonomous Oceanographic Data Collection}
+
+% "Former Title of the paper" (DG 08.06.2020)
+%\title{Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields} \runtitle{Excursion Probabilities for Informative Sampling}
 
@@
 \author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e3]{jo.eidsvik@ntnu.no}}
-\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e4]{david.ginsbourger@idiap.ch}}
+\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e4]{david.ginsbourger@stat.unibe.ch}}
 \and
@@
 \affiliation[t1]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
-\affiliation[t2]{Centre for Autonomous Marine Operations and Systems (AMOS).}
+\affiliation[t2]{Centre for Autonomous Marine Operations and Systems, NTNU.}
 \affiliation[t3]{Institute of Mathematical Statistics and Actuarial Science, University of Bern, Switzerland.}
@@
 \end{frontmatter}
-
 \section{Introduction}
 
+%NB: The subsections are used as a temporary instrument to highlight the proposed structure and should probably be abolished upon convergence to a stable version.
+
+\subsection{Oceanic data collection and spatial design of experiments}
+
+%\textcolor{blue}{NB: I started to match previously listed thoughts/ideas with pieces of the former intro.
+Will do more asap.
+Also, Section 3 I find excellent (and more precise than the intro) and start thinking that some bits might end up in the newer intro (while some others could land in a later section).}
+
+Monitoring the world's oceans has gained increased importance in light of the changing climate and increasing anthropogenic impact. 
+\begin{comment} % TO BE RE-INSERTED LATER IN THE ETXT?
+A central problem for understanding these factors is the lack of representative data with sufficient resolution.
+Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of sampling.
+By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
+This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
+\end{comment}
+Central to understanding the changes taking place in the upper
+water-column is knowledge of the bio-geophysical interaction driven by
+an agglomeration of physical forcings (e.g. wind, topography,
+bathymetry, tidal influences, etc.) and incipient micro-biology driven
+by planktonic and coastal anthropogenic input, such as pollution and
+agricultural runoff transported into the ocean by rivers and streams.
+These often result in a range of ecosystem-related phenomena such as
+blooms and plumes, with direct and indirect effects on society. \kc{we
+  should look for a citation here. Check with John Ryan}
+ One of the bottlenecks in the study of such phenomena lies however in the lack of obervational data with sufficient resolution.
+Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of data collection.
+Based on collected data and sometimes also of prior knowledge, models from spatial statistics can be used to make predictions of the state variables under consideration over regions of interest.
+In case of multiple continuous state variables being jointly studied relying on spatially scattered data, vector-valued random field models and associated prediction approaches such as co-Kriging (REFs) can be considered nowadays as part of the standard statistical toolkit in oceanographic research and beyond.
+%applications and beyond.
+ %By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
+This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
+
+Traditional data collection to understand natural and artificial phenomena at sea has typically been based on static buoys, floats, or ship-based methods,
+with significant logistical limitations that directly impact coverage
+and sampling resolution. Modern methods using satellite remote-sensing
+provide large-scale coverage but have limited resolution, are limited
+to sensing the surface, and are impacted by cloud cover. Numerical
+ocean models similarly find it challenging to provide detail at fine
+scale \citep{Lermusiaux:2006}, and also come with computation costs that can be practically limiting. 
+%limiting computing resources, no time for hard-core ocean model Navier-Stokes, hence GPs and approximations, etc.
+%
+\begin{figure}[!h] 
+  \centering 
+  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
+    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
+  \hfill
+  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
+    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
+  \caption{\ref{fig:envir1} Traditional ocean observation based on 
+    ship-based sampling has been augmented by autonomous
+    robotic vehicles. % and their interactions.  
+    AUV platforms are an integral part of this network being able to
+    reason and make decisions for efficient onboard adaptive sampling.
+    % using the sense-plan-act control approach to autonomous control. 
+    \ref{fig:nidelven} The interaction of river and ocean creates
+    processes that are challenging to map, where the combination of
+    statistics and robotics can play a vital role in enabling more
+    effective oceanographic observation.\kc{The inset ``AUV robot''
+      looks very tacky and ad-hoc; might be best to remove it. Change
+      ``ships'' to
+      ``Ship'', remove the aerial drone, since we didn't use it and
+      don't plan to in this work.}}
+  \label{fig:envir} \end{figure}
+%
+\begin{comment} % Somehow redundant with the parag following below
+Mobile robotic platforms \citep{Bellingham07} have contributed significantly to environmental monitoring and sampling.
+In particular, AUVs have advanced the state of in-situ sampling, and made robotics an integral part of ocean observation, filling parts of the undersampling gap \cite{rudnick03,rudnick18} (See Fig.
+\ref{fig:envir1}).
+\end{comment}
+%
+The advent of robust mobile robotic platforms \citep{Bellingham07} has
+resulted in significant contributions to environmental monitoring and
+sampling in the ocean. In particular, autonomous underwater vehicles (AUVs) have advanced the state of data collection and consequently have made robotics an integral part of ocean observation; %our 
+previous work by some of the authors of the present paper have contributed to this effort \citep{das11b,Das2015,fossuminformation,fossum18b}.
+
+As full numerical ocean models based on complex differential equations cannot be run onboard a marine robot with limited computational capacity, statistical models relying on collected data and random field assumptions appear quite relevant as means to guide AUV data collection trajectories.
+Our main research angle in the present work is to extend sequential design strategies from the world of spatial statistics and computer experiments to frameworks featuring both vector-valued observational data and experimental designs that may be constrained by the nature of measurement paths.
+\\
+
+Surveys with AUVs are usually limited to observations along fixed
+transects that are pre-scripted in mission plans created manually by a
+human operator. Missions can be specified operating on a scale of
+hundreds of meters to tens of kilometers depending on the scientific
+context. Faced with limited coverage capacity, a more effective
+approach is to instead use onboard algorithms to continuously
+evaluate, update, and refine future sampling locations (sense-plan-act
+cycle in Fig. \ref{fig:envir1}), making the information gathering
+\emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
+% This usually occurs on a spatial grid, called a waypoint graph. 
+In doing so, the space of sampling opportunities is still limited by
+the so-called waypoint graph, a discretization of the search domain, but the AUV has flexibility to modify its path at each waypoint based on in-situ sensing and measurements onboard \citep{py10,Rajan12,Rajan12b}.\\
+
+The work presented here is primarily inspired by a cast study 
+%application where an AUV is used to measure both temperature and salinity in 
+pertaining to the spatial characterization of a frontal system generated by
+river plumes. Fig. \ref{fig:nidelven} shows the survey area in
+Trondheim, Norway, where cold freshwater enters from the river,
+creating a strong gradient in both temperature and salinity. Because
+of the local topography and the Coriolis force the cold fresh water
+tends to flow to the east. Depending on the river discharge, tidal
+effects, wind, and temperature differences, this boundary often gets
+distorted. Initial knowledge about the location and evolution of these
+features are highly uncertain, making deterministic planning
+challenging.
+Here the goal is to use AUV measurements towards a better identification of freshwater versus oceanic water, with the help of a vector-valued random field model for both temperature and salinity.
+As we will present in the next sections, some of the original contributions inspired by this specific problem are actually applicable to further settings and application areas.
+\\
+
+The questions tackled here hence pertain to the broader area of spatial data collection and experimental design, yet with the particularity of estimating some regions of the domain, mostly excursion sets, being implicitly defined by a scarcely observed vector-valued random field.
+%because of limited resources to sample the large oceanographic domains, one must plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs.
+The acquired in-situ information must then be assimilated into the statistical model onboard the platform, so that it can be used to re-compute the sampling criterion and in this way inform decisions on where to sample next.
+\\
+Moreover, given constraints on AUV movements and the fact that surveya rely by design in such a framework on successive measurements spread along a trajectory, addressing corresponding design problems calls for particular sequential strategies, possibly also accounting for the finiteness of resources by anticipating multiple steps ahead thereby improving over myopic strategies.
+\\
+
+Here we aim to leverage and extent recent progress in uncertainty quantification and reduction for excursion sets of Gaussian random fields in order to address vector-valued observational cases under constrained design such as in the motivating application of designing AUV trajectories to better distinguish between freshwater from ocean waters by relying on temperature and salinity measurements.
+%approach will take substantial advantage of existing work in the field of excursion set estimation mainly dedicated to computer experiments in the scalar-valued case, and where space exploration can be performed without constraints regarding the distance between successive design points.
+Original contributions compared to recent related work in the framework of scalar-valued computer experiments include: 
+\begin{itemize}
+    \item Extension of two important Sequential Uncertainty Reduction criteria of (REFs) to homotopic or heterotopic vector-valued settings,
+    \item Investigation of myopic and multiple-step ahead batch-sequential strategies for optimizing trajectories with respect to such criteria,
+    \item Replicable experiments with synthetic test cases and original data sets with accompagnying codes 
+\end{itemize}
+Let us briefly review random field modelling and recent advances in targetd sequential design of experiments based on Gaussian Processes before detailing further existing literature tackling the considered problems, our proposed approach, as well as outlining the rest of the paper.
+
+\subsection{Random field modelling and targeted sequential design of experiments}
+
+\begin{itemize}
+    \item 
+\end{itemize}
+
+% From earlier version; to be relocated?
+%Hence statistical proxy models of the environment must be used.
+For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
+\subsection{Related work and proposed approach in a nutshell}
+
+\begin{itemize}
+    \item 
+\end{itemize}
+
+\textcolor{blue}{We might want to re-use some of the following (from earlier version):}
+
+A major contribution of our work is to derive closed-form results for
+this design criteria for situations where the underlying model is
+based on multivariate GPs. Another contribution is that we embed this
+approach in a sequential strategy to produce sampling algorithms, and
+apply this to a real-world scenario of autonomously sampling
+temperature and salinity gradients using an AUV, with the intention of
+characterizing the variability of oceanic and riverine waters in
+typical mixed coastal environments.
+
+Motivating examples relevant for ESs of multivariate processes are
+abundant. In medicine, doctors do not rely solely on a single symptom
+but must see several combined effects before making a diagnosis. In
+our context of environmental sampling, the salinity and temperature
+excursions of a river plume can similarly help characterize the
+underlying bio-geochemical processes
+\citep{hopkins2013detection,Pinto2018}, where sampling can be aimed
+towards reducing uncertainty in the temperature and salinity ES.
+Unlike what has been done in previous plume exploration studies, we
+emphasize the links to spatial statistical modeling and targeted
+multivariate sampling criteria for sequential sampling.
+
+Other statistical
+work in the oceanographic domain include \cite{wikle2013modern}
+focusing on hierarchical statistical models; \cite{sahu2008space},
+studying spatio-temporal models for sea surface temperature and
+salinity data; and \cite{mellucci2018oceanic} looking at the
+statistical prediction of features using an underwater glider.
+In this paper the focus is not on statistical modeling, but rather on statistical results and computations for efficient sampling designs.
+
+We focus on spatial characterization of a frontal system generated by
+river plumes. Fig. \ref{fig:nidelven} shows the survey area in
+Trondheim, Norway, where cold freshwater enters from the river,
+creating a strong gradient in both temperature and salinity. Because
+of the local topography and the Coriolis force the cold fresh water
+tends to flow to the east. Depending on the river discharge, tidal
+effects, wind, and temperature differences, this boundary often gets
+distorted. Initial knowledge about the location and evolution of these
+features are highly uncertain, making deterministic planning
+challenging.
+
+Adaptive in-situ AUV sampling of an evolving frontal feature has been
+explored in \cite{fronts11,Zhang2012,Pinto2018,costa19}. These
+approaches typically use a reactive-adaptive scheme, whereby
+exploration does not rely on a statistical model of the environment,
+but rather adaptation is based on closing the sensing and actuation
+loop. Myopic sampling, i.e. stage-wise selection of the path (on the
+waypoint graph), has been used for surveys
+\citep{singh2009efficient,Binney2013} that focus largely on reducing
+predictive variance or entropy. These criteria are widely adopted in
+the statistics literature on spatio-temporal design as well
+\cite{bueso1998state,zidek2019monitoring}, but variance and entropy
+reduction are independent of the actual data realizations under the
+assumptions of GP models, so it has limited
+decisional flexibility. The use of data-driven adaptive criteria
+was introduced to include more targeted sampling of regions of
+scientific interest in \cite{Low2009} and \cite{fossuminformation}. In
+this paper, we focus on mapping the river plume by rewarding designs
+that improve the classification of ES in temperature and salinity.
+
+\subsection{Outline of the paper and its appendices / supplemental material}
+
+\begin{itemize}
+    \item 
+\end{itemize}
+
+The remainder of this paper is organized as follows: 
+%Section \ref{sec:bg} provides select background on ocean sampling. 
+Section \ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for
+vector-valued GPs. Section \ref{sec:heuristics} builds on these
+assumptions when deriving the sequential design criteria for adaptive
+sampling. Section \ref{sec:simulations} discusses properties of the
+methods in simulation studies. Section \ref{sec:case_study}
+demonstrates the methodology used in field work characterizing a river
+plume and finally, Section \ref{sec:concl_disc} contains a summary and
+a discussion of future work.
+
+%\subsection{}
+
+%\begin{itemize}
+%    \item 
+%\end{itemize}
+
+\begin{comment}
+\section{Former Introduction (Before 08.06.2020)}
+
 Monitoring the world's oceans has gained increased importance in light
-of the changing climate and increasing anthropogenic impact. A central
-problem for understanding these factors is the lack of representative
-data with sufficient resolution. Most of this \emph{undersampling} can
-be attributed to the large spatio-temporal variations in which ocean
-processes transpire, prompting the need for effective means of
-sampling. By \emph{sampling}, we refer to the design of observational
-strategies in the spatial domain, where the use autonomous robotic
-platforms can be combined with statistical methods to pursue
-measurements with high scientific relevance. This combination, of
-statistical tools and robotic platforms, constitute a methodological
-basis for addressing the problem of efficient sampling of the ocean,
-which is at the core of this work.
+of the changing climate and increasing anthropogenic impact. 
+A central problem for understanding these factors is the lack of representative data with sufficient resolution.
+Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of sampling.
+By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
+This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
 
@@
 a discussion of future work.
-
+\end{comment}
+
+\begin{comment}
 \section{Ocean Processes and Observation}
@@
 
+\end{comment}
+
+\subsection{To be relocated elsewhere}
 The AUVs considered here are powered untethered platforms, that
@@
 
+\begin{comment}
 Surveys with AUVs are usually limited to observations along fixed
@@
 that improve the classification of ES in temperature and salinity.
+
+\end{comment}
 
@@
 
-Assuming now that $q$ observations are available, one wishes to
-predict $\gp[\x]$. The multivariate Kriging equations holds with size
+Assuming now that $q$ observations $\by_d$ made at $\bu_d$ are available, one wishes to predict $\gp[\x]$. The multivariate Kriging equations holds with size
 $p$ vector of conditional means:  
@@
 \end{equation}
-where $\bmu$ is size $p$ unconditional mean function of $Z$, $\by_d$
-is the realized vector of observed responses, and
-$\lambda(\x,\bu_d)=K^{-1}_y k(\bu_d, \bu)$. Here,
-the data covariance is
-$K_y=\mbox{Cov}(\bY_d,\bY_d)=G_d K(\bu_d,\bu_d)G_d^T+R_d$ while
-the cross-covariance is $\mbox{Cov}(\bY_d,Z(\x))=G_d k(\bu_d, \x)$, assuming the measurement
+where
+$\lambda(\x,\bu_d)=K^{-1}_y G_d k(\bu_d, \bu) $ with data covariance
+$K_y=\mbox{Cov}(\bY_d,\bY_d)=G_d K(\bu_d,\bu_d)G_d^T+R_d$ and cross-covariance $\mbox{Cov}(\bY_d,Z(\x))=G_d k(\bu_d, \x)$, assuming the measurement
 error is independent of the random fields.  The Kriging (cross-)covariance function can also be expressed in the same vein via
@@
   chevalier2014fast,stroh}, but not generalized to our situation with
-random sets for vector-valued GPs. Notably, the mean in equation
-\eqref{eq:cokrig}, is in the form of $\ba+B V$, where
+random sets for vector-valued GPs. Notably, the entry in Eq.
+\eqref{eq:post_ep} is in the form of $\ba+B V$, where
 $V \sim N(0_q, K_y (\bu_d,\bu_d))$,
-$\ba=(t_1,t_2)^t-\bmu(\x)+\lambda(\x,\bu_d)^T G_d \bmu(\bu_d)$
-and $B=\lambda(\x,\bu_d)^T$. The following then applies to give a
+$\ba=(t_1,t_2)^t-\bmu(\x)$
+and $B=\lambda(\x,\bu_d)^T$. The following proposition then applies to give a
 closed-form solution to EIBV calculation in this case.
@@
 \end{eqnarray}
-where $\lambda^{[s,s]}(\x,\bu_{d^{j,s}})=K^{-1}_{Y_{d^{j,s}}} k(\bu_{d^{j,s}},\x)$ and the data covariance is $K^{-1}_{Y_{d^{j,s}}}=G_{d^{j,s}} K^s(\bu_{d^{j,s}},\bu_{d^{j,s}})G^T_{d^{j,s}}+R_{d^{j,s}}$, with measurement model and design matrix defined as in equation \eqref{eq:meas}
+where $\lambda^{[s,s]}(\x,\bu_{d^{j,s}})=K^{-1}_{Y_{d^{j,s}}} G_{d^{j,s}} k(\bu_{d^{j,s}},\x)$ and the data covariance is $K^{-1}_{Y_{d^{j,s}}}=G_{d^{j,s}} K^s(\bu_{d^{j,s}},\bu_{d^{j,s}})G^T_{d^{j,s}}+R_{d^{j,s}}$, with measurement model as in Eq.
+\eqref{eq:meas}.
 
@@
 application. We hence focus on choosing designs for data modeled by
-equation \eqref{eq:meas}, to conduct efficient sequential uncertainty
-reduction according to the EIBV criterion.  Notably, compared with the
-static design $d$ in equation \eqref{eq:eibv}, the optimal sequential
+Eq. \eqref{eq:meas}, to conduct efficient sequential uncertainty
+reduction according to the EIBV criterion. Compared with the
+static design $d$ in Eq. \eqref{eq:eibv}, the optimal sequential
 design not only considers the best current design grid point, but also
@@
 the current best design. When data are collected at stage $s$, the GP model is
-updated as in Eqn. \eqref{eq:krigSUR}, and the mean and (cross) covariance are
+updated as in Eq. \eqref{eq:krigSUR}, and the mean and (cross) covariance are
 used to compute the next design for stage $s+1$, now anticipating what

### 2020-06-10 ans_to_reviewers/answers-to-reviewers.tex
Updates from Overleaf

@@
 
-\reply: We elaborated on space-time applications \cite{sigrist2015stochastic,richardson2017sparsity}. TO DO: 
-More complex sampling Look-ahead, going further. 3D sampling, requiring more computationally efficient spatial statistical models.
+\reply: In the revised 'Closing remarks', we elaborate a bit more on space-time modeling including relevant references \cite{sigrist2015stochastic,richardson2017sparsity} and computational approximations required for three-dimensional grid models, on using more complicated criteria for the data collection, and thoughts on approaches for finite horizon optimization in our context with references to approximate dynamic programming \cite{Powell:11}.
+
+In this context we brought up the issue of server communication for the AUV, which could be possible using wifi communication to a mother-ship at waypoints.
+First, this can enable more advanced modeling and computational routines.
+Also, this is a currently a hot topic in control engineering, and we believe there will be a need for statistical approaches for designing efficient sampling strategies for multiple robots in spatio-temporal domains.
 
@@
 
-\reply: This is changed in revised version.
+\reply: This is changed in the revised version.
 
@@
 
-\reply: We show EPs for myopic selection. Also, the accompanying code provides more cases.
+\reply: In the revised version we show EPs for myopic and look-ahead selection in this display.
+
+In the accompanying code there are also resources for displaying other outputs if the reader is interested in comparing results with for instance univariate excursion probabilities.
 

### 2020-06-11 ans_to_reviewers/answers-to-reviewers.tex
Updates from Overleaf

--- 2020-05-29 ans_to_reviewers/answers-to-reviewers.tex
@@
 
-\reply: We elaborated on space-time applications \cite{sigrist2015stochastic,richardson2017sparsity}. TO DO: 
-More complex sampling Look-ahead, going further. 3D sampling, requiring more computationally efficient spatial statistical models.
+\reply: In the revised 'Closing remarks', we elaborate more on computational approximations required for three-dimensional way-point graphs and space-time modeling including relevant references \citep{sigrist2015stochastic,richardson2017sparsity}, on using more complicated criteria for the data collection, and thoughts related to finite horizon optimization where there has been much creative work in the robotics literature including probabilistic road maps and rapidly-exploring random trees \citep{karaman2011sampling} whose statistical properties are not yet clear.
+
+We also strengthen the discussion on communication for the AUV.
+In doing so, we relate the topics of autonomy and communication to the modeling and computational assumptions that applies to this paper.
+With the possibilities of wifi communication to a mother-ship at waypoints, more advanced modeling and computational routines are possible, of course with new challenges related to design and planning for when and where this communication should occur.
+This topic has caught much interest in control engineering \citep{zolich2019survey}, and statistical evaluations could contribute in this setting.
 
@@
 
-\reply: We show EPs for myopic selection. Also, the accompanying code provides more cases.
+\reply: In the revised version we show EPs for a static survey line, myopic and look-ahead selection in three displays.
+
+In the accompanying code there are also resources for displaying other outputs if the reader is interested in comparing results with for instance univariate excursion probabilities.
 

### 2020-06-11 main_new.tex
Updates from Overleaf

@@
 
-% == Enable text degree
+% == Enable text degreehttps://www.overleaf.com/project/5a3268379ecbdc657d8767e8
 \usepackage{textcomp}
@@
 Monitoring the world's oceans has gained increased importance in light of the changing climate and increasing anthropogenic impact. 
-\begin{comment} % TO BE RE-INSERTED LATER IN THE ETXT?
+\begin{comment} % TO BE RE-INSERTED LATER IN THE TEXT?
 A central problem for understanding these factors is the lack of representative data with sufficient resolution.
@@
 Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of data collection.
+%
+ \textcolor{blue}{By \emph{sampling}, we refer here primarily to the design of observational strategies in the spatial domain %where the use autonomous robotic platforms can be combined with statistical methods
+ with the aim to pursue measurements with high scientific relevance}.
 Based on collected data and sometimes also of prior knowledge, models from spatial statistics can be used to make predictions of the state variables under consideration over regions of interest.
-In case of multiple continuous state variables being jointly studied relying on spatially scattered data, vector-valued random field models and associated prediction approaches such as co-Kriging (REFs) can be considered nowadays as part of the standard statistical toolkit in oceanographic research and beyond.
+In case of multiple continuous state variables being jointly studied relying on spatially scattered data, vector-valued random field models and associated prediction approaches such as co-Kriging \citep[See, e.g.,][]{Wackernagel2003} can be considered nowadays as part of the standard statistical toolkit in oceanographic research and beyond.
 %applications and beyond.
@@
 
-Traditional data collection to understand natural and artificial phenomena at sea has typically been based on static buoys, floats, or ship-based methods,
+Data collection to understand natural and artificial phenomena at sea has typically been based on static buoys, floats, or ship-based methods,
 with significant logistical limitations that directly impact coverage
@@
 As full numerical ocean models based on complex differential equations cannot be run onboard a marine robot with limited computational capacity, statistical models relying on collected data and random field assumptions appear quite relevant as means to guide AUV data collection trajectories.
-Our main research angle in the present work is to extend sequential design strategies from the world of spatial statistics and computer experiments to frameworks featuring both vector-valued observational data and experimental designs that may be constrained by the nature of measurement paths.
+Our main research angle in the present work is to extend sequential design strategies from the world of spatial statistics and computer experiments to frameworks featuring both vector-valued observational data and experimental designs that may be constrained by the nature of feasible trajectories.
 \\
@@
 
-The work presented here is primarily inspired by a cast study 
+The work presented here is primarily inspired by a case study 
 %application where an AUV is used to measure both temperature and salinity in 
@@
 \\
-Moreover, given constraints on AUV movements and the fact that surveya rely by design in such a framework on successive measurements spread along a trajectory, addressing corresponding design problems calls for particular sequential strategies, possibly also accounting for the finiteness of resources by anticipating multiple steps ahead thereby improving over myopic strategies.
+Moreover, given constraints on AUV movements and the fact that surveys rely by design in such a framework on successive measurements spread along a trajectory, addressing corresponding design problems calls for particular sequential strategies, possibly also accounting for the finiteness of resources by anticipating multiple steps ahead and thereby improving over myopic strategies.
 \\
 
-Here we aim to leverage and extent recent progress in uncertainty quantification and reduction for excursion sets of Gaussian random fields in order to address vector-valued observational cases under constrained design such as in the motivating application of designing AUV trajectories to better distinguish between freshwater from ocean waters by relying on temperature and salinity measurements.
+Here we aim to leverage and extent recent progress in uncertainty quantification and reduction for excursion sets of Gaussian random fields in order to address vector-valued observational cases under constrained design such as in the motivating application of designing AUV trajectories to better distinguish between freshwater from ocean waters, relying on temperature and salinity measurements.
 %approach will take substantial advantage of existing work in the field of excursion set estimation mainly dedicated to computer experiments in the scalar-valued case, and where space exploration can be performed without constraints regarding the distance between successive design points.
-Original contributions compared to recent related work in the framework of scalar-valued computer experiments include: 
+Original contributions compared to recent related work in the framework of scalar-valued (computer) experiments include: 
 \begin{itemize}
-    \item Extension of two important Sequential Uncertainty Reduction criteria of (REFs) to homotopic or heterotopic vector-valued settings,
+    \item Extension of two important Sequential Uncertainty Reduction criteria of \citep{Bect.etal2012,chevalier2014fast,bect2019} to homotopic or heterotopic vector-valued settings,
     \item Investigation of myopic and multiple-step ahead batch-sequential strategies for optimizing trajectories with respect to such criteria,
-    \item Replicable experiments with synthetic test cases and original data sets with accompagnying codes 
+    \item Replicable experiments with synthetic test cases and original data sets with accompagnying codes. 
 \end{itemize}
@@
 
-\begin{itemize}
-    \item 
-\end{itemize}
+ While random field modelling has been one of the main topics throughout the history of spatial statistics \citep{Krige1951a, Matheron1963, Stein1999, Adler.Taylor2007}, there has recently been a renewed interest for random field models in the context of static sequential experimental design, be it in the context of spatial data collection \citep{Mueller2007} or in computer experiments and machine learning \citep[See for instance][and references therein]{Santner.etal2003, Romero.etal2013, Beck2016}.
+As detailed in \cite{Ginsbourger2018}, Gaussian random field models have been used in particular as a basis to sequential design of comupter experiments dedicated to various goals such as global optimization but also set estimation.
+\cite{Picheny.etal2010} revisited the classical Integrated Mean Square Error criterion with an adaptation dedicated to contour line estimation, \cite{Bect.etal2012} focused on strategies to reduce uncertainties on volumes of excursion, i.e.
+on the measure of the domain of input parameters which associated (scalar) response level exceeds a prescribed threshold, followed up by \cite{chevalier2014fast} presenting computational contributions towards an efficient batch-sequential implementation of strategies introduced in the former.
+Meanwhile, rather than focusing on excursion volumes, some approaches were investigated in \cite{French.Sain2013,Chevalier.etal2013b,Bolin.Lindgren2015,Azzimonti.etal2016} that ambition to estimate sets themselves.
+Recently, sequential design of experiments for the conservative estimation of excursion sets based on Gaussian random field models was presented in \citep{Azzimonti.etal}.
+ %
+Surprisingly less attention has been dedicated to sequential strategies in the case of vector-valued observations to our knowledge.
+It has been acknowledge for several decades that co-Kriging models could be efficiently updated with vector-valued observations being assimilated sequentially
+\cite{Vargas-Guzman1999}, sequential strategies targeted to estimate prescribed features of vector-valued random fields are not in still in their infancy.
+We can refer for instance to \citep{LeGratiet.etal2015} for co-Kriging-based sequential design using fast cross-validation and dedicated to multi-fidelity computer codes.
+Related ideas are to be found in \cite{Poloczek2017} in the context of Multi-Information Source Optimization.
+As for extensions of Stepwise Uncertainty Reduction strategies, to our knowledge the only attempt so far to generalize excursion-related Stepwise Uncertainty Reduction criteria (such as reviewed above) in vector-valued settings pertains to the design of multi-fidelity simulations for fire safety, in the PhD thesis \citep[][p.82]{stroh}, where the outputs are mainly assumed independent yet the feasibility of treating the dependent case by appealing to higher-dimensional Gaussian CDF evaluations is mentioned.
 
@@
 For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
-\subsection{Related work and proposed approach in a nutshell}
-
-\begin{itemize}
-    \item 
-\end{itemize}
-
-\textcolor{blue}{We might want to re-use some of the following (from earlier version):}
+\subsection{Contributions, selected previous work on AUV sampling, and outline} 
+%work and proposed approach in a nutshell}
+%\begin{itemize}
+%    \item 
+%\end{itemize}
+%\textcolor{blue}{We might want to re-use some of the following (from earlier version):}
 
@@
 
+\begin{comment}
 \subsection{Outline of the paper and its appendices / supplemental material}
@@
 \end{itemize}
+\end{comment}
 
@@
 \end{comment}
-
-\subsection{To be relocated elsewhere}
-The AUVs considered here are powered untethered platforms, that
-operate at $1$-$3$ m/s in the upper water column, where they are free
-to move in six degrees of freedom (6 DOF). The in-water operation time
-capability depends on survey speed, payload sensors and navigation;
-typically this amounts to 8-48 hrs. AUVs typically use single-board
-computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA
-Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared
-cache, and 4 GB of 1600 MHz DRAM) for computation onboard.
 
@@
 \citep{py10,Rajan12,Rajan12b}, running an instance of the
-\textit{myopic} strategy from Section \ref{sec:myopic} to control the
+\textit{myopic} strategy from Section \ref{sec:myopic} to 
+control the
 AUV and decide between sampling locations.
+
+\textcolor{blue}{TO BE MERGED WITH THE PARA ABOVE? The AUVs considered here are powered untethered platforms, that
+operate at $1$-$3$ m/s in the upper water column, where they are free
+to move in six degrees of freedom (6 DOF). The in-water operation time
+capability depends on survey speed, payload sensors and navigation;
+typically this amounts to 8-48 hrs. AUVs typically use single-board
+computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA
+Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared
+cache, and 4 GB of 1600 MHz DRAM) for computation onboard.}
 
@@
 This work builds on a multidisciplinary effort combining statistical
-modeling with robotic surveying techniques for oceanographic
+methods with robotic surveying techniques for oceanographic
 applications. We show how observation practices can gain efficiency
-and accuracy from the development of statistical techniques to achieve
-more effective spatial monitoring. We further demonstrate the
+and accuracy from the development of statistical techniques for spatial monitoring. We further demonstrate the
 opportunities available for real-time multivariable spatial
-data-gathering and analysis onboard autonomous platforms, which
+data gathering and analysis onboard autonomous platforms, which
 statisticians can exploit to create general-purpose toolkits for
@@
 connected to the properties of water masses. The characterization of
-uncertainties in random sets is extended with new results for the EIBV
-uncertainty reduction achieved by spatial sampling design. This is
+uncertainties in random sets is extended with new results for the expected integrated Bernoulli variance
+reduction achieved by spatial sampling designs. This is
 provided in closed-form for the situation with a static design, and
 then extended to the adaptive situation. The sequential derivations
-provide new insights into efficient applications of adaptive sampling,
+provide new insights into efficient applications of adaptive data collection,
 as demonstrated in our application.
 
-The study did not consider any temporal effects, which would be
+The case study consider the upper water column in the river plume, represented by a two dimensional grid.
+Extensions to three-dimensional domains are not methodologically different, but the operation must likely approximate calculations by integrating terms only in the vicinity of the autonomous vehicle \citep{fossum18b}.
+We did not consider any temporal effects, which would be
 relevant on a larger time scale. We consider the extension to
@@
 advection-diffusion equations could be useful in this kind of modeling
-\citep{sigrist2015stochastic}. For more complex oceanographic
+\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex oceanographic
 phenomena, the methods will need to be extended to non-Gaussian
-phenomena, possibly feature-based mixtures of GPs which could still be
-run onboard and augmented by dynamical models. \kc{How about talking
-  about removing any grids or waypoints altogether?}
-
-The spatial-statistical design criterion, building on random sets is
-relevant in our setting with different water properties, and we show
-mathematical generality beyond the EIBV which is convenient to
-implement for onboard robotic planning. Such criteria could also be
-useful in other oceanographic settings as for algal-blooms, anoxic
-zones or open water fronts \cite{costa19}. Of course, other criteria
+phenomena, possibly feature-based mixtures of Gaussian processes which could still be
+run onboard and augmented by dynamical models.
+
+The spatial-statistical design criterion building on random sets is
+relevant in our setting with different water properties. 
+We show
+mathematical generality beyond the expected integrated Bernoulli variance, for instance that of volume uncertainties which is possibly more relevant, but requires more computational resources.
+Such criteria could be particularly useful in other oceanographic settings of algal-blooms, anoxic
+zones or open water fronts \cite{costa19}.
+Of course, other criteria
 are also relevant for instance, hybrid or multi-attribute criteria
@@
 downstream decision-making, with policy makers and regulators who need
-to make difficult decisions related to fish farming, other marine
-resource operations. Value of information analysis (VOI)
-\citep{Eidsvik:15} could be used in a similar vein as the IBV in the
-current work, as both analytic methods explore when, where and which
+to make difficult decisions related to fish farming or other marine
+resource operations, and value of information analysis
+\citep{Eidsvik:15} could be used to evaluate whether
 information is likely to result in improved decision-making.
 
-While more effort could be spent on approximating the look-ahead
-stages in the adaptive designs, it is perhaps more interesting to
+\kc{How about talking
+  about removing any grids or waypoints altogether?}
+
+In our context the myopic strategy perform rather well, and for computational reasons we did not go in depth on the dynamic program solution.
+There has been much creative work on finite horizon optimization in the robotics literature including probabilistic road maps and rapidly-exploring random trees \citep{karaman2011sampling}, but their statistical properties are not yet clear.
+It is equally interesting to
 explore the additional flexibility that can be gained by having
-multiple AUVs co-temporally exploring a spatial or spatio-temporal
+multiple vehicles co-temporally exploring a spatial or spatio-temporal
 domain \citep{ferreira2019advancing}. Such an approach would enable
@@
 to move in parallel to best capture the excursion set.
+The sometimes conflicting topics of autonomy and communication have caught much interest in control engineering lately \citep{zolich2019survey}, and statistical evaluations could contribute in the future.
+In our context, autonomy is highlighted, and this is reflected in our modeling and computational assumptions.
+With the possibilities of wifi communication to a mother-ship at waypoints, more advanced modeling and computational routines are of course possible, but with new challenges related to design and planning for when and where the communication should occur.
 

### 2020-06-12 ans_to_reviewers/answers-to-reviewers.tex
Updates from Overleaf

@@
 %----------------------------------------------------------------------
-\title{Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets
+\title{Autonomous oceanographic Sampling Designs Using Excursion Sets for Multivariate Gaussian Random Fields
   \\\vspace{5mm}
@@
 suggestions. They have all been considered in this revised
-version of this manuscript.
-We have worked hard on developing both theoretical results, computational resources, and finally on the presentation in the paper.
-To help us on these efforts, we included Cedric Traveletti as co-author because he has contributed significantly on the revised version of the paper.
+version of our manuscript.
+We have worked a lot on developing both theoretical results, computational resources, and on the presentation in the paper.
+To help us in these efforts, we included Cedric Traveletti as co-author because he has contributed significantly on the revised version of the paper.
 
-The principal revisions made to the manuscript are as follows and are done to address the main comments by Editors and Reviewers:
+The principal revisions made to the manuscript are as shown in this list, and they are done to address the main comments of Editors and Reviewers:
 
-\begin{enumerate}
-\item The motivation for excursion sets (ESs) and in particular that of {\bf{multivariate ESs}} has been strengthened.
+\begin{itemize}
+\item The motivation for excursion sets (ESs), and in particular that of {\bf{multivariate ESs}}, has been strengthened.
 Motivating examples are abundant in medicine where doctors do not solely rely on a single symptom but must see several combined effects before making a diagnosis.
-Another example is that of photosynthesis where plants and algae produce oxygen only when given enough of every component; light, carbon dioxide and water.
-Our motivation in this paper is that joint temperature and salinity excursions are proxies for several processes in the oceans.
-For instance, salmon lice larvae are known to depend on relatively high temperature and high salinities to survive.
+Our motivation in the current paper is that temperature and salinity are proxies for more complex biological variables in the ocean, and in several cases they must be jointly high or low for phenomena to occur.
+For instance, salmon lice larvae are known to depend on relatively high temperature and high salinity to survive.
 Salmon lice are parasites to salmon, and the bivariate ES for temperature and salinity is hence important to understand for sustainable activities in salmon fisheries.
 
-In the revised manuscript these aspects are largely rewritten in the Introduction, Background, expected integrated Bernoulli variance (EIBV) and Case Study.
-This change addresses the comments of AE1, R1 (3,) MORE.
+In the revised manuscript these aspects are largely rewritten in the Introduction, and partly also in Section 2 on the expected integrated Bernoulli variance (EIBV) and in the Case Study ({\bf{TO DO 6}}).
+This change addresses some comments of AE, Rev #1 in particular.
 
-In the revised version we motivate more for statisticians, in particular we emphasize the statistical design problem which is sought solved by AUV with its constraints on navigation and computing resources.
-
-\item The {\bf{statistical methodological contribution}} have been clarified and improved, including definitions of ESs and having more general results related to the EIBV for vector-valued processes, using marginal (one-location) entries and for volume calculations requiring multiple locations.
+\item The {\bf{statistical methodological contribution}} have been clarified and improved, including definitions of random sets and ESs, and having more general results than just the EIBV for the bi-variate case of temperature and salinity.
 The revised version highlights these novel statistical results as a main contribution of the paper.
 
-In the revised manuscript these aspects are largely rewritten in the main body containing novel results for EIBV for sums of marginal variances as well as for more general results inluding that of volumes of random sets, while the Appendix derives the closed-form results.
-This change addresses the comments of E2, AE1, R2 (7), R3 (1), MORE.
+Moreover, we emphasize the statistical design problem which is sought solved by AUV with its constraints on navigation and computing resources.
 
-\item We realized that a contribution of the current work is available code for statisticians who could try this themselves or even extend it for their own statistical cases.
-The computer code could of course also be useful to practitioners having related sampling design problems.
-The {\bf{simulation results and code}} are thoroughly re-worked.
-Open-source python code is now provided for all the examples used in the paper and also for simulation studies using more nuanced statistical models (non-stationary and non-isotropic).
+In this sense, the revised version should provide ample motivation for statisticians.
 
-The simulation study section is largely modified and it contains pointers to the associated open-source code. This change addresses the comments of R1 (2,3),  MORE.
+In the revised manuscript these aspects are largely rewritten in the Introcuction and in Sect 2 and 3 on the statistical theory.
+Here, the main body contains the novel results for the expected uncertainty reduction in random sets exemplified by the EIBV, while the Appendix holds the proofs of key results.
+This change addresses the some comments of Editor, Associate Editor, Rec #2 and Rec #3.
 
-\end{enumerate}
+\item We realized that a contribution of the current work is available code for practioners or statisticians who could try this themselves, or possibly extend it for their own cases.
+We have hence made a substantial effort in structuring {\bf{computer code for examples}} which are presented on a URL link to go with the paper.
+This open-source python code is also ready to run on more nuanced statistical models (non-stationary and non-isotropic).
+
+The simulation study section contains pointers to the associated open-source code. This change addresses the comments of Rev #1 in particular.
+
+\end{itemize}
 \par \vspace{1em}
@@
 
-\reply: The revised version has an entirely new Section 3 which includes general background for random sets and excursion sets.
-We added references, for instance \cite{adler2000excursion,Molchanov2005}.
+\reply: The revised version has an entirely new Section 2 which includes general background for random sets and excursion sets.
+We added references, for instance \cite{Adler.Taylor2007} on excursion sets and \cite{Molchanov2005} on random sets.
 

### 2020-06-12 UQonSets.tex
Updates from Overleaf

--- 2020-06-08 main_new.tex
@@
-\documentclass[aoas]{imsart}
-
-\input{commands.tex}
-\usepackage{amsmath}
-\usepackage{pstricks,pst-grad}
-\usepackage{graphicx}
-\usepackage{floatrow}
-\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
-\floatsetup[table]{capposition=top}
-\usepackage{subfigure}
-\usepackage[utf8]{inputenc}
-\usepackage{booktabs}                     % horizontal lines in tables
-\usepackage{comment}
-
-% == Enable text degree
-\usepackage{textcomp}
-
-\usepackage{amsthm,amsmath,natbib}
-\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
-
-% == Trygve Test
-\usepackage{color, colortbl}
-\definecolor{Gray}{gray}{0.9}
-
-% put your definitions there:
-\startlocaldefs
-\newcommand{\edcomment}[1]{{\color{green}{\{Editor: #1\}}}}
-\newcommand{\frevcomment}[1]{{\color{blue}{\{Rev 1: #1\}}}}
-\newcommand{\srevcomment}[1]{{\color{red}{\{Rev 2: #1\}}}}
-\newcommand{\trevcomment}[1]{{\color{violet}{\{Rev 3: #1\}}}}
-\endlocaldefs
-
-% New Marcos.
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\input{macros}
-
-\begin{document}
-
-\begin{frontmatter}
-
-% "New Title of the paper"
-\title{Autonomous Oceanographic Data Collection via Uncertainty Reduction for  Excursion Sets of Vector-valued Random Fields} 
-\runtitle{Autonomous Oceanographic Uncertainty Reduction}
-
-% "Former Title of the paper" (DG 08.06.2020)
-%\title{Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields} \runtitle{Excursion Probabilities for Informative Sampling}
-
-\begin{aug}
-\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
-\author{\fnms{Cédric} \snm{Travelletti}\thanksref{t3}, \corref{} \ead[label=e2]{cedric.travelletti@stat.unibe.ch}}
-\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e3]{jo.eidsvik@ntnu.no}}
-\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e4]{david.ginsbourger@idiap.ch}}
-\and
-\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e5]{kanna.rajan@fe.up.pt}}
-
-\affiliation[t1]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
-\affiliation[t2]{Centre for Autonomous Marine Operations and Systems, NTNU.}
-% (AMOS).} Is the acronym needed now that this affiliation appears once only?
-% Added NTNU, though. 
-\affiliation[t3]{Institute of Mathematical Statistics and Actuarial Science, University of Bern, Switzerland.}
-\affiliation[t4]{Department of Mathematical Sciences, NTNU.}
-\affiliation[t5]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
-
-\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
-\printead{e1}}
-\address{Cédric Travelletti\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
-Switzerland.
-\printead{e2}}
-\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e3}}
-\address{David Ginsbourger\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
-Switzerland.
-\printead{e4}}
-\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
-  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
-\printead{e5}}
-
-\runauthor{TO. Fossum et al.}
-\end{aug}
-
-\begin{abstract}
-
-  Improving and optimizing oceanographic sampling is a crucial task
-  for marine science and maritime management. Faced with limited
-  resources to understand processes in the water-column, the
-  combination of statistics and autonomous robotics provides new
-  opportunities for experimental design. In this work we develop
-  methods for efficient spatial sampling applied to the mapping of
-  coastal ocean processes by providing informative descriptions of
-  spatial characteristics of ocean phenomena. Specifically, we define
-  a design
-  criterion based on improved characterization of the uncertainty in
-  the excursions of vector-valued Gaussian random fields, and derive
-  tractable expressions for the expected Bernoulli variance reduction
-  in such a framework. We demonstrate how this criterion can be used
-  to prioritize sampling efforts at locations that are ambiguous,
-  making exploration more effective. We use simulations to study the
-  properties of methods and to compare them with state-of-the-art
-  approaches, followed by results from field deployments with an
-  autonomous underwater vehicle (AUV) as part of a study mapping the
-  boundary of a river plume. The results demonstrate the potential of
-  combining statistical methods and robotic platforms to effectively
-  inform and execute data-driven environmental sampling.
-
-%Motivated by the challenges related to efficient allocation of sampling resources in environmental sensing, the combination of Excursion Probabilities and Gaussian process modeling is explored for autonomous robotic sampling of ocean features; enabling information driven measures  sampling efforts to high-interest regions.
-These regions are usually characterized by gradients of measurable environmental variables, e.g., temperature or salinity gradients, on which EPs subsequently can be used...Correlation among samples and multivariate requirements are typical in environmental studies.
-
-\end{abstract}
-
-\begin{keyword}
-\kwd{Ocean Sampling}
-\kwd{Excursion Sets}
-\kwd{Gaussian Processes}
-\kwd{Experimental Design}
-\kwd{Autonomous robots}
-\kwd{Adaptive Information Gathering}
-\end{keyword}
-
-\end{frontmatter}
-\section{Wannabe new introduction structure}
-
-NB: The subsections are used as a temporary instrument to highlight the proposed structure and should probably be abolished upon convergence to a stable version.
-
-\subsection{Oceanic data collection and spatial design of experiments}
-
-\begin{itemize}
-    \item Oceanic data collection as a timely and crucial challenge 
-    \item Uneasy to measure state variables varying continuously across space
-    \item Some lend themselves well to random field modelling 
-    \item Where to make measurements is a question related to experimental design
-    \item Yet with AUVs measurements ought to be made along a trajectory 
-    \item Furthermore, in several problems such as mapping Fjord versus ocean water, the goal is to reduce uncertainties on regions implicitly defined by several state variables such as temperature and salinity.
-    \item Besides this, AUV are able to revise their trajectory after intermediate measurements, calling for experimental design strategies that are adaptive and ideally also account for multiple steps ahead.
-    \item All this in the framework where several responses are modelled within a vector-valued random field framework and under observation noise calls for sophisticated design criteria and computational strategies for their optimization.
-    \item Here we aim to leverage recent progress in sequential design of computer experiments and adapt them to the specifics of guiding AUV measurements for parcimoniously distinguishing fjord from ocean waters.
-    \item While our problem involves multivariate observational responses to be observed along a continuous trajectory, our approach will take substantial advantage of existing work in the field of excursion set estimation mainly dedicated to computer experiments in the scalar-valued case, and where space exploration can be performed without constraints regarding the distance between successive design points.
-    \item Let us briefly review random field modelling and recent advances in sequential design of experiments before detailing further our proposed approach and outline.  
-\end{itemize}
-
-\subsection{Random field modelling and sequential design of experiments}
-
-\begin{itemize}
-    \item 
-\end{itemize}
-
-\subsection{Proposed approach in a nutshell}
-
-\begin{itemize}
-    \item 
-\end{itemize}
-
-\subsection{Outline of the paper and its appendices / supplemental material}
-
-\begin{itemize}
-    \item 
-\end{itemize}
-
-%\subsection{}
-
-%\begin{itemize}
-%    \item 
-%\end{itemize}
-
-\section{Former Introduction (Before 08.06.2020)}
-
-Monitoring the world's oceans has gained increased importance in light
-of the changing climate and increasing anthropogenic impact. A central
-problem for understanding these factors is the lack of representative
-data with sufficient resolution. Most of this \emph{undersampling} can
-be attributed to the large spatio-temporal variations in which ocean
-processes transpire, prompting the need for effective means of
-sampling. By \emph{sampling}, we refer to the design of observational
-strategies in the spatial domain, where the use autonomous robotic
-platforms can be combined with statistical methods to pursue
-measurements with high scientific relevance. This combination, of
-statistical tools and robotic platforms, constitute a methodological
-basis for addressing the problem of efficient sampling of the ocean,
-which is at the core of this work.
-
-The
-advent of robust mobile robotic platforms \citep{Bellingham07} has
-resulted in significant contributions to environmental monitoring and
-sampling in the ocean. In particular, autonomous underwater vehicles (AUVs), have
-advanced the state of sampling and consequently have made robotics an
-integral part of ocean observation; our previous work has contributed
-to this effort \citep{das11b,Das2015,fossuminformation,fossum18b}.
-
-The potential that lies in using information-based sampling methods, which we
-highlight in this paper, is to do efficient environmental sampling and
-overcome common challenges as discussed next. First, full numerical ocean models based on complex
-differential equations cannot be run onboard a marine robot with
-limited computational capacity.
-% provide accurate results for online
-% operations because the onboard computers lack the computational power
-% required.
-Hence statistical proxy models of the environment must be used. For our purpose, we rely on Gaussian process (GP)
-representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
-Second, because of limited resources to sample the large oceanographic domains, one must plan for
-active learning during the operation, where a relevant criterion is
-used to extract the most valuable designs. The acquired in-situ
-information must then be assimilated into the statistical model
-onboard the platform, so that it can be used to re-compute the
-sampling criterion and in this way inform decisions on where to sample
-next. We use a criterion focusing on the prediction of excursion sets
-(ESs) for the multivariate spatial processes of interest. Third, the
-number of sampling designs is enormous, creating a trade-off between
-optimization and computability. One must often compromise in both
-development and practical operations, and we will rely on
-short-horizon optimization routines for sequential designs among
-nearest neighbors points in a spatial grid (waypoint graph).
-
-\kc{The para above needs a rewrite to me. It starts off by stating
-  there are challenges. But then jumps to models without stating the
-  connection with sampling, and then jumps into learning without
-  stating why learning is important. Typically the English in formal
-  text starts with a statement which is supported or clarified. In the
-  two cases above, the connections are simply not obvious. If this was
-  there in the first version of the m/s submitted, I'm not sure why
-  the reviewers didn't catch it. This is not just a stylistic issue;
-  it has to do with making sense and bring the reader/reviewer along.}
-
-\kc{Second, I don't see a motivation for GPs here and as I read
-  further into the text. Even for stats folks, are GPs the ``best''
-  ways to go? Are they obviously the tools one would use in
-  environmental sampling? And even if they are, I think we ought to
-  make the case for it strongly and generally, so this paper is
-  actually accessible and understandable to those not in the Spatial
-  Stats field. I would suggest adding a short para on GPs and why
-  they're important -- i.e. motivate the use of GPs early, so we can
-  freely make attributions to them later, as in the next section, for
-  example.}
-
-The imperfect understanding of causality coupled with the underlying
-fundamental uncertainty of the ocean implies that we need to make a
-large number of measurements. In addition, the use statistical models
-and methods to determine what and how phenomena form, is one way to
-increase the quality of predictive power aiding our understanding of
-upper water-column processes.
-
-With this in mind, we design AUV sampling strategies for reducing the spatially-integrated Bernoulli variance (IBV) in the ES
-indicator variables. 
-ES of random fields are thoroughly presented in \citep{adler2009random}. Recent work on ESs and associated excursion
-probabilities (EPs) connected to spatial statistics include
-\cite{french2013spatio,bolin2015excursion,french2016credible}. Our
-focus is on uncertainty characterization and how this is influenced by
-sampling. In this sense, our work is similar to
-\cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who
-describe analytical results for the IBV for univariate processes.
-
-A major contribution of our work is to derive closed-form results for
-this design criteria for situations where the underlying model is
-based on multivariate GPs. Another contribution is that we embed this
-approach in a sequential strategy to produce sampling algorithms, and
-apply this to a real-world scenario of autonomously sampling
-temperature and salinity gradients using an AUV, with the intention of
-characterizing the variability of oceanic and riverine waters in
-typical mixed coastal environments.
-
-Motivating examples relevant for ESs of multivariate processes are
-abundant. In medicine, doctors do not rely solely on a single symptom
-but must see several combined effects before making a diagnosis. In
-our context of environmental sampling, the salinity and temperature
-excursions of a river plume can similarly help characterize the
-underlying bio-geochemical processes
-\citep{hopkins2013detection,Pinto2018}, where sampling can be aimed
-towards reducing uncertainty in the temperature and salinity ES.
-Unlike what has been done in previous plume exploration studies, we
-emphasize the links to spatial statistical modeling and targeted
-multivariate sampling criteria for sequential sampling.
-
-Other statistical
-work in the oceanographic domain include \cite{wikle2013modern}
-focusing on hierarchical statistical models; \cite{sahu2008space},
-studying spatio-temporal models for sea surface temperature and
-salinity data; and \cite{mellucci2018oceanic} looking at the
-statistical prediction of features using an underwater glider.
-In this paper the focus is not on statistical modeling, but rather on statistical results and computations for efficient sampling designs.
-
-The remainder of this paper is organized as follows: Section
-\ref{sec:bg} provides select background on ocean sampling. Section
-\ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for
-vector-valued GPs. Section \ref{sec:heuristics} builds on these
-assumptions when deriving the sequential design criteria for adaptive
-sampling. Section \ref{sec:simulations} discusses properties of the
-methods in simulation studies. Section \ref{sec:case_study}
-demonstrates the methodology used in field work characterizing a river
-plume and finally, Section \ref{sec:concl_disc} contains a summary and
-a discussion of future work.
-
-\section{Ocean Processes and Observation}
-\label{sec:bg}
-
-Central to understanding the changes taking place in the upper
-water-column is knowledge of the bio-geophysical interaction driven by
-an agglomeration of physical forcings (e.g. wind, topography,
-bathymetry, tidal influences, etc.) and incipient micro-biology driven
-by planktonic and coastal anthropogenic input, such as pollution and
-agricultural runoff transported into the ocean by rivers and streams.
-These often result in a range of ecosystem-related phenomena such as
-blooms and plumes, with direct and indirect effects on society. \kc{we
-  should look for a citation here. Check with John Ryan}
-
-Traditional data collection to understand such phenomena at sea has
-typically been based on static buoys, floats, or ship-based methods,
-with significant logistical limitations that directly impact coverage
-and sampling resolution. Modern methods using satellite remote-sensing
-provide large-scale coverage but have limited resolution, are limited
-to sensing the surface, and are impacted by cloud cover. Numerical
-ocean models similarly find it challenging to provide detail at fine
-scale \citep{Lermusiaux:2006}. Mobile robotic platforms
-\citep{Bellingham07} have contributed significantly to environmental monitoring and sampling. In particular, AUVs have
-advanced the state of in-situ sampling, and made robotics an integral
-part of ocean observation, filling parts of the undersampling gap
-\cite{rudnick03,rudnick18}(See Fig. \ref{fig:envir1}).
-
-%Walter Munk referred to the 20\textsuperscript{th} century
-%as the \emph{century of undersampling} \citep{munk2002}, pointing out
-%a lack of sufficient sampling resolution in time and space as a
-%critical issue for characterizing ocean phenomena. Even today,
-%policymakers struggle with limited observability of severe
-%environmental problems; a prominent example is the recent (May 2018)
-%outbreak of harmful algal blooms in Northern Norway\footnote{Report
-%  from the Institute of Marine Research: https://bit.ly/2KZGJfI}, with
-%losses exceeding 12 million tons of salmon valued at \$80 million
-%(Norwegian Directorate of Fisheries). This is a case in point of how
-%adequate sampling efforts could have provided a more accurate
-%assessment and early warning of changing environmental conditions.
-
-\begin{figure}[!h] 
-  \centering 
-  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
-    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
-  \hfill
-  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
-    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
-  \caption{\ref{fig:envir1} Traditional ocean observation based on 
-    ship-based sampling has been augmented by autonomous
-    robotic vehicles. % and their interactions.  
-    AUV platforms are an integral part of this network being able to
-    reason and make decisions for efficient onboard adaptive sampling.
-    % using the sense-plan-act control approach to autonomous control. 
-    \ref{fig:nidelven} The interaction of river and ocean creates
-    processes that are challenging to map, where the combination of
-    statistics and robotics can play a vital role in enabling more
-    effective oceanographic observation.\kc{The inset ``AUV robot''
-      looks very tacky and ad-hoc; might be best to remove it. Change
-      ``ships'' to
-      ``Ship'', remove the aerial drone, since we didn't use it and
-      don't plan to in this work.}}
-  \label{fig:envir} \end{figure}
-
-The AUVs considered here are powered untethered platforms, that
-operate at $1$-$3$ m/s in the upper water column, where they are free
-to move in six degrees of freedom (6 DOF). The in-water operation time
-capability depends on survey speed, payload sensors and navigation;
-typically this amounts to 8-48 hrs. AUVs typically use single-board
-computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA
-Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared
-cache, and 4 GB of 1600 MHz DRAM) for computation onboard.
-
-Surveys with AUVs are usually limited to observations along fixed
-transects that are pre-scripted in mission plans created manually by a
-human operator. The mission can be specified operating on a scale of
-hundreds of meters to tens of kilometers depending on the scientific
-context. Faced with limited coverage capacity, a more effective
-approach is to instead use onboard algorithms to continuously
-evaluate, update, and refine future sampling locations (sense-plan-act
-cycle in Fig. \ref{fig:envir1}), making the information gathering
-\emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
-% This usually occurs on a spatial grid, called a waypoint graph. 
-In doing so, the space of sampling opportunities is still limited by
-the waypoint graph, but the AUV has flexibility to modify its path at
-each waypoint based on in-situ sensing and measurements onboard
-\citep{py10,Rajan12,Rajan12b}.
-
-We focus on spatial characterization of a frontal system generated by
-river plumes. Fig. \ref{fig:nidelven} shows the survey area in
-Trondheim, Norway, where cold freshwater enters from the river,
-creating a strong gradient in both temperature and salinity. Because
-of the local topography and the Coriolis force the cold fresh water
-tends to flow to the east. Depending on the river discharge, tidal
-effects, wind, and temperature differences, this boundary often gets
-distorted. Initial knowledge about the location and evolution of these
-features are highly uncertain, making deterministic planning
-challenging.
-
-Adaptive in-situ AUV sampling of an evolving frontal feature has been
-explored in \cite{fronts11,Zhang2012,Pinto2018,costa19}. These
-approaches typically use a reactive-adaptive scheme, whereby
-exploration does not rely on a statistical model of the environment,
-but rather adaptation is based on closing the sensing and actuation
-loop. Myopic sampling, i.e. stage-wise selection of the path (on the
-waypoint graph), has been used for surveys
-\citep{singh2009efficient,Binney2013} that focus largely on reducing
-predictive variance or entropy. These criteria are widely adopted in
-the statistics literature on spatio-temporal design as well
-\cite{bueso1998state,zidek2019monitoring}, but variance and entropy
-reduction are independent of the actual data realizations under the
-assumptions of GP models, so it has limited
-decisional flexibility. The use of data-driven adaptive criteria
-was introduced to include more targeted sampling of regions of
-scientific interest in \cite{Low2009} and \cite{fossuminformation}. In
-this paper, we focus on mapping the river plume by rewarding designs
-that improve the classification of ES in temperature and salinity.
-
-\section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
-\label{sec:ESEP}
-
-We define the random sets of interest in Section \ref{sec:ES}, while the closed-form expressions for the expected uncertainty reduction are derived in Section \ref{sec:EIBV}.
-Section \ref{Sec:UnivarEx} provides an instructive bivariate example relevant for sampling in the temperature and salinity application.
-
-\subsection{Excursion sets and Integrated Bernoulli variance}
-\label{sec:ES}
-
-Our objective is to leverage statistical tools onboard an autonomous
-robotic platform to characterize a river plume, focusing on spatial
-separation of cold freshwater from a river and warmer saline waters
-of a fjord. The characterization of uncertainties on the ES is a
-useful goal of this sampling strategy.
-
-In order to formalize the problem in mathematical terms, let us
-consider a multivariate (i.e., vector-valued) Gaussian random field
-$(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary
-set (typically in geostatistical studies, a geographical subset of interest) and the
-dimensionality of the output $\no \geq 1$ is arbitrary. We have $\no = 2$ in the example with temperature and salinity in the river plume.
-
-Assume furthermore that the set of interest
-is characterized in terms of a subset of values of the variables, say
-$\T \subset \R^{\no}$, and denote the corresponding set of index
-points by
-$$
-\es=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
-$$
-Our goal here is to develop approaches to quantify and improve the
-characterization of uncertainties on such sets $\es$, in the
-context of multivariate random fields.
-
-While some aspects of the developed approaches do not call for a
-specific form of $\T$, we will often stick for simplicity to the case
-of orthants
-($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
-$t_1,\dots, t_{\no} \in \R$) as this will allow efficient calculation
-of several key quantities. Note that changing some $\leq$ inequalities
-to $\geq$ ones would lead to immediate adaptations. In the application
-specific properties following the orthant assumption come into play
-\kc{This last phrase doesn't parse; is it ``..., the following
-  orthant....''?}: We let $t_1$ be a threshold for temperature and
-$t_2$ a threshold for salinity \kc{Might be easier to remember with
-  symbols 't' and 's' in some
-  form.}. The set of index point then defines
-$\mbox{ES} = \{\x \in D : \gp[\x,1] \leq t_1,\gp[\x,2] \leq t_2\}$.
-The associated EPs are
-$p(\x) = P(\gp[\x,1] \leq t_1,\gp[\x,2] \leq t_2)$,
-$\x \in \mathcal{M}$.
-
-For statistical generality, we will however keep the abstract notation
-with $\T$ and general $\es$ \kc{Again, previous phrase doesn't parse.}
-for most of the derivations. Note that assuming that $\gp$ has
-continuous trajectories (almost surely) and $T$ to be closed, then
-$\es$ can be cast as a Random Closed Set
-\citep{Molchanov2005}. %In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
-Let us in particular consider a locally finite (Borel) measure $\mes$
-on $D$ and investigate the probability distribution of $\mes(\es)$
-through its moments.
+Let us in particular consider a locally finite (Borel) measure $\mes$ on $D$ and investigate the probability distribution 
+of $\mes(\es)$ through its moments.
 
@@
 %Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
-	For a measurable random field $\gp$ and a locally finite
-        measure $\mes$ on $\mathcal{M}$, $\mes(\es)$ is a random
-        variable and for any $r\geq 1$, \begin{equation*}
-          \begin{split} \mathbb{E}[\mes(\es)^r]
-            &=\int_{\mathcal{M}^{r}} \jointExcuProb \productMeasure ,
-          \end{split} \end{equation*}
+For a measurable random field $\gp$ and a locally finite measure $\mes$ on $D$, $\mes(\es)$ is a random variable and for 
+any $r\geq 1$,
+\begin{equation*}
+\begin{split}
+\mathbb{E}[\mes(\es)^r]
+&=\int_{D^{r}} \jointExcuProb
+\productMeasure
+,
+\end{split}
+\end{equation*}
 
@@
 %Note that we will use boldface to denote concatenated variables. % TO RELOCATE
-	Here $\gp$ is defined on $\mathcal{M}$, and for
-	$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in \mathcal{M}^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
-	\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.\kc{Is $\no r$ a
-        product of the $\no$ from above? Is that obvious?}
+Here $\gp$ is defined on $D$, and for
+$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in D^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
+\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.
 \medskip
@@
 \end{align*}
-	where $\mathcal{N}_{\no r}(\cdot ; \meanUU, \covUU)$ is the Gaussian measure on $\mathbb{R}^{\no r}$ with mean $\meanUU$ and covariance matrix $\covUU$, respectively defined blockwise by  
+where $\mathcal{N}_{\no r}(\cdot ; \meanUU, \covUU)$ is the Gaussian measure on $\mathbb{R}^{\no r}$ with mean $\meanUU$ 
+and covariance matrix $\covUU$, respectively defined blockwise by
 %where both are defined blockwise by
@@
 \end{align*}
-	each of the $r\times r$ blocks of the latter matrix being itself a (cross-)covariance matrix of dimension $\no \times \no$. 
+each of the $r\times r$ blocks of the latter matrix being itself a (cross-)covariance matrix of dimension $\no \times 
+\no$.
 %dimensional Gaussian random vector.
-	Assuming further that $\covUU$ is non-singular, the probability of interest can be formulated in terms of the $\no r$-dimensional Gaussian probability density function 
+Assuming further that $\covUU$ is non-singular, the probability of interest can be formulated in terms of the $\no 
+r$-dimensional Gaussian probability density function
 $\varphi_{\no  r}(\cdot;~\meanUU, \covUU)$ as
@@
 In the particular orthant case with $\T=(-\infty, t_1] \times \dots \times (-\infty, t_{r}]$,
-	the latter probability directly writes \kc{What does
-          ``writes'' mean here? Connotes?} in terms of the multivariate Gaussian
+the latter probability directly writes in terms of the multivariate Gaussian
 cumulative distribution, % with mean $\meanUU$ and covariance matrix $\covUU$,
-	this time without requiring $\covUU$ %the latter 
+this time by the way without requiring $\covUU$ %the latter
 to be non-singular:
@@
 \end{propo}
-
-Proposition~\ref{propo1} enables one to calculate centered moments of $\mes(\es)$. 
-The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$
-is probably the most relevant of these for our purposes. We
-have 
-\begin{equation*}
-\begin{split}
-\emv
-&=\int_{\mathcal{M}^2} \mathbb{P}\left(
-\gp[u]\in T, \gp[v]\in T \right) 
-d\mes^{\otimes}(u, v)\\
-&-\left( \int_{\mathcal{M}} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
-\end{split}
-\end{equation*} 
-which evaluates in the excursion/sojourn case where
-$\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ to
-\begin{equation*} \begin{split}
-  \emv
-%\operatorname{Var}[\mes(\es)]
-&=\int_{\mathcal{M}^2} 
-\varPhi_{2\no}
-\left(
-(\bt, \bt); \bmu((u,v)), 
-K((u,v),(u,v))
-\right) 
-\
-\mathrm{d}\mes^{\otimes} %\mes 
-%\productMeasure 
-(u,v)\\
-&-\left( \int_{\mathcal{M}} \varPhi_{\no}\left(\bt;\bmu(u), K(u)\right) d\mes(u) \right)^2.
-\end{split}
-\end{equation*}
-This quantity requires  an integral over $D^2$. In
-contrast, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal}
-involves solely an integral on $\mathcal{M}$ and can be expanded in
-our present setting as follows 
-\begin{equation}\label{IBV}
-\begin{split}
-\operatorname{IBV} %(\es) %;\mes)
-&=\int_{\mathcal{M}}
-\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
-d\mes(u) \\
-&=\int_{\mathcal{M}}
-\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) 
--\left(\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) \right)^2
-\mathrm{d}\mes(u). 
-\end{split}
-\end{equation}
-
-\subsection{Expected Integrated Bernoulli Variance and Sampling Designs}
-\label{sec:EIBV}
-
-Our goal is to construct informative sampling designs, given limited
-resources \kc{information?}. In our domain, data is gathered by an AUV
-that can measure temperature and salinity at chosen locations on a
-defined waypoint graph. Such a discretized grid fosters reliable
-navigation and mission performance. The waypoints further enable a
-structure to the design problem in that, that sampling flexibility is
-limited to operation on the graph. % This gives lots of sampling
-% opportunities, and it is important to spend the sampling efforts
-% wisely.
-We focus on sampling designs that reduce the expected integrated
-Bernoulli variance (EIBV). Note that this sampling can be done in a
-static manner where the survey design is pre-planned, or in an
-adaptive setting. For the latter case, data can be processed
-sequentially and the updated model is then used to design and select
-the next waypoint graph to visit, which in turn are used again to
-update the model, and so on (Section \ref{sec:heuristics}). We next
-describe the data modeling assumptions for one stage and develop
-design criteria of semi-analytical forms relying on integrals of
-multivariate Gaussian cumulative distribution functions; we will
-appeal to some key intermediate results presented next.
-
-We denote a design by $d \in \mathcal{D}$, where the concatenation of designs with $q$ sampling locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$ is $\mathcal{D}$. The
-measurement model is defined by 
-\begin{equation}\label{eq:meas}
-    \bY_d=G_d \bz(\bu_d) + \epsilon, \hspace{3mm} \epsilon \sim N(0,R_d).
-\end{equation}
-Here, $\bz(\bu_d)$ is the size $pq$ vector of GP variables at the
-sampling locations, and $G_d$ is the design matrix which in principle
-allows for observing subsets of the vector-valued field at some of the
-$q$ locations (i.e. the matrix has fewer rows than columns). If the
-additive error has independent elements, the covariance matrix $R_d$
-is diagonal. In our setting temperature and salinity measurements are
-extracted from different sensors, so the independence assumption
-between variables is not unrealistic, but this is not necessarily the
-case for other kinds of measurements on the AUV. It is further common
-in geostatistics to assume spatially independent measurement errors,
-and we also do so. But our methodology for constructing sampling
-designs does not rely on these independence assumptions.
-
-For a given design $d$ the evaluation is based on the expected
-information contained in the data to influence the criterion.   
-Considering the EIBV criteria, we have
-\begin{eqnarray}\label{eq:eibv}
-    \mbox{EIBV} (d) &=& E_{\bY_d} \left( \int_{\mathcal{M}}
-p(\x;\by_d)(1-p(\x;\by_d))
-d\mes(u) \right), \nonumber \\
-p(\x;\by_d) &=& \mathbb{P}\left(\gp[\uu]\in T|\bY_d=\by_d \right)
-\end{eqnarray}
-where the expectation is over the planned data.  The optimal design
-for an AUV survey is then 
-\begin{equation}\label{crit}
-    d^* = \mbox{argmin}_{d \in \mathcal{D}} \mbox{EIBV}(d).
-\end{equation}
-For another criterion using random moments of random sets similar
-expectations would be required, but as we will see, the EIBV has a
-closed-form expression which is relatively fast to compute and hence
-useful for onboard design selection.
-
-Assuming now that $q$ observations are available, one wishes to
-predict $\gp[\x]$. The multivariate Kriging equations holds with size
-$p$ vector of conditional means:  
-\begin{equation}\label{eq:cokrig}
-\bbm(\x)=\bmu(\x)+\lambda(\x,\bu_d)^T \left( \by_d-G_d \bmu(\bu_d)\right),
-\end{equation}
-where $\bmu$ is size $p$ unconditional mean function of $Z$, $\by_d$
-is the realized vector of observed responses, and
-$\lambda(\x,\bu_d)=K^{-1}_y k(\bu_d, \bu)$. Here,
-the data covariance is
-$K_y=\mbox{Cov}(\bY_d,\bY_d)=G_d K(\bu_d,\bu_d)G_d^T+R_d$ while
-the cross-covariance is $\mbox{Cov}(\bY_d,Z(\x))=G_d k(\bu_d, \x)$, assuming the measurement
-error is independent of the random fields.  The Kriging (cross-)covariance function can also be expressed in the same vein via
-a $p \times p$ matrix 
-\begin{equation}\label{eq:krig:cov}
-\Sigma(\x,\x')=K(\x,\x')-\lambda(\x,\bu_d)^T K_y(\bu_d, \bu_d) \lambda(\x,\bu_d).
-\end{equation}
-
-Eq. \eqref{eq:cokrig} and \eqref{eq:krig:cov} are used to
-provide informative, updated statements about sets of interest. For
-our case with temperature and salinity, the conditional EPs, given
-measurements $\by_d$, can then be computed from
-\begin{eqnarray}\label{eq:post_ep}
- p(\x;\by_d) &=& P(Z_1(\x) \leq t_1, Z_{2}(\x) \leq t_2 |\bY_d=\by_d),  \nonumber \\
- &=& \Phi_2 ((t_1,t_2)^T-\bbm(\x);0_2, \Sigma(\x,\x)). 
-\end{eqnarray}
-
-A critical element in the derivation of a closed-form for the EIBV is
-that the conditional mean in Eq. \eqref{eq:cokrig} is a linear
-(affine) function of the data $\by_d$ and the covariance is not a
-function of the data. This in turn means that the probabilities in Eq.
-\eqref{eq:post_ep} involve inequality statements for linear
-combinations of Gaussian variables. Related closed-form solutions have
-been noted in similar contexts \citep{bhattacharjya2013value,
-  chevalier2014fast,stroh}, but not generalized to our situation with
-random sets for vector-valued GPs. Notably, the mean in equation
-\eqref{eq:cokrig}, is in the form of $\ba+B V$, where
-$V \sim N(0_q, K_y (\bu_d,\bu_d))$,
-$\ba=(t_1,t_2)^t-\bmu(\x)+\lambda(\x,\bu_d)^T G_d \bmu(\bu_d)$
-and $B=\lambda(\x,\bu_d)^T$. The following then applies to give a
-closed-form solution to EIBV calculation in this case.
-
-\begin{propo}
-	\label{propo2}
-	Let $p, q, h \geq 1$, $a \in \R^p$, $B \in \R^{p\times q}$,
-	and $\covN, \covV$ be two covariance matrices in
-	$\R^{p\times p}$ and $\R^{q\times q}$, respectively.
-	Then, for $V \sim \mathcal{N}_{q}(0_q, \covV)$,
-	\begin{equation*}
-	\mathbb{E}\left[ \varPhi_{p}\left( \ba + BV; 0_p,\covN \right)^h \right]
-	=
-	\varPhi_{ph}
-	\left(
-	\ba
-	;0_{ph},
-	\Sigma
-	\right),
-	\end{equation*}
-	where the vector $\ba \in \R^{p h}$ is defined as 
-	$\ba := 1_h\otimes a = 
-	\left(a, \dots , a
-	\right)'$
-	and the $p h\times p h$ covariance matrix is given by
-	$\Sigma := 
-	1_h 1_h'\otimes B\covV B' + I_h\otimes \covN$.
-\end{propo}
-The proof of Proposition \ref{propo2} is provided in the Appendix.
-
-In block-wise representation, the following relation holds for the
-covariance in Proposition \ref{propo2}: 
-\begin{align*}
-	\Sigma&=
-	\begin{pmatrix}
-	\covN & &\\
-	& \ddots &\\
-	&   & \covN
-	\end{pmatrix}
-	+
-	\begin{pmatrix}
-	B\covV B' & \dots & B\covV  B'\\
-	\vdots & & \vdots\\
-	B\covV B' & \dots & B\covV B'\\
-	\end{pmatrix}.
-\end{align*}
-The calculation of the multivariate Gaussian cumulative distribution
-is done effectively using code such as that of
-\cite{genz2009computation}.
-
-In the more general case, for instance that of volumes of random sets,
-the criterion will include couplings with different locations and the
-result must be generalized. For the case of multivariate monomials in
-orthant probabilities with thresholds affine in a common Gaussian
-vector \kc{This sentence is dangling. Doesn't have a conclusion.}.
-%Namely, for monomials of degree $k$, we have.
-
-\begin{propo}%[\textcolor{red}{Notation in progress, to be reviewed/revised}]
-	\label{propo3}
-	Let $g, p, q\geq 1$, $h_{1},\dots, h_{g}\geq 1$ with $H=\sum_{i=1}^g h_i$, $a_{i} \in \R^{p}$, $B_{i}\in \R^{p \times q}$, and covariance matrices $\covN_i \in \R^{p \times p}$ $(1\leq i \leq g)$.
-Then, for any covariance matrix $\covV \in \R^{q\times q}$ and $V\sim\mathcal{N}_{q}(0_q,\covV)$,
-	%	Let $g$ and $k$ be integers, and consider a $g$-terms monomial of degree $k$
-	%	\[
-	%	    \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \Sigma_{i}
-	%	    \right)^{h_i},
-	%    \]
-	%    where $\sum_{i=1}^g h_i = k$, each $a_i$, $B_i$ and $\Sigma_i$ is as in
-	%    \cref{propo2} and $V$ is a $q$-dimensional Gaussian vector distributed as
-	%    $V\sim\mathcal{N}(0,\covV)$.
-	%
-	% Then the expectation of the monomial is given by
-	\begin{equation}
-	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V;0_p, \covN_{i} \right)^{h_i} \right]
-	=
-	\varPhi_{p H}
-	\left(
-	\bm{a}
-	;0_{pH},
-	\mathbf{\Sigma}
-	\right),
-	\end{equation}
-	with $\bm{a}=(1_{h_1}\otimes a_1, \dots, 1_{h_g}\otimes a_{g}) \in \R^{p H}$
-	and $\mathbf{\Sigma}\in \R^{p H \times p H}$ is defined blockwise by $(\Sigma_{i,j})_{i,j \in \{1,\dots, g\}}$ where, for any $i,j \in \{1,\dots, g\}$, %the blocks are given by 
-	\begin{equation}
-	\Sigma_{i,j}=
-	(1_{h_{i}}1_{h_{j}}')\otimes (B_{i}\Sigma_{V} B_{j}') + \delta_{i,j}(I_{h_{i}}\otimes \covN_{i}) \in \R^{p h_{i} \times p h_{j}}.
-	\end{equation}
-	%with $\bm{a}=\sum_{j=1}^g \mathbf{1}_{J_{j}} \otimes a_j \in \R^{p \times k}$ 
-	%where $J_{j}=\{\sum_{i=1}^{j-1}h_i + 1, \dots, \sum_{i=1}^{j}h_i\}$,
-	%$\mathbf{1}_{I_{j}}$ denotes the vector of $\R^{k}$ with components equal to $1$ for indices in $I_{j}$ and $0$ else ($1\leq j \leq g$), 
-	%and $\mathbf{\Sigma}\in \R^{p k \times p k}$ is defined blockwise by 
-	%\begin{equation}
-	%\mathbf{\Sigma}_{\widetilde{J}_{j_{1}},\widetilde{J}_{j_{2}}}=
-	%(\mathbf{1}_{h_{j_1}}\mathbf{1}_{h_{j_2}}')\otimes (B_{j_{1}}\Sigma_{V} B_{j_2}') + \delta_{j_{1},j_{2}}(I_{j_{1}}\otimes \Sigma_{j_{1}}),
-	%\end{equation}
-	%where $\widetilde{J}_{j}=\{p\sum_{i=1}^{j-1}h_i + 1, \dots, p\sum_{i=1}^{j}h_i\}$ ($1\leq j \leq g$).
-\end{propo}
-The proof of Proposition \ref{propo2} is provided in the Appendix.
-
-\subsection{An Example of EIBV using Temperature and Salinity}
-\label{Sec:UnivarEx}
-
-% In this section, 
-We calculate EPs and the Bernoulli variance for a few bivariate
-Gaussian distributions for temperature and salinity to illustrate the concepts articulated above.
-The parameter inputs are made up, but they are in a realistic range as we have in the simulation study and real-world example below.
-The thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
-
-Fig. \ref{illus_bivarDens} shows contour plots of three different
-densities with increasing correlation $\gamma$ between temperature and
-salinity. 
-\begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.pdf}
-  \caption{Density contour plots with different correlations between
-    temperature and salinity. The densities have unit variance and the
-    thresholds are identical to the mean values $5^o C$ and
-    $30 mg/l$. X-axis is temperature and y-axis is salinity.}
-\label{illus_bivarDens}
-\end{figure}
-The displayed densities have unit standard deviations for both
-temperature and salinity, but we also study the effect of doubling the
-standard deviations.
-
-Table \ref{tab:sim_rhoab} shows the initial EPs and the associated
-Bernoulli variance (second row) for the examples indicated in Fig.
-\ref{illus_bivarDens}. The EPs increase with the correlation as there
-is a strong tendency to have concurrently low temperature and salinity. The Bernoulli variance is similarly large for high
-correlations. EPs and Bernoulli variances are the same for standard
-deviation $1$ or $2$, which implies that high variability in
-temperature and salinity is not captured in the $p(1-p)$ expression.
-
-\begin{table}[!h] \centering \caption{EP and Bernoulli variance for
-    different correlations and variances (top rows), and expected
-    Bernoulli variances for both temperature and salinity data $\by$ and 
-    temperature $y_2$ (bottom rows).}
-  \begin{tabular}{c|ccc|ccc}
- &\multicolumn{3}{c}{$\sigma_1=\sigma_2=1$} & \multicolumn{3}{c}{$\sigma_1=\sigma_2=2$} \\
-\hline
-Correlation $\gamma$ & 0.2 & 0.6 & 0.8 & 0.2 & 0.6 & 0.8 \\
-\hline
-$p$ & 0.28 & 0.35 & 0.40 & 0.28 & 0.35 & 0.40 \\ 
-$p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
-EIBV, Temp and Salinity data & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
-EIBV, Temperature data only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
-\hline
-\end{tabular}
-\label{tab:sim_rhoab}
-\end{table}
-
-Table \ref{tab:sim_rhoab} (bottom two rows) shows results of expected
-Bernoulli variance calculations. This is presented for a design
-gathering both data types, and for a design with temperature
-measurements alone. When both data are gathered, the measurement model is
-$(Y_{d,1},Y_{d,2})^t=(Z_1,Z_2)^t+\bepsilon$, with $\bepsilon \sim N(0,0.5^2I_2)$, while $Y_d=Z_1+\epsilon$, $\epsilon \sim N(0,0.5^2)$ when only temperature is measured.
-For this illustration, Table \ref{tab:sim_rhoab} shows that the
-expected Bernoulli variance gets lower with larger standard deviations
-$\sigma_1$ and $\sigma_2$ (right columns). The reduction of Bernoulli
-variance is largest for the cases with high correlation
-$\gamma$. Albeit smaller, there is also uncertainty reduction when
-only temperature is measured (bottom row), especially when temperature
-and salinity are highly correlated. When correlation is low
-($\gamma=0.2$), there is little information about salinity in the
-temperature data, and therefore less uncertainty reduction. In an
-application with fresh cold water from a river source, the temperature
-and salinity variables will not only be interdependent, but will also
-likely show dependence in the spatial dimension. This in turn will
-impact the design criteria when we evaluate the information measure by
-integrating over several locations (Section \ref{sec:simulations}).
-
-%\subsection{Expected classification criteria}
-
-%{\bf{I have not done anything here - skip this, I think}}.
-
-%Rather than minimizing expected variance one can aim at minimizing the classification probability of the excursion set: 
-%$\min [p,(1-p) ]$, see e.g. \cite{lilleborge2016information}.
-
-%Again, since the conditional mean is a linear in the data, we only have to look at the %relevant linear combination via $\bm_{\xi}=E(\bxi)$.
-%See also \cite{bhattacharjya2013value}.
-
-%\begin{equation}
-%E(\min \{ p,(1-p)\})=\int \min P(\xi_1 \leq t, \xi_2 \leq s),[1-P(\xi_1 \leq t, \xi_2 \leq s)] p(E(\bxi)=) dE(\bxi)=,
-%\end{equation}
-%for the complementary probabilities, this is again evaluated by the corner regions for the block, and $\Phi_4()$ evaluations are required.
-
-%In the end, this result is integrated over the spatial domain $x \in X$.
-
-\section{Sequential designs and heuristic path planning}\label{sec:heuristics}
-
-An important concept underlying this work is the potential of robotic
-sampling to adapt survey plans based on data and statistical methods
-in accordance with the sense-plan-act control loop
-(Fig. \ref{fig:envir1}). We now present a few adaptive strategies for
-robotic sampling that build on the notion of EIBV.
-
-\subsection{Optimal Sequential Design}
-\label{Optdes}
-
-Our adaptive AUV survey is split into many stages. The selection on
-where to sample is constrained by a triangular waypoint graph and
-defined by its nearest grid nodes.
-
-Let $d^{j,s}$ denote design number $j$ at stage $s$ of a sequential
-survey.
-For the AUV sampling situation, the choice of design entails a commitment to one of many possible navigation directions on the waypoint graph.
-Once the choice has been made, the AUV steers in the chosen direction and this entails sampling locations $\bu_{d^{j,s}}$ and data $\by_{d^{j,s}}$ is gathered.
-We let
-$\mathcal{Y}^{s-1} =\{(d^{*,1}, \by_{d^{*,1}}),\ldots, (d^{*,s-1},
-\by_{d^{*,s-1}}) \}$ denote the set of all data gathered until stage
-$s-1$ according to selected design $d^{*,1},\ldots,d^{*,s-1}$.
-
-The design and data define
-the conditional GP with mean
-$E(\gp[\x]|\mathcal{Y}^{s-1})$ and (cross) covariance matrix
-$\mbox{Var}(\gp[\x]|\mathcal{Y}^{s-1})$.
-These are efficiently updated in a sequential manner keeping track of the mean and covariance kernel of the GP over the stages.
-As extensions of Eq.
-\eqref{eq:cokrig} and \eqref{eq:krig:cov}, we have
-\begin{eqnarray}\label{eq:krigSUR}
-\mu^{s}(\x) &=&\mu^{s-1}(\x)+\lambda^{[s,s]}(\x,\bu_{d^{j,s}})^T (\by_{d^{j,s}}-G_{d^{j,s}} \mu(\x)), \nonumber \\
-k^{s}(\x,\x') &=& k^{s-1}(\x,\x')-\lambda^{[s,s]}(\x,\bu_{d^{j,s}})^T K_{Y_{d^{j,s}}} \lambda^{[s,s]}(\x',\bu_{d^{j,s}}),
-\end{eqnarray}
-where $\lambda^{[s,s]}(\x,\bu_{d^{j,s}})=K^{-1}_{Y_{d^{j,s}}} k(\bu_{d^{j,s}},\x)$ and the data covariance is $K^{-1}_{Y_{d^{j,s}}}=G_{d^{j,s}} K^s(\bu_{d^{j,s}},\bu_{d^{j,s}})G^T_{d^{j,s}}+R_{d^{j,s}}$, with measurement model and design matrix defined as in equation \eqref{eq:meas}
-
-At stage $s$, the robot must then choose design $d^{j,s}$ where measurements $\by_{d^{j,s}}$ should
-be gathered. For the triangularized waypoint
-graph all six possible
-directions must be evaluated to minimize the expected uncertainty of random sets, and in particular the EIBV for our 
-application. We hence focus on choosing designs for data modeled by
-equation \eqref{eq:meas}, to conduct efficient sequential uncertainty
-reduction according to the EIBV criterion.  Notably, compared with the
-static design $d$ in equation \eqref{eq:eibv}, the optimal sequential
-design not only considers the best current design grid point, but also
-how the data gathered at this node could inform future sampling at
-successive stages.
-
-%The optimal path
-%selection situation is depicted in Fig. %\ref{fig:PathSelOpt},
-%\begin{figure}[b!]
-%\centering
-%\includegraphics[width=0.85\textwidth]{Figures/sequent_select.pdf}
-%\caption{Optimal sequential path selection.}\label{fig:PathSelOpt}
-%\end{figure}
-%where design choices are indicated by squares while %data realizations are indicated by circles. 
-%terms this optimal design is defined as
-%\begin{equation}\label{opt_crit}
-%    \bd^* = \mbox{argmin}_{d^{j,1}} \left\{ \int_{\by^{j,1}} \mbox{argmin}_{d^{j,2}} \left\{ \int_{\by^{j,2}} \ldots \pi(\by^{j,2}|\by^{j,1}) d\by^{j,2} \right\} \pi(\by^{j,1}) d \by^{j,1} \right\},
-%\end{equation}
-%where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages. 
-The mathematical expression for the optimal design involves a series
-of intermixed maximizations over designs and integrals over data. In
-practice, the optimal solution is intractable because of the enormous
-growth over stages (see e.g. \cite{sucar2015probabilistic} and
-\cite{powell2016perspectives}).  Instead, we outline heuristic
-strategies.
-
-%The derived closed-form solutions for the EIBV are still important building blocks when constructing the sampling designs, as they will be used to score the different adaptive designs. % Efficient
-% calculation is important for using adaptive survey designs with
-% robotic platforms.
-
-\subsection{A Naive Sampling Strategy}
-\label{naive}
-
-The simplest heuristic for adaptive sampling is to choose the next
-sampling location based on current probabilities. One reduces the
-largest levels of uncertainty by selecting the design $j$ with
-$P(Z(\x_j) \in T)$ closest to $0.5$.  With temperature and salinity
-variables, this {\it{naive}} strategy selects the stage $s$ sampling
-design according to:
-\begin{eqnarray}\label{critNaive}
-    d^{*,s} &=& \mbox{argmin}_{j} |P(Z_1(\x_j) \leq t_1, Z_2(\x_j) \leq t_2 | \mathcal{Y}^{s-1})-0.5|, \nonumber
-\end{eqnarray}
-where $j$ is indexing all neighboring locations on the waypoint graph.
-
-This strategy does not account for the uncertainty in the GP
-variables, but is instead applicable only if one design (node) has EPs
-closer to $0.5$ (see Table \ref{tab:sim_rhoab}, row
-two). Additionally, this strategy does not account for spatial
-correlation. This strategy lacks memory of where it has been and where
-the uncertainty has been reduced, and therefore susceptible to local
-minima.
-
-\subsection{Myopic Path Planning}
-\label{sec:myopic}
-
-The myopic (greedy) strategy which we present here is optimal if we
-imagine taking only one more stage of measurements. This selection
-strategy is based on expected uncertainty reduction, but it is a
-heuristic because there is no anticipation of what the subsequent
-designs might offer beyond the first stage.
-
-Based on the currently available data $\mathcal{Y}^{s-1}$, we fit an
-updated GP model. Focusing on the EIBV criterion, at the next stage
-$s$, the selected design is:
-\begin{eqnarray}\label{critSEQ}
-    d^{*,s} &=& \mbox{argmin}_{j} \left\{ \int_{\mathcal{M}} E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( p(\x;\mathcal{Y}^{j,s})\left( 1-p(\x;\mathcal{Y}^{j,s})\right) \right) d\mes(u), \right\} \\
-    p(\x;\mathcal{Y}^{j,s}) &=& P(Z_u \in T |\by_{d^{j,s}},\mathcal{Y}^{s-1}). \nonumber
-\end{eqnarray}
-
-Note that this strategy gives a sequential conditional version of
-expression (\ref{eq:eibv}). Now $\mathcal{Y}^{s-1}$ is available, and
-the expectation is conditional to this data. A similar closed-form
-calculation for EIBV is hence applicable in Eq.  \eqref{critSEQ} using the updated
-GP model from stage $s-1$ as defined via Eq.
-\eqref{eq:krigSUR} in the results from Proposition \ref{propo2}.
-Once the data are collected for the best design $d^{*s}$ at stage $s$, the GP model is updated again.
-The mean and covariance are used to compute the next design at stage $s+1$, and so on.
-
-Even though this myopic strategy is non-anticipative, it still gives a
-reasonable approach for creating designs in many
-applications. Moreover, it is easily implemented, in our case, onboard
-an AUV, using an efficient data update of the GP model and the
-calculation of the closed-form EIBV expressions for each subsequent
-trajectory.
-
-\subsection{Look-ahead Trajectory Planning}
-\label{sec:LA}
-
-We now extend the myopic strategy to a look-ahead strategy. We consider two stages of
-measurements, and the strategy is then optimal in the sense that it accounts consistently for the expectations and minimizations in these two stages, but still without including any planning beyond those two steps.
-In addition to the
-next stage of measurements $\by^{j,s}$, this look-ahead strategy anticipates the subsequent design indexed $j_2$ with data
-$\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$. The
-selected design is:
-\begin{eqnarray}\label{critLA}
-    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( \mbox{argmin}_{j_2}  V_{j_2} \right) \right\}, \\
-V_{j_2} & = & \int_{\mathcal{M}} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1}) ( 1-p(\x;\mathcal{Y}^{j,s+1}) ) \right) d\mes(u), \nonumber \\
-    p(\x;\mathcal{Y}^{j,s+1}) &=& P(Z(\x) \in T |\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}). \nonumber
-\end{eqnarray}
-Here, $\mathcal{Y}^{j,s}=\{\by_{d^{j,s}},\mathcal{Y}^{s-1}\}$ and
-$\mathcal{Y}^{j,s+1}=\{\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}\}$
-represent the sets of data variables, and the expectations are
-calculated with respect to the conditional densities given data at
-that stage.  We solve the first expectation by Monte
-Carlo sampling of data $\by_{d^{j,s}}$ from its conditional
-distribution. For each of these data samples, the second expectation
-is solved using the closed-form expressions for EIBV using the results from Proposition \ref{propo2}, now with conditioning to the first stage data already going into the mean and covariances in Eq.
-\eqref{eq:krigSUR}.
-
-Even though the strategy looks at two sampling stages, it is only used to find
-the current best design. When data are collected at stage $s$, the GP model is
-updated as in Eqn. \eqref{eq:krigSUR}, and the mean and (cross) covariance are
-used to compute the next design for stage $s+1$, now anticipating what
-stage $s+1$ and $s+2$ could offer, and so on.
-This look-ahead approach is much more computationally demanding than
-the myopic strategy, so for practical implementation onboard the AUV this is reaching the limits allowed for calculation.
-We conduct some pruning of the least valuable design paths in the evaluation of Eq.
-\eqref{critLA}, to get a satisfactory computation time below one minute per waypoint.
-This means that we use the myopic
-strategy to rank the three best designs following the first stage alone, and
-for each of these preferred designs, we undertake the look-ahead calculations.
-
-\section{Simulation study}
-\label{sec:simulations}
-
-We now study the properties of the different static and sequential
-survey designs in a realistic case that emulates the spatial
-properties of a river plume interacting with coastal water.
-
-\subsection{Modeling}
-
-A GP model is used to represent the bivariate random fields of
-temperature and salinity. At location $\x \in \mathcal{M}$, the
-bivariate distribution is:
-\begin{equation}\label{gp_bar}
-\begin{bmatrix}Z_1(\x) \\
-Z_2(\x) \end{bmatrix}
-\sim N \left( 
-\begin{bmatrix} \mu_{1}(\x)\\
-\mu_{2}(\x)
-\end{bmatrix},\begin{bmatrix}
-\sigma_{1}^2 & \sigma_{1} \sigma_2 \gamma \\
-\sigma_{1} \sigma_2 \gamma  & \sigma_{2}^2 
-\end{bmatrix}
-\right).
-\end{equation}
-We specify the mean by $\mu_{1}(\x)=5.8 + 0.065 u_{\mbox{West}}$ and
-$\mu_{2}(\x)=29 + 0.1 u_{\mbox{West}}$, which means that the expected
-levels of both temperature and salinity increase linearly in the
-western direction.  This situation mimics that of a river mouth
-opening from the south, with the water masses pulled to the east by
-the Coriolis force \kc{This seems overly specific to the Nidelva use
-  case. Can it be generalized, after all that is the overall intent of
-  this manuscript?}.  The variances are set to $\sigma_1=0.25$ and
-$\sigma_2=0.6$, while the correlation is $0.6$.  The spatial
-correlation is of separable Matern type (see e.g. \cite{Cressie:11})
-so that
-\begin{equation}\label{v}
-\small
-\mbox{Corr} 
-\left(
-\begin{bmatrix}
-    Z_1(\x) \\
-    Z_2(\x) 
-    \end{bmatrix},
-    \begin{bmatrix}
-    Z_1(\x') \\
-    Z_2(\x') 
-    \end{bmatrix}
-    \right)
-    = \begin{bmatrix}
-1 & 0.6  \\
-0.6  & 1
-\end{bmatrix}(1+\phi h)\exp (-\phi h),
-\end{equation}
-where $h=\sqrt{||\x-\x'||^2}$ and $\phi=0.3$ indicates an effective
-correlation range of about $1200$ m.
-
-In the accompanying Python code, these modeling assumptions can be
-generalized to anisotropic covariance and changing variance levels
-across the spatial domain. One can hence, easily run the code to see
-if different models result in other sampling designs. Anisotropy and
-non-stationary variance are both relevant for the setting with river
-plumes, but in practice this requires more parameters to be
-specified. With extensive data and prior knowledge, one could also
-possibly fit and estimate parameters of more complex multivariate
-spatial covariance functions
-\citep{gneiting2010matern,genton2015cross}, but that is outside the
-scope of the current paper.
-
-We discretize the spatial domain $\mathcal{M}$ to a set of $n$ grid
-locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,n \}$, \kc{is the index
-  correct? TeX shows it as u or g for $\mathcal{M}$} where each cell
-has area $\Delta$; the same grid is used for the waypoint graph for
-possible design locations. The EIBV is approximated by sums over all
-grid cells.
-
-One realization of the random fields representing salinity and
-temperature is shown in Fig. \ref{fig:true_temp} and
-\ref{fig:true_sal}. The true ES is a realization from this model as
-shown in Fig. \ref{fig:ESet}.
-
-\begin{figure}[!h]
-  \centering
-  \subfigure[Simulated temperature.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/true_temp.pdf}\label{fig:true_temp}}
-  \hfill
-  \subfigure[Simulated salinity.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/true_sal.pdf}\label{fig:true_sal}}
-  \hfill
-  \subfigure[Excursion set given the realization of temperature and salinity.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_true.pdf}\label{fig:ESet}}
-  \hfill
-  \subfigure[Excursion probabilities given collected data using the
-  `static\_north" strategy.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
-  \caption{\ref{fig:true_temp} and \ref{fig:true_sal} show a
-    realization of the simulated temperature and salinity fields. The
-    associated ES is in \ref{fig:ESet}, while \ref{fig:Eprob} shows
-    the estimated EPs after performing data collection along the
-    static N-S survey line.}
-\label{fig:realisations}
-\end{figure}
-
-\subsection{Static and Sequential Sampling Designs}
-\label{sec:sampling_designs}
-
-%Three different designs are considered as indicated in Figure \ref{fig:stat_design}. 
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.65\textwidth]{Figures/Des3.eps}
-%\caption{Three different static survey designs plotted on the initial EP.}\label{fig:stat_design}
-%\end{figure}
-%In this display the designs are plotted along with the prior probability contours of the ES for the reference parameter inputs.
-
-We compare three different static designs denoted
-\textit{static\_north}, \textit{static\_east}, and
-\textit{static\_zigzag} with the three described sequential approaches
-\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
-sampling paths are pre-determined and cannot be altered and represent
-the pre-planned strategies used in most current AUV operational survey
-designs.
-
-For a fixed survey length, a closed-form expression for the EIBV is
-available as in Eq. \eqref{eq:eibv}. However, for the sequential
-approaches this is not the case. For comparison the properties are
-therefore evaluated using Monte Carlo sampling over several replicates
-of realizations from the model while conducting simulated sequential
-surveys for each one. Fig. \ref{fig:Eprob} shows the conditional EP,
-given data gathered along the north-south survey lines in these
-displays. In the Monte Carlo replicates, such results are repeatedly
-computed to approximate the EIBV. We also compare predictive
-performance measured by root mean square error (RMSE) for temperature
-and salinity estimates as well as also the variance reduction in these
-two variables. It is important to note that the objective function
-used by the agent \kc{First of 'agent'. Define it or change it
-  to 'robot'? Note further uses below.} is focused on reducing the
-EIBV, but we nevertheless expect that we will achieve good predictive
-performance for criteria such as RMSE as well. Another non-statistical
-criteria that is important for practical purposes is the computational
-time needed for the strategy, as this will impact the performance for
-an embedded system.
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
-%\caption{The equilateral waypoint graph used to discretize the
-%  trajectory choices over the $31\times31$ grid used to discretize the GP.}
-%\label{fig:wp_graph}
-%\end{figure}
-
-\begin{figure}[!h] 
-\centering 
-\subfigure[The waypoint graph.]{\includegraphics[width =
-0.49\textwidth]{Figures/sim/wp_graph_paper.pdf}\label{fig:wp_graph_a}}
-\hfill
-\subfigure[The waypoint graph in 3D.]{\includegraphics[width =
-0.49\textwidth]{Figures/sim/wp_graph_3d.pdf}\label{fig:wp_graph_b}}
-\caption{\ref{fig:wp_graph_a} The equilateral waypoint graph used to discretize the
-trajectory choices over the $31\times31$ grid used to discretize the GP.
-\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment.}
-\label{fig:wp_graph}
-\end{figure}
-
-Each strategy is conducted on an equilateral grid as shown in
-Fig. \ref{fig:wp_graph}. The sequential sampling agent starts at the
-center East-West coordinate at the southern end of the domain (node 53
-\kc{Maybe color this node differently and modify the caption?}). The
-AUV moves along edges in the waypoint graph while the vehicle makes
-measurements. The data is assimilated into the GP model before an
-evaluation of the next node to sample is conducted at the end of the
-edge.
-
-\subsection{Simulation Results}
-
-A total of 100 replicate simulations were conducted with all
-strategies. The results are shown in Fig. \ref{fig:sim_results}, where
-the different criteria are plotted as a function of survey
-distance. Fig. \ref{fig:avg_ev} shows the resulting drop in IBV for
-each of the six strategies. IBV reduction occurs most under the
-\textit{myopic} and \textit{look-ahead} strategies, each performing
-almost equally; this is expected as the two criteria
-(Eq. \eqref{critSEQ} and \eqref{critLA}) are sensitive to differences
-in IBV. The \textit{static\_north} design also does well here because
-the path is parallel to the boundary between the water masses.
-
-\begin{figure}[h!]
-  \centering
-  % \subfigure[Excursion set variance $E_{\by}(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
-  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
-  \hfill
-  \subfigure[RMSE between estimated field and truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
-  \hfill 
-  \subfigure[Explained variance $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
-  \hfill 
-  \subfigure[Computational time for inferencing.\kc{I see only 
-    lines associated with 4 variables showing in the graph. Where is
-    static\_north and static\_east?}]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}} 
-\caption{Simulation results from 100 replicate simulations for 10
-  sampling choices/stages on the grid. \kc{Vertical lines show large
-    variation in replicate results.}}  
-\label{fig:sim_results}
-\end{figure}
-
-Fig. \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop
-in RMSE and increase in explained variance, respectively. Both
-\textit{myopic} and \textit{look-ahead} strategies perform well here,
-but some of the \textit{static\_east} and \textit{static\_zigzag} also
-achieve good results because they are pre-determined to cover large
-parts of the domain without re-visitation. Sequential strategies
-targeting IBV will sometimes not reach similar coverage, as
-interesting data may draw the path into twists and turns. There is
-relatively large variety in the replicate results as indicated by the
-vertical lines. Nevertheless, the ordering of strategies is similar.
-
-\begin{figure}[!b]
-  \centering
-  \subfigure[Look-ahead]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_look-ahead.pdf}\label{fig:avg_look-ahead}}
-  \hfill
-  \subfigure[Myopic]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_myopic.pdf}\label{fig:avg_myopic}}
-  \hfill
-  \subfigure[Naive]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_naive.pdf}\label{fig:route_naive}}
-  \hfill
-  \subfigure[Static]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_static.pdf}\label{fig:route_static}}
-  \hfill
-  \caption{The overlaid route choices (strategies), superimposed on
-    the 100 replicate surveys with 10 sampling choices/stages. \kc{I
-      would reorder these figurs above based on the order the material
-      was discussed -- Static, naive, myopic and then Look-ahead. Also
-      do you really need a key associated with each subfigure, given
-      the caption? An abridged analysis as in the text, might also be
-      in order here in the caption. Also refer to
-      Fig. \ref{fig:realisations} as the reference, where the
-      temp/salinity fields are shown.}}
-\label{fig:route_choices}
-\end{figure}
-
-Fig. \ref{fig:avg_time} shows the computational effort: the
-\textit{naive} strategy is on par with the static designs, while the
-\textit{myopic} strategy is slower. The \textit{look-ahead} is even
-slower, reaching levels that are nearly impractical for execution on a
-vehicle. Some pruning of the graph is performed to improve the
-performance, such as ruling out repeated visitations and
-back-and-forth routes. Some of the intermediate results are also
-stored for longer planning horizons. Further pruning of branches or
-inclusion of other heuristics could be included for better
-performance. Then again, the inclusion of heuristics is likely a
-contributing factor for the \textit{look-ahead} strategy failing to
-outperform the \textit{myopic} strategy.
-
-In Fig. \ref{fig:route_choices}, the realized sampling paths for each
-of the sequential schemes and static designs are shown. The
-\textit{naive} strategy often gets stuck in the southern part of the
-domain because it is too focused on the probabilities near $0.5$. The
-\textit{myopic} strategy covers a wider domain than the naive or
-look-ahead. There are several reasons for this. First, a greedy
-approach will tend to put more emphasis on promising locations close
-to the agent, which may lead away from the centre. Second, as the
-agent evaluates the impact of locations further away (look-ahead)
-where assimilated data has less predictive power, the GP model (which
-is centered here) will act to restrict paths deviating from the
-central zone.
-
-We studied the sensitivity of the results by modifying the input
-parameters to have different correlations between temperature and
-salinity, standard deviations, and spatial correlation range.  In all
-runs, the \textit{myopic} and \textit{look-ahead} strategies perform
-the best in terms of realized IBV, and much better than
-\textit{naive}. The \textit{look-ahead} strategy seems to be
-substantially better than the \textit{myopic} design only for very
-small initial standard deviation or very large spatial correlation
-range. \textit{static\_north} continues to be the best static design
-for IBV, while \textit{static\_zigzag} is the best design for the
-other predictive performance measures, especially so with large
-spatial correlation range. We also ran simulation studies with only
-temperature data, and for realistic correlation levels between
-temperature and salinity, the IBV results are not much worse when only
-temperature data are available. In addition to the comparison made in
-Table \ref{tab:sim_rhoab}, the current setting includes spatial
-correlation and this likely reduces the additional influence of having
-bivariate data. However, it seems that having temperature data alone
-does a substantially worse job in terms of explained variance
-
-\begin{comment}
-\subsection{Performance Metrics}
-
-The tables below explore the sensitivity of the result to different input correlation values between temperature and salinity measured onboard the AUV, standard deviations, and spatial
-correlations (see table rows). We only compare results at the end of
-each survey. The expected ES variances are shown in Table
-\ref{tab:sim_res_ev}, results of RMSE are in Table
-\ref{tab:sim_res_rmse}, and explained variance is in Table
-\ref{tab:sim_res_r2}. In all runs, both the \textit{myopic} and
-\textit{look-ahead} strategies perform the best in terms of expected
-variance in the ES. The \textit{myopic} strategy
-performs slightly better than the \textit{look-ahead}. Part of the
-reason for this discrepancy can be traced to the act of pruning some
-of the paths for the \textit{look-ahead} strategy. The \textit{static\_north} design is clearly
-better than the other two static designs, the reason being that the
-others do not focus on the areas where the ES is really uncertain
-(centre locations are the most uncertain). By measuring temperature
-alone, the expected variances in the ES are only slightly higher than
-the combined temperature-salinity estimate. This occurs because the
-correlation of $0.6$ means that temperature observations provide
-information about salinity as well. For the other predictive measures,
-\textit{myopic} and \textit{look-ahead} strategies do well, clearly
-beating the \textit{naive} strategy.
-
-\begin{table}[!h]
-    \centering
-    \scalebox{0.87}{
-    \begin{tabular}{lrrrrrr}
-    \toprule
-    Parameter: $E_{\by}(p[1-p])$ &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
-    \midrule
-                    \rowcolor{Gray}
-ts. cor. low: 0.2  &  29.65 &   27.08 &       \textbf{26.44} &         32.33 &        40.31 &          36.23 \\
-ts. cor. high: 0.8 &  36.18 &   \textbf{28.42} &       30.30 &         32.99 &        37.18 &          36.04 \\
-                    \rowcolor{Gray}
-std. low: 0.1      &  18.26 &   15.79 &       \textbf{15.35} &         18.89 &        26.65 &          26.10 \\
-std. high: 0.5     &  47.01 &   \textbf{41.31} &       43.15 &         47.92 &        52.63 &          49.71 \\
-                    \rowcolor{Gray}
-cor. low: 0.8      &  44.42 &   \textbf{42.02} &       42.47 &         43.84 &        46.05 &          47.82 \\
-cor. high: 0.2     &  29.73 &   21.09 &       \textbf{20.43} &         27.91 &        37.60 &          34.35 \\
-                    \rowcolor{Gray}
-temp. only         &  38.25 &   29.16 &       \textbf{28.69} &         34.32 &        39.81 &          36.17 \\
-basecase           &  35.21 &   \textbf{28.56} &       28.94 &         33.50 &        39.84 &          39.26 \\
-    \bottomrule
-    \end{tabular}}
-  \caption{Simulation results for the final mean IBV (Eq. \eqref{two_partsK}).}
-    \label{tab:sim_res_ev}
-\end{table}
-\vspace{-0.4cm}
-\begin{table}[!h]
-    \centering
-    \scalebox{0.92}{
-        \begin{tabular}{lrrrrrr}
-        \toprule
-        Parameter: RMSE &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
-        \midrule
-                \rowcolor{Gray}
-ts. cor. low: 0.2  &   0.63 &    \textbf{0.57} &        0.59 &          0.62 &         0.57 &           0.59 \\
-ts. cor. high: 0.8 &   0.63 &    \textbf{0.59} &        0.61 &          0.64 &         0.60 &           0.60 \\
-                \rowcolor{Gray}
-std. low: 0.1      &   0.39 &    0.37 &        0.38 &          0.38 &         0.40 &           \textbf{0.35} \\
-std. high: 0.5     &   0.85 &    0.82 &        0.84 &          0.82 &         0.83 &           \textbf{0.81} \\  
-                \rowcolor{Gray}
-cor. low: 0.8      &   0.68 &    0.66 &        0.67 &          0.68 &         0.66 &           \textbf{0.65} \\
-cor. high: 0.2     &   0.58 &    0.54 &        0.55 &          0.58 &         0.54 &           \textbf{0.49} \\
-                \rowcolor{Gray}
-temp. only         &   0.65 &    0.63 &        0.62 &          0.65 &         \textbf{0.61} &           0.62 \\
-basecase           &  0.61 &   \textbf{0.56} &       \textbf{0.56} &         0.60 &        0.57 &          0.57 \\
-        \bottomrule
-        \end{tabular}}
-      \caption{Simulation results for the final mean RMSE ($\frac{1}{B} \sum_{b=1}^{B} \sqrt{(\bmu-\tilde{\bmu})^2}$).}
-    \label{tab:sim_res_rmse}
-\end{table}
-\vspace{-0.4cm}
-\begin{table}[!h]
-    \centering
-    \scalebox{0.95}{
-        \begin{tabular}{lrrrrrr}
-        \toprule
-        Parameter: $\bR^{2}$ &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
-        \midrule
-                        \rowcolor{Gray}
-ts. cor. low: 0.2  &  27.00 &   33.29 &       32.24 &         26.71 &        30.56 &          \textbf{33.50} \\
-ts. cor. high: 0.8 &  23.68 &   \textbf{33.87} &       32.42 &         26.72 &        30.73 &          33.61 \\
-                        \rowcolor{Gray}
-std. low: 0.1      &  25.38 &   30.80 &       30.29 &         26.09 &        28.75 &          \textbf{31.66} \\
-std. high: 0.5     &  26.11 &   34.22 &       32.48 &         26.95 &        31.77 &          \textbf{34.48} \\
-                        \rowcolor{Gray}
-cor. low: 0.8      &   8.80 &   11.50 &       11.23 &          9.56 &        12.31 &          \textbf{12.58} \\
-cor. high: 0.2     &  37.57 &   \textbf{48.68} &       48.11 &         39.61 &        41.98 &          48.04 \\
-                        \rowcolor{Gray}
-temp. only         &  11.67 &   \textbf{22.82} &       21.98 &         18.16 &        20.77 &          22.78 \\
-basecase           &  24.71 &   33.45 &       32.25 &         26.71 &        30.62 &          \textbf{33.54}\\
-        \bottomrule
-        \end{tabular}}
-      \caption{Simulation results for the final mean explained variance $\bR^{2}=100*(1-(\Sigma_{posterior}/\Sigma_{initial}))$).}
-    \label{tab:sim_res_r2}
-\end{table}
-
-\end{comment}
-
-\section{Case Study - Mapping a River Plume}
-\label{sec:case_study}
-
-To demonstrate the applicability of using multivariate EPs and the IBV
-to inform oceanographic sampling, we present a case study mapping a
-river plume with an AUV. The experiment was performed in Trondheim,
-Norway, surveying the Nidelva river (Fig. \ref{fig:nidelven}). The
-experiments were conducted in late Spring 2019, when there is still
-snow melting in the surrounding mountains so that the river water is
-substantially colder than the water in the fjord. The experiment was
-focused along the frontal zone that runs more or less parallel to the
-eastern shore as noted in Fig. \ref{fig:nidelven}.
-
-\subsection{Model Specification}
-\label{sec:exp_modeling}
-
-The statistical model parameters were specified based on a short
-preliminary survey where the AUV made an initial transect to determine
-the trends in environmental conditions and correlation
-structures. Based on the initial data, the trend parameters were
-estimated by linear regression, where both temperature and salinity
-are assumed to increase linearly, going west from the river
-mouth. Next, the residuals from the regression analysis were analyzed
-to study the fit of the GP model and to specify the covariance
-parameters.
-
-%\begin{figure}[!h] 
-% \centering 
-%\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
-%\caption{Data analysis from a preliminary trial experiment using the
-%  AUV. Left: Residual plot of temperature and salinity along with
-%  Gaussian contours. Middle: Empirical CDF (solid) of the quadratic form of
-%  the residuals along with the theoretical CDF (dashed) of the $\chi^2$
-%  distribution with two degrees of freedom. Right: Empirical variogram
-%  of the salinity and temperature data.} \label{fig:parest}
-%\end{figure}
-
-\begin{figure}[!h]
-  \centering
-  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_a.pdf}\label{fig:parest_a}}
-  \hfill
-  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_b.pdf}\label{fig:parest_b}}
-  \hfill
-  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_c.pdf}\label{fig:parest_c}}
-  \caption{Data analysis from a preliminary trial experiment using the
-    AUV. \ref{fig:parest_a} Residual plot of temperature and salinity
-    along with Gaussian contours. \ref{fig:parest_b} Empirical CDF
-    (solid) of the quadratic form of the residuals along with the
-    theoretical CDF (dashed) of the $\chi^2$ distribution with two
-    degrees of freedom. \ref{fig:parest_c} Empirical variogram of the
-    salinity and temperature data.}
-\label{fig:parest}
-\end{figure}
-
-Fig. \ref{fig:parest} summarizes diagnostic plots of this
-analysis. Fig. \ref{fig:parest_a} shows a cross-plot of temperature
-and salinity residuals after the westerly trends in both salinity and
-temperature are subtracted from the data. This scatter-plot of joint
-residuals indicates larger variability in salinity than in
-temperature, and a positive correlation ($0.5$) between the two
-variables. Based on the fitted bivariate Gaussian model (ellipses in
-Fig. \ref{fig:parest_a}), we can compute the modeled quadratic form of
-the residuals, and if the model is adequate they should be
-approximately $\chi^2_2$ distributed. Fig. \ref{fig:parest_b} shows
-the empirical cumulative distribution function (CDF) of these
-quadratic forms (solid) together with the theoretical CDF of the
-$\chi^2_2$ distribution. The modeled and theoretical curves are very
-similar, which indicates that the Gaussian model fits reasonably
-well. Even though there appears to be some clustering in both
-Fig. \ref{fig:parest_a} and \ref{fig:parest_b}, the bivariate
-diagnostic plots look reasonable and justify a Gaussian
-model. Fig. \ref{fig:parest_c} shows the empirical variogram of the
-scaled residuals for temperature and salinity. The decay is similar
-for the two, and seems to be negligible after about $150$ m.
-
-Based on the analysis in Fig. \ref{fig:parest}, the resulting
-parameters are given in Table \ref{tab:experiment_param}. The
-regression parameters shown here are scaled to represent the east and
-west boundaries of the domain as seen in the preliminary transect
-data, and the thresholds are intermediate values. These parameter
-values were then used in field trials where we explored the
-algorithm's ability to characterize the river plume front separating
-the river and fjord water masses, providing a spatial map of this
-boundary.
-
-%Mapping the spatial extent of a frontal zones is an important problem for studying many bio-physical interactions in the ocean.
-The frontal zone is determined by the boundary where plumes of sediments, nutrients, and possibly pollutants spreading from the river outlet meet and interact with adjacent coastal water.
-Due to the lower density the plumes spread on the surface, creating a front with an sharp gradient in both temperature and salinity.
-
-\begin{table}[!h]
-\centering
-\begin{tabular}{lrr}
-\toprule
-Parameter & Value & Source\\
-\midrule
-\rowcolor{Gray}
-Cross correlation temp. and sal. & 0.5 & AUV observations\\
-Temp. variance &  0.20 & AUV observations (variogram)\\
-\rowcolor{Gray}
-Sal. variance &  5.76 & AUV observations (variogram)\\
-Corr. range  & 0.15 km & AUV observations (variogram)\\
-\rowcolor{Gray}
-River temp. $T_{river}$ & $10.0\,^{\circ}\mathrm{C}$ & AUV observations\\
-Ocean temp. $T_{ocean}$ & $11.0\,^{\circ}\mathrm{C}$ & AUV observations\\
-\rowcolor{Gray}
-River sal. $S_{river}$ & $14.0$ g/kg & AUV observations\\
-Ocean sal. $S_{ocean}$ & $22.0$ g/kg & AUV observations\\
-\rowcolor{Gray}
-Threshold temp. & $10.5\,^{\circ}\mathrm{C}$ & $(T_{ocean}-T_{river})/2+T_{river}$\\
-Threshold sal. & $18.0$ g/kg & $(S_{ocean}-S_{river})/2+S_{river}$\\
-\rowcolor{Gray}
-\bottomrule
-\end{tabular}
-\caption{Model and threshold parameters from an initial AUV
-  survey. Observations were taken across the front while crossing from
-  fresh, cold river water to saline and warmer ocean waters. \kc{Might
-  be good to expand the terms and not have abbreviations in the table
-  above; there's plenty of space.}}
-\label{tab:experiment_param}
-\end{table}
-
-\subsection{Experimental Setup}
-
-The sampling locations were distributed over an equilateral grid, as
-shown in the grey-colored lattice in Fig. \ref{fig:map}. The robotic
-platform consisted of a Light AUV \citep{sousa2012lauv}
-(Fig. \ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49
-conductivity, temperature, and depth (CTD) sensor providing
-temperature and salinity
-measurements. %The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
-The sampling agent was built on top of the autonomous agent framework
-Teleo-Reactive EXecutive (\textit{T-REX})
-\citep{py10,Rajan12,Rajan12b}, running an instance of the
-\textit{myopic} strategy from Section \ref{sec:myopic} to control the
-AUV and decide between sampling locations.
-
-\begin{figure}[!h] 
-\centering 
-\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
-\caption{The commercially available Light Autonomous Underwater
-  Vehicle (LAUV) platform for upper water-column exploration used in
-  our experiments.}
-\label{fig:lauv}
-\end{figure}
-
-The sampling strategy was designed around the concept of visiting
-waypoints sequentially. Arriving at a desired waypoint with new
-measurements and an updated model, the AUV triggers the myopic
-strategy to evaluate the different design criteria (see
-Eq. \eqref{critSEQ}). The waypoint-and-path combination that is
-expected to reduce the IBV the most is selected, and upon arrival this
-procedure is then repeated. At each stage, it takes the AUV about 30
-seconds to evaluate the EIBV for all the possible waypoint-and-path
-alternatives.
-
-The AUV was set to start in the south-center part of the waypoint
-graph, with the previously outlined GP model of the environment
-(Section \ref{sec:exp_modeling}). A survey was set to take
-approximately 40 minutes, visiting 15 waypoints on the grid, with the
-AUV running near the surface to capture the plume. On its path from
-one waypoint to the next, the AUV gathered data regularly, and the GP
-model assimilated temperature and salinity data with an update
-frequency of 30 seconds, giving about three updates per stage.
-
-\begin{figure*}[!h]
-\centering
-\subfigure[AUV survey area]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
-\hspace{0.3cm}
-\subfigure[Temperature tracks]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/auv.pdf}\label{fig:res_both}}
-
-\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv1_es_sal_ep.pdf}\label{fig:res1}}
-\hspace{0.2cm}
-\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv4_es_sal_ep.pdf}\label{fig:res2}}
-
-%\subfigure[ES for Survey 1]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
-%\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
-\caption{Results from mapping the Nidelva river, Trondheim, Norway
-  over two survey missions. \ref{fig:map} shows an overview of the
-  survey area overlaid with the AUV path in black and dashed
-  line. Note the shaded region indicating a typical frontal
-  region. \ref{fig:res_both} shows the collected temperature data as
-  colored trails. Note waypoint 5 (WP5) which indicates where the two
-  surveys diverge. \ref{fig:res1} and \ref{fig:res2} shows the
-  collected salinity data overlaid on the final EP, which indicate the
-  AUVs statistical impression of the front. For both missions the
-  temperature and salinity data correspond with an indication of the
-  EP front. \kc{Might want to indicate when the surveys (dates) were
-    conducted so people can understand the variability in the EP with
-    respect to time.}}
-\label{fig:results}
-\end{figure*}
-
-\subsection{Results}
-
-Two surveys missions (1 and 2), were run successively from 11:00 AM to
-01:00 PM, with a short break in between. The resulting path of the
-selected waypoints are shown in the map in Fig. \ref{fig:map}, both
-within the expected frontal region (shaded pink). The recorded
-temperatures are shown as colored trails in Fig. \ref{fig:res_both},
-clearly indicating the temperature difference between fjord and
-riverine waters. The salinity data are then shown separately, overlaid
-with the estimated EP for each survey in Fig. \ref{fig:res1} and
-\ref{fig:res2}.
-
-Both surveys successfully estimated and navigated the separation zone,
-crossing the frontal boundary multiple times. As conditions changed
-slightly between the two surveys, the resulting path (after waypoint
-5) is shown to deviate. Survey 1 continued northwards, tracking the
-north-eastern portion of the front, while Survey 2 turned west,
-mapping the south-western region.
-
-The final predictions of the front location, represented by
-conditional EPs in Fig. \ref{fig:res1} and \ref{fig:res2} as dashed
-lines, correspond with one another. In both surveys they yield a
-picture of the front being to the west in the southern portions of the
-region and gradually bending off toward the north east. The amount of
-exploration done by Survey 1 is greater than Survey 2. In Survey 1,
-the AUV obtained more detail by going north from waypoint 5, while
-Survey 2, coming close to the survey area borders in the south-western
-corner, obtained a poorer understanding of the northern parts. A
-look-ahead strategy might identify and discourage such choices.
-
-%\begin{figure}[!h] 
-% \centering 
-%\includegraphics[width=0.48\textwidth]{Figures/envir_ocean.pdf}
-%\caption{Ocean observation is moving away from %single-ship sampling
-%towards more collaborative networked operations in %order to resolve
-%the numerous processes and their interaction.} %\label{fig:envir}
-%\end{figure}
-%\newpage
-\section{Closing remarks}\label{sec:concl_disc}
-
-This work builds on a multidisciplinary effort combining statistical
-modeling with robotic surveying techniques for oceanographic
-applications. We show how observation practices can gain efficiency
-and accuracy from the development of statistical techniques to achieve
-more effective spatial monitoring. We further demonstrate the
-opportunities available for real-time multivariable spatial
-data-gathering and analysis onboard autonomous platforms, which
-statisticians can exploit to create general-purpose toolkits for
-similar applications.
-
-In particular, we derive and show results for characterizing phenomena
-connected to the properties of water masses. The characterization of
-uncertainties in random sets is extended with new results for the EIBV
-uncertainty reduction achieved by spatial sampling design. This is
-provided in closed-form for the situation with a static design, and
-then extended to the adaptive situation. The sequential derivations
-provide new insights into efficient applications of adaptive sampling,
-as demonstrated in our application.
-
-The study did not consider any temporal effects, which would be
-relevant on a larger time scale. We consider the extension to
-spatio-temporal modeling as future work, and envision that
-advection-diffusion equations could be useful in this kind of modeling
-\citep{sigrist2015stochastic}. For more complex oceanographic
-phenomena, the methods will need to be extended to non-Gaussian
-phenomena, possibly feature-based mixtures of GPs which could still be
-run onboard and augmented by dynamical models. \kc{How about talking
-  about removing any grids or waypoints altogether?}
-
-The spatial-statistical design criterion, building on random sets is
-relevant in our setting with different water properties, and we show
-mathematical generality beyond the EIBV which is convenient to
-implement for onboard robotic planning. Such criteria could also be
-useful in other oceanographic settings as for algal-blooms, anoxic
-zones or open water fronts \cite{costa19}. Of course, other criteria
-are also relevant for instance, hybrid or multi-attribute criteria
-that could balance goals of exploration and exploitation in this
-situation. Equally, such techniques have significant use cases in
-downstream decision-making, with policy makers and regulators who need
-to make difficult decisions related to fish farming, other marine
-resource operations. Value of information analysis (VOI)
-\citep{Eidsvik:15} could be used in a similar vein as the IBV in the
-current work, as both analytic methods explore when, where and which
-information is likely to result in improved decision-making.
-
-While more effort could be spent on approximating the look-ahead
-stages in the adaptive designs, it is perhaps more interesting to
-explore the additional flexibility that can be gained by having
-multiple AUVs co-temporally exploring a spatial or spatio-temporal
-domain \citep{ferreira2019advancing}. Such an approach would enable
-concurrent sampling in different parts of the space, or opportunities
-to move in parallel to best capture the excursion set.
-
-\section*{Acknowledgements}
-
-TOF acknowledges support from the Centre for Autonomous Marine
-Operations and Systems
-(AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of
-Excellence, project number 223254, and the Applied Underwater Robotics
-Labortatory (AURLab). JE and KR acknowledge support from Norwegian
-research council (RCN), project number 305445. CT and DG acknowledge
-support from the Swiss National Science Foundation, project number
-178858.
-
-%\begin{supplement}
-%\sname{Supplement A}\label{suppA}
-%\stitle{Title of the Supplement A}
-%\slink[url]{http://www.e-publications.org/ims/support/dowload/imsart-ims.zip}
-%\sdescription{Dum esset rex in
-%accubitu suo, nardus mea dedit odorem suavitatis. Quoniam confortavit
-%seras portarum tuarum, benedixit filiis tuis in te. Qui posuit fines tuos}
-%\end{supplement}
-
-% == Adding references
-\footnotesize
-\bibliographystyle{imsart-nameyear}
-\bibliography{ref}
-
-\section*{Appendix}
-
-Proof of Proposition \ref{propo1}:
-
 \begin{proof}
@@
 
-Proof of Proposition \ref{propo2}:
-
-\begin{proof}
-	By definition of $\Phi_{p}$, for $N\sim \mathcal{N}_{p}(0_{p},\covN)$, 
-	$$
-	\mathbb{P}(N\leq a + BV | V)
-	=
-	\varPhi_{p}\left( a + BV; \covN \right).
-	$$
-	Now for $\varPhi_{p}\left( a + BV; \covN \right)^h$, provided that the probability space is sufficiently large to accomodate $h$ independent Gaussian random vectors $N_i\sim \mathcal{N}_{p}(0,\covN)$ (which is silently assumed here), using the former equality delivers
-	$$
-	\varPhi_{p}\left( a + BV; \covN \right)^h
-	=
-	\prod_{i=1}^h \mathbb{P}(N_i\leq a + BV | V).
-	$$
-	Now by independence of the $N_i$'s we obtain the joint conditional probability 
-	$$
-	\prod_{i=1}^h \mathbb{P}(N_i\leq a + BV | V)
-	=
-	\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV| V),
-	$$
-	whereof, by virtue of the law of total expectation, 
+Proposition~\ref{propo1} enables in turn calculating centred moments of $\mes(\es)$.
+The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ is probably the most relevant of them for the considered purposes, and hence writes
 \begin{equation*}
 \begin{split}
-	\mathbb{E}\left[ \varPhi_{p}\left( a + BV; \covN \right)^h \right]
-	&=\mathbb{E}\left[\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV| V)\right]\\
-	&=\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV)\\
-	&=\mathbb{P}(W_1 \leq a, \dots, W_h\leq a)\\
-	&=\varPhi_{ph}
-	\left(
-	1_{h} \otimes a
-	;
-	(1_{h}1_{h}')\otimes (B\Sigma_{V} B') + 
-	I_{h}\otimes \covN
-	\right),
+\emv
+&=\int_{D^2} \mathbb{P}\left(
+\gp[u]\in T, \gp[v]\in T \right)
+d\mes^{\otimes}(u, v)\\
+&-\left( \int_{D} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
 \end{split}
 \end{equation*}
-	where $\mathbf{W}=(W_1,\dots,W_h)$ with $W_i=N_i- BV \ (1\leq i \leq h)$
-	and the last line follows $\mathbf{W}$ forming a Gaussian vector (by global independence of the $N_i$'s and $V$) and from the definition of $\varPhi_{p h}$.
-The covariance matrix $\mathbf{\Sigma}$ of $\mathbf{W}$ is obtained by noting that $\operatorname{cov}(W_i,W_j)=B \covV B' + \delta_{ij} \covN \
-	(i,j \in \{1,\dots,h\})$.
-	%the latter concatenated Gaussian vector is obtained by straightfoward calculation. 
-\end{proof}
-
-Proof of Proposition \ref{propo3}.
-
-\begin{remark}
-	%    Note that the vector $\bm{a}$ is obtained by stacking $h_i$ times each $a_i$
-	%\begin{equation*}
-	%\bm{a} = \begin{pmatrix}
-	%    \bm{a}_1\\ \vdots\\ \bm{a}_g\end{pmatrix},~
-	%\bm{a}_i =
-	%\begin{pmatrix}
-	%    a_i\\ \vdots\\ a_i\end{pmatrix}
-	%\in\mathbb{R}^{l\times h_i}\end{equation*}
-	%and the covariance matrix is obtained by similar stacking
-	Using blockwise representation for the blocks themselves delivers %$\bm{\Sigma}$ can be expressed as follows
-	\begin{equation*}
-	%    \bm{\Sigma} = \begin{pmatrix}
-	%        \Sigma_{11} & & \Sigma_{1g}\\
-	%        & \ddots &\\
-	%        \Sigma_{g1}&   & \Sigma_{gg}
-	%\end{pmatrix},~
-	\Sigma_{ij} =
-	\begin{pmatrix}
-	B_i\Sigma_{V} B_j' & \dots & B_i\Sigma_{V} B_j'\\
-	\vdots & & \vdots\\
-	B_i\Sigma_{V} B_j' & \dots & B_i\Sigma_{V} B_j'\\
-	\end{pmatrix}
-	+
-	\delta_{ij}
-	\begin{pmatrix}
-	\covN_i & &\\
-	& \ddots &\\
-	&   & \covN_i
-	\end{pmatrix}%\in\mathbb{R}^{lh_i\times lh_j}
-	\end{equation*}
-	Here each $\Sigma_{ij}$ is made of $h_i$ times $h_j$
-	(vertically/horizontally) $p \times p$ sub-blocks, hence possesses $ph_i$ lines and $ph_j$ columns.
-\end{remark}
-
-%Proof TODO along with notation polish.
-
-\begin{proof}
-	The proof relies (again) heavily on the fact that, by definition of $\Phi_{p}$, for any covariance matrix $\covN \in \R^{p \times p}$, $a\in \R^p$, $B\in \R^{p \times q}$, and $N\sim \mathcal{N}_{p}(0_{p},\covN)$,
-	$$
-	\mathbb{P}(N\leq a + BV | V)
-	=
-	\varPhi_{p}\left( a + BV; \covN \right).
-	$$
-	In particular, for globally independent $N_{i,j} \sim \mathcal{N}_{p}(0_{p},\covN_i)$ $(1\leq j \leq h_i, 1\leq i \leq g)$, 
+which boils down in the excursion/sojourn case where $\T=(-\infty, t_1] \times
+\dots \times (-\infty, t_{\no}]$ to
 \begin{equation*}
 \begin{split}
-	\prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i}
-	&=
-	\prod_{i=1}^{g}
-	\prod_{j=1}^{h_{i}}
-	\mathbb{P}(N_{i,j}\leq a_i + B_i V | V)\\
-	&=\mathbb{P}(N_{1,1}\leq a_1 + B_1 V, \dots, N_{g,h_{g}}\leq a_g + B_g V | V),
+\emv
+%\operatorname{Var}[\mes(\es)]
+&=\int_{D^2}
+\varPhi_{2\no}
+\left(
+(\bt, \bt); \mu((u,v)),
+K((u,v),(u,v))
+\right)
+\
+\mathrm{d}\mes^{\otimes} %\mes
+%\productMeasure
+(u,v)\\
+&-\left( \int_{D} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
 \end{split}
 \end{equation*}
-	so that, by the law of total expectation, 
+%
+Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $D^2$. In 
+contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves 
+solely and integral on $D$ and can be expanded in our present settings as follows
 \begin{equation*}
 \begin{split}
-	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
-	=
-	\mathbb{P}(W_{1} \leq 1_{h_{1}} \otimes a_1, \dots, W_{g} \leq 1_{h_{g}} \otimes a_g)
+\operatorname{IBV} %(\es) %;\mes)
+&=\int_{D}
+\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
+d\mes(u) \\
+&=\int_{D}
+\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right)
+-\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
+\mathrm{d}\mes(u).
 \end{split}
 \end{equation*}
-	where 
-	$W_{1}=(N_{1,1}- B_1 V, \dots, N_{1,h_{1}}- B_1 V), 
-	W_{2}=(N_{2,1}- B_2 V, \dots, N_{2,h_{2}}- B_2 V), 
-	\dots, W_{g}=(N_{g,1}- B_g V, \dots, N_{g,h_{g}}- B_g V)$. Noting that $\mathbf{W}=(W_1,\dots, W_g)$ is a centred $p H$-dimensional Gaussian random vector, we finally obtain that 
-	\begin{equation*}
-	\begin{split}
-	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
-	=
-	\varPhi_{p H}\left(\bm{a};\mathbf{\Sigma}\right),
-	\end{split}
-	\end{equation*}
-	with $\bm{a}=(1_{h_{1}} \otimes a_1, \dots, 1_{h_{g}} \otimes a_g)$ and $\bm{\Sigma}=(\operatorname{cov}(W_i,W_j))_{i,j \in \{1,\dots, g\}}$.
-\end{proof}
-
-% AOS,AOAS: If there are supplements please fill:
-%\begin{supplement}[id=suppA]
-%  \sname{Supplement A}
-%  \stitle{Title}
-%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
-%  \sdatatype{.pdf}" 
-%  \sdescription{Some text}
-%\end{supplement}
-
-% === Not used
-% After reviews: - This is moved to Section 2.1.
-
-%Traditional data collection at sea has typically been based on static
-%buoys, Lagrangian floats, or ship-based methods, with significant
-%logistical limitations that directly impact coverage and sampling
-%resolution. Modern methods using satellite remote-sensing provide
-%large-scale coverage but have limited resolution, are limited to
-%sensing the surface of the ocean, and are impacted by cloud cover. The
-%advent of robust mobile robotic platforms \citep{Bellingham07} has
-%resulted in significant contributions to environmental monitoring and
-%sampling. In particular, autonomous underwater vehicles (AUVs), have
-%advanced the state of sampling and consequently have made robotics an
-%integral part of ocean observation; our previous work has contributed
-%to this effort \citep{das11b,Das2015,fossum18b,fossuminformation}. Other ¤ %statistical work in the oceanographic domain include \cite{wikle2013modern}
-%focusing on hierarchical statistical models; \cite{sahu2008space},
-%studying spatio-temporal models for sea surface temperature and
-%salinity data; and \cite{mellucci2018oceanic} looking at the
-%statistical prediction of features using an underwater glider.
-
-%\begin{itemize}
 %
-%\item Full numerical ocean models cannot provide accurate results
-%  online if run on robotic sensing platforms, as onboard computers
-%  cannot deliver the computational power required. Hence statistical
-%  proxy models of the environment must be used for learning where to sample.
-%
-%\item With limited available information about the state of the ocean, there is substantial value in reacting to
-%  information obtained from measurements taken in-situ. This
-%  acquired information must be assimilated into statistical models that
-%  can be used to inform decisions on where to sample
-%  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.
-%
-%\item For sampling problems related to environmental sensing, the
-%  number of choices (i.e. locations, trajectories, and candidate
-%  designs) is enormous, creating a
-%  trade-off between optimization (finding the most resource-efficient
-%  design to collect necessary data) and computability (arriving at a
-%  solution in reasonable time). To successfully resolve features, this
-%  trade-off has to be considered in development and practice.
-%
-%\end{itemize}
-
-%Addressing this, the combination of statistical tools and robotic platforms is a
-%natural symbiosis which enables information-based sensing. Central to
-%this is the ability to model spatially-correlated variables and
-%provide formal measures of uncertainty. Our formulation is based on
-%Gaussian Processes (GPs) as they allow efficient implementation 
-%and evaluation in real time onboard a robotic platform.
-
-%Sampling can, in this context, not simply be distributed evenly ---along simple transects or ``lawn-mover" patterns--- but must instead be prioritized to relevant regions to ensure it is cost-effective while providing adequate coverage and resolution of the area of scientific interest.
-
-%While the focus has often been on
-%biological and anthropogenic impact from micro-plastics to
-%pollution, biological oceanographers have focused intently on studying
-%micro-organisms at the base of the human food web. These organisms are
-%critically impacted by the changing dynamics in the upper water-column,
-%especially in coastal zones which are complex and often hard to observe
-%in space and time. By studying the bio-geochemical processes in the
-%upper water-column scientists can measure the impact of change, natural
-%or anthropomorphic, and provide an informed opinion to policy makers to
-%effect changes in preserving the environment. However, the challenge of 
-% The pressure on marine resources is growing and increased accuracy,
-% resolution, and persistent monitoring of the oceans is crucial for
-% long-term sustainable management.
-
-% A
-% sustained focus on prioritized and efficient data collection strategies
-% have therefore started to emerge. The advent of marine robotic
-% platforms, especially
-
-% provided means to execute this prioritization through the capacity of
-% autonomy and data-driven sampling, where data collection in principle
-% can be optimized. These capabilities have made
-
-% of the emerging sensing practice for ocean science, allowing scientists
-% to increase the observational efficiency and resolution beyond what was
-% previously possible. But how should a robotic platform, such as an AUV,
-% effectively prioritize and identify important regions for sampling? The
-% answer to this question relates to
-
-% , and
-% the application domain is clearly an arena where statisticians can
-% contribute.
-% There has recently been some statistical attention in oceanography:
-
-%From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest.
-Each phenomenon can in turn be characterized by a set of process specific conditions expressed through different measures of key environmental variables, such as temperature or salinity.
-One such measure is the gradient, that can be associated with a number of important processes, such as the vertical location of the thermocline and pycnocline, location of upwelling systems, vertical mixing, eddies, fronts, and currents \cite{sverdrup2006}, as well as distribution, growth, and accumulation of biological activity \cite{SatOceanSoci00, Ryan2014}.
-These gradients create boundaries separating the ocean into process specific regions which are of profound interest to both identify and map effectively.
-Quantification of gradient features is therefore a much needed competence in robotic sampling.
-\end{document}
+Next we investigate uncertainty reduction criteria related to those indicators.
+In order to be able to calculate them in semi-analytical forms relying on integrals of multivariate Gaussian cumulative 
+distribution functions, we will appeal to some key intermediate results presented next.

### 2020-06-12 main_new.tex
Updates from Overleaf

@@
   should look for a citation here. Check with John Ryan}
- One of the bottlenecks in the study of such phenomena lies however in the lack of obervational data with sufficient resolution.
+ One of the bottlenecks in the study of such phenomena lies however in the lack of observational data with sufficient resolution.
 Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of data collection.
@@
 challenging.
-Here the goal is to use AUV measurements towards a better identification of freshwater versus oceanic water, with the help of a vector-valued random field model for both temperature and salinity.
+Here the goal is to use AUV measurements towards a better description of the interface between freshwater versus oceanic water, with the help of a vector-valued random field model for both temperature and salinity.
 As we will present in the next sections, some of the original contributions inspired by this specific problem are actually applicable to further settings and application areas.
@@
 
+\input{section2}
+
+\begin{comment}
 \section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
@@
 %In the end, this result is integrated over the spatial domain $x \in X$.
+\end{comment}
 

### 2020-06-12 section2.tex
Updates from Overleaf

--- 2020-06-09 main_new.tex
@@
-\documentclass[aoas]{imsart}
-
-\input{commands.tex}
-\usepackage{amsmath}
-\usepackage{pstricks,pst-grad}
-\usepackage{graphicx}
-\usepackage{floatrow}
-\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
-\floatsetup[table]{capposition=top}
-\usepackage{subfigure}
-\usepackage[utf8]{inputenc}
-\usepackage{booktabs}                     % horizontal lines in tables
-\usepackage{comment}
-
-% == Enable text degree
-\usepackage{textcomp}
-
-\usepackage{amsthm,amsmath,natbib}
-\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
-
-% == Trygve Test
-\usepackage{color, colortbl}
-\definecolor{Gray}{gray}{0.9}
-
-% put your definitions there:
-\startlocaldefs
-\newcommand{\edcomment}[1]{{\color{green}{\{Editor: #1\}}}}
-\newcommand{\frevcomment}[1]{{\color{blue}{\{Rev 1: #1\}}}}
-\newcommand{\srevcomment}[1]{{\color{red}{\{Rev 2: #1\}}}}
-\newcommand{\trevcomment}[1]{{\color{violet}{\{Rev 3: #1\}}}}
-\endlocaldefs
-
-% New Marcos.
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\input{macros}
-
-\begin{document}
-
-\begin{frontmatter}
-
-% "New Title of the paper"
-\title{Autonomous Oceanographic Data Collection towards Estimating Excursion Sets of Vector-valued Random Fields} 
-\runtitle{Autonomous Oceanographic Data Collection}
-
-% "Former Title of the paper" (DG 08.06.2020)
-%\title{Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields} \runtitle{Excursion Probabilities for Informative Sampling}
-
-\begin{aug}
-\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
-\author{\fnms{Cédric} \snm{Travelletti}\thanksref{t3}, \corref{} \ead[label=e2]{cedric.travelletti@stat.unibe.ch}}
-\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e3]{jo.eidsvik@ntnu.no}}
-\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e4]{david.ginsbourger@stat.unibe.ch}}
-\and
-\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e5]{kanna.rajan@fe.up.pt}}
-
-\affiliation[t1]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
-\affiliation[t2]{Centre for Autonomous Marine Operations and Systems, NTNU.}
-\affiliation[t3]{Institute of Mathematical Statistics and Actuarial Science, University of Bern, Switzerland.}
-\affiliation[t4]{Department of Mathematical Sciences, NTNU.}
-\affiliation[t5]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
-
-\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
-\printead{e1}}
-\address{Cédric Travelletti\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
-Switzerland.
-\printead{e2}}
-\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e3}}
-\address{David Ginsbourger\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
-Switzerland.
-\printead{e4}}
-\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
-  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
-\printead{e5}}
-
-\runauthor{TO. Fossum et al.}
-\end{aug}
-
-\begin{abstract}
-
-  Improving and optimizing oceanographic sampling is a crucial task
-  for marine science and maritime management. Faced with limited
-  resources to understand processes in the water-column, the
-  combination of statistics and autonomous robotics provides new
-  opportunities for experimental design. In this work we develop
-  methods for efficient spatial sampling applied to the mapping of
-  coastal ocean processes by providing informative descriptions of
-  spatial characteristics of ocean phenomena. Specifically, we define
-  a design
-  criterion based on improved characterization of the uncertainty in
-  the excursions of vector-valued Gaussian random fields, and derive
-  tractable expressions for the expected Bernoulli variance reduction
-  in such a framework. We demonstrate how this criterion can be used
-  to prioritize sampling efforts at locations that are ambiguous,
-  making exploration more effective. We use simulations to study the
-  properties of methods and to compare them with state-of-the-art
-  approaches, followed by results from field deployments with an
-  autonomous underwater vehicle (AUV) as part of a study mapping the
-  boundary of a river plume. The results demonstrate the potential of
-  combining statistical methods and robotic platforms to effectively
-  inform and execute data-driven environmental sampling.
-
-%Motivated by the challenges related to efficient allocation of sampling resources in environmental sensing, the combination of Excursion Probabilities and Gaussian process modeling is explored for autonomous robotic sampling of ocean features; enabling information driven measures  sampling efforts to high-interest regions.
-These regions are usually characterized by gradients of measurable environmental variables, e.g., temperature or salinity gradients, on which EPs subsequently can be used...Correlation among samples and multivariate requirements are typical in environmental studies.
-
-\end{abstract}
-
-\begin{keyword}
-\kwd{Ocean Sampling}
-\kwd{Excursion Sets}
-\kwd{Gaussian Processes}
-\kwd{Experimental Design}
-\kwd{Autonomous robots}
-\kwd{Adaptive Information Gathering}
-\end{keyword}
-
-\end{frontmatter}
-\section{Introduction}
-
-%NB: The subsections are used as a temporary instrument to highlight the proposed structure and should probably be abolished upon convergence to a stable version.
-
-\subsection{Oceanic data collection and spatial design of experiments}
-
-%\textcolor{blue}{NB: I started to match previously listed thoughts/ideas with pieces of the former intro.
-Will do more asap.
-Also, Section 3 I find excellent (and more precise than the intro) and start thinking that some bits might end up in the newer intro (while some others could land in a later section).}
-
-Monitoring the world's oceans has gained increased importance in light of the changing climate and increasing anthropogenic impact. 
-\begin{comment} % TO BE RE-INSERTED LATER IN THE ETXT?
-A central problem for understanding these factors is the lack of representative data with sufficient resolution.
-Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of sampling.
-By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
-This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
-\end{comment}
-Central to understanding the changes taking place in the upper
-water-column is knowledge of the bio-geophysical interaction driven by
-an agglomeration of physical forcings (e.g. wind, topography,
-bathymetry, tidal influences, etc.) and incipient micro-biology driven
-by planktonic and coastal anthropogenic input, such as pollution and
-agricultural runoff transported into the ocean by rivers and streams.
-These often result in a range of ecosystem-related phenomena such as
-blooms and plumes, with direct and indirect effects on society. \kc{we
-  should look for a citation here. Check with John Ryan}
- One of the bottlenecks in the study of such phenomena lies however in the lack of obervational data with sufficient resolution.
-Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of data collection.
-Based on collected data and sometimes also of prior knowledge, models from spatial statistics can be used to make predictions of the state variables under consideration over regions of interest.
-In case of multiple continuous state variables being jointly studied relying on spatially scattered data, vector-valued random field models and associated prediction approaches such as co-Kriging (REFs) can be considered nowadays as part of the standard statistical toolkit in oceanographic research and beyond.
-%applications and beyond.
- %By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
-This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
-
-Traditional data collection to understand natural and artificial phenomena at sea has typically been based on static buoys, floats, or ship-based methods,
-with significant logistical limitations that directly impact coverage
-and sampling resolution. Modern methods using satellite remote-sensing
-provide large-scale coverage but have limited resolution, are limited
-to sensing the surface, and are impacted by cloud cover. Numerical
-ocean models similarly find it challenging to provide detail at fine
-scale \citep{Lermusiaux:2006}, and also come with computation costs that can be practically limiting. 
-%limiting computing resources, no time for hard-core ocean model Navier-Stokes, hence GPs and approximations, etc.
-%
-\begin{figure}[!h] 
-  \centering 
-  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
-    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
-  \hfill
-  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
-    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
-  \caption{\ref{fig:envir1} Traditional ocean observation based on 
-    ship-based sampling has been augmented by autonomous
-    robotic vehicles. % and their interactions.  
-    AUV platforms are an integral part of this network being able to
-    reason and make decisions for efficient onboard adaptive sampling.
-    % using the sense-plan-act control approach to autonomous control. 
-    \ref{fig:nidelven} The interaction of river and ocean creates
-    processes that are challenging to map, where the combination of
-    statistics and robotics can play a vital role in enabling more
-    effective oceanographic observation.\kc{The inset ``AUV robot''
-      looks very tacky and ad-hoc; might be best to remove it. Change
-      ``ships'' to
-      ``Ship'', remove the aerial drone, since we didn't use it and
-      don't plan to in this work.}}
-  \label{fig:envir} \end{figure}
-%
-\begin{comment} % Somehow redundant with the parag following below
-Mobile robotic platforms \citep{Bellingham07} have contributed significantly to environmental monitoring and sampling.
-In particular, AUVs have advanced the state of in-situ sampling, and made robotics an integral part of ocean observation, filling parts of the undersampling gap \cite{rudnick03,rudnick18} (See Fig.
-\ref{fig:envir1}).
-\end{comment}
-%
-The advent of robust mobile robotic platforms \citep{Bellingham07} has
-resulted in significant contributions to environmental monitoring and
-sampling in the ocean. In particular, autonomous underwater vehicles (AUVs) have advanced the state of data collection and consequently have made robotics an integral part of ocean observation; %our 
-previous work by some of the authors of the present paper have contributed to this effort \citep{das11b,Das2015,fossuminformation,fossum18b}.
-
-As full numerical ocean models based on complex differential equations cannot be run onboard a marine robot with limited computational capacity, statistical models relying on collected data and random field assumptions appear quite relevant as means to guide AUV data collection trajectories.
-Our main research angle in the present work is to extend sequential design strategies from the world of spatial statistics and computer experiments to frameworks featuring both vector-valued observational data and experimental designs that may be constrained by the nature of measurement paths.
-\\
-
-Surveys with AUVs are usually limited to observations along fixed
-transects that are pre-scripted in mission plans created manually by a
-human operator. Missions can be specified operating on a scale of
-hundreds of meters to tens of kilometers depending on the scientific
-context. Faced with limited coverage capacity, a more effective
-approach is to instead use onboard algorithms to continuously
-evaluate, update, and refine future sampling locations (sense-plan-act
-cycle in Fig. \ref{fig:envir1}), making the information gathering
-\emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
-% This usually occurs on a spatial grid, called a waypoint graph. 
-In doing so, the space of sampling opportunities is still limited by
-the so-called waypoint graph, a discretization of the search domain, but the AUV has flexibility to modify its path at each waypoint based on in-situ sensing and measurements onboard \citep{py10,Rajan12,Rajan12b}.\\
-
-The work presented here is primarily inspired by a cast study 
-%application where an AUV is used to measure both temperature and salinity in 
-pertaining to the spatial characterization of a frontal system generated by
-river plumes. Fig. \ref{fig:nidelven} shows the survey area in
-Trondheim, Norway, where cold freshwater enters from the river,
-creating a strong gradient in both temperature and salinity. Because
-of the local topography and the Coriolis force the cold fresh water
-tends to flow to the east. Depending on the river discharge, tidal
-effects, wind, and temperature differences, this boundary often gets
-distorted. Initial knowledge about the location and evolution of these
-features are highly uncertain, making deterministic planning
-challenging.
-Here the goal is to use AUV measurements towards a better identification of freshwater versus oceanic water, with the help of a vector-valued random field model for both temperature and salinity.
-As we will present in the next sections, some of the original contributions inspired by this specific problem are actually applicable to further settings and application areas.
-\\
-
-The questions tackled here hence pertain to the broader area of spatial data collection and experimental design, yet with the particularity of estimating some regions of the domain, mostly excursion sets, being implicitly defined by a scarcely observed vector-valued random field.
-%because of limited resources to sample the large oceanographic domains, one must plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs.
-The acquired in-situ information must then be assimilated into the statistical model onboard the platform, so that it can be used to re-compute the sampling criterion and in this way inform decisions on where to sample next.
-\\
-Moreover, given constraints on AUV movements and the fact that surveya rely by design in such a framework on successive measurements spread along a trajectory, addressing corresponding design problems calls for particular sequential strategies, possibly also accounting for the finiteness of resources by anticipating multiple steps ahead thereby improving over myopic strategies.
-\\
-
-Here we aim to leverage and extent recent progress in uncertainty quantification and reduction for excursion sets of Gaussian random fields in order to address vector-valued observational cases under constrained design such as in the motivating application of designing AUV trajectories to better distinguish between freshwater from ocean waters by relying on temperature and salinity measurements.
-%approach will take substantial advantage of existing work in the field of excursion set estimation mainly dedicated to computer experiments in the scalar-valued case, and where space exploration can be performed without constraints regarding the distance between successive design points.
-Original contributions compared to recent related work in the framework of scalar-valued computer experiments include: 
-\begin{itemize}
-    \item Extension of two important Sequential Uncertainty Reduction criteria of (REFs) to homotopic or heterotopic vector-valued settings,
-    \item Investigation of myopic and multiple-step ahead batch-sequential strategies for optimizing trajectories with respect to such criteria,
-    \item Replicable experiments with synthetic test cases and original data sets with accompagnying codes 
-\end{itemize}
-Let us briefly review random field modelling and recent advances in targetd sequential design of experiments based on Gaussian Processes before detailing further existing literature tackling the considered problems, our proposed approach, as well as outlining the rest of the paper.
-
-\subsection{Random field modelling and targeted sequential design of experiments}
-
-\begin{itemize}
-    \item 
-\end{itemize}
-
-% From earlier version; to be relocated?
-%Hence statistical proxy models of the environment must be used.
-For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
-\subsection{Related work and proposed approach in a nutshell}
-
-\begin{itemize}
-    \item 
-\end{itemize}
-
-\textcolor{blue}{We might want to re-use some of the following (from earlier version):}
-
-A major contribution of our work is to derive closed-form results for
-this design criteria for situations where the underlying model is
-based on multivariate GPs. Another contribution is that we embed this
-approach in a sequential strategy to produce sampling algorithms, and
-apply this to a real-world scenario of autonomously sampling
-temperature and salinity gradients using an AUV, with the intention of
-characterizing the variability of oceanic and riverine waters in
-typical mixed coastal environments.
-
-Motivating examples relevant for ESs of multivariate processes are
-abundant. In medicine, doctors do not rely solely on a single symptom
-but must see several combined effects before making a diagnosis. In
-our context of environmental sampling, the salinity and temperature
-excursions of a river plume can similarly help characterize the
-underlying bio-geochemical processes
-\citep{hopkins2013detection,Pinto2018}, where sampling can be aimed
-towards reducing uncertainty in the temperature and salinity ES.
-Unlike what has been done in previous plume exploration studies, we
-emphasize the links to spatial statistical modeling and targeted
-multivariate sampling criteria for sequential sampling.
-
-Other statistical
-work in the oceanographic domain include \cite{wikle2013modern}
-focusing on hierarchical statistical models; \cite{sahu2008space},
-studying spatio-temporal models for sea surface temperature and
-salinity data; and \cite{mellucci2018oceanic} looking at the
-statistical prediction of features using an underwater glider.
-In this paper the focus is not on statistical modeling, but rather on statistical results and computations for efficient sampling designs.
-
-We focus on spatial characterization of a frontal system generated by
-river plumes. Fig. \ref{fig:nidelven} shows the survey area in
-Trondheim, Norway, where cold freshwater enters from the river,
-creating a strong gradient in both temperature and salinity. Because
-of the local topography and the Coriolis force the cold fresh water
-tends to flow to the east. Depending on the river discharge, tidal
-effects, wind, and temperature differences, this boundary often gets
-distorted. Initial knowledge about the location and evolution of these
-features are highly uncertain, making deterministic planning
-challenging.
-
-Adaptive in-situ AUV sampling of an evolving frontal feature has been
-explored in \cite{fronts11,Zhang2012,Pinto2018,costa19}. These
-approaches typically use a reactive-adaptive scheme, whereby
-exploration does not rely on a statistical model of the environment,
-but rather adaptation is based on closing the sensing and actuation
-loop. Myopic sampling, i.e. stage-wise selection of the path (on the
-waypoint graph), has been used for surveys
-\citep{singh2009efficient,Binney2013} that focus largely on reducing
-predictive variance or entropy. These criteria are widely adopted in
-the statistics literature on spatio-temporal design as well
-\cite{bueso1998state,zidek2019monitoring}, but variance and entropy
-reduction are independent of the actual data realizations under the
-assumptions of GP models, so it has limited
-decisional flexibility. The use of data-driven adaptive criteria
-was introduced to include more targeted sampling of regions of
-scientific interest in \cite{Low2009} and \cite{fossuminformation}. In
-this paper, we focus on mapping the river plume by rewarding designs
-that improve the classification of ES in temperature and salinity.
-
-\subsection{Outline of the paper and its appendices / supplemental material}
-
-\begin{itemize}
-    \item 
-\end{itemize}
-
-The remainder of this paper is organized as follows: 
-%Section \ref{sec:bg} provides select background on ocean sampling. 
-Section \ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for
-vector-valued GPs. Section \ref{sec:heuristics} builds on these
-assumptions when deriving the sequential design criteria for adaptive
-sampling. Section \ref{sec:simulations} discusses properties of the
-methods in simulation studies. Section \ref{sec:case_study}
-demonstrates the methodology used in field work characterizing a river
-plume and finally, Section \ref{sec:concl_disc} contains a summary and
-a discussion of future work.
-
-%\subsection{}
-
-%\begin{itemize}
-%    \item 
-%\end{itemize}
-
-\begin{comment}
-\section{Former Introduction (Before 08.06.2020)}
-
-Monitoring the world's oceans has gained increased importance in light
-of the changing climate and increasing anthropogenic impact. 
-A central problem for understanding these factors is the lack of representative data with sufficient resolution.
-Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of sampling.
-By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
-This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
-
-The
-advent of robust mobile robotic platforms \citep{Bellingham07} has
-resulted in significant contributions to environmental monitoring and
-sampling in the ocean. In particular, autonomous underwater vehicles (AUVs), have
-advanced the state of sampling and consequently have made robotics an
-integral part of ocean observation; our previous work has contributed
-to this effort \citep{das11b,Das2015,fossuminformation,fossum18b}.
-
-The potential that lies in using information-based sampling methods, which we
-highlight in this paper, is to do efficient environmental sampling and
-overcome common challenges as discussed next. First, full numerical ocean models based on complex
-differential equations cannot be run onboard a marine robot with
-limited computational capacity.
-% provide accurate results for online
-% operations because the onboard computers lack the computational power
-% required.
-Hence statistical proxy models of the environment must be used. For our purpose, we rely on Gaussian process (GP)
-representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
-Second, because of limited resources to sample the large oceanographic domains, one must plan for
-active learning during the operation, where a relevant criterion is
-used to extract the most valuable designs. The acquired in-situ
-information must then be assimilated into the statistical model
-onboard the platform, so that it can be used to re-compute the
-sampling criterion and in this way inform decisions on where to sample
-next. We use a criterion focusing on the prediction of excursion sets
-(ESs) for the multivariate spatial processes of interest. Third, the
-number of sampling designs is enormous, creating a trade-off between
-optimization and computability. One must often compromise in both
-development and practical operations, and we will rely on
-short-horizon optimization routines for sequential designs among
-nearest neighbors points in a spatial grid (waypoint graph).
-
-\kc{The para above needs a rewrite to me. It starts off by stating
-  there are challenges. But then jumps to models without stating the
-  connection with sampling, and then jumps into learning without
-  stating why learning is important. Typically the English in formal
-  text starts with a statement which is supported or clarified. In the
-  two cases above, the connections are simply not obvious. If this was
-  there in the first version of the m/s submitted, I'm not sure why
-  the reviewers didn't catch it. This is not just a stylistic issue;
-  it has to do with making sense and bring the reader/reviewer along.}
-
-\kc{Second, I don't see a motivation for GPs here and as I read
-  further into the text. Even for stats folks, are GPs the ``best''
-  ways to go? Are they obviously the tools one would use in
-  environmental sampling? And even if they are, I think we ought to
-  make the case for it strongly and generally, so this paper is
-  actually accessible and understandable to those not in the Spatial
-  Stats field. I would suggest adding a short para on GPs and why
-  they're important -- i.e. motivate the use of GPs early, so we can
-  freely make attributions to them later, as in the next section, for
-  example.}
-
-The imperfect understanding of causality coupled with the underlying
-fundamental uncertainty of the ocean implies that we need to make a
-large number of measurements. In addition, the use statistical models
-and methods to determine what and how phenomena form, is one way to
-increase the quality of predictive power aiding our understanding of
-upper water-column processes.
-
-With this in mind, we design AUV sampling strategies for reducing the spatially-integrated Bernoulli variance (IBV) in the ES
-indicator variables. 
-ES of random fields are thoroughly presented in \citep{adler2009random}. Recent work on ESs and associated excursion
-probabilities (EPs) connected to spatial statistics include
-\cite{french2013spatio,bolin2015excursion,french2016credible}. Our
-focus is on uncertainty characterization and how this is influenced by
-sampling. In this sense, our work is similar to
-\cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who
-describe analytical results for the IBV for univariate processes.
-
-A major contribution of our work is to derive closed-form results for
-this design criteria for situations where the underlying model is
-based on multivariate GPs. Another contribution is that we embed this
-approach in a sequential strategy to produce sampling algorithms, and
-apply this to a real-world scenario of autonomously sampling
-temperature and salinity gradients using an AUV, with the intention of
-characterizing the variability of oceanic and riverine waters in
-typical mixed coastal environments.
-
-Motivating examples relevant for ESs of multivariate processes are
-abundant. In medicine, doctors do not rely solely on a single symptom
-but must see several combined effects before making a diagnosis. In
-our context of environmental sampling, the salinity and temperature
-excursions of a river plume can similarly help characterize the
-underlying bio-geochemical processes
-\citep{hopkins2013detection,Pinto2018}, where sampling can be aimed
-towards reducing uncertainty in the temperature and salinity ES.
-Unlike what has been done in previous plume exploration studies, we
-emphasize the links to spatial statistical modeling and targeted
-multivariate sampling criteria for sequential sampling.
-
-Other statistical
-work in the oceanographic domain include \cite{wikle2013modern}
-focusing on hierarchical statistical models; \cite{sahu2008space},
-studying spatio-temporal models for sea surface temperature and
-salinity data; and \cite{mellucci2018oceanic} looking at the
-statistical prediction of features using an underwater glider.
-In this paper the focus is not on statistical modeling, but rather on statistical results and computations for efficient sampling designs.
-
-The remainder of this paper is organized as follows: Section
-\ref{sec:bg} provides select background on ocean sampling. Section
-\ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for
-vector-valued GPs. Section \ref{sec:heuristics} builds on these
-assumptions when deriving the sequential design criteria for adaptive
-sampling. Section \ref{sec:simulations} discusses properties of the
-methods in simulation studies. Section \ref{sec:case_study}
-demonstrates the methodology used in field work characterizing a river
-plume and finally, Section \ref{sec:concl_disc} contains a summary and
-a discussion of future work.
-\end{comment}
-
-\begin{comment}
-\section{Ocean Processes and Observation}
-\label{sec:bg}
-
-Central to understanding the changes taking place in the upper
-water-column is knowledge of the bio-geophysical interaction driven by
-an agglomeration of physical forcings (e.g. wind, topography,
-bathymetry, tidal influences, etc.) and incipient micro-biology driven
-by planktonic and coastal anthropogenic input, such as pollution and
-agricultural runoff transported into the ocean by rivers and streams.
-These often result in a range of ecosystem-related phenomena such as
-blooms and plumes, with direct and indirect effects on society. \kc{we
-  should look for a citation here. Check with John Ryan}
-
-Traditional data collection to understand such phenomena at sea has
-typically been based on static buoys, floats, or ship-based methods,
-with significant logistical limitations that directly impact coverage
-and sampling resolution. Modern methods using satellite remote-sensing
-provide large-scale coverage but have limited resolution, are limited
-to sensing the surface, and are impacted by cloud cover. Numerical
-ocean models similarly find it challenging to provide detail at fine
-scale \citep{Lermusiaux:2006}. Mobile robotic platforms
-\citep{Bellingham07} have contributed significantly to environmental monitoring and sampling. In particular, AUVs have
-advanced the state of in-situ sampling, and made robotics an integral
-part of ocean observation, filling parts of the undersampling gap
-\cite{rudnick03,rudnick18}(See Fig. \ref{fig:envir1}).
-
-%Walter Munk referred to the 20\textsuperscript{th} century
-%as the \emph{century of undersampling} \citep{munk2002}, pointing out
-%a lack of sufficient sampling resolution in time and space as a
-%critical issue for characterizing ocean phenomena. Even today,
-%policymakers struggle with limited observability of severe
-%environmental problems; a prominent example is the recent (May 2018)
-%outbreak of harmful algal blooms in Northern Norway\footnote{Report
-%  from the Institute of Marine Research: https://bit.ly/2KZGJfI}, with
-%losses exceeding 12 million tons of salmon valued at \$80 million
-%(Norwegian Directorate of Fisheries). This is a case in point of how
-%adequate sampling efforts could have provided a more accurate
-%assessment and early warning of changing environmental conditions.
-
-\begin{figure}[!h] 
-  \centering 
-  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
-    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
-  \hfill
-  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
-    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
-  \caption{\ref{fig:envir1} Traditional ocean observation based on 
-    ship-based sampling has been augmented by autonomous
-    robotic vehicles. % and their interactions.  
-    AUV platforms are an integral part of this network being able to
-    reason and make decisions for efficient onboard adaptive sampling.
-    % using the sense-plan-act control approach to autonomous control. 
-    \ref{fig:nidelven} The interaction of river and ocean creates
-    processes that are challenging to map, where the combination of
-    statistics and robotics can play a vital role in enabling more
-    effective oceanographic observation.\kc{The inset ``AUV robot''
-      looks very tacky and ad-hoc; might be best to remove it. Change
-      ``ships'' to
-      ``Ship'', remove the aerial drone, since we didn't use it and
-      don't plan to in this work.}}
-  \label{fig:envir} \end{figure}
-
-\end{comment}
-
-\subsection{To be relocated elsewhere}
-The AUVs considered here are powered untethered platforms, that
-operate at $1$-$3$ m/s in the upper water column, where they are free
-to move in six degrees of freedom (6 DOF). The in-water operation time
-capability depends on survey speed, payload sensors and navigation;
-typically this amounts to 8-48 hrs. AUVs typically use single-board
-computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA
-Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared
-cache, and 4 GB of 1600 MHz DRAM) for computation onboard.
-
-\begin{comment}
-Surveys with AUVs are usually limited to observations along fixed
-transects that are pre-scripted in mission plans created manually by a
-human operator. The mission can be specified operating on a scale of
-hundreds of meters to tens of kilometers depending on the scientific
-context. Faced with limited coverage capacity, a more effective
-approach is to instead use onboard algorithms to continuously
-evaluate, update, and refine future sampling locations (sense-plan-act
-cycle in Fig. \ref{fig:envir1}), making the information gathering
-\emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
-% This usually occurs on a spatial grid, called a waypoint graph. 
-In doing so, the space of sampling opportunities is still limited by
-the waypoint graph, but the AUV has flexibility to modify its path at
-each waypoint based on in-situ sensing and measurements onboard
-\citep{py10,Rajan12,Rajan12b}.
-
-We focus on spatial characterization of a frontal system generated by
-river plumes. Fig. \ref{fig:nidelven} shows the survey area in
-Trondheim, Norway, where cold freshwater enters from the river,
-creating a strong gradient in both temperature and salinity. Because
-of the local topography and the Coriolis force the cold fresh water
-tends to flow to the east. Depending on the river discharge, tidal
-effects, wind, and temperature differences, this boundary often gets
-distorted. Initial knowledge about the location and evolution of these
-features are highly uncertain, making deterministic planning
-challenging.
-
-Adaptive in-situ AUV sampling of an evolving frontal feature has been
-explored in \cite{fronts11,Zhang2012,Pinto2018,costa19}. These
-approaches typically use a reactive-adaptive scheme, whereby
-exploration does not rely on a statistical model of the environment,
-but rather adaptation is based on closing the sensing and actuation
-loop. Myopic sampling, i.e. stage-wise selection of the path (on the
-waypoint graph), has been used for surveys
-\citep{singh2009efficient,Binney2013} that focus largely on reducing
-predictive variance or entropy. These criteria are widely adopted in
-the statistics literature on spatio-temporal design as well
-\cite{bueso1998state,zidek2019monitoring}, but variance and entropy
-reduction are independent of the actual data realizations under the
-assumptions of GP models, so it has limited
-decisional flexibility. The use of data-driven adaptive criteria
-was introduced to include more targeted sampling of regions of
-scientific interest in \cite{Low2009} and \cite{fossuminformation}. In
-this paper, we focus on mapping the river plume by rewarding designs
-that improve the classification of ES in temperature and salinity.
-
-\end{comment}
-
 \section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
@@
 
-\subsection{Excursion sets and Integrated Bernoulli variance}
-\label{sec:ES}
+\medskip
 
@@
 separation of cold freshwater from a river and warmer saline waters
-of a fjord. The characterization of uncertainties on the ES is a
-useful goal of this sampling strategy.
+of a fjord.
 
@@
 $(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary
-set (typically in geostatistical studies, a geographical subset of interest) and the
-dimensionality of the output $\no \geq 1$ is arbitrary. We have $\no = 2$ in the example with temperature and salinity in the river plume.
+set and the dimensionality of the output $\no \geq 1$ is arbitrary.
 
 Assume furthermore that the set of interest
-is characterized in terms of a subset of values of the variables, say
-$\T \subset \R^{\no}$, and denote the corresponding set of index
-points by
+is characterized in terms of a subset of values of the response variables, say
+$\T \subset \R^{\no}$, and denote the corresponding set of spatial locations by
 $$
@@
 of several key quantities. Note that changing some $\leq$ inequalities
-to $\geq$ ones would lead to immediate adaptations. In the application
-specific properties following the orthant assumption come into play
-\kc{This last phrase doesn't parse; is it ``..., the following
-  orthant....''?}: We let $t_1$ be a threshold for temperature and
-$t_2$ a threshold for salinity \kc{Might be easier to remember with
-  symbols 't' and 's' in some
-  form.}. The set of index point then defines
-$\mbox{ES} = \{\x \in D : \gp[\x,1] \leq t_1,\gp[\x,2] \leq t_2\}$.
-The associated EPs are
-$p(\x) = P(\gp[\x,1] \leq t_1,\gp[\x,2] \leq t_2)$,
-$\x \in \mathcal{M}$.
-
-For statistical generality, we will however keep the abstract notation
-with $\T$ and general $\es$ \kc{Again, previous phrase doesn't parse.}
-for most of the derivations. Note that assuming that $\gp$ has
-continuous trajectories (almost surely) and $T$ to be closed, then
-$\es$ can be cast as a Random Closed Set
+to $\geq$ ones would lead to immediate adaptations.
+
+If we assume that $\gp$ has
+continuous trajectories (almost surely) and $T$ is closed, then
+$\es$ becomes a Random Closed Set
 \citep{Molchanov2005}. %In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
-Let us in particular consider a locally finite (Borel) measure $\mes$
-on $D$ and investigate the probability distribution of $\mes(\es)$
-through its moments.
+\textcolor{red}{Peut etre supprimer } In what follows, we will be interested in characterizing the distribution of the volume of the excursion set $\es$ with respect to some (locally finite, Borel) measure $\mes$
+on $D$.
+
+\subsection{Background and Notation}
+\subsubsection{Co-kriging with respect to (batches) of generalized locations}
+
+To keep things as general as possible, we will consider performing measurements
+in batches, i.e. collecting several observations at a time. Moreover, we assume
+that given a site $s\in D$, one can choose which component(s) of
+$\gp[s]\in\mathbb{R}^{\no}$ one wants to observe.
+%observation of several components at a time being also allowed. 
+Denoting here for  $s\in D$ and $\ell \in \{1\dots,p\}$ the $\ell\text{th}$ component of $\gp[s]$ by $\gp[s,\ell]$ or 
+$\gp[x]$ with $x=(s,\ell)$, we see that $Z$ can actually also be thought of as a scalar-valued Gaussian random field 
+indexed by $D\times \{1\dots,p\}$. This simple change of representation allows in turn revisiting cokriging as kriging 
+in extended index set, and hence enjoying kriging results (such as batched-sequential update formulae, tackled in
+the next section) without unneccesary complications.
+%
+Let us refer to $x=(s,\ell)$ as (generalized) location and use boldface $s, \ell, x$ to denote batches of sites, 
+response indices, and generalized locations, respectively. A batch $\bm{x}=(x_1,\dots, x_q)$ of generalized locations 
+with $x_i=(s_i,\ell_i)$ $(1\ldots i \ldots q)$ is also by a slight abuse oftentimes denoted $\bm{x}=(\bm{s}, \bm{\ell})$ 
+for convenience.
+%
+We will also use concatenation for the field evaluated at a batch of generalized locations
+\begin{align*}
+\gp[\bm{x}]:=
+\left(\gp[s_1,\ell_1], ...,
+\gp[s_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
+\end{align*}
+%
+Assuming now that $n$ batches of observations are available, with respective sizes $q_1,\dots, q_n$, and that one wishes 
+to predict $\gp[\bm{x}]$ for some batch of $q\geq 1$ generalized locations $\bm{x}=(\bm{s}, \bm{\ell})$, the (simple) 
+cokriging mean would then amount to simple kriging with respect to a scalar-valued Gaussian random field indexed by 
+$D\times \{1\dots,p\}$ and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{s}, \bm{s}')_{\ell, \ell'}$, that is:
+%
+\begin{equation}
+\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})),
+\end{equation}
+where $\mu$ is $Z$'s initial mean function, $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of 
+observed responses of $Z$ at all considered generalized locations, and $\lambda_{[n]}(\bm{x})$ is a vector of weights 
+equal to $k(\bm{x}_{[n]}, \bm{x}_{[n]})^{-1} k(\bm{x}_{[n]}, \bm{x})$ with $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$, 
+$k(\bm{x}_{[n]}, \bm{x}_{[n]})$ being assumed non-singular throughout the presentation. The co-kriging %(conditional) 
+residual (cross-)covariance function (with respect to batches of generalized locations) can also be expressed in the same vein via
+%
+\begin{equation}
+k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T k(\bm{x}_{[n]}, \bm{x}_{[n]}) \lambda_{{[n]}}(\bm{x}').
+\end{equation}
+
+\subsubsection{Co-kriging update formulae}
+
+Let us now consider the case where co-kriging prediction of $Z$ was made with respect to $n$ batches $\bm{x}_i$ of generalized locations, concatenated again within
+$\bm{x}_n=(\bm{x}_1,\dots, \bm{x}_n)$, and one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
+%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
+Thanks to our representation of co-kriging in terms of simple kriging with respect to generalized locations, a strightforward adaptation of the batch-sequential kriging update formulae from \cite{Chevalier.etal2013a} delivers that
+% 
+\begin{equation}
+\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
+\end{equation}
+where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
+$\lambda_{[n+1]}(\bm{x})$ that corresponds to the kriging weigths associated with the last $q_{n+1}$ responses when 
+predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
+%\text{th}$ batch.
+%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
+Similarly, the updated co-kriging residual (cross-)covariance function then writes
+\begin{equation}
+k_{[n+1]}(\bm{x},\bm{x}')=k_{[n]}(\bm{x},\bm{x}')-\lambda_{[n+1,n+1]}(\bm{x})^T k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]}) \lambda_{{[n+1,n+1]}}(\bm{x}').
+\end{equation}
+%\medskip
+%
+Let us remark that, as noted in \cite{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
+extend to Universal Kriging in second-order settings and apply without Gaussian assumption.
+We will now see how the latter formulae are instrumental in deriving semi-analytical formulae for stepwise uncertainty reduction criteria for
+vector-valued random fields.
+
+\subsection{UQ on ES of vector-valued fields with volume moments and friends (focus on VV and IBV)}
+\input{UQonSets}
 
@@
 %In the end, this result is integrated over the spatial domain $x \in X$.
-
-\section{Sequential designs and heuristic path planning}\label{sec:heuristics}
-
-An important concept underlying this work is the potential of robotic
-sampling to adapt survey plans based on data and statistical methods
-in accordance with the sense-plan-act control loop
-(Fig. \ref{fig:envir1}). We now present a few adaptive strategies for
-robotic sampling that build on the notion of EIBV.
-
-\subsection{Optimal Sequential Design}
-\label{Optdes}
-
-Our adaptive AUV survey is split into many stages. The selection on
-where to sample is constrained by a triangular waypoint graph and
-defined by its nearest grid nodes.
-
-Let $d^{j,s}$ denote design number $j$ at stage $s$ of a sequential
-survey.
-For the AUV sampling situation, the choice of design entails a commitment to one of many possible navigation directions on the waypoint graph.
-Once the choice has been made, the AUV steers in the chosen direction and this entails sampling locations $\bu_{d^{j,s}}$ and data $\by_{d^{j,s}}$ is gathered.
-We let
-$\mathcal{Y}^{s-1} =\{(d^{*,1}, \by_{d^{*,1}}),\ldots, (d^{*,s-1},
-\by_{d^{*,s-1}}) \}$ denote the set of all data gathered until stage
-$s-1$ according to selected design $d^{*,1},\ldots,d^{*,s-1}$.
-
-The design and data define
-the conditional GP with mean
-$E(\gp[\x]|\mathcal{Y}^{s-1})$ and (cross) covariance matrix
-$\mbox{Var}(\gp[\x]|\mathcal{Y}^{s-1})$.
-These are efficiently updated in a sequential manner keeping track of the mean and covariance kernel of the GP over the stages.
-As extensions of Eq.
-\eqref{eq:cokrig} and \eqref{eq:krig:cov}, we have
-\begin{eqnarray}\label{eq:krigSUR}
-\mu^{s}(\x) &=&\mu^{s-1}(\x)+\lambda^{[s,s]}(\x,\bu_{d^{j,s}})^T (\by_{d^{j,s}}-G_{d^{j,s}} \mu(\x)), \nonumber \\
-k^{s}(\x,\x') &=& k^{s-1}(\x,\x')-\lambda^{[s,s]}(\x,\bu_{d^{j,s}})^T K_{Y_{d^{j,s}}} \lambda^{[s,s]}(\x',\bu_{d^{j,s}}),
-\end{eqnarray}
-where $\lambda^{[s,s]}(\x,\bu_{d^{j,s}})=K^{-1}_{Y_{d^{j,s}}} G_{d^{j,s}} k(\bu_{d^{j,s}},\x)$ and the data covariance is $K^{-1}_{Y_{d^{j,s}}}=G_{d^{j,s}} K^s(\bu_{d^{j,s}},\bu_{d^{j,s}})G^T_{d^{j,s}}+R_{d^{j,s}}$, with measurement model as in Eq.
-\eqref{eq:meas}.
-
-At stage $s$, the robot must then choose design $d^{j,s}$ where measurements $\by_{d^{j,s}}$ should
-be gathered. For the triangularized waypoint
-graph all six possible
-directions must be evaluated to minimize the expected uncertainty of random sets, and in particular the EIBV for our 
-application. We hence focus on choosing designs for data modeled by
-Eq. \eqref{eq:meas}, to conduct efficient sequential uncertainty
-reduction according to the EIBV criterion. Compared with the
-static design $d$ in Eq. \eqref{eq:eibv}, the optimal sequential
-design not only considers the best current design grid point, but also
-how the data gathered at this node could inform future sampling at
-successive stages.
-
-%The optimal path
-%selection situation is depicted in Fig. %\ref{fig:PathSelOpt},
-%\begin{figure}[b!]
-%\centering
-%\includegraphics[width=0.85\textwidth]{Figures/sequent_select.pdf}
-%\caption{Optimal sequential path selection.}\label{fig:PathSelOpt}
-%\end{figure}
-%where design choices are indicated by squares while %data realizations are indicated by circles. 
-%terms this optimal design is defined as
-%\begin{equation}\label{opt_crit}
-%    \bd^* = \mbox{argmin}_{d^{j,1}} \left\{ \int_{\by^{j,1}} \mbox{argmin}_{d^{j,2}} \left\{ \int_{\by^{j,2}} \ldots \pi(\by^{j,2}|\by^{j,1}) d\by^{j,2} \right\} \pi(\by^{j,1}) d \by^{j,1} \right\},
-%\end{equation}
-%where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages. 
-The mathematical expression for the optimal design involves a series
-of intermixed maximizations over designs and integrals over data. In
-practice, the optimal solution is intractable because of the enormous
-growth over stages (see e.g. \cite{sucar2015probabilistic} and
-\cite{powell2016perspectives}).  Instead, we outline heuristic
-strategies.
-
-%The derived closed-form solutions for the EIBV are still important building blocks when constructing the sampling designs, as they will be used to score the different adaptive designs. % Efficient
-% calculation is important for using adaptive survey designs with
-% robotic platforms.
-
-\subsection{A Naive Sampling Strategy}
-\label{naive}
-
-The simplest heuristic for adaptive sampling is to choose the next
-sampling location based on current probabilities. One reduces the
-largest levels of uncertainty by selecting the design $j$ with
-$P(Z(\x_j) \in T)$ closest to $0.5$.  With temperature and salinity
-variables, this {\it{naive}} strategy selects the stage $s$ sampling
-design according to:
-\begin{eqnarray}\label{critNaive}
-    d^{*,s} &=& \mbox{argmin}_{j} |P(Z_1(\x_j) \leq t_1, Z_2(\x_j) \leq t_2 | \mathcal{Y}^{s-1})-0.5|, \nonumber
-\end{eqnarray}
-where $j$ is indexing all neighboring locations on the waypoint graph.
-
-This strategy does not account for the uncertainty in the GP
-variables, but is instead applicable only if one design (node) has EPs
-closer to $0.5$ (see Table \ref{tab:sim_rhoab}, row
-two). Additionally, this strategy does not account for spatial
-correlation. This strategy lacks memory of where it has been and where
-the uncertainty has been reduced, and therefore susceptible to local
-minima.
-
-\subsection{Myopic Path Planning}
-\label{sec:myopic}
-
-The myopic (greedy) strategy which we present here is optimal if we
-imagine taking only one more stage of measurements. This selection
-strategy is based on expected uncertainty reduction, but it is a
-heuristic because there is no anticipation of what the subsequent
-designs might offer beyond the first stage.
-
-Based on the currently available data $\mathcal{Y}^{s-1}$, we fit an
-updated GP model. Focusing on the EIBV criterion, at the next stage
-$s$, the selected design is:
-\begin{eqnarray}\label{critSEQ}
-    d^{*,s} &=& \mbox{argmin}_{j} \left\{ \int_{\mathcal{M}} E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( p(\x;\mathcal{Y}^{j,s})\left( 1-p(\x;\mathcal{Y}^{j,s})\right) \right) d\mes(u), \right\} \\
-    p(\x;\mathcal{Y}^{j,s}) &=& P(Z_u \in T |\by_{d^{j,s}},\mathcal{Y}^{s-1}). \nonumber
-\end{eqnarray}
-
-Note that this strategy gives a sequential conditional version of
-expression (\ref{eq:eibv}). Now $\mathcal{Y}^{s-1}$ is available, and
-the expectation is conditional to this data. A similar closed-form
-calculation for EIBV is hence applicable in Eq.  \eqref{critSEQ} using the updated
-GP model from stage $s-1$ as defined via Eq.
-\eqref{eq:krigSUR} in the results from Proposition \ref{propo2}.
-Once the data are collected for the best design $d^{*s}$ at stage $s$, the GP model is updated again.
-The mean and covariance are used to compute the next design at stage $s+1$, and so on.
-
-Even though this myopic strategy is non-anticipative, it still gives a
-reasonable approach for creating designs in many
-applications. Moreover, it is easily implemented, in our case, onboard
-an AUV, using an efficient data update of the GP model and the
-calculation of the closed-form EIBV expressions for each subsequent
-trajectory.
-
-\subsection{Look-ahead Trajectory Planning}
-\label{sec:LA}
-
-We now extend the myopic strategy to a look-ahead strategy. We consider two stages of
-measurements, and the strategy is then optimal in the sense that it accounts consistently for the expectations and minimizations in these two stages, but still without including any planning beyond those two steps.
-In addition to the
-next stage of measurements $\by^{j,s}$, this look-ahead strategy anticipates the subsequent design indexed $j_2$ with data
-$\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$. The
-selected design is:
-\begin{eqnarray}\label{critLA}
-    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( \mbox{argmin}_{j_2}  V_{j_2} \right) \right\}, \\
-V_{j_2} & = & \int_{\mathcal{M}} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1}) ( 1-p(\x;\mathcal{Y}^{j,s+1}) ) \right) d\mes(u), \nonumber \\
-    p(\x;\mathcal{Y}^{j,s+1}) &=& P(Z(\x) \in T |\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}). \nonumber
-\end{eqnarray}
-Here, $\mathcal{Y}^{j,s}=\{\by_{d^{j,s}},\mathcal{Y}^{s-1}\}$ and
-$\mathcal{Y}^{j,s+1}=\{\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}\}$
-represent the sets of data variables, and the expectations are
-calculated with respect to the conditional densities given data at
-that stage.  We solve the first expectation by Monte
-Carlo sampling of data $\by_{d^{j,s}}$ from its conditional
-distribution. For each of these data samples, the second expectation
-is solved using the closed-form expressions for EIBV using the results from Proposition \ref{propo2}, now with conditioning to the first stage data already going into the mean and covariances in Eq.
-\eqref{eq:krigSUR}.
-
-Even though the strategy looks at two sampling stages, it is only used to find
-the current best design. When data are collected at stage $s$, the GP model is
-updated as in Eq. \eqref{eq:krigSUR}, and the mean and (cross) covariance are
-used to compute the next design for stage $s+1$, now anticipating what
-stage $s+1$ and $s+2$ could offer, and so on.
-This look-ahead approach is much more computationally demanding than
-the myopic strategy, so for practical implementation onboard the AUV this is reaching the limits allowed for calculation.
-We conduct some pruning of the least valuable design paths in the evaluation of Eq.
-\eqref{critLA}, to get a satisfactory computation time below one minute per waypoint.
-This means that we use the myopic
-strategy to rank the three best designs following the first stage alone, and
-for each of these preferred designs, we undertake the look-ahead calculations.
-
-\section{Simulation study}
-\label{sec:simulations}
-
-We now study the properties of the different static and sequential
-survey designs in a realistic case that emulates the spatial
-properties of a river plume interacting with coastal water.
-
-\subsection{Modeling}
-
-A GP model is used to represent the bivariate random fields of
-temperature and salinity. At location $\x \in \mathcal{M}$, the
-bivariate distribution is:
-\begin{equation}\label{gp_bar}
-\begin{bmatrix}Z_1(\x) \\
-Z_2(\x) \end{bmatrix}
-\sim N \left( 
-\begin{bmatrix} \mu_{1}(\x)\\
-\mu_{2}(\x)
-\end{bmatrix},\begin{bmatrix}
-\sigma_{1}^2 & \sigma_{1} \sigma_2 \gamma \\
-\sigma_{1} \sigma_2 \gamma  & \sigma_{2}^2 
-\end{bmatrix}
-\right).
-\end{equation}
-We specify the mean by $\mu_{1}(\x)=5.8 + 0.065 u_{\mbox{West}}$ and
-$\mu_{2}(\x)=29 + 0.1 u_{\mbox{West}}$, which means that the expected
-levels of both temperature and salinity increase linearly in the
-western direction.  This situation mimics that of a river mouth
-opening from the south, with the water masses pulled to the east by
-the Coriolis force \kc{This seems overly specific to the Nidelva use
-  case. Can it be generalized, after all that is the overall intent of
-  this manuscript?}.  The variances are set to $\sigma_1=0.25$ and
-$\sigma_2=0.6$, while the correlation is $0.6$.  The spatial
-correlation is of separable Matern type (see e.g. \cite{Cressie:11})
-so that
-\begin{equation}\label{v}
-\small
-\mbox{Corr} 
-\left(
-\begin{bmatrix}
-    Z_1(\x) \\
-    Z_2(\x) 
-    \end{bmatrix},
-    \begin{bmatrix}
-    Z_1(\x') \\
-    Z_2(\x') 
-    \end{bmatrix}
-    \right)
-    = \begin{bmatrix}
-1 & 0.6  \\
-0.6  & 1
-\end{bmatrix}(1+\phi h)\exp (-\phi h),
-\end{equation}
-where $h=\sqrt{||\x-\x'||^2}$ and $\phi=0.3$ indicates an effective
-correlation range of about $1200$ m.
-
-In the accompanying Python code, these modeling assumptions can be
-generalized to anisotropic covariance and changing variance levels
-across the spatial domain. One can hence, easily run the code to see
-if different models result in other sampling designs. Anisotropy and
-non-stationary variance are both relevant for the setting with river
-plumes, but in practice this requires more parameters to be
-specified. With extensive data and prior knowledge, one could also
-possibly fit and estimate parameters of more complex multivariate
-spatial covariance functions
-\citep{gneiting2010matern,genton2015cross}, but that is outside the
-scope of the current paper.
-
-We discretize the spatial domain $\mathcal{M}$ to a set of $n$ grid
-locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,n \}$, \kc{is the index
-  correct? TeX shows it as u or g for $\mathcal{M}$} where each cell
-has area $\Delta$; the same grid is used for the waypoint graph for
-possible design locations. The EIBV is approximated by sums over all
-grid cells.
-
-One realization of the random fields representing salinity and
-temperature is shown in Fig. \ref{fig:true_temp} and
-\ref{fig:true_sal}. The true ES is a realization from this model as
-shown in Fig. \ref{fig:ESet}.
-
-\begin{figure}[!h]
-  \centering
-  \subfigure[Simulated temperature.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/true_temp.pdf}\label{fig:true_temp}}
-  \hfill
-  \subfigure[Simulated salinity.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/true_sal.pdf}\label{fig:true_sal}}
-  \hfill
-  \subfigure[Excursion set given the realization of temperature and salinity.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_true.pdf}\label{fig:ESet}}
-  \hfill
-  \subfigure[Excursion probabilities given collected data using the
-  `static\_north" strategy.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
-  \caption{\ref{fig:true_temp} and \ref{fig:true_sal} show a
-    realization of the simulated temperature and salinity fields. The
-    associated ES is in \ref{fig:ESet}, while \ref{fig:Eprob} shows
-    the estimated EPs after performing data collection along the
-    static N-S survey line.}
-\label{fig:realisations}
-\end{figure}
-
-\subsection{Static and Sequential Sampling Designs}
-\label{sec:sampling_designs}
-
-%Three different designs are considered as indicated in Figure \ref{fig:stat_design}. 
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.65\textwidth]{Figures/Des3.eps}
-%\caption{Three different static survey designs plotted on the initial EP.}\label{fig:stat_design}
-%\end{figure}
-%In this display the designs are plotted along with the prior probability contours of the ES for the reference parameter inputs.
-
-We compare three different static designs denoted
-\textit{static\_north}, \textit{static\_east}, and
-\textit{static\_zigzag} with the three described sequential approaches
-\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
-sampling paths are pre-determined and cannot be altered and represent
-the pre-planned strategies used in most current AUV operational survey
-designs.
-
-For a fixed survey length, a closed-form expression for the EIBV is
-available as in Eq. \eqref{eq:eibv}. However, for the sequential
-approaches this is not the case. For comparison the properties are
-therefore evaluated using Monte Carlo sampling over several replicates
-of realizations from the model while conducting simulated sequential
-surveys for each one. Fig. \ref{fig:Eprob} shows the conditional EP,
-given data gathered along the north-south survey lines in these
-displays. In the Monte Carlo replicates, such results are repeatedly
-computed to approximate the EIBV. We also compare predictive
-performance measured by root mean square error (RMSE) for temperature
-and salinity estimates as well as also the variance reduction in these
-two variables. It is important to note that the objective function
-used by the agent \kc{First of 'agent'. Define it or change it
-  to 'robot'? Note further uses below.} is focused on reducing the
-EIBV, but we nevertheless expect that we will achieve good predictive
-performance for criteria such as RMSE as well. Another non-statistical
-criteria that is important for practical purposes is the computational
-time needed for the strategy, as this will impact the performance for
-an embedded system.
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
-%\caption{The equilateral waypoint graph used to discretize the
-%  trajectory choices over the $31\times31$ grid used to discretize the GP.}
-%\label{fig:wp_graph}
-%\end{figure}
-
-\begin{figure}[!h] 
-\centering 
-\subfigure[The waypoint graph.]{\includegraphics[width =
-0.49\textwidth]{Figures/sim/wp_graph_paper.pdf}\label{fig:wp_graph_a}}
-\hfill
-\subfigure[The waypoint graph in 3D.]{\includegraphics[width =
-0.49\textwidth]{Figures/sim/wp_graph_3d.pdf}\label{fig:wp_graph_b}}
-\caption{\ref{fig:wp_graph_a} The equilateral waypoint graph used to discretize the
-trajectory choices over the $31\times31$ grid used to discretize the GP.
-\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment.}
-\label{fig:wp_graph}
-\end{figure}
-
-Each strategy is conducted on an equilateral grid as shown in
-Fig. \ref{fig:wp_graph}. The sequential sampling agent starts at the
-center East-West coordinate at the southern end of the domain (node 53
-\kc{Maybe color this node differently and modify the caption?}). The
-AUV moves along edges in the waypoint graph while the vehicle makes
-measurements. The data is assimilated into the GP model before an
-evaluation of the next node to sample is conducted at the end of the
-edge.
-
-\subsection{Simulation Results}
-
-A total of 100 replicate simulations were conducted with all
-strategies. The results are shown in Fig. \ref{fig:sim_results}, where
-the different criteria are plotted as a function of survey
-distance. Fig. \ref{fig:avg_ev} shows the resulting drop in IBV for
-each of the six strategies. IBV reduction occurs most under the
-\textit{myopic} and \textit{look-ahead} strategies, each performing
-almost equally; this is expected as the two criteria
-(Eq. \eqref{critSEQ} and \eqref{critLA}) are sensitive to differences
-in IBV. The \textit{static\_north} design also does well here because
-the path is parallel to the boundary between the water masses.
-
-\begin{figure}[h!]
-  \centering
-  % \subfigure[Excursion set variance $E_{\by}(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
-  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
-  \hfill
-  \subfigure[RMSE between estimated field and truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
-  \hfill 
-  \subfigure[Explained variance $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
-  \hfill 
-  \subfigure[Computational time for inferencing.\kc{I see only 
-    lines associated with 4 variables showing in the graph. Where is
-    static\_north and static\_east?}]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}} 
-\caption{Simulation results from 100 replicate simulations for 10
-  sampling choices/stages on the grid. \kc{Vertical lines show large
-    variation in replicate results.}}  
-\label{fig:sim_results}
-\end{figure}
-
-Fig. \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop
-in RMSE and increase in explained variance, respectively. Both
-\textit{myopic} and \textit{look-ahead} strategies perform well here,
-but some of the \textit{static\_east} and \textit{static\_zigzag} also
-achieve good results because they are pre-determined to cover large
-parts of the domain without re-visitation. Sequential strategies
-targeting IBV will sometimes not reach similar coverage, as
-interesting data may draw the path into twists and turns. There is
-relatively large variety in the replicate results as indicated by the
-vertical lines. Nevertheless, the ordering of strategies is similar.
-
-\begin{figure}[!b]
-  \centering
-  \subfigure[Look-ahead]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_look-ahead.pdf}\label{fig:avg_look-ahead}}
-  \hfill
-  \subfigure[Myopic]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_myopic.pdf}\label{fig:avg_myopic}}
-  \hfill
-  \subfigure[Naive]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_naive.pdf}\label{fig:route_naive}}
-  \hfill
-  \subfigure[Static]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_static.pdf}\label{fig:route_static}}
-  \hfill
-  \caption{The overlaid route choices (strategies), superimposed on
-    the 100 replicate surveys with 10 sampling choices/stages. \kc{I
-      would reorder these figurs above based on the order the material
-      was discussed -- Static, naive, myopic and then Look-ahead. Also
-      do you really need a key associated with each subfigure, given
-      the caption? An abridged analysis as in the text, might also be
-      in order here in the caption. Also refer to
-      Fig. \ref{fig:realisations} as the reference, where the
-      temp/salinity fields are shown.}}
-\label{fig:route_choices}
-\end{figure}
-
-Fig. \ref{fig:avg_time} shows the computational effort: the
-\textit{naive} strategy is on par with the static designs, while the
-\textit{myopic} strategy is slower. The \textit{look-ahead} is even
-slower, reaching levels that are nearly impractical for execution on a
-vehicle. Some pruning of the graph is performed to improve the
-performance, such as ruling out repeated visitations and
-back-and-forth routes. Some of the intermediate results are also
-stored for longer planning horizons. Further pruning of branches or
-inclusion of other heuristics could be included for better
-performance. Then again, the inclusion of heuristics is likely a
-contributing factor for the \textit{look-ahead} strategy failing to
-outperform the \textit{myopic} strategy.
-
-In Fig. \ref{fig:route_choices}, the realized sampling paths for each
-of the sequential schemes and static designs are shown. The
-\textit{naive} strategy often gets stuck in the southern part of the
-domain because it is too focused on the probabilities near $0.5$. The
-\textit{myopic} strategy covers a wider domain than the naive or
-look-ahead. There are several reasons for this. First, a greedy
-approach will tend to put more emphasis on promising locations close
-to the agent, which may lead away from the centre. Second, as the
-agent evaluates the impact of locations further away (look-ahead)
-where assimilated data has less predictive power, the GP model (which
-is centered here) will act to restrict paths deviating from the
-central zone.
-
-We studied the sensitivity of the results by modifying the input
-parameters to have different correlations between temperature and
-salinity, standard deviations, and spatial correlation range.  In all
-runs, the \textit{myopic} and \textit{look-ahead} strategies perform
-the best in terms of realized IBV, and much better than
-\textit{naive}. The \textit{look-ahead} strategy seems to be
-substantially better than the \textit{myopic} design only for very
-small initial standard deviation or very large spatial correlation
-range. \textit{static\_north} continues to be the best static design
-for IBV, while \textit{static\_zigzag} is the best design for the
-other predictive performance measures, especially so with large
-spatial correlation range. We also ran simulation studies with only
-temperature data, and for realistic correlation levels between
-temperature and salinity, the IBV results are not much worse when only
-temperature data are available. In addition to the comparison made in
-Table \ref{tab:sim_rhoab}, the current setting includes spatial
-correlation and this likely reduces the additional influence of having
-bivariate data. However, it seems that having temperature data alone
-does a substantially worse job in terms of explained variance
-
-\begin{comment}
-\subsection{Performance Metrics}
-
-The tables below explore the sensitivity of the result to different input correlation values between temperature and salinity measured onboard the AUV, standard deviations, and spatial
-correlations (see table rows). We only compare results at the end of
-each survey. The expected ES variances are shown in Table
-\ref{tab:sim_res_ev}, results of RMSE are in Table
-\ref{tab:sim_res_rmse}, and explained variance is in Table
-\ref{tab:sim_res_r2}. In all runs, both the \textit{myopic} and
-\textit{look-ahead} strategies perform the best in terms of expected
-variance in the ES. The \textit{myopic} strategy
-performs slightly better than the \textit{look-ahead}. Part of the
-reason for this discrepancy can be traced to the act of pruning some
-of the paths for the \textit{look-ahead} strategy. The \textit{static\_north} design is clearly
-better than the other two static designs, the reason being that the
-others do not focus on the areas where the ES is really uncertain
-(centre locations are the most uncertain). By measuring temperature
-alone, the expected variances in the ES are only slightly higher than
-the combined temperature-salinity estimate. This occurs because the
-correlation of $0.6$ means that temperature observations provide
-information about salinity as well. For the other predictive measures,
-\textit{myopic} and \textit{look-ahead} strategies do well, clearly
-beating the \textit{naive} strategy.
-
-\begin{table}[!h]
-    \centering
-    \scalebox{0.87}{
-    \begin{tabular}{lrrrrrr}
-    \toprule
-    Parameter: $E_{\by}(p[1-p])$ &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
-    \midrule
-                    \rowcolor{Gray}
-ts. cor. low: 0.2  &  29.65 &   27.08 &       \textbf{26.44} &         32.33 &        40.31 &          36.23 \\
-ts. cor. high: 0.8 &  36.18 &   \textbf{28.42} &       30.30 &         32.99 &        37.18 &          36.04 \\
-                    \rowcolor{Gray}
-std. low: 0.1      &  18.26 &   15.79 &       \textbf{15.35} &         18.89 &        26.65 &          26.10 \\
-std. high: 0.5     &  47.01 &   \textbf{41.31} &       43.15 &         47.92 &        52.63 &          49.71 \\
-                    \rowcolor{Gray}
-cor. low: 0.8      &  44.42 &   \textbf{42.02} &       42.47 &         43.84 &        46.05 &          47.82 \\
-cor. high: 0.2     &  29.73 &   21.09 &       \textbf{20.43} &         27.91 &        37.60 &          34.35 \\
-                    \rowcolor{Gray}
-temp. only         &  38.25 &   29.16 &       \textbf{28.69} &         34.32 &        39.81 &          36.17 \\
-basecase           &  35.21 &   \textbf{28.56} &       28.94 &         33.50 &        39.84 &          39.26 \\
-    \bottomrule
-    \end{tabular}}
-  \caption{Simulation results for the final mean IBV (Eq. \eqref{two_partsK}).}
-    \label{tab:sim_res_ev}
-\end{table}
-\vspace{-0.4cm}
-\begin{table}[!h]
-    \centering
-    \scalebox{0.92}{
-        \begin{tabular}{lrrrrrr}
-        \toprule
-        Parameter: RMSE &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
-        \midrule
-                \rowcolor{Gray}
-ts. cor. low: 0.2  &   0.63 &    \textbf{0.57} &        0.59 &          0.62 &         0.57 &           0.59 \\
-ts. cor. high: 0.8 &   0.63 &    \textbf{0.59} &        0.61 &          0.64 &         0.60 &           0.60 \\
-                \rowcolor{Gray}
-std. low: 0.1      &   0.39 &    0.37 &        0.38 &          0.38 &         0.40 &           \textbf{0.35} \\
-std. high: 0.5     &   0.85 &    0.82 &        0.84 &          0.82 &         0.83 &           \textbf{0.81} \\  
-                \rowcolor{Gray}
-cor. low: 0.8      &   0.68 &    0.66 &        0.67 &          0.68 &         0.66 &           \textbf{0.65} \\
-cor. high: 0.2     &   0.58 &    0.54 &        0.55 &          0.58 &         0.54 &           \textbf{0.49} \\
-                \rowcolor{Gray}
-temp. only         &   0.65 &    0.63 &        0.62 &          0.65 &         \textbf{0.61} &           0.62 \\
-basecase           &  0.61 &   \textbf{0.56} &       \textbf{0.56} &         0.60 &        0.57 &          0.57 \\
-        \bottomrule
-        \end{tabular}}
-      \caption{Simulation results for the final mean RMSE ($\frac{1}{B} \sum_{b=1}^{B} \sqrt{(\bmu-\tilde{\bmu})^2}$).}
-    \label{tab:sim_res_rmse}
-\end{table}
-\vspace{-0.4cm}
-\begin{table}[!h]
-    \centering
-    \scalebox{0.95}{
-        \begin{tabular}{lrrrrrr}
-        \toprule
-        Parameter: $\bR^{2}$ &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
-        \midrule
-                        \rowcolor{Gray}
-ts. cor. low: 0.2  &  27.00 &   33.29 &       32.24 &         26.71 &        30.56 &          \textbf{33.50} \\
-ts. cor. high: 0.8 &  23.68 &   \textbf{33.87} &       32.42 &         26.72 &        30.73 &          33.61 \\
-                        \rowcolor{Gray}
-std. low: 0.1      &  25.38 &   30.80 &       30.29 &         26.09 &        28.75 &          \textbf{31.66} \\
-std. high: 0.5     &  26.11 &   34.22 &       32.48 &         26.95 &        31.77 &          \textbf{34.48} \\
-                        \rowcolor{Gray}
-cor. low: 0.8      &   8.80 &   11.50 &       11.23 &          9.56 &        12.31 &          \textbf{12.58} \\
-cor. high: 0.2     &  37.57 &   \textbf{48.68} &       48.11 &         39.61 &        41.98 &          48.04 \\
-                        \rowcolor{Gray}
-temp. only         &  11.67 &   \textbf{22.82} &       21.98 &         18.16 &        20.77 &          22.78 \\
-basecase           &  24.71 &   33.45 &       32.25 &         26.71 &        30.62 &          \textbf{33.54}\\
-        \bottomrule
-        \end{tabular}}
-      \caption{Simulation results for the final mean explained variance $\bR^{2}=100*(1-(\Sigma_{posterior}/\Sigma_{initial}))$).}
-    \label{tab:sim_res_r2}
-\end{table}
-
-\end{comment}
-
-\section{Case Study - Mapping a River Plume}
-\label{sec:case_study}
-
-To demonstrate the applicability of using multivariate EPs and the IBV
-to inform oceanographic sampling, we present a case study mapping a
-river plume with an AUV. The experiment was performed in Trondheim,
-Norway, surveying the Nidelva river (Fig. \ref{fig:nidelven}). The
-experiments were conducted in late Spring 2019, when there is still
-snow melting in the surrounding mountains so that the river water is
-substantially colder than the water in the fjord. The experiment was
-focused along the frontal zone that runs more or less parallel to the
-eastern shore as noted in Fig. \ref{fig:nidelven}.
-
-\subsection{Model Specification}
-\label{sec:exp_modeling}
-
-The statistical model parameters were specified based on a short
-preliminary survey where the AUV made an initial transect to determine
-the trends in environmental conditions and correlation
-structures. Based on the initial data, the trend parameters were
-estimated by linear regression, where both temperature and salinity
-are assumed to increase linearly, going west from the river
-mouth. Next, the residuals from the regression analysis were analyzed
-to study the fit of the GP model and to specify the covariance
-parameters.
-
-%\begin{figure}[!h] 
-% \centering 
-%\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
-%\caption{Data analysis from a preliminary trial experiment using the
-%  AUV. Left: Residual plot of temperature and salinity along with
-%  Gaussian contours. Middle: Empirical CDF (solid) of the quadratic form of
-%  the residuals along with the theoretical CDF (dashed) of the $\chi^2$
-%  distribution with two degrees of freedom. Right: Empirical variogram
-%  of the salinity and temperature data.} \label{fig:parest}
-%\end{figure}
-
-\begin{figure}[!h]
-  \centering
-  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_a.pdf}\label{fig:parest_a}}
-  \hfill
-  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_b.pdf}\label{fig:parest_b}}
-  \hfill
-  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_c.pdf}\label{fig:parest_c}}
-  \caption{Data analysis from a preliminary trial experiment using the
-    AUV. \ref{fig:parest_a} Residual plot of temperature and salinity
-    along with Gaussian contours. \ref{fig:parest_b} Empirical CDF
-    (solid) of the quadratic form of the residuals along with the
-    theoretical CDF (dashed) of the $\chi^2$ distribution with two
-    degrees of freedom. \ref{fig:parest_c} Empirical variogram of the
-    salinity and temperature data.}
-\label{fig:parest}
-\end{figure}
-
-Fig. \ref{fig:parest} summarizes diagnostic plots of this
-analysis. Fig. \ref{fig:parest_a} shows a cross-plot of temperature
-and salinity residuals after the westerly trends in both salinity and
-temperature are subtracted from the data. This scatter-plot of joint
-residuals indicates larger variability in salinity than in
-temperature, and a positive correlation ($0.5$) between the two
-variables. Based on the fitted bivariate Gaussian model (ellipses in
-Fig. \ref{fig:parest_a}), we can compute the modeled quadratic form of
-the residuals, and if the model is adequate they should be
-approximately $\chi^2_2$ distributed. Fig. \ref{fig:parest_b} shows
-the empirical cumulative distribution function (CDF) of these
-quadratic forms (solid) together with the theoretical CDF of the
-$\chi^2_2$ distribution. The modeled and theoretical curves are very
-similar, which indicates that the Gaussian model fits reasonably
-well. Even though there appears to be some clustering in both
-Fig. \ref{fig:parest_a} and \ref{fig:parest_b}, the bivariate
-diagnostic plots look reasonable and justify a Gaussian
-model. Fig. \ref{fig:parest_c} shows the empirical variogram of the
-scaled residuals for temperature and salinity. The decay is similar
-for the two, and seems to be negligible after about $150$ m.
-
-Based on the analysis in Fig. \ref{fig:parest}, the resulting
-parameters are given in Table \ref{tab:experiment_param}. The
-regression parameters shown here are scaled to represent the east and
-west boundaries of the domain as seen in the preliminary transect
-data, and the thresholds are intermediate values. These parameter
-values were then used in field trials where we explored the
-algorithm's ability to characterize the river plume front separating
-the river and fjord water masses, providing a spatial map of this
-boundary.
-
-%Mapping the spatial extent of a frontal zones is an important problem for studying many bio-physical interactions in the ocean.
-The frontal zone is determined by the boundary where plumes of sediments, nutrients, and possibly pollutants spreading from the river outlet meet and interact with adjacent coastal water.
-Due to the lower density the plumes spread on the surface, creating a front with an sharp gradient in both temperature and salinity.
-
-\begin{table}[!h]
-\centering
-\begin{tabular}{lrr}
-\toprule
-Parameter & Value & Source\\
-\midrule
-\rowcolor{Gray}
-Cross correlation temp. and sal. & 0.5 & AUV observations\\
-Temp. variance &  0.20 & AUV observations (variogram)\\
-\rowcolor{Gray}
-Sal. variance &  5.76 & AUV observations (variogram)\\
-Corr. range  & 0.15 km & AUV observations (variogram)\\
-\rowcolor{Gray}
-River temp. $T_{river}$ & $10.0\,^{\circ}\mathrm{C}$ & AUV observations\\
-Ocean temp. $T_{ocean}$ & $11.0\,^{\circ}\mathrm{C}$ & AUV observations\\
-\rowcolor{Gray}
-River sal. $S_{river}$ & $14.0$ g/kg & AUV observations\\
-Ocean sal. $S_{ocean}$ & $22.0$ g/kg & AUV observations\\
-\rowcolor{Gray}
-Threshold temp. & $10.5\,^{\circ}\mathrm{C}$ & $(T_{ocean}-T_{river})/2+T_{river}$\\
-Threshold sal. & $18.0$ g/kg & $(S_{ocean}-S_{river})/2+S_{river}$\\
-\rowcolor{Gray}
-\bottomrule
-\end{tabular}
-\caption{Model and threshold parameters from an initial AUV
-  survey. Observations were taken across the front while crossing from
-  fresh, cold river water to saline and warmer ocean waters. \kc{Might
-  be good to expand the terms and not have abbreviations in the table
-  above; there's plenty of space.}}
-\label{tab:experiment_param}
-\end{table}
-
-\subsection{Experimental Setup}
-
-The sampling locations were distributed over an equilateral grid, as
-shown in the grey-colored lattice in Fig. \ref{fig:map}. The robotic
-platform consisted of a Light AUV \citep{sousa2012lauv}
-(Fig. \ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49
-conductivity, temperature, and depth (CTD) sensor providing
-temperature and salinity
-measurements. %The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
-The sampling agent was built on top of the autonomous agent framework
-Teleo-Reactive EXecutive (\textit{T-REX})
-\citep{py10,Rajan12,Rajan12b}, running an instance of the
-\textit{myopic} strategy from Section \ref{sec:myopic} to control the
-AUV and decide between sampling locations.
-
-\begin{figure}[!h] 
-\centering 
-\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
-\caption{The commercially available Light Autonomous Underwater
-  Vehicle (LAUV) platform for upper water-column exploration used in
-  our experiments.}
-\label{fig:lauv}
-\end{figure}
-
-The sampling strategy was designed around the concept of visiting
-waypoints sequentially. Arriving at a desired waypoint with new
-measurements and an updated model, the AUV triggers the myopic
-strategy to evaluate the different design criteria (see
-Eq. \eqref{critSEQ}). The waypoint-and-path combination that is
-expected to reduce the IBV the most is selected, and upon arrival this
-procedure is then repeated. At each stage, it takes the AUV about 30
-seconds to evaluate the EIBV for all the possible waypoint-and-path
-alternatives.
-
-The AUV was set to start in the south-center part of the waypoint
-graph, with the previously outlined GP model of the environment
-(Section \ref{sec:exp_modeling}). A survey was set to take
-approximately 40 minutes, visiting 15 waypoints on the grid, with the
-AUV running near the surface to capture the plume. On its path from
-one waypoint to the next, the AUV gathered data regularly, and the GP
-model assimilated temperature and salinity data with an update
-frequency of 30 seconds, giving about three updates per stage.
-
-\begin{figure*}[!h]
-\centering
-\subfigure[AUV survey area]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
-\hspace{0.3cm}
-\subfigure[Temperature tracks]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/auv.pdf}\label{fig:res_both}}
-
-\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv1_es_sal_ep.pdf}\label{fig:res1}}
-\hspace{0.2cm}
-\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv4_es_sal_ep.pdf}\label{fig:res2}}
-
-%\subfigure[ES for Survey 1]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
-%\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
-\caption{Results from mapping the Nidelva river, Trondheim, Norway
-  over two survey missions. \ref{fig:map} shows an overview of the
-  survey area overlaid with the AUV path in black and dashed
-  line. Note the shaded region indicating a typical frontal
-  region. \ref{fig:res_both} shows the collected temperature data as
-  colored trails. Note waypoint 5 (WP5) which indicates where the two
-  surveys diverge. \ref{fig:res1} and \ref{fig:res2} shows the
-  collected salinity data overlaid on the final EP, which indicate the
-  AUVs statistical impression of the front. For both missions the
-  temperature and salinity data correspond with an indication of the
-  EP front. \kc{Might want to indicate when the surveys (dates) were
-    conducted so people can understand the variability in the EP with
-    respect to time.}}
-\label{fig:results}
-\end{figure*}
-
-\subsection{Results}
-
-Two surveys missions (1 and 2), were run successively from 11:00 AM to
-01:00 PM, with a short break in between. The resulting path of the
-selected waypoints are shown in the map in Fig. \ref{fig:map}, both
-within the expected frontal region (shaded pink). The recorded
-temperatures are shown as colored trails in Fig. \ref{fig:res_both},
-clearly indicating the temperature difference between fjord and
-riverine waters. The salinity data are then shown separately, overlaid
-with the estimated EP for each survey in Fig. \ref{fig:res1} and
-\ref{fig:res2}.
-
-Both surveys successfully estimated and navigated the separation zone,
-crossing the frontal boundary multiple times. As conditions changed
-slightly between the two surveys, the resulting path (after waypoint
-5) is shown to deviate. Survey 1 continued northwards, tracking the
-north-eastern portion of the front, while Survey 2 turned west,
-mapping the south-western region.
-
-The final predictions of the front location, represented by
-conditional EPs in Fig. \ref{fig:res1} and \ref{fig:res2} as dashed
-lines, correspond with one another. In both surveys they yield a
-picture of the front being to the west in the southern portions of the
-region and gradually bending off toward the north east. The amount of
-exploration done by Survey 1 is greater than Survey 2. In Survey 1,
-the AUV obtained more detail by going north from waypoint 5, while
-Survey 2, coming close to the survey area borders in the south-western
-corner, obtained a poorer understanding of the northern parts. A
-look-ahead strategy might identify and discourage such choices.
-
-%\begin{figure}[!h] 
-% \centering 
-%\includegraphics[width=0.48\textwidth]{Figures/envir_ocean.pdf}
-%\caption{Ocean observation is moving away from %single-ship sampling
-%towards more collaborative networked operations in %order to resolve
-%the numerous processes and their interaction.} %\label{fig:envir}
-%\end{figure}
-%\newpage
-\section{Closing remarks}\label{sec:concl_disc}
-
-This work builds on a multidisciplinary effort combining statistical
-modeling with robotic surveying techniques for oceanographic
-applications. We show how observation practices can gain efficiency
-and accuracy from the development of statistical techniques to achieve
-more effective spatial monitoring. We further demonstrate the
-opportunities available for real-time multivariable spatial
-data-gathering and analysis onboard autonomous platforms, which
-statisticians can exploit to create general-purpose toolkits for
-similar applications.
-
-In particular, we derive and show results for characterizing phenomena
-connected to the properties of water masses. The characterization of
-uncertainties in random sets is extended with new results for the EIBV
-uncertainty reduction achieved by spatial sampling design. This is
-provided in closed-form for the situation with a static design, and
-then extended to the adaptive situation. The sequential derivations
-provide new insights into efficient applications of adaptive sampling,
-as demonstrated in our application.
-
-The study did not consider any temporal effects, which would be
-relevant on a larger time scale. We consider the extension to
-spatio-temporal modeling as future work, and envision that
-advection-diffusion equations could be useful in this kind of modeling
-\citep{sigrist2015stochastic}. For more complex oceanographic
-phenomena, the methods will need to be extended to non-Gaussian
-phenomena, possibly feature-based mixtures of GPs which could still be
-run onboard and augmented by dynamical models. \kc{How about talking
-  about removing any grids or waypoints altogether?}
-
-The spatial-statistical design criterion, building on random sets is
-relevant in our setting with different water properties, and we show
-mathematical generality beyond the EIBV which is convenient to
-implement for onboard robotic planning. Such criteria could also be
-useful in other oceanographic settings as for algal-blooms, anoxic
-zones or open water fronts \cite{costa19}. Of course, other criteria
-are also relevant for instance, hybrid or multi-attribute criteria
-that could balance goals of exploration and exploitation in this
-situation. Equally, such techniques have significant use cases in
-downstream decision-making, with policy makers and regulators who need
-to make difficult decisions related to fish farming, other marine
-resource operations. Value of information analysis (VOI)
-\citep{Eidsvik:15} could be used in a similar vein as the IBV in the
-current work, as both analytic methods explore when, where and which
-information is likely to result in improved decision-making.
-
-While more effort could be spent on approximating the look-ahead
-stages in the adaptive designs, it is perhaps more interesting to
-explore the additional flexibility that can be gained by having
-multiple AUVs co-temporally exploring a spatial or spatio-temporal
-domain \citep{ferreira2019advancing}. Such an approach would enable
-concurrent sampling in different parts of the space, or opportunities
-to move in parallel to best capture the excursion set.
-
-\section*{Acknowledgements}
-
-TOF acknowledges support from the Centre for Autonomous Marine
-Operations and Systems
-(AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of
-Excellence, project number 223254, and the Applied Underwater Robotics
-Labortatory (AURLab). JE and KR acknowledge support from Norwegian
-research council (RCN), project number 305445. CT and DG acknowledge
-support from the Swiss National Science Foundation, project number
-178858.
-
-%\begin{supplement}
-%\sname{Supplement A}\label{suppA}
-%\stitle{Title of the Supplement A}
-%\slink[url]{http://www.e-publications.org/ims/support/dowload/imsart-ims.zip}
-%\sdescription{Dum esset rex in
-%accubitu suo, nardus mea dedit odorem suavitatis. Quoniam confortavit
-%seras portarum tuarum, benedixit filiis tuis in te. Qui posuit fines tuos}
-%\end{supplement}
-
-% == Adding references
-\footnotesize
-\bibliographystyle{imsart-nameyear}
-\bibliography{ref}
-
-\section*{Appendix}
-
-Proof of Proposition \ref{propo1}:
-
-\begin{proof}
-	That $\mes(\es)$ defines indeed a random variable follows from Fubini's theorem 
-	relying on the joint measurability of 
-	$(\x, \omega) \to \mathbbm{1}_{\es(\omega)}(\x)$, 
-	%= \mathbf{1}_{\xi_{\x}(\omega) \in T}$ and Fubini's theorem.  
-	itself inherited from the assumed measurability for  
-	$(\x, \omega) \to \gp[\x](\omega)$ and $T$, respectively. From there, following the steps of Robbins' theorem \cite{Robins1944}, we find that 
-	\begin{equation*}
-	\begin{split}
-	\mathbb{E}[\mes(\es)^{r}]
-	&=\mathbb{E}\left[\left(\int_{D} \mathbbm{1}_{\gp[u] \in T} ~d\mes(u) \right)^{r} \right]
-	=\mathbb{E}\left[ \prod_{i=1}^{r} \left(
-	\int_{D} \mathbbm{1}_{\gp[\uu^{(i)}] \in T} ~d\mes(\uu^{(i)}) 
-	\right) \right] \\
-	&
-	=
-	%\mathbb{E}\left[
-	%\int_{D^{r}}
-	%\prod_{i=1}^{r}
-	%\mathbf{1}_{\gp[\uu^{(i)}] \in T}
-	%\productMeasure
-	%\right]
-	%=
-	\mathbb{E}\left[
-	\int_{D^{r}}
-	\mathbbm{1}_{\gp[\uu^{(1)}] \in T,\dots, \gp[\uu^{(r)}]  \in T}
-	\productMeasure
-	\right]
-	%\\
-	%&=\int_{D^{r}} \mathbb{E}\left[\mathbf{1}_{\gp[\bm{u}] \in T^r} \right] \productMeasure \\
-	%&
-	=\int_{D^{r}}
-	\jointExcuProb 
-	\productMeasure.
-	\end{split}
-	\end{equation*}   
-	The rest consists in expliciting the probability of $T\times \dots \times T$ under the multivariate Gaussian distribution of  
-	$\left(\gp[u^{(1)}], \dots,  \gp[u^{(r)}] \right)$. 
-\end{proof}
-
-Proof of Proposition \ref{propo2}:
-
-\begin{proof}
-	By definition of $\Phi_{p}$, for $N\sim \mathcal{N}_{p}(0_{p},\covN)$, 
-	$$
-	\mathbb{P}(N\leq a + BV | V)
-	=
-	\varPhi_{p}\left( a + BV; \covN \right).
-	$$
-	Now for $\varPhi_{p}\left( a + BV; \covN \right)^h$, provided that the probability space is sufficiently large to accomodate $h$ independent Gaussian random vectors $N_i\sim \mathcal{N}_{p}(0,\covN)$ (which is silently assumed here), using the former equality delivers
-	$$
-	\varPhi_{p}\left( a + BV; \covN \right)^h
-	=
-	\prod_{i=1}^h \mathbb{P}(N_i\leq a + BV | V).
-	$$
-	Now by independence of the $N_i$'s we obtain the joint conditional probability 
-	$$
-	\prod_{i=1}^h \mathbb{P}(N_i\leq a + BV | V)
-	=
-	\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV| V),
-	$$
-	whereof, by virtue of the law of total expectation, 
-	\begin{equation*}
-	\begin{split}
-	\mathbb{E}\left[ \varPhi_{p}\left( a + BV; \covN \right)^h \right]
-	&=\mathbb{E}\left[\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV| V)\right]\\
-	&=\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV)\\
-	&=\mathbb{P}(W_1 \leq a, \dots, W_h\leq a)\\
-	&=\varPhi_{ph}
-	\left(
-	1_{h} \otimes a
-	;
-	(1_{h}1_{h}')\otimes (B\Sigma_{V} B') + 
-	I_{h}\otimes \covN
-	\right),
-	\end{split}
-	\end{equation*}
-	where $\mathbf{W}=(W_1,\dots,W_h)$ with $W_i=N_i- BV \ (1\leq i \leq h)$
-	and the last line follows $\mathbf{W}$ forming a Gaussian vector (by global independence of the $N_i$'s and $V$) and from the definition of $\varPhi_{p h}$.
-The covariance matrix $\mathbf{\Sigma}$ of $\mathbf{W}$ is obtained by noting that $\operatorname{cov}(W_i,W_j)=B \covV B' + \delta_{ij} \covN \
-	(i,j \in \{1,\dots,h\})$.
-	%the latter concatenated Gaussian vector is obtained by straightfoward calculation. 
-\end{proof}
-
-Proof of Proposition \ref{propo3}.
-
-\begin{remark}
-	%    Note that the vector $\bm{a}$ is obtained by stacking $h_i$ times each $a_i$
-	%\begin{equation*}
-	%\bm{a} = \begin{pmatrix}
-	%    \bm{a}_1\\ \vdots\\ \bm{a}_g\end{pmatrix},~
-	%\bm{a}_i =
-	%\begin{pmatrix}
-	%    a_i\\ \vdots\\ a_i\end{pmatrix}
-	%\in\mathbb{R}^{l\times h_i}\end{equation*}
-	%and the covariance matrix is obtained by similar stacking
-	Using blockwise representation for the blocks themselves delivers %$\bm{\Sigma}$ can be expressed as follows
-	\begin{equation*}
-	%    \bm{\Sigma} = \begin{pmatrix}
-	%        \Sigma_{11} & & \Sigma_{1g}\\
-	%        & \ddots &\\
-	%        \Sigma_{g1}&   & \Sigma_{gg}
-	%\end{pmatrix},~
-	\Sigma_{ij} =
-	\begin{pmatrix}
-	B_i\Sigma_{V} B_j' & \dots & B_i\Sigma_{V} B_j'\\
-	\vdots & & \vdots\\
-	B_i\Sigma_{V} B_j' & \dots & B_i\Sigma_{V} B_j'\\
-	\end{pmatrix}
-	+
-	\delta_{ij}
-	\begin{pmatrix}
-	\covN_i & &\\
-	& \ddots &\\
-	&   & \covN_i
-	\end{pmatrix}%\in\mathbb{R}^{lh_i\times lh_j}
-	\end{equation*}
-	Here each $\Sigma_{ij}$ is made of $h_i$ times $h_j$
-	(vertically/horizontally) $p \times p$ sub-blocks, hence possesses $ph_i$ lines and $ph_j$ columns.
-\end{remark}
-
-%Proof TODO along with notation polish.
-
-\begin{proof}
-	The proof relies (again) heavily on the fact that, by definition of $\Phi_{p}$, for any covariance matrix $\covN \in \R^{p \times p}$, $a\in \R^p$, $B\in \R^{p \times q}$, and $N\sim \mathcal{N}_{p}(0_{p},\covN)$,
-	$$
-	\mathbb{P}(N\leq a + BV | V)
-	=
-	\varPhi_{p}\left( a + BV; \covN \right).
-	$$
-	In particular, for globally independent $N_{i,j} \sim \mathcal{N}_{p}(0_{p},\covN_i)$ $(1\leq j \leq h_i, 1\leq i \leq g)$, 
-	\begin{equation*}
-	\begin{split}
-	\prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i}
-	&=
-	\prod_{i=1}^{g}
-	\prod_{j=1}^{h_{i}}
-	\mathbb{P}(N_{i,j}\leq a_i + B_i V | V)\\
-	&=\mathbb{P}(N_{1,1}\leq a_1 + B_1 V, \dots, N_{g,h_{g}}\leq a_g + B_g V | V),
-	\end{split} 
-	\end{equation*}
-	so that, by the law of total expectation, 
-	\begin{equation*}
-	\begin{split}
-	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
-	=
-	\mathbb{P}(W_{1} \leq 1_{h_{1}} \otimes a_1, \dots, W_{g} \leq 1_{h_{g}} \otimes a_g)
-	\end{split}
-	\end{equation*}
-	where 
-	$W_{1}=(N_{1,1}- B_1 V, \dots, N_{1,h_{1}}- B_1 V), 
-	W_{2}=(N_{2,1}- B_2 V, \dots, N_{2,h_{2}}- B_2 V), 
-	\dots, W_{g}=(N_{g,1}- B_g V, \dots, N_{g,h_{g}}- B_g V)$. Noting that $\mathbf{W}=(W_1,\dots, W_g)$ is a centred $p H$-dimensional Gaussian random vector, we finally obtain that 
-	\begin{equation*}
-	\begin{split}
-	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
-	=
-	\varPhi_{p H}\left(\bm{a};\mathbf{\Sigma}\right),
-	\end{split}
-	\end{equation*}
-	with $\bm{a}=(1_{h_{1}} \otimes a_1, \dots, 1_{h_{g}} \otimes a_g)$ and $\bm{\Sigma}=(\operatorname{cov}(W_i,W_j))_{i,j \in \{1,\dots, g\}}$.
-\end{proof}
-
-% AOS,AOAS: If there are supplements please fill:
-%\begin{supplement}[id=suppA]
-%  \sname{Supplement A}
-%  \stitle{Title}
-%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
-%  \sdatatype{.pdf}" 
-%  \sdescription{Some text}
-%\end{supplement}
-
-% === Not used
-% After reviews: - This is moved to Section 2.1.
-
-%Traditional data collection at sea has typically been based on static
-%buoys, Lagrangian floats, or ship-based methods, with significant
-%logistical limitations that directly impact coverage and sampling
-%resolution. Modern methods using satellite remote-sensing provide
-%large-scale coverage but have limited resolution, are limited to
-%sensing the surface of the ocean, and are impacted by cloud cover. The
-%advent of robust mobile robotic platforms \citep{Bellingham07} has
-%resulted in significant contributions to environmental monitoring and
-%sampling. In particular, autonomous underwater vehicles (AUVs), have
-%advanced the state of sampling and consequently have made robotics an
-%integral part of ocean observation; our previous work has contributed
-%to this effort \citep{das11b,Das2015,fossum18b,fossuminformation}. Other ¤ %statistical work in the oceanographic domain include \cite{wikle2013modern}
-%focusing on hierarchical statistical models; \cite{sahu2008space},
-%studying spatio-temporal models for sea surface temperature and
-%salinity data; and \cite{mellucci2018oceanic} looking at the
-%statistical prediction of features using an underwater glider.
-
-%\begin{itemize}
-%
-%\item Full numerical ocean models cannot provide accurate results
-%  online if run on robotic sensing platforms, as onboard computers
-%  cannot deliver the computational power required. Hence statistical
-%  proxy models of the environment must be used for learning where to sample.
-%
-%\item With limited available information about the state of the ocean, there is substantial value in reacting to
-%  information obtained from measurements taken in-situ. This
-%  acquired information must be assimilated into statistical models that
-%  can be used to inform decisions on where to sample
-%  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.
-%
-%\item For sampling problems related to environmental sensing, the
-%  number of choices (i.e. locations, trajectories, and candidate
-%  designs) is enormous, creating a
-%  trade-off between optimization (finding the most resource-efficient
-%  design to collect necessary data) and computability (arriving at a
-%  solution in reasonable time). To successfully resolve features, this
-%  trade-off has to be considered in development and practice.
-%
-%\end{itemize}
-
-%Addressing this, the combination of statistical tools and robotic platforms is a
-%natural symbiosis which enables information-based sensing. Central to
-%this is the ability to model spatially-correlated variables and
-%provide formal measures of uncertainty. Our formulation is based on
-%Gaussian Processes (GPs) as they allow efficient implementation 
-%and evaluation in real time onboard a robotic platform.
-
-%Sampling can, in this context, not simply be distributed evenly ---along simple transects or ``lawn-mover" patterns--- but must instead be prioritized to relevant regions to ensure it is cost-effective while providing adequate coverage and resolution of the area of scientific interest.
-
-%While the focus has often been on
-%biological and anthropogenic impact from micro-plastics to
-%pollution, biological oceanographers have focused intently on studying
-%micro-organisms at the base of the human food web. These organisms are
-%critically impacted by the changing dynamics in the upper water-column,
-%especially in coastal zones which are complex and often hard to observe
-%in space and time. By studying the bio-geochemical processes in the
-%upper water-column scientists can measure the impact of change, natural
-%or anthropomorphic, and provide an informed opinion to policy makers to
-%effect changes in preserving the environment. However, the challenge of 
-% The pressure on marine resources is growing and increased accuracy,
-% resolution, and persistent monitoring of the oceans is crucial for
-% long-term sustainable management.
-
-% A
-% sustained focus on prioritized and efficient data collection strategies
-% have therefore started to emerge. The advent of marine robotic
-% platforms, especially
-
-% provided means to execute this prioritization through the capacity of
-% autonomy and data-driven sampling, where data collection in principle
-% can be optimized. These capabilities have made
-
-% of the emerging sensing practice for ocean science, allowing scientists
-% to increase the observational efficiency and resolution beyond what was
-% previously possible. But how should a robotic platform, such as an AUV,
-% effectively prioritize and identify important regions for sampling? The
-% answer to this question relates to
-
-% , and
-% the application domain is clearly an arena where statisticians can
-% contribute.
-% There has recently been some statistical attention in oceanography:
-
-%From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest.
-Each phenomenon can in turn be characterized by a set of process specific conditions expressed through different measures of key environmental variables, such as temperature or salinity.
-One such measure is the gradient, that can be associated with a number of important processes, such as the vertical location of the thermocline and pycnocline, location of upwelling systems, vertical mixing, eddies, fronts, and currents \cite{sverdrup2006}, as well as distribution, growth, and accumulation of biological activity \cite{SatOceanSoci00, Ryan2014}.
-These gradients create boundaries separating the ocean into process specific regions which are of profound interest to both identify and map effectively.
-Quantification of gradient features is therefore a much needed competence in robotic sampling.
-\end{document}

### 2020-06-12 ans_to_reviewers/answers-to-reviewers.tex
Updates from Overleaf

@@
 \newcommand{\bZ}{ {\boldsymbol Z} }
-\newcommand{\vc}[1]{\mbox{\boldmath $#1$}}
+\newcommand{\vc}[1]{\mbox{\boldmath $1$}}
 \newcommand{\balph}{ {\boldsymbol \alpha} }
@@
 
-\newcommand{\edcomment}[1]{{\color{green}{\{Editor: #1\}}}}
-\newcommand{\frevcomment}[1]{{\color{blue}{\{Rev 1: #1\}}}}
-\newcommand{\srevcomment}[1]{{\color{red}{\{Rev 2: #1\}}}}
-\newcommand{\trevcomment}[1]{{\color{violet}{\{Rev 3: #1\}}}}
+\newcommand{\edcomment}[1]{{\color{green}{\{Editor: 1\}}}}
+\newcommand{\frevcomment}[1]{{\color{blue}{\{Rev 1: 1\}}}}
+\newcommand{\srevcomment}[1]{{\color{red}{\{Rev 2: 1\}}}}
+\newcommand{\trevcomment}[1]{{\color{violet}{\{Rev 3: 1\}}}}
 
@@
 
-In the revised manuscript these aspects are largely rewritten in the Introduction, and partly also in Section 2 on the expected integrated Bernoulli variance (EIBV) and in the Case Study ({\bf{TO DO 6}}).
-This change addresses some comments of AE, Rev #1 in particular.
+In the revised manuscript these aspects are largely rewritten in the Introduction, and partly also in Section 2 on the expected integrated Bernoulli variance (EIBV) and partly in the Case Study ({\bf{TO DO 6 AGAIN}}).
+This change addresses some key comments of AE, Rev 1 in particular.
 
@@
 
-In this sense, the revised version should provide ample motivation for statisticians.
+In this sense, the revised version should provide ample motivation for statisticians, while still maintaining strong links to the application for illustrating concepts.
 
-In the revised manuscript these aspects are largely rewritten in the Introcuction and in Sect 2 and 3 on the statistical theory.
-Here, the main body contains the novel results for the expected uncertainty reduction in random sets exemplified by the EIBV, while the Appendix holds the proofs of key results.
-This change addresses the some comments of Editor, Associate Editor, Rec #2 and Rec #3.
+In the revised manuscript these aspects are largely rewritten in the Introduction, in Sect 2 and 3, and on future work in the Closing remarks.
+The main body contains the novel statistical results for the expected uncertainty reduction in random sets exemplified by the EIBV, while the Appendix holds the proofs of key results.
+This change addresses the some comments of Editor, Associate Editor, Rev 1, Rev 2 and Rev 3.
 
-\item We realized that a contribution of the current work is available code for practioners or statisticians who could try this themselves, or possibly extend it for their own cases.
-We have hence made a substantial effort in structuring {\bf{computer code for examples}} which are presented on a URL link to go with the paper.
-This open-source python code is also ready to run on more nuanced statistical models (non-stationary and non-isotropic).
+\item We realized that a contribution of the current work is available code for practitioners or statisticians who could try this themselves, or possibly extend it for their own cases.
+We have hence made a substantial effort in structuring {\bf{computer code for examples}} which are presented on a URL link.
+(\url{https://cedrictravelletti.github.io/Submarine/} )
+ to go with the paper. This open-source python code is also ready to run on more nuanced statistical models (non-stationary and non-isotropic).
 
-The simulation study section contains pointers to the associated open-source code. This change addresses the comments of Rev #1 in particular.
+The simulation study section contains pointers to the associated open-source code. This change addresses the comments of Rev 1 and Rev 2 in particular.
 
@@
 \reply: The revised version has an entirely new Section 2 which includes general background for random sets and excursion sets.
-We added references, for instance \cite{Adler.Taylor2007} on excursion sets and \cite{Molchanov2005} on random sets.
+We added references, for instance \cite{adler2009random} on excursion sets and \cite{Molchanov2005} on random sets.
 
@@
 
-\reply: The assumptions are conditionally independent measurements of salinity and temperature.
-We believe this is realistic as they are gathered from different sensors.
+\reply: The assumption is that salinity and temperature are conditionally independent measurements.
+(In the latent Gaussian random field, the variables are dependent.) We believe this conditional independence assumption is realistic as the measurements are gathered from different sensors.
 We added a comment about this not necessarily being the case with other data that are extracted and processed by combining sensors.
-In the latent Gaussian random field, the variables are dependent.
+Then assumptions can be tested in a lab under idealized conditions.
+But it is difficult to do so for the AUV in the changing oceans.
 
@@
 
-\reply: We have changed the statement on page 9 to: ``$\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations."
+\reply: We have changed the statement on page 9 to: ``$\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_1$ and $r^2_2$, holding the measurement error variances of temperature and salinity observations."
 
@@
 
-\reply: we have made this change to the citations \cite{genton2015cross} and \cite{french2016credible} - page 30.
+\reply: We have made this change to the citations: \cite{genton2015cross} and \cite{french2016credible} - page 30.
 \end{itemize}
@@
 
-\reply: TO DO
+\reply: The motivation for an applied statistical audience is strengthened, while we have maintained strong links to the applications of Robotics, Ocean Science and AI.
+In particular in the Introduction and in Section 2 are now stronger on motivating multivariate ESs, and highlighting the need for new statistics in these marine robotics applications.
+See main comments early on in this reply letter.
 
@@
 The code is optimized for speed to allow fast computation using Pytorch (citation).
-The code is available here: (TO DO link).
+The code is available here: (\url{https://cedrictravelletti.github.io/Submarine/} )
 
@@
 \begin{quote}
-The AUVs, see Fig.
+"The AUVs, see Fig.
 9, considered here operate at $1$-$3$ m/s in the upper water column, where they are free to move in six degrees of freedom (6 DOF).
@@
 AUVs have limited computational capacity, as they typically use single-board computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM) to do computation.
+"
 \end{quote}
@@
 
-\reply: Code is improved in the revision and provided for open-source usage: LINK
+\reply: Code is improved in the revision and provided for open-source usage (\url{https://cedrictravelletti.github.io/Submarine/} )
 
@@
 
-\reply: In the revised version we show EPs for a static survey line, myopic and look-ahead selection in three displays.
+\reply: In the revised version we show EPs for a static survey line, myopic and look-ahead selection in three displays. (This is in the new Figure 3.)
 
-In the accompanying code there are also resources for displaying other outputs if the reader is interested in comparing results with for instance univariate excursion probabilities.
+In the accompanying code (\url{https://cedrictravelletti.github.io/Submarine/} )
+there are also resources for displaying other outputs if the reader is interested in comparing results with for instance univariate excursion probabilities.
 

### 2020-06-12 UQonSets.tex
Updates from Overleaf

@@
-Let us in particular consider a locally finite (Borel) measure $\mes$ on $D$ and investigate the probability distribution 
+We now introduce quantities that allow to quantify the uncertainty on the volume of the excursion set $\es$. Let $\mes$ be a 
+(locally finite, Borel) measure  on $\domain$. We want to investigate the probability distribution 
 of $\mes(\es)$ through its moments.
@@
 
-%That $\mu(\es)$ is a well-defined random variable follows from
-\begin{propo}
-    \label{propo1}
-%Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
-For a measurable random field $\gp$ and a locally finite measure $\mes$ on $D$, $\mes(\es)$ is a random variable and for 
-any $r\geq 1$,
-\begin{equation*}
-\begin{split}
-\mathbb{E}[\mes(\es)^r]
-&=\int_{D^{r}} \jointExcuProb
-\productMeasure
-,
-\end{split}
-\end{equation*}
-
-where the product measure is denoted as
-$\nu^{\otimes}:=\bigotimes_{i=1}^r \nu$.
-%Note that we will use boldface to denote concatenated variables. % TO RELOCATE
-Here $\gp$ is defined on $D$, and for
-$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in D^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
-\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.
-\medskip
-
-In the particular case where $\gp$ is a multivariate Gaussian random field
-%and denoting the mean and covariance matrix of $\gp[\bm{u}]$ by $\meanUU\in\mathbb{R}^{p\times r}$ and $\covUU$, respectively,
-we have
-\begin{align*}
-\jointExcuProb = \mathcal{N}_{\no r}(T^r; \meanUU, \covUU),
-\end{align*}
-where $\mathcal{N}_{\no r}(\cdot ; \meanUU, \covUU)$ is the Gaussian measure on $\mathbb{R}^{\no r}$ with mean $\meanUU$ 
-and covariance matrix $\covUU$, respectively defined blockwise by
-%where both are defined blockwise by
-\begin{align*}
-\meanUU&=\begin{pmatrix}\mu(u^{(1)})\\ \vdots\\ \mu(u^{(r)})\end{pmatrix}
-\in \mathbb{R}^{\no r}, \\
-\text{and } \covUU &= \begin{pmatrix}
-\cov(\gp[u^{(1)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(1)}],
-\gp[u^{(r)}])\\
-\vdots & & \vdots\\
-\cov(\gp[u^{(r)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(r)}],
-\gp[u^{(r)}])\\
-\end{pmatrix}\in \mathbb{R}^{pr\times pr},
-\end{align*}
-each of the $r\times r$ blocks of the latter matrix being itself a (cross-)covariance matrix of dimension $\no \times 
-\no$.
-%dimensional Gaussian random vector.
-Assuming further that $\covUU$ is non-singular, the probability of interest can be formulated in terms of the $\no 
-r$-dimensional Gaussian probability density function
-$\varphi_{\no  r}(\cdot;~\meanUU, \covUU)$ as
-%$\varphi_{\no \times r}$ as
-%$\bm{u}\in D^r$ such that $\gp[\bm{u}]$ be non-degenerate, the integrand can be expressed in terms of the $r \times \no$-dimensional Gaussian distribution, namely
-%cumulative distribution function, namely
-\begin{equation*}
-\begin{split}
-&\jointExcuProb
-=
-\int_{T^r} \varphi_{\no  r}\left(\bm{v};
-    ~\meanUU, \covUU\right)
-    \mathrm{d}\bm{v},
-\end{split}
-\end{equation*}
-In the particular orthant case with $\T=(-\infty, t_1] \times \dots \times (-\infty, t_{r}]$,
-the latter probability directly writes in terms of the multivariate Gaussian
-cumulative distribution, % with mean $\meanUU$ and covariance matrix $\covUU$,
-this time by the way without requiring $\covUU$ %the latter
-to be non-singular:
-\begin{equation*}
-\begin{split}
-\jointExcuProb
-&=
-\varPhi_{\no r}\left(\bm{t};~\meanUU, \covUU\right),
-\end{split}
-\end{equation*}
-where we have used the notations
-$t=(t_1,\dots,t_{\no}%,..., t_1, ..., t_p
-)\in\mathbb{R}^{\no}$, $1_{r}=(1,\dots,1)\in \R^{r}$, and
-$\bm{t}=1_{r}\otimes \bm{t}=(t_1,\dots,t_{\no},\dots,t_1,\dots,t_{\no})
-\in \R^{\no r}$.
-\end{propo}
-\begin{proof}
-That $\mes(\es)$ defines indeed a random variable follows from Fubini's theorem
-relying on the joint measurability of
-$(\x, \omega) \to \mathbbm{1}_{\es(\omega)}(\x)$,
-%= \mathbf{1}_{\xi_{\x}(\omega) \in T}$ and Fubini's theorem.
-itself inherited from the assumed measurability for
-$(\x, \omega) \to \gp[\x](\omega)$ and $T$, respectively. From there, following the steps of Robbins' theorem \cite{Robins1944}, we find that
-\begin{equation*}
-\begin{split}
-\mathbb{E}[\mes(\es)^{r}]
-&=\mathbb{E}\left[\left(\int_{D} \mathbbm{1}_{\gp[u] \in T} ~d\mes(u) \right)^{r} \right]
-=\mathbb{E}\left[ \prod_{i=1}^{r} \left(
-        \int_{D} \mathbbm{1}_{\gp[\uu^{(i)}] \in T} ~d\mes(\uu^{(i)})
-\right) \right] \\
-&
-=
-%\mathbb{E}\left[
-%\int_{D^{r}}
-%\prod_{i=1}^{r}
-%\mathbf{1}_{\gp[\uu^{(i)}] \in T}
-%\productMeasure
-%\right]
-%=
-\mathbb{E}\left[
-\int_{D^{r}}
-\mathbbm{1}_{\gp[\uu^{(1)}] \in T,\dots, \gp[\uu^{(r)}]  \in T}
-\productMeasure
-\right]
-%\\
-%&=\int_{D^{r}} \mathbb{E}\left[\mathbf{1}_{\gp[\bm{u}] \in T^r} \right] \productMeasure \\
-%&
-=\int_{D^{r}}
-\jointExcuProb
-\productMeasure.
-\end{split}
-\end{equation*}
-The rest consists in expliciting the probability of $T\times \dots \times T$ under the multivariate Gaussian distribution of
-$\left(\gp[u^{(1)}], \dots,  \gp[u^{(r)}] \right)$.
-\end{proof}
-
-Proposition~\ref{propo1} enables in turn calculating centred moments of $\mes(\es)$.
-The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ is probably the most relevant of them for the considered purposes, and hence writes
+Centred moments may be computed using Proposition~\ref{propo1} developed in the appendix. 
+In particular, it allows to write the excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ as an integral of the excursion probability
 \begin{equation*}
@@
 \emv
-&=\int_{D^2} \mathbb{P}\left(
+&=\int_{\domain^2} \mathbb{P}\left(
 \gp[u]\in T, \gp[v]\in T \right)
 d\mes^{\otimes}(u, v)\\
-&-\left( \int_{D} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
+&-\left( \int_{\domain} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
 \end{split}
@@
 %\operatorname{Var}[\mes(\es)]
-&=\int_{D^2}
+&=\int_{\domain^2}
 \varPhi_{2\no}
@@
 (u,v)\\
-&-\left( \int_{D} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
+&-\left( \int_{\domain} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
 \end{split}
@@
 %
-Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $D^2$. In 
+Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $\domain^2$. In 
 contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves 
-solely and integral on $D$ and can be expanded in our present settings as follows
+solely an integral on $\domain$ and can be expanded in our present settings as follows
 \begin{equation*}
@@
 \operatorname{IBV} %(\es) %;\mes)
-&=\int_{D}
+&=\int_{\domain}
 \mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
 d\mes(u) \\
-&=\int_{D}
+&=\int_{\domain}
 \varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right)
@@
 Next we investigate uncertainty reduction criteria related to those indicators.
-In order to be able to calculate them in semi-analytical forms relying on integrals of multivariate Gaussian cumulative 
-distribution functions, we will appeal to some key intermediate results presented next.

### 2020-06-12 appendix.tex
Updates from Overleaf

--- 2020-06-09 main_new.tex
@@
-\documentclass[aoas]{imsart}
-
-\input{commands.tex}
-\usepackage{amsmath}
-\usepackage{pstricks,pst-grad}
-\usepackage{graphicx}
-\usepackage{floatrow}
-\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
-\floatsetup[table]{capposition=top}
-\usepackage{subfigure}
-\usepackage[utf8]{inputenc}
-\usepackage{booktabs}                     % horizontal lines in tables
-\usepackage{comment}
-
-% == Enable text degree
-\usepackage{textcomp}
-
-\usepackage{amsthm,amsmath,natbib}
-\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
-
-% == Trygve Test
-\usepackage{color, colortbl}
-\definecolor{Gray}{gray}{0.9}
-
-% put your definitions there:
-\startlocaldefs
-\newcommand{\edcomment}[1]{{\color{green}{\{Editor: #1\}}}}
-\newcommand{\frevcomment}[1]{{\color{blue}{\{Rev 1: #1\}}}}
-\newcommand{\srevcomment}[1]{{\color{red}{\{Rev 2: #1\}}}}
-\newcommand{\trevcomment}[1]{{\color{violet}{\{Rev 3: #1\}}}}
-\endlocaldefs
-
-% New Marcos.
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\input{macros}
-
-\begin{document}
-
-\begin{frontmatter}
-
-% "New Title of the paper"
-\title{Autonomous Oceanographic Data Collection towards Estimating Excursion Sets of Vector-valued Random Fields} 
-\runtitle{Autonomous Oceanographic Data Collection}
-
-% "Former Title of the paper" (DG 08.06.2020)
-%\title{Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields} \runtitle{Excursion Probabilities for Informative Sampling}
-
-\begin{aug}
-\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
-\author{\fnms{Cédric} \snm{Travelletti}\thanksref{t3}, \corref{} \ead[label=e2]{cedric.travelletti@stat.unibe.ch}}
-\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e3]{jo.eidsvik@ntnu.no}}
-\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e4]{david.ginsbourger@stat.unibe.ch}}
-\and
-\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e5]{kanna.rajan@fe.up.pt}}
-
-\affiliation[t1]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
-\affiliation[t2]{Centre for Autonomous Marine Operations and Systems, NTNU.}
-\affiliation[t3]{Institute of Mathematical Statistics and Actuarial Science, University of Bern, Switzerland.}
-\affiliation[t4]{Department of Mathematical Sciences, NTNU.}
-\affiliation[t5]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
-
-\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
-\printead{e1}}
-\address{Cédric Travelletti\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
-Switzerland.
-\printead{e2}}
-\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e3}}
-\address{David Ginsbourger\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
-Switzerland.
-\printead{e4}}
-\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
-  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
-\printead{e5}}
-
-\runauthor{TO. Fossum et al.}
-\end{aug}
-
-\begin{abstract}
-
-  Improving and optimizing oceanographic sampling is a crucial task
-  for marine science and maritime management. Faced with limited
-  resources to understand processes in the water-column, the
-  combination of statistics and autonomous robotics provides new
-  opportunities for experimental design. In this work we develop
-  methods for efficient spatial sampling applied to the mapping of
-  coastal ocean processes by providing informative descriptions of
-  spatial characteristics of ocean phenomena. Specifically, we define
-  a design
-  criterion based on improved characterization of the uncertainty in
-  the excursions of vector-valued Gaussian random fields, and derive
-  tractable expressions for the expected Bernoulli variance reduction
-  in such a framework. We demonstrate how this criterion can be used
-  to prioritize sampling efforts at locations that are ambiguous,
-  making exploration more effective. We use simulations to study the
-  properties of methods and to compare them with state-of-the-art
-  approaches, followed by results from field deployments with an
-  autonomous underwater vehicle (AUV) as part of a study mapping the
-  boundary of a river plume. The results demonstrate the potential of
-  combining statistical methods and robotic platforms to effectively
-  inform and execute data-driven environmental sampling.
-
-%Motivated by the challenges related to efficient allocation of sampling resources in environmental sensing, the combination of Excursion Probabilities and Gaussian process modeling is explored for autonomous robotic sampling of ocean features; enabling information driven measures  sampling efforts to high-interest regions.
-These regions are usually characterized by gradients of measurable environmental variables, e.g., temperature or salinity gradients, on which EPs subsequently can be used...Correlation among samples and multivariate requirements are typical in environmental studies.
-
-\end{abstract}
-
-\begin{keyword}
-\kwd{Ocean Sampling}
-\kwd{Excursion Sets}
-\kwd{Gaussian Processes}
-\kwd{Experimental Design}
-\kwd{Autonomous robots}
-\kwd{Adaptive Information Gathering}
-\end{keyword}
-
-\end{frontmatter}
-\section{Introduction}
-
-%NB: The subsections are used as a temporary instrument to highlight the proposed structure and should probably be abolished upon convergence to a stable version.
-
-\subsection{Oceanic data collection and spatial design of experiments}
-
-%\textcolor{blue}{NB: I started to match previously listed thoughts/ideas with pieces of the former intro.
-Will do more asap.
-Also, Section 3 I find excellent (and more precise than the intro) and start thinking that some bits might end up in the newer intro (while some others could land in a later section).}
-
-Monitoring the world's oceans has gained increased importance in light of the changing climate and increasing anthropogenic impact. 
-\begin{comment} % TO BE RE-INSERTED LATER IN THE ETXT?
-A central problem for understanding these factors is the lack of representative data with sufficient resolution.
-Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of sampling.
-By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
-This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
-\end{comment}
-Central to understanding the changes taking place in the upper
-water-column is knowledge of the bio-geophysical interaction driven by
-an agglomeration of physical forcings (e.g. wind, topography,
-bathymetry, tidal influences, etc.) and incipient micro-biology driven
-by planktonic and coastal anthropogenic input, such as pollution and
-agricultural runoff transported into the ocean by rivers and streams.
-These often result in a range of ecosystem-related phenomena such as
-blooms and plumes, with direct and indirect effects on society. \kc{we
-  should look for a citation here. Check with John Ryan}
- One of the bottlenecks in the study of such phenomena lies however in the lack of obervational data with sufficient resolution.
-Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of data collection.
-Based on collected data and sometimes also of prior knowledge, models from spatial statistics can be used to make predictions of the state variables under consideration over regions of interest.
-In case of multiple continuous state variables being jointly studied relying on spatially scattered data, vector-valued random field models and associated prediction approaches such as co-Kriging (REFs) can be considered nowadays as part of the standard statistical toolkit in oceanographic research and beyond.
-%applications and beyond.
- %By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
-This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
-
-Traditional data collection to understand natural and artificial phenomena at sea has typically been based on static buoys, floats, or ship-based methods,
-with significant logistical limitations that directly impact coverage
-and sampling resolution. Modern methods using satellite remote-sensing
-provide large-scale coverage but have limited resolution, are limited
-to sensing the surface, and are impacted by cloud cover. Numerical
-ocean models similarly find it challenging to provide detail at fine
-scale \citep{Lermusiaux:2006}, and also come with computation costs that can be practically limiting. 
-%limiting computing resources, no time for hard-core ocean model Navier-Stokes, hence GPs and approximations, etc.
-%
-\begin{figure}[!h] 
-  \centering 
-  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
-    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
-  \hfill
-  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
-    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
-  \caption{\ref{fig:envir1} Traditional ocean observation based on 
-    ship-based sampling has been augmented by autonomous
-    robotic vehicles. % and their interactions.  
-    AUV platforms are an integral part of this network being able to
-    reason and make decisions for efficient onboard adaptive sampling.
-    % using the sense-plan-act control approach to autonomous control. 
-    \ref{fig:nidelven} The interaction of river and ocean creates
-    processes that are challenging to map, where the combination of
-    statistics and robotics can play a vital role in enabling more
-    effective oceanographic observation.\kc{The inset ``AUV robot''
-      looks very tacky and ad-hoc; might be best to remove it. Change
-      ``ships'' to
-      ``Ship'', remove the aerial drone, since we didn't use it and
-      don't plan to in this work.}}
-  \label{fig:envir} \end{figure}
-%
-\begin{comment} % Somehow redundant with the parag following below
-Mobile robotic platforms \citep{Bellingham07} have contributed significantly to environmental monitoring and sampling.
-In particular, AUVs have advanced the state of in-situ sampling, and made robotics an integral part of ocean observation, filling parts of the undersampling gap \cite{rudnick03,rudnick18} (See Fig.
-\ref{fig:envir1}).
-\end{comment}
-%
-The advent of robust mobile robotic platforms \citep{Bellingham07} has
-resulted in significant contributions to environmental monitoring and
-sampling in the ocean. In particular, autonomous underwater vehicles (AUVs) have advanced the state of data collection and consequently have made robotics an integral part of ocean observation; %our 
-previous work by some of the authors of the present paper have contributed to this effort \citep{das11b,Das2015,fossuminformation,fossum18b}.
-
-As full numerical ocean models based on complex differential equations cannot be run onboard a marine robot with limited computational capacity, statistical models relying on collected data and random field assumptions appear quite relevant as means to guide AUV data collection trajectories.
-Our main research angle in the present work is to extend sequential design strategies from the world of spatial statistics and computer experiments to frameworks featuring both vector-valued observational data and experimental designs that may be constrained by the nature of measurement paths.
-\\
-
-Surveys with AUVs are usually limited to observations along fixed
-transects that are pre-scripted in mission plans created manually by a
-human operator. Missions can be specified operating on a scale of
-hundreds of meters to tens of kilometers depending on the scientific
-context. Faced with limited coverage capacity, a more effective
-approach is to instead use onboard algorithms to continuously
-evaluate, update, and refine future sampling locations (sense-plan-act
-cycle in Fig. \ref{fig:envir1}), making the information gathering
-\emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
-% This usually occurs on a spatial grid, called a waypoint graph. 
-In doing so, the space of sampling opportunities is still limited by
-the so-called waypoint graph, a discretization of the search domain, but the AUV has flexibility to modify its path at each waypoint based on in-situ sensing and measurements onboard \citep{py10,Rajan12,Rajan12b}.\\
-
-The work presented here is primarily inspired by a cast study 
-%application where an AUV is used to measure both temperature and salinity in 
-pertaining to the spatial characterization of a frontal system generated by
-river plumes. Fig. \ref{fig:nidelven} shows the survey area in
-Trondheim, Norway, where cold freshwater enters from the river,
-creating a strong gradient in both temperature and salinity. Because
-of the local topography and the Coriolis force the cold fresh water
-tends to flow to the east. Depending on the river discharge, tidal
-effects, wind, and temperature differences, this boundary often gets
-distorted. Initial knowledge about the location and evolution of these
-features are highly uncertain, making deterministic planning
-challenging.
-Here the goal is to use AUV measurements towards a better identification of freshwater versus oceanic water, with the help of a vector-valued random field model for both temperature and salinity.
-As we will present in the next sections, some of the original contributions inspired by this specific problem are actually applicable to further settings and application areas.
-\\
-
-The questions tackled here hence pertain to the broader area of spatial data collection and experimental design, yet with the particularity of estimating some regions of the domain, mostly excursion sets, being implicitly defined by a scarcely observed vector-valued random field.
-%because of limited resources to sample the large oceanographic domains, one must plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs.
-The acquired in-situ information must then be assimilated into the statistical model onboard the platform, so that it can be used to re-compute the sampling criterion and in this way inform decisions on where to sample next.
-\\
-Moreover, given constraints on AUV movements and the fact that surveya rely by design in such a framework on successive measurements spread along a trajectory, addressing corresponding design problems calls for particular sequential strategies, possibly also accounting for the finiteness of resources by anticipating multiple steps ahead thereby improving over myopic strategies.
-\\
-
-Here we aim to leverage and extent recent progress in uncertainty quantification and reduction for excursion sets of Gaussian random fields in order to address vector-valued observational cases under constrained design such as in the motivating application of designing AUV trajectories to better distinguish between freshwater from ocean waters by relying on temperature and salinity measurements.
-%approach will take substantial advantage of existing work in the field of excursion set estimation mainly dedicated to computer experiments in the scalar-valued case, and where space exploration can be performed without constraints regarding the distance between successive design points.
-Original contributions compared to recent related work in the framework of scalar-valued computer experiments include: 
-\begin{itemize}
-    \item Extension of two important Sequential Uncertainty Reduction criteria of (REFs) to homotopic or heterotopic vector-valued settings,
-    \item Investigation of myopic and multiple-step ahead batch-sequential strategies for optimizing trajectories with respect to such criteria,
-    \item Replicable experiments with synthetic test cases and original data sets with accompagnying codes 
-\end{itemize}
-Let us briefly review random field modelling and recent advances in targetd sequential design of experiments based on Gaussian Processes before detailing further existing literature tackling the considered problems, our proposed approach, as well as outlining the rest of the paper.
-
-\subsection{Random field modelling and targeted sequential design of experiments}
-
-\begin{itemize}
-    \item 
-\end{itemize}
-
-% From earlier version; to be relocated?
-%Hence statistical proxy models of the environment must be used.
-For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
-\subsection{Related work and proposed approach in a nutshell}
-
-\begin{itemize}
-    \item 
-\end{itemize}
-
-\textcolor{blue}{We might want to re-use some of the following (from earlier version):}
-
-A major contribution of our work is to derive closed-form results for
-this design criteria for situations where the underlying model is
-based on multivariate GPs. Another contribution is that we embed this
-approach in a sequential strategy to produce sampling algorithms, and
-apply this to a real-world scenario of autonomously sampling
-temperature and salinity gradients using an AUV, with the intention of
-characterizing the variability of oceanic and riverine waters in
-typical mixed coastal environments.
-
-Motivating examples relevant for ESs of multivariate processes are
-abundant. In medicine, doctors do not rely solely on a single symptom
-but must see several combined effects before making a diagnosis. In
-our context of environmental sampling, the salinity and temperature
-excursions of a river plume can similarly help characterize the
-underlying bio-geochemical processes
-\citep{hopkins2013detection,Pinto2018}, where sampling can be aimed
-towards reducing uncertainty in the temperature and salinity ES.
-Unlike what has been done in previous plume exploration studies, we
-emphasize the links to spatial statistical modeling and targeted
-multivariate sampling criteria for sequential sampling.
-
-Other statistical
-work in the oceanographic domain include \cite{wikle2013modern}
-focusing on hierarchical statistical models; \cite{sahu2008space},
-studying spatio-temporal models for sea surface temperature and
-salinity data; and \cite{mellucci2018oceanic} looking at the
-statistical prediction of features using an underwater glider.
-In this paper the focus is not on statistical modeling, but rather on statistical results and computations for efficient sampling designs.
-
-We focus on spatial characterization of a frontal system generated by
-river plumes. Fig. \ref{fig:nidelven} shows the survey area in
-Trondheim, Norway, where cold freshwater enters from the river,
-creating a strong gradient in both temperature and salinity. Because
-of the local topography and the Coriolis force the cold fresh water
-tends to flow to the east. Depending on the river discharge, tidal
-effects, wind, and temperature differences, this boundary often gets
-distorted. Initial knowledge about the location and evolution of these
-features are highly uncertain, making deterministic planning
-challenging.
-
-Adaptive in-situ AUV sampling of an evolving frontal feature has been
-explored in \cite{fronts11,Zhang2012,Pinto2018,costa19}. These
-approaches typically use a reactive-adaptive scheme, whereby
-exploration does not rely on a statistical model of the environment,
-but rather adaptation is based on closing the sensing and actuation
-loop. Myopic sampling, i.e. stage-wise selection of the path (on the
-waypoint graph), has been used for surveys
-\citep{singh2009efficient,Binney2013} that focus largely on reducing
-predictive variance or entropy. These criteria are widely adopted in
-the statistics literature on spatio-temporal design as well
-\cite{bueso1998state,zidek2019monitoring}, but variance and entropy
-reduction are independent of the actual data realizations under the
-assumptions of GP models, so it has limited
-decisional flexibility. The use of data-driven adaptive criteria
-was introduced to include more targeted sampling of regions of
-scientific interest in \cite{Low2009} and \cite{fossuminformation}. In
-this paper, we focus on mapping the river plume by rewarding designs
-that improve the classification of ES in temperature and salinity.
-
-\subsection{Outline of the paper and its appendices / supplemental material}
-
-\begin{itemize}
-    \item 
-\end{itemize}
-
-The remainder of this paper is organized as follows: 
-%Section \ref{sec:bg} provides select background on ocean sampling. 
-Section \ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for
-vector-valued GPs. Section \ref{sec:heuristics} builds on these
-assumptions when deriving the sequential design criteria for adaptive
-sampling. Section \ref{sec:simulations} discusses properties of the
-methods in simulation studies. Section \ref{sec:case_study}
-demonstrates the methodology used in field work characterizing a river
-plume and finally, Section \ref{sec:concl_disc} contains a summary and
-a discussion of future work.
-
-%\subsection{}
-
-%\begin{itemize}
-%    \item 
-%\end{itemize}
-
-\begin{comment}
-\section{Former Introduction (Before 08.06.2020)}
-
-Monitoring the world's oceans has gained increased importance in light
-of the changing climate and increasing anthropogenic impact. 
-A central problem for understanding these factors is the lack of representative data with sufficient resolution.
-Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of sampling.
-By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
-This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
-
-The
-advent of robust mobile robotic platforms \citep{Bellingham07} has
-resulted in significant contributions to environmental monitoring and
-sampling in the ocean. In particular, autonomous underwater vehicles (AUVs), have
-advanced the state of sampling and consequently have made robotics an
-integral part of ocean observation; our previous work has contributed
-to this effort \citep{das11b,Das2015,fossuminformation,fossum18b}.
-
-The potential that lies in using information-based sampling methods, which we
-highlight in this paper, is to do efficient environmental sampling and
-overcome common challenges as discussed next. First, full numerical ocean models based on complex
-differential equations cannot be run onboard a marine robot with
-limited computational capacity.
-% provide accurate results for online
-% operations because the onboard computers lack the computational power
-% required.
-Hence statistical proxy models of the environment must be used. For our purpose, we rely on Gaussian process (GP)
-representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
-Second, because of limited resources to sample the large oceanographic domains, one must plan for
-active learning during the operation, where a relevant criterion is
-used to extract the most valuable designs. The acquired in-situ
-information must then be assimilated into the statistical model
-onboard the platform, so that it can be used to re-compute the
-sampling criterion and in this way inform decisions on where to sample
-next. We use a criterion focusing on the prediction of excursion sets
-(ESs) for the multivariate spatial processes of interest. Third, the
-number of sampling designs is enormous, creating a trade-off between
-optimization and computability. One must often compromise in both
-development and practical operations, and we will rely on
-short-horizon optimization routines for sequential designs among
-nearest neighbors points in a spatial grid (waypoint graph).
-
-\kc{The para above needs a rewrite to me. It starts off by stating
-  there are challenges. But then jumps to models without stating the
-  connection with sampling, and then jumps into learning without
-  stating why learning is important. Typically the English in formal
-  text starts with a statement which is supported or clarified. In the
-  two cases above, the connections are simply not obvious. If this was
-  there in the first version of the m/s submitted, I'm not sure why
-  the reviewers didn't catch it. This is not just a stylistic issue;
-  it has to do with making sense and bring the reader/reviewer along.}
-
-\kc{Second, I don't see a motivation for GPs here and as I read
-  further into the text. Even for stats folks, are GPs the ``best''
-  ways to go? Are they obviously the tools one would use in
-  environmental sampling? And even if they are, I think we ought to
-  make the case for it strongly and generally, so this paper is
-  actually accessible and understandable to those not in the Spatial
-  Stats field. I would suggest adding a short para on GPs and why
-  they're important -- i.e. motivate the use of GPs early, so we can
-  freely make attributions to them later, as in the next section, for
-  example.}
-
-The imperfect understanding of causality coupled with the underlying
-fundamental uncertainty of the ocean implies that we need to make a
-large number of measurements. In addition, the use statistical models
-and methods to determine what and how phenomena form, is one way to
-increase the quality of predictive power aiding our understanding of
-upper water-column processes.
-
-With this in mind, we design AUV sampling strategies for reducing the spatially-integrated Bernoulli variance (IBV) in the ES
-indicator variables. 
-ES of random fields are thoroughly presented in \citep{adler2009random}. Recent work on ESs and associated excursion
-probabilities (EPs) connected to spatial statistics include
-\cite{french2013spatio,bolin2015excursion,french2016credible}. Our
-focus is on uncertainty characterization and how this is influenced by
-sampling. In this sense, our work is similar to
-\cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who
-describe analytical results for the IBV for univariate processes.
-
-A major contribution of our work is to derive closed-form results for
-this design criteria for situations where the underlying model is
-based on multivariate GPs. Another contribution is that we embed this
-approach in a sequential strategy to produce sampling algorithms, and
-apply this to a real-world scenario of autonomously sampling
-temperature and salinity gradients using an AUV, with the intention of
-characterizing the variability of oceanic and riverine waters in
-typical mixed coastal environments.
-
-Motivating examples relevant for ESs of multivariate processes are
-abundant. In medicine, doctors do not rely solely on a single symptom
-but must see several combined effects before making a diagnosis. In
-our context of environmental sampling, the salinity and temperature
-excursions of a river plume can similarly help characterize the
-underlying bio-geochemical processes
-\citep{hopkins2013detection,Pinto2018}, where sampling can be aimed
-towards reducing uncertainty in the temperature and salinity ES.
-Unlike what has been done in previous plume exploration studies, we
-emphasize the links to spatial statistical modeling and targeted
-multivariate sampling criteria for sequential sampling.
-
-Other statistical
-work in the oceanographic domain include \cite{wikle2013modern}
-focusing on hierarchical statistical models; \cite{sahu2008space},
-studying spatio-temporal models for sea surface temperature and
-salinity data; and \cite{mellucci2018oceanic} looking at the
-statistical prediction of features using an underwater glider.
-In this paper the focus is not on statistical modeling, but rather on statistical results and computations for efficient sampling designs.
-
-The remainder of this paper is organized as follows: Section
-\ref{sec:bg} provides select background on ocean sampling. Section
-\ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for
-vector-valued GPs. Section \ref{sec:heuristics} builds on these
-assumptions when deriving the sequential design criteria for adaptive
-sampling. Section \ref{sec:simulations} discusses properties of the
-methods in simulation studies. Section \ref{sec:case_study}
-demonstrates the methodology used in field work characterizing a river
-plume and finally, Section \ref{sec:concl_disc} contains a summary and
-a discussion of future work.
-\end{comment}
-
-\begin{comment}
-\section{Ocean Processes and Observation}
-\label{sec:bg}
-
-Central to understanding the changes taking place in the upper
-water-column is knowledge of the bio-geophysical interaction driven by
-an agglomeration of physical forcings (e.g. wind, topography,
-bathymetry, tidal influences, etc.) and incipient micro-biology driven
-by planktonic and coastal anthropogenic input, such as pollution and
-agricultural runoff transported into the ocean by rivers and streams.
-These often result in a range of ecosystem-related phenomena such as
-blooms and plumes, with direct and indirect effects on society. \kc{we
-  should look for a citation here. Check with John Ryan}
-
-Traditional data collection to understand such phenomena at sea has
-typically been based on static buoys, floats, or ship-based methods,
-with significant logistical limitations that directly impact coverage
-and sampling resolution. Modern methods using satellite remote-sensing
-provide large-scale coverage but have limited resolution, are limited
-to sensing the surface, and are impacted by cloud cover. Numerical
-ocean models similarly find it challenging to provide detail at fine
-scale \citep{Lermusiaux:2006}. Mobile robotic platforms
-\citep{Bellingham07} have contributed significantly to environmental monitoring and sampling. In particular, AUVs have
-advanced the state of in-situ sampling, and made robotics an integral
-part of ocean observation, filling parts of the undersampling gap
-\cite{rudnick03,rudnick18}(See Fig. \ref{fig:envir1}).
-
-%Walter Munk referred to the 20\textsuperscript{th} century
-%as the \emph{century of undersampling} \citep{munk2002}, pointing out
-%a lack of sufficient sampling resolution in time and space as a
-%critical issue for characterizing ocean phenomena. Even today,
-%policymakers struggle with limited observability of severe
-%environmental problems; a prominent example is the recent (May 2018)
-%outbreak of harmful algal blooms in Northern Norway\footnote{Report
-%  from the Institute of Marine Research: https://bit.ly/2KZGJfI}, with
-%losses exceeding 12 million tons of salmon valued at \$80 million
-%(Norwegian Directorate of Fisheries). This is a case in point of how
-%adequate sampling efforts could have provided a more accurate
-%assessment and early warning of changing environmental conditions.
-
-\begin{figure}[!h] 
-  \centering 
-  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
-    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
-  \hfill
-  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
-    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
-  \caption{\ref{fig:envir1} Traditional ocean observation based on 
-    ship-based sampling has been augmented by autonomous
-    robotic vehicles. % and their interactions.  
-    AUV platforms are an integral part of this network being able to
-    reason and make decisions for efficient onboard adaptive sampling.
-    % using the sense-plan-act control approach to autonomous control. 
-    \ref{fig:nidelven} The interaction of river and ocean creates
-    processes that are challenging to map, where the combination of
-    statistics and robotics can play a vital role in enabling more
-    effective oceanographic observation.\kc{The inset ``AUV robot''
-      looks very tacky and ad-hoc; might be best to remove it. Change
-      ``ships'' to
-      ``Ship'', remove the aerial drone, since we didn't use it and
-      don't plan to in this work.}}
-  \label{fig:envir} \end{figure}
-
-\end{comment}
-
-\subsection{To be relocated elsewhere}
-The AUVs considered here are powered untethered platforms, that
-operate at $1$-$3$ m/s in the upper water column, where they are free
-to move in six degrees of freedom (6 DOF). The in-water operation time
-capability depends on survey speed, payload sensors and navigation;
-typically this amounts to 8-48 hrs. AUVs typically use single-board
-computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA
-Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared
-cache, and 4 GB of 1600 MHz DRAM) for computation onboard.
-
-\begin{comment}
-Surveys with AUVs are usually limited to observations along fixed
-transects that are pre-scripted in mission plans created manually by a
-human operator. The mission can be specified operating on a scale of
-hundreds of meters to tens of kilometers depending on the scientific
-context. Faced with limited coverage capacity, a more effective
-approach is to instead use onboard algorithms to continuously
-evaluate, update, and refine future sampling locations (sense-plan-act
-cycle in Fig. \ref{fig:envir1}), making the information gathering
-\emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
-% This usually occurs on a spatial grid, called a waypoint graph. 
-In doing so, the space of sampling opportunities is still limited by
-the waypoint graph, but the AUV has flexibility to modify its path at
-each waypoint based on in-situ sensing and measurements onboard
-\citep{py10,Rajan12,Rajan12b}.
-
-We focus on spatial characterization of a frontal system generated by
-river plumes. Fig. \ref{fig:nidelven} shows the survey area in
-Trondheim, Norway, where cold freshwater enters from the river,
-creating a strong gradient in both temperature and salinity. Because
-of the local topography and the Coriolis force the cold fresh water
-tends to flow to the east. Depending on the river discharge, tidal
-effects, wind, and temperature differences, this boundary often gets
-distorted. Initial knowledge about the location and evolution of these
-features are highly uncertain, making deterministic planning
-challenging.
-
-Adaptive in-situ AUV sampling of an evolving frontal feature has been
-explored in \cite{fronts11,Zhang2012,Pinto2018,costa19}. These
-approaches typically use a reactive-adaptive scheme, whereby
-exploration does not rely on a statistical model of the environment,
-but rather adaptation is based on closing the sensing and actuation
-loop. Myopic sampling, i.e. stage-wise selection of the path (on the
-waypoint graph), has been used for surveys
-\citep{singh2009efficient,Binney2013} that focus largely on reducing
-predictive variance or entropy. These criteria are widely adopted in
-the statistics literature on spatio-temporal design as well
-\cite{bueso1998state,zidek2019monitoring}, but variance and entropy
-reduction are independent of the actual data realizations under the
-assumptions of GP models, so it has limited
-decisional flexibility. The use of data-driven adaptive criteria
-was introduced to include more targeted sampling of regions of
-scientific interest in \cite{Low2009} and \cite{fossuminformation}. In
-this paper, we focus on mapping the river plume by rewarding designs
-that improve the classification of ES in temperature and salinity.
-
-\end{comment}
-
-\section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
-\label{sec:ESEP}
-
-We define the random sets of interest in Section \ref{sec:ES}, while the closed-form expressions for the expected uncertainty reduction are derived in Section \ref{sec:EIBV}.
-Section \ref{Sec:UnivarEx} provides an instructive bivariate example relevant for sampling in the temperature and salinity application.
-
-\subsection{Excursion sets and Integrated Bernoulli variance}
-\label{sec:ES}
-
-Our objective is to leverage statistical tools onboard an autonomous
-robotic platform to characterize a river plume, focusing on spatial
-separation of cold freshwater from a river and warmer saline waters
-of a fjord. The characterization of uncertainties on the ES is a
-useful goal of this sampling strategy.
-
-In order to formalize the problem in mathematical terms, let us
-consider a multivariate (i.e., vector-valued) Gaussian random field
-$(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary
-set (typically in geostatistical studies, a geographical subset of interest) and the
-dimensionality of the output $\no \geq 1$ is arbitrary. We have $\no = 2$ in the example with temperature and salinity in the river plume.
-
-Assume furthermore that the set of interest
-is characterized in terms of a subset of values of the variables, say
-$\T \subset \R^{\no}$, and denote the corresponding set of index
-points by
-$$
-\es=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
-$$
-Our goal here is to develop approaches to quantify and improve the
-characterization of uncertainties on such sets $\es$, in the
-context of multivariate random fields.
-
-While some aspects of the developed approaches do not call for a
-specific form of $\T$, we will often stick for simplicity to the case
-of orthants
-($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
-$t_1,\dots, t_{\no} \in \R$) as this will allow efficient calculation
-of several key quantities. Note that changing some $\leq$ inequalities
-to $\geq$ ones would lead to immediate adaptations. In the application
-specific properties following the orthant assumption come into play
-\kc{This last phrase doesn't parse; is it ``..., the following
-  orthant....''?}: We let $t_1$ be a threshold for temperature and
-$t_2$ a threshold for salinity \kc{Might be easier to remember with
-  symbols 't' and 's' in some
-  form.}. The set of index point then defines
-$\mbox{ES} = \{\x \in D : \gp[\x,1] \leq t_1,\gp[\x,2] \leq t_2\}$.
-The associated EPs are
-$p(\x) = P(\gp[\x,1] \leq t_1,\gp[\x,2] \leq t_2)$,
-$\x \in \mathcal{M}$.
-
-For statistical generality, we will however keep the abstract notation
-with $\T$ and general $\es$ \kc{Again, previous phrase doesn't parse.}
-for most of the derivations. Note that assuming that $\gp$ has
-continuous trajectories (almost surely) and $T$ to be closed, then
-$\es$ can be cast as a Random Closed Set
-\citep{Molchanov2005}. %In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
-Let us in particular consider a locally finite (Borel) measure $\mes$
-on $D$ and investigate the probability distribution of $\mes(\es)$
-through its moments.
-
-\medskip
-
-%That $\mu(\es)$ is a well-defined random variable follows from
 \begin{propo}
@@
 \end{propo}
-
-Proposition~\ref{propo1} enables one to calculate centered moments of $\mes(\es)$. 
-The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$
-is probably the most relevant of these for our purposes. We
-have 
-\begin{equation*}
-\begin{split}
-\emv
-&=\int_{\mathcal{M}^2} \mathbb{P}\left(
-\gp[u]\in T, \gp[v]\in T \right) 
-d\mes^{\otimes}(u, v)\\
-&-\left( \int_{\mathcal{M}} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
-\end{split}
-\end{equation*} 
-which evaluates in the excursion/sojourn case where
-$\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ to
-\begin{equation*} \begin{split}
-  \emv
-%\operatorname{Var}[\mes(\es)]
-&=\int_{\mathcal{M}^2} 
-\varPhi_{2\no}
-\left(
-(\bt, \bt); \bmu((u,v)), 
-K((u,v),(u,v))
-\right) 
-\
-\mathrm{d}\mes^{\otimes} %\mes 
-%\productMeasure 
-(u,v)\\
-&-\left( \int_{\mathcal{M}} \varPhi_{\no}\left(\bt;\bmu(u), K(u)\right) d\mes(u) \right)^2.
-\end{split}
-\end{equation*}
-This quantity requires  an integral over $D^2$. In
-contrast, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal}
-involves solely an integral on $\mathcal{M}$ and can be expanded in
-our present setting as follows 
-\begin{equation}\label{IBV}
-\begin{split}
-\operatorname{IBV} %(\es) %;\mes)
-&=\int_{\mathcal{M}}
-\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
-d\mes(u) \\
-&=\int_{\mathcal{M}}
-\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) 
--\left(\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) \right)^2
-\mathrm{d}\mes(u). 
-\end{split}
-\end{equation}
-
-\subsection{Expected Integrated Bernoulli Variance and Sampling Designs}
-\label{sec:EIBV}
-
-Our goal is to construct informative sampling designs, given limited
-resources \kc{information?}. In our domain, data is gathered by an AUV
-that can measure temperature and salinity at chosen locations on a
-defined waypoint graph. Such a discretized grid fosters reliable
-navigation and mission performance. The waypoints further enable a
-structure to the design problem in that, that sampling flexibility is
-limited to operation on the graph. % This gives lots of sampling
-% opportunities, and it is important to spend the sampling efforts
-% wisely.
-We focus on sampling designs that reduce the expected integrated
-Bernoulli variance (EIBV). Note that this sampling can be done in a
-static manner where the survey design is pre-planned, or in an
-adaptive setting. For the latter case, data can be processed
-sequentially and the updated model is then used to design and select
-the next waypoint graph to visit, which in turn are used again to
-update the model, and so on (Section \ref{sec:heuristics}). We next
-describe the data modeling assumptions for one stage and develop
-design criteria of semi-analytical forms relying on integrals of
-multivariate Gaussian cumulative distribution functions; we will
-appeal to some key intermediate results presented next.
-
-We denote a design by $d \in \mathcal{D}$, where the concatenation of designs with $q$ sampling locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$ is $\mathcal{D}$. The
-measurement model is defined by 
-\begin{equation}\label{eq:meas}
-    \bY_d=G_d \bz(\bu_d) + \epsilon, \hspace{3mm} \epsilon \sim N(0,R_d).
-\end{equation}
-Here, $\bz(\bu_d)$ is the size $pq$ vector of GP variables at the
-sampling locations, and $G_d$ is the design matrix which in principle
-allows for observing subsets of the vector-valued field at some of the
-$q$ locations (i.e. the matrix has fewer rows than columns). If the
-additive error has independent elements, the covariance matrix $R_d$
-is diagonal. In our setting temperature and salinity measurements are
-extracted from different sensors, so the independence assumption
-between variables is not unrealistic, but this is not necessarily the
-case for other kinds of measurements on the AUV. It is further common
-in geostatistics to assume spatially independent measurement errors,
-and we also do so. But our methodology for constructing sampling
-designs does not rely on these independence assumptions.
-
-For a given design $d$ the evaluation is based on the expected
-information contained in the data to influence the criterion.   
-Considering the EIBV criteria, we have
-\begin{eqnarray}\label{eq:eibv}
-    \mbox{EIBV} (d) &=& E_{\bY_d} \left( \int_{\mathcal{M}}
-p(\x;\by_d)(1-p(\x;\by_d))
-d\mes(u) \right), \nonumber \\
-p(\x;\by_d) &=& \mathbb{P}\left(\gp[\uu]\in T|\bY_d=\by_d \right)
-\end{eqnarray}
-where the expectation is over the planned data.  The optimal design
-for an AUV survey is then 
-\begin{equation}\label{crit}
-    d^* = \mbox{argmin}_{d \in \mathcal{D}} \mbox{EIBV}(d).
-\end{equation}
-For another criterion using random moments of random sets similar
-expectations would be required, but as we will see, the EIBV has a
-closed-form expression which is relatively fast to compute and hence
-useful for onboard design selection.
-
-Assuming now that $q$ observations $\by_d$ made at $\bu_d$ are available, one wishes to predict $\gp[\x]$. The multivariate Kriging equations holds with size
-$p$ vector of conditional means:  
-\begin{equation}\label{eq:cokrig}
-\bbm(\x)=\bmu(\x)+\lambda(\x,\bu_d)^T \left( \by_d-G_d \bmu(\bu_d)\right),
-\end{equation}
-where
-$\lambda(\x,\bu_d)=K^{-1}_y G_d k(\bu_d, \bu) $ with data covariance
-$K_y=\mbox{Cov}(\bY_d,\bY_d)=G_d K(\bu_d,\bu_d)G_d^T+R_d$ and cross-covariance $\mbox{Cov}(\bY_d,Z(\x))=G_d k(\bu_d, \x)$, assuming the measurement
-error is independent of the random fields.  The Kriging (cross-)covariance function can also be expressed in the same vein via
-a $p \times p$ matrix 
-\begin{equation}\label{eq:krig:cov}
-\Sigma(\x,\x')=K(\x,\x')-\lambda(\x,\bu_d)^T K_y(\bu_d, \bu_d) \lambda(\x,\bu_d).
-\end{equation}
-
-Eq. \eqref{eq:cokrig} and \eqref{eq:krig:cov} are used to
-provide informative, updated statements about sets of interest. For
-our case with temperature and salinity, the conditional EPs, given
-measurements $\by_d$, can then be computed from
-\begin{eqnarray}\label{eq:post_ep}
- p(\x;\by_d) &=& P(Z_1(\x) \leq t_1, Z_{2}(\x) \leq t_2 |\bY_d=\by_d),  \nonumber \\
- &=& \Phi_2 ((t_1,t_2)^T-\bbm(\x);0_2, \Sigma(\x,\x)). 
-\end{eqnarray}
-
-A critical element in the derivation of a closed-form for the EIBV is
-that the conditional mean in Eq. \eqref{eq:cokrig} is a linear
-(affine) function of the data $\by_d$ and the covariance is not a
-function of the data. This in turn means that the probabilities in Eq.
-\eqref{eq:post_ep} involve inequality statements for linear
-combinations of Gaussian variables. Related closed-form solutions have
-been noted in similar contexts \citep{bhattacharjya2013value,
-  chevalier2014fast,stroh}, but not generalized to our situation with
-random sets for vector-valued GPs. Notably, the entry in Eq.
-\eqref{eq:post_ep} is in the form of $\ba+B V$, where
-$V \sim N(0_q, K_y (\bu_d,\bu_d))$,
-$\ba=(t_1,t_2)^t-\bmu(\x)$
-and $B=\lambda(\x,\bu_d)^T$. The following proposition then applies to give a
-closed-form solution to EIBV calculation in this case.
-
-\begin{propo}
-	\label{propo2}
-	Let $p, q, h \geq 1$, $a \in \R^p$, $B \in \R^{p\times q}$,
-	and $\covN, \covV$ be two covariance matrices in
-	$\R^{p\times p}$ and $\R^{q\times q}$, respectively.
-	Then, for $V \sim \mathcal{N}_{q}(0_q, \covV)$,
-	\begin{equation*}
-	\mathbb{E}\left[ \varPhi_{p}\left( \ba + BV; 0_p,\covN \right)^h \right]
-	=
-	\varPhi_{ph}
-	\left(
-	\ba
-	;0_{ph},
-	\Sigma
-	\right),
-	\end{equation*}
-	where the vector $\ba \in \R^{p h}$ is defined as 
-	$\ba := 1_h\otimes a = 
-	\left(a, \dots , a
-	\right)'$
-	and the $p h\times p h$ covariance matrix is given by
-	$\Sigma := 
-	1_h 1_h'\otimes B\covV B' + I_h\otimes \covN$.
-\end{propo}
-The proof of Proposition \ref{propo2} is provided in the Appendix.
-
-In block-wise representation, the following relation holds for the
-covariance in Proposition \ref{propo2}: 
-\begin{align*}
-	\Sigma&=
-	\begin{pmatrix}
-	\covN & &\\
-	& \ddots &\\
-	&   & \covN
-	\end{pmatrix}
-	+
-	\begin{pmatrix}
-	B\covV B' & \dots & B\covV  B'\\
-	\vdots & & \vdots\\
-	B\covV B' & \dots & B\covV B'\\
-	\end{pmatrix}.
-\end{align*}
-The calculation of the multivariate Gaussian cumulative distribution
-is done effectively using code such as that of
-\cite{genz2009computation}.
-
-In the more general case, for instance that of volumes of random sets,
-the criterion will include couplings with different locations and the
-result must be generalized. For the case of multivariate monomials in
-orthant probabilities with thresholds affine in a common Gaussian
-vector \kc{This sentence is dangling. Doesn't have a conclusion.}.
-%Namely, for monomials of degree $k$, we have.
-
-\begin{propo}%[\textcolor{red}{Notation in progress, to be reviewed/revised}]
-	\label{propo3}
-	Let $g, p, q\geq 1$, $h_{1},\dots, h_{g}\geq 1$ with $H=\sum_{i=1}^g h_i$, $a_{i} \in \R^{p}$, $B_{i}\in \R^{p \times q}$, and covariance matrices $\covN_i \in \R^{p \times p}$ $(1\leq i \leq g)$.
-Then, for any covariance matrix $\covV \in \R^{q\times q}$ and $V\sim\mathcal{N}_{q}(0_q,\covV)$,
-	%	Let $g$ and $k$ be integers, and consider a $g$-terms monomial of degree $k$
-	%	\[
-	%	    \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \Sigma_{i}
-	%	    \right)^{h_i},
-	%    \]
-	%    where $\sum_{i=1}^g h_i = k$, each $a_i$, $B_i$ and $\Sigma_i$ is as in
-	%    \cref{propo2} and $V$ is a $q$-dimensional Gaussian vector distributed as
-	%    $V\sim\mathcal{N}(0,\covV)$.
-	%
-	% Then the expectation of the monomial is given by
-	\begin{equation}
-	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V;0_p, \covN_{i} \right)^{h_i} \right]
-	=
-	\varPhi_{p H}
-	\left(
-	\bm{a}
-	;0_{pH},
-	\mathbf{\Sigma}
-	\right),
-	\end{equation}
-	with $\bm{a}=(1_{h_1}\otimes a_1, \dots, 1_{h_g}\otimes a_{g}) \in \R^{p H}$
-	and $\mathbf{\Sigma}\in \R^{p H \times p H}$ is defined blockwise by $(\Sigma_{i,j})_{i,j \in \{1,\dots, g\}}$ where, for any $i,j \in \{1,\dots, g\}$, %the blocks are given by 
-	\begin{equation}
-	\Sigma_{i,j}=
-	(1_{h_{i}}1_{h_{j}}')\otimes (B_{i}\Sigma_{V} B_{j}') + \delta_{i,j}(I_{h_{i}}\otimes \covN_{i}) \in \R^{p h_{i} \times p h_{j}}.
-	\end{equation}
-	%with $\bm{a}=\sum_{j=1}^g \mathbf{1}_{J_{j}} \otimes a_j \in \R^{p \times k}$ 
-	%where $J_{j}=\{\sum_{i=1}^{j-1}h_i + 1, \dots, \sum_{i=1}^{j}h_i\}$,
-	%$\mathbf{1}_{I_{j}}$ denotes the vector of $\R^{k}$ with components equal to $1$ for indices in $I_{j}$ and $0$ else ($1\leq j \leq g$), 
-	%and $\mathbf{\Sigma}\in \R^{p k \times p k}$ is defined blockwise by 
-	%\begin{equation}
-	%\mathbf{\Sigma}_{\widetilde{J}_{j_{1}},\widetilde{J}_{j_{2}}}=
-	%(\mathbf{1}_{h_{j_1}}\mathbf{1}_{h_{j_2}}')\otimes (B_{j_{1}}\Sigma_{V} B_{j_2}') + \delta_{j_{1},j_{2}}(I_{j_{1}}\otimes \Sigma_{j_{1}}),
-	%\end{equation}
-	%where $\widetilde{J}_{j}=\{p\sum_{i=1}^{j-1}h_i + 1, \dots, p\sum_{i=1}^{j}h_i\}$ ($1\leq j \leq g$).
-\end{propo}
-The proof of Proposition \ref{propo2} is provided in the Appendix.
-
-\subsection{An Example of EIBV using Temperature and Salinity}
-\label{Sec:UnivarEx}
-
-% In this section, 
-We calculate EPs and the Bernoulli variance for a few bivariate
-Gaussian distributions for temperature and salinity to illustrate the concepts articulated above.
-The parameter inputs are made up, but they are in a realistic range as we have in the simulation study and real-world example below.
-The thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
-
-Fig. \ref{illus_bivarDens} shows contour plots of three different
-densities with increasing correlation $\gamma$ between temperature and
-salinity. 
-\begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.pdf}
-  \caption{Density contour plots with different correlations between
-    temperature and salinity. The densities have unit variance and the
-    thresholds are identical to the mean values $5^o C$ and
-    $30 mg/l$. X-axis is temperature and y-axis is salinity.}
-\label{illus_bivarDens}
-\end{figure}
-The displayed densities have unit standard deviations for both
-temperature and salinity, but we also study the effect of doubling the
-standard deviations.
-
-Table \ref{tab:sim_rhoab} shows the initial EPs and the associated
-Bernoulli variance (second row) for the examples indicated in Fig.
-\ref{illus_bivarDens}. The EPs increase with the correlation as there
-is a strong tendency to have concurrently low temperature and salinity. The Bernoulli variance is similarly large for high
-correlations. EPs and Bernoulli variances are the same for standard
-deviation $1$ or $2$, which implies that high variability in
-temperature and salinity is not captured in the $p(1-p)$ expression.
-
-\begin{table}[!h] \centering \caption{EP and Bernoulli variance for
-    different correlations and variances (top rows), and expected
-    Bernoulli variances for both temperature and salinity data $\by$ and 
-    temperature $y_2$ (bottom rows).}
-  \begin{tabular}{c|ccc|ccc}
- &\multicolumn{3}{c}{$\sigma_1=\sigma_2=1$} & \multicolumn{3}{c}{$\sigma_1=\sigma_2=2$} \\
-\hline
-Correlation $\gamma$ & 0.2 & 0.6 & 0.8 & 0.2 & 0.6 & 0.8 \\
-\hline
-$p$ & 0.28 & 0.35 & 0.40 & 0.28 & 0.35 & 0.40 \\ 
-$p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
-EIBV, Temp and Salinity data & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
-EIBV, Temperature data only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
-\hline
-\end{tabular}
-\label{tab:sim_rhoab}
-\end{table}
-
-Table \ref{tab:sim_rhoab} (bottom two rows) shows results of expected
-Bernoulli variance calculations. This is presented for a design
-gathering both data types, and for a design with temperature
-measurements alone. When both data are gathered, the measurement model is
-$(Y_{d,1},Y_{d,2})^t=(Z_1,Z_2)^t+\bepsilon$, with $\bepsilon \sim N(0,0.5^2I_2)$, while $Y_d=Z_1+\epsilon$, $\epsilon \sim N(0,0.5^2)$ when only temperature is measured.
-For this illustration, Table \ref{tab:sim_rhoab} shows that the
-expected Bernoulli variance gets lower with larger standard deviations
-$\sigma_1$ and $\sigma_2$ (right columns). The reduction of Bernoulli
-variance is largest for the cases with high correlation
-$\gamma$. Albeit smaller, there is also uncertainty reduction when
-only temperature is measured (bottom row), especially when temperature
-and salinity are highly correlated. When correlation is low
-($\gamma=0.2$), there is little information about salinity in the
-temperature data, and therefore less uncertainty reduction. In an
-application with fresh cold water from a river source, the temperature
-and salinity variables will not only be interdependent, but will also
-likely show dependence in the spatial dimension. This in turn will
-impact the design criteria when we evaluate the information measure by
-integrating over several locations (Section \ref{sec:simulations}).
-
-%\subsection{Expected classification criteria}
-
-%{\bf{I have not done anything here - skip this, I think}}.
-
-%Rather than minimizing expected variance one can aim at minimizing the classification probability of the excursion set: 
-%$\min [p,(1-p) ]$, see e.g. \cite{lilleborge2016information}.
-
-%Again, since the conditional mean is a linear in the data, we only have to look at the %relevant linear combination via $\bm_{\xi}=E(\bxi)$.
-%See also \cite{bhattacharjya2013value}.
-
-%\begin{equation}
-%E(\min \{ p,(1-p)\})=\int \min P(\xi_1 \leq t, \xi_2 \leq s),[1-P(\xi_1 \leq t, \xi_2 \leq s)] p(E(\bxi)=) dE(\bxi)=,
-%\end{equation}
-%for the complementary probabilities, this is again evaluated by the corner regions for the block, and $\Phi_4()$ evaluations are required.
-
-%In the end, this result is integrated over the spatial domain $x \in X$.
-
-\section{Sequential designs and heuristic path planning}\label{sec:heuristics}
-
-An important concept underlying this work is the potential of robotic
-sampling to adapt survey plans based on data and statistical methods
-in accordance with the sense-plan-act control loop
-(Fig. \ref{fig:envir1}). We now present a few adaptive strategies for
-robotic sampling that build on the notion of EIBV.
-
-\subsection{Optimal Sequential Design}
-\label{Optdes}
-
-Our adaptive AUV survey is split into many stages. The selection on
-where to sample is constrained by a triangular waypoint graph and
-defined by its nearest grid nodes.
-
-Let $d^{j,s}$ denote design number $j$ at stage $s$ of a sequential
-survey.
-For the AUV sampling situation, the choice of design entails a commitment to one of many possible navigation directions on the waypoint graph.
-Once the choice has been made, the AUV steers in the chosen direction and this entails sampling locations $\bu_{d^{j,s}}$ and data $\by_{d^{j,s}}$ is gathered.
-We let
-$\mathcal{Y}^{s-1} =\{(d^{*,1}, \by_{d^{*,1}}),\ldots, (d^{*,s-1},
-\by_{d^{*,s-1}}) \}$ denote the set of all data gathered until stage
-$s-1$ according to selected design $d^{*,1},\ldots,d^{*,s-1}$.
-
-The design and data define
-the conditional GP with mean
-$E(\gp[\x]|\mathcal{Y}^{s-1})$ and (cross) covariance matrix
-$\mbox{Var}(\gp[\x]|\mathcal{Y}^{s-1})$.
-These are efficiently updated in a sequential manner keeping track of the mean and covariance kernel of the GP over the stages.
-As extensions of Eq.
-\eqref{eq:cokrig} and \eqref{eq:krig:cov}, we have
-\begin{eqnarray}\label{eq:krigSUR}
-\mu^{s}(\x) &=&\mu^{s-1}(\x)+\lambda^{[s,s]}(\x,\bu_{d^{j,s}})^T (\by_{d^{j,s}}-G_{d^{j,s}} \mu(\x)), \nonumber \\
-k^{s}(\x,\x') &=& k^{s-1}(\x,\x')-\lambda^{[s,s]}(\x,\bu_{d^{j,s}})^T K_{Y_{d^{j,s}}} \lambda^{[s,s]}(\x',\bu_{d^{j,s}}),
-\end{eqnarray}
-where $\lambda^{[s,s]}(\x,\bu_{d^{j,s}})=K^{-1}_{Y_{d^{j,s}}} G_{d^{j,s}} k(\bu_{d^{j,s}},\x)$ and the data covariance is $K^{-1}_{Y_{d^{j,s}}}=G_{d^{j,s}} K^s(\bu_{d^{j,s}},\bu_{d^{j,s}})G^T_{d^{j,s}}+R_{d^{j,s}}$, with measurement model as in Eq.
-\eqref{eq:meas}.
-
-At stage $s$, the robot must then choose design $d^{j,s}$ where measurements $\by_{d^{j,s}}$ should
-be gathered. For the triangularized waypoint
-graph all six possible
-directions must be evaluated to minimize the expected uncertainty of random sets, and in particular the EIBV for our 
-application. We hence focus on choosing designs for data modeled by
-Eq. \eqref{eq:meas}, to conduct efficient sequential uncertainty
-reduction according to the EIBV criterion. Compared with the
-static design $d$ in Eq. \eqref{eq:eibv}, the optimal sequential
-design not only considers the best current design grid point, but also
-how the data gathered at this node could inform future sampling at
-successive stages.
-
-%The optimal path
-%selection situation is depicted in Fig. %\ref{fig:PathSelOpt},
-%\begin{figure}[b!]
-%\centering
-%\includegraphics[width=0.85\textwidth]{Figures/sequent_select.pdf}
-%\caption{Optimal sequential path selection.}\label{fig:PathSelOpt}
-%\end{figure}
-%where design choices are indicated by squares while %data realizations are indicated by circles. 
-%terms this optimal design is defined as
-%\begin{equation}\label{opt_crit}
-%    \bd^* = \mbox{argmin}_{d^{j,1}} \left\{ \int_{\by^{j,1}} \mbox{argmin}_{d^{j,2}} \left\{ \int_{\by^{j,2}} \ldots \pi(\by^{j,2}|\by^{j,1}) d\by^{j,2} \right\} \pi(\by^{j,1}) d \by^{j,1} \right\},
-%\end{equation}
-%where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages. 
-The mathematical expression for the optimal design involves a series
-of intermixed maximizations over designs and integrals over data. In
-practice, the optimal solution is intractable because of the enormous
-growth over stages (see e.g. \cite{sucar2015probabilistic} and
-\cite{powell2016perspectives}).  Instead, we outline heuristic
-strategies.
-
-%The derived closed-form solutions for the EIBV are still important building blocks when constructing the sampling designs, as they will be used to score the different adaptive designs. % Efficient
-% calculation is important for using adaptive survey designs with
-% robotic platforms.
-
-\subsection{A Naive Sampling Strategy}
-\label{naive}
-
-The simplest heuristic for adaptive sampling is to choose the next
-sampling location based on current probabilities. One reduces the
-largest levels of uncertainty by selecting the design $j$ with
-$P(Z(\x_j) \in T)$ closest to $0.5$.  With temperature and salinity
-variables, this {\it{naive}} strategy selects the stage $s$ sampling
-design according to:
-\begin{eqnarray}\label{critNaive}
-    d^{*,s} &=& \mbox{argmin}_{j} |P(Z_1(\x_j) \leq t_1, Z_2(\x_j) \leq t_2 | \mathcal{Y}^{s-1})-0.5|, \nonumber
-\end{eqnarray}
-where $j$ is indexing all neighboring locations on the waypoint graph.
-
-This strategy does not account for the uncertainty in the GP
-variables, but is instead applicable only if one design (node) has EPs
-closer to $0.5$ (see Table \ref{tab:sim_rhoab}, row
-two). Additionally, this strategy does not account for spatial
-correlation. This strategy lacks memory of where it has been and where
-the uncertainty has been reduced, and therefore susceptible to local
-minima.
-
-\subsection{Myopic Path Planning}
-\label{sec:myopic}
-
-The myopic (greedy) strategy which we present here is optimal if we
-imagine taking only one more stage of measurements. This selection
-strategy is based on expected uncertainty reduction, but it is a
-heuristic because there is no anticipation of what the subsequent
-designs might offer beyond the first stage.
-
-Based on the currently available data $\mathcal{Y}^{s-1}$, we fit an
-updated GP model. Focusing on the EIBV criterion, at the next stage
-$s$, the selected design is:
-\begin{eqnarray}\label{critSEQ}
-    d^{*,s} &=& \mbox{argmin}_{j} \left\{ \int_{\mathcal{M}} E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( p(\x;\mathcal{Y}^{j,s})\left( 1-p(\x;\mathcal{Y}^{j,s})\right) \right) d\mes(u), \right\} \\
-    p(\x;\mathcal{Y}^{j,s}) &=& P(Z_u \in T |\by_{d^{j,s}},\mathcal{Y}^{s-1}). \nonumber
-\end{eqnarray}
-
-Note that this strategy gives a sequential conditional version of
-expression (\ref{eq:eibv}). Now $\mathcal{Y}^{s-1}$ is available, and
-the expectation is conditional to this data. A similar closed-form
-calculation for EIBV is hence applicable in Eq.  \eqref{critSEQ} using the updated
-GP model from stage $s-1$ as defined via Eq.
-\eqref{eq:krigSUR} in the results from Proposition \ref{propo2}.
-Once the data are collected for the best design $d^{*s}$ at stage $s$, the GP model is updated again.
-The mean and covariance are used to compute the next design at stage $s+1$, and so on.
-
-Even though this myopic strategy is non-anticipative, it still gives a
-reasonable approach for creating designs in many
-applications. Moreover, it is easily implemented, in our case, onboard
-an AUV, using an efficient data update of the GP model and the
-calculation of the closed-form EIBV expressions for each subsequent
-trajectory.
-
-\subsection{Look-ahead Trajectory Planning}
-\label{sec:LA}
-
-We now extend the myopic strategy to a look-ahead strategy. We consider two stages of
-measurements, and the strategy is then optimal in the sense that it accounts consistently for the expectations and minimizations in these two stages, but still without including any planning beyond those two steps.
-In addition to the
-next stage of measurements $\by^{j,s}$, this look-ahead strategy anticipates the subsequent design indexed $j_2$ with data
-$\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$. The
-selected design is:
-\begin{eqnarray}\label{critLA}
-    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( \mbox{argmin}_{j_2}  V_{j_2} \right) \right\}, \\
-V_{j_2} & = & \int_{\mathcal{M}} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1}) ( 1-p(\x;\mathcal{Y}^{j,s+1}) ) \right) d\mes(u), \nonumber \\
-    p(\x;\mathcal{Y}^{j,s+1}) &=& P(Z(\x) \in T |\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}). \nonumber
-\end{eqnarray}
-Here, $\mathcal{Y}^{j,s}=\{\by_{d^{j,s}},\mathcal{Y}^{s-1}\}$ and
-$\mathcal{Y}^{j,s+1}=\{\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}\}$
-represent the sets of data variables, and the expectations are
-calculated with respect to the conditional densities given data at
-that stage.  We solve the first expectation by Monte
-Carlo sampling of data $\by_{d^{j,s}}$ from its conditional
-distribution. For each of these data samples, the second expectation
-is solved using the closed-form expressions for EIBV using the results from Proposition \ref{propo2}, now with conditioning to the first stage data already going into the mean and covariances in Eq.
-\eqref{eq:krigSUR}.
-
-Even though the strategy looks at two sampling stages, it is only used to find
-the current best design. When data are collected at stage $s$, the GP model is
-updated as in Eq. \eqref{eq:krigSUR}, and the mean and (cross) covariance are
-used to compute the next design for stage $s+1$, now anticipating what
-stage $s+1$ and $s+2$ could offer, and so on.
-This look-ahead approach is much more computationally demanding than
-the myopic strategy, so for practical implementation onboard the AUV this is reaching the limits allowed for calculation.
-We conduct some pruning of the least valuable design paths in the evaluation of Eq.
-\eqref{critLA}, to get a satisfactory computation time below one minute per waypoint.
-This means that we use the myopic
-strategy to rank the three best designs following the first stage alone, and
-for each of these preferred designs, we undertake the look-ahead calculations.
-
-\section{Simulation study}
-\label{sec:simulations}
-
-We now study the properties of the different static and sequential
-survey designs in a realistic case that emulates the spatial
-properties of a river plume interacting with coastal water.
-
-\subsection{Modeling}
-
-A GP model is used to represent the bivariate random fields of
-temperature and salinity. At location $\x \in \mathcal{M}$, the
-bivariate distribution is:
-\begin{equation}\label{gp_bar}
-\begin{bmatrix}Z_1(\x) \\
-Z_2(\x) \end{bmatrix}
-\sim N \left( 
-\begin{bmatrix} \mu_{1}(\x)\\
-\mu_{2}(\x)
-\end{bmatrix},\begin{bmatrix}
-\sigma_{1}^2 & \sigma_{1} \sigma_2 \gamma \\
-\sigma_{1} \sigma_2 \gamma  & \sigma_{2}^2 
-\end{bmatrix}
-\right).
-\end{equation}
-We specify the mean by $\mu_{1}(\x)=5.8 + 0.065 u_{\mbox{West}}$ and
-$\mu_{2}(\x)=29 + 0.1 u_{\mbox{West}}$, which means that the expected
-levels of both temperature and salinity increase linearly in the
-western direction.  This situation mimics that of a river mouth
-opening from the south, with the water masses pulled to the east by
-the Coriolis force \kc{This seems overly specific to the Nidelva use
-  case. Can it be generalized, after all that is the overall intent of
-  this manuscript?}.  The variances are set to $\sigma_1=0.25$ and
-$\sigma_2=0.6$, while the correlation is $0.6$.  The spatial
-correlation is of separable Matern type (see e.g. \cite{Cressie:11})
-so that
-\begin{equation}\label{v}
-\small
-\mbox{Corr} 
-\left(
-\begin{bmatrix}
-    Z_1(\x) \\
-    Z_2(\x) 
-    \end{bmatrix},
-    \begin{bmatrix}
-    Z_1(\x') \\
-    Z_2(\x') 
-    \end{bmatrix}
-    \right)
-    = \begin{bmatrix}
-1 & 0.6  \\
-0.6  & 1
-\end{bmatrix}(1+\phi h)\exp (-\phi h),
-\end{equation}
-where $h=\sqrt{||\x-\x'||^2}$ and $\phi=0.3$ indicates an effective
-correlation range of about $1200$ m.
-
-In the accompanying Python code, these modeling assumptions can be
-generalized to anisotropic covariance and changing variance levels
-across the spatial domain. One can hence, easily run the code to see
-if different models result in other sampling designs. Anisotropy and
-non-stationary variance are both relevant for the setting with river
-plumes, but in practice this requires more parameters to be
-specified. With extensive data and prior knowledge, one could also
-possibly fit and estimate parameters of more complex multivariate
-spatial covariance functions
-\citep{gneiting2010matern,genton2015cross}, but that is outside the
-scope of the current paper.
-
-We discretize the spatial domain $\mathcal{M}$ to a set of $n$ grid
-locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,n \}$, \kc{is the index
-  correct? TeX shows it as u or g for $\mathcal{M}$} where each cell
-has area $\Delta$; the same grid is used for the waypoint graph for
-possible design locations. The EIBV is approximated by sums over all
-grid cells.
-
-One realization of the random fields representing salinity and
-temperature is shown in Fig. \ref{fig:true_temp} and
-\ref{fig:true_sal}. The true ES is a realization from this model as
-shown in Fig. \ref{fig:ESet}.
-
-\begin{figure}[!h]
-  \centering
-  \subfigure[Simulated temperature.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/true_temp.pdf}\label{fig:true_temp}}
-  \hfill
-  \subfigure[Simulated salinity.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/true_sal.pdf}\label{fig:true_sal}}
-  \hfill
-  \subfigure[Excursion set given the realization of temperature and salinity.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_true.pdf}\label{fig:ESet}}
-  \hfill
-  \subfigure[Excursion probabilities given collected data using the
-  `static\_north" strategy.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
-  \caption{\ref{fig:true_temp} and \ref{fig:true_sal} show a
-    realization of the simulated temperature and salinity fields. The
-    associated ES is in \ref{fig:ESet}, while \ref{fig:Eprob} shows
-    the estimated EPs after performing data collection along the
-    static N-S survey line.}
-\label{fig:realisations}
-\end{figure}
-
-\subsection{Static and Sequential Sampling Designs}
-\label{sec:sampling_designs}
-
-%Three different designs are considered as indicated in Figure \ref{fig:stat_design}. 
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.65\textwidth]{Figures/Des3.eps}
-%\caption{Three different static survey designs plotted on the initial EP.}\label{fig:stat_design}
-%\end{figure}
-%In this display the designs are plotted along with the prior probability contours of the ES for the reference parameter inputs.
-
-We compare three different static designs denoted
-\textit{static\_north}, \textit{static\_east}, and
-\textit{static\_zigzag} with the three described sequential approaches
-\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
-sampling paths are pre-determined and cannot be altered and represent
-the pre-planned strategies used in most current AUV operational survey
-designs.
-
-For a fixed survey length, a closed-form expression for the EIBV is
-available as in Eq. \eqref{eq:eibv}. However, for the sequential
-approaches this is not the case. For comparison the properties are
-therefore evaluated using Monte Carlo sampling over several replicates
-of realizations from the model while conducting simulated sequential
-surveys for each one. Fig. \ref{fig:Eprob} shows the conditional EP,
-given data gathered along the north-south survey lines in these
-displays. In the Monte Carlo replicates, such results are repeatedly
-computed to approximate the EIBV. We also compare predictive
-performance measured by root mean square error (RMSE) for temperature
-and salinity estimates as well as also the variance reduction in these
-two variables. It is important to note that the objective function
-used by the agent \kc{First of 'agent'. Define it or change it
-  to 'robot'? Note further uses below.} is focused on reducing the
-EIBV, but we nevertheless expect that we will achieve good predictive
-performance for criteria such as RMSE as well. Another non-statistical
-criteria that is important for practical purposes is the computational
-time needed for the strategy, as this will impact the performance for
-an embedded system.
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
-%\caption{The equilateral waypoint graph used to discretize the
-%  trajectory choices over the $31\times31$ grid used to discretize the GP.}
-%\label{fig:wp_graph}
-%\end{figure}
-
-\begin{figure}[!h] 
-\centering 
-\subfigure[The waypoint graph.]{\includegraphics[width =
-0.49\textwidth]{Figures/sim/wp_graph_paper.pdf}\label{fig:wp_graph_a}}
-\hfill
-\subfigure[The waypoint graph in 3D.]{\includegraphics[width =
-0.49\textwidth]{Figures/sim/wp_graph_3d.pdf}\label{fig:wp_graph_b}}
-\caption{\ref{fig:wp_graph_a} The equilateral waypoint graph used to discretize the
-trajectory choices over the $31\times31$ grid used to discretize the GP.
-\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment.}
-\label{fig:wp_graph}
-\end{figure}
-
-Each strategy is conducted on an equilateral grid as shown in
-Fig. \ref{fig:wp_graph}. The sequential sampling agent starts at the
-center East-West coordinate at the southern end of the domain (node 53
-\kc{Maybe color this node differently and modify the caption?}). The
-AUV moves along edges in the waypoint graph while the vehicle makes
-measurements. The data is assimilated into the GP model before an
-evaluation of the next node to sample is conducted at the end of the
-edge.
-
-\subsection{Simulation Results}
-
-A total of 100 replicate simulations were conducted with all
-strategies. The results are shown in Fig. \ref{fig:sim_results}, where
-the different criteria are plotted as a function of survey
-distance. Fig. \ref{fig:avg_ev} shows the resulting drop in IBV for
-each of the six strategies. IBV reduction occurs most under the
-\textit{myopic} and \textit{look-ahead} strategies, each performing
-almost equally; this is expected as the two criteria
-(Eq. \eqref{critSEQ} and \eqref{critLA}) are sensitive to differences
-in IBV. The \textit{static\_north} design also does well here because
-the path is parallel to the boundary between the water masses.
-
-\begin{figure}[h!]
-  \centering
-  % \subfigure[Excursion set variance $E_{\by}(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
-  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
-  \hfill
-  \subfigure[RMSE between estimated field and truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
-  \hfill 
-  \subfigure[Explained variance $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
-  \hfill 
-  \subfigure[Computational time for inferencing.\kc{I see only 
-    lines associated with 4 variables showing in the graph. Where is
-    static\_north and static\_east?}]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}} 
-\caption{Simulation results from 100 replicate simulations for 10
-  sampling choices/stages on the grid. \kc{Vertical lines show large
-    variation in replicate results.}}  
-\label{fig:sim_results}
-\end{figure}
-
-Fig. \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop
-in RMSE and increase in explained variance, respectively. Both
-\textit{myopic} and \textit{look-ahead} strategies perform well here,
-but some of the \textit{static\_east} and \textit{static\_zigzag} also
-achieve good results because they are pre-determined to cover large
-parts of the domain without re-visitation. Sequential strategies
-targeting IBV will sometimes not reach similar coverage, as
-interesting data may draw the path into twists and turns. There is
-relatively large variety in the replicate results as indicated by the
-vertical lines. Nevertheless, the ordering of strategies is similar.
-
-\begin{figure}[!b]
-  \centering
-  \subfigure[Look-ahead]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_look-ahead.pdf}\label{fig:avg_look-ahead}}
-  \hfill
-  \subfigure[Myopic]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_myopic.pdf}\label{fig:avg_myopic}}
-  \hfill
-  \subfigure[Naive]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_naive.pdf}\label{fig:route_naive}}
-  \hfill
-  \subfigure[Static]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_static.pdf}\label{fig:route_static}}
-  \hfill
-  \caption{The overlaid route choices (strategies), superimposed on
-    the 100 replicate surveys with 10 sampling choices/stages. \kc{I
-      would reorder these figurs above based on the order the material
-      was discussed -- Static, naive, myopic and then Look-ahead. Also
-      do you really need a key associated with each subfigure, given
-      the caption? An abridged analysis as in the text, might also be
-      in order here in the caption. Also refer to
-      Fig. \ref{fig:realisations} as the reference, where the
-      temp/salinity fields are shown.}}
-\label{fig:route_choices}
-\end{figure}
-
-Fig. \ref{fig:avg_time} shows the computational effort: the
-\textit{naive} strategy is on par with the static designs, while the
-\textit{myopic} strategy is slower. The \textit{look-ahead} is even
-slower, reaching levels that are nearly impractical for execution on a
-vehicle. Some pruning of the graph is performed to improve the
-performance, such as ruling out repeated visitations and
-back-and-forth routes. Some of the intermediate results are also
-stored for longer planning horizons. Further pruning of branches or
-inclusion of other heuristics could be included for better
-performance. Then again, the inclusion of heuristics is likely a
-contributing factor for the \textit{look-ahead} strategy failing to
-outperform the \textit{myopic} strategy.
-
-In Fig. \ref{fig:route_choices}, the realized sampling paths for each
-of the sequential schemes and static designs are shown. The
-\textit{naive} strategy often gets stuck in the southern part of the
-domain because it is too focused on the probabilities near $0.5$. The
-\textit{myopic} strategy covers a wider domain than the naive or
-look-ahead. There are several reasons for this. First, a greedy
-approach will tend to put more emphasis on promising locations close
-to the agent, which may lead away from the centre. Second, as the
-agent evaluates the impact of locations further away (look-ahead)
-where assimilated data has less predictive power, the GP model (which
-is centered here) will act to restrict paths deviating from the
-central zone.
-
-We studied the sensitivity of the results by modifying the input
-parameters to have different correlations between temperature and
-salinity, standard deviations, and spatial correlation range.  In all
-runs, the \textit{myopic} and \textit{look-ahead} strategies perform
-the best in terms of realized IBV, and much better than
-\textit{naive}. The \textit{look-ahead} strategy seems to be
-substantially better than the \textit{myopic} design only for very
-small initial standard deviation or very large spatial correlation
-range. \textit{static\_north} continues to be the best static design
-for IBV, while \textit{static\_zigzag} is the best design for the
-other predictive performance measures, especially so with large
-spatial correlation range. We also ran simulation studies with only
-temperature data, and for realistic correlation levels between
-temperature and salinity, the IBV results are not much worse when only
-temperature data are available. In addition to the comparison made in
-Table \ref{tab:sim_rhoab}, the current setting includes spatial
-correlation and this likely reduces the additional influence of having
-bivariate data. However, it seems that having temperature data alone
-does a substantially worse job in terms of explained variance
-
-\begin{comment}
-\subsection{Performance Metrics}
-
-The tables below explore the sensitivity of the result to different input correlation values between temperature and salinity measured onboard the AUV, standard deviations, and spatial
-correlations (see table rows). We only compare results at the end of
-each survey. The expected ES variances are shown in Table
-\ref{tab:sim_res_ev}, results of RMSE are in Table
-\ref{tab:sim_res_rmse}, and explained variance is in Table
-\ref{tab:sim_res_r2}. In all runs, both the \textit{myopic} and
-\textit{look-ahead} strategies perform the best in terms of expected
-variance in the ES. The \textit{myopic} strategy
-performs slightly better than the \textit{look-ahead}. Part of the
-reason for this discrepancy can be traced to the act of pruning some
-of the paths for the \textit{look-ahead} strategy. The \textit{static\_north} design is clearly
-better than the other two static designs, the reason being that the
-others do not focus on the areas where the ES is really uncertain
-(centre locations are the most uncertain). By measuring temperature
-alone, the expected variances in the ES are only slightly higher than
-the combined temperature-salinity estimate. This occurs because the
-correlation of $0.6$ means that temperature observations provide
-information about salinity as well. For the other predictive measures,
-\textit{myopic} and \textit{look-ahead} strategies do well, clearly
-beating the \textit{naive} strategy.
-
-\begin{table}[!h]
-    \centering
-    \scalebox{0.87}{
-    \begin{tabular}{lrrrrrr}
-    \toprule
-    Parameter: $E_{\by}(p[1-p])$ &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
-    \midrule
-                    \rowcolor{Gray}
-ts. cor. low: 0.2  &  29.65 &   27.08 &       \textbf{26.44} &         32.33 &        40.31 &          36.23 \\
-ts. cor. high: 0.8 &  36.18 &   \textbf{28.42} &       30.30 &         32.99 &        37.18 &          36.04 \\
-                    \rowcolor{Gray}
-std. low: 0.1      &  18.26 &   15.79 &       \textbf{15.35} &         18.89 &        26.65 &          26.10 \\
-std. high: 0.5     &  47.01 &   \textbf{41.31} &       43.15 &         47.92 &        52.63 &          49.71 \\
-                    \rowcolor{Gray}
-cor. low: 0.8      &  44.42 &   \textbf{42.02} &       42.47 &         43.84 &        46.05 &          47.82 \\
-cor. high: 0.2     &  29.73 &   21.09 &       \textbf{20.43} &         27.91 &        37.60 &          34.35 \\
-                    \rowcolor{Gray}
-temp. only         &  38.25 &   29.16 &       \textbf{28.69} &         34.32 &        39.81 &          36.17 \\
-basecase           &  35.21 &   \textbf{28.56} &       28.94 &         33.50 &        39.84 &          39.26 \\
-    \bottomrule
-    \end{tabular}}
-  \caption{Simulation results for the final mean IBV (Eq. \eqref{two_partsK}).}
-    \label{tab:sim_res_ev}
-\end{table}
-\vspace{-0.4cm}
-\begin{table}[!h]
-    \centering
-    \scalebox{0.92}{
-        \begin{tabular}{lrrrrrr}
-        \toprule
-        Parameter: RMSE &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
-        \midrule
-                \rowcolor{Gray}
-ts. cor. low: 0.2  &   0.63 &    \textbf{0.57} &        0.59 &          0.62 &         0.57 &           0.59 \\
-ts. cor. high: 0.8 &   0.63 &    \textbf{0.59} &        0.61 &          0.64 &         0.60 &           0.60 \\
-                \rowcolor{Gray}
-std. low: 0.1      &   0.39 &    0.37 &        0.38 &          0.38 &         0.40 &           \textbf{0.35} \\
-std. high: 0.5     &   0.85 &    0.82 &        0.84 &          0.82 &         0.83 &           \textbf{0.81} \\  
-                \rowcolor{Gray}
-cor. low: 0.8      &   0.68 &    0.66 &        0.67 &          0.68 &         0.66 &           \textbf{0.65} \\
-cor. high: 0.2     &   0.58 &    0.54 &        0.55 &          0.58 &         0.54 &           \textbf{0.49} \\
-                \rowcolor{Gray}
-temp. only         &   0.65 &    0.63 &        0.62 &          0.65 &         \textbf{0.61} &           0.62 \\
-basecase           &  0.61 &   \textbf{0.56} &       \textbf{0.56} &         0.60 &        0.57 &          0.57 \\
-        \bottomrule
-        \end{tabular}}
-      \caption{Simulation results for the final mean RMSE ($\frac{1}{B} \sum_{b=1}^{B} \sqrt{(\bmu-\tilde{\bmu})^2}$).}
-    \label{tab:sim_res_rmse}
-\end{table}
-\vspace{-0.4cm}
-\begin{table}[!h]
-    \centering
-    \scalebox{0.95}{
-        \begin{tabular}{lrrrrrr}
-        \toprule
-        Parameter: $\bR^{2}$ &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
-        \midrule
-                        \rowcolor{Gray}
-ts. cor. low: 0.2  &  27.00 &   33.29 &       32.24 &         26.71 &        30.56 &          \textbf{33.50} \\
-ts. cor. high: 0.8 &  23.68 &   \textbf{33.87} &       32.42 &         26.72 &        30.73 &          33.61 \\
-                        \rowcolor{Gray}
-std. low: 0.1      &  25.38 &   30.80 &       30.29 &         26.09 &        28.75 &          \textbf{31.66} \\
-std. high: 0.5     &  26.11 &   34.22 &       32.48 &         26.95 &        31.77 &          \textbf{34.48} \\
-                        \rowcolor{Gray}
-cor. low: 0.8      &   8.80 &   11.50 &       11.23 &          9.56 &        12.31 &          \textbf{12.58} \\
-cor. high: 0.2     &  37.57 &   \textbf{48.68} &       48.11 &         39.61 &        41.98 &          48.04 \\
-                        \rowcolor{Gray}
-temp. only         &  11.67 &   \textbf{22.82} &       21.98 &         18.16 &        20.77 &          22.78 \\
-basecase           &  24.71 &   33.45 &       32.25 &         26.71 &        30.62 &          \textbf{33.54}\\
-        \bottomrule
-        \end{tabular}}
-      \caption{Simulation results for the final mean explained variance $\bR^{2}=100*(1-(\Sigma_{posterior}/\Sigma_{initial}))$).}
-    \label{tab:sim_res_r2}
-\end{table}
-
-\end{comment}
-
-\section{Case Study - Mapping a River Plume}
-\label{sec:case_study}
-
-To demonstrate the applicability of using multivariate EPs and the IBV
-to inform oceanographic sampling, we present a case study mapping a
-river plume with an AUV. The experiment was performed in Trondheim,
-Norway, surveying the Nidelva river (Fig. \ref{fig:nidelven}). The
-experiments were conducted in late Spring 2019, when there is still
-snow melting in the surrounding mountains so that the river water is
-substantially colder than the water in the fjord. The experiment was
-focused along the frontal zone that runs more or less parallel to the
-eastern shore as noted in Fig. \ref{fig:nidelven}.
-
-\subsection{Model Specification}
-\label{sec:exp_modeling}
-
-The statistical model parameters were specified based on a short
-preliminary survey where the AUV made an initial transect to determine
-the trends in environmental conditions and correlation
-structures. Based on the initial data, the trend parameters were
-estimated by linear regression, where both temperature and salinity
-are assumed to increase linearly, going west from the river
-mouth. Next, the residuals from the regression analysis were analyzed
-to study the fit of the GP model and to specify the covariance
-parameters.
-
-%\begin{figure}[!h] 
-% \centering 
-%\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
-%\caption{Data analysis from a preliminary trial experiment using the
-%  AUV. Left: Residual plot of temperature and salinity along with
-%  Gaussian contours. Middle: Empirical CDF (solid) of the quadratic form of
-%  the residuals along with the theoretical CDF (dashed) of the $\chi^2$
-%  distribution with two degrees of freedom. Right: Empirical variogram
-%  of the salinity and temperature data.} \label{fig:parest}
-%\end{figure}
-
-\begin{figure}[!h]
-  \centering
-  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_a.pdf}\label{fig:parest_a}}
-  \hfill
-  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_b.pdf}\label{fig:parest_b}}
-  \hfill
-  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_c.pdf}\label{fig:parest_c}}
-  \caption{Data analysis from a preliminary trial experiment using the
-    AUV. \ref{fig:parest_a} Residual plot of temperature and salinity
-    along with Gaussian contours. \ref{fig:parest_b} Empirical CDF
-    (solid) of the quadratic form of the residuals along with the
-    theoretical CDF (dashed) of the $\chi^2$ distribution with two
-    degrees of freedom. \ref{fig:parest_c} Empirical variogram of the
-    salinity and temperature data.}
-\label{fig:parest}
-\end{figure}
-
-Fig. \ref{fig:parest} summarizes diagnostic plots of this
-analysis. Fig. \ref{fig:parest_a} shows a cross-plot of temperature
-and salinity residuals after the westerly trends in both salinity and
-temperature are subtracted from the data. This scatter-plot of joint
-residuals indicates larger variability in salinity than in
-temperature, and a positive correlation ($0.5$) between the two
-variables. Based on the fitted bivariate Gaussian model (ellipses in
-Fig. \ref{fig:parest_a}), we can compute the modeled quadratic form of
-the residuals, and if the model is adequate they should be
-approximately $\chi^2_2$ distributed. Fig. \ref{fig:parest_b} shows
-the empirical cumulative distribution function (CDF) of these
-quadratic forms (solid) together with the theoretical CDF of the
-$\chi^2_2$ distribution. The modeled and theoretical curves are very
-similar, which indicates that the Gaussian model fits reasonably
-well. Even though there appears to be some clustering in both
-Fig. \ref{fig:parest_a} and \ref{fig:parest_b}, the bivariate
-diagnostic plots look reasonable and justify a Gaussian
-model. Fig. \ref{fig:parest_c} shows the empirical variogram of the
-scaled residuals for temperature and salinity. The decay is similar
-for the two, and seems to be negligible after about $150$ m.
-
-Based on the analysis in Fig. \ref{fig:parest}, the resulting
-parameters are given in Table \ref{tab:experiment_param}. The
-regression parameters shown here are scaled to represent the east and
-west boundaries of the domain as seen in the preliminary transect
-data, and the thresholds are intermediate values. These parameter
-values were then used in field trials where we explored the
-algorithm's ability to characterize the river plume front separating
-the river and fjord water masses, providing a spatial map of this
-boundary.
-
-%Mapping the spatial extent of a frontal zones is an important problem for studying many bio-physical interactions in the ocean.
-The frontal zone is determined by the boundary where plumes of sediments, nutrients, and possibly pollutants spreading from the river outlet meet and interact with adjacent coastal water.
-Due to the lower density the plumes spread on the surface, creating a front with an sharp gradient in both temperature and salinity.
-
-\begin{table}[!h]
-\centering
-\begin{tabular}{lrr}
-\toprule
-Parameter & Value & Source\\
-\midrule
-\rowcolor{Gray}
-Cross correlation temp. and sal. & 0.5 & AUV observations\\
-Temp. variance &  0.20 & AUV observations (variogram)\\
-\rowcolor{Gray}
-Sal. variance &  5.76 & AUV observations (variogram)\\
-Corr. range  & 0.15 km & AUV observations (variogram)\\
-\rowcolor{Gray}
-River temp. $T_{river}$ & $10.0\,^{\circ}\mathrm{C}$ & AUV observations\\
-Ocean temp. $T_{ocean}$ & $11.0\,^{\circ}\mathrm{C}$ & AUV observations\\
-\rowcolor{Gray}
-River sal. $S_{river}$ & $14.0$ g/kg & AUV observations\\
-Ocean sal. $S_{ocean}$ & $22.0$ g/kg & AUV observations\\
-\rowcolor{Gray}
-Threshold temp. & $10.5\,^{\circ}\mathrm{C}$ & $(T_{ocean}-T_{river})/2+T_{river}$\\
-Threshold sal. & $18.0$ g/kg & $(S_{ocean}-S_{river})/2+S_{river}$\\
-\rowcolor{Gray}
-\bottomrule
-\end{tabular}
-\caption{Model and threshold parameters from an initial AUV
-  survey. Observations were taken across the front while crossing from
-  fresh, cold river water to saline and warmer ocean waters. \kc{Might
-  be good to expand the terms and not have abbreviations in the table
-  above; there's plenty of space.}}
-\label{tab:experiment_param}
-\end{table}
-
-\subsection{Experimental Setup}
-
-The sampling locations were distributed over an equilateral grid, as
-shown in the grey-colored lattice in Fig. \ref{fig:map}. The robotic
-platform consisted of a Light AUV \citep{sousa2012lauv}
-(Fig. \ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49
-conductivity, temperature, and depth (CTD) sensor providing
-temperature and salinity
-measurements. %The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
-The sampling agent was built on top of the autonomous agent framework
-Teleo-Reactive EXecutive (\textit{T-REX})
-\citep{py10,Rajan12,Rajan12b}, running an instance of the
-\textit{myopic} strategy from Section \ref{sec:myopic} to control the
-AUV and decide between sampling locations.
-
-\begin{figure}[!h] 
-\centering 
-\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
-\caption{The commercially available Light Autonomous Underwater
-  Vehicle (LAUV) platform for upper water-column exploration used in
-  our experiments.}
-\label{fig:lauv}
-\end{figure}
-
-The sampling strategy was designed around the concept of visiting
-waypoints sequentially. Arriving at a desired waypoint with new
-measurements and an updated model, the AUV triggers the myopic
-strategy to evaluate the different design criteria (see
-Eq. \eqref{critSEQ}). The waypoint-and-path combination that is
-expected to reduce the IBV the most is selected, and upon arrival this
-procedure is then repeated. At each stage, it takes the AUV about 30
-seconds to evaluate the EIBV for all the possible waypoint-and-path
-alternatives.
-
-The AUV was set to start in the south-center part of the waypoint
-graph, with the previously outlined GP model of the environment
-(Section \ref{sec:exp_modeling}). A survey was set to take
-approximately 40 minutes, visiting 15 waypoints on the grid, with the
-AUV running near the surface to capture the plume. On its path from
-one waypoint to the next, the AUV gathered data regularly, and the GP
-model assimilated temperature and salinity data with an update
-frequency of 30 seconds, giving about three updates per stage.
-
-\begin{figure*}[!h]
-\centering
-\subfigure[AUV survey area]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
-\hspace{0.3cm}
-\subfigure[Temperature tracks]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/auv.pdf}\label{fig:res_both}}
-
-\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv1_es_sal_ep.pdf}\label{fig:res1}}
-\hspace{0.2cm}
-\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv4_es_sal_ep.pdf}\label{fig:res2}}
-
-%\subfigure[ES for Survey 1]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
-%\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
-\caption{Results from mapping the Nidelva river, Trondheim, Norway
-  over two survey missions. \ref{fig:map} shows an overview of the
-  survey area overlaid with the AUV path in black and dashed
-  line. Note the shaded region indicating a typical frontal
-  region. \ref{fig:res_both} shows the collected temperature data as
-  colored trails. Note waypoint 5 (WP5) which indicates where the two
-  surveys diverge. \ref{fig:res1} and \ref{fig:res2} shows the
-  collected salinity data overlaid on the final EP, which indicate the
-  AUVs statistical impression of the front. For both missions the
-  temperature and salinity data correspond with an indication of the
-  EP front. \kc{Might want to indicate when the surveys (dates) were
-    conducted so people can understand the variability in the EP with
-    respect to time.}}
-\label{fig:results}
-\end{figure*}
-
-\subsection{Results}
-
-Two surveys missions (1 and 2), were run successively from 11:00 AM to
-01:00 PM, with a short break in between. The resulting path of the
-selected waypoints are shown in the map in Fig. \ref{fig:map}, both
-within the expected frontal region (shaded pink). The recorded
-temperatures are shown as colored trails in Fig. \ref{fig:res_both},
-clearly indicating the temperature difference between fjord and
-riverine waters. The salinity data are then shown separately, overlaid
-with the estimated EP for each survey in Fig. \ref{fig:res1} and
-\ref{fig:res2}.
-
-Both surveys successfully estimated and navigated the separation zone,
-crossing the frontal boundary multiple times. As conditions changed
-slightly between the two surveys, the resulting path (after waypoint
-5) is shown to deviate. Survey 1 continued northwards, tracking the
-north-eastern portion of the front, while Survey 2 turned west,
-mapping the south-western region.
-
-The final predictions of the front location, represented by
-conditional EPs in Fig. \ref{fig:res1} and \ref{fig:res2} as dashed
-lines, correspond with one another. In both surveys they yield a
-picture of the front being to the west in the southern portions of the
-region and gradually bending off toward the north east. The amount of
-exploration done by Survey 1 is greater than Survey 2. In Survey 1,
-the AUV obtained more detail by going north from waypoint 5, while
-Survey 2, coming close to the survey area borders in the south-western
-corner, obtained a poorer understanding of the northern parts. A
-look-ahead strategy might identify and discourage such choices.
-
-%\begin{figure}[!h] 
-% \centering 
-%\includegraphics[width=0.48\textwidth]{Figures/envir_ocean.pdf}
-%\caption{Ocean observation is moving away from %single-ship sampling
-%towards more collaborative networked operations in %order to resolve
-%the numerous processes and their interaction.} %\label{fig:envir}
-%\end{figure}
-%\newpage
-\section{Closing remarks}\label{sec:concl_disc}
-
-This work builds on a multidisciplinary effort combining statistical
-modeling with robotic surveying techniques for oceanographic
-applications. We show how observation practices can gain efficiency
-and accuracy from the development of statistical techniques to achieve
-more effective spatial monitoring. We further demonstrate the
-opportunities available for real-time multivariable spatial
-data-gathering and analysis onboard autonomous platforms, which
-statisticians can exploit to create general-purpose toolkits for
-similar applications.
-
-In particular, we derive and show results for characterizing phenomena
-connected to the properties of water masses. The characterization of
-uncertainties in random sets is extended with new results for the EIBV
-uncertainty reduction achieved by spatial sampling design. This is
-provided in closed-form for the situation with a static design, and
-then extended to the adaptive situation. The sequential derivations
-provide new insights into efficient applications of adaptive sampling,
-as demonstrated in our application.
-
-The study did not consider any temporal effects, which would be
-relevant on a larger time scale. We consider the extension to
-spatio-temporal modeling as future work, and envision that
-advection-diffusion equations could be useful in this kind of modeling
-\citep{sigrist2015stochastic}. For more complex oceanographic
-phenomena, the methods will need to be extended to non-Gaussian
-phenomena, possibly feature-based mixtures of GPs which could still be
-run onboard and augmented by dynamical models. \kc{How about talking
-  about removing any grids or waypoints altogether?}
-
-The spatial-statistical design criterion, building on random sets is
-relevant in our setting with different water properties, and we show
-mathematical generality beyond the EIBV which is convenient to
-implement for onboard robotic planning. Such criteria could also be
-useful in other oceanographic settings as for algal-blooms, anoxic
-zones or open water fronts \cite{costa19}. Of course, other criteria
-are also relevant for instance, hybrid or multi-attribute criteria
-that could balance goals of exploration and exploitation in this
-situation. Equally, such techniques have significant use cases in
-downstream decision-making, with policy makers and regulators who need
-to make difficult decisions related to fish farming, other marine
-resource operations. Value of information analysis (VOI)
-\citep{Eidsvik:15} could be used in a similar vein as the IBV in the
-current work, as both analytic methods explore when, where and which
-information is likely to result in improved decision-making.
-
-While more effort could be spent on approximating the look-ahead
-stages in the adaptive designs, it is perhaps more interesting to
-explore the additional flexibility that can be gained by having
-multiple AUVs co-temporally exploring a spatial or spatio-temporal
-domain \citep{ferreira2019advancing}. Such an approach would enable
-concurrent sampling in different parts of the space, or opportunities
-to move in parallel to best capture the excursion set.
-
-\section*{Acknowledgements}
-
-TOF acknowledges support from the Centre for Autonomous Marine
-Operations and Systems
-(AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of
-Excellence, project number 223254, and the Applied Underwater Robotics
-Labortatory (AURLab). JE and KR acknowledge support from Norwegian
-research council (RCN), project number 305445. CT and DG acknowledge
-support from the Swiss National Science Foundation, project number
-178858.
-
-%\begin{supplement}
-%\sname{Supplement A}\label{suppA}
-%\stitle{Title of the Supplement A}
-%\slink[url]{http://www.e-publications.org/ims/support/dowload/imsart-ims.zip}
-%\sdescription{Dum esset rex in
-%accubitu suo, nardus mea dedit odorem suavitatis. Quoniam confortavit
-%seras portarum tuarum, benedixit filiis tuis in te. Qui posuit fines tuos}
-%\end{supplement}
-
-% == Adding references
-\footnotesize
-\bibliographystyle{imsart-nameyear}
-\bibliography{ref}
-
-\section*{Appendix}
-
-Proof of Proposition \ref{propo1}:
-
 \begin{proof}
@@
 
-Proof of Proposition \ref{propo2}:
+The propositions below provide formulae for computations of expectations of moments of multivariate gaussian CDFs.
+
+\begin{propo}
+    \label{propo2}
+Let $p, q, h \geq 1$, $a \in \R^p$, $B \in \R^{p\times q}$,
+and $\covN, \covV$ be two covariance matrices in
+$\R^{p\times p}$ and $\R^{q\times q}$, respectively.
+Then, for $V \sim \mathcal{N}_{q}(0_q, \covV)$,
+\begin{equation*}
+\mathbb{E}\left[ \varPhi_{p}\left( a + BV; \covN \right)^h \right]
+=
+\varPhi_{ph}
+\left(
+    \bm{a}
+;~
+\bm{\Sigma}
+\right),
+\end{equation*}
+where the vector $\bm{a} \in \R^{p h}$ is defined as
+$\bm{a} := 1_h\otimes a = 
+\left(a, \dots , a
+\right)'$
+ and the $p h\times p h$ covariance matrix is given by
+ $\bm{\Sigma} := 
+1_h 1_h'\otimes B\covV B' + I_h\otimes \covN$.
+\end{propo}
+
+\begin{remark}
+In blockwise representation, $\bm{\Sigma}$ can be expressed as follows:
+\begin{align*}
+% \bm{\Sigma}&=
+\begin{pmatrix}
+    \covN & &\\
+        & \ddots &\\
+        &   & \covN
+\end{pmatrix}
++
+\begin{pmatrix}
+B\covV B' & \dots & B\covV  B'\\
+\vdots & & \vdots\\
+B\covV B' & \dots & B\covV B'\\
+\end{pmatrix}
+\end{align*}
+\end{remark}
 
@@
 
-Proof of Proposition \ref{propo3}.
+We now generalize Proposition~\ref{propo2} to the case of multivariate monomials in orthant probabilities with thresholds affine in a common Gaussian vector.
+%Namely, for monomials of degree $k$, we have.
+
+\begin{propo}
+    \label{propo3}
+Let $g, p, q\geq 1$, $h_{1},\dots, h_{g}\geq 1$ with $H=\sum_{i=1}^g h_i$, $a_{i} \in \R^{p}$, $B_{i}\in \R^{p \times q}$, and covariance matrices $\covN_i \in \R^{p \times p}$ $(1\leq i \leq g)$.
+Then, for any covariance matrix $\covV \in \R^{q\times q}$ and $V\sim\mathcal{N}_{q}(0_q,\covV)$,
+    \begin{equation}
+    \mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
+    =
+\varPhi_{p H}
+\left(
+    \bm{a}
+;
+\mathbf{\Sigma}
+\right),
+\end{equation}
+with $\bm{a}=(1_{h_1}\otimes a_1, \dots, 1_{h_g}\otimes a_{g}) \in \R^{p H}$
+and $\mathbf{\Sigma}\in \R^{p H \times p H}$ is defined blockwise by $(\Sigma_{i,j})_{i,j \in \{1,\dots, g\}}$ where, for any $i,j \in \{1,\dots, g\}$, %the blocks are given by 
+\begin{equation}
+\Sigma_{i,j}=
+(1_{h_{i}}1_{h_{j}}')\otimes (B_{i}\Sigma_{V} B_{j}') + \delta_{i,j}(I_{h_{i}}\otimes \covN_{i}) \in \R^{p h_{i} \times p h_{j}}.
+\end{equation}
+\end{propo}
 
 \begin{remark}
-	%    Note that the vector $\bm{a}$ is obtained by stacking $h_i$ times each $a_i$
-	%\begin{equation*}
-	%\bm{a} = \begin{pmatrix}
-	%    \bm{a}_1\\ \vdots\\ \bm{a}_g\end{pmatrix},~
-	%\bm{a}_i =
-	%\begin{pmatrix}
-	%    a_i\\ \vdots\\ a_i\end{pmatrix}
-	%\in\mathbb{R}^{l\times h_i}\end{equation*}
-	%and the covariance matrix is obtained by similar stacking
 Using blockwise representation for the blocks themselves delivers %$\bm{\Sigma}$ can be expressed as follows
@@
 \end{remark}
-
-%Proof TODO along with notation polish.
 
@@
 
-% AOS,AOAS: If there are supplements please fill:
-%\begin{supplement}[id=suppA]
-%  \sname{Supplement A}
-%  \stitle{Title}
-%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
-%  \sdatatype{.pdf}" 
-%  \sdescription{Some text}
-%\end{supplement}
+Those two general results allow us to derive simple expressions for the expected effect of the inclusion of new datapoints on the $\ibv$ (Proposition \ref{propo_eibv}) and on the $\emv$ (Proposition \ref{propo_emv}) for which we provide proofs below.
 
-% === Not used
-% After reviews: - This is moved to Section 2.1.
+\begin{proof}{(Proposition \ref{propo_eibv})}
+Applying Tonelli-Fubini followed by the law of total
+expectation first delivers
+\begin{equation*}
+\begin{split}
+\eibv_{[n]}(\bm{x})
+&=\int_{D}
+\currentExp{\futureProba{\gp[\uu]\in
+        T}(1-\futureProba{\gp[\uu]\in T})} d\mes(u) \\
+&=\int_{D} \varPhi_{\no}\left(\bt;
+~\futureMean{\uu},
+\futureCov{u,u}\right) d\mes(u)\\
+&-\int_{D} \currentExp{
+    \varPhi_{\no}\left(\bt;
+    ~\futureMean{\uu},
+    \futureCov{u,u}\right)^2
+}
+d\mes(u), 
+\end{split}
+\end{equation*}
+%
+where $\futureCov{u,u}$ denotes the $\no \times \no$ covariance matrix between all $\no$
+responses at point $u$ conditional on the first $n+1$ observation batches.
+Now, by using co-kriging update formulae and our shortcut notation for the CDF of centred
+multivariate Gaussian vectors, we observe that
+\begin{equation*}
+\begin{split}
+%\mathbb{E}_{n}[\left(
+&\varPhi_{\no}\left(\bt;~\futureMean{\uu}, \futureCov{u, u}\right) 
+%\right)^2]
+\\
+=&
+\varPhi_{\no}\left(\bt-\futureMean{\uu}; \futureCov{u, u}\right) \\
+=&
+\varPhi_{\no}\left(\bt-\currentMean{\uu}-\lambda_{[n+1,n+1]}(u)^T(\gp[\bm{x}_{n+1}]
+-\currentMean{\bm{x}_{n+1}}), \futureCov{u, u}\right) \\
+=&
+\varPhi_{\no}\left(a + BV, \futureCov{u, u}\right),
+\end{split}
+\end{equation*}
+with $a=\bt-\currentMean{\uu}$, %+\lambda_{[n+1,n+1]}(u)^T\currentMean{\bm{x}_{n+1}}$,
+$B=-\lambda_{[n+1,n+1]}(u)^T$
+and $V=\gp[\bm{x}_{n+1}]-\currentMean{\bm{x}_{n+1}}$.
+Applying Proposition~\ref{propo2} then delivers that
+%
+\begin{equation*}
+\begin{split}
+&\currentExp{
+    \varPhi_{\no}\left(\bt;~\futureMean{\uu}, \futureCov{u, u}\right)^2 
+}
+%\\
+=\varPhi_{2\no}
+\left(
+\left(
+\begin{matrix}
+\bt-\currentMean{\uu}\\
+%\vdots \\ 
+%\vdots \\ 
+%\vdots \\ 
+\bt-\currentMean{\uu}
+\end{matrix}
+\right);
+\mathbf{\Sigma}_{[n]}(\uu)
+\right),
+\end{split}
+\end{equation*}
+with $\mathbf{\Sigma}_{[n]}(\uu)$ as in the formulation of the proposition. This completes the proof.
+\end{proof}
 
-%Traditional data collection at sea has typically been based on static
-%buoys, Lagrangian floats, or ship-based methods, with significant
-%logistical limitations that directly impact coverage and sampling
-%resolution. Modern methods using satellite remote-sensing provide
-%large-scale coverage but have limited resolution, are limited to
-%sensing the surface of the ocean, and are impacted by cloud cover. The
-%advent of robust mobile robotic platforms \citep{Bellingham07} has
-%resulted in significant contributions to environmental monitoring and
-%sampling. In particular, autonomous underwater vehicles (AUVs), have
-%advanced the state of sampling and consequently have made robotics an
-%integral part of ocean observation; our previous work has contributed
-%to this effort \citep{das11b,Das2015,fossum18b,fossuminformation}. Other ¤ %statistical work in the oceanographic domain include \cite{wikle2013modern}
-%focusing on hierarchical statistical models; \cite{sahu2008space},
-%studying spatio-temporal models for sea surface temperature and
-%salinity data; and \cite{mellucci2018oceanic} looking at the
-%statistical prediction of features using an underwater glider.
-
-%\begin{itemize}
+\begin{proof}{(Proposition \ref{propo_emv})}
+\begin{equation*}
+\begin{split}
+\eemv_{[n]}(\bm{x})
+%\operatorname{Var}[\mes(\es)]
+&=\int_{D^2} 
+\varPhi_{2\no}
+\left(
+(\bt, \bt); \mu((u,v)), 
+K((u,v),(u,v))
+\right) 
+\
+\mathrm{d}\mes^{\otimes} %\mes 
+%\productMeasure 
+(u,v)\\
+&-\currentExp{\left( \int_{D} \varPhi_{\no}\left(\bt;\mu_{[n+1]}(u), K_{[n+1]}(u)\right) d\mes(u) \right)^2}.
+\end{split}
+\end{equation*}
+The second term can be worked out in semi-analytical form by writing it as
 %
-%\item Full numerical ocean models cannot provide accurate results
-%  online if run on robotic sensing platforms, as onboard computers
-%  cannot deliver the computational power required. Hence statistical
-%  proxy models of the environment must be used for learning where to sample.
-%
-%\item With limited available information about the state of the ocean, there is substantial value in reacting to
-%  information obtained from measurements taken in-situ. This
-%  acquired information must be assimilated into statistical models that
-%  can be used to inform decisions on where to sample
-%  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.
-%
-%\item For sampling problems related to environmental sensing, the
-%  number of choices (i.e. locations, trajectories, and candidate
-%  designs) is enormous, creating a
-%  trade-off between optimization (finding the most resource-efficient
-%  design to collect necessary data) and computability (arriving at a
-%  solution in reasonable time). To successfully resolve features, this
-%  trade-off has to be considered in development and practice.
-%
-%\end{itemize}
-
-%Addressing this, the combination of statistical tools and robotic platforms is a
-%natural symbiosis which enables information-based sensing. Central to
-%this is the ability to model spatially-correlated variables and
-%provide formal measures of uncertainty. Our formulation is based on
-%Gaussian Processes (GPs) as they allow efficient implementation 
-%and evaluation in real time onboard a robotic platform.
-
-%Sampling can, in this context, not simply be distributed evenly ---along simple transects or ``lawn-mover" patterns--- but must instead be prioritized to relevant regions to ensure it is cost-effective while providing adequate coverage and resolution of the area of scientific interest.
-
-%While the focus has often been on
-%biological and anthropogenic impact from micro-plastics to
-%pollution, biological oceanographers have focused intently on studying
-%micro-organisms at the base of the human food web. These organisms are
-%critically impacted by the changing dynamics in the upper water-column,
-%especially in coastal zones which are complex and often hard to observe
-%in space and time. By studying the bio-geochemical processes in the
-%upper water-column scientists can measure the impact of change, natural
-%or anthropomorphic, and provide an informed opinion to policy makers to
-%effect changes in preserving the environment. However, the challenge of 
-% The pressure on marine resources is growing and increased accuracy,
-% resolution, and persistent monitoring of the oceans is crucial for
-% long-term sustainable management.
-
-% A
-% sustained focus on prioritized and efficient data collection strategies
-% have therefore started to emerge. The advent of marine robotic
-% platforms, especially
-
-% provided means to execute this prioritization through the capacity of
-% autonomy and data-driven sampling, where data collection in principle
-% can be optimized. These capabilities have made
-
-% of the emerging sensing practice for ocean science, allowing scientists
-% to increase the observational efficiency and resolution beyond what was
-% previously possible. But how should a robotic platform, such as an AUV,
-% effectively prioritize and identify important regions for sampling? The
-% answer to this question relates to
-
-% , and
-% the application domain is clearly an arena where statisticians can
-% contribute.
-% There has recently been some statistical attention in oceanography:
-
-%From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest.
-Each phenomenon can in turn be characterized by a set of process specific conditions expressed through different measures of key environmental variables, such as temperature or salinity.
-One such measure is the gradient, that can be associated with a number of important processes, such as the vertical location of the thermocline and pycnocline, location of upwelling systems, vertical mixing, eddies, fronts, and currents \cite{sverdrup2006}, as well as distribution, growth, and accumulation of biological activity \cite{SatOceanSoci00, Ryan2014}.
-These gradients create boundaries separating the ocean into process specific regions which are of profound interest to both identify and map effectively.
-Quantification of gradient features is therefore a much needed competence in robotic sampling.
-\end{document}
+\begin{equation*}
+\begin{split}
+%\eemv_{[n]}(\bm{x})
+%\operatorname{Var}[\mes(\es)]
+&\int_{D^2} 
+\currentExp{
+\varPhi_{\no}\left(\bt;\mu_{[n+1]}(u), K_{[n+1]}(u)\right)
+\varPhi_{\no}\left(\bt;\mu_{[n+1]}(v), K_{[n+1]}(v)\right)
+}
+\
+\mathrm{d}\mes^{\otimes} %\mes 
+%\productMeasure 
+(u,v)\\
+=&
+\int_{D^2} 
+\currentExp{
+    \varPhi_{\no}\left(a_1+B_1 V; C_1\right)
+    \varPhi_{\no}\left(a_2+B_2 V; C_2\right)
+}
+\
+\mathrm{d}\mes^{\otimes} %\mes 
+%\productMeasure 
+(u,v),\\
+\end{split}
+\end{equation*}
+and expanding the integrand (for fixed $n$, $\uu, \vv$) as
+\begin{equation*}
+\begin{split}
+&
+\currentExp{
+    \varPhi_{\no}\left(\bt;\mu_{[n+1]}(u), K_{[n+1]}(u)\right)
+    \varPhi_{\no}\left(\bt;\mu_{[n+1]}(v), K_{[n+1]}(v)\right)
+}\\
+=&
+\currentExp{
+    \varPhi_{\no}\left(a_1+B_1 V; C_1\right)
+    \varPhi_{\no}\left(a_2+B_2 V; C_2\right)
+}
+\end{split}
+\end{equation*}
+with $V=\gp[\bm{x}_{n+1}]-\currentMean{\bm{x}_{n+1}} \sim \mathcal{N}(0_{q_{n+1}},k_{[n]}(\bm{x}_{n+1},\bm{x}_{n+1}))$ and $a_1=\bt-\currentMean{\uu}$,
+$B_1=-\lambda_{[n+1,n+1]}(\uu)^T$, $a_2=\bt-\currentMean{\vv}$, $B_2=-\lambda_{[n+1,n+1]}(\vv)^T$.
+\end{proof}

### 2020-06-12 eibv.tex
Updates from Overleaf

This section is devoted to computing the expected effect of the inclusion of new observations on the $\emv$ and $\ibv$ of the excursion set $\es$. Let us consider the same setting as in Section \ref{sec:cokriging} where $n$ batches of measurements have already been performed, and let 
$\currentExp{.}$ and $\currentProba{.}$ denote conditional expectation
and probability conditional on the first $n$ batches of observations, respectively. We want to co pute the effect of the inclusion of a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and $\ibv$.

For uncertainty functional such as $\ibv$ and $\emv$, we use a similar indexing system
with respect to a sequence of batch, and further indicate when applicable the new batch
of generalized location between parentheses; e.g., $\emv_{[n+1]}(\bm{x})$ denotes the value
of $\emv$ (a random variable, seen from ``step'' $n$, i.e. when knowing $Z$ solely at
the $n$ first generalized location batches) when adding observations at $\bm{x}$ to those
already available at step $n$.

\begin{propo}
\label{propo_eibv}
Let $\eibv_{[n]}(\bm{x})=\currentExp{\ibv_{[n+1]}(\bm{x})}$ (where $\bm{x}$ plays
the role of $\bm{x}_{n+1}$) be the expected effect of adding an observation a $\bm{x}$ on the IBV.
Then we have
\begin{equation}
\begin{split}
\eibv_{[n]}(\bm{x})
&=\int_{D} \varPhi_{\no}\left(\bt;~\currentMean{\uu}, \currentCov{u, u}\right) d\mes(u)\\
&-\int_{D} \varPhi_{2\no}
\left(
\left(
\begin{matrix}
\bt-\currentMean{u}\\
\bt-\currentMean{u}
\end{matrix}
\right);
\mathbf{\Sigma}_{[n]}(\uu)
\right)
d\mes(u),
\end{split}
\end{equation}
where the matrix $\mathbf{\Sigma}_{[n]}(\uu)$ is defined as
\begin{equation*}
\begin{split}
\mathbf{\Sigma}_{[n]}(\uu)&=
\left(
\begin{matrix}
\currentCov{u, u} & \currentCov{u, u}-\futureCov{u, u}\\
\currentCov{u, u}-\futureCov{u, u} & \currentCov{u, u}
\end{matrix}
%\begin{matrix}
%\currentCov{u, u} & \Delta_{[n]}(\uu)\\
%\Delta_{[n]}(\uu) & \currentCov{u, u}
%\end{matrix}
\right).\\
\end{split}
\end{equation*}
\end{propo}

As for the expected excursion measure variance, a similar result may be derived.
\begin{propo}
\label{propo_emv}

\begin{equation*}
\begin{split}
\eemv_{[n]}(\bm{x})
&=\int_{D^2} 
\varPhi_{2\no}
\left(
(\bt, \bt); \mu((u,v)), 
K((u,v),(u,v))
\right) 
\
\mathrm{d}\mes^{\otimes} %\mes 
%\productMeasure 
(u,v)\\
&-\currentExp{
    \varPhi_{\no}\left(a_1+B_1 V; C_1\right)
    \varPhi_{\no}\left(a_2+B_2 V; C_2\right)
},
\end{split}
\end{equation*}

with $V=\gp[\bm{x}_{n+1}]-\currentMean{\bm{x}_{n+1}} \sim \mathcal{N}(0_{q_{n+1}},k_{[n]}(\bm{x}_{n+1},\bm{x}_{n+1}))$ and $a_1=\bt-\currentMean{\uu}$,
$B_1=-\lambda_{[n+1,n+1]}(\uu)^T$, $a_2=\bt-\currentMean{\vv}$, $B_2=-\lambda_{[n+1,n+1]}(\vv)^T$.

This integrand finally boils down to an evaluation of the $2\no$-dimensional Gaussian CDF by
Proposition~\ref{propo3}. Details are omitted here for brevity as we did not implement further
this criterion that requires numerical integration over $\domain^2$.
\end{propo}

Note that the calculation of the multivariate Gaussian cumulative distribution
may be done effectively using code such as that of
\cite{genz2009computation}.

\textcolor{red}{Do we really want to keep this?}
A critical element in the derivation of a closed-form for the EIBV is
that the conditional mean in the cokriging equations is a linear
(affine) function of the data $\by_d$ and the covariance is not a
function of the data. This in turn means that the probabilities in Eq.
\eqref{eq:post_ep} involve inequality statements for linear
combinations of Gaussian variables. Related closed-form solutions have
been noted in similar contexts \citep{bhattacharjya2013value,
  chevalier2014fast,stroh}, but not generalized to our situation with
random sets for vector-valued GPs.

### 2020-06-12 main_new.tex
Updates from Overleaf

@@
 \section*{Appendix}
-
-Proof of Proposition \ref{propo1}:
-
-\begin{proof}
-	That $\mes(\es)$ defines indeed a random variable follows from Fubini's theorem 
-	relying on the joint measurability of 
-	$(\x, \omega) \to \mathbbm{1}_{\es(\omega)}(\x)$, 
-	%= \mathbf{1}_{\xi_{\x}(\omega) \in T}$ and Fubini's theorem.  
-	itself inherited from the assumed measurability for  
-	$(\x, \omega) \to \gp[\x](\omega)$ and $T$, respectively. From there, following the steps of Robbins' theorem \cite{Robins1944}, we find that 
-	\begin{equation*}
-	\begin{split}
-	\mathbb{E}[\mes(\es)^{r}]
-	&=\mathbb{E}\left[\left(\int_{D} \mathbbm{1}_{\gp[u] \in T} ~d\mes(u) \right)^{r} \right]
-	=\mathbb{E}\left[ \prod_{i=1}^{r} \left(
-	\int_{D} \mathbbm{1}_{\gp[\uu^{(i)}] \in T} ~d\mes(\uu^{(i)}) 
-	\right) \right] \\
-	&
-	=
-	%\mathbb{E}\left[
-	%\int_{D^{r}}
-	%\prod_{i=1}^{r}
-	%\mathbf{1}_{\gp[\uu^{(i)}] \in T}
-	%\productMeasure
-	%\right]
-	%=
-	\mathbb{E}\left[
-	\int_{D^{r}}
-	\mathbbm{1}_{\gp[\uu^{(1)}] \in T,\dots, \gp[\uu^{(r)}]  \in T}
-	\productMeasure
-	\right]
-	%\\
-	%&=\int_{D^{r}} \mathbb{E}\left[\mathbf{1}_{\gp[\bm{u}] \in T^r} \right] \productMeasure \\
-	%&
-	=\int_{D^{r}}
-	\jointExcuProb 
-	\productMeasure.
-	\end{split}
-	\end{equation*}   
-	The rest consists in expliciting the probability of $T\times \dots \times T$ under the multivariate Gaussian distribution of  
-	$\left(\gp[u^{(1)}], \dots,  \gp[u^{(r)}] \right)$. 
-\end{proof}
-
-Proof of Proposition \ref{propo2}:
-
-\begin{proof}
-	By definition of $\Phi_{p}$, for $N\sim \mathcal{N}_{p}(0_{p},\covN)$, 
-	$$
-	\mathbb{P}(N\leq a + BV | V)
-	=
-	\varPhi_{p}\left( a + BV; \covN \right).
-	$$
-	Now for $\varPhi_{p}\left( a + BV; \covN \right)^h$, provided that the probability space is sufficiently large to accomodate $h$ independent Gaussian random vectors $N_i\sim \mathcal{N}_{p}(0,\covN)$ (which is silently assumed here), using the former equality delivers
-	$$
-	\varPhi_{p}\left( a + BV; \covN \right)^h
-	=
-	\prod_{i=1}^h \mathbb{P}(N_i\leq a + BV | V).
-	$$
-	Now by independence of the $N_i$'s we obtain the joint conditional probability 
-	$$
-	\prod_{i=1}^h \mathbb{P}(N_i\leq a + BV | V)
-	=
-	\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV| V),
-	$$
-	whereof, by virtue of the law of total expectation, 
-	\begin{equation*}
-	\begin{split}
-	\mathbb{E}\left[ \varPhi_{p}\left( a + BV; \covN \right)^h \right]
-	&=\mathbb{E}\left[\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV| V)\right]\\
-	&=\mathbb{P}(N_1\leq a + BV, \dots, N_h\leq a + BV)\\
-	&=\mathbb{P}(W_1 \leq a, \dots, W_h\leq a)\\
-	&=\varPhi_{ph}
-	\left(
-	1_{h} \otimes a
-	;
-	(1_{h}1_{h}')\otimes (B\Sigma_{V} B') + 
-	I_{h}\otimes \covN
-	\right),
-	\end{split}
-	\end{equation*}
-	where $\mathbf{W}=(W_1,\dots,W_h)$ with $W_i=N_i- BV \ (1\leq i \leq h)$
-	and the last line follows $\mathbf{W}$ forming a Gaussian vector (by global independence of the $N_i$'s and $V$) and from the definition of $\varPhi_{p h}$.
-The covariance matrix $\mathbf{\Sigma}$ of $\mathbf{W}$ is obtained by noting that $\operatorname{cov}(W_i,W_j)=B \covV B' + \delta_{ij} \covN \
-	(i,j \in \{1,\dots,h\})$.
-	%the latter concatenated Gaussian vector is obtained by straightfoward calculation. 
-\end{proof}
-
-Proof of Proposition \ref{propo3}.
-
-\begin{remark}
-	%    Note that the vector $\bm{a}$ is obtained by stacking $h_i$ times each $a_i$
-	%\begin{equation*}
-	%\bm{a} = \begin{pmatrix}
-	%    \bm{a}_1\\ \vdots\\ \bm{a}_g\end{pmatrix},~
-	%\bm{a}_i =
-	%\begin{pmatrix}
-	%    a_i\\ \vdots\\ a_i\end{pmatrix}
-	%\in\mathbb{R}^{l\times h_i}\end{equation*}
-	%and the covariance matrix is obtained by similar stacking
-	Using blockwise representation for the blocks themselves delivers %$\bm{\Sigma}$ can be expressed as follows
-	\begin{equation*}
-	%    \bm{\Sigma} = \begin{pmatrix}
-	%        \Sigma_{11} & & \Sigma_{1g}\\
-	%        & \ddots &\\
-	%        \Sigma_{g1}&   & \Sigma_{gg}
-	%\end{pmatrix},~
-	\Sigma_{ij} =
-	\begin{pmatrix}
-	B_i\Sigma_{V} B_j' & \dots & B_i\Sigma_{V} B_j'\\
-	\vdots & & \vdots\\
-	B_i\Sigma_{V} B_j' & \dots & B_i\Sigma_{V} B_j'\\
-	\end{pmatrix}
-	+
-	\delta_{ij}
-	\begin{pmatrix}
-	\covN_i & &\\
-	& \ddots &\\
-	&   & \covN_i
-	\end{pmatrix}%\in\mathbb{R}^{lh_i\times lh_j}
-	\end{equation*}
-	Here each $\Sigma_{ij}$ is made of $h_i$ times $h_j$
-	(vertically/horizontally) $p \times p$ sub-blocks, hence possesses $ph_i$ lines and $ph_j$ columns.
-\end{remark}
-
-%Proof TODO along with notation polish.
-
-\begin{proof}
-	The proof relies (again) heavily on the fact that, by definition of $\Phi_{p}$, for any covariance matrix $\covN \in \R^{p \times p}$, $a\in \R^p$, $B\in \R^{p \times q}$, and $N\sim \mathcal{N}_{p}(0_{p},\covN)$,
-	$$
-	\mathbb{P}(N\leq a + BV | V)
-	=
-	\varPhi_{p}\left( a + BV; \covN \right).
-	$$
-	In particular, for globally independent $N_{i,j} \sim \mathcal{N}_{p}(0_{p},\covN_i)$ $(1\leq j \leq h_i, 1\leq i \leq g)$, 
-	\begin{equation*}
-	\begin{split}
-	\prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i}
-	&=
-	\prod_{i=1}^{g}
-	\prod_{j=1}^{h_{i}}
-	\mathbb{P}(N_{i,j}\leq a_i + B_i V | V)\\
-	&=\mathbb{P}(N_{1,1}\leq a_1 + B_1 V, \dots, N_{g,h_{g}}\leq a_g + B_g V | V),
-	\end{split} 
-	\end{equation*}
-	so that, by the law of total expectation, 
-	\begin{equation*}
-	\begin{split}
-	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
-	=
-	\mathbb{P}(W_{1} \leq 1_{h_{1}} \otimes a_1, \dots, W_{g} \leq 1_{h_{g}} \otimes a_g)
-	\end{split}
-	\end{equation*}
-	where 
-	$W_{1}=(N_{1,1}- B_1 V, \dots, N_{1,h_{1}}- B_1 V), 
-	W_{2}=(N_{2,1}- B_2 V, \dots, N_{2,h_{2}}- B_2 V), 
-	\dots, W_{g}=(N_{g,1}- B_g V, \dots, N_{g,h_{g}}- B_g V)$. Noting that $\mathbf{W}=(W_1,\dots, W_g)$ is a centred $p H$-dimensional Gaussian random vector, we finally obtain that 
-	\begin{equation*}
-	\begin{split}
-	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \covN_{i} \right)^{h_i} \right]
-	=
-	\varPhi_{p H}\left(\bm{a};\mathbf{\Sigma}\right),
-	\end{split}
-	\end{equation*}
-	with $\bm{a}=(1_{h_{1}} \otimes a_1, \dots, 1_{h_{g}} \otimes a_g)$ and $\bm{\Sigma}=(\operatorname{cov}(W_i,W_j))_{i,j \in \{1,\dots, g\}}$.
-\end{proof}
+\input{appendix}
 

### 2020-06-12 section2.tex
Updates from Overleaf

@@
 
-We define the random sets of interest in Section \ref{sec:ES}, while the closed-form expressions for the expected uncertainty reduction are derived in Section \ref{sec:EIBV}.
-Section \ref{Sec:UnivarEx} provides an instructive bivariate example relevant for sampling in the temperature and salinity application.
+In section \ref{sec:bg_and_notation} we introduce the notation used in the rest of the paper, review (co-)kriging of multivariate gaussian random fields (GRFs) and present update formulae for cokriging.
+Uncertainty quantification (UQ) techniques on excursion sets of GRFs are presented in \ref{sec:set_uq}, in particular the Integrated Bernoulli Variance (IBV) and the excursion measure variance (EMV) are defined \textcolor{red}{and explicit expressions enabling their computation are derived.}
+Section \ref{sec:eibv} turns to the effect of new observations on EMV and IBV. Specifically, the expected effect of the 
+inclusion of a new set of observations on the EMV and IBV are considered and closed-form expressions for the expected 
+uncertainty reduction are derived. Those expression form the backbone of the uncertainty reduction strategies presented in
+Section \ref{sec:heuristics}.
+Finally, Section \ref{Sec:UnivarEx} illustrates the application of the above concepts on a bivariate example relevant for 
+sampling in the temperature and salinity application.
 
@@
 
-In order to formalize the problem in mathematical terms, let us
-consider a multivariate (i.e., vector-valued) Gaussian random field
-$(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary
-set and the dimensionality of the output $\no \geq 1$ is arbitrary.
+The temperature and salinity field over the domain of interest will be modeled as a bivariate gaussian random field.
+The \textcolor{red}{complement of the} river plume may then be characterized as an \textit{excursion set} of the field, that is, as the region of the domain for which temperature and salinity lie above some specified threshold.
+Our goal is then to guide the data collection process so as to improve the characterization of this set.
+The methods we develop are fairly general, and not limited to the special case of two-dimensional random fields.
+Hence, the rest of this section will consider generic excursion sets of gaussian random fields with an arbitrary number of output dimensions.
 
-Assume furthermore that the set of interest
-is characterized in terms of a subset of values of the response variables, say
-$\T \subset \R^{\no}$, and denote the corresponding set of spatial locations by
+\begin{figure}[h!] \centering
+  \includegraphics[width=0.99\textwidth]{Figures/example_excu_1.png}
+  \caption{Realization of a bivariate GRF and excursion set above some threshold. Joint excursion in red, excursion of a single variable in white.}
+\label{example_excu}
+\end{figure}
+
+Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
+and we are interested in recovering the set of locations $\es$ in the domain for which the response variables (the components of $\gp$) lie in some set of specified values $\T\subset \mathbb{R}^{\no}$:
 $$
-\es=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
+\es:=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
 $$
 Our goal here is to develop approaches to quantify and improve the
-characterization of uncertainties on such sets $\es$, in the
-context of multivariate random fields.
+characterization of uncertainties on $\es$. \textcolor{red}{Alread mention how observations are included.}
 
-While some aspects of the developed approaches do not call for a
+If we assume that $\gp$ has
+continuous trajectories (almost surely) and $T$ is closed, then
+$\es$ becomes a Random Closed Set
+\citep{Molchanov2005} and concepts from the theory of random sets will prove useful in characterizing the distribution of the volume of $\es$ under some measure defined on the domain.
+
+\textcolor{red}{Is this really needed?} Note that while some aspects of the developed approaches do not call for a
 specific form of $\T$, we will often stick for simplicity to the case
@@
 
-If we assume that $\gp$ has
-continuous trajectories (almost surely) and $T$ is closed, then
-$\es$ becomes a Random Closed Set
-\citep{Molchanov2005}. %In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
-\textcolor{red}{Peut etre supprimer } In what follows, we will be interested in characterizing the distribution of the volume of the excursion set $\es$ with respect to some (locally finite, Borel) measure $\mes$
-on $D$.
+\subsection{Background, Notation and Co-Kriging}
+\label{sec:bg_and_notation}
+Given a random field $\gp$ and (noisy) observations of some of its components at some points in the domain, one can predict the value of the field at some unobserved location $s\in \domain$ by using the conditional mean of $\gp[s]$, conditional on the data.
+This process is called (co-)kriging, and kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
+We will here present the most general form of cokriging, where one can include several observations (batch) at a time and 
+observations at a given location $s \in \domain$ may only include a subset of the components of $\gp[s]\in\mathbb{R}^{\no}$ 
+(heterotopic).
 
-\subsection{Background and Notation}
-\subsubsection{Co-kriging with respect to (batches) of generalized locations}
+\subsubsection{Notation}
+In order to work at this level of generality, some notational tricks are needed.
+Indeed, when considering a new observation, one shall in general specify which component of the field was observed and where it was observed.
+This leads us to introduce the concept of \textit{generalized location}.
 
-To keep things as general as possible, we will consider performing measurements
-in batches, i.e. collecting several observations at a time. Moreover, we assume
-that given a site $s\in D$, one can choose which component(s) of
-$\gp[s]\in\mathbb{R}^{\no}$ one wants to observe.
-%observation of several components at a time being also allowed. 
-Denoting here for  $s\in D$ and $\ell \in \{1\dots,p\}$ the $\ell\text{th}$ component of $\gp[s]$ by $\gp[s,\ell]$ or 
-$\gp[x]$ with $x=(s,\ell)$, we see that $Z$ can actually also be thought of as a scalar-valued Gaussian random field 
-indexed by $D\times \{1\dots,p\}$. This simple change of representation allows in turn revisiting cokriging as kriging 
-in extended index set, and hence enjoying kriging results (such as batched-sequential update formulae, tackled in
-the next section) without unneccesary complications.
-%
-Let us refer to $x=(s,\ell)$ as (generalized) location and use boldface $s, \ell, x$ to denote batches of sites, 
-response indices, and generalized locations, respectively. A batch $\bm{x}=(x_1,\dots, x_q)$ of generalized locations 
-with $x_i=(s_i,\ell_i)$ $(1\ldots i \ldots q)$ is also by a slight abuse oftentimes denoted $\bm{x}=(\bm{s}, \bm{\ell})$ 
-for convenience.
-%
-We will also use concatenation for the field evaluated at a batch of generalized locations
+Letting $\gp[s,\ell]$ denote the $\ell\text{-th}$ component of $\gp[s]$, $s\in \domain$ and $\ell \in \{1\dots,p\}$, we will call \textit{generalized location} the couple $x=(s,\ell)$.
+Given such a generalized location notation $x$,  the notation $\gp[x]$ will be used to denote $\gp[s,\ell]$.
+
+This slight change of notation will allow us to think of $\gp$ as a scalar-valued Gaussian random field 
+indexed by $D\times \{1\dots,p\}$, which will give the co-kriging equations a particularly simple form that parallels the one of univariate kriging.
+Due to the naturality of the concept of \textit{generalized location},
+we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $s \in \domain$.
+In the following, the letter $x$ will be reserved for generalized locations, while the letters $s$ and $\ell$ will be used for spatial locations and response indices respectively.
+
+It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the 
+one for a single observation, provided some notation adjustments are made. This motivates the following.
+
+Given a dataset consisting of  $q$ observations at spatial locations $s_i \in \domain$ and response indices $l_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we can concatenate the information needed to specify the dataset by using the notation
+\begin{align*}
+\bm{x}:=(\bm{s}, \bm{\ell}):= (x_1,\dots, x_q),~\text{with }x_i=(s_i,\ell_i).
+\end{align*}
+In general, boldface letters will be used to denote concatenated quantities corresponding to batches of observations. 
+We can then also compactly denote the values of the field at those different locations by
 \begin{align*}
@@
 %
+\subsubsection{Co-Kriging}
+\label{sec:cokriging}
 Assuming now that $n$ batches of observations are available, with respective sizes $q_1,\dots, q_n$, and that one wishes 
@@
 Let us now consider the case where co-kriging prediction of $Z$ was made with respect to $n$ batches $\bm{x}_i$ of generalized locations, concatenated again within
-$\bm{x}_n=(\bm{x}_1,\dots, \bm{x}_n)$, and one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
+$\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$, and one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
 %It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
@@
 
-\subsection{UQ on ES of vector-valued fields with volume moments and friends (focus on VV and IBV)}
+\subsection{Uncertainty Quantification on Excursion Sets of multivariate Gaussian Random Fields}
+\label{sec:set_uq}
 \input{UQonSets}
 
-\medskip
-
-%That $\mu(\es)$ is a well-defined random variable follows from
-\begin{propo}
-	\label{propo1}
-	%Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
-	For a measurable random field $\gp$ and a locally finite
-        measure $\mes$ on $\mathcal{M}$, $\mes(\es)$ is a random
-        variable and for any $r\geq 1$, \begin{equation*}
-          \begin{split} \mathbb{E}[\mes(\es)^r]
-            &=\int_{\mathcal{M}^{r}} \jointExcuProb \productMeasure ,
-          \end{split} \end{equation*}
-
-	where the product measure is denoted as
-	$\nu^{\otimes}:=\bigotimes_{i=1}^r \nu$. 
-	%Note that we will use boldface to denote concatenated variables. % TO RELOCATE
-	Here $\gp$ is defined on $\mathcal{M}$, and for
-	$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in \mathcal{M}^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
-	\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.\kc{Is $\no r$ a
-        product of the $\no$ from above? Is that obvious?}
-	\medskip
-
-	In the particular case where $\gp$ is a multivariate Gaussian random field 
-	%and denoting the mean and covariance matrix of $\gp[\bm{u}]$ by $\meanUU\in\mathbb{R}^{p\times r}$ and $\covUU$, respectively, 
-	we have 
-	\begin{align*}
-	\jointExcuProb = \mathcal{N}_{\no r}(T^r; \meanUU, \covUU),
-	\end{align*}
-	where $\mathcal{N}_{\no r}(\cdot ; \meanUU, \covUU)$ is the Gaussian measure on $\mathbb{R}^{\no r}$ with mean $\meanUU$ and covariance matrix $\covUU$, respectively defined blockwise by  
-	%where both are defined blockwise by 
-	\begin{align*}
-	\meanUU&=\begin{pmatrix}\mu(u^{(1)})\\ \vdots\\ \mu(u^{(r)})\end{pmatrix}
-	\in \mathbb{R}^{\no r}, \\
-	\text{and } \covUU &= \begin{pmatrix}
-	\cov(\gp[u^{(1)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(1)}],
-	\gp[u^{(r)}])\\
-	\vdots & & \vdots\\
-	\cov(\gp[u^{(r)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(r)}],
-	\gp[u^{(r)}])\\
-	\end{pmatrix}\in \mathbb{R}^{pr\times pr},
-	\end{align*}
-	each of the $r\times r$ blocks of the latter matrix being itself a (cross-)covariance matrix of dimension $\no \times \no$. 
-	%dimensional Gaussian random vector. 
-	Assuming further that $\covUU$ is non-singular, the probability of interest can be formulated in terms of the $\no r$-dimensional Gaussian probability density function 
-	$\varphi_{\no  r}(\cdot;~\meanUU, \covUU)$ as 
-	%$\varphi_{\no \times r}$ as
-	%$\bm{u}\in D^r$ such that $\gp[\bm{u}]$ be non-degenerate, the integrand can be expressed in terms of the $r \times \no$-dimensional Gaussian distribution, namely 
-	%cumulative distribution function, namely 
-	\begin{equation*}
-	\begin{split}
-	&\jointExcuProb
-	=
-	\int_{T^r} \varphi_{\no  r}\left(\bm{v};
-	~\meanUU, \covUU\right) 
-	\mathrm{d}\bm{v},
-	\end{split}
-	\end{equation*}
-	In the particular orthant case with $\T=(-\infty, t_1] \times \dots \times (-\infty, t_{r}]$,
-	the latter probability directly writes \kc{What does
-          ``writes'' mean here? Connotes?} in terms of the multivariate Gaussian
-	cumulative distribution, % with mean $\meanUU$ and covariance matrix $\covUU$,
-	this time without requiring $\covUU$ %the latter 
-	to be non-singular: 
-	\begin{equation*}
-	\begin{split}
-	\jointExcuProb 
-	&=
-	\varPhi_{\no r}\left(\bm{t};~\meanUU, \covUU\right),
-	\end{split}
-	\end{equation*}
-	where we have used the notations 
-	$t=(t_1,\dots,t_{\no}%,..., t_1, ..., t_p
-	)\in\mathbb{R}^{\no}$, $1_{r}=(1,\dots,1)\in \R^{r}$, and 
-	$\bm{t}=1_{r}\otimes \bm{t}=(t_1,\dots,t_{\no},\dots,t_1,\dots,t_{\no}) 
-	\in \R^{\no r}$.
-\end{propo}
-
-Proposition~\ref{propo1} enables one to calculate centered moments of $\mes(\es)$. 
-The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$
-is probably the most relevant of these for our purposes. We
-have 
-\begin{equation*}
-\begin{split}
-\emv
-&=\int_{\mathcal{M}^2} \mathbb{P}\left(
-\gp[u]\in T, \gp[v]\in T \right) 
-d\mes^{\otimes}(u, v)\\
-&-\left( \int_{\mathcal{M}} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
-\end{split}
-\end{equation*} 
-which evaluates in the excursion/sojourn case where
-$\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ to
-\begin{equation*} \begin{split}
-  \emv
-%\operatorname{Var}[\mes(\es)]
-&=\int_{\mathcal{M}^2} 
-\varPhi_{2\no}
-\left(
-(\bt, \bt); \bmu((u,v)), 
-K((u,v),(u,v))
-\right) 
-\
-\mathrm{d}\mes^{\otimes} %\mes 
-%\productMeasure 
-(u,v)\\
-&-\left( \int_{\mathcal{M}} \varPhi_{\no}\left(\bt;\bmu(u), K(u)\right) d\mes(u) \right)^2.
-\end{split}
-\end{equation*}
-This quantity requires  an integral over $D^2$. In
-contrast, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal}
-involves solely an integral on $\mathcal{M}$ and can be expanded in
-our present setting as follows 
-\begin{equation}\label{IBV}
-\begin{split}
-\operatorname{IBV} %(\es) %;\mes)
-&=\int_{\mathcal{M}}
-\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
-d\mes(u) \\
-&=\int_{\mathcal{M}}
-\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) 
--\left(\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) \right)^2
-\mathrm{d}\mes(u). 
-\end{split}
-\end{equation}
-
-\subsection{Expected Integrated Bernoulli Variance and Sampling Designs}
-\label{sec:EIBV}
-
-Our goal is to construct informative sampling designs, given limited
-resources \kc{information?}. In our domain, data is gathered by an AUV
-that can measure temperature and salinity at chosen locations on a
-defined waypoint graph. Such a discretized grid fosters reliable
-navigation and mission performance. The waypoints further enable a
-structure to the design problem in that, that sampling flexibility is
-limited to operation on the graph. % This gives lots of sampling
-% opportunities, and it is important to spend the sampling efforts
-% wisely.
-We focus on sampling designs that reduce the expected integrated
-Bernoulli variance (EIBV). Note that this sampling can be done in a
-static manner where the survey design is pre-planned, or in an
-adaptive setting. For the latter case, data can be processed
-sequentially and the updated model is then used to design and select
-the next waypoint graph to visit, which in turn are used again to
-update the model, and so on (Section \ref{sec:heuristics}). We next
-describe the data modeling assumptions for one stage and develop
-design criteria of semi-analytical forms relying on integrals of
-multivariate Gaussian cumulative distribution functions; we will
-appeal to some key intermediate results presented next.
-
-We denote a design by $d \in \mathcal{D}$, where the concatenation of designs with $q$ sampling locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$ is $\mathcal{D}$. The
-measurement model is defined by 
-\begin{equation}\label{eq:meas}
-    \bY_d=G_d \bz(\bu_d) + \epsilon, \hspace{3mm} \epsilon \sim N(0,R_d).
-\end{equation}
-Here, $\bz(\bu_d)$ is the size $pq$ vector of GP variables at the
-sampling locations, and $G_d$ is the design matrix which in principle
-allows for observing subsets of the vector-valued field at some of the
-$q$ locations (i.e. the matrix has fewer rows than columns). If the
-additive error has independent elements, the covariance matrix $R_d$
-is diagonal. In our setting temperature and salinity measurements are
-extracted from different sensors, so the independence assumption
-between variables is not unrealistic, but this is not necessarily the
-case for other kinds of measurements on the AUV. It is further common
-in geostatistics to assume spatially independent measurement errors,
-and we also do so. But our methodology for constructing sampling
-designs does not rely on these independence assumptions.
-
-For a given design $d$ the evaluation is based on the expected
-information contained in the data to influence the criterion.   
-Considering the EIBV criteria, we have
-\begin{eqnarray}\label{eq:eibv}
-    \mbox{EIBV} (d) &=& E_{\bY_d} \left( \int_{\mathcal{M}}
-p(\x;\by_d)(1-p(\x;\by_d))
-d\mes(u) \right), \nonumber \\
-p(\x;\by_d) &=& \mathbb{P}\left(\gp[\uu]\in T|\bY_d=\by_d \right)
-\end{eqnarray}
-where the expectation is over the planned data.  The optimal design
-for an AUV survey is then 
-\begin{equation}\label{crit}
-    d^* = \mbox{argmin}_{d \in \mathcal{D}} \mbox{EIBV}(d).
-\end{equation}
-For another criterion using random moments of random sets similar
-expectations would be required, but as we will see, the EIBV has a
-closed-form expression which is relatively fast to compute and hence
-useful for onboard design selection.
-
-Assuming now that $q$ observations $\by_d$ made at $\bu_d$ are available, one wishes to predict $\gp[\x]$. The multivariate Kriging equations holds with size
-$p$ vector of conditional means:  
-\begin{equation}\label{eq:cokrig}
-\bbm(\x)=\bmu(\x)+\lambda(\x,\bu_d)^T \left( \by_d-G_d \bmu(\bu_d)\right),
-\end{equation}
-where
-$\lambda(\x,\bu_d)=K^{-1}_y G_d k(\bu_d, \bu) $ with data covariance
-$K_y=\mbox{Cov}(\bY_d,\bY_d)=G_d K(\bu_d,\bu_d)G_d^T+R_d$ and cross-covariance $\mbox{Cov}(\bY_d,Z(\x))=G_d k(\bu_d, \x)$, assuming the measurement
-error is independent of the random fields.  The Kriging (cross-)covariance function can also be expressed in the same vein via
-a $p \times p$ matrix 
-\begin{equation}\label{eq:krig:cov}
-\Sigma(\x,\x')=K(\x,\x')-\lambda(\x,\bu_d)^T K_y(\bu_d, \bu_d) \lambda(\x,\bu_d).
-\end{equation}
-
-Eq. \eqref{eq:cokrig} and \eqref{eq:krig:cov} are used to
-provide informative, updated statements about sets of interest. For
-our case with temperature and salinity, the conditional EPs, given
-measurements $\by_d$, can then be computed from
-\begin{eqnarray}\label{eq:post_ep}
- p(\x;\by_d) &=& P(Z_1(\x) \leq t_1, Z_{2}(\x) \leq t_2 |\bY_d=\by_d),  \nonumber \\
- &=& \Phi_2 ((t_1,t_2)^T-\bbm(\x);0_2, \Sigma(\x,\x)). 
-\end{eqnarray}
-
-A critical element in the derivation of a closed-form for the EIBV is
-that the conditional mean in Eq. \eqref{eq:cokrig} is a linear
-(affine) function of the data $\by_d$ and the covariance is not a
-function of the data. This in turn means that the probabilities in Eq.
-\eqref{eq:post_ep} involve inequality statements for linear
-combinations of Gaussian variables. Related closed-form solutions have
-been noted in similar contexts \citep{bhattacharjya2013value,
-  chevalier2014fast,stroh}, but not generalized to our situation with
-random sets for vector-valued GPs. Notably, the entry in Eq.
-\eqref{eq:post_ep} is in the form of $\ba+B V$, where
-$V \sim N(0_q, K_y (\bu_d,\bu_d))$,
-$\ba=(t_1,t_2)^t-\bmu(\x)$
-and $B=\lambda(\x,\bu_d)^T$. The following proposition then applies to give a
-closed-form solution to EIBV calculation in this case.
-
-\begin{propo}
-	\label{propo2}
-	Let $p, q, h \geq 1$, $a \in \R^p$, $B \in \R^{p\times q}$,
-	and $\covN, \covV$ be two covariance matrices in
-	$\R^{p\times p}$ and $\R^{q\times q}$, respectively.
-	Then, for $V \sim \mathcal{N}_{q}(0_q, \covV)$,
-	\begin{equation*}
-	\mathbb{E}\left[ \varPhi_{p}\left( \ba + BV; 0_p,\covN \right)^h \right]
-	=
-	\varPhi_{ph}
-	\left(
-	\ba
-	;0_{ph},
-	\Sigma
-	\right),
-	\end{equation*}
-	where the vector $\ba \in \R^{p h}$ is defined as 
-	$\ba := 1_h\otimes a = 
-	\left(a, \dots , a
-	\right)'$
-	and the $p h\times p h$ covariance matrix is given by
-	$\Sigma := 
-	1_h 1_h'\otimes B\covV B' + I_h\otimes \covN$.
-\end{propo}
-The proof of Proposition \ref{propo2} is provided in the Appendix.
-
-In block-wise representation, the following relation holds for the
-covariance in Proposition \ref{propo2}: 
-\begin{align*}
-	\Sigma&=
-	\begin{pmatrix}
-	\covN & &\\
-	& \ddots &\\
-	&   & \covN
-	\end{pmatrix}
-	+
-	\begin{pmatrix}
-	B\covV B' & \dots & B\covV  B'\\
-	\vdots & & \vdots\\
-	B\covV B' & \dots & B\covV B'\\
-	\end{pmatrix}.
-\end{align*}
-The calculation of the multivariate Gaussian cumulative distribution
-is done effectively using code such as that of
-\cite{genz2009computation}.
-
-In the more general case, for instance that of volumes of random sets,
-the criterion will include couplings with different locations and the
-result must be generalized. For the case of multivariate monomials in
-orthant probabilities with thresholds affine in a common Gaussian
-vector \kc{This sentence is dangling. Doesn't have a conclusion.}.
-%Namely, for monomials of degree $k$, we have.
-
-\begin{propo}%[\textcolor{red}{Notation in progress, to be reviewed/revised}]
-	\label{propo3}
-	Let $g, p, q\geq 1$, $h_{1},\dots, h_{g}\geq 1$ with $H=\sum_{i=1}^g h_i$, $a_{i} \in \R^{p}$, $B_{i}\in \R^{p \times q}$, and covariance matrices $\covN_i \in \R^{p \times p}$ $(1\leq i \leq g)$.
-Then, for any covariance matrix $\covV \in \R^{q\times q}$ and $V\sim\mathcal{N}_{q}(0_q,\covV)$,
-	%	Let $g$ and $k$ be integers, and consider a $g$-terms monomial of degree $k$
-	%	\[
-	%	    \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \Sigma_{i}
-	%	    \right)^{h_i},
-	%    \]
-	%    where $\sum_{i=1}^g h_i = k$, each $a_i$, $B_i$ and $\Sigma_i$ is as in
-	%    \cref{propo2} and $V$ is a $q$-dimensional Gaussian vector distributed as
-	%    $V\sim\mathcal{N}(0,\covV)$.
-	%
-	% Then the expectation of the monomial is given by
-	\begin{equation}
-	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V;0_p, \covN_{i} \right)^{h_i} \right]
-	=
-	\varPhi_{p H}
-	\left(
-	\bm{a}
-	;0_{pH},
-	\mathbf{\Sigma}
-	\right),
-	\end{equation}
-	with $\bm{a}=(1_{h_1}\otimes a_1, \dots, 1_{h_g}\otimes a_{g}) \in \R^{p H}$
-	and $\mathbf{\Sigma}\in \R^{p H \times p H}$ is defined blockwise by $(\Sigma_{i,j})_{i,j \in \{1,\dots, g\}}$ where, for any $i,j \in \{1,\dots, g\}$, %the blocks are given by 
-	\begin{equation}
-	\Sigma_{i,j}=
-	(1_{h_{i}}1_{h_{j}}')\otimes (B_{i}\Sigma_{V} B_{j}') + \delta_{i,j}(I_{h_{i}}\otimes \covN_{i}) \in \R^{p h_{i} \times p h_{j}}.
-	\end{equation}
-	%with $\bm{a}=\sum_{j=1}^g \mathbf{1}_{J_{j}} \otimes a_j \in \R^{p \times k}$ 
-	%where $J_{j}=\{\sum_{i=1}^{j-1}h_i + 1, \dots, \sum_{i=1}^{j}h_i\}$,
-	%$\mathbf{1}_{I_{j}}$ denotes the vector of $\R^{k}$ with components equal to $1$ for indices in $I_{j}$ and $0$ else ($1\leq j \leq g$), 
-	%and $\mathbf{\Sigma}\in \R^{p k \times p k}$ is defined blockwise by 
-	%\begin{equation}
-	%\mathbf{\Sigma}_{\widetilde{J}_{j_{1}},\widetilde{J}_{j_{2}}}=
-	%(\mathbf{1}_{h_{j_1}}\mathbf{1}_{h_{j_2}}')\otimes (B_{j_{1}}\Sigma_{V} B_{j_2}') + \delta_{j_{1},j_{2}}(I_{j_{1}}\otimes \Sigma_{j_{1}}),
-	%\end{equation}
-	%where $\widetilde{J}_{j}=\{p\sum_{i=1}^{j-1}h_i + 1, \dots, p\sum_{i=1}^{j}h_i\}$ ($1\leq j \leq g$).
-\end{propo}
-The proof of Proposition \ref{propo2} is provided in the Appendix.
+\subsection{Expected Integrated Bernoulli Variance and Excursion Measure}
+\label{sec:eibv}
+\input{eibv}
 
@@
 The thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
+
+\begin{figure}[h!] \centering
+  \includegraphics[width=0.99\textwidth]{Figures/ebvs.png}
+  \caption{Expected Bernoulli variance for observation of first, second, or both components.}
+\label{illus_bivarDens}
+\end{figure}
 

### 2020-06-15 ans_to_reviewers/answers-to-reviewers.tex
Updates from Overleaf

@@
 \begin{itemize}
-\item The motivation for excursion sets (ESs), and in particular that of {\bf{multivariate ESs}}, has been strengthened.
-Motivating examples are abundant in medicine where doctors do not solely rely on a single symptom but must see several combined effects before making a diagnosis.
+\item The motivation for excursion sets (ESs), and in particular that of {\bf{multivariate ESs}}, has been strengthened in this revision.
+Motivating examples are abundant in medicine where doctors do not solely rely on a single symptom but must see several combined effects before making a diagnosis, for example.
 Our motivation in the current paper is that temperature and salinity are proxies for more complex biological variables in the ocean, and in several cases they must be jointly high or low for phenomena to occur.
@@
 
-In the revised manuscript these aspects are largely rewritten in the Introduction, and partly also in Section 2 on the expected integrated Bernoulli variance (EIBV) and partly in the Case Study ({\bf{TO DO 6 AGAIN}}).
-This change addresses some key comments of AE, Rev 1 in particular.
+In the revised manuscript these aspects are largely rewritten into the Introduction, and partly also in Section 2 on the expected integrated Bernoulli variance (EIBV), and partly in the Case Study ({\bf{TO DO 6 AGAIN}}).
+These changes address some key comments from AE and Rev.
+1 in particular.
 

### 2020-06-16 main_new.tex
updates

@@
   \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
-    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
+    0.49\textwidth]{Figures/envir.pdf}\label{fig:envir1}}
   \hfill

### 2020-06-16 main_new.tex
Updates from Overleaf

@@
 
-\section{Simulation study}
+\subsection{Simulation study}
 \label{sec:simulations}
 
-We now study the properties of the different static and sequential
-survey designs in a realistic case that emulates the spatial
-properties of a river plume interacting with coastal water.
-
-\subsection{Modeling}
-
-A GP model is used to represent the bivariate random fields of
-temperature and salinity. At location $\x \in \mathcal{M}$, the
-bivariate distribution is:
-\begin{equation}\label{gp_bar}
-\begin{bmatrix}Z_1(\x) \\
-Z_2(\x) \end{bmatrix}
-\sim N \left( 
-\begin{bmatrix} \mu_{1}(\x)\\
-\mu_{2}(\x)
-\end{bmatrix},\begin{bmatrix}
-\sigma_{1}^2 & \sigma_{1} \sigma_2 \gamma \\
-\sigma_{1} \sigma_2 \gamma  & \sigma_{2}^2 
-\end{bmatrix}
-\right).
-\end{equation}
-We specify the mean by $\mu_{1}(\x)=5.8 + 0.065 u_{\mbox{West}}$ and
-$\mu_{2}(\x)=29 + 0.1 u_{\mbox{West}}$, which means that the expected
-levels of both temperature and salinity increase linearly in the
-western direction.  This situation mimics that of a river mouth
-opening from the south, with the water masses pulled to the east by
-the Coriolis force \kc{This seems overly specific to the Nidelva use
-  case. Can it be generalized, after all that is the overall intent of
-  this manuscript?}.  The variances are set to $\sigma_1=0.25$ and
-$\sigma_2=0.6$, while the correlation is $0.6$.  The spatial
-correlation is of separable Matern type (see e.g. \cite{Cressie:11})
-so that
-\begin{equation}\label{v}
-\small
-\mbox{Corr} 
-\left(
-\begin{bmatrix}
-    Z_1(\x) \\
-    Z_2(\x) 
-    \end{bmatrix},
-    \begin{bmatrix}
-    Z_1(\x') \\
-    Z_2(\x') 
-    \end{bmatrix}
-    \right)
-    = \begin{bmatrix}
-1 & 0.6  \\
-0.6  & 1
-\end{bmatrix}(1+\phi h)\exp (-\phi h),
-\end{equation}
-where $h=\sqrt{||\x-\x'||^2}$ and $\phi=0.3$ indicates an effective
-correlation range of about $1200$ m.
-
-In the accompanying Python code, these modeling assumptions can be
-generalized to anisotropic covariance and changing variance levels
-across the spatial domain. One can hence, easily run the code to see
-if different models result in other sampling designs. Anisotropy and
-non-stationary variance are both relevant for the setting with river
-plumes, but in practice this requires more parameters to be
-specified. With extensive data and prior knowledge, one could also
-possibly fit and estimate parameters of more complex multivariate
-spatial covariance functions
-\citep{gneiting2010matern,genton2015cross}, but that is outside the
-scope of the current paper.
-
-We discretize the spatial domain $\mathcal{M}$ to a set of $n$ grid
-locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,n \}$, \kc{is the index
-  correct? TeX shows it as u or g for $\mathcal{M}$} where each cell
-has area $\Delta$; the same grid is used for the waypoint graph for
-possible design locations. The EIBV is approximated by sums over all
-grid cells.
-
-One realization of the random fields representing salinity and
-temperature is shown in Fig. \ref{fig:true_temp} and
-\ref{fig:true_sal}. The true ES is a realization from this model as
-shown in Fig. \ref{fig:ESet}.
-
-\begin{figure}[!h]
-  \centering
-  \subfigure[Simulated temperature.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/true_temp.pdf}\label{fig:true_temp}}
-  \hfill
-  \subfigure[Simulated salinity.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/true_sal.pdf}\label{fig:true_sal}}
-  \hfill
-  \subfigure[Excursion set given the realization of temperature and salinity.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_true.pdf}\label{fig:ESet}}
-  \hfill
-  \subfigure[Excursion probabilities given collected data using the
-  `static\_north" strategy.]{\includegraphics[height = 0.40\textwidth]{Figures/sim/es_ts_posterior.pdf}\label{fig:Eprob}}
-  \caption{\ref{fig:true_temp} and \ref{fig:true_sal} show a
-    realization of the simulated temperature and salinity fields. The
-    associated ES is in \ref{fig:ESet}, while \ref{fig:Eprob} shows
-    the estimated EPs after performing data collection along the
-    static N-S survey line.}
-\label{fig:realisations}
-\end{figure}
-
-\subsection{Static and Sequential Sampling Designs}
+\susubsection{Static and Sequential Sampling Designs}
 \label{sec:sampling_designs}
@@
 vertical lines. Nevertheless, the ordering of strategies is similar.
-
-\begin{figure}[!b]
-  \centering
-  \subfigure[Look-ahead]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_look-ahead.pdf}\label{fig:avg_look-ahead}}
-  \hfill
-  \subfigure[Myopic]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_myopic.pdf}\label{fig:avg_myopic}}
-  \hfill
-  \subfigure[Naive]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_naive.pdf}\label{fig:route_naive}}
-  \hfill
-  \subfigure[Static]{\includegraphics[height = 0.46\textwidth]{Figures/sim/route_static.pdf}\label{fig:route_static}}
-  \hfill
-  \caption{The overlaid route choices (strategies), superimposed on
-    the 100 replicate surveys with 10 sampling choices/stages. \kc{I
-      would reorder these figurs above based on the order the material
-      was discussed -- Static, naive, myopic and then Look-ahead. Also
-      do you really need a key associated with each subfigure, given
-      the caption? An abridged analysis as in the text, might also be
-      in order here in the caption. Also refer to
-      Fig. \ref{fig:realisations} as the reference, where the
-      temp/salinity fields are shown.}}
-\label{fig:route_choices}
-\end{figure}
 

### 2020-06-16 section2.tex
Updates from Overleaf

@@
 \label{sec:bg_and_notation}
-Given a random field $\gp$ and (noisy) observations of some of its components at some points in the domain, one can predict the value of the field at some unobserved location $s\in \domain$ by using the conditional mean of $\gp[s]$, conditional on the data.
+Given a random field $\gp$ and (noisy) observations of some of its components at some points in the domain, one can predict the value of the field at some unobserved location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$, conditional on the data.
 This process is called (co-)kriging, and kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
 We will here present the most general form of cokriging, where one can include several observations (batch) at a time and 
-observations at a given location $s \in \domain$ may only include a subset of the components of $\gp[s]\in\mathbb{R}^{\no}$ 
+observations at a given location $s \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ 
 (heterotopic).
@@
 
-Letting $\gp[s,\ell]$ denote the $\ell\text{-th}$ component of $\gp[s]$, $s\in \domain$ and $\ell \in \{1\dots,p\}$, we will call \textit{generalized location} the couple $x=(s,\ell)$.
-Given such a generalized location notation $x$,  the notation $\gp[x]$ will be used to denote $\gp[s,\ell]$.
+Letting $\gp[\spatloc,\ell]$ denote the $\ell\text{-th}$ component of $\gp[\spatloc]$, $\spatloc\in \domain$ and $\ell \in \{1\dots,p\}$, we will call \textit{generalized location} the couple $x=(\spatloc,\ell)$.
+Given such a generalized location notation $x$,  the notation $\gp[x]$ will be used to denote $\gp[\spatloc,\ell]$.
 
@@
 Due to the naturality of the concept of \textit{generalized location},
-we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $s \in \domain$.
-In the following, the letter $x$ will be reserved for generalized locations, while the letters $s$ and $\ell$ will be used for spatial locations and response indices respectively.
+we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $\spatloc \in \domain$.
+In the following, the letter $x$ will be reserved for generalized locations, while the letters $\spatloc$ and $\ell$ will be used for spatial locations and response indices respectively.
 
@@
 
-Given a dataset consisting of  $q$ observations at spatial locations $s_i \in \domain$ and response indices $l_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we can concatenate the information needed to specify the dataset by using the notation
+Given a dataset consisting of  $q$ observations at spatial locations $\spatloc_i \in \domain$ and response indices $l_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we can concatenate the information needed to specify the dataset by using the notation
 \begin{align*}
-\bm{x}:=(\bm{s}, \bm{\ell}):= (x_1,\dots, x_q),~\text{with }x_i=(s_i,\ell_i).
+\bm{x}:=(\bm{\spatloc}, \bm{\ell}):= (x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
 \end{align*}
@@
 \gp[\bm{x}]:=
-\left(\gp[s_1,\ell_1], ...,
-\gp[s_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
+\left(\gp[\spatloc_1,\ell_1], ...,
+\gp[\spatloc_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
 \end{align*}
@@
 Assuming now that $n$ batches of observations are available, with respective sizes $q_1,\dots, q_n$, and that one wishes 
-to predict $\gp[\bm{x}]$ for some batch of $q\geq 1$ generalized locations $\bm{x}=(\bm{s}, \bm{\ell})$, the (simple) 
+to predict $\gp[\bm{x}]$ for some batch of $q\geq 1$ generalized locations $\bm{x}=(\bm{\spatloc}, \bm{\ell})$, the (simple) 
 cokriging mean would then amount to simple kriging with respect to a scalar-valued Gaussian random field indexed by 
-$D\times \{1\dots,p\}$ and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{s}, \bm{s}')_{\ell, \ell'}$, that is:
+$D\times \{1\dots,p\}$ and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{\spatloc}, \bm{\spatloc}')_{\ell, \ell'}$, that is:
 %
@@
 The thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
-
-\begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/ebvs.png}
-  \caption{Expected Bernoulli variance for observation of first, second, or both components.}
-\label{illus_bivarDens}
-\end{figure}
 
@@
 
-%\subsection{Expected classification criteria}
-
-%{\bf{I have not done anything here - skip this, I think}}.
-
-%Rather than minimizing expected variance one can aim at minimizing the classification probability of the excursion set: 
-%$\min [p,(1-p) ]$, see e.g. \cite{lilleborge2016information}.
-
-%Again, since the conditional mean is a linear in the data, we only have to look at the %relevant linear combination via $\bm_{\xi}=E(\bxi)$.
-%See also \cite{bhattacharjya2013value}.
-
-%\begin{equation}
-%E(\min \{ p,(1-p)\})=\int \min P(\xi_1 \leq t, \xi_2 \leq s),[1-P(\xi_1 \leq t, \xi_2 \leq s)] p(E(\bxi)=) dE(\bxi)=,
-%\end{equation}
-%for the complementary probabilities, this is again evaluated by the corner regions for the block, and $\Phi_4()$ evaluations are required.
-
-%In the end, this result is integrated over the spatial domain $x \in X$.
+\begin{figure}[h!] \centering
+  \includegraphics[width=0.99\textwidth]{Figures/ebvs.png}
+  \caption{Expected Bernoulli variance for observation of first, second, or both components.}
+\label{illus_bivarDens}
+\end{figure}

### 2020-06-17 section2.tex
Updates from Overleaf

@@
 
-\subsection{An Example of EIBV using Temperature and Salinity}
+\subsection{Expected Bernoulli Variance on a two dimensional Example}
 \label{Sec:UnivarEx}
+We illustrate the EBV associated to different designs on a simple bivariate example.
+The example mimics our river plume application and hence the first and second component of the random field will be called \textit{temperature} and \textit{salinity} for illustrative purpose.
+Effects of the hyperparameters of the GRF prior on the excursion probabilities will also be studied.
+For simplicity, we begin with a \textit{pointwise} example, considering a single bivariate gaussian distribution (no spatiality).
 
-% In this section, 
-We calculate EPs and the Bernoulli variance for a few bivariate
-Gaussian distributions for temperature and salinity to illustrate the concepts articulated above.
-The parameter inputs are made up, but they are in a realistic range as we have in the simulation study and real-world example below.
-The thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
+\subsubsection{A pointwise study}
+Say we want to study the excursion probability of a bivariate gaussian, where the excursion set is defined as
+\[
+\Gamma := \lbrace \spatloc \in \domain: Z_{\spatloc, 1} \geq t_1, Z_{\spatloc, 2}\geq t_2\rbrace
+\]
+and the thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
 
@@
 
+\subsubsection{Including Spatiality}
+INTRODUCTORY SENTENCE.
+The class of GRF model we will consider in the article will generally have a linear trend
+\begin{align*}
+\mu(s)=\mathbb{E}\left[\begin{pmatrix}
+Z_{s, 1}\\ Z_{s, 2}
+\end{pmatrix}\right] &= \beta_0 + \beta_1 s
+\end{align*}
+with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. We will only consider covariance models of separable type
+\begin{align*}
+\textrm{Cov}\left(Z_{s, i}, Z_{u, j}\right) &= k(s, u) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_l^2,~ i=j\\
+   \gamma_0 \sigma_i \sigma_j,~i\neq j
+        \end{cases}
+\end{align*}
+where $k(., .)$ is one of the traditional spatial covariance kernel (exponential, Mat\'{e}rn, ...) and $\gamma$ defines the cross-covariance structure.
+
+In the accompanying Python code, these modeling assumptions can be
+generalized to anisotropic covariance and changing variance levels
+across the spatial domain. One can hence, easily run the code to see
+if different models result in other sampling designs. Anisotropy and
+non-stationary variance are both relevant for the setting with river
+plumes, but in practice this requires more parameters to be
+specified. With extensive data and prior knowledge, one could also
+possibly fit and estimate parameters of more complex multivariate
+spatial covariance functions
+\citep{gneiting2010matern,genton2015cross}, but that is outside the
+scope of the current paper.
+
+In the rest of this section, we will consider a GRF with mean and covariance structure as above and parameters
+\begin{align*}
+\beta_0 = \begin{pmatrix}
+5.8\\ 24.0
+\end{pmatrix}, ~ \beta_1 = \begin{pmatrix}
+0.0 & -4.0\\
+0.0 & -3.8
+\end{pmatrix},~ \sigma_1 = \sigma_2 = 2.25, ~ \gamma_0 = 0.2
+\end{align*}
+and the spatial covariance is given by a Mat\'{e}rn 3/2 kernel with unit variance and range parameter $\lambda=0.5$.
+One realization of this GRF is shown in Fig. \ref{example_excu}.
+
+We now study how the expected Bernoulli variance eq.\ref{eq:eibv} associated to data collection at a point changes if only one of the two components of the field is observed.
+We first draw a realization of the GRF defined above and use it as ground-truth to mimick the real data-collection process.
+A first data collection step is performed at the location in green, and the data is used to update the GRF model.
+The expected Bernoulli variance reduction (at each grid point) that would result from observing only one component, or both at the next grid point upwards is then computed an shown in Fig.\ref{fig:ebv_comp}.
+
 \begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/ebvs.png}
-  \caption{Expected Bernoulli variance for observation of first, second, or both components.}
-\label{illus_bivarDens}
+  \includegraphics[width=0.99\textwidth]{Figures/ebv_comp.png}
+  \caption{Pointwise Bernoulli variance reduction for observation of a single or both components of the random field at one location.
+Data collection locations in green.
+True excursion set in red.
+Places where only one response is above threshold are depicted in pink}
+\label{fig:ebv_comp}
 \end{figure}
+
+Note that plotting the Bernoulli variance reduction at each point might also be used to compare different data collection plans.
+For example, Fig.
+\ref{fig:ebv_comp} shows the expected Bernoulli variance reduction associated to a data collection plan along a straight upwards line (static north) and the one associated to a straight horizontal line (static east).
+Both expectations are computed according to the a priori distribution of the GRF (i.e.
+no observations have been included yet).
+
+\begin{figure}[h!] \centering
+  \includegraphics[width=0.99\textwidth]{Figures/ebv_north_vs_east.png}
+  \caption{Pointwise Bernoulli variance reduction for two different observation plans.
+Data collection locations in green.
+True excursion set is in red.
+Places where only one response is above threshold are depicted in pink.}
+\label{fig:ebv_north_vs_east}
+\end{figure}
+
+\begin{remark}
+In all the above examples, the spatial domain $\domain$ has been discretized to a set of $n$ grid
+locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,n \}$, where each cell
+has area $\Delta$; the same grid is used for the waypoint graph for
+possible design locations. The EIBV is approximated by sums over all
+grid cells.
+\end{remark}

### 2020-06-17 eibv.tex
Updates from Overleaf

@@
 already available at step $n$.
+
+\textcolor{red}{C.T.: This doesn't work, we cannot be that short on the definition of EIBV, it is highly unclear. I suggest we go for the following notation:
+$$\operatorname{IBV}(\bm{x}; \bm{y})
+$$ denotes the IBV under the current law of the field, conditioned on observing $\bm{y}$ at $\bm{x}$ (generalized, possibly batch observation). Then
+$$
+\operatorname{EIBV}(\bm{x}):=\mathbb{E}_{\bm{Y}}\left[\textrb{IBV}(\bm{x}; \bm{Y})\right]
+$$
+where $\bm{Y}$ is distributed according to the current law of $Z_{\bm{x}}$.}
+
+\textcolor{red}{
+The philosophy is the following: EIBV and EEMV are functionals of a probability distribution.
+Then, if not mentioned explicitly, the probability distribution will be taken to be the current law of the field.
+(This notation is highly efficient).}
+
+\textcolor{red}{When one want to consider several steps, one can use $\textrm{IBV}_{n}(\bm{x}; \bm{y})$ to denote the fact that the starting point (starting probability distribution) is the law of the field at the $n$-th step $\mathbb{P}_n$ to which conditioning on $\bm{y}$ at $\bm{x}$ is added.}
+
+\textcolor{red}{For a rigorous definition
+$$
+\operatorname{IBV}(\bm{x}; \bm{y}):=\int_{\domain}
+\mathbb{P}\left(\gp[\uu]\in T\mid Z_{\bm{x}}=\bm{y}\right)(1-\mathbb{P}\left(\gp[\uu]\in T \mid Z_{\bm{x}}=\bm{y}\right))
+d\mes(u) \\
+$$
+where $\mathbb{P}$ denotes the current law of the field.}
 

### 2020-06-18 main_new.tex
Updates from Overleaf

@@
 a discussion of future work.
-
-%\subsection{}
-
-%\begin{itemize}
-%    \item 
-%\end{itemize}
-
-\begin{comment}
-\section{Former Introduction (Before 08.06.2020)}
-
-Monitoring the world's oceans has gained increased importance in light
-of the changing climate and increasing anthropogenic impact. 
-A central problem for understanding these factors is the lack of representative data with sufficient resolution.
-Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of sampling.
-By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
-This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
-
-The
-advent of robust mobile robotic platforms \citep{Bellingham07} has
-resulted in significant contributions to environmental monitoring and
-sampling in the ocean. In particular, autonomous underwater vehicles (AUVs), have
-advanced the state of sampling and consequently have made robotics an
-integral part of ocean observation; our previous work has contributed
-to this effort \citep{das11b,Das2015,fossuminformation,fossum18b}.
-
-The potential that lies in using information-based sampling methods, which we
-highlight in this paper, is to do efficient environmental sampling and
-overcome common challenges as discussed next. First, full numerical ocean models based on complex
-differential equations cannot be run onboard a marine robot with
-limited computational capacity.
-% provide accurate results for online
-% operations because the onboard computers lack the computational power
-% required.
-Hence statistical proxy models of the environment must be used. For our purpose, we rely on Gaussian process (GP)
-representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
-Second, because of limited resources to sample the large oceanographic domains, one must plan for
-active learning during the operation, where a relevant criterion is
-used to extract the most valuable designs. The acquired in-situ
-information must then be assimilated into the statistical model
-onboard the platform, so that it can be used to re-compute the
-sampling criterion and in this way inform decisions on where to sample
-next. We use a criterion focusing on the prediction of excursion sets
-(ESs) for the multivariate spatial processes of interest. Third, the
-number of sampling designs is enormous, creating a trade-off between
-optimization and computability. One must often compromise in both
-development and practical operations, and we will rely on
-short-horizon optimization routines for sequential designs among
-nearest neighbors points in a spatial grid (waypoint graph).
-
-\kc{The para above needs a rewrite to me. It starts off by stating
-  there are challenges. But then jumps to models without stating the
-  connection with sampling, and then jumps into learning without
-  stating why learning is important. Typically the English in formal
-  text starts with a statement which is supported or clarified. In the
-  two cases above, the connections are simply not obvious. If this was
-  there in the first version of the m/s submitted, I'm not sure why
-  the reviewers didn't catch it. This is not just a stylistic issue;
-  it has to do with making sense and bring the reader/reviewer along.}
-
-\kc{Second, I don't see a motivation for GPs here and as I read
-  further into the text. Even for stats folks, are GPs the ``best''
-  ways to go? Are they obviously the tools one would use in
-  environmental sampling? And even if they are, I think we ought to
-  make the case for it strongly and generally, so this paper is
-  actually accessible and understandable to those not in the Spatial
-  Stats field. I would suggest adding a short para on GPs and why
-  they're important -- i.e. motivate the use of GPs early, so we can
-  freely make attributions to them later, as in the next section, for
-  example.}
-
-The imperfect understanding of causality coupled with the underlying
-fundamental uncertainty of the ocean implies that we need to make a
-large number of measurements. In addition, the use statistical models
-and methods to determine what and how phenomena form, is one way to
-increase the quality of predictive power aiding our understanding of
-upper water-column processes.
-
-With this in mind, we design AUV sampling strategies for reducing the spatially-integrated Bernoulli variance (IBV) in the ES
-indicator variables. 
-ES of random fields are thoroughly presented in \citep{adler2009random}. Recent work on ESs and associated excursion
-probabilities (EPs) connected to spatial statistics include
-\cite{french2013spatio,bolin2015excursion,french2016credible}. Our
-focus is on uncertainty characterization and how this is influenced by
-sampling. In this sense, our work is similar to
-\cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who
-describe analytical results for the IBV for univariate processes.
-
-A major contribution of our work is to derive closed-form results for
-this design criteria for situations where the underlying model is
-based on multivariate GPs. Another contribution is that we embed this
-approach in a sequential strategy to produce sampling algorithms, and
-apply this to a real-world scenario of autonomously sampling
-temperature and salinity gradients using an AUV, with the intention of
-characterizing the variability of oceanic and riverine waters in
-typical mixed coastal environments.
-
-Motivating examples relevant for ESs of multivariate processes are
-abundant. In medicine, doctors do not rely solely on a single symptom
-but must see several combined effects before making a diagnosis. In
-our context of environmental sampling, the salinity and temperature
-excursions of a river plume can similarly help characterize the
-underlying bio-geochemical processes
-\citep{hopkins2013detection,Pinto2018}, where sampling can be aimed
-towards reducing uncertainty in the temperature and salinity ES.
-Unlike what has been done in previous plume exploration studies, we
-emphasize the links to spatial statistical modeling and targeted
-multivariate sampling criteria for sequential sampling.
-
-Other statistical
-work in the oceanographic domain include \cite{wikle2013modern}
-focusing on hierarchical statistical models; \cite{sahu2008space},
-studying spatio-temporal models for sea surface temperature and
-salinity data; and \cite{mellucci2018oceanic} looking at the
-statistical prediction of features using an underwater glider.
-In this paper the focus is not on statistical modeling, but rather on statistical results and computations for efficient sampling designs.
-
-The remainder of this paper is organized as follows: Section
-\ref{sec:bg} provides select background on ocean sampling. Section
-\ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for
-vector-valued GPs. Section \ref{sec:heuristics} builds on these
-assumptions when deriving the sequential design criteria for adaptive
-sampling. Section \ref{sec:simulations} discusses properties of the
-methods in simulation studies. Section \ref{sec:case_study}
-demonstrates the methodology used in field work characterizing a river
-plume and finally, Section \ref{sec:concl_disc} contains a summary and
-a discussion of future work.
-\end{comment}
-
-\begin{comment}
-\section{Ocean Processes and Observation}
-\label{sec:bg}
-
-Central to understanding the changes taking place in the upper
-water-column is knowledge of the bio-geophysical interaction driven by
-an agglomeration of physical forcings (e.g. wind, topography,
-bathymetry, tidal influences, etc.) and incipient micro-biology driven
-by planktonic and coastal anthropogenic input, such as pollution and
-agricultural runoff transported into the ocean by rivers and streams.
-These often result in a range of ecosystem-related phenomena such as
-blooms and plumes, with direct and indirect effects on society. \kc{we
-  should look for a citation here. Check with John Ryan}
-
-Traditional data collection to understand such phenomena at sea has
-typically been based on static buoys, floats, or ship-based methods,
-with significant logistical limitations that directly impact coverage
-and sampling resolution. Modern methods using satellite remote-sensing
-provide large-scale coverage but have limited resolution, are limited
-to sensing the surface, and are impacted by cloud cover. Numerical
-ocean models similarly find it challenging to provide detail at fine
-scale \citep{Lermusiaux:2006}. Mobile robotic platforms
-\citep{Bellingham07} have contributed significantly to environmental monitoring and sampling. In particular, AUVs have
-advanced the state of in-situ sampling, and made robotics an integral
-part of ocean observation, filling parts of the undersampling gap
-\cite{rudnick03,rudnick18}(See Fig. \ref{fig:envir1}).
-
-%Walter Munk referred to the 20\textsuperscript{th} century
-%as the \emph{century of undersampling} \citep{munk2002}, pointing out
-%a lack of sufficient sampling resolution in time and space as a
-%critical issue for characterizing ocean phenomena. Even today,
-%policymakers struggle with limited observability of severe
-%environmental problems; a prominent example is the recent (May 2018)
-%outbreak of harmful algal blooms in Northern Norway\footnote{Report
-%  from the Institute of Marine Research: https://bit.ly/2KZGJfI}, with
-%losses exceeding 12 million tons of salmon valued at \$80 million
-%(Norwegian Directorate of Fisheries). This is a case in point of how
-%adequate sampling efforts could have provided a more accurate
-%assessment and early warning of changing environmental conditions.
-
-\begin{figure}[!h] 
-  \centering 
-  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
-    0.49\textwidth]{Figures/envir.pdf}\label{fig:envir1}}
-  \hfill
-  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
-    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
-  \caption{\ref{fig:envir1} Traditional ocean observation based on 
-    ship-based sampling has been augmented by autonomous
-    robotic vehicles. % and their interactions.  
-    AUV platforms are an integral part of this network being able to
-    reason and make decisions for efficient onboard adaptive sampling.
-    % using the sense-plan-act control approach to autonomous control. 
-    \ref{fig:nidelven} The interaction of river and ocean creates
-    processes that are challenging to map, where the combination of
-    statistics and robotics can play a vital role in enabling more
-    effective oceanographic observation.\kc{The inset ``AUV robot''
-      looks very tacky and ad-hoc; might be best to remove it. Change
-      ``ships'' to
-      ``Ship'', remove the aerial drone, since we didn't use it and
-      don't plan to in this work.}}
-  \label{fig:envir} \end{figure}
-
-\end{comment}
-
-\begin{comment}
-Surveys with AUVs are usually limited to observations along fixed
-transects that are pre-scripted in mission plans created manually by a
-human operator. The mission can be specified operating on a scale of
-hundreds of meters to tens of kilometers depending on the scientific
-context. Faced with limited coverage capacity, a more effective
-approach is to instead use onboard algorithms to continuously
-evaluate, update, and refine future sampling locations (sense-plan-act
-cycle in Fig. \ref{fig:envir1}), making the information gathering
-\emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
-% This usually occurs on a spatial grid, called a waypoint graph. 
-In doing so, the space of sampling opportunities is still limited by
-the waypoint graph, but the AUV has flexibility to modify its path at
-each waypoint based on in-situ sensing and measurements onboard
-\citep{py10,Rajan12,Rajan12b}.
-
-We focus on spatial characterization of a frontal system generated by
-river plumes. Fig. \ref{fig:nidelven} shows the survey area in
-Trondheim, Norway, where cold freshwater enters from the river,
-creating a strong gradient in both temperature and salinity. Because
-of the local topography and the Coriolis force the cold fresh water
-tends to flow to the east. Depending on the river discharge, tidal
-effects, wind, and temperature differences, this boundary often gets
-distorted. Initial knowledge about the location and evolution of these
-features are highly uncertain, making deterministic planning
-challenging.
-
-Adaptive in-situ AUV sampling of an evolving frontal feature has been
-explored in \cite{fronts11,Zhang2012,Pinto2018,costa19}. These
-approaches typically use a reactive-adaptive scheme, whereby
-exploration does not rely on a statistical model of the environment,
-but rather adaptation is based on closing the sensing and actuation
-loop. Myopic sampling, i.e. stage-wise selection of the path (on the
-waypoint graph), has been used for surveys
-\citep{singh2009efficient,Binney2013} that focus largely on reducing
-predictive variance or entropy. These criteria are widely adopted in
-the statistics literature on spatio-temporal design as well
-\cite{bueso1998state,zidek2019monitoring}, but variance and entropy
-reduction are independent of the actual data realizations under the
-assumptions of GP models, so it has limited
-decisional flexibility. The use of data-driven adaptive criteria
-was introduced to include more targeted sampling of regions of
-scientific interest in \cite{Low2009} and \cite{fossuminformation}. In
-this paper, we focus on mapping the river plume by rewarding designs
-that improve the classification of ES in temperature and salinity.
-
-\end{comment}
-
-\input{section2}
-
-\begin{comment}
+%\input{section2}
+
 \section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
@@
 
-We define the random sets of interest in Section \ref{sec:ES}, while the closed-form expressions for the expected uncertainty reduction are derived in Section \ref{sec:EIBV}.
-Section \ref{Sec:UnivarEx} provides an instructive bivariate example relevant for sampling in the temperature and salinity application.
-
-\subsection{Excursion sets and Integrated Bernoulli variance}
-\label{sec:ES}
-
-Our objective is to leverage statistical tools onboard an autonomous
-robotic platform to characterize a river plume, focusing on spatial
-separation of cold freshwater from a river and warmer saline waters
-of a fjord. The characterization of uncertainties on the ES is a
-useful goal of this sampling strategy.
-
-In order to formalize the problem in mathematical terms, let us
-consider a multivariate (i.e., vector-valued) Gaussian random field
-$(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary
-set (typically in geostatistical studies, a geographical subset of interest) and the
-dimensionality of the output $\no \geq 1$ is arbitrary. We have $\no = 2$ in the example with temperature and salinity in the river plume.
-
-Assume furthermore that the set of interest
-is characterized in terms of a subset of values of the variables, say
-$\T \subset \R^{\no}$, and denote the corresponding set of index
-points by
+In section \ref{sec:bg_and_notation} we introduce the notation used in the rest of the paper, review (co-)kriging of multivariate gaussian random fields (GRFs) and present update formulae for cokriging.
+Uncertainty quantification (UQ) techniques on excursion sets of GRFs are presented in \ref{sec:set_uq}, in particular the Integrated Bernoulli Variance (IBV) and the excursion measure variance (EMV) are defined and semi-analytical expressions easing their computation and optimization are derived.
+Section \ref{sec:eibv} turns to the  effect of new observations on EMV and IBV.
+Specifically, the expected effect of the inclusion of a new set of observations on the EMV and IBV are considered and further semi-analytical expressions for the expected
+uncertainty reduction are derived. The resulting formulae form the backbone of the uncertainty reduction strategies presented in Section \ref{sec:heuristics}.
+Finally, Section \ref{Sec:UnivarEx} illustrates the application of the above concepts on a bivariate example relevant for sampling in the temperature and salinity application.
+
+\begin{comment}
+\medskip
+
+%Our objective is to leverage statistical tools onboard an autonomous robotic platform to characterize a river plume, focusing on spatial separation of cold freshwater from a river and warmer saline waters of a fjord.
+
+The temperature and salinity field over the domain of interest will be modeled as a bivariate gaussian random field.
+The \textcolor{red}{complement of the} river plume may then be characterized as an \textit{excursion set} of the field, that is, as the region of the domain for which temperature and salinity lie above some specified threshold.
+Our goal is then to guide the data collection process so as to improve the characterization of this set.
+The methods we develop are fairly general, and not limited to the special case of two-dimensional random fields.
+Hence, the rest of this section will consider generic excursion sets of gaussian random fields with an arbitrary number of output dimensions.
+
+\begin{figure}[h!] \centering
+  \includegraphics[width=0.99\textwidth]{Figures/example_excu_1.png}
+  \caption{Realization of a bivariate GRF and excursion set above some threshold. Joint excursion in red, excursion of a single variable in white.}
+\label{example_excu}
+\end{figure}
+
+Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
+and we are interested in recovering the set of locations $\es$ in the domain for which the response variables (the components of $\gp$) lie in some set of specified values $\T\subset \mathbb{R}^{\no}$:
 $$
-\es=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
+\es:=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
 $$
 Our goal here is to develop approaches to quantify and improve the
-characterization of uncertainties on such sets $\es$, in the
-context of multivariate random fields.
-
-While some aspects of the developed approaches do not call for a
+characterization of uncertainties on $\es$. \textcolor{red}{Alread mention how observations are included.}
+
+If we assume that $\gp$ has
+continuous trajectories (almost surely) and $T$ is closed, then
+$\es$ becomes a Random Closed Set
+\citep{Molchanov2005} and concepts from the theory of random sets will prove useful in characterizing the distribution of the volume of $\es$ under some measure defined on the domain.
+
+\textcolor{red}{Is this really needed?} Note that while some aspects of the developed approaches do not call for a
 specific form of $\T$, we will often stick for simplicity to the case
@@
 of several key quantities. Note that changing some $\leq$ inequalities
-to $\geq$ ones would lead to immediate adaptations. In the application
-specific properties following the orthant assumption come into play
-\kc{This last phrase doesn't parse; is it ``..., the following
-  orthant....''?}: We let $t_1$ be a threshold for temperature and
-$t_2$ a threshold for salinity \kc{Might be easier to remember with
-  symbols 't' and 's' in some
-  form.}. The set of index point then defines
-$\mbox{ES} = \{\x \in D : \gp[\x,1] \leq t_1,\gp[\x,2] \leq t_2\}$.
-The associated EPs are
-$p(\x) = P(\gp[\x,1] \leq t_1,\gp[\x,2] \leq t_2)$,
-$\x \in \mathcal{M}$.
-
-For statistical generality, we will however keep the abstract notation
-with $\T$ and general $\es$ \kc{Again, previous phrase doesn't parse.}
-for most of the derivations. Note that assuming that $\gp$ has
-continuous trajectories (almost surely) and $T$ to be closed, then
-$\es$ can be cast as a Random Closed Set
-\citep{Molchanov2005}. %In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
-Let us in particular consider a locally finite (Borel) measure $\mes$
-on $D$ and investigate the probability distribution of $\mes(\es)$
-through its moments.
+to $\geq$ ones would lead to immediate adaptations.
+
+\end{comment}
+\subsection{Background, Notation and Co-Kriging}
+\label{sec:bg_and_notation}
+
+Throughout the rest of the article, we denote by $\gp$ a vector-valued random field indexed by some arbitrary domain $\domain$, and assume values of the field at any fixed location $\x \in \domain$, denoted $\gp[\x]$, to be a $\no$-variate random vector ($\no\geq 2$).
+In our leading river plume characterization application, $\domain$ is a prescribed domain in the fjord of Trondheim (for the purpose of our AUV application, a discretization of a $2$-dimensional domain at fixed depth is considered), while $\no=2$ and the two responses of interest stand for temperature and salinity.
+A bivariate gaussian random field model is assumed for $\gp$, see the two first panels of Figure~\ref{example_excu} for an illustration of a realization of such a vector-valued model (RK NOTATION).
+The right panel of this figure represents a by-product of interest derived from these realizations, namely regions i) in red, where both temperature and salinity are high (i.e., exceeding respective thresholds), standing for ocean water ii) in blue, where both temperature and salinity are low, standing for riverine water, and iii) in white, where one variable is above and the other below their respective thresholds, standing for mixed waters.
+
+\begin{figure}[h!] \centering
+  \includegraphics[width=0.99\textwidth]{Figures/example_excu_1.png}
+  \caption{Realization of a bivariate GRF and excursion set above some threshold. Joint excursion in red, excursion of a single variable in white.}
+\label{example_excu}
+\end{figure}
+
+%The methods we develop are fairly general, and not limited to the special case of two-dimensional random fields.
+Hence, the rest of this section will consider generic excursion sets of gaussian random fields with an arbitrary number of output dimensions.
+% Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
+Coming back to the more general settings of a $\no$-variate random field, 
+we are interested in recovering the set of locations $\es$ in the domain for which the response variables (the components of $\gp$) lie in some set of specified values $\T\subset \mathbb{R}^{\no}$, in other words \textit{the pre-image of $T$ by $\gp$}:
+$$
+\es:=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
+$$
+%
+%Our goal here is to develop approaches to quantify and improve the
+%characterization of uncertainties on $\es$. \textcolor{red}{Alread mention how observations are included.}
+
+If we assume that $\gp$ has
+continuous trajectories (almost surely) and $T$ is closed, then
+$\es$ becomes a Random Closed Set
+\citep{Molchanov2005} and concepts from the theory of random sets will prove useful in characterizing the distribution of the volume of $\es$ under some measure defined on the domain.
+%\textcolor{red}{Is this really needed?} 
+Note that while some aspects of the developed approaches do not call for a
+specific form of $\T$, we will often stick for simplicity to the case
+of orthants
+($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
+$t_1,\dots, t_{\no} \in \R$) as this will allow efficient calculation
+of several key quantities. Note that changing some $\leq$ inequalities
+to $\geq$ ones would lead to immediate adaptations.
 
@@
 
-%That $\mu(\es)$ is a well-defined random variable follows from
-\begin{propo}
-	\label{propo1}
-	%Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
-	For a measurable random field $\gp$ and a locally finite
-        measure $\mes$ on $\mathcal{M}$, $\mes(\es)$ is a random
-        variable and for any $r\geq 1$, \begin{equation*}
-          \begin{split} \mathbb{E}[\mes(\es)^r]
-            &=\int_{\mathcal{M}^{r}} \jointExcuProb \productMeasure ,
-          \end{split} \end{equation*}
-
-	where the product measure is denoted as
-	$\nu^{\otimes}:=\bigotimes_{i=1}^r \nu$. 
-	%Note that we will use boldface to denote concatenated variables. % TO RELOCATE
-	Here $\gp$ is defined on $\mathcal{M}$, and for
-	$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in \mathcal{M}^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
-	\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.\kc{Is $\no r$ a
-        product of the $\no$ from above? Is that obvious?}
+%\subsubsection{Notation}
+%In order to work at this level of generality, some notational tricks are needed.
+Indeed, when considering a new observation, one shall in general specify which component of the field was observed and where it was observed.
+This leads us to introduce the concept of \textit{generalized location}.
+
+Letting $\gp[\spatloc,\ell]$ denote the $\ell\text{-th}$ component of $\gp[\spatloc]$, $\spatloc\in \domain$ and $\ell \in \{1\dots,p\}$, we will call \textit{generalized location} the couple $x=(\spatloc,\ell)$.
+%Given such a generalized location notation $x$,  
+The notation $\gp[x]$ will be used to denote $\gp[\spatloc,\ell]$ 
+and
+%
+%This slight change of notation 
+will allow us to think of $\gp$ as a scalar-valued Gaussian random field 
+indexed by $D\times \{1\dots,p\}$, which will give the co-kriging equations a particularly simple form that parallels the one of univariate kriging.
+Due to the naturality of the concept of \textit{generalized location},
+we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $\spatloc \in \domain$.
+In the following, the letter $x$ will be reserved for generalized locations, while the letters $\spatloc$ and $\ell$ will be used for spatial locations and response indices respectively.
+
+It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the 
+one for a single observation, provided some notation adjustments are made. This motivates the following.
+
+Given a dataset consisting of  $q$ observations at spatial locations $\spatloc_i \in \domain$ and response indices $l_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we can concatenate the information needed to specify the dataset by using the notation
+\begin{align*}
+\bm{x}:=(\bm{\spatloc}, \bm{\ell}):= (x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
+\end{align*}
+In general, boldface letters will be used to denote concatenated quantities corresponding to batches of observations. 
+We can then also compactly denote the values of the field at those different locations by
+\begin{align*}
+\gp[\bm{x}]:=
+\left(\gp[\spatloc_1,\ell_1], ...,
+\gp[\spatloc_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
+\end{align*}
+%
+
+Given a random field $\gp$ and (noisy) observations of some of its components at some points in the domain, one can predict the value of the field at some unobserved location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$, conditional on the data.
+This process is called (co-)kriging, and kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
+We will here present the most general form of cokriging, where one can include several observations (batch) at a time and 
+observations at a given location $s \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ 
+(heterotopic).
+
+\subsubsection{Co-Kriging}
+\label{sec:cokriging}
+Assuming now that $n$ batches of observations are available, with respective sizes $q_1,\dots, q_n$, and that one wishes 
+to predict $\gp[\bm{x}]$ for some batch of $q\geq 1$ generalized locations $\bm{x}=(\bm{\spatloc}, \bm{\ell})$, the (simple) 
+cokriging mean would then amount to simple kriging with respect to a scalar-valued Gaussian random field indexed by 
+$D\times \{1\dots,p\}$ and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{\spatloc}, \bm{\spatloc}')_{\ell, \ell'}$, that is:
+%
+\begin{equation}
+\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})),
+\end{equation}
+where $\mu$ is $Z$'s initial mean function, $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of 
+observed responses of $Z$ at all considered generalized locations, and $\lambda_{[n]}(\bm{x})$ is a vector of weights 
+equal to $k(\bm{x}_{[n]}, \bm{x}_{[n]})^{-1} k(\bm{x}_{[n]}, \bm{x})$ with $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$, 
+$k(\bm{x}_{[n]}, \bm{x}_{[n]})$ being assumed non-singular throughout the presentation. The co-kriging %(conditional) 
+residual (cross-)covariance function (with respect to batches of generalized locations) can also be expressed in the same vein via
+%
+\begin{equation}
+k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T k(\bm{x}_{[n]}, \bm{x}_{[n]}) \lambda_{{[n]}}(\bm{x}').
+\end{equation}
+
+\subsubsection{Co-kriging update formulae}
+
+Let us now consider the case where co-kriging prediction of $Z$ was made with respect to $n$ batches $\bm{x}_i$ of generalized locations, concatenated again within
+$\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$, and one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
+%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
+Thanks to our representation of co-kriging in terms of simple kriging with respect to generalized locations, a strightforward adaptation of the batch-sequential kriging update formulae from \cite{Chevalier.etal2013a} delivers that
+% 
+\begin{equation}
+\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
+\end{equation}
+where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
+$\lambda_{[n+1]}(\bm{x})$ that corresponds to the kriging weigths associated with the last $q_{n+1}$ responses when 
+predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
+%\text{th}$ batch.
+%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
+Similarly, the updated co-kriging residual (cross-)covariance function then writes
+\begin{equation}
+k_{[n+1]}(\bm{x},\bm{x}')=k_{[n]}(\bm{x},\bm{x}')-\lambda_{[n+1,n+1]}(\bm{x})^T k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]}) \lambda_{{[n+1,n+1]}}(\bm{x}').
+\end{equation}
+%\medskip
+%
+Let us remark that, as noted in \cite{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
+extend to Universal Kriging in second-order settings and apply without Gaussian assumption.
+We will now see how the latter formulae are instrumental in deriving semi-analytical formulae for stepwise uncertainty reduction criteria for
+vector-valued random fields.
+
+\subsection{Uncertainty Quantification on Excursion Sets of multivariate Gaussian Random Fields}
+\label{sec:set_uq}
+We now introduce quantities that allow to quantify the uncertainty on the volume of the excursion set $\es$. Let $\mes$ be a 
+(locally finite, Borel) measure  on $\domain$. We want to investigate the probability distribution 
+of $\mes(\es)$ through its moments.
+
 \medskip
 
-	In the particular case where $\gp$ is a multivariate Gaussian random field 
-	%and denoting the mean and covariance matrix of $\gp[\bm{u}]$ by $\meanUU\in\mathbb{R}^{p\times r}$ and $\covUU$, respectively, 
-	we have 
-	\begin{align*}
-	\jointExcuProb = \mathcal{N}_{\no r}(T^r; \meanUU, \covUU),
-	\end{align*}
-	where $\mathcal{N}_{\no r}(\cdot ; \meanUU, \covUU)$ is the Gaussian measure on $\mathbb{R}^{\no r}$ with mean $\meanUU$ and covariance matrix $\covUU$, respectively defined blockwise by  
-	%where both are defined blockwise by 
-	\begin{align*}
-	\meanUU&=\begin{pmatrix}\mu(u^{(1)})\\ \vdots\\ \mu(u^{(r)})\end{pmatrix}
-	\in \mathbb{R}^{\no r}, \\
-	\text{and } \covUU &= \begin{pmatrix}
-	\cov(\gp[u^{(1)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(1)}],
-	\gp[u^{(r)}])\\
-	\vdots & & \vdots\\
-	\cov(\gp[u^{(r)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(r)}],
-	\gp[u^{(r)}])\\
-	\end{pmatrix}\in \mathbb{R}^{pr\times pr},
-	\end{align*}
-	each of the $r\times r$ blocks of the latter matrix being itself a (cross-)covariance matrix of dimension $\no \times \no$. 
-	%dimensional Gaussian random vector. 
-	Assuming further that $\covUU$ is non-singular, the probability of interest can be formulated in terms of the $\no r$-dimensional Gaussian probability density function 
-	$\varphi_{\no  r}(\cdot;~\meanUU, \covUU)$ as 
-	%$\varphi_{\no \times r}$ as
-	%$\bm{u}\in D^r$ such that $\gp[\bm{u}]$ be non-degenerate, the integrand can be expressed in terms of the $r \times \no$-dimensional Gaussian distribution, namely 
-	%cumulative distribution function, namely 
-	\begin{equation*}
-	\begin{split}
-	&\jointExcuProb
-	=
-	\int_{T^r} \varphi_{\no  r}\left(\bm{v};
-	~\meanUU, \covUU\right) 
-	\mathrm{d}\bm{v},
-	\end{split}
-	\end{equation*}
-	In the particular orthant case with $\T=(-\infty, t_1] \times \dots \times (-\infty, t_{r}]$,
-	the latter probability directly writes \kc{What does
-          ``writes'' mean here? Connotes?} in terms of the multivariate Gaussian
-	cumulative distribution, % with mean $\meanUU$ and covariance matrix $\covUU$,
-	this time without requiring $\covUU$ %the latter 
-	to be non-singular: 
-	\begin{equation*}
-	\begin{split}
-	\jointExcuProb 
-	&=
-	\varPhi_{\no r}\left(\bm{t};~\meanUU, \covUU\right),
-	\end{split}
-	\end{equation*}
-	where we have used the notations 
-	$t=(t_1,\dots,t_{\no}%,..., t_1, ..., t_p
-	)\in\mathbb{R}^{\no}$, $1_{r}=(1,\dots,1)\in \R^{r}$, and 
-	$\bm{t}=1_{r}\otimes \bm{t}=(t_1,\dots,t_{\no},\dots,t_1,\dots,t_{\no}) 
-	\in \R^{\no r}$.
-\end{propo}
-
-Proposition~\ref{propo1} enables one to calculate centered moments of $\mes(\es)$. 
-The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$
-is probably the most relevant of these for our purposes. We
-have 
+Centred moments may be computed using Proposition~\ref{propo1} developed in the appendix. 
+In particular, it allows to write the excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ as an integral of the excursion probability
 \begin{equation*}
@@
 \emv
-&=\int_{\mathcal{M}^2} \mathbb{P}\left(
+&=\int_{\domain^2} \mathbb{P}\left(
 \gp[u]\in T, \gp[v]\in T \right)
 d\mes^{\otimes}(u, v)\\
-&-\left( \int_{\mathcal{M}} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
+&-\left( \int_{\domain} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
 \end{split}
 \end{equation*}
-which evaluates in the excursion/sojourn case where
-$\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ to
-\begin{equation*} \begin{split}
+which boils down in the excursion/sojourn case where $\T=(-\infty, t_1] \times
+\dots \times (-\infty, t_{\no}]$ to
+\begin{equation*}
+\begin{split}
 \emv
 %\operatorname{Var}[\mes(\es)]
-&=\int_{\mathcal{M}^2} 
+&=\int_{\domain^2}
 \varPhi_{2\no}
 \left(
-(\bt, \bt); \bmu((u,v)), 
+(\bt, \bt); \mu((u,v)),
 K((u,v),(u,v))
@@
 (u,v)\\
-&-\left( \int_{\mathcal{M}} \varPhi_{\no}\left(\bt;\bmu(u), K(u)\right) d\mes(u) \right)^2.
+&-\left( \int_{\domain} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
 \end{split}
 \end{equation*}
-This quantity requires  an integral over $D^2$. In
-contrast, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal}
-involves solely an integral on $\mathcal{M}$ and can be expanded in
-our present setting as follows 
-\begin{equation}\label{IBV}
+%
+Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $\domain^2$. In 
+contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves 
+solely an integral on $\domain$ and can be expanded in our present settings as follows
+\begin{equation*}
 \begin{split}
 \operatorname{IBV} %(\es) %;\mes)
-&=\int_{\mathcal{M}}
+&=\int_{\domain}
 \mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
 d\mes(u) \\
-&=\int_{\mathcal{M}}
-\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) 
--\left(\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) \right)^2
+&=\int_{\domain}
+\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right)
+-\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
 \mathrm{d}\mes(u).
 \end{split}
+\end{equation*}
+%
+Next we investigate uncertainty reduction criteria related to those indicators.
+
+\subsection{Expected Integrated Bernoulli Variance and Excursion Measure}
+\label{sec:eibv}
+This section is devoted to computing the expected effect of the inclusion of new observations on the $\emv$ and $\ibv$ of the excursion set $\es$.
+Let us consider the same setting as in Section \ref{sec:cokriging} where $n$ batches of measurements have already been performed, and let
+$\currentExp{.}$ and $\currentProba{.}$ denote conditional expectation
+and probability conditional on the first $n$ batches of observations, respectively.
+We want to co pute the effect of the inclusion of a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and $\ibv$.
+
+For uncertainty functional such as $\ibv$ and $\emv$, we use a similar indexing system
+with respect to a sequence of batch, and further indicate when applicable the new batch
+of generalized location between parentheses; e.g., $\emv_{[n+1]}(\bm{x})$ denotes the value
+of $\emv$ (a random variable, seen from ``step'' $n$, i.e. when knowing $Z$ solely at
+the $n$ first generalized location batches) when adding observations at $\bm{x}$ to those
+already available at step $n$.
+
+\textcolor{red}{C.T.: This doesn't work, we cannot be that short on the definition of EIBV, it is highly unclear. I suggest we go for the following notation:
+$$\operatorname{IBV}(\bm{x}; \bm{y})
+$$ denotes the IBV under the current law of the field, conditioned on observing $\bm{y}$ at $\bm{x}$ (generalized, possibly batch observation). Then
+$$
+\operatorname{EIBV}(\bm{x}):=\mathbb{E}_{\bm{Y}}\left[\textrb{IBV}(\bm{x}; \bm{Y})\right]
+$$
+where $\bm{Y}$ is distributed according to the current law of $Z_{\bm{x}}$.}
+
+\textcolor{red}{
+The philosophy is the following: EIBV and EEMV are functionals of a probability distribution.
+Then, if not mentioned explicitly, the probability distribution will be taken to be the current law of the field.
+(This notation is highly efficient).}
+
+\textcolor{red}{When one want to consider several steps, one can use $\textrm{IBV}_{n}(\bm{x}; \bm{y})$ to denote the fact that the starting point (starting probability distribution) is the law of the field at the $n$-th step $\mathbb{P}_n$ to which conditioning on $\bm{y}$ at $\bm{x}$ is added.}
+
+\textcolor{red}{For a rigorous definition
+$$
+\operatorname{IBV}(\bm{x}; \bm{y}):=\int_{\domain}
+\mathbb{P}\left(\gp[\uu]\in T\mid Z_{\bm{x}}=\bm{y}\right)(1-\mathbb{P}\left(\gp[\uu]\in T \mid Z_{\bm{x}}=\bm{y}\right))
+d\mes(u) \\
+$$
+where $\mathbb{P}$ denotes the current law of the field.}
+
+\begin{propo}
+\label{propo_eibv}
+Let $\eibv_{[n]}(\bm{x})=\currentExp{\ibv_{[n+1]}(\bm{x})}$ (where $\bm{x}$ plays
+the role of $\bm{x}_{n+1}$) be the expected effect of adding an observation a $\bm{x}$ on the IBV.
+Then we have
+\begin{equation}
+\begin{split}
+\eibv_{[n]}(\bm{x})
+&=\int_{D} \varPhi_{\no}\left(\bt;~\currentMean{\uu}, \currentCov{u, u}\right) d\mes(u)\\
+&-\int_{D} \varPhi_{2\no}
+\left(
+\left(
+\begin{matrix}
+\bt-\currentMean{u}\\
+\bt-\currentMean{u}
+\end{matrix}
+\right);
+\mathbf{\Sigma}_{[n]}(\uu)
+\right)
+d\mes(u),
+\end{split}
 \end{equation}
-
-\subsection{Expected Integrated Bernoulli Variance and Sampling Designs}
-\label{sec:EIBV}
-
-Our goal is to construct informative sampling designs, given limited
-resources \kc{information?}. In our domain, data is gathered by an AUV
-that can measure temperature and salinity at chosen locations on a
-defined waypoint graph. Such a discretized grid fosters reliable
-navigation and mission performance. The waypoints further enable a
-structure to the design problem in that, that sampling flexibility is
-limited to operation on the graph. % This gives lots of sampling
-% opportunities, and it is important to spend the sampling efforts
-% wisely.
-We focus on sampling designs that reduce the expected integrated
-Bernoulli variance (EIBV). Note that this sampling can be done in a
-static manner where the survey design is pre-planned, or in an
-adaptive setting. For the latter case, data can be processed
-sequentially and the updated model is then used to design and select
-the next waypoint graph to visit, which in turn are used again to
-update the model, and so on (Section \ref{sec:heuristics}). We next
-describe the data modeling assumptions for one stage and develop
-design criteria of semi-analytical forms relying on integrals of
-multivariate Gaussian cumulative distribution functions; we will
-appeal to some key intermediate results presented next.
-
-We denote a design by $d \in \mathcal{D}$, where the concatenation of designs with $q$ sampling locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$ is $\mathcal{D}$. The
-measurement model is defined by 
-\begin{equation}\label{eq:meas}
-    \bY_d=G_d \bz(\bu_d) + \epsilon, \hspace{3mm} \epsilon \sim N(0,R_d).
-\end{equation}
-Here, $\bz(\bu_d)$ is the size $pq$ vector of GP variables at the
-sampling locations, and $G_d$ is the design matrix which in principle
-allows for observing subsets of the vector-valued field at some of the
-$q$ locations (i.e. the matrix has fewer rows than columns). If the
-additive error has independent elements, the covariance matrix $R_d$
-is diagonal. In our setting temperature and salinity measurements are
-extracted from different sensors, so the independence assumption
-between variables is not unrealistic, but this is not necessarily the
-case for other kinds of measurements on the AUV. It is further common
-in geostatistics to assume spatially independent measurement errors,
-and we also do so. But our methodology for constructing sampling
-designs does not rely on these independence assumptions.
-
-For a given design $d$ the evaluation is based on the expected
-information contained in the data to influence the criterion.   
-Considering the EIBV criteria, we have
-\begin{eqnarray}\label{eq:eibv}
-    \mbox{EIBV} (d) &=& E_{\bY_d} \left( \int_{\mathcal{M}}
-p(\x;\by_d)(1-p(\x;\by_d))
-d\mes(u) \right), \nonumber \\
-p(\x;\by_d) &=& \mathbb{P}\left(\gp[\uu]\in T|\bY_d=\by_d \right)
-\end{eqnarray}
-where the expectation is over the planned data.  The optimal design
-for an AUV survey is then 
-\begin{equation}\label{crit}
-    d^* = \mbox{argmin}_{d \in \mathcal{D}} \mbox{EIBV}(d).
-\end{equation}
-For another criterion using random moments of random sets similar
-expectations would be required, but as we will see, the EIBV has a
-closed-form expression which is relatively fast to compute and hence
-useful for onboard design selection.
-
-Assuming now that $q$ observations $\by_d$ made at $\bu_d$ are available, one wishes to predict $\gp[\x]$. The multivariate Kriging equations holds with size
-$p$ vector of conditional means:  
-\begin{equation}\label{eq:cokrig}
-\bbm(\x)=\bmu(\x)+\lambda(\x,\bu_d)^T \left( \by_d-G_d \bmu(\bu_d)\right),
-\end{equation}
-where
-$\lambda(\x,\bu_d)=K^{-1}_y G_d k(\bu_d, \bu) $ with data covariance
-$K_y=\mbox{Cov}(\bY_d,\bY_d)=G_d K(\bu_d,\bu_d)G_d^T+R_d$ and cross-covariance $\mbox{Cov}(\bY_d,Z(\x))=G_d k(\bu_d, \x)$, assuming the measurement
-error is independent of the random fields.  The Kriging (cross-)covariance function can also be expressed in the same vein via
-a $p \times p$ matrix 
-\begin{equation}\label{eq:krig:cov}
-\Sigma(\x,\x')=K(\x,\x')-\lambda(\x,\bu_d)^T K_y(\bu_d, \bu_d) \lambda(\x,\bu_d).
-\end{equation}
-
-Eq. \eqref{eq:cokrig} and \eqref{eq:krig:cov} are used to
-provide informative, updated statements about sets of interest. For
-our case with temperature and salinity, the conditional EPs, given
-measurements $\by_d$, can then be computed from
-\begin{eqnarray}\label{eq:post_ep}
- p(\x;\by_d) &=& P(Z_1(\x) \leq t_1, Z_{2}(\x) \leq t_2 |\bY_d=\by_d),  \nonumber \\
- &=& \Phi_2 ((t_1,t_2)^T-\bbm(\x);0_2, \Sigma(\x,\x)). 
-\end{eqnarray}
-
+where the matrix $\mathbf{\Sigma}_{[n]}(\uu)$ is defined as
+\begin{equation*}
+\begin{split}
+\mathbf{\Sigma}_{[n]}(\uu)&=
+\left(
+\begin{matrix}
+\currentCov{u, u} & \currentCov{u, u}-\futureCov{u, u}\\
+\currentCov{u, u}-\futureCov{u, u} & \currentCov{u, u}
+\end{matrix}
+%\begin{matrix}
+%\currentCov{u, u} & \Delta_{[n]}(\uu)\\
+%\Delta_{[n]}(\uu) & \currentCov{u, u}
+%\end{matrix}
+\right).\\
+\end{split}
+\end{equation*}
+\end{propo}
+
+As for the expected excursion measure variance, a similar result may be derived.
+\begin{propo}
+\label{propo_emv}
+
+\begin{equation*}
+\begin{split}
+\eemv_{[n]}(\bm{x})
+&=\int_{D^2} 
+\varPhi_{2\no}
+\left(
+(\bt, \bt); \mu((u,v)), 
+K((u,v),(u,v))
+\right) 
+\
+\mathrm{d}\mes^{\otimes} %\mes 
+%\productMeasure 
+(u,v)\\
+&-\currentExp{
+    \varPhi_{\no}\left(a_1+B_1 V; C_1\right)
+    \varPhi_{\no}\left(a_2+B_2 V; C_2\right)
+},
+\end{split}
+\end{equation*}
+
+with $V=\gp[\bm{x}_{n+1}]-\currentMean{\bm{x}_{n+1}} \sim \mathcal{N}(0_{q_{n+1}},k_{[n]}(\bm{x}_{n+1},\bm{x}_{n+1}))$ and $a_1=\bt-\currentMean{\uu}$,
+$B_1=-\lambda_{[n+1,n+1]}(\uu)^T$, $a_2=\bt-\currentMean{\vv}$, $B_2=-\lambda_{[n+1,n+1]}(\vv)^T$.
+
+This integrand finally boils down to an evaluation of the $2\no$-dimensional Gaussian CDF by
+Proposition~\ref{propo3}. Details are omitted here for brevity as we did not implement further
+this criterion that requires numerical integration over $\domain^2$.
+\end{propo}
+
+Note that the calculation of the multivariate Gaussian cumulative distribution
+may be done effectively using code such as that of
+\cite{genz2009computation}.
+
+\textcolor{red}{Do we really want to keep this?}
 A critical element in the derivation of a closed-form for the EIBV is
-that the conditional mean in Eq. \eqref{eq:cokrig} is a linear
+that the conditional mean in the cokriging equations is a linear
 (affine) function of the data $\by_d$ and the covariance is not a
@@
   chevalier2014fast,stroh}, but not generalized to our situation with
-random sets for vector-valued GPs. Notably, the entry in Eq.
-\eqref{eq:post_ep} is in the form of $\ba+B V$, where
-$V \sim N(0_q, K_y (\bu_d,\bu_d))$,
-$\ba=(t_1,t_2)^t-\bmu(\x)$
-and $B=\lambda(\x,\bu_d)^T$. The following proposition then applies to give a
-closed-form solution to EIBV calculation in this case.
-
-\begin{propo}
-	\label{propo2}
-	Let $p, q, h \geq 1$, $a \in \R^p$, $B \in \R^{p\times q}$,
-	and $\covN, \covV$ be two covariance matrices in
-	$\R^{p\times p}$ and $\R^{q\times q}$, respectively.
-	Then, for $V \sim \mathcal{N}_{q}(0_q, \covV)$,
-	\begin{equation*}
-	\mathbb{E}\left[ \varPhi_{p}\left( \ba + BV; 0_p,\covN \right)^h \right]
-	=
-	\varPhi_{ph}
-	\left(
-	\ba
-	;0_{ph},
-	\Sigma
-	\right),
-	\end{equation*}
-	where the vector $\ba \in \R^{p h}$ is defined as 
-	$\ba := 1_h\otimes a = 
-	\left(a, \dots , a
-	\right)'$
-	and the $p h\times p h$ covariance matrix is given by
-	$\Sigma := 
-	1_h 1_h'\otimes B\covV B' + I_h\otimes \covN$.
-\end{propo}
-The proof of Proposition \ref{propo2} is provided in the Appendix.
-
-In block-wise representation, the following relation holds for the
-covariance in Proposition \ref{propo2}: 
-\begin{align*}
-	\Sigma&=
-	\begin{pmatrix}
-	\covN & &\\
-	& \ddots &\\
-	&   & \covN
-	\end{pmatrix}
-	+
-	\begin{pmatrix}
-	B\covV B' & \dots & B\covV  B'\\
-	\vdots & & \vdots\\
-	B\covV B' & \dots & B\covV B'\\
-	\end{pmatrix}.
-\end{align*}
-The calculation of the multivariate Gaussian cumulative distribution
-is done effectively using code such as that of
-\cite{genz2009computation}.
-
-In the more general case, for instance that of volumes of random sets,
-the criterion will include couplings with different locations and the
-result must be generalized. For the case of multivariate monomials in
-orthant probabilities with thresholds affine in a common Gaussian
-vector \kc{This sentence is dangling. Doesn't have a conclusion.}.
-%Namely, for monomials of degree $k$, we have.
-
-\begin{propo}%[\textcolor{red}{Notation in progress, to be reviewed/revised}]
-	\label{propo3}
-	Let $g, p, q\geq 1$, $h_{1},\dots, h_{g}\geq 1$ with $H=\sum_{i=1}^g h_i$, $a_{i} \in \R^{p}$, $B_{i}\in \R^{p \times q}$, and covariance matrices $\covN_i \in \R^{p \times p}$ $(1\leq i \leq g)$.
-Then, for any covariance matrix $\covV \in \R^{q\times q}$ and $V\sim\mathcal{N}_{q}(0_q,\covV)$,
-	%	Let $g$ and $k$ be integers, and consider a $g$-terms monomial of degree $k$
-	%	\[
-	%	    \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V; \Sigma_{i}
-	%	    \right)^{h_i},
-	%    \]
-	%    where $\sum_{i=1}^g h_i = k$, each $a_i$, $B_i$ and $\Sigma_i$ is as in
-	%    \cref{propo2} and $V$ is a $q$-dimensional Gaussian vector distributed as
-	%    $V\sim\mathcal{N}(0,\covV)$.
-	%
-	% Then the expectation of the monomial is given by
-	\begin{equation}
-	\mathbb{E}\left[ \prod_{i=1}^{g} \varPhi_{p}\left(a_i + B_{i}V;0_p, \covN_{i} \right)^{h_i} \right]
-	=
-	\varPhi_{p H}
-	\left(
-	\bm{a}
-	;0_{pH},
-	\mathbf{\Sigma}
-	\right),
-	\end{equation}
-	with $\bm{a}=(1_{h_1}\otimes a_1, \dots, 1_{h_g}\otimes a_{g}) \in \R^{p H}$
-	and $\mathbf{\Sigma}\in \R^{p H \times p H}$ is defined blockwise by $(\Sigma_{i,j})_{i,j \in \{1,\dots, g\}}$ where, for any $i,j \in \{1,\dots, g\}$, %the blocks are given by 
-	\begin{equation}
-	\Sigma_{i,j}=
-	(1_{h_{i}}1_{h_{j}}')\otimes (B_{i}\Sigma_{V} B_{j}') + \delta_{i,j}(I_{h_{i}}\otimes \covN_{i}) \in \R^{p h_{i} \times p h_{j}}.
-	\end{equation}
-	%with $\bm{a}=\sum_{j=1}^g \mathbf{1}_{J_{j}} \otimes a_j \in \R^{p \times k}$ 
-	%where $J_{j}=\{\sum_{i=1}^{j-1}h_i + 1, \dots, \sum_{i=1}^{j}h_i\}$,
-	%$\mathbf{1}_{I_{j}}$ denotes the vector of $\R^{k}$ with components equal to $1$ for indices in $I_{j}$ and $0$ else ($1\leq j \leq g$), 
-	%and $\mathbf{\Sigma}\in \R^{p k \times p k}$ is defined blockwise by 
-	%\begin{equation}
-	%\mathbf{\Sigma}_{\widetilde{J}_{j_{1}},\widetilde{J}_{j_{2}}}=
-	%(\mathbf{1}_{h_{j_1}}\mathbf{1}_{h_{j_2}}')\otimes (B_{j_{1}}\Sigma_{V} B_{j_2}') + \delta_{j_{1},j_{2}}(I_{j_{1}}\otimes \Sigma_{j_{1}}),
-	%\end{equation}
-	%where $\widetilde{J}_{j}=\{p\sum_{i=1}^{j-1}h_i + 1, \dots, p\sum_{i=1}^{j}h_i\}$ ($1\leq j \leq g$).
-\end{propo}
-The proof of Proposition \ref{propo2} is provided in the Appendix.
-
-\subsection{An Example of EIBV using Temperature and Salinity}
+random sets for vector-valued GPs.
+
+\subsection{Expected Bernoulli Variance on a two dimensional Example}
 \label{Sec:UnivarEx}
-
-% In this section, 
-We calculate EPs and the Bernoulli variance for a few bivariate
-Gaussian distributions for temperature and salinity to illustrate the concepts articulated above.
-The parameter inputs are made up, but they are in a realistic range as we have in the simulation study and real-world example below.
-The thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
+We illustrate the EBV associated to different designs on a simple bivariate example.
+The example mimics our river plume application and hence the first and second component of the random field will be called \textit{temperature} and \textit{salinity} for illustrative purpose.
+Effects of the hyperparameters of the GRF prior on the excursion probabilities will also be studied.
+For simplicity, we begin with a \textit{pointwise} example, considering a single bivariate gaussian distribution (no spatiality).
+
+\subsubsection{A pointwise study}
+Say we want to study the excursion probability of a bivariate gaussian, where the excursion set is defined as
+\[
+\Gamma := \lbrace \spatloc \in \domain: Z_{\spatloc, 1} \geq t_1, Z_{\spatloc, 2}\geq t_2\rbrace
+\]
+and the thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
 
@@
 
-%\subsection{Expected classification criteria}
-
-%{\bf{I have not done anything here - skip this, I think}}.
-
-%Rather than minimizing expected variance one can aim at minimizing the classification probability of the excursion set: 
-%$\min [p,(1-p) ]$, see e.g. \cite{lilleborge2016information}.
-
-%Again, since the conditional mean is a linear in the data, we only have to look at the %relevant linear combination via $\bm_{\xi}=E(\bxi)$.
-%See also \cite{bhattacharjya2013value}.
-
-%\begin{equation}
-%E(\min \{ p,(1-p)\})=\int \min P(\xi_1 \leq t, \xi_2 \leq s),[1-P(\xi_1 \leq t, \xi_2 \leq s)] p(E(\bxi)=) dE(\bxi)=,
-%\end{equation}
-%for the complementary probabilities, this is again evaluated by the corner regions for the block, and $\Phi_4()$ evaluations are required.
-
-%In the end, this result is integrated over the spatial domain $x \in X$.
-\end{comment}
-
-\section{Sequential designs and heuristic path planning}\label{sec:heuristics}
-
-An important concept underlying this work is the potential of robotic
-sampling to adapt survey plans based on data and statistical methods
-in accordance with the sense-plan-act control loop
-(Fig. \ref{fig:envir1}). We now present a few adaptive strategies for
-robotic sampling that build on the notion of EIBV.
-
-\subsection{Optimal Sequential Design}
-\label{Optdes}
-
-Our adaptive AUV survey is split into many stages. The selection on
-where to sample is constrained by a triangular waypoint graph and
-defined by its nearest grid nodes.
-
-Let $d^{j,s}$ denote design number $j$ at stage $s$ of a sequential
-survey.
-For the AUV sampling situation, the choice of design entails a commitment to one of many possible navigation directions on the waypoint graph.
-Once the choice has been made, the AUV steers in the chosen direction and this entails sampling locations $\bu_{d^{j,s}}$ and data $\by_{d^{j,s}}$ is gathered.
-We let
-$\mathcal{Y}^{s-1} =\{(d^{*,1}, \by_{d^{*,1}}),\ldots, (d^{*,s-1},
-\by_{d^{*,s-1}}) \}$ denote the set of all data gathered until stage
-$s-1$ according to selected design $d^{*,1},\ldots,d^{*,s-1}$.
-
-The design and data define
-the conditional GP with mean
-$E(\gp[\x]|\mathcal{Y}^{s-1})$ and (cross) covariance matrix
-$\mbox{Var}(\gp[\x]|\mathcal{Y}^{s-1})$.
-These are efficiently updated in a sequential manner keeping track of the mean and covariance kernel of the GP over the stages.
-As extensions of Eq.
-\eqref{eq:cokrig} and \eqref{eq:krig:cov}, we have
-\begin{eqnarray}\label{eq:krigSUR}
-\mu^{s}(\x) &=&\mu^{s-1}(\x)+\lambda^{[s,s]}(\x,\bu_{d^{j,s}})^T (\by_{d^{j,s}}-G_{d^{j,s}} \mu(\x)), \nonumber \\
-k^{s}(\x,\x') &=& k^{s-1}(\x,\x')-\lambda^{[s,s]}(\x,\bu_{d^{j,s}})^T K_{Y_{d^{j,s}}} \lambda^{[s,s]}(\x',\bu_{d^{j,s}}),
-\end{eqnarray}
-where $\lambda^{[s,s]}(\x,\bu_{d^{j,s}})=K^{-1}_{Y_{d^{j,s}}} G_{d^{j,s}} k(\bu_{d^{j,s}},\x)$ and the data covariance is $K^{-1}_{Y_{d^{j,s}}}=G_{d^{j,s}} K^s(\bu_{d^{j,s}},\bu_{d^{j,s}})G^T_{d^{j,s}}+R_{d^{j,s}}$, with measurement model as in Eq.
-\eqref{eq:meas}.
-
-At stage $s$, the robot must then choose design $d^{j,s}$ where measurements $\by_{d^{j,s}}$ should
-be gathered. For the triangularized waypoint
-graph all six possible
-directions must be evaluated to minimize the expected uncertainty of random sets, and in particular the EIBV for our 
-application. We hence focus on choosing designs for data modeled by
-Eq. \eqref{eq:meas}, to conduct efficient sequential uncertainty
-reduction according to the EIBV criterion. Compared with the
-static design $d$ in Eq. \eqref{eq:eibv}, the optimal sequential
-design not only considers the best current design grid point, but also
-how the data gathered at this node could inform future sampling at
-successive stages.
-
-%The optimal path
-%selection situation is depicted in Fig. %\ref{fig:PathSelOpt},
-%\begin{figure}[b!]
-%\centering
-%\includegraphics[width=0.85\textwidth]{Figures/sequent_select.pdf}
-%\caption{Optimal sequential path selection.}\label{fig:PathSelOpt}
-%\end{figure}
-%where design choices are indicated by squares while %data realizations are indicated by circles. 
-%terms this optimal design is defined as
-%\begin{equation}\label{opt_crit}
-%    \bd^* = \mbox{argmin}_{d^{j,1}} \left\{ \int_{\by^{j,1}} \mbox{argmin}_{d^{j,2}} \left\{ \int_{\by^{j,2}} \ldots \pi(\by^{j,2}|\by^{j,1}) d\by^{j,2} \right\} \pi(\by^{j,1}) d \by^{j,1} \right\},
-%\end{equation}
-%where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages. 
-The mathematical expression for the optimal design involves a series
-of intermixed maximizations over designs and integrals over data. In
-practice, the optimal solution is intractable because of the enormous
-growth over stages (see e.g. \cite{sucar2015probabilistic} and
-\cite{powell2016perspectives}).  Instead, we outline heuristic
-strategies.
-
-%The derived closed-form solutions for the EIBV are still important building blocks when constructing the sampling designs, as they will be used to score the different adaptive designs. % Efficient
-% calculation is important for using adaptive survey designs with
-% robotic platforms.
-
-\subsection{A Naive Sampling Strategy}
-\label{naive}
-
-The simplest heuristic for adaptive sampling is to choose the next
-sampling location based on current probabilities. One reduces the
-largest levels of uncertainty by selecting the design $j$ with
-$P(Z(\x_j) \in T)$ closest to $0.5$.  With temperature and salinity
-variables, this {\it{naive}} strategy selects the stage $s$ sampling
-design according to:
-\begin{eqnarray}\label{critNaive}
-    d^{*,s} &=& \mbox{argmin}_{j} |P(Z_1(\x_j) \leq t_1, Z_2(\x_j) \leq t_2 | \mathcal{Y}^{s-1})-0.5|, \nonumber
-\end{eqnarray}
-where $j$ is indexing all neighboring locations on the waypoint graph.
-
-This strategy does not account for the uncertainty in the GP
-variables, but is instead applicable only if one design (node) has EPs
-closer to $0.5$ (see Table \ref{tab:sim_rhoab}, row
-two). Additionally, this strategy does not account for spatial
-correlation. This strategy lacks memory of where it has been and where
-the uncertainty has been reduced, and therefore susceptible to local
-minima.
-
-\subsection{Myopic Path Planning}
-\label{sec:myopic}
-
-The myopic (greedy) strategy which we present here is optimal if we
-imagine taking only one more stage of measurements. This selection
-strategy is based on expected uncertainty reduction, but it is a
-heuristic because there is no anticipation of what the subsequent
-designs might offer beyond the first stage.
-
-Based on the currently available data $\mathcal{Y}^{s-1}$, we fit an
-updated GP model. Focusing on the EIBV criterion, at the next stage
-$s$, the selected design is:
-\begin{eqnarray}\label{critSEQ}
-    d^{*,s} &=& \mbox{argmin}_{j} \left\{ \int_{\mathcal{M}} E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( p(\x;\mathcal{Y}^{j,s})\left( 1-p(\x;\mathcal{Y}^{j,s})\right) \right) d\mes(u), \right\} \\
-    p(\x;\mathcal{Y}^{j,s}) &=& P(Z_u \in T |\by_{d^{j,s}},\mathcal{Y}^{s-1}). \nonumber
-\end{eqnarray}
-
-Note that this strategy gives a sequential conditional version of
-expression (\ref{eq:eibv}). Now $\mathcal{Y}^{s-1}$ is available, and
-the expectation is conditional to this data. A similar closed-form
-calculation for EIBV is hence applicable in Eq.  \eqref{critSEQ} using the updated
-GP model from stage $s-1$ as defined via Eq.
-\eqref{eq:krigSUR} in the results from Proposition \ref{propo2}.
-Once the data are collected for the best design $d^{*s}$ at stage $s$, the GP model is updated again.
-The mean and covariance are used to compute the next design at stage $s+1$, and so on.
-
-Even though this myopic strategy is non-anticipative, it still gives a
-reasonable approach for creating designs in many
-applications. Moreover, it is easily implemented, in our case, onboard
-an AUV, using an efficient data update of the GP model and the
-calculation of the closed-form EIBV expressions for each subsequent
-trajectory.
-
-\subsection{Look-ahead Trajectory Planning}
-\label{sec:LA}
-
-We now extend the myopic strategy to a look-ahead strategy. We consider two stages of
-measurements, and the strategy is then optimal in the sense that it accounts consistently for the expectations and minimizations in these two stages, but still without including any planning beyond those two steps.
-In addition to the
-next stage of measurements $\by^{j,s}$, this look-ahead strategy anticipates the subsequent design indexed $j_2$ with data
-$\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$. The
-selected design is:
-\begin{eqnarray}\label{critLA}
-    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( \mbox{argmin}_{j_2}  V_{j_2} \right) \right\}, \\
-V_{j_2} & = & \int_{\mathcal{M}} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1}) ( 1-p(\x;\mathcal{Y}^{j,s+1}) ) \right) d\mes(u), \nonumber \\
-    p(\x;\mathcal{Y}^{j,s+1}) &=& P(Z(\x) \in T |\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}). \nonumber
-\end{eqnarray}
-Here, $\mathcal{Y}^{j,s}=\{\by_{d^{j,s}},\mathcal{Y}^{s-1}\}$ and
-$\mathcal{Y}^{j,s+1}=\{\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}\}$
-represent the sets of data variables, and the expectations are
-calculated with respect to the conditional densities given data at
-that stage.  We solve the first expectation by Monte
-Carlo sampling of data $\by_{d^{j,s}}$ from its conditional
-distribution. For each of these data samples, the second expectation
-is solved using the closed-form expressions for EIBV using the results from Proposition \ref{propo2}, now with conditioning to the first stage data already going into the mean and covariances in Eq.
-\eqref{eq:krigSUR}.
-
-Even though the strategy looks at two sampling stages, it is only used to find
-the current best design. When data are collected at stage $s$, the GP model is
-updated as in Eq. \eqref{eq:krigSUR}, and the mean and (cross) covariance are
-used to compute the next design for stage $s+1$, now anticipating what
-stage $s+1$ and $s+2$ could offer, and so on.
-This look-ahead approach is much more computationally demanding than
-the myopic strategy, so for practical implementation onboard the AUV this is reaching the limits allowed for calculation.
-We conduct some pruning of the least valuable design paths in the evaluation of Eq.
-\eqref{critLA}, to get a satisfactory computation time below one minute per waypoint.
-This means that we use the myopic
-strategy to rank the three best designs following the first stage alone, and
-for each of these preferred designs, we undertake the look-ahead calculations.
-
-\subsection{Simulation study}
-\label{sec:simulations}
-
-\susubsection{Static and Sequential Sampling Designs}
-\label{sec:sampling_designs}
-
-%Three different designs are considered as indicated in Figure \ref{fig:stat_design}. 
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.65\textwidth]{Figures/Des3.eps}
-%\caption{Three different static survey designs plotted on the initial EP.}\label{fig:stat_design}
-%\end{figure}
-%In this display the designs are plotted along with the prior probability contours of the ES for the reference parameter inputs.
-
-We compare three different static designs denoted
-\textit{static\_north}, \textit{static\_east}, and
-\textit{static\_zigzag} with the three described sequential approaches
-\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
-sampling paths are pre-determined and cannot be altered and represent
-the pre-planned strategies used in most current AUV operational survey
-designs.
-
-For a fixed survey length, a closed-form expression for the EIBV is
-available as in Eq. \eqref{eq:eibv}. However, for the sequential
-approaches this is not the case. For comparison the properties are
-therefore evaluated using Monte Carlo sampling over several replicates
-of realizations from the model while conducting simulated sequential
-surveys for each one. Fig. \ref{fig:Eprob} shows the conditional EP,
-given data gathered along the north-south survey lines in these
-displays. In the Monte Carlo replicates, such results are repeatedly
-computed to approximate the EIBV. We also compare predictive
-performance measured by root mean square error (RMSE) for temperature
-and salinity estimates as well as also the variance reduction in these
-two variables. It is important to note that the objective function
-used by the agent \kc{First of 'agent'. Define it or change it
-  to 'robot'? Note further uses below.} is focused on reducing the
-EIBV, but we nevertheless expect that we will achieve good predictive
-performance for criteria such as RMSE as well. Another non-statistical
-criteria that is important for practical purposes is the computational
-time needed for the strategy, as this will impact the performance for
-an embedded system.
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
-%\caption{The equilateral waypoint graph used to discretize the
-%  trajectory choices over the $31\times31$ grid used to discretize the GP.}
-%\label{fig:wp_graph}
-%\end{figure}
-
-\begin{figure}[!h] 
-\centering 
-\subfigure[The waypoint graph.]{\includegraphics[width =
-0.49\textwidth]{Figures/sim/wp_graph_paper.pdf}\label{fig:wp_graph_a}}
-\hfill
-\subfigure[The waypoint graph in 3D.]{\includegraphics[width =
-0.49\textwidth]{Figures/sim/wp_graph_3d.pdf}\label{fig:wp_graph_b}}
-\caption{\ref{fig:wp_graph_a} The equilateral waypoint graph used to discretize the
-trajectory choices over the $31\times31$ grid used to discretize the GP.
-\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment.}
-\label{fig:wp_graph}
+\subsubsection{Including Spatiality}
+INTRODUCTORY SENTENCE.
+The class of GRF model we will consider in the article will generally have a linear trend
+\begin{align*}
+\mu(s)=\mathbb{E}\left[\begin{pmatrix}
+Z_{s, 1}\\ Z_{s, 2}
+\end{pmatrix}\right] &= \beta_0 + \beta_1 s
+\end{align*}
+with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. We will only consider covariance models of separable type
+\begin{align*}
+\textrm{Cov}\left(Z_{s, i}, Z_{u, j}\right) &= k(s, u) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_l^2,~ i=j\\
+   \gamma_0 \sigma_i \sigma_j,~i\neq j
+        \end{cases}
+\end{align*}
+where $k(., .)$ is one of the traditional spatial covariance kernel (exponential, Mat\'{e}rn, ...) and $\gamma$ defines the cross-covariance structure.
+
+In the accompanying Python code, these modeling assumptions can be
+generalized to anisotropic covariance and changing variance levels
+across the spatial domain. One can hence, easily run the code to see
+if different models result in other sampling designs. Anisotropy and
+non-stationary variance are both relevant for the setting with river
+plumes, but in practice this requires more parameters to be
+specified. With extensive data and prior knowledge, one could also
+possibly fit and estimate parameters of more complex multivariate
+spatial covariance functions
+\citep{gneiting2010matern,genton2015cross}, but that is outside the
+scope of the current paper.
+
+In the rest of this section, we will consider a GRF with mean and covariance structure as above and parameters
+\begin{align*}
+\beta_0 = \begin{pmatrix}
+5.8\\ 24.0
+\end{pmatrix}, ~ \beta_1 = \begin{pmatrix}
+0.0 & -4.0\\
+0.0 & -3.8
+\end{pmatrix},~ \sigma_1 = \sigma_2 = 2.25, ~ \gamma_0 = 0.2
+\end{align*}
+and the spatial covariance is given by a Mat\'{e}rn 3/2 kernel with unit variance and range parameter $\lambda=0.5$.
+One realization of this GRF is shown in Fig. \ref{example_excu}.
+
+We now study how the expected Bernoulli variance eq.\ref{eq:eibv} associated to data collection at a point changes if only one of the two components of the field is observed.
+We first draw a realization of the GRF defined above and use it as ground-truth to mimick the real data-collection process.
+A first data collection step is performed at the location in green, and the data is used to update the GRF model.
+The expected Bernoulli variance reduction (at each grid point) that would result from observing only one component, or both at the next grid point upwards is then computed an shown in Fig.\ref{fig:ebv_comp}.
+
+\begin{figure}[h!] \centering
+  \includegraphics[width=0.99\textwidth]{Figures/ebv_comp.png}
+  \caption{Pointwise Bernoulli variance reduction for observation of a single or both components of the random field at one location.
+Data collection locations in green.
+True excursion set in red.
+Places where only one response is above threshold are depicted in pink}
+\label{fig:ebv_comp}
 \end{figure}
 
-Each strategy is conducted on an equilateral grid as shown in
-Fig. \ref{fig:wp_graph}. The sequential sampling agent starts at the
-center East-West coordinate at the southern end of the domain (node 53
-\kc{Maybe color this node differently and modify the caption?}). The
-AUV moves along edges in the waypoint graph while the vehicle makes
-measurements. The data is assimilated into the GP model before an
-evaluation of the next node to sample is conducted at the end of the
-edge.
-
-\subsection{Simulation Results}
-
-A total of 100 replicate simulations were conducted with all
-strategies. The results are shown in Fig. \ref{fig:sim_results}, where
-the different criteria are plotted as a function of survey
-distance. Fig. \ref{fig:avg_ev} shows the resulting drop in IBV for
-each of the six strategies. IBV reduction occurs most under the
-\textit{myopic} and \textit{look-ahead} strategies, each performing
-almost equally; this is expected as the two criteria
-(Eq. \eqref{critSEQ} and \eqref{critLA}) are sensitive to differences
-in IBV. The \textit{static\_north} design also does well here because
-the path is parallel to the boundary between the water masses.
-
-\begin{figure}[h!]
-  \centering
-  % \subfigure[Excursion set variance $E_{\by}(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
-  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
-  \hfill
-  \subfigure[RMSE between estimated field and truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
-  \hfill 
-  \subfigure[Explained variance $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
-  \hfill 
-  \subfigure[Computational time for inferencing.\kc{I see only 
-    lines associated with 4 variables showing in the graph. Where is
-    static\_north and static\_east?}]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}} 
-\caption{Simulation results from 100 replicate simulations for 10
-  sampling choices/stages on the grid. \kc{Vertical lines show large
-    variation in replicate results.}}  
-\label{fig:sim_results}
+Note that plotting the Bernoulli variance reduction at each point might also be used to compare different data collection plans.
+For example, Fig.
+\ref{fig:ebv_comp} shows the expected Bernoulli variance reduction associated to a data collection plan along a straight upwards line (static north) and the one associated to a straight horizontal line (static east).
+Both expectations are computed according to the a priori distribution of the GRF (i.e.
+no observations have been included yet).
+
+\begin{figure}[h!] \centering
+  \includegraphics[width=0.99\textwidth]{Figures/ebv_north_vs_east.png}
+  \caption{Pointwise Bernoulli variance reduction for two different observation plans.
+Data collection locations in green.
+True excursion set is in red.
+Places where only one response is above threshold are depicted in pink.}
+\label{fig:ebv_north_vs_east}
 \end{figure}
 
-Fig. \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop
-in RMSE and increase in explained variance, respectively. Both
-\textit{myopic} and \textit{look-ahead} strategies perform well here,
-but some of the \textit{static\_east} and \textit{static\_zigzag} also
-achieve good results because they are pre-determined to cover large
-parts of the domain without re-visitation. Sequential strategies
-targeting IBV will sometimes not reach similar coverage, as
-interesting data may draw the path into twists and turns. There is
-relatively large variety in the replicate results as indicated by the
-vertical lines. Nevertheless, the ordering of strategies is similar.
-
-Fig. \ref{fig:avg_time} shows the computational effort: the
-\textit{naive} strategy is on par with the static designs, while the
-\textit{myopic} strategy is slower. The \textit{look-ahead} is even
-slower, reaching levels that are nearly impractical for execution on a
-vehicle. Some pruning of the graph is performed to improve the
-performance, such as ruling out repeated visitations and
-back-and-forth routes. Some of the intermediate results are also
-stored for longer planning horizons. Further pruning of branches or
-inclusion of other heuristics could be included for better
-performance. Then again, the inclusion of heuristics is likely a
-contributing factor for the \textit{look-ahead} strategy failing to
-outperform the \textit{myopic} strategy.
-
-In Fig. \ref{fig:route_choices}, the realized sampling paths for each
-of the sequential schemes and static designs are shown. The
-\textit{naive} strategy often gets stuck in the southern part of the
-domain because it is too focused on the probabilities near $0.5$. The
-\textit{myopic} strategy covers a wider domain than the naive or
-look-ahead. There are several reasons for this. First, a greedy
-approach will tend to put more emphasis on promising locations close
-to the agent, which may lead away from the centre. Second, as the
-agent evaluates the impact of locations further away (look-ahead)
-where assimilated data has less predictive power, the GP model (which
-is centered here) will act to restrict paths deviating from the
-central zone.
-
-We studied the sensitivity of the results by modifying the input
-parameters to have different correlations between temperature and
-salinity, standard deviations, and spatial correlation range.  In all
-runs, the \textit{myopic} and \textit{look-ahead} strategies perform
-the best in terms of realized IBV, and much better than
-\textit{naive}. The \textit{look-ahead} strategy seems to be
-substantially better than the \textit{myopic} design only for very
-small initial standard deviation or very large spatial correlation
-range. \textit{static\_north} continues to be the best static design
-for IBV, while \textit{static\_zigzag} is the best design for the
-other predictive performance measures, especially so with large
-spatial correlation range. We also ran simulation studies with only
-temperature data, and for realistic correlation levels between
-temperature and salinity, the IBV results are not much worse when only
-temperature data are available. In addition to the comparison made in
-Table \ref{tab:sim_rhoab}, the current setting includes spatial
-correlation and this likely reduces the additional influence of having
-bivariate data. However, it seems that having temperature data alone
-does a substantially worse job in terms of explained variance
-
-\begin{comment}
-\subsection{Performance Metrics}
-
-The tables below explore the sensitivity of the result to different input correlation values between temperature and salinity measured onboard the AUV, standard deviations, and spatial
-correlations (see table rows). We only compare results at the end of
-each survey. The expected ES variances are shown in Table
-\ref{tab:sim_res_ev}, results of RMSE are in Table
-\ref{tab:sim_res_rmse}, and explained variance is in Table
-\ref{tab:sim_res_r2}. In all runs, both the \textit{myopic} and
-\textit{look-ahead} strategies perform the best in terms of expected
-variance in the ES. The \textit{myopic} strategy
-performs slightly better than the \textit{look-ahead}. Part of the
-reason for this discrepancy can be traced to the act of pruning some
-of the paths for the \textit{look-ahead} strategy. The \textit{static\_north} design is clearly
-better than the other two static designs, the reason being that the
-others do not focus on the areas where the ES is really uncertain
-(centre locations are the most uncertain). By measuring temperature
-alone, the expected variances in the ES are only slightly higher than
-the combined temperature-salinity estimate. This occurs because the
-correlation of $0.6$ means that temperature observations provide
-information about salinity as well. For the other predictive measures,
-\textit{myopic} and \textit{look-ahead} strategies do well, clearly
-beating the \textit{naive} strategy.
-
-\begin{table}[!h]
-    \centering
-    \scalebox{0.87}{
-    \begin{tabular}{lrrrrrr}
-    \toprule
-    Parameter: $E_{\by}(p[1-p])$ &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
-    \midrule
-                    \rowcolor{Gray}
-ts. cor. low: 0.2  &  29.65 &   27.08 &       \textbf{26.44} &         32.33 &        40.31 &          36.23 \\
-ts. cor. high: 0.8 &  36.18 &   \textbf{28.42} &       30.30 &         32.99 &        37.18 &          36.04 \\
-                    \rowcolor{Gray}
-std. low: 0.1      &  18.26 &   15.79 &       \textbf{15.35} &         18.89 &        26.65 &          26.10 \\
-std. high: 0.5     &  47.01 &   \textbf{41.31} &       43.15 &         47.92 &        52.63 &          49.71 \\
-                    \rowcolor{Gray}
-cor. low: 0.8      &  44.42 &   \textbf{42.02} &       42.47 &         43.84 &        46.05 &          47.82 \\
-cor. high: 0.2     &  29.73 &   21.09 &       \textbf{20.43} &         27.91 &        37.60 &          34.35 \\
-                    \rowcolor{Gray}
-temp. only         &  38.25 &   29.16 &       \textbf{28.69} &         34.32 &        39.81 &          36.17 \\
-basecase           &  35.21 &   \textbf{28.56} &       28.94 &         33.50 &        39.84 &          39.26 \\
-    \bottomrule
-    \end{tabular}}
-  \caption{Simulation results for the final mean IBV (Eq. \eqref{two_partsK}).}
-    \label{tab:sim_res_ev}
-\end{table}
-\vspace{-0.4cm}
-\begin{table}[!h]
-    \centering
-    \scalebox{0.92}{
-        \begin{tabular}{lrrrrrr}
-        \toprule
-        Parameter: RMSE &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
-        \midrule
-                \rowcolor{Gray}
-ts. cor. low: 0.2  &   0.63 &    \textbf{0.57} &        0.59 &          0.62 &         0.57 &           0.59 \\
-ts. cor. high: 0.8 &   0.63 &    \textbf{0.59} &        0.61 &          0.64 &         0.60 &           0.60 \\
-                \rowcolor{Gray}
-std. low: 0.1      &   0.39 &    0.37 &        0.38 &          0.38 &         0.40 &           \textbf{0.35} \\
-std. high: 0.5     &   0.85 &    0.82 &        0.84 &          0.82 &         0.83 &           \textbf{0.81} \\  
-                \rowcolor{Gray}
-cor. low: 0.8      &   0.68 &    0.66 &        0.67 &          0.68 &         0.66 &           \textbf{0.65} \\
-cor. high: 0.2     &   0.58 &    0.54 &        0.55 &          0.58 &         0.54 &           \textbf{0.49} \\
-                \rowcolor{Gray}
-temp. only         &   0.65 &    0.63 &        0.62 &          0.65 &         \textbf{0.61} &           0.62 \\
-basecase           &  0.61 &   \textbf{0.56} &       \textbf{0.56} &         0.60 &        0.57 &          0.57 \\
-        \bottomrule
-        \end{tabular}}
-      \caption{Simulation results for the final mean RMSE ($\frac{1}{B} \sum_{b=1}^{B} \sqrt{(\bmu-\tilde{\bmu})^2}$).}
-    \label{tab:sim_res_rmse}
-\end{table}
-\vspace{-0.4cm}
-\begin{table}[!h]
-    \centering
-    \scalebox{0.95}{
-        \begin{tabular}{lrrrrrr}
-        \toprule
-        Parameter: $\bR^{2}$ &  naive &  myopic &  look-ahead &  static\_north &  static\_east &  static\_zigzag \\
-        \midrule
-                        \rowcolor{Gray}
-ts. cor. low: 0.2  &  27.00 &   33.29 &       32.24 &         26.71 &        30.56 &          \textbf{33.50} \\
-ts. cor. high: 0.8 &  23.68 &   \textbf{33.87} &       32.42 &         26.72 &        30.73 &          33.61 \\
-                        \rowcolor{Gray}
-std. low: 0.1      &  25.38 &   30.80 &       30.29 &         26.09 &        28.75 &          \textbf{31.66} \\
-std. high: 0.5     &  26.11 &   34.22 &       32.48 &         26.95 &        31.77 &          \textbf{34.48} \\
-                        \rowcolor{Gray}
-cor. low: 0.8      &   8.80 &   11.50 &       11.23 &          9.56 &        12.31 &          \textbf{12.58} \\
-cor. high: 0.2     &  37.57 &   \textbf{48.68} &       48.11 &         39.61 &        41.98 &          48.04 \\
-                        \rowcolor{Gray}
-temp. only         &  11.67 &   \textbf{22.82} &       21.98 &         18.16 &        20.77 &          22.78 \\
-basecase           &  24.71 &   33.45 &       32.25 &         26.71 &        30.62 &          \textbf{33.54}\\
-        \bottomrule
-        \end{tabular}}
-      \caption{Simulation results for the final mean explained variance $\bR^{2}=100*(1-(\Sigma_{posterior}/\Sigma_{initial}))$).}
-    \label{tab:sim_res_r2}
-\end{table}
-
-\end{comment}
+\begin{remark}
+In all the above examples, the spatial domain $\domain$ has been discretized to a set of $n$ grid
+locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,n \}$, where each cell
+has area $\Delta$; the same grid is used for the waypoint graph for
+possible design locations. The EIBV is approximated by sums over all
+grid cells.
+\end{remark}
+
+\input{section3}
 

### 2020-06-18 section3.tex
Updates from Overleaf

\section{Sequential designs and heuristic path planning}\label{sec:heuristics}

The goal of this section is to present data collection strategies that aim at decreasing the uncertainty on the target excursion set $\es$.

\subsection{Background}
As before, let $\gp$ be a $\no$-variate random field on $\domain$. See Section \ref{sec:bg_and_notation} for a recap of the specific notation used to handle mean and covariance function of multivariate random fields.

In a sequential setting, the point of view is that $n$ data collection steps have already been performed and one wants to choose what data to collect next. After each step, the GRF model is updated using the cokriging equations \ref{eq:cokrig_mean} and \ref{eq:cokrig_cov} and we will denote by $\mathbb{P}_n$ the law of the field conditional on the data gathered up to (and including) stage $n$.

\begin{remark}
The type of data collected at each stage can be of any type (observation of all components of the field at a single location, observation of some components only but at several locations, ...) since the concept of \textit{generalized location} allows the cokriging equations to handle all those case by putting them on equal footage.
\end{remark}

Our goal is then to optimally select the next spatial location $\spatloc_{n+1}$ at which to observe the field (all of its component), where the meaning of optimal is yet to be defined. Note that here again, the concept of \textit{generalized location} allows one to consider settings where only one component of the field can be observed at each step and the startegy has to choose the spatial location as well as the component to observe (heterotopic case). In the same fashion, one could consider settings where several observations are allowed at each stage (batch case).

We will here restrict ourselves to the case where all components of the field are observed at one spatial location at each stage (isotopic case).

\subsection{Optimal Sequential Design}
\label{Optdes}

At stage $s$, the robot must then choose design $d^{j,s}$ where measurements $\by_{d^{j,s}}$ should
be gathered. For the triangularized waypoint
graph all six possible
directions must be evaluated to minimize the expected uncertainty of random sets, and in particular the EIBV for our 
application. We hence focus on choosing designs for data modeled by
Eq. \eqref{eq:meas}, to conduct efficient sequential uncertainty
reduction according to the EIBV criterion. Compared with the
static design $d$ in Eq. \eqref{eq:eibv}, the optimal sequential
design not only considers the best current design grid point, but also
how the data gathered at this node could inform future sampling at
successive stages.

The mathematical expression for the optimal design involves a series
of intermixed maximizations over designs and integrals over data. In
practice, the optimal solution is intractable because of the enormous
growth over stages (see e.g. \cite{sucar2015probabilistic} and
\cite{powell2016perspectives}).  Instead, we outline heuristic
strategies.

\textcolor{red}{Here introduce the fact that candidate points are always chosen in a set $\candidates$. Mention that we usually only allow a finite set. Mention that in the specific case of robot planning, we have a waypoint graph and we thus usually take $\candidates$ to be the set of current neighbouring nodes. Mention that yes, $\candidates$ normally depends on the current location (this is for robot, in general setting, depends on the location of the last observations), but we silently suppress this dependency for readability. Nevertheless, one should always remember that the current set of candidates depends on the current setting.}
\subsection{A Naive Sampling Strategy}
\label{naive}

The simplest heuristic for adaptive sampling is to choose the next
sampling location $\spatloc_{n+1}$ based on current excursion probabilities. At stage $\stage$, one reduces the
largest levels of uncertainty by selecting the location $\spatloc_{\stage + 1}$ with
$P_{\stage}(\gp_{\spatloc_{\stage + 1}} \in T)$ closest to $0.5$..

This strategy does not account for the uncertainty in the GP
variables, \textcolor{red}{(C.T.: do we need this sentence? does it mean anything?) but is instead applicable only if one design (node) has EPs
closer to $0.5$ (see Table \ref{tab:sim_rhoab}, row
two).}

Additionally, this strategy does not account for spatial
correlation. This strategy lacks memory of where it has been and where
the uncertainty has been reduced, and therefore susceptible to local
minima.

\subsection{Myopic Path Planning}
\label{sec:myopic}

The myopic (greedy) strategy which we present here is optimal if we
imagine taking only one more stage of measurements. This selection
strategy is based on expected uncertainty reduction, but it is a
heuristic because there is no anticipation of what the subsequent
designs might offer beyond the first stage.

Based on the currently available data the myopic strategy selects the location that leads to the biggest expected reduction in IBV:
\begin{criterion}[Myopic]
\begin{equation}\label{critSEQ}
    \spatloc_{\stage + 1} = \argmin_{\spatloc \in \candidates} \EIBV_{\stage}\left(\spatloc\right)
\end{equation}
\end{criterion}

The expected IBV might be efficiently computed for each candidate points using Proposition \ref{propo2}. Once the best location has been selected, a stage of observation is performed there, the GP model is updated, yielding a conditional law $\mathbb{P}_{\stage + 1}$ and the process is repeated.

Even though this myopic strategy is non-anticipative, it still gives a
reasonable approach for creating designs in many
applications. Moreover, it can be implemented in a way that is not too demanding on computational power, making it well-suited for embedding on autonomous mobile survey platform (see Section \ref{sec:case_study} for an example).

\subsection{Look-ahead Trajectory Planning}\label{sec:LA}

We now extend the myopic strategy to a look-ahead strategy. We consider two stages of
measurements, and the strategy is then optimal in the sense that it accounts consistently for the expectations and minimizations in these two stages, but still without including any planning beyond those two steps.
\midskip
Say we have collected data up to stage $n$. The principle of two-steps lookahead
strategies is to select the next observation location $\spatloc_{\stage + 1}$ as follows

  \begin{itemize}
      \item First, do as if the $\stage + 1$ observation had already been done and say
          we observed $y_{\stage + 1}$ at $\spatloc_{\stage + 1}$.
      \item If we were at this stage, then we would choose the next points by
          minimizing the expected IBV under the current ($\stage + 1$) conditional
          distribution
          \[
              \spatloc_{\stage + 2} = \argmin_{\spatloc \in \candidates}~ \EIBV_{\stage + 1}\left(\spatloc \mid \spatloc_{\stage + 1},
          y_{\spatloc + 1}\right)
          \]
      \item Now the above expression depends on the observed data $y_{\stage + 1}$,
          whose value we don't know beforehand. What we know is that
          at step $\stage$ it is distributed according to the conditional distribution
          $\mathbb{P}_{\stage}$. We can thus replace it by a random variable $Y$
          following the current conditional law and then average over it.
  \end{itemize}

  Let us now try to make the above more precise.
  \begin{criterion}[2-steps lookahead]
      The next observation location $\spatloc_{\stage + 1}$ is selected according to
      \begin{align}\label{critLA}
          \spatloc_{\stage + 1} = \argmin_{\spatloc \in \candidates}~ \mathbb{E}_{\stage}\left[\min_{\spatloc' \in
                  \candidates(\spatloc)} \EIBV_{\stage + 1}\left(\spatloc' \mid \spatloc,
      Y\right)\right]
      \end{align}
  where $Y$ is a random variable distributed according to the conditional
  law of $\gp_{\spatloc}$ at step $\stage$ (note it silently depends on the considered point
  $\spatloc$) and $\EIBV_{\stage +  1}(\spatloc'\mid \spatloc,Y)$ is the
  expected IBV with the expectation taken according to the conditional law $\mathbb{P}_{\stage}(\cdot\mid
  \spatloc, Y)$. Note that the dependence of the set of candidates on the current location has been made explicit for the second stage of measurements.
  \end{criterion}

\begin{remark}
In a practical setting, the first expectation might be computed by Monte
Carlo sampling of data $Y$ from its conditional
distribution. For each of these data samples, the second expectation
is solved using the closed-form expressions for EIBV provided b< Proposition \ref{propo2}, now with conditioning to the first stage data already going into the mean and covariances in Eq. \eqref{eq:krigSUR}. 
\end{remark}

\subsection{Simulation study}
\label{sec:simulations}

\susubsection{Static and Sequential Sampling Designs}
\label{sec:sampling_designs}

%Three different designs are considered as indicated in Figure \ref{fig:stat_design}. 
%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.65\textwidth]{Figures/Des3.eps}
%\caption{Three different static survey designs plotted on the initial EP.}\label{fig:stat_design}
%\end{figure}
%In this display the designs are plotted along with the prior probability contours of the ES for the reference parameter inputs.

We compare three different static designs denoted
\textit{static\_north}, \textit{static\_east}, and
\textit{static\_zigzag} with the three described sequential approaches
\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
sampling paths are pre-determined and cannot be altered and represent
the pre-planned strategies used in most current AUV operational survey
designs.

For a fixed survey length, a closed-form expression for the EIBV is
available as in Eq. \eqref{eq:eibv}. However, for the sequential
approaches this is not the case. For comparison the properties are
therefore evaluated using Monte Carlo sampling over several replicates
of realizations from the model while conducting simulated sequential
surveys for each one. Fig. \ref{fig:Eprob} shows the conditional EP,
given data gathered along the north-south survey lines in these
displays. In the Monte Carlo replicates, such results are repeatedly
computed to approximate the EIBV. We also compare predictive
performance measured by root mean square error (RMSE) for temperature
and salinity estimates as well as also the variance reduction in these
two variables. It is important to note that the objective function
used by the agent \kc{First of 'agent'. Define it or change it
  to 'robot'? Note further uses below.} is focused on reducing the
EIBV, but we nevertheless expect that we will achieve good predictive
performance for criteria such as RMSE as well. Another non-statistical
criteria that is important for practical purposes is the computational
time needed for the strategy, as this will impact the performance for
an embedded system.

%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
%\caption{The equilateral waypoint graph used to discretize the
%  trajectory choices over the $31\times31$ grid used to discretize the GP.}
%\label{fig:wp_graph}
%\end{figure}

\begin{figure}[!h] 
\centering 
\subfigure[The waypoint graph.]{\includegraphics[width =
0.49\textwidth]{Figures/sim/wp_graph_paper.pdf}\label{fig:wp_graph_a}}
\hfill
\subfigure[The waypoint graph in 3D.]{\includegraphics[width =
0.49\textwidth]{Figures/sim/wp_graph_3d.pdf}\label{fig:wp_graph_b}}
\caption{\ref{fig:wp_graph_a} The equilateral waypoint graph used to discretize the
trajectory choices over the $31\times31$ grid used to discretize the GP.
\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment.}
\label{fig:wp_graph}
\end{figure}

Each strategy is conducted on an equilateral grid as shown in
Fig. \ref{fig:wp_graph}. The sequential sampling agent starts at the
center East-West coordinate at the southern end of the domain (node 53
\kc{Maybe color this node differently and modify the caption?}). The
AUV moves along edges in the waypoint graph while the vehicle makes
measurements. The data is assimilated into the GP model before an
evaluation of the next node to sample is conducted at the end of the
edge.

\subsection{Simulation Results}

A total of 100 replicate simulations were conducted with all
strategies. The results are shown in Fig. \ref{fig:sim_results}, where
the different criteria are plotted as a function of survey
distance. Fig. \ref{fig:avg_ev} shows the resulting drop in IBV for
each of the six strategies. IBV reduction occurs most under the
\textit{myopic} and \textit{look-ahead} strategies, each performing
almost equally; this is expected as the two criteria
(Eq. \eqref{critSEQ} and \eqref{critLA}) are sensitive to differences
in IBV. The \textit{static\_north} design also does well here because
the path is parallel to the boundary between the water masses.

\begin{figure}[h!]
  \centering
  % \subfigure[Excursion set variance $E_{\by}(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
  \hfill
  \subfigure[RMSE between estimated field and truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
  \hfill 
  \subfigure[Explained variance $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
  \hfill 
  \subfigure[Computational time for inferencing.\kc{I see only 
    lines associated with 4 variables showing in the graph. Where is
    static\_north and static\_east?}]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}} 
\caption{Simulation results from 100 replicate simulations for 10
  sampling choices/stages on the grid. \kc{Vertical lines show large
    variation in replicate results.}}  
\label{fig:sim_results}
\end{figure}

Fig. \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop
in RMSE and increase in explained variance, respectively. Both
\textit{myopic} and \textit{look-ahead} strategies perform well here,
but some of the \textit{static\_east} and \textit{static\_zigzag} also
achieve good results because they are pre-determined to cover large
parts of the domain without re-visitation. Sequential strategies
targeting IBV will sometimes not reach similar coverage, as
interesting data may draw the path into twists and turns. There is
relatively large variety in the replicate results as indicated by the
vertical lines. Nevertheless, the ordering of strategies is similar.

Fig. \ref{fig:avg_time} shows the computational effort: the
\textit{naive} strategy is on par with the static designs, while the
\textit{myopic} strategy is slower. The \textit{look-ahead} is even
slower, reaching levels that are nearly impractical for execution on a
vehicle. Some pruning of the graph is performed to improve the
performance, such as ruling out repeated visitations and
back-and-forth routes. Some of the intermediate results are also
stored for longer planning horizons. Further pruning of branches or
inclusion of other heuristics could be included for better
performance. Then again, the inclusion of heuristics is likely a
contributing factor for the \textit{look-ahead} strategy failing to
outperform the \textit{myopic} strategy.

In Fig. \ref{fig:route_choices}, the realized sampling paths for each
of the sequential schemes and static designs are shown. The
\textit{naive} strategy often gets stuck in the southern part of the
domain because it is too focused on the probabilities near $0.5$. The
\textit{myopic} strategy covers a wider domain than the naive or
look-ahead. There are several reasons for this. First, a greedy
approach will tend to put more emphasis on promising locations close
to the agent, which may lead away from the centre. Second, as the
agent evaluates the impact of locations further away (look-ahead)
where assimilated data has less predictive power, the GP model (which
is centered here) will act to restrict paths deviating from the
central zone.

We studied the sensitivity of the results by modifying the input
parameters to have different correlations between temperature and
salinity, standard deviations, and spatial correlation range.  In all
runs, the \textit{myopic} and \textit{look-ahead} strategies perform
the best in terms of realized IBV, and much better than
\textit{naive}. The \textit{look-ahead} strategy seems to be
substantially better than the \textit{myopic} design only for very
small initial standard deviation or very large spatial correlation
range. \textit{static\_north} continues to be the best static design
for IBV, while \textit{static\_zigzag} is the best design for the
other predictive performance measures, especially so with large
spatial correlation range. We also ran simulation studies with only
temperature data, and for realistic correlation levels between
temperature and salinity, the IBV results are not much worse when only
temperature data are available. In addition to the comparison made in
Table \ref{tab:sim_rhoab}, the current setting includes spatial
correlation and this likely reduces the additional influence of having
bivariate data. However, it seems that having temperature data alone
does a substantially worse job in terms of explained variance

### 2020-06-18 section3_old.tex
Updates from Overleaf

--- 2020-06-18 section3.tex
@@
 
-The goal of this section is to present data collection strategies that aim at decreasing the uncertainty on the target excursion set $\es$.
+An important concept underlying this work is the potential of robotic
+sampling to adapt survey plans based on data and statistical methods
+in accordance with the sense-plan-act control loop
+(Fig. \ref{fig:envir1}). We now present a few adaptive strategies for
+robotic sampling that build on the notion of EIBV.
 
 \subsection{Background}
-As before, let $\gp$ be a $\no$-variate random field on $\domain$.
-See Section \ref{sec:bg_and_notation} for a recap of the specific notation used to handle mean and covariance function of multivariate random fields.
-
-In a sequential setting, the point of view is that $n$ data collection steps have already been performed and one wants to choose what data to collect next.
-After each step, the GRF model is updated using the cokriging equations \ref{eq:cokrig_mean} and \ref{eq:cokrig_cov} and we will denote by $\mathbb{P}_n$ the law of the field conditional on the data gathered up to (and including) stage $n$.
-
-\begin{remark}
-The type of data collected at each stage can be of any type (observation of all components of the field at a single location, observation of some components only but at several locations, ...) since the concept of \textit{generalized location} allows the cokriging equations to handle all those case by putting them on equal footage.
-\end{remark}
-
-Our goal is then to optimally select the next spatial location $\spatloc_{n+1}$ at which to observe the field (all of its component), where the meaning of optimal is yet to be defined.
-Note that here again, the concept of \textit{generalized location} allows one to consider settings where only one component of the field can be observed at each step and the startegy has to choose the spatial location as well as the component to observe (heterotopic case).
-In the same fashion, one could consider settings where several observations are allowed at each stage (batch case).
-
-We will here restrict ourselves to the case where all components of the field are observed at one spatial location at each stage (isotopic case).
 
@@
 \label{Optdes}
+
+Our adaptive AUV survey is split into many stages. The selection on
+where to sample is constrained by a triangular waypoint graph and
+defined by its nearest grid nodes.
+
+Let $d^{j,s}$ denote design number $j$ at stage $s$ of a sequential
+survey.
+For the AUV sampling situation, the choice of design entails a commitment to one of many possible navigation directions on the waypoint graph.
+Once the choice has been made, the AUV steers in the chosen direction and this entails sampling locations $\bu_{d^{j,s}}$ and data $\by_{d^{j,s}}$ is gathered.
+We let
+$\mathcal{Y}^{s-1} =\{(d^{*,1}, \by_{d^{*,1}}),\ldots, (d^{*,s-1},
+\by_{d^{*,s-1}}) \}$ denote the set of all data gathered until stage
+$s-1$ according to selected design $d^{*,1},\ldots,d^{*,s-1}$.
+
+The design and data define
+the conditional GP with mean
+$E(\gp[\x]|\mathcal{Y}^{s-1})$ and (cross) covariance matrix
+$\mbox{Var}(\gp[\x]|\mathcal{Y}^{s-1})$.
+These are efficiently updated in a sequential manner keeping track of the mean and covariance kernel of the GP over the stages.
+As extensions of Eq.
+\eqref{eq:cokrig} and \eqref{eq:krig:cov}, we have
+\begin{eqnarray}\label{eq:krigSUR}
+\mu^{s}(\x) &=&\mu^{s-1}(\x)+\lambda^{[s,s]}(\x,\bu_{d^{j,s}})^T (\by_{d^{j,s}}-G_{d^{j,s}} \mu(\x)), \nonumber \\
+k^{s}(\x,\x') &=& k^{s-1}(\x,\x')-\lambda^{[s,s]}(\x,\bu_{d^{j,s}})^T K_{Y_{d^{j,s}}} \lambda^{[s,s]}(\x',\bu_{d^{j,s}}),
+\end{eqnarray}
+where $\lambda^{[s,s]}(\x,\bu_{d^{j,s}})=K^{-1}_{Y_{d^{j,s}}} G_{d^{j,s}} k(\bu_{d^{j,s}},\x)$ and the data covariance is $K^{-1}_{Y_{d^{j,s}}}=G_{d^{j,s}} K^s(\bu_{d^{j,s}},\bu_{d^{j,s}})G^T_{d^{j,s}}+R_{d^{j,s}}$, with measurement model as in Eq.
+\eqref{eq:meas}.
 
@@
 
+%The optimal path
+%selection situation is depicted in Fig. %\ref{fig:PathSelOpt},
+%\begin{figure}[b!]
+%\centering
+%\includegraphics[width=0.85\textwidth]{Figures/sequent_select.pdf}
+%\caption{Optimal sequential path selection.}\label{fig:PathSelOpt}
+%\end{figure}
+%where design choices are indicated by squares while %data realizations are indicated by circles. 
+%terms this optimal design is defined as
+%\begin{equation}\label{opt_crit}
+%    \bd^* = \mbox{argmin}_{d^{j,1}} \left\{ \int_{\by^{j,1}} \mbox{argmin}_{d^{j,2}} \left\{ \int_{\by^{j,2}} \ldots \pi(\by^{j,2}|\by^{j,1}) d\by^{j,2} \right\} \pi(\by^{j,1}) d \by^{j,1} \right\},
+%\end{equation}
+%where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages. 
 The mathematical expression for the optimal design involves a series
@@
 
-\textcolor{red}{Here introduce the fact that candidate points are always chosen in a set $\candidates$.
-Mention that we usually only allow a finite set.
-Mention that in the specific case of robot planning, we have a waypoint graph and we thus usually take $\candidates$ to be the set of current neighbouring nodes.
-Mention that yes, $\candidates$ normally depends on the current location (this is for robot, in general setting, depends on the location of the last observations), but we silently suppress this dependency for readability.
-Nevertheless, one should always remember that the current set of candidates depends on the current setting.}
+%The derived closed-form solutions for the EIBV are still important building blocks when constructing the sampling designs, as they will be used to score the different adaptive designs. % Efficient
+% calculation is important for using adaptive survey designs with
+% robotic platforms.
+
 \subsection{A Naive Sampling Strategy}
@@
 The simplest heuristic for adaptive sampling is to choose the next
-sampling location $\spatloc_{n+1}$ based on current excursion probabilities. At stage $\stage$, one reduces the
-largest levels of uncertainty by selecting the location $\spatloc_{\stage + 1}$ with
-$P_{\stage}(\gp_{\spatloc_{\stage + 1}} \in T)$ closest to $0.5$..
+sampling location based on current probabilities. One reduces the
+largest levels of uncertainty by selecting the design $j$ with
+$P(Z(\x_j) \in T)$ closest to $0.5$.  With temperature and salinity
+variables, this {\it{naive}} strategy selects the stage $s$ sampling
+design according to:
+\begin{eqnarray}\label{critNaive}
+    d^{*,s} &=& \mbox{argmin}_{j} |P(Z_1(\x_j) \leq t_1, Z_2(\x_j) \leq t_2 | \mathcal{Y}^{s-1})-0.5|, \nonumber
+\end{eqnarray}
+where $j$ is indexing all neighboring locations on the waypoint graph.
 
 This strategy does not account for the uncertainty in the GP
-variables, \textcolor{red}{(C.T.: do we need this sentence? does it mean anything?) but is instead applicable only if one design (node) has EPs
+variables, but is instead applicable only if one design (node) has EPs
 closer to $0.5$ (see Table \ref{tab:sim_rhoab}, row
-two).}
-
-Additionally, this strategy does not account for spatial
+two). Additionally, this strategy does not account for spatial
 correlation. This strategy lacks memory of where it has been and where
@@
 
-Based on the currently available data the myopic strategy selects the location that leads to the biggest expected reduction in IBV:
-\begin{criterion}[Myopic]
-\begin{equation}\label{critSEQ}
-    \spatloc_{\stage + 1} = \argmin_{\spatloc \in \candidates} \EIBV_{\stage}\left(\spatloc\right)
-\end{equation}
-\end{criterion}
+Based on the currently available data $\mathcal{Y}^{s-1}$, we fit an
+updated GP model. Focusing on the EIBV criterion, at the next stage
+$s$, the selected design is:
+\begin{eqnarray}\label{critSEQ}
+    d^{*,s} &=& \mbox{argmin}_{j} \left\{ \int_{\mathcal{M}} E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( p(\x;\mathcal{Y}^{j,s})\left( 1-p(\x;\mathcal{Y}^{j,s})\right) \right) d\mes(u), \right\} \\
+    p(\x;\mathcal{Y}^{j,s}) &=& P(Z_u \in T |\by_{d^{j,s}},\mathcal{Y}^{s-1}). \nonumber
+\end{eqnarray}
 
-The expected IBV might be efficiently computed for each candidate points using Proposition \ref{propo2}.
-Once the best location has been selected, a stage of observation is performed there, the GP model is updated, yielding a conditional law $\mathbb{P}_{\stage + 1}$ and the process is repeated.
+Note that this strategy gives a sequential conditional version of
+expression (\ref{eq:eibv}). Now $\mathcal{Y}^{s-1}$ is available, and
+the expectation is conditional to this data. A similar closed-form
+calculation for EIBV is hence applicable in Eq.  \eqref{critSEQ} using the updated
+GP model from stage $s-1$ as defined via Eq.
+\eqref{eq:krigSUR} in the results from Proposition \ref{propo2}.
+Once the data are collected for the best design $d^{*s}$ at stage $s$, the GP model is updated again.
+The mean and covariance are used to compute the next design at stage $s+1$, and so on.
 
@@
 reasonable approach for creating designs in many
-applications.
-Moreover, it can be implemented in a way that is not too demanding on computational power, making it well-suited for embedding on autonomous mobile survey platform (see Section \ref{sec:case_study} for an example).
+applications. Moreover, it is easily implemented, in our case, onboard
+an AUV, using an efficient data update of the GP model and the
+calculation of the closed-form EIBV expressions for each subsequent
+trajectory.
 
-\subsection{Look-ahead Trajectory Planning}\label{sec:LA}
+\subsection{Look-ahead Trajectory Planning}
+\label{sec:LA}
 
@@
 measurements, and the strategy is then optimal in the sense that it accounts consistently for the expectations and minimizations in these two stages, but still without including any planning beyond those two steps.
-\midskip
-Say we have collected data up to stage $n$. The principle of two-steps lookahead
-strategies is to select the next observation location $\spatloc_{\stage + 1}$ as follows
+In addition to the
+next stage of measurements $\by^{j,s}$, this look-ahead strategy anticipates the subsequent design indexed $j_2$ with data
+$\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$. The
+selected design is:
+\begin{eqnarray}\label{critLA}
+    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( \mbox{argmin}_{j_2}  V_{j_2} \right) \right\}, \\
+V_{j_2} & = & \int_{\mathcal{M}} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1}) ( 1-p(\x;\mathcal{Y}^{j,s+1}) ) \right) d\mes(u), \nonumber \\
+    p(\x;\mathcal{Y}^{j,s+1}) &=& P(Z(\x) \in T |\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}). \nonumber
+\end{eqnarray}
+Here, $\mathcal{Y}^{j,s}=\{\by_{d^{j,s}},\mathcal{Y}^{s-1}\}$ and
+$\mathcal{Y}^{j,s+1}=\{\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}\}$
+represent the sets of data variables, and the expectations are
+calculated with respect to the conditional densities given data at
+that stage.  We solve the first expectation by Monte
+Carlo sampling of data $\by_{d^{j,s}}$ from its conditional
+distribution. For each of these data samples, the second expectation
+is solved using the closed-form expressions for EIBV using the results from Proposition \ref{propo2}, now with conditioning to the first stage data already going into the mean and covariances in Eq.
+\eqref{eq:krigSUR}.
 
-  \begin{itemize}
-      \item First, do as if the $\stage + 1$ observation had already been done and say
-          we observed $y_{\stage + 1}$ at $\spatloc_{\stage + 1}$.
-      \item If we were at this stage, then we would choose the next points by
-          minimizing the expected IBV under the current ($\stage + 1$) conditional
-          distribution
-          \[
-              \spatloc_{\stage + 2} = \argmin_{\spatloc \in \candidates}~ \EIBV_{\stage + 1}\left(\spatloc \mid \spatloc_{\stage + 1},
-          y_{\spatloc + 1}\right)
-          \]
-      \item Now the above expression depends on the observed data $y_{\stage + 1}$,
-          whose value we don't know beforehand. What we know is that
-          at step $\stage$ it is distributed according to the conditional distribution
-          $\mathbb{P}_{\stage}$. We can thus replace it by a random variable $Y$
-          following the current conditional law and then average over it.
-  \end{itemize}
-
-  Let us now try to make the above more precise.
-  \begin{criterion}[2-steps lookahead]
-      The next observation location $\spatloc_{\stage + 1}$ is selected according to
-      \begin{align}\label{critLA}
-          \spatloc_{\stage + 1} = \argmin_{\spatloc \in \candidates}~ \mathbb{E}_{\stage}\left[\min_{\spatloc' \in
-                  \candidates(\spatloc)} \EIBV_{\stage + 1}\left(\spatloc' \mid \spatloc,
-      Y\right)\right]
-      \end{align}
-  where $Y$ is a random variable distributed according to the conditional
-  law of $\gp_{\spatloc}$ at step $\stage$ (note it silently depends on the considered point
-  $\spatloc$) and $\EIBV_{\stage +  1}(\spatloc'\mid \spatloc,Y)$ is the
-  expected IBV with the expectation taken according to the conditional law $\mathbb{P}_{\stage}(\cdot\mid
-  \spatloc, Y)$. Note that the dependence of the set of candidates on the current location has been made explicit for the second stage of measurements.
-  \end{criterion}
-
-\begin{remark}
-In a practical setting, the first expectation might be computed by Monte
-Carlo sampling of data $Y$ from its conditional
-distribution. For each of these data samples, the second expectation
-is solved using the closed-form expressions for EIBV provided b< Proposition \ref{propo2}, now with conditioning to the first stage data already going into the mean and covariances in Eq.
-\eqref{eq:krigSUR}.
-\end{remark}
+Even though the strategy looks at two sampling stages, it is only used to find
+the current best design. When data are collected at stage $s$, the GP model is
+updated as in Eq. \eqref{eq:krigSUR}, and the mean and (cross) covariance are
+used to compute the next design for stage $s+1$, now anticipating what
+stage $s+1$ and $s+2$ could offer, and so on.
+This look-ahead approach is much more computationally demanding than
+the myopic strategy, so for practical implementation onboard the AUV this is reaching the limits allowed for calculation.
+We conduct some pruning of the least valuable design paths in the evaluation of Eq.
+\eqref{critLA}, to get a satisfactory computation time below one minute per waypoint.
+This means that we use the myopic
+strategy to rank the three best designs following the first stage alone, and
+for each of these preferred designs, we undertake the look-ahead calculations.
 

### 2020-06-19 main_new.tex
Updates from Overleaf

@@
 
+Motivated by the challenges related to efficient data collection strategies for the vast oceans, the combination of Gaussian process modeling is explored for autonomous robotic sampling of ocean features; enabling information driven measures  sampling efforts to high-interest regions.
+This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
+
 \subsection{Oceanic data collection and spatial design of experiments}
 
-%\textcolor{blue}{NB: I started to match previously listed thoughts/ideas with pieces of the former intro.
-Will do more asap.
-Also, Section 3 I find excellent (and more precise than the intro) and start thinking that some bits might end up in the newer intro (while some others could land in a later section).}
-
 Monitoring the world's oceans has gained increased importance in light of the changing climate and increasing anthropogenic impact.
-\begin{comment} % TO BE RE-INSERTED LATER IN THE TEXT?
-A central problem for understanding these factors is the lack of representative data with sufficient resolution.
-Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of sampling.
-By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
-This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
-\end{comment}
+
 Central to understanding the changes taking place in the upper
@@
 
-\begin{comment}
-\medskip
-
-%Our objective is to leverage statistical tools onboard an autonomous robotic platform to characterize a river plume, focusing on spatial separation of cold freshwater from a river and warmer saline waters of a fjord.
-
-The temperature and salinity field over the domain of interest will be modeled as a bivariate gaussian random field.
-The \textcolor{red}{complement of the} river plume may then be characterized as an \textit{excursion set} of the field, that is, as the region of the domain for which temperature and salinity lie above some specified threshold.
-Our goal is then to guide the data collection process so as to improve the characterization of this set.
-The methods we develop are fairly general, and not limited to the special case of two-dimensional random fields.
-Hence, the rest of this section will consider generic excursion sets of gaussian random fields with an arbitrary number of output dimensions.
-
-\begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/example_excu_1.png}
-  \caption{Realization of a bivariate GRF and excursion set above some threshold. Joint excursion in red, excursion of a single variable in white.}
-\label{example_excu}
-\end{figure}
-
-Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
-and we are interested in recovering the set of locations $\es$ in the domain for which the response variables (the components of $\gp$) lie in some set of specified values $\T\subset \mathbb{R}^{\no}$:
-$$
-\es:=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
-$$
-Our goal here is to develop approaches to quantify and improve the
-characterization of uncertainties on $\es$. \textcolor{red}{Alread mention how observations are included.}
-
-If we assume that $\gp$ has
-continuous trajectories (almost surely) and $T$ is closed, then
-$\es$ becomes a Random Closed Set
-\citep{Molchanov2005} and concepts from the theory of random sets will prove useful in characterizing the distribution of the volume of $\es$ under some measure defined on the domain.
-
-\textcolor{red}{Is this really needed?} Note that while some aspects of the developed approaches do not call for a
-specific form of $\T$, we will often stick for simplicity to the case
-of orthants
-($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
-$t_1,\dots, t_{\no} \in \R$) as this will allow efficient calculation
-of several key quantities. Note that changing some $\leq$ inequalities
-to $\geq$ ones would lead to immediate adaptations.
-
-\end{comment}
 \subsection{Background, Notation and Co-Kriging}
@@
 A bivariate gaussian random field model is assumed for $\gp$, see the two first panels of Figure~\ref{example_excu} for an illustration of a realization of such a vector-valued model (RK NOTATION).
-The right panel of this figure represents a by-product of interest derived from these realizations, namely regions i) in red, where both temperature and salinity are high (i.e., exceeding respective thresholds), standing for ocean water ii) in blue, where both temperature and salinity are low, standing for riverine water, and iii) in white, where one variable is above and the other below their respective thresholds, standing for mixed waters.
+The right panel of this figure represents a by-product of interest derived from these realizations, namely regions i) in red, where both temperature and salinity are high (i.e., exceeding respective thresholds), standing for ocean water ii) in pink, where both temperature and salinity are low, standing for riverine water, and iii) in white, where one variable is above and the other below their respective thresholds, standing for mixed waters.
 
@@
   \includegraphics[width=0.99\textwidth]{Figures/example_excu_1.png}
-  \caption{Realization of a bivariate GRF and excursion set above some threshold. Joint excursion in red, excursion of a single variable in white.}
+  \caption{Realization of a bivariate GRF and excursion set above some threshold. Joint excursion in red, excursion of a single variable in pink.}
 \label{example_excu}
@@
 %
+\textcolor{red}{(ADDENDUM, C.T.) In the same fashion, we will use $\mu(\bm{x})$ to denote the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ for the corresponding $q \times q$ covariance matrix.}
 
@@
 \label{sec:cokriging}
+\textcolor{red}{(C.T.) Why mention batches?
+We do not care.
+We have previously available data at $\bm{x}$ and want to predict at $\bm{x}$'.
+No need to introduce concepts we will never use.
+Moreover, this presentation suggest that to go from n to n+1 (batch-) observations, we have to condition on the whole $n$ dataset, instead of only including the last step.}
+
 Assuming now that $n$ batches of observations are available, with respective sizes $q_1,\dots, q_n$, and that one wishes 
@@
 %
-\begin{equation}
+\begin{equation}\label{eq:cokrig_mean}
 \mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})),
@@
 %
-\begin{equation}
+\begin{equation}\label{eq:cokrig_cov}
 k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T k(\bm{x}_{[n]}, \bm{x}_{[n]}) \lambda_{{[n]}}(\bm{x}').
@@
 Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $\domain^2$. In 
-contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves 
+contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{bect2019} involves 
 solely an integral on $\domain$ and can be expanded in our present settings as follows
@@
 and probability conditional on the first $n$ batches of observations, respectively.
-We want to co pute the effect of the inclusion of a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and $\ibv$.
+We want to compute the effect of the inclusion of a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and $\ibv$.
 
 For uncertainty functional such as $\ibv$ and $\emv$, we use a similar indexing system
-with respect to a sequence of batch, and further indicate when applicable the new batch
-of generalized location between parentheses; e.g., $\emv_{[n+1]}(\bm{x})$ denotes the value
-of $\emv$ (a random variable, seen from ``step'' $n$, i.e. when knowing $Z$ solely at
-the $n$ first generalized location batches) when adding observations at $\bm{x}$ to those
-already available at step $n$.
-
-\textcolor{red}{C.T.: This doesn't work, we cannot be that short on the definition of EIBV, it is highly unclear. I suggest we go for the following notation:
-$$\operatorname{IBV}(\bm{x}; \bm{y})
-$$ denotes the IBV under the current law of the field, conditioned on observing $\bm{y}$ at $\bm{x}$ (generalized, possibly batch observation). Then
-$$
-\operatorname{EIBV}(\bm{x}):=\mathbb{E}_{\bm{Y}}\left[\textrb{IBV}(\bm{x}; \bm{Y})\right]
-$$
-where $\bm{Y}$ is distributed according to the current law of $Z_{\bm{x}}$.}
+with respect to a sequence of batch and use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.
+
+\medskip
+
+In order to study the effect of the inclusion of a new data point, we let
+$
+\currentIBV(\bm{x}; \bm{y})
+$
+denote the IBV under the current law of the field, conditioned on observing $\bm{y}$ at $\bm{x}$ (generalized, possibly batch observation).
+The expected effect of a new observation on the IBV is then captured by
+\begin{equation}\label{def:eibv}
+    \currentEIBV(\bm{x}):=\mathbb{E}_{\stage}\left[\textrb{IBV}(\bm{x}; \bm{Y})\right]
+\end{equation}
+where $\bm{Y}$ is distributed according to the current law of $Z_{\bm{x}}$.
 
@@
 
-\textcolor{red}{When one want to consider several steps, one can use $\textrm{IBV}_{n}(\bm{x}; \bm{y})$ to denote the fact that the starting point (starting probability distribution) is the law of the field at the $n$-th step $\mathbb{P}_n$ to which conditioning on $\bm{y}$ at $\bm{x}$ is added.}
-
-\textcolor{red}{For a rigorous definition
-$$
-\operatorname{IBV}(\bm{x}; \bm{y}):=\int_{\domain}
-\mathbb{P}\left(\gp[\uu]\in T\mid Z_{\bm{x}}=\bm{y}\right)(1-\mathbb{P}\left(\gp[\uu]\in T \mid Z_{\bm{x}}=\bm{y}\right))
-d\mes(u) \\
-$$
-where $\mathbb{P}$ denotes the current law of the field.}
+We next present a result that allows to compute $\EIBV$ as an integral of CDFs of multivariate gaussians. This will prove useful when designing sequential uncertainty reduction strategies.
 
@@
 \label{propo_eibv}
-Let $\eibv_{[n]}(\bm{x})=\currentExp{\ibv_{[n+1]}(\bm{x})}$ (where $\bm{x}$ plays
-the role of $\bm{x}_{n+1}$) be the expected effect of adding an observation a $\bm{x}$ on the IBV.
-Then we have
 \begin{equation}
 \begin{split}
-\eibv_{[n]}(\bm{x})
+\currentEIBV(\bm{x})
 &=\int_{D} \varPhi_{\no}\left(\bt;~\currentMean{\uu}, \currentCov{u, u}\right) d\mes(u)\\
@@
 
-\textcolor{red}{Do we really want to keep this?}
-A critical element in the derivation of a closed-form for the EIBV is
-that the conditional mean in the cokriging equations is a linear
-(affine) function of the data $\by_d$ and the covariance is not a
-function of the data. This in turn means that the probabilities in Eq.
-\eqref{eq:post_ep} involve inequality statements for linear
-combinations of Gaussian variables. Related closed-form solutions have
-been noted in similar contexts \citep{bhattacharjya2013value,
-  chevalier2014fast,stroh}, but not generalized to our situation with
+\begin{remark}
+Propositions \ref{propo_eibv} and \ref{propo_emv} are twofold generalizations of  results from \cite{chevalier2014fast}: they extend previous results to the multivariate setting and also allow for the inclusion of batch or heterotopic observations through the concept of generalized locations.
+Related closed-form solutions have been noted in similar contexts \citep{bhattacharjya2013value,stroh}, but not generalized to our situation with
 random sets for vector-valued GPs.
+\end{remark}
 
@@
 
+% To BE COMMENTED (TEMPORARILY?)
 Fig. \ref{illus_bivarDens} shows contour plots of three different
@@
 standard deviations.
+% END TO BE COMMENTED (TEMPORARILY?)
 
@@
 \begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/ebv_comp.png}
+  \includegraphics[width=0.99\textwidth]{Figures/ebv_comp_2.png}
   \caption{Pointwise Bernoulli variance reduction for observation of a single or both components of the random field at one location.
@@
 \begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/ebv_north_vs_east.png}
+  \includegraphics[width=0.99\textwidth]{Figures/ebv_north_vs_east_2.png}
   \caption{Pointwise Bernoulli variance reduction for two different observation plans.

### 2020-06-19 section2_old.tex
Updates from Overleaf

--- 2020-06-17 section2.tex
@@
 \label{sec:set_uq}
-\input{UQonSets}
+We now introduce quantities that allow to quantify the uncertainty on the volume of the excursion set $\es$. Let $\mes$ be a 
+(locally finite, Borel) measure  on $\domain$. We want to investigate the probability distribution 
+of $\mes(\es)$ through its moments.
+
+\medskip
+
+Centred moments may be computed using Proposition~\ref{propo1} developed in the appendix. 
+In particular, it allows to write the excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ as an integral of the excursion probability
+\begin{equation*}
+\begin{split}
+\emv
+&=\int_{\domain^2} \mathbb{P}\left(
+\gp[u]\in T, \gp[v]\in T \right)
+d\mes^{\otimes}(u, v)\\
+&-\left( \int_{\domain} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
+\end{split}
+\end{equation*}
+which boils down in the excursion/sojourn case where $\T=(-\infty, t_1] \times
+\dots \times (-\infty, t_{\no}]$ to
+\begin{equation*}
+\begin{split}
+\emv
+%\operatorname{Var}[\mes(\es)]
+&=\int_{\domain^2}
+\varPhi_{2\no}
+\left(
+(\bt, \bt); \mu((u,v)),
+K((u,v),(u,v))
+\right)
+\
+\mathrm{d}\mes^{\otimes} %\mes
+%\productMeasure
+(u,v)\\
+&-\left( \int_{\domain} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
+\end{split}
+\end{equation*}
+%
+Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $\domain^2$. In 
+contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves 
+solely an integral on $\domain$ and can be expanded in our present settings as follows
+\begin{equation*}
+\begin{split}
+\operatorname{IBV} %(\es) %;\mes)
+&=\int_{\domain}
+\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
+d\mes(u) \\
+&=\int_{\domain}
+\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right)
+-\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
+\mathrm{d}\mes(u).
+\end{split}
+\end{equation*}
+%
+Next we investigate uncertainty reduction criteria related to those indicators.
 
@@
 \label{sec:eibv}
-\input{eibv}
+This section is devoted to computing the expected effect of the inclusion of new observations on the $\emv$ and $\ibv$ of the excursion set $\es$.
+Let us consider the same setting as in Section \ref{sec:cokriging} where $n$ batches of measurements have already been performed, and let
+$\currentExp{.}$ and $\currentProba{.}$ denote conditional expectation
+and probability conditional on the first $n$ batches of observations, respectively.
+We want to co pute the effect of the inclusion of a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and $\ibv$.
+
+For uncertainty functional such as $\ibv$ and $\emv$, we use a similar indexing system
+with respect to a sequence of batch, and further indicate when applicable the new batch
+of generalized location between parentheses; e.g., $\emv_{[n+1]}(\bm{x})$ denotes the value
+of $\emv$ (a random variable, seen from ``step'' $n$, i.e. when knowing $Z$ solely at
+the $n$ first generalized location batches) when adding observations at $\bm{x}$ to those
+already available at step $n$.
+
+\textcolor{red}{C.T.: This doesn't work, we cannot be that short on the definition of EIBV, it is highly unclear. I suggest we go for the following notation:
+$$\operatorname{IBV}(\bm{x}; \bm{y})
+$$ denotes the IBV under the current law of the field, conditioned on observing $\bm{y}$ at $\bm{x}$ (generalized, possibly batch observation). Then
+$$
+\operatorname{EIBV}(\bm{x}):=\mathbb{E}_{\bm{Y}}\left[\textrb{IBV}(\bm{x}; \bm{Y})\right]
+$$
+where $\bm{Y}$ is distributed according to the current law of $Z_{\bm{x}}$.}
+
+\textcolor{red}{
+The philosophy is the following: EIBV and EEMV are functionals of a probability distribution.
+Then, if not mentioned explicitly, the probability distribution will be taken to be the current law of the field.
+(This notation is highly efficient).}
+
+\textcolor{red}{When one want to consider several steps, one can use $\textrm{IBV}_{n}(\bm{x}; \bm{y})$ to denote the fact that the starting point (starting probability distribution) is the law of the field at the $n$-th step $\mathbb{P}_n$ to which conditioning on $\bm{y}$ at $\bm{x}$ is added.}
+
+\textcolor{red}{For a rigorous definition
+$$
+\operatorname{IBV}(\bm{x}; \bm{y}):=\int_{\domain}
+\mathbb{P}\left(\gp[\uu]\in T\mid Z_{\bm{x}}=\bm{y}\right)(1-\mathbb{P}\left(\gp[\uu]\in T \mid Z_{\bm{x}}=\bm{y}\right))
+d\mes(u) \\
+$$
+where $\mathbb{P}$ denotes the current law of the field.}
+
+\begin{propo}
+\label{propo_eibv}
+Let $\eibv_{[n]}(\bm{x})=\currentExp{\ibv_{[n+1]}(\bm{x})}$ (where $\bm{x}$ plays
+the role of $\bm{x}_{n+1}$) be the expected effect of adding an observation a $\bm{x}$ on the IBV.
+Then we have
+\begin{equation}
+\begin{split}
+\eibv_{[n]}(\bm{x})
+&=\int_{D} \varPhi_{\no}\left(\bt;~\currentMean{\uu}, \currentCov{u, u}\right) d\mes(u)\\
+&-\int_{D} \varPhi_{2\no}
+\left(
+\left(
+\begin{matrix}
+\bt-\currentMean{u}\\
+\bt-\currentMean{u}
+\end{matrix}
+\right);
+\mathbf{\Sigma}_{[n]}(\uu)
+\right)
+d\mes(u),
+\end{split}
+\end{equation}
+where the matrix $\mathbf{\Sigma}_{[n]}(\uu)$ is defined as
+\begin{equation*}
+\begin{split}
+\mathbf{\Sigma}_{[n]}(\uu)&=
+\left(
+\begin{matrix}
+\currentCov{u, u} & \currentCov{u, u}-\futureCov{u, u}\\
+\currentCov{u, u}-\futureCov{u, u} & \currentCov{u, u}
+\end{matrix}
+%\begin{matrix}
+%\currentCov{u, u} & \Delta_{[n]}(\uu)\\
+%\Delta_{[n]}(\uu) & \currentCov{u, u}
+%\end{matrix}
+\right).\\
+\end{split}
+\end{equation*}
+\end{propo}
+
+As for the expected excursion measure variance, a similar result may be derived.
+\begin{propo}
+\label{propo_emv}
+
+\begin{equation*}
+\begin{split}
+\eemv_{[n]}(\bm{x})
+&=\int_{D^2} 
+\varPhi_{2\no}
+\left(
+(\bt, \bt); \mu((u,v)), 
+K((u,v),(u,v))
+\right) 
+\
+\mathrm{d}\mes^{\otimes} %\mes 
+%\productMeasure 
+(u,v)\\
+&-\currentExp{
+    \varPhi_{\no}\left(a_1+B_1 V; C_1\right)
+    \varPhi_{\no}\left(a_2+B_2 V; C_2\right)
+},
+\end{split}
+\end{equation*}
+
+with $V=\gp[\bm{x}_{n+1}]-\currentMean{\bm{x}_{n+1}} \sim \mathcal{N}(0_{q_{n+1}},k_{[n]}(\bm{x}_{n+1},\bm{x}_{n+1}))$ and $a_1=\bt-\currentMean{\uu}$,
+$B_1=-\lambda_{[n+1,n+1]}(\uu)^T$, $a_2=\bt-\currentMean{\vv}$, $B_2=-\lambda_{[n+1,n+1]}(\vv)^T$.
+
+This integrand finally boils down to an evaluation of the $2\no$-dimensional Gaussian CDF by
+Proposition~\ref{propo3}. Details are omitted here for brevity as we did not implement further
+this criterion that requires numerical integration over $\domain^2$.
+\end{propo}
+
+Note that the calculation of the multivariate Gaussian cumulative distribution
+may be done effectively using code such as that of
+\cite{genz2009computation}.
+
+\textcolor{red}{Do we really want to keep this?}
+A critical element in the derivation of a closed-form for the EIBV is
+that the conditional mean in the cokriging equations is a linear
+(affine) function of the data $\by_d$ and the covariance is not a
+function of the data. This in turn means that the probabilities in Eq.
+\eqref{eq:post_ep} involve inequality statements for linear
+combinations of Gaussian variables. Related closed-form solutions have
+been noted in similar contexts \citep{bhattacharjya2013value,
+  chevalier2014fast,stroh}, but not generalized to our situation with
+random sets for vector-valued GPs.
 

### 2020-06-19 section3.tex
Updates from Overleaf

@@
 After each step, the GRF model is updated using the cokriging equations \ref{eq:cokrig_mean} and \ref{eq:cokrig_cov} and we will denote by $\mathbb{P}_n$ the law of the field conditional on the data gathered up to (and including) stage $n$.
+More generally, we will use subscripts to denote quantities conditional on the data available up to stage $\stage$.
+Hence, $\mu_{\stage}(\cdot)$ and $k_{\stage}(\cdot, \cdot)$ will denote conditional mean and covariance functions conditional on the first $\stage$ data collection stages.
 
@@
 
-Our goal is then to optimally select the next spatial location $\spatloc_{n+1}$ at which to observe the field (all of its component), where the meaning of optimal is yet to be defined.
-Note that here again, the concept of \textit{generalized location} allows one to consider settings where only one component of the field can be observed at each step and the startegy has to choose the spatial location as well as the component to observe (heterotopic case).
-In the same fashion, one could consider settings where several observations are allowed at each stage (batch case).
+Our goal is then to optimally select what observations to do next, where the meaning of optimal is yet to be defined.
+Note that here again, the concept of \textit{generalized location} allows one to consider settings where only one component of the field can be observed at each step and the strategy has to choose the spatial location as well as the component to observe (heterotopic case), or settings where several observations are allowed at each stage (batch case).
 
-We will here restrict ourselves to the case where all components of the field are observed at one spatial location at each stage (isotopic case).
+We will here restrict ourselves to the case where only one spatial location can be chosen at each stage and all components of the field are observed there (isotopic case), leaving the general case for future work.
 
@@
 
-At stage $s$, the robot must then choose design $d^{j,s}$ where measurements $\by_{d^{j,s}}$ should
-be gathered. For the triangularized waypoint
-graph all six possible
-directions must be evaluated to minimize the expected uncertainty of random sets, and in particular the EIBV for our 
-application. We hence focus on choosing designs for data modeled by
-Eq. \eqref{eq:meas}, to conduct efficient sequential uncertainty
-reduction according to the EIBV criterion. Compared with the
-static design $d$ in Eq. \eqref{eq:eibv}, the optimal sequential
-design not only considers the best current design grid point, but also
-how the data gathered at this node could inform future sampling at
-successive stages.
-
+\textcolor{red}{(C.T.) I suspect we need this paragraph so we cite people who have to be cited. How do we want to blend it with the rest? Personally I would get rid of this subsection.}
 The mathematical expression for the optimal design involves a series
@@
 measurements, and the strategy is then optimal in the sense that it accounts consistently for the expectations and minimizations in these two stages, but still without including any planning beyond those two steps.
-\midskip
+
+\medskip
+
 Say we have collected data up to stage $n$. The principle of two-steps lookahead
-strategies is to select the next observation location $\spatloc_{\stage + 1}$ as follows
+strategies is to select the next observation location $\spatloc_{\stage + 1}$ that would yield the biggest exepcted decrease in IBV if we were to (optimally) add one more observation after the one at $\spatloc_{\stage + 1}$.
+In order to formalize this concept, we need to be able to talk about the EIBV in the future (after observation $\stage + 1$ has been made).
+The definition \ref{def:eibv} of $\currentEIBV$ includes a subscript to stress that expectations are taken with respect the conditional law at stage $n$.
+We will extend this notion by writing $\currentEIBV(\cdot; u, y)$ to denote EIBV where expectations are taken conditional on the data available at $n$ an on an additional observation $y$ at $u$.
 
-  \begin{itemize}
-      \item First, do as if the $\stage + 1$ observation had already been done and say
-          we observed $y_{\stage + 1}$ at $\spatloc_{\stage + 1}$.
-      \item If we were at this stage, then we would choose the next points by
-          minimizing the expected IBV under the current ($\stage + 1$) conditional
-          distribution
-          \[
-              \spatloc_{\stage + 2} = \argmin_{\spatloc \in \candidates}~ \EIBV_{\stage + 1}\left(\spatloc \mid \spatloc_{\stage + 1},
-          y_{\spatloc + 1}\right)
-          \]
-      \item Now the above expression depends on the observed data $y_{\stage + 1}$,
-          whose value we don't know beforehand. What we know is that
-          at step $\stage$ it is distributed according to the conditional distribution
-          $\mathbb{P}_{\stage}$. We can thus replace it by a random variable $Y$
-          following the current conditional law and then average over it.
-  \end{itemize}
-
-  Let us now try to make the above more precise.
   \begin{criterion}[2-steps lookahead]
@@
           \spatloc_{\stage + 1} = \argmin_{\spatloc \in \candidates}~ \mathbb{E}_{\stage}\left[\min_{\spatloc' \in
-                  \candidates(\spatloc)} \EIBV_{\stage + 1}\left(\spatloc' \mid \spatloc,
+                  \candidates(\spatloc)} \EIBV_{\stage}\left(\spatloc' ; \spatloc,
       Y\right)\right]
@@
   law of $\gp_{\spatloc}$ at step $\stage$ (note it silently depends on the considered point
-  $\spatloc$) and $\EIBV_{\stage +  1}(\spatloc'\mid \spatloc,Y)$ is the
-  expected IBV with the expectation taken according to the conditional law $\mathbb{P}_{\stage}(\cdot\mid
-  \spatloc, Y)$. Note that the dependence of the set of candidates on the current location has been made explicit for the second stage of measurements.
+  $\spatloc$). Here, the dependence of the set of candidates on the current location has been made explicit for the second stage of measurements.
   \end{criterion}
@@
 distribution. For each of these data samples, the second expectation
-is solved using the closed-form expressions for EIBV provided b< Proposition \ref{propo2}, now with conditioning to the first stage data already going into the mean and covariances in Eq.
-\eqref{eq:krigSUR}.
+is solved using the closed-form expressions for EIBV provided by Proposition \ref{propo2}, now with conditioning on the first stage data already going into the mean and covariances in Eqs.
+\eqref{eq:cokrig_mean} and \eqref{eq:cokrig_cov}.
+\end{remark}
+
+\begin{remark}
+All our results have been formulated in a noiseless setting.
+That is, to us an observation $y$ at $\spatloc$ is a realization of $\gp_{\spatloc$.
+Generalization to a noisy setting where $y$ is a realization of $\gp_{\spatloc} + \epsilon$ with epsilon some noise process may be performed by including noise covariance matrices where necessary.
 \end{remark}

### 2020-06-20 main_new.tex
before running through from start.

@@
 % "New Title of the paper"
-\title{Autonomous Oceanographic Data Collection towards Estimating Excursion Sets of Vector-valued Random Fields} 
+\title{Autonomous Oceanographic Data Collection using Excursion Sets of Vector-valued Gaussian Random Fields} 
 \runtitle{Autonomous Oceanographic Data Collection}
@@
 
-Motivated by the challenges related to efficient data collection strategies for the vast oceans, the combination of Gaussian process modeling is explored for autonomous robotic sampling of ocean features; enabling information driven measures  sampling efforts to high-interest regions.
-This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
+Motivated by the challenges related to efficient data collection strategies for the vast oceans, the combination of spatial statistics, design of experiments and marine robotics is studied in this paper.
+The multidisciplinary efforts enable information-driven data collection in regions of high-interest regions.
 
@@
 %This slight change of notation 
-will allow us to think of $\gp$ as a scalar-valued Gaussian random field 
-indexed by $D\times \{1\dots,p\}$, which will give the co-kriging equations a particularly simple form that parallels the one of univariate kriging.
-Due to the naturality of the concept of \textit{generalized location},
-we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $\spatloc \in \domain$.
-In the following, the letter $x$ will be reserved for generalized locations, while the letters $\spatloc$ and $\ell$ will be used for spatial locations and response indices respectively.
+will allow us to think of $\gp$ as a scalar-valued %Gaussian 
+random field indexed by $D\times \{1\dots,p\}$, which will give the co-kriging equations a particularly simple form that parallels the ones of kriging for the scalar-valued case. 
+%Due to the naturality of the concept of \textit{generalized location},  we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $\spatloc \in \domain$.
+In the following, the letter $x$ will be reserved for generalized locations, while the letters $\spatloc$ and $\ell$ will be used for \textit{spatial locations} and \textit{response indices}, respectively.
 
 It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the 
-one for a single observation, provided some notation adjustments are made. This motivates the following.
-
-Given a dataset consisting of  $q$ observations at spatial locations $\spatloc_i \in \domain$ and response indices $l_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we can concatenate the information needed to specify the dataset by using the notation
+one for a single observation, provided some notation adjustments are made.
+This motivates the following.Given a dataset consisting of  $q$ observations at spatial locations $\spatloc_i \in \domain$ and response indices $\ell_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we can concatenate the information needed to specify the dataset by using the notation
 \begin{align*}
-\bm{x}:=(\bm{\spatloc}, \bm{\ell}):= (x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
+\bm{x}:=%(\bm{\spatloc}, \bm{\ell}):= 
+(x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
 \end{align*}
@@
 %
-\textcolor{red}{(ADDENDUM, C.T.) In the same fashion, we will use $\mu(\bm{x})$ to denote the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ for the corresponding $q \times q$ covariance matrix.}
-
-Given a random field $\gp$ and (noisy) observations of some of its components at some points in the domain, one can predict the value of the field at some unobserved location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$, conditional on the data.
+For a second order random field $(Z_{\spatloc})_{\spatloc \in \domain}$ with mean $\mu$ and matrix covariance function $K$, $\mu$ is naturally extended to $\domain \in \lbrace 1, ..., \no\rbrace$ into a function of $x=(\spatloc, \ell)$ and is further straightforwardly vectorized into a function of $\bm{x}$.
+As for $K$, it induces a covariance kernel $k$ on the set of extended locations via  $k((\spatloc, \ell),(\spatloc', \ell'))=K(\spatloc, \spatloc')_{\ell, \ell'}$.
+In vectorized/batch form, $k(\bm{x}, \bm{x}')$ then amounts to a matrix with numbers of lines and columns the numbers of generalized locations in the batches $\bm{x}$ in $\bm{x}'$, respectively.
+Such vectorized quantities will turn out to be quite useful in order to arrive at embarrassingly simple expressions for co-Kriging equations below.
+%\textcolor{red}{(ADDENDUM, C.T.) In the same fashion, we will use $\mu(\bm{x})$ to denote the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ for the corresponding $q \times q$ covariance matrix.}
+
+Now given a Gaussian random field $\gp$ and (noisy) observations of some of its components at some points in the domain, one can predict the value of the field at some unobserved location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$, conditional on the data.
 This process is called (co-)kriging, and kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
-We will here present the most general form of cokriging, where one can include several observations (batch) at a time and 
+We will here present a rather general form of cokriging, in the sense that it allows including several observations (batch) at a time and 
 observations at a given location $s \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ 
@@
 \label{sec:cokriging}
-\textcolor{red}{(C.T.) Why mention batches?
+% Adressed during a discussion on 19.06.2020
+%\textcolor{red}{(C.T.) Why mention batches?
 We do not care.
@@
 
-Assuming now that $n$ batches of observations are available, with respective sizes $q_1,\dots, q_n$, and that one wishes 
-to predict $\gp[\bm{x}]$ for some batch of $q\geq 1$ generalized locations $\bm{x}=(\bm{\spatloc}, \bm{\ell})$, the (simple) 
-cokriging mean would then amount to simple kriging with respect to a scalar-valued Gaussian random field indexed by 
-$D\times \{1\dots,p\}$ and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{\spatloc}, \bm{\spatloc}')_{\ell, \ell'}$, that is:
+Assuming that $n$ batches of observations are available, with respective sizes $q_1,\dots, q_n$, and that one wishes 
+to predict $\gp[\bm{x}]$ for some batch of $q\geq 1$ generalized locations $\bm{x}$, %\equiv(\bm{\spatloc}, \bm{\ell})$, 
+the (simple) cokriging mean would then amount to simple kriging with respect to a scalar-valued Gaussian random field indexed by 
+$D\times \{1\dots,p\}$:  
+%and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{\spatloc}, \bm{\spatloc}')_{\ell, \ell'}$, that is:
 %
@@
 \label{Sec:UnivarEx}
+
 We illustrate the EBV associated to different designs on a simple bivariate example.
@@
 \subsubsection{A pointwise study}
-Say we want to study the excursion probability of a bivariate gaussian, where the excursion set is defined as
-\[
-\Gamma := \lbrace \spatloc \in \domain: Z_{\spatloc, 1} \geq t_1, Z_{\spatloc, 2}\geq t_2\rbrace
-\]
-and the thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
+\textcolor{red}{(C.T.) What is EIBV in the table? How can we take an integral here? Also note that if we keep the ellipsoid plots, they should have Z's instead of Xi's}
+
+Say we want to study the excursion probability of a bivariate gaussian above some thresholds, where the thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
 
@@
 measurements alone. When both data are gathered, the measurement model is
-$(Y_{d,1},Y_{d,2})^t=(Z_1,Z_2)^t+\bepsilon$, with $\bepsilon \sim N(0,0.5^2I_2)$, while $Y_d=Z_1+\epsilon$, $\epsilon \sim N(0,0.5^2)$ when only temperature is measured.
+$(Y_1,Y_2)^t=(Z_1,Z_2)^t+\bepsilon$, with $\bepsilon \sim N(0,0.5^2I_2)$, while $Y_1=Z_1+\epsilon$, $\epsilon \sim N(0,0.5^2)$ when only temperature is measured.
 For this illustration, Table \ref{tab:sim_rhoab} shows that the
@@
 \subsubsection{Including Spatiality}
-INTRODUCTORY SENTENCE.
-The class of GRF model we will consider in the article will generally have a linear trend
+We now turn to an example involving a full-fledged gaussian random field (GRF). The class of GRF model we will consider in the article will generally have a linear trend
 \begin{align*}
 \mu(s)=\mathbb{E}\left[\begin{pmatrix}
-Z_{s, 1}\\ Z_{s, 2}
-\end{pmatrix}\right] &= \beta_0 + \beta_1 s
+Z_{\spatloc, 1}\\ Z_{\spatloc, 2}
+\end{pmatrix}\right] &= \beta_0 + \beta_1 \spatloc
 \end{align*}
-with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. We will only consider covariance models of separable type
+with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. In our examples, we will only consider covariance models of separable type
 \begin{align*}
-\textrm{Cov}\left(Z_{s, i}, Z_{u, j}\right) &= k(s, u) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_l^2,~ i=j\\
+\textrm{Cov}\left(Z_{\spatloc, i}, Z_{v, j}\right) &= k(\spatloc, v) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_l^2,~ i=j\\
    \gamma_0 \sigma_i \sigma_j,~i\neq j
@@
 
-In the accompanying Python code, these modeling assumptions can be
+In the accompanying Python code \textcolor{red}{(C.T.) Do we want to brand the code as a child of this project, or to say that we have implemented our experiments using a package that is available online (and not say too much)?}, these modeling assumptions can be
 generalized to anisotropic covariance and changing variance levels
-across the spatial domain. One can hence, easily run the code to see
-if different models result in other sampling designs. Anisotropy and
+across the spatial domain. Anisotropy and
 non-stationary variance are both relevant for the setting with river
@@
 
-We now study how the expected Bernoulli variance eq.\ref{eq:eibv} associated to data collection at a point changes if only one of the two components of the field is observed.
+We now study how the expected Bernoulli variance eq.\eqref{def:eibv} associated to data collection at a point changes if only one of the two components of the field is observed.
 We first draw a realization of the GRF defined above and use it as ground-truth to mimick the real data-collection process.
-A first data collection step is performed at the location in green, and the data is used to update the GRF model.
-The expected Bernoulli variance reduction (at each grid point) that would result from observing only one component, or both at the next grid point upwards is then computed an shown in Fig.\ref{fig:ebv_comp}.
+A first bunch of observations are done at the locations depicted in grey, and the data is used to update the GRF model.
+We then consider the green triangle as a potential next observation locations and plot the expected Bernoulli variance reduction (at each grid point) that would result from observing only one component of the field (temperature or salinity), or both at that point, see Fig.\ref{fig:ebv_comp}.
 
@@
 For example, Fig.
-\ref{fig:ebv_comp} shows the expected Bernoulli variance reduction associated to a data collection plan along a straight upwards line (static north) and the one associated to a straight horizontal line (static east).
+\ref{fig:ebv_north_vs_east} shows the expected Bernoulli variance reduction associated to a data collection plan along a straight upwards line (static north) and the one associated to a straight horizontal line (static east).
 Both expectations are computed according to the a priori distribution of the GRF (i.e.
@@
 
-\input{section3}
+\section{Sequential designs and heuristic path planning}\label{sec:heuristics}
+
+The goal of this section is to present data collection strategies that aim at decreasing the uncertainty on the target excursion set $\es$.
+
+\subsection{Background}
+As before, let $\gp$ be a $\no$-variate random field on $\domain$.
+See Section \ref{sec:bg_and_notation} for a recap of the specific notation used to handle mean and covariance function of multivariate random fields.
+
+In a sequential setting, the point of view is that $n$ data collection steps have already been performed and one wants to choose what data to collect next.
+After each step, the GRF model is updated using the cokriging equations \ref{eq:cokrig_mean} and \ref{eq:cokrig_cov} and we will denote by $\mathbb{P}_n$ the law of the field conditional on the data gathered up to (and including) stage $n$.
+More generally, we will use subscripts to denote quantities conditional on the data available up to stage $\stage$.
+Hence, $\mu_{\stage}(\cdot)$ and $k_{\stage}(\cdot, \cdot)$ will denote conditional mean and covariance functions conditional on the first $\stage$ data collection stages.
+
+\begin{remark}
+The type of data collected at each stage can be of any type (observation of all components of the field at a single location, observation of some components only but at several locations, ...) since the concept of \textit{generalized location} allows the cokriging equations to handle all those case by putting them on equal footage.
+\end{remark}
+
+Our goal is then to optimally select what observations to do next, where the meaning of optimal is yet to be defined.
+Note that here again, the concept of \textit{generalized location} allows one to consider settings where only one component of the field can be observed at each step and the strategy has to choose the spatial location as well as the component to observe (heterotopic case), or settings where several observations are allowed at each stage (batch case).
+
+We will here restrict ourselves to the case where only one spatial location can be chosen at each stage and all components of the field are observed there (isotopic case), leaving the general case for future work.
+
+\subsection{Optimal Sequential Design}
+\label{Optdes}
+
+\textcolor{red}{(C.T.) I suspect we need this paragraph so we cite people who have to be cited. How do we want to blend it with the rest? Personally I would get rid of this subsection.}
+The mathematical expression for the optimal design involves a series
+of intermixed maximizations over designs and integrals over data. In
+practice, the optimal solution is intractable because of the enormous
+growth over stages (see e.g. \cite{sucar2015probabilistic} and
+\cite{powell2016perspectives}).  Instead, we outline heuristic
+strategies.
+
+\textcolor{red}{Here introduce the fact that candidate points are always chosen in a set $\candidates$.
+Mention that we usually only allow a finite set.
+Mention that in the specific case of robot planning, we have a waypoint graph and we thus usually take $\candidates$ to be the set of current neighbouring nodes.
+Mention that yes, $\candidates$ normally depends on the current location (this is for robot, in general setting, depends on the location of the last observations), but we silently suppress this dependency for readability.
+Nevertheless, one should always remember that the current set of candidates depends on the current setting.}
+\subsection{A Naive Sampling Strategy}
+\label{naive}
+
+A very simple heuristic for adaptive sampling is to greedily evaluate $Z$ at the location with current excursion probability closest to $\frac{1}{2}$.While extremely easy to implement, this strategy can lead to spend too many iterations in boundary regions regardless of the the possible effect of sampling at the considered point on the future conditional distribution of $Z$.
+Working in moderate dimension and having a relatively rough discretization such as the considered waypoint graph can be beneficial by helping not being stuck for too long in poorly informative regions, yet it is not really desirable to be tributary of properties of the domain discretization in order to guarantee a fair amount of exploration.
+% REFS?
+
+\subsection{Myopic Path Planning}
+\label{sec:myopic}
+
+The myopic (greedy) strategy which we present here is optimal if we
+imagine taking only one more stage of measurements. This selection
+strategy is based on expected uncertainty reduction, but it is a
+heuristic because there is no anticipation of what the subsequent
+designs might offer beyond the first stage.
+
+Based on the currently available data the myopic strategy selects the location that leads to the biggest expected reduction in IBV:
+\begin{criterion}[Myopic]
+The next observation location $\spatloc_{\stage + 1}$ is chosen among the minimizers in $\candidates$ of the criterion:
+\begin{equation}\label{critSEQ}
+     C_{\text{myopic}}(u) = \EIBV_{\stage}\left(\spatloc\right)
+\end{equation}
+\end{criterion}
+
+The expected IBV might be efficiently computed for each candidate points using Proposition \ref{propo2}.
+Once the best location has been selected, a stage of observation is performed there, the GP model is updated, yielding a conditional law $\mathbb{P}_{\stage + 1}$ and the process is repeated.
+
+Even though this myopic strategy is non-anticipative, it still gives a
+reasonable approach for creating designs in many
+applications.
+Moreover, it can be implemented in a way that is not too demanding on computational power, making it well-suited for embedding on autonomous mobile survey platform (see Section \ref{sec:case_study} for an example).
+
+\subsection{Look-ahead Trajectory Planning}\label{sec:LA}
+
+We now extend the myopic strategy to a look-ahead strategy. We consider two stages of
+measurements, and the strategy is then optimal in the sense that it accounts consistently for the expectations and minimizations in these two stages, but still without including any planning beyond those two steps.
+
+\medskip
+
+Say we have collected data up to stage $n$. The principle of two-steps lookahead
+strategies is to select the next observation location $\spatloc_{\stage + 1}$ that would yield the biggest exepcted decrease in IBV if we were to (optimally) add one more observation after the one at $\spatloc_{\stage + 1}$.
+In order to formalize this concept, we need to be able to talk about the EIBV in the future (after observation $\stage + 1$ has been made).
+The definition \ref{def:eibv} of $\currentEIBV$ includes a subscript to stress that expectations are taken with respect the conditional law at stage $n$.
+We will extend this notion by writing $\currentEIBV(\cdot; u, y)$ to denote EIBV where expectations are taken conditional on the data available at $n$ an on an additional observation $y$ at $u$.
+
+  \begin{criterion}[2-steps lookahead]
+      The next observation location $\spatloc_{\stage + 1}$ is chosen among the minimizers in $\candidates$ of the criterion
+      \begin{align}\label{critLA}
+          C_{\text{2-steps}}(u) &= \mathbb{E}_{\stage}\left[\min_{\spatloc' \in
+                  \candidates(\spatloc)} \EIBV_{\stage}\left(\spatloc' ; \spatloc,
+      Y\right)\right]
+      \end{align}
+  where $Y$ is a random variable distributed according to the conditional
+  law of $\gp_{\spatloc}$ at step $\stage$ (note it silently depends on the considered point
+  $\spatloc$). Here, the dependence of the set of candidates on the current location has been made explicit for the second stage of measurements.
+  \end{criterion}
+
+\begin{remark}
+In a practical setting, the first expectation might be computed by Monte
+Carlo sampling of data $Y$ from its conditional
+distribution. For each of these data samples, the second expectation
+is solved using the closed-form expressions for EIBV provided by Proposition \ref{propo2}, now with conditioning on the first stage data already going into the mean and covariances in Eqs.
+\eqref{eq:cokrig_mean} and \eqref{eq:cokrig_cov}.
+\end{remark}
+
+\begin{remark}
+All our results have been formulated in a noiseless setting.
+That is, to us an observation $y$ at $\spatloc$ is a realization of $\gp_{\spatloc$.
+Generalization to a noisy setting where $y$ is a realization of $\gp_{\spatloc} + \epsilon$ with epsilon some noise process may be performed by including noise covariance matrices where necessary.
+\end{remark}
+
+\subsection{Simulation study}
+\label{sec:simulations}
+
+\susubsection{Static and Sequential Sampling Designs}
+\label{sec:sampling_designs}
+
+%Three different designs are considered as indicated in Figure \ref{fig:stat_design}. 
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=0.65\textwidth]{Figures/Des3.eps}
+%\caption{Three different static survey designs plotted on the initial EP.}\label{fig:stat_design}
+%\end{figure}
+%In this display the designs are plotted along with the prior probability contours of the ES for the reference parameter inputs.
+
+We compare three different static designs denoted
+\textit{static\_north}, \textit{static\_east}, and
+\textit{static\_zigzag} with the three described sequential approaches
+\textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
+sampling paths are pre-determined and cannot be altered and represent
+the pre-planned strategies used in most current AUV operational survey
+designs.
+
+For a fixed survey length, a closed-form expression for the EIBV is
+available as in Propositoin \ref{propo_eibv}
+. However, for the sequential
+approaches this is not the case. For comparison the properties are
+therefore evaluated using Monte Carlo sampling over several replicates
+of realizations from the model while conducting simulated sequential
+surveys for each one. We also compare predictive
+performance measured by root mean square error (RMSE) for temperature
+and salinity estimates as well as also the variance reduction in these
+two variables. It is important to note that the objective function
+used by the agent \kc{First of 'agent'. Define it or change it
+  to 'robot'? Note further uses below.} is focused on reducing the
+EIBV, but we nevertheless expect that we will achieve good predictive
+performance for criteria such as RMSE as well. Another non-statistical
+criteria that is important for practical purposes is the computational
+time needed for the strategy, as this will impact the performance for
+an embedded system.
+
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
+%\caption{The equilateral waypoint graph used to discretize the
+%  trajectory choices over the $31\times31$ grid used to discretize the GP.}
+%\label{fig:wp_graph}
+%\end{figure}
+
+\begin{figure}[!h] 
+\centering 
+\subfigure[The waypoint graph.]{\includegraphics[width =
+0.49\textwidth]{Figures/sim/wp_graph_paper.pdf}\label{fig:wp_graph_a}}
+\hfill
+\subfigure[The waypoint graph in 3D.]{\includegraphics[width =
+0.49\textwidth]{Figures/sim/wp_graph_3d.pdf}\label{fig:wp_graph_b}}
+\caption{\ref{fig:wp_graph_a} The equilateral waypoint graph used to discretize the
+trajectory choices over the $31\times31$ grid used to discretize the GP.
+\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment.}
+\label{fig:wp_graph}
+\end{figure}
+
+Each strategy is conducted on an equilateral grid as shown in
+Fig. \ref{fig:wp_graph}. The sequential sampling agent starts at the
+center East-West coordinate at the southern end of the domain (node 53
+\kc{Maybe color this node differently and modify the caption?}). The
+AUV moves along edges in the waypoint graph while the vehicle makes
+measurements. The data is assimilated into the GP model before an
+evaluation of the next node to sample is conducted at the end of the
+edge.
+
+\subsection{Simulation Results}
+
+A total of 100 replicate simulations were conducted with all
+strategies. The results are shown in Fig. \ref{fig:sim_results}, where
+the different criteria are plotted as a function of survey
+distance. Fig. \ref{fig:avg_ev} shows the resulting drop in IBV for
+each of the six strategies. IBV reduction occurs most under the
+\textit{myopic} and \textit{look-ahead} strategies, each performing
+almost equally; this is expected as the two criteria
+(Eq. \eqref{critSEQ} and \eqref{critLA}) are sensitive to differences
+in IBV. The \textit{static\_north} design also does well here because
+the path is parallel to the boundary between the water masses.
+
+\begin{figure}[h!]
+  \centering
+  % \subfigure[Excursion set variance $E_{\by}(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
+  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
+  \hfill
+  \subfigure[RMSE between estimated field and truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
+  \hfill 
+  \subfigure[Explained variance $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
+  \hfill 
+  \subfigure[Computational time for inferencing.\kc{I see only 
+    lines associated with 4 variables showing in the graph. Where is
+    static\_north and static\_east?}]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}} 
+\caption{Simulation results from 100 replicate simulations for 10
+  sampling choices/stages on the grid. \kc{Vertical lines show large
+    variation in replicate results.}}  
+\label{fig:sim_results}
+\end{figure}
+
+Fig. \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop
+in RMSE and increase in explained variance, respectively. Both
+\textit{myopic} and \textit{look-ahead} strategies perform well here,
+but some of the \textit{static\_east} and \textit{static\_zigzag} also
+achieve good results because they are pre-determined to cover large
+parts of the domain without re-visitation. Sequential strategies
+targeting IBV will sometimes not reach similar coverage, as
+interesting data may draw the path into twists and turns. There is
+relatively large variety in the replicate results as indicated by the
+vertical lines. Nevertheless, the ordering of strategies is similar.
+
+Fig. \ref{fig:avg_time} shows the computational effort: the
+\textit{naive} strategy is on par with the static designs, while the
+\textit{myopic} strategy is slower. The \textit{look-ahead} is even
+slower, reaching levels that are nearly impractical for execution on a
+vehicle. Some pruning of the graph is performed to improve the
+performance, such as ruling out repeated visitations and
+back-and-forth routes. Some of the intermediate results are also
+stored for longer planning horizons. Further pruning of branches or
+inclusion of other heuristics could be included for better
+performance. Then again, the inclusion of heuristics is likely a
+contributing factor for the \textit{look-ahead} strategy failing to
+outperform the \textit{myopic} strategy.
+
+In Fig. \ref{fig:route_choices}, the realized sampling paths for each
+of the sequential schemes and static designs are shown. The
+\textit{naive} strategy often gets stuck in the southern part of the
+domain because it is too focused on the probabilities near $0.5$. The
+\textit{myopic} strategy covers a wider domain than the naive or
+look-ahead. There are several reasons for this. First, a greedy
+approach will tend to put more emphasis on promising locations close
+to the agent, which may lead away from the centre. Second, as the
+agent evaluates the impact of locations further away (look-ahead)
+where assimilated data has less predictive power, the GP model (which
+is centered here) will act to restrict paths deviating from the
+central zone.
+
+We studied the sensitivity of the results by modifying the input
+parameters to have different correlations between temperature and
+salinity, standard deviations, and spatial correlation range.  In all
+runs, the \textit{myopic} and \textit{look-ahead} strategies perform
+the best in terms of realized IBV, and much better than
+\textit{naive}. The \textit{look-ahead} strategy seems to be
+substantially better than the \textit{myopic} design only for very
+small initial standard deviation or very large spatial correlation
+range. \textit{static\_north} continues to be the best static design
+for IBV, while \textit{static\_zigzag} is the best design for the
+other predictive performance measures, especially so with large
+spatial correlation range. We also ran simulation studies with only
+temperature data, and for realistic correlation levels between
+temperature and salinity, the IBV results are not much worse when only
+temperature data are available. In addition to the comparison made in
+Table \ref{tab:sim_rhoab}, the current setting includes spatial
+correlation and this likely reduces the additional influence of having
+bivariate data. However, it seems that having temperature data alone
+does a substantially worse job in terms of explained variance
 

### 2020-06-20 section3_archive_June19.tex
before running through from start.

--- 2020-06-19 section3.tex
@@
 
-This strategy does not account for the uncertainty in the GP
-variables, \textcolor{red}{(C.T.: do we need this sentence? does it mean anything?) but is instead applicable only if one design (node) has EPs
-closer to $0.5$ (see Table \ref{tab:sim_rhoab}, row
-two).}
+%This strategy does not account for the uncertainty in the GP variables, \textcolor{red}{(C.T.: do we need this sentence?
+does it mean anything?) but is instead applicable only if one design (node) has EPs closer to $0.5$ (see Table \ref{tab:sim_rhoab}, row two).}
 
-Additionally, this strategy does not account for spatial
-correlation. This strategy lacks memory of where it has been and where
-the uncertainty has been reduced, and therefore susceptible to local
-minima.
+%Additionally, this strategy does not account for spatial correlation.
+This strategy lacks memory of where it has been and where the uncertainty has been reduced, and therefore susceptible to local minima.
 
@@
 \begin{criterion}[Myopic]
+The next observation location $\spatloc_{\stage + 1}$ is chosen among the minimizers in $\candidates$ of the criterion:
 \begin{equation}\label{critSEQ}
-    \spatloc_{\stage + 1} = \argmin_{\spatloc \in \candidates} \EIBV_{\stage}\left(\spatloc\right)
+     C_{\text{myopic}}(u) = \EIBV_{\stage}\left(\spatloc\right)
 \end{equation}
@@
   \begin{criterion}[2-steps lookahead]
-      The next observation location $\spatloc_{\stage + 1}$ is selected according to
+      The next observation location $\spatloc_{\stage + 1}$ is chosen among the minimizers in $\candidates$ of the criterion
       \begin{align}\label{critLA}
-          \spatloc_{\stage + 1} = \argmin_{\spatloc \in \candidates}~ \mathbb{E}_{\stage}\left[\min_{\spatloc' \in
+          C_{\text{2-steps}}(u) &= \mathbb{E}_{\stage}\left[\min_{\spatloc' \in
                   \candidates(\spatloc)} \EIBV_{\stage}\left(\spatloc' ; \spatloc,

### 2020-06-20 main_new.tex
Modified Intro.

@@
 Monitoring the world's oceans has gained increased importance in light of the changing climate and increasing anthropogenic impact.
-
-Central to understanding the changes taking place in the upper
-water-column is knowledge of the bio-geophysical interaction driven by
+Central to understanding the changes taking place in the upper water-column is knowledge of the bio-geophysical interaction driven by
 an agglomeration of physical forcings (e.g. wind, topography,
@@
 by planktonic and coastal anthropogenic input, such as pollution and
-agricultural runoff transported into the ocean by rivers and streams.
+agricultural runoff transported into the ocean by rivers.
 These often result in a range of ecosystem-related phenomena such as
@@
 Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of data collection.
-%
- \textcolor{blue}{By \emph{sampling}, we refer here primarily to the design of observational strategies in the spatial domain %where the use autonomous robotic platforms can be combined with statistical methods
- with the aim to pursue measurements with high scientific relevance}.
-Based on collected data and sometimes also of prior knowledge, models from spatial statistics can be used to make predictions of the state variables under consideration over regions of interest.
-In case of multiple continuous state variables being jointly studied relying on spatially scattered data, vector-valued random field models and associated prediction approaches such as co-Kriging \citep[See, e.g.,][]{Wackernagel2003} can be considered nowadays as part of the standard statistical toolkit in oceanographic research and beyond.
+By \emph{sampling}, we refer here primarily to the design of observational strategies in the spatial domain with the aim to pursue measurements with high scientific relevance.
+Models from spatial statistics and methods for design of experiments can clearly contribute in this oceanographic sampling challenge.
+
 %applications and beyond.  
@@
 
-Data collection to understand natural and artificial phenomena at sea has typically been based on static buoys, floats, or ship-based methods,
+Data collection at sea has typically been based on static buoys, floats, or ship-based methods,
 with significant logistical limitations that directly impact coverage
@@
 ocean models similarly find it challenging to provide detail at fine
-scale \citep{Lermusiaux:2006}, and also come with computation costs that can be practically limiting. 
-%limiting computing resources, no time for hard-core ocean model Navier-Stokes, hence GPs and approximations, etc.
-%
+scale \citep{Lermusiaux:2006}, and also come with computation costs that can be practically limiting. The advent of robust mobile robotic platforms \citep{Bellingham07} has
+resulted in significant contributions to environmental monitoring and
+sampling in the ocean (Fig.
+\ref{fig:envir1}).
+In particular, autonomous underwater vehicles (AUVs) have advanced the state of data collection and consequently have made robotics an integral part of ocean observation \citep{das11b,Das2015,fossuminformation,fossum18b}.
+
 \begin{figure}[!h] 
@@
   \label{fig:envir} \end{figure}
-%
-\begin{comment} % Somehow redundant with the parag following below
-Mobile robotic platforms \citep{Bellingham07} have contributed significantly to environmental monitoring and sampling.
-In particular, AUVs have advanced the state of in-situ sampling, and made robotics an integral part of ocean observation, filling parts of the undersampling gap \cite{rudnick03,rudnick18} (See Fig.
-\ref{fig:envir1}).
-\end{comment}
-%
-The advent of robust mobile robotic platforms \citep{Bellingham07} has
-resulted in significant contributions to environmental monitoring and
-sampling in the ocean. In particular, autonomous underwater vehicles (AUVs) have advanced the state of data collection and consequently have made robotics an integral part of ocean observation; %our 
-previous work by some of the authors of the present paper have contributed to this effort \citep{das11b,Das2015,fossuminformation,fossum18b}.
-
-As full numerical ocean models based on complex differential equations cannot be run onboard a marine robot with limited computational capacity, statistical models relying on collected data and random field assumptions appear quite relevant as means to guide AUV data collection trajectories.
-Our main research angle in the present work is to extend sequential design strategies from the world of spatial statistics and computer experiments to frameworks featuring both vector-valued observational data and experimental designs that may be constrained by the nature of feasible trajectories.
-\\
 
@@
 evaluate, update, and refine future sampling locations (sense-plan-act
-cycle in Fig. \ref{fig:envir1}), making the information gathering
-\emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
-% This usually occurs on a spatial grid, called a waypoint graph. 
+cycle in Fig. \ref{fig:envir1}) {\bf{Jo: no longer in this Display? Bring back.}}, making the data collection
+\emph{adaptive}.
 In doing so, the space of sampling opportunities is still limited by
-the so-called waypoint graph, a discretization of the search domain, but the AUV has flexibility to modify its path at each waypoint based on in-situ sensing and measurements onboard \citep{py10,Rajan12,Rajan12b}.\\
+the so-called waypoint graph, which forms a discretization of the search domain where the AUV can navigate, but the AUV can now modify its path at each waypoint based on in-situ measurements and calculations onboard \citep{py10,Rajan12,Rajan12b}.
+Full numerical ocean models based on complex differential equations cannot be run onboard the AUV with limited computational capacity, and statistical models relying on random field assumptions are relevant as means to effectively update the onboard model from in-situ data, and to guide AUV data collection trajectories.
 
 The work presented here is primarily inspired by a case study 
-%application where an AUV is used to measure both temperature and salinity in 
-pertaining to the spatial characterization of a frontal system generated by
-river plumes. Fig. \ref{fig:nidelven} shows the survey area in
+pertaining to using an AUV for spatial characterization of a frontal system generated by
+a river plume. Fig. \ref{fig:nidelven} shows the survey area in
 Trondheim, Norway, where cold freshwater enters from the river,
@@
 of the local topography and the Coriolis force the cold fresh water
-tends to flow to the east. Depending on the river discharge, tidal
-effects, wind, and temperature differences, this boundary often gets
-distorted. Initial knowledge about the location and evolution of these
-features are highly uncertain, making deterministic planning
-challenging.
-Here the goal is to use AUV measurements towards a better description of the interface between freshwater versus oceanic water, with the help of a vector-valued random field model for both temperature and salinity.
-As we will present in the next sections, some of the original contributions inspired by this specific problem are actually applicable to further settings and application areas.
-\\
-
-The questions tackled here hence pertain to the broader area of spatial data collection and experimental design, yet with the particularity of estimating some regions of the domain, mostly excursion sets, being implicitly defined by a scarcely observed vector-valued random field.
-%because of limited resources to sample the large oceanographic domains, one must plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs.
-The acquired in-situ information must then be assimilated into the statistical model onboard the platform, so that it can be used to re-compute the sampling criterion and in this way inform decisions on where to sample next.
-\\
-Moreover, given constraints on AUV movements and the fact that surveys rely by design in such a framework on successive measurements spread along a trajectory, addressing corresponding design problems calls for particular sequential strategies, possibly also accounting for the finiteness of resources by anticipating multiple steps ahead and thereby improving over myopic strategies.
-\\
-
-Here we aim to leverage and extent recent progress in uncertainty quantification and reduction for excursion sets of Gaussian random fields in order to address vector-valued observational cases under constrained design such as in the motivating application of designing AUV trajectories to better distinguish between freshwater from ocean waters, relying on temperature and salinity measurements.
+tends to flow to the east. Depending on the variations in river discharge, tidal
+effects, coastal current and wind, this boundary often gets
+distorted, and knowledge about its location is highly uncertain, making deterministic planning
+challenging. Here the goal is to use AUV measurements for improved description of the interface between freshwater and oceanic water. 
+It is often not possible to sample the biological variables of main interest in such AUV operations, but standard instruments provide temperature and salinity measurements which serve as proxy variables for the underlying biological oceanography phenomenon.
+With the help of a vector-valued random field model for temperature and salinity, one can hence aim to describe the plume to a reasonable extent.
+The goal of plume characterization is in this way related to that of estimating some regions of the domain, typically excursion sets (ESs), being implicitly defined by the vector-valued random field.
+In our context of environmental sampling, the joint salinity and temperature
+excursions of a river plume help characterize the
+underlying bio-geochemical processes
+\citep{hopkins2013detection,Pinto2018}.
+Motivating examples for ESs of multivariate processes are also abundant in other contexts, for instance in medicine, where doctors do not rely solely on a single symptom but must see several combined effects before making a diagnosis.
+
+The questions tackled here hence pertain to the broader area of spatial data collection and experimental designs for vector-valued random field. 
+%Because of limited resources to sample the large oceanographic domains, one must plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs.
+Given the operational constraints on AUV movements and the fact that surveys rely on successive measurements along a trajectory, addressing corresponding design problems calls for particular sequential strategies.
+Our main research angle in the present work is to extend sequential design strategies from the world of spatial statistics and computer experiments to the setting of both vector-valued observational data and experimental designs for feasible robotic trajectories.
+We leverage and extend recent progress in expected uncertainty reduction for ESs of Gaussian random fields (GRFs) in order to address this research problem.
 %approach will take substantial advantage of existing work in the field of excursion set estimation mainly dedicated to computer experiments in the scalar-valued case, and where space exploration can be performed without constraints regarding the distance between successive design points.
-Original contributions compared to recent related work in the framework of scalar-valued (computer) experiments include: 
-\begin{itemize}
-    \item Extension of two important Sequential Uncertainty Reduction criteria of \citep{Bect.etal2012,chevalier2014fast,bect2019} to homotopic or heterotopic vector-valued settings,
-    \item Investigation of myopic and multiple-step ahead batch-sequential strategies for optimizing trajectories with respect to such criteria,
-    \item Replicable experiments with synthetic test cases and original data sets with accompagnying codes. 
-\end{itemize}
-Let us briefly review random field modelling and recent advances in targetd sequential design of experiments based on Gaussian Processes before detailing further existing literature tackling the considered problems, our proposed approach, as well as outlining the rest of the paper.
-
-\subsection{Random field modelling and targeted sequential design of experiments}
-
- While random field modelling has been one of the main topics throughout the history of spatial statistics \citep{Krige1951a, Matheron1963, Stein1999, Adler.Taylor2007}, there has recently been a renewed interest for random field models in the context of static sequential experimental design, be it in the context of spatial data collection \citep{Mueller2007} or in computer experiments and machine learning \citep[See for instance][and references therein]{Santner.etal2003, Romero.etal2013, Beck2016}.
-As detailed in \cite{Ginsbourger2018}, Gaussian random field models have been used in particular as a basis to sequential design of comupter experiments dedicated to various goals such as global optimization but also set estimation.
-\cite{Picheny.etal2010} revisited the classical Integrated Mean Square Error criterion with an adaptation dedicated to contour line estimation, \cite{Bect.etal2012} focused on strategies to reduce uncertainties on volumes of excursion, i.e.
-on the measure of the domain of input parameters which associated (scalar) response level exceeds a prescribed threshold, followed up by \cite{chevalier2014fast} presenting computational contributions towards an efficient batch-sequential implementation of strategies introduced in the former.
-Meanwhile, rather than focusing on excursion volumes, some approaches were investigated in \cite{French.Sain2013,Chevalier.etal2013b,Bolin.Lindgren2015,Azzimonti.etal2016} that ambition to estimate sets themselves.
-Recently, sequential design of experiments for the conservative estimation of excursion sets based on Gaussian random field models was presented in \citep{Azzimonti.etal}.
- %
-Surprisingly less attention has been dedicated to sequential strategies in the case of vector-valued observations to our knowledge.
-It has been acknowledge for several decades that co-Kriging models could be efficiently updated with vector-valued observations being assimilated sequentially
-\cite{Vargas-Guzman1999}, sequential strategies targeted to estimate prescribed features of vector-valued random fields are not in still in their infancy.
-We can refer for instance to \citep{LeGratiet.etal2015} for co-Kriging-based sequential design using fast cross-validation and dedicated to multi-fidelity computer codes.
-Related ideas are to be found in \cite{Poloczek2017} in the context of Multi-Information Source Optimization.
-As for extensions of Stepwise Uncertainty Reduction strategies, to our knowledge the only attempt so far to generalize excursion-related Stepwise Uncertainty Reduction criteria (such as reviewed above) in vector-valued settings pertains to the design of multi-fidelity simulations for fire safety, in the PhD thesis \citep[][p.82]{stroh}, where the outputs are mainly assumed independent yet the feasibility of treating the dependent case by appealing to higher-dimensional Gaussian CDF evaluations is mentioned.
+We briefly review recent advances in targeted sequential design of experiments based on GRFs before detailing further existing literature handling related AUV sampling problems, our contributions, and outlining the rest of the paper.
+
+\subsection{Random field modeling and targeted sequential design of experiments}
+
+ While random field modeling has been one of the main topics throughout the history of spatial statistics \citep{Krige1951a,Stein1999}, even for vector-valued random field models with associated prediction approaches such as co-Kriging \citep[See, e.g.,][]{Wackernagel2003}, there has lately been a renewed interest for random field models in the context of static or sequential experimental design, be it in the context of spatial data collection \citep{Mueller2007} or in computer experiments \citep{Santner.etal2003}.
+As detailed in \cite{Ginsbourger2018}, GRF models have been used in particular as a basis to sequential design of computer experiments dedicated to various goals such as global optimization and set estimation\cite{Adler.Taylor2007}.
+Of particular relevance to our context, \cite{Bect.etal2012} focused on strategies to reduce uncertainties on volumes of excursion exceeding a prescribed threshold, while \cite{chevalier2014fast} contributed with a computationally efficient implementation of sequential strategies for this challenge.
+Rather than focusing on excursion volumes, approaches were investigated in \cite{French.Sain2013,Chevalier.etal2013b,Bolin.Lindgren2015,Azzimonti.etal2016} with ambitions of estimating sets themselves.
+Recently, sequential designs of experiments for the conservative estimation of ESs based on GRF models were presented in \citep{Azzimonti.etal}.
+
+Surprisingly less attention has been dedicated to sequential strategies in the case of vector-valued observations.
+It has been long acknowledged that co-Kriging could be efficiently used for assimilating sequentially data
+\citep{Vargas-Guzman1999}, but sequential strategies for estimate features of vector-valued random fields are still in their infancy.
+\cite{LeGratiet.etal2015} used co-Kriging-based sequential designs to multi-fidelity computer codes and \cite{Poloczek2017} used related ideas for multi-information source optimization, but not for ES like we do here.
+More relevant to our setting, the PhD thesis \citep[][p.82]{stroh} mentions general possibilities of stepwise uncertainty reduction strategies for ESs in the context of designing simulations, but the outputs are mainly assumed independent.
 
@@
 For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
-\subsection{Contributions, selected previous work on AUV sampling, and outline} 
+\subsection{Relation to previous AUV sampling work, contributions and outline} 
 %work and proposed approach in a nutshell}
@@
 %\textcolor{blue}{We might want to re-use some of the following (from earlier version):}
-
-A major contribution of our work is to derive closed-form results for
-this design criteria for situations where the underlying model is
-based on multivariate GPs. Another contribution is that we embed this
-approach in a sequential strategy to produce sampling algorithms, and
-apply this to a real-world scenario of autonomously sampling
-temperature and salinity gradients using an AUV, with the intention of
-characterizing the variability of oceanic and riverine waters in
-typical mixed coastal environments.
-
-Motivating examples relevant for ESs of multivariate processes are
-abundant. In medicine, doctors do not rely solely on a single symptom
-but must see several combined effects before making a diagnosis. In
-our context of environmental sampling, the salinity and temperature
-excursions of a river plume can similarly help characterize the
-underlying bio-geochemical processes
-\citep{hopkins2013detection,Pinto2018}, where sampling can be aimed
-towards reducing uncertainty in the temperature and salinity ES.
-Unlike what has been done in previous plume exploration studies, we
-emphasize the links to spatial statistical modeling and targeted
-multivariate sampling criteria for sequential sampling.
 
@@
 statistical prediction of features using an underwater glider.
-In this paper the focus is not on statistical modeling, but rather on statistical results and computations for efficient sampling designs.
-
-We focus on spatial characterization of a frontal system generated by
-river plumes. Fig. \ref{fig:nidelven} shows the survey area in
-Trondheim, Norway, where cold freshwater enters from the river,
-creating a strong gradient in both temperature and salinity. Because
-of the local topography and the Coriolis force the cold fresh water
-tends to flow to the east. Depending on the river discharge, tidal
-effects, wind, and temperature differences, this boundary often gets
-distorted. Initial knowledge about the location and evolution of these
-features are highly uncertain, making deterministic planning
-challenging.
+In this paper the focus is not on statistical modeling, but rather on statistical properties of computations for efficient data collection.
+In this setting, we combine novel possibilities in marine robotics with spatial statistics and experimental design to provide useful AUV sampling designs.
 
@@
 the statistics literature on spatio-temporal design as well
-\cite{bueso1998state,zidek2019monitoring}, but variance and entropy
+\citep{bueso1998state,zidek2019monitoring}, but variance and entropy
 reduction are independent of the actual data realizations under the
 assumptions of GP models, so it has limited
-decisional flexibility. The use of data-driven adaptive criteria
+flexibility for active adaptation of trajectories based on what is measured. 
+The use of data-driven adaptive criteria
 was introduced to include more targeted sampling of regions of
-scientific interest in \cite{Low2009} and \cite{fossuminformation}. In
-this paper, we focus on mapping the river plume by rewarding designs
-that improve the classification of ES in temperature and salinity.
-
-\begin{comment}
-\subsection{Outline of the paper and its appendices / supplemental material}
-
+scientific interest in \cite{Low2009} and \cite{fossuminformation}.
+
+The main contributions of the current paper are
 \begin{itemize}
-    \item 
+	\item Extending expected uncertainty reduction criteria to vector-valued settings. 
+	\item Closed form expressions for the expected integrated Bernoulli variance (IBV) of the excursions in GRFs.
+	\item Algorithms for myopic and multiple-step ahead sequential strategies for optimizing AUV sampling with respect to the mentioned criteria.
+	\item Replicable experiments on synthetic test cases with accompanying code.
+	\item Results of field trials running myopic strategies onboard an AUV for the characterization of a river plume.
 \end{itemize}
-\end{comment}
 
@@
 %Section \ref{sec:bg} provides select background on ocean sampling. 
-Section \ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for
+Section \ref{sec:ESEP} defines ESs, excursion probabilities (EPs), and the design criteria connected to the IBV for excursions of 
 vector-valued GPs. Section \ref{sec:heuristics} builds on these
 assumptions when deriving the sequential design criteria for adaptive
-sampling. Section \ref{sec:simulations} discusses properties of the
-methods in simulation studies. Section \ref{sec:case_study}
+sampling. In both sections properties of the
+methods are studied using simulations. Section \ref{sec:case_study}
 demonstrates the methodology used in field work characterizing a river
-plume and finally, Section \ref{sec:concl_disc} contains a summary and
+plume. Section \ref{sec:concl_disc} contains a summary and
 a discussion of future work.
@@
 
-\section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
+\section{Quantifying uncertainty on ESs implicitly defined by GRFs}
 \label{sec:ESEP}
 
-In section \ref{sec:bg_and_notation} we introduce the notation used in the rest of the paper, review (co-)kriging of multivariate gaussian random fields (GRFs) and present update formulae for cokriging.
-Uncertainty quantification (UQ) techniques on excursion sets of GRFs are presented in \ref{sec:set_uq}, in particular the Integrated Bernoulli Variance (IBV) and the excursion measure variance (EMV) are defined and semi-analytical expressions easing their computation and optimization are derived.
+In section \ref{sec:bg_and_notation} we introduce the notation used in the rest of the paper, review (co-)kriging of multivariate GRFs and present update formulae for cokriging.
+Uncertainty quantification (UQ) techniques on excursion sets of GRFs are presented in \ref{sec:set_uq}, in particular the IBV and the excursion measure variance (EMV) are defined and semi-analytical expressions easing their computation and optimization are derived.
 Section \ref{sec:eibv} turns to the  effect of new observations on EMV and IBV.

### 2020-06-21 main_new.tex
Revised Sect 2

@@
   \centering 
-  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
+  \subfigure[Illustration of a range of ocean sensing opportunities.]{\includegraphics[width =
     0.49\textwidth]{Figures/envir.pdf}\label{fig:envir1}}
@@
     statistics and robotics can play a vital role in enabling more
-    effective oceanographic observation.\kc{The inset ``AUV robot''
-      looks very tacky and ad-hoc; might be best to remove it. Change
-      ``ships'' to
-      ``Ship'', remove the aerial drone, since we didn't use it and
-      don't plan to in this work.}}
+    effective oceanographic observation.}
   \label{fig:envir} \end{figure}
@@
 approach is to instead use onboard algorithms to continuously
-evaluate, update, and refine future sampling locations (sense-plan-act
-cycle in Fig. \ref{fig:envir1}) {\bf{Jo: no longer in this Display? Bring back.}}, making the data collection
+evaluate, update, and refine future sampling locations, making the data collection
 \emph{adaptive}.
@@
 	\item Algorithms for myopic and multiple-step ahead sequential strategies for optimizing AUV sampling with respect to the mentioned criteria.
-	\item Replicable experiments on synthetic test cases with accompanying code.
+	\item Replicable experiments on synthetic test cases with accompanying code: \url{https://github.com/CedricTravelletti/MESLAS}
 	\item Results of field trials running myopic strategies onboard an AUV for the characterization of a river plume.
@@
 
-\section{Quantifying uncertainty on ESs implicitly defined by GRFs}
+\section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
 \label{sec:ESEP}
 
-In section \ref{sec:bg_and_notation} we introduce the notation used in the rest of the paper, review (co-)kriging of multivariate GRFs and present update formulae for cokriging.
-Uncertainty quantification (UQ) techniques on excursion sets of GRFs are presented in \ref{sec:set_uq}, in particular the IBV and the excursion measure variance (EMV) are defined and semi-analytical expressions easing their computation and optimization are derived.
-Section \ref{sec:eibv} turns to the  effect of new observations on EMV and IBV.
-Specifically, the expected effect of the inclusion of a new set of observations on the EMV and IBV are considered and further semi-analytical expressions for the expected
-uncertainty reduction are derived. The resulting formulae form the backbone of the uncertainty reduction strategies presented in Section \ref{sec:heuristics}.
-Finally, Section \ref{Sec:UnivarEx} illustrates the application of the above concepts on a bivariate example relevant for sampling in the temperature and salinity application.
+Section \ref{sec:bg_and_notation} introduces notation and co-Kriging equations of multivariate GRFs.
+Section \ref{sec:set_uq} presents uncertainty quantification (UQ) techniques on ESs of GRFs, in particular the IBV and the excursion measure variance (EMV).
+Section \ref{sec:eibv} turns to the  effect of new observations on EMV and IBV, and semi-analytical expected EMV and IBV over these observations are derived. 
+%The resulting formulae form the backbone of the expected sequenatil uncertainty reduction strategies presented in Section \ref{sec:heuristics}.
+Section \ref{Sec:UnivarEx} illustrates the concepts on a bivariate example relevant for sampling in the temperature and salinity application.
 
@@
 
-Throughout the rest of the article, we denote by $\gp$ a vector-valued random field indexed by some arbitrary domain $\domain$, and assume values of the field at any fixed location $\x \in \domain$, denoted $\gp[\x]$, to be a $\no$-variate random vector ($\no\geq 2$).
-In our leading river plume characterization application, $\domain$ is a prescribed domain in the fjord of Trondheim (for the purpose of our AUV application, a discretization of a $2$-dimensional domain at fixed depth is considered), while $\no=2$ and the two responses of interest stand for temperature and salinity.
-A bivariate gaussian random field model is assumed for $\gp$, see the two first panels of Figure~\ref{example_excu} for an illustration of a realization of such a vector-valued model (RK NOTATION).
-The right panel of this figure represents a by-product of interest derived from these realizations, namely regions i) in red, where both temperature and salinity are high (i.e., exceeding respective thresholds), standing for ocean water ii) in pink, where both temperature and salinity are low, standing for riverine water, and iii) in white, where one variable is above and the other below their respective thresholds, standing for mixed waters.
+We denote by $\gp$ a vector-valued random field indexed by some arbitrary domain $\domain$, and assume values of the field at any fixed location $\x \in \domain$, denoted $\gp[\x]$, to be a $\no$-variate random vector ($\no\geq 2$).
+In the river plume characterization case, $\domain$ is a prescribed domain in the fjord of Trondheim (for the purpose of our AUV application, a discretization of a $2$-dimensional domain at fixed depth is considered), and $\no=2$ with responses of temperature and salinity.
+A bivariate GRF model is assumed for $\gp$.
+To motivate concepts, Figure~\ref{example_excu}(left and middle) shows a realization of such a vector-valued GRF on $\domain=(0,1)^2$.
+Figure~\ref{example_excu}(right) represents a by-product of interest derived from these realizations, namely regions i) in red, where both temperature and salinity are high (i.e., exceeding respective thresholds), indicative of ocean water ii) in white, where both temperature and salinity are low, indicative of riverine water, and iii) in pink, where one variable is above and the other below their respective thresholds, indicative of mixed waters.
 
@@
   \includegraphics[width=0.99\textwidth]{Figures/example_excu_1.png}
-  \caption{Realization of a bivariate GRF and excursion set above some threshold. Joint excursion in red, excursion of a single variable in pink.}
+  \caption{Realization of a bivariate GRF and excursion set above some threshold ({\bf{JE: are means around 5 and 30 here, so same as later in the example?}} ).
+Joint excursion in red, excursion of a single variable in pink.}
 \label{example_excu}
@@
 % Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
-Coming back to the more general settings of a $\no$-variate random field, 
-we are interested in recovering the set of locations $\es$ in the domain for which the response variables (the components of $\gp$) lie in some set of specified values $\T\subset \mathbb{R}^{\no}$, in other words \textit{the pre-image of $T$ by $\gp$}:
+For the general setting of a $\no$-variate random field, 
+we are interested in recovering the set of locations $\es$ in the domain for which the components of $\gp$ lie in some set of specified values $\T\subset \mathbb{R}^{\no}$, in other words \textit{the pre-image of $T$ by $\gp$}:
 $$
@@
 
-Letting $\gp[\spatloc,\ell]$ denote the $\ell\text{-th}$ component of $\gp[\spatloc]$, $\spatloc\in \domain$ and $\ell \in \{1\dots,p\}$, we will call \textit{generalized location} the couple $x=(\spatloc,\ell)$.
+Letting $\gp[\spatloc,\ell]$ denote the $\ell\text{-th}$ component of $\gp[\spatloc]$, and $\ell \in \{1\dots,p\}$, we use the term \textit{generalized location} for the couple $x=(\spatloc,\ell)$. 
 %Given such a generalized location notation $x$,  
@@
 %This slight change of notation 
-will allow us to think of $\gp$ as a scalar-valued %Gaussian 
-random field indexed by $D\times \{1\dots,p\}$, which will give the co-kriging equations a particularly simple form that parallels the ones of kriging for the scalar-valued case. 
+will allow us to think of $\gp$ as a scalar-valued Gaussian random field 
+indexed by $\domain \times \{1\dots,p\}$, {\bf{Jo: I used $\domain$ instead of $D$!}}, which will give the co-kriging equations a particularly simple form that parallels the one of univariate kriging.
 %Due to the naturality of the concept of \textit{generalized location}, we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $\spatloc \in \domain$.
-In the following, the letter $x$ will be reserved for generalized locations, while the letters $\spatloc$ and $\ell$ will be used for \textit{spatial locations} and \textit{response indices}, respectively.
-
-It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the 
-one for a single observation, provided some notation adjustments are made.
-This motivates the following.Given a dataset consisting of  $q$ observations at spatial locations $\spatloc_i \in \domain$ and response indices $\ell_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we can concatenate the information needed to specify the dataset by using the notation
+In the following, the letter $x$ will be reserved for generalized locations, while
+The letters $\spatloc$ and $\ell$ will be used for spatial locations and response indices respectively.
+
+%It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the one for a single observation, provided some notation adjustments are made.
+This motivates the following.
+
+Given a dataset consisting of  $q$ observations (a batch) at spatial locations $\spatloc_i \in \domain$ and response indices $l_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we use boldface letters to denote concatenated quantities;
 \begin{align*}
-\bm{x}:=%(\bm{\spatloc}, \bm{\ell}):= 
-(x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
+\bm{x}:=(\bm{\spatloc}, \bm{\ell}):= (x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
 \end{align*}
-In general, boldface letters will be used to denote concatenated quantities corresponding to batches of observations. 
-We can then also compactly denote the values of the field at those different locations by
+%In general,  corresponding to batches of observations. 
+The field values of batch observations are denoted by
 \begin{align*}
@@
 \end{align*}
-%
-For a second order random field $(Z_{\spatloc})_{\spatloc \in \domain}$ with mean $\mu$ and matrix covariance function $K$, $\mu$ is naturally extended to $\domain \in \lbrace 1, ..., \no\rbrace$ into a function of $x=(\spatloc, \ell)$ and is further straightforwardly vectorized into a function of $\bm{x}$.
-As for $K$, it induces a covariance kernel $k$ on the set of extended locations via  $k((\spatloc, \ell),(\spatloc', \ell'))=K(\spatloc, \spatloc')_{\ell, \ell'}$.
-In vectorized/batch form, $k(\bm{x}, \bm{x}')$ then amounts to a matrix with numbers of lines and columns the numbers of generalized locations in the batches $\bm{x}$ in $\bm{x}'$, respectively.
-Such vectorized quantities will turn out to be quite useful in order to arrive at embarrassingly simple expressions for co-Kriging equations below.
-%\textcolor{red}{(ADDENDUM, C.T.) In the same fashion, we will use $\mu(\bm{x})$ to denote the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ for the corresponding $q \times q$ covariance matrix.}
-
-Now given a Gaussian random field $\gp$ and (noisy) observations of some of its components at some points in the domain, one can predict the value of the field at some unobserved location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$, conditional on the data.
-This process is called (co-)kriging, and kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
-We will here present a rather general form of cokriging, in the sense that it allows including several observations (batch) at a time and 
-observations at a given location $s \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ 
+In the same fashion, $\mu(\bm{x})$ denotes the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ the $q \times q$ covariance matrix.
+
+Given a random field $\gp$ and (noisy) observations of some of its components at selected points in the domain, one can predict the value of the field at location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$ and the associated covariance .
+This process is called co-Kriging.% , and kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
+We present a general form of cokriging, where observations at locations $s \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ 
 (heterotopic).
 
-\subsubsection{Co-Kriging}
-\label{sec:cokriging}
-% Adressed during a discussion on 19.06.2020
-%\textcolor{red}{(C.T.) Why mention batches?
-We do not care.
-We have previously available data at $\bm{x}$ and want to predict at $\bm{x}$'.
-No need to introduce concepts we will never use.
-Moreover, this presentation suggest that to go from n to n+1 (batch-) observations, we have to condition on the whole $n$ dataset, instead of only including the last step.}
-
-Assuming that $n$ batches of observations are available, with respective sizes $q_1,\dots, q_n$, and that one wishes 
-to predict $\gp[\bm{x}]$ for some batch of $q\geq 1$ generalized locations $\bm{x}$, %\equiv(\bm{\spatloc}, \bm{\ell})$, 
-the (simple) cokriging mean would then amount to simple kriging with respect to a scalar-valued Gaussian random field indexed by 
-$D\times \{1\dots,p\}$:  
-%and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{\spatloc}, \bm{\spatloc}')_{\ell, \ell'}$, that is:
+Assuming that one wishes 
+to predict $\gp[\bm{x}]$ at $q\geq 1$ generalized locations $\bm{x}=(\bm{\spatloc}, \bm{\ell})$, 
+given $n$ batched of observations $\mathbf{z}_{[n]}$ at $\sum_{i=1}^n q_i$ generalized locations $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$.
+The co-Kriging mean then amounts to simple kriging with respect to a scalar-valued GRF indexed by $\domain \times \{1\dots,p\}$: % and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{\spatloc}, \bm{\spatloc}')_{\ell, \ell'}$, that is:
 %
@@
 \end{equation}
-where $\mu$ is $Z$'s initial mean function, $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of 
-observed responses of $Z$ at all considered generalized locations, and $\lambda_{[n]}(\bm{x})$ is a vector of weights 
-equal to $k(\bm{x}_{[n]}, \bm{x}_{[n]})^{-1} k(\bm{x}_{[n]}, \bm{x})$ with $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$, 
-$k(\bm{x}_{[n]}, \bm{x}_{[n]})$ being assumed non-singular throughout the presentation. The co-kriging %(conditional) 
-residual (cross-)covariance function (with respect to batches of generalized locations) can also be expressed in the same vein via
+%where $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of observed responses of $Z$ at all considered generalized locations, and 
+where $\lambda_{[n]}(\bm{x})$ is a vector of weights 
+equal to $\left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right]^{-1} k(\bm{x}_{[n]}, \bm{x})$ with matrix $R(\bm{x}_{[n]}, \bm{x}_{[n]})$ holding the observation noise covariance at the generalized locations.
+%$k(\bm{x}_{[n]}, \bm{x}_{[n]})$ being assumed non-singular throughout the presentation. 
+The associated co-kriging %(conditional) 
+residual cross-covariance function can be expressed in the same vein via
 %
 \begin{equation}\label{eq:cokrig_cov}
-k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T k(\bm{x}_{[n]}, \bm{x}_{[n]}) \lambda_{{[n]}}(\bm{x}').
+k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T \left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n]}}(\bm{x}').
 \end{equation}
 
-\subsubsection{Co-kriging update formulae}
-
-Let us now consider the case where co-kriging prediction of $Z$ was made with respect to $n$ batches $\bm{x}_i$ of generalized locations, concatenated again within
-$\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$, and one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
+%\subsubsection{Co-kriging update formulae}
+
+In our case it is important to update the GRF model efficiently onboard the vehicle after every batch of obervations.
+Our setting is like that of a spatial Kalman filter with static model parameters, and we provide the co-Kriging equations for this case.
+Because of our representation of co-kriging in terms of simple kriging with respect to generalized locations, the formulae is a straightforward adaptation of the batch-sequential Kriging update formulae from \cite{Chevalier.etal2013a}.
+The formulae are instrumental in deriving semi-analytical results for stepwise expected uncertainty reduction criteria for vector-valued random fields in the subsequent sections.
+
+Given the observations from $n$ batches, one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
+The mean update formulae is
 %It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
-Thanks to our representation of co-kriging in terms of simple kriging with respect to generalized locations, a strightforward adaptation of the batch-sequential kriging update formulae from \cite{Chevalier.etal2013a} delivers that
+% delivers that
 % 
-\begin{equation}
+\begin{equation}\label{eq:meanCoK}
 \mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
@@
 where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
-$\lambda_{[n+1]}(\bm{x})$ that corresponds to the kriging weigths associated with the last $q_{n+1}$ responses when 
+$\lambda_{[n+1]}(\bm{x})$ that corresponds to the Kriging weigths associated with the last $q_{n+1}$ responses when 
 predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
@@
 %, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
-Similarly, the updated co-kriging residual (cross-)covariance function then writes
-\begin{equation}
+The updated co-Kriging residual cross-covariance function is
+\begin{equation}\label{eq:varCoK}
 k_{[n+1]}(\bm{x},\bm{x}')=k_{[n]}(\bm{x},\bm{x}')-\lambda_{[n+1,n+1]}(\bm{x})^T k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]}) \lambda_{{[n+1,n+1]}}(\bm{x}').
@@
 %
-Let us remark that, as noted in \cite{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
+We remark that, as noted in \cite{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
 extend to Universal Kriging in second-order settings and apply without Gaussian assumption.
-We will now see how the latter formulae are instrumental in deriving semi-analytical formulae for stepwise uncertainty reduction criteria for
-vector-valued random fields.
-
-\subsection{Uncertainty Quantification on Excursion Sets of multivariate Gaussian Random Fields}
-\label{sec:set_uq}
-We now introduce quantities that allow to quantify the uncertainty on the volume of the excursion set $\es$. Let $\mes$ be a 
+
+\subsection{UQ on ESs of multivariate GRFs}\label{sec:set_uq}
+We now introduce quantities that allow UQ on the volume of the ES $\es$. Let $\mes$ be a 
 (locally finite, Borel) measure  on $\domain$. We want to investigate the probability distribution 
 of $\mes(\es)$ through its moments.
-
-\medskip
-
 Centred moments may be computed using Proposition~\ref{propo1} developed in the appendix. 
-In particular, it allows to write the excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ as an integral of the excursion probability
+In particular, the excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ is an integral of the EP
 \begin{equation*}
@@
 \end{equation*}
-which boils down in the excursion/sojourn case where $\T=(-\infty, t_1] \times
-\dots \times (-\infty, t_{\no}]$ to
+which in the excursion/sojourn case where $\T=(-\infty, t_1] \times
+\dots \times (-\infty, t_{\no}]$ is
 \begin{equation*}
@@
 \end{equation*}
+where $\varPhi_{\no}$ denotes the $\no$-variate Gaussian cumulative distribution function (CDF), which can be calculated using for instance the code of \cite{genz2009computation}.
+
 %
-Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $\domain^2$. In 
-contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{bect2019} involves 
-solely an integral on $\domain$ and can be expanded in our present settings as follows
+Note that this quantity requires the solution of an integral over $\domain^2$. In 
+contrast, and still like in the scalar-valued case, the IBV of \cite{bect2019} involves 
+solely an integral on $\domain$ and can be expanded as 
 \begin{equation*}
@@
 %
-Next we investigate uncertainty reduction criteria related to those indicators.
-
-\subsection{Expected Integrated Bernoulli Variance and Excursion Measure}
+
+\subsection{Expected IBV and EMV}
 \label{sec:eibv}
-This section is devoted to computing the expected effect of the inclusion of new observations on the $\emv$ and $\ibv$ of the excursion set $\es$.
-Let us consider the same setting as in Section \ref{sec:cokriging} where $n$ batches of measurements have already been performed, and let
+We compute the expected effect of the inclusion of new observations on the $\emv$ and $\ibv$ of the excursion set $\es$.
+Let us consider the same setting as in Eqn.
+\eqref{eq:meanCoK} and \eqref{eq:varCoK}, and let
 $\currentExp{.}$ and $\currentProba{.}$ denote conditional expectation
 and probability conditional on the first $n$ batches of observations, respectively.
-We want to compute the effect of the inclusion of a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and $\ibv$.
-
-For uncertainty functional such as $\ibv$ and $\emv$, we use a similar indexing system
-with respect to a sequence of batch and use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.
+We want to compute the effect of including a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and $\ibv$.
+We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.
 
@@
 $
-denote the IBV under the current law of the field, conditioned on observing $\bm{y}$ at $\bm{x}$ (generalized, possibly batch observation).
+denote the expected IBV under the current law of the field, conditioned on observing $\bm{y}$ at $\bm{x}$ (generalized, possibly batch observation).
 The expected effect of a new observation on the IBV is then captured by
 \begin{equation}\label{def:eibv}
-    \currentEIBV(\bm{x}):=\mathbb{E}_{\stage}\left[\textrb{IBV}(\bm{x}; \bm{Y})\right]
+    \currentEIBV(\bm{x}):=\mathbb{E}_{\stage}\left[\mbox{IBV}(\bm{x}; \bm{Y})\right]
 \end{equation}
-where $\bm{Y}$ is distributed according to the current law of $Z_{\bm{x}}$.
-
-\textcolor{red}{
-The philosophy is the following: EIBV and EEMV are functionals of a probability distribution.
-Then, if not mentioned explicitly, the probability distribution will be taken to be the current law of the field.
-(This notation is highly efficient).}
-
-We next present a result that allows to compute $\EIBV$ as an integral of CDFs of multivariate gaussians. This will prove useful when designing sequential uncertainty reduction strategies.
+where $\bm{Y}$ is distributed according to the current law of $Z_{\bm{x}}$ and with independent noise having covariance matrix $R(\bm{x},\bm{x})$.
+
+We next present a result that allows efficient computation of $\EIBV$ as an integral of CDFs of multivariate gaussians.
+This will prove useful when designing sequential expected uncertainty reduction strategies.
 
@@
 
-As for the expected excursion measure variance, a similar result may be derived.
+{\bf{Jo: Could we get the noise covariance $R$ in here? }}
+
+As for the expected EMV, a similar result may be derived.
 \begin{propo}
@@
 
-Note that the calculation of the multivariate Gaussian cumulative distribution
-may be done effectively using code such as that of
-\cite{genz2009computation}.
-
-\begin{remark}
+We remarks that 
 Propositions \ref{propo_eibv} and \ref{propo_emv} are twofold generalizations of  results from \cite{chevalier2014fast}: they extend previous results to the multivariate setting and also allow for the inclusion of batch or heterotopic observations through the concept of generalized locations.
@@
 random sets for vector-valued GPs.
-\end{remark}
-
-\subsection{Expected Bernoulli Variance on a two dimensional Example}
+
+\subsection{EBV for a bivariate example}
 \label{Sec:UnivarEx}
-
-We illustrate the EBV associated to different designs on a simple bivariate example.
-The example mimics our river plume application and hence the first and second component of the random field will be called \textit{temperature} and \textit{salinity} for illustrative purpose.
+We illustrate the EBV associated with different designs on a simple bivariate example that mimics our river plume application.
+The first and second component of the random field are \textit{temperature} and \textit{salinity}.
 Effects of the hyperparameters of the GRF prior on the excursion probabilities will also be studied.
-For simplicity, we begin with a \textit{pointwise} example, considering a single bivariate gaussian distribution (no spatiality).
+For simplicity, we begin with a \textit{single-location} example, considering a bivariate Gaussian distribution without any spatial elements.
 
 \subsubsection{A pointwise study}
-\textcolor{red}{(C.T.) What is EIBV in the table? How can we take an integral here? Also note that if we keep the ellipsoid plots, they should have Z's instead of Xi's}
-
-Say we want to study the excursion probability of a bivariate gaussian above some thresholds, where the thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
+For this bivariate Gaussian, the ES is defined as
+\[
+\Gamma := \lbrace \spatloc \in \domain: Z_{\spatloc, 1} \geq t_1, Z_{\spatloc, 2}\geq t_2\rbrace
+\]
+with thresholds set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity.
+We play with the temperature and salinity correlation and variances to study the effect on the EP and EBV.
 
@@
 \begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.pdf}
+  \includegraphics[width=0.99\textwidth]{Figures/illus_bivarZ.eps}
   \caption{Density contour plots with different correlations between
@@
 \ref{illus_bivarDens}. The EPs increase with the correlation as there
-is a strong tendency to have concurrently low temperature and salinity. The Bernoulli variance is similarly large for high
-correlations. EPs and Bernoulli variances are the same for standard
+is a strong tendency to have concurrently low temperature and salinity. Note that EPs and Bernoulli variances are the same for standard
 deviation $1$ or $2$, which implies that high variability in
@@
 $p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
-EIBV, Temp and Salinity data & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
-EIBV, Temperature data only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
+EBV, Temp and Salinity data & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
+EBV, Temperature data only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
 \hline
@@
 
-Table \ref{tab:sim_rhoab} (bottom two rows) shows results of expected
-Bernoulli variance calculations. This is presented for a design
+Table \ref{tab:sim_rhoab} (bottom two rows) shows results of EBV calculations. This is presented for a design
 gathering both data types, and for a design with temperature
 measurements alone. When both data are gathered, the measurement model is
-$(Y_1,Y_2)^t=(Z_1,Z_2)^t+\bepsilon$, with $\bepsilon \sim N(0,0.5^2I_2)$, while $Y_1=Z_1+\epsilon$, $\epsilon \sim N(0,0.5^2)$ when only temperature is measured.
+$(Y_{d,1},Y_{d,2})^t=(Z_1,Z_2)^t+\bepsilon$, with $\bepsilon \sim N(0,0.5^2I_2)$, while $Y_d=Z_1+\epsilon$, $\epsilon \sim N(0,0.5^2)$ when only temperature is measured.
 For this illustration, Table \ref{tab:sim_rhoab} shows that the
-expected Bernoulli variance gets lower with larger standard deviations
-$\sigma_1$ and $\sigma_2$ (right columns). The reduction of Bernoulli
+EBV gets lower with larger standard deviations
+$\sigma_1$ and $\sigma_2$ (right columns). The expected reduction of Bernoulli
 variance is largest for the cases with high correlation
@@
 ($\gamma=0.2$), there is little information about salinity in the
-temperature data, and therefore less uncertainty reduction. In an
-application with fresh cold water from a river source, the temperature
-and salinity variables will not only be interdependent, but will also
-likely show dependence in the spatial dimension. This in turn will
-impact the design criteria when we evaluate the information measure by
-integrating over several locations (Section \ref{sec:simulations}).
-
-\subsubsection{Including Spatiality}
-We now turn to an example involving a full-fledged gaussian random field (GRF). The class of GRF model we will consider in the article will generally have a linear trend
+temperature data, and therefore less uncertainty reduction.
+
+\subsubsection{Including Spatial elements}
+We extend this example to a full-fledged GRF model. 
+The class of GRF model we will consider in the article will generally have a linear trend
 \begin{align*}
 \mu(s)=\mathbb{E}\left[\begin{pmatrix}
-Z_{\spatloc, 1}\\ Z_{\spatloc, 2}
-\end{pmatrix}\right] &= \beta_0 + \beta_1 \spatloc
+Z_{s, 1}\\ Z_{s, 2}
+\end{pmatrix}\right] &= \beta_0 + \beta_1 s
 \end{align*}
-with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. In our examples, we will only consider covariance models of separable type
+with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. We will only consider covariance models of separable type
 \begin{align*}
-\textrm{Cov}\left(Z_{\spatloc, i}, Z_{v, j}\right) &= k(\spatloc, v) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_l^2,~ i=j\\
-   \gamma_0 \sigma_i \sigma_j,~i\neq j
+\textrm{Cov}\left(Z_{s, i}, Z_{u, j}\right) &= k(s, u) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_i^2,~ i=j\\
+   \gamma_0 \sigma_i \sigma_j,~i\neq j,
         \end{cases}
 \end{align*}
-where $k(., .)$ is one of the traditional spatial covariance kernel (exponential, Mat\'{e}rn, ...) and $\gamma$ defines the cross-covariance structure.
-
-In the accompanying Python code \textcolor{red}{(C.T.) Do we want to brand the code as a child of this project, or to say that we have implemented our experiments using a package that is available online (and not say too much)?}, these modeling assumptions can be
+where $k(., .)$ is specified as a Mat\'{e}rn covariance kernel in the spatial domain, i.e.
+$(1+\phi h) \exp(-\phi h)$ where $h$ is the Euclidean distance between locations $u$ and $u'$, and $\gamma$ defines the cross-covariance structure.
+In the accompanying Python code ({\bf{give URL here, or only in bullet point list in Introduction}}, these modeling assumptions can be
 generalized to anisotropic covariance and changing variance levels
-across the spatial domain. Anisotropy and
+across the spatial domain. One can hence, easily run the code to see
+if different models result in other sampling designs. Anisotropy and
 non-stationary variance are both relevant for the setting with river
@@
 \end{align*}
-and the spatial covariance is given by a Mat\'{e}rn 3/2 kernel with unit variance and range parameter $\lambda=0.5$.
+and the spatial covariance is given by a Mat\'{e}rn 3/2 kernel with unit variance and range parameter $\phi=0.5$.
+{\bf{Jo: Is this the decay parameter in formulae above, or are you trying to calculate some kind of effective spatial range?}}
 One realization of this GRF is shown in Fig. \ref{example_excu}.
 
-We now study how the expected Bernoulli variance eq.\eqref{def:eibv} associated to data collection at a point changes if only one of the two components of the field is observed.
-We first draw a realization of the GRF defined above and use it as ground-truth to mimick the real data-collection process.
-A first bunch of observations are done at the locations depicted in grey, and the data is used to update the GRF model.
-We then consider the green triangle as a potential next observation locations and plot the expected Bernoulli variance reduction (at each grid point) that would result from observing only one component of the field (temperature or salinity), or both at that point, see Fig.\ref{fig:ebv_comp}.
+We now study how the EBV associated with data collection at a point changes if only one of the two components of the field is observed.
+We first draw a realization of the GRF defined above and use it as ground-truth to mimic the real data-collection process.
+A first data collection step is performed at the location in green, and the data is used to update the GRF model.
+The EBV reduction (at each grid point) that would result from observing only one component, or both at the next grid point upwards is then computed an shown in Fig.\ref{fig:ebv_comp}.
 
@@
 
-Note that plotting the Bernoulli variance reduction at each point might also be used to compare different data collection plans.
+Note that plotting the Bernoulli variance reduction at each point might also be used to compare different pre-scripted data collection plans.
 For example, Fig.
-\ref{fig:ebv_north_vs_east} shows the expected Bernoulli variance reduction associated to a data collection plan along a straight upwards line (static north) and the one associated to a straight horizontal line (static east).
-Both expectations are computed according to the a priori distribution of the GRF (i.e.
-no observations have been included yet).
+\ref{fig:ebv_comp} shows the EBV reduction associated to a data collection plan along a straight upwards line (static north) and the one associated to a straight horizontal line (static east).
+Such calculations for pre-planned surveys are done in just a single batch.
 
@@
 
-\begin{remark}
-In all the above examples, the spatial domain $\domain$ has been discretized to a set of $n$ grid
-locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,n \}$, where each cell
-has area $\Delta$; the same grid is used for the waypoint graph for
+In all examples, the spatial domain $\domain$ is discretized to a set of $N$ grid
+locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,N \}$, where each cell
+has area $\Delta$; the same grid as used for the waypoint graph for
 possible design locations. The EIBV is approximated by sums over all
 grid cells.
-\end{remark}
 
@@
 All our results have been formulated in a noiseless setting.
-That is, to us an observation $y$ at $\spatloc$ is a realization of $\gp_{\spatloc$.
+That is, to us an observation $y$ at $\spatloc$ is a realization of $\gp_{\spatloc}$.
 Generalization to a noisy setting where $y$ is a realization of $\gp_{\spatloc} + \epsilon$ with epsilon some noise process may be performed by including noise covariance matrices where necessary.
@@
 
-\susubsection{Static and Sequential Sampling Designs}
+\subsubsection{Static and Sequential Sampling Designs}
 \label{sec:sampling_designs}

### 2020-06-21 main_new.tex
Updates from Overleaf

@@
  While random field modeling has been one of the main topics throughout the history of spatial statistics \citep{Krige1951a,Stein1999}, even for vector-valued random field models with associated prediction approaches such as co-Kriging \citep[See, e.g.,][]{Wackernagel2003}, there has lately been a renewed interest for random field models in the context of static or sequential experimental design, be it in the context of spatial data collection \citep{Mueller2007} or in computer experiments \citep{Santner.etal2003}.
-As detailed in \cite{Ginsbourger2018}, GRF models have been used in particular as a basis to sequential design of computer experiments dedicated to various goals such as global optimization and set estimation\cite{Adler.Taylor2007}.
+As detailed in \cite{Ginsbourger2018}, GRF models have been used in particular as a basis to sequential design of computer experiments dedicated to various goals such as global optimization and set estimation \citep{Adler.Taylor2007}.
 Of particular relevance to our context, \cite{Bect.etal2012} focused on strategies to reduce uncertainties on volumes of excursion exceeding a prescribed threshold, while \cite{chevalier2014fast} contributed with a computationally efficient implementation of sequential strategies for this challenge.
@@
 %where $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of observed responses of $Z$ at all considered generalized locations, and 
-where $\lambda_{[n]}(\bm{x})$ is a vector of weights 
-equal to $\left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right]^{-1} k(\bm{x}_{[n]}, \bm{x})$ with matrix $R(\bm{x}_{[n]}, \bm{x}_{[n]})$ holding the observation noise covariance at the generalized locations.
+where $\lambda_{[n]}(\bm{x})=\left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right]^{-1} k(\bm{x}_{[n]}, \bm{x})$ is a vector of weights and matrix $R(\bm{x}_{[n]}, \bm{x}_{[n]})$ holds the observation noise covariance at the generalized locations.
 %$k(\bm{x}_{[n]}, \bm{x}_{[n]})$ being assumed non-singular throughout the presentation. 
@@
 The updated co-Kriging residual cross-covariance function is
-\begin{equation}\label{eq:varCoK}
-k_{[n+1]}(\bm{x},\bm{x}')=k_{[n]}(\bm{x},\bm{x}')-\lambda_{[n+1,n+1]}(\bm{x})^T k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]}) \lambda_{{[n+1,n+1]}}(\bm{x}').
-\end{equation}
+\begin{eqnarray}\label{eq:varCoK}
+k_{[n+1]}(\bm{x},\bm{x}') &=& k_{[n]}(\bm{x},\bm{x}') -c(\bm{x},\bm{x}',\bm{x}_{[n]}), \\
+c(\bm{x},\bm{x}',\bm{x}_{[n]})&=&\lambda_{[n+1,n+1]}(\bm{x})^T \left[k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})+R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n+1,n+1]}}(\bm{x}'). \nonumber
+\end{eqnarray}
 %\medskip
@@
 
-We remarks that 
+We remark that 
 Propositions \ref{propo_eibv} and \ref{propo_emv} are twofold generalizations of  results from \cite{chevalier2014fast}: they extend previous results to the multivariate setting and also allow for the inclusion of batch or heterotopic observations through the concept of generalized locations.
+A key element for the understanding of the propositions is that the conditional co-Kriging means entering in the EPs depend linearly on (batch) observations.
+The conditional equality expressions thus become linear combinations of Gaussian variables whose mean and covariance are easily calculated.
 Related closed-form solutions have been noted in similar contexts \citep{bhattacharjya2013value,stroh}, but not generalized to our situation with
@@
 Data collection locations in green.
-True excursion set in red.
-Places where only one response is above threshold are depicted in pink}
+{\bf{Jo: left display not explained.}} Places where only one response is above threshold are depicted in pink}
 \label{fig:ebv_comp}
@@
 Data collection locations in green.
-True excursion set is in red.
-Places where only one response is above threshold are depicted in pink.}
+True ES is in red (left).
+Places where only one response is above threshold are depicted in pink.
+{\bf{Jo: is the left display wrong?}}}
 \label{fig:ebv_north_vs_east}
@@
 
-The goal of this section is to present data collection strategies that aim at decreasing the uncertainty on the target excursion set $\es$.
+The goal of this section is to present sequential data collection strategies that aim at reducing the expected uncertainty on the target ES $\es$.
 
 \subsection{Background}
-As before, let $\gp$ be a $\no$-variate random field on $\domain$.
+%As before, let $\gp$ be a $\no$-variate random field on $\domain$.
 See Section \ref{sec:bg_and_notation} for a recap of the specific notation used to handle mean and covariance function of multivariate random fields.
 
-In a sequential setting, the point of view is that $n$ data collection steps have already been performed and one wants to choose what data to collect next.
-After each step, the GRF model is updated using the cokriging equations \ref{eq:cokrig_mean} and \ref{eq:cokrig_cov} and we will denote by $\mathbb{P}_n$ the law of the field conditional on the data gathered up to (and including) stage $n$.
-More generally, we will use subscripts to denote quantities conditional on the data available up to stage $\stage$.
-Hence, $\mu_{\stage}(\cdot)$ and $k_{\stage}(\cdot, \cdot)$ will denote conditional mean and covariance functions conditional on the first $\stage$ data collection stages.
-
-\begin{remark}
-The type of data collected at each stage can be of any type (observation of all components of the field at a single location, observation of some components only but at several locations, ...) since the concept of \textit{generalized location} allows the cokriging equations to handle all those case by putting them on equal footage.
-\end{remark}
-
-Our goal is then to optimally select what observations to do next, where the meaning of optimal is yet to be defined.
-Note that here again, the concept of \textit{generalized location} allows one to consider settings where only one component of the field can be observed at each step and the strategy has to choose the spatial location as well as the component to observe (heterotopic case), or settings where several observations are allowed at each stage (batch case).
-
-We will here restrict ourselves to the case where only one spatial location can be chosen at each stage and all components of the field are observed there (isotopic case), leaving the general case for future work.
-
-\subsection{Optimal Sequential Design}
-\label{Optdes}
-
-\textcolor{red}{(C.T.) I suspect we need this paragraph so we cite people who have to be cited. How do we want to blend it with the rest? Personally I would get rid of this subsection.}
-The mathematical expression for the optimal design involves a series
+In a sequential point of view, $n$ data collection steps have already been performed and one wants to choose what data to collect next.
+After each step, the GRF model is updated using the co-Kriging equations \ref{eq:meanCoK} and \ref{eq:varCoK}.
+The design evaluations are based on the conditional expectation
+$\currentExp{.}$ from the law of the field conditional on the data gathered up to (and including) stage $n$.
+
+Note that the type of data collected at each stage can be of various type (all components of the field at a single location, only some components at a subset of the selected locations, etc.) since the concept of \textit{generalized location} allows the co-Kriging equations to handle all those case by putting them on equal footage.
+In general, a design strategy must choose the spatial location as well as the components to observe (heterotopic case), or where several observations are allowed at each stage (batch case).
+For the case with an AUV exploring the river plume, we limit scope to choosing one of the neighboring spatial location (waypoints) at each stage, and all components (temperature and salinity) of the field are observed there (isotopic case).
+The candidate points at this stage are denoted $\candidates$ as defined from the 6 directions (apart from edges) in the waypoint graph.
+The set $\candidates$ depends on the current location, but for readability we suppress this in the notation.
+
+The mathematical expression for the optimal designs in this sequential setting involves a series
 of intermixed maximizations over designs and integrals over data. In
@@
 
-\textcolor{red}{Here introduce the fact that candidate points are always chosen in a set $\candidates$.
-Mention that we usually only allow a finite set.
-Mention that in the specific case of robot planning, we have a waypoint graph and we thus usually take $\candidates$ to be the set of current neighbouring nodes.
-Mention that yes, $\candidates$ normally depends on the current location (this is for robot, in general setting, depends on the location of the last observations), but we silently suppress this dependency for readability.
-Nevertheless, one should always remember that the current set of candidates depends on the current setting.}
 \subsection{A Naive Sampling Strategy}
@@
 
-A very simple heuristic for adaptive sampling is to greedily evaluate $Z$ at the location with current excursion probability closest to $\frac{1}{2}$.While extremely easy to implement, this strategy can lead to spend too many iterations in boundary regions regardless of the the possible effect of sampling at the considered point on the future conditional distribution of $Z$.
-Working in moderate dimension and having a relatively rough discretization such as the considered waypoint graph can be beneficial by helping not being stuck for too long in poorly informative regions, yet it is not really desirable to be tributary of properties of the domain discretization in order to guarantee a fair amount of exploration.
-% REFS?
+A very simple heuristic for adaptive sampling is to observe $Z$ at the location in $\candidates$ with current EP closest to $\frac{1}{2}$.
+While extremely easy to implement, this strategy can lead to spending many stages in boundary regions regardless of the possible effect of sampling at the considered point for the future conditional distribution of $Z$.
+The strategy does not account for the expected reduction in uncertainty, and it is not considering the integrated effect over other locations.
 
@@
 The myopic (greedy) strategy which we present here is optimal if we
-imagine taking only one more stage of measurements. This selection
-strategy is based on expected uncertainty reduction, but it is a
-heuristic because there is no anticipation of what the subsequent
+imagine taking only one more stage of measurements. It does not anticipate what the subsequent
 designs might offer beyond the first stage.
-
-Based on the currently available data the myopic strategy selects the location that leads to the biggest expected reduction in IBV:
+Based on the currently available data the myopic strategy selects the location that leads to the biggest reduction in EIBV:
 \begin{criterion}[Myopic]
@@
 
-The expected IBV might be efficiently computed for each candidate points using Proposition \ref{propo2}.
-Once the best location has been selected, a stage of observation is performed there, the GP model is updated, yielding a conditional law $\mathbb{P}_{\stage + 1}$ and the process is repeated.
+The EIBV is efficiently computed for each of the candidate points $\candidates$ using Proposition \ref{propo2}.
+Once the best location has been selected, a stage of observation is performed there, the GRF model is updated, yielding a conditional law $\mathbb{P}_{\stage + 1}$ and the process is repeated.
 
@@
 reasonable approach for creating designs in many
-applications.
-Moreover, it can be implemented in a way that is not too demanding on computational power, making it well-suited for embedding on autonomous mobile survey platform (see Section \ref{sec:case_study} for an example).
+applications. Moreover, it can be implemented in a way that is not too demanding on computational power, making it well-suited for embedding on the AUV.
 
@@
 
-\medskip
-
-Say we have collected data up to stage $n$. The principle of two-steps lookahead
-strategies is to select the next observation location $\spatloc_{\stage + 1}$ that would yield the biggest exepcted decrease in IBV if we were to (optimally) add one more observation after the one at $\spatloc_{\stage + 1}$.
-In order to formalize this concept, we need to be able to talk about the EIBV in the future (after observation $\stage + 1$ has been made).
-The definition \ref{def:eibv} of $\currentEIBV$ includes a subscript to stress that expectations are taken with respect the conditional law at stage $n$.
-We will extend this notion by writing $\currentEIBV(\cdot; u, y)$ to denote EIBV where expectations are taken conditional on the data available at $n$ an on an additional observation $y$ at $u$.
+Assume data have been collected up to stage $n$. The principle of two-steps lookahead
+strategies is to select the next observation location $\spatloc_{\stage + 1}$ that yields the biggest reduction in EIBV if we were to (optimally) add one more observation after the one at $\spatloc_{\stage + 1}$.
+In order to formalize this concept, we must extend the notation for EIBV in the future (after observation $\stage + 1$ has been made).
+We let $\currentEIBV(\cdot; u, y)$ to denote the EIBV where expectations are taken conditional on the data available at $n$ and on an additional observation $y$ at $u$.
 
@@
   where $Y$ is a random variable distributed according to the conditional
-  law of $\gp_{\spatloc}$ at step $\stage$ (note it silently depends on the considered point
-  $\spatloc$). Here, the dependence of the set of candidates on the current location has been made explicit for the second stage of measurements.
+  law of $\gp_{\spatloc}$ at step $\stage$. The dependence of the set of candidates on the current location has been made explicit for the second stage of measurements.
   \end{criterion}
 
-\begin{remark}
-In a practical setting, the first expectation might be computed by Monte
+In a practical setting, the first expectation can be computed by Monte
 Carlo sampling of data $Y$ from its conditional
 distribution. For each of these data samples, the second expectation
-is solved using the closed-form expressions for EIBV provided by Proposition \ref{propo2}, now with conditioning on the first stage data already going into the mean and covariances in Eqs.
-\eqref{eq:cokrig_mean} and \eqref{eq:cokrig_cov}.
-\end{remark}
-
-\begin{remark}
-All our results have been formulated in a noiseless setting.
-That is, to us an observation $y$ at $\spatloc$ is a realization of $\gp_{\spatloc}$.
-Generalization to a noisy setting where $y$ is a realization of $\gp_{\spatloc} + \epsilon$ with epsilon some noise process may be performed by including noise covariance matrices where necessary.
-\end{remark}
+is solved using the closed-form expressions for EIBV provided by Proposition \ref{propo2}, now with conditioning on the first stage data already going into the co-Kriging updating equations.
 
@@
 \textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
-sampling paths are pre-determined and cannot be altered and represent
-the pre-planned strategies used in most current AUV operational survey
-designs.
-
+AUV sampling paths are pre-scripted and cannot be altered.
 For a fixed survey length, a closed-form expression for the EIBV is
-available as in Propositoin \ref{propo_eibv}
-. However, for the sequential
+available as in Proposition \ref{propo_eibv}. However, for the sequential
 approaches this is not the case. For comparison the properties are
@@
 performance measured by root mean square error (RMSE) for temperature
-and salinity estimates as well as also the variance reduction in these
+and salinity estimates as well as the variance reduction in these
 two variables. It is important to note that the objective function
-used by the agent \kc{First of 'agent'. Define it or change it
-  to 'robot'? Note further uses below.} is focused on reducing the
+used by the robot. is focused on reducing the
 EIBV, but we nevertheless expect that we will achieve good predictive
 performance for criteria such as RMSE as well. Another non-statistical
-criteria that is important for practical purposes is the computational
+criterion that is important for practical purposes is the computational
 time needed for the strategy, as this will impact the performance for
@@
 \caption{\ref{fig:wp_graph_a} The equilateral waypoint graph used to discretize the
-trajectory choices over the $31\times31$ grid used to discretize the GP.
+trajectory choices over the $31\times31$ grid used to discretize the GRF. The AUV is set to start in node $53$.
 \ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment.}
@@
 Each strategy is conducted on an equilateral grid as shown in
-Fig. \ref{fig:wp_graph}. The sequential sampling agent starts at the
-center East-West coordinate at the southern end of the domain (node 53
-\kc{Maybe color this node differently and modify the caption?}). The
-AUV moves along edges in the waypoint graph while the vehicle makes
-measurements. The data is assimilated into the GP model before an
-evaluation of the next node to sample is conducted at the end of the
-edge.
+Fig. \ref{fig:wp_graph}. The AUV starts at the
+center East-West coordinate at the southern end of the domain (node 53). The
+AUV moves along edges in the waypoint graph while collecting data which are assimilated into the GRF model onboard. This is used in the
+evaluation of the next node to sample.
+The procedure is run for $10$ stages. A total of 100 replicate simulations were conducted with all
+strategies.
 
@@
 
-A total of 100 replicate simulations were conducted with all
-strategies. The results are shown in Fig. \ref{fig:sim_results}, where
+The results of the replicate runs are shown in Fig. \ref{fig:sim_results}, where
 the different criteria are plotted as a function of survey
-distance. Fig. \ref{fig:avg_ev} shows the resulting drop in IBV for
+distance. Fig. \ref{fig:avg_ev} shows the resulting drop in realized IBV for
 each of the six strategies. IBV reduction occurs most under the
@@
 but some of the \textit{static\_east} and \textit{static\_zigzag} also
-achieve good results because they are pre-determined to cover large
+achieve good results because they are cover large
 parts of the domain without re-visitation. Sequential strategies
 targeting IBV will sometimes not reach similar coverage, as
-interesting data may draw the path into twists and turns. There is
+interesting data may draw the AUV path into twists and turns. There is
 relatively large variety in the replicate results as indicated by the
@@
 \textit{naive} strategy is on par with the static designs, while the
-\textit{myopic} strategy is slower. The \textit{look-ahead} is even
+\textit{myopic} strategy is slower because it evaluates expected values for all candidate directions at the waypoints.
+But it is still able to do so in reasonable time, which enables real-world applicability.
+The \textit{look-ahead} is even
 slower, reaching levels that are nearly impractical for execution on a
 vehicle. Some pruning of the graph is performed to improve the
-performance, such as ruling out repeated visitations and
-back-and-forth routes. Some of the intermediate results are also
-stored for longer planning horizons. Further pruning of branches or
+performance, such as ruling out repeated visitations. Further pruning of branches or
 inclusion of other heuristics could be included for better
-performance. Then again, the inclusion of heuristics is likely a
+performance. Then again, the inclusion of such heuristics is likely a
 contributing factor for the \textit{look-ahead} strategy failing to
@@
 
-In Fig. \ref{fig:route_choices}, the realized sampling paths for each
-of the sequential schemes and static designs are shown. The
-\textit{naive} strategy often gets stuck in the southern part of the
-domain because it is too focused on the probabilities near $0.5$. The
-\textit{myopic} strategy covers a wider domain than the naive or
-look-ahead. There are several reasons for this. First, a greedy
-approach will tend to put more emphasis on promising locations close
-to the agent, which may lead away from the centre. Second, as the
-agent evaluates the impact of locations further away (look-ahead)
-where assimilated data has less predictive power, the GP model (which
-is centered here) will act to restrict paths deviating from the
-central zone.
+%In Fig. \ref{fig:route_choices}, the realized sampling paths for each of the sequential schemes and static designs are shown. The
+%\textit{naive} strategy often gets stuck in the southern part of the
+%domain because it is too focused on the probabilities near $0.5$. The
+%\textit{myopic} strategy covers a wider domain than the naive or
+%look-ahead. There are several reasons for this. First, a greedy
+%approach will tend to put more emphasis on promising locations close
+%to the agent, which may lead away from the centre. Second, as the
+%agent evaluates the impact of locations further away (look-ahead)
+%where assimilated data has less predictive power, the GP model (which
+%is centered here) will act to restrict paths deviating from the
+%central zone.
 
@@
 small initial standard deviation or very large spatial correlation
-range. \textit{static\_north} continues to be the best static design
-for IBV, while \textit{static\_zigzag} is the best design for the
-other predictive performance measures, especially so with large
-spatial correlation range. We also ran simulation studies with only
+range. 
+%\textit{static\_north} continues to be the best static design
+%for IBV, while \textit{static\_zigzag} is the best design for the
+%other predictive performance measures, especially so with large
+%spatial correlation range. 
+We also ran simulation studies with only
 temperature data, and for realistic correlation levels between
@@
 Table \ref{tab:sim_rhoab}, the current setting includes spatial
-correlation and this likely reduces the additional influence of having
+correlation and this likely reduces the added effect of having
 bivariate data. However, it seems that having temperature data alone
-does a substantially worse job in terms of explained variance
+does a substantially worse job in terms of explained variance.
 
@@
 information is likely to result in improved decision-making.
+We also foresee opportunities related to more nuanced design of expfor multivariate processes using the notion of generalized locations.
 

### 2020-06-21 main_new.tex
Sect 3 revised in main_new.tex.

@@
 information is likely to result in improved decision-making.
-We also foresee opportunities related to more nuanced design of expfor multivariate processes using the notion of generalized locations.
+We also foresee opportunities related to design of experiments for multivariate processes using our notion of generalized locations.
+For instance, this is likely to be useful for internet-of-things applications or computer experiments context, where some observations or evaluations are rather inexpensive, while others are much more costly and must only be done when they are really valuable.
 

### 2020-06-22 main_new.tex
updates

@@
   Improving and optimizing oceanographic sampling is a crucial task
-  for marine science and maritime management. Faced with limited
-  resources to understand processes in the water-column, the
-  combination of statistics and autonomous robotics provides new
+  for marine science and maritime resource management. Faced with
+  limited resources in understanding processes in the water-column,
+  the combination of statistics and autonomous systems provide new
   opportunities for experimental design. In this work we develop
@@
   spatial characteristics of ocean phenomena. Specifically, we define
-  a design
-  criterion based on improved characterization of the uncertainty in
-  the excursions of vector-valued Gaussian random fields, and derive
-  tractable expressions for the expected Bernoulli variance reduction
-  in such a framework. We demonstrate how this criterion can be used
-  to prioritize sampling efforts at locations that are ambiguous,
-  making exploration more effective. We use simulations to study the
-  properties of methods and to compare them with state-of-the-art
-  approaches, followed by results from field deployments with an
-  autonomous underwater vehicle (AUV) as part of a study mapping the
-  boundary of a river plume. The results demonstrate the potential of
-  combining statistical methods and robotic platforms to effectively
-  inform and execute data-driven environmental sampling.
+  a design criterion based on improved characterization of the
+  uncertainty in the excursions of vector-valued Gaussian random
+  fields, and derive tractable expressions for the expected Bernoulli
+  variance reduction in such a framework. We demonstrate how this
+  criterion can be used to prioritize sampling efforts at locations
+  that are ambiguous, making exploration more effective. We use
+  simulations to study the properties of methods and to compare them
+  with state-of-the-art approaches, followed by results from field
+  deployments with an autonomous underwater vehicle (AUV) as part of a
+  study mapping the boundary of a river plume. The results demonstrate
+  the potential of combining statistical methods and robotic platforms
+  to effectively inform and execute data-driven environmental
+  sampling.
 
@@
 
-Motivated by the challenges related to efficient data collection strategies for the vast oceans, the combination of spatial statistics, design of experiments and marine robotics is studied in this paper.
-The multidisciplinary efforts enable information-driven data collection in regions of high-interest regions.
+Motivated by the challenges related to efficient data collection
+strategies for our vast oceans, we combine spatial statistics, design
+of experiments and marine robotics in this work.  The
+multidisciplinary efforts enable information-driven data collection in
+regions of high-interest.
 
@@
 
-Monitoring the world's oceans has gained increased importance in light of the changing climate and increasing anthropogenic impact.
-Central to understanding the changes taking place in the upper water-column is knowledge of the bio-geophysical interaction driven by
-an agglomeration of physical forcings (e.g. wind, topography,
-bathymetry, tidal influences, etc.) and incipient micro-biology driven
-by planktonic and coastal anthropogenic input, such as pollution and
-agricultural runoff transported into the ocean by rivers.
-These often result in a range of ecosystem-related phenomena such as
-blooms and plumes, with direct and indirect effects on society. \kc{we
-  should look for a citation here. Check with John Ryan}
- One of the bottlenecks in the study of such phenomena lies however in the lack of observational data with sufficient resolution.
-Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of data collection.
-By \emph{sampling}, we refer here primarily to the design of observational strategies in the spatial domain with the aim to pursue measurements with high scientific relevance.
-Models from spatial statistics and methods for design of experiments can clearly contribute in this oceanographic sampling challenge.
+Monitoring the world's oceans has gained increased importance in light
+of the changing climate and increasing anthropogenic impact. Central
+to understanding the changes taking place in the upper water-column is
+knowledge of the bio-geophysical interaction driven by an
+agglomeration of physical forcings (e.g. wind, topography, bathymetry,
+tidal influences, etc.) and incipient micro-biology driven by
+planktonic and coastal anthropogenic input, such as pollution and
+agricultural runoff transported into the ocean by rivers.  These often
+result in a range of ecosystem-related phenomena such as blooms and
+plumes, with direct and indirect effects on society. \kc{we should
+  look for a citation here. Check with John Ryan} One of the
+bottlenecks in the study of such phenomena lies however in the lack of
+observational data with sufficient resolution. Most of this
+\emph{undersampling} can be attributed to the large spatio-temporal
+variations in which ocean processes transpire, prompting the need for
+effective means of data collection.  By \emph{sampling}, we refer here
+primarily to the design of observational strategies in the spatial
+domain with the aim to pursue measurements with high scientific
+relevance.  Spatial statistics and methods for experimental design can
+clearly contribute to this sampling challenge.
 
@@
 
-Data collection at sea has typically been based on static buoys, floats, or ship-based methods,
-with significant logistical limitations that directly impact coverage
-and sampling resolution. Modern methods using satellite remote-sensing
-provide large-scale coverage but have limited resolution, are limited
-to sensing the surface, and are impacted by cloud cover. Numerical
-ocean models similarly find it challenging to provide detail at fine
-scale \citep{Lermusiaux:2006}, and also come with computation costs that can be practically limiting. The advent of robust mobile robotic platforms \citep{Bellingham07} has
+Data collection at sea has typically been based on static buoys,
+floats, or ship-based methods, with significant logistical limitations
+that directly impact coverage and sampling resolution. Modern methods
+using satellite remote-sensing provide large-scale coverage but have
+limited resolution, are limited to sensing the surface, and are
+impacted by cloud cover. Numerical ocean models similarly find it
+challenging to provide detail at fine scale \citep{Lermusiaux:2006},
+and also come with computational costs that can be limiting. The
+advent of robust mobile robotic platforms \citep{Bellingham07} has
 resulted in significant contributions to environmental monitoring and
-sampling in the ocean (Fig.
-\ref{fig:envir1}).
-In particular, autonomous underwater vehicles (AUVs) have advanced the state of data collection and consequently have made robotics an integral part of ocean observation \citep{das11b,Das2015,fossuminformation,fossum18b}.
+sampling in the ocean (Fig. \ref{fig:envir1}). In particular,
+autonomous underwater vehicles (AUVs) have advanced the state of data
+collection and consequently have made robotics an integral part of
+ocean observation \citep{das11b,Das2015,fossuminformation,fossum18b}.
 
@@
 approach is to instead use onboard algorithms to continuously
-evaluate, update, and refine future sampling locations, making the data collection
-\emph{adaptive}.
-In doing so, the space of sampling opportunities is still limited by
-the so-called waypoint graph, which forms a discretization of the search domain where the AUV can navigate, but the AUV can now modify its path at each waypoint based on in-situ measurements and calculations onboard \citep{py10,Rajan12,Rajan12b}.
-Full numerical ocean models based on complex differential equations cannot be run onboard the AUV with limited computational capacity, and statistical models relying on random field assumptions are relevant as means to effectively update the onboard model from in-situ data, and to guide AUV data collection trajectories.
+evaluate, update, and refine future sampling locations, making the
+data collection \emph{adaptive}.  In doing so, the space of sampling
+opportunities is still limited by a waypoint graph, which forms a
+discretization of the search domain where the AUV can navigate;
+however the AUV can now modify its path at each waypoint based on
+in-situ measurements and calculations onboard using onboard
+deliberation \citep{py10,Rajan12,Rajan12b}.  Full numerical ocean
+models based on complex differential equations cannot be run onboard
+the AUV with limited computational capacity, and statistical models
+relying on random field assumptions are relevant as a means to
+effectively update the onboard model from in-situ data, and to guide
+AUV data collection trajectories.
 
 The work presented here is primarily inspired by a case study
-pertaining to using an AUV for spatial characterization of a frontal system generated by
-a river plume. Fig. \ref{fig:nidelven} shows the survey area in
-Trondheim, Norway, where cold freshwater enters from the river,
-creating a strong gradient in both temperature and salinity. Because
-of the local topography and the Coriolis force the cold fresh water
-tends to flow to the east. Depending on the variations in river discharge, tidal
-effects, coastal current and wind, this boundary often gets
-distorted, and knowledge about its location is highly uncertain, making deterministic planning
-challenging. Here the goal is to use AUV measurements for improved description of the interface between freshwater and oceanic water. 
-It is often not possible to sample the biological variables of main interest in such AUV operations, but standard instruments provide temperature and salinity measurements which serve as proxy variables for the underlying biological oceanography phenomenon.
-With the help of a vector-valued random field model for temperature and salinity, one can hence aim to describe the plume to a reasonable extent.
-The goal of plume characterization is in this way related to that of estimating some regions of the domain, typically excursion sets (ESs), being implicitly defined by the vector-valued random field.
-In our context of environmental sampling, the joint salinity and temperature
-excursions of a river plume help characterize the
+pertaining to using an AUV for spatial characterization of a frontal
+system generated by a river plume. Fig. \ref{fig:nidelven} shows the
+survey area in Trondheim, Norway, where cold freshwater enters from
+the river, creating a strong gradient in both temperature and
+salinity. Because of the local topography and the Coriolis force the
+cold fresh water tends to flow east. Depending on the variations in
+river discharge, tidal effects, coastal current and wind, this
+boundary often gets distorted, and knowledge about its location is
+highly uncertain, making deterministic planning challenging. The goal
+is therefore, to use AUV measurements for improved description of the
+interface between fresh and oceanic waters.  It is often not possible
+to sample the biological variables of fundamental interest in such AUV
+operations, but off-the-shelf instruments provide temperature and
+salinity measurements which serve as proxys for the underlying
+biological phenomenon. With the help of a vector-valued random field
+model for temperature and salinity, one can then aim to describe the
+plume.  The goal of plume characterization, in this way, relates to
+that of estimating some regions of the domain, typically excursion
+sets (ESs), when implicitly defined by the vector-valued random field.
+In our context of environmental sampling, the joint salinity and
+temperature excursions of a river plume help characterize the
 underlying bio-geochemical processes
-\citep{hopkins2013detection,Pinto2018}.
-Motivating examples for ESs of multivariate processes are also abundant in other contexts, for instance in medicine, where doctors do not rely solely on a single symptom but must see several combined effects before making a diagnosis.
-
-The questions tackled here hence pertain to the broader area of spatial data collection and experimental designs for vector-valued random field. 
+\citep{hopkins2013detection,Pinto2018}. Motivating examples for ESs of
+multivariate processes are also abundant in other contexts, for
+instance in medicine, where doctors do not rely solely on a single
+symptom but must see several combined effects before making a
+diagnosis.
+
+The questions tackled here hence pertain to the broader area of
+spatial data collection and experimental designs for vector-valued
+random fields.
 %Because of limited resources to sample the large oceanographic domains, one must plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs.
-Given the operational constraints on AUV movements and the fact that surveys rely on successive measurements along a trajectory, addressing corresponding design problems calls for particular sequential strategies.
-Our main research angle in the present work is to extend sequential design strategies from the world of spatial statistics and computer experiments to the setting of both vector-valued observational data and experimental designs for feasible robotic trajectories.
-We leverage and extend recent progress in expected uncertainty reduction for ESs of Gaussian random fields (GRFs) in order to address this research problem.
+Given the operational constraints on AUV movements and the fact that
+surveys rely on successive measurements along a trajectory, addressing
+corresponding design problems calls for sequential strategies.  Our
+main research angle in the present work is to extend sequential design
+strategies from the world of spatial statistics and computer
+experiments to the setting of both vector-valued observational data
+and experimental designs for feasible robotic trajectories. We
+leverage and extend recent progress in expected uncertainty reduction
+for ESs of Gaussian random fields (GRFs) in order to address this
+research
+problem.
 %approach will take substantial advantage of existing work in the field of excursion set estimation mainly dedicated to computer experiments in the scalar-valued case, and where space exploration can be performed without constraints regarding the distance between successive design points.
-We briefly review recent advances in targeted sequential design of experiments based on GRFs before detailing further existing literature handling related AUV sampling problems, our contributions, and outlining the rest of the paper.
+We briefly review recent advances in targeted sequential design of
+experiments based on GRFs before detailing other literature related to
+AUV sampling and our contributions prior to outlining the rest of the
+paper.
 
@@
 
- While random field modeling has been one of the main topics throughout the history of spatial statistics \citep{Krige1951a,Stein1999}, even for vector-valued random field models with associated prediction approaches such as co-Kriging \citep[See, e.g.,][]{Wackernagel2003}, there has lately been a renewed interest for random field models in the context of static or sequential experimental design, be it in the context of spatial data collection \citep{Mueller2007} or in computer experiments \citep{Santner.etal2003}.
-As detailed in \cite{Ginsbourger2018}, GRF models have been used in particular as a basis to sequential design of computer experiments dedicated to various goals such as global optimization and set estimation \citep{Adler.Taylor2007}.
-Of particular relevance to our context, \cite{Bect.etal2012} focused on strategies to reduce uncertainties on volumes of excursion exceeding a prescribed threshold, while \cite{chevalier2014fast} contributed with a computationally efficient implementation of sequential strategies for this challenge.
-Rather than focusing on excursion volumes, approaches were investigated in \cite{French.Sain2013,Chevalier.etal2013b,Bolin.Lindgren2015,Azzimonti.etal2016} with ambitions of estimating sets themselves.
-Recently, sequential designs of experiments for the conservative estimation of ESs based on GRF models were presented in \citep{Azzimonti.etal}.
-
-Surprisingly less attention has been dedicated to sequential strategies in the case of vector-valued observations.
-It has been long acknowledged that co-Kriging could be efficiently used for assimilating sequentially data
-\citep{Vargas-Guzman1999}, but sequential strategies for estimate features of vector-valued random fields are still in their infancy.
-\cite{LeGratiet.etal2015} used co-Kriging-based sequential designs to multi-fidelity computer codes and \cite{Poloczek2017} used related ideas for multi-information source optimization, but not for ES like we do here.
-More relevant to our setting, the PhD thesis \citep[][p.82]{stroh} mentions general possibilities of stepwise uncertainty reduction strategies for ESs in the context of designing simulations, but the outputs are mainly assumed independent.
+While random field modeling has been one of the main topics throughout
+the history of spatial statistics \citep{Krige1951a,Stein1999}, even
+for vector-valued random field models with associated prediction
+approaches such as co-Kriging \citep[See, e.g.,][]{Wackernagel2003},
+there has lately been a renewed interest for random field models in
+the context of static or sequential experimental design, be it in the
+context of spatial data collection \citep{Mueller2007} or in
+simulation experiments \citep{Santner.etal2003}. As detailed in
+\cite{Ginsbourger2018}, GRF models have been used in particular as a
+basis to sequential design of simulations dedicated to various goals
+such as global optimization and set estimation
+\citep{Adler.Taylor2007}. Of particular relevance to our context,
+\cite{Bect.etal2012} focuses on strategies to reduce uncertainty in
+volumes of excursion exceeding a prescribed threshold, while
+\cite{chevalier2014fast} contributes a computationally efficient
+implementation of sequential strategies. Rather than focusing on
+excursion volumes, approaches were investigated in
+\cite{French.Sain2013,Chevalier.etal2013b,Bolin.Lindgren2015,Azzimonti.etal2016}
+with ambitions of estimating sets themselves. Recently, sequential
+designs of experiments for the conservative estimation of ESs based on
+GRF models were presented in \citep{Azzimonti.etal}.
+
+Surprisingly less attention has been dedicated to sequential
+strategies in the case of vector-valued observations. It has been long
+acknowledged that co-Kriging could be used efficiently for
+assimilating sequential data \citep{Vargas-Guzman1999}, but sequential
+strategies for estimating features of vector-valued random fields are
+still in their infancy. \cite{LeGratiet.etal2015} used co-Kriging
+based sequential designs to multi-fidelity computer codes and
+\cite{Poloczek2017} used related ideas for multi-information source
+optimization, but not for ES's like we do here. More relevant to our
+setting, the PhD thesis \citep[][p.82]{stroh} mentions general
+possibilities of stepwise uncertainty reduction strategies for ES's in
+the context of designing simulations, but the outputs are mainly
+assumed independent.
 
@@
 For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
-\subsection{Relation to previous AUV sampling work, contributions and outline} 
-%work and proposed approach in a nutshell}
-%\begin{itemize}
-%    \item 
-%\end{itemize}
-%\textcolor{blue}{We might want to re-use some of the following (from earlier version):}
-
-Other statistical
-work in the oceanographic domain include \cite{wikle2013modern}
-focusing on hierarchical statistical models; \cite{sahu2008space},
-studying spatio-temporal models for sea surface temperature and
-salinity data; and \cite{mellucci2018oceanic} looking at the
-statistical prediction of features using an underwater glider.
-In this paper the focus is not on statistical modeling, but rather on statistical properties of computations for efficient data collection.
-In this setting, we combine novel possibilities in marine robotics with spatial statistics and experimental design to provide useful AUV sampling designs.
+\subsection{Previous work in AUV sampling}
+
+Other statistical work in the oceanographic domain include
+\cite{wikle2013modern} focusing on hierarchical statistical models,
+\cite{sahu2008space} studying spatio-temporal models for sea surface
+temperature and salinity data and \cite{mellucci2018oceanic} looking
+at the statistical prediction of features using an underwater glider.
+In this paper the focus is not on statistical modeling, but rather on
+statistical properties of computation for efficient data
+collection. In this setting, we combine novel possibilities in marine
+robotics with spatial statistics and experimental design to provide
+useful AUV sampling designs.
 
 Adaptive in-situ AUV sampling of an evolving frontal feature has been
-explored in \cite{fronts11,Zhang2012,Pinto2018,costa19}. These
+explored in \cite{fronts11,Smith2016,Pinto2018,costa19}. These
 approaches typically use a reactive-adaptive scheme, whereby
@@
 the statistics literature on spatio-temporal design as well
-\citep{bueso1998state,zidek2019monitoring}, but variance and entropy
-reduction are independent of the actual data realizations under the
-assumptions of GP models, so it has limited
-flexibility for active adaptation of trajectories based on what is measured. 
-The use of data-driven adaptive criteria
-was introduced to include more targeted sampling of regions of
-scientific interest in \cite{Low2009} and \cite{fossuminformation}.
-
-The main contributions of the current paper are
+\citep{bueso1998state,zidek2019monitoring}. However, variance and
+entropy reduction are independent of the actual data realizations
+under the assumptions of GP models, so it has limited flexibility for
+active adaptation of trajectories based on what is measured.  The use
+of data-driven adaptive criteria was introduced to include more
+targeted sampling of regions of scientific interest in \cite{Low2009}
+and \cite{fossuminformation}.
+
+The primary contributions of this work are:
+
 \begin{itemize}
-	\item Extending expected uncertainty reduction criteria to vector-valued settings. 
-	\item Closed form expressions for the expected integrated Bernoulli variance (IBV) of the excursions in GRFs.
-	\item Algorithms for myopic and multiple-step ahead sequential strategies for optimizing AUV sampling with respect to the mentioned criteria.
-	\item Replicable experiments on synthetic test cases with accompanying code: \url{https://github.com/CedricTravelletti/MESLAS}
-	\item Results of field trials running myopic strategies onboard an AUV for the characterization of a river plume.
+\item Extending expected uncertainty reduction criteria to
+  vector-valued settings.  
+\item Closed form expressions for the expected integrated Bernoulli
+  variance (IBV) of the excursions in GRFs. 
+\item Algorithms for myopic and multiple-step ahead sequential
+  strategies for optimizing AUV sampling with respect to the mentioned
+  criteria. 
+\item Replicable experiments on synthetic test cases with accompanying
+  code~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}}
+\item Results of field trials running myopic strategies onboard an AUV
+  for the characterization of a river plume. 
 \end{itemize}
@@
 %Section \ref{sec:bg} provides select background on ocean sampling. 
-Section \ref{sec:ESEP} defines ESs, excursion probabilities (EPs), and the design criteria connected to the IBV for excursions of 
+Section \ref{sec:ESEP} defines ESs, excursion probabilities (EPs), and
+the design criteria connected to the IBV for excursions of
 vector-valued GPs. Section \ref{sec:heuristics} builds on these
 assumptions when deriving the sequential design criteria for adaptive
-sampling. In both sections properties of the
-methods are studied using simulations. Section \ref{sec:case_study}
-demonstrates the methodology used in field work characterizing a river
-plume. Section \ref{sec:concl_disc} contains a summary and
-a discussion of future work.
+sampling. In both sections properties of the methods are studied using
+simulations. Section \ref{sec:case_study} demonstrates the methodology
+used in field work characterizing a river plume. Section
+\ref{sec:concl_disc} contains a summary and a discussion of future
+work.
 %\input{section2}
@@
 
-Section \ref{sec:bg_and_notation} introduces notation and co-Kriging equations of multivariate GRFs.
-Section \ref{sec:set_uq} presents uncertainty quantification (UQ) techniques on ESs of GRFs, in particular the IBV and the excursion measure variance (EMV).
-Section \ref{sec:eibv} turns to the  effect of new observations on EMV and IBV, and semi-analytical expected EMV and IBV over these observations are derived. 
+Section \ref{sec:bg_and_notation} introduces notation and co-Kriging
+equations of multivariate GRFs.  Section \ref{sec:set_uq} presents
+uncertainty quantification (UQ) techniques on ESs of GRFs, in
+particular the IBV and the excursion measure variance (EMV).  Section
+\ref{sec:eibv} turns to the effect of new observations on EMV and IBV,
+and semi-analytical expected EMV and IBV over these observations are
+derived.
 %The resulting formulae form the backbone of the expected sequenatil uncertainty reduction strategies presented in Section \ref{sec:heuristics}.
-Section \ref{Sec:UnivarEx} illustrates the concepts on a bivariate example relevant for sampling in the temperature and salinity application.
+Section \ref{Sec:UnivarEx} illustrates the concepts on a bivariate
+example relevant for sampling temperature and salinity in our domain.
 
@@
 
-We denote by $\gp$ a vector-valued random field indexed by some arbitrary domain $\domain$, and assume values of the field at any fixed location $\x \in \domain$, denoted $\gp[\x]$, to be a $\no$-variate random vector ($\no\geq 2$).
-In the river plume characterization case, $\domain$ is a prescribed domain in the fjord of Trondheim (for the purpose of our AUV application, a discretization of a $2$-dimensional domain at fixed depth is considered), and $\no=2$ with responses of temperature and salinity.
-A bivariate GRF model is assumed for $\gp$.
-To motivate concepts, Figure~\ref{example_excu}(left and middle) shows a realization of such a vector-valued GRF on $\domain=(0,1)^2$.
-Figure~\ref{example_excu}(right) represents a by-product of interest derived from these realizations, namely regions i) in red, where both temperature and salinity are high (i.e., exceeding respective thresholds), indicative of ocean water ii) in white, where both temperature and salinity are low, indicative of riverine water, and iii) in pink, where one variable is above and the other below their respective thresholds, indicative of mixed waters.
+We denote by $\gp$ a vector-valued random field indexed by some
+arbitrary domain $\domain$, and assume values of the field at any
+fixed location $\x \in \domain$, denoted $\gp[\x]$, to be a
+$\no$-variate random vector ($\no\geq 2$). In the river plume
+characterization case, $\domain$ is a prescribed domain in the fjord
+of Trondheim (for the purpose of our AUV application, a discretization
+of a $2$-dimensional domain at fixed depth is considered), and $\no=2$
+with responses of temperature and salinity. A bivariate GRF model is
+assumed for $\gp$. To motivate concepts,
+Figure~\ref{example_excu}(left and middle) shows a realization of such
+a vector-valued GRF on
+$\domain=(0,1)^2$. Figure~\ref{example_excu}(right) represents a
+by-product of interest derived from these realizations, namely regions
+i) in red, where both temperature and salinity are high (i.e.,
+exceeding respective thresholds), indicative of ocean water ii) in
+white, where both temperature and salinity are low, indicative of
+riverine water, and iii) in pink, where one variable is above and the
+other below their respective thresholds, indicative of mixed waters.
 

### 2020-06-22 main_new.tex
updates

@@
 $\no$-variate random vector ($\no\geq 2$). In the river plume
-characterization case, $\domain$ is a prescribed domain in the fjord
-of Trondheim (for the purpose of our AUV application, a discretization
-of a $2$-dimensional domain at fixed depth is considered), and $\no=2$
-with responses of temperature and salinity. A bivariate GRF model is
-assumed for $\gp$. To motivate concepts,
-Figure~\ref{example_excu}(left and middle) shows a realization of such
-a vector-valued GRF on
-$\domain=(0,1)^2$. Figure~\ref{example_excu}(right) represents a
-by-product of interest derived from these realizations, namely regions
-i) in red, where both temperature and salinity are high (i.e.,
-exceeding respective thresholds), indicative of ocean water ii) in
-white, where both temperature and salinity are low, indicative of
-riverine water, and iii) in pink, where one variable is above and the
-other below their respective thresholds, indicative of mixed waters.
+characterization case, $\domain$ is a prescribed domain in
+Trondheimsfjord, Norway (for the purpose of our AUV application, a
+discretization of a $2$-dimensional domain at fixed depth is
+considered), and $\no=2$ with responses of temperature and salinity. A
+bivariate GRF model is assumed for $\gp$. To motivate concepts,
+Fig.~\ref{example_excu} (left and middle \kc{If you split this figure
+  in its parts and add labels, it would be more elegant and precise to
+  explicitly refer to them.}) shows a realization of such a
+vector-valued GRF on $\domain=(0,1)^2$. Fig.~\ref{example_excu} (right
+\kc{See above}) represents a by-product of interest derived from these
+realizations, namely regions i) in red, where both temperature and
+salinity are high (i.e., exceeding respective thresholds), indicative
+of ocean water ii) in white, where both temperature and salinity are
+low, indicative of riverine water, and iii) in pink, where one
+variable is above and the other below their respective thresholds,
+indicative of mixed waters.
 
@@
   \includegraphics[width=0.99\textwidth]{Figures/example_excu_1.png}
-  \caption{Realization of a bivariate GRF and excursion set above some threshold ({\bf{JE: are means around 5 and 30 here, so same as later in the example?}} ).
-Joint excursion in red, excursion of a single variable in pink.}
+  \caption{Realization of a bivariate GRF and excursion set above some
+    threshold ({\kc{JE: are means around 5 and 30 here, so same as
+        later in the example?}} ). Joint excursion in red, excursion
+    of a single variable in pink.}
 \label{example_excu}
@@
 
-If we assume that $\gp$ has
-continuous trajectories (almost surely) and $T$ is closed, then
-$\es$ becomes a Random Closed Set
-\citep{Molchanov2005} and concepts from the theory of random sets will prove useful in characterizing the distribution of the volume of $\es$ under some measure defined on the domain.
+If we assume that $\gp$ has continuous trajectories (almost surely
+\kc{This seems very handwavy; would delete}) and $T$ is closed, then
+$\es$ becomes a Random Closed Set \citep{Molchanov2005} and concepts
+from the theory of random sets will prove useful in characterizing the
+distribution of the volume of $\es$ under some measure defined on the
+domain.
 %\textcolor{red}{Is this really needed?} 

### 2020-06-22 main_new.tex
Updates from Overleaf

@@
 % "New Title of the paper"
-\title{Autonomous Oceanographic Data Collection using Excursion Sets of Vector-valued Gaussian Random Fields} 
+\title{Autonomous Oceanographic Data Collection Estimating Excursion Sets of Vector-valued Gaussian Random Fields} 
 \runtitle{Autonomous Oceanographic Data Collection}
-
-% "Former Title of the paper" (DG 08.06.2020)
-%\title{Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields} \runtitle{Excursion Probabilities for Informative Sampling}
 
@@
 \begin{abstract}
-
-  Improving and optimizing oceanographic sampling is a crucial task
-  for marine science and maritime resource management. Faced with
-  limited resources in understanding processes in the water-column,
-  the combination of statistics and autonomous systems provide new
-  opportunities for experimental design. In this work we develop
-  methods for efficient spatial sampling applied to the mapping of
-  coastal ocean processes by providing informative descriptions of
-  spatial characteristics of ocean phenomena. Specifically, we define
-  a design criterion based on improved characterization of the
+Improving and optimizing oceanographic sampling is a crucial task for marine science and maritime resource management.
+Faced with limited resources in understanding processes in the water-column, the combination of statistics and autonomous systems provide new opportunities for experimental design.
+In this work we develop methods for efficient spatial sampling applied to the mapping of coastal ocean processes by providing informative descriptions of spatial characteristics of ocean phenomena.
+Specifically, we define a design criterion based on
  uncertainty in the excursions of vector-valued Gaussian random
-  fields, and derive tractable expressions for the expected Bernoulli
-  variance reduction in such a framework. We demonstrate how this
-  criterion can be used to prioritize sampling efforts at locations
-  that are ambiguous, making exploration more effective. We use
-  simulations to study the properties of methods and to compare them
-  with state-of-the-art approaches, followed by results from field
-  deployments with an autonomous underwater vehicle (AUV) as part of a
-  study mapping the boundary of a river plume. The results demonstrate
-  the potential of combining statistical methods and robotic platforms
-  to effectively inform and execute data-driven environmental
-  sampling.
+ fields, and derive tractable expressions for the expected Bernoulli variance reduction in such a framework.
+We demonstrate how this criterion can be used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective.
+We use simulations to study the properties of methods and to compare them with state-of-the-art approaches, followed by results from field
+ deployments with an autonomous underwater vehicle (AUV) as part of a study mapping the boundary of a river plume.
+The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.
 
@@
 
-If we assume that $\gp$ has continuous trajectories (almost surely
-\kc{This seems very handwavy; would delete}) and $T$ is closed, then
-$\es$ becomes a Random Closed Set \citep{Molchanov2005} and concepts
-from the theory of random sets will prove useful in characterizing the
-distribution of the volume of $\es$ under some measure defined on the
-domain.
+If we assume that $\gp$ has continuous trajectories 
+%(almost surely \kc{This seems very handwavy; would delete}) 
+and $T$ is closed, then
+$\es$ becomes a Random Closed Set \citep{Molchanov2005} and concepts from the theory of random sets will prove useful to study $\es$. 
+%under some measure defined on the domain.
 %\textcolor{red}{Is this really needed?} 
@@
 
-Letting $\gp[\spatloc,\ell]$ denote the $\ell\text{-th}$ component of $\gp[\spatloc]$, and $\ell \in \{1\dots,p\}$, we use the term \textit{generalized location} for the couple $x=(\spatloc,\ell)$. 
+Letting $\gp[\spatloc,\ell]$ denote the $\ell\text{-th}$ component of $\gp[\spatloc]$ ($1\leq \ell\leq \no$), we use the term \textit{generalized location} for the couple $x=(\spatloc,\ell)$. 
 %Given such a generalized location notation $x$,  
@@
 %This slight change of notation 
-will allow us to think of $\gp$ as a scalar-valued Gaussian random field 
-indexed by $\domain \times \{1\dots,p\}$, {\bf{Jo: I used $\domain$ instead of $D$!}}, which will give the co-kriging equations a particularly simple form that parallels the one of univariate kriging.
+will allow us to think of $\gp$ as a scalar-valued random field indexed by $\domain \times \{1\dots,p\}$, which will give the co-kriging equations a particularly simple form that parallels the one of univariate kriging.
 %Due to the naturality of the concept of \textit{generalized location}, we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $\spatloc \in \domain$.
@@
 The letters $\spatloc$ and $\ell$ will be used for spatial locations and response indices respectively.
-
-%It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the one for a single observation, provided some notation adjustments are made.
-This motivates the following.
-
-Given a dataset consisting of  $q$ observations (a batch) at spatial locations $\spatloc_i \in \domain$ and response indices $l_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we use boldface letters to denote concatenated quantities;
+%
+% Replacement by overlooked revised chunks from June 19th
+%\begin{comment}
+%Given a dataset consisting of  $q$ observations (a batch) at spatial locations $\spatloc_i \in \domain$ and response indices $l_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we use boldface letters to denote concatenated quantities;
+%\begin{align*}
+%\bm{x}:=(\bm{\spatloc}, \bm{\ell}):= (x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
+%\end{align*}
+%In general,  corresponding to batches of observations. 
+%The field values of batch observations are denoted by
+%\begin{align*}
+%\gp[\bm{x}]:=
+%\left(\gp[\spatloc_1,\ell_1], ...,
+%\gp[\spatloc_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
+%\end{align*}
+%In the same fashion, $\mu(\bm{x})$ denotes the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ the $q \times q$ covariance matrix.
+
+%Given a random field $\gp$ and (noisy) observations of some of its components at selected points in the domain, one can predict the value of the field at location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$ and the associated covariance .
+This process is called co-Kriging.% , and kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
+%We present a general form of cokriging, where observations at locations $s \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ (heterotopic).
+
+%Assuming that one wishes 
+%to predict $\gp[\bm{x}]$ at $q\geq 1$ generalized locations $\bm{x}=(\bm{\spatloc}, \bm{\ell})$, 
+%given $n$ batched of observations $\mathbf{z}_{[n]}$ at $\sum_{i=1}^n q_i$ generalized locations $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$.
+%The co-Kriging mean then amounts to simple kriging with respect to a scalar-valued GRF indexed by $\domain \times \{1\dots,p\}$: % and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{\spatloc}, \bm{\spatloc}')_{\ell, \ell'}$, that is:
+%
+%\begin{equation}\label{eq:cokrig_mean}
+%\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})),
+%\end{equation}
+%where $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of observed responses of $Z$ at all considered generalized locations, and 
+%where $\lambda_{[n]}(\bm{x})=\left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right]^{-1} k(\bm{x}_{[n]}, \bm{x})$ is a vector of weights and matrix $R(\bm{x}_{[n]}, \bm{x}_{[n]})$ holds the observation noise covariance at the generalized locations.
+%$k(\bm{x}_{[n]}, \bm{x}_{[n]})$ being assumed non-singular throughout the presentation. 
+%The associated co-kriging %(conditional) 
+%residual cross-covariance function can be expressed in the same vein via
+%
+%\begin{equation}\label{eq:cokrig_cov}
+%k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T \left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n]}}(\bm{x}').
+%\end{equation}
+
+%\subsubsection{Co-kriging update formulae}
+
+%In our case it is important to update the GRF model efficiently onboard the vehicle after every batch of obervations.
+Our setting is like that of a spatial Kalman filter with static model parameters, and we provide the co-Kriging equations for this case.
+Because of our representation of co-kriging in terms of simple kriging with respect to generalized locations, the formulae is a straightforward adaptation of the batch-sequential Kriging update formulae from \cite{Chevalier.etal2013a}.
+The formulae are instrumental in deriving semi-analytical results for stepwise expected uncertainty reduction criteria for vector-valued random fields in the subsequent sections.
+
+%Given the observations from $n$ batches, one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
+The mean update formulae is
+%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
+% delivers that
+% 
+%\begin{equation}\label{eq:meanCoK}
+%\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
+%\end{equation}
+%where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
+%$\lambda_{[n+1]}(\bm{x})$ that corresponds to the Kriging weigths associated with the last $q_{n+1}$ responses when 
+%predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
+%\text{th}$ batch.
+%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
+%The updated co-Kriging residual cross-covariance function is
+%\begin{eqnarray}\label{eq:varCoK}
+%k_{[n+1]}(\bm{x},\bm{x}') &=& k_{[n]}(\bm{x},\bm{x}') -c(\bm{x},\bm{x}',\bm{x}_{[n]}), \\
+%c(\bm{x},\bm{x}',\bm{x}_{[n]})&=&\lambda_{[n+1,n+1]}(\bm{x})^T \left[k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})+R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n+1,n+1]}}(\bm{x}'). \nonumber
+%\end{eqnarray}
+%\medskip
+%
+%We remark that, as noted in \cite{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
+%extend to Universal Kriging in second-order settings and apply without Gaussian assumption. 
+%\end{comment}
+%
+%It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the 
+%one for a single observation, provided some notation adjustments are made. This motivates the following.
+Furthemore, boldface letters will be used to denote concatenated quantities corresponding to batches of observations.
+Given a dataset consisting of  $q$ observations at spatial locations $\bm{\spatloc}=(\spatloc_1,\dots,\spatloc_q) \in \domain^q$ and response indices $\bm{\ell}=(\ell_1,\dots, \ell_q)\in \lbrace 1, ..., \no\rbrace^q$, %$i=1, ..., q$,
+we use the concatenated notation
 \begin{align*}
-\bm{x}:=(\bm{\spatloc}, \bm{\ell}):= (x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
+\bm{x}:=%(\bm{\spatloc}, \bm{\ell}):= 
+(x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
 \end{align*}
-%In general,  corresponding to batches of observations. 
-The field values of batch observations are denoted by
+We also compactly denote the field values at those different locations by
 \begin{align*}
@@
 \end{align*}
-In the same fashion, $\mu(\bm{x})$ denotes the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ the $q \times q$ covariance matrix.
-
-Given a random field $\gp$ and (noisy) observations of some of its components at selected points in the domain, one can predict the value of the field at location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$ and the associated covariance .
-This process is called co-Kriging.% , and kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
-We present a general form of cokriging, where observations at locations $s \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ 
+%
+For a second order random field $(Z_{\spatloc})_{\spatloc \in \domain}$ with mean $\mu$ and matrix covariance function $K$, $\mu$ is naturally extended to $\domain \in \lbrace 1, ..., \no\rbrace$ into a function of $x=(\spatloc, \ell)$ and is further straightforwardly vectorized into a function of $\bm{x}$.
+As for $K$, it induces a covariance kernel $k$ on the set of extended locations via  $k((\spatloc, \ell),(\spatloc', \ell'))=K(\spatloc, \spatloc')_{\ell, \ell'}$.
+In vectorized/batch form, $k(\bm{x}, \bm{x}')$ then amounts to a matrix with numbers of lines and columns the numbers of generalized locations in the batches $\bm{x}$ in $\bm{x}'$, respectively.
+Such vectorized quantities will turn out to be quite useful in order to arrive at embarrassingly simple expressions for co-Kriging equations below.
+%\textcolor{red}{(ADDENDUM, C.T.) In the same fashion, we will use $\mu(\bm{x})$ to denote the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ for the corresponding $q \times q$ covariance matrix.}
+
+%\medskip
+
+Now given a GRF $\gp$ and observations of some of its components at some points in the domain, one can predict the value of the field at some unobserved location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$, conditional on the data.
+This coincides with the method called (co-)kriging, and (co-)kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
+We will here present a rather general form of cokriging, in the sense that it allows including several observations (batch) at a time and 
+observations at a given location $u \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ 
 (heterotopic).
 
-Assuming that one wishes 
-to predict $\gp[\bm{x}]$ at $q\geq 1$ generalized locations $\bm{x}=(\bm{\spatloc}, \bm{\ell})$, 
-given $n$ batched of observations $\mathbf{z}_{[n]}$ at $\sum_{i=1}^n q_i$ generalized locations $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$.
-The co-Kriging mean then amounts to simple kriging with respect to a scalar-valued GRF indexed by $\domain \times \{1\dots,p\}$: % and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{\spatloc}, \bm{\spatloc}')_{\ell, \ell'}$, that is:
+Assuming that $n$ batches of observations are available, with respective sizes $q_1,\dots, q_n$, and that one wishes 
+to predict $\gp[\bm{x}]$ for some batch of $q\geq 1$ generalized locations $\bm{x}$, %\equiv(\bm{\spatloc}, \bm{\ell})$, 
+the (simple) cokriging mean then amounts to kriging with respect to a scalar-valued GRF indexed by 
+$\domain\times \{1\dots,p\}$:  
 %
@@
 \end{equation}
-%where $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of observed responses of $Z$ at all considered generalized locations, and 
-where $\lambda_{[n]}(\bm{x})=\left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right]^{-1} k(\bm{x}_{[n]}, \bm{x})$ is a vector of weights and matrix $R(\bm{x}_{[n]}, \bm{x}_{[n]})$ holds the observation noise covariance at the generalized locations.
-%$k(\bm{x}_{[n]}, \bm{x}_{[n]})$ being assumed non-singular throughout the presentation. 
-The associated co-kriging %(conditional) 
-residual cross-covariance function can be expressed in the same vein via
+where $\mu$ is $Z$'s initial mean function, $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of observed (noisy) responses of $Z$ at all considered generalized locations, and $\lambda_{[n]}(\bm{x})$ is a vector of weights
+equal to $\widetilde{\bm{K}_{[0,n]}}^{-1} k(\bm{x}_{[n]}, \bm{x})$ with $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$ and $\widetilde{\bm{K}_{[0,n]}}=k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]}$, assumed non-singular throughout the presentation, where $\Delta_{[n]}$ stands for the covariance matrix of Gaussian-distributed noise assumed to have affected measurements up to batch $n$.
+The co-kriging %(conditional)
+residual (cross-)covariance function (with respect to batches of generalized locations) can also be expressed in the same vein via
 %
 \begin{equation}\label{eq:cokrig_cov}
-k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T \left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n]}}(\bm{x}').
+k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T \widetilde{\bm{K}_{[0,n]}} \lambda_{{[n]}}(\bm{x}').
 \end{equation}
@@
 
-In our case it is important to update the GRF model efficiently onboard the vehicle after every batch of obervations.
-Our setting is like that of a spatial Kalman filter with static model parameters, and we provide the co-Kriging equations for this case.
-Because of our representation of co-kriging in terms of simple kriging with respect to generalized locations, the formulae is a straightforward adaptation of the batch-sequential Kriging update formulae from \cite{Chevalier.etal2013a}.
-The formulae are instrumental in deriving semi-analytical results for stepwise expected uncertainty reduction criteria for vector-valued random fields in the subsequent sections.
-
-Given the observations from $n$ batches, one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
-The mean update formulae is
+Let us now consider the case where co-kriging prediction of $Z$ was made with respect to $n$ batches $\bm{x}_i$ of generalized locations, concatenated again within
+$\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$, and one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
 %It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
-% delivers that
+Thanks to our representation of co-kriging in terms of simple kriging with respect to generalized locations, a strightforward adaptation of the batch-sequential kriging update formulae from \citep{Chevalier.etal2013a} delivers that
 % 
@@
 where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
-$\lambda_{[n+1]}(\bm{x})$ that corresponds to the Kriging weigths associated with the last $q_{n+1}$ responses when 
+$\lambda_{[n+1]}(\bm{x})$ that corresponds to the kriging weigths associated with the last $q_{n+1}$ responses when 
 predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
@@
 %, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
-The updated co-Kriging residual cross-covariance function is
-\begin{eqnarray}\label{eq:varCoK}
-k_{[n+1]}(\bm{x},\bm{x}') &=& k_{[n]}(\bm{x},\bm{x}') -c(\bm{x},\bm{x}',\bm{x}_{[n]}), \\
-c(\bm{x},\bm{x}',\bm{x}_{[n]})&=&\lambda_{[n+1,n+1]}(\bm{x})^T \left[k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})+R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n+1,n+1]}}(\bm{x}'). \nonumber
-\end{eqnarray}
+Similarly, the updated co-kriging residual (cross-)covariance function then writes
+\begin{equation}\label{eq:varCoK}
+k_{[n+1]}(\bm{x},\bm{x}')=k_{[n]}(\bm{x},\bm{x}')-\lambda_{[n+1,n+1]}(\bm{x})^T 
+\widetilde{\bm{K}_{[n,n+1]}}
+%k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})
+\lambda_{{[n+1,n+1]}}(\bm{x}'),
+\end{equation}
 %\medskip
 %
-We remark that, as noted in \cite{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
+where $\widetilde{\bm{K}_{[n,n+1]}}=k_{[n]}(\bm{x}_{n+1}, \bm{x}_{n+1})+\Delta_{n+1}$.
+Let us remark that, as noted in \citep{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
 extend to Universal Kriging in second-order settings and apply without Gaussian assumption.
+We will now see how the latter formulae are instrumental in deriving semi-analytical formulae for stepwise uncertainty reduction criteria for
+vector-valued random fields.
 
@@
 \end{equation}
-where $\bm{Y}$ is distributed according to the current law of $Z_{\bm{x}}$ and with independent noise having covariance matrix $R(\bm{x},\bm{x})$.
+where $\bm{Y}$ is distributed according to the current law of $Z_{\bm{x}}$ and with independent noise having covariance matrix $\Delta_n$.
 
@@
 
-{\bf{Jo: Could we get the noise covariance $R$ in here? }}
-
 As for the expected EMV, a similar result may be derived.
@@
 Related closed-form solutions have been noted in similar contexts \citep{bhattacharjya2013value,stroh}, but not generalized to our situation with
-random sets for vector-valued GPs.
-
-\subsection{EBV for a bivariate example}
+random sets for vector-valued GRFs.
+
+\subsection{Expected Bernoulli Variance on a two dimensional Example}
 \label{Sec:UnivarEx}
-We illustrate the EBV associated with different designs on a simple bivariate example that mimics our river plume application.
-The first and second component of the random field are \textit{temperature} and \textit{salinity}.
+
+We illustrate the EBV associated to different designs on a simple bivariate example.
+The example mimics our river plume application and hence the first and second component of the random field will be called \textit{temperature} and \textit{salinity} for illustrative purpose.
 Effects of the hyperparameters of the GRF prior on the excursion probabilities will also be studied.
-For simplicity, we begin with a \textit{single-location} example, considering a bivariate Gaussian distribution without any spatial elements.
+For simplicity, we begin with a \textit{pointwise} example, considering a single bivariate gaussian distribution (no spatiality).
 
 \subsubsection{A pointwise study}
-For this bivariate Gaussian, the ES is defined as
-\[
-\Gamma := \lbrace \spatloc \in \domain: Z_{\spatloc, 1} \geq t_1, Z_{\spatloc, 2}\geq t_2\rbrace
-\]
-with thresholds set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ mg/l for salinity.
-We play with the temperature and salinity correlation and variances to study the effect on the EP and EBV.
+
+Say we want to study the excursion probability of a bivariate gaussian above some thresholds, where the thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ g/kg for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
 
@@
 \begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/illus_bivarZ.eps}
+  \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
   \caption{Density contour plots with different correlations between
@@
     thresholds are identical to the mean values $5^o C$ and
-    $30 mg/l$. X-axis is temperature and y-axis is salinity.}
+    $30$ g/kg.}
 \label{illus_bivarDens}
@@
 \ref{illus_bivarDens}. The EPs increase with the correlation as there
-is a strong tendency to have concurrently low temperature and salinity. Note that EPs and Bernoulli variances are the same for standard
+is a strong tendency to have concurrently low temperature and salinity. The Bernoulli variance is similarly large for high
+correlations. EPs and Bernoulli variances are the same for standard
 deviation $1$ or $2$, which implies that high variability in
@@
 $p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
-EBV, Temp and Salinity data & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
-EBV, Temperature data only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
+EIBV, Temp and Salinity data & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
+EIBV, Temperature data only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
 \hline
@@
 
-Table \ref{tab:sim_rhoab} (bottom two rows) shows results of EBV calculations. This is presented for a design
+Table \ref{tab:sim_rhoab} (bottom two rows) shows results of expected
+Bernoulli variance calculations. This is presented for a design
 gathering both data types, and for a design with temperature
 measurements alone. When both data are gathered, the measurement model is
-$(Y_{d,1},Y_{d,2})^t=(Z_1,Z_2)^t+\bepsilon$, with $\bepsilon \sim N(0,0.5^2I_2)$, while $Y_d=Z_1+\epsilon$, $\epsilon \sim N(0,0.5^2)$ when only temperature is measured.
+$(Y_1,Y_2)^t=(Z_1,Z_2)^t+\bepsilon$, with $\bepsilon \sim N(0,0.5^2I_2)$, while $Y_1=Z_1+\epsilon$, $\epsilon \sim N(0,0.5^2)$ when only temperature is measured.
 For this illustration, Table \ref{tab:sim_rhoab} shows that the
-EBV gets lower with larger standard deviations
-$\sigma_1$ and $\sigma_2$ (right columns). The expected reduction of Bernoulli
+expected Bernoulli variance gets lower with larger standard deviations
+$\sigma_1$ and $\sigma_2$ (right columns). The reduction of Bernoulli
 variance is largest for the cases with high correlation
@@
 ($\gamma=0.2$), there is little information about salinity in the
-temperature data, and therefore less uncertainty reduction.
-
-\subsubsection{Including Spatial elements}
-We extend this example to a full-fledged GRF model. 
-The class of GRF model we will consider in the article will generally have a linear trend
+temperature data, and therefore less uncertainty reduction. In an
+application with fresh cold water from a river source, the temperature
+and salinity variables will not only be interdependent, but will also
+likely show dependence in the spatial dimension. This in turn will
+impact the design criteria when we evaluate the information measure by
+integrating over several locations (Section \ref{sec:simulations}).
+
+\subsubsection{Including Spatiality}
+We now turn to an example involving a full-fledged gaussian random field (GRF). The class of GRF model we will consider in the article will generally have a linear trend
 \begin{align*}
 \mu(s)=\mathbb{E}\left[\begin{pmatrix}
-Z_{s, 1}\\ Z_{s, 2}
-\end{pmatrix}\right] &= \beta_0 + \beta_1 s
+Z_{\spatloc, 1}\\ Z_{\spatloc, 2}
+\end{pmatrix}\right] &= \beta_0 + \beta_1 \spatloc
 \end{align*}
-with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. We will only consider covariance models of separable type
+with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. In our examples, we will only consider covariance models of separable type
 \begin{align*}
-\textrm{Cov}\left(Z_{s, i}, Z_{u, j}\right) &= k(s, u) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_i^2,~ i=j\\
-   \gamma_0 \sigma_i \sigma_j,~i\neq j,
+\textrm{Cov}\left(Z_{\spatloc, i}, Z_{v, j}\right) &= k(\spatloc, v) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_l^2,~ i=j\\
+   \gamma_0 \sigma_i \sigma_j,~i\neq j
         \end{cases}
 \end{align*}
-where $k(., .)$ is specified as a Mat\'{e}rn covariance kernel in the spatial domain, i.e.
-$(1+\phi h) \exp(-\phi h)$ where $h$ is the Euclidean distance between locations $u$ and $u'$, and $\gamma$ defines the cross-covariance structure.
-In the accompanying Python code ({\bf{give URL here, or only in bullet point list in Introduction}}, these modeling assumptions can be
-generalized to anisotropic covariance and changing variance levels
-across the spatial domain. One can hence, easily run the code to see
-if different models result in other sampling designs. Anisotropy and
+where $k(., .)$ is one of the traditional spatial covariance kernel (exponential, Mat\'{e}rn, ...) and $\gamma$ defines the cross-covariance structure.
+
+In the accompanying Python examples taking place within the MESLAS toolbox (URL), these modeling assumptions can be
+generalized to anisotropic covariance and changing variance levels across the spatial domain. Anisotropy and
 non-stationary variance are both relevant for the setting with river
@@
 \end{align*}
-and the spatial covariance is given by a Mat\'{e}rn 3/2 kernel with unit variance and range parameter $\phi=0.5$.
-{\bf{Jo: Is this the decay parameter in formulae above, or are you trying to calculate some kind of effective spatial range?}}
+and the spatial covariance is given by a Mat\'{e}rn 3/2 kernel with unit variance and range parameter $\lambda=0.5$.
 One realization of this GRF is shown in Fig. \ref{example_excu}.
 
-We now study how the EBV associated with data collection at a point changes if only one of the two components of the field is observed.
-We first draw a realization of the GRF defined above and use it as ground-truth to mimic the real data-collection process.
-A first data collection step is performed at the location in green, and the data is used to update the GRF model.
-The EBV reduction (at each grid point) that would result from observing only one component, or both at the next grid point upwards is then computed an shown in Fig.\ref{fig:ebv_comp}.
+We now study how the expected Bernoulli variance eq.\eqref{def:eibv} associated to data collection at a point changes if only one of the two components of the field is observed.
+We first draw a realization of the GRF defined above and use it as ground-truth to mimick the real data-collection process.
+A first bunch of observations are done at the locations depicted in grey, and the data is used to update the GRF model.
+We then consider the green triangle as a potential next observation locations and plot the expected Bernoulli variance reduction (at each grid point) that would result from observing only one component of the field (temperature or salinity), or both at that point, see Fig.\ref{fig:ebv_comp}.
 
@@
 Data collection locations in green.
-{\bf{Jo: left display not explained.}} Places where only one response is above threshold are depicted in pink}
+True excursion set in red.
+Places where only one response is above threshold are depicted in pink}
 \label{fig:ebv_comp}
@@
 
-Note that plotting the Bernoulli variance reduction at each point might also be used to compare different pre-scripted data collection plans.
+Note that plotting the Bernoulli variance reduction at each point might also be used to compare different data collection plans.
 For example, Fig.
-\ref{fig:ebv_comp} shows the EBV reduction associated to a data collection plan along a straight upwards line (static north) and the one associated to a straight horizontal line (static east).
-Such calculations for pre-planned surveys are done in just a single batch.
+\ref{fig:ebv_north_vs_east} shows the expected Bernoulli variance reduction associated to a data collection plan along a straight upwards line (static north) and the one associated to a straight horizontal line (static east).
+Both expectations are computed according to the a priori distribution of the GRF (i.e.
+no observations have been included yet).
 
@@
 Data collection locations in green.
-True ES is in red (left).
-Places where only one response is above threshold are depicted in pink.
-{\bf{Jo: is the left display wrong?}}}
+True excursion set is in red.
+Places where only one response is above threshold are depicted in pink.}
 \label{fig:ebv_north_vs_east}
@@
 
-In all examples, the spatial domain $\domain$ is discretized to a set of $N$ grid
+%\begin{remark}
+In all the computed examples, the spatial domain $\domain$ has been discretized to a set of $N$ grid
 locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,N \}$, where each cell
-has area $\Delta$; the same grid as used for the waypoint graph for
+has area $\delta$; the same grid is used for the waypoint graph for
 possible design locations. The EIBV is approximated by sums over all
 grid cells.
+%\end{remark}
 
@@
 the trends in environmental conditions and correlation
-structures. Based on the initial data, the trend parameters were
-estimated by linear regression, where both temperature and salinity
-are assumed to increase linearly, going west from the river
-mouth. Next, the residuals from the regression analysis were analyzed
-to study the fit of the GP model and to specify the covariance
-parameters.
+structures. Based on the initial data, the mean was specified by linear regression, where both temperature and salinity
+were assumed to increase with the west location. Next, the residuals from the regression analysis were analyzed to specify the covariance parameters of the GRF model.
 
@@
 temperature are subtracted from the data. This scatter-plot of joint
-residuals indicates larger variability in salinity than in
+residuals indicates larger variability in salinity than
 temperature, and a positive correlation ($0.5$) between the two
@@
 approximately $\chi^2_2$ distributed. Fig. \ref{fig:parest_b} shows
-the empirical cumulative distribution function (CDF) of these
+the empirical CDF of these
 quadratic forms (solid) together with the theoretical CDF of the
@@
 similar, which indicates that the Gaussian model fits reasonably
-well. Even though there appears to be some clustering in both
-Fig. \ref{fig:parest_a} and \ref{fig:parest_b}, the bivariate
-diagnostic plots look reasonable and justify a Gaussian
-model. Fig. \ref{fig:parest_c} shows the empirical variogram of the
+well. Fig. \ref{fig:parest_c} shows the empirical variogram of the
 scaled residuals for temperature and salinity. The decay is similar
 for the two, and seems to be negligible after about $150$ m.
+The working assumption of a separable covariance function is hence not unreasonable.
 
@@
 algorithm's ability to characterize the river plume front separating
-the river and fjord water masses, providing a spatial map of this
-boundary.
+the river and fjord water masses.
 
@@
 \rowcolor{Gray}
-Cross correlation temp. and sal. & 0.5 & AUV observations\\
-Temp. variance &  0.20 & AUV observations (variogram)\\
+Cross correlation temperature and salinity & 0.5 & AUV observations\\
+Temperature variance &  0.20 & AUV observations (variogram)\\
 \rowcolor{Gray}
-Sal. variance &  5.76 & AUV observations (variogram)\\
-Corr. range  & 0.15 km & AUV observations (variogram)\\
+Salinity variance &  5.76 & AUV observations (variogram)\\
+Correlation range  & 0.15 km & AUV observations (variogram)\\
 \rowcolor{Gray}
-River temp. $T_{river}$ & $10.0\,^{\circ}\mathrm{C}$ & AUV observations\\
-Ocean temp. $T_{ocean}$ & $11.0\,^{\circ}\mathrm{C}$ & AUV observations\\
+River temperature  & $10.0\,^{\circ}\mathrm{C}$ & AUV observations\\
+Ocean temperature $T_{ocean}$ & $11.0\,^{\circ}\mathrm{C}$ & AUV observations\\
 \rowcolor{Gray}
-River sal. $S_{river}$ & $14.0$ g/kg & AUV observations\\
-Ocean sal. $S_{ocean}$ & $22.0$ g/kg & AUV observations\\
+River salinity $S_{river}$ & $14.0$ g/kg & AUV observations\\
+Ocean salinity $S_{ocean}$ & $22.0$ g/kg & AUV observations\\
 \rowcolor{Gray}
-Threshold temp. & $10.5\,^{\circ}\mathrm{C}$ & $(T_{ocean}-T_{river})/2+T_{river}$\\
-Threshold sal. & $18.0$ g/kg & $(S_{ocean}-S_{river})/2+S_{river}$\\
+Threshold in temperature & $10.5\,^{\circ}\mathrm{C}$ & User specified \\
+Threshold in salinity & $18.0$ g/kg & User specified \\
 \rowcolor{Gray}
@@
   survey. Observations were taken across the front while crossing from
-  fresh, cold river water to saline and warmer ocean waters. \kc{Might
-  be good to expand the terms and not have abbreviations in the table
-  above; there's plenty of space.}}
+  fresh, cold river water to saline and warmer ocean waters.}
 \label{tab:experiment_param}
@@
 
-The sampling locations were distributed over an equilateral grid, as
-shown in the grey-colored lattice in Fig. \ref{fig:map}. The robotic
-platform consisted of a Light AUV \citep{sousa2012lauv}
+A Light AUV \citep{sousa2012lauv}
 (Fig. \ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49
-conductivity, temperature, and depth (CTD) sensor providing
-temperature and salinity
-measurements. %The accuracy of the CTD instrument is $\pm 0.0003$ S/m (conductivity) and $\pm0.002\,^{\circ}\mathrm{C}$ (temperature).
-The sampling agent was built on top of the autonomous agent framework
+conductivity, temperature, and depth (CTD) sensor was used to provide
+salinity and temperature measurements.
+We assume that the measurements are conditionally independent because the salinity is extracted from the conductivity instrument which is different from the temperature instrument.
+We specify variance $0.25^2$ for both errors, which is based on a middle ground between the nugget effect in the empirical variogram and the sensor specifications.
+
+The AUV is a powered untethered platform that
+operates at $1$-$3$ m/s in the upper water column. The AUV uses a multicore GPU NVIDIA
+Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared
+cache, and 4 GB of 1600 MHz DRAM) for computation onboard.
+The sampling algorithm was built on top of the autonomous agent framework
 Teleo-Reactive EXecutive (\textit{T-REX})
-\citep{py10,Rajan12,Rajan12b}, running an instance of the
-\textit{myopic} strategy from Section \ref{sec:myopic} to 
-control the
-AUV and decide between sampling locations.
-
-\textcolor{blue}{TO BE MERGED WITH THE PARA ABOVE? The AUVs considered here are powered untethered platforms, that
-operate at $1$-$3$ m/s in the upper water column, where they are free
-to move in six degrees of freedom (6 DOF). The in-water operation time
-capability depends on survey speed, payload sensors and navigation;
-typically this amounts to 8-48 hrs. AUVs typically use single-board
-computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA
-Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared
-cache, and 4 GB of 1600 MHz DRAM) for computation onboard.}
+\citep{py10,Rajan12,Rajan12b}.
 
@@
 
-The sampling strategy was designed around the concept of visiting
-waypoints sequentially. Arriving at a desired waypoint with new
-measurements and an updated model, the AUV triggers the myopic
-strategy to evaluate the different design criteria (see
-Eq. \eqref{critSEQ}). The waypoint-and-path combination that is
-expected to reduce the IBV the most is selected, and upon arrival this
-procedure is then repeated. At each stage, it takes the AUV about 30
-seconds to evaluate the EIBV for all the possible waypoint-and-path
+The AUV was running a
+\textit{myopic} strategy to decide between sampling locations on the waypoint graph distributed over an equilateral grid, as shown in the grey-colored lattice in Fig. \ref{fig:map}. 
+At each stage, it takes the AUV about 30
+seconds to evaluate the EIBV for all the possible waypoint
 alternatives.
-
 The AUV was set to start in the south-center part of the waypoint
-graph, with the previously outlined GP model of the environment
-(Section \ref{sec:exp_modeling}). A survey was set to take
+graph. A survey was set to take
 approximately 40 minutes, visiting 15 waypoints on the grid, with the
 AUV running near the surface to capture the plume. On its path from
-one waypoint to the next, the AUV gathered data regularly, and the GP
-model assimilated temperature and salinity data with an update
-frequency of 30 seconds, giving about three updates per stage.
+one waypoint to the next, the AUV collected data with an update
+frequency of 30 seconds, giving three measurements per batch in the updates at each stage.
 
@@
   temperature and salinity data correspond with an indication of the
-  EP front. \kc{Might want to indicate when the surveys (dates) were
-    conducted so people can understand the variability in the EP with
-    respect to time.}}
+  EP front. About 2 hours time separated the two runs.}
 \label{fig:results}
@@
 
-Two surveys missions (1 and 2), were run successively from 11:00 AM to
-01:00 PM, with a short break in between. The resulting path of the
+Two survey missions (1 and 2), were run successively, with a short break in between. The resulting path of the
 selected waypoints are shown in the map in Fig. \ref{fig:map}, both
@@
 region and gradually bending off toward the north east. The amount of
-exploration done by Survey 1 is greater than Survey 2. In Survey 1,
-the AUV obtained more detail by going north from waypoint 5, while
-Survey 2, coming close to the survey area borders in the south-western
-corner, obtained a poorer understanding of the northern parts. A
-look-ahead strategy might identify and discourage such choices.
-
-%\begin{figure}[!h] 
-% \centering 
-%\includegraphics[width=0.48\textwidth]{Figures/envir_ocean.pdf}
-%\caption{Ocean observation is moving away from %single-ship sampling
-%towards more collaborative networked operations in %order to resolve
-%the numerous processes and their interaction.} %\label{fig:envir}
-%\end{figure}
-%\newpage
+exploration done by Survey 1 which turned north is greater than Survey 2 which was coming close to the survey area borders in the south-western
+corner.
+
 \section{Closing remarks}\label{sec:concl_disc}

### 2020-06-22 main_new.tex
Updates from Overleaf

@@
 \begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/example_excu_1.png}
+  \includegraphics[width=0.99\textwidth]{Figures/example_excu_4.png}
   \caption{Realization of a bivariate GRF and excursion set above some
@@
 where $\mu$ is $Z$'s initial mean function, $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of observed (noisy) responses of $Z$ at all considered generalized locations, and $\lambda_{[n]}(\bm{x})$ is a vector of weights
-equal to $\widetilde{\bm{K}_{[0,n]}}^{-1} k(\bm{x}_{[n]}, \bm{x})$ with $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$ and $\widetilde{\bm{K}_{[0,n]}}=k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]}$, assumed non-singular throughout the presentation, where $\Delta_{[n]}$ stands for the covariance matrix of Gaussian-distributed noise assumed to have affected measurements up to batch $n$.
+equal to 
+$$\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)^{-1} k(\bm{x}_{[n]}, \bm{x})
+$$
+with $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$ and where $\Delta_{[n]}$ stands for the covariance matrix of Gaussian-distributed noise assumed to have affected measurements up to batch $n$.
+The matrix in parenthesis in the above equation will be assumed non-singular throughout the presentation.
 The co-kriging %(conditional)
@@
 \begin{equation}\label{eq:cokrig_cov}
-k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T \widetilde{\bm{K}_{[0,n]}} \lambda_{{[n]}}(\bm{x}').
+k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T 
+\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)
+\lambda_{{[n]}}(\bm{x}').
 \end{equation}

### 2020-06-22 main_new.tex
Updates from Overleaf

@@
 Similarly, the updated co-kriging residual (cross-)covariance function then writes
-\begin{equation}\label{eq:varCoK}
-k_{[n+1]}(\bm{x},\bm{x}')=k_{[n]}(\bm{x},\bm{x}')-\lambda_{[n+1,n+1]}(\bm{x})^T 
-\widetilde{\bm{K}_{[n,n+1]}}
+\begin{align}\label{eq:varCoK}
+k_{[n+1]}(\bm{x},\bm{x}') & = k_{[n]}(\bm{x},\bm{x}')\\
+ \nonumber - & \lambda_{[n+1,n+1]}(\bm{x})^T 
+\left(k_{[n]}(\bm{x}_{n+1}, \bm{x}_{n+1})+\Delta_{n+1}\right)
 %k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})
 \lambda_{{[n+1,n+1]}}(\bm{x}'),
-\end{equation}
+\end{align}
 %\medskip
 %
-where $\widetilde{\bm{K}_{[n,n+1]}}=k_{[n]}(\bm{x}_{n+1}, \bm{x}_{n+1})+\Delta_{n+1}$.
 Let us remark that, as noted in \citep{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
@@
 \eemv_{[n]}(\bm{x})
-&=\int_{D^2} 
+&=\int_{\domain^2} 
 \varPhi_{2\no}
@@
 (u,v)\\
-&-\currentExp{
-    \varPhi_{\no}\left(a_1+B_1 V; C_1\right)
-    \varPhi_{\no}\left(a_2+B_2 V; C_2\right)
-},
+&-\int_{\domain^2}\currentExp{
+    \varPhi_{\no}\left(a_1+B_1 V; K_{[n+1]}(u)\right)
+    \varPhi_{\no}\left(a_2+B_2 V; K_{[n+1]}(v)\right)
+}\mathrm{d}\mes^{\otimes}(u, v),
 \end{split}
@@
 
-with $V=\gp[\bm{x}_{n+1}]-\currentMean{\bm{x}_{n+1}} \sim \mathcal{N}(0_{q_{n+1}},k_{[n]}(\bm{x}_{n+1},\bm{x}_{n+1}))$ and $a_1=\bt-\currentMean{\uu}$,
+with $V=\gp[\bm{x}]-\currentMean{\bm{x}} \sim \mathcal{N}(0_{q_{n+1}},k_{[n]}(\bm{x},\bm{x}))$ and $a_1=\bt-\currentMean{\uu}$,
 $B_1=-\lambda_{[n+1,n+1]}(\uu)^T$, $a_2=\bt-\currentMean{\vv}$, $B_2=-\lambda_{[n+1,n+1]}(\vv)^T$.
@@
 \begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/ebv_comp_2.png}
+  \includegraphics[width=0.99\textwidth]{Figures/ebv_comp_3.png}
   \caption{Pointwise Bernoulli variance reduction for observation of a single or both components of the random field at one location.
@@
   \hfill 
-  \subfigure[Computational time for inferencing.\kc{I see only 
-    lines associated with 4 variables showing in the graph. Where is
-    static\_north and static\_east?}]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}} 
+  \subfigure[Computational time for inference (some of the lines overlap).]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}} 
 \caption{Simulation results from 100 replicate simulations for 10
-  sampling choices/stages on the grid. \kc{Vertical lines show large
-    variation in replicate results.}}  
+  sampling choices/stages on the grid. Vertical lines show variation in replicate results.}  
 \label{fig:sim_results}

### 2020-06-23 main_new.tex
updates

@@
 \begin{abstract}
-Improving and optimizing oceanographic sampling is a crucial task for marine science and maritime resource management.
-Faced with limited resources in understanding processes in the water-column, the combination of statistics and autonomous systems provide new opportunities for experimental design.
-In this work we develop methods for efficient spatial sampling applied to the mapping of coastal ocean processes by providing informative descriptions of spatial characteristics of ocean phenomena.
-Specifically, we define a design criterion based on
- uncertainty in the excursions of vector-valued Gaussian random
- fields, and derive tractable expressions for the expected Bernoulli variance reduction in such a framework.
-We demonstrate how this criterion can be used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective.
-We use simulations to study the properties of methods and to compare them with state-of-the-art approaches, followed by results from field
- deployments with an autonomous underwater vehicle (AUV) as part of a study mapping the boundary of a river plume.
-The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.
+  Improving and optimizing oceanographic sampling is a crucial task
+  for marine science and maritime resource management. Faced with
+  limited resources in understanding processes in the water-column,
+  the combination of statistics and autonomous systems provide new
+  opportunities for experimental design. In this work we develop
+  methods for efficient spatial sampling applied to the mapping of
+  coastal ocean processes by providing informative descriptions of
+  spatial characteristics of ocean phenomena. Specifically, we define
+  a design criterion based on uncertainty in the excursions of
+  vector-valued Gaussian random fields, and derive tractable
+  expressions for the expected Bernoulli variance reduction in such a
+  framework. We demonstrate how this criterion can be used to
+  prioritize sampling efforts at locations that are ambiguous, making
+  exploration more effective. We use simulations to study the
+  properties of methods and to compare them with state-of-the-art
+  approaches, followed by results from field deployments with an
+  autonomous underwater vehicle (AUV) as part of a study mapping the
+  boundary of a river plume. The results demonstrate the potential of
+  combining statistical methods and robotic platforms to effectively
+  inform and execute data-driven environmental sampling.
 
@@
 % Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
-For the general setting of a $\no$-variate random field, 
-we are interested in recovering the set of locations $\es$ in the domain for which the components of $\gp$ lie in some set of specified values $\T\subset \mathbb{R}^{\no}$, in other words \textit{the pre-image of $T$ by $\gp$}:
+For the general setting of a $\no$-variate random field, we are
+interested in recovering the set of locations $\es$ in the domain for
+which the components of $\gp$ lie in some set of specified values
+$\T\subset \mathbb{R}^{\no}$; in other words \textit{the pre-image of
+  $T$ by $\gp$}:
 $$
@@
 %(almost surely \kc{This seems very handwavy; would delete}) 
-and $T$ is closed, then
-$\es$ becomes a Random Closed Set \citep{Molchanov2005} and concepts from the theory of random sets will prove useful to study $\es$. 
+and $T$ is closed, then $\es$ becomes a Random Closed Set
+\citep{Molchanov2005} and concepts from the theory of random sets will
+prove useful to study $\es$.
 %under some measure defined on the domain.
 %\textcolor{red}{Is this really needed?} 
-Note that while some aspects of the developed approaches do not call for a
-specific form of $\T$, we will often stick for simplicity to the case
-of orthants
+Note that while some aspects of the developed approaches do not call
+for a specific form of $\T$, we will often, for purposes of
+simplicity, stay with the case of orthants
 ($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
@@
 
-Letting $\gp[\spatloc,\ell]$ denote the $\ell\text{-th}$ component of $\gp[\spatloc]$ ($1\leq \ell\leq \no$), we use the term \textit{generalized location} for the couple $x=(\spatloc,\ell)$. 
+Letting $\gp[\spatloc,\ell]$ denote the $\ell\text{-th}$ component of
+$\gp[\spatloc]$ ($1\leq \ell\leq \no$), we use the term
+\textit{generalized location} for the couple $x=(\spatloc,\ell)$.
 %Given such a generalized location notation $x$,  
@@
 %one for a single observation, provided some notation adjustments are made. This motivates the following.
-Furthemore, boldface letters will be used to denote concatenated quantities corresponding to batches of observations.
-Given a dataset consisting of  $q$ observations at spatial locations $\bm{\spatloc}=(\spatloc_1,\dots,\spatloc_q) \in \domain^q$ and response indices $\bm{\ell}=(\ell_1,\dots, \ell_q)\in \lbrace 1, ..., \no\rbrace^q$, %$i=1, ..., q$,
+Furthemore, boldface letters will be used to denote concatenated
+quantities corresponding to batches of observations.  Given a dataset
+consisting of
+$q$ observations at spatial locations
+$\bm{\spatloc}=(\spatloc_1,\dots,\spatloc_q) \in
+\domain^q$ and response indices $\bm{\ell}=(\ell_1,\dots, \ell_q)\in
+\lbrace 1, ..., \no\rbrace^q$, %$i=1, ..., q$,
 we use the concatenated notation
@@
 %
-For a second order random field $(Z_{\spatloc})_{\spatloc \in \domain}$ with mean $\mu$ and matrix covariance function $K$, $\mu$ is naturally extended to $\domain \in \lbrace 1, ..., \no\rbrace$ into a function of $x=(\spatloc, \ell)$ and is further straightforwardly vectorized into a function of $\bm{x}$.
-As for $K$, it induces a covariance kernel $k$ on the set of extended locations via  $k((\spatloc, \ell),(\spatloc', \ell'))=K(\spatloc, \spatloc')_{\ell, \ell'}$.
-In vectorized/batch form, $k(\bm{x}, \bm{x}')$ then amounts to a matrix with numbers of lines and columns the numbers of generalized locations in the batches $\bm{x}$ in $\bm{x}'$, respectively.
-Such vectorized quantities will turn out to be quite useful in order to arrive at embarrassingly simple expressions for co-Kriging equations below.
+For a second order random field $(Z_{\spatloc})_{\spatloc \in
+  \domain}$ with mean $\mu$ and matrix covariance function $K$,
+$\mu$ is naturally extended to $\domain \in \lbrace 1, ...,
+\no\rbrace$ into a function of $x=(\spatloc,
+\ell)$ and is further straightforwardly vectorized into a function of
+$\bm{x}$. As for $K$, it induces a covariance kernel
+$k$ on the set of extended locations via $k((\spatloc,
+\ell),(\spatloc', \ell'))=K(\spatloc, \spatloc')_{\ell,
+  \ell'}$. In vectorized/batch form, $k(\bm{x},
+\bm{x}')$ then amounts to a matrix with numbers of lines and columns
+the numbers of generalized locations in the batches
+$\bm{x}$ in
+$\bm{x}'$, respectively \kc{This last sentence doesn't parse.}. Such
+vectorized quantities turn out to be useful in order to arrive at
+simple expressions for co-Kriging equations below.
 %\textcolor{red}{(ADDENDUM, C.T.) In the same fashion, we will use $\mu(\bm{x})$ to denote the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ for the corresponding $q \times q$ covariance matrix.}
@@
 
-Now given a GRF $\gp$ and observations of some of its components at some points in the domain, one can predict the value of the field at some unobserved location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$, conditional on the data.
-This coincides with the method called (co-)kriging, and (co-)kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
-We will here present a rather general form of cokriging, in the sense that it allows including several observations (batch) at a time and 
-observations at a given location $u \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ 
-(heterotopic).
-
-Assuming that $n$ batches of observations are available, with respective sizes $q_1,\dots, q_n$, and that one wishes 
-to predict $\gp[\bm{x}]$ for some batch of $q\geq 1$ generalized locations $\bm{x}$, %\equiv(\bm{\spatloc}, \bm{\ell})$, 
-the (simple) cokriging mean then amounts to kriging with respect to a scalar-valued GRF indexed by 
-$\domain\times \{1\dots,p\}$:  
+Given a GRF $\gp$ and observations of some of its components at some
+points in the domain, one can predict the value of the field at some
+unobserved location $\spatloc\in \domain$ by using the conditional
+mean of $\gp[\spatloc]$, conditional on the data \kc{...conditional
+  mean on the data?}. This coincides with (co-)kriging equations,
+which tell us precisely how to compute conditional means and
+covariances conditional on an arbitrary dataset.  We will present a
+general form of co-kriging, in the sense that it allows inclusion of
+several observations (batch) at a time; observations at a given
+location $u \in \domain$ may only include a subset of the components
+of $\gp[\spatloc]\in\mathbb{R}^{\no}$ (heterotopic).
+
+Assuming that $n$ batches of observations are available with sizes
+$q_1,\dots, q_n$, and that one wishes to predict $\gp[\bm{x}]$ for
+some batch of $q\geq 1$ generalized locations
+$\bm{x}$, %\equiv(\bm{\spatloc}, \bm{\ell})$,
+the simple cokriging mean then amounts to kriging with respect to a
+scalar-valued GRF indexed by $\domain\times \{1\dots,p\}$:
 %
@@
 \end{equation}
-where $\mu$ is $Z$'s initial mean function, $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of observed (noisy) responses of $Z$ at all considered generalized locations, and $\lambda_{[n]}(\bm{x})$ is a vector of weights
-equal to 
+where $\mu$ is $Z$'s initial mean function,
+$\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n
+q_i$)-dimensional vector of observed (noisy) responses of
+$Z$ at all considered generalized locations, and
+$\lambda_{[n]}(\bm{x})$ is a vector of weights equal to
 $$\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)^{-1} k(\bm{x}_{[n]}, \bm{x})
 $$
-with $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$ and where $\Delta_{[n]}$ stands for the covariance matrix of Gaussian-distributed noise assumed to have affected measurements up to batch $n$.
-The matrix in parenthesis in the above equation will be assumed non-singular throughout the presentation.
-The co-kriging %(conditional)
-residual (cross-)covariance function (with respect to batches of generalized locations) can also be expressed in the same vein via
+with $\bm{x}_{[n]}=(\bm{x}_1,\dots,
+\bm{x}_n)$ and where
+$\Delta_{[n]}$ stands for the covariance matrix of
+Gaussian-distributed noise assumed to have affected measurements up to
+batch
+$n$. The matrix in parenthesis will be assumed to be non-singular
+throughout the presentation. The co-kriging %(conditional)
+residual (cross-)covariance function (with respect to batches of
+generalized locations) can also be expressed in the same vein via
 %
@@
 
-Let us now consider the case where co-kriging prediction of $Z$ was made with respect to $n$ batches $\bm{x}_i$ of generalized locations, concatenated again within
-$\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$, and one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
+Let us now consider the case where a co-kriging prediction of $Z$ was
+made with respect to $n$ batches $\bm{x}_i$ of generalized locations,
+concatenated again within $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$,
+and one wishes to update the prediction by incorporating a new vector
+of observations $\mathbf{z}_{n+1}$ measured at a batch of
+$q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
 %It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
-Thanks to our representation of co-kriging in terms of simple kriging with respect to generalized locations, a strightforward adaptation of the batch-sequential kriging update formulae from \citep{Chevalier.etal2013a} delivers that
+Thanks to our representation of co-kriging in terms of simple kriging
+with respect to generalized locations, a strightforward adaptation of
+the batch-sequential kriging update formulae from
+\citep{Chevalier.etal2013a} suggests that
 % 
@@
 \end{equation}
-where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
-$\lambda_{[n+1]}(\bm{x})$ that corresponds to the kriging weigths associated with the last $q_{n+1}$ responses when 
-predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
+where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional
+sub-vector extracted from $\lambda_{[n+1]}(\bm{x})$ that corresponds
+to the kriging weigths associated with the last $q_{n+1}$ responses
+when predicting at $\bm{x}$ relying on all measurements until batch
+$(n+1)$.
 %\text{th}$ batch.
 %, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
-Similarly, the updated co-kriging residual (cross-)covariance function then writes
+Similarly, the updated co-kriging residual (cross-)covariance function
+then becomes
 \begin{align}\label{eq:varCoK}
@@
 %
-Let us remark that, as noted in \citep{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
-extend to Universal Kriging in second-order settings and apply without Gaussian assumption.
-We will now see how the latter formulae are instrumental in deriving semi-analytical formulae for stepwise uncertainty reduction criteria for
-vector-valued random fields.
-
-\subsection{UQ on ESs of multivariate GRFs}\label{sec:set_uq}
-We now introduce quantities that allow UQ on the volume of the ES $\es$. Let $\mes$ be a 
-(locally finite, Borel) measure  on $\domain$. We want to investigate the probability distribution 
-of $\mes(\es)$ through its moments.
-Centred moments may be computed using Proposition~\ref{propo1} developed in the appendix. 
-In particular, the excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ is an integral of the EP
+As noted in \citep{Chevalier2015} in the case of scalar-valued fields,
+these update formulae naturally extend to Universal Kriging in
+second-order settings and apply without Gaussian assumptions. We will
+now see how the latter formulae are instrumental in deriving
+semi-analytical formulae for stepwise uncertainty reduction criteria
+for vector-valued random fields.
+
+\subsection{Uncertainty Quantification on ESs of multivariate GRFs}
+\label{sec:set_uq}
+
+We now introduce quantities that allow UQ on the volume of the ES
+$\es$. Let $\mes$ be a (locally finite, Borel) measure on
+$\domain$. We want to investigate the probability distribution of
+$\mes(\es)$ through its moments.  Centred moments may be computed
+using Proposition~\ref{propo1} developed in the appendix.  In
+particular, the excursion measure variance
+$\emv = \operatorname{Var}[\mes(\es)]$ is an integral of the excursion
+probabilities (EP)
+
 \begin{equation*}
@@
 \end{equation*}
-where $\varPhi_{\no}$ denotes the $\no$-variate Gaussian cumulative distribution function (CDF), which can be calculated using for instance the code of \cite{genz2009computation}.
-
-%
-Note that this quantity requires the solution of an integral over $\domain^2$. In 
-contrast, and still like in the scalar-valued case, the IBV of \cite{bect2019} involves 
-solely an integral on $\domain$ and can be expanded as 
+
+where $\varPhi_{\no}$ denotes the $\no$-variate Gaussian cumulative
+distribution function (CDF), which can be calculated, using for
+instance, the code of \cite{genz2009computation}.
+
+Note that this quantity requires the solution of an integral over
+$\domain^2$. In contrast, and as in the scalar-valued case, the IBV of
+\cite{bect2019} involves solely an integral on $\domain$ and can be
+expanded as
 \begin{equation*}
@@
 \label{sec:eibv}
-We compute the expected effect of the inclusion of new observations on the $\emv$ and $\ibv$ of the excursion set $\es$.
-Let us consider the same setting as in Eqn.
-\eqref{eq:meanCoK} and \eqref{eq:varCoK}, and let
-$\currentExp{.}$ and $\currentProba{.}$ denote conditional expectation
-and probability conditional on the first $n$ batches of observations, respectively.
-We want to compute the effect of including a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and $\ibv$.
-We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.
+
+We compute the expected effect of the inclusion of new observations on
+the $\emv$ and $\ibv$ of the excursion set $\es$. Let us consider the
+same setting as in Eqn. \eqref{eq:meanCoK} and \eqref{eq:varCoK}, and
+let $\currentExp{.}$ and $\currentProba{.}$ denote conditional
+expectation and probability conditional on the first $n$ batches of
+observations, respectively. We want to compute the effect of including
+a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and
+$\ibv$. We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the
+conditional law $\mathbb{P}_{\stage}$.
 
@@
 
-In order to study the effect of the inclusion of a new data point, we let
-$
-\currentIBV(\bm{x}; \bm{y})
-$
-denote the expected IBV under the current law of the field, conditioned on observing $\bm{y}$ at $\bm{x}$ (generalized, possibly batch observation).
-The expected effect of a new observation on the IBV is then captured by
+In order to study the effect of the inclusion of a new data point, we
+let $ \currentIBV(\bm{x}; \bm{y}) $ denote the expected IBV under the
+current law of the field \kc{Is this a legitimate expression 'law of
+  the field'?}, conditioned on observing $\bm{y}$ at $\bm{x}$
+(generalized, possibly batch observation). The expected effect of a
+new observation on the IBV is then shown by
 \begin{equation}\label{def:eibv}
@@
 \end{equation}
-where $\bm{Y}$ is distributed according to the current law of $Z_{\bm{x}}$ and with independent noise having covariance matrix $\Delta_n$.
-
-We next present a result that allows efficient computation of $\EIBV$ as an integral of CDFs of multivariate gaussians.
-This will prove useful when designing sequential expected uncertainty reduction strategies.
+where $\bm{Y}$ is distributed according to the current law of
+$Z_{\bm{x}}$ and with independent noise having covariance matrix
+$\Delta_n$.
+
+We next present a result that allows efficient computation of $\EIBV$
+as an integral of CDFs of multivariate gaussians. This will prove
+useful when designing sequential expected uncertainty reduction
+strategies.
 
@@
 
-This integrand finally boils down to an evaluation of the $2\no$-dimensional Gaussian CDF by
-Proposition~\ref{propo3}. Details are omitted here for brevity as we did not implement further
-this criterion that requires numerical integration over $\domain^2$.
+This integrand finally boils down to an evaluation of the
+$2\no$-dimensional Gaussian CDF by Proposition~\ref{propo3}. Details
+are omitted here for brevity as we do not implement this criterion
+further and which requires numerical integration over $\domain^2$.
 \end{propo}
 
-We remark that 
-Propositions \ref{propo_eibv} and \ref{propo_emv} are twofold generalizations of  results from \cite{chevalier2014fast}: they extend previous results to the multivariate setting and also allow for the inclusion of batch or heterotopic observations through the concept of generalized locations.
-A key element for the understanding of the propositions is that the conditional co-Kriging means entering in the EPs depend linearly on (batch) observations.
-The conditional equality expressions thus become linear combinations of Gaussian variables whose mean and covariance are easily calculated.
-Related closed-form solutions have been noted in similar contexts \citep{bhattacharjya2013value,stroh}, but not generalized to our situation with
-random sets for vector-valued GRFs.
+We remark that Propositions \ref{propo_eibv} and \ref{propo_emv} are
+twofold generalizations of results from \cite{chevalier2014fast}: they
+extend previous results to the multivariate setting and also allow for
+the inclusion of batch or heterotopic observations through the concept
+of generalized locations.  A key element for understanding of these
+propositions is that the conditional co-Kriging means entering in the
+EPs depend linearly on (batch) observations. The conditional equality
+expressions thus become linear combinations of Gaussian variables
+whose mean and covariance are easily calculated.  Related closed-form
+solutions have been noted in similar contexts
+\citep{bhattacharjya2013value,stroh}, but not generalized to our
+situation with random sets for vector-valued GRFs.
 
@@
 
-We illustrate the EBV associated to different designs on a simple bivariate example.
-The example mimics our river plume application and hence the first and second component of the random field will be called \textit{temperature} and \textit{salinity} for illustrative purpose.
-Effects of the hyperparameters of the GRF prior on the excursion probabilities will also be studied.
-For simplicity, we begin with a \textit{pointwise} example, considering a single bivariate gaussian distribution (no spatiality).
+We illustrate the EBV associated with different designs on a simple
+bivariate example. This mimics our river plume application and hence
+the first and second component of the random field will be called
+\textit{temperature} and \textit{salinity} for illustrative
+purpose. Effects of the hyperparameters of the GRF prior on the
+excursion probabilities will also be studied. For simplicity, we begin
+with a \textit{pointwise} example, considering a single bivariate
+gaussian distribution (i.e. no spatiality).
 
@@
 
-Say we want to study the excursion probability of a bivariate gaussian above some thresholds, where the thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ g/kg for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.
-
-% To BE COMMENTED (TEMPORARILY?)
-Fig. \ref{illus_bivarDens} shows contour plots of three different
-densities with increasing correlation $\gamma$ between temperature and
-salinity. 
-\begin{figure}[h!] \centering
+Say we want to study the excursion probability of a bivariate gaussian
+above some threshold, where the thresholds are set equal to the mean;
+$\mu_1=t_1=5^o C$ for temperature and $\mu_2=t_2=30$ g/kg for
+salinity, and we play with the temperature and salinity correlation
+and variances to study the effect on the EP and expected Bernoulli
+variance.
+
+\begin{figure}[!b] 
+\centering
 \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
@@
 \end{figure}
-The displayed densities have unit standard deviations for both
-temperature and salinity, but we also study the effect of doubling the
-standard deviations.
-% END TO BE COMMENTED (TEMPORARILY?)
+
+Fig. \ref{illus_bivarDens} shows contour plots of three different
+densities with increasing correlation $\gamma$ between temperature and
+salinity. The displayed densities have unit standard deviations for
+both temperature and salinity, but we also study the effect of
+doubling the standard deviations.
 
@@
 \ref{illus_bivarDens}. The EPs increase with the correlation as there
-is a strong tendency to have concurrently low temperature and salinity. The Bernoulli variance is similarly large for high
+is a strong tendency to have concurrently low temperature and
+salinity. The Bernoulli variance is similarly large for high
 correlations. EPs and Bernoulli variances are the same for standard
@@
 
-\begin{table}[!h] \centering \caption{EP and Bernoulli variance for
+\begin{table}[!t] \centering \caption{EP and Bernoulli variance for
     different correlations and variances (top rows), and expected
-    Bernoulli variances for both temperature and salinity data $\by$ and 
-    temperature $y_2$ (bottom rows).}
+    Bernoulli variances for both temperature and salinity data $\by$
+    and temperature $y_2$ (bottom rows). \kc{$\mu_1$  or $\by$? And
+      whatever happened to make the terms more accessible with $t$ and
+    $s$ as we had before?}}
   \begin{tabular}{c|ccc|ccc}
@@
 $p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
-EIBV, Temp and Salinity data & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
-EIBV, Temperature data only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
+EIBV, Temperature and Salinity & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
+EIBV, Temperature only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
 \hline
@@
 
-Table \ref{tab:sim_rhoab} (bottom two rows) shows results of expected
-Bernoulli variance calculations. This is presented for a design
-gathering both data types, and for a design with temperature
-measurements alone. When both data are gathered, the measurement model is
-$(Y_1,Y_2)^t=(Z_1,Z_2)^t+\bepsilon$, with $\bepsilon \sim N(0,0.5^2I_2)$, while $Y_1=Z_1+\epsilon$, $\epsilon \sim N(0,0.5^2)$ when only temperature is measured.
-For this illustration, Table \ref{tab:sim_rhoab} shows that the
-expected Bernoulli variance gets lower with larger standard deviations
-$\sigma_1$ and $\sigma_2$ (right columns). The reduction of Bernoulli
-variance is largest for the cases with high correlation
-$\gamma$. Albeit smaller, there is also uncertainty reduction when
-only temperature is measured (bottom row), especially when temperature
-and salinity are highly correlated. When correlation is low
-($\gamma=0.2$), there is little information about salinity in the
-temperature data, and therefore less uncertainty reduction. In an
-application with fresh cold water from a river source, the temperature
-and salinity variables will not only be interdependent, but will also
-likely show dependence in the spatial dimension. This in turn will
-impact the design criteria when we evaluate the information measure by
-integrating over several locations (Section \ref{sec:simulations}).
+The bottom two rows of Table \ref{tab:sim_rhoab} show results of
+expected Bernoulli variance. This is presented for a design gathering
+both data types, and for a design with temperature measurements
+alone. When both data are gathered, the measurement model is
+$(Y_1,Y_2)^t=(Z_1,Z_2)^t+\bepsilon$, with
+$\bepsilon \sim N(0,0.5^2I_2)$, while $Y_1=Z_1+\epsilon$,
+$\epsilon \sim N(0,0.5^2)$ when only temperature is measured.  For
+this illustration, Table \ref{tab:sim_rhoab} shows that the expected
+Bernoulli variance gets lower with larger standard deviations
+$\sigma_1$ and $\sigma_2$. The reduction of Bernoulli variance is
+largest for the cases with high correlation $\gamma$. Albeit smaller,
+there is also uncertainty reduction when only temperature is measured
+(bottom row), especially when temperature and salinity are highly
+correlated. When correlation is low ($\gamma=0.2$), there is little
+information about salinity in the temperature data, and therefore less
+uncertainty reduction. In an application with fresh cold water from a
+river source, the temperature and salinity variables will not only be
+interdependent, but will also likely show dependence in the spatial
+dimension. This in turn will impact the design criteria when we
+evaluate the information measure by integrating over several locations
+(Section \ref{sec:simulations}).
 
 \subsubsection{Including Spatiality}
-We now turn to an example involving a full-fledged gaussian random field (GRF). The class of GRF model we will consider in the article will generally have a linear trend
+
+We now turn to an example involving a full-fledged gaussian random
+field (GRF). The class of GRF model we will consider will generally
+have a linear trend
+
 \begin{align*}
@@
 \end{align*}
-where $k(., .)$ is one of the traditional spatial covariance kernel (exponential, Mat\'{e}rn, ...) and $\gamma$ defines the cross-covariance structure.
-
-In the accompanying Python examples taking place within the MESLAS toolbox (URL), these modeling assumptions can be
-generalized to anisotropic covariance and changing variance levels across the spatial domain. Anisotropy and
+where $k(., .)$ is one of the traditional spatial covariance kernel
+(exponential, Mat\'{e}rn, ...) and $\gamma$ defines the
+cross-covariance structure.
+
+In the accompanying \kc{Where? On Github? If so say so.} Python
+examples taking place within the MESLAS toolbox (URL \kc{Note}), these
+modeling assumptions can be generalized to anisotropic covariance and
+changing variance levels across the spatial domain. Anisotropy and
 non-stationary variance are both relevant for the setting with river
@@
 
-In the rest of this section, we will consider a GRF with mean and covariance structure as above and parameters
+In the rest of this section, we will consider a GRF with mean and
+covariance structure as above and parameters
 \begin{align*}
@@
 
-We now study how the expected Bernoulli variance eq.\eqref{def:eibv} associated to data collection at a point changes if only one of the two components of the field is observed.
-We first draw a realization of the GRF defined above and use it as ground-truth to mimick the real data-collection process.
-A first bunch of observations are done at the locations depicted in grey, and the data is used to update the GRF model.
-We then consider the green triangle as a potential next observation locations and plot the expected Bernoulli variance reduction (at each grid point) that would result from observing only one component of the field (temperature or salinity), or both at that point, see Fig.\ref{fig:ebv_comp}.
+We now study how the expected Bernoulli variance eq.\eqref{def:eibv}
+associated with data collection at a point changes if only one of the
+two components of the field is observed. We first draw a realization
+of the GRF defined above and use it as ground-truth to mimick the real
+data-collection process. A first set of observations are done at the
+locations depicted in grey (see Fig.~\ref{fig:ebv_comp}), and the data
+is used to update the GRF model. We then consider the green triangle
+as a potential next observation location and plot the expected
+Bernoulli variance reduction (at each grid point \kc{I don't recall we
+  have been explicit about grids before this mention. It might be
+  useful to recall that here.}) that would result from observing only
+one component of the field (temperature or salinity), or both at that
+point.
 
@@
   \includegraphics[width=0.99\textwidth]{Figures/ebv_comp_3.png}
-  \caption{Pointwise Bernoulli variance reduction for observation of a single or both components of the random field at one location.
-Data collection locations in green.
-True excursion set in red.
-Places where only one response is above threshold are depicted in pink}
+  \caption{Pointwise Bernoulli variance reduction for observation of a
+    single or both components of the random field at one
+    location. Data collection locations in green. True excursion set
+    in red. Places where only one response is above threshold are
+    depicted in pink}
 \label{fig:ebv_comp}
@@
 
-Note that plotting the Bernoulli variance reduction at each point might also be used to compare different data collection plans.
-For example, Fig.
-\ref{fig:ebv_north_vs_east} shows the expected Bernoulli variance reduction associated to a data collection plan along a straight upwards line (static north) and the one associated to a straight horizontal line (static east).
-Both expectations are computed according to the a priori distribution of the GRF (i.e.
-no observations have been included yet).
+Note that plotting the Bernoulli variance reduction at each point
+might also be used to compare different data collection plans. For
+example, Fig. \ref{fig:ebv_north_vs_east} shows the expected Bernoulli
+variance reduction associated with a data collection plan along a
+vertical line (static north) and one associated with a horizontal
+(static east). Both expectations are computed according to the
+a-priori distribution of the GRF (i.e. no observations have been
+included yet).
 
@@
   \includegraphics[width=0.99\textwidth]{Figures/ebv_north_vs_east_2.png}
-  \caption{Pointwise Bernoulli variance reduction for two different observation plans.
-Data collection locations in green.
-True excursion set is in red.
-Places where only one response is above threshold are depicted in pink.}
+  \caption{Pointwise Bernoulli variance reduction for two different
+    observation plans. Data collection locations in green. True
+    excursion set is in red. Places where only one response is above
+    threshold are depicted in pink.}
 \label{fig:ebv_north_vs_east}
@@
 
-%\begin{remark}
-In all the computed examples, the spatial domain $\domain$ has been discretized to a set of $N$ grid
-locations $\mathcal{M}_g = \{\x_i, i=1,\ldots,N \}$, where each cell
-has area $\delta$; the same grid is used for the waypoint graph for
-possible design locations. The EIBV is approximated by sums over all
-grid cells.
-%\end{remark}
+In all the computed examples, the spatial domain $\domain$ has been
+discretized to a set of $N$ grid locations
+$\mathcal{M}_g = \{\x_i, i=1,\ldots,N \}$, where each cell has area
+$\delta$; the same grid is used for the waypoint graph for possible
+design locations. The EIBV is approximated by sums over all grid
+cells. \kc{This should be promoted up a couple of paras and would
+  address my last comment above.}
 
@@
 
-The goal of this section is to present sequential data collection strategies that aim at reducing the expected uncertainty on the target ES $\es$.
+The goal of this section is to present sequential data collection
+strategies that aim at reducing the expected uncertainty on the target
+ES $\es$.
 
@@
 
-In a sequential point of view, $n$ data collection steps have already been performed and one wants to choose what data to collect next.
-After each step, the GRF model is updated using the co-Kriging equations \ref{eq:meanCoK} and \ref{eq:varCoK}.
-The design evaluations are based on the conditional expectation
-$\currentExp{.}$ from the law of the field conditional on the data gathered up to (and including) stage $n$.
-
-Note that the type of data collected at each stage can be of various type (all components of the field at a single location, only some components at a subset of the selected locations, etc.) since the concept of \textit{generalized location} allows the co-Kriging equations to handle all those case by putting them on equal footage.
-In general, a design strategy must choose the spatial location as well as the components to observe (heterotopic case), or where several observations are allowed at each stage (batch case).
-For the case with an AUV exploring the river plume, we limit scope to choosing one of the neighboring spatial location (waypoints) at each stage, and all components (temperature and salinity) of the field are observed there (isotopic case).
-The candidate points at this stage are denoted $\candidates$ as defined from the 6 directions (apart from edges) in the waypoint graph.
-The set $\candidates$ depends on the current location, but for readability we suppress this in the notation.
-
-The mathematical expression for the optimal designs in this sequential setting involves a series
-of intermixed maximizations over designs and integrals over data. In
-practice, the optimal solution is intractable because of the enormous
-growth over stages (see e.g. \cite{sucar2015probabilistic} and
-\cite{powell2016perspectives}).  Instead, we outline heuristic
-strategies.
+From a sequential point of view, $n$ data collection steps have
+already been performed and one wants to choose what data to collect
+next. After each step, the GRF model is updated using co-Kriging
+equations \ref{eq:meanCoK} and \ref{eq:varCoK}. The design evaluations
+are based on the conditional expectation $\currentExp{.}$ from the law
+of the field, conditional on the data gathered up to (and including)
+stage $n$.
+
+Note that the type of data collected at each stage can be of various
+type (all components of the field at a single location, only some
+components at a subset of the selected locations, etc.) since the
+concept of \textit{generalized location} allows the co-Kriging
+equations to handle all those case by putting them on equal footing.
+In general, a design strategy must choose the spatial location as well
+as the components to observe (heterotopic), or where several
+observations are allowed at each stage (batch).  For the case with an
+AUV exploring the river plume, we limit our scope to choosing one of
+the neighboring spatial location (waypoints) at each stage, and all
+components (temperature and salinity) of the field are observed there
+(isotopic). The candidate points at this stage are denoted
+$\candidates$ as defined from the 6 directions (apart from edges) in
+the waypoint graph (see Fig.~\ref{fig:wp_graph_a} for an example). The
+set $\candidates$ depends on the current location, but for readability
+we suppress this in the notation.
+
+The mathematical expression for the optimal design in this sequential
+setting involves a series of intermixed maximizations over designs and
+integrals over data. In practice, the optimal solution is intractable
+because of the enormous growth over stages (see
+e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}).
+Instead, we outline heuristic strategies.
 
@@
 
-A very simple heuristic for adaptive sampling is to observe $Z$ at the location in $\candidates$ with current EP closest to $\frac{1}{2}$.
-While extremely easy to implement, this strategy can lead to spending many stages in boundary regions regardless of the possible effect of sampling at the considered point for the future conditional distribution of $Z$.
-The strategy does not account for the expected reduction in uncertainty, and it is not considering the integrated effect over other locations.
+A simple heuristic for adaptive sampling is to observe $Z$ at the
+location in $\candidates$ with current EP closest to
+$\frac{1}{2}$. While easy to implement, this strategy can lead to
+spending many stages in boundary regions regardless of the possible
+effect of sampling at the considered point for the future conditional
+distribution of $Z$. The strategy does not account for the expected
+reduction in uncertainty, and it does not consider having an
+integrated effect over other locations.
 
@@
 The myopic (greedy) strategy which we present here is optimal if we
-imagine taking only one more stage of measurements. It does not anticipate what the subsequent
-designs might offer beyond the first stage.
-Based on the currently available data the myopic strategy selects the location that leads to the biggest reduction in EIBV:
+imagine taking only one more stage of measurements; it does not
+anticipate what the subsequent designs might offer beyond the first
+stage.  Based on the currently available data the myopic strategy
+selects the location that leads to the biggest reduction in EIBV:
 \begin{criterion}[Myopic]
-The next observation location $\spatloc_{\stage + 1}$ is chosen among the minimizers in $\candidates$ of the criterion:
+The next observation location $\spatloc_{\stage + 1}$ is chosen among
+the minimizers in $\candidates$ of the criterion: 
 \begin{equation}\label{critSEQ}
@@
 
-The EIBV is efficiently computed for each of the candidate points $\candidates$ using Proposition \ref{propo2}.
-Once the best location has been selected, a stage of observation is performed there, the GRF model is updated, yielding a conditional law $\mathbb{P}_{\stage + 1}$ and the process is repeated.
-
-Even though this myopic strategy is non-anticipative, it still gives a
-reasonable approach for creating designs in many
-applications. Moreover, it can be implemented in a way that is not too demanding on computational power, making it well-suited for embedding on the AUV.
+The EIBV is efficiently computed for each of the candidate points
+$\candidates$ using Proposition \ref{propo2}. Once the best location
+has been selected, a stage of observation is performed and the GRF
+model is updated, yielding a conditional law $\mathbb{P}_{\stage + 1}$
+after which the process is repeated.
+
+Even though this myopic strategy is non-anticipatory, it still
+provides a reasonable approach for creating designs in many
+applications. Moreover, it can be implemented without any demands on
+computational power, making it well-suited for embedding on an AUV.
 
@@
 
-We now extend the myopic strategy to a look-ahead strategy. We consider two stages of
-measurements, and the strategy is then optimal in the sense that it accounts consistently for the expectations and minimizations in these two stages, but still without including any planning beyond those two steps.
-
-Assume data have been collected up to stage $n$. The principle of two-steps lookahead
-strategies is to select the next observation location $\spatloc_{\stage + 1}$ that yields the biggest reduction in EIBV if we were to (optimally) add one more observation after the one at $\spatloc_{\stage + 1}$.
-In order to formalize this concept, we must extend the notation for EIBV in the future (after observation $\stage + 1$ has been made).
-We let $\currentEIBV(\cdot; u, y)$ to denote the EIBV where expectations are taken conditional on the data available at $n$ and on an additional observation $y$ at $u$.
-
-  \begin{criterion}[2-steps lookahead]
+We now extend the myopic strategy by considering two stages of
+measurements, which is optimal in that it accounts consistently for
+the expectations and minimizations in these two stages, but
+nevertheless without including any planning beyond those two steps.
+
+Assume data have been collected up to stage $n$. The principle of
+two-step lookahead is to select the next observation location
+$\spatloc_{\stage + 1}$ that yields the biggest reduction in EIBV if
+we were to (optimally) add one more observation after the one at
+$\spatloc_{\stage + 1}$. In order to formalize this concept, we must
+extend the notation for EIBV in the future (after observation
+$\stage + 1$ has been made). We let $\currentEIBV(\cdot; u, y)$ to
+denote the EIBV where expectations are taken conditional on the data
+available at $n$ and on an additional observation $y$ at $u$.
+
+\begin{criterion}[2-step lookahead]
       The next observation location $\spatloc_{\stage + 1}$ is chosen among the minimizers in $\candidates$ of the criterion
@@
       \end{align}
-  where $Y$ is a random variable distributed according to the conditional
-  law of $\gp_{\spatloc}$ at step $\stage$. The dependence of the set of candidates on the current location has been made explicit for the second stage of measurements.
+      where $Y$ is a random variable distributed according to the
+      conditional law of $\gp_{\spatloc}$ at step $\stage$ with the
+      dependence of the set of candidates on the current location
+      having been made explicit for the second stage of measurements.
     \end{criterion}
 
-In a practical setting, the first expectation can be computed by Monte
-Carlo sampling of data $Y$ from its conditional
-distribution. For each of these data samples, the second expectation
-is solved using the closed-form expressions for EIBV provided by Proposition \ref{propo2}, now with conditioning on the first stage data already going into the co-Kriging updating equations.
+    In a practical setting, the first expectation can be computed by
+    Monte Carlo sampling of data $Y$ from its conditional
+    distribution. For each of these data samples, the second
+    expectation is solved using the closed-form expressions for EIBV
+    provided by Proposition \ref{propo2}, now with conditioning on the
+    first stage data already going into the co-Kriging updating
+    equations.
 
@@
 
-\begin{figure}[!h] 
+\begin{figure}[!b] 
 \centering 
@@
 Each strategy is conducted on an equilateral grid as shown in
-Fig. \ref{fig:wp_graph}. The AUV starts at the
-center East-West coordinate at the southern end of the domain (node 53). The
-AUV moves along edges in the waypoint graph while collecting data which are assimilated into the GRF model onboard. This is used in the
-evaluation of the next node to sample.
-The procedure is run for $10$ stages. A total of 100 replicate simulations were conducted with all
-strategies.
+Fig. \ref{fig:wp_graph}. The AUV starts at the center East-West
+coordinate at the southern end of the domain (node 53). It then moves
+along edges in the waypoint graph while collecting data which are
+assimilated into the onboard GRF model. This is used in the evaluation
+of the next node to sample.  The procedure is run for $10$ stages. A
+total of 100 replicate simulations were conducted with all strategies.
 
@@
 
-The results of the replicate runs are shown in Fig. \ref{fig:sim_results}, where
-the different criteria are plotted as a function of survey
-distance. Fig. \ref{fig:avg_ev} shows the resulting drop in realized IBV for
-each of the six strategies. IBV reduction occurs most under the
-\textit{myopic} and \textit{look-ahead} strategies, each performing
-almost equally; this is expected as the two criteria
-(Eq. \eqref{critSEQ} and \eqref{critLA}) are sensitive to differences
-in IBV. The \textit{static\_north} design also does well here because
-the path is parallel to the boundary between the water masses.
-
-\begin{figure}[h!]
+The results of the replicate runs are shown in
+Fig. \ref{fig:sim_results}, where the different criteria are plotted
+as a function of survey distance. Fig. \ref{fig:avg_ev} shows the
+resulting drop in realized IBV for each of the six strategies. IBV
+reduction occurs most under the \textit{myopic} and
+\textit{look-ahead} strategies, each performing almost equally; this
+is expected as the two criteria (Eq. \eqref{critSEQ} and
+\eqref{critLA}) are sensitive to differences in IBV. The
+\textit{static\_north} design also does well here because the path is
+parallel to the boundary between the water masses.
+
+\begin{figure}[!h]
   \centering
@@
   \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
-  \hfill
-  \subfigure[RMSE between estimated field and truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
-  \hfill 
-  \subfigure[Explained variance $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
-  \hfill 
-  \subfigure[Computational time for inference (some of the lines overlap).]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}} 
+  \hfill \subfigure[RMSE between estimated field and
+  truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
+  \hfill \subfigure[Explained variance
+  $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
+  \hfill \subfigure[Computational time for inference (some lines
+  overlap \kc{Which ones? As figure with 6 keys, one sees only 4. This
+    needs to be fixed or explained -- this was a repeat from an
+    earlier
+    feedback.}).]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}}
 \caption{Simulation results from 100 replicate simulations for 10

### 2020-06-23 main_new.tex
updates

@@
 resulted in significant contributions to environmental monitoring and
-sampling in the ocean (Fig. \ref{fig:envir1}). In particular,
+sampling in the ocean (Fig.~\ref{fig:envir1}). In particular,
 autonomous underwater vehicles (AUVs) have advanced the state of data
@@
 pertaining to using an AUV for spatial characterization of a frontal
-system generated by a river plume. Fig. \ref{fig:nidelven} shows the
+system generated by a river plume. Fig.~\ref{fig:nidelven} shows the
 survey area in Trondheim, Norway, where cold freshwater enters from
@@
 
-Fig. \ref{illus_bivarDens} shows contour plots of three different
+Fig.~\ref{illus_bivarDens} shows contour plots of three different
 densities with increasing correlation $\gamma$ between temperature and
@@
 Table \ref{tab:sim_rhoab} shows the initial EPs and the associated
-Bernoulli variance (second row) for the examples indicated in Fig.
-\ref{illus_bivarDens}. The EPs increase with the correlation as there
-is a strong tendency to have concurrently low temperature and
+Bernoulli variance (second row) for the examples indicated in
+Fig.~\ref{illus_bivarDens}. The EPs increase with the correlation as
+there is a strong tendency to have concurrently low temperature and
 salinity. The Bernoulli variance is similarly large for high
@@
 and the spatial covariance is given by a Mat\'{e}rn 3/2 kernel with unit variance and range parameter $\lambda=0.5$.
-One realization of this GRF is shown in Fig. \ref{example_excu}.
+One realization of this GRF is shown in Fig.~\ref{example_excu}.
 
@@
 Each strategy is conducted on an equilateral grid as shown in
-Fig. \ref{fig:wp_graph}. The AUV starts at the center East-West
+Fig.~\ref{fig:wp_graph}. The AUV starts at the center East-West
 coordinate at the southern end of the domain (node 53). It then moves
@@
 The results of the replicate runs are shown in
-Fig. \ref{fig:sim_results}, where the different criteria are plotted
-as a function of survey distance. Fig. \ref{fig:avg_ev} shows the
+Fig.~\ref{fig:sim_results}, where the different criteria are plotted
+as a function of survey distance. Fig.~\ref{fig:avg_ev} shows the
 resulting drop in realized IBV for each of the six strategies. IBV
@@
 
-Fig. \ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop
+Fig.~\ref{fig:avg_rmse} and \ref{fig:avg_r2} show the resulting drop
 in RMSE and increase in explained variance, respectively. Both
@@
 but some of the \textit{static\_east} and \textit{static\_zigzag} also
-achieve good results because they are cover large
-parts of the domain without re-visitation. Sequential strategies
-targeting IBV will sometimes not reach similar coverage, as
-interesting data may draw the AUV path into twists and turns. There is
-relatively large variety in the replicate results as indicated by the
-vertical lines. Nevertheless, the ordering of strategies is similar.
-
-Fig. \ref{fig:avg_time} shows the computational effort: the
+achieve good results because they cover large parts of the domain
+without re-visitation. Sequential strategies targeting IBV will
+sometimes not reach similar coverage, as interesting data may draw the
+AUV into twists and turns. There is a relatively large variety in the
+replicate results as indicated by the vertical lines. Nevertheless,
+the ordering of strategies is similar.
+
+Fig.~\ref{fig:avg_time} shows the computational effort: the
 \textit{naive} strategy is on par with the static designs, while the
-\textit{myopic} strategy is slower because it evaluates expected values for all candidate directions at the waypoints.
-But it is still able to do so in reasonable time, which enables real-world applicability.
-The \textit{look-ahead} is even
-slower, reaching levels that are nearly impractical for execution on a
-vehicle. Some pruning of the graph is performed to improve the
-performance, such as ruling out repeated visitations. Further pruning of branches or
-inclusion of other heuristics could be included for better
-performance. Then again, the inclusion of such heuristics is likely a
-contributing factor for the \textit{look-ahead} strategy failing to
-outperform the \textit{myopic} strategy.
+\textit{myopic} strategy is slower because it evaluates expected
+values for all candidate directions at the waypoints. But it is still
+able to do so in reasonable time, which allows for real-world
+applicability. The \textit{look-ahead} is slower, reaching levels that
+are nearly impractical for execution on a vehicle. Some pruning of the
+graph is performed to improve the performance, such as ruling out
+repeated visitations. Further pruning of branches or inclusion of
+other heuristics could be included for better performance. Then again,
+the inclusion of such heuristics is likely a contributing factor for
+the \textit{look-ahead} strategy failing to outperform the
+\textit{myopic} strategy.
 
@@
 substantially better than the \textit{myopic} design only for very
-small initial standard deviation or very large spatial correlation
+small initial standard deviations or very large spatial correlation
 range. 
@@
 %spatial correlation range. 
-We also ran simulation studies with only
-temperature data, and for realistic correlation levels between
-temperature and salinity, the IBV results are not much worse when only
-temperature data are available. In addition to the comparison made in
-Table \ref{tab:sim_rhoab}, the current setting includes spatial
+We also ran simulation studies with only temperature data, and for
+realistic correlation levels between temperature and salinity, the IBV
+results are not much worse when only temperature data are
+available. In addition to the comparison made in
+Table~\ref{tab:sim_rhoab}, the current setting includes spatial
 correlation and this likely reduces the added effect of having
@@
 river plume with an AUV. The experiment was performed in Trondheim,
-Norway, surveying the Nidelva river (Fig. \ref{fig:nidelven}). The
+Norway, surveying the Nidelva river (Fig.~\ref{fig:nidelven}). The
 experiments were conducted in late Spring 2019, when there is still
@@
 focused along the frontal zone that runs more or less parallel to the
-eastern shore as noted in Fig. \ref{fig:nidelven}.
+eastern shore as noted in Fig.~\ref{fig:nidelven}.
 
@@
 
-\begin{figure}[!h]
+\begin{figure}[!b]
   \centering
@@
 
-Fig. \ref{fig:parest} summarizes diagnostic plots of this
-analysis. Fig. \ref{fig:parest_a} shows a cross-plot of temperature
+Fig.~\ref{fig:parest} summarizes diagnostic plots of this
+analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
 and salinity residuals after the westerly trends in both salinity and
 temperature are subtracted from the data. This scatter-plot of joint
-residuals indicates larger variability in salinity than
-temperature, and a positive correlation ($0.5$) between the two
-variables. Based on the fitted bivariate Gaussian model (ellipses in
-Fig. \ref{fig:parest_a}), we can compute the modeled quadratic form of
+residuals indicates larger variability in salinity than temperature,
+and a positive correlation ($0.5$) between the two variables. Based on
+the fitted bivariate Gaussian model (ellipses in
+Fig.~\ref{fig:parest_a}), we can compute the modeled quadratic form of
 the residuals, and if the model is adequate they should be
-approximately $\chi^2_2$ distributed. Fig. \ref{fig:parest_b} shows
-the empirical CDF of these
-quadratic forms (solid) together with the theoretical CDF of the
-$\chi^2_2$ distribution. The modeled and theoretical curves are very
-similar, which indicates that the Gaussian model fits reasonably
-well. Fig. \ref{fig:parest_c} shows the empirical variogram of the
-scaled residuals for temperature and salinity. The decay is similar
-for the two, and seems to be negligible after about $150$ m.
-The working assumption of a separable covariance function is hence not unreasonable.
-
-Based on the analysis in Fig. \ref{fig:parest}, the resulting
+approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
+the empirical CDF of these quadratic forms (solid) together with the
+theoretical CDF of the $\chi^2_2$ distribution. The modeled and
+theoretical curves are very similar, which indicates that the Gaussian
+model fits reasonably well. Fig.~\ref{fig:parest_c} shows the
+empirical variogram of the scaled residuals for temperature and
+salinity. The decay is similar for the two, and seems to be negligible
+after about $150$ m.  The working assumption of a separable covariance
+function is hence not unreasonable.
+
+Based on the analysis in Fig.~\ref{fig:parest}, the resulting
 parameters are given in Table \ref{tab:experiment_param}. The
@@
 
-A Light AUV \citep{sousa2012lauv}
-(Fig. \ref{fig:lauv}) equipped with a 16 Hz Seabird Fastcat-49
-conductivity, temperature, and depth (CTD) sensor was used to provide
-salinity and temperature measurements.
-We assume that the measurements are conditionally independent because the salinity is extracted from the conductivity instrument which is different from the temperature instrument.
-We specify variance $0.25^2$ for both errors, which is based on a middle ground between the nugget effect in the empirical variogram and the sensor specifications.
-
-The AUV is a powered untethered platform that
-operates at $1$-$3$ m/s in the upper water column. The AUV uses a multicore GPU NVIDIA
-Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared
-cache, and 4 GB of 1600 MHz DRAM) for computation onboard.
-The sampling algorithm was built on top of the autonomous agent framework
-Teleo-Reactive EXecutive (\textit{T-REX})
-\citep{py10,Rajan12,Rajan12b}.
+A Light AUV \citep{sousa2012lauv} (Fig.~\ref{fig:lauv}) equipped with
+a 16 Hz Seabird Fastcat-49 conductivity, temperature, and depth (CTD)
+sensor was used to provide salinity and temperature measurements.  The
+AUV is a powered untethered platform that operates at $1$-$3$ m/s in
+the upper water column. It has a multicore GPU NVIDIA Jetson TX1
+(quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4
+GB of 1600 MHz DRAM) for computation onboard.  The sampling algorithm
+was built on top of the autonomous Teleo-Reactive EXecutive
+(\textit{T-REX}) framework \citep{py10,Rajan12,Rajan12b}. We assume
+that the measurements are conditionally independent because the
+salinity is extracted from the conductivity sensor which is different
+from the temperature sensor. We specify variance $0.25^2$ for both
+errors, which is based on a middle ground between the nugget effect in
+the empirical variogram and the sensor specifications.
 
@@
 
-The AUV was running a
-\textit{myopic} strategy to decide between sampling locations on the waypoint graph distributed over an equilateral grid, as shown in the grey-colored lattice in Fig. \ref{fig:map}. 
-At each stage, it takes the AUV about 30
-seconds to evaluate the EIBV for all the possible waypoint
-alternatives.
-The AUV was set to start in the south-center part of the waypoint
-graph. A survey was set to take
-approximately 40 minutes, visiting 15 waypoints on the grid, with the
-AUV running near the surface to capture the plume. On its path from
-one waypoint to the next, the AUV collected data with an update
-frequency of 30 seconds, giving three measurements per batch in the updates at each stage.
+The AUV was running a \textit{myopic} strategy to decide between
+sampling locations on the waypoint graph distributed over an
+equilateral grid, as shown in the grey-colored lattice in
+Fig.~\ref{fig:map}.  At each stage, it takes the AUV about 30 seconds
+to evaluate the EIBV for all the possible waypoint alternatives.  It
+was set to start in the south-center part of the waypoint graph. A
+survey was set to take approximately 40 minutes, visiting 15 waypoints
+on the grid, with the vehicle running near the surface to capture the
+plume. On its path from one waypoint to the next, the AUV collects
+data with an update frequency of 30 seconds, giving three measurements
+per batch in the updates at each stage.
 
@@
 \subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv4_es_sal_ep.pdf}\label{fig:res2}}
-
 %\subfigure[ES for Survey 1]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
@@
 
-Two survey missions (1 and 2), were run successively, with a short break in between. The resulting path of the
-selected waypoints are shown in the map in Fig. \ref{fig:map}, both
-within the expected frontal region (shaded pink). The recorded
-temperatures are shown as colored trails in Fig. \ref{fig:res_both},
-clearly indicating the temperature difference between fjord and
-riverine waters. The salinity data are then shown separately, overlaid
-with the estimated EP for each survey in Fig. \ref{fig:res1} and
-\ref{fig:res2}.
+Two survey missions (1 and 2), were run successively, with a short
+break in between. The resulting path of the selected waypoints are
+shown in the map in Fig.~\ref{fig:map}, both within the expected
+frontal region (shaded pink). The recorded temperatures are shown as
+colored trails in Fig.~\ref{fig:res_both}, clearly indicating the
+temperature difference between fjord and riverine waters. The salinity
+data are then shown separately, overlaid with the estimated EP for
+each survey in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2}.
 
@@
 crossing the frontal boundary multiple times. As conditions changed
-slightly between the two surveys, the resulting path (after waypoint
-5) is shown to deviate. Survey 1 continued northwards, tracking the
-north-eastern portion of the front, while Survey 2 turned west,
-mapping the south-western region.
+slightly between the two surveys, the resulting trajectory (after
+waypoint 5) is shown to deviate. Survey 1 continued northwards,
+tracking the north-eastern portion of the front, while Survey 2 turned
+west, mapping the south-western region.
 
 The final predictions of the front location, represented by
-conditional EPs in Fig. \ref{fig:res1} and \ref{fig:res2} as dashed
-lines, correspond with one another. In both surveys they yield a
-picture of the front being to the west in the southern portions of the
-region and gradually bending off toward the north east. The amount of
-exploration done by Survey 1 which turned north is greater than Survey 2 which was coming close to the survey area borders in the south-western
-corner.
-
-\section{Closing remarks}\label{sec:concl_disc}
+conditional EPs in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2} as
+dashed lines, correspond with one another. In both surveys they yield
+a picture of the front being to the west in the southern portions of
+the region and gradually bending off toward the north east. The amount
+of exploration done by Survey 1 which turned north is greater than
+Survey 2 which was coming close to the survey area borders in the
+south-western corner.
+
+\section{Closing remarks}
+\label{sec:concl_disc}
 
 This work builds on a multidisciplinary effort combining statistical
-methods with robotic surveying techniques for oceanographic
+methods with robotic survey techniques for oceanographic
 applications. We show how observation practices can gain efficiency
-and accuracy from the development of statistical techniques for spatial monitoring. We further demonstrate the
-opportunities available for real-time multivariable spatial
-data gathering and analysis onboard autonomous platforms, which
-statisticians can exploit to create general-purpose toolkits for
-similar applications.
+and accuracy from the development of statistical techniques for
+spatial monitoring. We further demonstrate the opportunities available
+for real-time multivariable spatial data gathering and analysis on
+autonomous platforms, which statisticians can exploit to create
+general-purpose toolkits for similar applications.
 
@@
 connected to the properties of water masses. The characterization of
-uncertainties in random sets is extended with new results for the expected integrated Bernoulli variance
-reduction achieved by spatial sampling designs. This is
-provided in closed-form for the situation with a static design, and
-then extended to the adaptive situation. The sequential derivations
-provide new insights into efficient applications of adaptive data collection,
-as demonstrated in our application.
-
-The case study consider the upper water column in the river plume, represented by a two dimensional grid.
-Extensions to three-dimensional domains are not methodologically different, but the operation must likely approximate calculations by integrating terms only in the vicinity of the autonomous vehicle \citep{fossum18b}.
-We did not consider any temporal effects, which would be
-relevant on a larger time scale. We consider the extension to
-spatio-temporal modeling as future work, and envision that
-advection-diffusion equations could be useful in this kind of modeling
-\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex oceanographic
-phenomena, the methods will need to be extended to non-Gaussian
-phenomena, possibly feature-based mixtures of Gaussian processes which could still be
-run onboard and augmented by dynamical models.
+uncertainties in random sets is extended with new results for the
+expected integrated Bernoulli variance reduction achieved by spatial
+sampling designs. This is provided in closed-form for the situation
+with a static design, and then extended to the adaptive situation. The
+sequential derivations provide new insights into efficient
+applications of adaptive data collection, as demonstrated in our
+application.
+
+The case study considers the upper water column in the river plume,
+represented by a two dimensional grid. Extensions to three-dimensional
+domains are not methodologically different, but the operation must
+likely approximate calculations by integrating terms only in the
+vicinity of the autonomous vehicle \citep{fossum18b}. We did not
+consider any temporal effects, which would be relevant on a larger
+time scale. We consider the extension to spatio-temporal modeling as
+future work, and envision that advection-diffusion equations could be
+useful in this kind of modeling
+\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
+oceanographic phenomena, the methods will need to be extended to
+non-Gaussian phenomena, possibly feature-based mixtures of Gaussian
+processes which could still be run onboard and augmented by dynamical
+models.
 
 The spatial-statistical design criterion building on random sets is
-relevant in our setting with different water properties. 
-We show
-mathematical generality beyond the expected integrated Bernoulli variance, for instance that of volume uncertainties which is possibly more relevant, but requires more computational resources.
-Such criteria could be particularly useful in other oceanographic settings of algal-blooms, anoxic
-zones or open water fronts \cite{costa19}.
-Of course, other criteria
-are also relevant for instance, hybrid or multi-attribute criteria
-that could balance goals of exploration and exploitation in this
-situation. Equally, such techniques have significant use cases in
-downstream decision-making, with policy makers and regulators who need
-to make difficult decisions related to fish farming or other marine
-resource operations, and value of information analysis
-\citep{Eidsvik:15} could be used to evaluate whether
-information is likely to result in improved decision-making.
-We also foresee opportunities related to design of experiments for multivariate processes using our notion of generalized locations.
-For instance, this is likely to be useful for internet-of-things applications or computer experiments context, where some observations or evaluations are rather inexpensive, while others are much more costly and must only be done when they are really valuable.
+relevant in our setting with different water properties.  We show
+mathematical generality beyond the expected integrated Bernoulli
+variance, for instance that of volume uncertainties which is possibly
+more relevant, but requires more computational resources.  Such
+criteria could be particularly useful in other oceanographic settings
+related to mapping of algal-blooms, anoxic zones or open water fronts
+\cite{costa19}.  Of course, other criteria are also relevant for
+instance, hybrid or multi-attribute criteria that could balance goals
+of exploration and exploitation in this situation. Equally, such
+techniques have significant use cases in downstream decision-making,
+with policy makers and regulators who need to make difficult decisions
+related to fish farming or other marine resource operations. Value of
+information analysis \citep{Eidsvik:15} could be used to evaluate
+whether information is likely to result in improved decision-making,
+in such a context.  We also foresee opportunities related to design of
+experiments for multivariate processes using our notion of generalized
+locations. For instance, this is likely to be useful for
+internet-of-things applications or computer experiments where some
+observations or evaluations are rather inexpensive, while others 
+must only be done when they are really valuable.
 
@@
 
-In our context the myopic strategy perform rather well, and for computational reasons we did not go in depth on the dynamic program solution.
-There has been much creative work on finite horizon optimization in the robotics literature including probabilistic road maps and rapidly-exploring random trees \citep{karaman2011sampling}, but their statistical properties are not yet clear.
-It is equally interesting to
-explore the additional flexibility that can be gained by having
-multiple vehicles co-temporally exploring a spatial or spatio-temporal
-domain \citep{ferreira2019advancing}. Such an approach would enable
-concurrent sampling in different parts of the space, or opportunities
-to move in parallel to best capture the excursion set.
-The sometimes conflicting topics of autonomy and communication have caught much interest in control engineering lately \citep{zolich2019survey}, and statistical evaluations could contribute in the future.
-In our context, autonomy is highlighted, and this is reflected in our modeling and computational assumptions.
-With the possibilities of wifi communication to a mother-ship at waypoints, more advanced modeling and computational routines are of course possible, but with new challenges related to design and planning for when and where the communication should occur.
+In our context the myopic strategy performs well, and for
+computational reasons we did not go in depth on the dynamic program
+solution. There has been much creative work on finite horizon
+optimization in the robotics literature including probabilistic road
+maps and rapidly-exploring random trees \citep{karaman2011sampling},
+but their statistical properties are not yet clear.  It is equally
+interesting to explore the additional flexibility that can be gained
+by having multiple vehicles co-temporally exploring a spatial or
+spatio-temporal domain \citep{ferreira2019advancing}. Such an approach
+would enable concurrent sampling in different parts of the space, or
+opportunities to move in parallel to best capture the excursion set.
+The sometimes conflicting topics of autonomy and communication have
+caught much interest in control engineering lately
+\citep{zolich2019survey}, and statistical evaluations could contribute
+in the future. In our context, autonomy is highlighted, and this is
+reflected in our modeling and computational assumptions. With the
+possibilities of wifi communication to a mother-ship at waypoints,
+more advanced modeling and computational routines are of course
+possible, but with new challenges related to design and planning for
+when and where the communication should occur.
 

### 2020-06-23 main_new.tex
Updates from Overleaf

@@
 result in a range of ecosystem-related phenomena such as blooms and
-plumes, with direct and indirect effects on society. \kc{we should
-  look for a citation here. Check with John Ryan} One of the
+plumes, with direct and indirect effects on society \citep{ryan2017}. One of the
 bottlenecks in the study of such phenomena lies however in the lack of

### 2020-06-25 ans_to_reviewers/answers-to-reviewers.tex
Updates from Overleaf

@@
 
-We are grateful to the reviewers for their comments and
+We are grateful to the editors and reviewers for their comments and
 suggestions. They have all been considered in this revised
@@
 We have worked a lot on developing both theoretical results, computational resources, and on the presentation in the paper.
-To help us in these efforts, we included Cedric Traveletti as co-author because he has contributed significantly on the revised version of the paper.
+To help us in these efforts, we included Cedric Travelletti as co-author because he has contributed significantly on the revised version of the paper.
 
@@
 
-\begin{itemize}
+\begin{enumerate}
+\item The revision has lead to {\bf{structural changes}} as follows: \\
+    i) previous section 1 and 2 are combined to one introduction which should help the narrative and highlight the main contributions of the work. \\
+    ii) merged the data assimilation part with equations for coKriging to clarify that this work is on multivariate GRFs. \\
+    iii) Split the simulation study in two and used these as to directly motivate the statistical concepts of uncertainty reduction (Sect 2) and adaptive sampling strategies (Sect 3).
+
 \item The motivation for excursion sets (ESs), and in particular that of {\bf{multivariate ESs}}, has been strengthened in this revision.
@@
 Our motivation in the current paper is that temperature and salinity are proxies for more complex biological variables in the ocean, and in several cases they must be jointly high or low for phenomena to occur.
-For instance, salmon lice larvae are known to depend on relatively high temperature and high salinity to survive.
-Salmon lice are parasites to salmon, and the bivariate ES for temperature and salinity is hence important to understand for sustainable activities in salmon fisheries.
 
-In the revised manuscript these aspects are largely rewritten into the Introduction, and partly also in Section 2 on the expected integrated Bernoulli variance (EIBV), and partly in the Case Study ({\bf{TO DO 6 AGAIN}}).
-These changes address some key comments from AE and Rev.
-1 in particular.
+In the revised manuscript these aspects are largely rewritten into the Introduction, and partly also in Section 2. \\
+This change addresses comments of the Associate Editor and Rev. 1 in particular.
 
@@
 The revised version highlights these novel statistical results as a main contribution of the paper.
-
-Moreover, we emphasize the statistical design problem which is sought solved by AUV with its constraints on navigation and computing resources.
-
+We emphasize the statistical design problem which is sought solved by the AUV with its constraints on navigation and computing resources.
 In this sense, the revised version should provide ample motivation for statisticians, while still maintaining strong links to the application for illustrating concepts.
 
-In the revised manuscript these aspects are largely rewritten in the Introduction, in Sect 2 and 3, and on future work in the Closing remarks.
-The main body contains the novel statistical results for the expected uncertainty reduction in random sets exemplified by the EIBV, while the Appendix holds the proofs of key results.
-This change addresses the some comments of Editor, Associate Editor, Rev 1, Rev 2 and Rev 3.
+In the revised manuscript these aspects are largely rewritten in the Introduction, in Section 2 and 3, and on in the Closing remarks.
+The main body contains the novel statistical results for the expected uncertainty reduction in random sets exemplified by the EIBV, while the Appendix holds the proofs.
+\\
+This change addresses comments of the Editor, Associate Editor, Rev. 1, Rev. 2 and Rev. 3.
 
-\item We realized that a contribution of the current work is available code for practitioners or statisticians who could try this themselves, or possibly extend it for their own cases.
+\item We realize that a contribution of the current work is available code for practitioners or statisticians who could try this themselves, or possibly extend it for their own cases.
 We have hence made a substantial effort in structuring {\bf{computer code for examples}} which are presented on a URL link.
@@
 
-The simulation study section contains pointers to the associated open-source code. This change addresses the comments of Rev 1 and Rev 2 in particular.
+The Introduction section contains pointers to the open-source code. \\
+This change addresses comments of Rev. 1 and Rev. 2 in particular.
 
-\end{itemize}
+\end{enumerate}
 \par \vspace{1em}
@@
 %======================================================================
-All changes requested from the editor are listed below and answered on a point-by-point basis.
+All comments of the Editor are listed below and answered on a point-by-point basis.
 
@@
 
-\reply: The comment by you, AE and reviewers made us realize that the initial submission had a motivation which was too applied.
-We were not selling the paper contributions in early enough, and it was not targeted sufficiently to professional statisticians.
-In the revised version we motivate more for statisticians.
-
-Our actions are to i) rewrite Introduction and Background sections to catch our intended audience and highlight our contributions, ii) restructure the presentation to highlight both statistical contributions as well as the impact such technologies and applied questions could provide for statisticians.
-(See also comment 1 in the list of main changes above.)
+\reply: Based on the comment by you, AE and reviewers, we have restructured the Introduction and later parts of the paper to be more readable for professional statisticians.
+While doing so, we still strive to highlight how statistics can learn from new technologies like marine robotics and how robotics can benefit from new statistical developments.
 
@@
 
-\reply: We realize that we did not stress the statistical novelty sufficiently in the submitted version.
-The main statistical novelty is in the closed-form calculation for the design criteria of the expected integrated Bernoulli variance (EIBV) associated with excursion sets (ESs) for multivariate processes.
-In the revised version this is extended to volumes (going beyond single location).
-We believe that part of the problem with the initial submission was the organization of these new results in the structure of the paper.
-This is hence re-structured in the revision.
-
-Our actions are to i) extend the closed-form calculations to a criterion working on ES volume calculations (going beyond single location), ii) restructure the presentation to work directly on $p$-variate random fields and hence generalize the presentation, iii) restructure the closed-form results to be presented in the main body, but with proofs in the new Appendix.
+\reply: We realize that we did not stress the statistical novelty sufficiently in the initial version.
+The main statistical novelty is closed-form calculations for the design expected integrated Bernoulli variance (EIBV) for multivariate processes.
+In the revised version has clarified the statistical contributions via re-structured sections.
+Moreover, we extend the calculations to more general criteria (such as volumes, going beyond single location).
+The main closed-form results are presented in the main body, but with proofs in the new Appendix.
 
@@
 
-\reply: At this spatial scale there is often not so much data to infer a detailed statistical model, and this is the reason for our parsimonious construction for the real-data case.
-In other related situations, for instance that of larger fjord, ocean fronts, which are at a larger scale, there would usually be more data available for instance in the form of satellite data or from ensembles of computer model simulations.
-For such cases, we agree that it would make sense to embed more nuanced models, for instance that of larger variances at certain locations and stronger correlations in certain directions, but for our field trial here we attempted to have as few parameters as possible and had no reason to assume more complicated forms.
+\reply: At this spatial scale there is often not so much data to infer a detailed statistical model, and this is the reason for our parsimonious model in the real-data case.
+In other related situations, for instance that of larger fjord or ocean fronts, which are at a larger scale, there would usually be more data available for instance in the form of satellite data or from computer model simulations.
+(We added references to related papers that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here.) For such cases, we agree that it would make sense to embed more nuanced models onboard the AUV, for instance that of non-stationary models or anisotropic random fields, but for our field trial here we attempted to have as few parameters.
 
-That said, in the updated simulation study, which we have generalized and made open-source there are opportunities and examples of more complex covariance models (non-stationary and non-isotropic), but we do not conduct any inference for parameters in these models because it would require high resolution data from ocean models or satellites.
-
-Our actions include i) a better description of our parameter choices for the synthetic case and the specification for the real-data case, and ii) brief notes of the extended opportunities available for the associated open-access code, iii) references to related papers (including one of our own) that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here.
-MORE
+That said, in the updated simulation study, which we have generalized and made open-source, there are demos using more complex covariance models (non-stationary variance and non-isotropic random fields).
 
@@
 
-\reply: Based on the comments by yourself, AE and referees, we realize that the motivation part of the paper was too long.
-
-Our actions include a much shorter Section 2 with background. We now aim to engage readers earlier on with the applied statistical focus and methodological contribution of our work.
+\reply: Based on the comments by yourself, Associate Editor and referees, we realize that the motivation part of the paper was too long.
+We have merged the previous Introduction and Section 2 to the current Introduction, and we cut redundant parts in this process.
+We now aim to engage readers earlier, particularly so with applied statistical focus and the methodological contributions of our work.
 
@@
 %======================================================================
-All changes requested from the editor are listed below and answered on a point-by-point basis.
+All comments of the Associate Editor are listed below and answered on a point-by-point basis.
 
@@
 
-General: The inferential problem considered is the paper is that of estimating the “boundary” of a plume.
+General: The inferential problem considered in the paper is that of estimating the “boundary” of a plume.
 However, the authors never define exactly what a “boundary” is.
@@
 
-\reply: The revised version has an entirely new Section 2 which includes general background for random sets and excursion sets.
-We added references, for instance \cite{adler2009random} on excursion sets and \cite{Molchanov2005} on random sets.
+\reply: In the revised version we have strengthened the motivation for focusing on ESs and using the EIBV criteria.
+The revised version has an entirely rewritten Introduction, and Section 2 now includes general background for random sets and excursion sets.
+We added references, for instance \cite{adler2009random} on ESs and \cite{Molchanov2005} on random sets.
+The revised version provides results that go beyond the EIBV (with proofs in the new Appendix), but the EIBV is used in experiments because it is computationally tractable onboard the AUV.
 
@@
 
-\reply: We agree that they are not necessary for the methodological contributions, so moved this specification to the simulation study and example. 
-In the open-source that is now available, we implement non-stationary and non-isotropic covariance models.
-These represent situations that have relevance for our case.
+\reply: We agree that they are not necessary for the methodological contributions, so we now bring up this model specification when needed for the examples. 
+In the open-source code that is now available, we implement non-stationary and non-isotropic covariance models.
+These represent situations that have some relevance for our case.
 The code we post along with the paper facilitate testing so users can input more complex models to see how it influences designs.
@@
 \reply: The assumption is that salinity and temperature are conditionally independent measurements.
-(In the latent Gaussian random field, the variables are dependent.) We believe this conditional independence assumption is realistic as the measurements are gathered from different sensors.
-We added a comment about this not necessarily being the case with other data that are extracted and processed by combining sensors.
-Then assumptions can be tested in a lab under idealized conditions.
-But it is difficult to do so for the AUV in the changing oceans.
+(In the latent Gaussian random field, the variables are dependent.) We believe this conditional independence assumption is realistic as the measurements are gathered by a CTD sensor using different instruments for conductivity (transformed to salinity) and for temperature.
+We added a comment about this not necessarily being the case with other data that are extracted and processed by combining various sensor readings. 
+These assumptions can be tested in a lab under idealized conditions, but it is difficult to do so in the changing ocean conditions.
 
@@
 
-\reply: The AUV and the sampling happens regularly at 1 Hz, the collected measurements are then assimilated to this regular grid.
-The sampling density is therefore higher than the grid resolution.
-Still, there is a limitation on restricting the sampling locations to finite number of fixed locations.
-In order to handle the computational load and navigation resources available on the AUV this simplification is necessary.
+\reply: The sampling capabilities are restricted by the sensors onboard the AUV, its computational load and navigation resources.
+The sampling of temperature and salinity variables can happen regularly at 1 Hz.
+The sampling density can therefore be higher than the grid resolution.
+Still, in order to handle the computational load and navigation resources available on the AUV this simplification of data assimilation at the waypoints is necessary.
 % Comment Trygve: Should we add some more here?
@@
 
-\reply: We clarify this. TO DO!
+\reply: We clarify this in the revised document. The design is restricted to the waypoint graph with about 3 samples per batch.
 
@@
 
-\reply: We clarified the parameter specification. TO CHECK.
+\reply: We clarified the parameter specification.
+In practice the linear regression trend is fit from the levels at the boundary of the waypoint graph (in our case riverine water and ocean water).
+For the field experiment (Section 4) we clarified this specification.
 
-In a larger scale this could be improved by having auxiliary data for instance from ocean models or satellite data. (Mentioned under further work in closing remarks.)
+At larger scales this prior trend could be more nuanced because auxiliary data can then be available from ocean models or satellite observations.
+(This is mentioned under further work in closing remarks.)
 
@@
 
-\reply: This part has been rewritten. TO DO.
+\reply: This part has been rewritten.
 
@@
 
-\reply: We have changed the statement on page 9 to: ``$\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_1$ and $r^2_2$, holding the measurement error variances of temperature and salinity observations."
+\reply: This part has been rewritten.
 
@@
 
-\reply: We have changed this in the text - page 19.
+\reply: We have changed this in the text.
 
@@
 
-\reply: We have made this change to the citations: \cite{genton2015cross} and \cite{french2016credible} - page 30.
+\reply: We have made this change to the citations: \cite{genton2015cross} and \cite{french2016credible}.
 \end{itemize}
@@
 %=====================================================================
-All changes requested from Reviewer 1 are listed below and answered on a point-by-point basis.
+All comments of Reviewer 1 are listed below and answered on a point-by-point basis.
 
@@
 
-\reply: Thank you. We have improved and documented the code which is now open-source available.
+\reply: Thank you.
 
@@
 
-The things that stand out as novel in this paper are: the need for a se- quential design for driving an autonomous under water vehicle; a way of characterizing water plumes that are washed down from the land into the sea.
+The things that stand out as novel in this paper are: the need for a sequential design for driving an autonomous under water vehicle; a way of characterizing water plumes that are washed down from the land into the sea.
 Papers co-authored by one of the authors (Ginsbourger) of this paper inevitably means some overlap of ideas with the other papers, although the two step ahead greedy design algorithm seems novel.
@@
 
-\reply: We have improved the augmented the discussion on other design criteria and why the ES related IBV is used in our case.
+\reply: We have improved and augmented the discussion on other design criteria and why the ES related EIBV is used in our case.
+The paper also contains statistical results that go beyond this criterion (with proofs in the new Appendix), but the EIBV is used in practice because it is computationally tractable onboard the AUV.
 In this part, we cite the handbook reference on design \cite{zidek2019monitoring} and the paper we believe that contains a comment from Jon Lee \citep{bueso1998state}.
@@
 
-Overall the paper is remarkably free of a need for editorial improve- ments. I will attach a marked up copy indicating a few errors as well as suggestions for minor changes.
+Overall the paper is remarkably free of a need for editorial improvements. I will attach a marked up copy indicating a few errors as well as suggestions for minor changes.
 
-\reply: We incorporated the suggested minor changes.
+\reply: Thanks. We incorporated the suggested minor changes. But note that large parts of the paper have been rewritten (Point 1. in the start of this reply letter).
 
@@
 
-\reply: The motivation for an applied statistical audience is strengthened, while we have maintained strong links to the applications of Robotics, Ocean Science and AI.
-In particular in the Introduction and in Section 2 are now stronger on motivating multivariate ESs, and highlighting the need for new statistics in these marine robotics applications.
-See main comments early on in this reply letter.
+\reply: The motivation for an applied statistical audience is strengthened, while we have maintained strong links to the application of marine robotics.
+The Introduction and Section 2 are now stronger on motivating multivariate ESs, and highlighting the need for new statistics in the marine robotics applications.
+(See Point 2 and 3 in the start of this reply letter.)
 
@@
 That suggests including the relevant code for running the analyses, especially the simulation studies.
-An alterna- tive journal to a statistical one might be more appropriate from that perspective.
+An alternative journal to a statistical one might be more appropriate from that perspective.
 
@@
 This includes commented code examples for exploration and analysis of the material presented in the paper.
-The code is optimized for speed to allow fast computation using Pytorch (citation).
-The code is available here: (\url{https://cedrictravelletti.github.io/Submarine/} )
+The Python code allow fast calculation and replicable cases.
+It is also possible to run for more general statistical model than that shown in the paper (non-stationary variance and anisotropic random fields).
+The code is available here: \url{https://cedrictravelletti.github.io/Submarine/}.
 
@@
 
+Whatever the intended audience, the paper would benefit a lot from being shortened. I slogged through the first six papers before I discovered what the paper was to be about on page 6:\\
+”In this paper, with a focus on mapping the river plume, we reward the designs taht improve the classification of water masses as a means to characterize the frontal zone and thereby the extent of the river plume.”\\
+Later we find on page 7:
+”Our goal is to construct adaptive sampling strategies based on the continuous evolution of the EPs.”\\
+These sentences should have been put right at the beginning after suitable rewording along with the “In this work...” paragraph on page 3.
+If statisticians are the intended readers, deleting much of the first five pages would improve the paper.
+There are other places where technical material can be condensed without much if any loss.
+I will mention some later.
 On shortening the paper and moving important sentences earlier.
@@
 
-\reply: TO DO
+\reply: We have restructured the paper. See Point 1. at the start of this reply letter.
 
@@
 
-\reply: We added reference to \cite{adler2009random} on random sets and motivated this topic for environmental sensing. See our general comment 1 on top.
+\reply: We added reference to \cite{adler2009random} on random sets and motivated this topic for our application and more general settings.
+See our general point 2.
+and 3.
+at the start of this reply letter.
 
@@
 
-\reply: We improved our statistical contribution. See our general comment 2 on top.
+\reply: The design criteria for multivariate random fields are new.
+We further improved our statistical contribution by developing more general results (with proofs in Appendix).
+There are finally new elements in tailoring this sequential design methods in the context of the AUV application with its computational and navigation limitations.
+See our general comment 2 and 3 on top.
 
@@
 \reply: The traditional entropy calculations for this GP aim to reduce the log determinant of the updated covariance matrix at the grid locations.
-It will not vary with the data realizations, so in this setting this information criteria will always give the same path.
+It will not vary with the data realizations, so in our setting this information criteria will always give the same path.
 The EIBV criterion is aligned with our ES motivation for this study.
@@
 We see this as a clear benefit of the approach compared with entropy-based calculations.
-Still, sampling approaches using hybrid criteria or multi-attribute criteria (Multi-objective optimization and Pareto fronts) might balance goals in such situations, but we leave that for future research (closing remarks).
+Still, sampling approaches using hybrid criteria or multi-attribute criteria might balance goals in such situations, but we leave that for future research.
+(This is briefly mentioned as further work in Closing remarks.)
 
@@
 
-\reply: We focus on point-wise predictions in the criterion.
-Our trends are simple, so another covariance might be overfitting too much to data.
-We could try to fit different ranges to temperature and salinity, while maintaining separability.
-Could also refer to non-separable models, but pulls focus away from the current paper's focus.
-There are even tests for separability, but not worth so much, what do you do if you reject?
-Spatio-temporal variations are more interesting here, e.g.
-\cite{sigrist2015stochastic,richardson2017sparsity}.
-Ocean models, though rarely available on the scale of river plumes, might be highly useful for training - more complex trends and covariance model could then maybe be fit from the data.
+\reply: At this spatial scale there is often not so much data to infer a detailed statistical model, and this is the reason for our parsimonious model in the real-data case.
+In other related situations, for instance that of larger fjord or ocean fronts, which are at a larger scale, there would usually be more data available for instance in the form of satellite data or from computer model simulations.
+(We added references to related papers that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here.) That said, we do provide some validation of the model via residual analysis.
+Moreover, the main contribution of this paper is on design more than modeling.
+In the online resources with open-source code, there are demos using more complex covariance models (non-stationary variance and non-isotropic random fields).
+
+We agree that more complex models would matter more if we step up to a criterion such as volume calculations.
 
@@
 
-\reply: TO DO LA improved in revision! + code. %Also studied benefits of look-ahead strategies in sequential sampling in \cite{martinelli2013dynamic}.
+\reply: We have chosen not to explore any re-estimation of parameters.
+The resources onboard do not allow such computationally demanding evaluations.
+Overall, we believe the capabilities and limitations of the AUV are clarified in the revised version of the paper.
 
@@
 
-\reply: This is a simple model for plume temperature and salinity variation.
-However, it is not common to have ocean model data or satellite data at this resolution, which would be required for training.
-
-We have chosen to set few parameters, which also means that it is relatively easy to tune them at a break in the operation, if so is necessary.
+\reply: We agree that this is a simple model for plume temperature and salinity variation.
+However, it is not common to have ocean model data or satellite data at this spatial resolution, which would be required for training.
+So rather than overfitting, we have chosen a parsimonious model.
 
@@
 
-\reply: As pointed out above, we believe that spatio-temporal variations would be a more realistic extension, especially if repeated satellite data or ocean modeling is available.
-But this is likely relevant at a larger scale, such as that of fronts in a fjord or ocean interface.
-The methodological approach for sampling designs would however still be the same and this is the main contribution of the current paper.
+\reply: There are ongoing efforts to provide high-resolution ocean models which would be required for such detailed model specification.
+But as pointed out above, we believe that spatio-temporal variations would be a more realistic extension, especially if repeated satellite data or ocean modeling is available.
+The methodological approach for sampling designs, which is the main contribution of the current paper, would however still be the transferable.
 
@@
 \reply: We avoid the term' joint excursion sets' in the revised version.
-Instead we have introduced a mo more general random sets notation, but link to the ES for temperature and salinity whenever it is appropriate to exemplify concepts and show results.
+The revised version build on a more general notation, but link to the ES for temperature and salinity whenever it is appropriate to exemplify concepts and show results.
 
@@
 
-\reply: Preliminary runs set these trends. In a larger scale, one could here use ocean model inputs or satellite data.
+\reply: Preliminary runs set these trends. We have clarified the description in the paper. In a larger scale, one could here use ocean model inputs or satellite data.
 
@@
 
-\reply: We removed this Figure. It took much space and it is sufficient to explain the optimal design in text as we did. 
+\reply: We removed this Figure in the revision.
+
 \vspace{1em}
@@
 
-\reply: The revised Section 3 contains a general formulation for random sets, which is then exemplified by our excursion sets for the multivariate process.
-In the Introduction we motivate the excursion set for multivariate processes for this application, because it is important to find zones with joint high values of many variables.
-In the revised version we believe that the sets are introduced both more in the line with the practical application as well as for statistical novelty.
+\reply: The revised Section 2 contains a general formulation for random sets, which is then exemplified by our ESs for the multivariate process.
+In the Introduction and Section 2 we motivate the ES for multivariate processes for this application.
+In many applications it is important to find zones with joint high values of many variables.
+
 \vspace{1em}
@@
 I doubt that most statisticians would know that phrase.\par
-\reply: TO DO. Clarify. 
+\reply: The waypoint graph is clarified in the revised version.  
 \vspace{1em}
@@
 
-\reply: We use $Z$ in the revised version.
+\reply: We use $Z$ for the random field in the revised version.
 \vspace{1em}
@@
 Statisticians will know these things and users will not care.\par
-\reply: We collapsed the previous sections 3, 4 and 5 to one new section 3 containing the statistical novelty on EIBV closed-form expressions. The proofs are in a new Appendix.
+\reply: We collapsed the previous sections 3, 4 and 5 to new Section 2 and 3 containing the statistical novelty on EIBV closed-form expressions. The proofs are in a new Appendix.
 
@@
 %======================================================================
-All changes requested from Reviewer 2 are listed below and answered on a point-by-point basis.
+All comments of Reviewer 2 are listed below and answered on a point-by-point basis.
 
@@
 
-\reply: We have added more information about AUVs and their computational capabilities in Section 2.1 (Background on Ocean Processes and Observations) to address this. 
-\begin{quote}
-"The AUVs, see Fig.
-9, considered here operate at $1$-$3$ m/s in the upper water column, where they are free to move in six degrees of freedom (6 DOF).
-The in-water operation time capacity depends on survey speed, payload sensors and navigation.
-Generally this amounts to 8-24 hrs.
-operation for smaller crafts, and 24-48 hrs.
-for larger crafts.
-AUVs have limited computational capacity, as they typically use single-board computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM) to do computation.
-"
-\end{quote}
+\reply: We have clarified information about AUVs and their computational capabilities in Section 4. 
+It is a multicore GPU NVIDIA Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM). 
 \vspace{1em}
@@
 
-\reply: Maximum speed and payload are limiting for coverage.
-Generally this amounts to 8-24 hrs.
-operation for smaller crafts, and 24-48 hrs.
-for larger crafts.
+\reply: Maximum speed and payload are limiting coverage.
+This ranges to minutes to 24h for larger crafts.
 Time constraints for operations are typically related to costs of ship time and man-hours during operation.
-This information is contained in the changes made in point 2.1.1 above.
 \vspace{1em}
@@
 
-\reply: AUVs are free to move in 6 degrees of freedom on the triangulated graph.
-They are essentially flying under water.
-However, one usually always proceed forward while pitching up and down.
-We have clarified this in the updated section shown in point 2.1.1.
-This is also shown in some of the figures (Fig.1 and Fig.
-4) as well as the waypoint graph.
+\reply: AUVs as in our application are free to move in 6 degrees of freedom on the triangulated graph.
+This is shown in Fig.1 and Fig.
+4 with the waypoint graph.
+(They are essentially flying under water.
+In oceanographic application they also perform yo-yo movements in depth.)
 \vspace{1em}
@@
 
-\reply: The coverage available is usually limited on battery capacity, which also determines the maximum number of points to sample.
-The important question here is "How do one prioritize the correct points to visit?", which is central to the development of the IBV and the application shown in the paper.
+\reply: The coverage available is usually limited by man-hours to follow operations, but for large-time operation also on battery capacity.
+The important question here is "How do one prioritize the correct points to visit?", which is central to the development of the EIBV and the application shown in the paper.
 \vspace{1em}
@@
 \reply: The waypoint grid size $n$ should be small enough to resolve the process that is of interest and to avoid aliasing.
+Usually it is in the 100s or 1000s.
 A guiding rule of thumb provided by the the Nyquist theorem is to sample at least twice in time for the shortest significant record period, and twice in space for the shortest significant length to either resolve or eliminate (by filtering) scales of oceanic variability shorter than those being studied.
@@
 \reply: In the revised version the notation has been improved.
+There is limited data for training in our case, so there is a linear regression trend while covariance parameters have been set constant across locations.
 
@@
 
-\reply: In our setting the temperature and salinity data are from different sensors and it is naturally to say they are conditionally independent measurements.
-That said, for other measurements, they might be extracted or processed from common sensor readings and would then be conditionally dependent.
-We mentioned this briefly in the data model description.
+\reply: The temperature and salinity are dependent in the GRF model.
+But the temperature and salinity data are collected from different instruments, and we assume that they are conditionally independent, given the GRF variables.
+That said, other measurements might be extracted or processed from common sensor readings and would then be conditionally dependent.
+We mentioned this briefly in the model description.
 
@@
 
-\reply: Only the surface is considered here, so computational requirements are not mainly driven by the size of the domain, but rather by the number of combinations for sampling designs (computational complexity).
-No 3D with depth, which would require simplified spatial models like fixed-rank-kriging or similar approaches.
+\reply: The update formula assimilates relatively modest data sizes at each waypoint.
+Moreover, only the near-surface grid is considered here (two dimensional spatial domain).
+This means that the computational requirements are mainly driven by the number of combinations for sampling designs, not the size of the domain or the data dimension.
+We still agree that computational approximations, such as fixed-rank-kriging, would be useful for larger-size three dimensional applications.
 
@@
 
-\reply: We have generalized the closed-form solutions.
-(See point 2 at the start of reply letter.) The main body should contain results of general interest as well as provide insight about the statistical novelty.
+\reply: These parts of the paper are largely rewritten.
+We have generalized the closed-form solutions.
+(See point 2 and 3 at the start of reply letter.) The main body should contain results of general interest as well as provide insight about the statistical novelty.
 The proofs are now in Appendix with a succinct notation.
@@
 It seems to be related to the standardization that results in $\bZ$?
-Please clarify.
-
-\reply: We have generalized the closed-form solutions. (See reply above.)
-
-\item{\revcom  :}\label{r2c8}
-
-p. 11, Is $\bpi(\bm)$ the density associated with the distribution in Eq. (14)?
-
-\reply: We have generalized the closed-form solutions. (See reply above.)
-
-\item{\revcom :}\label{r2c9}
-
+Please clarify.\\
+p. 11, Is $\bpi(\bm)$ the density associated with the distribution in Eq. (14)? \\
 p.
@@
 (18).
-
-\reply: We have generalized the closed-form solutions. (See above.)
-
-\item{\revcom :}\label{r2c10}
-
+\\
 p.
 11, It is stated that $\bZ$ is chosen to be independent of $\bm$, so presumably the joint CDF $F_{z,m}=F_{Z}F_m$ which would certainly be convenient, but is not obvious to me why this would be true.
-Instead, it appears that $\bZ$ is chosen to be a pivotal quantity to get the simple CDF in Eq. (20). Please clarify.
-
-\reply: We have generalized the closed-form solutions. (See reply above.)
-
-\item{\revcom :}\label{r2c11}
-
+Instead, it appears that $\bZ$ is chosen to be a pivotal quantity to get the simple CDF in Eq. (20). Please clarify.\\
 p. 11, Eq. (21). The trailing $p(\bm)$ should be $\bpi(\bm)$. Should the right side of the equality be squared?
@@
 
-\reply: In the revised version we provide closed-form solutions to expected uncertainty reduction on volumes of random sets.
-However, for real-world application this involves many more terms in the evaluation criteria, so for the application, we still used marginals.
+\reply: In the revised version we provide closed-form solutions to more general criteria (beyond the EIBV), such as volume calculations involving coupling between locations.
+But as you mention, the real-world application has computational limitations, and we still use marginals in the examples.
 
@@
 
-\reply: TO DO
+\reply: We have removed this figure because it confused the explanation. The online resources contain movies of sampling paths, and the code allows users to generate sampling results by themselves.
 
@@
 
-\reply: Code is improved in the revision and provided for open-source usage (\url{https://cedrictravelletti.github.io/Submarine/} )
+\reply: Code is improved in the revision and provided for open-source usage: \url{https://cedrictravelletti.github.io/Submarine/}.
 
@@
 
-\reply:
+\reply: Modified in the revised version.
 
@@
 
-We also strengthen the discussion on communication for the AUV.
+We have also strengthened the discussion on communication for the AUV.
 In doing so, we relate the topics of autonomy and communication to the modeling and computational assumptions that applies to this paper.
@@
 
-\reply: In the revised version we show EPs for a static survey line, myopic and look-ahead selection in three displays. (This is in the new Figure 3.)
+\reply: These plots are changed in the revised version.
 
 In the accompanying code (\url{https://cedrictravelletti.github.io/Submarine/} )
-there are also resources for displaying other outputs if the reader is interested in comparing results with for instance univariate excursion probabilities.
+there are showing various sampling results and for displaying other outputs if the reader is interested in comparing results with for instance univariate excursion probabilities.
 

### 2020-06-25 main_new.tex
Updates from Overleaf

@@
   approaches, followed by results from field deployments with an
-  autonomous underwater vehicle (AUV) as part of a study mapping the
+  autonomous underwater vehicle as part of a study mapping the
   boundary of a river plume. The results demonstrate the potential of
@@
     ship-based sampling has been augmented by autonomous
-    robotic vehicles. % and their interactions.  
-    AUV platforms are an integral part of this network being able to
-    reason and make decisions for efficient onboard adaptive sampling.
+    robotic vehicles such as AUVs. % and their interactions.  
+    %AUV platforms are an integral part of this network being able to reason and make decisions for efficient onboard adaptive sampling.
     % using the sense-plan-act control approach to autonomous control. 
@@
 %Hence statistical proxy models of the environment must be used.
-For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
+For our purpose, we rely on Gaussian Random Field (GRF) representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
 \subsection{Previous work in AUV sampling}
@@
 In this paper the focus is not on statistical modeling, but rather on
-statistical properties of computation for efficient data
-collection. In this setting, we combine novel possibilities in marine
+statistical properties of computations for efficient data
+collection. We combine novel possibilities in marine
 robotics with spatial statistics and experimental design to provide
@@
 entropy reduction are independent of the actual data realizations
-under the assumptions of GP models, so it has limited flexibility for
+under the assumptions of GRF models, so it has limited flexibility for
 active adaptation of trajectories based on what is measured.  The use
@@
 \begin{itemize}
-\item Extending expected uncertainty reduction criteria to
-  vector-valued settings.  
-\item Closed form expressions for the expected integrated Bernoulli
+\item Extending uncertainty reduction criteria to
+  vector-valued cases.  
+\item Closed-form expressions for the expected integrated Bernoulli
   variance (IBV) of the excursions in GRFs. 
@@
   criteria. 
-\item Replicable experiments on synthetic test cases with accompanying
+\item Replicable experiments on synthetic cases with accompanying
   code~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}}
-\item Results of field trials running myopic strategies onboard an AUV
-  for the characterization of a river plume. 
+\item Results from full-scale field trials running myopic strategies onboard an AUV for the characterization of a river plume. 
 \end{itemize}
@@
 The remainder of this paper is organized as follows:
-%Section \ref{sec:bg} provides select background on ocean sampling. 
 Section \ref{sec:ESEP} defines ESs, excursion probabilities (EPs), and
 the design criteria connected to the IBV for excursions of
-vector-valued GPs. Section \ref{sec:heuristics} builds on these
+vector-valued GRFs. Section \ref{sec:heuristics} builds on these
 assumptions when deriving the sequential design criteria for adaptive
@@
 work.
-%\input{section2}
 
@@
 Section \ref{Sec:UnivarEx} illustrates the concepts on a bivariate
-example relevant for sampling temperature and salinity in our domain.
-
-\subsection{Background, Notation and Co-Kriging}
+example relevant for temperature and salinity in our case.
+
+\subsection{Background, Notation and co-Kriging}
 \label{sec:bg_and_notation}
@@
 bivariate GRF model is assumed for $\gp$. To motivate concepts,
-Fig.~\ref{example_excu} (left and middle \kc{If you split this figure
-  in its parts and add labels, it would be more elegant and precise to
-  explicitly refer to them.}) shows a realization of such a
-vector-valued GRF on $\domain=(0,1)^2$. Fig.~\ref{example_excu} (right
-\kc{See above}) represents a by-product of interest derived from these
-realizations, namely regions i) in red, where both temperature and
-salinity are high (i.e., exceeding respective thresholds), indicative
-of ocean water ii) in white, where both temperature and salinity are
-low, indicative of riverine water, and iii) in pink, where one
-variable is above and the other below their respective thresholds,
-indicative of mixed waters.
-
-\begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/example_excu_4.png}
-  \caption{Realization of a bivariate GRF and excursion set above some
-    threshold ({\kc{JE: are means around 5 and 30 here, so same as
-        later in the example?}} ). Joint excursion in red, excursion
-    of a single variable in pink.}
+Fig.~\ref{fig:real_temp} and \ref{fig:real_sal} shows a realization of such a
+vector-valued GRF on $\domain=(0,1)^2$. Fig.~\ref{fig:jointex_roi} represents a by-product of interest derived from these
+realizations, namely regions: i) in red, where both temperature and salinity are high (i.e., exceeding respective thresholds), indicative of ocean water, ii) in white, where both temperature and salinity are low, indicative of riverine water, and iii) in light-red, where one variable is above and the other below their respective thresholds, indicative of mixed waters.
+
+\begin{figure}[!b] 
+\centering 
+\subfigure[Temperature.]{\includegraphics[
+height=0.27\textwidth,keepaspectratio]{Figures/intro_temp.png}\label{fig:real_temp}}
+\subfigure[Salinity.]{\includegraphics[
+height=0.27\textwidth,keepaspectratio]{Figures/intro_sal.png}\label{fig:real_sal}}
+\subfigure[Regions of interest.]{\includegraphics[
+height=0.27\textwidth,keepaspectratio]{Figures/intro_excu.png}\label{fig:jointex_roi}}
+\caption{Realization of a bivariate GRF (display (a) and (b)) and excursion set above some threshold (c). Joint excursion in red and excursion of a single variable in light-red.}
 \label{example_excu}
@@
 $$
-%
-%Our goal here is to develop approaches to quantify and improve the
-%characterization of uncertainties on $\es$. \textcolor{red}{Alread mention how observations are included.}
-
 If we assume that $\gp$ has continuous trajectories 
-%(almost surely \kc{This seems very handwavy; would delete}) 
 and $T$ is closed, then $\es$ becomes a Random Closed Set
@@
 
-\medskip
-
-%\subsubsection{Notation}
-%In order to work at this level of generality, some notational tricks are needed.
-Indeed, when considering a new observation, one shall in general specify which component of the field was observed and where it was observed.
-This leads us to introduce the concept of \textit{generalized location}.
-
 Letting $\gp[\spatloc,\ell]$ denote the $\ell\text{-th}$ component of
@@
 %This slight change of notation 
-will allow us to think of $\gp$ as a scalar-valued random field indexed by $\domain \times \{1\dots,p\}$, which will give the co-kriging equations a particularly simple form that parallels the one of univariate kriging.
+will allow us to think of $\gp$ as a scalar-valued random field indexed by $\domain \times \{1\dots,p\}$, which will give the co-Kriging equations a particularly simple form that parallels the one of univariate Kriging.
 %Due to the naturality of the concept of \textit{generalized location}, we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $\spatloc \in \domain$.
@@
 %\end{align*}
-%In general,  corresponding to batches of observations. 
-%The field values of batch observations are denoted by
+%In general,  corresponding to batches of observations. %The field values of batch observations are denoted by
 %\begin{align*}
@@
 %In the same fashion, $\mu(\bm{x})$ denotes the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ the $q \times q$ covariance matrix.
-
 %Given a random field $\gp$ and (noisy) observations of some of its components at selected points in the domain, one can predict the value of the field at location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$ and the associated covariance .
@@
 %We present a general form of cokriging, where observations at locations $s \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ (heterotopic).
-
 %Assuming that one wishes 
@@
 %$k(\bm{x}_{[n]}, \bm{x}_{[n]})$ being assumed non-singular throughout the presentation. 
-%The associated co-kriging %(conditional) 
+%The associated co-Kriging %(conditional) 
 %residual cross-covariance function can be expressed in the same vein via
@@
 %\end{equation}
-
-%\subsubsection{Co-kriging update formulae}
-
+%\subsubsection{co-Kriging update formulae}
 %In our case it is important to update the GRF model efficiently onboard the vehicle after every batch of obervations.
 Our setting is like that of a spatial Kalman filter with static model parameters, and we provide the co-Kriging equations for this case.
-Because of our representation of co-kriging in terms of simple kriging with respect to generalized locations, the formulae is a straightforward adaptation of the batch-sequential Kriging update formulae from \cite{Chevalier.etal2013a}.
+Because of our representation of co-Kriging in terms of simple kriging with respect to generalized locations, the formulae is a straightforward adaptation of the batch-sequential Kriging update formulae from \cite{Chevalier.etal2013a}.
 The formulae are instrumental in deriving semi-analytical results for stepwise expected uncertainty reduction criteria for vector-valued random fields in the subsequent sections.
-
 %Given the observations from $n$ batches, one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
@@
 \end{align*}
-%
+
 For a second order random field $(Z_{\spatloc})_{\spatloc \in
@@
   \ell'}$. In vectorized/batch form, $k(\bm{x},
-\bm{x}')$ then amounts to a matrix with numbers of lines and columns
-the numbers of generalized locations in the batches
-$\bm{x}$ in
-$\bm{x}'$, respectively \kc{This last sentence doesn't parse.}. Such
+\bm{x}')$ then amounts to a matrix with numbers of lines and columns equal to
+the numbers of generalized locations in
+$\bm{x}$ and
+$\bm{x}'$, respectively. Such
 vectorized quantities turn out to be useful in order to arrive at
-simple expressions for co-Kriging equations below.
+simple expressions for the co-Kriging equations below.
 %\textcolor{red}{(ADDENDUM, C.T.) In the same fashion, we will use $\mu(\bm{x})$ to denote the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ for the corresponding $q \times q$ covariance matrix.}
 
-%\medskip
-
-Given a GRF $\gp$ and observations of some of its components at some
-points in the domain, one can predict the value of the field at some
+Given a GRF $\gp$ and observations of some of its components at locations in the domain, one can predict the value of the field at some
 unobserved location $\spatloc\in \domain$ by using the conditional
-mean of $\gp[\spatloc]$, conditional on the data \kc{...conditional
-  mean on the data?}. This coincides with (co-)kriging equations,
-which tell us precisely how to compute conditional means and
-covariances conditional on an arbitrary dataset.  We will present a
-general form of co-kriging, in the sense that it allows inclusion of
-several observations (batch) at a time; observations at a given
+mean of $\gp[\spatloc]$, conditional on the data. This coincides with co-Kriging equations,
+which tell us precisely how to compute conditional means and covariances.  We will present a
+general form of co-Kriging, in the sense that it allows inclusion of
+several (batch) observations at a time; observations at a given
 location $u \in \domain$ may only include a subset of the components
@@
 $\bm{x}$, %\equiv(\bm{\spatloc}, \bm{\ell})$,
-the simple cokriging mean then amounts to kriging with respect to a
+the simple co-Kriging mean then amounts to Kriging with respect to a
 scalar-valued GRF indexed by $\domain\times \{1\dots,p\}$:
@@
 \begin{equation}\label{eq:cokrig_mean}
-\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})),
+\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})).
 \end{equation}
-where $\mu$ is $Z$'s initial mean function,
-$\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n
+Here, $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n
 q_i$)-dimensional vector of observed (noisy) responses of
@@
 \bm{x}_n)$ and where
-$\Delta_{[n]}$ stands for the covariance matrix of
+$\Delta_{[n]}$ is the covariance matrix of
 Gaussian-distributed noise assumed to have affected measurements up to
 batch
-$n$. The matrix in parenthesis will be assumed to be non-singular
-throughout the presentation. The co-kriging %(conditional)
-residual (cross-)covariance function (with respect to batches of
-generalized locations) can also be expressed in the same vein via
+$n$.
+For our applications with salinity and temperature observations, this matrix is diagonal because we assume conditionally independent sensor readings, but it might not be diagonal with other types of combined measurements.
+The matrix in parenthesis will be assumed to be non-singular
+throughout the presentation. The associated co-Kriging %(conditional)
+residual (cross-)covariance function can also be expressed in the same vein via
 %
@@
 
-%\subsubsection{Co-kriging update formulae}
-
-Let us now consider the case where a co-kriging prediction of $Z$ was
-made with respect to $n$ batches $\bm{x}_i$ of generalized locations,
-concatenated again within $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$,
+%\subsubsection{co-Kriging update formulae}
+
+Let us now consider the case where a co-Kriging prediction of $Z$ was
+made with respect to $n$ batches of generalized locations, concatenated again within $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$,
 and one wishes to update the prediction by incorporating a new vector
@@
 %It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
-Thanks to our representation of co-kriging in terms of simple kriging
+Thanks to our representation of co-Kriging in terms of simple Kriging
 with respect to generalized locations, a strightforward adaptation of
-the batch-sequential kriging update formulae from
+the batch-sequential Kriging update formulae from
 \citep{Chevalier.etal2013a} suggests that
@@
 sub-vector extracted from $\lambda_{[n+1]}(\bm{x})$ that corresponds
-to the kriging weigths associated with the last $q_{n+1}$ responses
+to the Kriging weights for the last $q_{n+1}$ responses
 when predicting at $\bm{x}$ relying on all measurements until batch
@@
 %, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
-Similarly, the updated co-kriging residual (cross-)covariance function
-then becomes
+The associated co-Kriging residual (cross-)covariance function is
 \begin{align}\label{eq:varCoK}
@@
 As noted in \citep{Chevalier2015} in the case of scalar-valued fields,
-these update formulae naturally extend to Universal Kriging in
+these update formulae naturally extend to universal Kriging in
 second-order settings and apply without Gaussian assumptions. We will
 now see how the latter formulae are instrumental in deriving
-semi-analytical formulae for stepwise uncertainty reduction criteria
+semi-analytical formulae for step-wise uncertainty reduction criteria
 for vector-valued random fields.
@@
 $\domain$. We want to investigate the probability distribution of
-$\mes(\es)$ through its moments.  Centred moments may be computed
+$\mes(\es)$ through its moments.  Centered moments may be computed
 using Proposition~\ref{propo1} developed in the appendix.  In
-particular, the excursion measure variance
-$\emv = \operatorname{Var}[\mes(\es)]$ is an integral of the excursion
-probabilities (EP)
-
+particular, as an integral over EPs, the 
+$\emv = \operatorname{Var}[\mes(\es)]$ is:
 \begin{equation*}
@@
 \end{equation*}
-
 where $\varPhi_{\no}$ denotes the $\no$-variate Gaussian cumulative
-distribution function (CDF), which can be calculated, using for
-instance, the code of \cite{genz2009computation}.
+distribution function (CDF) numerically \citep{genz2009computation}.
 
 Note that this quantity requires the solution of an integral over
-$\domain^2$. In contrast, and as in the scalar-valued case, the IBV of
+$\domain^2$. In contrast, the IBV of
 \cite{bect2019} involves solely an integral on $\domain$ and can be
@@
 We compute the expected effect of the inclusion of new observations on
-the $\emv$ and $\ibv$ of the excursion set $\es$. Let us consider the
+the $\emv$ and $\ibv$ of the ES $\es$. Let us consider the
 same setting as in Eqn. \eqref{eq:meanCoK} and \eqref{eq:varCoK}, and
@@
 expectation and probability conditional on the first $n$ batches of
-observations, respectively. We want to compute the effect of including
-a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and
-$\ibv$. We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the
-conditional law $\mathbb{P}_{\stage}$.
+observations, respectively. %We want to compute the effect of including a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and $\ibv$. 
+We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.
 
@@
 let $ \currentIBV(\bm{x}; \bm{y}) $ denote the expected IBV under the
-current law of the field \kc{Is this a legitimate expression 'law of
-  the field'?}, conditioned on observing $\bm{y}$ at $\bm{x}$
+current law of the field, conditioned on observing $\bm{y}$ at $\bm{x}$
 (generalized, possibly batch observation). The expected effect of a
-new observation on the IBV is then shown by
+new observation on the IBV is then 
 \begin{equation}\label{def:eibv}
-    \currentEIBV(\bm{x}):=\mathbb{E}_{\stage}\left[\mbox{IBV}(\bm{x}; \bm{Y})\right]
+    \currentEIBV(\bm{x}):=\mathbb{E}_{\stage}\left[\mbox{IBV}(\bm{x}; \bm{Y})\right],
 \end{equation}
@@
 We next present a result that allows efficient computation of $\EIBV$
-as an integral of CDFs of multivariate gaussians. This will prove
-useful when designing sequential expected uncertainty reduction
-strategies.
+as an integral of CDFs of the multivariate Gaussian distribution. This will prove
+useful when designing sequential expected uncertainty reduction strategies.
 
@@
 \end{equation*}
-
 with $V=\gp[\bm{x}]-\currentMean{\bm{x}} \sim \mathcal{N}(0_{q_{n+1}},k_{[n]}(\bm{x},\bm{x}))$ and $a_1=\bt-\currentMean{\uu}$,
@@
 
-This integrand finally boils down to an evaluation of the
-$2\no$-dimensional Gaussian CDF by Proposition~\ref{propo3}. Details
+This integrand boils down to an evaluation of the
+$2\no$-dimensional Gaussian CDF by Proposition~\ref{propo3} (Appendix). Details
 are omitted here for brevity as we do not implement this criterion
-further and which requires numerical integration over $\domain^2$.
+further because it requires numerical integration over $\domain^2$.
 \end{propo}
@@
 the inclusion of batch or heterotopic observations through the concept
-of generalized locations.  A key element for understanding of these
-propositions is that the conditional co-Kriging means entering in the
+of generalized locations.  A key element for understanding these
+propositions is that the conditional co-Kriging mean entering in the
 EPs depend linearly on (batch) observations. The conditional equality
@@
 the first and second component of the random field will be called
-\textit{temperature} and \textit{salinity} for illustrative
-purpose. Effects of the hyperparameters of the GRF prior on the
-excursion probabilities will also be studied. For simplicity, we begin
-with a \textit{pointwise} example, considering a single bivariate
-gaussian distribution (i.e. no spatiality).
+\textit{temperature} and \textit{salinity}. We begin
+with a \textit{pointwise} example, considering a single bivariate Gaussian distribution (i.e. no spatial elements).
 
@@
 
-Say we want to study the excursion probability of a bivariate gaussian
+Say we want to study the excursion probability of a bivariate Gaussian
 above some threshold, where the thresholds are set equal to the mean;
@@
 salinity, and we play with the temperature and salinity correlation
-and variances to study the effect on the EP and expected Bernoulli
-variance.
+and variances to study the effect on the EP and expected Bernoulli variance.
 
@@
 \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
-\caption{Density contour plots with different correlations between
-  temperature and salinity. The densities have unit variance and the
-  thresholds are identical to the mean values $5^o C$ and
+\caption{Density contour plots with increasing correlations between
+  temperature and salinity. The densities have unit variance and
+  thresholds identical to the mean values $5^o C$ and
   $30$ g/kg.}
@@
 Fig.~\ref{illus_bivarDens}. The EPs increase with the correlation as
-there is a strong tendency to have concurrently low temperature and
-salinity. The Bernoulli variance is similarly large for high
+there is a strong tendency to have jointly low or high temperature and
+salinity. The Bernoulli variance is similarly larger for high
 correlations. EPs and Bernoulli variances are the same for standard
@@
 
-\begin{table}[!t] \centering \caption{EP and Bernoulli variance for
-    different correlations and variances (top rows), and expected
-    Bernoulli variances for both temperature and salinity data $\by$
-    and temperature $y_2$ (bottom rows). \kc{$\mu_1$  or $\by$? And
-      whatever happened to make the terms more accessible with $t$ and
-    $s$ as we had before?}}
+\begin{table}[!t] \centering \caption{EP and Bernoulli variance for different correlations and variances (top rows), and expected Bernoulli variances for both temperature and salinity data, and only temperature data (bottom rows).}
   \begin{tabular}{c|ccc|ccc}
@@
 $p(1-p)$ & 0.20 & 0.23 & 0.24 & 0.20 & 0.23 & 0.24 \\ 
-EIBV, Temperature and Salinity & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
-EIBV, Temperature only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
+EBV, Temperature and Salinity & 0.092 & 0.089 & 0.085 & 0.052 & 0.051 & 0.049 \\ 
+EBV, Temperature only & 0.151 & 0.138 & 0.123 & 0.137 & 0.114 & 0.093 \\ 
 \hline
@@
 \end{table}
-
-The bottom two rows of Table \ref{tab:sim_rhoab} show results of
-expected Bernoulli variance. This is presented for a design gathering
+The bottom two rows of Table \ref{tab:sim_rhoab} show 
+expected Bernoulli variance results. This is presented for a design gathering
 both data types, and for a design with temperature measurements
@@
 Bernoulli variance gets lower with larger standard deviations
-$\sigma_1$ and $\sigma_2$. The reduction of Bernoulli variance is
+$\sigma_1$ and $\sigma_2$. The expected reduction of Bernoulli variance is
 largest for the cases with high correlation $\gamma$. Albeit smaller,
@@
 information about salinity in the temperature data, and therefore less
-uncertainty reduction. In an application with fresh cold water from a
-river source, the temperature and salinity variables will not only be
-interdependent, but will also likely show dependence in the spatial
-dimension. This in turn will impact the design criteria when we
-evaluate the information measure by integrating over several locations
-(Section \ref{sec:simulations}).
+uncertainty reduction. 
+%In an application with fresh cold water from a river source, the temperature and salinity variables will not only be interdependent, but will also likely show dependence in the spatial dimension.
+This in turn will impact the design criteria when we
+% evaluate the information measure by integrating over several locations (Section \ref{sec:simulations}).
 
@@
 We now turn to an example involving a full-fledged gaussian random
-field (GRF). The class of GRF model we will consider will generally
-have a linear trend
-
+field (GRF). The class of GRF model we consider have a linear trend
 \begin{align*}
@@
 \end{align*}
-with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. In our examples, we will only consider covariance models of separable type
+with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. In our examples, we only consider separable covariance models;
 \begin{align*}
-\textrm{Cov}\left(Z_{\spatloc, i}, Z_{v, j}\right) &= k(\spatloc, v) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_l^2,~ i=j\\
-   \gamma_0 \sigma_i \sigma_j,~i\neq j
+\textrm{Cov}\left(Z_{\spatloc, i}, Z_{v, j}\right) &= k(\spatloc, v) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_i^2,~ i=j\\
+   \gamma \sigma_i \sigma_j,~i\neq j,
         \end{cases}
@@
 where $k(., .)$ is one of the traditional spatial covariance kernel
-(exponential, Mat\'{e}rn, ...) and $\gamma$ defines the
+(exponential, Mat\'{e}rn, etc.) and $\gamma$ defines the
 cross-covariance structure. 
-
-In the accompanying \kc{Where? On Github? If so say so.} Python
-examples taking place within the MESLAS toolbox (URL \kc{Note}), these
-modeling assumptions can be generalized to anisotropic covariance and
+In the accompanying Python
+examples taking place within the MESLAS toolbox~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}}, these
+modeling assumptions can however be relaxed to anisotropic covariance and
 changing variance levels across the spatial domain. Anisotropy and
@@
 plumes, but in practice this requires more parameters to be
-specified. With extensive data and prior knowledge, one could also
-possibly fit and estimate parameters of more complex multivariate
+specified. With extensive satellite data or prior knowledge from high-resolution ocean models, one could also possibly fit more complex multivariate
 spatial covariance functions
@@
 \end{align*}
-and the spatial covariance is given by a Mat\'{e}rn 3/2 kernel with unit variance and range parameter $\lambda=0.5$.
+and the spatial covariance has unit variance and a Mat\'{e}rn 3/2 kernel, $(1+\eta h)\exp (-\eta h)$, for Euclidean distance $h$ and decay parameter $\eta=0.5$.
 One realization of this GRF is shown in Fig.~\ref{example_excu}.
@@
 two components of the field is observed. We first draw a realization
-of the GRF defined above and use it as ground-truth to mimick the real
+of the GRF defined above and use it as ground-truth to mimic the real
 data-collection process. A first set of observations are done at the
@@
 as a potential next observation location and plot the expected
-Bernoulli variance reduction (at each grid point \kc{I don't recall we
-  have been explicit about grids before this mention. It might be
-  useful to recall that here.}) that would result from observing only
-one component of the field (temperature or salinity), or both at that
-point.
-
-\begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/ebv_comp_3.png}
+Bernoulli variance reduction (at each grid point/nodes in the the waypoint graph) that would result from observing only one component of the field (temperature or salinity), or both at that point.
+
+\begin{figure}[!b] 
+\centering 
+\subfigure[Regions of interest.]{\includegraphics[
+height=0.21\textwidth,keepaspectratio]{Figures/ebv_comp_excu.png}\label{fig:ebv_comp_excu}}
+\subfigure[Temperature.]{\includegraphics[
+height=0.22\textwidth,keepaspectratio]{Figures/ebv_comp_temp.png}\label{fig:ebv_comp_temp}}
+\subfigure[Salinity.]{\includegraphics[
+height=0.22\textwidth,keepaspectratio]{Figures/ebv_comp_sal.png}\label{fig:ebv_comp_sal}}
+\subfigure[Both.]{\includegraphics[
+height=0.22\textwidth,keepaspectratio]{Figures/ebv_comp_both.png}\label{fig:ebv_comp_both}}
 \caption{Pointwise Bernoulli variance reduction for observation of a
@@
     in red. Places where only one response is above threshold are
-    depicted in pink}
+    depicted in pink. EBV reduction associated to observing one or both responses at the green location are shown in \ref{fig:ebv_comp_temp}, \ref{fig:ebv_comp_sal} and \ref{fig:ebv_comp_both}.}
 \label{fig:ebv_comp}
@@
 
-\begin{figure}[h!] \centering
-  \includegraphics[width=0.99\textwidth]{Figures/ebv_north_vs_east_2.png}
-  \caption{Pointwise Bernoulli variance reduction for two different
-    observation plans. Data collection locations in green. True
-    excursion set is in red. Places where only one response is above
-    threshold are depicted in pink.}
+\begin{figure}[!b] 
+\centering 
+\subfigure[Excursion probability.]{\includegraphics[
+height=0.25\textwidth,keepaspectratio]{Figures/ebv_static_excu.png}\label{fig:ebv_static_excu}}
+\subfigure[Static north design.]{\includegraphics[
+height=0.26\textwidth,keepaspectratio]{Figures/ebv_static_north.png}\label{fig:ebv_static_north}}
+\subfigure[Static east design.]{\includegraphics[
+height=0.26\textwidth,keepaspectratio]{Figures/ebv_static_east.png}\label{fig:ebv_static_east}}
+  \caption{Pointwise Bernoulli variance reduction for two different observation plans.
+The prior excursion probability is shown in \ref{fig:ebv_static_excu}.
+EBV reduction for each design shown in \ref{fig:ebv_static_north} and \ref{fig:ebv_static_east}.}
 \label{fig:ebv_north_vs_east}
@@
 Assume data have been collected up to stage $n$. The principle of
-two-step lookahead is to select the next observation location
+two-step look-ahead is to select the next observation location
 $\spatloc_{\stage + 1}$ that yields the biggest reduction in EIBV if
@@
 
-\begin{criterion}[2-step lookahead]
+\begin{criterion}[2-step look-ahead]
       The next observation location $\spatloc_{\stage + 1}$ is chosen among the minimizers in $\candidates$ of the criterion
@@
   $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
-  \hfill \subfigure[Computational time for inference (some lines
-  overlap \kc{Which ones? As figure with 6 keys, one sees only 4. This
-    needs to be fixed or explained -- this was a repeat from an
-    earlier
-    feedback.}).]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}}
-\caption{Simulation results from 100 replicate simulations for 10
-  sampling choices/stages on the grid. Vertical lines show variation in replicate results.}  
+  \hfill \subfigure[Computational time for inference (the lines for \textit{naive}, \textit{static\_north}, \textit{static\_east}, and \textit{static\_zigzag}
+  overlap).]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}}
+\caption{Simulation results from 100 replicate simulations for 10 sampling choices/stages on the grid. Vertical lines show variation in replicate results.}  
 \label{fig:sim_results}
@@
 the trends in environmental conditions and correlation
-structures. Based on the initial data, the mean was specified by linear regression, where both temperature and salinity
-were assumed to increase with the west location. Next, the residuals from the regression analysis were analyzed to specify the covariance parameters of the GRF model.
+structures.
+Based on the initial runs we get a reasonable idea of the temperature and salinity of river and ocean waters, and also specify the mean by linear regression, where both temperature and salinity
+were both assumed to increase linearly with the west location on the waypoint graph.
+Next, the residuals from the regression analysis were analyzed to specify the covariance parameters of the GRF model.
 
@@
 analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
-and salinity residuals after the westerly trends in both salinity and
+and salinity residuals after the regression mean values of salinity and
 temperature are subtracted from the data. This scatter-plot of joint
@@
 and a positive correlation ($0.5$) between the two variables. Based on
-the fitted bivariate Gaussian model (ellipses in
-Fig.~\ref{fig:parest_a}), we can compute the modeled quadratic form of
-the residuals, and if the model is adequate they should be
-approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
-the empirical CDF of these quadratic forms (solid) together with the
-theoretical CDF of the $\chi^2_2$ distribution. The modeled and
-theoretical curves are very similar, which indicates that the Gaussian
-model fits reasonably well. Fig.~\ref{fig:parest_c} shows the
-empirical variogram of the scaled residuals for temperature and
-salinity. The decay is similar for the two, and seems to be negligible
+the fitted bivariate covariance model (ellipses in
+Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
+the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
+the empirical CDF of the quadratic forms (solid) together with the
+theoretical CDF of the $\chi^2_2$ distribution (dashed). The modeled and
+theoretical curves are rather similar, which indicates that the Gaussian
+model with constant spatial variance and correlation fits reasonably well. Fig.~\ref{fig:parest_c} shows the
+empirical variogram of the residuals for temperature and salinity. The decay is similar for the two, and seems to be negligible
 after about $150$ m.  The working assumption of a separable covariance
@@
 Fig.~\ref{fig:map}.  At each stage, it takes the AUV about 30 seconds
-to evaluate the EIBV for all the possible waypoint alternatives.  It
+to assimilate data and evaluate the EIBV for all the possible waypoint alternatives.  It
 was set to start in the south-center part of the waypoint graph. A
@@
 processes which could still be run onboard and augmented by dynamical
-models.
+models. If high-resolution ocean model or satelite data are available, it is hopefully possible to fit such more complex statistical models.
 
@@
 
-\kc{How about talking
-  about removing any grids or waypoints altogether?}
-
 In our context the myopic strategy performs well, and for
@@
 maps and rapidly-exploring random trees \citep{karaman2011sampling},
-but their statistical properties are not yet clear.  It is equally
+but their statistical properties are not yet clear.  
+In some cases it is also limiting to use the waypoint graph, and it would be beneficial to allow more continuous updates and navigation at the highest frequency allowed by the AUV.
+It is equally
 interesting to explore the additional flexibility that can be gained

### 2020-06-25 main_new.tex
Updates from Overleaf

@@
   vector-valued Gaussian random fields, and derive tractable
-  expressions for the expected Bernoulli variance reduction in such a
+  expressions for the expected integrated Bernoulli variance reduction in such a
   framework. We demonstrate how this criterion can be used to
@@
 
-\subsection{Expected Bernoulli Variance on a two dimensional Example}
+\subsection{Expected Bernoulli variance for a two dimensional Example}
 \label{Sec:UnivarEx}
 
-We illustrate the EBV associated with different designs on a simple
+We illustrate the expected Bernoulli variance (EBV) associated with different designs on a simple
 bivariate example. This mimics our river plume application and hence
@@
 salinity, and we play with the temperature and salinity correlation
-and variances to study the effect on the EP and expected Bernoulli variance.
+and variances to study the effect on the EP and EBV.
 
@@
 salinity. The Bernoulli variance is similarly larger for high
-correlations. EPs and Bernoulli variances are the same for standard
-deviation $1$ or $2$, which implies that high variability in
+correlations. EPs and Bernoulli variances are the same for temperature and salinity standard
+deviations $\sigma_1$ and $\sigma_2$, which implies that high variability in
 temperature and salinity is not captured in the $p(1-p)$ expression.
 
-\begin{table}[!t] \centering \caption{EP and Bernoulli variance for different correlations and variances (top rows), and expected Bernoulli variances for both temperature and salinity data, and only temperature data (bottom rows).}
+\begin{table}[!t] \centering \caption{EP and Bernoulli variance for different correlations and variances (top rows), and EBVs for both temperature and salinity data, and only temperature data (bottom rows).}
   \begin{tabular}{c|ccc|ccc}
@@
 The bottom two rows of Table \ref{tab:sim_rhoab} show 
-expected Bernoulli variance results. This is presented for a design gathering
+EBV results. This is presented for a design gathering
 both data types, and for a design with temperature measurements
@@
 this illustration, Table \ref{tab:sim_rhoab} shows that the expected
-Bernoulli variance gets lower with larger standard deviations
-$\sigma_1$ and $\sigma_2$. The expected reduction of Bernoulli variance is
-largest for the cases with high correlation $\gamma$. Albeit smaller,
+Bernoulli variance gets smaller with larger standard deviations. The expected reduction of Bernoulli variance is
+further largest for the cases with high correlation $\gamma$. Albeit smaller,
 there is also uncertainty reduction when only temperature is measured
@@
 
-We now turn to an example involving a full-fledged gaussian random
-field (GRF). The class of GRF model we consider have a linear trend
+We now turn to an example involving a full-fledged GRF. We consider a linear trend
 \begin{align*}
@@
 Z_{\spatloc, 1}\\ Z_{\spatloc, 2}
-\end{pmatrix}\right] &= \beta_0 + \beta_1 \spatloc
+\end{pmatrix}\right] &= \beta_0 + \beta_1 \spatloc,
 \end{align*}
@@
 \end{align*}
-where $k(., .)$ is one of the traditional spatial covariance kernel
-(exponential, Mat\'{e}rn, etc.) and $\gamma$ defines the
-cross-covariance structure. 
+where an isotropic Mat\'{e}rn 3/2 kernel $(1+\eta h)\exp (-\eta h)$ is used, for Euclidean distance $h$.
 In the accompanying Python
@@
 modeling assumptions can however be relaxed to anisotropic covariance and
-changing variance levels across the spatial domain. Anisotropy and
-non-stationary variance are both relevant for the setting with river
-plumes, but in practice this requires more parameters to be
-specified. With extensive satellite data or prior knowledge from high-resolution ocean models, one could also possibly fit more complex multivariate
+changing variance levels across the spatial domain. Both extensions are relevant for the setting with river
+plumes, but in practice this requires more parameters to be specified.
+With extensive satellite data or prior knowledge from high-resolution ocean models, one could also possibly fit more complex multivariate
 spatial covariance functions
@@
 
-In the rest of this section, we will consider a GRF with mean and
+In the rest of this section, we consider a GRF with mean and
 covariance structure as above and parameters
@@
 0.0 & -3.8
-\end{pmatrix},~ \sigma_1 = \sigma_2 = 2.25, ~ \gamma_0 = 0.2
+\end{pmatrix},~ \sigma_1 = 2.5,~ \sigma_2 = 2.25, ~ \gamma = 0.2,
 \end{align*}
-and the spatial covariance has unit variance and a Mat\'{e}rn 3/2 kernel, $(1+\eta h)\exp (-\eta h)$, for Euclidean distance $h$ and decay parameter $\eta=0.5$.
+and kernel parameter $\eta=3.5$.
 One realization of this GRF is shown in Fig.~\ref{example_excu}.
 
-We now study how the expected Bernoulli variance eq.\eqref{def:eibv}
+We now study how the EBV [Eqn.\eqref{def:eibv}]
 associated with data collection at a point changes if only one of the
@@
 is used to update the GRF model. We then consider the green triangle
-as a potential next observation location and plot the expected
-Bernoulli variance reduction (at each grid point/nodes in the the waypoint graph) that would result from observing only one component of the field (temperature or salinity), or both at that point.
+as a potential next observation location and plot the EBV reduction (at each grid node in the waypoint graph) that would result from observing only one component of the field (temperature or salinity), or both at that point.
 
@@
 
-Note that plotting the Bernoulli variance reduction at each point
+Note that plotting the EBV reduction at each point
 might also be used to compare different data collection plans. For
-example, Fig. \ref{fig:ebv_north_vs_east} shows the expected Bernoulli
-variance reduction associated with a data collection plan along a
-vertical line (static north) and one associated with a horizontal
+example, Fig. \ref{fig:ebv_north_vs_east} shows the EBV reduction associated with a data collection plan along a vertical line (static north) and one associated with a horizontal
 (static east). Both expectations are computed according to the
@@
   \caption{Pointwise Bernoulli variance reduction for two different observation plans.
-The prior excursion probability is shown in \ref{fig:ebv_static_excu}.
+The prior EP is shown in \ref{fig:ebv_static_excu}.
 EBV reduction for each design shown in \ref{fig:ebv_static_north} and \ref{fig:ebv_static_east}.}

### 2020-06-25 appendix.tex
Updates from Overleaf

@@
 \mathbb{E}[\mes(\es)^r]
-&=\int_{D^{r}} \jointExcuProb
+&=\int_{\domain^{r}} \jointExcuProb
 \productMeasure
@@
 \mathbb{E}[\mes(\es)^{r}]
-&=\mathbb{E}\left[\left(\int_{D} \mathbbm{1}_{\gp[u] \in T} ~d\mes(u) \right)^{r} \right]
+&=\mathbb{E}\left[\left(\int_{\domain} \mathbbm{1}_{\gp[u] \in T} ~d\mes(u) \right)^{r} \right]
 =\mathbb{E}\left[ \prod_{i=1}^{r} \left(
-        \int_{D} \mathbbm{1}_{\gp[\uu^{(i)}] \in T} ~d\mes(\uu^{(i)})
+        \int_{\domain} \mathbbm{1}_{\gp[\uu^{(i)}] \in T} ~d\mes(\uu^{(i)})
 \right) \right] \\
@@
 \mathbb{E}\left[
-\int_{D^{r}}
+\int_{\domain^{r}}
 \mathbbm{1}_{\gp[\uu^{(1)}] \in T,\dots, \gp[\uu^{(r)}]  \in T}
@@
 %&
-=\int_{D^{r}}
+=\int_{\domain^{r}}
 \jointExcuProb
@@
 
-Those two general results allow us to derive simple expressions for the expected effect of the inclusion of new datapoints on the $\ibv$ (Proposition \ref{propo_eibv}) and on the $\emv$ (Proposition \ref{propo_emv}) for which we provide proofs below.
+Those two general results allow us to derive simple expressions for the expected effect of the inclusion of new datapoints on the $\IBV$ (Proposition \ref{propo_eibv}) and on the $\EMV$ (Proposition \ref{propo_emv}) for which we provide proofs below.
 
@@
 \eibv_{[n]}(\bm{x})
-&=\int_{D}
+&=\int_{\domain}
 \currentExp{\futureProba{\gp[\uu]\in
         T}(1-\futureProba{\gp[\uu]\in T})} d\mes(u) \\
-&=\int_{D} \varPhi_{\no}\left(\bt;
+&=\int_{\domain} \varPhi_{\no}\left(\bt;
 ~\futureMean{\uu},
 \futureCov{u,u}\right) d\mes(u)\\
-&-\int_{D} \currentExp{
+&-\int_{\domain} \currentExp{
     \varPhi_{\no}\left(\bt;
@@
 \right);
-\mathbf{\Sigma}_{[n]}(\uu)
+\mathbf{\Sigma}_{[\stage]}(\uu)
 \right),
@@
 \begin{split}
-\eemv_{[n]}(\bm{x})
+\currentEEMV(\bm{x})
 %\operatorname{Var}[\mes(\es)]
-&=\int_{D^2} 
+&=\int_{\domain^2} 
 \varPhi_{2\no}
@@
 (u,v)\\
-&-\currentExp{\left( \int_{D} \varPhi_{\no}\left(\bt;\mu_{[n+1]}(u), K_{[n+1]}(u)\right) d\mes(u) \right)^2}.
-\end{split}
-\end{equation*}
-The second term can be worked out in semi-analytical form by writing it as
-%
-\begin{equation*}
-\begin{split}
-%\eemv_{[n]}(\bm{x})
-%\operatorname{Var}[\mes(\es)]
-&\int_{D^2} 
+&-\int_{\domain^2} 
 \currentExp{
-\varPhi_{\no}\left(\bt;\mu_{[n+1]}(u), K_{[n+1]}(u)\right)
-\varPhi_{\no}\left(\bt;\mu_{[n+1]}(v), K_{[n+1]}(v)\right)
+\varPhi_{\no}\left(\bt; \futureMean{\uu}, K_{[n+1]}(\uu, \uu)\right)
+\varPhi_{\no}\left(\bt; \futureMean{\vv}, K_{[n+1]}(\vv, \vv)\right)
 }
@@
 %\productMeasure 
-(u,v)\\
-=&
-\int_{D^2} 
-\currentExp{
-    \varPhi_{\no}\left(a_1+B_1 V; C_1\right)
-    \varPhi_{\no}\left(a_2+B_2 V; C_2\right)
-}
-\
-\mathrm{d}\mes^{\otimes} %\mes 
-%\productMeasure 
-(u,v),\\
+(u,v)
 \end{split}
 \end{equation*}
-and expanding the integrand (for fixed $n$, $\uu, \vv$) as
-\begin{equation*}
-\begin{split}
-&
-\currentExp{
-    \varPhi_{\no}\left(\bt;\mu_{[n+1]}(u), K_{[n+1]}(u)\right)
-    \varPhi_{\no}\left(\bt;\mu_{[n+1]}(v), K_{[n+1]}(v)\right)
-}\\
-=&
-\currentExp{
-    \varPhi_{\no}\left(a_1+B_1 V; C_1\right)
-    \varPhi_{\no}\left(a_2+B_2 V; C_2\right)
-}
-\end{split}
-\end{equation*}
-with $V=\gp[\bm{x}_{n+1}]-\currentMean{\bm{x}_{n+1}} \sim \mathcal{N}(0_{q_{n+1}},k_{[n]}(\bm{x}_{n+1},\bm{x}_{n+1}))$ and $a_1=\bt-\currentMean{\uu}$,
-$B_1=-\lambda_{[n+1,n+1]}(\uu)^T$, $a_2=\bt-\currentMean{\vv}$, $B_2=-\lambda_{[n+1,n+1]}(\vv)^T$.
+and the proof follows by applying Proposition \ref{propo3}
+with
+$$V=\gp[\bm{x}_{n+1}]-\currentMean{\bm{x}_{n+1}} \sim \mathcal{N}(0_{q_{n+1}},k_{[n]}(\bm{x}_{n+1},\bm{x}_{n+1}))$$
+and $a_1=\bt-\currentMean{\uu}$,
+$B_1=-\lambda_{[n+1,n+1]}(\uu)^T$, $a_2=\bt-\currentMean{\vv}$, $B_2=-\lambda_{[n+1,n+1]}(\vv)^T$, $C_1 = \currentCov{\uu, \uu}$, $C_2 = \currentCov{\vv, \vv}$.
 \end{proof}

### 2020-06-25 main_new.tex
Updates from Overleaf

@@
 particular, as an integral over EPs, the 
-$\emv = \operatorname{Var}[\mes(\es)]$ is:
+$\EMV = \operatorname{Var}[\mes(\es)]$ is:
 \begin{equation*}
 \begin{split}
-\emv
+\EMV
 &=\int_{\domain^2} \mathbb{P}\left(
@@
 \begin{split}
-\emv
+\EMV
 %\operatorname{Var}[\mes(\es)]
@@
 We compute the expected effect of the inclusion of new observations on
-the $\emv$ and $\ibv$ of the ES $\es$. Let us consider the
+the $\EMV$ and $\IBV$ of the ES $\es$. Let us consider the
 same setting as in Eqn. \eqref{eq:meanCoK} and \eqref{eq:varCoK}, and
@@
 \begin{equation}\label{def:eibv}
-    \currentEIBV(\bm{x}):=\mathbb{E}_{\stage}\left[\mbox{IBV}(\bm{x}; \bm{Y})\right],
+    \currentEIBV(\bm{x}):=\currentExp{\mbox{IBV}(\bm{x}; \bm{Y})},
 \end{equation}
@@
 \currentEIBV(\bm{x})
-&=\int_{D} \varPhi_{\no}\left(\bt;~\currentMean{\uu}, \currentCov{u, u}\right) d\mes(u)\\
-&-\int_{D} \varPhi_{2\no}
+&=\int_{\domain} \varPhi_{\no}\left(\bt;~\currentMean{\uu}, \currentCov{u, u}\right) d\mes(u)\\
+&-\int_{\domain} \varPhi_{2\no}
 \left(
@@
 \begin{split}
-\eemv_{[n]}(\bm{x})
+\currentEEMV(\bm{x})
 &=\int_{\domain^2} 
@@
 \left(
-(\bt, \bt); \mu((u,v)), 
+(\bt, \bt); ~ \mu((u,v)), 
 K((u,v),(u,v))
@@
 (u,v)\\
-&-\int_{\domain^2}\currentExp{
-    \varPhi_{\no}\left(a_1+B_1 V; K_{[n+1]}(u)\right)
-    \varPhi_{\no}\left(a_2+B_2 V; K_{[n+1]}(v)\right)
-}\mathrm{d}\mes^{\otimes}(u, v),
+&-\int_{\domain^2} \varPhi_{2\no}
+\left(
+\left(
+\begin{matrix}
+\bt-\currentMean{\uu}\\
+\bt-\currentMean{\vv}
+\end{matrix}
+\right);~
+\mathbf{\tilde{\Sigma}}_{[\stage]}(\uu, \vv)
+\right)
+\mathrm{d}\mes^{\otimes} %\mes 
+%\productMeasure 
+(u,v)
 \end{split}
 \end{equation*}
-with $V=\gp[\bm{x}]-\currentMean{\bm{x}} \sim \mathcal{N}(0_{q_{n+1}},k_{[n]}(\bm{x},\bm{x}))$ and $a_1=\bt-\currentMean{\uu}$,
-$B_1=-\lambda_{[n+1,n+1]}(\uu)^T$, $a_2=\bt-\currentMean{\vv}$, $B_2=-\lambda_{[n+1,n+1]}(\vv)^T$.
-
-This integrand boils down to an evaluation of the
-$2\no$-dimensional Gaussian CDF by Proposition~\ref{propo3} (Appendix). Details
-are omitted here for brevity as we do not implement this criterion
-further because it requires numerical integration over $\domain^2$.
+where the matrix $\mathbf{\tilde{\Sigma}}_{[n]}(\uu, \vv)$ is defined blockwise as
+\begin{equation*}
+\begin{split}
+\mathbf{\tilde{\Sigma}}_{[n]}(\uu, \vv)&=
+\left(
+\begin{matrix}
+\tilde{\Sigma}_{u,u} & \tilde{\Sigma}_{u,v}\\
+\tilde{\Sigma}_{v,u} & \tilde{\Sigma}_{v,v}
+\end{matrix}\right)
+\end{split}
+\end{equation*}
+with blocks
+\begin{equation*}
+\begin{split}
+\tilde{\Sigma}_{u, v} &= \lambda_{[n+1,n+1]}(\uu)^T k_{[n]}(\bm{x},\bm{x}) \lambda_{[n+1,n+1]}(\vv)\\
+\tilde{\Sigma}_{u, u} &= \lambda_{[n+1,n+1]}(\uu)^T k_{[n]}(\bm{x},\bm{x}) \lambda_{[n+1,n+1]}(\uu) + \futureCov{u, u}
+\end{split}
+\end{equation*}
+and the other blocks are defined by following the same logic, exchanging the role of $\uu$ and $\vv$.
 \end{propo}

### 2020-06-26 ans_to_reviewers/answers-to-reviewers.tex
Updates from Overleaf

@@
 %----------------------------------------------------------------------
-\title{Autonomous oceanographic Sampling Designs Using Excursion Sets for Multivariate Gaussian Random Fields
+\title{Autonomous Oceanographic Data Collection. Learning Excursion Sets of Vector-valued Gaussian Random Fields
   \\\vspace{5mm}
@@
 version of our manuscript.
-We have worked a lot on developing both theoretical results, computational resources, and on the presentation in the paper.
+We have worked a lot on developing both theoretical results, computational resources, and on the paper presentation.
 To help us in these efforts, we included Cedric Travelletti as co-author because he has contributed significantly on the revised version of the paper.
@@
 \item The revision has lead to {\bf{structural changes}} as follows: \\
-    i) previous section 1 and 2 are combined to one introduction which should help the narrative and highlight the main contributions of the work. \\
-    ii) merged the data assimilation part with equations for coKriging to clarify that this work is on multivariate GRFs. \\
+    i) previous Section 1 and 2 are combined to one Introduction which should help the narrative and highlight the main contributions of the work. \\
+    ii) merged the data assimilation part with equations for co-Kriging to clarify that this work is on multivariate GRFs. \\
     iii) Split the simulation study in two and used these as to directly motivate the statistical concepts of uncertainty reduction (Sect 2) and adaptive sampling strategies (Sect 3).
+    iv) Present mathematical contributions succintly in Section 2 and 3, while the mathematical proofs are in Appendix.
 
@@
 
-In the revised manuscript these aspects are largely rewritten into the Introduction, and partly also in Section 2. \\
-This change addresses comments of the Associate Editor and Rev. 1 in particular.
+In the revised manuscript these aspects are largely rewritten, in particular in the Introduction. \\
+This change addresses comments of Associate Editor and Rev. 1 in particular.
 
@@
 
-In the revised manuscript these aspects are largely rewritten in the Introduction, in Section 2 and 3, and on in the Closing remarks.
+In the revised manuscript these aspects are largely rewritten in the Introduction, in Section 2 and 3, and in the Closing remarks.
 The main body contains the novel statistical results for the expected uncertainty reduction in random sets exemplified by the EIBV, while the Appendix holds the proofs.
 \\
-This change addresses comments of the Editor, Associate Editor, Rev. 1, Rev. 2 and Rev. 3.
+This change addresses comments of Editor, Associate Editor, Rev. 1, Rev. 2 and Rev. 3.
 
@@
 
-The Introduction section contains pointers to the open-source code. \\
+The Introduction and simulations in Section 3 contain pointers to the open-source code. \\
 This change addresses comments of Rev. 1 and Rev. 2 in particular.
@@
 \reply: Based on the comment by you, AE and reviewers, we have restructured the Introduction and later parts of the paper to be more readable for professional statisticians.
-While doing so, we still strive to highlight how statistics can learn from new technologies like marine robotics and how robotics can benefit from new statistical developments.
+While doing so, we still highlight how statistics can learn from new technologies like marine robotics and how robotics can benefit from new statistical developments.
 
@@
 \reply: We realize that we did not stress the statistical novelty sufficiently in the initial version.
-The main statistical novelty is closed-form calculations for the design expected integrated Bernoulli variance (EIBV) for multivariate processes.
-In the revised version has clarified the statistical contributions via re-structured sections.
-Moreover, we extend the calculations to more general criteria (such as volumes, going beyond single location).
+The main statistical novelty is closed-form calculations for the design criterion of expected integrated Bernoulli variance (EIBV) for multivariate processes.
+In the revised version, we clarified the statistical contributions via re-structured sections.
+Moreover, we extend the statistical calculations to more general design criteria (such as volumes, going beyond single location).
 The main closed-form results are presented in the main body, but with proofs in the new Appendix.
@@
 In other related situations, for instance that of larger fjord or ocean fronts, which are at a larger scale, there would usually be more data available for instance in the form of satellite data or from computer model simulations.
-(We added references to related papers that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here.) For such cases, we agree that it would make sense to embed more nuanced models onboard the AUV, for instance that of non-stationary models or anisotropic random fields, but for our field trial here we attempted to have as few parameters.
+For such cases, we agree that it would make sense to embed more nuanced models onboard the AUV, for instance that of non-stationary models or anisotropic random fields, but for our field trial here we attempted to have as few parameters.
 
@@
 
-\reply: Based on the comments by yourself, Associate Editor and referees, we realize that the motivation part of the paper was too long.
+\reply: Based on the comments by yourself, AE and referees, we realize that the motivation part of the paper was too long.
 We have merged the previous Introduction and Section 2 to the current Introduction, and we cut redundant parts in this process.

### 2020-06-26 main_new.tex
Updates from Overleaf

@@
 % "New Title of the paper"
-\title{Autonomous Oceanographic Data Collection Estimating Excursion Sets of Vector-valued Gaussian Random Fields} 
+\title{Autonomous Oceanographic Data Collection. Learning Excursion Sets of Vector-valued Gaussian Random Fields} 
 \runtitle{Autonomous Oceanographic Data Collection}
@@
   the combination of statistics and autonomous systems provide new
-  opportunities for experimental design. In this work we develop
-  methods for efficient spatial sampling applied to the mapping of
-  coastal ocean processes by providing informative descriptions of
-  spatial characteristics of ocean phenomena. Specifically, we define
-  a design criterion based on uncertainty in the excursions of
-  vector-valued Gaussian random fields, and derive tractable
-  expressions for the expected integrated Bernoulli variance reduction in such a
-  framework. We demonstrate how this criterion can be used to
-  prioritize sampling efforts at locations that are ambiguous, making
-  exploration more effective. We use simulations to study the
-  properties of methods and to compare them with state-of-the-art
-  approaches, followed by results from field deployments with an
-  autonomous underwater vehicle as part of a study mapping the
-  boundary of a river plume. The results demonstrate the potential of
-  combining statistical methods and robotic platforms to effectively
-  inform and execute data-driven environmental sampling.
+  opportunities for experimental design. 
+  %
+  %In this work we develop methods for efficient spatial sampling applied to the mapping of coastal ocean processes by providing informative descriptions of spatial characteristics of ocean phenomena.
+  %
+  % REPLACED BY 
+  %
+  In this work we develop efficient spatial sampling methods for characterizing regions defined by simultaneous exceedances above prescribed thresholds of several responses, with an application focus on mapping coastal ocean phenomena based on temperature and salinity measurements.
+  % EN REPLACEMENT
+  Specifically, we define a design criterion based on uncertainty in the excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected integrated Bernoulli variance reduction in such a framework.
+We demonstrate how this criterion can be used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective.
+  We use simulations to study and compare properties of the considered approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a study mapping the boundary of a river plume.
+The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.
 
@@
 domain with the aim to pursue measurements with high scientific
-relevance.  Spatial statistics and methods for experimental design can
-clearly contribute to this sampling challenge.
+relevance.  Models and methods from spatial statistics and experimental design can clearly contribute to this sampling challenge.
 
@@
 system generated by a river plume. Fig.~\ref{fig:nidelven} shows the
-survey area in Trondheim, Norway, where cold freshwater enters from
-the river, creating a strong gradient in both temperature and
+survey area in Trondheim, Norway, where cold freshwater enters the fjord from a river, creating a strong gradient in both temperature and
 salinity. Because of the local topography and the Coriolis force the
@@
 highly uncertain, making deterministic planning challenging. The goal
-is therefore, to use AUV measurements for improved description of the
+is therefore to use AUV measurements for improved description of the
 interface between fresh and oceanic waters.  It is often not possible
@@
 multivariate processes are also abundant in other contexts, for
-instance in medicine, where doctors do not rely solely on a single
+instance in medicine, where physicians do not rely solely on a single
 symptom but must see several combined effects before making a
@@
 The questions tackled here hence pertain to the broader area of
-spatial data collection and experimental designs for vector-valued
-random fields.
+spatial data collection for vector-valued random fields.
 %Because of limited resources to sample the large oceanographic domains, one must plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs.
@@
 
-While random field modeling has been one of the main topics throughout
-the history of spatial statistics \citep{Krige1951a,Stein1999}, even
+While random field modeling has been one of the main topics throughout the history of spatial statistics \citep{Krige1951a,Stein1999}, even
 for vector-valued random field models with associated prediction
@@
 there has lately been a renewed interest for random field models in
-the context of static or sequential experimental design, be it in the
-context of spatial data collection \citep{Mueller2007} or in
+the context of static or sequential experimental design, be it in the context of spatial data collection \citep{Mueller2007} or in
 simulation experiments \citep{Santner.etal2003}. As detailed in
@@
 basis to sequential design of simulations dedicated to various goals
-such as global optimization and set estimation
-\citep{Adler.Taylor2007}. Of particular relevance to our context,
-\cite{Bect.etal2012} focuses on strategies to reduce uncertainty in
-volumes of excursion exceeding a prescribed threshold, while
-\cite{chevalier2014fast} contributes a computationally efficient
-implementation of sequential strategies. Rather than focusing on
-excursion volumes, approaches were investigated in
+such as global optimization and set estimation. Of particular relevance to our context,
+\cite{Bect.etal2012} focuses on strategies to reduce uncertainties related to volumes of excursion exceeding a prescribed threshold, while \cite{chevalier2014fast} concentrates on making the latter strategies computationally efficient and batch-sequential.
+Rather than focusing on excursion volumes, approaches were investigated in
 \cite{French.Sain2013,Chevalier.etal2013b,Bolin.Lindgren2015,Azzimonti.etal2016}
@@
 strategies in the case of vector-valued observations. It has been long
-acknowledged that co-Kriging could be used efficiently for
-assimilating sequential data \citep{Vargas-Guzman1999}, but sequential
+acknowledged that co-Kriging could be updated efficiently in the context of sequential data assimilation \citep{Vargas-Guzman1999}, but sequential
 strategies for estimating features of vector-valued random fields are
@@
 possibilities of stepwise uncertainty reduction strategies for ES's in
-the context of designing simulations, but the outputs are mainly
+the context of designing fire simulations, yet outputs are mainly
 assumed independent.
@@
 at the statistical prediction of features using an underwater glider.
-In this paper the focus is not on statistical modeling, but rather on
-statistical properties of computations for efficient data
-collection. We combine novel possibilities in marine
+In this paper the main focus is not on statistical modeling per se, but rather on statistical principles and computations underlying efficient data collection. We combine novel possibilities in marine
 robotics with spatial statistics and experimental design to provide
@@
 the statistics literature on spatio-temporal design as well
-\citep{bueso1998state,zidek2019monitoring}. However, variance and
-entropy reduction are independent of the actual data realizations
-under the assumptions of GRF models, so it has limited flexibility for
-active adaptation of trajectories based on what is measured.  The use
+\citep{bueso1998state,zidek2019monitoring}.
+However, response variance and entropy being depending only in GRF models on measurement locations and not on response values, criteria based on them only tend to have limited flexibility for
+active adaptation of trajectories based on measurement values.  The use
 of data-driven adaptive criteria was introduced to include more
@@
 
-\section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
+\section{Quantifying uncertainty on Excursion Sets implicitly defined by GRFs}
 \label{sec:ESEP}
@@
 Fig.~\ref{fig:real_temp} and \ref{fig:real_sal} shows a realization of such a
-vector-valued GRF on $\domain=(0,1)^2$. Fig.~\ref{fig:jointex_roi} represents a by-product of interest derived from these
+vector-valued GRF on $\domain=[0,1]^2$. Fig.~\ref{fig:jointex_roi} represents a by-product of interest derived from these
 realizations, namely regions: i) in red, where both temperature and salinity are high (i.e., exceeding respective thresholds), indicative of ocean water, ii) in white, where both temperature and salinity are low, indicative of riverine water, and iii) in light-red, where one variable is above and the other below their respective thresholds, indicative of mixed waters.
@@
 the $\EMV$ and $\IBV$ of the ES $\es$. Let us consider the
-same setting as in Eqn. \eqref{eq:meanCoK} and \eqref{eq:varCoK}, and
+same setting as in Eq. \eqref{eq:meanCoK} and \eqref{eq:varCoK}, and
 let $\currentExp{.}$ and $\currentProba{.}$ denote conditional
@@
 \begin{matrix}
-\tilde{\Sigma}_{u,u} & \tilde{\Sigma}_{u,v}\\
-\tilde{\Sigma}_{v,u} & \tilde{\Sigma}_{v,v}
+\tilde{\Sigma}_{1,1}(\uu, \uu) & \tilde{\Sigma}_{1,2}(\uu, \vv)\\
+\tilde{\Sigma}_{2,1}(\vv, \uu) & \tilde{\Sigma}_{2,2}(\vv, \vv)
 \end{matrix}\right)
@@
 \end{equation*}
-with blocks
+with blocks given by
 \begin{equation*}
 \begin{split}
-\tilde{\Sigma}_{u, v} &= \lambda_{[n+1,n+1]}(\uu)^T k_{[n]}(\bm{x},\bm{x}) \lambda_{[n+1,n+1]}(\vv)\\
-\tilde{\Sigma}_{u, u} &= \lambda_{[n+1,n+1]}(\uu)^T k_{[n]}(\bm{x},\bm{x}) \lambda_{[n+1,n+1]}(\uu) + \futureCov{u, u}
+\tilde{\Sigma}_{i,j}(u, v) &= \lambda_{[n+1,n+1]}(\uu)^T k_{[n]}(\bm{x},\bm{x}) \lambda_{[n+1,n+1]}(\vv) + \delta_{i,j}\futureCov{\uu, \vv}
 \end{split}
@@
 
-We illustrate the expected Bernoulli variance (EBV) associated with different designs on a simple
+We illustrate the expected Bernoulli variance (EBV) associated with different designs on a
 bivariate example. This mimics our river plume application and hence
@@
 
-We now turn to an example involving a full-fledged GRF. We consider a linear trend
+We now turn to an example involving a full-fledged GRF. The statistical model we consider has a linear trend
 \begin{align*}
@@
 One realization of this GRF is shown in Fig.~\ref{example_excu}.
-
-We now study how the EBV [Eqn.\eqref{def:eibv}]
+In the computed examples, the spatial domain $\domain$ is
+discretized to a set of $N$ grid locations
+$\mathcal{M}_g = \{\x_i, i=1,\ldots,N \}$, where each cell has area
+$\delta$; the same grid is used for the waypoint graph for possible
+design locations. The EIBV is approximated by sums over all grid
+cells.
+
+We now study how the EBV [Eq.\eqref{def:eibv}]
 associated with data collection at a point changes if only one of the
@@
 
-In all the computed examples, the spatial domain $\domain$ has been
-discretized to a set of $N$ grid locations
-$\mathcal{M}_g = \{\x_i, i=1,\ldots,N \}$, where each cell has area
-$\delta$; the same grid is used for the waypoint graph for possible
-design locations. The EIBV is approximated by sums over all grid
-cells. \kc{This should be promoted up a couple of paras and would
-  address my last comment above.}
-
 \section{Sequential designs and heuristic path planning}\label{sec:heuristics}
 
-The goal of this section is to present sequential data collection
+We present sequential data collection
 strategies that aim at reducing the expected uncertainty on the target
@@
 already been performed and one wants to choose what data to collect
-next. After each step, the GRF model is updated using co-Kriging
-equations \ref{eq:meanCoK} and \ref{eq:varCoK}. The design evaluations
+next. The design evaluations
 are based on the conditional expectation $\currentExp{.}$ from the law
-of the field, conditional on the data gathered up to (and including)
+of the field, conditional on all data available at
 stage $n$.
+Once the best design at stage $n$
+has been selected, the data are collected and the GRF
+model is updated using co-Kriging
+Eq. \ref{eq:meanCoK} and \ref{eq:varCoK}, yielding a conditional law $\mathbb{P}_{\stage + 1}$
+after which the process is repeated.
 
@@
 type (all components of the field at a single location, only some
-components at a subset of the selected locations, etc.) since the
-concept of \textit{generalized location} allows the co-Kriging
-equations to handle all those case by putting them on equal footing.
+components at a subset of selected locations, etc.) because of the
+concept of \textit{generalized location} in the co-Kriging expressions.
 In general, a design strategy must choose the spatial location as well
@@
 the neighboring spatial location (waypoints) at each stage, and all
-components (temperature and salinity) of the field are observed there
+components (temperature and salinity) of the field are observed
 (isotopic). The candidate points at this stage are denoted
 $\candidates$ as defined from the 6 directions (apart from edges) in
-the waypoint graph (see Fig.~\ref{fig:wp_graph_a} for an example). The
+the waypoint graph (see Fig.~\ref{fig:wp_graph_a}). The
 set $\candidates$ depends on the current location, but for readability
@@
 because of the enormous growth over stages (see
-e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}).
+e.g. \cite{powell2016perspectives}).
 Instead, we outline heuristic strategies.
@@
 The EIBV is efficiently computed for each of the candidate points
-$\candidates$ using Proposition \ref{propo2}. Once the best location
-has been selected, a stage of observation is performed and the GRF
-model is updated, yielding a conditional law $\mathbb{P}_{\stage + 1}$
-after which the process is repeated.
-
+$\candidates$ using Proposition \ref{propo2}. 
 Even though this myopic strategy is non-anticipatory, it still
 provides a reasonable approach for creating designs in many
-applications. Moreover, it can be implemented without any demands on
+applications. Moreover, it can be implemented without too much demand on
 computational power, making it well-suited for embedding on an AUV.
@@
 measurements, which is optimal in that it accounts consistently for
-the expectations and minimizations in these two stages, but
-nevertheless without including any planning beyond those two steps.
-
-Assume data have been collected up to stage $n$. The principle of
-two-step look-ahead is to select the next observation location
-$\spatloc_{\stage + 1}$ that yields the biggest reduction in EIBV if
-we were to (optimally) add one more observation after the one at
-$\spatloc_{\stage + 1}$. In order to formalize this concept, we must
+the expectations and minimizations in these two stages, but no anticipation beyond that.
+
+The principle of
+two-step look-ahead is to select the next observation location $\spatloc_{\stage + 1}$ that yields the biggest reduction in EIBV if
+we were to (optimally) add one more observation after that again. In order to formalize this concept, we must
 extend the notation for EIBV in the future (after observation
-$\stage + 1$ has been made). We let $\currentEIBV(\cdot; u, y)$ to
+$\stage + 1$ has been made). We let $\currentEIBV(\cdot; u, y)$
 denote the EIBV where expectations are taken conditional on the data
-available at $n$ and on an additional observation $y$ at $u$.
+available at stage $n$ and on an additional observation $y$ at $u$ at stage $n+1$.
 
@@
       \end{align}
-      where $Y$ is a random variable distributed according to the
-      conditional law of $\gp_{\spatloc}$ at step $\stage$ with the
+      where $Y$ is the random data realization of $\gp_{\spatloc}$ according to its
+      conditional law at step $\stage$ with the
       dependence of the set of candidates on the current location
@@
 available as in Proposition \ref{propo_eibv}. However, for the sequential
-approaches this is not the case. For comparison the properties are
+approaches this is not the case. For comparison, the properties are
 therefore evaluated using Monte Carlo sampling over several replicates
@@
 two variables. It is important to note that the objective function
-used by the robot. is focused on reducing the
+used by the AUV is focused on reducing the
 EIBV, but we nevertheless expect that we will achieve good predictive
@@
 criterion that is important for practical purposes is the computational
-time needed for the strategy, as this will impact the performance for
-an embedded system.
+time needed for the strategy.
 
@@
 Each strategy is conducted on an equilateral grid as shown in
-Fig.~\ref{fig:wp_graph}. The AUV starts at the center East-West
-coordinate at the southern end of the domain (node 53). It then moves
+Fig.~\ref{fig:wp_graph}. The AUV starts at the center coordinate at the southern end of the domain (node 53). It then moves
 along edges in the waypoint graph while collecting data which are
-assimilated into the onboard GRF model. This is used in the evaluation
+assimilated onboard to update the GRF model. This is used in the evaluation
 of the next node to sample.  The procedure is run for $10$ stages. A
@@
 resulting drop in realized IBV for each of the six strategies. IBV
-reduction occurs most under the \textit{myopic} and
+reduction is largest for the \textit{myopic} and
 \textit{look-ahead} strategies, each performing almost equally; this
@@
 able to do so in reasonable time, which allows for real-world
-applicability. The \textit{look-ahead} is slower, reaching levels that
-are nearly impractical for execution on a vehicle. Some pruning of the
+applicability. The \textit{look-ahead} strategy is much slower, reaching levels that
+are nearly impractical for execution on an AUV. Some pruning of the
 graph is performed to improve the performance, such as ruling out
@@
 Table~\ref{tab:sim_rhoab}, the current setting includes spatial
-correlation and this likely reduces the added effect of having
-bivariate data. However, it seems that having temperature data alone
+correlation and this likely strengthen the effect of temperature information. However, it seems that having temperature data alone
 does a substantially worse job in terms of explained variance.
@@
 experiments were conducted in late Spring 2019, when there is still
-snow melting in the surrounding mountains so that the river water is
-substantially colder than the water in the fjord. The experiment was
+snow melting in the surrounding mountains so that the river water is colder than the water in the fjord. The experiment was
 focused along the frontal zone that runs more or less parallel to the
-eastern shore as noted in Fig.~\ref{fig:nidelven}.
+eastern shore.
 
@@
 structures.
-Based on the initial runs we get a reasonable idea of the temperature and salinity of river and ocean waters, and also specify the mean by linear regression, where both temperature and salinity
-were both assumed to increase linearly with the west location on the waypoint graph.
+Based on the initial runs we get a reasonable idea of the temperature and salinity of river and ocean waters, and also specify the trend by linear regression, where both temperature and salinity were assumed to increase linearly with the west coordinate.
 Next, the residuals from the regression analysis were analyzed to specify the covariance parameters of the GRF model.
-
 %\begin{figure}[!h] 
@@
 %\end{figure}
-
 \begin{figure}[!b]
@@
 \end{figure}
-
 Fig.~\ref{fig:parest} summarizes diagnostic plots of this
@@
 This work builds on a multidisciplinary effort combining statistical
-methods with robotic survey techniques for oceanographic
+methods with robotic surveying techniques for oceanographic
 applications. We show how observation practices can gain efficiency
@@
 spatial monitoring. We further demonstrate the opportunities available
-for real-time multivariable spatial data gathering and analysis on
+for real-time multivariate spatial data gathering and analysis on
 autonomous platforms, which statisticians can exploit to create
@@
 connected to the properties of water masses. The characterization of
-uncertainties in random sets is extended with new results for the
+uncertainties in random sets is extended in the vector-valued case with new results for the
 expected integrated Bernoulli variance reduction achieved by spatial
-sampling designs. This is provided in closed-form for the situation
+sampling designs. This is provided in semi-analytical form for the situation
 with a static design, and then extended to the adaptive situation. The
@@
 domains are not methodologically different, but the operation must
-likely approximate calculations by integrating terms only in the
+likely approximate calculations by concentrating numerical integration efforts on terms in the
 vicinity of the autonomous vehicle \citep{fossum18b}. We did not
@@
 processes which could still be run onboard and augmented by dynamical
-models. If high-resolution ocean model or satelite data are available, it is hopefully possible to fit such more complex statistical models.
+models. If high-resolution ocean model or satellite data are available, it is hopefully possible to fit such more complex statistical models.
 
@@
 TOF acknowledges support from the Centre for Autonomous Marine
-Operations and Systems
-(AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of
-Excellence, project number 223254, and the Applied Underwater Robotics
-Labortatory (AURLab). JE and KR acknowledge support from Norwegian
-research council (RCN), project number 305445. CT and DG acknowledge
+Operations and Systems (AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of
+Excellence, project number 223254, and the Applied Underwater Robotics Labortatory (AURLab). CT and DG acknowledge
 support from the Swiss National Science Foundation, project number
-178858.
+178858. JE and KR acknowledge support from Norwegian
+research council (RCN), project number 305445.
+DG would like to aknowledge support of Idiap Research Institute, where he had is main affiliation when the first version of this work was submitted.
+The authors would also like to thank Niklas Linde (University of Lausanne) for pointing out the relevance of revisiting the proposed approach for estimating volumes of mixing zones, e.g.
+in hydrology.
 

### 2020-06-26 ans_to_reviewers/answers-to-reviewers.tex
Updates from Overleaf

@@
 The revised version has an entirely rewritten Introduction, and Section 2 now includes general background for random sets and excursion sets.
-We added references, for instance \cite{adler2009random} on ESs and \cite{Molchanov2005} on random sets.
-The revised version provides results that go beyond the EIBV (with proofs in the new Appendix), but the EIBV is used in experiments because it is computationally tractable onboard the AUV.
+(We added references, for instance \cite{Molchanov2005} on random sets.) The revised version provides results that go beyond the EIBV (with proofs in the new Appendix), but the EIBV is used in experiments because it is computationally tractable onboard the AUV.
 
@@
 
-\reply: We agree that they are not necessary for the methodological contributions, so we now bring up this model specification when needed for the examples. 
-In the open-source code that is now available, we implement non-stationary and non-isotropic covariance models.
-These represent situations that have some relevance for our case.
+\reply: We agree that the detailed statistical modeling assumptions are not necessary for the methodological contributions, so we now bring up this model specification when needed for the examples. 
+In the open-source code that is now available, we also implement non-stationary and anisotropic covariance models.
+These represent situations that have some relevance for our river plume case.
 The code we post along with the paper facilitate testing so users can input more complex models to see how it influences designs.
@@
 
-\reply: We clarify this in the revised document. The design is restricted to the waypoint graph with about 3 samples per batch.
+\reply: We clarified this in the revised document. The design is restricted to the waypoint graph with about 3 samples per batch.
 
@@
 \reply: We clarified the parameter specification.
-In practice the linear regression trend is fit from the levels at the boundary of the waypoint graph (in our case riverine water and ocean water).
+In practice the trend captures the temperature and salinity levels on the domain defined by the waypoint graph which is constructed to cover both riverine water and ocean water.
 For the field experiment (Section 4) we clarified this specification.
@@
 
-\reply: We have made this change to the citations: \cite{genton2015cross} and \cite{french2016credible}.
+\reply: We have made this change to the citations: \cite{genton2015cross} and \cite{french2013spatio}.
 \end{itemize}
@@
 
-\reply: We have improved and augmented the discussion on other design criteria and why the ES related EIBV is used in our case.
+\reply: We have improved and augmented the discussion on other design criteria and why the ES related EIBV criterion is used in our case.
 The paper also contains statistical results that go beyond this criterion (with proofs in the new Appendix), but the EIBV is used in practice because it is computationally tractable onboard the AUV.
@@
 
-\reply: We added reference to \cite{adler2009random} on random sets and motivated this topic for our application and more general settings.
-See our general point 2.
+\reply: We added reference \cite{Molchanov2005} on random sets and motivated this topic for our application and more general settings.
+See our general Point 2.
 and 3.
@@
 There are finally new elements in tailoring this sequential design methods in the context of the AUV application with its computational and navigation limitations.
-See our general comment 2 and 3 on top.
+See our general Point 2 and 3 on top.
 
@@
 
-\reply: The traditional entropy calculations for this GP aim to reduce the log determinant of the updated covariance matrix at the grid locations.
+\reply: The traditional entropy calculations for this Gaussian Random Field model aim to reduce the log determinant of the updated covariance matrix at the grid locations.
 It will not vary with the data realizations, so in our setting this information criteria will always give the same path.
@@
 Still, sampling approaches using hybrid criteria or multi-attribute criteria might balance goals in such situations, but we leave that for future research.
-(This is briefly mentioned as further work in Closing remarks.)
+(This is mentioned as further work in Closing remarks.)
 
@@
 In other related situations, for instance that of larger fjord or ocean fronts, which are at a larger scale, there would usually be more data available for instance in the form of satellite data or from computer model simulations.
-(We added references to related papers that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here.) That said, we do provide some validation of the model via residual analysis.
+That said, we do provide some validation of the model via residual analysis.
 Moreover, the main contribution of this paper is on design more than modeling.
-In the online resources with open-source code, there are demos using more complex covariance models (non-stationary variance and non-isotropic random fields).
+Finally, the online resources with open-source code have demos using more complex covariance models (non-stationary variance and anisotropic random fields).
 
@@
 \reply: We have chosen not to explore any re-estimation of parameters.
-The resources onboard do not allow such computationally demanding evaluations.
+The resources onboard the AUV do not allow such computationally demanding evaluations.
 Overall, we believe the capabilities and limitations of the AUV are clarified in the revised version of the paper.
@@
 \reply: We agree that this is a simple model for plume temperature and salinity variation.
-However, it is not common to have ocean model data or satellite data at this spatial resolution, which would be required for training.
+However, it is not common to have ocean model data or satellite data at this spatial resolution, which would be required for detailed model specification.
 So rather than overfitting, we have chosen a parsimonious model.
@@
 
-\reply: There are ongoing efforts to provide high-resolution ocean models which would be required for such detailed model specification.
-But as pointed out above, we believe that spatio-temporal variations would be a more realistic extension, especially if repeated satellite data or ocean modeling is available.
+\reply: High-resolution ocean models would be required for such detailed model specification, and it might be doable with large efforts in careful numerical implementation.
+But we believe that spatio-temporal variations would be a more realistic extension.
 The methodological approach for sampling designs, which is the main contribution of the current paper, would however still be the transferable.
@@
 \reply: We avoid the term' joint excursion sets' in the revised version.
-The revised version build on a more general notation, but link to the ES for temperature and salinity whenever it is appropriate to exemplify concepts and show results.
+The revised version build on a more general notation around random sets and EIBV for multivariate random fields, but we link to the ES for temperature and salinity whenever it is appropriate to exemplify concepts and show results.
 
@@
 
-\reply: Preliminary runs set these trends. We have clarified the description in the paper. In a larger scale, one could here use ocean model inputs or satellite data.
+\reply: A preliminary run was used to set these trends. We have clarified the description in the paper. At a larger scale, one could here use ocean model inputs or satellite data.
 
@@
 In many applications it is important to find zones with joint high values of many variables.
+In our context, the temperature and salinity are proxies for biological variables which are not easily measured.
 
@@
 Statisticians will know these things and users will not care.\par
-\reply: We collapsed the previous sections 3, 4 and 5 to new Section 2 and 3 containing the statistical novelty on EIBV closed-form expressions. The proofs are in a new Appendix.
+
+\reply: We collapsed the previous Section 3, 4 and 5 to new Section 2 containing the statistical novelty on EIBV closed-form expressions. The proofs are in the new Appendix.
 
@@
 
-\reply: We have clarified information about AUVs and their computational capabilities in Section 4. 
+\reply: We have clarified the description about AUVs and their computational capabilities in Section 4. 
 It is a multicore GPU NVIDIA Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM). 
@@
 \reply: Maximum speed and payload are limiting coverage.
-This ranges to minutes to 24h for larger crafts.
-Time constraints for operations are typically related to costs of ship time and man-hours during operation.
+For AUV operations in general, this ranges from minutes to 24h for larger crafts.
+Time constraints for operations are case specific, and also typically related to costs of ship time and man-hours during operation.
 \vspace{1em}
@@
 
-\reply: AUVs as in our application are free to move in 6 degrees of freedom on the triangulated graph.
-This is shown in Fig.1 and Fig.
+\reply: The AUV used in our application is free to move in 6 degrees of freedom on the triangulated graph.
+This is shown in Fig.
 4 with the waypoint graph.
 (They are essentially flying under water.
-In oceanographic application they also perform yo-yo movements in depth.)
+In larger-scale oceanographic applications they also perform yo-yo movements in depth.)
 \vspace{1em}
@@
 \reply: The coverage available is usually limited by man-hours to follow operations, but for large-time operation also on battery capacity.
-The important question here is "How do one prioritize the correct points to visit?", which is central to the development of the EIBV and the application shown in the paper.
+The important question here is in any case "How do one prioritize the correct points to visit?", which is central to the development of designs, and we believe the EIBV which we develop here is a useful criterion in several settings.
 \vspace{1em}
@@
 
-\reply: The waypoint grid size $n$ should be small enough to resolve the process that is of interest and to avoid aliasing.
+\reply: The waypoint grid size should be small enough to resolve the process that is of interest and to avoid aliasing.
 Usually it is in the 100s or 1000s.
-A guiding rule of thumb provided by the the Nyquist theorem is to sample at least twice in time for the shortest significant record period, and twice in space for the shortest significant length to either resolve or eliminate (by filtering) scales of oceanic variability shorter than those being studied.
+A guiding rule of thumb provided by the Nyquist theorem is to sample at least twice in time for the shortest significant record period, and twice in space for the shortest significant length to either resolve or eliminate (by filtering) scales of oceanic variability shorter than those being studied.
 In practice, this means mapping at an adequate spatial resolution faster than significant changes – in the phenomena – occur.
@@
 
-\reply: As is common in such field trials with AUVs, we gathered data in an initial run that also tests the depth and bearings navigation systems.
-From this data, the model parameters were tuned and the AUV conducted adaptive sampling using the optimization strategy.
+\reply: As is common in such field trials with AUVs, we gathered data in an initial run that is also used to test the depth and bearings navigation systems on the AUV.
+(With the low salinity in the river this is not always trivial as the AUVs can sink.) From this initial data, the model parameters were tuned and the AUV started conducting adaptive sampling using the optimization strategy.
 \vspace{1em}
@@
 \reply: We agree that this is a simplified model.
-However, one would need high-resolution ocean model data or satellite information to fit a more complex model.
-But in the accompanying code we run cases with more complex covariances, so the interested user can test if / how it matters for the design.
+However, one would need ocean model data or satellite information to fit a more complex model, and this is often not available at our spatial resolution.
+However, in the accompanying Python code we run cases with more complex covariances (non-stationary variance and anisotropic covariance), so the interested user can test if / how it matters for the design.
 We believe spatio-temporal models would be a very interesting next step and present that as further work.
@@
 
-\reply: The temperature and salinity are dependent in the GRF model.
-But the temperature and salinity data are collected from different instruments, and we assume that they are conditionally independent, given the GRF variables.
+\reply: The temperature and salinity are dependent in the Gaussian Random Field field model.
+But the temperature and salinity data are collected from different instruments, and we assume that they are conditionally independent, given the random field variables.
 That said, other measurements might be extracted or processed from common sensor readings and would then be conditionally dependent.
-We mentioned this briefly in the model description.
+We mentioned this briefly in the model assumptions descriptions.
 
@@
 \reply: The update formula assimilates relatively modest data sizes at each waypoint.
-Moreover, only the near-surface grid is considered here (two dimensional spatial domain).
+Moreover, only the near-surface grid is considered here (two dimensional spatial random field model).
 This means that the computational requirements are mainly driven by the number of combinations for sampling designs, not the size of the domain or the data dimension.
@@
 \reply: These parts of the paper are largely rewritten.
-We have generalized the closed-form solutions.
-(See point 2 and 3 at the start of reply letter.) The main body should contain results of general interest as well as provide insight about the statistical novelty.
-The proofs are now in Appendix with a succinct notation.
+We have also generalized the closed-form solutions.
+(See point 2 and 3 at the start of reply letter.) The main body contains results of general interest and provide insight about the statistical novelty.
+The proofs are in Appendix with a succinct notation.
 
@@
 \reply: In the revised version we provide closed-form solutions to more general criteria (beyond the EIBV), such as volume calculations involving coupling between locations.
-But as you mention, the real-world application has computational limitations, and we still use marginals in the examples.
+But as you mention, the real-world application has computational limitations, and we still use marginal distributions and EIBV in the examples.
 
@@
 
-\reply: We have removed this figure because it confused the explanation. The online resources contain movies of sampling paths, and the code allows users to generate sampling results by themselves.
+\reply: We have removed this figure because it confused the explanation.
+The online resources contain more results, including sampling paths, and this code allows users to generate sampling results by themselves.
 
@@
 
-\item[2.15.1] “closed-form” is sometimes written as “closed-form” (e.g., p. 7, p. 18)
+\item[2.15.1] “closed-form” is sometimes written as “closed form” (e.g., p. 7, p. 18)
 
@@
 
-\reply: In the revised 'Closing remarks', we elaborate more on computational approximations required for three-dimensional way-point graphs and space-time modeling including relevant references \citep{sigrist2015stochastic,richardson2017sparsity}, on using more complicated criteria for the data collection, and thoughts related to finite horizon optimization where there has been much creative work in the robotics literature including probabilistic road maps and rapidly-exploring random trees \citep{karaman2011sampling} whose statistical properties are not yet clear.
+\reply: In the revised 'Closing remarks', we elaborate more on: \\
+i) computational approximations required for three-dimensional way-point graphs and space-time modeling including relevant references \citep{sigrist2015stochastic,richardson2017sparsity} \\ 
+ii) using more complicated criteria for the data collection \\
+iii)
+finite horizon optimization where there has been much creative work in the robotics literature including probabilistic road maps and rapidly-exploring random trees \citep{karaman2011sampling} whose statistical properties are not yet clear.
 
-We have also strengthened the discussion on communication for the AUV.
+We have further strengthened the discussion on communication for the AUV.
 In doing so, we relate the topics of autonomy and communication to the modeling and computational assumptions that applies to this paper.
@@
 In the accompanying code (\url{https://cedrictravelletti.github.io/Submarine/} )
-there are showing various sampling results and for displaying other outputs if the reader is interested in comparing results with for instance univariate excursion probabilities.
+one can visualize various sampling results and also display other outputs such as instance univariate excursion probabilities.
 

### 2020-06-27 main_new.tex
updates

@@
 \usepackage{cleveref}
+\usepackage{xcolor}
 \input{macros}
@@
 % "New Title of the paper"
-\title{Autonomous Oceanographic Data Collection. Learning Excursion Sets of Vector-valued Gaussian Random Fields} 
+\title{Autonomous Oceanographic Data Collection. Learning Excursion
+  Sets of Vector-valued Gaussian Random Fields \\or\\
+Learning Excursion Sets of Vector-valued Gaussian Random Fields for
+Autonomous Ocean Sampling
+}
+
 \runtitle{Autonomous Oceanographic Data Collection}
@@
   %
-  In this work we develop efficient spatial sampling methods for characterizing regions defined by simultaneous exceedances above prescribed thresholds of several responses, with an application focus on mapping coastal ocean phenomena based on temperature and salinity measurements.
+  In this work we develop efficient spatial sampling methods for
+  characterizing regions defined by simultaneous exceedances above
+  prescribed thresholds of several responses, with an application
+  focus on mapping coastal ocean phenomena based on temperature and
+  salinity measurements.
   % EN REPLACEMENT
-  Specifically, we define a design criterion based on uncertainty in the excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected integrated Bernoulli variance reduction in such a framework.
-We demonstrate how this criterion can be used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective.
-  We use simulations to study and compare properties of the considered approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a study mapping the boundary of a river plume.
-The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.
+  Specifically, we define a design criterion based on uncertainty in
+  the excursions of vector-valued Gaussian random fields, and derive
+  tractable expressions for the expected integrated Bernoulli variance
+  reduction in such a framework. We demonstrate how this criterion can
+  be used to prioritize sampling efforts at locations that are
+  ambiguous, making exploration more effective.  We use simulations to
+  study and compare properties of the considered approaches, followed
+  by results from field deployments with an autonomous underwater
+  vehicle as part of a study mapping the boundary of a river
+  plume. The results demonstrate the potential of combining
+  statistical methods and robotic platforms to effectively inform and
+  execute data-driven environmental sampling.
 
@@
 \begin{keyword}
-\kwd{Ocean Sampling}
 \kwd{Excursion Sets}
@@
 \kwd{Autonomous robots}
+\kwd{Ocean Sampling}
 \kwd{Adaptive Information Gathering}
@@
 at the statistical prediction of features using an underwater glider.
-In this paper the main focus is not on statistical modeling per se, but rather on statistical principles and computations underlying efficient data collection. We combine novel possibilities in marine
-robotics with spatial statistics and experimental design to provide
-useful AUV sampling designs.
+In this work the main focus is not on statistical modeling per se, but
+rather on statistical principles and computation underlying efficient
+data collection. We combine novel possibilities in marine robotics
+with spatial statistics and experimental design to provide useful AUV
+sampling designs.
 
@@
 \item Replicable experiments on synthetic cases with accompanying
-  code~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}}
+  code
 \item Results from full-scale field trials running myopic strategies onboard an AUV for the characterization of a river plume. 
@@
 
-Given a GRF $\gp$ and observations of some of its components at locations in the domain, one can predict the value of the field at some
-unobserved location $\spatloc\in \domain$ by using the conditional
-mean of $\gp[\spatloc]$, conditional on the data. This coincides with co-Kriging equations,
-which tell us precisely how to compute conditional means and covariances.  We will present a
-general form of co-Kriging, in the sense that it allows inclusion of
-several (batch) observations at a time; observations at a given
-location $u \in \domain$ may only include a subset of the components
-of $\gp[\spatloc]\in\mathbb{R}^{\no}$ (heterotopic).
+Given a GRF $\gp$ and observations of some of its components at
+locations in the domain, one can predict the value of the field at
+some unobserved location $\spatloc\in \domain$ by using the
+conditional mean of $\gp[\spatloc]$, conditional on the data. This
+coincides with co-Kriging equations, which tell us precisely how to
+compute conditional means and covariances.  We will present a general
+form of co-Kriging, in the sense that it allows inclusion of several
+(batch) observations at a time; observations at a given location
+$u \in \domain$ may only include a subset of the components of
+$\gp[\spatloc]\in\mathbb{R}^{\no}$ (heterotopic).
 
@@
 \bm{x}_n)$ and where
-$\Delta_{[n]}$ is the covariance matrix of
-Gaussian-distributed noise assumed to have affected measurements up to
-batch
-$n$.
-For our applications with salinity and temperature observations, this matrix is diagonal because we assume conditionally independent sensor readings, but it might not be diagonal with other types of combined measurements.
-The matrix in parenthesis will be assumed to be non-singular
-throughout the presentation. The associated co-Kriging %(conditional)
-residual (cross-)covariance function can also be expressed in the same vein via
+$\Delta_{[n]}$ is the covariance matrix of Gaussian-distributed noise
+assumed to have affected measurements up to batch
+$n$. For our applications with salinity and temperature observations,
+this matrix is diagonal because we assume conditionally independent
+sensor readings, but it might not be diagonal with other types of
+combined measurements.  The matrix in parenthesis will be assumed to
+be non-singular throughout the presentation. The associated
+co-Kriging %(conditional)
+residual (cross-)covariance function can also be expressed in the same
+vein via
 %
@@
 
-We illustrate the expected Bernoulli variance (EBV) associated with different designs on a
-bivariate example. This mimics our river plume application and hence
-the first and second component of the random field will be called
-\textit{temperature} and \textit{salinity}. We begin
-with a \textit{pointwise} example, considering a single bivariate Gaussian distribution (i.e. no spatial elements).
+We illustrate the expected Bernoulli variance (EBV) associated with
+different designs on a bivariate example. This mimics our river plume
+application and hence the first and second component of the random
+field will be called \textit{temperature} and \textit{salinity}. We
+begin with a \textit{pointwise} example, considering a single
+bivariate Gaussian distribution (i.e. no spatial elements).
 
@@
 salinity. The Bernoulli variance is similarly larger for high
-correlations. EPs and Bernoulli variances are the same for temperature and salinity standard
-deviations $\sigma_1$ and $\sigma_2$, which implies that high variability in
-temperature and salinity is not captured in the $p(1-p)$ expression.
-
-\begin{table}[!t] \centering \caption{EP and Bernoulli variance for different correlations and variances (top rows), and EBVs for both temperature and salinity data, and only temperature data (bottom rows).}
+correlations. EPs and Bernoulli variances are the same for temperature
+and salinity standard deviations $\sigma_1$ and $\sigma_2$, which
+implies that high variability in temperature and salinity is not
+captured in the $p(1-p)$ expression.
+
+\begin{table}[!t] \centering \caption{EP and Bernoulli variance for
+    different correlations and variances (top rows), and EBVs for both
+    temperature and salinity data, and only temperature data (bottom
+    rows).}
   \begin{tabular}{c|ccc|ccc}
@@
 \end{table}
-The bottom two rows of Table \ref{tab:sim_rhoab} show 
-EBV results. This is presented for a design gathering
-both data types, and for a design with temperature measurements
-alone. When both data are gathered, the measurement model is
+The bottom two rows of Table \ref{tab:sim_rhoab} show EBV
+results. This is presented for a design gathering both data types, and
+for a design with temperature measurements alone. When both data are
+gathered, the measurement model is
 $(Y_1,Y_2)^t=(Z_1,Z_2)^t+\bepsilon$, with
@@
 this illustration, Table \ref{tab:sim_rhoab} shows that the expected
-Bernoulli variance gets smaller with larger standard deviations. The expected reduction of Bernoulli variance is
-further largest for the cases with high correlation $\gamma$. Albeit smaller,
-there is also uncertainty reduction when only temperature is measured
-(bottom row), especially when temperature and salinity are highly
-correlated. When correlation is low ($\gamma=0.2$), there is little
-information about salinity in the temperature data, and therefore less
-uncertainty reduction. 
+Bernoulli variance gets smaller with larger standard deviations. The
+expected reduction of Bernoulli variance is further largest for the
+cases with high correlation $\gamma$. Albeit smaller, there is also
+uncertainty reduction when only temperature is measured (bottom row),
+especially when temperature and salinity are highly correlated. When
+correlation is low ($\gamma=0.2$), there is little information about
+salinity in the temperature data, and therefore less uncertainty
+reduction.
 %In an application with fresh cold water from a river source, the temperature and salinity variables will not only be interdependent, but will also likely show dependence in the spatial dimension.
@@
 \end{align*}
-where an isotropic Mat\'{e}rn 3/2 kernel $(1+\eta h)\exp (-\eta h)$ is used, for Euclidean distance $h$.
-In the accompanying Python
-examples taking place within the MESLAS toolbox~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}}, these
-modeling assumptions can however be relaxed to anisotropic covariance and
-changing variance levels across the spatial domain. Both extensions are relevant for the setting with river
-plumes, but in practice this requires more parameters to be specified.
-With extensive satellite data or prior knowledge from high-resolution ocean models, one could also possibly fit more complex multivariate
-spatial covariance functions
+where an isotropic Mat\'{e}rn 3/2 kernel $(1+\eta h)\exp (-\eta h)$ is
+used, for Euclidean distance $h$.  In the accompanying Python examples
+taking place within the MESLAS
+toolbox~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}},
+these modeling assumptions can however be relaxed to anisotropic
+covariance and changing variance levels across the spatial
+domain. Both extensions are relevant for the setting with river
+plumes, but in practice this requires more parameters to be
+specified. With extensive satellite data or prior knowledge from
+high-resolution ocean models, one could also possibly fit more complex
+multivariate spatial covariance functions
 \citep{gneiting2010matern,genton2015cross}, but that is outside the
-scope of the current paper.
+scope of the current work.
 
@@
 
-We now study how the EBV [Eq.\eqref{def:eibv}]
-associated with data collection at a point changes if only one of the
-two components of the field is observed. We first draw a realization
-of the GRF defined above and use it as ground-truth to mimic the real
-data-collection process. A first set of observations are done at the
-locations depicted in grey (see Fig.~\ref{fig:ebv_comp}), and the data
-is used to update the GRF model. We then consider the green triangle
-as a potential next observation location and plot the EBV reduction (at each grid node in the waypoint graph) that would result from observing only one component of the field (temperature or salinity), or both at that point.
+We now study how the EBV [Eq.\eqref{def:eibv}] associated with data
+collection at a point changes if only one of the two components of the
+field is observed. We first draw a realization of the GRF defined
+above and use it as ground-truth to mimic the real data-collection
+process. A first set of observations are done at the locations
+depicted in grey (see Fig.~\ref{fig:ebv_comp}), and the data is used
+to update the GRF model. We then consider the green triangle as a
+potential next observation location and plot the EBV reduction (at
+each grid node in the waypoint graph) that would result from observing
+only one component of the field (temperature or salinity), or both at
+that point.
 
@@
 \caption{Pointwise Bernoulli variance reduction for observation of a
-    single or both components of the random field at one
-    location. Data collection locations in green. True excursion set
-    in red. Places where only one response is above threshold are
-    depicted in pink. EBV reduction associated to observing one or both responses at the green location are shown in \ref{fig:ebv_comp_temp}, \ref{fig:ebv_comp_sal} and \ref{fig:ebv_comp_both}.}
+  single or both components of the random field at one location. Data
+  collection locations in green. True excursion set in red. Places
+  where only one response is above threshold are depicted in pink. EBV
+  reduction associated to observing one or both responses at the green
+  location are shown in \ref{fig:ebv_comp_temp},
+  \ref{fig:ebv_comp_sal} and \ref{fig:ebv_comp_both}.}
 \label{fig:ebv_comp}
@@
 
-Note that plotting the EBV reduction at each point
-might also be used to compare different data collection plans. For
-example, Fig. \ref{fig:ebv_north_vs_east} shows the EBV reduction associated with a data collection plan along a vertical line (static north) and one associated with a horizontal
-(static east). Both expectations are computed according to the
-a-priori distribution of the GRF (i.e. no observations have been
-included yet).
+Note that plotting the EBV reduction at each point might also be used
+to compare different data collection plans. For example,
+Fig. \ref{fig:ebv_north_vs_east} shows the EBV reduction associated
+with a data collection plan along a vertical line (static north) and
+one associated with a horizontal (static east). Both expectations are
+computed according to the a-priori distribution of the GRF (i.e. no
+observations have been included yet).
 
@@
 height=0.26\textwidth,keepaspectratio]{Figures/ebv_static_east.png}\label{fig:ebv_static_east}}
-  \caption{Pointwise Bernoulli variance reduction for two different observation plans.
-The prior EP is shown in \ref{fig:ebv_static_excu}.
-EBV reduction for each design shown in \ref{fig:ebv_static_north} and \ref{fig:ebv_static_east}.}
+\caption{Pointwise Bernoulli variance reduction for two different
+  observation plans. The prior EP is shown in
+  \ref{fig:ebv_static_excu}. EBV reduction for each design shown in
+  \ref{fig:ebv_static_north} and \ref{fig:ebv_static_east}.}
 \label{fig:ebv_north_vs_east}
@@
 
-\section{Sequential designs and heuristic path planning}\label{sec:heuristics}
+\section{Sequential designs and heuristic path planning}
+\label{sec:heuristics}
 
@@
 
-\subsection{Look-ahead Trajectory Planning}\label{sec:LA}
+\subsection{Look-ahead Trajectory Planning}
+\label{sec:LA}
 
@@
 
-\subsection{Simulation study}
+\subsection{Simulation studies}
 \label{sec:simulations}
@@
 Each strategy is conducted on an equilateral grid as shown in
-Fig.~\ref{fig:wp_graph}. The AUV starts at the center coordinate at the southern end of the domain (node 53). It then moves
-along edges in the waypoint graph while collecting data which are
-assimilated onboard to update the GRF model. This is used in the evaluation
-of the next node to sample.  The procedure is run for $10$ stages. A
-total of 100 replicate simulations were conducted with all strategies.
+Fig.~\ref{fig:wp_graph}. The AUV starts at the center coordinate at
+the southern end of the domain (node 53). It then moves along edges in
+the waypoint graph while collecting data which are assimilated onboard
+to update the GRF model. This is used in the evaluation of the next
+node to sample.  The procedure is run for $10$ stages. A total of 100
+replicate simulations were conducted with all strategies.
 
@@
   overlap).]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}}
-\caption{Simulation results from 100 replicate simulations for 10 sampling choices/stages on the grid. Vertical lines show variation in replicate results.}  
+  \caption{Simulation results from 100 replicate simulations for 10
+    sampling choices/stages on the grid. Vertical lines show variation
+    in replicate results.}
 \label{fig:sim_results}
@@
 This work builds on a multidisciplinary effort combining statistical
-methods with robotic surveying techniques for oceanographic
-applications. We show how observation practices can gain efficiency
-and accuracy from the development of statistical techniques for
-spatial monitoring. We further demonstrate the opportunities available
-for real-time multivariate spatial data gathering and analysis on
-autonomous platforms, which statisticians can exploit to create
-general-purpose toolkits for similar applications.
-
-In particular, we derive and show results for characterizing phenomena
-connected to the properties of water masses. The characterization of
-uncertainties in random sets is extended in the vector-valued case with new results for the
-expected integrated Bernoulli variance reduction achieved by spatial
-sampling designs. This is provided in semi-analytical form for the situation
-with a static design, and then extended to the adaptive situation. The
-sequential derivations provide new insights into efficient
-applications of adaptive data collection, as demonstrated in our
-application.
+methods with robotic sampling for oceanographic applications. We show
+how observation practices can gain efficiency and accuracy from
+statistical techniques for spatial monitoring and demonstrate the the
+need for real-time multivariate spatial sampling on autonomous
+platforms.% , which statisticians can exploit to create general-purpose
+% toolkits for similar applications.
+
+In particular, we derive and show results for a real-world domain
+characterizing water mass properties. The characterization of
+uncertainties in random sets is extended in the vector-valued case
+with new results for the expected integrated Bernoulli variance
+reduction achieved by spatial sampling designs. This is provided in
+semi-analytical form for static designs, and then extended to the
+adaptive situations. The sequential derivations provide new insights
+into efficient applications of adaptive data collection, as
+demonstrated in our application.
 
@@
 represented by a two dimensional grid. Extensions to three-dimensional
-domains are not methodologically different, but the operation must
-likely approximate calculations by concentrating numerical integration efforts on terms in the
-vicinity of the autonomous vehicle \citep{fossum18b}. We did not
+domains are not methodologically different, but likely approximate
+calculations by concentrating numerical integration on terms in the
+vicinity of the autonomous vehicle \citep{fossum18b}. While we did not
 consider any temporal effects, which would be relevant on a larger
-time scale. We consider the extension to spatio-temporal modeling as
-future work, and envision that advection-diffusion equations could be
-useful in this kind of modeling
+time scale, we do consider the extension to spatio-temporal modeling
+and envision that advection-diffusion equations could be useful in
+such modeling
 \citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
@@
 processes which could still be run onboard and augmented by dynamical
-models. If high-resolution ocean model or satellite data are available, it is hopefully possible to fit such more complex statistical models.
-
-The spatial-statistical design criterion building on random sets is
+models. If high-resolution ocean model or satellite data are
+available, it is hopefully possible to fit such more complex
+statistical models. \kc{'Fit' as in modeling fit, or fit for
+  augmentation with dynamical models? But also note ocean models are
+  unlikely to be able to run onboard, so a simple phrase should be
+  added to that effect.}
+
+The spatio-statistical design criterion building on random sets is
 relevant in our setting with different water properties.  We show
 mathematical generality beyond the expected integrated Bernoulli
-variance, for instance that of volume uncertainties which is possibly
-more relevant, but requires more computational resources.  Such
-criteria could be particularly useful in other oceanographic settings
-related to mapping of algal-blooms, anoxic zones or open water fronts
-\cite{costa19}.  Of course, other criteria are also relevant for
-instance, hybrid or multi-attribute criteria that could balance goals
-of exploration and exploitation in this situation. Equally, such
+variance, for instance, that of volume uncertainties which is possibly
+more relevant, but one that requires more computational resources.
+Such criteria could be particularly useful in other oceanographic
+settings related to mapping of algal-blooms, anoxic zones or open
+water fronts \cite{costa19}.  Other criteria could also be relevant,
+for instance, hybrid or multi-attribute criteria that could balance
+goals of exploration and exploitation in this situation. Equally, such
 techniques have significant use cases in downstream decision-making,
 with policy makers and regulators who need to make difficult decisions
-related to fish farming or other marine resource operations. Value of
-information analysis \citep{Eidsvik:15} could be used to evaluate
-whether information is likely to result in improved decision-making,
-in such a context.  We also foresee opportunities related to design of
+related to aquaculture or other marine resources. Value of information
+analysis \citep{Eidsvik:15} could be used to evaluate whether
+information is likely to result in improved decision-making, in such a
+context.  We also foresee opportunities related to design of
 experiments for multivariate processes using our notion of generalized
-locations. For instance, this is likely to be useful for
-internet-of-things applications or computer experiments where some
-observations or evaluations are rather inexpensive, while others 
-must only be done when they are really valuable.
-
-In our context the myopic strategy performs well, and for
-computational reasons we did not go in depth on the dynamic program
-solution. There has been much creative work on finite horizon
+locations.
+
+In our context the myopic strategy performs well, and due to
+computational constraints we did not go in depth on the dynamic
+programming solutions. There has been much work on finite horizon
 optimization in the robotics literature including probabilistic road
 maps and rapidly-exploring random trees \citep{karaman2011sampling},
-but their statistical properties are not yet clear.  
-In some cases it is also limiting to use the waypoint graph, and it would be beneficial to allow more continuous updates and navigation at the highest frequency allowed by the AUV.
-It is equally
+but their statistical properties are unclear.  In some cases it is
+also limiting to use a waypoint graph, and would be beneficial to
+allow more continuous updates and navigation at the highest frequency
+possible given limitations being onboard an AUV.  It is equally
 interesting to explore the additional flexibility that can be gained
@@
 opportunities to move in parallel to best capture the excursion set.
-The sometimes conflicting topics of autonomy and communication have
-caught much interest in control engineering lately
-\citep{zolich2019survey}, and statistical evaluations could contribute
-in the future. In our context, autonomy is highlighted, and this is
-reflected in our modeling and computational assumptions. With the
-possibilities of wifi communication to a mother-ship at waypoints,
-more advanced modeling and computational routines are of course
-possible, but with new challenges related to design and planning for
-when and where the communication should occur.
+The value of information related to when and what to communicate (to
+shore or to other vehicles) is also an interesting thrust for research
+and likely to be useful for internet-of-things applications or
+computer experiments where some observations or evaluations are rather
+inexpensive, while others must only be done when they are really
+valuable.
+
+% The sometimes conflicting topics of autonomy and communication have
+% caught much interest in control engineering lately
+% \citep{zolich2019survey}, and statistical evaluations could contribute
+% in the future. In our context, autonomy is highlighted, and this is
+% reflected in our modeling and computational assumptions. With the
+% possibilities of wifi communication to a research vessel at waypoints,
+% more advanced modeling and computational routines are possible, but
+% with new challenges related to design and planning for when and where
+% the communication should occur.
 
@@
 TOF acknowledges support from the Centre for Autonomous Marine
-Operations and Systems (AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of
-Excellence, project number 223254, and the Applied Underwater Robotics Labortatory (AURLab). CT and DG acknowledge
-support from the Swiss National Science Foundation, project number
-178858. JE and KR acknowledge support from Norwegian
-research council (RCN), project number 305445.
-DG would like to aknowledge support of Idiap Research Institute, where he had is main affiliation when the first version of this work was submitted.
-The authors would also like to thank Niklas Linde (University of Lausanne) for pointing out the relevance of revisiting the proposed approach for estimating volumes of mixing zones, e.g.
-in hydrology.
+Operations and Systems (AMOS), Center of Excellence, project number
+223254, and the Applied Underwater Robotics Labortatory (AURLab). CT
+and DG acknowledge support from the Swiss National Science Foundation,
+project number 178858. JE and KR acknowledge support from Norwegian
+research council (RCN), project number 305445. DG would like to
+aknowledge support of Idiap Research Institute, his primary
+affiliation in an early version of this work.  The authors would also
+like to thank Niklas Linde of the University of Lausanne for providing
+constructive feedback on a version of the manuscript.
+
+%  for pointing out the relevance of revisiting the proposed
+% approach for estimating volumes of mixing zones, e.g. in hydrology.
 

### 2020-06-27 main_new.tex
Updates from Overleaf

@@
 \end{equation*}
-with blocks given by
+with blocks given, for $i,j\in \{1,2\}$ and $u,v\in \domain$, by
 \begin{equation*}
 \begin{split}
-\tilde{\Sigma}_{i,j}(u, v) &= \lambda_{[n+1,n+1]}(\uu)^T k_{[n]}(\bm{x},\bm{x}) \lambda_{[n+1,n+1]}(\vv) + \delta_{i,j}\futureCov{\uu, \vv}
+\tilde{\Sigma}_{i,j}(u, v) &= \lambda_{[n+1,n+1]}(\uu)^T k_{[n]}(\bm{x},\bm{x}) \lambda_{[n+1,n+1]}(\vv) + \delta_{i,j}\futureCov{\uu, \vv}.
 \end{split}
 \end{equation*}
-and the other blocks are defined by following the same logic, exchanging the role of $\uu$ and $\vv$.
+%and the other blocks are defined by following the same logic, exchanging the role of $\uu$ and $\vv$.
 \end{propo}
@@
 research council (RCN), project number 305445. DG would like to
-aknowledge support of Idiap Research Institute, his primary
-affiliation in an early version of this work.  The authors would also
-like to thank Niklas Linde of the University of Lausanne for providing
-constructive feedback on a version of the manuscript.
+aknowledge support of Idiap Research Institute, his primary affiliation in an early version of this manuscript.  The authors would also
+like to thank Niklas Linde of the University of Lausanne for providing constructive feedback on this work.
 

### 2020-06-29 appendix.tex
Updates from Overleaf

@@
 %Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
-For a measurable random field $\gp$ and a locally finite measure $\mes$ on $D$, $\mes(\es)$ is a random variable and for 
+For a measurable random field $\gp$ and a locally finite measure $\mes$ on $\domain$, $\mes(\es)$ is a random variable and for 
 any $r\geq 1$,
@@
 itself inherited from the assumed measurability for
-$(\x, \omega) \to \gp[\x](\omega)$ and $T$, respectively. From there, following the steps of Robbins' theorem \cite{Robins1944}, we find that
+$(\x, \omega) \to \gp[\x](\omega)$ and $T$, respectively. From there, following the steps of Robbins' theorem \cite{Robbins1944}, we find that
 \begin{equation*}
@@
 Let $p, q, h \geq 1$, $a \in \R^p$, $B \in \R^{p\times q}$,
-and $\covN, \covV$ be two covariance matrices in
+and $\covN$, $\covV$ be two covariance matrices in
 $\R^{p\times p}$ and $\R^{q\times q}$, respectively.

### 2020-06-29 main_new.tex
Updates from Overleaf

@@
 % "New Title of the paper"
-\title{Autonomous Oceanographic Data Collection. Learning Excursion
-  Sets of Vector-valued Gaussian Random Fields \\or\\
+\title{%Autonomous Oceanographic Data Collection. Learning Excursion
+%  Sets of Vector-valued Gaussian Random Fields \\or\\
 Learning Excursion Sets of Vector-valued Gaussian Random Fields for
@@
 research council (RCN), project number 305445. DG would like to
-aknowledge support of Idiap Research Institute, his primary affiliation in an early version of this manuscript.  The authors would also
-like to thank Niklas Linde of the University of Lausanne for providing constructive feedback on this work.
+aknowledge support of Idiap Research Institute, his primary affiliation in an early version of this manuscript.
+The authors would also like to thank Niklas Linde of the University of Lausanne for providing constructive feedback on a version of the manuscript, and finally members of the AURLab for help with AUV deployments
 

### 2020-06-29 ans_to_reviewers/answers-to-reviewers.tex
Updates from Overleaf

@@
 To help us in these efforts, we included Cedric Travelletti as co-author because he has contributed significantly on the revised version of the paper.
+\\
 
@@
 \item The revision has lead to {\bf{structural changes}} as follows: \\
-    i) previous Section 1 and 2 are combined to one Introduction which should help the narrative and highlight the main contributions of the work. \\
-    ii) merged the data assimilation part with equations for co-Kriging to clarify that this work is on multivariate GRFs. \\
+    i) Section 1 and 2 are now combined into one single \emph{Introduction} which should help the narrative and highlight the main contributions of the work. \\
+    ii) The data assimilation part is now merged with equations for co-Kriging to clarify that this work is on multivariate GRFs. \\
     iii) Split the simulation study in two and used these as to directly motivate the statistical concepts of uncertainty reduction (Sect 2) and adaptive sampling strategies (Sect 3).
-    iv) Present mathematical contributions succintly in Section 2 and 3, while the mathematical proofs are in Appendix.
+    iv) Present mathematical contributions succinctly in Section 2 and 3, while the mathematical proofs are in Appendix.
 
@@
 Motivating examples are abundant in medicine where doctors do not solely rely on a single symptom but must see several combined effects before making a diagnosis, for example.
-Our motivation in the current paper is that temperature and salinity are proxies for more complex biological variables in the ocean, and in several cases they must be jointly high or low for phenomena to occur.
+Our motivation in the current paper is that temperature and salinity are proxies/indicators for more complex biological variables in the ocean, and in several cases they must be jointly high or low for phenomena to occur.
 
@@
 
-\item The {\bf{statistical methodological contribution}} have been clarified and improved, including definitions of random sets and ESs, and having more general results than just the EIBV for the bi-variate case of temperature and salinity.
+\item The {\bf{statistical methodological contribution}} have been clarified and improved, including definitions of random sets and ESs.
+This also includes more general results than previously which just included the EIBV for the bi-variate case of temperature and salinity.
 The revised version highlights these novel statistical results as a main contribution of the paper.

### 2020-06-29 cover_letter/cover_letter.tex
Updates from Overleaf

--- 2019-10-03 cover_letter/cover_letter.tex
@@
 
-Please see our manuscript \textbf{Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets} submitted for possible publication in the \emph{Annals of Applied Statistics} (AOAS).\\*
+We thank the editors and reviewers for providing feedback on our manuscript \textbf{AOAS1910-027}.
+The feedback have been insightful and constructive.
+Our response is attached along with the updated manuscript itself.
+We have carefully considered all the feedback in this revised version of the manuscript, which we believe is a significantly improved version of our work.
 
-This work focuses on tackling the challenges related to oceanographic sampling, where observations are expensive, few, and far apart. A key question in this regard is
+After extensive discussion we collectively felt the need to not only recruit an additional co-author, but also to change the title to 'Learning Excursion Sets of Vector-valued Gaussian Random Fields for Autonomous Ocean Sampling', which we believe better reflects the current manuscript and the valuable reviewer and editorial feedback.
 
-\begin{quote}
-\textit{where, when, and how should we gather information in order to effectively characterize a phenomenon in the ocean?}  
-\end{quote}{}
-
-The research presented aims to address this central question by combining statistical models and methodology with autonomous robotic sensing platforms.
-A central effort is the introduction of informative descriptions of spatial characteristics of ocean phenomena that greatly improve how sequential data collection can be conducted.
-Specifically, we present a design criterion based on reducing uncertainty in the joint excursions of random fields, and derive analytical expressions for the expected Bernoulli variance reduction in the excursion set of spatial temperature and salinity variables.
-The proposed methodology is deployed and tested in a full-scale deployment where an autonomous underwater vehicle is used to map a river plume.
-
-This interdisciplinary study demonstrates the potential of statistics applied to support autonomous robotic decision making, representing methods and approaches with a substantial potential for further integration, which are necessary and critical to understand our changing oceans.
+We look forward to receiving any further feedback at the earliest.
 \vspace{1em}

### 2020-06-30 main_new.tex
Updates from Overleaf

@@
 aknowledge support of Idiap Research Institute, his primary affiliation in an early version of this manuscript.
-The authors would also like to thank Niklas Linde of the University of Lausanne for providing constructive feedback on a version of the manuscript, and finally members of the AURLab for help with AUV deployments
+The authors would also like to thank Niklas Linde of the University of Lausanne for providing constructive feedback about this work, and finally members of the AURLab for help with AUV deployments
 

### 2020-07-01 cover_letter/cover_letter.tex
Updated cover letter

@@
 The feedback have been insightful and constructive.
+Inspired by comments by the referees regarding the scope of the paper, we have spent substantial energy polishing its statistical character and also made an effort keeping the oceanograhic detail compact, yet sufficiently informative for the paper to be readable also by more application-driven research communities (in maritime sciences but also for instance in robotics and AI).
+From the statistical and mathematical side, we have tried our best to revise the formalism in such a way that further contributions can be built on top of ours, and also that we valorize some secondary findings that we have made along the way and that could be useful to researchers and practitioners from spatial statistics and beyond.
+This include a very modular (and conceptually appealing) way to present batch-sequential co-Kriging update formulae in heterotopic settings, generalities about excursion measures/volumes of (measurable) random fields, and also technical results towards the calculation of Stepwise Uncertainty Reduction in the vector-valued case.
+In particular, generalizing our former results enabled use to tackle in the revised version an arbitrary number of responses, as well as to calculate the expected excursion measure variance (on top of the expected integrated Bernoulli variance).\\
+
 Our response is attached along with the updated manuscript itself.
-We have carefully considered all the feedback in this revised version of the manuscript, which we believe is a significantly improved version of our work.
+We have carefully considered all the feedback in this revised version of the manuscript, which we believe is a significantly improved version of our work.\\
 

### 2020-07-01 main_new.tex
Updated cover letter

@@
 
-% == Enable text degreehttps://www.overleaf.com/project/5a3268379ecbdc657d8767e8
+% == Enable text 
 \usepackage{textcomp}
-
 \usepackage{amsthm,amsmath,natbib}
@@
 
-\runtitle{Autonomous Oceanographic Data Collection}
+\runtitle{Learning Excursion Sets for Autonomous Oceanographic Data Collection}
 
@@
 
-\subsection{Simulation Results}
+\subsubsection{Simulation Results}
 
@@
 time scale, we do consider the extension to spatio-temporal modeling
-and envision that advection-diffusion equations could be useful in
-such modeling
-\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
+and envision that advection-diffusion equations could be useful \citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
 oceanographic phenomena, the methods will need to be extended to
@@
 processes which could still be run onboard and augmented by dynamical
-models. If high-resolution ocean model or satellite data are
-available, it is hopefully possible to fit such more complex
-statistical models. \kc{'Fit' as in modeling fit, or fit for
-  augmentation with dynamical models? But also note ocean models are
-  unlikely to be able to run onboard, so a simple phrase should be
-  added to that effect.}
+models. Running large numerical models onboard a robotic vehicle is not currently possible, but high-resolution ocean model or satellite data can be used to fit a more complex statistical model.
 

### 2020-07-01 cover_letter/cover_letter.tex
Updates from Overleaf

@@
 The feedback have been insightful and constructive.
-Inspired by comments by the referees regarding the scope of the paper, we have spent substantial energy polishing its statistical character and also made an effort keeping the oceanograhic detail compact, yet sufficiently informative for the paper to be readable also by more application-driven research communities (in maritime sciences but also for instance in robotics and AI).
+Inspired by comments by the referees regarding the scope of the paper, we have spent substantial energy polishing its statistical character and also made an effort keeping the oceanographic detail compact, yet sufficiently informative for the paper to be readable also by more application-driven research communities (in maritime sciences but also for instance in robotics and AI).
 From the statistical and mathematical side, we have tried our best to revise the formalism in such a way that further contributions can be built on top of ours, and also that we valorize some secondary findings that we have made along the way and that could be useful to researchers and practitioners from spatial statistics and beyond.

### 2020-07-01 main_new.tex
updates

@@
 time scale, we do consider the extension to spatio-temporal modeling
-and envision that advection-diffusion equations could be useful \citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
+and envision that advection-diffusion equations could be useful
+\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
 oceanographic phenomena, the methods will need to be extended to
 non-Gaussian phenomena, possibly feature-based mixtures of Gaussian
-processes which could still be run onboard and augmented by dynamical
-models. Running large numerical models onboard a robotic vehicle is not currently possible, but high-resolution ocean model or satellite data can be used to fit a more complex statistical model.
+processes which could potentially be run onboard augmented by
+dynamical models. Running numerical models onboard a robotic vehicle
+is currently infeasible, but high-resolution ocean models or remote
+sensing data can be used to fit a more complex statistical model
+\cite{davidson19}.
 
@@
 research council (RCN), project number 305445. DG would like to
-aknowledge support of Idiap Research Institute, his primary affiliation in an early version of this manuscript.
-The authors would also like to thank Niklas Linde of the University of Lausanne for providing constructive feedback about this work, and finally members of the AURLab for help with AUV deployments
+aknowledge support of Idiap Research Institute, his primary
+affiliation in an early version of this manuscript. The authors would
+also like to thank Niklas Linde of the University of Lausanne for
+providing constructive feedback about this work, and members of the
+NTNU AURLab for help with AUV deployments
 

### 2020-07-03 main_new.tex
Updates from Overleaf

@@
 
-\runtitle{Learning Excursion Sets for Autonomous Oceanographic Data Collection}
+\runtitle{Learning Excursion Sets for Autonomous Data Collection}
 
@@
 
-\begin{figure}[!b] 
+\begin{figure}[ht] 
 \centering 
@@
 height=0.26\textwidth,keepaspectratio]{Figures/ebv_static_east.png}\label{fig:ebv_static_east}}
-\caption{Pointwise Bernoulli variance reduction for two different
-  observation plans. The prior EP is shown in
+\hspace{2em}
+\subfigure[Excursion realization.]{\includegraphics[
+height=0.26\textwidth,keepaspectratio]{Figures/real_myopic.png}\label{fig:real_myopic}}
+\hspace{2em}
+\subfigure[Route taken by myopic strategy.]{\includegraphics[
+height=0.26\textwidth,keepaspectratio]{Figures/route_myopic.png}\label{fig:route_myopic}}
+\hspace{2em}
+\subfigure[Directional reduction of EBV.]{\includegraphics[
+height=0.26\textwidth,keepaspectratio]{Figures/compass_myopic.png}\label{fig:compass_myopic}}
+\caption{Pointwise Bernoulli variance reduction for two static and the myopic observation strategies. The prior EP is shown in
   \ref{fig:ebv_static_excu}. EBV reduction for each design shown in
-  \ref{fig:ebv_static_north} and \ref{fig:ebv_static_east}.}
+  \ref{fig:ebv_static_north} and \ref{fig:ebv_static_east}.
+The true excursion for a realization of temperature and salinity is shown in \ref{fig:real_myopic}.
+The route taken by the myopic observation strategy is shown in \ref{fig:route_myopic}.
+The different alternative route choices and their expected EBV, illustrated as a compass from the perspective of the AUV, is shown in \ref{fig:compass_myopic}.
+Southward or $270 ^{\circ}$ is indicated by the thick and light color to be the best alternative direction.}
 \label{fig:ebv_north_vs_east}
@@
 \textit{static\_north}, \textit{static\_east}, and
-\textit{static\_zigzag} with the three described sequential approaches
+\textit{static\_zigzag} (a version of \textit{static\_north} where with some east-west transitions in a zigzag pattern) with the three described sequential approaches
 \textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
@@
 approaches this is not the case. For comparison, the properties are
-therefore evaluated using Monte Carlo sampling over several replicates
+therefore evaluated using Monte Carlo integration over several replicates
 of realizations from the model while conducting simulated sequential
-surveys for each one. We also compare predictive
+surveys for each one. An example of such a realization with a \textit{myopic} strategy is shown in Fig.~\ref{fig:route_myopic}.
+
+We also compare predictive
 performance measured by root mean square error (RMSE) for temperature
@@
 is expected as the two criteria (Eq. \eqref{critSEQ} and
-\eqref{critLA}) are sensitive to differences in IBV. The
-\textit{static\_north} design also does well here because the path is
+\eqref{critLA}) are sensitive to differences in IBV. The \textit{static\_north} design also does well here because the path is
 parallel to the boundary between the water masses.

### 2021-01-02 ans_to_reviewers/ans-to-2nd-rev.tex
Updates from Overleaf

\documentclass[a4paper]{article}
\topmargin	-10mm
\oddsidemargin	0mm
\textwidth	150mm
\textheight	230mm

\newcommand{\ba}{ {\boldsymbol a} }
\newcommand{\bA}{ {\boldsymbol A} }
\newcommand{\bb}{ {\boldsymbol b} }
\newcommand{\bB}{ {\boldsymbol B} }
\newcommand{\bc}{ {\boldsymbol c} }
\newcommand{\bC}{ {\boldsymbol C} }
\newcommand{\bd}{ {\boldsymbol d} }
\newcommand{\bD}{ {\boldsymbol D} }
\newcommand{\be}{ {\boldsymbol e} }
\newcommand{\bE}{ {\boldsymbol E} }
\newcommand{\boldf}{ {\boldsymbol f} }
\newcommand{\bF}{ {\boldsymbol F} }
\newcommand{\bg}{ {\boldsymbol g} }
\newcommand{\bG}{ {\boldsymbol G} }
\newcommand{\bh}{ {\boldsymbol h} }
\newcommand{\bH}{ {\boldsymbol H} }
\newcommand{\bi}{ {\boldsymbol i} }
\newcommand{\bI}{ {\boldsymbol I} }
\newcommand{\bj}{ {\boldsymbol j} }
\newcommand{\bJ}{ {\boldsymbol J} }
\newcommand{\bk}{ {\boldsymbol k} }
\newcommand{\bK}{ {\boldsymbol K} }
\newcommand{\bl}{ {\boldsymbol l} }
\newcommand{\bL}{ {\boldsymbol L} }
\newcommand{\bm}{ {\boldsymbol m} }
\newcommand{\bM}{ {\boldsymbol M} }
\newcommand{\bn}{ {\boldsymbol n} }
\newcommand{\bN}{ {\boldsymbol N} }
\newcommand{\bo}{ {\boldsymbol o} }
\newcommand{\bO}{ {\boldsymbol O} }
\newcommand{\bp}{ {\boldsymbol p} }
\newcommand{\bP}{ {\boldsymbol P} }
\newcommand{\bq}{ {\boldsymbol q} }
\newcommand{\bQ}{ {\boldsymbol Q} }
\newcommand{\br}{ {\boldsymbol r} }
\newcommand{\bR}{ {\boldsymbol R} }
\newcommand{\bs}{ {\boldsymbol s} }
\newcommand{\bS}{ {\boldsymbol S} }
\newcommand{\bt}{ {\boldsymbol t} }
\newcommand{\bT}{ {\boldsymbol T} }
\newcommand{\bu}{ {\boldsymbol u} }
\newcommand{\bU}{ {\boldsymbol U} }
\newcommand{\bv}{ {\boldsymbol v} }
\newcommand{\bV}{ {\boldsymbol V} }
\newcommand{\bw}{ {\boldsymbol w} }
\newcommand{\bW}{ {\boldsymbol W} }
\newcommand{\bx}{ {\boldsymbol x} }
\newcommand{\bX}{ {\boldsymbol X} }
\newcommand{\by}{ {\boldsymbol y} }
\newcommand{\bY}{ {\boldsymbol Y} }
\newcommand{\bz}{ {\boldsymbol z} }
\newcommand{\bZ}{ {\boldsymbol Z} }
\newcommand{\vc}[1]{\mbox{\boldmath $1$}}
\newcommand{\balph}{ {\boldsymbol \alpha} }
\newcommand{\balpha}{ {\boldsymbol \alpha} }
\newcommand{\bbet}{ {\boldsymbol \beta} }
\newcommand{\bbeta}{ {\boldsymbol \beta} }
\newcommand{\bgam}{ {\boldsymbol \gamma} }
\newcommand{\bgamma}{ {\boldsymbol \gamma} }
\newcommand{\bGamma}{ {\boldsymbol \Gamma} }
\newcommand{\bdelta}{ {\boldsymbol \delta} }
\newcommand{\bDelta}{ {\boldsymbol \Delta} }
\newcommand{\beps}{ {\boldsymbol \epsilon} }
\newcommand{\bepsilon}{ {\boldsymbol \epsilon} }
\newcommand{\bphi}{ {\boldsymbol \phi} }
\newcommand{\bPhi}{ {\boldsymbol \Phi} }
\newcommand{\bpi}{ {\boldsymbol \pi} }
\newcommand{\bpsi}{ {\boldsymbol \psi} }
\newcommand{\bkap}{ {\boldsymbol \kappa} }
\newcommand{\bkappa}{ {\boldsymbol \kappa} }
\newcommand{\bKappa}{ {\boldsymbol \Kappa} }
\newcommand{\blam}{ {\boldsymbol \lambda} }
\newcommand{\blambda}{ {\boldsymbol \lambda} }
\newcommand{\bLambda}{ {\boldsymbol \Lambda} }
\newcommand{\bmu}{ {\boldsymbol \mu} }
\newcommand{\bMu}{ {\boldsymbol \Mu} }
\newcommand{\bet}{ {\boldsymbol \eta} }
\newcommand{\bome}{ {\boldsymbol \omega} }
\newcommand{\bomega}{ {\boldsymbol \omega} }
\newcommand{\bOmega}{ {\boldsymbol \Omega} }
\newcommand{\bnabla}{ {\boldsymbol \nabla} }
\newcommand{\brho}{ {\boldsymbol \rho} }
\newcommand{\bsigma}{ {\boldsymbol \sigma} }
\newcommand{\bSig}{ {\boldsymbol \Sigma} }
\newcommand{\bSigma}{ {\boldsymbol \Sigma} }
\newcommand{\btheta}{ {\boldsymbol \theta} }
\newcommand{\bTheta}{ {\boldsymbol \Theta} }
\newcommand{\bzeta}{ {\boldsymbol \zeta} }
\newcommand{\bPsi}{ {\boldsymbol \Psi} }
\newcommand{\btau}{ {\boldsymbol \tau} }
\newcommand{\bzero}{ {\boldsymbol 0} }
\newcommand{\bones}{ {\boldsymbol 1} }
\newcommand{\given}{\,|\,}
\newcommand{\sS}{{\cal S}}
\newcommand{\Ss}{{\cal S}}
\newcommand{\Field}{{\cal F}}
\newcommand{\colsp}{{\cal C}}
\newcommand{\nullsp}{{\cal N}}
\newcommand{\rowsp}{{\cal R}}
\newcommand{\tildeC}{\tilde{C}}
\newcommand{\tildeK}{\tilde{K}}
\newcommand{\tildew}{\tilde{w}}
\newcommand{\tildebw}{\tilde{\bw}}
\newcommand{\tildebW}{\tilde{\bW}}
\newcommand{\calC}{{\cal C}}
\newcommand{\calcbC}{{\bf {\cal C}}}

\def\rx{{\texttt{T-REX\ }}}
\def\rxe{{\texttt{T-REX}}}
\def\du{{\texttt{DUNE}\ }}
\def\due{{\texttt{DUNE}}}

% == Maths
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage[normalem]{ulem} %to strike the words

\newcounter{reviewer}
\def\revcom{\textbf{Reviewer comment}}
\def\aecom{\textbf{Associate Editor comment }}
\def\reply{\textbf{Reply}}
\def\action{\textbf{Action}}

\newlist{answers}{enumerate}{10}
\setlist[answers]{leftmargin=*, label=\arabic{reviewer}.\arabic*}

\newcommand{\edcomment}[1]{{\color{green}{\{Editor: 1\}}}}
\newcommand{\frevcomment}[1]{{\color{blue}{\{Rev 1: 1\}}}}
\newcommand{\srevcomment}[1]{{\color{red}{\{Rev 2: 1\}}}}
\newcommand{\trevcomment}[1]{{\color{violet}{\{Rev 3: 1\}}}}

\usepackage{amsthm,amsmath,natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

\begin{document}

%----------------------------------------------------------------------
\title{Autonomous Oceanographic Data Collection. Learning Excursion Sets of Vector-valued Gaussian Random Fields
  \\\vspace{5mm}
 Authors response to 2$^{nd}$ review
  \\\vspace{5mm}
\small{Manuscript: AOAS1910-027}}
\author{ }

\date{\today}

\maketitle

We are grateful to the editors and reviewers for their comments and
suggestions. They have all been considered in this revised
version of our manuscript.\\

\par \vspace{1em}

%======================================================================
\section*{Editor}
%======================================================================
All comments of the Editor are listed below and answered on a point-by-point basis.

\vspace{5mm}
\noindent {\bf{1. Sampling limitations:}} You explained in your response file (R1S3.pdf) the limitations of the sampling design (comment #4 from the AE's previous review). I agree with the AE (bullet 1, R2AE2.pdf) that some (perhaps slightly abbreviated) version of your response on p4 (middle) would be important to include; viz.

\begin{quote}
"The sampling capabilities are restricted by the sensors onboard
the AUV, the computational load and navigation resources. The sampling
of temperature and salinity variables can be at 1Hz (higher than the
grid resolution), but, due to computational and navigation resources
available on the AUV, this simplification of data assimilation at the
waypoints is necessary."   
\end{quote}

\reply:

\vspace{5mm}
\noindent {\bf{2. Variance of errors:}} In view of the apparent nugget effects of 0.42 and 0.7 for salinity
and temperature, respectively, in Fig 9(c) (p21), I confess to being
a little puzzled myself by this statement on p22:
"We specify variance $0.25^2$ for both errors, which is based on a
middle ground between the nugget effect in the empirical variogram
and the sensor specifications."  Is 0.0625 really a "middle ground"?
Or, did you have other reasons for choosing 0.0625 not stated here?
In any event, I agree with the AE (bullet 5): "either give a better
justification of your choice of 0.0625 or, if that’s not possible,
then redo the analysis with a more justifiable choice."

\vspace{5mm}
\reply:

\vspace{5mm}
\noindent {\bf{3. Exposition:}} Apart from corrections noted by the AE (bullets 2,3,4,6), Referee 1
offered suggestions on shortening. Neither the AE nor I feel that re-structuring is needed.  In fact, some of the Referee's proposed changes would not be even desirable, from my perspective: I would not want you to "Consider putting the background material in an appendix to go as supplementary material" because, in my opinion, this background material is essential. And the simulation section (design and results) is already quite efficient. I include the Referee's report (R2R5.pdf) primarily for your information (the remarks from Ehrenberg are worth reading).

\vspace{5mm}
\reply:

\vspace{5mm}
\noindent {\bf{4. Figures:}} Like most journals, AOAS discourages the use of color in figures,
unless absolutely essential for reader comprehension.  I believe
that color is important for Figures 1, 2, 4, 11 and perhaps 5 and 6.
But color is not even mentioned in the caption for Figure 3; your
different line types likely will suffice in Figures 8 and 9; and
probably nothing will be lost if Figures 7 & 10 were black-and-white.
Color is very expensive to print, so we appreciate your attention
to this cost-saving measure.

\vspace{5mm}
\reply:

\vspace{5mm}
\noindent {\bf{Other comments:}} In your revision, you may want to include in the
acknowledgements, p26, the conscientious efforts of the anonymous
reviewers, whose comments led to this favorable outcome.

\vspace{5mm}
\reply:

%======================================================================
 \section*{Associate Editor}
%======================================================================
All comments of the Associate Editor are listed below and answered on a point-by-point basis.

\setcounter{reviewer}{1}

\vspace{5mm}
\noindent \aecom \textbf{1}:

Bla bla bla.

\reply:

\vspace{1em}

% == Adding references
\footnotesize
\bibliographystyle{apalike}
\bibliography{ref}

%=====================================================================
\end{document}
%=====================================================================

### 2021-01-02 main_new.tex
Updates from Overleaf

@@
-\documentclass[aoas]{imsart}
+\documentclass[aoas,preprint]{imsart}
+\setattribute{journal}{name}{}
 
@@
 \subsubsection{Including Spatiality}
+\label{sec:including_spatiality}
 
@@
 height=0.26\textwidth,keepaspectratio]{Figures/ebv_static_east.png}\label{fig:ebv_static_east}}
-\hspace{2em}
-\subfigure[Excursion realization.]{\includegraphics[
-height=0.26\textwidth,keepaspectratio]{Figures/real_myopic.png}\label{fig:real_myopic}}
-\hspace{2em}
-\subfigure[Route taken by myopic strategy.]{\includegraphics[
-height=0.26\textwidth,keepaspectratio]{Figures/route_myopic.png}\label{fig:route_myopic}}
-\hspace{2em}
-\subfigure[Directional reduction of EBV.]{\includegraphics[
-height=0.26\textwidth,keepaspectratio]{Figures/compass_myopic.png}\label{fig:compass_myopic}}
-\caption{Pointwise Bernoulli variance reduction for two static and the myopic observation strategies. The prior EP is shown in
+\caption{Pointwise Bernoulli variance reduction for two different static designs (later noted as \textit{static\_north} and \textit{static\_east}). The prior EP is shown in
   \ref{fig:ebv_static_excu}. EBV reduction for each design shown in
-  \ref{fig:ebv_static_north} and \ref{fig:ebv_static_east}.
-The true excursion for a realization of temperature and salinity is shown in \ref{fig:real_myopic}.
-The route taken by the myopic observation strategy is shown in \ref{fig:route_myopic}.
-The different alternative route choices and their expected EBV, illustrated as a compass from the perspective of the AUV, is shown in \ref{fig:compass_myopic}.
-Southward or $270 ^{\circ}$ is indicated by the thick and light color to be the best alternative direction.}
+  \ref{fig:ebv_static_north} and \ref{fig:ebv_static_east}.}
 \label{fig:ebv_north_vs_east}
@@
 computational power, making it well-suited for embedding on an AUV.
+
+\begin{figure}[ht] 
+\centering 
+\subfigure[Excursion realization.]{\includegraphics[
+height=0.26\textwidth,keepaspectratio]{Figures/ebv_myopic_excu.png}\label{fig:real_myopic}}
+\hspace{2em}
+\subfigure[BV reduction.]{\includegraphics[
+height=0.26\textwidth,keepaspectratio]{Figures/ebv_myopic_ebv.png}\label{fig:route_myopic}}
+\hspace{2em}
+\subfigure[Expected reduction of BV.]{\includegraphics[
+height=0.26\textwidth,keepaspectratio]{Figures/ebv_myopic_radar.png}\label{fig:compass_myopic}}
+\caption{Example run of the myopic strategy on a given realisation of the GRF model from \ref{sec:including_spatiality}.
+Past observation locations are depicted in green and current position of the AUV in cyan.
+The expected EBV reduction associated to data collection at neighbouring nodes of the current location is shown in \ref{fig:compass_myopic}.
+The thick and light color indicate the north-east node at $30 ^{\circ}$ to be the best possible choice.}
+\label{fig:ebv_myopic}
+\end{figure}
 

### 2021-01-02 main_web.tex
Updates from Overleaf

--- 2021-01-02 main_new.tex
@@
 
-% == Enable text 
 \usepackage{textcomp}
 \usepackage{amsthm,amsmath,natbib}
-\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
-
-% == Trygve Test
+%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
 \usepackage{color, colortbl}
@@
 
-% put your definitions there:
 \startlocaldefs
@@
 
-% New Marcos.
 \usepackage{amssymb}
@@
 
-% "New Title of the paper"
-\title{%Autonomous Oceanographic Data Collection. Learning Excursion
-%  Sets of Vector-valued Gaussian Random Fields \\or\\
+\title{
 Learning Excursion Sets of Vector-valued Gaussian Random Fields for
@@
   opportunities for experimental design.
-  %
-  %In this work we develop methods for efficient spatial sampling applied to the mapping of coastal ocean processes by providing informative descriptions of spatial characteristics of ocean phenomena.
-  %
-  % REPLACED BY 
-  %
   In this work we develop efficient spatial sampling methods for
@@
   salinity measurements.
-  % EN REPLACEMENT
   Specifically, we define a design criterion based on uncertainty in
@@
   execute data-driven environmental sampling.
-
-%Motivated by the challenges related to efficient allocation of sampling resources in environmental sensing, the combination of Excursion Probabilities and Gaussian process modeling is explored for autonomous robotic sampling of ocean features; enabling information driven measures  sampling efforts to high-interest regions.
-These regions are usually characterized by gradients of measurable environmental variables, e.g., temperature or salinity gradients, on which EPs subsequently can be used...Correlation among samples and multivariate requirements are typical in environmental studies.
-
 \end{abstract}
@@
 \section{Introduction}
-
-%NB: The subsections are used as a temporary instrument to highlight the proposed structure and should probably be abolished upon convergence to a stable version.
 
@@
 
-%applications and beyond.  
- %By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance.
-This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.
-
 Data collection at sea has typically been based on static buoys,
@@
   \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
-    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
+    0.49\textwidth]{Figures/web/river_proccess.png}\label{fig:nidelven}}
   \caption{\ref{fig:envir1} Traditional ocean observation based on 
     ship-based sampling has been augmented by autonomous
-    robotic vehicles such as AUVs. % and their interactions.  
-    %AUV platforms are an integral part of this network being able to reason and make decisions for efficient onboard adaptive sampling.
-    % using the sense-plan-act control approach to autonomous control. 
+    robotic vehicles such as AUVs. 
     \ref{fig:nidelven} The interaction of river and ocean creates
@@
 spatial data collection for vector-valued random fields.
-%Because of limited resources to sample the large oceanographic domains, one must plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs.
 Given the operational constraints on AUV movements and the fact that
@@
 for ESs of Gaussian random fields (GRFs) in order to address this
-research
-problem.
-%approach will take substantial advantage of existing work in the field of excursion set estimation mainly dedicated to computer experiments in the scalar-valued case, and where space exploration can be performed without constraints regarding the distance between successive design points.
+research problem. 
 We briefly review recent advances in targeted sequential design of
@@
 
-% From earlier version; to be relocated?
-%Hence statistical proxy models of the environment must be used.
-For our purpose, we rely on Gaussian Random Field (GRF) representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes.
 \subsection{Previous work in AUV sampling}
@@
 derived.
-%The resulting formulae form the backbone of the expected sequenatil uncertainty reduction strategies presented in Section \ref{sec:heuristics}.
 Section \ref{Sec:UnivarEx} illustrates the concepts on a bivariate
@@
 
-%The methods we develop are fairly general, and not limited to the special case of two-dimensional random fields.
-Hence, the rest of this section will consider generic excursion sets of gaussian random fields with an arbitrary number of output dimensions.
-% Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
 For the general setting of a $\no$-variate random field, we are
@@
 prove useful to study $\es$.
-%under some measure defined on the domain.
-%\textcolor{red}{Is this really needed?} 
 Note that while some aspects of the developed approaches do not call
@@
 %
-%This slight change of notation 
 will allow us to think of $\gp$ as a scalar-valued random field indexed by $\domain \times \{1\dots,p\}$, which will give the co-Kriging equations a particularly simple form that parallels the one of univariate Kriging.
-%Due to the naturality of the concept of \textit{generalized location}, we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $\spatloc \in \domain$.
-In the following, the letter $x$ will be reserved for generalized locations, while
 The letters $\spatloc$ and $\ell$ will be used for spatial locations and response indices respectively.
 %
-% Replacement by overlooked revised chunks from June 19th
-%\begin{comment}
-%Given a dataset consisting of  $q$ observations (a batch) at spatial locations $\spatloc_i \in \domain$ and response indices $l_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we use boldface letters to denote concatenated quantities;
-%\begin{align*}
-%\bm{x}:=(\bm{\spatloc}, \bm{\ell}):= (x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
-%\end{align*}
-%In general,  corresponding to batches of observations. %The field values of batch observations are denoted by
-%\begin{align*}
-%\gp[\bm{x}]:=
-%\left(\gp[\spatloc_1,\ell_1], ...,
-%\gp[\spatloc_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
-%\end{align*}
-%In the same fashion, $\mu(\bm{x})$ denotes the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ the $q \times q$ covariance matrix.
-%Given a random field $\gp$ and (noisy) observations of some of its components at selected points in the domain, one can predict the value of the field at location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$ and the associated covariance .
-This process is called co-Kriging.% , and kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
-%We present a general form of cokriging, where observations at locations $s \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ (heterotopic).
-%Assuming that one wishes 
-%to predict $\gp[\bm{x}]$ at $q\geq 1$ generalized locations $\bm{x}=(\bm{\spatloc}, \bm{\ell})$, 
-%given $n$ batched of observations $\mathbf{z}_{[n]}$ at $\sum_{i=1}^n q_i$ generalized locations $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$.
-%The co-Kriging mean then amounts to simple kriging with respect to a scalar-valued GRF indexed by $\domain \times \{1\dots,p\}$: % and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{\spatloc}, \bm{\spatloc}')_{\ell, \ell'}$, that is:
-%
-%\begin{equation}\label{eq:cokrig_mean}
-%\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})),
-%\end{equation}
-%where $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of observed responses of $Z$ at all considered generalized locations, and 
-%where $\lambda_{[n]}(\bm{x})=\left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right]^{-1} k(\bm{x}_{[n]}, \bm{x})$ is a vector of weights and matrix $R(\bm{x}_{[n]}, \bm{x}_{[n]})$ holds the observation noise covariance at the generalized locations.
-%$k(\bm{x}_{[n]}, \bm{x}_{[n]})$ being assumed non-singular throughout the presentation. 
-%The associated co-Kriging %(conditional) 
-%residual cross-covariance function can be expressed in the same vein via
-%
-%\begin{equation}\label{eq:cokrig_cov}
-%k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T \left[ k(\bm{x}_{[n]}, \bm{x}_{[n]}) + R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n]}}(\bm{x}').
-%\end{equation}
-%\subsubsection{co-Kriging update formulae}
-%In our case it is important to update the GRF model efficiently onboard the vehicle after every batch of obervations.
-Our setting is like that of a spatial Kalman filter with static model parameters, and we provide the co-Kriging equations for this case.
-Because of our representation of co-Kriging in terms of simple kriging with respect to generalized locations, the formulae is a straightforward adaptation of the batch-sequential Kriging update formulae from \cite{Chevalier.etal2013a}.
-The formulae are instrumental in deriving semi-analytical results for stepwise expected uncertainty reduction criteria for vector-valued random fields in the subsequent sections.
-%Given the observations from $n$ batches, one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
-The mean update formulae is
-%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
-% delivers that
-% 
-%\begin{equation}\label{eq:meanCoK}
-%\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
-%\end{equation}
-%where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
-%$\lambda_{[n+1]}(\bm{x})$ that corresponds to the Kriging weigths associated with the last $q_{n+1}$ responses when 
-%predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
-%\text{th}$ batch.
-%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
-%The updated co-Kriging residual cross-covariance function is
-%\begin{eqnarray}\label{eq:varCoK}
-%k_{[n+1]}(\bm{x},\bm{x}') &=& k_{[n]}(\bm{x},\bm{x}') -c(\bm{x},\bm{x}',\bm{x}_{[n]}), \\
-%c(\bm{x},\bm{x}',\bm{x}_{[n]})&=&\lambda_{[n+1,n+1]}(\bm{x})^T \left[k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})+R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n+1,n+1]}}(\bm{x}'). \nonumber
-%\end{eqnarray}
-%\medskip
-%
-%We remark that, as noted in \cite{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
-%extend to Universal Kriging in second-order settings and apply without Gaussian assumption. 
-%\end{comment}
-%
-%It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the 
-%one for a single observation, provided some notation adjustments are made. This motivates the following.
 Furthemore, boldface letters will be used to denote concatenated
@@
 simple expressions for the co-Kriging equations below.
-%\textcolor{red}{(ADDENDUM, C.T.) In the same fashion, we will use $\mu(\bm{x})$ to denote the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ for the corresponding $q \times q$ covariance matrix.}
 
@@
 some batch of $q\geq 1$ generalized locations
-$\bm{x}$, %\equiv(\bm{\spatloc}, \bm{\ell})$,
-the simple co-Kriging mean then amounts to Kriging with respect to a
+$\bm{x}$, the simple co-Kriging mean then amounts to Kriging with respect to a
 scalar-valued GRF indexed by $\domain\times \{1\dots,p\}$:
@@
 
-%\subsubsection{co-Kriging update formulae}
-
 Let us now consider the case where a co-Kriging prediction of $Z$ was
@@
 $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
-%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
 Thanks to our representation of co-Kriging in terms of simple Kriging
@@
 to the Kriging weights for the last $q_{n+1}$ responses
-when predicting at $\bm{x}$ relying on all measurements until batch
-$(n+1)$.
-%\text{th}$ batch.
-%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
+when predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
 The associated co-Kriging residual (cross-)covariance function is
@@
 \end{align}
-%\medskip
 %
@@
 \EMV
-%\operatorname{Var}[\mes(\es)]
 &=\int_{\domain^2}
@@
 \
-\mathrm{d}\mes^{\otimes} %\mes
-%\productMeasure
+\mathrm{d}\mes^{\otimes} 
 (u,v)\\
@@
 expectation and probability conditional on the first $n$ batches of
-observations, respectively. %We want to compute the effect of including a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and $\ibv$. 
-We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.
+observations, respectively. We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.
 
@@
 \end{matrix}
-%\begin{matrix}
-%\currentCov{u, u} & \Delta_{[n]}(\uu)\\
-%\Delta_{[n]}(\uu) & \currentCov{u, u}
-%\end{matrix}
 \right).\\
@@
 \
-\mathrm{d}\mes^{\otimes} %\mes 
-%\productMeasure 
+\mathrm{d}\mes^{\otimes} 
 (u,v)\\
@@
 \right)
-\mathrm{d}\mes^{\otimes} %\mes 
-%\productMeasure 
+\mathrm{d}\mes^{\otimes} 
 (u,v)
@@
 \end{equation*}
-%and the other blocks are defined by following the same logic, exchanging the role of $\uu$ and $\vv$.
 \end{propo}
@@
 \centering
-\includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
+\includegraphics[width=0.99\textwidth]{Figures/illus_bivar-eps-converted-to.pdf}%illus_bivar.epse}
 \caption{Density contour plots with increasing correlations between
@@
 correlation is low ($\gamma=0.2$), there is little information about
-salinity in the temperature data, and therefore less uncertainty
-reduction.
-%In an application with fresh cold water from a river source, the temperature and salinity variables will not only be interdependent, but will also likely show dependence in the spatial dimension.
-This in turn will impact the design criteria when we
-% evaluate the information measure by integrating over several locations (Section \ref{sec:simulations}).
+salinity in the temperature data, and therefore less uncertainty reduction.
 
@@
 \subsection{Background}
-%As before, let $\gp$ be a $\no$-variate random field on $\domain$.
-See Section \ref{sec:bg_and_notation} for a recap of the specific notation used to handle mean and covariance function of multivariate random fields.
 
@@
 height=0.26\textwidth,keepaspectratio]{Figures/ebv_myopic_radar.png}\label{fig:compass_myopic}}
-\caption{Example run of the myopic strategy on a given realisation of the GRF model from \ref{sec:including_spatiality}.
-Past observation locations are depicted in green and current position of the AUV in cyan.
-The expected EBV reduction associated to data collection at neighbouring nodes of the current location is shown in \ref{fig:compass_myopic}.
-The thick and light color indicate the north-east node at $30 ^{\circ}$ to be the best possible choice.}
+\caption{Example run of the myopic strategy on a realisation of the GRF
+model from \ref{sec:including_spatiality}. Reduction in Bernoulli variance
+compared to the prior is shown in \ref{fig:route_myopic}, with past observation
+locations in green and current AUV position in cyan. The expected IBV reduction
+associated to data collection at neighbouring nodes of the current location is
+shown in \ref{fig:compass_myopic}. The thick and light color indicates the node at $30 ^{\circ}$ to be the best possible choice.}
 \label{fig:ebv_myopic}
@@
 
-%Three different designs are considered as indicated in Figure \ref{fig:stat_design}. 
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.65\textwidth]{Figures/Des3.eps}
-%\caption{Three different static survey designs plotted on the initial EP.}\label{fig:stat_design}
-%\end{figure}
-%In this display the designs are plotted along with the prior probability contours of the ES for the reference parameter inputs.
-
 We compare three different static designs denoted
@@
 of realizations from the model while conducting simulated sequential
-surveys for each one. An example of such a realization with a \textit{myopic} strategy is shown in Fig.~\ref{fig:route_myopic}.
+surveys for each one. An example of such a realization with a \textit{myopic}
+strategy is shown in Fig.~\ref{fig:ebv_myopic}.
 
@@
 
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
-%\caption{The equilateral waypoint graph used to discretize the
-%  trajectory choices over the $31\times31$ grid used to discretize the GP.}
-%\label{fig:wp_graph}
-%\end{figure}
-
 \begin{figure}[!b] 
@@
 \subfigure[The waypoint graph.]{\includegraphics[width =
-0.49\textwidth]{Figures/sim/wp_graph_paper.pdf}\label{fig:wp_graph_a}}
+0.49\textwidth]{Figures/web/wp_graph_paper.jpeg}\label{fig:wp_graph_a}}
 \hfill
@@
   \centering
-  % \subfigure[Excursion set variance $E_{\by}(p[1-p])$.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
-  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_EV.pdf}}
+  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/web/avg_EV.jpeg}}
   \hfill \subfigure[RMSE between estimated field and
@@
 
-%In Fig. \ref{fig:route_choices}, the realized sampling paths for each of the sequential schemes and static designs are shown. The
-%\textit{naive} strategy often gets stuck in the southern part of the
-%domain because it is too focused on the probabilities near $0.5$. The
-%\textit{myopic} strategy covers a wider domain than the naive or
-%look-ahead. There are several reasons for this. First, a greedy
-%approach will tend to put more emphasis on promising locations close
-%to the agent, which may lead away from the centre. Second, as the
-%agent evaluates the impact of locations further away (look-ahead)
-%where assimilated data has less predictive power, the GP model (which
-%is centered here) will act to restrict paths deviating from the
-%central zone.
-
 We studied the sensitivity of the results by modifying the input
@@
 range. 
-%\textit{static\_north} continues to be the best static design
-%for IBV, while \textit{static\_zigzag} is the best design for the
-%other predictive performance measures, especially so with large
-%spatial correlation range. 
 We also ran simulation studies with only temperature data, and for
@@
 Next, the residuals from the regression analysis were analyzed to specify the covariance parameters of the GRF model.
-%\begin{figure}[!h] 
-% \centering 
-%\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
-%\caption{Data analysis from a preliminary trial experiment using the
-%  AUV. Left: Residual plot of temperature and salinity along with
-%  Gaussian contours. Middle: Empirical CDF (solid) of the quadratic form of
-%  the residuals along with the theoretical CDF (dashed) of the $\chi^2$
-%  distribution with two degrees of freedom. Right: Empirical variogram
-%  of the salinity and temperature data.} \label{fig:parest}
-%\end{figure}
 \begin{figure}[!b]
   \centering
-  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_a.pdf}\label{fig:parest_a}}
+  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_a.pdf}\label{fig:parest_a}}
   \hfill
-  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_b.pdf}\label{fig:parest_b}}
+  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_b.pdf}\label{fig:parest_b}}
   \hfill
-  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_c.pdf}\label{fig:parest_c}}
+  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_c.pdf}\label{fig:parest_c}}
   \caption{Data analysis from a preliminary trial experiment using the
@@
 the river and fjord water masses.
-
-%Mapping the spatial extent of a frontal zones is an important problem for studying many bio-physical interactions in the ocean.
-The frontal zone is determined by the boundary where plumes of sediments, nutrients, and possibly pollutants spreading from the river outlet meet and interact with adjacent coastal water.
-Due to the lower density the plumes spread on the surface, creating a front with an sharp gradient in both temperature and salinity.
 
@@
 \centering 
-\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
+\includegraphics[width=0.98\textwidth]{Figures/web/harald.jpeg}
 \caption{The commercially available Light Autonomous Underwater
@@
 \centering
-\subfigure[AUV survey area]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/alt_map.eps}\label{fig:map}}
+\subfigure[AUV survey area]{\includegraphics[height=0.41\textwidth]{Figures/web/alt_map.jpeg}\label{fig:map}}
 \hspace{0.3cm}
-\subfigure[Temperature tracks]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/auv.pdf}\label{fig:res_both}}
-
-\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv1_es_sal_ep.pdf}\label{fig:res1}}
+\subfigure[Temperature tracks]{\includegraphics[height=0.41\textwidth]{Figures/web/auv.jpeg}\label{fig:res_both}}
+
+\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/web/auv1_es_sal_ep.jpeg}\label{fig:res1}}
 \hspace{0.2cm}
-\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv4_es_sal_ep.pdf}\label{fig:res2}}
-%\subfigure[ES for Survey 1]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
-%\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
+\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/web/auv4_es_sal_ep.jpeg}\label{fig:res2}}
 \caption{Results from mapping the Nidelva river, Trondheim, Norway
@@
 need for real-time multivariate spatial sampling on autonomous
-platforms.% , which statisticians can exploit to create general-purpose
-% toolkits for similar applications.
+platforms.
 
@@
 
-% The sometimes conflicting topics of autonomy and communication have
-% caught much interest in control engineering lately
-% \citep{zolich2019survey}, and statistical evaluations could contribute
-% in the future. In our context, autonomy is highlighted, and this is
-% reflected in our modeling and computational assumptions. With the
-% possibilities of wifi communication to a research vessel at waypoints,
-% more advanced modeling and computational routines are possible, but
-% with new challenges related to design and planning for when and where
-% the communication should occur.
-
 \section*{Acknowledgements}
@@
 
-%  for pointing out the relevance of revisiting the proposed
-% approach for estimating volumes of mixing zones, e.g. in hydrology.
-
-%\begin{supplement}
-%\sname{Supplement A}\label{suppA}
-%\stitle{Title of the Supplement A}
-%\slink[url]{http://www.e-publications.org/ims/support/dowload/imsart-ims.zip}
-%\sdescription{Dum esset rex in
-%accubitu suo, nardus mea dedit odorem suavitatis. Quoniam confortavit
-%seras portarum tuarum, benedixit filiis tuis in te. Qui posuit fines tuos}
-%\end{supplement}
-
-% == Adding references
 \footnotesize
@@
 \input{appendix}
-
-% AOS,AOAS: If there are supplements please fill:
-%\begin{supplement}[id=suppA]
-%  \sname{Supplement A}
-%  \stitle{Title}
-%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
-%  \sdatatype{.pdf}" 
-%  \sdescription{Some text}
-%\end{supplement}
-
-% === Not used
-% After reviews: - This is moved to Section 2.1.
-
-%Traditional data collection at sea has typically been based on static
-%buoys, Lagrangian floats, or ship-based methods, with significant
-%logistical limitations that directly impact coverage and sampling
-%resolution. Modern methods using satellite remote-sensing provide
-%large-scale coverage but have limited resolution, are limited to
-%sensing the surface of the ocean, and are impacted by cloud cover. The
-%advent of robust mobile robotic platforms \citep{Bellingham07} has
-%resulted in significant contributions to environmental monitoring and
-%sampling. In particular, autonomous underwater vehicles (AUVs), have
-%advanced the state of sampling and consequently have made robotics an
-%integral part of ocean observation; our previous work has contributed
-%to this effort \citep{das11b,Das2015,fossum18b,fossuminformation}. Other ¤ %statistical work in the oceanographic domain include \cite{wikle2013modern}
-%focusing on hierarchical statistical models; \cite{sahu2008space},
-%studying spatio-temporal models for sea surface temperature and
-%salinity data; and \cite{mellucci2018oceanic} looking at the
-%statistical prediction of features using an underwater glider.
-
-%\begin{itemize}
-%
-%\item Full numerical ocean models cannot provide accurate results
-%  online if run on robotic sensing platforms, as onboard computers
-%  cannot deliver the computational power required. Hence statistical
-%  proxy models of the environment must be used for learning where to sample.
-%
-%\item With limited available information about the state of the ocean, there is substantial value in reacting to
-%  information obtained from measurements taken in-situ. This
-%  acquired information must be assimilated into statistical models that
-%  can be used to inform decisions on where to sample
-%  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.
-%
-%\item For sampling problems related to environmental sensing, the
-%  number of choices (i.e. locations, trajectories, and candidate
-%  designs) is enormous, creating a
-%  trade-off between optimization (finding the most resource-efficient
-%  design to collect necessary data) and computability (arriving at a
-%  solution in reasonable time). To successfully resolve features, this
-%  trade-off has to be considered in development and practice.
-%
-%\end{itemize}
-
-%Addressing this, the combination of statistical tools and robotic platforms is a
-%natural symbiosis which enables information-based sensing. Central to
-%this is the ability to model spatially-correlated variables and
-%provide formal measures of uncertainty. Our formulation is based on
-%Gaussian Processes (GPs) as they allow efficient implementation 
-%and evaluation in real time onboard a robotic platform.
-
-%Sampling can, in this context, not simply be distributed evenly ---along simple transects or ``lawn-mover" patterns--- but must instead be prioritized to relevant regions to ensure it is cost-effective while providing adequate coverage and resolution of the area of scientific interest.
-
-%While the focus has often been on
-%biological and anthropogenic impact from micro-plastics to
-%pollution, biological oceanographers have focused intently on studying
-%micro-organisms at the base of the human food web. These organisms are
-%critically impacted by the changing dynamics in the upper water-column,
-%especially in coastal zones which are complex and often hard to observe
-%in space and time. By studying the bio-geochemical processes in the
-%upper water-column scientists can measure the impact of change, natural
-%or anthropomorphic, and provide an informed opinion to policy makers to
-%effect changes in preserving the environment. However, the challenge of 
-% The pressure on marine resources is growing and increased accuracy,
-% resolution, and persistent monitoring of the oceans is crucial for
-% long-term sustainable management.
-
-% A
-% sustained focus on prioritized and efficient data collection strategies
-% have therefore started to emerge. The advent of marine robotic
-% platforms, especially
-
-% provided means to execute this prioritization through the capacity of
-% autonomy and data-driven sampling, where data collection in principle
-% can be optimized. These capabilities have made
-
-% of the emerging sensing practice for ocean science, allowing scientists
-% to increase the observational efficiency and resolution beyond what was
-% previously possible. But how should a robotic platform, such as an AUV,
-% effectively prioritize and identify important regions for sampling? The
-% answer to this question relates to
-
-% , and
-% the application domain is clearly an arena where statisticians can
-% contribute.
-% There has recently been some statistical attention in oceanography:
-
-%From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest.
-Each phenomenon can in turn be characterized by a set of process specific conditions expressed through different measures of key environmental variables, such as temperature or salinity.
-One such measure is the gradient, that can be associated with a number of important processes, such as the vertical location of the thermocline and pycnocline, location of upwelling systems, vertical mixing, eddies, fronts, and currents \cite{sverdrup2006}, as well as distribution, growth, and accumulation of biological activity \cite{SatOceanSoci00, Ryan2014}.
-These gradients create boundaries separating the ocean into process specific regions which are of profound interest to both identify and map effectively.
-Quantification of gradient features is therefore a much needed competence in robotic sampling.
 \end{document}

### 2021-01-04 ans_to_reviewers/ans-to-2nd-rev.tex
Updates from Overleaf

@@
 \vspace{5mm}
-\noindent \aecom \textbf{1}:
+\noindent \aecom \textbf{1 - Reply to AE comment 4}: 
+T hank you for explaining the sampling limitations to me in your reply.
+However, I think these limitations should also be noted in the narrative, and the sub-optimality of the resulting trajectory should be acknowledged.
 
-Bla bla bla.
+\vspace{5mm}
+\reply:
 
+\vspace{5mm}
+\noindent \aecom \textbf{2 - Page 2, line 15}:  I think it would be more appropriate to refer to this as sampling design rather than sampling.
+
+\vspace{5mm}
+\reply:
+
+\vspace{5mm}
+\noindent \aecom \textbf{3 - Page 4, line 3}:  Replace “proxys” with the correct plural “proxies”
+
+\vspace{5mm}
+\reply:
+
+\vspace{5mm}
+\noindent \aecom \textbf{4 - Page 5, line 16:}  Change  “the  PhD  thesis  (Stroh,  2018,  p.  82)”  to ”Stroh (2018, p. 82)”
+
+\vspace{5mm}
+\reply:
+
+\vspace{5mm}
+\noindent \aecom \textbf{5 - Page 22, line 8:}  From Figure 9(c), the nugget effect in the empirical variogram appears to be approximately 0.4 for salinity and 0.7 for temperature.
+Even if the sensor specifications were essentially 0, it would be a stretch to say that a variance of 0.0625 was a “middle ground” between the empirical nugget effect and the sensor specifications.
+So either give a better justification of your choice of 0.0625 or, if that’snot possible, then redo the analysis with a more justifiable choice.
+
+\vspace{5mm}
+\reply:
+
+\vspace{5mm}
+\noindent \aecom \textbf{6 - Page 22, line 5:}  Insert “triangular” immediately after “equilateral”
+
+\vspace{5mm}
 \reply:

### 2021-01-05 ans_to_reviewers/ans-to-2nd-rev.tex
Updates from Overleaf

@@
 
-\reply:
+\reply: We agree and have addressed this point in \aecom \textbf{1 - Reply to AE comment 4}.
 
@@
 \vspace{5mm}
-\reply:
+\reply: Line type will be tricky for Fig. 8, when lines are very close together especially in Fig.8b, and Fig.8c.
 
@@
 \noindent \aecom \textbf{1 - Reply to AE comment 4}: 
-T hank you for explaining the sampling limitations to me in your reply.
+I hank you for explaining the sampling limitations to me in your reply.
 However, I think these limitations should also be noted in the narrative, and the sub-optimality of the resulting trajectory should be acknowledged.
@@
 \vspace{5mm}
-\reply:
+\reply: We agree that this should be mentioned in the text. We have added the following text in the Closing remarks section at page 25.
+
+\begin{quote}
+    "The sampling capabilities are restricted by the sensors onboard the AUV, the computational load and navigation resources.
+The sampling of temperature and salinity variables can be much higher than the grid resolution, but, due to computational and navigation resources available on the AUV, assimilation into a discretized fixed grid is necessary.``
+\end{quote}
 
@@
 \vspace{5mm}
-\reply:
+\reply: We have removed the word sampling from the sentence, with made it more clear and readable. See the change below:
+
+\begin{quote}
+    "Models and methods from spatial statistics and experimental design can clearly contribute to this challenge.``
+\end{quote}
 
@@
 \vspace{5mm}
-\reply:
+\reply: Changed to “proxies” on page 4.
 
@@
 \vspace{5mm}
-\reply:
+\reply: We have changed the reference accordingly in the text at Page 5.
 

### 2021-01-05 main_new.tex
Updates from Overleaf

@@
 domain with the aim to pursue measurements with high scientific
-relevance.  Models and methods from spatial statistics and experimental design can clearly contribute to this sampling challenge.
+relevance. Models and methods from spatial statistics and experimental design can clearly contribute to this challenge.
 
@@
 operations, but off-the-shelf instruments provide temperature and
-salinity measurements which serve as proxys for the underlying
+salinity measurements which serve as proxies for the underlying
 biological phenomenon. With the help of a vector-valued random field
@@
 optimization, but not for ES's like we do here. More relevant to our
-setting, the PhD thesis \citep[][p.82]{stroh} mentions general
-possibilities of stepwise uncertainty reduction strategies for ES's in
+setting, \citet[p.82]{stroh} mentions general
+possibilities of step-wise uncertainty reduction strategies for ES's in
 the context of designing fire simulations, yet outputs are mainly
@@
 and envision that advection-diffusion equations could be useful
-\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
+\citep{sigrist2015stochastic,richardson2017sparsity}.
+The sampling capabilities are restricted by the sensors onboard the AUV, the computational load and navigation resources.
+The sampling of temperature and salinity variables can be much higher than the grid resolution, but, due to computational and navigation resources
+available on the AUV, assimilation into a discretized fixed grid is necessary. For more complex
 oceanographic phenomena, the methods will need to be extended to

### 2021-01-07 ans_to_reviewers/ans-to-2nd-rev.tex
Updates from Overleaf

@@
 
-\reply: We agree and have addressed this point in \aecom \textbf{1 - Reply to AE comment 4}.
+\reply: We agree and have addressed this point in \aecom \textbf{1}.
 
@@
 \vspace{5mm}
-\reply:
+\reply: We have elaborated on the specification of nugget in the revised version. We have addressed this point in \aecom \textbf{5}.
 
@@
 \vspace{5mm}
-\reply:
+\reply: We agree that the remarks of Ehrenberg are very useful!
+(It should be curriculum for graduate students.) Regarding the structure, we seriously considered your suggestions about the background in supplementary material, but in the end thought it was too much of a reduction to go through with.
+It would be very difficult to grasp the applied contribution of our paper if this background part was in the supplemental material, and only some readers take the extra effort of reading supplemental material.
+Based on the report of the reviewer, we did however go through the background and cut a few unnecessary and lengthy formulations in the revised version.
 
@@
 \vspace{5mm}
-\reply: Line type will be tricky for Fig. 8, when lines are very close together especially in Fig.8b, and Fig.8c.
+\reply: We agree on all our points here, except Fig.8.
+Line type will be tricky for Fig.
+8, when lines are very close together especially in Fig.8b, and Fig.8c.
+So we ask to consider that this figure is kept in color.
+Fig.
+4, 5, and 6 have been made with a black-and-white compatible color-map, so they can be converted to black-and-white.
 
@@
 \vspace{5mm}
-\reply:
+\reply: We agree that this effort should be noted. We have added the following statement to the acknowledgements.
+
+\begin{quote}
+"We thank the anonymous reviewers for their careful reading of our manuscript and their many insightful comments and suggestions.``
+\end{quote}
 
@@
 \noindent \aecom \textbf{1 - Reply to AE comment 4}: 
-I hank you for explaining the sampling limitations to me in your reply.
+I thank you for explaining the sampling limitations to me in your reply.
 However, I think these limitations should also be noted in the narrative, and the sub-optimality of the resulting trajectory should be acknowledged.
@@
 \vspace{5mm}
-\reply:
+\reply: We have elaborated on the specification of nugget in the revised version.
+The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, and in related data-sets we have not seen any nugget effect when measuring temperature and salinity in water samples.
+We believe that one challenge with the initial sampling line is that it was rather short and that it did no samples at very short distance.
+We made an empirical variogram from the actual survey data, which are shown below in Figure \ref{empVar}, and these curves support our statements of a smaller nugget.
+\begin{figure}[!b] 
+\centering
+\includegraphics[width=0.99\textwidth]{Figures/field-trials/empVAR.png}
+\caption{Empirial variogram from the two runs. }
+\label{empVar}
+\end{figure}
+
+One reason to keep a significant nugget effect here is to account for the dynamic water masses that are not captured in our model, but we expect this to be very small in our limited survey time.
+Then again, between the two surveys, there is a change in the survey 1 and 2 paths caused by slightly different measurements at the two times of surveying.
+Another reason for including a nugget is the uncertainty in the AUV location which means it samples might not be allocated to its true position.
+
+We further checked the sensitivity of the nugget in the sequential algorithm.
+This was done by running synthetic cases of sampling paths and data up to a certain stage, but then updating the model with different nugget effects in the Gaussian random field model.
+Figure X shows sampling paths and the radar plots for the AUV direction choice at one steps (top vs bottom) for two different nuggets.
+Straighter lines when the nugget is large.
+
+In the revised manuscript we wrote another sentence on Fig 11: 'xx'
 
@@
 \vspace{5mm}
-\reply:
+\reply: We have improved our
 

### 2021-01-07 main_web.tex
Updates from Overleaf

@@
 optimization, but not for ES's like we do here. More relevant to our
-setting, the PhD thesis \citep[][p.82]{stroh} mentions general
-possibilities of stepwise uncertainty reduction strategies for ES's in
+setting, \citet[p.82]{stroh} mentions general mentions general
+possibilities of step-wise uncertainty reduction strategies for ES's in
 the context of designing fire simulations, yet outputs are mainly
@@
 
-The statistical model parameters were specified based on a short
+The statistical model parameters were specified based on experience as well as data gathered in a short
 preliminary survey where the AUV made an initial transect to determine
-the trends in environmental conditions and correlation
+the trends in environmental conditions, variance and correlation
 structures.
-Based on the initial runs we get a reasonable idea of the temperature and salinity of river and ocean waters, and also specify the trend by linear regression, where both temperature and salinity were assumed to increase linearly with the west coordinate.
+This gives a reasonable idea of the temperature and salinity of river and ocean waters.
+Both temperature and salinity were assumed to increase linearly with the west coordinate.
 Next, the residuals from the regression analysis were analyzed to specify the covariance parameters of the GRF model.
-\begin{figure}[!b]
-  \centering
-  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_a.pdf}\label{fig:parest_a}}
-  \hfill
-  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_b.pdf}\label{fig:parest_b}}
-  \hfill
-  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_c.pdf}\label{fig:parest_c}}
-  \caption{Data analysis from a preliminary trial experiment using the
-    AUV. \ref{fig:parest_a} Residual plot of temperature and salinity
-    along with Gaussian contours. \ref{fig:parest_b} Empirical CDF
-    (solid) of the quadratic form of the residuals along with the
-    theoretical CDF (dashed) of the $\chi^2$ distribution with two
-    degrees of freedom. \ref{fig:parest_c} Empirical variogram of the
-    salinity and temperature data.}
-\label{fig:parest}
-\end{figure}
-Fig.~\ref{fig:parest} summarizes diagnostic plots of this
-analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
-and salinity residuals after the regression mean values of salinity and
-temperature are subtracted from the data. This scatter-plot of joint
-residuals indicates larger variability in salinity than temperature,
-and a positive correlation ($0.5$) between the two variables. Based on
-the fitted bivariate covariance model (ellipses in
-Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
-the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
-the empirical CDF of the quadratic forms (solid) together with the
-theoretical CDF of the $\chi^2_2$ distribution (dashed). The modeled and
-theoretical curves are rather similar, which indicates that the Gaussian
-model with constant spatial variance and correlation fits reasonably well. Fig.~\ref{fig:parest_c} shows the
-empirical variogram of the residuals for temperature and salinity. The decay is similar for the two, and seems to be negligible
-after about $150$ m.  The working assumption of a separable covariance
-function is hence not unreasonable.
-
-Based on the analysis in Fig.~\ref{fig:parest}, the resulting
+The resulting
 parameters are given in Table \ref{tab:experiment_param}. The
@@
 from the temperature sensor. We specify variance $0.25^2$ for both
-errors, which is based on a middle ground between the nugget effect in
-the empirical variogram and the sensor specifications.
+errors.
+This is based on prior experience indicating that there can be some noise in temperature and salinity readings mostly because of dynamic water masses and uncertainty in AUV position.
+The initial survey transect was not sampled very densely so a reliable extraction of a nugget parameter from those data was not possible.
 
@@
 
+\begin{figure}[!b]
+  \centering
+  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_aN.eps}\label{fig:parest_a}}
+  \hfill
+  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_bN.eps}\label{fig:parest_b}}
+  \hfill
+  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_cN.eps}\label{fig:parest_c}}
+  \caption{Data analysis from from AUV experiments. \ref{fig:parest_a} Residual plot of temperature and salinity
+    along with specified Gaussian contours. \ref{fig:parest_b} Empirical CDF
+    (solid) of the quadratic form of the residuals along with the
+    modeled theoretical CDF (dashed) of the $\chi^2$ distribution with two
+    degrees of freedom. \ref{fig:parest_c} Scaled empirical variograms of the
+    salinity and temperature data.}
+\label{fig:parest}
+\end{figure}
+
+Of course, the initial 
+Fig.~\ref{fig:parest} summarizes diagnostic plots of this
+analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
+and salinity residuals after the regression mean values of salinity and
+temperature are subtracted from the data. This scatter-plot of joint
+residuals indicates that the correlation between the two variables might be slightly larger than the initially specified $0.5$ and that uncertainty in temperature was set a bit too high. Based on
+the fitted bivariate covariance model (ellipses in
+Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
+the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
+the empirical CDF of the quadratic forms (solid) together with the
+theoretical CDF of the $\chi^2_2$ distribution (dashed). The empirical and
+theoretical curves are not very different, which indicates that the specified Gaussian
+model fits reasonably well. Fig.~\ref{fig:parest_c} shows the
+empirical variogram of the scaled residuals for temperature and salinity. The decay is similar for the two, and seems to be negligible
+after about $150$ m as was specified. The working assumption of a separable covariance
+function is hence not unreasonable.
+
+Overall, the spatial Gaussian model representation is of course a simplification of the changing water conditions, and even within the model formulation, the parameter specification could likely have been improved.
+However, we do not expect minor changes to have caused much difference in sampling paths.
+
 \section{Closing remarks}
@@
 and envision that advection-diffusion equations could be useful
-\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
+\citep{sigrist2015stochastic,richardson2017sparsity}.
+The sampling capabilities are restricted by the sensors onboard the AUV, the computational load and navigation resources.
+The sampling of temperature and salinity variables can be much higher than the grid resolution, but, due to computational and navigation resources
+available on the AUV, assimilation into a discretized fixed grid is necessary.For more complex
 oceanographic phenomena, the methods will need to be extended to
@@
 Operations and Systems (AMOS), Center of Excellence, project number
-223254, and the Applied Underwater Robotics Labortatory (AURLab). CT
+223254, Nansen Legacy Program, project number 276730, and the Applied Underwater Robotics Labortatory (AURLab). CT
 and DG acknowledge support from the Swiss National Science Foundation,
@@
 aknowledge support of Idiap Research Institute, his primary
-affiliation in an early version of this manuscript. The authors would
-also like to thank Niklas Linde of the University of Lausanne for
-providing constructive feedback about this work, and members of the
-NTNU AURLab for help with AUV deployments
+affiliation in an early version of this manuscript.
+We thank the anonymous    reviewers for their careful reading of our manuscript and their many insightful comments and suggestions.
+The authors would also like to thank Niklas Linde of the University of Lausanne for providing constructive feedback about this work, and members of the NTNU AURLab for help with AUV deployments
 

### 2021-01-08 ans_to_reviewers/ans-to-2nd-rev.tex
Updates from Overleaf

@@
 
+\usepackage{float}
+\usepackage{cleveref}
+
 \newcounter{reviewer}
@@
 
-We are grateful to the editors and reviewers for their comments and
-suggestions. They have all been considered in this revised
-version of our manuscript.\\
+We are grateful to the editors and reviewers for their comments and suggestions in this 2$^{nd}$ review.
+They have all been considered in this revised version of our manuscript.
+The changes in the manuscript are reflected in the points shown below.\\
 
@@
 \begin{quote}
-"The sampling capabilities are restricted by the sensors onboard
-the AUV, the computational load and navigation resources. The sampling
-of temperature and salinity variables can be at 1Hz (higher than the
-grid resolution), but, due to computational and navigation resources
-available on the AUV, this simplification of data assimilation at the
-waypoints is necessary."   
+"The sampling capabilities are restricted by the sensors onboard the AUV, the computational load, and navigation resources.
+The sampling of temperature and salinity variables can be much higher than the grid resolution, but, due to computational and navigation resources available on the AUV, assimilation into a discretized fixed grid is necessary.``
 \end{quote}
@@
 (It should be curriculum for graduate students.) Regarding the structure, we seriously considered your suggestions about the background in supplementary material, but in the end thought it was too much of a reduction to go through with.
-It would be very difficult to grasp the applied contribution of our paper if this background part was in the supplemental material, and only some readers take the extra effort of reading supplemental material.
+It would be very difficult to grasp the applied contribution of our paper if the background parts were in the supplemental material, and only some readers take the extra effort of reading that.
 Based on the report of the reviewer, we did however go through the background and cut a few unnecessary and lengthy formulations in the revised version.
+{\bf{to go through for typos, redundant text, etc}}
 
@@
 4, 5, and 6 have been made with a black-and-white compatible color-map, so they can be converted to black-and-white.
+{\bf{to check all}}
 
@@
 \reply: We have elaborated on the specification of nugget in the revised version.
-The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, and in related data-sets we have not seen any nugget effect when measuring temperature and salinity in water samples.
-We believe that one challenge with the initial sampling line is that it was rather short and that it did no samples at very short distance.
-We made an empirical variogram from the actual survey data, which are shown below in Figure \ref{empVar}, and these curves support our statements of a smaller nugget.
-\begin{figure}[!b] 
+The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, and in related data-sets we have not seen much of a nugget effect for temperature and salinity in water samples close in space and time.
+The initial sampling transect had no samples at very short distance, and it is hence not possible to see the nugget effect in the empirical variogram.
+We made an empirical variogram from the actual survey data, where we have some samples close in space, see Figure \ref{empVar}.
+These curves support our statements of a smaller nugget.
+\begin{figure}[H] 
 \centering
-\includegraphics[width=0.99\textwidth]{Figures/field-trials/empVAR.png}
-\caption{Empirial variogram from the two runs. }
+\includegraphics[width=0.6\textwidth]{Figures/field-trials/empVAR.png}
+\caption{Empirical variogram of the scaled variables from the two runs. }
 \label{empVar}
@@
 
-One reason to keep a significant nugget effect here is to account for the dynamic water masses that are not captured in our model, but we expect this to be very small in our limited survey time.
+One reason to keep a significant nugget effect here is to account for the dynamic water masses that are not captured in our model, but we expect this to be very small given our limited survey time, as was the case in \cite{fossuminformation}.
 Then again, between the two surveys, there is a change in the survey 1 and 2 paths caused by slightly different measurements at the two times of surveying.
-Another reason for including a nugget is the uncertainty in the AUV location which means it samples might not be allocated to its true position.
+There could also be some uncertainty in the AUV location leading to samples not allocated to its true position.
 
-We further checked the sensitivity of the nugget in the sequential algorithm.
-This was done by running synthetic cases of sampling paths and data up to a certain stage, but then updating the model with different nugget effects in the Gaussian random field model.
-Figure X shows sampling paths and the radar plots for the AUV direction choice at one steps (top vs bottom) for two different nuggets.
-Straighter lines when the nugget is large.
+We further checked the sensitivity of the nugget specification in the sequential sampling algorithm.
+This was done by running synthetic cases of sampling paths and data up to a certain stage for different measurement noise variance levels.
+At the end, the posterior excursion probability and the chosen sampling path are compared, as show in Figure \ref{fig:nugget_comparison}.
+We see moderate changes in the path patterns
+and overall, paths tend to be smoother for large nugget effect.
+Indeed, with small nugget, the AUV reacts more to the observed temperature and salinity levels, and this leads to more small turns in its path. 
+We note that though the sampling paths can be different, the posterior excursion probabilities are similar and will thus lead to similar estimates for the excursion set.
+We also remark that the differences in Figure \ref{fig:nugget_comparison} would get smaller if we would let the sampling continue (here stopped for presentation purposes).
 
-In the revised manuscript we wrote another sentence on Fig 11: 'xx'
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale=0.52]{ans_to_reviewers/merged_paths_2.png}
+    \caption{Comparison of sampling paths and posterior excursion probability for different values of the measurement noise standard deviation (from left to right: $0.1$, $0.25$ and $0.5$).
+The sampling has been stopped after 200 steps.}
+    \label{fig:nugget_comparison}
+\end{figure}
+
+\newpage
+
+In the revised manuscript we wrote another sentence on Fig 11: 'The AUV paths in both runs are relatively winding in the sense that the vehicle reacts much to the data.
+In our experience from simulated runs, the paths tend to be smoother when we use a larger nugget effect.
+Still, the estimated nugget from the actual trial data is about $0.1$ which is close to the specified level.'
 
@@
 \vspace{5mm}
-\reply: We have improved our
+\reply: We have improved our {\bf{to do}}
 

### 2021-01-08 main_web.tex
Updates from Overleaf

@@
 south-western corner.
+The AUV paths in both runs are relatively winding in the sense that the vehicle reacts much to the data.
+In our experience from simulated runs, the paths tend to be smoother when we use a larger nugget effect.
+But here the estimated nugget from the actual trial data is close to the specified level.
 
@@
 \citep{sigrist2015stochastic,richardson2017sparsity}.
-The sampling capabilities are restricted by the sensors onboard the AUV, the computational load and navigation resources.
+The sampling capabilities are restricted by the sensors onboard the AUV, the computational load, and navigation resources.
 The sampling of temperature and salinity variables can be much higher than the grid resolution, but, due to computational and navigation resources
-available on the AUV, assimilation into a discretized fixed grid is necessary.For more complex
+available on the AUV, assimilation into a discretized fixed grid is necessary. For more complex
 oceanographic phenomena, the methods will need to be extended to

### 2021-01-08 main_web.tex
updates

@@
 \and
-\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e5]{kanna.rajan@fe.up.pt}}
+\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5,t6}. \ead[label=e5]{kanna.rajan@fe.up.pt}}
 
@@
 \affiliation[t5]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
+\affiliation[t6]{SIFT LLC, Minneapolis, Minnesota, United States.}
 
@@
 
-While random field modeling has been one of the main topics throughout the history of spatial statistics \citep{Krige1951a,Stein1999}, even
+While random field modeling has been one of the main topics throughout
+the history of spatial statistics \citep{Krige1951a,Stein1999}, even
 for vector-valued random field models with associated prediction
@@
 there has lately been a renewed interest for random field models in
-the context of static or sequential experimental design, be it in the context of spatial data collection \citep{Mueller2007} or in
+the context of static or sequential experimental design, be it in the
+context of spatial data collection \citep{Mueller2007} or in
 simulation experiments \citep{Santner.etal2003}. As detailed in
@@
 basis to sequential design of simulations dedicated to various goals
-such as global optimization and set estimation. Of particular relevance to our context,
-\cite{Bect.etal2012} focuses on strategies to reduce uncertainties related to volumes of excursion exceeding a prescribed threshold, while \cite{chevalier2014fast} concentrates on making the latter strategies computationally efficient and batch-sequential.
-Rather than focusing on excursion volumes, approaches were investigated in
+such as global optimization and set estimation. Of particular
+relevance to our context, \cite{Bect.etal2012} focuses on strategies
+to reduce uncertainties related to volumes of excursion exceeding a
+prescribed threshold, while \cite{chevalier2014fast} concentrates on
+making the latter strategies computationally efficient and
+batch-sequential. Rather than focusing on excursion volumes,
+approaches were investigated in
 \cite{French.Sain2013,Chevalier.etal2013b,Bolin.Lindgren2015,Azzimonti.etal2016}
@@
 strategies in the case of vector-valued observations. It has been long
-acknowledged that co-Kriging could be updated efficiently in the context of sequential data assimilation \citep{Vargas-Guzman1999}, but sequential
-strategies for estimating features of vector-valued random fields are
-still in their infancy. \cite{LeGratiet.etal2015} used co-Kriging
-based sequential designs to multi-fidelity computer codes and
-\cite{Poloczek2017} used related ideas for multi-information source
-optimization, but not for ES's like we do here. More relevant to our
-setting, \citet[p.82]{stroh} mentions general mentions general
-possibilities of step-wise uncertainty reduction strategies for ES's in
-the context of designing fire simulations, yet outputs are mainly
-assumed independent.
+acknowledged that co-Kriging could be updated efficiently in the
+context of sequential data assimilation \citep{Vargas-Guzman1999}, but
+sequential strategies for estimating features of vector-valued random
+fields are still in their infancy. \cite{LeGratiet.etal2015} used
+co-Kriging based sequential designs to multi-fidelity computer codes
+and \cite{Poloczek2017} used related ideas for multi-information
+source optimization, but not for ES's like we do here. More relevant
+to our setting, \citet[p.82]{stroh} mentions general possibilities of
+step-wise uncertainty reduction strategies for ES's in the context of
+designing fire simulations, yet outputs are mainly assumed
+independent.
 

### 2021-01-08 main_web.tex
updates

@@
 sensor was used to provide salinity and temperature measurements.  The
-AUV is a powered untethered platform that operates at $1$-$3$ m/s in
+AUV is a powered, untethered platform that operates at $1$-$3$ m/s in
 the upper water column. It has a multicore GPU NVIDIA Jetson TX1
@@
 from the temperature sensor. We specify variance $0.25^2$ for both
-errors.
-This is based on prior experience indicating that there can be some noise in temperature and salinity readings mostly because of dynamic water masses and uncertainty in AUV position.
-The initial survey transect was not sampled very densely so a reliable extraction of a nugget parameter from those data was not possible.
+errors. This is based on prior experience indicating that there can be
+some noise in temperature and salinity readings mostly because of
+dynamic water masses and uncertainty in AUV position. The initial
+survey transect was not sampled very densely so a reliable extraction
+of a nugget parameter from those data was not possible.
 
@@
 Survey 2 which was coming close to the survey area borders in the
-south-western corner.
-The AUV paths in both runs are relatively winding in the sense that the vehicle reacts much to the data.
-In our experience from simulated runs, the paths tend to be smoother when we use a larger nugget effect.
-But here the estimated nugget from the actual trial data is close to the specified level.
+south-western corner.  The AUV paths in both runs are relatively
+winding in the sense that the vehicle reacts to the data. In our
+experience from simulated runs, the paths tend to be smoother when we
+use a larger nugget effect. But here the estimated nugget from the
+actual trial data is close to the specified level.
 
@@
 
-Of course, the initial 
 Fig.~\ref{fig:parest} summarizes diagnostic plots of this
 analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
-and salinity residuals after the regression mean values of salinity and
-temperature are subtracted from the data. This scatter-plot of joint
-residuals indicates that the correlation between the two variables might be slightly larger than the initially specified $0.5$ and that uncertainty in temperature was set a bit too high. Based on
+and salinity residuals after the regression mean values of salinity
+and temperature are subtracted from the data. This scatter-plot of
+joint residuals indicates that the correlation between the two
+variables might be slightly larger than the initially specified $0.5$
+and that uncertainty in temperature was set a bit too high. Based on
 the fitted bivariate covariance model (ellipses in
 Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
-the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
+the residuals, and if the model is adequate they should be
+approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
 the empirical CDF of the quadratic forms (solid) together with the
-theoretical CDF of the $\chi^2_2$ distribution (dashed). The empirical and
-theoretical curves are not very different, which indicates that the specified Gaussian
-model fits reasonably well. Fig.~\ref{fig:parest_c} shows the
-empirical variogram of the scaled residuals for temperature and salinity. The decay is similar for the two, and seems to be negligible
-after about $150$ m as was specified. The working assumption of a separable covariance
-function is hence not unreasonable.
-
-Overall, the spatial Gaussian model representation is of course a simplification of the changing water conditions, and even within the model formulation, the parameter specification could likely have been improved.
-However, we do not expect minor changes to have caused much difference in sampling paths.
+theoretical CDF of the $\chi^2_2$ distribution (dashed). The empirical
+and theoretical curves are not very different, which indicates that
+the specified Gaussian model fits reasonably
+well. Fig.~\ref{fig:parest_c} shows the empirical variogram of the
+scaled residuals for temperature and salinity. The decay is similar
+for the two, and seems to be negligible after about $150$ m as was
+specified. The working assumption of a separable covariance function
+is hence not unreasonable.
+
+Overall, the spatial Gaussian model representation is a simplification
+of the changing water conditions, and even within the model
+formulation, the parameter specification could likely have been
+improved.  However, we do not expect minor changes to have caused much
+difference in sampling paths.
 
@@
 and envision that advection-diffusion equations could be useful
-\citep{sigrist2015stochastic,richardson2017sparsity}.
-The sampling capabilities are restricted by the sensors onboard the AUV, the computational load, and navigation resources.
-The sampling of temperature and salinity variables can be much higher than the grid resolution, but, due to computational and navigation resources
-available on the AUV, assimilation into a discretized fixed grid is necessary. For more complex
-oceanographic phenomena, the methods will need to be extended to
-non-Gaussian phenomena, possibly feature-based mixtures of Gaussian
-processes which could potentially be run onboard augmented by
-dynamical models. Running numerical models onboard a robotic vehicle
-is currently infeasible, but high-resolution ocean models or remote
-sensing data can be used to fit a more complex statistical model
-\cite{davidson19}.
+\citep{sigrist2015stochastic,richardson2017sparsity}. The sampling
+capabilities are restricted by the sensors onboard the AUV, the
+computational load, and navigation resources. The sampling of
+temperature and salinity variables can be much higher than the grid
+resolution, but, due to computational and navigation resources
+available on the AUV, assimilation into a discretized fixed grid is
+necessary. For more complex oceanographic phenomena, the methods will
+need to be extended to non-Gaussian phenomena, possibly feature-based
+mixtures of Gaussian processes which could potentially be run onboard
+augmented by dynamical models. Running numerical models onboard a
+robotic vehicle is currently infeasible, but high-resolution ocean
+models or remote sensing data can be used to fit a more complex
+statistical model \cite{davidson19}.
 
@@
 Operations and Systems (AMOS), Center of Excellence, project number
-223254, Nansen Legacy Program, project number 276730, and the Applied Underwater Robotics Labortatory (AURLab). CT
-and DG acknowledge support from the Swiss National Science Foundation,
-project number 178858. JE and KR acknowledge support from Norwegian
-research council (RCN), project number 305445. DG would like to
-aknowledge support of Idiap Research Institute, his primary
-affiliation in an early version of this manuscript.
-We thank the anonymous    reviewers for their careful reading of our manuscript and their many insightful comments and suggestions.
-The authors would also like to thank Niklas Linde of the University of Lausanne for providing constructive feedback about this work, and members of the NTNU AURLab for help with AUV deployments
+223254, Nansen Legacy Program, project number 276730, and the Applied
+Underwater Robotics Labortatory (AURLab). CT and DG acknowledge
+support from the Swiss National Science Foundation, project number
+178858. JE and KR acknowledge support from Norwegian research council
+(RCN), project number 305445. DG would like to aknowledge support of
+Idiap Research Institute, his primary affiliation in an early version
+of this manuscript. We thank the anonymous reviewers for their careful
+reading of our manuscript and their many insightful comments and
+suggestions. The authors would also like to thank Niklas Linde of the
+University of Lausanne for providing constructive feedback about this
+work, and members of the NTNU AURLab for help with AUV deployments.
 

### 2021-01-12 main_web.tex
Updates from Overleaf

@@
 is therefore to use AUV measurements for improved description of the
-interface between fresh and oceanic waters.  It is often not possible
-to sample the biological variables of fundamental interest in such AUV
-operations, but off-the-shelf instruments provide temperature and
+interface between fresh and oceanic waters, which in turn enables more direct sampling of biological dynamics. It is often not possible
+to sample these biological variables directly, but off-the-shelf instruments provide temperature and
 salinity measurements which serve as proxies for the underlying
@@
 plume. The goal of plume characterization, in this way, relates to
-that of estimating some regions of the domain, typically excursion
+estimating some regions of the domain, typically excursion
 sets (ESs), when implicitly defined by the vector-valued random field.
+The goal of plume characterization will in this setting relate to estimating an excursion set (ES), implicitly defined over the vector-valued random field.
 In our context of environmental sampling, the joint salinity and
@@
 rather on statistical principles and computation underlying efficient
-data collection. We combine novel possibilities in marine robotics
-with spatial statistics and experimental design to provide useful AUV
-sampling designs.
+data collection. We combine marine robotics
+with spatial statistics and experimental design to provide rewarding AUV
+sampling designs yielding scientific relevant data.
 
@@
 \citep{bueso1998state,zidek2019monitoring}.
-However, response variance and entropy being depending only in GRF models on measurement locations and not on response values, criteria based on them only tend to have limited flexibility for
-active adaptation of trajectories based on measurement values.  The use
+However, in GRF models response variance and entropy only depend on measurement locations and not on response values.
+Such criteria tend to have limited flexibility for
+active adaptation compared to trajectories based on measurement values. The use
 of data-driven adaptive criteria was introduced to include more
@@
 %
-Furthemore, boldface letters will be used to denote concatenated
+Furthermore, boldface letters will be used to denote concatenated
 quantities corresponding to batches of observations.  Given a dataset
@@
 $q_1,\dots, q_n$, and that one wishes to predict $\gp[\bm{x}]$ for
-some batch of $q\geq 1$ generalized locations
+a batch of $q\geq 1$ generalized locations
 $\bm{x}$, the simple co-Kriging mean then amounts to Kriging with respect to a
@@
 As noted in \citep{Chevalier2015} in the case of scalar-valued fields,
-these update formulae naturally extend to universal Kriging in
+the update formulae naturally extends to universal Kriging in
 second-order settings and apply without Gaussian assumptions. We will
@@
 consider any temporal effects, which would be relevant on a larger
-time scale, we do consider the extension to spatio-temporal modeling
-and envision that advection-diffusion equations could be useful
+time scale. We do consider the extension to spatio-temporal modeling
+and envision that advection-diffusion equations could be a point of departure
 \citep{sigrist2015stochastic,richardson2017sparsity}. The sampling

### 2021-01-15 ans_to_reviewers/ans-to-2nd-rev.tex
Updates from Overleaf

@@
 \vspace{5mm}
-\reply: We have elaborated on the specification of nugget in the revised version. We have addressed this point in \aecom \textbf{5}.
+\reply: We have elaborated on the specification of the nugget effect in the revised version.
+In the revised version we include a plot with analysis of survey sampling residuals at the end of the paper.
+This shows that the specified model is reasonable, but it is of course a simplification of the complex ocean.
+We believe this part fits as an honest discusion at the end of the results section.
+We have addressed this point further in \aecom \textbf{5}.
 
@@
 \reply: We agree that the remarks of Ehrenberg are very useful!
-(It should be curriculum for graduate students.) Regarding the structure, we seriously considered your suggestions about the background in supplementary material, but in the end thought it was too much of a reduction to go through with.
+Regarding the structure, we seriously considered your suggestions about the background in supplementary material, but in the end thought it was too much of a reduction to go through with.
 It would be very difficult to grasp the applied contribution of our paper if the background parts were in the supplemental material, and only some readers take the extra effort of reading that.
-Based on the report of the reviewer, we did however go through the background and cut a few unnecessary and lengthy formulations in the revised version.
-{\bf{to go through for typos, redundant text, etc}}
+Based on the report of the reviewer, we did however go through the background and cut a few unnecessary formulations in the revised version.
 
@@
 4, 5, and 6 have been made with a black-and-white compatible color-map, so they can be converted to black-and-white.
-{\bf{to check all}}
 
@@
 \vspace{5mm}
-\reply: We have removed the word sampling from the sentence, with made it more clear and readable. See the change below:
+\reply: We have removed the word sampling from the sentence. It now reads:
 
@@
 Even if the sensor specifications were essentially 0, it would be a stretch to say that a variance of 0.0625 was a “middle ground” between the empirical nugget effect and the sensor specifications.
-So either give a better justification of your choice of 0.0625 or, if that’snot possible, then redo the analysis with a more justifiable choice.
+So either give a better justification of your choice of 0.0625 or, if that’s not possible, then redo the analysis with a more justifiable choice.
 
@@
 \reply: We have elaborated on the specification of nugget in the revised version.
-The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, and in related data-sets we have not seen much of a nugget effect for temperature and salinity in water samples close in space and time.
-The initial sampling transect had no samples at very short distance, and it is hence not possible to see the nugget effect in the empirical variogram.
-We made an empirical variogram from the actual survey data, where we have some samples close in space, see Figure \ref{empVar}.
-These curves support our statements of a smaller nugget.
-\begin{figure}[H] 
+The initial sampling transect had no samples at very short distance, and these data are hence not reliable for extracting the nugget effect.
+The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, and in related data-sets.
+However, based on our experience \citep{fossuminformation}, there is more variability than this in temperature and salinity measurements in water samples close in space and time.
+This variability is partly due to dynamic water masses but also uncertainty in the AUV position.
+\begin{figure}[h] 
 \centering
@@
 
-One reason to keep a significant nugget effect here is to account for the dynamic water masses that are not captured in our model, but we expect this to be very small given our limited survey time, as was the case in \cite{fossuminformation}.
+In the revised manuscript, we first clarified the specification of model parameters, and we then made residual plots and empirical variograms for the actual survey data (Figure 11 in the revised paper).
+In the two surveys, we have some samples close in space and it is possible to extract the nugget effect.
+These curves support our statements of a small but non-zero nugget.
+
+%One reason to keep a significant nugget effect here is to account for the dynamic water masses that are not captured in our model, but we expect this to be very small given our limited survey time, as was the case in \citep{fossuminformation}.
 Then again, between the two surveys, there is a change in the survey 1 and 2 paths caused by slightly different measurements at the two times of surveying.
@@
 This was done by running synthetic cases of sampling paths and data up to a certain stage for different measurement noise variance levels.
-At the end, the posterior excursion probability and the chosen sampling path are compared, as show in Figure \ref{fig:nugget_comparison}.
-We see moderate changes in the path patterns
-and overall, paths tend to be smoother for large nugget effect.
+We see moderate changes in the sampling path patterns, but paths tend to be smoother for large nugget effect, as illustrated in Figure \ref{fig:nugget_comparison}.
 Indeed, with small nugget, the AUV reacts more to the observed temperature and salinity levels, and this leads to more small turns in its path. 
-We note that though the sampling paths can be different, the posterior excursion probabilities are similar and will thus lead to similar estimates for the excursion set.
-We also remark that the differences in Figure \ref{fig:nugget_comparison} would get smaller if we would let the sampling continue (here stopped for presentation purposes).
+We note that even though the sampling paths can be different, the posterior EPs in Figure \ref{fig:nugget_comparison} are similar and will thus lead to similar estimates for the excursion set.
+%We also remark that the differences in Figure \ref{fig:nugget_comparison} would get smaller if the sampling continues (here stopped for presentation purposes).
 
-\begin{figure}[H]
+\begin{figure}[h]
     \centering
     \includegraphics[scale=0.52]{ans_to_reviewers/merged_paths_2.png}
-    \caption{Comparison of sampling paths and posterior excursion probability for different values of the measurement noise standard deviation (from left to right: $0.1$, $0.25$ and $0.5$).
+    \caption{Comparison of sampling paths and posterior EPs for different values of the measurement noise standard deviation (from left to right: $0.1$ and $0.5$).
 The sampling has been stopped after 200 steps.}
@@
 \end{figure}
-
-\newpage
-
-In the revised manuscript we wrote another sentence on Fig 11: 'The AUV paths in both runs are relatively winding in the sense that the vehicle reacts much to the data.
-In our experience from simulated runs, the paths tend to be smoother when we use a larger nugget effect.
-Still, the estimated nugget from the actual trial data is about $0.1$ which is close to the specified level.'
+In the revised manuscript we wrote another sentence on Fig 11: 'The AUV paths in both runs are relatively
+winding in the sense that the vehicle reacts to the data. In our
+experience from simulations, the paths tend to be smoother when we
+use larger measurement error variances for the AUV temperature and salinity observations. '
 
@@
 \vspace{5mm}
-\reply: We have improved our {\bf{to do}}
+\reply: We have inserted this as suggested.
 

### 2021-01-15 main_web.tex
Updates from Overleaf

@@
   \hfill
-  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
+  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway (looking north).]{\includegraphics[width =
     0.49\textwidth]{Figures/web/river_proccess.png}\label{fig:nidelven}}
@@
 models based on complex differential equations cannot be run onboard
-the AUV with limited computational capacity, and statistical models
+the AUV with limited computational capacity. Instead, statistical models
 relying on random field assumptions are relevant as a means to
@@
 salinity measurements which serve as proxies for the underlying
-biological phenomenon. With the help of a vector-valued random field
+biological phenomena. With the help of a vector-valued random field
 model for temperature and salinity, one can then aim to describe the
-plume. The goal of plume characterization, in this way, relates to
-estimating some regions of the domain, typically excursion
-sets (ESs), when implicitly defined by the vector-valued random field.
-The goal of plume characterization will in this setting relate to estimating an excursion set (ES), implicitly defined over the vector-valued random field.
+plume. The goal of plume characterization will in this setting relate to that of estimating an excursion set (ES), which is implicitly defined over the vector-valued random field. 
 In our context of environmental sampling, the joint salinity and
@@
 such as global optimization and set estimation. Of particular
-relevance to our context, \cite{Bect.etal2012} focuses on strategies
+relevance to our context, \cite{Bect.etal2012} focus on strategies
 to reduce uncertainties related to volumes of excursion exceeding a
-prescribed threshold, while \cite{chevalier2014fast} concentrates on
+prescribed threshold, while \cite{chevalier2014fast} concentrate on
 making the latter strategies computationally efficient and
@@
 designs of experiments for the conservative estimation of ESs based on
-GRF models were presented in \citep{Azzimonti.etal}.
+GRF models were presented in \cite{Azzimonti.etal}.
 
@@
 sequential strategies for estimating features of vector-valued random
-fields are still in their infancy. \cite{LeGratiet.etal2015} used
+fields are still in their infancy. \cite{LeGratiet.etal2015} use
 co-Kriging based sequential designs to multi-fidelity computer codes
-and \cite{Poloczek2017} used related ideas for multi-information
-source optimization, but not for ES's like we do here. More relevant
+and \cite{Poloczek2017} use related ideas for multi-information
+source optimization, but not for ESs like we do here. More relevant
 to our setting, \citet[p.82]{stroh} mentions general possibilities of
-step-wise uncertainty reduction strategies for ES's in the context of
+step-wise uncertainty reduction strategies for ESs in the context of
 designing fire simulations, yet outputs are mainly assumed
@@
 \cite{sahu2008space} studying spatio-temporal models for sea surface
-temperature and salinity data and \cite{mellucci2018oceanic} looking
+temperature and salinity data, and \cite{mellucci2018oceanic} looking
 at the statistical prediction of features using an underwater glider.
 In this work the main focus is not on statistical modeling per se, but
-rather on statistical principles and computation underlying efficient
+rather on statistical principles and computations underlying efficient
 data collection. We combine marine robotics
@@
 However, in GRF models response variance and entropy only depend on measurement locations and not on response values.
-Such criteria tend to have limited flexibility for
-active adaptation compared to trajectories based on measurement values. The use
+This means that there is limited flexibility for
+active adaptation of trajectories based on measurement values. The use
 of data-driven adaptive criteria was introduced to include more
@@
 considered), and $\no=2$ with responses of temperature and salinity. A
-bivariate GRF model is assumed for $\gp$. To motivate concepts,
-Fig.~\ref{fig:real_temp} and \ref{fig:real_sal} shows a realization of such a
+multivariate GRF model is assumed for $\gp$. To motivate concepts,
+Fig.~\ref{fig:real_temp} and \ref{fig:real_sal} show a realization of such a
 vector-valued GRF on $\domain=[0,1]^2$. Fig.~\ref{fig:jointex_roi} represents a by-product of interest derived from these
@@
 height=0.27\textwidth,keepaspectratio]{Figures/intro_excu.png}\label{fig:jointex_roi}}
-\caption{Realization of a bivariate GRF (display (a) and (b)) and excursion set above some threshold (c). Joint excursion in red and excursion of a single variable in light-red.}
+\caption{Realization of a bivariate GRF (display (a) and (b)) and ES above some threshold (c). Joint excursion in red and excursion of a single variable in light-red.}
 \label{example_excu}
@@
 If we assume that $\gp$ has continuous trajectories 
-and $T$ is closed, then $\es$ becomes a Random Closed Set
+and $T$ is closed, then $\es$ becomes a random closed set
 \citep{Molchanov2005} and concepts from the theory of random sets will
@@
 of several key quantities. Note that changing some $\leq$ inequalities
-to $\geq$ ones would lead to immediate adaptations.
+to $\geq$ would lead to immediate adaptations.
 
@@
 some unobserved location $\spatloc\in \domain$ by using the
-conditional mean of $\gp[\spatloc]$, conditional on the data. This
+conditional mean of $\gp[\spatloc]$, given the data. This
 coincides with co-Kriging equations, which tell us precisely how to
@@
 the batch-sequential Kriging update formulae from
-\citep{Chevalier.etal2013a} suggests that
+\cite{Chevalier.etal2013a} suggest that
 % 
@@
 to the Kriging weights for the last $q_{n+1}$ responses
-when predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
+when predicting at $\bm{x}$ relying on all measurements until batch $n+1$.
 The associated co-Kriging residual (cross-)covariance function is
@@
 %
-As noted in \citep{Chevalier2015} in the case of scalar-valued fields,
+As noted in \cite{Chevalier2015} in the case of scalar-valued fields,
 the update formulae naturally extends to universal Kriging in
@@
 
-\subsection{Uncertainty Quantification on ESs of multivariate GRFs}
+\subsection{UQ on ESs of multivariate GRFs}
 \label{sec:set_uq}
@@
 \end{equation*}
-which in the excursion/sojourn case where $\T=(-\infty, t_1] \times
+which in the excursion/sojourn case with $\T=(-\infty, t_1] \times
 \dots \times (-\infty, t_{\no}]$ is
@@
 let $\currentExp{.}$ and $\currentProba{.}$ denote conditional
-expectation and probability conditional on the first $n$ batches of
-observations, respectively. We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.
+expectation and probability given the first $n$ batches of observations. We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.
 
@@
 $Z_{\bm{x}}$ and with independent noise having covariance matrix
-$\Delta_n$.
+$\Delta_{n+1}$.
 
@@
 different designs on a bivariate example. This mimics our river plume
-application and hence the first and second component of the random
+application and hence the components of the random
 field will be called \textit{temperature} and \textit{salinity}. We
@@
 
-Say we want to study the excursion probability of a bivariate Gaussian
-above some threshold, where the thresholds are set equal to the mean;
+Say we want to study the EP of a bivariate Gaussian
+vector variable below some thresholds. The thresholds are set equal to the mean;
 $\mu_1=t_1=5^o C$ for temperature and $\mu_2=t_2=30$ g/kg for
-salinity, and we play with the temperature and salinity correlation
+salinity. We play with the temperature and salinity correlation
 and variances to study the effect on the EP and EBV.
@@
 salinity. The displayed densities have unit standard deviations for
-both temperature and salinity, but we also study the effect of
+both temperature and salinity, but we will also study the effect of
 doubling the standard deviations.
@@
 
-\subsubsection{Including Spatiality}
+\subsubsection{Going Spatial}
 \label{sec:including_spatiality}
@@
 and kernel parameter $\eta=3.5$.
-One realization of this GRF is shown in Fig.~\ref{example_excu}.
-In the computed examples, the spatial domain $\domain$ is
+One realization of this GRF is shown in Fig.~\ref{example_excu}. Thresholds are $t_1=2.3$ and $t_2=22$.
+In these computer examples, the spatial domain $\domain$ is
 discretized to a set of $N$ grid locations
@@
 
-We now study how the EBV [Eq.\eqref{def:eibv}] associated with data
+We now study how the EBV associated with data
 collection at a point changes if only one of the two components of the
@@
   collection locations in green. True excursion set in red. Places
-  where only one response is above threshold are depicted in pink. EBV
+  where only one response is above threshold are depicted in light-red. EBV
   reduction associated to observing one or both responses at the green
@@
 model is updated using co-Kriging
-Eq. \ref{eq:meanCoK} and \ref{eq:varCoK}, yielding a conditional law $\mathbb{P}_{\stage + 1}$
+Eq. \ref{eq:meanCoK} and \ref{eq:varCoK}, yielding a conditional law $\mathbb{P}_{[\stage+1]}$
 after which the process is repeated.
 
-Note that the type of data collected at each stage can be of various
-type (all components of the field at a single location, only some
+Note that the data collected at each stage can be of various
+types (all components of the field at a single location, only some
 components at a subset of selected locations, etc.) because of the
@@
 AUV exploring the river plume, we limit our scope to choosing one of
-the neighboring spatial location (waypoints) at each stage, and all
+the neighboring spatial locations (waypoints) at each stage, and all
 components (temperature and salinity) of the field are observed
@@
 setting involves a series of intermixed maximizations over designs and
-integrals over data. In practice, the optimal solution is intractable
+expected values over data outcomes. In practice, the optimal solution is intractable
 because of the enormous growth over stages (see
@@
 \begin{equation}\label{critSEQ}
-     C_{\text{myopic}}(u) = \EIBV_{\stage}\left(\spatloc\right)
+     C_{\text{myopic}}(u) = \EIBV_{[\stage]}\left(\spatloc\right)
 \end{equation}
@@
       \begin{align}\label{critLA}
-          C_{\text{2-steps}}(u) &= \mathbb{E}_{\stage}\left[\min_{\spatloc' \in
-                  \candidates(\spatloc)} \EIBV_{\stage}\left(\spatloc' ; \spatloc,
+          C_{\text{2-steps}}(u) &= \mathbb{E}_{[\stage]}\left[\min_{\spatloc' \in
+                  \candidates(\spatloc)} \EIBV_{[\stage]}\left(\spatloc' ; \spatloc,
       Y\right)\right]
@@
 
-Each strategy is conducted on an equilateral grid as shown in
+Each strategy is conducted on an equilateral triangular grid as shown in
 Fig.~\ref{fig:wp_graph}. The AUV starts at the center coordinate at
@@
 To demonstrate the applicability of using multivariate EPs and the IBV
-to inform oceanographic sampling, we present a case study mapping a
-river plume with an AUV. The experiment was performed in Trondheim,
-Norway, surveying the Nidelva river (Fig.~\ref{fig:nidelven}). The
-experiments were conducted in late Spring 2019, when there is still
-snow melting in the surrounding mountains so that the river water is colder than the water in the fjord. The experiment was
-focused along the frontal zone that runs more or less parallel to the
+to inform oceanographic sampling, we present a case study of using an AUV to map the Nidelva river plume in Trondheim, Norway (Fig.~\ref{fig:nidelven}). The
+experiments were conducted in late Spring 2019, when there was still
+snow melting in the surrounding mountains so that the river water was colder than the water in the fjord. The experiment was
+focused along the frontal zone that ran more or less parallel to the
 eastern shore.
@@
 Both temperature and salinity were assumed to increase linearly with the west coordinate.
-Next, the residuals from the regression analysis were analyzed to specify the covariance parameters of the GRF model.
+Next, empirical correlation and variograms of the residuals from the regression analysis were analyzed to specify the temperature and salinity correlation, variances and spatial correlation decay parameter of the GRF model.
 The resulting
@@
 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4
-GB of 1600 MHz DRAM) for computation onboard.  The sampling algorithm
+GB of 1600 MHz DRAM) for computations onboard.  The sampling algorithm
 was built on top of the autonomous Teleo-Reactive EXecutive
@@
 from the temperature sensor. We specify variance $0.25^2$ for both
-errors. This is based on prior experience indicating that there can be
-some noise in temperature and salinity readings mostly because of
-dynamic water masses and uncertainty in AUV position. The initial
-survey transect was not sampled very densely so a reliable extraction
-of a nugget parameter from those data was not possible.
+errors. These are reasonable error levels based on prior experience \citep{fossuminformation} indicating that there can be
+some noise in temperature and salinity data because of
+dynamic water masses and uncertainty in the AUV position. The initial
+survey transect was not sampled very densely, and a reliable extraction
+of nugget effects that could be representative of the measurement noise variance parameters was not possible.
 
@@
 sampling locations on the waypoint graph distributed over an
-equilateral grid, as shown in the grey-colored lattice in
-Fig.~\ref{fig:map}.  At each stage, it takes the AUV about 30 seconds
+equilateral triangular grid, as shown in the grey-colored lattice in
+Fig.~\ref{fig:map}.  At each stage, it took the AUV about 30 seconds
 to assimilate data and evaluate the EIBV for all the possible waypoint alternatives.  It
 was set to start in the south-center part of the waypoint graph. A
-survey was set to take approximately 40 minutes, visiting 15 waypoints
+survey took approximately 40 minutes, visiting 15 waypoints
 on the grid, with the vehicle running near the surface to capture the
-plume. On its path from one waypoint to the next, the AUV collects
+plume. On its path from one waypoint to the next, the AUV collected
 data with an update frequency of 30 seconds, giving three measurements
@@
 shown in the map in Fig.~\ref{fig:map}, both within the expected
-frontal region (shaded pink). The recorded temperatures are shown as
+frontal region (shaded light-red). The recorded temperatures are shown as
 colored trails in Fig.~\ref{fig:res_both}, clearly indicating the
@@
 winding in the sense that the vehicle reacts to the data. In our
-experience from simulated runs, the paths tend to be smoother when we
-use a larger nugget effect. But here the estimated nugget from the
-actual trial data is close to the specified level.
+experience from simulations, the paths tend to be smoother when we
+use larger measurement error variances for the AUV temperature and salinity observations. 
+%But here the estimated nugget from the actual trial data is close to the specified level.
 
@@
   \hfill
-  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_bN.eps}\label{fig:parest_b}}
+  \subfigure[Quadratic form CDFs.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_bN.eps}\label{fig:parest_b}}
   \hfill
@@
 joint residuals indicates that the correlation between the two
-variables might be slightly larger than the initially specified $0.5$
-and that uncertainty in temperature was set a bit too high. Based on
+variables might be slightly larger than the initially specified $0.5$ correlation,
+and that initial uncertainty in temperature was set a bit too high. Based on
 the fitted bivariate covariance model (ellipses in
@@
 the specified Gaussian model fits reasonably
-well. Fig.~\ref{fig:parest_c} shows the empirical variogram of the
+well. Fig.~\ref{fig:parest_c} shows the empirical variograms of the
 scaled residuals for temperature and salinity. The decay is similar
@@
 specified. The working assumption of a separable covariance function
-is hence not unreasonable.
-
+is hence not unreasonable. There appears to be a slight nugget effect, but not much different from what was set as the measurement noise variance.
 Overall, the spatial Gaussian model representation is a simplification
-of the changing water conditions, and even within the model
-formulation, the parameter specification could likely have been
-improved.  However, we do not expect minor changes to have caused much
-difference in sampling paths.
+of the changing water conditions, and even within our working assumptions,
+the parameter specification could likely have been
+improved. However, we do not expect that minor changes in parameters would have caused much
+difference in AUV sampling paths.
 
@@
 models or remote sensing data can be used to fit a more complex
-statistical model \cite{davidson19}.
+statistical model \citep{davidson19}.
 
@@
 settings related to mapping of algal-blooms, anoxic zones or open
-water fronts \cite{costa19}.  Other criteria could also be relevant,
+water fronts \citep{costa19}.  Other criteria could also be relevant,
 for instance, hybrid or multi-attribute criteria that could balance

### 2021-01-19 ans_to_reviewers/ans-to-2nd-rev.tex
Updates from Overleaf

@@
 \noindent {\bf{1.
-Sampling limitations:}} You explained in your response file (R1S3.pdf) the limitations of the sampling design (comment #4 from the AE's previous review).
+Sampling limitations:}} You explained in your response file (R1S3.pdf) the limitations of the sampling design (comment 4 from the AE's previous review).
 I agree with the AE (bullet 1, R2AE2.pdf) that some (perhaps slightly abbreviated) version of your response on p4 (middle) would be important to include; viz.
@@
 different line types likely will suffice in Figures 8 and 9; and
-probably nothing will be lost if Figures 7 & 10 were black-and-white.
+probably nothing will be lost if Figures 7 and 10 were black-and-white.
 Color is very expensive to print, so we appreciate your attention
@@
 So we ask to consider that this figure is kept in color.
+Revised Fig.
+3 and 11 are in black-and-white.
 Fig.

### 2021-01-19 main_web.tex
Updates from Overleaf

@@
 \centering
-\includegraphics[width=0.99\textwidth]{Figures/illus_bivar-eps-converted-to.pdf}%illus_bivar.epse}
+\includegraphics[width=0.99\textwidth]{Figures/illus_bivarNEW.eps}%illus_bivar.epse}
 \caption{Density contour plots with increasing correlations between
@@
   \centering
-  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_aN.eps}\label{fig:parest_a}}
+  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/res_diag_aNEW.eps}\label{fig:parest_a}}
   \hfill
-  \subfigure[Quadratic form CDFs.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_bN.eps}\label{fig:parest_b}}
+  \subfigure[Quadratic form CDFs.]{\includegraphics[width = 0.32\textwidth]{Figures/res_diag_bNEW.eps}\label{fig:parest_b}}
   \hfill
-  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_cN.eps}\label{fig:parest_c}}
-  \caption{Data analysis from from AUV experiments. \ref{fig:parest_a} Residual plot of temperature and salinity
+  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/res_diag_cNEW.eps}\label{fig:parest_c}}
+  \caption{Data analysis from AUV experiments. \ref{fig:parest_a} Residual plot of temperature and salinity
     along with specified Gaussian contours. \ref{fig:parest_b} Empirical CDF
@@
     modeled theoretical CDF (dashed) of the $\chi^2$ distribution with two
-    degrees of freedom. \ref{fig:parest_c} Scaled empirical variograms of the
+    degrees of freedom. \ref{fig:parest_c} Scaled empirical variograms of the survey
     salinity and temperature data.}

### 2021-01-22 ans_to_reviewers/ans-to-2nd-rev.tex
Updates from Overleaf

@@
 
-We are grateful to the editors and reviewers for their comments and suggestions in this 2$^{nd}$ review.
+We are grateful to the editors and reviewers for their comments and suggestions.
 They have all been considered in this revised version of our manuscript.
@@
 
-\reply: We agree and have addressed this point in \aecom \textbf{1}.
+\reply: We agree, and we have addressed this point as you suggest. For more details, see \aecom \textbf{1}.
 
@@
 \vspace{5mm}
-\reply: We have elaborated on the specification of the nugget effect in the revised version.
-In the revised version we include a plot with analysis of survey sampling residuals at the end of the paper.
-This shows that the specified model is reasonable, but it is of course a simplification of the complex ocean.
-We believe this part fits as an honest discusion at the end of the results section.
+\reply: We have revised the text specification of model parameters.
+In the revised version we have a plot with analysis of survey sampling residuals at the end of the paper.
+This shows that the specified nugget effect and other parameters are reasonable, but it is of course a simplification of the complex ocean.
+We believe this part fits as an honest discussion at the end of the results section.
 We have addressed this point further in \aecom \textbf{5}.
@@
 \vspace{5mm}
-\reply: We agree on all our points here, except Fig.8.
+\reply: We agree on all your points here, except Fig.8.
 Line type will be tricky for Fig.
@@
 \vspace{5mm}
-\reply: We have elaborated on the specification of nugget in the revised version.
-The initial sampling transect had no samples at very short distance, and these data are hence not reliable for extracting the nugget effect.
-The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, and in related data-sets.
-However, based on our experience \citep{fossuminformation}, there is more variability than this in temperature and salinity measurements in water samples close in space and time.
+\reply: We have revised the text specification of model parameters in Sect.
+4.1 and 4.2.
+The initial survey transect had no samples at very short distance, and these data are hence not reliable for extracting the nugget effect.
+The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, but based on our experience \citep{fossuminformation}, there is more variability than this in temperature and salinity measurements in water samples close in space and time.
 This variability is partly due to dynamic water masses but also uncertainty in the AUV position.
@@
 
-In the revised manuscript, we first clarified the specification of model parameters, and we then made residual plots and empirical variograms for the actual survey data (Figure 11 in the revised paper).
+In the revised manuscript, we made residual plots and empirical variograms for the actual survey data in Sect 4.3 (Figure 11 in the revised paper).
 In the two surveys, we have some samples close in space and it is possible to extract the nugget effect.
-These curves support our statements of a small but non-zero nugget.
+These curves support our statements of a small but non-zero nugget (see Figure \ref{empVar}).
 
@@
 We see moderate changes in the sampling path patterns, but paths tend to be smoother for large nugget effect, as illustrated in Figure \ref{fig:nugget_comparison}.
-Indeed, with small nugget, the AUV reacts more to the observed temperature and salinity levels, and this leads to more small turns in its path. 
-We note that even though the sampling paths can be different, the posterior EPs in Figure \ref{fig:nugget_comparison} are similar and will thus lead to similar estimates for the excursion set.
+Indeed, with small nugget (left display), the AUV reacts more to the observed temperature and salinity levels, and this leads to more small turns in its path.
+
+We note that even though the sampling paths can be different, the posterior EPs in Figure \ref{fig:nugget_comparison} are similar and will thus lead to similar estimates for the ES.
 %We also remark that the differences in Figure \ref{fig:nugget_comparison} would get smaller if the sampling continues (here stopped for presentation purposes).
@@
 \end{figure}
-In the revised manuscript we wrote another sentence on Fig 11: 'The AUV paths in both runs are relatively
+In the revised manuscript we wrote two sentences around this sensitivity near Fig 10: 'The AUV paths in both runs are relatively
 winding in the sense that the vehicle reacts to the data. In our

### 2021-01-22 main_web.tex
Updates from Overleaf

@@
 
-\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
+\address{Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
 \printead{e1}}
@@
   \centering
-  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/res_diag_aNEW.eps}\label{fig:parest_a}}
+  \subfigure[Residual plot.]{\includegraphics[height = 0.26\textwidth]{Figures/res_diag_aNEW.eps}\label{fig:parest_a}}
   \hfill
-  \subfigure[Quadratic form CDFs.]{\includegraphics[width = 0.32\textwidth]{Figures/res_diag_bNEW.eps}\label{fig:parest_b}}
+  \subfigure[Quadratic form CDFs.]{\includegraphics[height = 0.26\textwidth]{Figures/res_diag_bNEW.eps}\label{fig:parest_b}}
   \hfill
-  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/res_diag_cNEW.eps}\label{fig:parest_c}}
+  \subfigure[Empirical variogram.]{\includegraphics[height = 0.252\textwidth]{Figures/res_diag_cNEW.eps}\label{fig:parest_c}}
   \caption{Data analysis from AUV experiments. \ref{fig:parest_a} Residual plot of temperature and salinity
@@
 
-Fig.~\ref{fig:parest} summarizes diagnostic plots of this
+Fig.~\ref{fig:parest} summarizes with diagnostic plots of this
 analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature

### 2021-03-08 main_web.tex
Updates from Overleaf

@@
-\documentclass[aoas,preprint]{imsart}
+\documentclass[aoas]{imsart}
 \setattribute{journal}{name}{}
@@
 $\mes(\es)$ through its moments.  Centered moments may be computed
-using Proposition~\ref{propo1} developed in the appendix.  In
+using Proposition~\ref{propo1} developed in the Supplementary Material \ref{suppA}.  In
 particular, as an integral over EPs, the 
@@
 178858. JE and KR acknowledge support from Norwegian research council
-(RCN), project number 305445. DG would like to aknowledge support of
+(RCN), project number 305445. DG would like to acknowledge support of
 Idiap Research Institute, his primary affiliation in an early version
@@
 
+\begin{supplement}
+%\sname{Supplement A}
+\stitle{Proof for proposition 1 \& 2}
+\slink[doi]{COMPLETED BY THE TYPESETTER}
+\sdatatype{.pdf}
+\sdescription{Development of proofs and nugget sensitivity.}
+\end{supplement}
+
 \footnotesize

### 2021-03-08 supplementary_material.tex
Updates from Overleaf

--- 2020-06-29 appendix.tex
@@
+\documentclass[aoas]{imsart}
+\setattribute{journal}{name}{}
+
+\input{commands.tex}
+\usepackage{amsmath}
+\usepackage{pstricks,pst-grad}
+\usepackage{graphicx}
+\usepackage{floatrow}
+\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
+\floatsetup[table]{capposition=top}
+\usepackage{subfigure}
+\usepackage[utf8]{inputenc}
+\usepackage{booktabs}                     % horizontal lines in tables
+\usepackage{comment}
+
+\usepackage{textcomp}
+\usepackage{amsthm,amsmath,natbib}
+%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+\usepackage{color, colortbl}
+\definecolor{Gray}{gray}{0.9}
+
+\usepackage{amssymb}
+\usepackage{hyperref}
+\usepackage{cleveref}
+\usepackage{xcolor}
+\input{macros}
+
+\begin{document}
+
+\begin{frontmatter}
+
+\title{Supplementary Materials for Learning Excursion Sets of Vector-valued Gaussian Random Fields for
+Autonomous Ocean Sampling}
+
+\begin{aug}
+\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
+\author{\fnms{Cédric} \snm{Travelletti}\thanksref{t3}, \corref{} \ead[label=e2]{cedric.travelletti@stat.unibe.ch}}
+\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e3]{jo.eidsvik@ntnu.no}}
+\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e4]{david.ginsbourger@stat.unibe.ch}}
+\and
+\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5,t6}. \ead[label=e5]{kanna.rajan@fe.up.pt}}
+
+\affiliation[t1]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
+\affiliation[t2]{Centre for Autonomous Marine Operations and Systems, NTNU.}
+\affiliation[t3]{Institute of Mathematical Statistics and Actuarial Science, University of Bern, Switzerland.}
+\affiliation[t4]{Department of Mathematical Sciences, NTNU.}
+\affiliation[t5]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}
+\affiliation[t6]{SIFT LLC, Minneapolis, Minnesota, United States.}
+
+\address{Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
+\printead{e1}}
+\address{Cédric Travelletti\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
+Switzerland.
+\printead{e2}}
+\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e3}}
+\address{David Ginsbourger\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
+Switzerland.
+\printead{e4}}
+\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
+  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
+\printead{e5}}
+
+\runauthor{TO. Fossum et al.}
+\end{aug}
+\end{frontmatter}
+
+\section{Derivation of proofs 1 \& 2}
+
 \begin{propo}
@@
 %Note that we will use boldface to denote concatenated variables. % TO RELOCATE
-Here $\gp$ is defined on $D$, and for
-$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in D^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
+Here $\gp$ is defined on $\domain$, and for
+$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in \domain^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
 \gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.
@@
 
-\begin{proof}{(Proposition \ref{propo_emv})}
+\begin{proof}{(Proposition 2)}
 \begin{equation*}
@@
 \end{equation*}
-and the proof follows by applying Proposition \ref{propo3}
+and the proof follows by applying Proposition 5
 with
@@
 \end{proof}
+\end{document}

### 2021-03-09 supplementary_material.tex
Updates from Overleaf

@@
 
-\section{Derivation of proofs 1 \& 2}
+\textbf{The PDF file includes:} Supplementary materials to the manuscript ``Learning Excursion Sets of Vector-valued Gaussian Random Fields for
+Autonomous Ocean Sampling" by Fossum et. al 2021.
 
+\section{Derivation of proofs}
+\setcounter{propo}{2}
 \begin{propo}
@@
 
-Those two general results allow us to derive simple expressions for the expected effect of the inclusion of new datapoints on the $\IBV$ (Proposition \ref{propo_eibv}) and on the $\EMV$ (Proposition \ref{propo_emv}) for which we provide proofs below.
+Those two general results allow us to derive simple expressions for the expected effect of the inclusion of new datapoints on the $\IBV$ (Proposition 1) and on the $\EMV$ (Proposition 2) for which we provide proofs below.
 
-\begin{proof}{(Proposition \ref{propo_eibv})}
+\begin{proof}{(Proposition 1)}
 Applying Tonelli-Fubini followed by the law of total
@@
 \end{equation*}
-and the proof follows by applying Proposition 5
+and the proof follows by applying Proposition \ref{propo3}
 with
@@
 \end{proof}
+
+\section{Discussion on nugget sensitivity}
+
+The initial survey transect had no samples at very short distance, and these data are hence not reliable for extracting the nugget effect.
+The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, but based on our experience \citep{fossuminformation}, there is more variability than this in temperature and salinity measurements in water samples close in space and time.
+This variability is partly due to dynamic water masses but also uncertainty in the AUV position.
+\begin{figure}[h] 
+\centering
+\includegraphics[width=0.6\textwidth]{Figures/field-trials/empVAR.png}
+\caption{Empirical variogram of the scaled variables from the two runs. }
+\label{empVar}
+\end{figure}
+
+In the revised manuscript, we made residual plots and empirical variograms for the actual survey data in Sect 4.3 (Figure 11 in the revised paper).
+In the two surveys, we have some samples close in space and it is possible to extract the nugget effect.
+These curves support our statements of a small but non-zero nugget (see Figure \ref{empVar}).
+
+We further checked the sensitivity of the nugget specification in the sequential sampling algorithm.
+This was done by running synthetic cases of sampling paths and data up to a certain stage for different measurement noise variance levels.
+We see moderate changes in the sampling path patterns, but paths tend to be smoother for large nugget effect, as illustrated in Figure \ref{fig:nugget_comparison}.
+Indeed, with small nugget (left display), the AUV reacts more to the observed temperature and salinity levels, and this leads to more small turns in its path.
+
+We note that even though the sampling paths can be different, the posterior EPs in Figure \ref{fig:nugget_comparison} are similar and will thus lead to similar estimates for the ES.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.41]{ans_to_reviewers/merged_paths_2.png}
+    \caption{Comparison of sampling paths and posterior EPs for different values of the measurement noise standard deviation (from left to right: $0.1$ and $0.5$).
+The sampling has been stopped after 200 steps.}
+    \label{fig:nugget_comparison}
+\end{figure}
+
+The AUV paths in both runs are relatively
+winding in the sense that the vehicle reacts to the data. In our
+experience from simulations, the paths tend to be smoother when we
+use larger measurement error variances for the AUV temperature and salinity observations.
+
+\footnotesize
+\bibliographystyle{imsart-nameyear}
+\bibliography{ref}
+
 \end{document}

### 2021-03-10 supplementary_material.tex
Updates from Overleaf

@@
 
-The initial survey transect had no samples at very short distance, and these data are hence not reliable for extracting the nugget effect.
+In the field experiments, the initial survey transect had no samples at very short distance, and these data are hence not reliable for extracting the nugget effect.
 The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, but based on our experience \citep{fossuminformation}, there is more variability than this in temperature and salinity measurements in water samples close in space and time.
@@
 
-In the revised manuscript, we made residual plots and empirical variograms for the actual survey data in Sect 4.3 (Figure 11 in the revised paper).
+We made residual plots and empirical variograms for the actual survey data (Figure 11 in the main body of the paper).
 In the two surveys, we have some samples close in space and it is possible to extract the nugget effect.
@@
 We see moderate changes in the sampling path patterns, but paths tend to be smoother for large nugget effect, as illustrated in Figure \ref{fig:nugget_comparison}.
-Indeed, with small nugget (left display), the AUV reacts more to the observed temperature and salinity levels, and this leads to more small turns in its path.
+Indeed, with small nugget (left display), the AUV reacts more to the observed temperature and salinity levels, and this leads to a more winding path with many small turns.
 
@@
 
-The AUV paths in both runs are relatively
-winding in the sense that the vehicle reacts to the data. In our
-experience from simulations, the paths tend to be smoother when we
-use larger measurement error variances for the AUV temperature and salinity observations.
-
 \footnotesize

### 2021-03-10 main_web.tex
Updates from Overleaf

@@
   limited resources in understanding processes in the water-column,
-  the combination of statistics and autonomous systems provide new
+  the combination of statistics and autonomous systems provides new
   opportunities for experimental design.
@@
 Motivated by the challenges related to efficient data collection
-strategies for our vast oceans, we combine spatial statistics, design
-of experiments and marine robotics in this work.  The
+strategies for our vast oceans, we combine spatial statistics, 
+experiment design, and marine robotics in this work.  The
 multidisciplinary efforts enable information-driven data collection in
-regions of high-interest.
+regions of high interest.
 
@@
 of the changing climate and increasing anthropogenic impact. Central
-to understanding the changes taking place in the upper water-column is
-knowledge of the bio-geophysical interaction driven by an
+to understanding the changes taking place in the upper water column is
+knowledge of bio-geophysical interactions driven by an
 agglomeration of physical forcings (e.g. wind, topography, bathymetry,
-tidal influences, etc.) and incipient micro-biology driven by
+tidal influences) and incipient micro-biology driven by
 planktonic and coastal anthropogenic input, such as pollution and
@@
 plumes, with direct and indirect effects on society \citep{ryan2017}. One of the
-bottlenecks in the study of such phenomena lies however in the lack of
+bottlenecks in the study of such phenomena, however, lies in the lack of
 observational data with sufficient resolution. Most of this
@@
     0.49\textwidth]{Figures/web/river_proccess.png}\label{fig:nidelven}}
-  \caption{\ref{fig:envir1} Traditional ocean observation based on 
+  \caption{\ref{fig:envir1} Traditional ocean observation using 
     ship-based sampling has been augmented by autonomous
@@
     \ref{fig:nidelven} The interaction of river and ocean creates
-    processes that are challenging to map, where the combination of
+    processes that are challenging to map, yet the combination of
     statistics and robotics can play a vital role in enabling more
@@
 discretization of the search domain where the AUV can navigate;
-however the AUV can now modify its path at each waypoint based on
+however, the AUV can now modify its path at each waypoint based on
 in-situ measurements and calculations onboard using onboard
@@
 survey area in Trondheim, Norway, where cold freshwater enters the fjord from a river, creating a strong gradient in both temperature and
-salinity. Because of the local topography and the Coriolis force the
-cold fresh water tends to flow east. Depending on the variations in
+salinity. Because of the local topography and the Coriolis force, the
+cold freshwater tends to flow east. Depending on the variations in
 river discharge, tidal effects, coastal current and wind, this
@@
 plume. The goal of plume characterization will in this setting relate to that of estimating an excursion set (ES), which is implicitly defined over the vector-valued random field. 
-In our context of environmental sampling, the joint salinity and
+In the context of environmental sampling, the joint salinity and
 temperature excursions of a river plume help characterize the
@@
 
-The questions tackled here hence pertain to the broader area of
+The questions tackled here pertain to the broader area of
 spatial data collection for vector-valued random fields.
@@
 experiments based on GRFs before detailing other literature related to
-AUV sampling and our contributions prior to outlining the rest of the
-paper.
+AUV sampling and our contributions prior to describing the results of this experiment.
 
@@
 
-While random field modeling has been one of the main topics throughout
-the history of spatial statistics \citep{Krige1951a,Stein1999}, even
+While random field modeling has been one of the main topics explored throughout
+the history of spatial statistics \citep{Krige1951a,Stein1999} — even
 for vector-valued random field models with associated prediction
-approaches such as co-Kriging \citep[See, e.g.,][]{Wackernagel2003},
+approaches such as co-Kriging \citep[See, e.g.,][]{Wackernagel2003} —
 there has lately been a renewed interest for random field models in
@@
 \cite{Ginsbourger2018}, GRF models have been used in particular as a
-basis to sequential design of simulations dedicated to various goals
+basis for the sequential design of simulations dedicated to various goals
 such as global optimization and set estimation. Of particular
@@
 \cite{French.Sain2013,Chevalier.etal2013b,Bolin.Lindgren2015,Azzimonti.etal2016}
-with ambitions of estimating sets themselves. Recently, sequential
+with ambitions to estimate sets themselves. Recently, sequential
 designs of experiments for the conservative estimation of ESs based on
@@
 fields are still in their infancy. \cite{LeGratiet.etal2015} use
-co-Kriging based sequential designs to multi-fidelity computer codes
+co-Kriging-based sequential designs to multi-fidelity computer codes
 and \cite{Poloczek2017} use related ideas for multi-information
@@
 Other statistical work in the oceanographic domain include
-\cite{wikle2013modern} focusing on hierarchical statistical models,
-\cite{sahu2008space} studying spatio-temporal models for sea surface
-temperature and salinity data, and \cite{mellucci2018oceanic} looking
+\cite{wikle2013modern}, focusing on hierarchical statistical models,
+\cite{sahu2008space}, studying spatio-temporal models for sea surface
+temperature and salinity data, and \cite{mellucci2018oceanic}, looking
 at the statistical prediction of features using an underwater glider.
-In this work the main focus is not on statistical modeling per se, but
+In this work, the main focus is not on statistical modeling per se, but
 rather on statistical principles and computations underlying efficient
@@
 with spatial statistics and experimental design to provide rewarding AUV
-sampling designs yielding scientific relevant data.
+sampling designs yielding  relevant scientific data.
 
@@
 \citep{bueso1998state,zidek2019monitoring}.
-However, in GRF models response variance and entropy only depend on measurement locations and not on response values.
+However, in GRF models, response variance and entropy only depend on measurement locations and not on response values.
 This means that there is limited flexibility for
@@
 \item Replicable experiments on synthetic cases with accompanying
-  code
+  code.
 \item Results from full-scale field trials running myopic strategies onboard an AUV for the characterization of a river plume. 
@@
 assumptions when deriving the sequential design criteria for adaptive
-sampling. In both sections properties of the methods are studied using
+sampling. In both sections, properties of the methods are studied using
 simulations. Section \ref{sec:case_study} demonstrates the methodology
@@
 \begin{supplement}
-%\sname{Supplement A}
-\stitle{Proof for proposition 1 \& 2}
+\stitle{Supplementary proofs and discussion on nugget sensitivity}
 \slink[doi]{COMPLETED BY THE TYPESETTER}
@@
 
-\section*{Appendix}
-\input{appendix}
+% Content moved to supplementary material (separate document/tex file).
+% \section*{Appendix}
+% \input{appendix}
 \end{document}

### 2021-03-11 main_web.tex
Updates from Overleaf

@@
 considered), and $\no=2$ with responses of temperature and salinity. A
-multivariate GRF model is assumed for $\gp$. To motivate concepts,
+multivariate GRF model is assumed for $\gp$. To illustrate these concepts,
 Fig.~\ref{fig:real_temp} and \ref{fig:real_sal} show a realization of such a
@@
 \centering 
-\subfigure[Temperature.]{\includegraphics[
+\subfigure[Temperature]{\includegraphics[
 height=0.27\textwidth,keepaspectratio]{Figures/intro_temp.png}\label{fig:real_temp}}
-\subfigure[Salinity.]{\includegraphics[
+\subfigure[Salinity]{\includegraphics[
 height=0.27\textwidth,keepaspectratio]{Figures/intro_sal.png}\label{fig:real_sal}}
-\subfigure[Regions of interest.]{\includegraphics[
+\subfigure[Regions of interest]{\includegraphics[
 height=0.27\textwidth,keepaspectratio]{Figures/intro_excu.png}\label{fig:jointex_roi}}
@@
 will allow us to think of $\gp$ as a scalar-valued random field indexed by $\domain \times \{1\dots,p\}$, which will give the co-Kriging equations a particularly simple form that parallels the one of univariate Kriging.
-The letters $\spatloc$ and $\ell$ will be used for spatial locations and response indices respectively.
+The letters $\spatloc$ and $\ell$ will be used for spatial locations and response indices, respectively.
 %
@@
 compute conditional means and covariances.  We will present a general
-form of co-Kriging, in the sense that it allows inclusion of several
+form of co-Kriging, in the sense that it allows for the inclusion of several
 (batch) observations at a time; observations at a given location

### 2021-03-11 main_web.tex
Updates from Overleaf

@@
 $$\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)^{-1} k(\bm{x}_{[n]}, \bm{x})
-$$
+$$,
 with $\bm{x}_{[n]}=(\bm{x}_1,\dots,
@@
 sensor readings, but it might not be diagonal with other types of
-combined measurements.  The matrix in parenthesis will be assumed to
+combined measurements.  The matrix in parentheses will be assumed to
 be non-singular throughout the presentation. The associated
@@
 
-Let us now consider the case where a co-Kriging prediction of $Z$ was
+Let us now consider the case where a co-Kriging prediction of $Z$ is
 made with respect to $n$ batches of generalized locations, concatenated again within $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$,
@@
 \begin{equation}\label{eq:meanCoK}
-\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
+\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})).
 \end{equation}
-where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional
+In this instance,  $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional
 sub-vector extracted from $\lambda_{[n+1]}(\bm{x})$ that corresponds
@@
 %k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})
-\lambda_{{[n+1,n+1]}}(\bm{x}'),
+\lambda_{{[n+1,n+1]}}(\bm{x}').
 \end{align}
 %
-As noted in \cite{Chevalier2015} in the case of scalar-valued fields,
-the update formulae naturally extends to universal Kriging in
+As noted in \cite{Chevalier2015}, in the case of scalar-valued fields,
+the update formulae naturally extend to universal Kriging in
 second-order settings and apply without Gaussian assumptions. We will
@@
 $\domain^2$. In contrast, the IBV of
-\cite{bect2019} involves solely an integral on $\domain$ and can be
+\cite{bect2019} solely involves an integral on $\domain$ and can be
 expanded as
@@
 let $\currentExp{.}$ and $\currentProba{.}$ denote conditional
-expectation and probability given the first $n$ batches of observations. We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.
+expectation and probability, respectively, given the first $n$ batches of observations. We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.
 
@@
 \end{equation*}
-with blocks given, for $i,j\in \{1,2\}$ and $u,v\in \domain$, by
+with blocks given for $i,j\in \{1,2\}$ and $u,v\in \domain$, by
 \begin{equation*}
@@
 Say we want to study the EP of a bivariate Gaussian
-vector variable below some thresholds. The thresholds are set equal to the mean;
+vector variable below some thresholds. The thresholds are set equal to the mean:
 $\mu_1=t_1=5^o C$ for temperature and $\mu_2=t_2=30$ g/kg for
@@
 this illustration, Table \ref{tab:sim_rhoab} shows that the expected
-Bernoulli variance gets smaller with larger standard deviations. The
-expected reduction of Bernoulli variance is further largest for the
+Bernoulli variance gets smaller with larger standard deviations. Furthermore, the
+expected reduction of Bernoulli variance is largest for the
 cases with high correlation $\gamma$. Albeit smaller, there is also
@@
 \end{align*}
-with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. In our examples, we only consider separable covariance models;
+where $\beta_0$ is a two dimensional vector and $\beta_1$ is a $2\times 2$ matrix. In our examples, we only consider separable covariance models:
 \begin{align*}
@@
 where an isotropic Mat\'{e}rn 3/2 kernel $(1+\eta h)\exp (-\eta h)$ is
-used, for Euclidean distance $h$.  In the accompanying Python examples
+used for Euclidean distance $h$.  In the accompanying Python examples
 taking place within the MESLAS
 toolbox~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}},
-these modeling assumptions can however be relaxed to anisotropic
+these modeling assumptions can, however, be relaxed to anisotropic
 covariance and changing variance levels across the spatial

### 2021-03-12 main_web.tex
Updates from Overleaf

@@
 Monitoring the world's oceans has gained increased importance in light
-of the changing climate and increasing anthropogenic impact. Central
+of the changing climate and escalating anthropogenic impact. Central
 to understanding the changes taking place in the upper water column is
@@
 using satellite remote-sensing provide large-scale coverage but have
-limited resolution, are limited to sensing the surface, and are
+limited resolution, are constrained to sensing the surface, and are
 impacted by cloud cover. Numerical ocean models similarly find it
@@
 Surveys with AUVs are usually limited to observations along fixed
-transects that are pre-scripted in mission plans created manually by a
-human operator. Missions can be specified operating on a scale of
+transects that are pre scripted in mission plans created manually by a
+human operator. Missions can operate on a scale of a scale of
 hundreds of meters to tens of kilometers depending on the scientific
@@
 salinity. Because of the local topography and the Coriolis force, the
-cold freshwater tends to flow east. Depending on the variations in
-river discharge, tidal effects, coastal current and wind, this
-boundary often gets distorted, and knowledge about its location is
-highly uncertain, making deterministic planning challenging. The goal
-is therefore to use AUV measurements for improved description of the
+cold freshwater tends to flow east.
+Depending on the variations in river discharge, tidal effects, coastal current, and wind, this boundary often gets distorted.
+As a result, knowledge about its location is highly uncertain, making deterministic planning challenging.
+The goal is therefore to use AUV measurements for improved description of the
 interface between fresh and oceanic waters, which in turn enables more direct sampling of biological dynamics. It is often not possible
@@
 fields are still in their infancy. \cite{LeGratiet.etal2015} use
-co-Kriging-based sequential designs to multi-fidelity computer codes
+co-Kriging-based sequential designs to develop multi-fidelity computer codes
 and \cite{Poloczek2017} use related ideas for multi-information
@@
 waypoint graph), has been used for surveys
-\citep{singh2009efficient,Binney2013} that focus largely on reducing
-predictive variance or entropy. These criteria are widely adopted in
+ that focus largely on reducing
+predictive variance or entropy \citep{singh2009efficient,Binney2013}. These criteria are widely adopted in
 the statistics literature on spatio-temporal design as well
@@
 vector-valued GRF on $\domain=[0,1]^2$. Fig.~\ref{fig:jointex_roi} represents a by-product of interest derived from these
-realizations, namely regions: i) in red, where both temperature and salinity are high (i.e., exceeding respective thresholds), indicative of ocean water, ii) in white, where both temperature and salinity are low, indicative of riverine water, and iii) in light-red, where one variable is above and the other below their respective thresholds, indicative of mixed waters.
+realizations, namely regions: i) in red, where both temperature and salinity are high (i.e., exceeding respective thresholds), indicative of ocean water, ii) in white, where both temperature and salinity are low, indicative of riverine water, and iii) in pink, where one variable is above and the other below their respective thresholds, indicative of mixed waters.
 
@@
 height=0.27\textwidth,keepaspectratio]{Figures/intro_excu.png}\label{fig:jointex_roi}}
-\caption{Realization of a bivariate GRF (display (a) and (b)) and ES above some threshold (c). Joint excursion in red and excursion of a single variable in light-red.}
+\caption{Realization of a bivariate GRF (display (a) and (b)) and ES above some threshold (c). Joint excursion in red and excursion of a single variable in pink.}
 \label{example_excu}
@@
 $\lambda_{[n]}(\bm{x})$ is a vector of weights equal to
-$$\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)^{-1} k(\bm{x}_{[n]}, \bm{x})
-$$,
+$$\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)^{-1} k(\bm{x}_{[n]}, \bm{x}),
+$$
 with $\bm{x}_{[n]}=(\bm{x}_1,\dots,
@@
 
-\subsection{UQ on ESs of multivariate GRFs}
+\subsection{Uncertainty quantification on ESs of multivariate GRFs}
 \label{sec:set_uq}
@@
 $\mu_1=t_1=5^o C$ for temperature and $\mu_2=t_2=30$ g/kg for
-salinity. We play with the temperature and salinity correlation
+salinity. We alter the temperature and salinity correlation
 and variances to study the effect on the EP and EBV.
@@
 potential next observation location and plot the EBV reduction (at
-each grid node in the waypoint graph) that would result from observing
-only one component of the field (temperature or salinity), or both at
-that point.
+each grid node in the waypoint graph) that would result from observing either one or both components of the field (temperature or salinity) at that point.
 
@@
 \centering 
-\subfigure[Regions of interest.]{\includegraphics[
+\subfigure[Regions of interest]{\includegraphics[
 height=0.21\textwidth,keepaspectratio]{Figures/ebv_comp_excu.png}\label{fig:ebv_comp_excu}}
-\subfigure[Temperature.]{\includegraphics[
+\subfigure[Temperature]{\includegraphics[
 height=0.22\textwidth,keepaspectratio]{Figures/ebv_comp_temp.png}\label{fig:ebv_comp_temp}}
-\subfigure[Salinity.]{\includegraphics[
+\subfigure[Salinity]{\includegraphics[
 height=0.22\textwidth,keepaspectratio]{Figures/ebv_comp_sal.png}\label{fig:ebv_comp_sal}}
-\subfigure[Both.]{\includegraphics[
+\subfigure[Both]{\includegraphics[
 height=0.22\textwidth,keepaspectratio]{Figures/ebv_comp_both.png}\label{fig:ebv_comp_both}}
@@
   collection locations in green. True excursion set in red. Places
-  where only one response is above threshold are depicted in light-red. EBV
+  where only one response is above threshold are depicted in pink. EBV
   reduction associated to observing one or both responses at the green
@@
 one associated with a horizontal (static east). Both expectations are
-computed according to the a-priori distribution of the GRF (i.e. no
+computed according to the a priori distribution of the GRF (i.e. no
 observations have been included yet).
@@
 \centering 
-\subfigure[Excursion probability.]{\includegraphics[
+\subfigure[Excursion probability]{\includegraphics[
 height=0.25\textwidth,keepaspectratio]{Figures/ebv_static_excu.png}\label{fig:ebv_static_excu}}
-\subfigure[Static north design.]{\includegraphics[
+\subfigure[Static north design]{\includegraphics[
 height=0.26\textwidth,keepaspectratio]{Figures/ebv_static_north.png}\label{fig:ebv_static_north}}
-\subfigure[Static east design.]{\includegraphics[
+\subfigure[Static east design]{\includegraphics[
 height=0.26\textwidth,keepaspectratio]{Figures/ebv_static_east.png}\label{fig:ebv_static_east}}
@@
 model is updated using co-Kriging
-Eq. \ref{eq:meanCoK} and \ref{eq:varCoK}, yielding a conditional law $\mathbb{P}_{[\stage+1]}$
+Eq. \ref{eq:meanCoK} and \ref{eq:varCoK}, yielding a conditional law $\mathbb{P}_{[\stage+1]}$,
 after which the process is repeated.
@@
 anticipate what the subsequent designs might offer beyond the first
-stage.  Based on the currently available data the myopic strategy
+stage.  Based on the currently available data, the myopic strategy
 selects the location that leads to the biggest reduction in EIBV:
@@
 \centering 
-\subfigure[Excursion realization.]{\includegraphics[
+\subfigure[Excursion realization]{\includegraphics[
 height=0.26\textwidth,keepaspectratio]{Figures/ebv_myopic_excu.png}\label{fig:real_myopic}}
 \hspace{2em}
-\subfigure[BV reduction.]{\includegraphics[
+\subfigure[BV reduction]{\includegraphics[
 height=0.26\textwidth,keepaspectratio]{Figures/ebv_myopic_ebv.png}\label{fig:route_myopic}}
 \hspace{2em}
-\subfigure[Expected reduction of BV.]{\includegraphics[
+\subfigure[Expected reduction of BV]{\includegraphics[
 height=0.26\textwidth,keepaspectratio]{Figures/ebv_myopic_radar.png}\label{fig:compass_myopic}}
@@
 associated to data collection at neighbouring nodes of the current location is
-shown in \ref{fig:compass_myopic}. The thick and light color indicates the node at $30 ^{\circ}$ to be the best possible choice.}
+shown in \ref{fig:compass_myopic}. The thick and light colored line indicates the node at $30 ^{\circ}$ to be the best possible choice.}
 \label{fig:ebv_myopic}
@@
 We now extend the myopic strategy by considering two stages of
-measurements, which is optimal in that it accounts consistently for
-the expectations and minimizations in these two stages, but no anticipation beyond that.
+measurements; this is optimal in that it accounts consistently for
+the expectations and minimizations in these two stages, but does not anticipate beyond that.
 
@@
 two-step look-ahead is to select the next observation location $\spatloc_{\stage + 1}$ that yields the biggest reduction in EIBV if
-we were to (optimally) add one more observation after that again. In order to formalize this concept, we must
+we were to (optimally) add one more observation. In order to formalize this concept, we must
 extend the notation for EIBV in the future (after observation
@@
                   \candidates(\spatloc)} \EIBV_{[\stage]}\left(\spatloc' ; \spatloc,
-      Y\right)\right]
+      Y\right)\right],
       \end{align}
       where $Y$ is the random data realization of $\gp_{\spatloc}$ according to its
-      conditional law at step $\stage$ with the
+      conditional law at step $\stage$, with the
       dependence of the set of candidates on the current location
@@
 \textit{static\_north}, \textit{static\_east}, and
-\textit{static\_zigzag} (a version of \textit{static\_north} where with some east-west transitions in a zigzag pattern) with the three described sequential approaches
+\textit{static\_zigzag} (a version of \textit{static\_north} with some east-west transitions in a zigzag pattern) with the three described sequential approaches
 \textit{naive}, \textit{myopic}, and \textit{look-ahead}. The static
@@
 approaches this is not the case. For comparison, the properties are
-therefore evaluated using Monte Carlo integration over several replicates
+evaluated using Monte Carlo integration over several replicates
 of realizations from the model while conducting simulated sequential

### 2021-03-12 main_web.tex
Updates from Overleaf

@@
 \centering 
-\subfigure[The waypoint graph.]{\includegraphics[width =
+\subfigure[The waypoint graph]{\includegraphics[width =
 0.49\textwidth]{Figures/web/wp_graph_paper.jpeg}\label{fig:wp_graph_a}}
 \hfill
-\subfigure[The waypoint graph in 3D.]{\includegraphics[width =
+\subfigure[The waypoint graph in 3D]{\includegraphics[width =
 0.49\textwidth]{Figures/sim/wp_graph_3d.pdf}\label{fig:wp_graph_b}}
@@
 trajectory choices over the $31\times31$ grid used to discretize the GRF. The AUV is set to start in node $53$.
-\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment.}
+\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment}
 \label{fig:wp_graph}
@@
 to update the GRF model. This is used in the evaluation of the next
-node to sample.  The procedure is run for $10$ stages. A total of 100
+node to sample.  The procedure is run for $10$ stages. In this experiment, a total of 100
 replicate simulations were conducted with all strategies.
@@
   \centering
-  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/web/avg_EV.jpeg}}
+  \subfigure[IBV]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/web/avg_EV.jpeg}}
   \hfill \subfigure[RMSE between estimated field and
-  truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
+  truth]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
   \hfill \subfigure[Explained variance
-  $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
+  $\bR^{2}$]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
   \hfill \subfigure[Computational time for inference (the lines for \textit{naive}, \textit{static\_north}, \textit{static\_east}, and \textit{static\_zigzag}
-  overlap).]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}}
+  overlap)]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}}
   \caption{Simulation results from 100 replicate simulations for 10
@@
 \textit{myopic} strategy is slower because it evaluates expected
-values for all candidate directions at the waypoints. But it is still
+values for all candidate directions at the waypoints; however, it is still
 able to do so in reasonable time, which allows for real-world
@@
 substantially better than the \textit{myopic} design only for very
-small initial standard deviations or very large spatial correlation
+small initial standard deviations or a very large spatial correlation
 range. 
 We also ran simulation studies with only temperature data, and for
-realistic correlation levels between temperature and salinity, the IBV
+realistic correlation levels between temperature and salinity; the IBV
 results are not much worse when only temperature data are
@@
 Table~\ref{tab:sim_rhoab}, the current setting includes spatial
-correlation and this likely strengthen the effect of temperature information. However, it seems that having temperature data alone
-does a substantially worse job in terms of explained variance.
+correlation, which likely strengthened the effect of temperature information. However, it seems that temperature data alone
+perform substantially worse in terms of explained variance.
 
@@
 structures.
-This gives a reasonable idea of the temperature and salinity of river and ocean waters.
+This provided a reasonable idea of the temperature and salinity of river and ocean waters.
 Both temperature and salinity were assumed to increase linearly with the west coordinate.
-Next, empirical correlation and variograms of the residuals from the regression analysis were analyzed to specify the temperature and salinity correlation, variances and spatial correlation decay parameter of the GRF model.
+Next, empirical correlation and variograms of the residuals from the regression analysis were analyzed to specify the temperature and salinity correlation, variances, and spatial correlation decay parameter of the GRF model.
 The resulting
@@
 \subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/web/auv4_es_sal_ep.jpeg}\label{fig:res2}}
-\caption{Results from mapping the Nidelva river, Trondheim, Norway
+\caption{Results from mapping the Nidelva river (Trondheim, Norway)
   over two survey missions. \ref{fig:map} shows an overview of the
   survey area overlaid with the AUV path in black and dashed
-  line. Note the shaded region indicating a typical frontal
+  lines. Note the shaded region indicating a typical frontal
   region. \ref{fig:res_both} shows the collected temperature data as
-  colored trails. Note waypoint 5 (WP5) which indicates where the two
-  surveys diverge. \ref{fig:res1} and \ref{fig:res2} shows the
+  colored trails. Note waypoint 5 (WP5), which indicates where the two
+  surveys diverge. \ref{fig:res1} and \ref{fig:res2} show the
   collected salinity data overlaid on the final EP, which indicate the
-  AUVs statistical impression of the front. For both missions the
+  AUV's statistical impression of the front. For both missions, the
   temperature and salinity data correspond with an indication of the
@@
 
-Two survey missions (1 and 2), were run successively, with a short
+Two survey missions (1 and 2) were run successively, with a short
 break in between. The resulting path of the selected waypoints are
@@
 slightly between the two surveys, the resulting trajectory (after
-waypoint 5) is shown to deviate. Survey 1 continued northwards,
+waypoint 5) also deviated. Survey 1 continued northwards,
 tracking the north-eastern portion of the front, while Survey 2 turned
@@
 the region and gradually bending off toward the north east. The amount
-of exploration done by Survey 1 which turned north is greater than
-Survey 2 which was coming close to the survey area borders in the
+of exploration done by Survey 1, which turned north, is greater than
+Survey 2, which came close to the survey area borders in the
 south-western corner. The AUV paths in both runs are relatively
@@
   \centering
-  \subfigure[Residual plot.]{\includegraphics[height = 0.26\textwidth]{Figures/res_diag_aNEW.eps}\label{fig:parest_a}}
+  \subfigure[Residual plot]{\includegraphics[height = 0.26\textwidth]{Figures/res_diag_aNEW.eps}\label{fig:parest_a}}
   \hfill
-  \subfigure[Quadratic form CDFs.]{\includegraphics[height = 0.26\textwidth]{Figures/res_diag_bNEW.eps}\label{fig:parest_b}}
+  \subfigure[Quadratic form CDFs]{\includegraphics[height = 0.26\textwidth]{Figures/res_diag_bNEW.eps}\label{fig:parest_b}}
   \hfill
-  \subfigure[Empirical variogram.]{\includegraphics[height = 0.252\textwidth]{Figures/res_diag_cNEW.eps}\label{fig:parest_c}}
+  \subfigure[Empirical variogram]{\includegraphics[height = 0.252\textwidth]{Figures/res_diag_cNEW.eps}\label{fig:parest_c}}
   \caption{Data analysis from AUV experiments. \ref{fig:parest_a} Residual plot of temperature and salinity
     along with specified Gaussian contours. \ref{fig:parest_b} Empirical CDF
-    (solid) of the quadratic form of the residuals along with the
-    modeled theoretical CDF (dashed) of the $\chi^2$ distribution with two
+    (solid line) of the quadratic form of the residuals along with the
+    modeled theoretical CDF (dashed line) of the $\chi^2$ distribution with two
     degrees of freedom. \ref{fig:parest_c} Scaled empirical variograms of the survey
@@
 
-Fig.~\ref{fig:parest} summarizes with diagnostic plots of this
+Fig.~\ref{fig:parest} summarizes diagnostic plots of this
 analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
 and salinity residuals after the regression mean values of salinity
-and temperature are subtracted from the data. This scatter-plot of
+and temperature are subtracted from the data. This scatter plot of
 joint residuals indicates that the correlation between the two
@@
 Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
-the residuals, and if the model is adequate they should be
+the residuals, and if the model is adequate, they should be
 approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
-the empirical CDF of the quadratic forms (solid) together with the
-theoretical CDF of the $\chi^2_2$ distribution (dashed). The empirical
+the empirical CDF of the quadratic forms (solid line) together with the
+theoretical CDF of the $\chi^2_2$ distribution (dashed line). The empirical
 and theoretical curves are not very different, which indicates that
@@
 scaled residuals for temperature and salinity. The decay is similar
-for the two, and seems to be negligible after about $150$ m as was
+for the two, and seems to be negligible after about $150$ m, as was
 specified. The working assumption of a separable covariance function
-is hence not unreasonable. There appears to be a slight nugget effect, but not much different from what was set as the measurement noise variance.
+is therefore not unreasonable. There appears to be a slight nugget effect, though it is not much different from what was set as the measurement noise variance.
 Overall, the spatial Gaussian model representation is a simplification
@@
 how observation practices can gain efficiency and accuracy from
-statistical techniques for spatial monitoring and demonstrate the the
+statistical techniques for spatial monitoring and demonstrate the
 need for real-time multivariate spatial sampling on autonomous
@@
 The case study considers the upper water column in the river plume,
-represented by a two dimensional grid. Extensions to three-dimensional
+represented by a two-dimensional grid. Extensions to three-dimensional
 domains are not methodologically different, but likely approximate
@@
 vicinity of the autonomous vehicle \citep{fossum18b}. While we did not
-consider any temporal effects, which would be relevant on a larger
+consider any temporal effects, they would be relevant on a larger
 time scale. We do consider the extension to spatio-temporal modeling
@@
 need to be extended to non-Gaussian phenomena, possibly feature-based
-mixtures of Gaussian processes which could potentially be run onboard
+mixtures of Gaussian processes which could potentially be run onboard and
 augmented by dynamical models. Running numerical models onboard a
@@
 for instance, hybrid or multi-attribute criteria that could balance
-goals of exploration and exploitation in this situation. Equally, such
+goals of exploration and exploitation in this situation. Similarly, such
 techniques have significant use cases in downstream decision-making,
@@
 analysis \citep{Eidsvik:15} could be used to evaluate whether
-information is likely to result in improved decision-making, in such a
+information is likely to result in improved decision-making in such a
 context.  We also foresee opportunities related to design of
@@
 In our context the myopic strategy performs well, and due to
-computational constraints we did not go in depth on the dynamic
+computational constraints, we did not go in-depth on the dynamic
 programming solutions. There has been much work on finite horizon
-optimization in the robotics literature including probabilistic road
+optimization in the robotics literature, including probabilistic road
 maps and rapidly-exploring random trees \citep{karaman2011sampling},
@@
 The value of information related to when and what to communicate (to
-shore or to other vehicles) is also an interesting thrust for research
+shore or to other vehicles) is also an interesting avenue for future research
 and likely to be useful for internet-of-things applications or

### 2021-03-15 main_web.tex
Updates from Overleaf

@@
 performance for criteria such as RMSE as well. Another non-statistical
-criterion that is important for practical purposes is the computational
-time needed for the strategy.
+criterion that is important for practical purposes is the computational time needed for the strategy (i.e. avoiding longer periods of halting during execution).
 
@@
 applicability. The \textit{look-ahead} strategy is much slower, reaching levels that
-are nearly impractical for execution on an AUV. Some pruning of the
+are largely impractical, for execution on an AUV. Some pruning of the
 graph is performed to improve the performance, such as ruling out
@@
 the upper water column. It has a multicore GPU NVIDIA Jetson TX1
-(quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4
-GB of 1600 MHz DRAM) for computations onboard.  The sampling algorithm
+(quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM) for computations onboard.  The sampling algorithm
 was built on top of the autonomous Teleo-Reactive EXecutive
@@
 shown in the map in Fig.~\ref{fig:map}, both within the expected
-frontal region (shaded light-red). The recorded temperatures are shown as
+frontal region (shaded pink). The recorded temperatures are shown as
 colored trails in Fig.~\ref{fig:res_both}, clearly indicating the
@@
 conditional EPs in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2} as
-dashed lines, correspond with one another. In both surveys they yield
-a picture of the front being to the west in the southern portions of
+dashed lines, correspond with one another. In both surveys, the predictions yield a picture of the front being to the west in the southern portions of
 the region and gradually bending off toward the north east. The amount
@@
 The spatio-statistical design criterion building on random sets is
-relevant in our setting with different water properties.  We show
-mathematical generality beyond the expected integrated Bernoulli
-variance, for instance, that of volume uncertainties which is possibly
-more relevant, but one that requires more computational resources.
+relevant in our setting with different water properties.
+We show mathematical generality beyond the expected integrated Bernoulli variance; for example, we demonstrated the generality of volume uncertainties, which is possibly more relevant to sampling strategies in the ocean, but which requires more computational resources.
 Such criteria could be particularly useful in other oceanographic

### 2021-03-16 main_web.tex
Updates from Overleaf

@@
 strategies for our vast oceans, we combine spatial statistics, 
-experiment design, and marine robotics in this work.  The
+experimental design, and marine robotics in this work.  The
 multidisciplinary efforts enable information-driven data collection in
@@
 Surveys with AUVs are usually limited to observations along fixed
-transects that are pre scripted in mission plans created manually by a
-human operator. Missions can operate on a scale of a scale of
+transects that are pre-scripted in mission plans created manually by a
+human operator. Missions can operate on a scale of
 hundreds of meters to tens of kilometers depending on the scientific
@@
 $\mes(\es)$ through its moments.  Centered moments may be computed
-using Proposition~\ref{propo1} developed in the Supplementary Material \ref{suppA}.  In
+using Proposition 3 developed in the Supplementary Material \citep{suppA}. In
 particular, as an integral over EPs, the 
@@
 
-\subsection{Expected IBV and EMV}
+\subsection{Expected IBV (EIBV) and EMV}
 \label{sec:eibv}
@@
 The EIBV is efficiently computed for each of the candidate points
-$\candidates$ using Proposition \ref{propo2}. 
+$\candidates$ using Proposition 4 developed in the Supplementary Material \citep{suppA}. 
 Even though this myopic strategy is non-anticipatory, it still
@@
 expectation is solved using the closed-form expressions for EIBV
-    provided by Proposition \ref{propo2}, now with conditioning on the
+provided by Proposition 4 developed in the Supplementary Material \citep{suppA}, now with conditioning on the
 first stage data already going into the co-Kriging updating
@@
 improved. However, we do not expect that minor changes in parameters would have caused much
-difference in AUV sampling paths.
+difference in AUV sampling paths. (See also Supplementary Material \citep{suppA}.)
 

### 2021-03-16 main_web.tex
Updates from Overleaf

@@
 %
-will allow us to think of $\gp$ as a scalar-valued random field indexed by $\domain \times \{1\dots,p\}$, which will give the co-Kriging equations a particularly simple form that parallels the one of univariate Kriging.
+will allow us to think of $\gp$ as a scalar-valued random field indexed by $\domain \times \{1\dots,p\}$, which will give the co-Kriging equations a particularly simple form that parallels the form of univariate Kriging.
 The letters $\spatloc$ and $\ell$ will be used for spatial locations and response indices, respectively.
@@
 used by the AUV is focused on reducing the
-EIBV, but we nevertheless expect that we will achieve good predictive
-performance for criteria such as RMSE as well. Another non-statistical
+EIBV, but we nevertheless anticipate that we will achieve good predictive
+performance for other criteria such as RMSE as well. Another non-statistical
 criterion that is important for practical purposes is the computational time needed for the strategy (i.e. avoiding longer periods of halting during execution).
